### 2017-12-20 Primer_Semestre/Poster_Dic17/Poster_eng.tex
primer commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Landscape Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}
\usepackage[utf8]{inputenc}
%\usepackage[spanish, mexico]{babel}
\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\usepackage{natbib}
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors
\usepackage{pgfgantt}
\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{graphicx} % Required for including images
\graphicspath{{figures/}} % Location of the graphics files
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{amsfonts, amsmath, amsthm, amssymb} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures

\begin{document}

%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into three boxes:
% The first is 55% wide and houses the title, subtitle, names and university/organization
% The second is 25% wide and houses contact information
% The third is 19% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[b]{0.45\linewidth}
\begin{flushleft}
\LARGE \color{NavyBlue} \textbf{ Acquiring causal models via reinforcement learning} \color{Black}\\ % Title
\Large\textit{From reinforcement learning to computational psychology}\\ % Subtitle
%\small \textbf{Mauricio Gonzalez Soto}\\ % Author(s)
%\small \textbf{Asesor: Dr. Hugo Jair Escalante}\\
%Email: \texttt{mauricio@inaoep.mx}
\end{flushleft}
\end{minipage}
%
\begin{minipage}[b]{0.20\linewidth}
\begin{center}
\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Comentarios/Figures/logos_2.png}
\end{center}
\end{minipage}
%
\begin{minipage}[b]{0.30\linewidth}
\begin{center}
\large \textbf{Mauricio Gonzalez Soto}\\ % Author(s)
\large \textbf{Asesor: Dr. Hugo Jair Escalante}\\
Email: \texttt{mauricio@inaoep.mx}
\end{center}
%\includegraphics[scale=0.2]{/Users/MauricioGS1/INAOE/Comentarios/Figures/inaoe_logo.jpg}
\end{minipage}

\vspace{0.3cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{3} % This is how many columns your poster will be broken into, a poster with many figures may benefit from less columns whereas a text-heavy poster benefits from more

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\color{Navy} % Navy color for the abstract

\begin{abstract}
Reinforcement Learning algorithms have seen a huge development in recent years. Standard algorithms only provide action-outcome mappings which are difficult to use in order to make high-level reasoning about the environment. The question adressed here is how can an agent learn a casual model in terms of his direct experience with the environment. It is proposed to develop a learning-by-interaction algorithm that also learns a causal model which is used on-line to make decisions. 
\end{abstract}
%----------------------------------------------------------------------------------------
%	INTRODUCCIÓN
%----------------------------------------------------------------------------------------
\color{SaddleBrown} % SaddleBrown color for the introduction
\section*{Introduction}
While playing a videogame or learning how to drive, human beings are forced to \textbf{make decisions} and to \textbf{actively intervene} in the world.\\
\\
This active \textbf{decision-making} process modifies the state of the world, which allows us to learn about \textbf{causal relations} that hold in our environment\\
\\
We ask ourselves how an \textbf{intelligent agent} can acquire and \textbf{use causal knowledge} while engaging with its environment to achieve some task.\\
\\
The goal is to learn a \textbf{causal decision model} and not only an action-outcome mapping.
%\begin{itemize}
%\item Videogames: after making some decision and we later will use that knowledge to predict outcomes while facing similar situations  (\cite{meder2008inferring}, \cite{hagmayer2009decision}).
%\item Driving: \cite{danks2014unifying} gives the example that while learning how to drive, we infer that moving the steering wheel will \textit{cause} a reacion in the car.
%\end{itemize}
%----------------------------------------------------------------------------------------
%	TRABAJO RELACIONADO
%----------------------------------------------------------------------------------------

\color{DarkSlateGray} % DarkSlateGray color for the rest of the content

\section*{Related work}
\subsection*{General Background}
\begin{itemize}
\item Causal graphical models (\cite{pearl2009causality}, \cite{koller2009probabilistic}): Causal semantics over probabilistic graphical models.
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_example.png}
\end{center}
\item Reinforcement Learning (\cite{sutton1998reinforcement}, \cite{gershman2015reinforcement}): An agent learns through exploration and interaction with environment.
\begin{center}
\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/RL.png}\footnote{Image taken from \cite{sutton1998reinforcement}}
\end{center}
\end{itemize}
\subsection*{Where are we standing?}
\begin{itemize}
\item Causal decision theory (\cite{joyce1999foundations}): Making decisions using causal knowledge; i.e. I know that $A$ \textit{causes} $B$, so it can be predicted what will happen if $A$ is chosen, or not, over other options. 
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_Decision.png}\footnote{Image taken from \cite{hagmayer2013repeated}}
\end{center}
\item Causal learning through sequential decision making (\cite{hagmayer2013repeated}): Causal learning in humans has been studied in the context of sequental decision-making
\begin{center}
\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Hagmayer_rats.png}\footnote{Image taken from \cite{hagmayer2013repeated}}
\end{center}
\item On-line learning in causal models(\cite{wellen2012learning}), \cite{garnelo2016towards}):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/LPL_model.png}\footnote{Image taken from \cite{wellen2012learning}}
\end{center}
\item Decision problems and causal models: \cite{dawid2002influence} show how causal Bayesian Nets can be augmented with decision nodes as to model decision problems that use causal information.
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Influence_diagram.png}\footnote{Image taken from \cite{dawid2002influence}}
\end{center}
\end{itemize}
%And we can think about these areas relating in the following way:
%\begin{center}
%\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/diagrama_eng.png}
%\end{center}

%----------------------------------------------------------------------------------------
%	MOTIVACIÓN
%----------------------------------------------------------------------------------------

\section*{Motivation}
It has been shown by \cite{garnelo2016towards} that causal structure can be learnt from the interaction of an agent with his environment while it learns to perform some task. \\
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Garnelo.png}\footnote{Image taken from \cite{garnelo2016towards}}
\end{center}
Instead of a fully expresive casual graphical model, they use first-order logic which is somehow restrictive while showing the importance of having some semantics capable of managing causal structure.\\
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Graph_vs_logic.png}
\end{center}
Meanwhile \cite{wellen2012learning} show how a causal graphical model can be learned on-line by a decision maker through local prediction-error learning.\\ 
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/LPL_example_2.png}
\end{center}
From the human psychology side \cite{hagmayer2013repeated} show that our intuitions aren't flawed, because human beings acquire and use causal knowledge while engaging in simple decision-making.
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_Feedback.png}
\end{center}
%\\
%\cite{hagmayer2013repeated} asked what would happen with an agent that must learn some particular task instead of only attempting to directly learn causal structure. \\
Each decision made by the agent can be interpreted as an intervention in his environment (\cite{hagmayer2009decision}) which provides information about the world's causal structure.\\
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Golf.png}
\end{center}
Learning causal structure will allow the agent to make better decisions because it will be able to consider \textbf{consequences} of his actions, and \textbf{not only reacting} to rewards.\\
\begin{center}
\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/jail.png}
\end{center}
Therefore, we ask ourselves how an agent would learn a causal model in terms of his actions and their consequences? How would this knowledge could be used to make better decisions in the future?
\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out
%----------------------------------------------------------------------------------------
%	PROPUESTA
%----------------------------------------------------------------------------------------
\section*{Proposal}
\subsection*{Research question}
Can \textbf{Reinforcement Learning} methods capture a \textbf{causal model} of the environment rather than just an action-outcome mapping as to make \textbf{better decisions}?
\subsection*{Objectives}
\begin{itemize}
\item Defining how to \textbf{learn} causal relations in an \textbf{on-line} setting in terms of actions taken by an agent and their outcomes.
\item Defining how to \textbf{incorporate} the current causal knowledge in order to choose \textbf{better actions}.
\item Defining a way to \textbf{evaluate} how the causal knowledge was used in the decision-making process. We do not want to lose the \textbf{interpretability} of causal graphs.
\item Defining how to \textbf{extract} the acquired causal knowledge to use it in other domains.
\end{itemize}
\subsection*{Proposed Methodology}
My proposal is about incorporating a \textbf{causal model} inside a sequential decision making problem in which an agent has to maximize some reward by \textbf{selecting actions}.\\
\begin{center}
\includegraphics[scale=0.9]{/Users/MauricioGS1/INAOE/Comentarios/Figures/choose_actions.png}
\end{center}
The agent will \textbf{interact} with its environment and receive a \textbf{feedback} signal based on the actions taken by him, which will modify his current causal knowledge.\\
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/cry.png}
\end{center}
%Using this information, as well as the \textbf{new state} of the world, the internal \textbf{causal model} of the agent will be modified (if needed), either adding or deleting edges or modifying the \textit{causal strength} associated.\\
%\\
Before the next step is taken, the current \textbf{causal knowledge} will be taken into account when \textbf{choosing the next action} together with the current value of each posible action.\\ 
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/baby.png}
\end{center}
This new action and its outcome will again modify the causal knowledge, which will improve the next action and so on.
\begin{center}
\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Proposed.png}
\end{center}
\subsection*{Evaluation}
How is this better? Why should we care about learning anything extra if the task is achieved?\\
\\
Causal structures learnt by the agent can be evaluated if compared with an external \textit{real} causal model, or comparing with the initial \textit{prior} causal graph as made by  \cite{hagmayer2013repeated}.
\subsection*{Expected contributions}
\begin{itemize}
\item A framework in which an agent set to learn some task by action-reward can learn causal structure about its environment.
\item An algorithm that learns causal structure in an on-line way.
\item A decision-making procedure that uses current causal knowledge together with action-value estimates.
\end{itemize}
\subsection*{Expected time-line}
\begin{center}
\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Gantt.png}
\end{center}
%------------------------------------------------------------------------------------------
% BIBLIOGRAPHY
%------------------------------------------------------------------------------------------
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Comentarios/Bibliografia.bib}
\end{multicols}
 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------
%\begin{minipage}[b]{0.55\linewidth}
%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
%\bibliographystyle{apalike}
%\bibliography{/Users/MauricioGS1/INAOE/Comentarios/Bibliografia.bib}
%\end{minipage}
%\begin{minipage}[b]{0.50\linewidth}
%\subsection*{Expected contribution}
%\begin{itemize}
%\item Build a framework in which an agent set to learn some task by action-reward can learn causal structure about its environment.
%\item Algorithm that learn causal structure in an on-line way.
%\item Using causal knowledge in each step to have better insights in the next decision.
%\end{itemize}
%\subsection*{Expected time-line}
%\begin{itemize}
%\item May 2018: Formal proposal.
%\item August 2018: Proposal defense.
%\%item December 2018: On-line causal learning module
%\end{itemize}
%\end{minipage}
\end{document}

### 2017-12-20 Primer_Semestre/Propuesta/Propuesta.tex
primer commit

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{natbib}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Survey}
\author{Mauricio Gonzalez Soto}
\begin{document}
\nocite{*}
\maketitle
\tableofcontents
\newpage
\section{Introducción}
El problema al cual nos enfrentamos consiste en el aprendizaje de relaciones causales a partir de datos. Los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos. Al tratar de averiguar relaciones causales, es común contar con información previa sobre el fenómeno dentro del cual se quiere investigar causalidad, por lo que recurrimos a la estadística bayesiana para el manejo de esta información inicial.\\
\\
La primera parte de este resumen consiste en Redes Bayesianas, que son objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red. Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
\\
Veremos que al intentar obtener relaciones causales a partir de información proveniente del mundo real necesitamos obtener las entidades que se relacionan entre sí de manera causal, por lo que haremos un repaso de técnicas utilizadas en extracción de relaciones.\\
\\
Por otro lado, en la literatura proveniente de Psicología Computacional y Ciencias Cognitivas, se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro. Se sabe también que el conocimiento causal adquirido es utilizado para mejorar las futuras decisiones tomadas.\\
\\
Finalmente, se discutirá un poco sobre Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno. Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal. Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones. La pregunta se convierte entonces en cómo obtener e incorporar conocimiento causal en un problema de aprendizaje por refuerzo. Además, uno se pregunta si es posible utilizar este conocimiento causal obtenido al interactuar sobre el mundo para realizar otro tipo de inferencias causales en ciertos dominios de interés.
\section{Causalidad}
\subsection{Intro}
Saber si es posible hablar de relaciones causa-efecto es un tema polémico; los métodos probabilistas tradicionales sólo codifican asociaciones estadísticas, pero la idea
La diferencia entre modelos probabilistas y modelos causales se da cuando es de interés tomar acciones que modifiquen el entorno y no nada más observar los valores de las variables. \\
Según \cite{pearl2014probabilistic}, las dos preguntas fundamentales en Causalidad son:
\begin{itemize}
\item ¿Qué evidencia empírica es necesaria para hacer inferencias válidas sobre relaciones causa-efecto?
\item Dado que estamos dispuestos a aceptar información causal sobre un fenómeno, ¿qué inferencias podemos obtener a partir de esta información, y cómo?
\end{itemize}
En cuanto al segundo punto, Pearl considera que hay tres tipos de preguntas a considerar
\begin{itemize}
\item Predicciones: si sabemos que una persona ingirió alimentos de dudosa procedencia, ¿se enfermará?
\item Intervenciones: si le damos cierto tratamiento a un paciente, ¿mejorará?
\item Contrafactuales: ¿el paciente del punto uno se habría enfermado de todos modos si no hubiera consumido esos alimentos?
\end{itemize}
Pearl argumenta que aunque causalidad suele entenderse como afirmaciones necesarias, estas son hechas en contexto de incertidumbre, por lo que cualquier teoría causal debe estar en términos de probabilidades.
\subsection{Un poco de historia}
Aristóteles definió 4 \textit{causas} para una cosa, una de ellas llamada \textit{causa eficiente} que estaba definida como aquello que hace la cosa.\\
\\
En 1690, el filósofo inglés John Locke propuso como definición de causa “aquello que produce alguna idea simple o compleja”, y lo que es producido lo definió como efecto.\\
\\
\cite{holland1986statistics} resalta la importante diferencia en que Aristóteles hizo énfasis en las causas de una cosa, pero no en el efecto de estas. \\
\\
Fue David Hume quien se considera que hiz una importante contribución, y él definió la causalidad como una relación entre experiencias más que una entre hechos. Argumentó que no es verificable empíricamente que la causa produzca el efecto, sino sólo que el evento llamado \textit{causa} es invariablemente seguido por aquella experiencia llamada \textit{efecto}.\\
\\
Hume definió 3 criterios básicos para causalidad:
\begin{itemize}
\item Contiguidad espacio-temporal: tanto causa y efecto tienen lugar en una entidad común (espacio y tiempo).
\item Sucesión temporal: La causa tiene que preceder temporalmente al efecto. 
\item Conjunción constante: Causa y efecto ocurren juntos, o no ocurren juntos.
\end{itemize}
John Stuart Mill argumentó tajantemente que observación sin experimentos no puede probar causalidad. Mill identificó cuatro métodos generales para la identificación de relaciones causales. 
\subsection{Modelo de Rubin}
El modelo de Rubin parte de una población de \textit{unidades} $U$. En general, estas unidades son los objetos de estudio sobre las cuales actuan las causas. Para hacer más sencilla la descripción del modelo, supondremos que sólo existen dos causas (o niveles de tratamiento) denotadas por $t$ y $c$. Sea $S$ una variable indicadora de la causa a la cual cada unidad $u \in U$ fue expuesta. Además, consideramos dos variables respuesta $Y_t$, $Y_c$. $Y_t(u)$ reoresenta el valor de la respuesta si la unidad $u$ es expuesta a $t$ y análogo para $c$.\\
\\
\begin{defi}
Se define el efecto de la causa $t$ sobre $u$ relativo a la causa $c$ como
\[ Y_t(u) - Y_c(u). \]
Esta expresión se lee de la siguiente manera: $t$ causa el efecto $Y_t(u) - Y_c(u)$ en la unidad $u$ (relativo al control $c$).
\end{defi}
Notemos que es imposible observar simultáneamente el efecto de $t$ y de $c$ sobre una misma unidad $u$ y a esto \cite{holland1986statistics} le llama el Problema Fundamental de la Inferencia Causal. El problema, según explica Holland, radica en la palabra \textit{observar} pues no podemos saber simultáneamente el efecto de exponer a un enfermo en particular a un nuevo tratamiento y al mismo tiempo el efecto de no exponerlo.
\subsection{Modelo de Suppes}
En su libro “Probabilistic Theory of Causality”, (\cite{suppes1970probabilistic}), Patrick Suppes intenta mejorar el análisis de Hume sobre causalidad; en particular, la condición de conjunción constante. Para Suppes, no importa qué sean las causas y los efectos, sino que se puedan expresar como eventos que ocurren bajo ciertas condiciones. Suppes supone que todas las causas necesariamente preceden en el tiempo a sus efectos. Define una causa \textit{prima facie} de un evento como un evento que precede temporalemnte a este y que está asociado de manera positiva con él. Define como una causa espuria a una causa prima facie tal que es condicionalmente independiente del efecto dada otra causa la cual es anterior en el tiempo a la causa prima facie y que está asociada con el evento dada la causa. Una causa genuina es una causa prima facie que no es espuria.
\begin{defi}
Si $r < s$ representan valores de tiempo, se dice que el evento $C_r$ es una causa prima facie del evento $E_s$ si
\[ P(E_s | C_r ) > P(E_s). \]
\end{defi}
\begin{defi}
Un evento $C_r$ se dice que es una causa espuria de un eventi $E_s$ si $C_r$ es una causa prima facie de él y para alguna $q < r < s$ existe un evento $D_q$ tal que 
\[P(E_s | C_r, D_q) = P (E_s | D_q) \]
y además,
\[P(E_s | C_r , D_q) \geq P(E_s | C_r). \]
\end{defi}
Por lo tanto,
\begin{defi}
Para tiempos $r < s$, un evento $C_r$ se dice que es una causa genuina de un evento $E_s$ si $C_r$ es una causa prima facie de $E_s$, pero no es una causa espuria.
\end{defi}
Notemos que Suppes está definiendo la causa de un efecto, más que el efecto de una causa. Además, no hay restricciones a la naturaleza de una causa más allá de que sea un evento que ocurre antes en el tiempo. El modelo de Suppes no puede expresar el efecto de una causa en particular, sólo comportamientos promedio (\cite{holland1986statistics}).\\
\\
En palabras de Holland, la causalidad de Suppes no es más que correlación entre causa y efecto que no puede ser eliminada al parcializar causas legítimas.
\subsection{Modelos Causales en Medicina}
Un caso de aplicación en medicina es determinar que ciertas bacterias causan ciertas enfermedades. Existe un conjunto de postulados planteados por Koch para decidir cuándo un micro-organismo está implicado en una enfermedad. Aunque no existe un planteamiento formal, estos pueden expresarse como (\cite{yerushalmy1959methodology}):
\begin{itemize}
\item El organismo debe estar presente en todos los casos de la enfermedad supuestamente provocada por él.
\item El organismo debe aislarse de los pacientes y cultivarse en estado puro.
\item El organismo cultivado en estado puro debe reproducir el padecimiento. 
\end{itemize}
Estos postulados aseguran que no haya datos que apoyen una hipótesis nula de cero efecto causal. \\
\\
Por otro lado, Hill, quien fue de los primeros en estudiar la asociación entre fumar y cáncer, encontró 9 factores útiles a la hora de decidir que una asociación observada es causal. 
\subsection{Causalidad de Granger en Economía}
La causalidad de Granger se utiliza para series de tiempo, lo cual es de alta importancia en Economía. \cite{holland1986statistics} opina que las ideas esenciales de Granger no están limitadas a series de tiempo y que pueden ser expresadas en términos de variables que actuan sobre unidades de una población.\\
\\
En la teoría de Granger, una variable causa otra variable; es decir ( \cite{holland1986statistics}), las realizaciones de una variable mejoran las predicciones hechas sobre la otra variable. La única parte en la que se utilizan series de tiempo en esta definición es para hablar de valores de las variables definidos hasta un cierto tiempo $t$.\\
\\
Si $X,Y,Z$ son variables aleatorias definidas sobre cierta población, se dice que $X$ no es una causa de Granger de $Y$ si $X$ y $Y$ son condicionalmente independientes dado $Z$. Entonces, $X$ es una causa de Granger de $Y$ si distintos valores de $X$ llevan a distintas distribuciones predictivas de $Y$ dados $X$ y $Z$; esto es, $X$ ayuda a predecir $Y$ aun cuando $Z$ es tomada en cuenta. La no-causalidad de Granger es parecida a las causas espurias de Suppes pues ambas consideran que no se pueda predecir un evento dada cierta información.\\
\\
En opinión de Holland, la causalidad de Granger falla de la misma manera que la de Suppes, pues a la luz de nueva información una causa Granger puede dejar de serlo.
\subsection{Modelos Causales en Ciencias Sociales}
\subsection{Aprendizaje de Modelos Causales a partir de datos}
En los métodos de estado del arte de Modelos Causales Funcionales suele hacerse los siguientes supuestos (\cite{goudet2017learning}):
\begin{itemize}
\item Suficiencia Causal (CSA)
\item Propiedad Markoviana (CMA)
\item Independencias Condicionales
\item Fiabilidad causal (CFA)
\end{itemize}
\subsubsection{Aprender el CPDAG}
Existen tres familias de métodos para extraer el grafo: métodos de restricciones, métodos de score y métodos híbridos.\\
\\
Los métodos basados en restricciones explotan las independencias condicionaes para identificar todas las v-estructuras. El algoritmo más conocido es el Algoritmo PC (\cite{spirtes2000causation}). El algoritmo PC primero construye un esqueleto basado en independencias condicionales; luego, identifica las v-estructuras y posteriormente orienta las aristas. El algoritmo FCI (\cite{spirtes1999algorithm}) extiende el algoritmo PC al relajar la suficiencia causal e incorporar variables latentes. El algoritmo RFCI de \cite{colombo2012learning} es más rápido que FCI y maneja datos de alta dimensionalidad. La debilidad de estos métodos, según \cite{goudet2017learning} es que dependen demasiado en pruebas de independencia, que dependen a su vez de la cantidad disponible de datos.\\
\\
Los métodos basados en score exploran el espacio de CPDAG's y minimizan un score global. La exploración se realiza en términos de operaciones sobre grafos, por ejemplo añadir o remover aristas. El algoritmo Greedy Equivalent Search (GES) de \cite{chickering2002optimal} encuentra la estructura óptima en términos del BIC.\\
\\
Los métodos híbridos combinan las técnicas anteriores. Por ejemplo el Max-Min Hill Climbing Algorithm de \cite{tsamardinos2006max} primero construye un esqueleto utilizando pruebas de independencia condicional, y luego le da orientación a las aristas utilizando un método greedy sobre un score bayesiano.
\subsubsection{Explotando la asimetría entre causa y efecto}
Los métodos anteriores no toman en cuenta la asimetría entre causas y efectos. La intuición detrás (\cite{goudet2017learning}) está dada por el siguiente ejemplo: Si consideramos el modelo causal funcional $Y=X+U$, entonces los métodos anteriores basados en grafos no pueden darle orientación a la arista $X - Y$, pues tanto $X \rightarrow Y$ como $Y \rightarrow X$ son Markov equivalentes.\\
\\
\cite{stegle2010probabilistic} infiere la dirección causal al construir dos modelos Bayesianos generativos: uno para $X \rightarrow Y$ y otro para $Y \rightarrow X$. La dirección causal es determinada a partir del modelo generativo que mejor ajuste a los datos. Otros métodos, como el propuesto por \cite{sgouritsa2015inference} se basan en que si $X \rightarrow Y$, entonces la probabilidad marginal de la causa $P(X)$ es independiente del mecanismo causal $P(X | Y)$; por lo tanto, estimar $P(Y | X)$ a partir de $P(X)$ no debería poderse, pero estimar $P(X | Y)$ a partir de $P(Y)$ sí debería ser posible.
\subsubsection{Como problema de Aprendizaje}
Inspirados en los concursos de Kaggle de \cite{Guyon2013}, \cite{lopez2015towards} plantean el problema de inferencia causal como un problema de clasificación de medidas de probabilidad. Los datos provistos por el concurso son 16,200 parejas de variables $(X_i,Y_i)$. Cada par consta de una realización $(x_i,y_i)$ y una etiqueta $l_i$ que indica el valor de la verdadera relación causal; es decir, $l_i$ indica uno de los siguientes casos: $X \rightarrow Y,$ $Y \rightarrow X$, $X_i \perp Y_i$, $X \leftrightarrow Y$.\\
\\
Notemos que este método tiene como punto débil que depende de la representatividad de la muestra de entrenamiento y que supone la existencia de una distribución madre
\subsubsection{Modelos Causales Funcionales Neuronales}
\cite{goudet2017learning} proponen un esquema que consiste en tres pasos: primero, identificar el esqueleto según propiedades de independencia estándar; luego, orientar las aristas utilizando métodos bivariados estándar y finalmente hacer una optimización sobre todo el grafo examinando relaciones locales. Este último paso se realiza al aprender estos modelos locales como redes neuronales generativas; es decir, se modela la distribución $P(X | X_{Pa_i}, U_i)$ de un nodo condicional en sus padres (causas) con una red neuronal.\\
\\
Formalmente, se modela cada ecuación
\[X_i = f_i (X_{Pa_i},U_i) \]
como una red neuronal generativa global, en la cual cada función $f_i$ es modelada por una red neuronal generativa con una capa oculta.\\
\\
Encontrar un modelo causal de este tipo requiere llevar a cabo dos tareas: estimar la estructura (el grafo) de los padres de cada variable, y las redes neuronales $f$
\subsection{Correlación y causalidad}
Dos variables aleatorias $X$,$Y$ pueden estar correlacionadas en distintos casos; por ejemplo, si $X$ causa $Y$, o si tanto $X$ como $Y$ son causadas por otra variable desconocida $V$. Si e conoce la existencia de $V$ y además podemos observarla. 
\section{Modelos gráficos causales}
Sabemos que en una red bayesiana los arcos, aunque dirigidos, no necesariamente tienen una interpretación directa, lo único que importa es que el grafo capture adecuadamente las relaciones de dependencia entre variables aleatorias. Nos gustaría que fuera el caso que $X \to Y$ significara $X$ \textit{causa} Y. Resulta que existe una familia de modelos causales en los cuales se \textit{interviene} en el mundo. La idea clave es utilizar \textit{intervenciones ideales}, las cuales se denotan $do(Z=z)$ que significa forzar a la variable aleatoria $Z$ a tomar el valor $z$, denotado de manera simple como $do(z)$. Esta operación corresponde al caso en que un agente directamente modificó el estado del mundo y no sólo observó los valores de este de manera pasiva; por esta razón es muy importante tener en mente que es distinto $P(Y | do(z), X=x)$ que $P(Y | Z=z, X=x)$. Las preguntas causales de interés son planteadas de la forma $P(X | do(z))$ y son llamadas \textit{quries intervencionales}. Este tipo de intervenciones surgen de manera natural en muchos problemas; por ejemplo, “si el paciente recibe este tratamiento, cuál es la probabilidad de sobrevivir la enfermedad” puede ser planteada como $P(sobreivir | do(tratamiento))$.\\
\\
Formalmente, un modelo gráfico causal (\cite{koller2009probabilistic}, \cite{pearl2014causality}) tiene la misma forma que una red Bayesiana; es decir, consiste en una red acíclica dirigida que codifica una distribución entre un conjunto de variables aleatorias. La diferencia radica en la interpretación de las flechas, pues cada variable aleatoria está gobernada por un modelo causal en función de sus padres en el grafo. Una intervención consiste en forzar uno de los nodos a tomar un valor con probabilidad 1 y a recortar todas las flechas que entran en ese nodo.\\
\\
Es importante que se cumpla que si dos variables están correlacionadas, entonces una es padre causal de la otra y que no falten padres causales de ninguna variable.\\
\\
Para poder contestar pregunta acerca de intervenciones, se necesita introducir el concepto de \textit{grafo mutilado}
\begin{defi}
Sea $\mathcal{B}$ una red sobre un conjunto de variables aleatorias y sea $Z_1=z_1,...,Z_k=z_k$ una instanciación de estas variables denotada por $\mathbf{Z}=\mathbf{z}$ . Se define el grafo mutilado $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$ como:
\begin{itemize}
\item Cada nodo $Z_i \in \mathbf{Z}$ no tiene padres en $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$
\item El modelo local de probabilidad condicional para $Z_i \in \mathbf{Z}$ es tal que asignal probabilidad uno al valor $z_i$ y cero a cualquier otro valor.
\item No se altera ningún otro nodo $X_i \notin \mathbf{Z}$
\end{itemize}
\end{defi}
Entonces, utilizando este concepto, tenemos que para un modelo causal $\mathbf{C}$ una intervención $P_{\mathcal{C}}(Y|do(z)$ es realmente una \textit{querie} probabilista estándar, pero en el grafo mutilado dado por $\mathcal{C}_{Z=z}$; es decir
\[ P_{\mathcal{C}}(Y|do(z),x) = P_{\mathcal{C}_{Z=z}}(Y|x).  \] 
\subsection{Identificabilidad de modelos}
Aun a pesar de desconocer la totalidad de las variables latentes causales, existen casos en los que es posible responder a preguntas causales en modelos que involucran variables latentes utilizando sólo variables observadas; este es el caso de los modelos identificables. Existen formas de simplificar queries intervencionales y convertirlas en queries que sólo requieren de datos observacionales pues existen equivalencias entre unas y otras bajo ciertas condiciones.\\
\\
\cite{peters2012identifiability} exploran las condiciones para obtener identificabilidad en modelos causales funcionales. 
\subsection{Aprendizaje de modelos gráficos causales}
Según \cite{koller2009probabilistic}, el problema de aprender un modelo causal a partir de datos puede dividirse en varios ejes:
\begin{itemize}
\item En primer lugar, definir qué se entiende por un modelo causal, pues si estamos trabajando con modelos gráficos causales entonces el problema se convierte en aprender redes bayesianas, pero con la diferencia de la interpretación que se da a las flechas; por otro lado, en modelos causales funcionales se tienen parametrizaciones mucho más complejas para las variables de respuesta, por lo que se necesitan modelos más complejos como el modelo neuronal presentado anteriormente.
\item Un segundo eje es determinar si ya se conoce la estructura y sólo se necesita aprender la parametrización, o si también debe ser aprendida la parametrización.
\item En tercer lugar, es el tipo de datos: observados o intervencionales. Datos observados son aquellos que provienen de alguna distribución generadora y datos intervencionales es en aquellos en los que se intervino el modelo y se observó la respuesta de este. Los datos intervencionales son difíciles de obtener en la práctica, o incluso pudieran ser inmorales o ilegales.
\item Finalmente, cómo tratar los factores de confusión (confounding factors). Uno puede suponer que simplemente no existen y en estos casos el problema se vuelve relativamente sencillo.
\end{itemize}
\subsubsection{Aprender sin factores de confusión}
El problema de aprender modelos gráficos causales sin factores de confusión es esencialmente el mismo que aprender redes bayesianas (\cite{koller2009probabilistic}), pero hay que hacer algunas suposiciones importantes, pues existen siempre varios DAG's que representen la misma distribución, pero que tendrían distintas interpretaciones causales. Una de las suposiciones necesarias para poder aprender modelos causales a través de datos observacionales es la llamada Propiedad Causal de Markov, que es un análogo a la Propiedad de Markov para redes Bayesianas; y su formulación es la siguiente (\cite{koller2009probabilistic}): \textit{cada variable es condicionalmente independiente de sus no-efectos dadas sus causas directas.} El otro supuesto necesario es que las únicas independencias condicionales sean aquellas que surgen de la d-separación en el grafo causal correspondiente (\textit{faithfulness}).\\
\\
Aun con estas suposiciones, lo más que puede aprenderse es una clase de equivalencias para el grafo causal y para llevar esto a cabo existen dos familias de métodos: basados en restricciones y basados en score y son los mismos que se utilizan para aprender el CPDAG en modelos causales funcionales.
\subsubsection{Aprender de datos intervencionales}
En el caso en que se tienen datos provenientes de intervenciones, lo cual es lo más común en el contexto de experimentos científicos. Se utilizan métodos basados en score para aprender de este tipo de datos.
\subsubsection{Aprender con variables latentes}
En el caso en el que suponemos que existen variables latentes, estamos pensando que los datos fueron generados por una distribución que es \textit{parecida}, o parcial, a la verdadera. De nuevo, existen familias de métodos basadas en restricciones y métodos basados en scores.
\subsection{Redes Bayesianas causales: aplicación en genética}
\cite{su2013using} intentan aprender la estructura de las relaciones entre un conjunto de variables relacionadas a un estudio de cáncer (\cite{karagas1998design},\cite{karagas2004incidence}). Específicamente, los autores analizan sólo 11 variables relacionadas a un solo gen. Identifican relaciones interesantes, como por ejemplo la variable “fumador” como padre causal de la variable “cancer”, aunque no apareció relación entre la variable cancer y un gen que se sabe está relacionado con el cáncer, pero esto se debe a la manera en la que fueron seleccionados los pacientes.
\section{Modelos causales funcionales y ecuaciones estructurales}
En otras áreas, como las ingenierías y la física, las relaciones causales se expresan como relaciones funcionales determinísticas en donde cualquier componente aleatorio se introduce en términos de variables no observadas. Esto corresponde (\cite{pearl2009causality}) a la noción Laplaciana de lo aleatorio. En este contexto, un modelo causal funcional consiste en un conjunto de ecuaciones de la forma
\[ X_i = f_i (X_{Pa_i},U_i), \textrm{  } i=1,...n\]
donde $Pa_i$ corresponde al conjunto de variables que directamente determinan a la variable con índice $i$ y donde $U_i$ representa cualquier error de medición o de aspectos no observados.\\
\\
Dado un modelo causal de esta forma, podemos obtener un diagrama de la siguiente manera: para cada variable pintamos un nodo, y desde cada elemento de $Pa_i$ dibujamos una flecha hacia el nodo que corresponde a $X_i$. A este diagrama se le llama \textit{diagrama causal}. Si cumple que el diagrama es acíclico, se le conoce como semi-Markoviano. Si, además de ser acíclico, los términos de error son independientes entonces se dice que es Markoviano. El siguiente Teorema muestra la conexión entre el diagrama causal y la distribución de probabilidad asociada. 
\begin{teo}
Todo modelo causal Markoviano induce una distribución $P(X_1,...,X_n)$ que satisface que cada variable $X_i$ es independiente de todos sus no-descendientes dado sus padres $Pa_i$.
\end{teo}
Sabemos que dos DAG's que tienen el mismo esqueleto y las mismas v-estructuras son equivalentes, y a la clase de equivalencia le llamaremos CPDAG (\cite{goudet2017learning})
\subsection{Intervenciones y efectos causales en modelos funcionales}
Los modelos funcionales causales proveen una manera de estimar cómo cambiaría la distribución en respuesta a cambios externos. Lo que se hace en estos modelos es representar las intervenciones como alteraciones a un conjunto de funciones. El efecto total de la intervención se calcularía modificando las ecuaciones correspondientes y con este nuevo modelo calcular las nuevas probabilidades. Denotamos una intervención como $do(X=x)$. Se dice que una variable $X_i$ es causa directa de $X_j$ si
\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
\subsection{Contrafactuales en modelos funcionales}

\section{Extracción de relaciones causales a partir de texto}
\subsection{Extracción basada en patrones gramaticales}
\cite{khoo2000extracting} desarrollan un método que permite identificar y extrar relaciones causa-efecto que se encuentra expresada de manera explícita en \textit{abstracts} de artículos médicos. Los autores deciden utilizar el dominio médico pues al ser importantes las relaciones causales es más probable que estas aparezcan explícitamente. Los autores primero identifican los \textit{marcadores lingüísticos} y con esto construyen un conjunto de \textit{patrones causales}. Además, cada oración es pasada por un \textit{parse tree} el cual genera una representación de la estructura sintáctica de la oración.\\
\\
Como ejemplo, consideremos la oración \textit{“Paclitaxel was well tolerated and resulted in a significant clinical response in this patient”}. Después de la creación del árbol sintáctico y del patrón gramatical, se obtiene que la causa presente en la oración es \textit{paclitaxel} y el efecto es \textit{a significant clinical response in this patient}.\\
\\
Analizando 200 abstracts que provienen de 4 áreas médicas (depresión, esquizofrenía, enfermedades cardiacas y SIDA), construyen los patrones causales. El proceso de extracción de información consiste en hacer un \textit{match} entre los patrones causales y el arbol sintáctico. Al aplicar los patrones causales obtenidos a un nuevo conjunto de abstracts, obtienen un 51 por ciento de accuracy al extraer causas y un 58 por ciento de accuracy al extraer efectos.\\
\\
\cite{girju2002text} presentan un método semi-automático para la detección de patrones causales en texto, es semi-automático pues los patrones son descubiertos de manera automática, pero la validación no lo es. El algoritmo de detección de patrones léxico-sintácticos que  se refieren a causalidad consiste en dos partes. La primera parte descubre patrones léxico sintácticos que pueden expresar relaciones causales, y la segunda valida estos patrones según restricciones semánticas. Se enfocan en encontrar patrones de la forma (frase verbo causal frase) donde el verbo es un verbo causal simple en los cuales el verbo se refiere únicamente a la causa; por ejemplo, \textit{Temblores generan tsunamis}. Como las frases obtenidas pueden ser muy largas, pueden aportar información no deseada. Para resolver esto \cite{girju2003automatic} utiliza WordNet para reducir el tamaño de las frases obtenidas.\\
\\
\cite{sanchez2004acquiring} desarrollan un sistema para la adquisición de conocimiento causal a partir de texto. El sistema propuesto identifica oraciones que especifican de manera explícita relaciones causales y extrae de ellas los patrones causales. El sistema incorpora información gramatical tal como conectores, conjunciones, disjunciones y negaciones y codifican todo esto en redes bayesianas.\\
\\
Para la detección de patrones causales, los autores se enfocan en trabajar con oraciones en las cuales ambos \textit{eventos} aparezcan como frases (noun phrases, ¿traducción?). Un ejemplo de patrones causales identificados por el sistema es \textit{“Anemia are caused by excessive hemolysis”}.
\\
Para la construcción de las redes, se toma como input un texto \textit{tokenizado}, etiquetado con \textit{part-of-speech tagging} y \textit{parsed} para encontrar expresiones verbales. Posteriormente, el sistema estima las probabilidades condicionales de los efectos dadas las causas mediante el conteo de frecuencias de eventos similares.\\
\\
Para la evaluación, hicieron una evaluación subjetiva que consistió en comparar la estructura de la red generada por el sistema y la estructura creada manualmente al observar los patrones causales en el texto. Se analizaron 7 textos y en general se encontró un empate entre las redes con una precisión de 60 por ciento, pero sin ningún valor de Recall, por lo que no es fácil interpretar sus resultados. Según \cite{bojduj2009extraction}, una de las razones por las cuales se ubtuvo baja precisión es que basar los nodos en las frecuencias no es una buena idea, pues la cantidad de texto disponible en general es tan grande que estas frecuencias no pueden ser representativas.\\
\\
\cite{bojduj2009extraction} menciona que muchos de estos métodos están más enfocados en cómo extraer las relaciones que en cómo usarlas. Es importante no sólo poder concluir que dos términos están relacionados causamente, sino cómo.
\subsection{Extracción basada en patrones de co-ocurrencias}
En vez de utilizar patrones gramaticales, \cite{saito2007extraction} y \cite{perrin2008global} analizan la co-ocurrencia de términos estadísticos para construir redes a partir de texto. Ellos definen un “statistical term” como \textit{“Statistical terms are expression of the measurements of statistics to watch the movements of phenomena; birth rates, public approval rating of the Cabinet and so on.”}; es decir, un término que sea importante a un sistema. El enfoque de su trabajo es construir redes causales entre los statistical terms. Lo que hacen es primero extraer relaciones de co-ocurrencia entre términos estadísticos y posteriormente extraer relaciones causales entre ellos. Los autores se enfocan en texto japonés, el cual cumple que tiene ciertos patrones en los sufijos utilizados, y estos patrones son utilizados al extraer los términos estadísticos. Luegom estos términos son clasificados según su nivel de abstracción. Posteriormente, construyen una red de términos estadísticos al considerar a dos de ellos co-ocurrentes si aparecen juntos en un párrafo. Se remueven los términos que son co-ocurrentes con muchos otros términos.\\
\\
Los autores evaluan de manera separada cada una de las etapas: extracción de términos, clasificación de términos, la construcción de redes y la clasificación de redes.
\\
La mayor desventaja de esto es que la dirección de la relación causal no puede ser inferida de la co-ocurrencia.\\
\\
Por otro lado, \cite{sakai2008cause} extrajeron términos de artículos financieros (en japonés) y luego considerarón que estos términos afectaban el desempeño de los negocios basados en co-ocurrencias con expresiones clave como “es bueno”. Su definición de causalidad depende de las expresiones clave.
\subsection{Modelos Gráficos para extracción de relaciones}
A partir de un artículo que trata específicamente sobre una entidad claramente indentificable (la biografía de Benito Juarez por ejemplo) \cite{yu2010jointly} intentan identificar entidades y descubrir relaciones semánticas entre ellas. Resulta evidente que en el documento aparecerán otras entidades, pero se hace el supuesto de que no están relacionadas entre sí. En este trabajo, todas las clases entidades están pre-definidas así como las relaciones; esto es, se sabe que en el documento aparecerán años, lugares, países y que una entidad de tipo persona habrá nacido en un país en cierto año. Para formular el problema, se tiene $\mathbf{x} = \{ x_1,...,x_N \}$ una sucesión observada de \textit{tokens}. Denotamos como $s_p$ la entidad principal. Sea $ \mathbf{s}=\{ s_1,...,s_L\}$ una segmentación en la cual cada $s_i$ es una tripleta $\{ \alpha_i, \beta_i,y_i \}$ con $\alpha_i$ es una posición inicial, $\beta_i$ posición final y $y_i$ una etiqueta asignada.\\
\\
Sea $r_{pn}$ la asignación de relaciones entre la principal $s_p$ y un candidato a entidad secundaria $s_n$ y sea $\mathbf{r}$ el conjunto de asignación de relaciones para la $\mathbf{x}$. Dada una observación $\mathbf{x}$, queremos encontrar $\mathbf{y}^\ast$ tal que
\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
donde $\mathbf{y}=\{ \mathbf{r},\mathbf{s} \}$.\\
\\
Se define una distribución condicional conjunta para $\mathbf{s},\mathbf{x},\mathbf{r}$ en un modelo grafico no dirigido $\mathcal{G}$ tal que se pueden dividir los factores de $\mathcal{G}$ en tres grupos: $\phi^S$, $\phi^R$, $\phi^\nabla$, que son el potencial de segmentación, el potencial de relaciones, y el potencial conjunto de segmentación-relación. \\
\\
Por el Teorema de Hammersley-Clifford, la condicional conjunta $P(\mathbf{y} | \mathbf{x})$ se factoriza como
\[ P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \prod_{C_S} \phi^S (i,\mathbf{s},\mathbf{x}) \prod_{C_R}\phi^R(r_{pm},r_{pn},\mathbf{r}) \prod_{C_\nabla} \phi^\nabla(s_p,s_j,\mathbf{r}) \]
Se hace el supuesto de que las funciones potencial se factorizan según un conjunto de atributos y un conjunto de pesos correspondientes.\\
\\
Incorporando estos features, tenemos que 
\[P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \exp \left( \sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r) \right) \]
\\
Dada una muestra $\mathcal{D}= (\mathbf{x}_i, \mathbf{y}_i)_{i=1}^N$ queremos estimar los parámetros $(\lambda_k,\mu_w,\nu_t)$.\\
\\
La log-verosimilitud regularizada es
\[ \mathcal{L} = \log [\Phi(r,s,x)] - \log [Z(x)] - \sum_{k=1}^K \frac{\lambda_k^2}{2 \sigma_\lambda^2} - \sum_{w=1}^W \frac{\mu_w^2}{2 \sigma_w^2} - \sum_{t=1}^T \frac{\nu_t^2}{2 \sigma_\nu^2} \]
Donde,
\[ \Phi(r,s,x) = \exp (\sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r)) \]
Tenemos que $\mathcal{L}$ es concava por lo que es fácil maximizarla.\\
\\
Para calcular la Asignacion Más Probable, hayue encontrar
\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
La cual no se puede resolver de manera exacta, se necesitan métodos aproximados. El algoritmo que se propone en el artículo se llama Collective Iterative Classification y la idea detrás de este es decodificar las variables ocultas objetivo basados en asignar etiquetas a las variables muestreadas. Esto se hace mediante un proceso iterativo de dos pasos; primero, en una etapa de bootstrapping se predice una etiqueta inicial para una $x_i$ dado el modelo ya entrenado. Luego, en la segunda etapa, conocida como clasificación iterativa, se re-estima la asignación a $x_i$ varias veces, tomándolas a partir de la asignación inicial.\\
\\
Los autores prueban su algoritmo con datos provenientes de Wikipedia: sus datos consisten en 441 páginas de Wikipedia. De estos datos, fueron etiquetados a mano 7740 entidades en 8 categorías. Además, se extrajeron 4700 relaciones de las cuales las más frecuentes fueron trabajo, visió, nació, miembro de, nació en día.\\
\\
El modelo superó los existentes tomando como medida de calidad la medida F. 
\section{Métodos Estadísticos Para Extracción de Relaciones Causales}
\subsection{Statistical Relational Learning}
En el contexto de inferencia causal, se estudian las interacciones entre variables que puedan causar efectos en la otra. Para esto, es necesario identificar los objetos y las relaciones entre ellos. Para esto, el área de Statistical Relational Learning ofrece una serie de métodos que modelan dependencias entre entidades relacionales. En esta sección veremos distintas técnicas existentes en esta área y sus aplicaciones a inferencia causal.\\
\\
Un dominio de interés consiste en objetos (entidades), sus atributos, y las relaciones entre sí de los objetos. El modelo DAPER de \cite{heckerman2004probabilistic} consiste en clases de entidades, clases de relaciones, clases de atributos y clases arco.
\subsubsection{Aplicación de Campos Aleatorios Condicionales a Extraccion de informacion}
\cite{ramani2005consolidating} analizaron 753,459 abstracts de journals médicos, de los cuales extrajeron 6580 interacciones entre 3,737 proteínas. Para ello, utilizaron un algoritmo de tres etapas: la primera etapa, consiste en identificar nombres de proteínas mediante un clasificador basado en Conditional Random Fields; la segunda, identificar interacciones a través de co-ocurrencias; posteriormente, filtrar estas interacciones con un clasificador Bayesiano para obtener interacciones válidas. Con esto, se obtiene una red que consiste en 31609 interacciones entre 7748 proteínas\\
\\
Para la etapa de identificación de nombres, se utilizó un Conditional Random field, cuyo desempeño fue evaluado en un conjunto de 200 abstracts médicos etiquetados y en 750 abstracts etiquetados a mano. Para extraer los nombres de proteínas, se utilizó este clasificador en el conjunto de 753,459 abstracts de Medline que contuvieran la palabra “human”.\\
\\
Para obtener las interacciones, se midió la co-citación entre los nombres de proteínas y luego se enriquecieron estos pares con interacciones físicas (reales) entre proteínas utilizando un filtro Bayesiano. Primero, se contó el número de veces que en un abstract aparece un par de proteínas, y con eso se calculó la probabilidad de co-citación. Con esto se obtienen 15,000 interacciones, pero se pueden dar co-citaciones por otras razones además de las relaciones reales, se aplicó un filtro bayesiano que mide la verosimilitud de los abstracts que citan un par de proteínas para discutir sus interacciones reales. El clasificador asigna un \textit{score} a cada uno de los abstracts que citan esas proteínas. Utilizando un umbral de score, se obtienen 6,580 interacciones entre 3,737 proteínas.\\
\\
Al combinar estas interacciones con otras 26,280 interacciones provenientes de otras fuentes, se obtiene el total mencionado.\\
\\
Por otro lado, \cite{culotta2006integrating} plantean el problema de extracción de relaciones como un problema de etiquetado de secuencias, algo similar a lo realizado por \cite{yu2010jointly}.   \cite{culotta2006integrating} definen un modelo de la forma
\[ P_{\Lambda} ( y | x ) =\frac{1}{Z} \prod_{c \in C} \varphi_c ( y_c , x_c, \Lambda ),\]
donde las $\varphi$ son potenciales parametrizados por $\Lambda$.\\
\\
Los autores tomaron 1,127 párrafos de 271 artículos biográficos de Wikipedia y etiquetaron 4,701 instancias de relaciones. Para identificar a las entidades importantes dentro de un artículo, utilizaron el hecho de que estas suelen tener un \textit{hyperlink}
\subsubsection{Aplicación de RMN's a Extraccion de informacion}
\cite{bunescu2007statistical} utilizan Relational Markov Networks para mejorar el desempeño de los CRF's en el problema de identificar nombres de proteínas en abstracts de journals médicos. \\
\\
La técnica que se propone es que dada una colección de documentos $D$, asociemos a dada documento $d \in D$ un conjunto de entidades $d.E$. Cada entidad$e \in d.E$ queda caracterizada por un conjunto de atributos booleanos $e.F$. A cada documento se le asocia un grafo de factores, el cual es un grafo bipartita que contiene dos tipos de nodos: nodos variables, que corresponden a las etiquetas de todas las entidades candidato en el documento y nodos potenciales, que modelan las correlaciones entre dos o más atributos. Para cada correlación posible,  un nodo potencial es crreado tal que está ligado a todos los nodos variables involucrados.\\
\\
Prueban su algoritmo en dos conjuntos de datos previamente etiquetados a mano. El primer conjunto consiste en 200 abstracts y el segundo en 225. Ambos conjuntos de datos fueron previamente etiquetados a mano, además de etiquetados para POS. En ambos conjuntos de datos, existe una mejora ligera respecto a los CRF's en términos de la medida F.

\subsubsection{Aprendizaje de Modelos Causales a partir de Datos Relacionales}
\cite{maier2013reasoning} definen la \textit{abstract ground graph} (AGG) y se desarrolla la noción de d-separación relacional. A partir de estas nociones teóricas, \cite{maier2013sound} extienden el algoritmo PC, el cual identifica todas las posibles orientaciones de dependencias causales, a datos relacionales. Utilizando la d-separación relacional, introducen restricciones que orienta dependencias bivariadas hasta en un 72$\%$ más. Se prueba teóricamente que con esta nueva regla, además de extensiones a datos relacionales del algoritmo PC, se obtiene un método \textit{sound and complete} para la extracción de relaciones causales. Este algoritmo nuevo se llama RCD (relational causal discovery).\\
\\
\cite{lee2016learning} notan que la demostración de \cite{maier2013reasoning} requiere que la AGG sea un DAG quie represente exactamente todas las flechas que aaparecerían en todas las posibles ground graphs, y en un trabajo anterior (\cite{lee2015lifted}) habían notado que existen casos en los cuales la d-separación de la AGG no captura las independencias condicionales que se cumplen en el modelo relacional causal. En su artículo, proponen un algoritmo basado en condiciones más débiles llamado RCD-light.

\subsection{Inferencia Bayesiana No Paramétrica}
La rama no-paramétrica de la inferencia bayesiana puede considerarse como una extensión a espacios infinito-dimensionales de la inferencia bayesiana clásica. Esto permite tener modelos que permiten capturar una complejidad mucho mayor en los datos. El objeto principal de la inferencia bayesiana no paramétrica es el Proceso Dirichlet, que es una extensión a espacios infinito-dimensionales de la distribución Dirichlet clásica.
\subsubsection{Proceso Dirichlet}
\begin{defi}
Sea $H$ una medida finita definida sobre un espacio $\mathcal{X}$ y $\alpha \in \mathbb{R}$. Una medida aleatoria $G$ sobre $\mathcal{X}$ se dice que es un Proceso Dirichlet de parámetros $(\alpha,H)$ si para cada partición finita medible $\{ A_1 , ..., A_k \}$ de $\mathcal{X}$ se tiene que la distribución conjunta de $(G(A_1),...,G(A_k))$ es Dirichlet de parámetros $(\alpha H(A_1),...,\alpha H(B_k))$. A la medida $H$ se le conoce como medida base y a $\alpha$ se le conoce como parámetro de concentración. Se denota como
\[ G \sim \textrm{ DP}(\alpha,H).\]
\end{defi}
Los parámetors $H$ y $\alpha$ juegan roles intuitivos en la definición. La medida base se puede entender, básicamente, como la \textit{media} del Proceso; es decir, para cada conjunto medible $A \subseteq \mathcal{X}$ se tiene que $\mathbb{E}[G(A)]=H(A)$. Por otro lado, el parámetro de concentración es un tipo de \textit{varianza inversa}, pues $V[G(A)]=H(A)(1-H(A))/(\alpha+1)$. Es decir, mientras más grande sea $\alpha$, entonces menor es la varianza y el DP tenderá a concentrarse más en torno a $H$.\\
\subsubsection{Distribución Predictiva y urnas de Blackwell-McQueen}
Consideremos $G \sim \textrm{ DP}(\alpha,H)$ y $\theta_1,... \sim G$. Como $\theta_{n+1} | G, \theta_1,...,\theta_n \sim G$, entonces para $A \subseteq \mathcal{X}$ tenemos que
\begin{eqnarray*}
P(\theta_{n+1} \in A | \theta_1,...,\theta_n ) &=& \mathbb{E}[G(A) | \theta_1,...,\theta_n]\\
                                                                    &=& \frac{1}{\alpha + n} \left(\alpha H(A) + \sum_{i=1}^n \delta_{\theta_i}(A) \right)
\end{eqnarray*}
Por lo tanto, marginalizando $G$, vemos que la distribución posterior base dados $\theta_1,...,\theta_n$ es la distribución predictiva de $\theta_{n+1}$.
\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum^n \delta_{\theta_i} \right) \]
La secuencia de distribuciones predictivas para $\theta_1, \theta_2,...$ se conoce como el esquema de urnas de Blackwell-McQueen; la interpretación es la siguiente: cada valor en $\mathcal{X}$ es un color, y cada realización $\theta \sim G$ son pelotas y el valor obtenido es el color de la pelota. Además, tenemos una urna en la cual se han colocado las bolas observadas. Al principio, no existen bolas en la urna y muestreamos un color de $H$; es decir, $\theta_1 \sim H$, pintamos la bola con ese color y la echamos en la urna. Luego, en el paso $n+1$, tomaremos un nuevo color con probabilidad $\alpha / (\alpha + n)$ o tomaremos un nuevo color, la pintaremos así y la echaremos en la urna con probabilidad $n / (n+\alpha)$.
\subsubsection{Proceso Dirichlet como Prior en problemas de Clustering}
La propiedad de que $G \sim DP$ sea discreto ofrece aplicaciones de clusterización. Por el momento, supongamos que $H$ es suave. Entonces, como los valores de las realizaciones son repetidos, sean $\theta^\ast_1,...,\theta^\ast_m$ los valores únicos de las realizaciones $\theta_1,...,\theta_n$ y sea $n_k$ el número de repeticiones de $\theta^\ast_k$. Entonces, la distribución predictiva se puede escribir como
\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum_{k=1}^m n_k \delta_{\theta^\ast_k} \right). \]
Esto es, el valor de $\theta^\ast_k$ aparecerá en $\theta_{n+1}$ con una probabilidad proporcional al número de veces que ha aparecido. Entonces, mientras más grade sea $n_k$, mayor será la probabilidad de que se vuelva aun mayor. Notemos que los valores únicos de las realizaciones $\theta_1,...,\theta_n$ inducen una partición aleatoria del conjunto $[n]=\{1,...,n\}$ si pensamos el cluster $k$ como aquel que contiene las $\theta_i$ que toman el valor $\theta^\ast_k$, tenemos que clusters grandes crecen más rápido.\\
\\
Podemos invertir este proceso y empezar con una distribución sobre particiones.
\subsubsection{Dirichlet Enhanced Relational Learning}
\cite{xu2005dirichlet} aplican un modelo jerárquico Bayesiano no-paramétrico para la extracción de relaciones. La ventaja de usar un modelo jerárquico es que los parámetros pueden estar \textit{personalizados} para entidades. En este trabajo se aplica el modelo al contexto médico, en el cual se tienen como entidades hospitales, pacientes, diagnósticos y procedimientos. Se sabe que la existencia de un diagnóstico o un procedimiento es dependiente de las características de un paciente; por otro lado, las características de los hospitales se modelan como inciertas.\\
\\
Un modelo Bayesiano consiste en un parámetro $\theta$ y una asignación inicial de probabilidad sobre este, la cual refleja nuestra incertidumbre inicial. Supondremos que esta distribución inicial cuenta con hiperparámetros $h$, de modo que sea $P(\theta | h)$. La verosimilitud de los datos $P(D|\theta)$ refleja cómo son generados estos dado el parámetro. EN el caso en el cual se asume que los datos provienen de distintos grupos es natural pensar en asignar un parámetro por grupo, lo cual lleva a la definición de modelos jerárquicos, en los cuales el modelo tiene la siguiente fforma
\[ P(h) \prod P(\theta_i | h )P(D_i | theta_i). \]
Si se aplicara el modelo DAPER al contexto médico, se tiene que los parámetros e hiperparámetros que especifican distribuciones condicionales como atributos globales. Esto implica que las probabilidades de un procedimiento son idénticas para todos los pacientes con la misma dolencia inicial. Luego, los procedimientos son modelados de manera independiente de manera que procedimientos anteriores en una persona no afectlan la selección de uno nuevo. Estas son suposiciones no realistas.\\
\\
Para mejorar esta situación, se asume que la probabilidad de un procedimiento es una distribución multinomial con un prior Dirichlet. En particular, para cada paciente, se asume que un vector individual de parámetros $\theta_{s|pc,pa}$ especifica la probabilidad de un procedimiento para un paciente $pa$ con dolencia inicial $pc$. Este vector es generado por una distribución Dirichlet de parámetros $h_{pc}=\{ \tau_{pc},\alpha_{pc} \}$, la cual puede escribirse como
\[ \textrm{Dir}(\theta_{\cdot | pc, pa} | \tau_{pc},\alpha_{pc}) = \frac{1}{C} \prod_{k=1}^K \theta^{\tau_{pc} \alpha_{k,pc} - 1}_{k,s | pc, pa}.  \]
En un modelo jerárquico bayesiano cada paciente obtiene sus propias probabilidades de procedimiento y comparte con los demás un parámetro a priori.\\
\\
En más de un caso, se necesita asumir que la prior tiene una forma muy flexible, por lo que es necesario recurrir a priors no paramétricas.\\
\\
Podemos tomar la distribución a priori como una realización del Proceso Dirichlet; es decir, 
\[ G_{pc} \sim \textrm{ DP }(G_0,\alpha_0) \]
donde $G_0$ es la medida base y $\alpha_0$ el parámetro de concentración, el cual especifica nuestro nivel de certidumbre sobre la prior.\\
\\
Luego, el parámetro $\theta_{\cdot | pc, pa}$ es una realización de $G_{pc}$.\\
\\
Otra manera de escribir el Proceso Dirichlet es mediante su representación \textit{stick breaking} (\cite{teh2011dirichlet}) según la cual
\[ G_{pc} = \sum \pi_{l,pc} \delta_{\theta^\ast_{l,pc}} \textrm{ ;  } \theta^\ast_{l,pc} \sim G_0  \]
\[ \pi'_{l,pc} \sim \textrm{ Beta }(1,\alpha_0) \textrm{  ;  } \pi_{l,pc} = \pi'_{l,pc} \prod_{k=1}^{l-1} (1 - \pi'_{k,pc}).\]
Notemos que aunque la distribución base pueda ser continua, el DP es siempre discreto (con probabilidad 1).\\
\\
Tradicionalmente, el aprendizaje en el contexto Bayesiano no paramétrico se lleva a cabo via muestreo de Gibbs, cuyas variaciones más comunes son la urna de Polya o el Proceso del Restaurante Chino (\cite{teh2011dirichlet}). En el artículo, se lleva a cabo un proceso más eficiente (\cite{Yu:2004:NHB:1008992.1009053}.)\\
\\
El objetivo es estimar $G_{pc}$ para cada posible $pc$ en la base de datos, utilizando la verosimilitud marginal:
\[ \hat{G}_{pc} = \textrm{ argmax DP }(G | G_0, \alpha_0) \prod_{\{pa\}_{pc}} \int \textrm{ Multinomial }(\{pr\}_{pa}| \theta_{\cdot | pc,pa}) d\theta_{\cdot | pc,pa}  \]
donde $\{pa\}_{pc}$ es el conjunto de pacientes tales que tienen la misma dolencia inicial y $\{pr\}_{pa}$ es el conjunto de procedimientos para el paciente $pa$. Resulta imposible calcular la verosimilitud, por lo que se realiza una aproximación de campo medio:
\[ G_{pc} \approx \sum^{|{pa}_{pc}|} \pi_{pa,pc} \delta_{\theta^\ast_{\cdot | pa,pc}}. \]
El proceso de aprendizaje se divide en dos etapas: 
\begin{itemize}
\item Calcular la ubicación de $ \theta^\ast_{\cdot | pa,pc}$
\item Estimar los pesos $\pi_{pa,pc}$
\end{itemize}
Para el primer paso, se utiliza la aproximación MAP de $\theta_{\cdot | pa,pc}$ que es:
\[ \theta^{MAP}_{\cdot | pa,pc} = \textrm{ arg max }P( \theta_{\cdot | pa,pc} | \{pr\}_{pa})\]
Y se utiliza como prior la Dirichlet mencionada antes. \\
\\
Para el segundo paso, es necesario hacer la siguiente suposición:
\[ \hat{P}( \theta_{\cdot | pa,pc} | \{pr\}_{pa}) \approx q_{pa}(\theta_{\cdot | pc,pa}) = \sum_{\{\tilde{pa}\}_{pc}} \xi_{\tilde{pa},pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}} \]
donde $\xi_{\tilde{pa},pc,pa} \geq 0 $ son parámetros variacionales tales que $\sum_{pa} \xi_{\tilde{pa},pc,pa} = 1$.\\
\\
Como paso E se obtiene que:
\[  \xi^t_{\tilde{pa},pc,pa} = \frac{P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}}) \hat{G}^{(t)}_{pc}(\delta_{\theta^{MAP}_{\cdot | pc,pa}}))}{\sum_{\tilde{pa}}P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}})\hat{G}^{(t)}_{pc}(\theta^{MAP}_{\cdot | pc,pa}))}. \]
Y el paso M queda como
\[ \hat{G}^{(t+1)}_{pc}(\theta_{\cdot,pc,pa}) = \frac{\alpha_0 G_0(\theta_{\cdot,pc,pa}) + \sum_{\{pa\}_{pc}} \xi_{pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}}}{\alpha_0 + |\{pa\}_{pc}|} \]

\subsubsection{Learning Systems of Concepts with an Infinite Relational Model}
Suponiendo que tenemos una o más relaciones que involucran uno o más tipos, \cite{kemp2006learning} definen el Infinite Relational Model, el cual busca partir cada tipo (type) en clusters, de manera que una buena partición permita que relaciones entre entidades se puedan predecir a partir de la asignación de clusters. Por ejemplo, si tenemos $m$ relaciones que involucran $n$ tipos, denotamos $R^i$ la $i$-ésima relación, $T^j$ la $j$-ésima relación y $z^j$ un vector de asignación de clusters. Lo que se busca es inferir la asignación de clusters, y ultimadamente la posterior $P(z^1,...,z^n | R^1, ... , R^m)$. Para especificar esta distribución, se define un modelo generativo para las relaciones y la asignación de clusters
\[ P(R^1,...,R^m, z^1,...,z^n) = \prod_{i=1}^m P(R^i | z^1 , ... , z^n) \prod_{j=1}^n P(z^j). \]
Estamos asumiendo que las relaciones son condicionalmente independientes dada la asignación de clusterss, y que la asignación de clusters es independiente entre tipos. Para que el IRM sea capaz de descubrir el número de clusters en el tipo $T$, se utiliza un prior que asigna probabilidad a todas las posibles particiones de $T$. Un prior razonable debería descubrir sólo tantos clusters como sugeridos por los datos. A partir de trabajos previos en Bayesiana No paramétrica (\cite{rasmussen2000infinite}) utiliza una distribución sobre particiones inducida por un CRP  (Chinese Restaurant Process, \cite{teh2011dirichlet}).\\
\\
La intuición detrás de este proceso, según describe \cite{teh2011dirichlet} es que en un restaurante chino hay un número infinito de mesas, cada una de las cuales puede sentar un número infinito de clientes. El primer cliente entra y se sienta en la primera mesa, el segundo cliente entra y decide si sentarse en la primera mesa o en otra; en general, el $n$-ésimo cliente se sienta en una mesa ocupada con una probabilidad proporcional al número de clientes sentados ahí.\\
\\
Entonces, la manera de constuir clusters utilizando el CRP es partir de un cluster con un solo objeto, e ir añadiendo objetos de modo que la probabilidad de caer en cierto cluster sea proporcional a los elementos en él. De esta manera, la distribución de clusters para el objeto $i$, dada la asignaclión de los objetos $1,...,i-1$ es:
\[
  P(z_i = a | z_1, ... , z_{i-1}) =
  \begin{cases}
                                   \frac{n_a}{i-1+\gamma} & \text{if $n_a > 0$} \\
                                   \frac{\gamma}{i-1+\gamma} & \text{$a$ es un nuevo cluster} 
  \end{cases}
\]
Donde $n_a$ es el número de objetos asignado al cluster $a$ y $\gamma$ es un parámetro. La distribución en $z$ inducida por el CRP es intercambiable; es decir, no importa el orden en que los objetos llegan a los clusters. Como nuevos objetos siempre pueden ser asignados, el IRM efectivamente tiene acceso a un número infinito de clusters, de ahí el nombre del Infinite Relational Model.\\
\\
Para generar Relaciones a partir de las asignaciones de clusters se asume que las relaciones son binarias, pero se puede extender a datos continuos y frecuencias. Como ejemplo, veremos primero el caso con un solo tipo $T$ y una relación $R: T \times T \to \{0,1 \}$. $T$ puede ser, por ejemplo, un conjunto de personas y $R$ una relación de amistad. El modelo generativo para este problema es
\begin{eqnarray*}
z | \gamma &\sim& CRP(\gamma)\\
\eta(a,b) | \beta &\sim& Beta(\beta,\beta)\\
R(i,j) | z, \eta &\sim& Bernoulli(\eta(z_i,z_j))
\end{eqnarray*}
Esto es, estamos suponiendo que la relación $R$ está generada a partir de dos estructuras latentes: una partición $z$ y una matriz de parámetros $\eta$. La entrada $R(i,j)$ se genera al lanzar una moneda con sesgo $\eta(z_i,z_j)$ donde $z_i,z_j$ son las asignaciones de cluster de las entidades $i$,$j$. El parámetro $\eta(a,b)$ especifica la probabilidad de que exista un link entre las entidades $i \in a$, $j \in b$. El IRM invierte ese modelo para descubrir la $z$ y la $\eta$ que mejor explican $R$.\\
\\
Estamos asumiendo que la tendencia de una entidad a participar en relaciones está dada por su asignación de cluster.\\
\\
Para hacer inferencia, notamos que en el ejemplo anterior se utilizó una prior conjugada, es sencillo calcular $P(R|z)$
\[ P(R|z)= \prod_{a,b} \frac{Beta(m(a,b) + \beta, \bar{m}(a,b)+\beta}{Beta(\beta,\beta)} \]
donde $m(a,b)$ es el número de pares $(i,j)$ tales que $i \in a$, $j \in b$ y $R(i,j)=1$ mientras que $\bar{m}(a,b)$ son las parejas donde $R(i,j)=0.$\\
\\
Para muestrear de la posterior $P(z | R ) \propto P(R|z)P(z)$ se pueden utilizar técnicas MCMC o al buscar la moda de la distribución.\\
\\
Para la evaluación del modelo, los autores generan datos sintéticos. Consideran conjuntos de datos de tres formas distintas; el primer sistema, $S1$, tiene sólo dos tipos $T1, T2$ y una sola relación $R:T1 \times T2 \to \{0,1 \}$. El sistema $S2$ consiste en 4 tipos y tres relaciones binarias mientras que $S3$ consiste de 3 tipos y una sola relación ternaria. Resulta que en este contexto, el modelo captura el número de clusters de manera precisa. \\
\\
Este método es excelente para clusterizar de manera no paramétrica un conjunto de datos relacionales, pero requiere identificar previamente las entidades y las relaciones. 
\subsubsection{Learning Infinite Hidden Relational Models}
\cite{xu2006learning} extienden la expresividad de los modelos relacionales al introducir para cada entidad una variable latente con un número infinito de estados que proviene de un Dirichlet Process Mixture. Debido a que el número de features de las cuales un atributo puede depender, resulta más conveniente  introducir para cada entidad una variable latente que es padre de los atributos de la entidad y de los atributos de las relaciones en las que participa la entidad; a esto se le conoce como hidden relationship model.\\
\\
Como cada entidad puede tener un distinto número de estados en sus variables latentes, es deseable que este número de estados latentes se determine de manera automática según los datos; esto es posible utilizando una mezcla de Proceso Dirichlet, que son modelos de mezcla con un número infinito de componentes de mezcla, pero que según los datos la complejidad se ajusta de manera automática. A la combinación de un hidden relational model con un DPm se le conoce como Infinite Hidden Relationship Model que generaliza el modelado jerárquico bayesiano noparamétrico al caso de modelos relacionales.\\
\\
Como aplicación, los autores evalúan su modelo en un conjunto de datos sobre películas y buscan predecir las preferencias de los usuarios. Para este conjuto de datos, las clases de entidades son Usuario y Película
\subsection{Modelos Jerárquicos Bayesianos}
Los Modelos Jerárquicos Bayesianos (HBM) son modelos probabilistas que están definidos por \textit{niveles}, en los cuales cada nivel representa un mayor grado de abstracción. Un ejemplo sencillo de esto puede ser un modelo de regresión de ingreso a nivel nacional en el cual existan parámetros por estado y por municipio.\\
\\
En el contexto causal, lo que está detrás de un HBM es que al aprender causalidad se incorpora conocimiento teórico previo y datos observables (\cite{griffiths2005structure}, \cite{griffiths2009theory}). Esto es, un HBM es una distribución de probabilidad entre supuestos teóricos, modelos causales y datos (\cite{hagmayer2013hierarchical}).\\
\\
\cite{lucas2010learning} proponen un framework que incorpora inferencia basada en covarianzas así como conocimiento adquirido. Se enfocan en aprender  \textit{forma funcional} de las relaciones causales. Las contribución principal del artículo es: se muestra que el problema de aprender la forma funcional de una relación causal puede ser formalizada utilizando modelos jerárquicos bayesianos; además, se muestran una serie de experimentos que ponen a prueba las predicciones cualitativas hechas por este modelo.\\
\\
\cite{hagmayer2013hierarchical} presentan un overview general sobre los modelos jerárquicos bayesianos en el contexto causal. En el artículo exponen las ventajas de utilizar estos modelos para realizar inferencia causal, así como sus limitaciones. Una limitación mencionada es que dependen considerablemente de conocimiento previo. Otra, que existe evidencia que la representación que las personas tienen de modeos causales no corresponde a modelos gráficos causales (redes bayesianas) pues existen experimentos en los cuales la propiedad Markoviana no se cumple. Finalmente, argumentan que los HBM's y las redes bayesianas no modelan los procesos cognitivos subyacentes al apredizaje causal, aunque existen evidencias de que las personas sí llevan a cabo un tipo de actualización Bayesiana

\section{Aprendizaje por Refuerzo}
El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno. Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente. El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente. A su vez, el ambiente devuelve información al agente en términos de una recompensa o ganancia según el estado en el que se encuentre. El objetivo del agente es encontrar una serie de acciones que lo lleven a lograr su objetivo final. La función de recompensa codifica qué es lo que se desea que el agente aprenda, pues esta función devuelve un valor para cada estado y con esto el agente asignará un valor a cada estado según la ganancia a largo plazo que un este representa.\\
\\
La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$. El objetivo del agente es encontrar la política que maximice la recompensa.\\
\\
La regla general que se sigue para la definición del ambiente (\cite{sutton1998reinforcement}) es considerar que todo aquello que no pueda ser explícitamente alterado por el agente pertenece al ambiente. La frontera agente-ambiente tiene que ver con los límites de lo que el agente puede controlar, no sobre lo que puede \textit{conocer}.
Veremos ahora con detalle las partes que conforman esta dinámica. 
\subsubsection{Objetivos y Recompensas}
Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada. Por ejemplo, un agente que queremos que aprenda a jugar ajedrez debe ser premiado por ganar los juegos, no por lograr sub-objetivos como comer más piezas del oponente o tomar control del centro del tablero (\cite{sutton1998reinforcement}), pues podría suceder que el agente encuentre cómo lograr esto sin ganar el juego necesariamente. Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
\subsubsection{Definición de Cadenas de Markov}
Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
\[P(X_{t+1} \in A | X_0=x_0, X_1=x_1 , ... , X_t=x_t) = P(X_{t+1} \in A | X_t = x_t).\]
Estos procesos quedan caracterizados por sus probabilidades de transición $p^t_{ij}=P(X_{t+1}=j | X_{t}=i)$. Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
\subsubsection{Procesos de Decisión Markovianos con recompensa}
Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$. En respuesta a la acción tomada, se recibe una recompensa $r_t = f(a_t)$ para alguna función $f$ desconocida, la cual puede ser aleatoria o determinista.
Los procesos de decision Markovianos con recompensa (\cite{Puterman:1994:MDP:528623}) están definidos por un espacio de estados $\mathcal{S}$, un espacio de acciones posibles $\mathcal{A}$, una función de transición $f$ y la función recompensa $r$. La función transición puede ser estocástica o determinística.

\subsubsection{Función valor}
Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado. Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política. Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$. Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$. La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante. Formalmente, para un proceso de decisión Markoviano, definimos la \textit{función valor} (value function en la literatura) para un estado $s$, y relativa a una política $\pi$, como el retorno esperado si se comienza en $s$ y a partir de ahí se sigue la política $\pi$
\[ V^{\pi}(s) = \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]. \]
De manera similar, definimos la \textit{función acción-valor} $Q^{\pi}(s,a)$ es una valoración de tomar la acción $a$ estandoen el estado $s$, y después seguir la \textit{política} $\pi$
\[ Q^\pi(s,a) = \mathbb{E}[R_t | s_t =s, a_t = a]. \]
Por la ley de los grandes números, podemos estimar ambas funciones $V^{\pi}$ y $Q^\pi$ si se promedian los retornos recibidos al llegar a cada estado con el número de veces que se ha llegado al estado. Esta es la idea detrás de los llamados Métodos Monte-Carlo que serán expuestos en una sección posterior.\\
Notemos una propiedad interesante de la función $V^{\pi}$:
\begin{eqnarray*}
V^{\pi}(s) &=& \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]\\
               &=& \mathbb{E}_{\pi} \left[ \sum_{k=0}^\infty \gamma^k r_{t+k+1} | s_t = s \right]\\
               &=& \mathbb{E}_{\pi} \left[ r_{t+1} + \gamma \sum_{k=0}^\infty \gamma^k r_{t+k+2} | s_t = s \right]\\
               &=& \sum_a \pi(s,a) \sum_{s'} P(s_{t+1}=s' | s_t = s, a_t = a) \left[ \mathbb{E}[r_{t+1}|s_t=s,a_t=a,s_{t+1}=s'] + \gamma \mathbb{E} \left[ \sum \gamma^k r_{t+k+2} | s_{t+1}=s' \right]\right]\\
               &=& \sum_a \pi(s,a) \sum_{s'} \mathcal{P}_{ss'}^a \left[ \mathcal{R}_{ss'}^a + \gamma V^{\pi}(s') \right]
\end{eqnarray*}
A esta última ecuación se le conoce como la Ecuación de Bellman para $V^\pi$.\\
\\
También para la función $Q^{\pi}(s,a)$ se tiene una ecuación de Bellman:
\[ Q^{\pi}(s,a)=\mathbb{E}_{s',a'}[r+ \gamma Q^{\pi}(s',a') | s',a']. \]
Además, la función $V^\pi$ puede ser obtenida a partir de la función $Q^{\pi}(s,a)$ de la siguiente manera:
\[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
\subsubsection{Soluciones y Optimalidad}
Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa. Resulta que es posible definir un orden parcial entre políticas en términos de la función valor, pues $\pi \geq \pi'$ si $V^{\pi}(s) \geq V^{\pi'}(s)$ para todo estado $s$. Para procesos de decisión Markovianos finitos existe siempre una política mejor que las demas, a la cual llamaremos \textit{política óptima}. De hecho, pueden existir varias, pero todas ellas comparten la misma función valor, denotada como $V^\ast$. Si para una política $\pi^\ast$ se tiene que es mejor que todas las demás, entonces podemos describir su función valor como
\[  V^{\pi^\ast} = \max_{\pi} V^{\pi}(s).\]
Por supuesto, $V^\ast = V^{\pi^\ast}$
\\
Análogamente, las políticas óptimas tienen la misma función $Q$ óptima, por lo que definimos \textit{función accion-valor óptima} como el retorno esperado de tomar la acción $a$ en el estado $s$ y a partir de ahí seguir una política óptima:
\[ Q^\ast (s,a) = \max_\pi Q^{\pi}(s,a) = Q^{\pi^\ast}(s,a). \]
Los métodos modernos de aprendizaje por refuerzo intentan aprender directamente la función valor o la función $Q$, y una vez Q y una vez que tenemos $Q^{\pi^\ast}$ podemos definir una política óptima de la siguiente manera:
\[ \pi^\ast = \text{argmax}_a Q^\ast (s,a). \]
Además, utilizando la ecuación de Bellman vimos que se podía escribir $Q$ en términos de $V$, y por lo tanto podemos escribir $Q^\ast$ en términos de $V^\ast$ como
\[ Q^\ast (s,a) = \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
La ecuación de Bellman puede escribirse sin hacer referencia a ninguna política de la siguiente manera:
\[ V^\ast (s) = \max_a \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
La función óptima $ Q^\ast$ lo que hace es maximizar sobre todas las decisiones:
\begin{eqnarray*}
Q^\ast &=& r_{t+1} + \gamma \max_{a_{t+1}} r_{t+2} + \gamma^2 \max_{a_{t+2}} r_{t+3}...\\
&=& r_{t+1} + \gamma \max_{a_{t+1}} Q^{\ast} (s_{t+1},a_{t+1}).
\end{eqnarray*}
Formalmente:
\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
\subsection{Métodos básicos de solución: Model-based}
\subsubsection{Programación Dinámica}
Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
				\begin{itemize}
					\item Evaluación de una política: Teniendo una política cualquiera $\pi$, ¿cómo obtener la función $V^\pi$?\\
					         Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
					         \[ V^\pi (s) = \mathbb{E}[r + \gamma V^\pi (s') | s_t = s]. \]
					          Pero esto nos deja con el mismo problema, pues tenemos de ambos lados una función desconocida. Para problemas de decisión finitos, este es un sistema de ecuaciones con tantas ecuaciones como estados, lo cual puede ser costoso de resolver. Consideramos entonces un método iterativo: 
					         \[ V_{k+1} (s) = \pi(s) \sum_{s'} \mathcal{P}_{s,s'}^a [ \mathcal{R}_{s,s'}^a + \gamma V_k(s')], \]
					         donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$. Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.
					\item Mejora de una política: Dada una política $\pi$ y estando en un estado $s$ nos preguntamos qué pasaría si en vez de escoger la acción $\pi(s)$ escogieramos cualquier otra acción $a$. La consecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$. Entonces, la clave es comparar esto con $V^{\pi}$. De ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$. Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.
					 Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
					\[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
					Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.
					\item Iteración de políticas: Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
					\[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
					pues cada política está garantizado que es mejor que la anterior.
					\item Iteración de valor: El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite. Podemos truncar la etapa de policy evaluation.
					\[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
					\item Iteración de política generalizada (GPI): Notar que la iteración de políticas consiste en dos procesos que interactuan entre sí, pero de manera que cada proceso se hace de manera completa antes de comenzar el otro lleva a preguntarse si esto puede hacerse de manera más eficiente. Se conoce como iteración de políticas generalizada (GPI) a la idea general de que los procesos de evaluación y mejora de política interactuen entre sí.
					\item Limitaciones: Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales. Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
				\end{itemize}
\subsection{Métodos básicos de solución: Model-free}
Métodos que no necesitan un modelo completo del sistema, sólo necesitan experiencia. Encontramos los métodos Monte Carlo que simulan las transiciones y los estados, y los métodos de diferencias temporales, que van actualizando iterativamente el valor de los estados observados, el más famoso de estos es el TD($\lambda$).
\subsubsection{Métodos Monte-Carlo} 
Los métodos Monte-Carlo no requieren un modelo del ambiente y son conceptualmente sencillos, pero no son adecuados para hacer actualizaciones incrementales. Estos métodos requieren \textit{experiencia} en vez de un modelo completo; es decir, requieren secuencias de estados, acciones y recompensas, ya sean estos obtenidos de simulaciones o de interacción \textit{on-line} con el ambiente. Los métodos Monte-Carlo suponen que la experiencia proviene de eventos episódicos, es decir que terminan siempre de una u otra forma y es al final de estos episodios que se alteran las estimaciones de valor así como las políticas. En este sentido, los métodos Monte-Carlo son incrementales por episodio, pero no paso a paso dentro de cada episodio. A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general. La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
\begin{itemize}
\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real. Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$. El algoritmo \textit{every-visit} estima $V^\pi (s)$ como las recompensas promedio de todas las visitas al estado $s$ en todos los episodios; por otro lado, el algoritmo \textit{first-visit} promedia sólo lo obtenido al visitar por primera vez un estado en todos los episodios. Ambos métodos convergen a $V^\pi$.
\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado. Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política. Entonces, es importante tener un método para estimar $Q^\ast$.
\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita. El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios. 
\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio. A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
\[ \pi_0 \xrightarrow{E} Q^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} Q^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} Q^\ast.  \]
Bajo el supuesto de episodios infinitos y visita a todas las parejas estado-acción se tiene que los métodos Monte-Carlo encuentran $Q^{\pi_k}$ de manera exacta para $\pi_k$.\\
Para hacer la mejora de una política, se escoge de manera determinística la acción $a$ en la cual se alcanza el máximo de la función $Q$; es decir,
\[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
							\begin{itemize}
									\item On-Policy: intentan evaluar mejorar la política que es actualmente utilizada para la toma de decisiones.
									\item Off-Policy: Evaluan o mejoran una política distinta
							\end{itemize}
\end{itemize}
\subsubsection{Diferencias Temporales}
Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos. Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
\begin{itemize}
\item Evaluación de política en Diferencias Temporales
\item TD(0)
\item SARSA: Es un método \textit{on-policy}
\item Q-Learning: Es un método \textit{off-Policy}
\item Métodos Actor-Critic
\end{itemize}
\section{Deep Reinforcement Learning}
\subsection{Mejoras a los métodos básicos}
	\begin{itemize}
		\item Programación Dinámica con aproximación de funciones.
		\item Cuando el espacio de estados es muy grande, no es factible utilizar una representación en una \textit{look up table}, por lo que es necesario utilizar otro tipo de aproximadores (\cite{busoniu2010reinforcement})
	\end{itemize}
\subsection{Deep Learning}
	\begin{itemize}
		\item Redes neuronales profundas como aproximadores generales.
		\item Redes Convolucionales para poder tomar como input los pixeles de un juego directamente.
	\end{itemize}
\subsection{Value based Reinforcement Learning:}
	\begin{itemize}
	   \item Representar la función valor mediante una \textit{Q-network} con pesos $w$:
	   \[Q(s,a,w) \approx Q^\ast(s,a) \]
	   \item Q-learning:
        \begin{itemize}	  
        	\item  Sabemos que los valores óptimos deben satisfacer la ecuación de Bellman:
        	\[ Q^\ast(s,a)=\mathbb{E}_{s'}[r + \gamma \max_{a'} Q^\ast(s',a')|s,a]. \]
        	\item Tratamos el término $r + \gamma \max_{a'} Q^\ast(s',a',w)$ como el objetivo.
        	\item Minimizar error cuadrático medio mediante Gradiente Estocástico:
        	\[ I = (r + \gamma \max_{a'} Q^\ast(s',a',w) - Q(s,a,w))^2. \]
        	\item Converge a $Q^\ast$ utilizando una representación tabular.
        	\item Diverge al utilizar redes neuronales debido a la alta correlación entre muestras y la no-estacionareidad del objetivo.
        	\item Lo resolvemos con Experience Replay (\cite{Lin1992}),  el cual consiste en guardar todas las experiencias de modo que al entrenar la red tomar sólo mini-batches aleatorios  en vez de utilizar la transición más reciente.
	    \end{itemize}
	\item Aplicaciones de DQN.
		\begin{itemize}
			\item \cite{mnih2013playing}. Paper en el que muestran cómo una computadora aprende a jugar Atari 2600 a través de sólo observar los pixeles del juego y obteniendo una 																		recompensa conforme el score aumentaba. Utilizando la misma arquitectura, se pudieron aprender 7 juegos
			\item \cite{mnih2015human}. En este paper, aumentaron a 49 juegos aprendidos. Utilizan una Deep Q-Network, que consiste en aproximar la función Q mediante una red 																			neuronal que toma como input los estados (las pantallas del juego) y devuelve como output el valor Q para cada posible acción. En este caso, 																			utilizaron una red convolucional con 3 capas de convolución y 2 capas totalmente conectadas.
	  \end{itemize}
\item Mejoras posteriores a DQN
   	   \begin{itemize}
   	   		\item Double DQN.
   	   		\item Prioritized Replay.
   	   		\item Duelling Networks.
   	   \end{itemize}
    	 \end{itemize}  	   
\subsection{Policy Based DRL}
Resulta que Deep Q-Learning no es un gran algoritmo, y en su lugar, \cite{DBLP:journals/corr/MnihBMGLHSK16} proponen utilizar directamente Policy-Gradients.
La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
\[ a = \pi (a|s,u) \textrm{ o } a = \pi (s,u). \]
En este contexto, definimos una función objetivo $L(\textbf{u})$ como:
		\[ L(\textbf{u}) = \mathbb{E}[r_1 + \gamma r_2 + \gamma^2 r_3 + ... |  \pi(\cdot,\textbf{u})]. \]
Optimizamos con Stochastic Gradient Descent. El gradiente de una policy estocástica está dado $\pi (a|s,\textbf{u})$ por:
		\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[  \frac{\partial \log \pi(a|s,\textbf{u})}{\partial \textbf{u}}Q^\pi(s,a)\right].\]
Por otro lado, el gradiente de una policy determinística $a=\pi(s)$ está dado por:
\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[ \frac{\partial Q^\pi (s,a)}{\partial a} \frac{\partial a}{\partial \textbf{u}} \right]. \]
 Los métodos \textit{actor-only}, donde \textit{actor} se utiliza como sinónimo de \textit{policy} utilizan familias parametrizadas de policies. Las ventajas de tener una policy parametrizada es que podemos generar un espectro continuo de accciones, pero los métodos de optimización (policy gradients) sufren de alta varianza en sus estimaciones del gradiente, lo cual genera un lento aprendizaje (\cite{grondman2012survey},\cite{konda2003onactor},\cite{richter2007natural},\cite{boyan2002technical},\cite{baxter2001infinite},\cite{berenji2003convergent}).\\
 \\
Los métodos \textit{critic-only}, donde \textit{critic} se utiliza como sinónimo de \textit{value function}, que utilizan diferencias temporales tienen una menor varianza en las estimaciones del retorno esperado (\cite{boyan2002technical},\cite{berenji2003convergent},\cite{sutton1988learning}). Una manera de derivar una policy en estos escenarios es seleccionando greedy actions; es decir, acciones praa las cuales la value funtion indique que el retorno esperado es máximo, lo cual puede requerir un alto costo computacional. Lo usual es discretizar el espacio continuo de acciones.
Los métodos \textit{actor-critic}, combinan las vetajas de ambos mundos. Mientras que las familias parametrizadas tienen la ventaja de utilizar acciones continuas, los métodos critic-only proveen de conocimiento de baja varianza sobre el desempeño; es decir, la estimación del retorno esperado del crític-only permite al actor-only actualizar los gradientes con información de menor varianza, lo cual acelera todo el proceso. La menor varianza es intercambiada por un mayor sesgo al inicio del aprendizaje (\cite{berenji2003convergent}). \\
\\
Los métodos de Policy Gradients optimizan directamente la recompensa acumulada y pueden utilizarse junto con aproximadores no-lineales como las redes neuronales. Este método presenta dos grandes retos: el número de muestras de entrenamiento requeridos y la dificultad de obtener mejoras estables debido a la no-estacionariedad de los datos.  \cite{DBLP:journals/corr/SchulmanMLJA15} atacan el primer problema se puede atacar utilizando value functions para reducir la varianza de los estimados. El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
\\
 \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada. Es un algoritmo similar a Policy Gradients y resulta útil para optimizar políticas con redes neuronales. Buen performance en juegos de Atari que utilizan la pantalla como input.

\section{Reinforcement Learning y Causalidad}
Aprendemos causalidad mediante intervenciones, interacciones, etc. Reinforcement Learning es aprender a realizar tareas mediante interacción. ¿Habrá algo en medio?
\subsection{Conocimiento causal y observabilidad parcial}
Los algoritmos actuales de reinforcement Learning suponen que el problema puede modelarse dentro de un contexto Markoviano de decisión con recompensas. Si el supuesto markoviano se elimina, los algoritmos podrían dejar no funcionar, como muestran en \cite{singh1994learning}. Además, todo el desarrollo teórico utiliza la propiedad Markoviana.\\
\\
Una manera en la que la propiedad Markoviana puede perderse es si el agente no cuenta con información completa sobre el estado; es decir, al igual que en los HMM, suponemos que el ambiente es Markoviano, pero el agente no tiene acceso a él de modo que los datos no le \textit{parecen} Markovianos. A diferencia de los HMM, no existe un procedimiento computacionalmente tratable para los POMDP. El problema es que una vez que se obtienen las estimaciones del valor para cada estado, se debe hacer programación dinámica en espacios continuos y esto es infactible salvo para casos pequeños\\
\\
La manera de generalizar el caso markoviano al caso no observado es considerar un MDP subyacente con estados $\mathcal{S}=\{ s_1,...,s_N\}$ y probabilidades de transición $\mathcal{P}^a_{s,s'}$. Luego, en cada tiempo $t$, el agente observa un mensaje $m_t$ el cual proviene de una distribución $P(m | s_t)$ desconocida.\\
\\
En \cite{jaakkola1995reinforcement} se describe un procedimiento que evita estimar el valor de los estados y trabaja directamente en el espacio de políticas estocásticas. El procedimiento es \textit{directo}, a diferencia de los métodos \textit{indirectos} en los cuales primero se aprenden los parámetros del MDP y luego se utiliza programación dinámica para obtener la política. El algoritmo propuesto consiste en dos partes: primero, se utiliza un método Monte-Carlo para calcular un análogo de los Q-values y luego se utiliza un paso de policy-improvement.\\
\\
Desde un punto de vista de psicología, una manear de entender la inferencia de estados ocultos es en términos de causas latentes (\cite{gershman2015reinforcement}). Siguiendo esta idea, es evidente que un agente de RL que opera en el mundo real no sólo debe hacer inferencia de estados ocultos, sino que debe poder descubrir los estados ocultos que están detrás del mundo que éste observa. Esto es una forma de aprendizaje estructural latente (\cite{gershman2010learning}).\\
\\
Un primer acercamiento a este problema puede hacerse mediante técnicas de inferencia Bayesiana no paramétrica. Por ejemplo, tanto \cite{gershman2010learning} como \cite{gershman2012exploring} utilizan priors no paramétricas para hacer inferencia sobre causas latentes. Un modelo causal latente es una buena representación de la estructura causal verdadera bajo condicionamiento Pavloviano, pues es el experimentador quien es una causa latente.\\
\\
En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
\subsection{Towards Deep Symbolic Reinforcement Learning}
Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
\\
\cite{garnelo2016towards} proponen una arquitectura híbrida que, combina redes neuronales con inteligencia artificial simbólica. El modelo propuesto consiste en combinar una red neuronal que como entrada directamente lo que se recibe del ambiente y transforma estos datos crudos en una representación simbólica basada en lógica de primer orden; esta representación simbólica es posteriormente utilizada para escoger la acción a tomar.\\
\\
Se utiliza un sencillo juego para probar la arquitectura y en este contexto la red neuronal debe generar representaciones simbólicas de secuencias de estados en el juego en las cuales el flujo de pixeles queda codificado en términos de objetos, tipos y sus interaccioes. Luego, con esto se aprenden las políticas.\\
\\
El primer paso que hacen es utilizar un autoencoder convolucional para generar, de manera no supervisada, un conjunto de símbolos que representarán los objetos en cada escena. Luego, para cada pixel, obtienen la feature con mayor activación y se toman los más altos. De esta manera se obtiene un pixel representativo para cada objeto a los cuales posteriormente se le asigna un tipo simbólico según el autoencoder.\\
\\
Luego, una vez que se tienen estos símbolos de bajo nivel para cada cuadro del juego, se necesita rastrearlos a lo largo del tiempo. Utilizan una verosimilitud que es función de proximidad espacial y el tipo de transiciones. Para considerar la información de la dinámica de los cuadros utilizan la diferencia entre cuadros y las posiciones relativas entre objetos.\\
\\
Para la parte de aprendizaje, entrenan una función $Q$ por separado para cada interacción entre objetos con actualización
\[Q^{ij}(s_t^{ij},a_t) \leftarrow  Q^{ij}(s_t^{ij},a_t) + \alpha \left( r_{t+1} + \gamma (\max_{a} Q^{ij} (s_{t+1}^{ij},a) - Q^{ij}(s_t^{ij},a_t)) \right). \]
\\
Se prueba el framework sobre varias variantes de un sencillo videojuego. \\
\\
Llama la atención como estos autores utilizan maquinaria de lo más reciente (deep learning, autoencoders) para luego obtener una representación en términos de lógica de primer orden. Este approach ha quedado claro que ya quedó atrás por lo que debería buscarse algo que sustituya este enfoque lógico. En particular, pudiera ser esto un modelo gráfico causal el cual vaya aprendiéndose conforme el agente de RL aprende sobre su entorno.

\subsection{Conocimiento causal y toma de decisiones secuenciales: acciones como intervenciones}
En cuanto al aprendizaje \textit{on-line} de modelos causales, encontramos en los trabajos de \cite{wellen2012learning} y de \cite{kummerfeld2013tracking} acercamientos tangenciales, o un punto de inicio, al aprendizaje de modelos causales al vuelo.\\
\\
Teorías de cómo las personas usan datos de covariación para aprender modelos gráfios causales pueden dividirse en enfoques racionales y heurísticos (según  \cite{wellen2012learning}). El enfoque racional modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana. Estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
\\
Por otro lado, el enfoque heurístico propone que las personas reaccionan a distintos estímulos que los hacen moficiar sus hipótesis causales de una manera no necesariamente racional. El modelo de efecto simple de \cite{waldmann2008causal} supone que las personas se enfocan en evaluar una sola relación causal a la vez. Es un modelo que aprende la magnitud de la relación causal. Si una variable supera cierto umbral, entonces la persona acepta esa relación causal y la incorpora a su bagaje de relaciones.\\
\\
El modelo propuesto por \cite{wellen2012learning} sigue esta línea del efecto simple, con la diferencia de que este modelo sí intenta descubrir el proceso cognitivo subyacente, pues el aprender magnitudes causales es un proceso puramente computacional. El modelo de efecto simple puede verse como una implementación del modelo de \cite{danks2003dynamical}. Además, según narran \cite{wellen2012learning}, el modelo de efecto simple no explica cómo la persona utiliza las magnitudes causales para determinar si existe o no una relación causal. En el modelo propuesto, \cite{wellen2012learning}, incorporan un procedimiento de decisión para determinar si existen o no relaciones causales.\\
\\
El modelo LPL (\cite{wellen2012learning}) comienza con una estructura causal inicial, la cual consiste en un grafo acíclico dirigido el cual representa las creencias iniciales de un individuo. Posteriormente, el modelo altera esta estructura al añadir o remover una sola arista, de manera que el problema se reduce a evaluar relaciones causales entre dos variables. Este modelo genera estimaciones de la fuerza de la relación causal para cada posible relación causa-efecto que no sea eliminada a priori.\\
\\
Supongamos que un individuo cree a priori que $B$ causa $E$, y está tratando de averiguar si $C$ causa $E$. A diferencia del modelo de efecto simple, las estimaciones de la fuerza causal están generadas por un mecanismo similar a los filtros de partículas Bayesianos
En particular, a partir del conocimiento causal inicial del individuo, el modelo LPL comienza con $n$ partículas. Estas partículas $\{V_C^1,...,V_C^n \}$ representan las creencias actuales del individuo sobre si $C$ causa $E$. Paralelamente, existen $\{ V_B^1,...V_B^n \}$ para $B$. Se define la \textit{capa} $i$ de partículas para el efecto $E$ como la $i$-ésima partícula de cada causa conocida y posible para $E$. En el ejemplo, tendríamos que esta capa es $\{ V_B^i, V_C^i \}$. Luego, el modelo LPL actualiza cada capa de manera independiente de la siguiente manera
\[ V_C^{i+1} = V_C^{i,t} + \alpha ( \textrm{observado} - \textrm{esperado}), \]
Donde “observado” tiene el valor de 1 si el efecto sucede y 0 en caso contrario. Por otro lado, “esperado” es el valor esperado del efecto, calculado según la forma funcional de la relación causa-efecto, el cual es calculado por separado para cada causa potencial en una capa. El valor esperado de $E$ (para la capa $i$ y la causa potencial $C$) es
\[ \textrm{esperado} = \prod_{k} (1-V_k^{t-1})(1 - \prod_j (1- V_j^{t-1}).\]
Conforme el individuo observa más datos, las partículas convergen a los valores verdaderos de los parámetros y las estimaciones para cada capa convergen. Este procedimiento no toma en cuenta aspectos estructurales, los cuales se modifican según un criterio de decisión. Lo que hace el modelo es una prueba $t$ en la cual la hipótesis nula es que las partículas provienen de una distribución con media cero. El criterio de decisión es el siguiente: si actualmente no existe una arista, y la prueba $t$ rechaza la hipótesis nula, entonces se añade la arista. Si existe una arista, y la prueba no es significativa, entonces se elimina la arista existente. Alterar una arista afectará las futuras estimaciones del valor “esperado” para otras causas.\\
\\
\cite{hagmayer2013hierarchical} llaman la atención sobre que los modelos actuales de Decisiones (maximum expected utility) no asumen que los tomadores de decisiones aprenden sobre la estructura causal del problema. Por otro lado, las teorías de aprendizaje causal proponen que un tomador de decisiones guían sus acciones a través de conocimiento causal previamente obtenido. El modelo de \cite{sloman2006causal} extiende las teorías de aprendizaje causal al contexto de toma de decisiones, la idea básica consiste en que los tomadores de decisiones utilizan la información disponible de su entorno para inducir un modelo causal del problema de decisión al que se enfrentan. Se sabe que en los seres humanos el conocimiento causal guía la interpretación de información probabilista y que los humanos somos capaces de aprender la estructura causal del entorno a través de una serie de (traducir cues) (\cite{lagnado2007beyond}). La pregunta fundamental que plantean \cite{hagmayer2008causal} es entender qué es lo que aprenden las personas al tomar decisiones de manera secuencial de tal forma que se busque maximizar cierta recompensa a largo plazo y donde las acciones alteran el entorno. Al exponerse una persona a estas situaciones de decisión secuencial se obtiene una retroalimentación de cómo las acciones se relacionaban con la obtención de recompensas. Las distintas teorías sobre toma de decisiones bajo incertidumbre suponen distintas representaciones para esta retroalimentación. Hagmayer propone agrupar estas distintas representaciones en tres niveles:
\begin{itemize}
\item Nivel 1: Opciones, resultados y recompensas. Sólo los valores esperados de las opciones posibles afectan la toma de decisiones. Se conoce como \textit{habit learning} en la literatura de aprendizaje en animales.
\item Nivel 2: Opciones, relaciones causales de estas opciones a los resultados, relaciones causales entre variables de resultado y relaciones de estas con las recompensas.  Aquí entra lo que se conoce como \textit{goal-directed learning}. La diferencia con el nivel 1 es la capacidad de distinguir entre acciones y los valores asociados a estas.
\item Nivel 3: Opciones, relaciones causales entre variables de resultado, relaciones causales entre estas variables y relaciones de estas con las recompensas. Estos son los modelos causales. Las distintas teorías sobre estos modelos suponen que los humanos (o el agente que aprende) adquieren estos modelos mediante procesos de aprendizaje. Algunas teorías suponen que la estructura causal del entorno se obtiene a través de \textit{claves}, como la propuesta por \cite{lagnado2007beyond}. Por otro lado, otros autores opinan que las personas utilizan propiedades estadísticas de los datos observados, este sería el caso de las teorías de \cite{gopnik2004theory} y de \cite{griffiths2005structure}.
\end{itemize}
Aunque las teorías que entran en el nivel 2 son sensibles a las correlaciones entre las respuestas obtenidas, sólo en el nivel 3 se hace una distinción entre causalidad y correlación.\\
\\
Los autores hacen énfasis en que aunque las personas sí utilicen modelos causales al tomar decisiones simples, estas decisiones suelen ser sobre escenarios hipotéticos, pero no se han hecho muchos estudios sobre adquisición de conocimiento causal al enfrentarse a problemas de decisión secuencial con el objetivo de maximizar alguna ganancia a largo plazo. Además, en los estudios realizados, se ha visto que las personas tienen la capacidad de inducir modelos causales a partir de claves en el entorno; se sabe además que a partir de intervenciones activas se toman mejores decisiones que considerando sólo datos observados de manera pasiva y que las personas tienen la capacidad de integrar relaciones causales obtenidas de manera separada. Pero en estos estudios las personas tenían como objetivo descubrir de manera directa estructuras causales. No se sabe bien hasta qué punto las personas adquieren conocimiento causal al perseguir objetivos que no sean descubrir directamente estructuras causales a partir del ambiente.\\
\\
Los autores realizaron tres experimentos para explorar las representaciones que las personas hacen de las relaciones causales que adquieren al tomar decisiones para maximizar algún objetivo. En el primer experimento, el objetivo era estudiar cómo las creencias causales de las personas alteran la toma de decisiones y cómo los participantes utilizaban los resultados de sus decisiones para revisar el modelo causal. El experimento consistía en dos etapas de toma de decisiones, la segunda siendo una etapa de prueba. En ambas etapas los participantes debían escoger entre una serie de opciones de modo que maximizaran ganancias. Se le sugirió a cada participante una posible cadena causal. En la primera etapa se les proveía retroalimentación de las consecuencias de sus decisiones, lo cual no sucedió en la segunda etapa sino que ellos mismos debían inferir posibles relaciones causales. Se concluyó que los participantes encontraron estructura causal en el problema y la utilizaron para tomar futuras decisiones.\\
\\
En el segundo experimento y tercer experimento, a los participantes se les mostró una serie de hipótesis causales para ver si inducían de manera espontánea un modelo causal de la situación.\\
\\
Lo que estos experimentos muestran, en conclusión de los autores, es que al intentar maximizar una ganancia a largo plazo, las personas adquieren conocimiento causal del entorno, el cual afecta futuras decisiones. Mencionan que estos experimentos ponen en tela de juicio los modelos estándar de toma de decisiones, pues utilizando sólo valores esperados no se llegarían a los mismos resultados.\\
\\
Como consecuencia de sus estudios, se obtiene que los métodos basados en los niveles 1 y 2, que es donde caerían los métodos tradicionales de aprendizaje por refuerzo no son suficientes para una adecuada toma de decisiones en un sistema causal, pues estas formulaciones no cuentan con la expresividad suficiente para capturar las relaciones causales. Esto confirma la necesidad de pensar en modelos que permitan aprender modelos causales a la par que un agente aprende de su entorno mediante interacción y exploración.\\
\\
En cuanto a la formulación de un problema de decisión en sistemas causales, ya desde \cite{dawid2002influence} tenemos una idea de cómo aumentar redes bayesianas con nodos de decisión y variables de estrategia para el modelado causal en problemas de decisión.\\
\\
\cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta. 
\section{Propuesta}
\subsection{Pregunta de Investigación}
¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
\subsection{Problema}
\subsection{Objetivos}
\begin{itemize}
\item Aprendizaje causal en línea a partir de acciones y recompensas obtenidas por un agente.
\item Evaluación del modelo causal.
\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
\item Exportar conocimiento causal obtenido para hacer inferencia causal.
\end{itemize}
\subsection{Propuesta metodológica}
Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en incorporar un modelo de aprendizaje causal dentro a un problema de toma de decisiones. En el contexto de aprendizaje por refuerzo, tendremos un agente que, al tener como objetivo aprender una tarea mediante interacciones secuenciales, aprenda con cada paso un poco sobre la \textit{estructura causal} de su entorno. Cada decisión que el agente toma (o acción que lleva a cabo) puede verse como una intervención en el ambiente y que proporciona información sobre la estructura causal de este. Esta información obtenida, será utilizada para mejorar las decisiones que toma y para aprender otros aspectos estructurales (causales) sobre su ambiente. La intuición detrás de esto es que al aprender mediante una toma de decisiones repetida, decisiones que intervienen directamente sobre el entorno, se aprenden también relaciones causales; por ejemplo, al aprender a conducir se aprende implícitamente que ciertos movimientos del volante \textit{causan} ciertos cambios en el estado del automovil (\cite{danks2014unifying}). Otro ejemplo es en el contexto de videojuegos, en el cual al tomar ciertas decisiones se conoce el impacto que estas provocan en el estado del juego, conocimiento que a su vez es utilizado para tomar mejores decisiones en un futuro y que maximicen las recompensas u objetivos a largo plazo. No sólo hay intuición detrás de esto, pues se sabe que los seres humanos utilizan el conocimiento causal que la experiencia les provee para realizar predicciones para intervenciones en el mundo no realizadas aun (\cite{meder2008inferring}, \cite{hagmayer2009decision})\\
%\\
%De manera formal, consideremos un fenómeno sobre el cual se quieran hacer inferencias causales y sobre el cual poseeemos un conjunto de datos. Supongamos que los datos tienen la forma $%(\mathbf{X},\mathbf{Y})$ donde $\mathbf{X}$ son un conjunto de atributos, así como acciones llevadas a cabo, y $\mathbf{Y}$ son los efectos observados. Entonces, es necesario %\textit{traducir} estos datos a la forma $(s \in \mathcal{S},a \in \mathcal{A},r \in \mathcal{R})$ de modo que las características, atributos, etc del problema original sean mapeados a estados, las acciones tomadas, a acciones en el mundo de RL, y las respuestas obtenidas, al reward y cambio de estado otorgado por el ambiente. Evidentemente esta no es una tarea trivial, pues es necesario también definir un \textit{task} adecuado para el agente que le permita interactuar con su entorno de manera tal que aprenda, o descubra, las relaciones causales sobre los datos originales. La definición del task a resolver implica definir una función recompensa; además, una transición entre estados es importante. Notemos que todos estos elementos deben ser definidos de manera que el agente interactue con su mundo de tal forma que lo que \textit{aprenda} nos sirva para identificar relaciones causales.\\
%\\
%Como ejemplo de lo anterior, podríamos tener un conjunto de datos de la forma $\mathbf{X} = (s,t)$ donde $s$ es enfermedad, $t$, tratamiento otorgado y $\mathbf{Y}=\{ \text{curado}, \text{ no curado} \}$. Definimos un agente de RL tal que los estados que recibe del ambiente son las características del paciente, la acción que toma es otorgar o no el medicamento, y la recompensa es $1$ si el paciente se mejoró. La tarea que este agente tiene que lograr es maximizar el número de pacientes curados; notemos que la tarea del agente, o el agente mismo, no tienen por qué tener interpretación en términos del problema original; es decir, no tiene por qué existir una interpretación del agente en términos de un médico, un paciente, etc. Una vez que el agente aprenda a resolver esta tarea, utilizaremos lo aprendido por él de la siguiente manera: si para un paciente del conjunto original de datos el agente propone el tratamiento, entonces es causal que el tratamiento mejoró al paciente.\\
\\
El trabajo de \cite{garnelo2016towards} nos muestra que efectivamente pueden aprenderse aspectos estructurales sobre el entorno mientras el agente aprende, incluso logrando así mejorar su desempeño. Además, muestra la importancia de tener una semántica que permita hablar de la estructura del entorno. Ellos utilizaron lógica de primer orden, pero un modelo gráfico sería más expresivo y adaptable.\\
\\
A su vez, el trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones. Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo. \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones. Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.\\
\\
\begin{center}
\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/diagrama.png}
\end{center}
\newpage
\bibliographystyle{apalike}
\bibliography{Bibliografia}
\end{document}

### 2017-12-20 Primer_Semestre/Resumenes/Resumen_IRM.tex
primer commit

--- 2017-12-20 Primer_Semestre/Propuesta/Propuesta.tex
@@
 \usepackage[margin=1.5cm]{geometry}
-\usepackage{natbib}
-%\bibliographystyle{stylename}
 %\usepackage{fancyhdr}
@@
 \newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Survey}
+\title{Resumen: Learning Systems of Concepts with an Infinite Relational Model - Kemp, Tenenbaum, Griffiths}
 \author{Mauricio Gonzalez Soto}
@@
 \maketitle
-\tableofcontents
-\newpage
 \section{Introducción}
-El problema al cual nos enfrentamos consiste en el aprendizaje de relaciones causales a partir de datos.
-Los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.
-Al tratar de averiguar relaciones causales, es común contar con información previa sobre el fenómeno dentro del cual se quiere investigar causalidad, por lo que recurrimos a la estadística bayesiana para el manejo de esta información inicial.\\
+Si cierto dominio consiste en varios tipos de entidades,
+\section{Infinite Relational Models}
+Suponiendo que tenemos una o más relaciones que involucran uno o más tipos.
+El IRM busca partir cada tipo (type) en clusters, de manera que una buena partición permita que relaciones entre entidades se puedan predecir a partir de la asignación de clusters.\\
 \\
-La primera parte de este resumen consiste en Redes Bayesianas, que son objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
-Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
-\\
-Veremos que al intentar obtener relaciones causales a partir de información proveniente del mundo real necesitamos obtener las entidades que se relacionan entre sí de manera causal, por lo que haremos un repaso de técnicas utilizadas en extracción de relaciones.\\
-\\
-Por otro lado, en la literatura proveniente de Psicología Computacional y Ciencias Cognitivas, se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
-Se sabe también que el conocimiento causal adquirido es utilizado para mejorar las futuras decisiones tomadas.\\
-\\
-Finalmente, se discutirá un poco sobre Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.
-Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.
-Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones.
-La pregunta se convierte entonces en cómo obtener e incorporar conocimiento causal en un problema de aprendizaje por refuerzo.
-Además, uno se pregunta si es posible utilizar este conocimiento causal obtenido al interactuar sobre el mundo para realizar otro tipo de inferencias causales en ciertos dominios de interés.
-\section{Causalidad}
-\subsection{Intro}
-Saber si es posible hablar de relaciones causa-efecto es un tema polémico; los métodos probabilistas tradicionales sólo codifican asociaciones estadísticas, pero la idea
-La diferencia entre modelos probabilistas y modelos causales se da cuando es de interés tomar acciones que modifiquen el entorno y no nada más observar los valores de las variables. \\
-Según \cite{pearl2014probabilistic}, las dos preguntas fundamentales en Causalidad son:
-\begin{itemize}
-\item ¿Qué evidencia empírica es necesaria para hacer inferencias válidas sobre relaciones causa-efecto?
-\item Dado que estamos dispuestos a aceptar información causal sobre un fenómeno, ¿qué inferencias podemos obtener a partir de esta información, y cómo?
-\end{itemize}
-En cuanto al segundo punto, Pearl considera que hay tres tipos de preguntas a considerar
-\begin{itemize}
-\item Predicciones: si sabemos que una persona ingirió alimentos de dudosa procedencia, ¿se enfermará?
-\item Intervenciones: si le damos cierto tratamiento a un paciente, ¿mejorará?
-\item Contrafactuales: ¿el paciente del punto uno se habría enfermado de todos modos si no hubiera consumido esos alimentos?
-\end{itemize}
-Pearl argumenta que aunque causalidad suele entenderse como afirmaciones necesarias, estas son hechas en contexto de incertidumbre, por lo que cualquier teoría causal debe estar en términos de probabilidades.
-\subsection{Un poco de historia}
-Aristóteles definió 4 \textit{causas} para una cosa, una de ellas llamada \textit{causa eficiente} que estaba definida como aquello que hace la cosa.\\
-\\
-En 1690, el filósofo inglés John Locke propuso como definición de causa “aquello que produce alguna idea simple o compleja”, y lo que es producido lo definió como efecto.\\
-\\
-\cite{holland1986statistics} resalta la importante diferencia en que Aristóteles hizo énfasis en las causas de una cosa, pero no en el efecto de estas. \\
-\\
-Fue David Hume quien se considera que hiz una importante contribución, y él definió la causalidad como una relación entre experiencias más que una entre hechos.
-Argumentó que no es verificable empíricamente que la causa produzca el efecto, sino sólo que el evento llamado \textit{causa} es invariablemente seguido por aquella experiencia llamada \textit{efecto}.\\
-\\
-Hume definió 3 criterios básicos para causalidad:
-\begin{itemize}
-\item Contiguidad espacio-temporal: tanto causa y efecto tienen lugar en una entidad común (espacio y tiempo).
-\item Sucesión temporal: La causa tiene que preceder temporalmente al efecto. 
-\item Conjunción constante: Causa y efecto ocurren juntos, o no ocurren juntos.
-\end{itemize}
-John Stuart Mill argumentó tajantemente que observación sin experimentos no puede probar causalidad. Mill identificó cuatro métodos generales para la identificación de relaciones causales. 
-\subsection{Modelo de Rubin}
-El modelo de Rubin parte de una población de \textit{unidades} $U$.
-En general, estas unidades son los objetos de estudio sobre las cuales actuan las causas.
-Para hacer más sencilla la descripción del modelo, supondremos que sólo existen dos causas (o niveles de tratamiento) denotadas por $t$ y $c$.
-Sea $S$ una variable indicadora de la causa a la cual cada unidad $u \in U$ fue expuesta.
-Además, consideramos dos variables respuesta $Y_t$, $Y_c$.
-$Y_t(u)$ reoresenta el valor de la respuesta si la unidad $u$ es expuesta a $t$ y análogo para $c$.\\
-\\
-\begin{defi}
-Se define el efecto de la causa $t$ sobre $u$ relativo a la causa $c$ como
-\[ Y_t(u) - Y_c(u). \]
-Esta expresión se lee de la siguiente manera: $t$ causa el efecto $Y_t(u) - Y_c(u)$ en la unidad $u$ (relativo al control $c$).
-\end{defi}
-Notemos que es imposible observar simultáneamente el efecto de $t$ y de $c$ sobre una misma unidad $u$ y a esto \cite{holland1986statistics} le llama el Problema Fundamental de la Inferencia Causal.
-El problema, según explica Holland, radica en la palabra \textit{observar} pues no podemos saber simultáneamente el efecto de exponer a un enfermo en particular a un nuevo tratamiento y al mismo tiempo el efecto de no exponerlo.
-\subsection{Modelo de Suppes}
-En su libro “Probabilistic Theory of Causality”, (\cite{suppes1970probabilistic}), Patrick Suppes intenta mejorar el análisis de Hume sobre causalidad; en particular, la condición de conjunción constante.
-Para Suppes, no importa qué sean las causas y los efectos, sino que se puedan expresar como eventos que ocurren bajo ciertas condiciones.
-Suppes supone que todas las causas necesariamente preceden en el tiempo a sus efectos.
-Define una causa \textit{prima facie} de un evento como un evento que precede temporalemnte a este y que está asociado de manera positiva con él.
-Define como una causa espuria a una causa prima facie tal que es condicionalmente independiente del efecto dada otra causa la cual es anterior en el tiempo a la causa prima facie y que está asociada con el evento dada la causa.
-Una causa genuina es una causa prima facie que no es espuria.
-\begin{defi}
-Si $r < s$ representan valores de tiempo, se dice que el evento $C_r$ es una causa prima facie del evento $E_s$ si
-\[ P(E_s | C_r ) > P(E_s). \]
-\end{defi}
-\begin{defi}
-Un evento $C_r$ se dice que es una causa espuria de un eventi $E_s$ si $C_r$ es una causa prima facie de él y para alguna $q < r < s$ existe un evento $D_q$ tal que 
-\[P(E_s | C_r, D_q) = P (E_s | D_q) \]
-y además,
-\[P(E_s | C_r , D_q) \geq P(E_s | C_r). \]
-\end{defi}
-Por lo tanto,
-\begin{defi}
-Para tiempos $r < s$, un evento $C_r$ se dice que es una causa genuina de un evento $E_s$ si $C_r$ es una causa prima facie de $E_s$, pero no es una causa espuria.
-\end{defi}
-Notemos que Suppes está definiendo la causa de un efecto, más que el efecto de una causa.
-Además, no hay restricciones a la naturaleza de una causa más allá de que sea un evento que ocurre antes en el tiempo.
-El modelo de Suppes no puede expresar el efecto de una causa en particular, sólo comportamientos promedio (\cite{holland1986statistics}).\\
-\\
-En palabras de Holland, la causalidad de Suppes no es más que correlación entre causa y efecto que no puede ser eliminada al parcializar causas legítimas.
-\subsection{Modelos Causales en Medicina}
-Un caso de aplicación en medicina es determinar que ciertas bacterias causan ciertas enfermedades.
-Existe un conjunto de postulados planteados por Koch para decidir cuándo un micro-organismo está implicado en una enfermedad.
-Aunque no existe un planteamiento formal, estos pueden expresarse como (\cite{yerushalmy1959methodology}):
-\begin{itemize}
-\item El organismo debe estar presente en todos los casos de la enfermedad supuestamente provocada por él.
-\item El organismo debe aislarse de los pacientes y cultivarse en estado puro.
-\item El organismo cultivado en estado puro debe reproducir el padecimiento. 
-\end{itemize}
-Estos postulados aseguran que no haya datos que apoyen una hipótesis nula de cero efecto causal. \\
-\\
-Por otro lado, Hill, quien fue de los primeros en estudiar la asociación entre fumar y cáncer, encontró 9 factores útiles a la hora de decidir que una asociación observada es causal. 
-\subsection{Causalidad de Granger en Economía}
-La causalidad de Granger se utiliza para series de tiempo, lo cual es de alta importancia en Economía.
-\cite{holland1986statistics} opina que las ideas esenciales de Granger no están limitadas a series de tiempo y que pueden ser expresadas en términos de variables que actuan sobre unidades de una población.\\
-\\
-En la teoría de Granger, una variable causa otra variable; es decir ( \cite{holland1986statistics}), las realizaciones de una variable mejoran las predicciones hechas sobre la otra variable.
-La única parte en la que se utilizan series de tiempo en esta definición es para hablar de valores de las variables definidos hasta un cierto tiempo $t$.\\
-\\
-Si $X,Y,Z$ son variables aleatorias definidas sobre cierta población, se dice que $X$ no es una causa de Granger de $Y$ si $X$ y $Y$ son condicionalmente independientes dado $Z$.
-Entonces, $X$ es una causa de Granger de $Y$ si distintos valores de $X$ llevan a distintas distribuciones predictivas de $Y$ dados $X$ y $Z$; esto es, $X$ ayuda a predecir $Y$ aun cuando $Z$ es tomada en cuenta.
-La no-causalidad de Granger es parecida a las causas espurias de Suppes pues ambas consideran que no se pueda predecir un evento dada cierta información.\\
-\\
-En opinión de Holland, la causalidad de Granger falla de la misma manera que la de Suppes, pues a la luz de nueva información una causa Granger puede dejar de serlo.
-\subsection{Modelos Causales en Ciencias Sociales}
-\subsection{Aprendizaje de Modelos Causales a partir de datos}
-En los métodos de estado del arte de Modelos Causales Funcionales suele hacerse los siguientes supuestos (\cite{goudet2017learning}):
-\begin{itemize}
-\item Suficiencia Causal (CSA)
-\item Propiedad Markoviana (CMA)
-\item Independencias Condicionales
-\item Fiabilidad causal (CFA)
-\end{itemize}
-\subsubsection{Aprender el CPDAG}
-Existen tres familias de métodos para extraer el grafo: métodos de restricciones, métodos de score y métodos híbridos.\\
-\\
-Los métodos basados en restricciones explotan las independencias condicionaes para identificar todas las v-estructuras.
-El algoritmo más conocido es el Algoritmo PC (\cite{spirtes2000causation}).
-El algoritmo PC primero construye un esqueleto basado en independencias condicionales; luego, identifica las v-estructuras y posteriormente orienta las aristas.
-El algoritmo FCI (\cite{spirtes1999algorithm}) extiende el algoritmo PC al relajar la suficiencia causal e incorporar variables latentes.
-El algoritmo RFCI de \cite{colombo2012learning} es más rápido que FCI y maneja datos de alta dimensionalidad.
-La debilidad de estos métodos, según \cite{goudet2017learning} es que dependen demasiado en pruebas de independencia, que dependen a su vez de la cantidad disponible de datos.\\
-\\
-Los métodos basados en score exploran el espacio de CPDAG's y minimizan un score global.
-La exploración se realiza en términos de operaciones sobre grafos, por ejemplo añadir o remover aristas.
-El algoritmo Greedy Equivalent Search (GES) de \cite{chickering2002optimal} encuentra la estructura óptima en términos del BIC.\\
-\\
-Los métodos híbridos combinan las técnicas anteriores.
-Por ejemplo el Max-Min Hill Climbing Algorithm de \cite{tsamardinos2006max} primero construye un esqueleto utilizando pruebas de independencia condicional, y luego le da orientación a las aristas utilizando un método greedy sobre un score bayesiano.
-\subsubsection{Explotando la asimetría entre causa y efecto}
-Los métodos anteriores no toman en cuenta la asimetría entre causas y efectos.
-La intuición detrás (\cite{goudet2017learning}) está dada por el siguiente ejemplo: Si consideramos el modelo causal funcional $Y=X+U$, entonces los métodos anteriores basados en grafos no pueden darle orientación a la arista $X - Y$, pues tanto $X \rightarrow Y$ como $Y \rightarrow X$ son Markov equivalentes.\\
-\\
-\cite{stegle2010probabilistic} infiere la dirección causal al construir dos modelos Bayesianos generativos: uno para $X \rightarrow Y$ y otro para $Y \rightarrow X$.
-La dirección causal es determinada a partir del modelo generativo que mejor ajuste a los datos.
-Otros métodos, como el propuesto por \cite{sgouritsa2015inference} se basan en que si $X \rightarrow Y$, entonces la probabilidad marginal de la causa $P(X)$ es independiente del mecanismo causal $P(X | Y)$; por lo tanto, estimar $P(Y | X)$ a partir de $P(X)$ no debería poderse, pero estimar $P(X | Y)$ a partir de $P(Y)$ sí debería ser posible.
-\subsubsection{Como problema de Aprendizaje}
-Inspirados en los concursos de Kaggle de \cite{Guyon2013}, \cite{lopez2015towards} plantean el problema de inferencia causal como un problema de clasificación de medidas de probabilidad.
-Los datos provistos por el concurso son 16,200 parejas de variables $(X_i,Y_i)$.
-Cada par consta de una realización $(x_i,y_i)$ y una etiqueta $l_i$ que indica el valor de la verdadera relación causal; es decir, $l_i$ indica uno de los siguientes casos: $X \rightarrow Y,$ $Y \rightarrow X$, $X_i \perp Y_i$, $X \leftrightarrow Y$.\\
-\\
-Notemos que este método tiene como punto débil que depende de la representatividad de la muestra de entrenamiento y que supone la existencia de una distribución madre
-\subsubsection{Modelos Causales Funcionales Neuronales}
-\cite{goudet2017learning} proponen un esquema que consiste en tres pasos: primero, identificar el esqueleto según propiedades de independencia estándar; luego, orientar las aristas utilizando métodos bivariados estándar y finalmente hacer una optimización sobre todo el grafo examinando relaciones locales.
-Este último paso se realiza al aprender estos modelos locales como redes neuronales generativas; es decir, se modela la distribución $P(X | X_{Pa_i}, U_i)$ de un nodo condicional en sus padres (causas) con una red neuronal.\\
-\\
-Formalmente, se modela cada ecuación
-\[X_i = f_i (X_{Pa_i},U_i) \]
-como una red neuronal generativa global, en la cual cada función $f_i$ es modelada por una red neuronal generativa con una capa oculta.\\
-\\
-Encontrar un modelo causal de este tipo requiere llevar a cabo dos tareas: estimar la estructura (el grafo) de los padres de cada variable, y las redes neuronales $f$
-\subsection{Correlación y causalidad}
-Dos variables aleatorias $X$,$Y$ pueden estar correlacionadas en distintos casos; por ejemplo, si $X$ causa $Y$, o si tanto $X$ como $Y$ son causadas por otra variable desconocida $V$.
-Si e conoce la existencia de $V$ y además podemos observarla.
-\section{Modelos gráficos causales}
-Sabemos que en una red bayesiana los arcos, aunque dirigidos, no necesariamente tienen una interpretación directa, lo único que importa es que el grafo capture adecuadamente las relaciones de dependencia entre variables aleatorias.
-Nos gustaría que fuera el caso que $X \to Y$ significara $X$ \textit{causa} Y.
-Resulta que existe una familia de modelos causales en los cuales se \textit{interviene} en el mundo.
-La idea clave es utilizar \textit{intervenciones ideales}, las cuales se denotan $do(Z=z)$ que significa forzar a la variable aleatoria $Z$ a tomar el valor $z$, denotado de manera simple como $do(z)$.
-Esta operación corresponde al caso en que un agente directamente modificó el estado del mundo y no sólo observó los valores de este de manera pasiva; por esta razón es muy importante tener en mente que es distinto $P(Y | do(z), X=x)$ que $P(Y | Z=z, X=x)$.
-Las preguntas causales de interés son planteadas de la forma $P(X | do(z))$ y son llamadas \textit{quries intervencionales}.
-Este tipo de intervenciones surgen de manera natural en muchos problemas; por ejemplo, “si el paciente recibe este tratamiento, cuál es la probabilidad de sobrevivir la enfermedad” puede ser planteada como $P(sobreivir | do(tratamiento))$.\\
-\\
-Formalmente, un modelo gráfico causal (\cite{koller2009probabilistic}, \cite{pearl2014causality}) tiene la misma forma que una red Bayesiana; es decir, consiste en una red acíclica dirigida que codifica una distribución entre un conjunto de variables aleatorias.
-La diferencia radica en la interpretación de las flechas, pues cada variable aleatoria está gobernada por un modelo causal en función de sus padres en el grafo.
-Una intervención consiste en forzar uno de los nodos a tomar un valor con probabilidad 1 y a recortar todas las flechas que entran en ese nodo.\\
-\\
-Es importante que se cumpla que si dos variables están correlacionadas, entonces una es padre causal de la otra y que no falten padres causales de ninguna variable.\\
-\\
-Para poder contestar pregunta acerca de intervenciones, se necesita introducir el concepto de \textit{grafo mutilado}
-\begin{defi}
-Sea $\mathcal{B}$ una red sobre un conjunto de variables aleatorias y sea $Z_1=z_1,...,Z_k=z_k$ una instanciación de estas variables denotada por $\mathbf{Z}=\mathbf{z}$ .
-Se define el grafo mutilado $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$ como:
-\begin{itemize}
-\item Cada nodo $Z_i \in \mathbf{Z}$ no tiene padres en $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$
-\item El modelo local de probabilidad condicional para $Z_i \in \mathbf{Z}$ es tal que asignal probabilidad uno al valor $z_i$ y cero a cualquier otro valor.
-\item No se altera ningún otro nodo $X_i \notin \mathbf{Z}$
-\end{itemize}
-\end{defi}
-Entonces, utilizando este concepto, tenemos que para un modelo causal $\mathbf{C}$ una intervención $P_{\mathcal{C}}(Y|do(z)$ es realmente una \textit{querie} probabilista estándar, pero en el grafo mutilado dado por $\mathcal{C}_{Z=z}$; es decir
-\[ P_{\mathcal{C}}(Y|do(z),x) = P_{\mathcal{C}_{Z=z}}(Y|x).  \] 
-\subsection{Identificabilidad de modelos}
-Aun a pesar de desconocer la totalidad de las variables latentes causales, existen casos en los que es posible responder a preguntas causales en modelos que involucran variables latentes utilizando sólo variables observadas; este es el caso de los modelos identificables.
-Existen formas de simplificar queries intervencionales y convertirlas en queries que sólo requieren de datos observacionales pues existen equivalencias entre unas y otras bajo ciertas condiciones.\\
-\\
-\cite{peters2012identifiability} exploran las condiciones para obtener identificabilidad en modelos causales funcionales. 
-\subsection{Aprendizaje de modelos gráficos causales}
-Según \cite{koller2009probabilistic}, el problema de aprender un modelo causal a partir de datos puede dividirse en varios ejes:
-\begin{itemize}
-\item En primer lugar, definir qué se entiende por un modelo causal, pues si estamos trabajando con modelos gráficos causales entonces el problema se convierte en aprender redes bayesianas, pero con la diferencia de la interpretación que se da a las flechas; por otro lado, en modelos causales funcionales se tienen parametrizaciones mucho más complejas para las variables de respuesta, por lo que se necesitan modelos más complejos como el modelo neuronal presentado anteriormente.
-\item Un segundo eje es determinar si ya se conoce la estructura y sólo se necesita aprender la parametrización, o si también debe ser aprendida la parametrización.
-\item En tercer lugar, es el tipo de datos: observados o intervencionales.
-Datos observados son aquellos que provienen de alguna distribución generadora y datos intervencionales es en aquellos en los que se intervino el modelo y se observó la respuesta de este.
-Los datos intervencionales son difíciles de obtener en la práctica, o incluso pudieran ser inmorales o ilegales.
-\item Finalmente, cómo tratar los factores de confusión (confounding factors). Uno puede suponer que simplemente no existen y en estos casos el problema se vuelve relativamente sencillo.
-\end{itemize}
-\subsubsection{Aprender sin factores de confusión}
-El problema de aprender modelos gráficos causales sin factores de confusión es esencialmente el mismo que aprender redes bayesianas (\cite{koller2009probabilistic}), pero hay que hacer algunas suposiciones importantes, pues existen siempre varios DAG's que representen la misma distribución, pero que tendrían distintas interpretaciones causales.
-Una de las suposiciones necesarias para poder aprender modelos causales a través de datos observacionales es la llamada Propiedad Causal de Markov, que es un análogo a la Propiedad de Markov para redes Bayesianas; y su formulación es la siguiente (\cite{koller2009probabilistic}): \textit{cada variable es condicionalmente independiente de sus no-efectos dadas sus causas directas.} El otro supuesto necesario es que las únicas independencias condicionales sean aquellas que surgen de la d-separación en el grafo causal correspondiente (\textit{faithfulness}).\\
-\\
-Aun con estas suposiciones, lo más que puede aprenderse es una clase de equivalencias para el grafo causal y para llevar esto a cabo existen dos familias de métodos: basados en restricciones y basados en score y son los mismos que se utilizan para aprender el CPDAG en modelos causales funcionales.
-\subsubsection{Aprender de datos intervencionales}
-En el caso en que se tienen datos provenientes de intervenciones, lo cual es lo más común en el contexto de experimentos científicos.
-Se utilizan métodos basados en score para aprender de este tipo de datos.
-\subsubsection{Aprender con variables latentes}
-En el caso en el que suponemos que existen variables latentes, estamos pensando que los datos fueron generados por una distribución que es \textit{parecida}, o parcial, a la verdadera.
-De nuevo, existen familias de métodos basadas en restricciones y métodos basados en scores.
-\subsection{Redes Bayesianas causales: aplicación en genética}
-\cite{su2013using} intentan aprender la estructura de las relaciones entre un conjunto de variables relacionadas a un estudio de cáncer (\cite{karagas1998design},\cite{karagas2004incidence}).
-Específicamente, los autores analizan sólo 11 variables relacionadas a un solo gen.
-Identifican relaciones interesantes, como por ejemplo la variable “fumador” como padre causal de la variable “cancer”, aunque no apareció relación entre la variable cancer y un gen que se sabe está relacionado con el cáncer, pero esto se debe a la manera en la que fueron seleccionados los pacientes.
-\section{Modelos causales funcionales y ecuaciones estructurales}
-En otras áreas, como las ingenierías y la física, las relaciones causales se expresan como relaciones funcionales determinísticas en donde cualquier componente aleatorio se introduce en términos de variables no observadas.
-Esto corresponde (\cite{pearl2009causality}) a la noción Laplaciana de lo aleatorio.
-En este contexto, un modelo causal funcional consiste en un conjunto de ecuaciones de la forma
-\[ X_i = f_i (X_{Pa_i},U_i), \textrm{  } i=1,...n\]
-donde $Pa_i$ corresponde al conjunto de variables que directamente determinan a la variable con índice $i$ y donde $U_i$ representa cualquier error de medición o de aspectos no observados.\\
-\\
-Dado un modelo causal de esta forma, podemos obtener un diagrama de la siguiente manera: para cada variable pintamos un nodo, y desde cada elemento de $Pa_i$ dibujamos una flecha hacia el nodo que corresponde a $X_i$.
-A este diagrama se le llama \textit{diagrama causal}.
-Si cumple que el diagrama es acíclico, se le conoce como semi-Markoviano.
-Si, además de ser acíclico, los términos de error son independientes entonces se dice que es Markoviano.
-El siguiente Teorema muestra la conexión entre el diagrama causal y la distribución de probabilidad asociada.
-\begin{teo}
-Todo modelo causal Markoviano induce una distribución $P(X_1,...,X_n)$ que satisface que cada variable $X_i$ es independiente de todos sus no-descendientes dado sus padres $Pa_i$.
-\end{teo}
-Sabemos que dos DAG's que tienen el mismo esqueleto y las mismas v-estructuras son equivalentes, y a la clase de equivalencia le llamaremos CPDAG (\cite{goudet2017learning})
-\subsection{Intervenciones y efectos causales en modelos funcionales}
-Los modelos funcionales causales proveen una manera de estimar cómo cambiaría la distribución en respuesta a cambios externos.
-Lo que se hace en estos modelos es representar las intervenciones como alteraciones a un conjunto de funciones.
-El efecto total de la intervención se calcularía modificando las ecuaciones correspondientes y con este nuevo modelo calcular las nuevas probabilidades.
-Denotamos una intervención como $do(X=x)$.
-Se dice que una variable $X_i$ es causa directa de $X_j$ si
-\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
-\subsection{Contrafactuales en modelos funcionales}
-
-\section{Extracción de relaciones causales a partir de texto}
-\subsection{Extracción basada en patrones gramaticales}
-\cite{khoo2000extracting} desarrollan un método que permite identificar y extrar relaciones causa-efecto que se encuentra expresada de manera explícita en \textit{abstracts} de artículos médicos.
-Los autores deciden utilizar el dominio médico pues al ser importantes las relaciones causales es más probable que estas aparezcan explícitamente.
-Los autores primero identifican los \textit{marcadores lingüísticos} y con esto construyen un conjunto de \textit{patrones causales}.
-Además, cada oración es pasada por un \textit{parse tree} el cual genera una representación de la estructura sintáctica de la oración.\\
-\\
-Como ejemplo, consideremos la oración \textit{“Paclitaxel was well tolerated and resulted in a significant clinical response in this patient”}.
-Después de la creación del árbol sintáctico y del patrón gramatical, se obtiene que la causa presente en la oración es \textit{paclitaxel} y el efecto es \textit{a significant clinical response in this patient}.\\
-\\
-Analizando 200 abstracts que provienen de 4 áreas médicas (depresión, esquizofrenía, enfermedades cardiacas y SIDA), construyen los patrones causales.
-El proceso de extracción de información consiste en hacer un \textit{match} entre los patrones causales y el arbol sintáctico.
-Al aplicar los patrones causales obtenidos a un nuevo conjunto de abstracts, obtienen un 51 por ciento de accuracy al extraer causas y un 58 por ciento de accuracy al extraer efectos.\\
-\\
-\cite{girju2002text} presentan un método semi-automático para la detección de patrones causales en texto, es semi-automático pues los patrones son descubiertos de manera automática, pero la validación no lo es.
-El algoritmo de detección de patrones léxico-sintácticos que  se refieren a causalidad consiste en dos partes.
-La primera parte descubre patrones léxico sintácticos que pueden expresar relaciones causales, y la segunda valida estos patrones según restricciones semánticas.
-Se enfocan en encontrar patrones de la forma (frase verbo causal frase) donde el verbo es un verbo causal simple en los cuales el verbo se refiere únicamente a la causa; por ejemplo, \textit{Temblores generan tsunamis}.
-Como las frases obtenidas pueden ser muy largas, pueden aportar información no deseada.
-Para resolver esto \cite{girju2003automatic} utiliza WordNet para reducir el tamaño de las frases obtenidas.\\
-\\
-\cite{sanchez2004acquiring} desarrollan un sistema para la adquisición de conocimiento causal a partir de texto.
-El sistema propuesto identifica oraciones que especifican de manera explícita relaciones causales y extrae de ellas los patrones causales.
-El sistema incorpora información gramatical tal como conectores, conjunciones, disjunciones y negaciones y codifican todo esto en redes bayesianas.\\
-\\
-Para la detección de patrones causales, los autores se enfocan en trabajar con oraciones en las cuales ambos \textit{eventos} aparezcan como frases (noun phrases, ¿traducción?).
-Un ejemplo de patrones causales identificados por el sistema es \textit{“Anemia are caused by excessive hemolysis”}.
-\\
-Para la construcción de las redes, se toma como input un texto \textit{tokenizado}, etiquetado con \textit{part-of-speech tagging} y \textit{parsed} para encontrar expresiones verbales.
-Posteriormente, el sistema estima las probabilidades condicionales de los efectos dadas las causas mediante el conteo de frecuencias de eventos similares.\\
-\\
-Para la evaluación, hicieron una evaluación subjetiva que consistió en comparar la estructura de la red generada por el sistema y la estructura creada manualmente al observar los patrones causales en el texto.
-Se analizaron 7 textos y en general se encontró un empate entre las redes con una precisión de 60 por ciento, pero sin ningún valor de Recall, por lo que no es fácil interpretar sus resultados.
-Según \cite{bojduj2009extraction}, una de las razones por las cuales se ubtuvo baja precisión es que basar los nodos en las frecuencias no es una buena idea, pues la cantidad de texto disponible en general es tan grande que estas frecuencias no pueden ser representativas.\\
-\\
-\cite{bojduj2009extraction} menciona que muchos de estos métodos están más enfocados en cómo extraer las relaciones que en cómo usarlas.
-Es importante no sólo poder concluir que dos términos están relacionados causamente, sino cómo.
-\subsection{Extracción basada en patrones de co-ocurrencias}
-En vez de utilizar patrones gramaticales, \cite{saito2007extraction} y \cite{perrin2008global} analizan la co-ocurrencia de términos estadísticos para construir redes a partir de texto.
-Ellos definen un “statistical term” como \textit{“Statistical terms are expression of the measurements of statistics to watch the movements of phenomena; birth rates, public approval rating of the Cabinet and so on.”}; es decir, un término que sea importante a un sistema.
-El enfoque de su trabajo es construir redes causales entre los statistical terms.
-Lo que hacen es primero extraer relaciones de co-ocurrencia entre términos estadísticos y posteriormente extraer relaciones causales entre ellos.
-Los autores se enfocan en texto japonés, el cual cumple que tiene ciertos patrones en los sufijos utilizados, y estos patrones son utilizados al extraer los términos estadísticos.
-Luegom estos términos son clasificados según su nivel de abstracción.
-Posteriormente, construyen una red de términos estadísticos al considerar a dos de ellos co-ocurrentes si aparecen juntos en un párrafo.
-Se remueven los términos que son co-ocurrentes con muchos otros términos.\\
-\\
-Los autores evaluan de manera separada cada una de las etapas: extracción de términos, clasificación de términos, la construcción de redes y la clasificación de redes.
-\\
-La mayor desventaja de esto es que la dirección de la relación causal no puede ser inferida de la co-ocurrencia.\\
-\\
-Por otro lado, \cite{sakai2008cause} extrajeron términos de artículos financieros (en japonés) y luego considerarón que estos términos afectaban el desempeño de los negocios basados en co-ocurrencias con expresiones clave como “es bueno”.
-Su definición de causalidad depende de las expresiones clave.
-\subsection{Modelos Gráficos para extracción de relaciones}
-A partir de un artículo que trata específicamente sobre una entidad claramente indentificable (la biografía de Benito Juarez por ejemplo) \cite{yu2010jointly} intentan identificar entidades y descubrir relaciones semánticas entre ellas.
-Resulta evidente que en el documento aparecerán otras entidades, pero se hace el supuesto de que no están relacionadas entre sí.
-En este trabajo, todas las clases entidades están pre-definidas así como las relaciones; esto es, se sabe que en el documento aparecerán años, lugares, países y que una entidad de tipo persona habrá nacido en un país en cierto año.
-Para formular el problema, se tiene $\mathbf{x} = \{ x_1,...,x_N \}$ una sucesión observada de \textit{tokens}.
-Denotamos como $s_p$ la entidad principal.
-Sea $ \mathbf{s}=\{ s_1,...,s_L\}$ una segmentación en la cual cada $s_i$ es una tripleta $\{ \alpha_i, \beta_i,y_i \}$ con $\alpha_i$ es una posición inicial, $\beta_i$ posición final y $y_i$ una etiqueta asignada.\\
-\\
-Sea $r_{pn}$ la asignación de relaciones entre la principal $s_p$ y un candidato a entidad secundaria $s_n$ y sea $\mathbf{r}$ el conjunto de asignación de relaciones para la $\mathbf{x}$.
-Dada una observación $\mathbf{x}$, queremos encontrar $\mathbf{y}^\ast$ tal que
-\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
-donde $\mathbf{y}=\{ \mathbf{r},\mathbf{s} \}$.\\
-\\
-Se define una distribución condicional conjunta para $\mathbf{s},\mathbf{x},\mathbf{r}$ en un modelo grafico no dirigido $\mathcal{G}$ tal que se pueden dividir los factores de $\mathcal{G}$ en tres grupos: $\phi^S$, $\phi^R$, $\phi^\nabla$, que son el potencial de segmentación, el potencial de relaciones, y el potencial conjunto de segmentación-relación.
-\\
-\\
-Por el Teorema de Hammersley-Clifford, la condicional conjunta $P(\mathbf{y} | \mathbf{x})$ se factoriza como
-\[ P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \prod_{C_S} \phi^S (i,\mathbf{s},\mathbf{x}) \prod_{C_R}\phi^R(r_{pm},r_{pn},\mathbf{r}) \prod_{C_\nabla} \phi^\nabla(s_p,s_j,\mathbf{r}) \]
-Se hace el supuesto de que las funciones potencial se factorizan según un conjunto de atributos y un conjunto de pesos correspondientes.\\
-\\
-Incorporando estos features, tenemos que 
-\[P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \exp \left( \sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r) \right) \]
-\\
-Dada una muestra $\mathcal{D}= (\mathbf{x}_i, \mathbf{y}_i)_{i=1}^N$ queremos estimar los parámetros $(\lambda_k,\mu_w,\nu_t)$.\\
-\\
-La log-verosimilitud regularizada es
-\[ \mathcal{L} = \log [\Phi(r,s,x)] - \log [Z(x)] - \sum_{k=1}^K \frac{\lambda_k^2}{2 \sigma_\lambda^2} - \sum_{w=1}^W \frac{\mu_w^2}{2 \sigma_w^2} - \sum_{t=1}^T \frac{\nu_t^2}{2 \sigma_\nu^2} \]
-Donde,
-\[ \Phi(r,s,x) = \exp (\sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r)) \]
-Tenemos que $\mathcal{L}$ es concava por lo que es fácil maximizarla.\\
-\\
-Para calcular la Asignacion Más Probable, hayue encontrar
-\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
-La cual no se puede resolver de manera exacta, se necesitan métodos aproximados.
-El algoritmo que se propone en el artículo se llama Collective Iterative Classification y la idea detrás de este es decodificar las variables ocultas objetivo basados en asignar etiquetas a las variables muestreadas.
-Esto se hace mediante un proceso iterativo de dos pasos; primero, en una etapa de bootstrapping se predice una etiqueta inicial para una $x_i$ dado el modelo ya entrenado.
-Luego, en la segunda etapa, conocida como clasificación iterativa, se re-estima la asignación a $x_i$ varias veces, tomándolas a partir de la asignación inicial.\\
-\\
-Los autores prueban su algoritmo con datos provenientes de Wikipedia: sus datos consisten en 441 páginas de Wikipedia.
-De estos datos, fueron etiquetados a mano 7740 entidades en 8 categorías.
-Además, se extrajeron 4700 relaciones de las cuales las más frecuentes fueron trabajo, visió, nació, miembro de, nació en día.\\
-\\
-El modelo superó los existentes tomando como medida de calidad la medida F. 
-\section{Métodos Estadísticos Para Extracción de Relaciones Causales}
-\subsection{Statistical Relational Learning}
-En el contexto de inferencia causal, se estudian las interacciones entre variables que puedan causar efectos en la otra.
-Para esto, es necesario identificar los objetos y las relaciones entre ellos.
-Para esto, el área de Statistical Relational Learning ofrece una serie de métodos que modelan dependencias entre entidades relacionales.
-En esta sección veremos distintas técnicas existentes en esta área y sus aplicaciones a inferencia causal.\\
-\\
-Un dominio de interés consiste en objetos (entidades), sus atributos, y las relaciones entre sí de los objetos.
-El modelo DAPER de \cite{heckerman2004probabilistic} consiste en clases de entidades, clases de relaciones, clases de atributos y clases arco.
-\subsubsection{Aplicación de Campos Aleatorios Condicionales a Extraccion de informacion}
-\cite{ramani2005consolidating} analizaron 753,459 abstracts de journals médicos, de los cuales extrajeron 6580 interacciones entre 3,737 proteínas.
-Para ello, utilizaron un algoritmo de tres etapas: la primera etapa, consiste en identificar nombres de proteínas mediante un clasificador basado en Conditional Random Fields; la segunda, identificar interacciones a través de co-ocurrencias; posteriormente, filtrar estas interacciones con un clasificador Bayesiano para obtener interacciones válidas.
-Con esto, se obtiene una red que consiste en 31609 interacciones entre 7748 proteínas\\
-\\
-Para la etapa de identificación de nombres, se utilizó un Conditional Random field, cuyo desempeño fue evaluado en un conjunto de 200 abstracts médicos etiquetados y en 750 abstracts etiquetados a mano.
-Para extraer los nombres de proteínas, se utilizó este clasificador en el conjunto de 753,459 abstracts de Medline que contuvieran la palabra “human”.\\
-\\
-Para obtener las interacciones, se midió la co-citación entre los nombres de proteínas y luego se enriquecieron estos pares con interacciones físicas (reales) entre proteínas utilizando un filtro Bayesiano.
-Primero, se contó el número de veces que en un abstract aparece un par de proteínas, y con eso se calculó la probabilidad de co-citación.
-Con esto se obtienen 15,000 interacciones, pero se pueden dar co-citaciones por otras razones además de las relaciones reales, se aplicó un filtro bayesiano que mide la verosimilitud de los abstracts que citan un par de proteínas para discutir sus interacciones reales.
-El clasificador asigna un \textit{score} a cada uno de los abstracts que citan esas proteínas.
-Utilizando un umbral de score, se obtienen 6,580 interacciones entre 3,737 proteínas.\\
-\\
-Al combinar estas interacciones con otras 26,280 interacciones provenientes de otras fuentes, se obtiene el total mencionado.\\
-\\
-Por otro lado, \cite{culotta2006integrating} plantean el problema de extracción de relaciones como un problema de etiquetado de secuencias, algo similar a lo realizado por \cite{yu2010jointly}.
-\cite{culotta2006integrating} definen un modelo de la forma
-\[ P_{\Lambda} ( y | x ) =\frac{1}{Z} \prod_{c \in C} \varphi_c ( y_c , x_c, \Lambda ),\]
-donde las $\varphi$ son potenciales parametrizados por $\Lambda$.\\
-\\
-Los autores tomaron 1,127 párrafos de 271 artículos biográficos de Wikipedia y etiquetaron 4,701 instancias de relaciones.
-Para identificar a las entidades importantes dentro de un artículo, utilizaron el hecho de que estas suelen tener un \textit{hyperlink}
-\subsubsection{Aplicación de RMN's a Extraccion de informacion}
-\cite{bunescu2007statistical} utilizan Relational Markov Networks para mejorar el desempeño de los CRF's en el problema de identificar nombres de proteínas en abstracts de journals médicos. \\
-\\
-La técnica que se propone es que dada una colección de documentos $D$, asociemos a dada documento $d \in D$ un conjunto de entidades $d.E$.
-Cada entidad$e \in d.E$ queda caracterizada por un conjunto de atributos booleanos $e.F$.
-A cada documento se le asocia un grafo de factores, el cual es un grafo bipartita que contiene dos tipos de nodos: nodos variables, que corresponden a las etiquetas de todas las entidades candidato en el documento y nodos potenciales, que modelan las correlaciones entre dos o más atributos.
-Para cada correlación posible,  un nodo potencial es crreado tal que está ligado a todos los nodos variables involucrados.\\
-\\
-Prueban su algoritmo en dos conjuntos de datos previamente etiquetados a mano.
-El primer conjunto consiste en 200 abstracts y el segundo en 225.
-Ambos conjuntos de datos fueron previamente etiquetados a mano, además de etiquetados para POS.
-En ambos conjuntos de datos, existe una mejora ligera respecto a los CRF's en términos de la medida F.
-
-\subsubsection{Aprendizaje de Modelos Causales a partir de Datos Relacionales}
-\cite{maier2013reasoning} definen la \textit{abstract ground graph} (AGG) y se desarrolla la noción de d-separación relacional.
-A partir de estas nociones teóricas, \cite{maier2013sound} extienden el algoritmo PC, el cual identifica todas las posibles orientaciones de dependencias causales, a datos relacionales.
-Utilizando la d-separación relacional, introducen restricciones que orienta dependencias bivariadas hasta en un 72$\%$ más.
-Se prueba teóricamente que con esta nueva regla, además de extensiones a datos relacionales del algoritmo PC, se obtiene un método \textit{sound and complete} para la extracción de relaciones causales.
-Este algoritmo nuevo se llama RCD (relational causal discovery).\\
-\\
-\cite{lee2016learning} notan que la demostración de \cite{maier2013reasoning} requiere que la AGG sea un DAG quie represente exactamente todas las flechas que aaparecerían en todas las posibles ground graphs, y en un trabajo anterior (\cite{lee2015lifted}) habían notado que existen casos en los cuales la d-separación de la AGG no captura las independencias condicionales que se cumplen en el modelo relacional causal.
-En su artículo, proponen un algoritmo basado en condiciones más débiles llamado RCD-light.
-
-\subsection{Inferencia Bayesiana No Paramétrica}
-La rama no-paramétrica de la inferencia bayesiana puede considerarse como una extensión a espacios infinito-dimensionales de la inferencia bayesiana clásica.
-Esto permite tener modelos que permiten capturar una complejidad mucho mayor en los datos.
-El objeto principal de la inferencia bayesiana no paramétrica es el Proceso Dirichlet, que es una extensión a espacios infinito-dimensionales de la distribución Dirichlet clásica.
-\subsubsection{Proceso Dirichlet}
-\begin{defi}
-Sea $H$ una medida finita definida sobre un espacio $\mathcal{X}$ y $\alpha \in \mathbb{R}$.
-Una medida aleatoria $G$ sobre $\mathcal{X}$ se dice que es un Proceso Dirichlet de parámetros $(\alpha,H)$ si para cada partición finita medible $\{ A_1 , ..., A_k \}$ de $\mathcal{X}$ se tiene que la distribución conjunta de $(G(A_1),...,G(A_k))$ es Dirichlet de parámetros $(\alpha H(A_1),...,\alpha H(B_k))$.
-A la medida $H$ se le conoce como medida base y a $\alpha$ se le conoce como parámetro de concentración.
-Se denota como
-\[ G \sim \textrm{ DP}(\alpha,H).\]
-\end{defi}
-Los parámetors $H$ y $\alpha$ juegan roles intuitivos en la definición.
-La medida base se puede entender, básicamente, como la \textit{media} del Proceso; es decir, para cada conjunto medible $A \subseteq \mathcal{X}$ se tiene que $\mathbb{E}[G(A)]=H(A)$.
-Por otro lado, el parámetro de concentración es un tipo de \textit{varianza inversa}, pues $V[G(A)]=H(A)(1-H(A))/(\alpha+1)$.
-Es decir, mientras más grande sea $\alpha$, entonces menor es la varianza y el DP tenderá a concentrarse más en torno a $H$.\\
-\subsubsection{Distribución Predictiva y urnas de Blackwell-McQueen}
-Consideremos $G \sim \textrm{ DP}(\alpha,H)$ y $\theta_1,... \sim G$. Como $\theta_{n+1} | G, \theta_1,...,\theta_n \sim G$, entonces para $A \subseteq \mathcal{X}$ tenemos que
-\begin{eqnarray*}
-P(\theta_{n+1} \in A | \theta_1,...,\theta_n ) &=& \mathbb{E}[G(A) | \theta_1,...,\theta_n]\\
-                                                                    &=& \frac{1}{\alpha + n} \left(\alpha H(A) + \sum_{i=1}^n \delta_{\theta_i}(A) \right)
-\end{eqnarray*}
-Por lo tanto, marginalizando $G$, vemos que la distribución posterior base dados $\theta_1,...,\theta_n$ es la distribución predictiva de $\theta_{n+1}$.
-\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum^n \delta_{\theta_i} \right) \]
-La secuencia de distribuciones predictivas para $\theta_1, \theta_2,...$ se conoce como el esquema de urnas de Blackwell-McQueen; la interpretación es la siguiente: cada valor en $\mathcal{X}$ es un color, y cada realización $\theta \sim G$ son pelotas y el valor obtenido es el color de la pelota.
-Además, tenemos una urna en la cual se han colocado las bolas observadas.
-Al principio, no existen bolas en la urna y muestreamos un color de $H$; es decir, $\theta_1 \sim H$, pintamos la bola con ese color y la echamos en la urna.
-Luego, en el paso $n+1$, tomaremos un nuevo color con probabilidad $\alpha / (\alpha + n)$ o tomaremos un nuevo color, la pintaremos así y la echaremos en la urna con probabilidad $n / (n+\alpha)$.
-\subsubsection{Proceso Dirichlet como Prior en problemas de Clustering}
-La propiedad de que $G \sim DP$ sea discreto ofrece aplicaciones de clusterización.
-Por el momento, supongamos que $H$ es suave.
-Entonces, como los valores de las realizaciones son repetidos, sean $\theta^\ast_1,...,\theta^\ast_m$ los valores únicos de las realizaciones $\theta_1,...,\theta_n$ y sea $n_k$ el número de repeticiones de $\theta^\ast_k$.
-Entonces, la distribución predictiva se puede escribir como
-\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum_{k=1}^m n_k \delta_{\theta^\ast_k} \right). \]
-Esto es, el valor de $\theta^\ast_k$ aparecerá en $\theta_{n+1}$ con una probabilidad proporcional al número de veces que ha aparecido.
-Entonces, mientras más grade sea $n_k$, mayor será la probabilidad de que se vuelva aun mayor.
-Notemos que los valores únicos de las realizaciones $\theta_1,...,\theta_n$ inducen una partición aleatoria del conjunto $[n]=\{1,...,n\}$ si pensamos el cluster $k$ como aquel que contiene las $\theta_i$ que toman el valor $\theta^\ast_k$, tenemos que clusters grandes crecen más rápido.\\
-\\
-Podemos invertir este proceso y empezar con una distribución sobre particiones.
-\subsubsection{Dirichlet Enhanced Relational Learning}
-\cite{xu2005dirichlet} aplican un modelo jerárquico Bayesiano no-paramétrico para la extracción de relaciones.
-La ventaja de usar un modelo jerárquico es que los parámetros pueden estar \textit{personalizados} para entidades.
-En este trabajo se aplica el modelo al contexto médico, en el cual se tienen como entidades hospitales, pacientes, diagnósticos y procedimientos.
-Se sabe que la existencia de un diagnóstico o un procedimiento es dependiente de las características de un paciente; por otro lado, las características de los hospitales se modelan como inciertas.\\
-\\
-Un modelo Bayesiano consiste en un parámetro $\theta$ y una asignación inicial de probabilidad sobre este, la cual refleja nuestra incertidumbre inicial.
-Supondremos que esta distribución inicial cuenta con hiperparámetros $h$, de modo que sea $P(\theta | h)$.
-La verosimilitud de los datos $P(D|\theta)$ refleja cómo son generados estos dado el parámetro.
-EN el caso en el cual se asume que los datos provienen de distintos grupos es natural pensar en asignar un parámetro por grupo, lo cual lleva a la definición de modelos jerárquicos, en los cuales el modelo tiene la siguiente fforma
-\[ P(h) \prod P(\theta_i | h )P(D_i | theta_i). \]
-Si se aplicara el modelo DAPER al contexto médico, se tiene que los parámetros e hiperparámetros que especifican distribuciones condicionales como atributos globales.
-Esto implica que las probabilidades de un procedimiento son idénticas para todos los pacientes con la misma dolencia inicial.
-Luego, los procedimientos son modelados de manera independiente de manera que procedimientos anteriores en una persona no afectlan la selección de uno nuevo.
-Estas son suposiciones no realistas.\\
-\\
-Para mejorar esta situación, se asume que la probabilidad de un procedimiento es una distribución multinomial con un prior Dirichlet.
-En particular, para cada paciente, se asume que un vector individual de parámetros $\theta_{s|pc,pa}$ especifica la probabilidad de un procedimiento para un paciente $pa$ con dolencia inicial $pc$.
-Este vector es generado por una distribución Dirichlet de parámetros $h_{pc}=\{ \tau_{pc},\alpha_{pc} \}$, la cual puede escribirse como
-\[ \textrm{Dir}(\theta_{\cdot | pc, pa} | \tau_{pc},\alpha_{pc}) = \frac{1}{C} \prod_{k=1}^K \theta^{\tau_{pc} \alpha_{k,pc} - 1}_{k,s | pc, pa}.  \]
-En un modelo jerárquico bayesiano cada paciente obtiene sus propias probabilidades de procedimiento y comparte con los demás un parámetro a priori.\\
-\\
-En más de un caso, se necesita asumir que la prior tiene una forma muy flexible, por lo que es necesario recurrir a priors no paramétricas.\\
-\\
-Podemos tomar la distribución a priori como una realización del Proceso Dirichlet; es decir, 
-\[ G_{pc} \sim \textrm{ DP }(G_0,\alpha_0) \]
-donde $G_0$ es la medida base y $\alpha_0$ el parámetro de concentración, el cual especifica nuestro nivel de certidumbre sobre la prior.\\
-\\
-Luego, el parámetro $\theta_{\cdot | pc, pa}$ es una realización de $G_{pc}$.\\
-\\
-Otra manera de escribir el Proceso Dirichlet es mediante su representación \textit{stick breaking} (\cite{teh2011dirichlet}) según la cual
-\[ G_{pc} = \sum \pi_{l,pc} \delta_{\theta^\ast_{l,pc}} \textrm{ ;  } \theta^\ast_{l,pc} \sim G_0  \]
-\[ \pi'_{l,pc} \sim \textrm{ Beta }(1,\alpha_0) \textrm{  ;  } \pi_{l,pc} = \pi'_{l,pc} \prod_{k=1}^{l-1} (1 - \pi'_{k,pc}).\]
-Notemos que aunque la distribución base pueda ser continua, el DP es siempre discreto (con probabilidad 1).\\
-\\
-Tradicionalmente, el aprendizaje en el contexto Bayesiano no paramétrico se lleva a cabo via muestreo de Gibbs, cuyas variaciones más comunes son la urna de Polya o el Proceso del Restaurante Chino (\cite{teh2011dirichlet}).
-En el artículo, se lleva a cabo un proceso más eficiente (\cite{Yu:2004:NHB:1008992.1009053}.)\\
-\\
-El objetivo es estimar $G_{pc}$ para cada posible $pc$ en la base de datos, utilizando la verosimilitud marginal:
-\[ \hat{G}_{pc} = \textrm{ argmax DP }(G | G_0, \alpha_0) \prod_{\{pa\}_{pc}} \int \textrm{ Multinomial }(\{pr\}_{pa}| \theta_{\cdot | pc,pa}) d\theta_{\cdot | pc,pa}  \]
-donde $\{pa\}_{pc}$ es el conjunto de pacientes tales que tienen la misma dolencia inicial y $\{pr\}_{pa}$ es el conjunto de procedimientos para el paciente $pa$.
-Resulta imposible calcular la verosimilitud, por lo que se realiza una aproximación de campo medio:
-\[ G_{pc} \approx \sum^{|{pa}_{pc}|} \pi_{pa,pc} \delta_{\theta^\ast_{\cdot | pa,pc}}. \]
-El proceso de aprendizaje se divide en dos etapas: 
-\begin{itemize}
-\item Calcular la ubicación de $ \theta^\ast_{\cdot | pa,pc}$
-\item Estimar los pesos $\pi_{pa,pc}$
-\end{itemize}
-Para el primer paso, se utiliza la aproximación MAP de $\theta_{\cdot | pa,pc}$ que es:
-\[ \theta^{MAP}_{\cdot | pa,pc} = \textrm{ arg max }P( \theta_{\cdot | pa,pc} | \{pr\}_{pa})\]
-Y se utiliza como prior la Dirichlet mencionada antes. \\
-\\
-Para el segundo paso, es necesario hacer la siguiente suposición:
-\[ \hat{P}( \theta_{\cdot | pa,pc} | \{pr\}_{pa}) \approx q_{pa}(\theta_{\cdot | pc,pa}) = \sum_{\{\tilde{pa}\}_{pc}} \xi_{\tilde{pa},pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}} \]
-donde $\xi_{\tilde{pa},pc,pa} \geq 0 $ son parámetros variacionales tales que $\sum_{pa} \xi_{\tilde{pa},pc,pa} = 1$.\\
-\\
-Como paso E se obtiene que:
-\[  \xi^t_{\tilde{pa},pc,pa} = \frac{P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}}) \hat{G}^{(t)}_{pc}(\delta_{\theta^{MAP}_{\cdot | pc,pa}}))}{\sum_{\tilde{pa}}P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}})\hat{G}^{(t)}_{pc}(\theta^{MAP}_{\cdot | pc,pa}))}.
-\]
-Y el paso M queda como
-\[ \hat{G}^{(t+1)}_{pc}(\theta_{\cdot,pc,pa}) = \frac{\alpha_0 G_0(\theta_{\cdot,pc,pa}) + \sum_{\{pa\}_{pc}} \xi_{pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}}}{\alpha_0 + |\{pa\}_{pc}|} \]
-
-\subsubsection{Learning Systems of Concepts with an Infinite Relational Model}
-Suponiendo que tenemos una o más relaciones que involucran uno o más tipos, \cite{kemp2006learning} definen el Infinite Relational Model, el cual busca partir cada tipo (type) en clusters, de manera que una buena partición permita que relaciones entre entidades se puedan predecir a partir de la asignación de clusters.
 Por ejemplo, si tenemos $m$ relaciones que involucran $n$ tipos, denotamos $R^i$ la $i$-ésima relación, $T^j$ la $j$-ésima relación y $z^j$ un vector de asignación de clusters.
@@
 \[ P(R^1,...,R^m, z^1,...,z^n) = \prod_{i=1}^m P(R^i | z^1 , ... , z^n) \prod_{j=1}^n P(z^j). \]
-Estamos asumiendo que las relaciones son condicionalmente independientes dada la asignación de clusterss, y que la asignación de clusters es independiente entre tipos.
+Estamos asumiendo qye las relaciones son condicionalmente independientes dada la asignación de clusterss, y que la asignación de clusters es independiente entre tipos.
+\section{Generación de clusters}
 Para que el IRM sea capaz de descubrir el número de clusters en el tipo $T$, se utiliza un prior que asigna probabilidad a todas las posibles particiones de $T$.
@@
 \\
-La intuición detrás de este proceso, según describe \cite{teh2011dirichlet} es que en un restaurante chino hay un número infinito de mesas, cada una de las cuales puede sentar un número infinito de clientes.
+La idea es (\cite{teh2011dirichlet}) que en un restaurante chino hay un número infinito de mesas, cada una de las cuales puede sentar un número infinito de clientes.
 El primer cliente entra y se sienta en la primera mesa, el segundo cliente entra y decide si sentarse en la primera mesa o en otra; en general, el $n$-ésimo cliente se sienta en una mesa ocupada con una probabilidad proporcional al número de clientes sentados ahí.\\
@@
 La distribución en $z$ inducida por el CRP es intercambiable; es decir, no importa el orden en que los objetos llegan a los clusters.
-Como nuevos objetos siempre pueden ser asignados, el IRM efectivamente tiene acceso a un número infinito de clusters, de ahí el nombre del Infinite Relational Model.\\
-\\
-Para generar Relaciones a partir de las asignaciones de clusters se asume que las relaciones son binarias, pero se puede extender a datos continuos y frecuencias.
+Como nuevos objetos siempre pueden ser asignados, el IRM efectivamente tiene acceso a un número infinito de clusters, de ahí el nombre del Infinite Relational Model.
+\section{Generación de relaciones a partir de clusters}
+Se asume que las relaciones son binarias, pero se puede extender a datos continuos y frecuencias.
 Como ejemplo, veremos primero el caso con un solo tipo $T$ y una relación $R: T \times T \to \{0,1 \}$.
@@
 \\
-Estamos asumiendo que la tendencia de una entidad a participar en relaciones está dada por su asignación de cluster.\\
-\\
-Para hacer inferencia, notamos que en el ejemplo anterior se utilizó una prior conjugada, es sencillo calcular $P(R|z)$
+Estamos asumiendo que la tendencia de una entidad a participar en relaciones está dada por su asignación de cluster.
+\section{Inferencia}
+Como en el ejemplo anterior se utilizó una prior conjugada, es sencillo calcular $P(R|z)$
 \[ P(R|z)= \prod_{a,b} \frac{Beta(m(a,b) + \beta, \bar{m}(a,b)+\beta}{Beta(\beta,\beta)} \]
@@
 \\
-Para muestrear de la posterior $P(z | R ) \propto P(R|z)P(z)$ se pueden utilizar técnicas MCMC o al buscar la moda de la distribución.\\
-\\
-Para la evaluación del modelo, los autores generan datos sintéticos.
-Consideran conjuntos de datos de tres formas distintas; el primer sistema, $S1$, tiene sólo dos tipos $T1, T2$ y una sola relación $R:T1 \times T2 \to \{0,1 \}$.
-El sistema $S2$ consiste en 4 tipos y tres relaciones binarias mientras que $S3$ consiste de 3 tipos y una sola relación ternaria.
-Resulta que en este contexto, el modelo captura el número de clusters de manera precisa.
-\\
-\\
-Este método es excelente para clusterizar de manera no paramétrica un conjunto de datos relacionales, pero requiere identificar previamente las entidades y las relaciones. 
-\subsubsection{Learning Infinite Hidden Relational Models}
-\cite{xu2006learning} extienden la expresividad de los modelos relacionales al introducir para cada entidad una variable latente con un número infinito de estados que proviene de un Dirichlet Process Mixture.
-Debido a que el número de features de las cuales un atributo puede depender, resulta más conveniente  introducir para cada entidad una variable latente que es padre de los atributos de la entidad y de los atributos de las relaciones en las que participa la entidad; a esto se le conoce como hidden relationship model.\\
-\\
-Como cada entidad puede tener un distinto número de estados en sus variables latentes, es deseable que este número de estados latentes se determine de manera automática según los datos; esto es posible utilizando una mezcla de Proceso Dirichlet, que son modelos de mezcla con un número infinito de componentes de mezcla, pero que según los datos la complejidad se ajusta de manera automática.
-A la combinación de un hidden relational model con un DPm se le conoce como Infinite Hidden Relationship Model que generaliza el modelado jerárquico bayesiano noparamétrico al caso de modelos relacionales.\\
-\\
-Como aplicación, los autores evalúan su modelo en un conjunto de datos sobre películas y buscan predecir las preferencias de los usuarios.
-Para este conjuto de datos, las clases de entidades son Usuario y Película
-\subsection{Modelos Jerárquicos Bayesianos}
-Los Modelos Jerárquicos Bayesianos (HBM) son modelos probabilistas que están definidos por \textit{niveles}, en los cuales cada nivel representa un mayor grado de abstracción.
-Un ejemplo sencillo de esto puede ser un modelo de regresión de ingreso a nivel nacional en el cual existan parámetros por estado y por municipio.\\
-\\
-En el contexto causal, lo que está detrás de un HBM es que al aprender causalidad se incorpora conocimiento teórico previo y datos observables (\cite{griffiths2005structure}, \cite{griffiths2009theory}).
-Esto es, un HBM es una distribución de probabilidad entre supuestos teóricos, modelos causales y datos (\cite{hagmayer2013hierarchical}).\\
-\\
-\cite{lucas2010learning} proponen un framework que incorpora inferencia basada en covarianzas así como conocimiento adquirido.
-Se enfocan en aprender  \textit{forma funcional} de las relaciones causales.
-Las contribución principal del artículo es: se muestra que el problema de aprender la forma funcional de una relación causal puede ser formalizada utilizando modelos jerárquicos bayesianos; además, se muestran una serie de experimentos que ponen a prueba las predicciones cualitativas hechas por este modelo.\\
-\\
-\cite{hagmayer2013hierarchical} presentan un overview general sobre los modelos jerárquicos bayesianos en el contexto causal.
-En el artículo exponen las ventajas de utilizar estos modelos para realizar inferencia causal, así como sus limitaciones.
-Una limitación mencionada es que dependen considerablemente de conocimiento previo.
-Otra, que existe evidencia que la representación que las personas tienen de modeos causales no corresponde a modelos gráficos causales (redes bayesianas) pues existen experimentos en los cuales la propiedad Markoviana no se cumple.
-Finalmente, argumentan que los HBM's y las redes bayesianas no modelan los procesos cognitivos subyacentes al apredizaje causal, aunque existen evidencias de que las personas sí llevan a cabo un tipo de actualización Bayesiana
-
-\section{Aprendizaje por Refuerzo}
-El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno.
-Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
-El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente.
-A su vez, el ambiente devuelve información al agente en términos de una recompensa o ganancia según el estado en el que se encuentre.
-El objetivo del agente es encontrar una serie de acciones que lo lleven a lograr su objetivo final.
-La función de recompensa codifica qué es lo que se desea que el agente aprenda, pues esta función devuelve un valor para cada estado y con esto el agente asignará un valor a cada estado según la ganancia a largo plazo que un este representa.\\
-\\
-La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
-El objetivo del agente es encontrar la política que maximice la recompensa.\\
-\\
-La regla general que se sigue para la definición del ambiente (\cite{sutton1998reinforcement}) es considerar que todo aquello que no pueda ser explícitamente alterado por el agente pertenece al ambiente.
-La frontera agente-ambiente tiene que ver con los límites de lo que el agente puede controlar, no sobre lo que puede \textit{conocer}.
-Veremos ahora con detalle las partes que conforman esta dinámica. 
-\subsubsection{Objetivos y Recompensas}
-Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada.
-Por ejemplo, un agente que queremos que aprenda a jugar ajedrez debe ser premiado por ganar los juegos, no por lograr sub-objetivos como comer más piezas del oponente o tomar control del centro del tablero (\cite{sutton1998reinforcement}), pues podría suceder que el agente encuentre cómo lograr esto sin ganar el juego necesariamente.
-Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
-\subsubsection{Definición de Cadenas de Markov}
-Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
-\[P(X_{t+1} \in A | X_0=x_0, X_1=x_1 , ... , X_t=x_t) = P(X_{t+1} \in A | X_t = x_t).\]
-Estos procesos quedan caracterizados por sus probabilidades de transición $p^t_{ij}=P(X_{t+1}=j | X_{t}=i)$.
-Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
-\subsubsection{Procesos de Decisión Markovianos con recompensa}
-Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$.
-En respuesta a la acción tomada, se recibe una recompensa $r_t = f(a_t)$ para alguna función $f$ desconocida, la cual puede ser aleatoria o determinista.
-Los procesos de decision Markovianos con recompensa (\cite{Puterman:1994:MDP:528623}) están definidos por un espacio de estados $\mathcal{S}$, un espacio de acciones posibles $\mathcal{A}$, una función de transición $f$ y la función recompensa $r$.
-La función transición puede ser estocástica o determinística.
-
-\subsubsection{Función valor}
-Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado.
-Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política.
-Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$.
-Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$.
-La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante.
-Formalmente, para un proceso de decisión Markoviano, definimos la \textit{función valor} (value function en la literatura) para un estado $s$, y relativa a una política $\pi$, como el retorno esperado si se comienza en $s$ y a partir de ahí se sigue la política $\pi$
-\[ V^{\pi}(s) = \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]. \]
-De manera similar, definimos la \textit{función acción-valor} $Q^{\pi}(s,a)$ es una valoración de tomar la acción $a$ estandoen el estado $s$, y después seguir la \textit{política} $\pi$
-\[ Q^\pi(s,a) = \mathbb{E}[R_t | s_t =s, a_t = a]. \]
-Por la ley de los grandes números, podemos estimar ambas funciones $V^{\pi}$ y $Q^\pi$ si se promedian los retornos recibidos al llegar a cada estado con el número de veces que se ha llegado al estado.
-Esta es la idea detrás de los llamados Métodos Monte-Carlo que serán expuestos en una sección posterior.\\
-Notemos una propiedad interesante de la función $V^{\pi}$:
-\begin{eqnarray*}
-V^{\pi}(s) &=& \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ \sum_{k=0}^\infty \gamma^k r_{t+k+1} | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ r_{t+1} + \gamma \sum_{k=0}^\infty \gamma^k r_{t+k+2} | s_t = s \right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} P(s_{t+1}=s' | s_t = s, a_t = a) \left[ \mathbb{E}[r_{t+1}|s_t=s,a_t=a,s_{t+1}=s'] + \gamma \mathbb{E} \left[ \sum \gamma^k r_{t+k+2} | s_{t+1}=s' \right]\right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} \mathcal{P}_{ss'}^a \left[ \mathcal{R}_{ss'}^a + \gamma V^{\pi}(s') \right]
-\end{eqnarray*}
-A esta última ecuación se le conoce como la Ecuación de Bellman para $V^\pi$.\\
-\\
-También para la función $Q^{\pi}(s,a)$ se tiene una ecuación de Bellman:
-\[ Q^{\pi}(s,a)=\mathbb{E}_{s',a'}[r+ \gamma Q^{\pi}(s',a') | s',a']. \]
-Además, la función $V^\pi$ puede ser obtenida a partir de la función $Q^{\pi}(s,a)$ de la siguiente manera:
-\[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
-\subsubsection{Soluciones y Optimalidad}
-Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa.
-Resulta que es posible definir un orden parcial entre políticas en términos de la función valor, pues $\pi \geq \pi'$ si $V^{\pi}(s) \geq V^{\pi'}(s)$ para todo estado $s$.
-Para procesos de decisión Markovianos finitos existe siempre una política mejor que las demas, a la cual llamaremos \textit{política óptima}.
-De hecho, pueden existir varias, pero todas ellas comparten la misma función valor, denotada como $V^\ast$.
-Si para una política $\pi^\ast$ se tiene que es mejor que todas las demás, entonces podemos describir su función valor como
-\[  V^{\pi^\ast} = \max_{\pi} V^{\pi}(s).\]
-Por supuesto, $V^\ast = V^{\pi^\ast}$
-\\
-Análogamente, las políticas óptimas tienen la misma función $Q$ óptima, por lo que definimos \textit{función accion-valor óptima} como el retorno esperado de tomar la acción $a$ en el estado $s$ y a partir de ahí seguir una política óptima:
-\[ Q^\ast (s,a) = \max_\pi Q^{\pi}(s,a) = Q^{\pi^\ast}(s,a). \]
-Los métodos modernos de aprendizaje por refuerzo intentan aprender directamente la función valor o la función $Q$, y una vez Q y una vez que tenemos $Q^{\pi^\ast}$ podemos definir una política óptima de la siguiente manera:
-\[ \pi^\ast = \text{argmax}_a Q^\ast (s,a). \]
-Además, utilizando la ecuación de Bellman vimos que se podía escribir $Q$ en términos de $V$, y por lo tanto podemos escribir $Q^\ast$ en términos de $V^\ast$ como
-\[ Q^\ast (s,a) = \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La ecuación de Bellman puede escribirse sin hacer referencia a ninguna política de la siguiente manera:
-\[ V^\ast (s) = \max_a \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La función óptima $ Q^\ast$ lo que hace es maximizar sobre todas las decisiones:
-\begin{eqnarray*}
-Q^\ast &=& r_{t+1} + \gamma \max_{a_{t+1}} r_{t+2} + \gamma^2 \max_{a_{t+2}} r_{t+3}...\\
-&=& r_{t+1} + \gamma \max_{a_{t+1}} Q^{\ast} (s_{t+1},a_{t+1}).
-\end{eqnarray*}
-Formalmente:
-\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
-\subsection{Métodos básicos de solución: Model-based}
-\subsubsection{Programación Dinámica}
-Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
-				\begin{itemize}
-					\item Evaluación de una política: Teniendo una política cualquiera $\pi$, ¿cómo obtener la función $V^\pi$?\\
-					         Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
-					         \[ V^\pi (s) = \mathbb{E}[r + \gamma V^\pi (s') | s_t = s]. \]
-					          Pero esto nos deja con el mismo problema, pues tenemos de ambos lados una función desconocida.
-Para problemas de decisión finitos, este es un sistema de ecuaciones con tantas ecuaciones como estados, lo cual puede ser costoso de resolver.
-Consideramos entonces un método iterativo:
-					         \[ V_{k+1} (s) = \pi(s) \sum_{s'} \mathcal{P}_{s,s'}^a [ \mathcal{R}_{s,s'}^a + \gamma V_k(s')], \]
-					         donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$.
-Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.
-					\item Mejora de una política: Dada una política $\pi$ y estando en un estado $s$ nos preguntamos qué pasaría si en vez de escoger la acción $\pi(s)$ escogieramos cualquier otra acción $a$.
-La consecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
-Entonces, la clave es comparar esto con $V^{\pi}$.
-De ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
-Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.
-					 Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
-					\[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
-					Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.
-					\item Iteración de políticas: Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
-					\[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
-					pues cada política está garantizado que es mejor que la anterior.
-					\item Iteración de valor: El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite. Podemos truncar la etapa de policy evaluation.
-					\[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
-					\item Iteración de política generalizada (GPI): Notar que la iteración de políticas consiste en dos procesos que interactuan entre sí, pero de manera que cada proceso se hace de manera completa antes de comenzar el otro lleva a preguntarse si esto puede hacerse de manera más eficiente.
-Se conoce como iteración de políticas generalizada (GPI) a la idea general de que los procesos de evaluación y mejora de política interactuen entre sí.
-					\item Limitaciones: Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
-Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
-				\end{itemize}
-\subsection{Métodos básicos de solución: Model-free}
-Métodos que no necesitan un modelo completo del sistema, sólo necesitan experiencia.
-Encontramos los métodos Monte Carlo que simulan las transiciones y los estados, y los métodos de diferencias temporales, que van actualizando iterativamente el valor de los estados observados, el más famoso de estos es el TD($\lambda$).
-\subsubsection{Métodos Monte-Carlo} 
-Los métodos Monte-Carlo no requieren un modelo del ambiente y son conceptualmente sencillos, pero no son adecuados para hacer actualizaciones incrementales.
-Estos métodos requieren \textit{experiencia} en vez de un modelo completo; es decir, requieren secuencias de estados, acciones y recompensas, ya sean estos obtenidos de simulaciones o de interacción \textit{on-line} con el ambiente.
-Los métodos Monte-Carlo suponen que la experiencia proviene de eventos episódicos, es decir que terminan siempre de una u otra forma y es al final de estos episodios que se alteran las estimaciones de valor así como las políticas.
-En este sentido, los métodos Monte-Carlo son incrementales por episodio, pero no paso a paso dentro de cada episodio.
-A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general.
-La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
-\begin{itemize}
-\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
-Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$.
-El algoritmo \textit{every-visit} estima $V^\pi (s)$ como las recompensas promedio de todas las visitas al estado $s$ en todos los episodios; por otro lado, el algoritmo \textit{first-visit} promedia sólo lo obtenido al visitar por primera vez un estado en todos los episodios.
-Ambos métodos convergen a $V^\pi$.
-\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
-Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política.
-Entonces, es importante tener un método para estimar $Q^\ast$.
-\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
-El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.
-\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
-A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
-\[ \pi_0 \xrightarrow{E} Q^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} Q^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} Q^\ast.  \]
-Bajo el supuesto de episodios infinitos y visita a todas las parejas estado-acción se tiene que los métodos Monte-Carlo encuentran $Q^{\pi_k}$ de manera exacta para $\pi_k$.\\
-Para hacer la mejora de una política, se escoge de manera determinística la acción $a$ en la cual se alcanza el máximo de la función $Q$; es decir,
-\[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
-\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
-\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
-							\begin{itemize}
-									\item On-Policy: intentan evaluar mejorar la política que es actualmente utilizada para la toma de decisiones.
-									\item Off-Policy: Evaluan o mejoran una política distinta
-							\end{itemize}
-\end{itemize}
-\subsubsection{Diferencias Temporales}
-Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos.
-Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
-\begin{itemize}
-\item Evaluación de política en Diferencias Temporales
-\item TD(0)
-\item SARSA: Es un método \textit{on-policy}
-\item Q-Learning: Es un método \textit{off-Policy}
-\item Métodos Actor-Critic
-\end{itemize}
-\section{Deep Reinforcement Learning}
-\subsection{Mejoras a los métodos básicos}
-	\begin{itemize}
-		\item Programación Dinámica con aproximación de funciones.
-		\item Cuando el espacio de estados es muy grande, no es factible utilizar una representación en una \textit{look up table}, por lo que es necesario utilizar otro tipo de aproximadores (\cite{busoniu2010reinforcement})
-	\end{itemize}
-\subsection{Deep Learning}
-	\begin{itemize}
-		\item Redes neuronales profundas como aproximadores generales.
-		\item Redes Convolucionales para poder tomar como input los pixeles de un juego directamente.
-	\end{itemize}
-\subsection{Value based Reinforcement Learning:}
-	\begin{itemize}
-	   \item Representar la función valor mediante una \textit{Q-network} con pesos $w$:
-	   \[Q(s,a,w) \approx Q^\ast(s,a) \]
-	   \item Q-learning:
-        \begin{itemize}	  
-        	\item  Sabemos que los valores óptimos deben satisfacer la ecuación de Bellman:
-        	\[ Q^\ast(s,a)=\mathbb{E}_{s'}[r + \gamma \max_{a'} Q^\ast(s',a')|s,a]. \]
-        	\item Tratamos el término $r + \gamma \max_{a'} Q^\ast(s',a',w)$ como el objetivo.
-        	\item Minimizar error cuadrático medio mediante Gradiente Estocástico:
-        	\[ I = (r + \gamma \max_{a'} Q^\ast(s',a',w) - Q(s,a,w))^2. \]
-        	\item Converge a $Q^\ast$ utilizando una representación tabular.
-        	\item Diverge al utilizar redes neuronales debido a la alta correlación entre muestras y la no-estacionareidad del objetivo.
-        	\item Lo resolvemos con Experience Replay (\cite{Lin1992}),  el cual consiste en guardar todas las experiencias de modo que al entrenar la red tomar sólo mini-batches aleatorios  en vez de utilizar la transición más reciente.
-	    \end{itemize}
-	\item Aplicaciones de DQN.
-		\begin{itemize}
-			\item \cite{mnih2013playing}.
-Paper en el que muestran cómo una computadora aprende a jugar Atari 2600 a través de sólo observar los pixeles del juego y obteniendo una 																		recompensa conforme el score aumentaba.
-Utilizando la misma arquitectura, se pudieron aprender 7 juegos
-			\item \cite{mnih2015human}.
-En este paper, aumentaron a 49 juegos aprendidos.
-Utilizan una Deep Q-Network, que consiste en aproximar la función Q mediante una red 																			neuronal que toma como input los estados (las pantallas del juego) y devuelve como output el valor Q para cada posible acción.
-En este caso, 																			utilizaron una red convolucional con 3 capas de convolución y 2 capas totalmente conectadas.
-	  \end{itemize}
-\item Mejoras posteriores a DQN
-   	   \begin{itemize}
-   	   		\item Double DQN.
-   	   		\item Prioritized Replay.
-   	   		\item Duelling Networks.
-   	   \end{itemize}
-    	 \end{itemize}  	   
-\subsection{Policy Based DRL}
-Resulta que Deep Q-Learning no es un gran algoritmo, y en su lugar, \cite{DBLP:journals/corr/MnihBMGLHSK16} proponen utilizar directamente Policy-Gradients.
-La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
-\[ a = \pi (a|s,u) \textrm{ o } a = \pi (s,u). \]
-En este contexto, definimos una función objetivo $L(\textbf{u})$ como:
-		\[ L(\textbf{u}) = \mathbb{E}[r_1 + \gamma r_2 + \gamma^2 r_3 + ... |  \pi(\cdot,\textbf{u})]. \]
-Optimizamos con Stochastic Gradient Descent. El gradiente de una policy estocástica está dado $\pi (a|s,\textbf{u})$ por:
-		\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[  \frac{\partial \log \pi(a|s,\textbf{u})}{\partial \textbf{u}}Q^\pi(s,a)\right].\]
-Por otro lado, el gradiente de una policy determinística $a=\pi(s)$ está dado por:
-\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[ \frac{\partial Q^\pi (s,a)}{\partial a} \frac{\partial a}{\partial \textbf{u}} \right]. \]
- Los métodos \textit{actor-only}, donde \textit{actor} se utiliza como sinónimo de \textit{policy} utilizan familias parametrizadas de policies.
-Las ventajas de tener una policy parametrizada es que podemos generar un espectro continuo de accciones, pero los métodos de optimización (policy gradients) sufren de alta varianza en sus estimaciones del gradiente, lo cual genera un lento aprendizaje (\cite{grondman2012survey},\cite{konda2003onactor},\cite{richter2007natural},\cite{boyan2002technical},\cite{baxter2001infinite},\cite{berenji2003convergent}).\\
- \\
-Los métodos \textit{critic-only}, donde \textit{critic} se utiliza como sinónimo de \textit{value function}, que utilizan diferencias temporales tienen una menor varianza en las estimaciones del retorno esperado (\cite{boyan2002technical},\cite{berenji2003convergent},\cite{sutton1988learning}).
-Una manera de derivar una policy en estos escenarios es seleccionando greedy actions; es decir, acciones praa las cuales la value funtion indique que el retorno esperado es máximo, lo cual puede requerir un alto costo computacional.
-Lo usual es discretizar el espacio continuo de acciones.
-Los métodos \textit{actor-critic}, combinan las vetajas de ambos mundos.
-Mientras que las familias parametrizadas tienen la ventaja de utilizar acciones continuas, los métodos critic-only proveen de conocimiento de baja varianza sobre el desempeño; es decir, la estimación del retorno esperado del crític-only permite al actor-only actualizar los gradientes con información de menor varianza, lo cual acelera todo el proceso.
-La menor varianza es intercambiada por un mayor sesgo al inicio del aprendizaje (\cite{berenji2003convergent}).
-\\
-\\
-Los métodos de Policy Gradients optimizan directamente la recompensa acumulada y pueden utilizarse junto con aproximadores no-lineales como las redes neuronales.
-Este método presenta dos grandes retos: el número de muestras de entrenamiento requeridos y la dificultad de obtener mejoras estables debido a la no-estacionariedad de los datos.
-\cite{DBLP:journals/corr/SchulmanMLJA15} atacan el primer problema se puede atacar utilizando value functions para reducir la varianza de los estimados.
-El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
-\\
- \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada.
-Es un algoritmo similar a Policy Gradients y resulta útil para optimizar políticas con redes neuronales.
-Buen performance en juegos de Atari que utilizan la pantalla como input.
-
-\section{Reinforcement Learning y Causalidad}
-Aprendemos causalidad mediante intervenciones, interacciones, etc. Reinforcement Learning es aprender a realizar tareas mediante interacción. ¿Habrá algo en medio?
-\subsection{Conocimiento causal y observabilidad parcial}
-Los algoritmos actuales de reinforcement Learning suponen que el problema puede modelarse dentro de un contexto Markoviano de decisión con recompensas.
-Si el supuesto markoviano se elimina, los algoritmos podrían dejar no funcionar, como muestran en \cite{singh1994learning}.
-Además, todo el desarrollo teórico utiliza la propiedad Markoviana.\\
-\\
-Una manera en la que la propiedad Markoviana puede perderse es si el agente no cuenta con información completa sobre el estado; es decir, al igual que en los HMM, suponemos que el ambiente es Markoviano, pero el agente no tiene acceso a él de modo que los datos no le \textit{parecen} Markovianos.
-A diferencia de los HMM, no existe un procedimiento computacionalmente tratable para los POMDP.
-El problema es que una vez que se obtienen las estimaciones del valor para cada estado, se debe hacer programación dinámica en espacios continuos y esto es infactible salvo para casos pequeños\\
-\\
-La manera de generalizar el caso markoviano al caso no observado es considerar un MDP subyacente con estados $\mathcal{S}=\{ s_1,...,s_N\}$ y probabilidades de transición $\mathcal{P}^a_{s,s'}$.
-Luego, en cada tiempo $t$, el agente observa un mensaje $m_t$ el cual proviene de una distribución $P(m | s_t)$ desconocida.\\
-\\
-En \cite{jaakkola1995reinforcement} se describe un procedimiento que evita estimar el valor de los estados y trabaja directamente en el espacio de políticas estocásticas.
-El procedimiento es \textit{directo}, a diferencia de los métodos \textit{indirectos} en los cuales primero se aprenden los parámetros del MDP y luego se utiliza programación dinámica para obtener la política.
-El algoritmo propuesto consiste en dos partes: primero, se utiliza un método Monte-Carlo para calcular un análogo de los Q-values y luego se utiliza un paso de policy-improvement.\\
-\\
-Desde un punto de vista de psicología, una manear de entender la inferencia de estados ocultos es en términos de causas latentes (\cite{gershman2015reinforcement}).
-Siguiendo esta idea, es evidente que un agente de RL que opera en el mundo real no sólo debe hacer inferencia de estados ocultos, sino que debe poder descubrir los estados ocultos que están detrás del mundo que éste observa.
-Esto es una forma de aprendizaje estructural latente (\cite{gershman2010learning}).\\
-\\
-Un primer acercamiento a este problema puede hacerse mediante técnicas de inferencia Bayesiana no paramétrica.
-Por ejemplo, tanto \cite{gershman2010learning} como \cite{gershman2012exploring} utilizan priors no paramétricas para hacer inferencia sobre causas latentes.
-Un modelo causal latente es una buena representación de la estructura causal verdadera bajo condicionamiento Pavloviano, pues es el experimentador quien es una causa latente.\\
-\\
-En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
-\subsection{Towards Deep Symbolic Reinforcement Learning}
-Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
-\\
-\cite{garnelo2016towards} proponen una arquitectura híbrida que, combina redes neuronales con inteligencia artificial simbólica.
-El modelo propuesto consiste en combinar una red neuronal que como entrada directamente lo que se recibe del ambiente y transforma estos datos crudos en una representación simbólica basada en lógica de primer orden; esta representación simbólica es posteriormente utilizada para escoger la acción a tomar.\\
-\\
-Se utiliza un sencillo juego para probar la arquitectura y en este contexto la red neuronal debe generar representaciones simbólicas de secuencias de estados en el juego en las cuales el flujo de pixeles queda codificado en términos de objetos, tipos y sus interaccioes.
-Luego, con esto se aprenden las políticas.\\
-\\
-El primer paso que hacen es utilizar un autoencoder convolucional para generar, de manera no supervisada, un conjunto de símbolos que representarán los objetos en cada escena.
-Luego, para cada pixel, obtienen la feature con mayor activación y se toman los más altos.
-De esta manera se obtiene un pixel representativo para cada objeto a los cuales posteriormente se le asigna un tipo simbólico según el autoencoder.\\
-\\
-Luego, una vez que se tienen estos símbolos de bajo nivel para cada cuadro del juego, se necesita rastrearlos a lo largo del tiempo.
-Utilizan una verosimilitud que es función de proximidad espacial y el tipo de transiciones.
-Para considerar la información de la dinámica de los cuadros utilizan la diferencia entre cuadros y las posiciones relativas entre objetos.\\
-\\
-Para la parte de aprendizaje, entrenan una función $Q$ por separado para cada interacción entre objetos con actualización
-\[Q^{ij}(s_t^{ij},a_t) \leftarrow  Q^{ij}(s_t^{ij},a_t) + \alpha \left( r_{t+1} + \gamma (\max_{a} Q^{ij} (s_{t+1}^{ij},a) - Q^{ij}(s_t^{ij},a_t)) \right). \]
-\\
-Se prueba el framework sobre varias variantes de un sencillo videojuego. \\
-\\
-Llama la atención como estos autores utilizan maquinaria de lo más reciente (deep learning, autoencoders) para luego obtener una representación en términos de lógica de primer orden.
-Este approach ha quedado claro que ya quedó atrás por lo que debería buscarse algo que sustituya este enfoque lógico.
-En particular, pudiera ser esto un modelo gráfico causal el cual vaya aprendiéndose conforme el agente de RL aprende sobre su entorno.
-
-\subsection{Conocimiento causal y toma de decisiones secuenciales: acciones como intervenciones}
-En cuanto al aprendizaje \textit{on-line} de modelos causales, encontramos en los trabajos de \cite{wellen2012learning} y de \cite{kummerfeld2013tracking} acercamientos tangenciales, o un punto de inicio, al aprendizaje de modelos causales al vuelo.\\
-\\
-Teorías de cómo las personas usan datos de covariación para aprender modelos gráfios causales pueden dividirse en enfoques racionales y heurísticos (según  \cite{wellen2012learning}).
-El enfoque racional modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana.
-Estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
-\\
-Por otro lado, el enfoque heurístico propone que las personas reaccionan a distintos estímulos que los hacen moficiar sus hipótesis causales de una manera no necesariamente racional.
-El modelo de efecto simple de \cite{waldmann2008causal} supone que las personas se enfocan en evaluar una sola relación causal a la vez.
-Es un modelo que aprende la magnitud de la relación causal.
-Si una variable supera cierto umbral, entonces la persona acepta esa relación causal y la incorpora a su bagaje de relaciones.\\
-\\
-El modelo propuesto por \cite{wellen2012learning} sigue esta línea del efecto simple, con la diferencia de que este modelo sí intenta descubrir el proceso cognitivo subyacente, pues el aprender magnitudes causales es un proceso puramente computacional.
-El modelo de efecto simple puede verse como una implementación del modelo de \cite{danks2003dynamical}.
-Además, según narran \cite{wellen2012learning}, el modelo de efecto simple no explica cómo la persona utiliza las magnitudes causales para determinar si existe o no una relación causal.
-En el modelo propuesto, \cite{wellen2012learning}, incorporan un procedimiento de decisión para determinar si existen o no relaciones causales.\\
-\\
-El modelo LPL (\cite{wellen2012learning}) comienza con una estructura causal inicial, la cual consiste en un grafo acíclico dirigido el cual representa las creencias iniciales de un individuo.
-Posteriormente, el modelo altera esta estructura al añadir o remover una sola arista, de manera que el problema se reduce a evaluar relaciones causales entre dos variables.
-Este modelo genera estimaciones de la fuerza de la relación causal para cada posible relación causa-efecto que no sea eliminada a priori.\\
-\\
-Supongamos que un individuo cree a priori que $B$ causa $E$, y está tratando de averiguar si $C$ causa $E$.
-A diferencia del modelo de efecto simple, las estimaciones de la fuerza causal están generadas por un mecanismo similar a los filtros de partículas Bayesianos
-En particular, a partir del conocimiento causal inicial del individuo, el modelo LPL comienza con $n$ partículas.
-Estas partículas $\{V_C^1,...,V_C^n \}$ representan las creencias actuales del individuo sobre si $C$ causa $E$.
-Paralelamente, existen $\{ V_B^1,...V_B^n \}$ para $B$.
-Se define la \textit{capa} $i$ de partículas para el efecto $E$ como la $i$-ésima partícula de cada causa conocida y posible para $E$.
-En el ejemplo, tendríamos que esta capa es $\{ V_B^i, V_C^i \}$.
-Luego, el modelo LPL actualiza cada capa de manera independiente de la siguiente manera
-\[ V_C^{i+1} = V_C^{i,t} + \alpha ( \textrm{observado} - \textrm{esperado}), \]
-Donde “observado” tiene el valor de 1 si el efecto sucede y 0 en caso contrario.
-Por otro lado, “esperado” es el valor esperado del efecto, calculado según la forma funcional de la relación causa-efecto, el cual es calculado por separado para cada causa potencial en una capa.
-El valor esperado de $E$ (para la capa $i$ y la causa potencial $C$) es
-\[ \textrm{esperado} = \prod_{k} (1-V_k^{t-1})(1 - \prod_j (1- V_j^{t-1}).\]
-Conforme el individuo observa más datos, las partículas convergen a los valores verdaderos de los parámetros y las estimaciones para cada capa convergen.
-Este procedimiento no toma en cuenta aspectos estructurales, los cuales se modifican según un criterio de decisión.
-Lo que hace el modelo es una prueba $t$ en la cual la hipótesis nula es que las partículas provienen de una distribución con media cero.
-El criterio de decisión es el siguiente: si actualmente no existe una arista, y la prueba $t$ rechaza la hipótesis nula, entonces se añade la arista.
-Si existe una arista, y la prueba no es significativa, entonces se elimina la arista existente.
-Alterar una arista afectará las futuras estimaciones del valor “esperado” para otras causas.\\
-\\
-\cite{hagmayer2013hierarchical} llaman la atención sobre que los modelos actuales de Decisiones (maximum expected utility) no asumen que los tomadores de decisiones aprenden sobre la estructura causal del problema.
-Por otro lado, las teorías de aprendizaje causal proponen que un tomador de decisiones guían sus acciones a través de conocimiento causal previamente obtenido.
-El modelo de \cite{sloman2006causal} extiende las teorías de aprendizaje causal al contexto de toma de decisiones, la idea básica consiste en que los tomadores de decisiones utilizan la información disponible de su entorno para inducir un modelo causal del problema de decisión al que se enfrentan.
-Se sabe que en los seres humanos el conocimiento causal guía la interpretación de información probabilista y que los humanos somos capaces de aprender la estructura causal del entorno a través de una serie de (traducir cues) (\cite{lagnado2007beyond}).
-La pregunta fundamental que plantean \cite{hagmayer2008causal} es entender qué es lo que aprenden las personas al tomar decisiones de manera secuencial de tal forma que se busque maximizar cierta recompensa a largo plazo y donde las acciones alteran el entorno.
-Al exponerse una persona a estas situaciones de decisión secuencial se obtiene una retroalimentación de cómo las acciones se relacionaban con la obtención de recompensas.
-Las distintas teorías sobre toma de decisiones bajo incertidumbre suponen distintas representaciones para esta retroalimentación.
-Hagmayer propone agrupar estas distintas representaciones en tres niveles:
-\begin{itemize}
-\item Nivel 1: Opciones, resultados y recompensas.
-Sólo los valores esperados de las opciones posibles afectan la toma de decisiones.
-Se conoce como \textit{habit learning} en la literatura de aprendizaje en animales.
-\item Nivel 2: Opciones, relaciones causales de estas opciones a los resultados, relaciones causales entre variables de resultado y relaciones de estas con las recompensas.
-Aquí entra lo que se conoce como \textit{goal-directed learning}.
-La diferencia con el nivel 1 es la capacidad de distinguir entre acciones y los valores asociados a estas.
-\item Nivel 3: Opciones, relaciones causales entre variables de resultado, relaciones causales entre estas variables y relaciones de estas con las recompensas.
-Estos son los modelos causales.
-Las distintas teorías sobre estos modelos suponen que los humanos (o el agente que aprende) adquieren estos modelos mediante procesos de aprendizaje.
-Algunas teorías suponen que la estructura causal del entorno se obtiene a través de \textit{claves}, como la propuesta por \cite{lagnado2007beyond}.
-Por otro lado, otros autores opinan que las personas utilizan propiedades estadísticas de los datos observados, este sería el caso de las teorías de \cite{gopnik2004theory} y de \cite{griffiths2005structure}.
-\end{itemize}
-Aunque las teorías que entran en el nivel 2 son sensibles a las correlaciones entre las respuestas obtenidas, sólo en el nivel 3 se hace una distinción entre causalidad y correlación.\\
-\\
-Los autores hacen énfasis en que aunque las personas sí utilicen modelos causales al tomar decisiones simples, estas decisiones suelen ser sobre escenarios hipotéticos, pero no se han hecho muchos estudios sobre adquisición de conocimiento causal al enfrentarse a problemas de decisión secuencial con el objetivo de maximizar alguna ganancia a largo plazo.
-Además, en los estudios realizados, se ha visto que las personas tienen la capacidad de inducir modelos causales a partir de claves en el entorno; se sabe además que a partir de intervenciones activas se toman mejores decisiones que considerando sólo datos observados de manera pasiva y que las personas tienen la capacidad de integrar relaciones causales obtenidas de manera separada.
-Pero en estos estudios las personas tenían como objetivo descubrir de manera directa estructuras causales.
-No se sabe bien hasta qué punto las personas adquieren conocimiento causal al perseguir objetivos que no sean descubrir directamente estructuras causales a partir del ambiente.\\
-\\
-Los autores realizaron tres experimentos para explorar las representaciones que las personas hacen de las relaciones causales que adquieren al tomar decisiones para maximizar algún objetivo.
-En el primer experimento, el objetivo era estudiar cómo las creencias causales de las personas alteran la toma de decisiones y cómo los participantes utilizaban los resultados de sus decisiones para revisar el modelo causal.
-El experimento consistía en dos etapas de toma de decisiones, la segunda siendo una etapa de prueba.
-En ambas etapas los participantes debían escoger entre una serie de opciones de modo que maximizaran ganancias.
-Se le sugirió a cada participante una posible cadena causal.
-En la primera etapa se les proveía retroalimentación de las consecuencias de sus decisiones, lo cual no sucedió en la segunda etapa sino que ellos mismos debían inferir posibles relaciones causales.
-Se concluyó que los participantes encontraron estructura causal en el problema y la utilizaron para tomar futuras decisiones.\\
-\\
-En el segundo experimento y tercer experimento, a los participantes se les mostró una serie de hipótesis causales para ver si inducían de manera espontánea un modelo causal de la situación.\\
-\\
-Lo que estos experimentos muestran, en conclusión de los autores, es que al intentar maximizar una ganancia a largo plazo, las personas adquieren conocimiento causal del entorno, el cual afecta futuras decisiones.
-Mencionan que estos experimentos ponen en tela de juicio los modelos estándar de toma de decisiones, pues utilizando sólo valores esperados no se llegarían a los mismos resultados.\\
-\\
-Como consecuencia de sus estudios, se obtiene que los métodos basados en los niveles 1 y 2, que es donde caerían los métodos tradicionales de aprendizaje por refuerzo no son suficientes para una adecuada toma de decisiones en un sistema causal, pues estas formulaciones no cuentan con la expresividad suficiente para capturar las relaciones causales.
-Esto confirma la necesidad de pensar en modelos que permitan aprender modelos causales a la par que un agente aprende de su entorno mediante interacción y exploración.\\
-\\
-En cuanto a la formulación de un problema de decisión en sistemas causales, ya desde \cite{dawid2002influence} tenemos una idea de cómo aumentar redes bayesianas con nodos de decisión y variables de estrategia para el modelado causal en problemas de decisión.\\
-\\
-\cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta. 
-\section{Propuesta}
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-\subsection{Objetivos}
-\begin{itemize}
-\item Aprendizaje causal en línea a partir de acciones y recompensas obtenidas por un agente.
-\item Evaluación del modelo causal.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
-\end{itemize}
-\subsection{Propuesta metodológica}
-Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en incorporar un modelo de aprendizaje causal dentro a un problema de toma de decisiones.
-En el contexto de aprendizaje por refuerzo, tendremos un agente que, al tener como objetivo aprender una tarea mediante interacciones secuenciales, aprenda con cada paso un poco sobre la \textit{estructura causal} de su entorno.
-Cada decisión que el agente toma (o acción que lleva a cabo) puede verse como una intervención en el ambiente y que proporciona información sobre la estructura causal de este.
-Esta información obtenida, será utilizada para mejorar las decisiones que toma y para aprender otros aspectos estructurales (causales) sobre su ambiente.
-La intuición detrás de esto es que al aprender mediante una toma de decisiones repetida, decisiones que intervienen directamente sobre el entorno, se aprenden también relaciones causales; por ejemplo, al aprender a conducir se aprende implícitamente que ciertos movimientos del volante \textit{causan} ciertos cambios en el estado del automovil (\cite{danks2014unifying}).
-Otro ejemplo es en el contexto de videojuegos, en el cual al tomar ciertas decisiones se conoce el impacto que estas provocan en el estado del juego, conocimiento que a su vez es utilizado para tomar mejores decisiones en un futuro y que maximicen las recompensas u objetivos a largo plazo.
-No sólo hay intuición detrás de esto, pues se sabe que los seres humanos utilizan el conocimiento causal que la experiencia les provee para realizar predicciones para intervenciones en el mundo no realizadas aun (\cite{meder2008inferring}, \cite{hagmayer2009decision})\\
-%\\
-%De manera formal, consideremos un fenómeno sobre el cual se quieran hacer inferencias causales y sobre el cual poseeemos un conjunto de datos.
-Supongamos que los datos tienen la forma $%(\mathbf{X},\mathbf{Y})$ donde $\mathbf{X}$ son un conjunto de atributos, así como acciones llevadas a cabo, y $\mathbf{Y}$ son los efectos observados.
-Entonces, es necesario %\textit{traducir} estos datos a la forma $(s \in \mathcal{S},a \in \mathcal{A},r \in \mathcal{R})$ de modo que las características, atributos, etc del problema original sean mapeados a estados, las acciones tomadas, a acciones en el mundo de RL, y las respuestas obtenidas, al reward y cambio de estado otorgado por el ambiente.
-Evidentemente esta no es una tarea trivial, pues es necesario también definir un \textit{task} adecuado para el agente que le permita interactuar con su entorno de manera tal que aprenda, o descubra, las relaciones causales sobre los datos originales.
-La definición del task a resolver implica definir una función recompensa; además, una transición entre estados es importante.
-Notemos que todos estos elementos deben ser definidos de manera que el agente interactue con su mundo de tal forma que lo que \textit{aprenda} nos sirva para identificar relaciones causales.\\
-%\\
-%Como ejemplo de lo anterior, podríamos tener un conjunto de datos de la forma $\mathbf{X} = (s,t)$ donde $s$ es enfermedad, $t$, tratamiento otorgado y $\mathbf{Y}=\{ \text{curado}, \text{ no curado} \}$.
-Definimos un agente de RL tal que los estados que recibe del ambiente son las características del paciente, la acción que toma es otorgar o no el medicamento, y la recompensa es $1$ si el paciente se mejoró.
-La tarea que este agente tiene que lograr es maximizar el número de pacientes curados; notemos que la tarea del agente, o el agente mismo, no tienen por qué tener interpretación en términos del problema original; es decir, no tiene por qué existir una interpretación del agente en términos de un médico, un paciente, etc.
-Una vez que el agente aprenda a resolver esta tarea, utilizaremos lo aprendido por él de la siguiente manera: si para un paciente del conjunto original de datos el agente propone el tratamiento, entonces es causal que el tratamiento mejoró al paciente.\\
-\\
-El trabajo de \cite{garnelo2016towards} nos muestra que efectivamente pueden aprenderse aspectos estructurales sobre el entorno mientras el agente aprende, incluso logrando así mejorar su desempeño.
-Además, muestra la importancia de tener una semántica que permita hablar de la estructura del entorno.
-Ellos utilizaron lógica de primer orden, pero un modelo gráfico sería más expresivo y adaptable.\\
-\\
-A su vez, el trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
-Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
-\cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.\\
-\\
-\begin{center}
-\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/diagrama.png}
-\end{center}
-\newpage
+Para muestrear de la posterior $P(z | R ) \propto P(R|z)P(z)$ se pueden utilizar técnicas MCMC o al buscar la moda de la distribución.
+\section{Aplicación}
+\subsection{Objetos y atributos}
+Aunque el IRM está pensado para datos relacionales, también puede utilizarse para encontrar estructura en datos objeto-atributo.
+Notemos que cualquier matriz de atributos puede verse como una relación $R: T^1 \times T^2 \to \{0,1 \}$ entre un set de objetos $T^1$ y uno de atributos $T^2$.
+También es importante notar que si se decide no clusterizar los atributos, sino asumir que cada uno se genera independientemente sobre las particiones entonces se obtiene el infinite mixture model de Rasmussen.
+\subsection{Ontologías}
+Se analizó una red semántica de 135 conceptos y 49 predicados binarios.
+Los conceptos son, por ejemplo “enfermedad”, “diagnóstico”, etc.
+Los predicados contienen verbos como “complica”, “causa”, “afecta”.
+Se aplicó el IRM a la relación ternaria $R: T^1 \times T^1 \times T^2 \to \{0,1\}$ donde $T^1$ es conjunto de conceptos, $T^2$ conjunto de atributos.
 \bibliographystyle{apalike}

### 2017-12-20 Primer_Semestre/Resumenes/Resumen_Infinite_Hidden_Relational_Models.tex
primer commit

--- 2017-12-20 Primer_Semestre/Resumenes/Resumen_IRM.tex
@@
 \newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Resumen: Learning Systems of Concepts with an Infinite Relational Model - Kemp, Tenenbaum, Griffiths}
+\title{Resumen: Learning Infinite Hidden Relational Models - Xu, Tresp, Kriegel}
 \author{Mauricio Gonzalez Soto}
@@
 \maketitle
-\section{Introducción}
-Si cierto dominio consiste en varios tipos de entidades,
-\section{Infinite Relational Models}
-Suponiendo que tenemos una o más relaciones que involucran uno o más tipos.
-El IRM busca partir cada tipo (type) en clusters, de manera que una buena partición permita que relaciones entre entidades se puedan predecir a partir de la asignación de clusters.\\
-\\
-Por ejemplo, si tenemos $m$ relaciones que involucran $n$ tipos, denotamos $R^i$ la $i$-ésima relación, $T^j$ la $j$-ésima relación y $z^j$ un vector de asignación de clusters.
-Lo que se busca es inferir la asignación de clusters, y ultimadamente la posterior $P(z^1,...,z^n | R^1, ...
-, R^m)$.
-Para especificar esta distribución, se define un modelo generativo para las relaciones y la asignación de clusters
-\[ P(R^1,...,R^m, z^1,...,z^n) = \prod_{i=1}^m P(R^i | z^1 , ... , z^n) \prod_{j=1}^n P(z^j). \]
-Estamos asumiendo qye las relaciones son condicionalmente independientes dada la asignación de clusterss, y que la asignación de clusters es independiente entre tipos.
-\section{Generación de clusters}
-Para que el IRM sea capaz de descubrir el número de clusters en el tipo $T$, se utiliza un prior que asigna probabilidad a todas las posibles particiones de $T$.
-Un prior razonable debería descubrir sólo tantos clusters como sugeridos por los datos.
-A partir de trabajos previos en Bayesiana No paramétrica (\cite{rasmussen2000infinite}) utiliza una distribución sobre particiones inducida por un CRP  (Chinese Restaurant Process, \cite{teh2011dirichlet}).\\
-\\
-La idea es (\cite{teh2011dirichlet}) que en un restaurante chino hay un número infinito de mesas, cada una de las cuales puede sentar un número infinito de clientes.
-El primer cliente entra y se sienta en la primera mesa, el segundo cliente entra y decide si sentarse en la primera mesa o en otra; en general, el $n$-ésimo cliente se sienta en una mesa ocupada con una probabilidad proporcional al número de clientes sentados ahí.\\
-\\
-Entonces, la manera de constuir clusters utilizando el CRP es partir de un cluster con un solo objeto, e ir añadiendo objetos de modo que la probabilidad de caer en cierto cluster sea proporcional a los elementos en él.
-De esta manera, la distribución de clusters para el objeto $i$, dada la asignaclión de los objetos $1,...,i-1$ es:
-\[
-  P(z_i = a | z_1, ... , z_{i-1}) =
-  \begin{cases}
-                                   \frac{n_a}{i-1+\gamma} & \text{if $n_a > 0$} \\
-                                   \frac{\gamma}{i-1+\gamma} & \text{$a$ es un nuevo cluster} 
-  \end{cases}
-\]
-Donde $n_a$ es el número de objetos asignado al cluster $a$ y $\gamma$ es un parámetro.
-La distribución en $z$ inducida por el CRP es intercambiable; es decir, no importa el orden en que los objetos llegan a los clusters.
-Como nuevos objetos siempre pueden ser asignados, el IRM efectivamente tiene acceso a un número infinito de clusters, de ahí el nombre del Infinite Relational Model.
-\section{Generación de relaciones a partir de clusters}
-Se asume que las relaciones son binarias, pero se puede extender a datos continuos y frecuencias.
-Como ejemplo, veremos primero el caso con un solo tipo $T$ y una relación $R: T \times T \to \{0,1 \}$.
-$T$ puede ser, por ejemplo, un conjunto de personas y $R$ una relación de amistad.
-El modelo generativo para este problema es
-\begin{eqnarray*}
-z | \gamma &\sim& CRP(\gamma)\\
-\eta(a,b) | \beta &\sim& Beta(\beta,\beta)\\
-R(i,j) | z, \eta &\sim& Bernoulli(\eta(z_i,z_j))
-\end{eqnarray*}
-Esto es, estamos suponiendo que la relación $R$ está generada a partir de dos estructuras latentes: una partición $z$ y una matriz de parámetros $\eta$.
-La entrada $R(i,j)$ se genera al lanzar una moneda con sesgo $\eta(z_i,z_j)$ donde $z_i,z_j$ son las asignaciones de cluster de las entidades $i$,$j$.
-El parámetro $\eta(a,b)$ especifica la probabilidad de que exista un link entre las entidades $i \in a$, $j \in b$.
-El IRM invierte ese modelo para descubrir la $z$ y la $\eta$ que mejor explican $R$.\\
-\\
-Estamos asumiendo que la tendencia de una entidad a participar en relaciones está dada por su asignación de cluster.
-\section{Inferencia}
-Como en el ejemplo anterior se utilizó una prior conjugada, es sencillo calcular $P(R|z)$
-\[ P(R|z)= \prod_{a,b} \frac{Beta(m(a,b) + \beta, \bar{m}(a,b)+\beta}{Beta(\beta,\beta)} \]
-donde $m(a,b)$ es el número de pares $(i,j)$ tales que $i \in a$, $j \in b$ y $R(i,j)=1$ mientras que $\bar{m}(a,b)$ son las parejas donde $R(i,j)=0.$\\
-\\
-Para muestrear de la posterior $P(z | R ) \propto P(R|z)P(z)$ se pueden utilizar técnicas MCMC o al buscar la moda de la distribución.
-\section{Aplicación}
-\subsection{Objetos y atributos}
-Aunque el IRM está pensado para datos relacionales, también puede utilizarse para encontrar estructura en datos objeto-atributo.
-Notemos que cualquier matriz de atributos puede verse como una relación $R: T^1 \times T^2 \to \{0,1 \}$ entre un set de objetos $T^1$ y uno de atributos $T^2$.
-También es importante notar que si se decide no clusterizar los atributos, sino asumir que cada uno se genera independientemente sobre las particiones entonces se obtiene el infinite mixture model de Rasmussen.
-\subsection{Ontologías}
-Se analizó una red semántica de 135 conceptos y 49 predicados binarios.
-Los conceptos son, por ejemplo “enfermedad”, “diagnóstico”, etc.
-Los predicados contienen verbos como “complica”, “causa”, “afecta”.
-Se aplicó el IRM a la relación ternaria $R: T^1 \times T^1 \times T^2 \to \{0,1\}$ donde $T^1$ es conjunto de conceptos, $T^2$ conjunto de atributos.
+\section{Introduccion}
+La selección de modelos para la estructura en un sistema relacional es difícil debido al número exponencial de \textit{features} de las cuales un atributo puede depender.
+Es mejor introducir, para cada entidad, una variable latente que sólo sea padre de los otros atributos de la entidad y de los atributos de las relaciones.
+Como cada clase de entidades puede tener un distinto número de estados en sus variables lattentes, y este número varía según la información, preferimos que el modelo determine el número de estados latentes de manera automática; esto es posible utilizando modelos de mezcla con el proceso Dirichlet, los cuales se pueden interpretar como modelos de mezcla con un número infinito de componentes de mezcla, pero en los cuales el modelo, basado en los datos, reduce la complejidad a un número finito de componentes
 \bibliographystyle{apalike}

### 2017-12-20 Primer_Semestre/Resumenes/Resumen_Jointly_Identifying_entities_and_extracting_relationships_via_graphical_models_Yu_Lam.tex
primer commit

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Resumen: Jointly Identifying}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\section{Intro}
En este artículo se estudia el problema de reconocimiento de entidades y extracción de relaciones.
\begin{itemize}
\item El documento básico es un artículo, el cual suponemos que trata sobre una entidad, llamada la entidad principal
\item El documento menciona entidades secundarias que están relacionadas a la principal, pero no entre sí
\item Se trabaja con entidades pre-definidas así como relaciones
\end{itemize}
\section{Formulación del problema}
\begin{itemize}
\item Sea $\mathbf{x} = \{ x_1,...,x_N \}$ una sucesión observada de \textit{tokens}.
\item Sea $s_p$ la entidad principal, que suponemos es claramente identificable
\item Sea $ \mathbf{s}=\{ s_1,...,s_L\}$ una segmentación, donde cada $s_i$ es una tripleta $\{ \alpha_i, \beta_i,y_i \}$ dode $\alpha_i$ es una posición inicial, $\beta_i$ posición final y $y_i$ una etiqueta asignada 
\item Debe cumplirse que $0 \leq \alpha_i \leq \beta_i \leq |x|$ y $\alpha_{i+1} = \beta_i + 1$
\item Sea $r_{pn}$ la asignación de relaciones entre la principal $s_p$ y un candidato a entidad secundaria $s_n$ y sea $\mathbf{r}$ el conjunto de asignación de relaciones para la $\mathbf{x}$
\item Sea $\mathbf{y}=\{ \mathbf{r},\mathbf{s} \}$
\item Dada una observación $\mathbf{x}$, queremos encontrar $\mathbf{y}^\ast$ tal que
\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
\end{itemize}
\section{Modelo propuesto}
\begin{itemize}
\item Sean $L$ y $M$ el número de segmentos y el número de relaciones para una secuencia observada $\mathbf{x}$
\item Se define una distribución condicional conjunta para $\mathbf{s},\mathbf{x},\mathbf{r}$ en un modelo grafico no dirigido $\mathcal{G}$
\item Se pueden dividir los factores de $\mathcal{G}$ en tres grupos: $\phi^S$, $\phi^R$, $\phi^\nabla$
\item Respectivamente, el potencial de segmentación, el potencial de relaciones, y el potencial conjunto de segmentación-relación.
\item La función potencial $\phi^S(i,\mathbf{s},\mathbf{x})$ modela la segmentación 
\item El potencial $\phi^R(r_{pm},r_{pn},\mathbf{r})$ representa dependencias entre relaciones en $\mathbf{r}$
\item El potencial conjunto $\phi^\nabla(s_p,s_j,\mathbf{r})$ captura interacciones entre segmentaciones para entidades secundarias
\item Por el Teorema de Hammersley-Clifford, la condicional conjunta $P(\mathbf{y} | \mathbf{x})$ se factoriza como
\[ P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \prod_{C_S} \phi^S (i,\mathbf{s},\mathbf{x}) \prod_{C_R}\phi^R(r_{pm},r_{pn},\mathbf{r}) \prod_{C_\nabla} \phi^\nabla(s_p,s_j,\mathbf{r}) \]
\item Tenemos que la función $Z$ es el normalizante
\[ Z(x)= \sum_{\mathbf{y}} P(\mathbf{y} | \mathbf{x}). \]
\item Se hace el supuesto de que las funciones potencial se factorizan según un set de features y correspondientes pesos
\item Con más detalle,
\[  \phi^S (i,\mathbf{s},\mathbf{x}) = \exp( \sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k (i,\mathbf{s},\mathbf{x}))\]
\item Además, haremos que la función $g_k$ dependa del segmento $s_i$, del anterior $s_{i-1}$ y de toda la observación $\mathbb{x}$ de modo que
\[  g_k (i,\mathbf{s},\mathbf{x}) = g_k(s_{i-1},s_i, \mathbf{x})\]
\item De la misma manera para $\phi^R$ y $\phi^\nabla$ tenemos que
\[ \phi^R(r_{pm},r_{pn},\mathbf{r}) = \exp ( \sum_{m,n}^M  \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},\mathbf{r})) \]
\item 
\[ \phi^\nabla(s_p,s_j,\mathbf{r}) =\exp(\sum_{j=1}^L \sum_{t=1}^T \nu_t h_t (s_p,s_j,\mathbf{r} ))\]
\item Entonces, tenemos que 
\[P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \exp \left( \sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r) \right) \]
\end{itemize}
\section{Aprender los parámetros}
\begin{itemize}
\item Dada una muestra $\mathcal{D}= (\mathbf{x}_i, \mathbf{y}_i)_{i=1}^N$ queremos estimar los parámetros $(\lambda_k,\mu_w,\nu_t)$
\item La log-verosimilitud regularizada es
\[ \mathcal{L} = \log [\Phi(r,s,x)] - \log [Z(x)] - \sum_{k=1}^K \frac{\lambda_k^2}{2 \sigma_\lambda^2} - \sum_{w=1}^W \frac{\mu_w^2}{2 \sigma_w^2} - \sum_{t=1}^T \frac{\nu_t^2}{2 \sigma_\nu^2} \]
\item Donde,
\[ \Phi(r,s,x) = \exp (\sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r)) \]
\item Tenemos que $\mathcal{L}$ es concava por lo que es fácil maximizarla
\end{itemize}
\section{Asignación más probable}
\begin{itemize}
\item El objetivo es encontrar
\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
\item No se puede resolver de manera exacta, se necesitan métodos aproximados
\item Se propone decodificar las variables ocultas objetivo basados en asignar etiquetas a las variables muestreadas.
\item Se hace en dos pasos: Primero, se predice una etiqueta inicial para una $x_i$ dado el modelo ya entrenado
\item En segundo lugar, se re-estima la asignación a $x_i$
\end{itemize}
\end{document}

### 2017-12-20 Primer_Semestre/Resumenes/Resumen_On_learning_causal_models_from_relational_data.tex
primer commit

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Resumen: On Learning Causal Models From Relational Data - Lee, Honavar}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\section{Previo}
\begin{itemize}
\item Se estudian Modelos Relacionales Causales (RCM) a través de sus contrapartes relacionales \textit{adjacenty-faithfulness} y \textit{orientation-faithfulness}, lo cual resulta en un enfoque sencillo para identificar un subconjunto de queries relacionales requeridas para determinar la estructura de un RCM utilizando criterios de d-separación sobre un grafo acíclico no-dirigido “desenrrollado” que representa el RCM
\item Maier et al (2010) introdujeron RPC, que extiende el modelo PC (Spirtes, 2000) al caso de datos relacionales para aprender relaciones causales. 
\item RPC utiliza el modelo DAPER (directed acyclic probabilistic entity-relationship) que a su vez extiende el modelo entity-relationship (Chen 1976) para incorporar dependencias probabilisticas
\item Maier et al (2013) demostraron que RPC no es completo para aprender modelos causales a partir de datos relacionales, e introdujeron Relational Causal Discovery (RCD) como alternativa
\item RCD utiliza un enfoque de restricciones
\item Maier et al introdujeron d-separación relacional, que es la contraparte relacional de la d-separación tradicional sobre grafos
\item Se introduce además la Abstract Ground Graph
\item La prueba del buen funcionamiento de RCD requiere que la AGG represente exactamente todos los vértices que podrían aparecer en todas las instancias el RCM
\item Resulta que existen casos en los cuales la d-separación en la AGG no implica las relaciones de independencia condicional que ocurren en el RCM
\end{itemize}
\section{Notación}
\begin{itemize}
\item Un esquema relacional $\mathcal{S}$ es una tupla $(\mathbf{E},\mathbf{R},\mathbf{A},c)$ formada por un conjunto de  clases de entidad $\mathbf{E}$, un conjunto de clases de relaciones $\mathbf{R}$, atributos $\mathbf{A}$ y una función $c : R \times E \to \{una, muchas \}$.
\item Denotamos como $\mathbf{I} = \mathbf{E} \cup \mathbf{R}$ y como $I_X$ un objeto $I$ con atributo $X$
\item Un esqueleto relacional $\sigma \in \Sigma_{\mathcal{S}}$ es una instanciación del esquema relacional $\mathcal{S}$ y está representado por un grafo de entidades y relaciones, donde $\Sigma_{\mathcal{S}}$ representa todos las posibles instancias de $\mathcal{S}$.
\item Denotamos como $\sigma(I)$ un conjunto de elementos en $\sigma$ de clase $I \in \mathbb{I}$.
\item Un modelo relacional causal (RCM) denotado como $\mathcal{M}$ consiste en un conjunto de relaciones causales $\mathbf{D}$ en el cual causas y efectos están relacionados dado un esquema relacional $\mathcal{S}$
\item Un camino relacional $P$ es una sucesión en la cual se alternan clase de entidad y clase de realcion; $P=[I_j,...,I_k]$. A $I_j$ se le conoce como clase base, y a $I_k$ como clase terminal. Un camino relacional corresponde a una caminata sobre el esquema y muestra cómo la clase terminal está relacionada a la clase base.
\item Una variable relacional $P.X$ consiste en un camino relacional $P$ y una clase de atributos $X$ de la clase terminal de $P$
\item Se dice que una variable relacional es canónica si la longitud de su camino relacional es 1.
\item Una dependencia relacional especifica causa y efecto; es de la forma $[I_j,...,I_k].Y \to [I_j].X$; es decir, causa y efecto comparten la misma clase base, y el efecto es canónico.
\item Ejemplo, “el éxito de un producto depende de las habilidades de los empleados que lo desarrollan” se representa como
\[ [Producto, Desarrollan, Empleados].Habilidades \to [Producto].Exito \]
\item Se dice que un RCM es acíclico si existe un órden parcial sobre las clases de atributos $\mathbf{A}$. Un RCM aciclico no permite dependencias que conectan una clase de atributos con ella misma
\item Un ground graph es una instancia del RCM dado un esqueleto; es decir, se obtiene de interpretar las causas de las dependencias en el RCM sobre el esqueleto al utilizar los conjuntos terminales de cada uno de los elementos del esqueleto. 
\item Dado un esqueleto relacional $\sigma$, el conjunto terminal de un camino relacional $P$ dado un elemnto base $b \in \sigma(P)$, que se denota $P | _b$ está formado por los elementos alcanzables a partir de $b$ cuando se recorre el esqueleto a lo largo de $P$ sin tocar un elemento dos o más veces.
\item Denotamos como $\mathcal{GG}_{\mathcal{M}_\sigma}$ la ground graph de un RCM $\mathcal{M}$
\item En $\mathcal{GG}_{\mathcal{M}_\sigma}$ existe un vértice $i_j . X \to i_j . Y$ si y sólo si existe una dependencia $[I_k,...,I_j].X \to [I_k].Y$ tal que $i_j \in \sigma(I_j)$ es alcanzable desde $i_k \in \sigma(I_k)$ a lo largo del camino relacional $[I_k,...,I_j]$
\end{itemize}
\section{Independencia Condicional en un RCM}
\begin{itemize}
\item Un RCM $\mathcal{M}$ puede entenderse como un modelo meta-causal definido sobre un esquema $\mathcal{S}$. Dado un esqueleto $\sigma$ del esquema, el RCM es instanciado en un ground graph, que corresponde a un grafo acíclico no dirigido
\item Dados los atributos y valores de los elementos y la estructura del esqueleto, las relaciones de independencia condicional que se cumplen en los datos son equivalentes a aquellas relaciones que cumplen la d-separación sobre $\mathcal{GG}_{\mathcal{M}_\sigma}$
\item La contraparte relacional de la d-separación en $\mathcal{M}$ se reduce a la d-separación tradicional sobre todas las instancias de la siguiente manera:
\item Si $\mathbf{U},\mathbf{V},\mathbf{W}$ son conjuntos disjuntos de variables relacionales. Entonces, para un modelo relacional $\mathcal{M}$, $\mathbf{U}$ y $\mathbf{V}$ están relacionalmente d-separadas por $\mathbf{W}$ si y sólo si para cada esqueleto $\sigma$, $U | _b$ y $V | _b$ están d-separados por $W | _b$ en el ground graph para todo $b \in \sigma(B)$.
\end{itemize}
\section{Grafo desenrrollado de un RCM}
\begin{itemize}
\item Dado un RCM $\mathcal{M}=(\mathcal{S},\mathbf{D})$, su grafo desenrrollado es un DAG denotado $\mathcal{G}_\mathcal{M}$ en el cual los vértices son variables relacionales de $\mathcal{S}$ y existe un vértice $P.X \to Q.Y$ si y sólo si existe una dependencia $R.X \to [I_Y].Y \in \mathbf{D}$
\end{itemize}
\section{Fidelidad de la representación}
\begin{itemize}
\item Decimos que una distribución de probabilidad es bien representada (faithfully) por un DAG si las independencias condicionales están bien representadas por el grafo. 
\item Resulta que un RCM no es fiel a sus AGG, pero se cumplen dos nociones más débiles: fidelidad de adjacencia y de orientación
\item Lema 2: si $U,V$ son dos variables relacionales con la misma clase base y son adyacentes en $\mathcal{G}_\mathcal{M}$, entonces son condicionalmente dependientes en cualquier subconjunto de $\mathcal{V}_B - \{U,V\}$, donde $\mathcal{V}_B$ es el conjunto de todas las variables relacionales con clase base $B$
\item Lema: Si $U,V,W$ son variables relacionales distintas (con misma clase base $B$) y se cumple (en $\mathcal{G}$) que $U$ y $W$ están conectadas a $V$, pero no entre sí (unshielded triple), entonces: si $U \to V \leftarrow W$ (unshielded collider), entonces $U$ y $W$ son dependientes dado cualquier subconjunto de variables relacionales que contenga a $V$; o, en caso contrario $U$ y $W$ son dependientes.
\end{itemize}
\section{Aprendiendo un RCM}
La idea es desarrollar un algoritmo que identifique dependencias no-dirigidas y las oriente utilizando los unshielded collider en el DAG. 
\begin{itemize}
\item Sea $D = P.X \to [I_Y].Y$, denotamos al reverso de $P$ como $\tilde{P}$ y $\tilde{D}= \tilde{P}.Y \to [I_X].X$
\item Decimos que una dependencia es no-dirigida si $D$ y $\tilde{D}$ son válidas
\item Se dice que un grafo es parcialmente acíclico dirigido si no existen ciclos dirigidos, pero los vértices pueden ser dirigidos o no.
\item Denotamos $X \prec Y$ si existe un camino dirigido de $X$ a $Y$
\item Proposición 1: Sean $[B].X$ y $Q.Y$ variables relacionales distintas con misma clase base $B$ tales que $Y$ no es descendiente de $X$ (en el orden parcial inducido por $\mathcal{M}$). Entonces, $[B].X$ y $Q.Y$ están relacionalmente d-separados por $Pa([B].X)$ y sólo si $Q.Y \to [B].X$ o $\tilde{Q}.X \to [I_Y].Y$ no pertenece a $\mathcal{M}$. 
\end{itemize}
\subsection{Fase 1: Identificar dependencias no-dirigidas}
Los algoritmos basados en independencias condicionales para aprender modelos causales empiezan enumerando todos los posibles dependencias candidato (Spirtes, Glymour, Scheines 2000). A diferencia el caso proposicional, en el cual el número de variables es fijo y finito, en el caso relacional las variables relacionales son infinitas, por lo que es imposible enumerarlas. (Maier et al. 2013) asume que el numero dependencias en un RCM es finito, y que la longitud del camino más largo en el grafo es conocido.
\begin{itemize}
\item Lema: Sea $D = P.X \to [I_Y].Y$. Entonces, $P.X$ es condicionalmente independiente de $[I_Y].Y$ dados los padres de $[I_Y].Y$ en $\mathcal{M}$ o $\tilde{P}.Y$ es condicionalmente independiente de $[I_X].X$ dados los padres de $[I_X].X$ si y sólo si $D$ y $\tilde{D}$ no están al mismo tiempo en $\mathbb{D}$.
\end{itemize}
\subsection{Fase 2: Orientar las dependencias utilizando independencias condicionales}
Sea $\mathcal{G}_{\tilde{\mathcal{M}}}$ el grafo desenrrollado de $\mathcal{M}=(\mathcal{S},\hat{D})$ con $\hat{\mathbf{D}}$ que se obtiene de la Fase 1. Vamos a utilizar el Lema 2 para orientar las dependencias no dirigidas. El siguiente Lema muestra cómo detectar colliders.
\begin{itemize}
\item Sean $(U,V,W)$ un “unshielded triple” en $\mathcal{G}_{\tilde{\mathcal{M}}}$. Si existe un conjunto separador tal que $U$ y $W$ sean condicionalmente independientes dado $S$ en $\mathcal{M}$ y $V$ no pertenece a $\mathbf{S}$, entonces $U \to V \leftarrow W$ en $\mathcal{G}_\mathcal{M}$.
\end{itemize}
Resulta que $\mathcal{G}_\mathcal{\hat{M}}$ es un grafo infinito, por lo que no podemos aplicar directamente detección de colisionadores. Pero resulta que para cada tripleta unshielded existe una tripleta representativa tal que orientarla sea equivalente a orientar las tripletas en $\mathcal{G}_\mathcal{\hat{M}}$
\begin{itemize}
\item Lema 5: Si $(P'.X,Q'.Y,R'.Z)$ es una tripleta unshielded en $\mathcal{G}_\mathcal{\hat{M}}$, existe una tripleta representativa $([I_X].X,Q.Y,R.Z)$ en $\mathcal{G}_\mathcal{M}$.
\end{itemize}
La existencia de estas tripletas representativas permite darle orientación a las dependencias relacionales en los colisionadores unshielded del RCM sin la necesidad de buscar estas tripletas unshielded sobre las AGG.\\
\\
Ahora, combinando el Lema 5 y la Proposición 1 tenemos que
\begin{itemize}
\item Corolario 1: Sea $([I_X].X, Q.Y,R.Z)$ una tripleta unshielded en $\mathcal{G}_\mathcal{\hat{M}}$ tal que $X$ no precede a $Z$ en el orden parcial del grafo. Entonces, $([I_X].X)$ es condiciomalmente independiente de $R.Z$ dados los padres de $[I_X].X$
\end{itemize}
Como la existencia de un colisionador unshielded $([I_X].X, Q.Y,R.Z)$ imploca la existencia de otro, $([I_Z].Z,\tilde{D}_2.Y,\tilde{R}.X)$ donde $D_2.Z \to [I_Y].Y$ está en $\mathbb{\hat{D}}$, entonces uno puede darle orientación a las dependencias entre $X$ y $Y$ y entre $X$ y $Z$ sin importar el orden entre $X$ y $Z$
\subsection{Fase 3}
La Fase 2 no sólo orienta dependencias que forman colisionadores unshielded, sino que además impone restricciones sobre pares de dependencias que forman tripletas unshielded. Lo que se hace ahora es traducir la información expresada utilizando variables relacionales a información descrita utilizando los atributos. Primero, representamos no-colisionadores unshielded como colisionadores unshielded sobre atributos.\\
\\
Se introduce una \textit{class dependency graph} $\mathcal{G}_{\pi}$ sobre el conjunto de clases de atributos $\mathbf{A}$ en la cual existe un vértice que une $X$ con $Y$ si existe una dependencia entre $X$ y $Y$
\section{Algoritmo}
\begin{itemize}
\item \textbf{Input:} Un esquema relacional $\mathcal{S}$, una distribución $p$, un tamaño máximo de pasos $h$.
\item \textbf{Salida} un modelo relacional causal parcialmente dirigido
\item 1 Inicializar $D$ con las dependencias candidato
\item $d=0$
\item Repetir
	\begin{itemize}
		\item For $D = U \to V \in \mathbf{D}$ do
			\begin{itemize}
				\item Si $U$ y $V$ son condicionalmente independientes dado $S \subset Pa(V)-\{U\}$ entonces remover $\{D,\tilde{D} \}$ de $\mathbf{D}$
			\end{itemize}
		\item $d=d+1$
	\end{itemize}
\item Hasta $| Pa(V) | < d$
\item $\mathcal{N}=\emptyset$, inicializar $\mathcal{G}_\pi$ a partir de $\mathbf{D}$ y $\mathcal{K}$
\item Aplicar reglas de orientación den $\mathcal{G}_\pi$
\item Para  tripletas representativas $([I_X].X,Q.Y,R.Z)$ hacer:
	\begin{itemize}
		\item Si $X \prec Z$, continuar
		\item Si $X - Y$ y $Y - Z$ están orientadas, continuar
		\item Si $(X,Y,Z) \in \mathcal{N}$ o $X \leftarrow Y$ o $Y \to Z$, continuar
		\item Si $[I_X].X$ ind cond dado $S \subseteq Pa([I_X].X$ entonces
			\begin{itemize}
				\item Si $Q.Y \notin S$, entonces orientar $X \to Y$, $Z \to Y$
				\item Si $X=Z$, orientar $Y \to X$
				\item En caso contrario, añadir $(X,Y,X)$ a $\mathcal{N}$
			\end{itemize}
		\item $X \to Z$
		\item Aplicar reglas de orientación
	\end{itemize}
\item Orientar $\mathbf{D}$
\end{itemize}
\bibliographystyle{apalike}
\bibliography{Bibliografia}
\end{document}

### 2017-12-20 Segundo_Semestre/Propuesta/Propuesta.tex
primer commit

--- 2017-12-20 Primer_Semestre/Propuesta/Propuesta.tex
@@
 \subsection{Contrafactuales en modelos funcionales}
-
-\section{Extracción de relaciones causales a partir de texto}
-\subsection{Extracción basada en patrones gramaticales}
-\cite{khoo2000extracting} desarrollan un método que permite identificar y extrar relaciones causa-efecto que se encuentra expresada de manera explícita en \textit{abstracts} de artículos médicos.
-Los autores deciden utilizar el dominio médico pues al ser importantes las relaciones causales es más probable que estas aparezcan explícitamente.
-Los autores primero identifican los \textit{marcadores lingüísticos} y con esto construyen un conjunto de \textit{patrones causales}.
-Además, cada oración es pasada por un \textit{parse tree} el cual genera una representación de la estructura sintáctica de la oración.\\
-\\
-Como ejemplo, consideremos la oración \textit{“Paclitaxel was well tolerated and resulted in a significant clinical response in this patient”}.
-Después de la creación del árbol sintáctico y del patrón gramatical, se obtiene que la causa presente en la oración es \textit{paclitaxel} y el efecto es \textit{a significant clinical response in this patient}.\\
-\\
-Analizando 200 abstracts que provienen de 4 áreas médicas (depresión, esquizofrenía, enfermedades cardiacas y SIDA), construyen los patrones causales.
-El proceso de extracción de información consiste en hacer un \textit{match} entre los patrones causales y el arbol sintáctico.
-Al aplicar los patrones causales obtenidos a un nuevo conjunto de abstracts, obtienen un 51 por ciento de accuracy al extraer causas y un 58 por ciento de accuracy al extraer efectos.\\
-\\
-\cite{girju2002text} presentan un método semi-automático para la detección de patrones causales en texto, es semi-automático pues los patrones son descubiertos de manera automática, pero la validación no lo es.
-El algoritmo de detección de patrones léxico-sintácticos que  se refieren a causalidad consiste en dos partes.
-La primera parte descubre patrones léxico sintácticos que pueden expresar relaciones causales, y la segunda valida estos patrones según restricciones semánticas.
-Se enfocan en encontrar patrones de la forma (frase verbo causal frase) donde el verbo es un verbo causal simple en los cuales el verbo se refiere únicamente a la causa; por ejemplo, \textit{Temblores generan tsunamis}.
-Como las frases obtenidas pueden ser muy largas, pueden aportar información no deseada.
-Para resolver esto \cite{girju2003automatic} utiliza WordNet para reducir el tamaño de las frases obtenidas.\\
-\\
-\cite{sanchez2004acquiring} desarrollan un sistema para la adquisición de conocimiento causal a partir de texto.
-El sistema propuesto identifica oraciones que especifican de manera explícita relaciones causales y extrae de ellas los patrones causales.
-El sistema incorpora información gramatical tal como conectores, conjunciones, disjunciones y negaciones y codifican todo esto en redes bayesianas.\\
-\\
-Para la detección de patrones causales, los autores se enfocan en trabajar con oraciones en las cuales ambos \textit{eventos} aparezcan como frases (noun phrases, ¿traducción?).
-Un ejemplo de patrones causales identificados por el sistema es \textit{“Anemia are caused by excessive hemolysis”}.
-\\
-Para la construcción de las redes, se toma como input un texto \textit{tokenizado}, etiquetado con \textit{part-of-speech tagging} y \textit{parsed} para encontrar expresiones verbales.
-Posteriormente, el sistema estima las probabilidades condicionales de los efectos dadas las causas mediante el conteo de frecuencias de eventos similares.\\
-\\
-Para la evaluación, hicieron una evaluación subjetiva que consistió en comparar la estructura de la red generada por el sistema y la estructura creada manualmente al observar los patrones causales en el texto.
-Se analizaron 7 textos y en general se encontró un empate entre las redes con una precisión de 60 por ciento, pero sin ningún valor de Recall, por lo que no es fácil interpretar sus resultados.
-Según \cite{bojduj2009extraction}, una de las razones por las cuales se ubtuvo baja precisión es que basar los nodos en las frecuencias no es una buena idea, pues la cantidad de texto disponible en general es tan grande que estas frecuencias no pueden ser representativas.\\
-\\
-\cite{bojduj2009extraction} menciona que muchos de estos métodos están más enfocados en cómo extraer las relaciones que en cómo usarlas.
-Es importante no sólo poder concluir que dos términos están relacionados causamente, sino cómo.
-\subsection{Extracción basada en patrones de co-ocurrencias}
-En vez de utilizar patrones gramaticales, \cite{saito2007extraction} y \cite{perrin2008global} analizan la co-ocurrencia de términos estadísticos para construir redes a partir de texto.
-Ellos definen un “statistical term” como \textit{“Statistical terms are expression of the measurements of statistics to watch the movements of phenomena; birth rates, public approval rating of the Cabinet and so on.”}; es decir, un término que sea importante a un sistema.
-El enfoque de su trabajo es construir redes causales entre los statistical terms.
-Lo que hacen es primero extraer relaciones de co-ocurrencia entre términos estadísticos y posteriormente extraer relaciones causales entre ellos.
-Los autores se enfocan en texto japonés, el cual cumple que tiene ciertos patrones en los sufijos utilizados, y estos patrones son utilizados al extraer los términos estadísticos.
-Luegom estos términos son clasificados según su nivel de abstracción.
-Posteriormente, construyen una red de términos estadísticos al considerar a dos de ellos co-ocurrentes si aparecen juntos en un párrafo.
-Se remueven los términos que son co-ocurrentes con muchos otros términos.\\
-\\
-Los autores evaluan de manera separada cada una de las etapas: extracción de términos, clasificación de términos, la construcción de redes y la clasificación de redes.
-\\
-La mayor desventaja de esto es que la dirección de la relación causal no puede ser inferida de la co-ocurrencia.\\
-\\
-Por otro lado, \cite{sakai2008cause} extrajeron términos de artículos financieros (en japonés) y luego considerarón que estos términos afectaban el desempeño de los negocios basados en co-ocurrencias con expresiones clave como “es bueno”.
-Su definición de causalidad depende de las expresiones clave.
-\subsection{Modelos Gráficos para extracción de relaciones}
-A partir de un artículo que trata específicamente sobre una entidad claramente indentificable (la biografía de Benito Juarez por ejemplo) \cite{yu2010jointly} intentan identificar entidades y descubrir relaciones semánticas entre ellas.
-Resulta evidente que en el documento aparecerán otras entidades, pero se hace el supuesto de que no están relacionadas entre sí.
-En este trabajo, todas las clases entidades están pre-definidas así como las relaciones; esto es, se sabe que en el documento aparecerán años, lugares, países y que una entidad de tipo persona habrá nacido en un país en cierto año.
-Para formular el problema, se tiene $\mathbf{x} = \{ x_1,...,x_N \}$ una sucesión observada de \textit{tokens}.
-Denotamos como $s_p$ la entidad principal.
-Sea $ \mathbf{s}=\{ s_1,...,s_L\}$ una segmentación en la cual cada $s_i$ es una tripleta $\{ \alpha_i, \beta_i,y_i \}$ con $\alpha_i$ es una posición inicial, $\beta_i$ posición final y $y_i$ una etiqueta asignada.\\
-\\
-Sea $r_{pn}$ la asignación de relaciones entre la principal $s_p$ y un candidato a entidad secundaria $s_n$ y sea $\mathbf{r}$ el conjunto de asignación de relaciones para la $\mathbf{x}$.
-Dada una observación $\mathbf{x}$, queremos encontrar $\mathbf{y}^\ast$ tal que
-\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
-donde $\mathbf{y}=\{ \mathbf{r},\mathbf{s} \}$.\\
-\\
-Se define una distribución condicional conjunta para $\mathbf{s},\mathbf{x},\mathbf{r}$ en un modelo grafico no dirigido $\mathcal{G}$ tal que se pueden dividir los factores de $\mathcal{G}$ en tres grupos: $\phi^S$, $\phi^R$, $\phi^\nabla$, que son el potencial de segmentación, el potencial de relaciones, y el potencial conjunto de segmentación-relación.
-\\
-\\
-Por el Teorema de Hammersley-Clifford, la condicional conjunta $P(\mathbf{y} | \mathbf{x})$ se factoriza como
-\[ P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \prod_{C_S} \phi^S (i,\mathbf{s},\mathbf{x}) \prod_{C_R}\phi^R(r_{pm},r_{pn},\mathbf{r}) \prod_{C_\nabla} \phi^\nabla(s_p,s_j,\mathbf{r}) \]
-Se hace el supuesto de que las funciones potencial se factorizan según un conjunto de atributos y un conjunto de pesos correspondientes.\\
-\\
-Incorporando estos features, tenemos que 
-\[P(\mathbf{y} | \mathbf{x}) = \frac{1}{Z(\mathbf{x})} \exp \left( \sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r) \right) \]
-\\
-Dada una muestra $\mathcal{D}= (\mathbf{x}_i, \mathbf{y}_i)_{i=1}^N$ queremos estimar los parámetros $(\lambda_k,\mu_w,\nu_t)$.\\
-\\
-La log-verosimilitud regularizada es
-\[ \mathcal{L} = \log [\Phi(r,s,x)] - \log [Z(x)] - \sum_{k=1}^K \frac{\lambda_k^2}{2 \sigma_\lambda^2} - \sum_{w=1}^W \frac{\mu_w^2}{2 \sigma_w^2} - \sum_{t=1}^T \frac{\nu_t^2}{2 \sigma_\nu^2} \]
-Donde,
-\[ \Phi(r,s,x) = \exp (\sum_{i=1}^{|s|} \sum_{k=1}^K \lambda_k g_k(i,s,x) + \sum_{m,n}^M \sum_{w=1}^W \mu_w q_w (r_{pm},r_{pn},r) + \sum_{j=1}^L \sum_{t=1}^T  \nu_t h_t(s_p,s_j,r)) \]
-Tenemos que $\mathcal{L}$ es concava por lo que es fácil maximizarla.\\
-\\
-Para calcular la Asignacion Más Probable, hayue encontrar
-\[ \mathbf{y}^\ast = \textrm{ arg max }_y p(\mathbf{y}| \mathbf{x})\]
-La cual no se puede resolver de manera exacta, se necesitan métodos aproximados.
-El algoritmo que se propone en el artículo se llama Collective Iterative Classification y la idea detrás de este es decodificar las variables ocultas objetivo basados en asignar etiquetas a las variables muestreadas.
-Esto se hace mediante un proceso iterativo de dos pasos; primero, en una etapa de bootstrapping se predice una etiqueta inicial para una $x_i$ dado el modelo ya entrenado.
-Luego, en la segunda etapa, conocida como clasificación iterativa, se re-estima la asignación a $x_i$ varias veces, tomándolas a partir de la asignación inicial.\\
-\\
-Los autores prueban su algoritmo con datos provenientes de Wikipedia: sus datos consisten en 441 páginas de Wikipedia.
-De estos datos, fueron etiquetados a mano 7740 entidades en 8 categorías.
-Además, se extrajeron 4700 relaciones de las cuales las más frecuentes fueron trabajo, visió, nació, miembro de, nació en día.\\
-\\
-El modelo superó los existentes tomando como medida de calidad la medida F. 
-\section{Métodos Estadísticos Para Extracción de Relaciones Causales}
-\subsection{Statistical Relational Learning}
-En el contexto de inferencia causal, se estudian las interacciones entre variables que puedan causar efectos en la otra.
-Para esto, es necesario identificar los objetos y las relaciones entre ellos.
-Para esto, el área de Statistical Relational Learning ofrece una serie de métodos que modelan dependencias entre entidades relacionales.
-En esta sección veremos distintas técnicas existentes en esta área y sus aplicaciones a inferencia causal.\\
-\\
-Un dominio de interés consiste en objetos (entidades), sus atributos, y las relaciones entre sí de los objetos.
-El modelo DAPER de \cite{heckerman2004probabilistic} consiste en clases de entidades, clases de relaciones, clases de atributos y clases arco.
-\subsubsection{Aplicación de Campos Aleatorios Condicionales a Extraccion de informacion}
-\cite{ramani2005consolidating} analizaron 753,459 abstracts de journals médicos, de los cuales extrajeron 6580 interacciones entre 3,737 proteínas.
-Para ello, utilizaron un algoritmo de tres etapas: la primera etapa, consiste en identificar nombres de proteínas mediante un clasificador basado en Conditional Random Fields; la segunda, identificar interacciones a través de co-ocurrencias; posteriormente, filtrar estas interacciones con un clasificador Bayesiano para obtener interacciones válidas.
-Con esto, se obtiene una red que consiste en 31609 interacciones entre 7748 proteínas\\
-\\
-Para la etapa de identificación de nombres, se utilizó un Conditional Random field, cuyo desempeño fue evaluado en un conjunto de 200 abstracts médicos etiquetados y en 750 abstracts etiquetados a mano.
-Para extraer los nombres de proteínas, se utilizó este clasificador en el conjunto de 753,459 abstracts de Medline que contuvieran la palabra “human”.\\
-\\
-Para obtener las interacciones, se midió la co-citación entre los nombres de proteínas y luego se enriquecieron estos pares con interacciones físicas (reales) entre proteínas utilizando un filtro Bayesiano.
-Primero, se contó el número de veces que en un abstract aparece un par de proteínas, y con eso se calculó la probabilidad de co-citación.
-Con esto se obtienen 15,000 interacciones, pero se pueden dar co-citaciones por otras razones además de las relaciones reales, se aplicó un filtro bayesiano que mide la verosimilitud de los abstracts que citan un par de proteínas para discutir sus interacciones reales.
-El clasificador asigna un \textit{score} a cada uno de los abstracts que citan esas proteínas.
-Utilizando un umbral de score, se obtienen 6,580 interacciones entre 3,737 proteínas.\\
-\\
-Al combinar estas interacciones con otras 26,280 interacciones provenientes de otras fuentes, se obtiene el total mencionado.\\
-\\
-Por otro lado, \cite{culotta2006integrating} plantean el problema de extracción de relaciones como un problema de etiquetado de secuencias, algo similar a lo realizado por \cite{yu2010jointly}.
-\cite{culotta2006integrating} definen un modelo de la forma
-\[ P_{\Lambda} ( y | x ) =\frac{1}{Z} \prod_{c \in C} \varphi_c ( y_c , x_c, \Lambda ),\]
-donde las $\varphi$ son potenciales parametrizados por $\Lambda$.\\
-\\
-Los autores tomaron 1,127 párrafos de 271 artículos biográficos de Wikipedia y etiquetaron 4,701 instancias de relaciones.
-Para identificar a las entidades importantes dentro de un artículo, utilizaron el hecho de que estas suelen tener un \textit{hyperlink}
-\subsubsection{Aplicación de RMN's a Extraccion de informacion}
-\cite{bunescu2007statistical} utilizan Relational Markov Networks para mejorar el desempeño de los CRF's en el problema de identificar nombres de proteínas en abstracts de journals médicos. \\
-\\
-La técnica que se propone es que dada una colección de documentos $D$, asociemos a dada documento $d \in D$ un conjunto de entidades $d.E$.
-Cada entidad$e \in d.E$ queda caracterizada por un conjunto de atributos booleanos $e.F$.
-A cada documento se le asocia un grafo de factores, el cual es un grafo bipartita que contiene dos tipos de nodos: nodos variables, que corresponden a las etiquetas de todas las entidades candidato en el documento y nodos potenciales, que modelan las correlaciones entre dos o más atributos.
-Para cada correlación posible,  un nodo potencial es crreado tal que está ligado a todos los nodos variables involucrados.\\
-\\
-Prueban su algoritmo en dos conjuntos de datos previamente etiquetados a mano.
-El primer conjunto consiste en 200 abstracts y el segundo en 225.
-Ambos conjuntos de datos fueron previamente etiquetados a mano, además de etiquetados para POS.
-En ambos conjuntos de datos, existe una mejora ligera respecto a los CRF's en términos de la medida F.
-
-\subsubsection{Aprendizaje de Modelos Causales a partir de Datos Relacionales}
-\cite{maier2013reasoning} definen la \textit{abstract ground graph} (AGG) y se desarrolla la noción de d-separación relacional.
-A partir de estas nociones teóricas, \cite{maier2013sound} extienden el algoritmo PC, el cual identifica todas las posibles orientaciones de dependencias causales, a datos relacionales.
-Utilizando la d-separación relacional, introducen restricciones que orienta dependencias bivariadas hasta en un 72$\%$ más.
-Se prueba teóricamente que con esta nueva regla, además de extensiones a datos relacionales del algoritmo PC, se obtiene un método \textit{sound and complete} para la extracción de relaciones causales.
-Este algoritmo nuevo se llama RCD (relational causal discovery).\\
-\\
-\cite{lee2016learning} notan que la demostración de \cite{maier2013reasoning} requiere que la AGG sea un DAG quie represente exactamente todas las flechas que aaparecerían en todas las posibles ground graphs, y en un trabajo anterior (\cite{lee2015lifted}) habían notado que existen casos en los cuales la d-separación de la AGG no captura las independencias condicionales que se cumplen en el modelo relacional causal.
-En su artículo, proponen un algoritmo basado en condiciones más débiles llamado RCD-light.
-
-\subsection{Inferencia Bayesiana No Paramétrica}
-La rama no-paramétrica de la inferencia bayesiana puede considerarse como una extensión a espacios infinito-dimensionales de la inferencia bayesiana clásica.
-Esto permite tener modelos que permiten capturar una complejidad mucho mayor en los datos.
-El objeto principal de la inferencia bayesiana no paramétrica es el Proceso Dirichlet, que es una extensión a espacios infinito-dimensionales de la distribución Dirichlet clásica.
-\subsubsection{Proceso Dirichlet}
-\begin{defi}
-Sea $H$ una medida finita definida sobre un espacio $\mathcal{X}$ y $\alpha \in \mathbb{R}$.
-Una medida aleatoria $G$ sobre $\mathcal{X}$ se dice que es un Proceso Dirichlet de parámetros $(\alpha,H)$ si para cada partición finita medible $\{ A_1 , ..., A_k \}$ de $\mathcal{X}$ se tiene que la distribución conjunta de $(G(A_1),...,G(A_k))$ es Dirichlet de parámetros $(\alpha H(A_1),...,\alpha H(B_k))$.
-A la medida $H$ se le conoce como medida base y a $\alpha$ se le conoce como parámetro de concentración.
-Se denota como
-\[ G \sim \textrm{ DP}(\alpha,H).\]
-\end{defi}
-Los parámetors $H$ y $\alpha$ juegan roles intuitivos en la definición.
-La medida base se puede entender, básicamente, como la \textit{media} del Proceso; es decir, para cada conjunto medible $A \subseteq \mathcal{X}$ se tiene que $\mathbb{E}[G(A)]=H(A)$.
-Por otro lado, el parámetro de concentración es un tipo de \textit{varianza inversa}, pues $V[G(A)]=H(A)(1-H(A))/(\alpha+1)$.
-Es decir, mientras más grande sea $\alpha$, entonces menor es la varianza y el DP tenderá a concentrarse más en torno a $H$.\\
-\subsubsection{Distribución Predictiva y urnas de Blackwell-McQueen}
-Consideremos $G \sim \textrm{ DP}(\alpha,H)$ y $\theta_1,... \sim G$. Como $\theta_{n+1} | G, \theta_1,...,\theta_n \sim G$, entonces para $A \subseteq \mathcal{X}$ tenemos que
-\begin{eqnarray*}
-P(\theta_{n+1} \in A | \theta_1,...,\theta_n ) &=& \mathbb{E}[G(A) | \theta_1,...,\theta_n]\\
-                                                                    &=& \frac{1}{\alpha + n} \left(\alpha H(A) + \sum_{i=1}^n \delta_{\theta_i}(A) \right)
-\end{eqnarray*}
-Por lo tanto, marginalizando $G$, vemos que la distribución posterior base dados $\theta_1,...,\theta_n$ es la distribución predictiva de $\theta_{n+1}$.
-\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum^n \delta_{\theta_i} \right) \]
-La secuencia de distribuciones predictivas para $\theta_1, \theta_2,...$ se conoce como el esquema de urnas de Blackwell-McQueen; la interpretación es la siguiente: cada valor en $\mathcal{X}$ es un color, y cada realización $\theta \sim G$ son pelotas y el valor obtenido es el color de la pelota.
-Además, tenemos una urna en la cual se han colocado las bolas observadas.
-Al principio, no existen bolas en la urna y muestreamos un color de $H$; es decir, $\theta_1 \sim H$, pintamos la bola con ese color y la echamos en la urna.
-Luego, en el paso $n+1$, tomaremos un nuevo color con probabilidad $\alpha / (\alpha + n)$ o tomaremos un nuevo color, la pintaremos así y la echaremos en la urna con probabilidad $n / (n+\alpha)$.
-\subsubsection{Proceso Dirichlet como Prior en problemas de Clustering}
-La propiedad de que $G \sim DP$ sea discreto ofrece aplicaciones de clusterización.
-Por el momento, supongamos que $H$ es suave.
-Entonces, como los valores de las realizaciones son repetidos, sean $\theta^\ast_1,...,\theta^\ast_m$ los valores únicos de las realizaciones $\theta_1,...,\theta_n$ y sea $n_k$ el número de repeticiones de $\theta^\ast_k$.
-Entonces, la distribución predictiva se puede escribir como
-\[ \theta_{n+1} | \theta_1,...,\theta_n \sim \frac{1}{\alpha + n} \left( \alpha H + \sum_{k=1}^m n_k \delta_{\theta^\ast_k} \right). \]
-Esto es, el valor de $\theta^\ast_k$ aparecerá en $\theta_{n+1}$ con una probabilidad proporcional al número de veces que ha aparecido.
-Entonces, mientras más grade sea $n_k$, mayor será la probabilidad de que se vuelva aun mayor.
-Notemos que los valores únicos de las realizaciones $\theta_1,...,\theta_n$ inducen una partición aleatoria del conjunto $[n]=\{1,...,n\}$ si pensamos el cluster $k$ como aquel que contiene las $\theta_i$ que toman el valor $\theta^\ast_k$, tenemos que clusters grandes crecen más rápido.\\
-\\
-Podemos invertir este proceso y empezar con una distribución sobre particiones.
-\subsubsection{Dirichlet Enhanced Relational Learning}
-\cite{xu2005dirichlet} aplican un modelo jerárquico Bayesiano no-paramétrico para la extracción de relaciones.
-La ventaja de usar un modelo jerárquico es que los parámetros pueden estar \textit{personalizados} para entidades.
-En este trabajo se aplica el modelo al contexto médico, en el cual se tienen como entidades hospitales, pacientes, diagnósticos y procedimientos.
-Se sabe que la existencia de un diagnóstico o un procedimiento es dependiente de las características de un paciente; por otro lado, las características de los hospitales se modelan como inciertas.\\
-\\
-Un modelo Bayesiano consiste en un parámetro $\theta$ y una asignación inicial de probabilidad sobre este, la cual refleja nuestra incertidumbre inicial.
-Supondremos que esta distribución inicial cuenta con hiperparámetros $h$, de modo que sea $P(\theta | h)$.
-La verosimilitud de los datos $P(D|\theta)$ refleja cómo son generados estos dado el parámetro.
-EN el caso en el cual se asume que los datos provienen de distintos grupos es natural pensar en asignar un parámetro por grupo, lo cual lleva a la definición de modelos jerárquicos, en los cuales el modelo tiene la siguiente fforma
-\[ P(h) \prod P(\theta_i | h )P(D_i | theta_i). \]
-Si se aplicara el modelo DAPER al contexto médico, se tiene que los parámetros e hiperparámetros que especifican distribuciones condicionales como atributos globales.
-Esto implica que las probabilidades de un procedimiento son idénticas para todos los pacientes con la misma dolencia inicial.
-Luego, los procedimientos son modelados de manera independiente de manera que procedimientos anteriores en una persona no afectlan la selección de uno nuevo.
-Estas son suposiciones no realistas.\\
-\\
-Para mejorar esta situación, se asume que la probabilidad de un procedimiento es una distribución multinomial con un prior Dirichlet.
-En particular, para cada paciente, se asume que un vector individual de parámetros $\theta_{s|pc,pa}$ especifica la probabilidad de un procedimiento para un paciente $pa$ con dolencia inicial $pc$.
-Este vector es generado por una distribución Dirichlet de parámetros $h_{pc}=\{ \tau_{pc},\alpha_{pc} \}$, la cual puede escribirse como
-\[ \textrm{Dir}(\theta_{\cdot | pc, pa} | \tau_{pc},\alpha_{pc}) = \frac{1}{C} \prod_{k=1}^K \theta^{\tau_{pc} \alpha_{k,pc} - 1}_{k,s | pc, pa}.  \]
-En un modelo jerárquico bayesiano cada paciente obtiene sus propias probabilidades de procedimiento y comparte con los demás un parámetro a priori.\\
-\\
-En más de un caso, se necesita asumir que la prior tiene una forma muy flexible, por lo que es necesario recurrir a priors no paramétricas.\\
-\\
-Podemos tomar la distribución a priori como una realización del Proceso Dirichlet; es decir, 
-\[ G_{pc} \sim \textrm{ DP }(G_0,\alpha_0) \]
-donde $G_0$ es la medida base y $\alpha_0$ el parámetro de concentración, el cual especifica nuestro nivel de certidumbre sobre la prior.\\
-\\
-Luego, el parámetro $\theta_{\cdot | pc, pa}$ es una realización de $G_{pc}$.\\
-\\
-Otra manera de escribir el Proceso Dirichlet es mediante su representación \textit{stick breaking} (\cite{teh2011dirichlet}) según la cual
-\[ G_{pc} = \sum \pi_{l,pc} \delta_{\theta^\ast_{l,pc}} \textrm{ ;  } \theta^\ast_{l,pc} \sim G_0  \]
-\[ \pi'_{l,pc} \sim \textrm{ Beta }(1,\alpha_0) \textrm{  ;  } \pi_{l,pc} = \pi'_{l,pc} \prod_{k=1}^{l-1} (1 - \pi'_{k,pc}).\]
-Notemos que aunque la distribución base pueda ser continua, el DP es siempre discreto (con probabilidad 1).\\
-\\
-Tradicionalmente, el aprendizaje en el contexto Bayesiano no paramétrico se lleva a cabo via muestreo de Gibbs, cuyas variaciones más comunes son la urna de Polya o el Proceso del Restaurante Chino (\cite{teh2011dirichlet}).
-En el artículo, se lleva a cabo un proceso más eficiente (\cite{Yu:2004:NHB:1008992.1009053}.)\\
-\\
-El objetivo es estimar $G_{pc}$ para cada posible $pc$ en la base de datos, utilizando la verosimilitud marginal:
-\[ \hat{G}_{pc} = \textrm{ argmax DP }(G | G_0, \alpha_0) \prod_{\{pa\}_{pc}} \int \textrm{ Multinomial }(\{pr\}_{pa}| \theta_{\cdot | pc,pa}) d\theta_{\cdot | pc,pa}  \]
-donde $\{pa\}_{pc}$ es el conjunto de pacientes tales que tienen la misma dolencia inicial y $\{pr\}_{pa}$ es el conjunto de procedimientos para el paciente $pa$.
-Resulta imposible calcular la verosimilitud, por lo que se realiza una aproximación de campo medio:
-\[ G_{pc} \approx \sum^{|{pa}_{pc}|} \pi_{pa,pc} \delta_{\theta^\ast_{\cdot | pa,pc}}. \]
-El proceso de aprendizaje se divide en dos etapas: 
-\begin{itemize}
-\item Calcular la ubicación de $ \theta^\ast_{\cdot | pa,pc}$
-\item Estimar los pesos $\pi_{pa,pc}$
-\end{itemize}
-Para el primer paso, se utiliza la aproximación MAP de $\theta_{\cdot | pa,pc}$ que es:
-\[ \theta^{MAP}_{\cdot | pa,pc} = \textrm{ arg max }P( \theta_{\cdot | pa,pc} | \{pr\}_{pa})\]
-Y se utiliza como prior la Dirichlet mencionada antes. \\
-\\
-Para el segundo paso, es necesario hacer la siguiente suposición:
-\[ \hat{P}( \theta_{\cdot | pa,pc} | \{pr\}_{pa}) \approx q_{pa}(\theta_{\cdot | pc,pa}) = \sum_{\{\tilde{pa}\}_{pc}} \xi_{\tilde{pa},pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}} \]
-donde $\xi_{\tilde{pa},pc,pa} \geq 0 $ son parámetros variacionales tales que $\sum_{pa} \xi_{\tilde{pa},pc,pa} = 1$.\\
-\\
-Como paso E se obtiene que:
-\[  \xi^t_{\tilde{pa},pc,pa} = \frac{P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}}) \hat{G}^{(t)}_{pc}(\delta_{\theta^{MAP}_{\cdot | pc,pa}}))}{\sum_{\tilde{pa}}P(\{pr\}_{\tilde{pa}} | \delta_{\theta^{MAP}_{\cdot | pc,pa}})\hat{G}^{(t)}_{pc}(\theta^{MAP}_{\cdot | pc,pa}))}.
-\]
-Y el paso M queda como
-\[ \hat{G}^{(t+1)}_{pc}(\theta_{\cdot,pc,pa}) = \frac{\alpha_0 G_0(\theta_{\cdot,pc,pa}) + \sum_{\{pa\}_{pc}} \xi_{pc,pa} \delta_{\theta^{MAP}_{\cdot | pc,pa}}}{\alpha_0 + |\{pa\}_{pc}|} \]
-
-\subsubsection{Learning Systems of Concepts with an Infinite Relational Model}
-Suponiendo que tenemos una o más relaciones que involucran uno o más tipos, \cite{kemp2006learning} definen el Infinite Relational Model, el cual busca partir cada tipo (type) en clusters, de manera que una buena partición permita que relaciones entre entidades se puedan predecir a partir de la asignación de clusters.
-Por ejemplo, si tenemos $m$ relaciones que involucran $n$ tipos, denotamos $R^i$ la $i$-ésima relación, $T^j$ la $j$-ésima relación y $z^j$ un vector de asignación de clusters.
-Lo que se busca es inferir la asignación de clusters, y ultimadamente la posterior $P(z^1,...,z^n | R^1, ...
-, R^m)$.
-Para especificar esta distribución, se define un modelo generativo para las relaciones y la asignación de clusters
-\[ P(R^1,...,R^m, z^1,...,z^n) = \prod_{i=1}^m P(R^i | z^1 , ... , z^n) \prod_{j=1}^n P(z^j). \]
-Estamos asumiendo que las relaciones son condicionalmente independientes dada la asignación de clusterss, y que la asignación de clusters es independiente entre tipos.
-Para que el IRM sea capaz de descubrir el número de clusters en el tipo $T$, se utiliza un prior que asigna probabilidad a todas las posibles particiones de $T$.
-Un prior razonable debería descubrir sólo tantos clusters como sugeridos por los datos.
-A partir de trabajos previos en Bayesiana No paramétrica (\cite{rasmussen2000infinite}) utiliza una distribución sobre particiones inducida por un CRP  (Chinese Restaurant Process, \cite{teh2011dirichlet}).\\
-\\
-La intuición detrás de este proceso, según describe \cite{teh2011dirichlet} es que en un restaurante chino hay un número infinito de mesas, cada una de las cuales puede sentar un número infinito de clientes.
-El primer cliente entra y se sienta en la primera mesa, el segundo cliente entra y decide si sentarse en la primera mesa o en otra; en general, el $n$-ésimo cliente se sienta en una mesa ocupada con una probabilidad proporcional al número de clientes sentados ahí.\\
-\\
-Entonces, la manera de constuir clusters utilizando el CRP es partir de un cluster con un solo objeto, e ir añadiendo objetos de modo que la probabilidad de caer en cierto cluster sea proporcional a los elementos en él.
-De esta manera, la distribución de clusters para el objeto $i$, dada la asignaclión de los objetos $1,...,i-1$ es:
-\[
-  P(z_i = a | z_1, ... , z_{i-1}) =
-  \begin{cases}
-                                   \frac{n_a}{i-1+\gamma} & \text{if $n_a > 0$} \\
-                                   \frac{\gamma}{i-1+\gamma} & \text{$a$ es un nuevo cluster} 
-  \end{cases}
-\]
-Donde $n_a$ es el número de objetos asignado al cluster $a$ y $\gamma$ es un parámetro.
-La distribución en $z$ inducida por el CRP es intercambiable; es decir, no importa el orden en que los objetos llegan a los clusters.
-Como nuevos objetos siempre pueden ser asignados, el IRM efectivamente tiene acceso a un número infinito de clusters, de ahí el nombre del Infinite Relational Model.\\
-\\
-Para generar Relaciones a partir de las asignaciones de clusters se asume que las relaciones son binarias, pero se puede extender a datos continuos y frecuencias.
-Como ejemplo, veremos primero el caso con un solo tipo $T$ y una relación $R: T \times T \to \{0,1 \}$.
-$T$ puede ser, por ejemplo, un conjunto de personas y $R$ una relación de amistad.
-El modelo generativo para este problema es
-\begin{eqnarray*}
-z | \gamma &\sim& CRP(\gamma)\\
-\eta(a,b) | \beta &\sim& Beta(\beta,\beta)\\
-R(i,j) | z, \eta &\sim& Bernoulli(\eta(z_i,z_j))
-\end{eqnarray*}
-Esto es, estamos suponiendo que la relación $R$ está generada a partir de dos estructuras latentes: una partición $z$ y una matriz de parámetros $\eta$.
-La entrada $R(i,j)$ se genera al lanzar una moneda con sesgo $\eta(z_i,z_j)$ donde $z_i,z_j$ son las asignaciones de cluster de las entidades $i$,$j$.
-El parámetro $\eta(a,b)$ especifica la probabilidad de que exista un link entre las entidades $i \in a$, $j \in b$.
-El IRM invierte ese modelo para descubrir la $z$ y la $\eta$ que mejor explican $R$.\\
-\\
-Estamos asumiendo que la tendencia de una entidad a participar en relaciones está dada por su asignación de cluster.\\
-\\
-Para hacer inferencia, notamos que en el ejemplo anterior se utilizó una prior conjugada, es sencillo calcular $P(R|z)$
-\[ P(R|z)= \prod_{a,b} \frac{Beta(m(a,b) + \beta, \bar{m}(a,b)+\beta}{Beta(\beta,\beta)} \]
-donde $m(a,b)$ es el número de pares $(i,j)$ tales que $i \in a$, $j \in b$ y $R(i,j)=1$ mientras que $\bar{m}(a,b)$ son las parejas donde $R(i,j)=0.$\\
-\\
-Para muestrear de la posterior $P(z | R ) \propto P(R|z)P(z)$ se pueden utilizar técnicas MCMC o al buscar la moda de la distribución.\\
-\\
-Para la evaluación del modelo, los autores generan datos sintéticos.
-Consideran conjuntos de datos de tres formas distintas; el primer sistema, $S1$, tiene sólo dos tipos $T1, T2$ y una sola relación $R:T1 \times T2 \to \{0,1 \}$.
-El sistema $S2$ consiste en 4 tipos y tres relaciones binarias mientras que $S3$ consiste de 3 tipos y una sola relación ternaria.
-Resulta que en este contexto, el modelo captura el número de clusters de manera precisa.
-\\
-\\
-Este método es excelente para clusterizar de manera no paramétrica un conjunto de datos relacionales, pero requiere identificar previamente las entidades y las relaciones. 
-\subsubsection{Learning Infinite Hidden Relational Models}
-\cite{xu2006learning} extienden la expresividad de los modelos relacionales al introducir para cada entidad una variable latente con un número infinito de estados que proviene de un Dirichlet Process Mixture.
-Debido a que el número de features de las cuales un atributo puede depender, resulta más conveniente  introducir para cada entidad una variable latente que es padre de los atributos de la entidad y de los atributos de las relaciones en las que participa la entidad; a esto se le conoce como hidden relationship model.\\
-\\
-Como cada entidad puede tener un distinto número de estados en sus variables latentes, es deseable que este número de estados latentes se determine de manera automática según los datos; esto es posible utilizando una mezcla de Proceso Dirichlet, que son modelos de mezcla con un número infinito de componentes de mezcla, pero que según los datos la complejidad se ajusta de manera automática.
-A la combinación de un hidden relational model con un DPm se le conoce como Infinite Hidden Relationship Model que generaliza el modelado jerárquico bayesiano noparamétrico al caso de modelos relacionales.\\
-\\
-Como aplicación, los autores evalúan su modelo en un conjunto de datos sobre películas y buscan predecir las preferencias de los usuarios.
-Para este conjuto de datos, las clases de entidades son Usuario y Película
-\subsection{Modelos Jerárquicos Bayesianos}
-Los Modelos Jerárquicos Bayesianos (HBM) son modelos probabilistas que están definidos por \textit{niveles}, en los cuales cada nivel representa un mayor grado de abstracción.
-Un ejemplo sencillo de esto puede ser un modelo de regresión de ingreso a nivel nacional en el cual existan parámetros por estado y por municipio.\\
-\\
-En el contexto causal, lo que está detrás de un HBM es que al aprender causalidad se incorpora conocimiento teórico previo y datos observables (\cite{griffiths2005structure}, \cite{griffiths2009theory}).
-Esto es, un HBM es una distribución de probabilidad entre supuestos teóricos, modelos causales y datos (\cite{hagmayer2013hierarchical}).\\
-\\
-\cite{lucas2010learning} proponen un framework que incorpora inferencia basada en covarianzas así como conocimiento adquirido.
-Se enfocan en aprender  \textit{forma funcional} de las relaciones causales.
-Las contribución principal del artículo es: se muestra que el problema de aprender la forma funcional de una relación causal puede ser formalizada utilizando modelos jerárquicos bayesianos; además, se muestran una serie de experimentos que ponen a prueba las predicciones cualitativas hechas por este modelo.\\
-\\
-\cite{hagmayer2013hierarchical} presentan un overview general sobre los modelos jerárquicos bayesianos en el contexto causal.
-En el artículo exponen las ventajas de utilizar estos modelos para realizar inferencia causal, así como sus limitaciones.
-Una limitación mencionada es que dependen considerablemente de conocimiento previo.
-Otra, que existe evidencia que la representación que las personas tienen de modeos causales no corresponde a modelos gráficos causales (redes bayesianas) pues existen experimentos en los cuales la propiedad Markoviana no se cumple.
-Finalmente, argumentan que los HBM's y las redes bayesianas no modelan los procesos cognitivos subyacentes al apredizaje causal, aunque existen evidencias de que las personas sí llevan a cabo un tipo de actualización Bayesiana
 

### 2017-12-20 Segundo_Semestre/Propuesta/Resumenes.tex
primer commit

--- 2017-12-20 Primer_Semestre/Propuesta/Propuesta.tex
@@
 \maketitle
-\tableofcontents
-\newpage
-\section{Introducción}
-El problema al cual nos enfrentamos consiste en el aprendizaje de relaciones causales a partir de datos.
-Los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.
-Al tratar de averiguar relaciones causales, es común contar con información previa sobre el fenómeno dentro del cual se quiere investigar causalidad, por lo que recurrimos a la estadística bayesiana para el manejo de esta información inicial.\\
-\\
-La primera parte de este resumen consiste en Redes Bayesianas, que son objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
-Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
-\\
-Veremos que al intentar obtener relaciones causales a partir de información proveniente del mundo real necesitamos obtener las entidades que se relacionan entre sí de manera causal, por lo que haremos un repaso de técnicas utilizadas en extracción de relaciones.\\
-\\
-Por otro lado, en la literatura proveniente de Psicología Computacional y Ciencias Cognitivas, se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
-Se sabe también que el conocimiento causal adquirido es utilizado para mejorar las futuras decisiones tomadas.\\
-\\
-Finalmente, se discutirá un poco sobre Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.
-Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.
-Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones.
-La pregunta se convierte entonces en cómo obtener e incorporar conocimiento causal en un problema de aprendizaje por refuerzo.
-Además, uno se pregunta si es posible utilizar este conocimiento causal obtenido al interactuar sobre el mundo para realizar otro tipo de inferencias causales en ciertos dominios de interés.
-\section{Causalidad}
-\subsection{Intro}
-Saber si es posible hablar de relaciones causa-efecto es un tema polémico; los métodos probabilistas tradicionales sólo codifican asociaciones estadísticas, pero la idea
-La diferencia entre modelos probabilistas y modelos causales se da cuando es de interés tomar acciones que modifiquen el entorno y no nada más observar los valores de las variables. \\
-Según \cite{pearl2014probabilistic}, las dos preguntas fundamentales en Causalidad son:
-\begin{itemize}
-\item ¿Qué evidencia empírica es necesaria para hacer inferencias válidas sobre relaciones causa-efecto?
-\item Dado que estamos dispuestos a aceptar información causal sobre un fenómeno, ¿qué inferencias podemos obtener a partir de esta información, y cómo?
-\end{itemize}
-En cuanto al segundo punto, Pearl considera que hay tres tipos de preguntas a considerar
-\begin{itemize}
-\item Predicciones: si sabemos que una persona ingirió alimentos de dudosa procedencia, ¿se enfermará?
-\item Intervenciones: si le damos cierto tratamiento a un paciente, ¿mejorará?
-\item Contrafactuales: ¿el paciente del punto uno se habría enfermado de todos modos si no hubiera consumido esos alimentos?
-\end{itemize}
-Pearl argumenta que aunque causalidad suele entenderse como afirmaciones necesarias, estas son hechas en contexto de incertidumbre, por lo que cualquier teoría causal debe estar en términos de probabilidades.
-\subsection{Un poco de historia}
-Aristóteles definió 4 \textit{causas} para una cosa, una de ellas llamada \textit{causa eficiente} que estaba definida como aquello que hace la cosa.\\
-\\
-En 1690, el filósofo inglés John Locke propuso como definición de causa “aquello que produce alguna idea simple o compleja”, y lo que es producido lo definió como efecto.\\
-\\
-\cite{holland1986statistics} resalta la importante diferencia en que Aristóteles hizo énfasis en las causas de una cosa, pero no en el efecto de estas. \\
-\\
-Fue David Hume quien se considera que hiz una importante contribución, y él definió la causalidad como una relación entre experiencias más que una entre hechos.
-Argumentó que no es verificable empíricamente que la causa produzca el efecto, sino sólo que el evento llamado \textit{causa} es invariablemente seguido por aquella experiencia llamada \textit{efecto}.\\
-\\
-Hume definió 3 criterios básicos para causalidad:
-\begin{itemize}
-\item Contiguidad espacio-temporal: tanto causa y efecto tienen lugar en una entidad común (espacio y tiempo).
-\item Sucesión temporal: La causa tiene que preceder temporalmente al efecto. 
-\item Conjunción constante: Causa y efecto ocurren juntos, o no ocurren juntos.
-\end{itemize}
-John Stuart Mill argumentó tajantemente que observación sin experimentos no puede probar causalidad. Mill identificó cuatro métodos generales para la identificación de relaciones causales. 
-\subsection{Modelo de Rubin}
-El modelo de Rubin parte de una población de \textit{unidades} $U$.
-En general, estas unidades son los objetos de estudio sobre las cuales actuan las causas.
-Para hacer más sencilla la descripción del modelo, supondremos que sólo existen dos causas (o niveles de tratamiento) denotadas por $t$ y $c$.
-Sea $S$ una variable indicadora de la causa a la cual cada unidad $u \in U$ fue expuesta.
-Además, consideramos dos variables respuesta $Y_t$, $Y_c$.
-$Y_t(u)$ reoresenta el valor de la respuesta si la unidad $u$ es expuesta a $t$ y análogo para $c$.\\
-\\
-\begin{defi}
-Se define el efecto de la causa $t$ sobre $u$ relativo a la causa $c$ como
-\[ Y_t(u) - Y_c(u). \]
-Esta expresión se lee de la siguiente manera: $t$ causa el efecto $Y_t(u) - Y_c(u)$ en la unidad $u$ (relativo al control $c$).
-\end{defi}
-Notemos que es imposible observar simultáneamente el efecto de $t$ y de $c$ sobre una misma unidad $u$ y a esto \cite{holland1986statistics} le llama el Problema Fundamental de la Inferencia Causal.
-El problema, según explica Holland, radica en la palabra \textit{observar} pues no podemos saber simultáneamente el efecto de exponer a un enfermo en particular a un nuevo tratamiento y al mismo tiempo el efecto de no exponerlo.
-\subsection{Modelo de Suppes}
-En su libro “Probabilistic Theory of Causality”, (\cite{suppes1970probabilistic}), Patrick Suppes intenta mejorar el análisis de Hume sobre causalidad; en particular, la condición de conjunción constante.
-Para Suppes, no importa qué sean las causas y los efectos, sino que se puedan expresar como eventos que ocurren bajo ciertas condiciones.
-Suppes supone que todas las causas necesariamente preceden en el tiempo a sus efectos.
-Define una causa \textit{prima facie} de un evento como un evento que precede temporalemnte a este y que está asociado de manera positiva con él.
-Define como una causa espuria a una causa prima facie tal que es condicionalmente independiente del efecto dada otra causa la cual es anterior en el tiempo a la causa prima facie y que está asociada con el evento dada la causa.
-Una causa genuina es una causa prima facie que no es espuria.
-\begin{defi}
-Si $r < s$ representan valores de tiempo, se dice que el evento $C_r$ es una causa prima facie del evento $E_s$ si
-\[ P(E_s | C_r ) > P(E_s). \]
-\end{defi}
-\begin{defi}
-Un evento $C_r$ se dice que es una causa espuria de un eventi $E_s$ si $C_r$ es una causa prima facie de él y para alguna $q < r < s$ existe un evento $D_q$ tal que 
-\[P(E_s | C_r, D_q) = P (E_s | D_q) \]
-y además,
-\[P(E_s | C_r , D_q) \geq P(E_s | C_r). \]
-\end{defi}
-Por lo tanto,
-\begin{defi}
-Para tiempos $r < s$, un evento $C_r$ se dice que es una causa genuina de un evento $E_s$ si $C_r$ es una causa prima facie de $E_s$, pero no es una causa espuria.
-\end{defi}
-Notemos que Suppes está definiendo la causa de un efecto, más que el efecto de una causa.
-Además, no hay restricciones a la naturaleza de una causa más allá de que sea un evento que ocurre antes en el tiempo.
-El modelo de Suppes no puede expresar el efecto de una causa en particular, sólo comportamientos promedio (\cite{holland1986statistics}).\\
-\\
-En palabras de Holland, la causalidad de Suppes no es más que correlación entre causa y efecto que no puede ser eliminada al parcializar causas legítimas.
-\subsection{Modelos Causales en Medicina}
-Un caso de aplicación en medicina es determinar que ciertas bacterias causan ciertas enfermedades.
-Existe un conjunto de postulados planteados por Koch para decidir cuándo un micro-organismo está implicado en una enfermedad.
-Aunque no existe un planteamiento formal, estos pueden expresarse como (\cite{yerushalmy1959methodology}):
-\begin{itemize}
-\item El organismo debe estar presente en todos los casos de la enfermedad supuestamente provocada por él.
-\item El organismo debe aislarse de los pacientes y cultivarse en estado puro.
-\item El organismo cultivado en estado puro debe reproducir el padecimiento. 
-\end{itemize}
-Estos postulados aseguran que no haya datos que apoyen una hipótesis nula de cero efecto causal. \\
-\\
-Por otro lado, Hill, quien fue de los primeros en estudiar la asociación entre fumar y cáncer, encontró 9 factores útiles a la hora de decidir que una asociación observada es causal. 
-\subsection{Causalidad de Granger en Economía}
-La causalidad de Granger se utiliza para series de tiempo, lo cual es de alta importancia en Economía.
-\cite{holland1986statistics} opina que las ideas esenciales de Granger no están limitadas a series de tiempo y que pueden ser expresadas en términos de variables que actuan sobre unidades de una población.\\
-\\
-En la teoría de Granger, una variable causa otra variable; es decir ( \cite{holland1986statistics}), las realizaciones de una variable mejoran las predicciones hechas sobre la otra variable.
-La única parte en la que se utilizan series de tiempo en esta definición es para hablar de valores de las variables definidos hasta un cierto tiempo $t$.\\
-\\
-Si $X,Y,Z$ son variables aleatorias definidas sobre cierta población, se dice que $X$ no es una causa de Granger de $Y$ si $X$ y $Y$ son condicionalmente independientes dado $Z$.
-Entonces, $X$ es una causa de Granger de $Y$ si distintos valores de $X$ llevan a distintas distribuciones predictivas de $Y$ dados $X$ y $Z$; esto es, $X$ ayuda a predecir $Y$ aun cuando $Z$ es tomada en cuenta.
-La no-causalidad de Granger es parecida a las causas espurias de Suppes pues ambas consideran que no se pueda predecir un evento dada cierta información.\\
-\\
-En opinión de Holland, la causalidad de Granger falla de la misma manera que la de Suppes, pues a la luz de nueva información una causa Granger puede dejar de serlo.
-\subsection{Modelos Causales en Ciencias Sociales}
-\subsection{Aprendizaje de Modelos Causales a partir de datos}
-En los métodos de estado del arte de Modelos Causales Funcionales suele hacerse los siguientes supuestos (\cite{goudet2017learning}):
-\begin{itemize}
-\item Suficiencia Causal (CSA)
-\item Propiedad Markoviana (CMA)
-\item Independencias Condicionales
-\item Fiabilidad causal (CFA)
-\end{itemize}
-\subsubsection{Aprender el CPDAG}
-Existen tres familias de métodos para extraer el grafo: métodos de restricciones, métodos de score y métodos híbridos.\\
-\\
-Los métodos basados en restricciones explotan las independencias condicionaes para identificar todas las v-estructuras.
-El algoritmo más conocido es el Algoritmo PC (\cite{spirtes2000causation}).
-El algoritmo PC primero construye un esqueleto basado en independencias condicionales; luego, identifica las v-estructuras y posteriormente orienta las aristas.
-El algoritmo FCI (\cite{spirtes1999algorithm}) extiende el algoritmo PC al relajar la suficiencia causal e incorporar variables latentes.
-El algoritmo RFCI de \cite{colombo2012learning} es más rápido que FCI y maneja datos de alta dimensionalidad.
-La debilidad de estos métodos, según \cite{goudet2017learning} es que dependen demasiado en pruebas de independencia, que dependen a su vez de la cantidad disponible de datos.\\
-\\
-Los métodos basados en score exploran el espacio de CPDAG's y minimizan un score global.
-La exploración se realiza en términos de operaciones sobre grafos, por ejemplo añadir o remover aristas.
-El algoritmo Greedy Equivalent Search (GES) de \cite{chickering2002optimal} encuentra la estructura óptima en términos del BIC.\\
-\\
-Los métodos híbridos combinan las técnicas anteriores.
-Por ejemplo el Max-Min Hill Climbing Algorithm de \cite{tsamardinos2006max} primero construye un esqueleto utilizando pruebas de independencia condicional, y luego le da orientación a las aristas utilizando un método greedy sobre un score bayesiano.
-\subsubsection{Explotando la asimetría entre causa y efecto}
-Los métodos anteriores no toman en cuenta la asimetría entre causas y efectos.
-La intuición detrás (\cite{goudet2017learning}) está dada por el siguiente ejemplo: Si consideramos el modelo causal funcional $Y=X+U$, entonces los métodos anteriores basados en grafos no pueden darle orientación a la arista $X - Y$, pues tanto $X \rightarrow Y$ como $Y \rightarrow X$ son Markov equivalentes.\\
-\\
-\cite{stegle2010probabilistic} infiere la dirección causal al construir dos modelos Bayesianos generativos: uno para $X \rightarrow Y$ y otro para $Y \rightarrow X$.
-La dirección causal es determinada a partir del modelo generativo que mejor ajuste a los datos.
-Otros métodos, como el propuesto por \cite{sgouritsa2015inference} se basan en que si $X \rightarrow Y$, entonces la probabilidad marginal de la causa $P(X)$ es independiente del mecanismo causal $P(X | Y)$; por lo tanto, estimar $P(Y | X)$ a partir de $P(X)$ no debería poderse, pero estimar $P(X | Y)$ a partir de $P(Y)$ sí debería ser posible.
-\subsubsection{Como problema de Aprendizaje}
-Inspirados en los concursos de Kaggle de \cite{Guyon2013}, \cite{lopez2015towards} plantean el problema de inferencia causal como un problema de clasificación de medidas de probabilidad.
-Los datos provistos por el concurso son 16,200 parejas de variables $(X_i,Y_i)$.
-Cada par consta de una realización $(x_i,y_i)$ y una etiqueta $l_i$ que indica el valor de la verdadera relación causal; es decir, $l_i$ indica uno de los siguientes casos: $X \rightarrow Y,$ $Y \rightarrow X$, $X_i \perp Y_i$, $X \leftrightarrow Y$.\\
-\\
-Notemos que este método tiene como punto débil que depende de la representatividad de la muestra de entrenamiento y que supone la existencia de una distribución madre
-\subsubsection{Modelos Causales Funcionales Neuronales}
-\cite{goudet2017learning} proponen un esquema que consiste en tres pasos: primero, identificar el esqueleto según propiedades de independencia estándar; luego, orientar las aristas utilizando métodos bivariados estándar y finalmente hacer una optimización sobre todo el grafo examinando relaciones locales.
-Este último paso se realiza al aprender estos modelos locales como redes neuronales generativas; es decir, se modela la distribución $P(X | X_{Pa_i}, U_i)$ de un nodo condicional en sus padres (causas) con una red neuronal.\\
-\\
-Formalmente, se modela cada ecuación
-\[X_i = f_i (X_{Pa_i},U_i) \]
-como una red neuronal generativa global, en la cual cada función $f_i$ es modelada por una red neuronal generativa con una capa oculta.\\
-\\
-Encontrar un modelo causal de este tipo requiere llevar a cabo dos tareas: estimar la estructura (el grafo) de los padres de cada variable, y las redes neuronales $f$
-\subsection{Correlación y causalidad}
-Dos variables aleatorias $X$,$Y$ pueden estar correlacionadas en distintos casos; por ejemplo, si $X$ causa $Y$, o si tanto $X$ como $Y$ son causadas por otra variable desconocida $V$.
-Si e conoce la existencia de $V$ y además podemos observarla.
-\section{Modelos gráficos causales}
-Sabemos que en una red bayesiana los arcos, aunque dirigidos, no necesariamente tienen una interpretación directa, lo único que importa es que el grafo capture adecuadamente las relaciones de dependencia entre variables aleatorias.
-Nos gustaría que fuera el caso que $X \to Y$ significara $X$ \textit{causa} Y.
-Resulta que existe una familia de modelos causales en los cuales se \textit{interviene} en el mundo.
-La idea clave es utilizar \textit{intervenciones ideales}, las cuales se denotan $do(Z=z)$ que significa forzar a la variable aleatoria $Z$ a tomar el valor $z$, denotado de manera simple como $do(z)$.
-Esta operación corresponde al caso en que un agente directamente modificó el estado del mundo y no sólo observó los valores de este de manera pasiva; por esta razón es muy importante tener en mente que es distinto $P(Y | do(z), X=x)$ que $P(Y | Z=z, X=x)$.
-Las preguntas causales de interés son planteadas de la forma $P(X | do(z))$ y son llamadas \textit{quries intervencionales}.
-Este tipo de intervenciones surgen de manera natural en muchos problemas; por ejemplo, “si el paciente recibe este tratamiento, cuál es la probabilidad de sobrevivir la enfermedad” puede ser planteada como $P(sobreivir | do(tratamiento))$.\\
-\\
-Formalmente, un modelo gráfico causal (\cite{koller2009probabilistic}, \cite{pearl2014causality}) tiene la misma forma que una red Bayesiana; es decir, consiste en una red acíclica dirigida que codifica una distribución entre un conjunto de variables aleatorias.
-La diferencia radica en la interpretación de las flechas, pues cada variable aleatoria está gobernada por un modelo causal en función de sus padres en el grafo.
-Una intervención consiste en forzar uno de los nodos a tomar un valor con probabilidad 1 y a recortar todas las flechas que entran en ese nodo.\\
-\\
-Es importante que se cumpla que si dos variables están correlacionadas, entonces una es padre causal de la otra y que no falten padres causales de ninguna variable.\\
-\\
-Para poder contestar pregunta acerca de intervenciones, se necesita introducir el concepto de \textit{grafo mutilado}
-\begin{defi}
-Sea $\mathcal{B}$ una red sobre un conjunto de variables aleatorias y sea $Z_1=z_1,...,Z_k=z_k$ una instanciación de estas variables denotada por $\mathbf{Z}=\mathbf{z}$ .
-Se define el grafo mutilado $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$ como:
-\begin{itemize}
-\item Cada nodo $Z_i \in \mathbf{Z}$ no tiene padres en $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$
-\item El modelo local de probabilidad condicional para $Z_i \in \mathbf{Z}$ es tal que asignal probabilidad uno al valor $z_i$ y cero a cualquier otro valor.
-\item No se altera ningún otro nodo $X_i \notin \mathbf{Z}$
-\end{itemize}
-\end{defi}
-Entonces, utilizando este concepto, tenemos que para un modelo causal $\mathbf{C}$ una intervención $P_{\mathcal{C}}(Y|do(z)$ es realmente una \textit{querie} probabilista estándar, pero en el grafo mutilado dado por $\mathcal{C}_{Z=z}$; es decir
-\[ P_{\mathcal{C}}(Y|do(z),x) = P_{\mathcal{C}_{Z=z}}(Y|x).  \] 
-\subsection{Identificabilidad de modelos}
-Aun a pesar de desconocer la totalidad de las variables latentes causales, existen casos en los que es posible responder a preguntas causales en modelos que involucran variables latentes utilizando sólo variables observadas; este es el caso de los modelos identificables.
-Existen formas de simplificar queries intervencionales y convertirlas en queries que sólo requieren de datos observacionales pues existen equivalencias entre unas y otras bajo ciertas condiciones.\\
-\\
-\cite{peters2012identifiability} exploran las condiciones para obtener identificabilidad en modelos causales funcionales. 
-\subsection{Aprendizaje de modelos gráficos causales}
-Según \cite{koller2009probabilistic}, el problema de aprender un modelo causal a partir de datos puede dividirse en varios ejes:
-\begin{itemize}
-\item En primer lugar, definir qué se entiende por un modelo causal, pues si estamos trabajando con modelos gráficos causales entonces el problema se convierte en aprender redes bayesianas, pero con la diferencia de la interpretación que se da a las flechas; por otro lado, en modelos causales funcionales se tienen parametrizaciones mucho más complejas para las variables de respuesta, por lo que se necesitan modelos más complejos como el modelo neuronal presentado anteriormente.
-\item Un segundo eje es determinar si ya se conoce la estructura y sólo se necesita aprender la parametrización, o si también debe ser aprendida la parametrización.
-\item En tercer lugar, es el tipo de datos: observados o intervencionales.
-Datos observados son aquellos que provienen de alguna distribución generadora y datos intervencionales es en aquellos en los que se intervino el modelo y se observó la respuesta de este.
-Los datos intervencionales son difíciles de obtener en la práctica, o incluso pudieran ser inmorales o ilegales.
-\item Finalmente, cómo tratar los factores de confusión (confounding factors). Uno puede suponer que simplemente no existen y en estos casos el problema se vuelve relativamente sencillo.
-\end{itemize}
-\subsubsection{Aprender sin factores de confusión}
-El problema de aprender modelos gráficos causales sin factores de confusión es esencialmente el mismo que aprender redes bayesianas (\cite{koller2009probabilistic}), pero hay que hacer algunas suposiciones importantes, pues existen siempre varios DAG's que representen la misma distribución, pero que tendrían distintas interpretaciones causales.
-Una de las suposiciones necesarias para poder aprender modelos causales a través de datos observacionales es la llamada Propiedad Causal de Markov, que es un análogo a la Propiedad de Markov para redes Bayesianas; y su formulación es la siguiente (\cite{koller2009probabilistic}): \textit{cada variable es condicionalmente independiente de sus no-efectos dadas sus causas directas.} El otro supuesto necesario es que las únicas independencias condicionales sean aquellas que surgen de la d-separación en el grafo causal correspondiente (\textit{faithfulness}).\\
-\\
-Aun con estas suposiciones, lo más que puede aprenderse es una clase de equivalencias para el grafo causal y para llevar esto a cabo existen dos familias de métodos: basados en restricciones y basados en score y son los mismos que se utilizan para aprender el CPDAG en modelos causales funcionales.
-\subsubsection{Aprender de datos intervencionales}
-En el caso en que se tienen datos provenientes de intervenciones, lo cual es lo más común en el contexto de experimentos científicos.
-Se utilizan métodos basados en score para aprender de este tipo de datos.
-\subsubsection{Aprender con variables latentes}
-En el caso en el que suponemos que existen variables latentes, estamos pensando que los datos fueron generados por una distribución que es \textit{parecida}, o parcial, a la verdadera.
-De nuevo, existen familias de métodos basadas en restricciones y métodos basados en scores.
-\subsection{Redes Bayesianas causales: aplicación en genética}
-\cite{su2013using} intentan aprender la estructura de las relaciones entre un conjunto de variables relacionadas a un estudio de cáncer (\cite{karagas1998design},\cite{karagas2004incidence}).
-Específicamente, los autores analizan sólo 11 variables relacionadas a un solo gen.
-Identifican relaciones interesantes, como por ejemplo la variable “fumador” como padre causal de la variable “cancer”, aunque no apareció relación entre la variable cancer y un gen que se sabe está relacionado con el cáncer, pero esto se debe a la manera en la que fueron seleccionados los pacientes.
-\section{Modelos causales funcionales y ecuaciones estructurales}
-En otras áreas, como las ingenierías y la física, las relaciones causales se expresan como relaciones funcionales determinísticas en donde cualquier componente aleatorio se introduce en términos de variables no observadas.
-Esto corresponde (\cite{pearl2009causality}) a la noción Laplaciana de lo aleatorio.
-En este contexto, un modelo causal funcional consiste en un conjunto de ecuaciones de la forma
-\[ X_i = f_i (X_{Pa_i},U_i), \textrm{  } i=1,...n\]
-donde $Pa_i$ corresponde al conjunto de variables que directamente determinan a la variable con índice $i$ y donde $U_i$ representa cualquier error de medición o de aspectos no observados.\\
-\\
-Dado un modelo causal de esta forma, podemos obtener un diagrama de la siguiente manera: para cada variable pintamos un nodo, y desde cada elemento de $Pa_i$ dibujamos una flecha hacia el nodo que corresponde a $X_i$.
-A este diagrama se le llama \textit{diagrama causal}.
-Si cumple que el diagrama es acíclico, se le conoce como semi-Markoviano.
-Si, además de ser acíclico, los términos de error son independientes entonces se dice que es Markoviano.
-El siguiente Teorema muestra la conexión entre el diagrama causal y la distribución de probabilidad asociada.
-\begin{teo}
-Todo modelo causal Markoviano induce una distribución $P(X_1,...,X_n)$ que satisface que cada variable $X_i$ es independiente de todos sus no-descendientes dado sus padres $Pa_i$.
-\end{teo}
-Sabemos que dos DAG's que tienen el mismo esqueleto y las mismas v-estructuras son equivalentes, y a la clase de equivalencia le llamaremos CPDAG (\cite{goudet2017learning})
-\subsection{Intervenciones y efectos causales en modelos funcionales}
-Los modelos funcionales causales proveen una manera de estimar cómo cambiaría la distribución en respuesta a cambios externos.
-Lo que se hace en estos modelos es representar las intervenciones como alteraciones a un conjunto de funciones.
-El efecto total de la intervención se calcularía modificando las ecuaciones correspondientes y con este nuevo modelo calcular las nuevas probabilidades.
-Denotamos una intervención como $do(X=x)$.
-Se dice que una variable $X_i$ es causa directa de $X_j$ si
-\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
-\subsection{Contrafactuales en modelos funcionales}
-
 \section{Extracción de relaciones causales a partir de texto}
@@
 Finalmente, argumentan que los HBM's y las redes bayesianas no modelan los procesos cognitivos subyacentes al apredizaje causal, aunque existen evidencias de que las personas sí llevan a cabo un tipo de actualización Bayesiana
-
-\section{Aprendizaje por Refuerzo}
-El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno.
-Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
-El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente.
-A su vez, el ambiente devuelve información al agente en términos de una recompensa o ganancia según el estado en el que se encuentre.
-El objetivo del agente es encontrar una serie de acciones que lo lleven a lograr su objetivo final.
-La función de recompensa codifica qué es lo que se desea que el agente aprenda, pues esta función devuelve un valor para cada estado y con esto el agente asignará un valor a cada estado según la ganancia a largo plazo que un este representa.\\
-\\
-La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
-El objetivo del agente es encontrar la política que maximice la recompensa.\\
-\\
-La regla general que se sigue para la definición del ambiente (\cite{sutton1998reinforcement}) es considerar que todo aquello que no pueda ser explícitamente alterado por el agente pertenece al ambiente.
-La frontera agente-ambiente tiene que ver con los límites de lo que el agente puede controlar, no sobre lo que puede \textit{conocer}.
-Veremos ahora con detalle las partes que conforman esta dinámica. 
-\subsubsection{Objetivos y Recompensas}
-Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada.
-Por ejemplo, un agente que queremos que aprenda a jugar ajedrez debe ser premiado por ganar los juegos, no por lograr sub-objetivos como comer más piezas del oponente o tomar control del centro del tablero (\cite{sutton1998reinforcement}), pues podría suceder que el agente encuentre cómo lograr esto sin ganar el juego necesariamente.
-Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
-\subsubsection{Definición de Cadenas de Markov}
-Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
-\[P(X_{t+1} \in A | X_0=x_0, X_1=x_1 , ... , X_t=x_t) = P(X_{t+1} \in A | X_t = x_t).\]
-Estos procesos quedan caracterizados por sus probabilidades de transición $p^t_{ij}=P(X_{t+1}=j | X_{t}=i)$.
-Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
-\subsubsection{Procesos de Decisión Markovianos con recompensa}
-Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$.
-En respuesta a la acción tomada, se recibe una recompensa $r_t = f(a_t)$ para alguna función $f$ desconocida, la cual puede ser aleatoria o determinista.
-Los procesos de decision Markovianos con recompensa (\cite{Puterman:1994:MDP:528623}) están definidos por un espacio de estados $\mathcal{S}$, un espacio de acciones posibles $\mathcal{A}$, una función de transición $f$ y la función recompensa $r$.
-La función transición puede ser estocástica o determinística.
-
-\subsubsection{Función valor}
-Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado.
-Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política.
-Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$.
-Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$.
-La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante.
-Formalmente, para un proceso de decisión Markoviano, definimos la \textit{función valor} (value function en la literatura) para un estado $s$, y relativa a una política $\pi$, como el retorno esperado si se comienza en $s$ y a partir de ahí se sigue la política $\pi$
-\[ V^{\pi}(s) = \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]. \]
-De manera similar, definimos la \textit{función acción-valor} $Q^{\pi}(s,a)$ es una valoración de tomar la acción $a$ estandoen el estado $s$, y después seguir la \textit{política} $\pi$
-\[ Q^\pi(s,a) = \mathbb{E}[R_t | s_t =s, a_t = a]. \]
-Por la ley de los grandes números, podemos estimar ambas funciones $V^{\pi}$ y $Q^\pi$ si se promedian los retornos recibidos al llegar a cada estado con el número de veces que se ha llegado al estado.
-Esta es la idea detrás de los llamados Métodos Monte-Carlo que serán expuestos en una sección posterior.\\
-Notemos una propiedad interesante de la función $V^{\pi}$:
-\begin{eqnarray*}
-V^{\pi}(s) &=& \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ \sum_{k=0}^\infty \gamma^k r_{t+k+1} | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ r_{t+1} + \gamma \sum_{k=0}^\infty \gamma^k r_{t+k+2} | s_t = s \right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} P(s_{t+1}=s' | s_t = s, a_t = a) \left[ \mathbb{E}[r_{t+1}|s_t=s,a_t=a,s_{t+1}=s'] + \gamma \mathbb{E} \left[ \sum \gamma^k r_{t+k+2} | s_{t+1}=s' \right]\right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} \mathcal{P}_{ss'}^a \left[ \mathcal{R}_{ss'}^a + \gamma V^{\pi}(s') \right]
-\end{eqnarray*}
-A esta última ecuación se le conoce como la Ecuación de Bellman para $V^\pi$.\\
-\\
-También para la función $Q^{\pi}(s,a)$ se tiene una ecuación de Bellman:
-\[ Q^{\pi}(s,a)=\mathbb{E}_{s',a'}[r+ \gamma Q^{\pi}(s',a') | s',a']. \]
-Además, la función $V^\pi$ puede ser obtenida a partir de la función $Q^{\pi}(s,a)$ de la siguiente manera:
-\[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
-\subsubsection{Soluciones y Optimalidad}
-Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa.
-Resulta que es posible definir un orden parcial entre políticas en términos de la función valor, pues $\pi \geq \pi'$ si $V^{\pi}(s) \geq V^{\pi'}(s)$ para todo estado $s$.
-Para procesos de decisión Markovianos finitos existe siempre una política mejor que las demas, a la cual llamaremos \textit{política óptima}.
-De hecho, pueden existir varias, pero todas ellas comparten la misma función valor, denotada como $V^\ast$.
-Si para una política $\pi^\ast$ se tiene que es mejor que todas las demás, entonces podemos describir su función valor como
-\[  V^{\pi^\ast} = \max_{\pi} V^{\pi}(s).\]
-Por supuesto, $V^\ast = V^{\pi^\ast}$
-\\
-Análogamente, las políticas óptimas tienen la misma función $Q$ óptima, por lo que definimos \textit{función accion-valor óptima} como el retorno esperado de tomar la acción $a$ en el estado $s$ y a partir de ahí seguir una política óptima:
-\[ Q^\ast (s,a) = \max_\pi Q^{\pi}(s,a) = Q^{\pi^\ast}(s,a). \]
-Los métodos modernos de aprendizaje por refuerzo intentan aprender directamente la función valor o la función $Q$, y una vez Q y una vez que tenemos $Q^{\pi^\ast}$ podemos definir una política óptima de la siguiente manera:
-\[ \pi^\ast = \text{argmax}_a Q^\ast (s,a). \]
-Además, utilizando la ecuación de Bellman vimos que se podía escribir $Q$ en términos de $V$, y por lo tanto podemos escribir $Q^\ast$ en términos de $V^\ast$ como
-\[ Q^\ast (s,a) = \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La ecuación de Bellman puede escribirse sin hacer referencia a ninguna política de la siguiente manera:
-\[ V^\ast (s) = \max_a \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La función óptima $ Q^\ast$ lo que hace es maximizar sobre todas las decisiones:
-\begin{eqnarray*}
-Q^\ast &=& r_{t+1} + \gamma \max_{a_{t+1}} r_{t+2} + \gamma^2 \max_{a_{t+2}} r_{t+3}...\\
-&=& r_{t+1} + \gamma \max_{a_{t+1}} Q^{\ast} (s_{t+1},a_{t+1}).
-\end{eqnarray*}
-Formalmente:
-\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
-\subsection{Métodos básicos de solución: Model-based}
-\subsubsection{Programación Dinámica}
-Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
-				\begin{itemize}
-					\item Evaluación de una política: Teniendo una política cualquiera $\pi$, ¿cómo obtener la función $V^\pi$?\\
-					         Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
-					         \[ V^\pi (s) = \mathbb{E}[r + \gamma V^\pi (s') | s_t = s]. \]
-					          Pero esto nos deja con el mismo problema, pues tenemos de ambos lados una función desconocida.
-Para problemas de decisión finitos, este es un sistema de ecuaciones con tantas ecuaciones como estados, lo cual puede ser costoso de resolver.
-Consideramos entonces un método iterativo:
-					         \[ V_{k+1} (s) = \pi(s) \sum_{s'} \mathcal{P}_{s,s'}^a [ \mathcal{R}_{s,s'}^a + \gamma V_k(s')], \]
-					         donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$.
-Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.
-					\item Mejora de una política: Dada una política $\pi$ y estando en un estado $s$ nos preguntamos qué pasaría si en vez de escoger la acción $\pi(s)$ escogieramos cualquier otra acción $a$.
-La consecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
-Entonces, la clave es comparar esto con $V^{\pi}$.
-De ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
-Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.
-					 Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
-					\[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
-					Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.
-					\item Iteración de políticas: Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
-					\[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
-					pues cada política está garantizado que es mejor que la anterior.
-					\item Iteración de valor: El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite. Podemos truncar la etapa de policy evaluation.
-					\[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
-					\item Iteración de política generalizada (GPI): Notar que la iteración de políticas consiste en dos procesos que interactuan entre sí, pero de manera que cada proceso se hace de manera completa antes de comenzar el otro lleva a preguntarse si esto puede hacerse de manera más eficiente.
-Se conoce como iteración de políticas generalizada (GPI) a la idea general de que los procesos de evaluación y mejora de política interactuen entre sí.
-					\item Limitaciones: Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
-Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
-				\end{itemize}
-\subsection{Métodos básicos de solución: Model-free}
-Métodos que no necesitan un modelo completo del sistema, sólo necesitan experiencia.
-Encontramos los métodos Monte Carlo que simulan las transiciones y los estados, y los métodos de diferencias temporales, que van actualizando iterativamente el valor de los estados observados, el más famoso de estos es el TD($\lambda$).
-\subsubsection{Métodos Monte-Carlo} 
-Los métodos Monte-Carlo no requieren un modelo del ambiente y son conceptualmente sencillos, pero no son adecuados para hacer actualizaciones incrementales.
-Estos métodos requieren \textit{experiencia} en vez de un modelo completo; es decir, requieren secuencias de estados, acciones y recompensas, ya sean estos obtenidos de simulaciones o de interacción \textit{on-line} con el ambiente.
-Los métodos Monte-Carlo suponen que la experiencia proviene de eventos episódicos, es decir que terminan siempre de una u otra forma y es al final de estos episodios que se alteran las estimaciones de valor así como las políticas.
-En este sentido, los métodos Monte-Carlo son incrementales por episodio, pero no paso a paso dentro de cada episodio.
-A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general.
-La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
-\begin{itemize}
-\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
-Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$.
-El algoritmo \textit{every-visit} estima $V^\pi (s)$ como las recompensas promedio de todas las visitas al estado $s$ en todos los episodios; por otro lado, el algoritmo \textit{first-visit} promedia sólo lo obtenido al visitar por primera vez un estado en todos los episodios.
-Ambos métodos convergen a $V^\pi$.
-\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
-Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política.
-Entonces, es importante tener un método para estimar $Q^\ast$.
-\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
-El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.
-\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
-A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
-\[ \pi_0 \xrightarrow{E} Q^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} Q^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} Q^\ast.  \]
-Bajo el supuesto de episodios infinitos y visita a todas las parejas estado-acción se tiene que los métodos Monte-Carlo encuentran $Q^{\pi_k}$ de manera exacta para $\pi_k$.\\
-Para hacer la mejora de una política, se escoge de manera determinística la acción $a$ en la cual se alcanza el máximo de la función $Q$; es decir,
-\[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
-\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
-\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
-							\begin{itemize}
-									\item On-Policy: intentan evaluar mejorar la política que es actualmente utilizada para la toma de decisiones.
-									\item Off-Policy: Evaluan o mejoran una política distinta
-							\end{itemize}
-\end{itemize}
-\subsubsection{Diferencias Temporales}
-Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos.
-Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
-\begin{itemize}
-\item Evaluación de política en Diferencias Temporales
-\item TD(0)
-\item SARSA: Es un método \textit{on-policy}
-\item Q-Learning: Es un método \textit{off-Policy}
-\item Métodos Actor-Critic
-\end{itemize}
-\section{Deep Reinforcement Learning}
-\subsection{Mejoras a los métodos básicos}
-	\begin{itemize}
-		\item Programación Dinámica con aproximación de funciones.
-		\item Cuando el espacio de estados es muy grande, no es factible utilizar una representación en una \textit{look up table}, por lo que es necesario utilizar otro tipo de aproximadores (\cite{busoniu2010reinforcement})
-	\end{itemize}
-\subsection{Deep Learning}
-	\begin{itemize}
-		\item Redes neuronales profundas como aproximadores generales.
-		\item Redes Convolucionales para poder tomar como input los pixeles de un juego directamente.
-	\end{itemize}
-\subsection{Value based Reinforcement Learning:}
-	\begin{itemize}
-	   \item Representar la función valor mediante una \textit{Q-network} con pesos $w$:
-	   \[Q(s,a,w) \approx Q^\ast(s,a) \]
-	   \item Q-learning:
-        \begin{itemize}	  
-        	\item  Sabemos que los valores óptimos deben satisfacer la ecuación de Bellman:
-        	\[ Q^\ast(s,a)=\mathbb{E}_{s'}[r + \gamma \max_{a'} Q^\ast(s',a')|s,a]. \]
-        	\item Tratamos el término $r + \gamma \max_{a'} Q^\ast(s',a',w)$ como el objetivo.
-        	\item Minimizar error cuadrático medio mediante Gradiente Estocástico:
-        	\[ I = (r + \gamma \max_{a'} Q^\ast(s',a',w) - Q(s,a,w))^2. \]
-        	\item Converge a $Q^\ast$ utilizando una representación tabular.
-        	\item Diverge al utilizar redes neuronales debido a la alta correlación entre muestras y la no-estacionareidad del objetivo.
-        	\item Lo resolvemos con Experience Replay (\cite{Lin1992}),  el cual consiste en guardar todas las experiencias de modo que al entrenar la red tomar sólo mini-batches aleatorios  en vez de utilizar la transición más reciente.
-	    \end{itemize}
-	\item Aplicaciones de DQN.
-		\begin{itemize}
-			\item \cite{mnih2013playing}.
-Paper en el que muestran cómo una computadora aprende a jugar Atari 2600 a través de sólo observar los pixeles del juego y obteniendo una 																		recompensa conforme el score aumentaba.
-Utilizando la misma arquitectura, se pudieron aprender 7 juegos
-			\item \cite{mnih2015human}.
-En este paper, aumentaron a 49 juegos aprendidos.
-Utilizan una Deep Q-Network, que consiste en aproximar la función Q mediante una red 																			neuronal que toma como input los estados (las pantallas del juego) y devuelve como output el valor Q para cada posible acción.
-En este caso, 																			utilizaron una red convolucional con 3 capas de convolución y 2 capas totalmente conectadas.
-	  \end{itemize}
-\item Mejoras posteriores a DQN
-   	   \begin{itemize}
-   	   		\item Double DQN.
-   	   		\item Prioritized Replay.
-   	   		\item Duelling Networks.
-   	   \end{itemize}
-    	 \end{itemize}  	   
-\subsection{Policy Based DRL}
-Resulta que Deep Q-Learning no es un gran algoritmo, y en su lugar, \cite{DBLP:journals/corr/MnihBMGLHSK16} proponen utilizar directamente Policy-Gradients.
-La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
-\[ a = \pi (a|s,u) \textrm{ o } a = \pi (s,u). \]
-En este contexto, definimos una función objetivo $L(\textbf{u})$ como:
-		\[ L(\textbf{u}) = \mathbb{E}[r_1 + \gamma r_2 + \gamma^2 r_3 + ... |  \pi(\cdot,\textbf{u})]. \]
-Optimizamos con Stochastic Gradient Descent. El gradiente de una policy estocástica está dado $\pi (a|s,\textbf{u})$ por:
-		\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[  \frac{\partial \log \pi(a|s,\textbf{u})}{\partial \textbf{u}}Q^\pi(s,a)\right].\]
-Por otro lado, el gradiente de una policy determinística $a=\pi(s)$ está dado por:
-\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[ \frac{\partial Q^\pi (s,a)}{\partial a} \frac{\partial a}{\partial \textbf{u}} \right]. \]
- Los métodos \textit{actor-only}, donde \textit{actor} se utiliza como sinónimo de \textit{policy} utilizan familias parametrizadas de policies.
-Las ventajas de tener una policy parametrizada es que podemos generar un espectro continuo de accciones, pero los métodos de optimización (policy gradients) sufren de alta varianza en sus estimaciones del gradiente, lo cual genera un lento aprendizaje (\cite{grondman2012survey},\cite{konda2003onactor},\cite{richter2007natural},\cite{boyan2002technical},\cite{baxter2001infinite},\cite{berenji2003convergent}).\\
- \\
-Los métodos \textit{critic-only}, donde \textit{critic} se utiliza como sinónimo de \textit{value function}, que utilizan diferencias temporales tienen una menor varianza en las estimaciones del retorno esperado (\cite{boyan2002technical},\cite{berenji2003convergent},\cite{sutton1988learning}).
-Una manera de derivar una policy en estos escenarios es seleccionando greedy actions; es decir, acciones praa las cuales la value funtion indique que el retorno esperado es máximo, lo cual puede requerir un alto costo computacional.
-Lo usual es discretizar el espacio continuo de acciones.
-Los métodos \textit{actor-critic}, combinan las vetajas de ambos mundos.
-Mientras que las familias parametrizadas tienen la ventaja de utilizar acciones continuas, los métodos critic-only proveen de conocimiento de baja varianza sobre el desempeño; es decir, la estimación del retorno esperado del crític-only permite al actor-only actualizar los gradientes con información de menor varianza, lo cual acelera todo el proceso.
-La menor varianza es intercambiada por un mayor sesgo al inicio del aprendizaje (\cite{berenji2003convergent}).
-\\
-\\
-Los métodos de Policy Gradients optimizan directamente la recompensa acumulada y pueden utilizarse junto con aproximadores no-lineales como las redes neuronales.
-Este método presenta dos grandes retos: el número de muestras de entrenamiento requeridos y la dificultad de obtener mejoras estables debido a la no-estacionariedad de los datos.
-\cite{DBLP:journals/corr/SchulmanMLJA15} atacan el primer problema se puede atacar utilizando value functions para reducir la varianza de los estimados.
-El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
-\\
- \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada.
-Es un algoritmo similar a Policy Gradients y resulta útil para optimizar políticas con redes neuronales.
-Buen performance en juegos de Atari que utilizan la pantalla como input.
-
-\section{Reinforcement Learning y Causalidad}
-Aprendemos causalidad mediante intervenciones, interacciones, etc. Reinforcement Learning es aprender a realizar tareas mediante interacción. ¿Habrá algo en medio?
-\subsection{Conocimiento causal y observabilidad parcial}
-Los algoritmos actuales de reinforcement Learning suponen que el problema puede modelarse dentro de un contexto Markoviano de decisión con recompensas.
-Si el supuesto markoviano se elimina, los algoritmos podrían dejar no funcionar, como muestran en \cite{singh1994learning}.
-Además, todo el desarrollo teórico utiliza la propiedad Markoviana.\\
-\\
-Una manera en la que la propiedad Markoviana puede perderse es si el agente no cuenta con información completa sobre el estado; es decir, al igual que en los HMM, suponemos que el ambiente es Markoviano, pero el agente no tiene acceso a él de modo que los datos no le \textit{parecen} Markovianos.
-A diferencia de los HMM, no existe un procedimiento computacionalmente tratable para los POMDP.
-El problema es que una vez que se obtienen las estimaciones del valor para cada estado, se debe hacer programación dinámica en espacios continuos y esto es infactible salvo para casos pequeños\\
-\\
-La manera de generalizar el caso markoviano al caso no observado es considerar un MDP subyacente con estados $\mathcal{S}=\{ s_1,...,s_N\}$ y probabilidades de transición $\mathcal{P}^a_{s,s'}$.
-Luego, en cada tiempo $t$, el agente observa un mensaje $m_t$ el cual proviene de una distribución $P(m | s_t)$ desconocida.\\
-\\
-En \cite{jaakkola1995reinforcement} se describe un procedimiento que evita estimar el valor de los estados y trabaja directamente en el espacio de políticas estocásticas.
-El procedimiento es \textit{directo}, a diferencia de los métodos \textit{indirectos} en los cuales primero se aprenden los parámetros del MDP y luego se utiliza programación dinámica para obtener la política.
-El algoritmo propuesto consiste en dos partes: primero, se utiliza un método Monte-Carlo para calcular un análogo de los Q-values y luego se utiliza un paso de policy-improvement.\\
-\\
-Desde un punto de vista de psicología, una manear de entender la inferencia de estados ocultos es en términos de causas latentes (\cite{gershman2015reinforcement}).
-Siguiendo esta idea, es evidente que un agente de RL que opera en el mundo real no sólo debe hacer inferencia de estados ocultos, sino que debe poder descubrir los estados ocultos que están detrás del mundo que éste observa.
-Esto es una forma de aprendizaje estructural latente (\cite{gershman2010learning}).\\
-\\
-Un primer acercamiento a este problema puede hacerse mediante técnicas de inferencia Bayesiana no paramétrica.
-Por ejemplo, tanto \cite{gershman2010learning} como \cite{gershman2012exploring} utilizan priors no paramétricas para hacer inferencia sobre causas latentes.
-Un modelo causal latente es una buena representación de la estructura causal verdadera bajo condicionamiento Pavloviano, pues es el experimentador quien es una causa latente.\\
-\\
-En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
-\subsection{Towards Deep Symbolic Reinforcement Learning}
-Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
-\\
-\cite{garnelo2016towards} proponen una arquitectura híbrida que, combina redes neuronales con inteligencia artificial simbólica.
-El modelo propuesto consiste en combinar una red neuronal que como entrada directamente lo que se recibe del ambiente y transforma estos datos crudos en una representación simbólica basada en lógica de primer orden; esta representación simbólica es posteriormente utilizada para escoger la acción a tomar.\\
-\\
-Se utiliza un sencillo juego para probar la arquitectura y en este contexto la red neuronal debe generar representaciones simbólicas de secuencias de estados en el juego en las cuales el flujo de pixeles queda codificado en términos de objetos, tipos y sus interaccioes.
-Luego, con esto se aprenden las políticas.\\
-\\
-El primer paso que hacen es utilizar un autoencoder convolucional para generar, de manera no supervisada, un conjunto de símbolos que representarán los objetos en cada escena.
-Luego, para cada pixel, obtienen la feature con mayor activación y se toman los más altos.
-De esta manera se obtiene un pixel representativo para cada objeto a los cuales posteriormente se le asigna un tipo simbólico según el autoencoder.\\
-\\
-Luego, una vez que se tienen estos símbolos de bajo nivel para cada cuadro del juego, se necesita rastrearlos a lo largo del tiempo.
-Utilizan una verosimilitud que es función de proximidad espacial y el tipo de transiciones.
-Para considerar la información de la dinámica de los cuadros utilizan la diferencia entre cuadros y las posiciones relativas entre objetos.\\
-\\
-Para la parte de aprendizaje, entrenan una función $Q$ por separado para cada interacción entre objetos con actualización
-\[Q^{ij}(s_t^{ij},a_t) \leftarrow  Q^{ij}(s_t^{ij},a_t) + \alpha \left( r_{t+1} + \gamma (\max_{a} Q^{ij} (s_{t+1}^{ij},a) - Q^{ij}(s_t^{ij},a_t)) \right). \]
-\\
-Se prueba el framework sobre varias variantes de un sencillo videojuego. \\
-\\
-Llama la atención como estos autores utilizan maquinaria de lo más reciente (deep learning, autoencoders) para luego obtener una representación en términos de lógica de primer orden.
-Este approach ha quedado claro que ya quedó atrás por lo que debería buscarse algo que sustituya este enfoque lógico.
-En particular, pudiera ser esto un modelo gráfico causal el cual vaya aprendiéndose conforme el agente de RL aprende sobre su entorno.
-
-\subsection{Conocimiento causal y toma de decisiones secuenciales: acciones como intervenciones}
-En cuanto al aprendizaje \textit{on-line} de modelos causales, encontramos en los trabajos de \cite{wellen2012learning} y de \cite{kummerfeld2013tracking} acercamientos tangenciales, o un punto de inicio, al aprendizaje de modelos causales al vuelo.\\
-\\
-Teorías de cómo las personas usan datos de covariación para aprender modelos gráfios causales pueden dividirse en enfoques racionales y heurísticos (según  \cite{wellen2012learning}).
-El enfoque racional modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana.
-Estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
-\\
-Por otro lado, el enfoque heurístico propone que las personas reaccionan a distintos estímulos que los hacen moficiar sus hipótesis causales de una manera no necesariamente racional.
-El modelo de efecto simple de \cite{waldmann2008causal} supone que las personas se enfocan en evaluar una sola relación causal a la vez.
-Es un modelo que aprende la magnitud de la relación causal.
-Si una variable supera cierto umbral, entonces la persona acepta esa relación causal y la incorpora a su bagaje de relaciones.\\
-\\
-El modelo propuesto por \cite{wellen2012learning} sigue esta línea del efecto simple, con la diferencia de que este modelo sí intenta descubrir el proceso cognitivo subyacente, pues el aprender magnitudes causales es un proceso puramente computacional.
-El modelo de efecto simple puede verse como una implementación del modelo de \cite{danks2003dynamical}.
-Además, según narran \cite{wellen2012learning}, el modelo de efecto simple no explica cómo la persona utiliza las magnitudes causales para determinar si existe o no una relación causal.
-En el modelo propuesto, \cite{wellen2012learning}, incorporan un procedimiento de decisión para determinar si existen o no relaciones causales.\\
-\\
-El modelo LPL (\cite{wellen2012learning}) comienza con una estructura causal inicial, la cual consiste en un grafo acíclico dirigido el cual representa las creencias iniciales de un individuo.
-Posteriormente, el modelo altera esta estructura al añadir o remover una sola arista, de manera que el problema se reduce a evaluar relaciones causales entre dos variables.
-Este modelo genera estimaciones de la fuerza de la relación causal para cada posible relación causa-efecto que no sea eliminada a priori.\\
-\\
-Supongamos que un individuo cree a priori que $B$ causa $E$, y está tratando de averiguar si $C$ causa $E$.
-A diferencia del modelo de efecto simple, las estimaciones de la fuerza causal están generadas por un mecanismo similar a los filtros de partículas Bayesianos
-En particular, a partir del conocimiento causal inicial del individuo, el modelo LPL comienza con $n$ partículas.
-Estas partículas $\{V_C^1,...,V_C^n \}$ representan las creencias actuales del individuo sobre si $C$ causa $E$.
-Paralelamente, existen $\{ V_B^1,...V_B^n \}$ para $B$.
-Se define la \textit{capa} $i$ de partículas para el efecto $E$ como la $i$-ésima partícula de cada causa conocida y posible para $E$.
-En el ejemplo, tendríamos que esta capa es $\{ V_B^i, V_C^i \}$.
-Luego, el modelo LPL actualiza cada capa de manera independiente de la siguiente manera
-\[ V_C^{i+1} = V_C^{i,t} + \alpha ( \textrm{observado} - \textrm{esperado}), \]
-Donde “observado” tiene el valor de 1 si el efecto sucede y 0 en caso contrario.
-Por otro lado, “esperado” es el valor esperado del efecto, calculado según la forma funcional de la relación causa-efecto, el cual es calculado por separado para cada causa potencial en una capa.
-El valor esperado de $E$ (para la capa $i$ y la causa potencial $C$) es
-\[ \textrm{esperado} = \prod_{k} (1-V_k^{t-1})(1 - \prod_j (1- V_j^{t-1}).\]
-Conforme el individuo observa más datos, las partículas convergen a los valores verdaderos de los parámetros y las estimaciones para cada capa convergen.
-Este procedimiento no toma en cuenta aspectos estructurales, los cuales se modifican según un criterio de decisión.
-Lo que hace el modelo es una prueba $t$ en la cual la hipótesis nula es que las partículas provienen de una distribución con media cero.
-El criterio de decisión es el siguiente: si actualmente no existe una arista, y la prueba $t$ rechaza la hipótesis nula, entonces se añade la arista.
-Si existe una arista, y la prueba no es significativa, entonces se elimina la arista existente.
-Alterar una arista afectará las futuras estimaciones del valor “esperado” para otras causas.\\
-\\
-\cite{hagmayer2013hierarchical} llaman la atención sobre que los modelos actuales de Decisiones (maximum expected utility) no asumen que los tomadores de decisiones aprenden sobre la estructura causal del problema.
-Por otro lado, las teorías de aprendizaje causal proponen que un tomador de decisiones guían sus acciones a través de conocimiento causal previamente obtenido.
-El modelo de \cite{sloman2006causal} extiende las teorías de aprendizaje causal al contexto de toma de decisiones, la idea básica consiste en que los tomadores de decisiones utilizan la información disponible de su entorno para inducir un modelo causal del problema de decisión al que se enfrentan.
-Se sabe que en los seres humanos el conocimiento causal guía la interpretación de información probabilista y que los humanos somos capaces de aprender la estructura causal del entorno a través de una serie de (traducir cues) (\cite{lagnado2007beyond}).
-La pregunta fundamental que plantean \cite{hagmayer2008causal} es entender qué es lo que aprenden las personas al tomar decisiones de manera secuencial de tal forma que se busque maximizar cierta recompensa a largo plazo y donde las acciones alteran el entorno.
-Al exponerse una persona a estas situaciones de decisión secuencial se obtiene una retroalimentación de cómo las acciones se relacionaban con la obtención de recompensas.
-Las distintas teorías sobre toma de decisiones bajo incertidumbre suponen distintas representaciones para esta retroalimentación.
-Hagmayer propone agrupar estas distintas representaciones en tres niveles:
-\begin{itemize}
-\item Nivel 1: Opciones, resultados y recompensas.
-Sólo los valores esperados de las opciones posibles afectan la toma de decisiones.
-Se conoce como \textit{habit learning} en la literatura de aprendizaje en animales.
-\item Nivel 2: Opciones, relaciones causales de estas opciones a los resultados, relaciones causales entre variables de resultado y relaciones de estas con las recompensas.
-Aquí entra lo que se conoce como \textit{goal-directed learning}.
-La diferencia con el nivel 1 es la capacidad de distinguir entre acciones y los valores asociados a estas.
-\item Nivel 3: Opciones, relaciones causales entre variables de resultado, relaciones causales entre estas variables y relaciones de estas con las recompensas.
-Estos son los modelos causales.
-Las distintas teorías sobre estos modelos suponen que los humanos (o el agente que aprende) adquieren estos modelos mediante procesos de aprendizaje.
-Algunas teorías suponen que la estructura causal del entorno se obtiene a través de \textit{claves}, como la propuesta por \cite{lagnado2007beyond}.
-Por otro lado, otros autores opinan que las personas utilizan propiedades estadísticas de los datos observados, este sería el caso de las teorías de \cite{gopnik2004theory} y de \cite{griffiths2005structure}.
-\end{itemize}
-Aunque las teorías que entran en el nivel 2 son sensibles a las correlaciones entre las respuestas obtenidas, sólo en el nivel 3 se hace una distinción entre causalidad y correlación.\\
-\\
-Los autores hacen énfasis en que aunque las personas sí utilicen modelos causales al tomar decisiones simples, estas decisiones suelen ser sobre escenarios hipotéticos, pero no se han hecho muchos estudios sobre adquisición de conocimiento causal al enfrentarse a problemas de decisión secuencial con el objetivo de maximizar alguna ganancia a largo plazo.
-Además, en los estudios realizados, se ha visto que las personas tienen la capacidad de inducir modelos causales a partir de claves en el entorno; se sabe además que a partir de intervenciones activas se toman mejores decisiones que considerando sólo datos observados de manera pasiva y que las personas tienen la capacidad de integrar relaciones causales obtenidas de manera separada.
-Pero en estos estudios las personas tenían como objetivo descubrir de manera directa estructuras causales.
-No se sabe bien hasta qué punto las personas adquieren conocimiento causal al perseguir objetivos que no sean descubrir directamente estructuras causales a partir del ambiente.\\
-\\
-Los autores realizaron tres experimentos para explorar las representaciones que las personas hacen de las relaciones causales que adquieren al tomar decisiones para maximizar algún objetivo.
-En el primer experimento, el objetivo era estudiar cómo las creencias causales de las personas alteran la toma de decisiones y cómo los participantes utilizaban los resultados de sus decisiones para revisar el modelo causal.
-El experimento consistía en dos etapas de toma de decisiones, la segunda siendo una etapa de prueba.
-En ambas etapas los participantes debían escoger entre una serie de opciones de modo que maximizaran ganancias.
-Se le sugirió a cada participante una posible cadena causal.
-En la primera etapa se les proveía retroalimentación de las consecuencias de sus decisiones, lo cual no sucedió en la segunda etapa sino que ellos mismos debían inferir posibles relaciones causales.
-Se concluyó que los participantes encontraron estructura causal en el problema y la utilizaron para tomar futuras decisiones.\\
-\\
-En el segundo experimento y tercer experimento, a los participantes se les mostró una serie de hipótesis causales para ver si inducían de manera espontánea un modelo causal de la situación.\\
-\\
-Lo que estos experimentos muestran, en conclusión de los autores, es que al intentar maximizar una ganancia a largo plazo, las personas adquieren conocimiento causal del entorno, el cual afecta futuras decisiones.
-Mencionan que estos experimentos ponen en tela de juicio los modelos estándar de toma de decisiones, pues utilizando sólo valores esperados no se llegarían a los mismos resultados.\\
-\\
-Como consecuencia de sus estudios, se obtiene que los métodos basados en los niveles 1 y 2, que es donde caerían los métodos tradicionales de aprendizaje por refuerzo no son suficientes para una adecuada toma de decisiones en un sistema causal, pues estas formulaciones no cuentan con la expresividad suficiente para capturar las relaciones causales.
-Esto confirma la necesidad de pensar en modelos que permitan aprender modelos causales a la par que un agente aprende de su entorno mediante interacción y exploración.\\
-\\
-En cuanto a la formulación de un problema de decisión en sistemas causales, ya desde \cite{dawid2002influence} tenemos una idea de cómo aumentar redes bayesianas con nodos de decisión y variables de estrategia para el modelado causal en problemas de decisión.\\
-\\
-\cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta. 
-\section{Propuesta}
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-\subsection{Objetivos}
-\begin{itemize}
-\item Aprendizaje causal en línea a partir de acciones y recompensas obtenidas por un agente.
-\item Evaluación del modelo causal.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
-\end{itemize}
-\subsection{Propuesta metodológica}
-Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en incorporar un modelo de aprendizaje causal dentro a un problema de toma de decisiones.
-En el contexto de aprendizaje por refuerzo, tendremos un agente que, al tener como objetivo aprender una tarea mediante interacciones secuenciales, aprenda con cada paso un poco sobre la \textit{estructura causal} de su entorno.
-Cada decisión que el agente toma (o acción que lleva a cabo) puede verse como una intervención en el ambiente y que proporciona información sobre la estructura causal de este.
-Esta información obtenida, será utilizada para mejorar las decisiones que toma y para aprender otros aspectos estructurales (causales) sobre su ambiente.
-La intuición detrás de esto es que al aprender mediante una toma de decisiones repetida, decisiones que intervienen directamente sobre el entorno, se aprenden también relaciones causales; por ejemplo, al aprender a conducir se aprende implícitamente que ciertos movimientos del volante \textit{causan} ciertos cambios en el estado del automovil (\cite{danks2014unifying}).
-Otro ejemplo es en el contexto de videojuegos, en el cual al tomar ciertas decisiones se conoce el impacto que estas provocan en el estado del juego, conocimiento que a su vez es utilizado para tomar mejores decisiones en un futuro y que maximicen las recompensas u objetivos a largo plazo.
-No sólo hay intuición detrás de esto, pues se sabe que los seres humanos utilizan el conocimiento causal que la experiencia les provee para realizar predicciones para intervenciones en el mundo no realizadas aun (\cite{meder2008inferring}, \cite{hagmayer2009decision})\\
-%\\
-%De manera formal, consideremos un fenómeno sobre el cual se quieran hacer inferencias causales y sobre el cual poseeemos un conjunto de datos.
-Supongamos que los datos tienen la forma $%(\mathbf{X},\mathbf{Y})$ donde $\mathbf{X}$ son un conjunto de atributos, así como acciones llevadas a cabo, y $\mathbf{Y}$ son los efectos observados.
-Entonces, es necesario %\textit{traducir} estos datos a la forma $(s \in \mathcal{S},a \in \mathcal{A},r \in \mathcal{R})$ de modo que las características, atributos, etc del problema original sean mapeados a estados, las acciones tomadas, a acciones en el mundo de RL, y las respuestas obtenidas, al reward y cambio de estado otorgado por el ambiente.
-Evidentemente esta no es una tarea trivial, pues es necesario también definir un \textit{task} adecuado para el agente que le permita interactuar con su entorno de manera tal que aprenda, o descubra, las relaciones causales sobre los datos originales.
-La definición del task a resolver implica definir una función recompensa; además, una transición entre estados es importante.
-Notemos que todos estos elementos deben ser definidos de manera que el agente interactue con su mundo de tal forma que lo que \textit{aprenda} nos sirva para identificar relaciones causales.\\
-%\\
-%Como ejemplo de lo anterior, podríamos tener un conjunto de datos de la forma $\mathbf{X} = (s,t)$ donde $s$ es enfermedad, $t$, tratamiento otorgado y $\mathbf{Y}=\{ \text{curado}, \text{ no curado} \}$.
-Definimos un agente de RL tal que los estados que recibe del ambiente son las características del paciente, la acción que toma es otorgar o no el medicamento, y la recompensa es $1$ si el paciente se mejoró.
-La tarea que este agente tiene que lograr es maximizar el número de pacientes curados; notemos que la tarea del agente, o el agente mismo, no tienen por qué tener interpretación en términos del problema original; es decir, no tiene por qué existir una interpretación del agente en términos de un médico, un paciente, etc.
-Una vez que el agente aprenda a resolver esta tarea, utilizaremos lo aprendido por él de la siguiente manera: si para un paciente del conjunto original de datos el agente propone el tratamiento, entonces es causal que el tratamiento mejoró al paciente.\\
-\\
-El trabajo de \cite{garnelo2016towards} nos muestra que efectivamente pueden aprenderse aspectos estructurales sobre el entorno mientras el agente aprende, incluso logrando así mejorar su desempeño.
-Además, muestra la importancia de tener una semántica que permita hablar de la estructura del entorno.
-Ellos utilizaron lógica de primer orden, pero un modelo gráfico sería más expresivo y adaptable.\\
-\\
-A su vez, el trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
-Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
-\cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.\\
-\\
-\begin{center}
-\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/diagrama.png}
-\end{center}
-\newpage
 \bibliographystyle{apalike}

### 2017-12-20 Primer_Semestre/BurnOut/lib/load_data.R

@@ -1,0 +1,10 @@
+setwd("~/datos_Adi")
+data<-read.csv("diabetes_marked_datos_sim.csv")
+
+data$paciente<-as.factor(data$paciente)
+tapply(data$presc.code, data$paciente,rle)
+by(data,data$paciente,rle)
+
+test<-c(1, 1, 1, 2, 3, 4, 4, 4, 4, 5, 5, 6, 6, 6, 6)
+
+duraciones<-rle(test)[1]

### 2018-01-08 Segundo_Semestre/Propuesta/Propuesta.tex
Ordenando la intro

@@
 \section{Introducción}
-El problema al cual nos enfrentamos consiste en el aprendizaje de relaciones causales a partir de datos.
-Los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.
-Al tratar de averiguar relaciones causales, es común contar con información previa sobre el fenómeno dentro del cual se quiere investigar causalidad, por lo que recurrimos a la estadística bayesiana para el manejo de esta información inicial.\\
+El problema que estamos tratando de atacar consiste en descubrir relaciones causales que se encuentren presentes en diversos fenómenos.
+El enfoque que se propone es descubrir estas relaciones mediante \textit{interacción} con el entorno.
+En el contexto de inferencia causal, los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.\\
 \\
-La primera parte de este resumen consiste en Redes Bayesianas, que son objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
+La primera parte de este resumen consiste en el estudio clásico de relaciones causales, el cual depende fuertemente de Redes Bayesianas, las cualesson objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
 Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
 \\
-Veremos que al intentar obtener relaciones causales a partir de información proveniente del mundo real necesitamos obtener las entidades que se relacionan entre sí de manera causal, por lo que haremos un repaso de técnicas utilizadas en extracción de relaciones.\\
+Después, veremos que existe un área del aprendizaje de máquina llamado Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.\\
 \\
-Por otro lado, en la literatura proveniente de Psicología Computacional y Ciencias Cognitivas, se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
-Se sabe también que el conocimiento causal adquirido es utilizado para mejorar las futuras decisiones tomadas.\\
+Además, las intuiciones iniciales sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
+Se sabe también que el conocimiento causal adquirido puede ser utilizado para mejorar las futuras decisiones tomadas.\\
 \\
-Finalmente, se discutirá un poco sobre Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.
 Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.

### 2018-01-24 Segundo_Semestre/Propuesta/Propuesta.tex
Añadiendo sobre teorema vNM

@@
 
+\section{Teoría de la Decisión}
+\subsection{von Neumann-Morgenstern}
+Según narra \cite{gilboa2009decision}, fue Daniel Bernoulli quién habló por primera vez de la maximización de la utildiad esperada; esto, al estudiar la \textit{paradoja de San Petersburgo}, la cual consiste en lo siguiente: un tomador de decisiones se enfrente a una serie de lanzamientos consecutivos de una moneda justa.
+Se considera un juego basado en los resultados de estas monedas, el cual terminará la primera vez que la moneda arroje una cara, y si esto ocurriera en el $n$-ésimo lanzamiento, el concursante recibirá $2^n$ pesos.
+La \textit{paradoja} consiste en que la mayoría de las personas, según registró Bernoulli, sólo está dispuesta a pagar una cantidad finita de dinero por entrar a este juego, cuyo valor esperado es infinito.
+La paradoja fue resuelta cuando Bernoulli se dio cuenta que las personas estaban maximizando el valor esperado de una \textit{función de utilidad}, pero no directamente el valor esperado del dinero.
+Resulta que si esta función es logarítmica, entonces la utilidad esperada es finita.
+\\
+\\
+El Teorema original de von Neumann y Morgenstern está formulado en términos de árboles de decisión y loterías compuestas.
+Más tarde \cite{jensen1967introduction} lo formuló de manera más compacta al suponer que las loterías compuestas son simplificadas de acuerdo a la fórmula de Bayes.\\
+\\
+Sea $X$ un conjunto de \textit{alternativas}, los objetos a elegir son loterías de soprte finito; de manera formal, definimos
+\[ L = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
+Una mezcla de loterías se puede definir de la siguiente manera: se toma una probabilidad $\alpha$ y se define una lotería $l^\alpha = \alpha P + (1- \alpha) Q$.\\
+\\
+Consideremos una relación  $\succeq L \times L$. El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva
+\item A2. Continuidad: Para toda $P,Q,R \in L$ tales que $P \succ Q \succ R$ existen $\alpha$ y $\beta$ tales que:
+\[  \alpha P + (1-\alpha) R \succ Q \succ \beta P + (1-\beta) R\]
+\item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si $\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R$.
+
+\end{itemize}
+\subsection{De Finnetti}
+\subsection{Savage}
 \section{Aprendizaje por Refuerzo}
@@
 \end{itemize}
-\section{Deep Reinforcement Learning}
-\subsection{Mejoras a los métodos básicos}
-	\begin{itemize}
-		\item Programación Dinámica con aproximación de funciones.
-		\item Cuando el espacio de estados es muy grande, no es factible utilizar una representación en una \textit{look up table}, por lo que es necesario utilizar otro tipo de aproximadores (\cite{busoniu2010reinforcement})
-	\end{itemize}
-\subsection{Deep Learning}
-	\begin{itemize}
-		\item Redes neuronales profundas como aproximadores generales.
-		\item Redes Convolucionales para poder tomar como input los pixeles de un juego directamente.
-	\end{itemize}
 \subsection{Value based Reinforcement Learning:}
@@
 \\
-Teorías de cómo las personas usan datos de covariación para aprender modelos gráfios causales pueden dividirse en enfoques racionales y heurísticos (según  \cite{wellen2012learning}).
-El enfoque racional modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana.
-Estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
+Según \cite{wellen2012learning} es posible dividir las teorías de cómo las personas usan información estadística para aprender modelos causales en enfoques racionales y heurísticos.
+El enfoque primero modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana; estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
 \\
@@
 \subsection{Propuesta metodológica}
-Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en incorporar un modelo de aprendizaje causal dentro a un problema de toma de decisiones.
-En el contexto de aprendizaje por refuerzo, tendremos un agente que, al tener como objetivo aprender una tarea mediante interacciones secuenciales, aprenda con cada paso un poco sobre la \textit{estructura causal} de su entorno.
-Cada decisión que el agente toma (o acción que lleva a cabo) puede verse como una intervención en el ambiente y que proporciona información sobre la estructura causal de este.
-Esta información obtenida, será utilizada para mejorar las decisiones que toma y para aprender otros aspectos estructurales (causales) sobre su ambiente.
+Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en diseñar un agente que incorpore un modelo causal dentro a un problema de toma de decisiones.
+Este agente buscará aprender un modelo causal de su entorno al mismo tiempo que aprende una \textit{política}.
+El agente aprenderá con cada paso un poco sobre la \textit{estructura causal} de su entorno.\\
+\\
+Cada decisión que el agente toma (acción que lleva a cabo) altera el estado del mundo.
+La información obtenida a partir de cada acción, será utilizada para actualizar el conocimiento causal que se tenga hasta el momento, esto con el fin mejorar las decisiones que toma y para aprender otros aspectos estructurales (causales) sobre su ambiente.
 La intuición detrás de esto es que al aprender mediante una toma de decisiones repetida, decisiones que intervienen directamente sobre el entorno, se aprenden también relaciones causales; por ejemplo, al aprender a conducir se aprende implícitamente que ciertos movimientos del volante \textit{causan} ciertos cambios en el estado del automovil (\cite{danks2014unifying}).
@@
 No sólo hay intuición detrás de esto, pues se sabe que los seres humanos utilizan el conocimiento causal que la experiencia les provee para realizar predicciones para intervenciones en el mundo no realizadas aun (\cite{meder2008inferring}, \cite{hagmayer2009decision})\\
-%\\
-%De manera formal, consideremos un fenómeno sobre el cual se quieran hacer inferencias causales y sobre el cual poseeemos un conjunto de datos.
-Supongamos que los datos tienen la forma $%(\mathbf{X},\mathbf{Y})$ donde $\mathbf{X}$ son un conjunto de atributos, así como acciones llevadas a cabo, y $\mathbf{Y}$ son los efectos observados.
-Entonces, es necesario %\textit{traducir} estos datos a la forma $(s \in \mathcal{S},a \in \mathcal{A},r \in \mathcal{R})$ de modo que las características, atributos, etc del problema original sean mapeados a estados, las acciones tomadas, a acciones en el mundo de RL, y las respuestas obtenidas, al reward y cambio de estado otorgado por el ambiente.
-Evidentemente esta no es una tarea trivial, pues es necesario también definir un \textit{task} adecuado para el agente que le permita interactuar con su entorno de manera tal que aprenda, o descubra, las relaciones causales sobre los datos originales.
-La definición del task a resolver implica definir una función recompensa; además, una transición entre estados es importante.
-Notemos que todos estos elementos deben ser definidos de manera que el agente interactue con su mundo de tal forma que lo que \textit{aprenda} nos sirva para identificar relaciones causales.\\
-%\\
-%Como ejemplo de lo anterior, podríamos tener un conjunto de datos de la forma $\mathbf{X} = (s,t)$ donde $s$ es enfermedad, $t$, tratamiento otorgado y $\mathbf{Y}=\{ \text{curado}, \text{ no curado} \}$.
-Definimos un agente de RL tal que los estados que recibe del ambiente son las características del paciente, la acción que toma es otorgar o no el medicamento, y la recompensa es $1$ si el paciente se mejoró.
-La tarea que este agente tiene que lograr es maximizar el número de pacientes curados; notemos que la tarea del agente, o el agente mismo, no tienen por qué tener interpretación en términos del problema original; es decir, no tiene por qué existir una interpretación del agente en términos de un médico, un paciente, etc.
-Una vez que el agente aprenda a resolver esta tarea, utilizaremos lo aprendido por él de la siguiente manera: si para un paciente del conjunto original de datos el agente propone el tratamiento, entonces es causal que el tratamiento mejoró al paciente.\\
 \\
@@
 \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.\\
-\\
+Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.
 \begin{center}
@@
 \end{center}
+\subsection{Pasos a seguir}
+\begin{itemize}
+\item ¿Por qué esto sería mejor que trabajo de Garnelo?
+\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? 
+\item Tengo una política aprendida, ¿cómo obtengo info causal?
+\item Exploración vs explotación causal: ¿cómo decido si seguir la info causal o tomar acciones exploratorias para descubrir relaciones causales? 
+\item ¿Un dominio de interés? 
+\end{itemize}
+\subsection{Caso de un MDP finito}
+En el caso de un MDP finito completamente especificado; es decir, en el cual conocemos las probabilidades de transición y las recompensas.
+Un modelo causal sobre este dominio opera sobre las acciones, pues son estas las que \textbf{\textit{suponemos}} que \textit{causan} algo; esto es,
+\[ a \xrightarrow[]{\text{causa}} \textrm{estado del mundo}. \]
+Entonces, en un modelo gráfico causal necesitaremos tener \textit{nodos de estado}, así como \textit{nodos de acciones} (\textit{causal decision graphs}).\\
+\\
+Es claro que los estados del mundo afectarán la siguiente decisión, pero no causan nada por sí mismos; en todo caso, el estado afeca la decisión pero sólo a través de la política (el valor del estado).\\
+\\
+El objetivo, ahora, es aprender una política que incorpore conocimiento causal, el cual a su vez fue aprendido mediante interacción.\\
+\\
+¿El valor de un estado debe cambiar en el caso causal?
+Para que el valor de un estado se vea afectado por lo que pueda causar el agente a partir de ese estado se necesita algún tipo de función de utilidad, ¿no?
+No lo sé...
 \newpage

### 2018-01-30 Segundo_Semestre/Propuesta/Propuesta.tex
cambios

@@
 \item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si $\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R$.
-
 \end{itemize}
+\begin{teo}
+Una relación $\succeq L \times L$ satisface los axiomas A1 - A3 si y sólo sí existe una función $u: X \to \mathbb{R}$ tal que para toda $P,Q \in L$ se tiene que $P \succeq Q$ si y sólo si 
+\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
+\end{teo}
+Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
 \subsection{De Finnetti}
+¿Cuál es la definición de probabilidad subjetiva?
+\cite{gilboa2009decision} sugiere que para saber qué son debemos explorar cómo se usan estas, y una posible manera de uso es al maximizar una utilidad esperara y esto es lo que hace el Teorema de DeFinnetti.\\
+\\
+De Finetti consideró apuestas sobre \textit{estados del mundo}.
+Consideremos que existen $1,...,n$ estados en el mundo.
+Una apuesta es una función que toma valores en el espacio de estados y los mapea a valores monetarios; es decir, una apuesta $x$ es un vector de $n$ elementos en donde $x_i$ es la ganancia en el caso de que la apuesta resulte en el estado $i$ como el ganador.
+Consideramos una relación $\succeq \subseteq X \times X$.
+Y consideremos los siguientes axiomas:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva.
+\item A2. Continuidad: Para cada $x \in X$, los conjuntos $\{ y | x \succ y \}$ y $\{ y | y \succ x \}$ son abiertos.
+\item A3. Aditividad: Para toda $x,y,z \in X$, $x \succeq y$ si y sólo si $x+z \succeq y+z$.
+\item A4 Monotonicidad: Para toda $x,y \in X$, si se cumple que $x_i \geq y_i$ entonces se tiene que $x \succeq y$.
+\item A5 No-trivialidad: Existen $x,y \in X$ tales que $x \succ y$.
+\end{itemize}
+Notemos que el axioma A3 sólo aplica bajo un esquema de neutralidad al riesgo.
+\begin{teo}
+La relación $\succeq \subseteq X \times X$ satisface A1-A5 si y sólo si existe un vector de probabilidad $p$ de tamaño $n$ tal que para toda $x,y \in X$ se cumple que
+\[ x \succeq y \textrm { si y sólo si } px \geq py. \]
+\end{teo}
+Entonces, la definición de probabilidad subjetiva sería aquella probabilidad utilizada por un tomador de decisiones para escoger una acción mediante el criterio de utilidad esperado.
 \subsection{Savage}
+El Teorema de von Neumann-Morgenstern indica cómo obtener una función de utilidad si las probabilidades están bien definidas; el Teorema de De Finnetti dice cómo obtener probabilidades si conocemos las utiidades de un tomador de decisiones.
+El Teorema de Savage obtiene al mismo tiempo probabilidades y utlidades.
+Su modelo asume sólo estados del mundo y sus resultados.
+Por lo tanto, los axiomas propuestos por Savage son bastante más abstractos y menos interpretables, pero se obtiene la ventaja de que sus axiomas son equivalentes a la existencia de una función de utilidad y de una medida de probabilidad (subjetiva) tales que las decisiones se llevan a cabo al maximizar utilidad esperada respecto a esta función de utilidad y esta medida de probabilidad.\\
+\subsubsection{Estados, resultados y acciones}
 \section{Aprendizaje por Refuerzo}

### 2018-01-30 Segundo_Semestre/Seminario/Semana2/Circles_State_of_Art.tex
Circulos de mi state of the art segun Bundy

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{natbib}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{State of The Art Circles}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\section{Outer Circle}
\begin{itemize}
\item (\cite{suppes1970probabilistic})
\item \cite{holland1986statistics}
\item \cite{sutton1998reinforcement}
\item Causal decision theory (\cite{joyce1999foundations}): Making decisions using causal knowledge; i.e. I know that $A$ \textit{causes} $B$, so it can be predicted what will happen if $A$ is chosen, or not, over other options. 
\item (\cite{spirtes2000causation})
\item cite{pearl2009causality},
\item Causal graphical models ( \cite{koller2009probabilistic}): Causal semantics over probabilistic graphical models.
\item Decision Theory Under Uncertainty (\cite{gilboa2009decision}).
\item \cite{pearl2014probabilistic}
\item  \cite{danks2014unifying}
\item \cite{gershman2015reinforcement}
\end{itemize}
\section{Middle Circle}
\begin{itemize}
\item \cite{baxter2001infinite}
\item \cite{boyan2002technical}
\item Decision problems and causal models: \cite{dawid2002influence} show how causal Bayesian Nets can be augmented with decision nodes as to model decision problems that use causal information.
\item \cite{konda2003onactor}
\item \cite{richter2007natural}
\item \cite{hagmayer2009decision}
\item \cite{meder2009role}
\item \cite{meder2010observing}
\item On-line learning in causal models(\cite{wellen2012learning})):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop.
\item \cite{grondman2012survey}
\item Policy Gradients  \cite{DBLP:journals/corr/MnihBMGLHSK16}
\item \cite{lopez2015towards}
\item \cite{NIPS2016_6234}
\item \cite{NIPS2016_6266}
\item \cite{NIPS2016_6059}
\item \cite{NIPS2016_6233}
\item \cite{NIPS2016_6042}
\item \cite{DBLP:journals/corr/SchulmanMLJA15}
\item \cite{goudet2017learning}
\item \cite{firoiu2017beating}
\item \cite{leibo2017multi}
\item \cite{amarjyoti2017deep}
\item \cite{NIPS2017_7019}
\item \cite{NIPS2017_7084}
\item \cite{NIPS2017_7134}
\item \cite{NIPS2017_6954}
\item \cite{NIPS2017_7134}
\item \cite{NIPS2017_6950}
\item \cite{NIPS2017_7007}
\item \cite{NIPS2017_6874}
\item \cite{NIPS2017_7123}
\item \cite{NIPS2017_7017}
\item \cite{NIPS2017_7223}
\item \cite{NIPS2017_6875}
\item \cite{DBLP:journals/corr/HuangT17}
\item \cite{guo2014deep}
\item \cite{DBLP:journals/corr/LevineFDA15}
\item \cite{DBLP:journals/corr/GuLSL16}
\item \cite{DBLP:journals/corr/ArulkumaranDSB16}
\item \cite{DBLP:journals/corr/ZahavyBM16}
\item \cite{DBLP:journals/corr/ChenDHSSA16}

\end{itemize}
\section{Inner Circle}
\begin{itemize}
\item Causal learning through sequential decision making (\cite{hagmayer2013repeated}).
\item \cite{NIPS2016_6195}
\item \cite{garnelo2016towards}
\end{itemize}
\newpage
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
\end{document}

### 2018-01-30 Segundo_Semestre/Seminario/Semana2/Circles_State_of_Art.tex
Añadiendo los de multiagent

@@
 \item \cite{grondman2012survey}
-\item Policy Gradients  \cite{DBLP:journals/corr/MnihBMGLHSK16}
+\item Policy Gradients  \cite{DBLP:journals/corr/MnihBMGLHSK16}: the authors of the original DQN paper who have shown Policy Gradients to work better than Q Learning when tuned well.
+PG is preferred because it is end-to-end: there’s an explicit policy and a principled approach that directly optimizes the expected reward.
 \item \cite{lopez2015towards}
@@
 \item \cite{DBLP:journals/corr/ChenDHSSA16}
-
+\item \cite{goudet2017learning}
+\item \cite{leibo2017multi}
+\item \cite{de2006learning}
+\item \cite{busoniu2008comprehensive}
+\item \cite{nowe2012game}
+\item \cite{bucsoniu2010multi}
+\item \cite{shoham2007if}
+\item \cite{panait2005cooperative}
+\item \cite{hernandez2017survey}
 \end{itemize}

### 2018-02-01 Segundo_Semestre/Propuesta/Propuesta.tex
Añadiendo subseccion de motivacion en la parte de propuesta

@@
 En el contexto de inferencia causal, los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.\\
+\\
+\cite{lake2017building} argumentan que los sistemas inteligentes actuales, a pesar de sus impresionantes logros, aun están lejos de alcanzar una verdadera inteligencia; uno de los requisitos para esto es que puedan construir modelos causales que sean interpretables, y no sólo reconocer patrones estadísticos.
 \\
@@
 Si e conoce la existencia de $V$ y además podemos observarla.
+
+\section{Modelos Gráficos Probabilistas}
+
 \section{Modelos gráficos causales}
@@
 \begin{itemize}
-\item Evaluación de política en Diferencias Temporales
-\item TD(0)
-\item SARSA: Es un método \textit{on-policy}
-\item Q-Learning: Es un método \textit{off-Policy}
-\item Métodos Actor-Critic
+\item Evaluación de política en Diferencias Temporales.
+\item TD(0).
+\item SARSA: Es un método \textit{on-policy}.
+\item Q-Learning: Es un método \textit{off-Policy}.
+\item Métodos Actor-Critic.
 \end{itemize}
-\subsection{Value based Reinforcement Learning:}
+\subsection{Value based Reinforcement Learning: Deep Q-Networks}
 	\begin{itemize}
-	   \item Representar la función valor mediante una \textit{Q-network} con pesos $w$:
-	   \[Q(s,a,w) \approx Q^\ast(s,a) \]
+	\item Los sorprendentes avances de Deep Learning en visión computacional llevaron de manera natural a plantear la pregunta de cómo utilizar estas nuevas técnicas para el aprendizaje por refuerzo. 
+	\item Las aplicaciones estándar de DL requerían grandes volúmenes de datos etiquetados, además de tener muestras independientes, pero en Reinforcement learning esto es un problema pues las secuencias de estados suelen estar altamente correlacionados.
+	\item \cite{mnih2013playing} proponen una “Deep Q-Network”, que consiste en una implementación de Q-learning con aproximación de funciones (\cite{sutton1998reinforcement}) en donde el aproximador es una red neuronal convolucional.
+	  \item Utilizando esta técnica fue como \cite{mnih2013playing} lograron que un agente aprendiera siete juegos de Atari 2600 utilizando como input únicamente la pantalla del juego. 
+	  \item Estimar la función $Q$ de manera iterativa es muy costoso, pues además de requerir un modelo completo del sistema cada secuencia de estados se utiliza por separado para las estimaciones.
+	  \item Por esto, es mejor utilizar alguna función aproximadora para poder generalizar. 
+	   \item En este caso, se pretende aproximar la función valor mediante una red neuronal con pesos $\theta$:
+	   \[Q(s,a,\theta) \approx Q^\ast(s,a). \]
+	   \item Esta red se entrena al minimizar una sucesión de funciones $L_i$ que está dada por:
+	   \[ L_i (\theta_i) = \mathbb{E}[(y_i - Q(s,a;\theta)^2], \textrm{ donde  }y_i = \mathbb{E}[r + \gamma \max_{a'} Q(s',a';\theta) | s,a]. \]
+	   \item Notemos que los objetivos $y_i$ dependen de los pesos de la red neuronal. 
+	   \item El gradiente de $L_i (\theta_i)$ está dado por:
+	   \[ \nabla_{\theta_i} L_i (\theta_i) = \mathbb{E}[(r + \gamma \max_{a'} Q(s',a';\theta_{i-1}) - Q(s,a;\theta_i)) \nabla_{\theta_i} Q(s,a;\theta_i).] \]
+	   \item Para optimizar la función $L$ se utiliza gradiente estocástico.
+	   \item Notemos que este es un método \textit{model-free} y \textit{off-policy}.
+	   \item Notemos que las secuencias consecutivas de estados y acciones están fuertemente correlacionadas, por lo que el aprendizaje podría ser ineficiente.
+	   \item Además, al aprender \textit{on-policy}, los parámetros actuales determinan los siguientes estados sobre los cuales se entrenará, por lo que pueden surgir loops que afecten la convergencia del aprendizaje.
+	   \item Para resolver estos problemas se utiliza Experience Replay (\cite{Lin1992}). Esta técnica consiste en guardar las experiencias del agente para cada tiempo $t$; esto es 
+	   \[ e_t = (s_t, a_t, r_t, s_{t+1}). \]
+	   \item Cada $e_t$ se almacena en un conjunto $\mathcal{D} = e_1 , ... , e_N$.
+	   \item En un loop interno del algoritmo, se tomarán aleatoriamente muestras de experiencia las cuales serán utilizadas para la actualización de pesos. 
 	   \item Q-learning:
-        \begin{itemize}	  
-        	\item  Sabemos que los valores óptimos deben satisfacer la ecuación de Bellman:
-        	\[ Q^\ast(s,a)=\mathbb{E}_{s'}[r + \gamma \max_{a'} Q^\ast(s',a')|s,a]. \]
-        	\item Tratamos el término $r + \gamma \max_{a'} Q^\ast(s',a',w)$ como el objetivo.
-        	\item Minimizar error cuadrático medio mediante Gradiente Estocástico:
-        	\[ I = (r + \gamma \max_{a'} Q^\ast(s',a',w) - Q(s,a,w))^2. \]
-        	\item Converge a $Q^\ast$ utilizando una representación tabular.
-        	\item Diverge al utilizar redes neuronales debido a la alta correlación entre muestras y la no-estacionareidad del objetivo.
-        	\item Lo resolvemos con Experience Replay (\cite{Lin1992}),  el cual consiste en guardar todas las experiencias de modo que al entrenar la red tomar sólo mini-batches aleatorios  en vez de utilizar la transición más reciente.
-	    \end{itemize}
 	   \item Aplicaciones de DQN.
@@
 	  \end{itemize}
-\item Mejoras posteriores a DQN
+	      	 \end{itemize}  	
+\subsection{Mejoras posteriores a DQN}
    	   \begin{itemize}
@@
    	   \end{itemize}
-    	 \end{itemize}  	   
-\subsection{Policy Based DRL}
-Resulta que Deep Q-Learning no es un gran algoritmo, y en su lugar, \cite{DBLP:journals/corr/MnihBMGLHSK16} proponen utilizar directamente Policy-Gradients.
+\subsection{Método Asíncronos}  
+\subsection{Policy Based DRL: Policy Gradients}
+Resulta que Deep Q-Learning no es un gran algoritmo, incluso los autores del paper original (\cite{DBLP:journals/corr/MnihBMGLHSK16}) mencionan sus desventajas. 
+Otra idea es atacar directamente la política.
 La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
@@
 \cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta.
-\section{Propuesta}
+
+\subsection{Experimentos secuenciales: Causal Bandits}
+
+\section{Mi Propuesta}
 \subsection{Pregunta de Investigación}
@@
 \begin{itemize}
-\item Aprendizaje causal en línea a partir de acciones y recompensas obtenidas por un agente.
-\item Evaluación del modelo causal.
+\item Plantear un modelo causal dado un ambiente.
+\item Aprendizaje estructural del modelo causal.
+\item Evaluación del modelo causa (queries significativas)l.
 \item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
@@
 \end{itemize}
-\subsection{Propuesta metodológica}
-Dado el contexto proporcionado en las secciones anteriores, la propuesta consiste en diseñar un agente que incorpore un modelo causal dentro a un problema de toma de decisiones.
+\subsection{Motivación}
+Tomando en cuenta el contenido de la secciones anteriores, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción
 Este agente buscará aprender un modelo causal de su entorno al mismo tiempo que aprende una \textit{política}.
-El agente aprenderá con cada paso un poco sobre la \textit{estructura causal} de su entorno.\\
+Con cada acción que toma, el agente aprenderá  un poco sobre la \textit{estructura causal} de su entorno y utilizará este conocimiento para la futura toma de decisiones.\\
 \\
@@
 \\
+El trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
+Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
+\\
+\\
+\cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
+\\
+\\
 El trabajo de \cite{garnelo2016towards} nos muestra que efectivamente pueden aprenderse aspectos estructurales sobre el entorno mientras el agente aprende, incluso logrando así mejorar su desempeño.
 Además, muestra la importancia de tener una semántica que permita hablar de la estructura del entorno.
-Ellos utilizaron lógica de primer orden, pero un modelo gráfico sería más expresivo y adaptable.\\
+Ellos utilizaron lógica de primer orden, pues están más interesados en hacer razonamientos \textit{de alto nivel} sobre el entorno, que involocura más que las relaciones causales, y por esto ellos se comprometen con lógica de primer orden, que al restringirnos al caso causal se queda corta pues los modelos gráficos probabilistas son más expresivos, además de modulares y montónicos.\\
 \\
-A su vez, el trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
-Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
-\cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Entonces, queda abierta la pregunta de qué pasaría con un agente que se enfrente a un problema de aprendizaje por refuerzo, queremos saber cómo este agente aprenderá paso a paso un modelo causal y lo utilizará para tomar mejores decisiones.
+\cite{lattimoreNIPS2016} explora la pregunta de cómo escoger \textit{la mejor intervención} dado un modelo causal a través de una serie de intervenciones y observaciones secuenciales; por ejemplo, en el caso en el que un granjero quiera maximizar sus cosechas, y sabe que ese rendimiento está afectado por el uso de fertilizante, la humedad y la exposición al sol, pero sólo tiene los recursos para llevar a cabo sólo una de estas modificaciones en cada temporada.
+El algoritmo de Lattimore le dirá, después de $T$ observaciones, cuál es la mejor intervención a llevar a cabo con el fin de maximizar esa cantidad.
+Este podría ser un primer paso hacia responder la pregunta \textit{¿dado un modelo causal, qué acción tomar?}
+\\
+\cite{ortega2014generalized} exploran un esquema de muestreo llamado Muestreo de Thompson, en el cual un agente muestrea una de entre varias acciones de acuerdo a la probabilidad subjetiva que tenga el agente de que esta sea la mejor acción.
+Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).
+\subsection{La propuesta}
+Diseñar un agente que tenga como objetivo aprender a llevar a cabo una tarea.
+El aprendizaje se dará por interacción con el ambiente y además contará con un modelo causal que vaya aprendiendo paso a paso y del cual se apoyará para  tomar mejores decisiones.
 \begin{center}
@@
 \item ¿Por qué esto sería mejor que trabajo de Garnelo?
-\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? 
-\item Tengo una política aprendida, ¿cómo obtengo info causal?
+\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? Aquí ayuda el artículo de \cite{lattimoreNIPS2016}
+\item Tengo una política aprendida, ¿cómo descubro estructura causal? Aquí puede ayudar el artículo de \cite{ortega2014generalized}
+\item ¿Qué aprendo primero?
+\item ¿Cómo le doy más peso a la información explicativa que a la información reactiva?
 \item Exploración vs explotación causal: ¿cómo decido si seguir la info causal o tomar acciones exploratorias para descubrir relaciones causales? 
@@
 \end{itemize}
-\subsection{Caso de un MDP finito}
+\subsection{Ejemplo: Caso de un MDP finito}
 En el caso de un MDP finito completamente especificado; es decir, en el cual conocemos las probabilidades de transición y las recompensas.
@@
 \bibliography{Bibliografia}
+\appendix
+\section{Teoría de Probabilidad}
+\section{Estadística Bayesiana}
+
 \end{document}

### 2018-02-05 Segundo_Semestre/Propuesta/Propuesta.tex
quitando bullets de programación dinámica

@@
 Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
-				\begin{itemize}
-					\item Evaluación de una política: Teniendo una política cualquiera $\pi$, ¿cómo obtener la función $V^\pi$?\\
+La idea fundamental en programación dinámica es poder asignar a cada estado del sistema un valor numérico que represente el retorno esperado, a largo plazo, de ese estado; es decir, estamos valorando cada estado según lo que esperemos obtener de él a muy largo plazo.
+\\
+\\
+Recordemos que una política $\pi$ especifica una manera de actuar en cada estado; es decir, a cada estado del sistema le asocia una acción.
+Si ya conocemos una política cualquiera, la pregunta natural es ¿cómo obtener la función $V^\pi$?\\
+\\
 Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
@@
 donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$.
-Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.
-					\item Mejora de una política: Dada una política $\pi$ y estando en un estado $s$ nos preguntamos qué pasaría si en vez de escoger la acción $\pi(s)$ escogieramos cualquier otra acción $a$.
-La consecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
-Entonces, la clave es comparar esto con $V^{\pi}$.
-De ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
-Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.
+Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.\\
+\\
+Sabemos que además de la función $V$ tenemos a la función $Q$, la cual valora parejas estado-acción.
+Utilizando esta función, podemos preguntarnos qué pasaría si en vez de escoger la acción $a=\pi(s)$ escogieramos cualquier otra acción $a$.
+Claramente, onsecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
+Entonces, si esto resulta ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
+Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.\\
+\\
 Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
 \[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
-					Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.
-					\item Iteración de políticas: Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
+Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.\\
+\\
+Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
 \[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
 pues cada política está garantizado que es mejor que la anterior.
-					\item Iteración de valor: El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite. Podemos truncar la etapa de policy evaluation.
+El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite.
+Una opción es truncar la etapa de policy evaluación y hacer
 \[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
-					\item Iteración de política generalizada (GPI): Notar que la iteración de políticas consiste en dos procesos que interactuan entre sí, pero de manera que cada proceso se hace de manera completa antes de comenzar el otro lleva a preguntarse si esto puede hacerse de manera más eficiente.
-Se conoce como iteración de políticas generalizada (GPI) a la idea general de que los procesos de evaluación y mejora de política interactuen entre sí.
-					\item Limitaciones: Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
+Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
 Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
-				\end{itemize}
 \subsection{Métodos básicos de solución: Model-free}

### 2018-02-05 Segundo_Semestre/Seminario/Semana3/Presentacion.Rmd
inner circle and introduction

---
title: "Presentation"
author: "Mauricio Gonzalez"
date: "5/2/2018"
output: ioslides_presentation
bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
---

# Introduction
## Intuition
- While playing a videogame or learning how to drive, human beings are forced to *make decisions* and to *actively intervene* in the world.
- This active *decision-making* process modifies the state of the world, which allows us to learn about *causal relations* that hold in our environment
- We ask ourselves how an *intelligent agent* can acquire and *use causal knowledge* while engaging with its environment to achieve some task.

# State of the Art
## Main groups and people
* Causality
    + Carnegie Mellon (Spirtes)
    + UC Los Angeles (Pearl)
    + MIT (Joshua Tenenbaum)
* Reinforcement Learning
    + Google DeepMind (David Silver, Alex Graves, V. Mnih, Shanahan, Garnelo)
    + Oxford (Nando De Freitas, Shimon Whiteson)
    + Cambridge (Ghahramani)
    + Alberta University (Richard Sutton, Csaba Szepesvári)
    + UC Berkeley (Pieter Abbeel)
    + Carnegie Mellon ()
    + Imperial College (Murray Shanahan, Marta Garnelo)
    + University College London

----

* Cognitive Science
    + Carnegie Mellon (David Danks)
    + MIT (Josh Tenenbaum)
    + UC Berkeley (Tom Griffiths)
    + Edinburgh University (Andy Clark)
    + Göttingen Institut für Psychologie (York Hagmayer)
    + Max Planck Institute for Human Development (Bjorn Meder)

## Outer Circle

Reference                  | Summary                                |
------------------------   |------------------------                |
[holland1986statistics]    | |
[@pearl1988probabilistic]  | Book on Networks for probabilistic reasoning|
[@spirtes2000causation]    | Classic General book on Causality      |
[@joyce1999foundations]    | Decision Theory and causal information |
[@pearl2009causality]      | Book on Causality and Graphical models |
[@koller2009probabilistic] | Book on Probabilistic Graphical Models |

----

Reference                    | Summary                                    |
-----------------------------|--------------------------------------------|
[@sutton1998reinforcement]   | Classic textbook on Reinforcement Learning |
[@van2012reinforcement]      | General review                             |
[@gershman2015reinforcement] | Reinforcement Learning in the Brain        |
[@li2017deep]                | Overview of Deep Reinforcement Learning    |

----

Reference            | Summary                                                                      |
---------------------|------------------------------------------------------------------------------|
[@krynski2007role]   |
[@danks2014unifying] | Cognitive functions can be modelled as operations on graphical models        |
[@lake2017building]  | Conditions for machines to be considered intelligent: Causal learning is one |

## Middle Circle
Reference | Summary |
----------|---------|
[@dawid2002influence] | |
[@hagmayer2009decision] | |
[@meder2009role] | |
[@meder2010observing] | |
[@kemp2010learning] | |
[@wellen2012learning] | |
[@mnih2013playing] | Introducing Deep Q-Network |
[@silver2014deterministic] | |
[@mnih2015human] | |
[@DBLP:journals/corr/MnihBMGLHSK16]
[@o2016pgq] | |

## Inner Circle

Reference | Summary |
----------|---------|
[@eberhardt2008causal] ||
[@eberhardt2012number] ||
[@hauser2012two] ||
[@hagmayer2013repeated] ||
[@ortega2014generalized] ||
[@alon2015online] ||
[@lattimoreNIPS2016] ||
[@garnelo2016towards] ||

### 2018-02-07 Segundo_Semestre/Seminario/Semana3/Presentacion.Rmd
Cambios en la presentacion

@@
 date: "5/2/2018"
-output: ioslides_presentation
+output:
+  ioslides_presentation: default
+  slidy_presentation: default
 bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
@@
 
+## Fields involved
+- Learning by interaction.
+- Sequential Decision Making.
+- Causal learning.
+
+# Research Problem
+## Statement
+- Acquiring causal kwnoledge via interaction with the environment and using this knowledge to guide a decision-making process.
+
+## Relevance
+- Current RL methods are purely reactive.
+- A lot of training time is required.
+- Do not allow high-level reasoning.
+
+## Feasibility
+- We know how to learn by interaction (from Reinforcement Learning)
+- We know that causal learning can be acquired in human beings through sequential decision making (from Computational Psychology).
+- We know that *best action* can be selected from causal data (from sequential experimental design).
+
+----
+
+- Therefore, the elements exist and can be knitted together, although it will not be an easy task. 
+- The elements may exist, and although core ideas are shared in the fine details are far away between them.
+- The Devil is in the details.
+
 # State of the Art
@@
     + MIT (Joshua Tenenbaum)
+    + Max Planck Institute for Intelligent Systems (Schölkopf)
+    + University of Copenhagen  (Jonas Peters)
+    + Edinburgh (Albrecht, Subramanian Ramamoorthy)
+
+----
+
 * Reinforcement Learning
@@
     + UC Berkeley (Pieter Abbeel)
-    + Carnegie Mellon ()
+    + Carnegie Mellon (Salakhutdinov)
     + Imperial College (Murray Shanahan, Marta Garnelo)
-    + University College London
+    + University College London (Peter Dayan, Watkins, Dawid)
 
@@
 ## Outer Circle
-
 Reference                  | Summary                                |
 ------------------------   |------------------------                |
-[holland1986statistics]    | |
+[@holland1986statistics]   | General review of different views on Causality |
 [@pearl1988probabilistic]  | Book on Networks for probabilistic reasoning|
+[@heckerman1995decision] ||
 [@spirtes2000causation]    | Classic General book on Causality      |
@@
 [@koller2009probabilistic] | Book on Probabilistic Graphical Models |
+[@lopez2015towards]        | Discriminative data-driven framework for distinguishing direction of causality |
+[@goudet2017learning]      | Learning to generate data with neural networks |
 
@@
 -----------------------------|--------------------------------------------|
+[@watkins1992q]              | Q-learning algorithm                       |
 [@sutton1998reinforcement]   | Classic textbook on Reinforcement Learning |
+[@szepesvari2010algorithms]  | State of the Art Algorithms                | 
 [@van2012reinforcement]      | General review                             |
@@
 ----------|---------|
-[@dawid2002influence] | |
+[@dawid2002influence] | Influence Diagrams|
+[@eberhardt2008causal] ||
 [@hagmayer2009decision] | |
@@
 [@kemp2010learning] | |
-[@wellen2012learning] | |
-[@mnih2013playing] | Introducing Deep Q-Network |
-[@silver2014deterministic] | |
-[@mnih2015human] | |
-[@DBLP:journals/corr/MnihBMGLHSK16]
-[@o2016pgq] | |
+[@wellen2012learning] | learning causal structure in local maner via precition-error loop |
+[@eberhardt2012number] ||
 
-## Inner Circle
+----
 
@@
 ----------|---------|
-[@eberhardt2008causal] ||
-[@eberhardt2012number] ||
+[@mnih2013playing] | Introducing Deep Q-Network |
+[@silver2014deterministic] | Deterministic Policy Gradients |
+[@mnih2015human] |  Nature report on Deep Q-Networks |
+[@DBLP:journals/corr/SchulmanLMJA15] | Trust Region Policy Optimization |
+[@DBLP:journals/corr/MnihBMGLHSK16] | Asynchronous Methods for Deep Reinforcement Learning |
+[@van2016deep] ||
+[@lakkarajuNIPS2016] ||
+[@toulisNIPS2016] ||
+[@kulkarniNIPS2016] | Hierarchical Deep Reinforcement Learning |
+[@foersterNIPS2016] | Learning to Communicate with Deep Multi-Agent Reinforcement Learning |
+[@lanctot2017] | Game-Theoretic Approach to Multiagent Reinforcement Learning  |
+[@goudet2017learning] | Functional Causal Models with Generative Neural Networks |
+[@rahmanian2017] | Online Dynamic Programming  |
+[@louizos2017] |Causal Effect Inference with Deep Latent-Variable Models  |
+[@christiano2017] |Deep Reinforcement Learning from Human Preferences|
+[@o2016pgq] | Combining policy gradient and q-learning |
+[@kirsch2017mdp | Environments |
+[@zhang2017deeper] ||
+
+## Inner Circle
+Reference | Summary |
+----------|---------|
 [@hauser2012two] ||
-[@hagmayer2013repeated] ||
-[@ortega2014generalized] ||
-[@alon2015online] ||
-[@lattimoreNIPS2016] ||
-[@garnelo2016towards] ||
+[@hagmayer2013repeated] | How repeated decision making is guided by causal beliefs. |
+[@ortega2014generalized] | Thompson sampling to solve sequential adaptive control when policy is known. |
+[@alon2015online] | When feedback is specified by graph, the graph structure controls the difficulty of learning. |
+[@bramley2015staying] ||
+[@lattimoreNIPS2016] | Causal Bandits: Given a causal model, choose best intervention. |
+[@albrecht2016exploiting] ||
+[@garnelo2016towards] | Reinforcement Learning framework that builds symbolic representation of environment. |
+[@innes2018reasoning] | An agent which, starting unaware of factors on which an optimal policy depends, learns optimal behaviour for single-stage decision problems via direct experience and advice from a domain expert.|
+[@pearl2018theoretical] ||
+
+## End
+- Thnks.

### 2018-02-07 Segundo_Semestre/Seminario/Semana2/Circles_State_of_Art.tex
Añadiendo mas

@@
 \item (\cite{spirtes2000causation})
-\item cite{pearl2009causality},
+\item \cite{pearl2009causality},
 \item Causal graphical models ( \cite{koller2009probabilistic}): Causal semantics over probabilistic graphical models.
 \item Decision Theory Under Uncertainty (\cite{gilboa2009decision}).
-\item \cite{pearl2014probabilistic}
+\item \cite{pearl1988probabilistic}
+\item \cite{van2012reinforcement}
 \item  \cite{danks2014unifying}
 \item \cite{gershman2015reinforcement}
+\item \cite{lopez2015towards}
+\item \cite{lake2017building}
 \end{itemize}
@@
 \begin{itemize}
+\item \cite{watkins1992q}
 \item \cite{baxter2001infinite}
@@
 \item \cite{konda2003onactor}
+\item \cite{sanjana2003bayesian}
+\item \cite{panait2005cooperative}
+\item \cite{de2006learning}
+\item \cite{shoham2007if}
 \item \cite{richter2007natural}
+\item \cite{krynski2007role}
+\item \cite{gopnik2007bayesian}
+\item \cite{busoniu2008comprehensive}
 \item \cite{hagmayer2009decision}
@@
 \item \cite{meder2010observing}
+\item \cite{bucsoniu2010multi}
+\item \cite{kemp2010learning}
+\item \cite{hasselt2010double}
 \item On-line learning in causal models(\cite{wellen2012learning})):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop.
 \item \cite{grondman2012survey}
+\item \cite{nowe2012game}
+\item \cite{mnih2013playing}
+\item \cite{guo2014deep}
+\item \cite{silver2014deterministic}
 \item Policy Gradients  \cite{DBLP:journals/corr/MnihBMGLHSK16}: the authors of the original DQN paper who have shown Policy Gradients to work better than Q Learning when tuned well.
@@
 \item \cite{lopez2015towards}
-\item \cite{NIPS2016_6234}
-\item \cite{NIPS2016_6266}
-\item \cite{NIPS2016_6059}
-\item \cite{NIPS2016_6233}
-\item \cite{NIPS2016_6042}
+\item \cite{krynski2007role} Human reasoning under uncertainty naturally operates over causa mental models and statistical data supports correct Bayesian inference only when they can be incorporated into a causal model.
+\item \cite{van2016deep}
+\item \cite{lakkarajuNIPS2016}
+\item \cite{lakkarajuNIPS2016}
+\item \cite{toulisNIPS2016}
+\item \cite{kulkarniNIPS2016}
+\item \cite{foersterNIPS2016}
 \item \cite{DBLP:journals/corr/SchulmanMLJA15}
@@
 \item \cite{amarjyoti2017deep}
-\item \cite{NIPS2017_7019}
-\item \cite{NIPS2017_7084}
-\item \cite{NIPS2017_7134}
-\item \cite{NIPS2017_6954}
-\item \cite{NIPS2017_7134}
-\item \cite{NIPS2017_6950}
-\item \cite{NIPS2017_7007}
-\item \cite{NIPS2017_6874}
-\item \cite{NIPS2017_7123}
-\item \cite{NIPS2017_7017}
-\item \cite{NIPS2017_7223}
-\item \cite{NIPS2017_6875}
+\item \cite{nguyen2017}
+\item \cite{wei2017}
+\item \cite{wang2017}
+\item \cite{swaminathan2017}
+\item \cite{papini2017}
+\item \cite{lanctot2017}
+\item \cite{ding2017}
+\item \cite{vanseijen2017}
+\item \cite{christiano2017}
+\item \cite{louizos2017}
+\item \cite{rahmanian2017}
 \item \cite{DBLP:journals/corr/HuangT17}
-\item \cite{guo2014deep}
 \item \cite{DBLP:journals/corr/LevineFDA15}
@@
 \item \cite{goudet2017learning}
+
 \item \cite{leibo2017multi}
-\item \cite{de2006learning}
-\item \cite{busoniu2008comprehensive}
-\item \cite{nowe2012game}
-\item \cite{bucsoniu2010multi}
-\item \cite{shoham2007if}
-\item \cite{panait2005cooperative}
 \item \cite{hernandez2017survey}
+\item \cite{li2017deep}
+\item \cite{zhang2017deeper}
+\item \cite{albrecht2017autonomous}
+\item \cite{pearl2018theoretical}
 \end{itemize}
@@
 \begin{itemize}
+\item \cite{eberhardt2008causal} plantea el problema de \textit{descubrimiento causal} como un juego entre un científico vs la naturaleza en el cual la naturaleza intenta mantener ocultos sus secretos y el científico intenta descubrirlos mientras minimiza un costo.
+\item \cite{eberhardt2012number}
+\item \cite{hauser2012two}
 \item Causal learning through sequential decision making (\cite{hagmayer2013repeated}).
-\item \cite{NIPS2016_6195}
+\item \cite{ortega2014generalized} 
+\item \cite{bramley2015staying}
+\item \cite{alon2015online} Estudian problemas de aprendizaje on-line en los cuales en cada ronda un jugador escoje una acción y recibe un \textit{loss} previamente especificado.
+Estudian las \textit{feedback graphs} que básicamente dicen si al ejecutar acción $i$ vemos el loss asociado a acción $j$ y estudian cómo la estructura de este grafo afecta el aprendizaje.
+Es principalmente un paper de estructura de grafos y el \textit{aprendizaje} se define en términos de Teoría de Juegos.
+\item \cite{lattimoreNIPS2016} Quieren descubrir la mejor intervención de modo que se minimice el \textit{regret} después de $T$ rondas.
+\item \cite{albrecht2016exploiting}
 \item \cite{garnelo2016towards}
+\item \cite{innes2018reasoning}
 \end{itemize}

### 2018-02-08 Segundo_Semestre/Propuesta/Propuesta.tex
Cambios en la seccion de motivacion según el reporte del research problem que entregare

@@
 \subsection{Value based Reinforcement Learning: Deep Q-Networks}
-	\begin{itemize}
-	\item Los sorprendentes avances de Deep Learning en visión computacional llevaron de manera natural a plantear la pregunta de cómo utilizar estas nuevas técnicas para el aprendizaje por refuerzo. 
-	\item Las aplicaciones estándar de DL requerían grandes volúmenes de datos etiquetados, además de tener muestras independientes, pero en Reinforcement learning esto es un problema pues las secuencias de estados suelen estar altamente correlacionados.
-	\item \cite{mnih2013playing} proponen una “Deep Q-Network”, que consiste en una implementación de Q-learning con aproximación de funciones (\cite{sutton1998reinforcement}) en donde el aproximador es una red neuronal convolucional.
-	  \item Utilizando esta técnica fue como \cite{mnih2013playing} lograron que un agente aprendiera siete juegos de Atari 2600 utilizando como input únicamente la pantalla del juego. 
-	  \item Estimar la función $Q$ de manera iterativa es muy costoso, pues además de requerir un modelo completo del sistema cada secuencia de estados se utiliza por separado para las estimaciones.
-	  \item Por esto, es mejor utilizar alguna función aproximadora para poder generalizar. 
-	   \item En este caso, se pretende aproximar la función valor mediante una red neuronal con pesos $\theta$:
+Los sorprendentes avances de Deep Learning en visión computacional llevaron de manera natural a plantear la pregunta de cómo utilizar estas nuevas técnicas para el aprendizaje por refuerzo. 
+Las aplicaciones estándar de DL requerían grandes volúmenes de datos etiquetados, además de tener muestras independientes, pero en Reinforcement learning esto es un problema pues las secuencias de estados suelen estar altamente correlacionados.\\
+\cite{mnih2013playing} introducen “Deep Q-Network”, que consiste en una implementación de Q-learning con aproximación de funciones (\cite{sutton1998reinforcement}) en donde el aproximador es una red neuronal convolucional.
+Utilizando esta técnica fue como \cite{mnih2013playing} lograron que un agente aprendiera siete juegos de Atari 2600 utilizando como input únicamente la pantalla del juego.
+\\
+\\
+Estimar la función $Q$ de manera iterativa es muy costoso, pues además de requerir un modelo completo del sistema cada secuencia de estados se utiliza por separado para las estimaciones.
+Por esto, es mejor utilizar alguna función aproximadora para poder generalizar. En este caso, se pretende aproximar la función valor mediante una red neuronal con pesos $\theta$:
 	   \[Q(s,a,\theta) \approx Q^\ast(s,a). \]
-	   \item Esta red se entrena al minimizar una sucesión de funciones $L_i$ que está dada por:
+Esta red se entrena al minimizar una sucesión de funciones $L_i$ que está dada por:
 	   \[ L_i (\theta_i) = \mathbb{E}[(y_i - Q(s,a;\theta)^2], \textrm{ donde  }y_i = \mathbb{E}[r + \gamma \max_{a'} Q(s',a';\theta) | s,a]. \]
-	   \item Notemos que los objetivos $y_i$ dependen de los pesos de la red neuronal. 
-	   \item El gradiente de $L_i (\theta_i)$ está dado por:
+Notemos que los objetivos $y_i$ dependen de los pesos de la red neuronal. El gradiente de $L_i (\theta_i)$ está dado por:
 	   \[ \nabla_{\theta_i} L_i (\theta_i) = \mathbb{E}[(r + \gamma \max_{a'} Q(s',a';\theta_{i-1}) - Q(s,a;\theta_i)) \nabla_{\theta_i} Q(s,a;\theta_i).] \]
-	   \item Para optimizar la función $L$ se utiliza gradiente estocástico.
-	   \item Notemos que este es un método \textit{model-free} y \textit{off-policy}.
-	   \item Notemos que las secuencias consecutivas de estados y acciones están fuertemente correlacionadas, por lo que el aprendizaje podría ser ineficiente.
-	   \item Además, al aprender \textit{on-policy}, los parámetros actuales determinan los siguientes estados sobre los cuales se entrenará, por lo que pueden surgir loops que afecten la convergencia del aprendizaje.
-	   \item Para resolver estos problemas se utiliza Experience Replay (\cite{Lin1992}). Esta técnica consiste en guardar las experiencias del agente para cada tiempo $t$; esto es 
+Para optimizar la función $L$ se utiliza gradiente estocástico; este es un método \textit{model-free} y \textit{off-policy}.\\
+\subsubsection{Experience Replay}
+Notemos que las secuencias consecutivas de estados y acciones están fuertemente correlacionadas, por lo que el aprendizaje podría ser ineficiente.
+Además, al aprender \textit{on-policy}, los parámetros actuales determinan los siguientes estados sobre los cuales se entrenará, por lo que pueden surgir loops que afecten la convergencia del aprendizaje.
+Para resolver estos problemas se utiliza una técnica conocida como Experience Replay (\cite{Lin1992}); esta técnica consiste en guardar las experiencias del agente para cada tiempo $t$; esto es 
 	   \[ e_t = (s_t, a_t, r_t, s_{t+1}). \]
-	   \item Cada $e_t$ se almacena en un conjunto $\mathcal{D} = e_1 , ... , e_N$.
-	   \item En un loop interno del algoritmo, se tomarán aleatoriamente muestras de experiencia las cuales serán utilizadas para la actualización de pesos. 
-	   \item Q-learning:
-	   \item Aplicaciones de DQN.
+Cada $e_t$ se almacena en un conjunto $\mathcal{D} = e_1 , ...
+, e_N$ y en un loop interno del algoritmo, se tomarán aleatoriamente muestras de experiencia las cuales serán utilizadas para la actualización de pesos.
+\subsubsection{Algoritmo completo}
+Deep Q-learning con Experience Replay:
+\subsubsection{Aplicaciones de DQN}
 		\begin{itemize}
@@
 	     \end{itemize}
-	      	 \end{itemize}  	
 \subsection{Mejoras posteriores a DQN}
-   	   \begin{itemize}
-   	   		\item Double DQN.
-   	   		\item Prioritized Replay.
-   	   		\item Duelling Networks.
-   	   \end{itemize}
+\subsubsection{Double DQN}
+\cite{hasselt2010double}, \cite{van2016deep} propuso una “Double Deep Q-Network”. Recordemos que en DeepQ-learning las actualizaciones que se hacen son:
+\[ \theta_{t+1} = \theta_t + \alpha (r_{t+1} + \gamma \max_{a} Q(s_{t+1},a;\theta_t)) \nabla_\theta Q(s_t,a_t;\theta_t). \]
+Por lo tanto, es más probable escoger acciones que tienen un alto valor, lo que resulta en valores demasiado optimistas.
+Lo que se propone en \cite{van2016deep} es evaluar la política greedy utilizando la red on-line, pero utilizar la red objetivo para estimar el valor.
+Debe hacerse el siguiente cambio en la actualización:
+\[ r_{t+1}  + \gamma  Q (s_{t+1}, \textrm_{ arg max}_a Q(s_{t+1},a_t,\theta_t); \theta^{-}\]
+donde $\theta$ es el parámetro de la red on-line y $\theta^{-}$ el parámetro de la red objetivo
+\subsubsection{Prioritized Replay}
+
+\subsubsection{Duelling Networks}
 \subsection{Método Asíncronos}  
-\subsection{Policy Based DRL: Policy Gradients}
+\subsection{Policy Based DRL}
+\subsubsection{Actor-Critic}
+\subsubsection{Policy Gradients}
 Resulta que Deep Q-Learning no es un gran algoritmo, incluso los autores del paper original (\cite{DBLP:journals/corr/MnihBMGLHSK16}) mencionan sus desventajas. 
@@
 El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
-\\
+\subsubsection{Trust Region Policy Optimization}
  \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada.
@@
 Buen performance en juegos de Atari que utilizan la pantalla como input.
-
 \section{Reinforcement Learning y Causalidad}
@@
 \subsection{Motivación}
-Tomando en cuenta el contenido de la secciones anteriores, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción
+Tomando el cuenta los trabajos mencionados en la sección anterior, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
 Este agente buscará aprender un modelo causal de su entorno al mismo tiempo que aprende una \textit{política}.
@@
 Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
+De la misma manera, \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
+Por lo tanto, es interesante preguntar si un agente que busque maximizar una recompensa pueda aprender modelos causales mediante la interacción con su entorno.
 \\
-\\
-\cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-\\
-\\
-El trabajo de \cite{garnelo2016towards} nos muestra que efectivamente pueden aprenderse aspectos estructurales sobre el entorno mientras el agente aprende, incluso logrando así mejorar su desempeño.
-Además, muestra la importancia de tener una semántica que permita hablar de la estructura del entorno.
+Un agente que interactua con su entorno con el fin de maximizar una recompensa efectivamente puede aprender aspectos estructurales sobre el entorno, como nos muestra \cite{garnelo2016towards}, quien además señala que el agente logra mejorar su desempeño.
+Además, muestran la importancia de tener una semántica que permita hablar de la estructura del entorno.
 Ellos utilizaron lógica de primer orden, pues están más interesados en hacer razonamientos \textit{de alto nivel} sobre el entorno, que involocura más que las relaciones causales, y por esto ellos se comprometen con lógica de primer orden, que al restringirnos al caso causal se queda corta pues los modelos gráficos probabilistas son más expresivos, además de modulares y montónicos.\\
 \\
-\cite{lattimoreNIPS2016} explora la pregunta de cómo escoger \textit{la mejor intervención} dado un modelo causal a través de una serie de intervenciones y observaciones secuenciales; por ejemplo, en el caso en el que un granjero quiera maximizar sus cosechas, y sabe que ese rendimiento está afectado por el uso de fertilizante, la humedad y la exposición al sol, pero sólo tiene los recursos para llevar a cabo sólo una de estas modificaciones en cada temporada.
+Mientras que \cite{garnelo2016towards} utiliza una representación lógica para escoger una acción, \cite{lattimoreNIPS2016} explora la pregunta de cómo escoger \textit{la mejor intervención} dado un modelo causal a través de una serie de intervenciones y observaciones secuenciales; por ejemplo, en el caso en el que un granjero quiera maximizar sus cosechas, y sabe que ese rendimiento está afectado por el uso de fertilizante, la humedad y la exposición al sol, pero sólo tiene los recursos para llevar a cabo sólo una de estas modificaciones en cada temporada.
 El algoritmo de Lattimore le dirá, después de $T$ observaciones, cuál es la mejor intervención a llevar a cabo con el fin de maximizar esa cantidad.
@@
 \\
-\cite{ortega2014generalized} exploran un esquema de muestreo llamado Muestreo de Thompson, en el cual un agente muestrea una de entre varias acciones de acuerdo a la probabilidad subjetiva que tenga el agente de que esta sea la mejor acción.
+Por otro lado, en el caso inverso \cite{ortega2014generalized} exploran un esquema de muestreo llamado Muestreo de Thompson, en el cual un agente muestrea una de entre varias acciones de acuerdo a la probabilidad subjetiva que tenga el agente de que esta sea la mejor acción.
 Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).

### 2018-02-08 Segundo_Semestre/Seminario/Semana3/Research_Problem.tex
Añadiendo seccion de motivacion parecida a mi documento propuesta y recortando estado del arte

--- 2018-02-08 Segundo_Semestre/Propuesta/Propuesta.tex
@@
 \newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Survey}
+\title{Research Problem}
 \author{Mauricio Gonzalez Soto}
 \begin{document}
-\nocite{*}
+%\nocite{*}
 \maketitle
-\tableofcontents
-\newpage
 \section{Introducción}
 El problema que estamos tratando de atacar consiste en descubrir relaciones causales que se encuentren presentes en diversos fenómenos.
-El enfoque que se propone es descubrir estas relaciones mediante \textit{interacción} con el entorno.
-En el contexto de inferencia causal, los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.\\
+El enfoque que se propone es descubrir estas relaciones mediante \textit{interacción} con el entorno.\\
 \\
-\cite{lake2017building} argumentan que los sistemas inteligentes actuales, a pesar de sus impresionantes logros, aun están lejos de alcanzar una verdadera inteligencia; uno de los requisitos para esto es que puedan construir modelos causales que sean interpretables, y no sólo reconocer patrones estadísticos.
+\cite{lake2017building} argumentan que los sistemas inteligentes actuales, a pesar de sus impresionantes logros, aun están lejos de alcanzar una verdadera inteligencia; uno de los requisitos para esto es que puedan construir modelos causales que sean interpretables, y no sólo reconocer patrones estadísticos.\\
 \\
-La primera parte de este resumen consiste en el estudio clásico de relaciones causales, el cual depende fuertemente de Redes Bayesianas, las cualesson objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
-Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
-\\
-Después, veremos que existe un área del aprendizaje de máquina llamado Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.\\
-\\
-Además, las intuiciones iniciales sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
+ las intuiciones iniciales sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
 Se sabe también que el conocimiento causal adquirido puede ser utilizado para mejorar las futuras decisiones tomadas.\\
@@
 Además, uno se pregunta si es posible utilizar este conocimiento causal obtenido al interactuar sobre el mundo para realizar otro tipo de inferencias causales en ciertos dominios de interés.
-\section{Causalidad}
-\subsection{Intro}
-Saber si es posible hablar de relaciones causa-efecto es un tema polémico; los métodos probabilistas tradicionales sólo codifican asociaciones estadísticas, pero la idea
-La diferencia entre modelos probabilistas y modelos causales se da cuando es de interés tomar acciones que modifiquen el entorno y no nada más observar los valores de las variables. \\
-Según \cite{pearl2014probabilistic}, las dos preguntas fundamentales en Causalidad son:
+
+\section{Círculos del State of The Art}
+\subsection{Outer Circle}
 \begin{itemize}
-\item ¿Qué evidencia empírica es necesaria para hacer inferencias válidas sobre relaciones causa-efecto?
-\item Dado que estamos dispuestos a aceptar información causal sobre un fenómeno, ¿qué inferencias podemos obtener a partir de esta información, y cómo?
+\item (\cite{suppes1970probabilistic})
+\item \cite{holland1986statistics}
+\item \cite{sutton1998reinforcement}
+\item Causal decision theory (\cite{joyce1999foundations}): Toma de decisiones bajo incertidumbre utilizando información causal; es decir; sé que $A$ \textit{causa} $B$, entonces puedo predecir qué pasará si $A$ es escogido, u observado.
+\item (\cite{spirtes2000causation})
+\item \cite{pearl2009causality},
+\item Causal graphical models ( \cite{koller2009probabilistic}): Semántica causal sobre modelos gráficos probabilistas.
+\item Libro de Teoría de la Decisión bajo incertidumbre (\cite{gilboa2009decision}).
+\item \cite{pearl1988probabilistic}, Libro clásico sobre uso de grafos para representar distribuciones.
+\item \cite{van2012reinforcement}, Estado del arte de aprendizaje por refuerzo.
+\item  \cite{danks2014unifying} Representar modelos mentales como modelos gráficos probabilistas; en particular, el problema de inducción causal. 
+\item \cite{gershman2015reinforcement}. Aprendizaje por Refuerzo en el Cerebro.
+\item \cite{lopez2015towards}. Hacia una teoría de aprendizaje para relaciones causa-efecto.
+\item \cite{lake2017building}. Condiciones que, idealmente, deberían cumplir algoritmos para que se pueda decir que piensan \textit{como} los humanos,
 \end{itemize}
-En cuanto al segundo punto, Pearl considera que hay tres tipos de preguntas a considerar
+\subsection{Middle Circle}
 \begin{itemize}
-\item Predicciones: si sabemos que una persona ingirió alimentos de dudosa procedencia, ¿se enfermará?
-\item Intervenciones: si le damos cierto tratamiento a un paciente, ¿mejorará?
-\item Contrafactuales: ¿el paciente del punto uno se habría enfermado de todos modos si no hubiera consumido esos alimentos?
+\item Decision problems and causal models: \cite{dawid2002influence} show how causal Bayesian Nets can be augmented with decision nodes as to model decision problems that use causal information.
+\item \cite{hagmayer2009decision}
+\item \cite{meder2009role}
+\item \cite{meder2010observing}
+\item \cite{kemp2010learning}
+\item On-line learning in causal models(\cite{wellen2012learning})):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop.
+
+\item \cite{lopez2015towards}
+\item \cite{krynski2007role} Human reasoning under uncertainty naturally operates over causa mental models and statistical data supports correct Bayesian inference only when they can be incorporated into a causal model.
+\item \cite{van2016deep}
+\item \cite{goudet2017learning}
+\item \cite{leibo2017multi}
+\item \cite{hernandez2017survey}
+\item \cite{li2017deep}. Overview general de Deep Reinforcement Learning
+
 \end{itemize}
-Pearl argumenta que aunque causalidad suele entenderse como afirmaciones necesarias, estas son hechas en contexto de incertidumbre, por lo que cualquier teoría causal debe estar en términos de probabilidades.
-\subsection{Un poco de historia}
-Aristóteles definió 4 \textit{causas} para una cosa, una de ellas llamada \textit{causa eficiente} que estaba definida como aquello que hace la cosa.\\
-\\
-En 1690, el filósofo inglés John Locke propuso como definición de causa “aquello que produce alguna idea simple o compleja”, y lo que es producido lo definió como efecto.\\
-\\
-\cite{holland1986statistics} resalta la importante diferencia en que Aristóteles hizo énfasis en las causas de una cosa, pero no en el efecto de estas. \\
-\\
-Fue David Hume quien se considera que hiz una importante contribución, y él definió la causalidad como una relación entre experiencias más que una entre hechos.
-Argumentó que no es verificable empíricamente que la causa produzca el efecto, sino sólo que el evento llamado \textit{causa} es invariablemente seguido por aquella experiencia llamada \textit{efecto}.\\
-\\
-Hume definió 3 criterios básicos para causalidad:
+\subsection{Inner Circle}
 \begin{itemize}
-\item Contiguidad espacio-temporal: tanto causa y efecto tienen lugar en una entidad común (espacio y tiempo).
-\item Sucesión temporal: La causa tiene que preceder temporalmente al efecto. 
-\item Conjunción constante: Causa y efecto ocurren juntos, o no ocurren juntos.
+\item \cite{eberhardt2008causal} plantea el problema de \textit{descubrimiento causal} como un juego entre un científico vs la naturaleza en el cual la naturaleza intenta mantener ocultos sus secretos y el científico intenta descubrirlos mientras minimiza un costo.
+\item \cite{eberhardt2012number}
+\item \cite{hauser2012two}
+\item Causal learning through sequential decision making (\cite{hagmayer2013repeated}).
+\item \cite{ortega2014generalized} 
+\item \cite{bramley2015staying}
+\item \cite{alon2015online} Estudian problemas de aprendizaje on-line en los cuales en cada ronda un jugador escoje una acción y recibe un \textit{loss} previamente especificado.
+Estudian las \textit{feedback graphs} que básicamente dicen si al ejecutar acción $i$ vemos el loss asociado a acción $j$ y estudian cómo la estructura de este grafo afecta el aprendizaje.
+Es principalmente un paper de estructura de grafos y el \textit{aprendizaje} se define en términos de Teoría de Juegos.
+\item \cite{lattimoreNIPS2016} Quieren descubrir la mejor intervención de modo que se minimice el \textit{regret} después de $T$ rondas.
+\item \cite{zhang2017deeper} Una vuelta a Experience Replay
+\item \cite{albrecht2017autonomous}
+\item \cite{albrecht2016exploiting}
+\item \cite{garnelo2016towards}
+\item \cite{pearl2018theoretical}
+\item \cite{innes2018reasoning}. 
 \end{itemize}
-John Stuart Mill argumentó tajantemente que observación sin experimentos no puede probar causalidad. Mill identificó cuatro métodos generales para la identificación de relaciones causales. 
-\subsection{Modelo de Rubin}
-El modelo de Rubin parte de una población de \textit{unidades} $U$.
-En general, estas unidades son los objetos de estudio sobre las cuales actuan las causas.
-Para hacer más sencilla la descripción del modelo, supondremos que sólo existen dos causas (o niveles de tratamiento) denotadas por $t$ y $c$.
-Sea $S$ una variable indicadora de la causa a la cual cada unidad $u \in U$ fue expuesta.
-Además, consideramos dos variables respuesta $Y_t$, $Y_c$.
-$Y_t(u)$ reoresenta el valor de la respuesta si la unidad $u$ es expuesta a $t$ y análogo para $c$.\\
-\\
-\begin{defi}
-Se define el efecto de la causa $t$ sobre $u$ relativo a la causa $c$ como
-\[ Y_t(u) - Y_c(u). \]
-Esta expresión se lee de la siguiente manera: $t$ causa el efecto $Y_t(u) - Y_c(u)$ en la unidad $u$ (relativo al control $c$).
-\end{defi}
-Notemos que es imposible observar simultáneamente el efecto de $t$ y de $c$ sobre una misma unidad $u$ y a esto \cite{holland1986statistics} le llama el Problema Fundamental de la Inferencia Causal.
-El problema, según explica Holland, radica en la palabra \textit{observar} pues no podemos saber simultáneamente el efecto de exponer a un enfermo en particular a un nuevo tratamiento y al mismo tiempo el efecto de no exponerlo.
-\subsection{Modelo de Suppes}
-En su libro “Probabilistic Theory of Causality”, (\cite{suppes1970probabilistic}), Patrick Suppes intenta mejorar el análisis de Hume sobre causalidad; en particular, la condición de conjunción constante.
-Para Suppes, no importa qué sean las causas y los efectos, sino que se puedan expresar como eventos que ocurren bajo ciertas condiciones.
-Suppes supone que todas las causas necesariamente preceden en el tiempo a sus efectos.
-Define una causa \textit{prima facie} de un evento como un evento que precede temporalemnte a este y que está asociado de manera positiva con él.
-Define como una causa espuria a una causa prima facie tal que es condicionalmente independiente del efecto dada otra causa la cual es anterior en el tiempo a la causa prima facie y que está asociada con el evento dada la causa.
-Una causa genuina es una causa prima facie que no es espuria.
-\begin{defi}
-Si $r < s$ representan valores de tiempo, se dice que el evento $C_r$ es una causa prima facie del evento $E_s$ si
-\[ P(E_s | C_r ) > P(E_s). \]
-\end{defi}
-\begin{defi}
-Un evento $C_r$ se dice que es una causa espuria de un eventi $E_s$ si $C_r$ es una causa prima facie de él y para alguna $q < r < s$ existe un evento $D_q$ tal que 
-\[P(E_s | C_r, D_q) = P (E_s | D_q) \]
-y además,
-\[P(E_s | C_r , D_q) \geq P(E_s | C_r). \]
-\end{defi}
-Por lo tanto,
-\begin{defi}
-Para tiempos $r < s$, un evento $C_r$ se dice que es una causa genuina de un evento $E_s$ si $C_r$ es una causa prima facie de $E_s$, pero no es una causa espuria.
-\end{defi}
-Notemos que Suppes está definiendo la causa de un efecto, más que el efecto de una causa.
-Además, no hay restricciones a la naturaleza de una causa más allá de que sea un evento que ocurre antes en el tiempo.
-El modelo de Suppes no puede expresar el efecto de una causa en particular, sólo comportamientos promedio (\cite{holland1986statistics}).\\
-\\
-En palabras de Holland, la causalidad de Suppes no es más que correlación entre causa y efecto que no puede ser eliminada al parcializar causas legítimas.
-\subsection{Modelos Causales en Medicina}
-Un caso de aplicación en medicina es determinar que ciertas bacterias causan ciertas enfermedades.
-Existe un conjunto de postulados planteados por Koch para decidir cuándo un micro-organismo está implicado en una enfermedad.
-Aunque no existe un planteamiento formal, estos pueden expresarse como (\cite{yerushalmy1959methodology}):
-\begin{itemize}
-\item El organismo debe estar presente en todos los casos de la enfermedad supuestamente provocada por él.
-\item El organismo debe aislarse de los pacientes y cultivarse en estado puro.
-\item El organismo cultivado en estado puro debe reproducir el padecimiento. 
-\end{itemize}
-Estos postulados aseguran que no haya datos que apoyen una hipótesis nula de cero efecto causal. \\
-\\
-Por otro lado, Hill, quien fue de los primeros en estudiar la asociación entre fumar y cáncer, encontró 9 factores útiles a la hora de decidir que una asociación observada es causal. 
-\subsection{Causalidad de Granger en Economía}
-La causalidad de Granger se utiliza para series de tiempo, lo cual es de alta importancia en Economía.
-\cite{holland1986statistics} opina que las ideas esenciales de Granger no están limitadas a series de tiempo y que pueden ser expresadas en términos de variables que actuan sobre unidades de una población.\\
-\\
-En la teoría de Granger, una variable causa otra variable; es decir ( \cite{holland1986statistics}), las realizaciones de una variable mejoran las predicciones hechas sobre la otra variable.
-La única parte en la que se utilizan series de tiempo en esta definición es para hablar de valores de las variables definidos hasta un cierto tiempo $t$.\\
-\\
-Si $X,Y,Z$ son variables aleatorias definidas sobre cierta población, se dice que $X$ no es una causa de Granger de $Y$ si $X$ y $Y$ son condicionalmente independientes dado $Z$.
-Entonces, $X$ es una causa de Granger de $Y$ si distintos valores de $X$ llevan a distintas distribuciones predictivas de $Y$ dados $X$ y $Z$; esto es, $X$ ayuda a predecir $Y$ aun cuando $Z$ es tomada en cuenta.
-La no-causalidad de Granger es parecida a las causas espurias de Suppes pues ambas consideran que no se pueda predecir un evento dada cierta información.\\
-\\
-En opinión de Holland, la causalidad de Granger falla de la misma manera que la de Suppes, pues a la luz de nueva información una causa Granger puede dejar de serlo.
-\subsection{Modelos Causales en Ciencias Sociales}
-\subsection{Aprendizaje de Modelos Causales a partir de datos}
-En los métodos de estado del arte de Modelos Causales Funcionales suele hacerse los siguientes supuestos (\cite{goudet2017learning}):
-\begin{itemize}
-\item Suficiencia Causal (CSA)
-\item Propiedad Markoviana (CMA)
-\item Independencias Condicionales
-\item Fiabilidad causal (CFA)
-\end{itemize}
-\subsubsection{Aprender el CPDAG}
-Existen tres familias de métodos para extraer el grafo: métodos de restricciones, métodos de score y métodos híbridos.\\
-\\
-Los métodos basados en restricciones explotan las independencias condicionaes para identificar todas las v-estructuras.
-El algoritmo más conocido es el Algoritmo PC (\cite{spirtes2000causation}).
-El algoritmo PC primero construye un esqueleto basado en independencias condicionales; luego, identifica las v-estructuras y posteriormente orienta las aristas.
-El algoritmo FCI (\cite{spirtes1999algorithm}) extiende el algoritmo PC al relajar la suficiencia causal e incorporar variables latentes.
-El algoritmo RFCI de \cite{colombo2012learning} es más rápido que FCI y maneja datos de alta dimensionalidad.
-La debilidad de estos métodos, según \cite{goudet2017learning} es que dependen demasiado en pruebas de independencia, que dependen a su vez de la cantidad disponible de datos.\\
-\\
-Los métodos basados en score exploran el espacio de CPDAG's y minimizan un score global.
-La exploración se realiza en términos de operaciones sobre grafos, por ejemplo añadir o remover aristas.
-El algoritmo Greedy Equivalent Search (GES) de \cite{chickering2002optimal} encuentra la estructura óptima en términos del BIC.\\
-\\
-Los métodos híbridos combinan las técnicas anteriores.
-Por ejemplo el Max-Min Hill Climbing Algorithm de \cite{tsamardinos2006max} primero construye un esqueleto utilizando pruebas de independencia condicional, y luego le da orientación a las aristas utilizando un método greedy sobre un score bayesiano.
-\subsubsection{Explotando la asimetría entre causa y efecto}
-Los métodos anteriores no toman en cuenta la asimetría entre causas y efectos.
-La intuición detrás (\cite{goudet2017learning}) está dada por el siguiente ejemplo: Si consideramos el modelo causal funcional $Y=X+U$, entonces los métodos anteriores basados en grafos no pueden darle orientación a la arista $X - Y$, pues tanto $X \rightarrow Y$ como $Y \rightarrow X$ son Markov equivalentes.\\
-\\
-\cite{stegle2010probabilistic} infiere la dirección causal al construir dos modelos Bayesianos generativos: uno para $X \rightarrow Y$ y otro para $Y \rightarrow X$.
-La dirección causal es determinada a partir del modelo generativo que mejor ajuste a los datos.
-Otros métodos, como el propuesto por \cite{sgouritsa2015inference} se basan en que si $X \rightarrow Y$, entonces la probabilidad marginal de la causa $P(X)$ es independiente del mecanismo causal $P(X | Y)$; por lo tanto, estimar $P(Y | X)$ a partir de $P(X)$ no debería poderse, pero estimar $P(X | Y)$ a partir de $P(Y)$ sí debería ser posible.
-\subsubsection{Como problema de Aprendizaje}
-Inspirados en los concursos de Kaggle de \cite{Guyon2013}, \cite{lopez2015towards} plantean el problema de inferencia causal como un problema de clasificación de medidas de probabilidad.
-Los datos provistos por el concurso son 16,200 parejas de variables $(X_i,Y_i)$.
-Cada par consta de una realización $(x_i,y_i)$ y una etiqueta $l_i$ que indica el valor de la verdadera relación causal; es decir, $l_i$ indica uno de los siguientes casos: $X \rightarrow Y,$ $Y \rightarrow X$, $X_i \perp Y_i$, $X \leftrightarrow Y$.\\
-\\
-Notemos que este método tiene como punto débil que depende de la representatividad de la muestra de entrenamiento y que supone la existencia de una distribución madre
-\subsubsection{Modelos Causales Funcionales Neuronales}
-\cite{goudet2017learning} proponen un esquema que consiste en tres pasos: primero, identificar el esqueleto según propiedades de independencia estándar; luego, orientar las aristas utilizando métodos bivariados estándar y finalmente hacer una optimización sobre todo el grafo examinando relaciones locales.
-Este último paso se realiza al aprender estos modelos locales como redes neuronales generativas; es decir, se modela la distribución $P(X | X_{Pa_i}, U_i)$ de un nodo condicional en sus padres (causas) con una red neuronal.\\
-\\
-Formalmente, se modela cada ecuación
-\[X_i = f_i (X_{Pa_i},U_i) \]
-como una red neuronal generativa global, en la cual cada función $f_i$ es modelada por una red neuronal generativa con una capa oculta.\\
-\\
-Encontrar un modelo causal de este tipo requiere llevar a cabo dos tareas: estimar la estructura (el grafo) de los padres de cada variable, y las redes neuronales $f$
-\subsection{Correlación y causalidad}
-Dos variables aleatorias $X$,$Y$ pueden estar correlacionadas en distintos casos; por ejemplo, si $X$ causa $Y$, o si tanto $X$ como $Y$ son causadas por otra variable desconocida $V$.
-Si e conoce la existencia de $V$ y además podemos observarla.
 
-\section{Modelos Gráficos Probabilistas}
-
-\section{Modelos gráficos causales}
-Sabemos que en una red bayesiana los arcos, aunque dirigidos, no necesariamente tienen una interpretación directa, lo único que importa es que el grafo capture adecuadamente las relaciones de dependencia entre variables aleatorias.
-Nos gustaría que fuera el caso que $X \to Y$ significara $X$ \textit{causa} Y.
-Resulta que existe una familia de modelos causales en los cuales se \textit{interviene} en el mundo.
-La idea clave es utilizar \textit{intervenciones ideales}, las cuales se denotan $do(Z=z)$ que significa forzar a la variable aleatoria $Z$ a tomar el valor $z$, denotado de manera simple como $do(z)$.
-Esta operación corresponde al caso en que un agente directamente modificó el estado del mundo y no sólo observó los valores de este de manera pasiva; por esta razón es muy importante tener en mente que es distinto $P(Y | do(z), X=x)$ que $P(Y | Z=z, X=x)$.
-Las preguntas causales de interés son planteadas de la forma $P(X | do(z))$ y son llamadas \textit{quries intervencionales}.
-Este tipo de intervenciones surgen de manera natural en muchos problemas; por ejemplo, “si el paciente recibe este tratamiento, cuál es la probabilidad de sobrevivir la enfermedad” puede ser planteada como $P(sobreivir | do(tratamiento))$.\\
-\\
-Formalmente, un modelo gráfico causal (\cite{koller2009probabilistic}, \cite{pearl2014causality}) tiene la misma forma que una red Bayesiana; es decir, consiste en una red acíclica dirigida que codifica una distribución entre un conjunto de variables aleatorias.
-La diferencia radica en la interpretación de las flechas, pues cada variable aleatoria está gobernada por un modelo causal en función de sus padres en el grafo.
-Una intervención consiste en forzar uno de los nodos a tomar un valor con probabilidad 1 y a recortar todas las flechas que entran en ese nodo.\\
-\\
-Es importante que se cumpla que si dos variables están correlacionadas, entonces una es padre causal de la otra y que no falten padres causales de ninguna variable.\\
-\\
-Para poder contestar pregunta acerca de intervenciones, se necesita introducir el concepto de \textit{grafo mutilado}
-\begin{defi}
-Sea $\mathcal{B}$ una red sobre un conjunto de variables aleatorias y sea $Z_1=z_1,...,Z_k=z_k$ una instanciación de estas variables denotada por $\mathbf{Z}=\mathbf{z}$ .
-Se define el grafo mutilado $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$ como:
-\begin{itemize}
-\item Cada nodo $Z_i \in \mathbf{Z}$ no tiene padres en $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$
-\item El modelo local de probabilidad condicional para $Z_i \in \mathbf{Z}$ es tal que asignal probabilidad uno al valor $z_i$ y cero a cualquier otro valor.
-\item No se altera ningún otro nodo $X_i \notin \mathbf{Z}$
-\end{itemize}
-\end{defi}
-Entonces, utilizando este concepto, tenemos que para un modelo causal $\mathbf{C}$ una intervención $P_{\mathcal{C}}(Y|do(z)$ es realmente una \textit{querie} probabilista estándar, pero en el grafo mutilado dado por $\mathcal{C}_{Z=z}$; es decir
-\[ P_{\mathcal{C}}(Y|do(z),x) = P_{\mathcal{C}_{Z=z}}(Y|x).  \] 
-\subsection{Identificabilidad de modelos}
-Aun a pesar de desconocer la totalidad de las variables latentes causales, existen casos en los que es posible responder a preguntas causales en modelos que involucran variables latentes utilizando sólo variables observadas; este es el caso de los modelos identificables.
-Existen formas de simplificar queries intervencionales y convertirlas en queries que sólo requieren de datos observacionales pues existen equivalencias entre unas y otras bajo ciertas condiciones.\\
-\\
-\cite{peters2012identifiability} exploran las condiciones para obtener identificabilidad en modelos causales funcionales. 
-\subsection{Aprendizaje de modelos gráficos causales}
-Según \cite{koller2009probabilistic}, el problema de aprender un modelo causal a partir de datos puede dividirse en varios ejes:
-\begin{itemize}
-\item En primer lugar, definir qué se entiende por un modelo causal, pues si estamos trabajando con modelos gráficos causales entonces el problema se convierte en aprender redes bayesianas, pero con la diferencia de la interpretación que se da a las flechas; por otro lado, en modelos causales funcionales se tienen parametrizaciones mucho más complejas para las variables de respuesta, por lo que se necesitan modelos más complejos como el modelo neuronal presentado anteriormente.
-\item Un segundo eje es determinar si ya se conoce la estructura y sólo se necesita aprender la parametrización, o si también debe ser aprendida la parametrización.
-\item En tercer lugar, es el tipo de datos: observados o intervencionales.
-Datos observados son aquellos que provienen de alguna distribución generadora y datos intervencionales es en aquellos en los que se intervino el modelo y se observó la respuesta de este.
-Los datos intervencionales son difíciles de obtener en la práctica, o incluso pudieran ser inmorales o ilegales.
-\item Finalmente, cómo tratar los factores de confusión (confounding factors). Uno puede suponer que simplemente no existen y en estos casos el problema se vuelve relativamente sencillo.
-\end{itemize}
-\subsubsection{Aprender sin factores de confusión}
-El problema de aprender modelos gráficos causales sin factores de confusión es esencialmente el mismo que aprender redes bayesianas (\cite{koller2009probabilistic}), pero hay que hacer algunas suposiciones importantes, pues existen siempre varios DAG's que representen la misma distribución, pero que tendrían distintas interpretaciones causales.
-Una de las suposiciones necesarias para poder aprender modelos causales a través de datos observacionales es la llamada Propiedad Causal de Markov, que es un análogo a la Propiedad de Markov para redes Bayesianas; y su formulación es la siguiente (\cite{koller2009probabilistic}): \textit{cada variable es condicionalmente independiente de sus no-efectos dadas sus causas directas.} El otro supuesto necesario es que las únicas independencias condicionales sean aquellas que surgen de la d-separación en el grafo causal correspondiente (\textit{faithfulness}).\\
-\\
-Aun con estas suposiciones, lo más que puede aprenderse es una clase de equivalencias para el grafo causal y para llevar esto a cabo existen dos familias de métodos: basados en restricciones y basados en score y son los mismos que se utilizan para aprender el CPDAG en modelos causales funcionales.
-\subsubsection{Aprender de datos intervencionales}
-En el caso en que se tienen datos provenientes de intervenciones, lo cual es lo más común en el contexto de experimentos científicos.
-Se utilizan métodos basados en score para aprender de este tipo de datos.
-\subsubsection{Aprender con variables latentes}
-En el caso en el que suponemos que existen variables latentes, estamos pensando que los datos fueron generados por una distribución que es \textit{parecida}, o parcial, a la verdadera.
-De nuevo, existen familias de métodos basadas en restricciones y métodos basados en scores.
-\subsection{Redes Bayesianas causales: aplicación en genética}
-\cite{su2013using} intentan aprender la estructura de las relaciones entre un conjunto de variables relacionadas a un estudio de cáncer (\cite{karagas1998design},\cite{karagas2004incidence}).
-Específicamente, los autores analizan sólo 11 variables relacionadas a un solo gen.
-Identifican relaciones interesantes, como por ejemplo la variable “fumador” como padre causal de la variable “cancer”, aunque no apareció relación entre la variable cancer y un gen que se sabe está relacionado con el cáncer, pero esto se debe a la manera en la que fueron seleccionados los pacientes.
-\section{Modelos causales funcionales y ecuaciones estructurales}
-En otras áreas, como las ingenierías y la física, las relaciones causales se expresan como relaciones funcionales determinísticas en donde cualquier componente aleatorio se introduce en términos de variables no observadas.
-Esto corresponde (\cite{pearl2009causality}) a la noción Laplaciana de lo aleatorio.
-En este contexto, un modelo causal funcional consiste en un conjunto de ecuaciones de la forma
-\[ X_i = f_i (X_{Pa_i},U_i), \textrm{  } i=1,...n\]
-donde $Pa_i$ corresponde al conjunto de variables que directamente determinan a la variable con índice $i$ y donde $U_i$ representa cualquier error de medición o de aspectos no observados.\\
-\\
-Dado un modelo causal de esta forma, podemos obtener un diagrama de la siguiente manera: para cada variable pintamos un nodo, y desde cada elemento de $Pa_i$ dibujamos una flecha hacia el nodo que corresponde a $X_i$.
-A este diagrama se le llama \textit{diagrama causal}.
-Si cumple que el diagrama es acíclico, se le conoce como semi-Markoviano.
-Si, además de ser acíclico, los términos de error son independientes entonces se dice que es Markoviano.
-El siguiente Teorema muestra la conexión entre el diagrama causal y la distribución de probabilidad asociada.
-\begin{teo}
-Todo modelo causal Markoviano induce una distribución $P(X_1,...,X_n)$ que satisface que cada variable $X_i$ es independiente de todos sus no-descendientes dado sus padres $Pa_i$.
-\end{teo}
-Sabemos que dos DAG's que tienen el mismo esqueleto y las mismas v-estructuras son equivalentes, y a la clase de equivalencia le llamaremos CPDAG (\cite{goudet2017learning})
-\subsection{Intervenciones y efectos causales en modelos funcionales}
-Los modelos funcionales causales proveen una manera de estimar cómo cambiaría la distribución en respuesta a cambios externos.
-Lo que se hace en estos modelos es representar las intervenciones como alteraciones a un conjunto de funciones.
-El efecto total de la intervención se calcularía modificando las ecuaciones correspondientes y con este nuevo modelo calcular las nuevas probabilidades.
-Denotamos una intervención como $do(X=x)$.
-Se dice que una variable $X_i$ es causa directa de $X_j$ si
-\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
-\subsection{Contrafactuales en modelos funcionales}
-
-\section{Teoría de la Decisión}
-\subsection{von Neumann-Morgenstern}
-Según narra \cite{gilboa2009decision}, fue Daniel Bernoulli quién habló por primera vez de la maximización de la utildiad esperada; esto, al estudiar la \textit{paradoja de San Petersburgo}, la cual consiste en lo siguiente: un tomador de decisiones se enfrente a una serie de lanzamientos consecutivos de una moneda justa.
-Se considera un juego basado en los resultados de estas monedas, el cual terminará la primera vez que la moneda arroje una cara, y si esto ocurriera en el $n$-ésimo lanzamiento, el concursante recibirá $2^n$ pesos.
-La \textit{paradoja} consiste en que la mayoría de las personas, según registró Bernoulli, sólo está dispuesta a pagar una cantidad finita de dinero por entrar a este juego, cuyo valor esperado es infinito.
-La paradoja fue resuelta cuando Bernoulli se dio cuenta que las personas estaban maximizando el valor esperado de una \textit{función de utilidad}, pero no directamente el valor esperado del dinero.
-Resulta que si esta función es logarítmica, entonces la utilidad esperada es finita.
-\\
-\\
-El Teorema original de von Neumann y Morgenstern está formulado en términos de árboles de decisión y loterías compuestas.
-Más tarde \cite{jensen1967introduction} lo formuló de manera más compacta al suponer que las loterías compuestas son simplificadas de acuerdo a la fórmula de Bayes.\\
-\\
-Sea $X$ un conjunto de \textit{alternativas}, los objetos a elegir son loterías de soprte finito; de manera formal, definimos
-\[ L = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
-Una mezcla de loterías se puede definir de la siguiente manera: se toma una probabilidad $\alpha$ y se define una lotería $l^\alpha = \alpha P + (1- \alpha) Q$.\\
-\\
-Consideremos una relación  $\succeq L \times L$. El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas:
-\begin{itemize}
-\item A1. Órden débil: $\succeq$ es completa y transitiva
-\item A2. Continuidad: Para toda $P,Q,R \in L$ tales que $P \succ Q \succ R$ existen $\alpha$ y $\beta$ tales que:
-\[  \alpha P + (1-\alpha) R \succ Q \succ \beta P + (1-\beta) R\]
-\item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si $\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R$.
-\end{itemize}
-\begin{teo}
-Una relación $\succeq L \times L$ satisface los axiomas A1 - A3 si y sólo sí existe una función $u: X \to \mathbb{R}$ tal que para toda $P,Q \in L$ se tiene que $P \succeq Q$ si y sólo si 
-\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
-\end{teo}
-Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
-\subsection{De Finnetti}
-¿Cuál es la definición de probabilidad subjetiva?
-\cite{gilboa2009decision} sugiere que para saber qué son debemos explorar cómo se usan estas, y una posible manera de uso es al maximizar una utilidad esperara y esto es lo que hace el Teorema de DeFinnetti.\\
-\\
-De Finetti consideró apuestas sobre \textit{estados del mundo}.
-Consideremos que existen $1,...,n$ estados en el mundo.
-Una apuesta es una función que toma valores en el espacio de estados y los mapea a valores monetarios; es decir, una apuesta $x$ es un vector de $n$ elementos en donde $x_i$ es la ganancia en el caso de que la apuesta resulte en el estado $i$ como el ganador.
-Consideramos una relación $\succeq \subseteq X \times X$.
-Y consideremos los siguientes axiomas:
-\begin{itemize}
-\item A1. Órden débil: $\succeq$ es completa y transitiva.
-\item A2. Continuidad: Para cada $x \in X$, los conjuntos $\{ y | x \succ y \}$ y $\{ y | y \succ x \}$ son abiertos.
-\item A3. Aditividad: Para toda $x,y,z \in X$, $x \succeq y$ si y sólo si $x+z \succeq y+z$.
-\item A4 Monotonicidad: Para toda $x,y \in X$, si se cumple que $x_i \geq y_i$ entonces se tiene que $x \succeq y$.
-\item A5 No-trivialidad: Existen $x,y \in X$ tales que $x \succ y$.
-\end{itemize}
-Notemos que el axioma A3 sólo aplica bajo un esquema de neutralidad al riesgo.
-\begin{teo}
-La relación $\succeq \subseteq X \times X$ satisface A1-A5 si y sólo si existe un vector de probabilidad $p$ de tamaño $n$ tal que para toda $x,y \in X$ se cumple que
-\[ x \succeq y \textrm { si y sólo si } px \geq py. \]
-\end{teo}
-Entonces, la definición de probabilidad subjetiva sería aquella probabilidad utilizada por un tomador de decisiones para escoger una acción mediante el criterio de utilidad esperado.
-\subsection{Savage}
-El Teorema de von Neumann-Morgenstern indica cómo obtener una función de utilidad si las probabilidades están bien definidas; el Teorema de De Finnetti dice cómo obtener probabilidades si conocemos las utiidades de un tomador de decisiones.
-El Teorema de Savage obtiene al mismo tiempo probabilidades y utlidades.
-Su modelo asume sólo estados del mundo y sus resultados.
-Por lo tanto, los axiomas propuestos por Savage son bastante más abstractos y menos interpretables, pero se obtiene la ventaja de que sus axiomas son equivalentes a la existencia de una función de utilidad y de una medida de probabilidad (subjetiva) tales que las decisiones se llevan a cabo al maximizar utilidad esperada respecto a esta función de utilidad y esta medida de probabilidad.\\
-\subsubsection{Estados, resultados y acciones}
-\section{Aprendizaje por Refuerzo}
-El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno.
-Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
-El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente.
-A su vez, el ambiente devuelve información al agente en términos de una recompensa o ganancia según el estado en el que se encuentre.
-El objetivo del agente es encontrar una serie de acciones que lo lleven a lograr su objetivo final.
-La función de recompensa codifica qué es lo que se desea que el agente aprenda, pues esta función devuelve un valor para cada estado y con esto el agente asignará un valor a cada estado según la ganancia a largo plazo que un este representa.\\
-\\
-La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
-El objetivo del agente es encontrar la política que maximice la recompensa.\\
-\\
-La regla general que se sigue para la definición del ambiente (\cite{sutton1998reinforcement}) es considerar que todo aquello que no pueda ser explícitamente alterado por el agente pertenece al ambiente.
-La frontera agente-ambiente tiene que ver con los límites de lo que el agente puede controlar, no sobre lo que puede \textit{conocer}.
-Veremos ahora con detalle las partes que conforman esta dinámica. 
-\subsubsection{Objetivos y Recompensas}
-Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada.
-Por ejemplo, un agente que queremos que aprenda a jugar ajedrez debe ser premiado por ganar los juegos, no por lograr sub-objetivos como comer más piezas del oponente o tomar control del centro del tablero (\cite{sutton1998reinforcement}), pues podría suceder que el agente encuentre cómo lograr esto sin ganar el juego necesariamente.
-Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
-\subsubsection{Definición de Cadenas de Markov}
-Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
-\[P(X_{t+1} \in A | X_0=x_0, X_1=x_1 , ... , X_t=x_t) = P(X_{t+1} \in A | X_t = x_t).\]
-Estos procesos quedan caracterizados por sus probabilidades de transición $p^t_{ij}=P(X_{t+1}=j | X_{t}=i)$.
-Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
-\subsubsection{Procesos de Decisión Markovianos con recompensa}
-Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$.
-En respuesta a la acción tomada, se recibe una recompensa $r_t = f(a_t)$ para alguna función $f$ desconocida, la cual puede ser aleatoria o determinista.
-Los procesos de decision Markovianos con recompensa (\cite{Puterman:1994:MDP:528623}) están definidos por un espacio de estados $\mathcal{S}$, un espacio de acciones posibles $\mathcal{A}$, una función de transición $f$ y la función recompensa $r$.
-La función transición puede ser estocástica o determinística.
-
-\subsubsection{Función valor}
-Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado.
-Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política.
-Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$.
-Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$.
-La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante.
-Formalmente, para un proceso de decisión Markoviano, definimos la \textit{función valor} (value function en la literatura) para un estado $s$, y relativa a una política $\pi$, como el retorno esperado si se comienza en $s$ y a partir de ahí se sigue la política $\pi$
-\[ V^{\pi}(s) = \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]. \]
-De manera similar, definimos la \textit{función acción-valor} $Q^{\pi}(s,a)$ es una valoración de tomar la acción $a$ estandoen el estado $s$, y después seguir la \textit{política} $\pi$
-\[ Q^\pi(s,a) = \mathbb{E}[R_t | s_t =s, a_t = a]. \]
-Por la ley de los grandes números, podemos estimar ambas funciones $V^{\pi}$ y $Q^\pi$ si se promedian los retornos recibidos al llegar a cada estado con el número de veces que se ha llegado al estado.
-Esta es la idea detrás de los llamados Métodos Monte-Carlo que serán expuestos en una sección posterior.\\
-Notemos una propiedad interesante de la función $V^{\pi}$:
-\begin{eqnarray*}
-V^{\pi}(s) &=& \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ \sum_{k=0}^\infty \gamma^k r_{t+k+1} | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ r_{t+1} + \gamma \sum_{k=0}^\infty \gamma^k r_{t+k+2} | s_t = s \right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} P(s_{t+1}=s' | s_t = s, a_t = a) \left[ \mathbb{E}[r_{t+1}|s_t=s,a_t=a,s_{t+1}=s'] + \gamma \mathbb{E} \left[ \sum \gamma^k r_{t+k+2} | s_{t+1}=s' \right]\right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} \mathcal{P}_{ss'}^a \left[ \mathcal{R}_{ss'}^a + \gamma V^{\pi}(s') \right]
-\end{eqnarray*}
-A esta última ecuación se le conoce como la Ecuación de Bellman para $V^\pi$.\\
-\\
-También para la función $Q^{\pi}(s,a)$ se tiene una ecuación de Bellman:
-\[ Q^{\pi}(s,a)=\mathbb{E}_{s',a'}[r+ \gamma Q^{\pi}(s',a') | s',a']. \]
-Además, la función $V^\pi$ puede ser obtenida a partir de la función $Q^{\pi}(s,a)$ de la siguiente manera:
-\[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
-\subsubsection{Soluciones y Optimalidad}
-Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa.
-Resulta que es posible definir un orden parcial entre políticas en términos de la función valor, pues $\pi \geq \pi'$ si $V^{\pi}(s) \geq V^{\pi'}(s)$ para todo estado $s$.
-Para procesos de decisión Markovianos finitos existe siempre una política mejor que las demas, a la cual llamaremos \textit{política óptima}.
-De hecho, pueden existir varias, pero todas ellas comparten la misma función valor, denotada como $V^\ast$.
-Si para una política $\pi^\ast$ se tiene que es mejor que todas las demás, entonces podemos describir su función valor como
-\[  V^{\pi^\ast} = \max_{\pi} V^{\pi}(s).\]
-Por supuesto, $V^\ast = V^{\pi^\ast}$
-\\
-Análogamente, las políticas óptimas tienen la misma función $Q$ óptima, por lo que definimos \textit{función accion-valor óptima} como el retorno esperado de tomar la acción $a$ en el estado $s$ y a partir de ahí seguir una política óptima:
-\[ Q^\ast (s,a) = \max_\pi Q^{\pi}(s,a) = Q^{\pi^\ast}(s,a). \]
-Los métodos modernos de aprendizaje por refuerzo intentan aprender directamente la función valor o la función $Q$, y una vez Q y una vez que tenemos $Q^{\pi^\ast}$ podemos definir una política óptima de la siguiente manera:
-\[ \pi^\ast = \text{argmax}_a Q^\ast (s,a). \]
-Además, utilizando la ecuación de Bellman vimos que se podía escribir $Q$ en términos de $V$, y por lo tanto podemos escribir $Q^\ast$ en términos de $V^\ast$ como
-\[ Q^\ast (s,a) = \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La ecuación de Bellman puede escribirse sin hacer referencia a ninguna política de la siguiente manera:
-\[ V^\ast (s) = \max_a \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La función óptima $ Q^\ast$ lo que hace es maximizar sobre todas las decisiones:
-\begin{eqnarray*}
-Q^\ast &=& r_{t+1} + \gamma \max_{a_{t+1}} r_{t+2} + \gamma^2 \max_{a_{t+2}} r_{t+3}...\\
-&=& r_{t+1} + \gamma \max_{a_{t+1}} Q^{\ast} (s_{t+1},a_{t+1}).
-\end{eqnarray*}
-Formalmente:
-\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
-\subsection{Métodos básicos de solución: Model-based}
-\subsubsection{Programación Dinámica}
-Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
-La idea fundamental en programación dinámica es poder asignar a cada estado del sistema un valor numérico que represente el retorno esperado, a largo plazo, de ese estado; es decir, estamos valorando cada estado según lo que esperemos obtener de él a muy largo plazo.
-\\
-\\
-Recordemos que una política $\pi$ especifica una manera de actuar en cada estado; es decir, a cada estado del sistema le asocia una acción.
-Si ya conocemos una política cualquiera, la pregunta natural es ¿cómo obtener la función $V^\pi$?\\
-\\
-Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
-\[ V^\pi (s) = \mathbb{E}[r + \gamma V^\pi (s') | s_t = s]. \]
-Pero esto nos deja con el mismo problema, pues tenemos de ambos lados una función desconocida.
-Para problemas de decisión finitos, este es un sistema de ecuaciones con tantas ecuaciones como estados, lo cual puede ser costoso de resolver.
-Consideramos entonces un método iterativo:
- \[ V_{k+1} (s) = \pi(s) \sum_{s'} \mathcal{P}_{s,s'}^a [ \mathcal{R}_{s,s'}^a + \gamma V_k(s')], \]
-donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$.
-Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.\\
-\\
-Sabemos que además de la función $V$ tenemos a la función $Q$, la cual valora parejas estado-acción.
-Utilizando esta función, podemos preguntarnos qué pasaría si en vez de escoger la acción $a=\pi(s)$ escogieramos cualquier otra acción $a$.
-Claramente, onsecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
-Entonces, si esto resulta ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
-Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.\\
-\\
-Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
-\[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
-Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.\\
-\\
-Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
-\[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
-pues cada política está garantizado que es mejor que la anterior.
-El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite.
-Una opción es truncar la etapa de policy evaluación y hacer
-\[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
-Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
-Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
-\subsection{Métodos básicos de solución: Model-free}
-Métodos que no necesitan un modelo completo del sistema, sólo necesitan experiencia.
-Encontramos los métodos Monte Carlo que simulan las transiciones y los estados, y los métodos de diferencias temporales, que van actualizando iterativamente el valor de los estados observados, el más famoso de estos es el TD($\lambda$).
-\subsubsection{Métodos Monte-Carlo} 
-Los métodos Monte-Carlo no requieren un modelo del ambiente y son conceptualmente sencillos, pero no son adecuados para hacer actualizaciones incrementales.
-Estos métodos requieren \textit{experiencia} en vez de un modelo completo; es decir, requieren secuencias de estados, acciones y recompensas, ya sean estos obtenidos de simulaciones o de interacción \textit{on-line} con el ambiente.
-Los métodos Monte-Carlo suponen que la experiencia proviene de eventos episódicos, es decir que terminan siempre de una u otra forma y es al final de estos episodios que se alteran las estimaciones de valor así como las políticas.
-En este sentido, los métodos Monte-Carlo son incrementales por episodio, pero no paso a paso dentro de cada episodio.
-A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general.
-La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
-\begin{itemize}
-\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
-Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$.
-El algoritmo \textit{every-visit} estima $V^\pi (s)$ como las recompensas promedio de todas las visitas al estado $s$ en todos los episodios; por otro lado, el algoritmo \textit{first-visit} promedia sólo lo obtenido al visitar por primera vez un estado en todos los episodios.
-Ambos métodos convergen a $V^\pi$.
-\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
-Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política.
-Entonces, es importante tener un método para estimar $Q^\ast$.
-\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
-El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.
-\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
-A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
-\[ \pi_0 \xrightarrow{E} Q^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} Q^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} Q^\ast.  \]
-Bajo el supuesto de episodios infinitos y visita a todas las parejas estado-acción se tiene que los métodos Monte-Carlo encuentran $Q^{\pi_k}$ de manera exacta para $\pi_k$.\\
-Para hacer la mejora de una política, se escoge de manera determinística la acción $a$ en la cual se alcanza el máximo de la función $Q$; es decir,
-\[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
-\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
-\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
-		\begin{itemize}
-				\item On-Policy: intentan evaluar mejorar la política que es actualmente utilizada para la toma de decisiones.
-				\item Off-Policy: Evaluan o mejoran una política distinta
-		\end{itemize}
-\end{itemize}
-\subsubsection{Diferencias Temporales}
-Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos.
-Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
-\begin{itemize}
-\item Evaluación de política en Diferencias Temporales.
-\item TD(0).
-\item SARSA: Es un método \textit{on-policy}.
-\item Q-Learning: Es un método \textit{off-Policy}.
-\item Métodos Actor-Critic.
-\end{itemize}
-\subsection{Value based Reinforcement Learning: Deep Q-Networks}
-Los sorprendentes avances de Deep Learning en visión computacional llevaron de manera natural a plantear la pregunta de cómo utilizar estas nuevas técnicas para el aprendizaje por refuerzo. 
-Las aplicaciones estándar de DL requerían grandes volúmenes de datos etiquetados, además de tener muestras independientes, pero en Reinforcement learning esto es un problema pues las secuencias de estados suelen estar altamente correlacionados.\\
-\cite{mnih2013playing} introducen “Deep Q-Network”, que consiste en una implementación de Q-learning con aproximación de funciones (\cite{sutton1998reinforcement}) en donde el aproximador es una red neuronal convolucional.
-Utilizando esta técnica fue como \cite{mnih2013playing} lograron que un agente aprendiera siete juegos de Atari 2600 utilizando como input únicamente la pantalla del juego.
-\\
-\\
-Estimar la función $Q$ de manera iterativa es muy costoso, pues además de requerir un modelo completo del sistema cada secuencia de estados se utiliza por separado para las estimaciones.
-Por esto, es mejor utilizar alguna función aproximadora para poder generalizar. En este caso, se pretende aproximar la función valor mediante una red neuronal con pesos $\theta$:
-	   \[Q(s,a,\theta) \approx Q^\ast(s,a). \]
-Esta red se entrena al minimizar una sucesión de funciones $L_i$ que está dada por:
-	   \[ L_i (\theta_i) = \mathbb{E}[(y_i - Q(s,a;\theta)^2], \textrm{ donde  }y_i = \mathbb{E}[r + \gamma \max_{a'} Q(s',a';\theta) | s,a]. \]
-Notemos que los objetivos $y_i$ dependen de los pesos de la red neuronal. El gradiente de $L_i (\theta_i)$ está dado por:
-	   \[ \nabla_{\theta_i} L_i (\theta_i) = \mathbb{E}[(r + \gamma \max_{a'} Q(s',a';\theta_{i-1}) - Q(s,a;\theta_i)) \nabla_{\theta_i} Q(s,a;\theta_i).] \]
-Para optimizar la función $L$ se utiliza gradiente estocástico; este es un método \textit{model-free} y \textit{off-policy}.\\
-\subsubsection{Experience Replay}
-Notemos que las secuencias consecutivas de estados y acciones están fuertemente correlacionadas, por lo que el aprendizaje podría ser ineficiente.
-Además, al aprender \textit{on-policy}, los parámetros actuales determinan los siguientes estados sobre los cuales se entrenará, por lo que pueden surgir loops que afecten la convergencia del aprendizaje.
-Para resolver estos problemas se utiliza una técnica conocida como Experience Replay (\cite{Lin1992}); esta técnica consiste en guardar las experiencias del agente para cada tiempo $t$; esto es 
-	   \[ e_t = (s_t, a_t, r_t, s_{t+1}). \]
-Cada $e_t$ se almacena en un conjunto $\mathcal{D} = e_1 , ...
-, e_N$ y en un loop interno del algoritmo, se tomarán aleatoriamente muestras de experiencia las cuales serán utilizadas para la actualización de pesos.
-\subsubsection{Algoritmo completo}
-Deep Q-learning con Experience Replay:
-\subsubsection{Aplicaciones de DQN}
-		\begin{itemize}
-			\item \cite{mnih2013playing}.
-Paper en el que muestran cómo una computadora aprende a jugar Atari 2600 a través de sólo observar los pixeles del juego y obteniendo una 																		recompensa conforme el score aumentaba.
-Utilizando la misma arquitectura, se pudieron aprender 7 juegos
-			\item \cite{mnih2015human}.
-En este paper, aumentaron a 49 juegos aprendidos.
-Utilizan una Deep Q-Network, que consiste en aproximar la función Q mediante una red 																			neuronal que toma como input los estados (las pantallas del juego) y devuelve como output el valor Q para cada posible acción.
-En este caso, 																			utilizaron una red convolucional con 3 capas de convolución y 2 capas totalmente conectadas.
-	     \end{itemize}
-\subsection{Mejoras posteriores a DQN}
-\subsubsection{Double DQN}
-\cite{hasselt2010double}, \cite{van2016deep} propuso una “Double Deep Q-Network”. Recordemos que en DeepQ-learning las actualizaciones que se hacen son:
-\[ \theta_{t+1} = \theta_t + \alpha (r_{t+1} + \gamma \max_{a} Q(s_{t+1},a;\theta_t)) \nabla_\theta Q(s_t,a_t;\theta_t). \]
-Por lo tanto, es más probable escoger acciones que tienen un alto valor, lo que resulta en valores demasiado optimistas.
-Lo que se propone en \cite{van2016deep} es evaluar la política greedy utilizando la red on-line, pero utilizar la red objetivo para estimar el valor.
-Debe hacerse el siguiente cambio en la actualización:
-\[ r_{t+1}  + \gamma  Q (s_{t+1}, \textrm_{ arg max}_a Q(s_{t+1},a_t,\theta_t); \theta^{-}\]
-donde $\theta$ es el parámetro de la red on-line y $\theta^{-}$ el parámetro de la red objetivo
-\subsubsection{Prioritized Replay}
-
-\subsubsection{Duelling Networks}
-\subsection{Método Asíncronos}  
-\subsection{Policy Based DRL}
-\subsubsection{Actor-Critic}
-\subsubsection{Policy Gradients}
-Resulta que Deep Q-Learning no es un gran algoritmo, incluso los autores del paper original (\cite{DBLP:journals/corr/MnihBMGLHSK16}) mencionan sus desventajas. 
-Otra idea es atacar directamente la política.
-La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
-\[ a = \pi (a|s,u) \textrm{ o } a = \pi (s,u). \]
-En este contexto, definimos una función objetivo $L(\textbf{u})$ como:
-		\[ L(\textbf{u}) = \mathbb{E}[r_1 + \gamma r_2 + \gamma^2 r_3 + ... |  \pi(\cdot,\textbf{u})]. \]
-Optimizamos con Stochastic Gradient Descent. El gradiente de una policy estocástica está dado $\pi (a|s,\textbf{u})$ por:
-		\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[  \frac{\partial \log \pi(a|s,\textbf{u})}{\partial \textbf{u}}Q^\pi(s,a)\right].\]
-Por otro lado, el gradiente de una policy determinística $a=\pi(s)$ está dado por:
-\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[ \frac{\partial Q^\pi (s,a)}{\partial a} \frac{\partial a}{\partial \textbf{u}} \right]. \]
- Los métodos \textit{actor-only}, donde \textit{actor} se utiliza como sinónimo de \textit{policy} utilizan familias parametrizadas de policies.
-Las ventajas de tener una policy parametrizada es que podemos generar un espectro continuo de accciones, pero los métodos de optimización (policy gradients) sufren de alta varianza en sus estimaciones del gradiente, lo cual genera un lento aprendizaje (\cite{grondman2012survey},\cite{konda2003onactor},\cite{richter2007natural},\cite{boyan2002technical},\cite{baxter2001infinite},\cite{berenji2003convergent}).\\
- \\
-Los métodos \textit{critic-only}, donde \textit{critic} se utiliza como sinónimo de \textit{value function}, que utilizan diferencias temporales tienen una menor varianza en las estimaciones del retorno esperado (\cite{boyan2002technical},\cite{berenji2003convergent},\cite{sutton1988learning}).
-Una manera de derivar una policy en estos escenarios es seleccionando greedy actions; es decir, acciones praa las cuales la value funtion indique que el retorno esperado es máximo, lo cual puede requerir un alto costo computacional.
-Lo usual es discretizar el espacio continuo de acciones.
-Los métodos \textit{actor-critic}, combinan las vetajas de ambos mundos.
-Mientras que las familias parametrizadas tienen la ventaja de utilizar acciones continuas, los métodos critic-only proveen de conocimiento de baja varianza sobre el desempeño; es decir, la estimación del retorno esperado del crític-only permite al actor-only actualizar los gradientes con información de menor varianza, lo cual acelera todo el proceso.
-La menor varianza es intercambiada por un mayor sesgo al inicio del aprendizaje (\cite{berenji2003convergent}).
-\\
-\\
-Los métodos de Policy Gradients optimizan directamente la recompensa acumulada y pueden utilizarse junto con aproximadores no-lineales como las redes neuronales.
-Este método presenta dos grandes retos: el número de muestras de entrenamiento requeridos y la dificultad de obtener mejoras estables debido a la no-estacionariedad de los datos.
-\cite{DBLP:journals/corr/SchulmanMLJA15} atacan el primer problema se puede atacar utilizando value functions para reducir la varianza de los estimados.
-El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
-\subsubsection{Trust Region Policy Optimization}
- \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada.
-Es un algoritmo similar a Policy Gradients y resulta útil para optimizar políticas con redes neuronales.
-Buen performance en juegos de Atari que utilizan la pantalla como input.
-\section{Reinforcement Learning y Causalidad}
-Aprendemos causalidad mediante intervenciones, interacciones, etc. Reinforcement Learning es aprender a realizar tareas mediante interacción. ¿Habrá algo en medio?
-\subsection{Conocimiento causal y observabilidad parcial}
-Los algoritmos actuales de reinforcement Learning suponen que el problema puede modelarse dentro de un contexto Markoviano de decisión con recompensas.
-Si el supuesto markoviano se elimina, los algoritmos podrían dejar no funcionar, como muestran en \cite{singh1994learning}.
-Además, todo el desarrollo teórico utiliza la propiedad Markoviana.\\
-\\
-Una manera en la que la propiedad Markoviana puede perderse es si el agente no cuenta con información completa sobre el estado; es decir, al igual que en los HMM, suponemos que el ambiente es Markoviano, pero el agente no tiene acceso a él de modo que los datos no le \textit{parecen} Markovianos.
-A diferencia de los HMM, no existe un procedimiento computacionalmente tratable para los POMDP.
-El problema es que una vez que se obtienen las estimaciones del valor para cada estado, se debe hacer programación dinámica en espacios continuos y esto es infactible salvo para casos pequeños\\
-\\
-La manera de generalizar el caso markoviano al caso no observado es considerar un MDP subyacente con estados $\mathcal{S}=\{ s_1,...,s_N\}$ y probabilidades de transición $\mathcal{P}^a_{s,s'}$.
-Luego, en cada tiempo $t$, el agente observa un mensaje $m_t$ el cual proviene de una distribución $P(m | s_t)$ desconocida.\\
-\\
-En \cite{jaakkola1995reinforcement} se describe un procedimiento que evita estimar el valor de los estados y trabaja directamente en el espacio de políticas estocásticas.
-El procedimiento es \textit{directo}, a diferencia de los métodos \textit{indirectos} en los cuales primero se aprenden los parámetros del MDP y luego se utiliza programación dinámica para obtener la política.
-El algoritmo propuesto consiste en dos partes: primero, se utiliza un método Monte-Carlo para calcular un análogo de los Q-values y luego se utiliza un paso de policy-improvement.\\
-\\
-Desde un punto de vista de psicología, una manear de entender la inferencia de estados ocultos es en términos de causas latentes (\cite{gershman2015reinforcement}).
-Siguiendo esta idea, es evidente que un agente de RL que opera en el mundo real no sólo debe hacer inferencia de estados ocultos, sino que debe poder descubrir los estados ocultos que están detrás del mundo que éste observa.
-Esto es una forma de aprendizaje estructural latente (\cite{gershman2010learning}).\\
-\\
-Un primer acercamiento a este problema puede hacerse mediante técnicas de inferencia Bayesiana no paramétrica.
-Por ejemplo, tanto \cite{gershman2010learning} como \cite{gershman2012exploring} utilizan priors no paramétricas para hacer inferencia sobre causas latentes.
-Un modelo causal latente es una buena representación de la estructura causal verdadera bajo condicionamiento Pavloviano, pues es el experimentador quien es una causa latente.\\
-\\
-En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
-\subsection{Towards Deep Symbolic Reinforcement Learning}
-Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
-\\
-\cite{garnelo2016towards} proponen una arquitectura híbrida que, combina redes neuronales con inteligencia artificial simbólica.
-El modelo propuesto consiste en combinar una red neuronal que como entrada directamente lo que se recibe del ambiente y transforma estos datos crudos en una representación simbólica basada en lógica de primer orden; esta representación simbólica es posteriormente utilizada para escoger la acción a tomar.\\
-\\
-Se utiliza un sencillo juego para probar la arquitectura y en este contexto la red neuronal debe generar representaciones simbólicas de secuencias de estados en el juego en las cuales el flujo de pixeles queda codificado en términos de objetos, tipos y sus interaccioes.
-Luego, con esto se aprenden las políticas.\\
-\\
-El primer paso que hacen es utilizar un autoencoder convolucional para generar, de manera no supervisada, un conjunto de símbolos que representarán los objetos en cada escena.
-Luego, para cada pixel, obtienen la feature con mayor activación y se toman los más altos.
-De esta manera se obtiene un pixel representativo para cada objeto a los cuales posteriormente se le asigna un tipo simbólico según el autoencoder.\\
-\\
-Luego, una vez que se tienen estos símbolos de bajo nivel para cada cuadro del juego, se necesita rastrearlos a lo largo del tiempo.
-Utilizan una verosimilitud que es función de proximidad espacial y el tipo de transiciones.
-Para considerar la información de la dinámica de los cuadros utilizan la diferencia entre cuadros y las posiciones relativas entre objetos.\\
-\\
-Para la parte de aprendizaje, entrenan una función $Q$ por separado para cada interacción entre objetos con actualización
-\[Q^{ij}(s_t^{ij},a_t) \leftarrow  Q^{ij}(s_t^{ij},a_t) + \alpha \left( r_{t+1} + \gamma (\max_{a} Q^{ij} (s_{t+1}^{ij},a) - Q^{ij}(s_t^{ij},a_t)) \right). \]
-\\
-Se prueba el framework sobre varias variantes de un sencillo videojuego. \\
-\\
-Llama la atención como estos autores utilizan maquinaria de lo más reciente (deep learning, autoencoders) para luego obtener una representación en términos de lógica de primer orden.
-Este approach ha quedado claro que ya quedó atrás por lo que debería buscarse algo que sustituya este enfoque lógico.
-En particular, pudiera ser esto un modelo gráfico causal el cual vaya aprendiéndose conforme el agente de RL aprende sobre su entorno.
-
-\subsection{Conocimiento causal y toma de decisiones secuenciales: acciones como intervenciones}
-En cuanto al aprendizaje \textit{on-line} de modelos causales, encontramos en los trabajos de \cite{wellen2012learning} y de \cite{kummerfeld2013tracking} acercamientos tangenciales, o un punto de inicio, al aprendizaje de modelos causales al vuelo.\\
-\\
-Según \cite{wellen2012learning} es posible dividir las teorías de cómo las personas usan información estadística para aprender modelos causales en enfoques racionales y heurísticos.
-El enfoque primero modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana; estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
-\\
-Por otro lado, el enfoque heurístico propone que las personas reaccionan a distintos estímulos que los hacen moficiar sus hipótesis causales de una manera no necesariamente racional.
-El modelo de efecto simple de \cite{waldmann2008causal} supone que las personas se enfocan en evaluar una sola relación causal a la vez.
-Es un modelo que aprende la magnitud de la relación causal.
-Si una variable supera cierto umbral, entonces la persona acepta esa relación causal y la incorpora a su bagaje de relaciones.\\
-\\
-El modelo propuesto por \cite{wellen2012learning} sigue esta línea del efecto simple, con la diferencia de que este modelo sí intenta descubrir el proceso cognitivo subyacente, pues el aprender magnitudes causales es un proceso puramente computacional.
-El modelo de efecto simple puede verse como una implementación del modelo de \cite{danks2003dynamical}.
-Además, según narran \cite{wellen2012learning}, el modelo de efecto simple no explica cómo la persona utiliza las magnitudes causales para determinar si existe o no una relación causal.
-En el modelo propuesto, \cite{wellen2012learning}, incorporan un procedimiento de decisión para determinar si existen o no relaciones causales.\\
-\\
-El modelo LPL (\cite{wellen2012learning}) comienza con una estructura causal inicial, la cual consiste en un grafo acíclico dirigido el cual representa las creencias iniciales de un individuo.
-Posteriormente, el modelo altera esta estructura al añadir o remover una sola arista, de manera que el problema se reduce a evaluar relaciones causales entre dos variables.
-Este modelo genera estimaciones de la fuerza de la relación causal para cada posible relación causa-efecto que no sea eliminada a priori.\\
-\\
-Supongamos que un individuo cree a priori que $B$ causa $E$, y está tratando de averiguar si $C$ causa $E$.
-A diferencia del modelo de efecto simple, las estimaciones de la fuerza causal están generadas por un mecanismo similar a los filtros de partículas Bayesianos
-En particular, a partir del conocimiento causal inicial del individuo, el modelo LPL comienza con $n$ partículas.
-Estas partículas $\{V_C^1,...,V_C^n \}$ representan las creencias actuales del individuo sobre si $C$ causa $E$.
-Paralelamente, existen $\{ V_B^1,...V_B^n \}$ para $B$.
-Se define la \textit{capa} $i$ de partículas para el efecto $E$ como la $i$-ésima partícula de cada causa conocida y posible para $E$.
-En el ejemplo, tendríamos que esta capa es $\{ V_B^i, V_C^i \}$.
-Luego, el modelo LPL actualiza cada capa de manera independiente de la siguiente manera
-\[ V_C^{i+1} = V_C^{i,t} + \alpha ( \textrm{observado} - \textrm{esperado}), \]
-Donde “observado” tiene el valor de 1 si el efecto sucede y 0 en caso contrario.
-Por otro lado, “esperado” es el valor esperado del efecto, calculado según la forma funcional de la relación causa-efecto, el cual es calculado por separado para cada causa potencial en una capa.
-El valor esperado de $E$ (para la capa $i$ y la causa potencial $C$) es
-\[ \textrm{esperado} = \prod_{k} (1-V_k^{t-1})(1 - \prod_j (1- V_j^{t-1}).\]
-Conforme el individuo observa más datos, las partículas convergen a los valores verdaderos de los parámetros y las estimaciones para cada capa convergen.
-Este procedimiento no toma en cuenta aspectos estructurales, los cuales se modifican según un criterio de decisión.
-Lo que hace el modelo es una prueba $t$ en la cual la hipótesis nula es que las partículas provienen de una distribución con media cero.
-El criterio de decisión es el siguiente: si actualmente no existe una arista, y la prueba $t$ rechaza la hipótesis nula, entonces se añade la arista.
-Si existe una arista, y la prueba no es significativa, entonces se elimina la arista existente.
-Alterar una arista afectará las futuras estimaciones del valor “esperado” para otras causas.\\
-\\
-\cite{hagmayer2013hierarchical} llaman la atención sobre que los modelos actuales de Decisiones (maximum expected utility) no asumen que los tomadores de decisiones aprenden sobre la estructura causal del problema.
-Por otro lado, las teorías de aprendizaje causal proponen que un tomador de decisiones guían sus acciones a través de conocimiento causal previamente obtenido.
-El modelo de \cite{sloman2006causal} extiende las teorías de aprendizaje causal al contexto de toma de decisiones, la idea básica consiste en que los tomadores de decisiones utilizan la información disponible de su entorno para inducir un modelo causal del problema de decisión al que se enfrentan.
-Se sabe que en los seres humanos el conocimiento causal guía la interpretación de información probabilista y que los humanos somos capaces de aprender la estructura causal del entorno a través de una serie de (traducir cues) (\cite{lagnado2007beyond}).
-La pregunta fundamental que plantean \cite{hagmayer2008causal} es entender qué es lo que aprenden las personas al tomar decisiones de manera secuencial de tal forma que se busque maximizar cierta recompensa a largo plazo y donde las acciones alteran el entorno.
-Al exponerse una persona a estas situaciones de decisión secuencial se obtiene una retroalimentación de cómo las acciones se relacionaban con la obtención de recompensas.
-Las distintas teorías sobre toma de decisiones bajo incertidumbre suponen distintas representaciones para esta retroalimentación.
-Hagmayer propone agrupar estas distintas representaciones en tres niveles:
-\begin{itemize}
-\item Nivel 1: Opciones, resultados y recompensas.
-Sólo los valores esperados de las opciones posibles afectan la toma de decisiones.
-Se conoce como \textit{habit learning} en la literatura de aprendizaje en animales.
-\item Nivel 2: Opciones, relaciones causales de estas opciones a los resultados, relaciones causales entre variables de resultado y relaciones de estas con las recompensas.
-Aquí entra lo que se conoce como \textit{goal-directed learning}.
-La diferencia con el nivel 1 es la capacidad de distinguir entre acciones y los valores asociados a estas.
-\item Nivel 3: Opciones, relaciones causales entre variables de resultado, relaciones causales entre estas variables y relaciones de estas con las recompensas.
-Estos son los modelos causales.
-Las distintas teorías sobre estos modelos suponen que los humanos (o el agente que aprende) adquieren estos modelos mediante procesos de aprendizaje.
-Algunas teorías suponen que la estructura causal del entorno se obtiene a través de \textit{claves}, como la propuesta por \cite{lagnado2007beyond}.
-Por otro lado, otros autores opinan que las personas utilizan propiedades estadísticas de los datos observados, este sería el caso de las teorías de \cite{gopnik2004theory} y de \cite{griffiths2005structure}.
-\end{itemize}
-Aunque las teorías que entran en el nivel 2 son sensibles a las correlaciones entre las respuestas obtenidas, sólo en el nivel 3 se hace una distinción entre causalidad y correlación.\\
-\\
-Los autores hacen énfasis en que aunque las personas sí utilicen modelos causales al tomar decisiones simples, estas decisiones suelen ser sobre escenarios hipotéticos, pero no se han hecho muchos estudios sobre adquisición de conocimiento causal al enfrentarse a problemas de decisión secuencial con el objetivo de maximizar alguna ganancia a largo plazo.
-Además, en los estudios realizados, se ha visto que las personas tienen la capacidad de inducir modelos causales a partir de claves en el entorno; se sabe además que a partir de intervenciones activas se toman mejores decisiones que considerando sólo datos observados de manera pasiva y que las personas tienen la capacidad de integrar relaciones causales obtenidas de manera separada.
-Pero en estos estudios las personas tenían como objetivo descubrir de manera directa estructuras causales.
-No se sabe bien hasta qué punto las personas adquieren conocimiento causal al perseguir objetivos que no sean descubrir directamente estructuras causales a partir del ambiente.\\
-\\
-Los autores realizaron tres experimentos para explorar las representaciones que las personas hacen de las relaciones causales que adquieren al tomar decisiones para maximizar algún objetivo.
-En el primer experimento, el objetivo era estudiar cómo las creencias causales de las personas alteran la toma de decisiones y cómo los participantes utilizaban los resultados de sus decisiones para revisar el modelo causal.
-El experimento consistía en dos etapas de toma de decisiones, la segunda siendo una etapa de prueba.
-En ambas etapas los participantes debían escoger entre una serie de opciones de modo que maximizaran ganancias.
-Se le sugirió a cada participante una posible cadena causal.
-En la primera etapa se les proveía retroalimentación de las consecuencias de sus decisiones, lo cual no sucedió en la segunda etapa sino que ellos mismos debían inferir posibles relaciones causales.
-Se concluyó que los participantes encontraron estructura causal en el problema y la utilizaron para tomar futuras decisiones.\\
-\\
-En el segundo experimento y tercer experimento, a los participantes se les mostró una serie de hipótesis causales para ver si inducían de manera espontánea un modelo causal de la situación.\\
-\\
-Lo que estos experimentos muestran, en conclusión de los autores, es que al intentar maximizar una ganancia a largo plazo, las personas adquieren conocimiento causal del entorno, el cual afecta futuras decisiones.
-Mencionan que estos experimentos ponen en tela de juicio los modelos estándar de toma de decisiones, pues utilizando sólo valores esperados no se llegarían a los mismos resultados.\\
-\\
-Como consecuencia de sus estudios, se obtiene que los métodos basados en los niveles 1 y 2, que es donde caerían los métodos tradicionales de aprendizaje por refuerzo no son suficientes para una adecuada toma de decisiones en un sistema causal, pues estas formulaciones no cuentan con la expresividad suficiente para capturar las relaciones causales.
-Esto confirma la necesidad de pensar en modelos que permitan aprender modelos causales a la par que un agente aprende de su entorno mediante interacción y exploración.\\
-\\
-En cuanto a la formulación de un problema de decisión en sistemas causales, ya desde \cite{dawid2002influence} tenemos una idea de cómo aumentar redes bayesianas con nodos de decisión y variables de estrategia para el modelado causal en problemas de decisión.\\
-\\
-\cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta.
-
-\subsection{Experimentos secuenciales: Causal Bandits}
-
-\section{Mi Propuesta}
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-\subsection{Objetivos}
-\begin{itemize}
-\item Plantear un modelo causal dado un ambiente.
-\item Aprendizaje estructural del modelo causal.
-\item Evaluación del modelo causa (queries significativas)l.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
-\end{itemize}
-\subsection{Motivación}
+\section{Motivación}
 Tomando el cuenta los trabajos mencionados en la sección anterior, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
@@
 Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).
-\subsection{La propuesta}
-Diseñar un agente que tenga como objetivo aprender a llevar a cabo una tarea.
-El aprendizaje se dará por interacción con el ambiente y además contará con un modelo causal que vaya aprendiendo paso a paso y del cual se apoyará para  tomar mejores decisiones.
-\begin{center}
-\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/diagrama.png}
-\end{center}
-\subsection{Pasos a seguir}
+
+\section{Problema de investigación}
+\subsection{Enunciado}
+Aprendizaje de modelos causales a través de la interaccion con el entorno y el uso de este conocimiento para la toma de mejores decisiones en problemas de decisiones secuenciales.
+\subsection{Relevancia}
+Los métodos actuales de Aprendizaje por Refuerzo (RL) son puramente reactivos (\cite{garnelo2016towards}), además de necesitar de mucho tiempo de entrenamiento, pues los métodos usuales requieren de observar muchas parejas acción-estado (\cite{sutton1998reinforcement}).
+Las representaciones del entorno normalmente utilizadas, en términos de Procesos de Decisión Markovianos no permiten hacer razonamiento de alto nivel sobre este entorno (\cite{innes2018reasoning}, \cite{garnelo2016towards}), por lo que aprender estructuras causales permitiran a un agente tomar mejores decisiones pues podrá considerar consecuencias de sus acciones en vez de sólo estar reaccionando a recompensas.
+\subsection{Factibilidad}
+Del trabajo desarrollado en el área de RL sabemos que es posible que un agente autónomo aprenda una enorme variedad de tareas mediante interacción; ejemplos de esto los tenemos en los trabajos de \cite{mnih2013playing}, \cite{mnih2015human}, \cite{silver2016mastering}, \cite{silver2017mastering}, \cite{impala2018}.
+\\
+\\
+Por otro lado, sabemos que los seres humanos adquieren conocimiento causal de su entorno en el caso de problemas de decisión secuenciales (\cite{hagmayer2009decision}, \cite{hagmayer2013repeated}).\\
+\\
+Además, gracias al trabajo de \cite{lattimoreNIPS2016} se sabe que es posible escoger acciones dado un modelo causal.\\
+\\
+Por lo tanto, los diversos elementos existen (en sus ideas de fondo) y pueden ser unidos, aunque no será una tarea trivial. 
+\section{Estado del arte}
+\subsection{Grupos principales}
+\subsubsection{Causalidad}
 \begin{itemize}
-\item ¿Por qué esto sería mejor que trabajo de Garnelo?
-\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? Aquí ayuda el artículo de \cite{lattimoreNIPS2016}
-\item Tengo una política aprendida, ¿cómo descubro estructura causal? Aquí puede ayudar el artículo de \cite{ortega2014generalized}
-\item ¿Qué aprendo primero?
-\item ¿Cómo le doy más peso a la información explicativa que a la información reactiva?
-\item Exploración vs explotación causal: ¿cómo decido si seguir la info causal o tomar acciones exploratorias para descubrir relaciones causales? 
-\item ¿Un dominio de interés? 
+\item Carnegie Mellon (Spirtes)
+\item UC Los Angeles (Pearl)
+\item MIT (Josh Tenenbaum)
+\item Max Planck Institute for Intelligent Systems (Schölkopf)
+\item University of Copenhagen  (Jonas Peters)
+\item Edinburgh (Albrecht, Subramanian Ramamoorthy)
 \end{itemize}
-\subsection{Ejemplo: Caso de un MDP finito}
-En el caso de un MDP finito completamente especificado; es decir, en el cual conocemos las probabilidades de transición y las recompensas.
-Un modelo causal sobre este dominio opera sobre las acciones, pues son estas las que \textbf{\textit{suponemos}} que \textit{causan} algo; esto es,
-\[ a \xrightarrow[]{\text{causa}} \textrm{estado del mundo}. \]
-Entonces, en un modelo gráfico causal necesitaremos tener \textit{nodos de estado}, así como \textit{nodos de acciones} (\textit{causal decision graphs}).\\
-\\
-Es claro que los estados del mundo afectarán la siguiente decisión, pero no causan nada por sí mismos; en todo caso, el estado afeca la decisión pero sólo a través de la política (el valor del estado).\\
-\\
-El objetivo, ahora, es aprender una política que incorpore conocimiento causal, el cual a su vez fue aprendido mediante interacción.\\
-\\
-¿El valor de un estado debe cambiar en el caso causal?
-Para que el valor de un estado se vea afectado por lo que pueda causar el agente a partir de ese estado se necesita algún tipo de función de utilidad, ¿no?
-No lo sé...
+\subsubsection{Reinforcement Learning}
+\begin{itemize}
+\item Google DeepMind (David Silver, Alex Graves, V. Mnih, Murray Shanahan, Marta Garnelo)
+\item Oxford (Nando De Freitas, Shimon Whiteson)
+\item Cambridge (Ghahramani)
+\item  Alberta University (Richard Sutton, Csaba Szepesvári)
+\item UC Berkeley (Pieter Abbeel)
+\item Carnegie Mellon (Salakhutdinov)
+\item Imperial College (Murray Shanahan, Marta Garnelo)
+\item University College London Gatsby Unit (Peter Dayan, Watkins, Dawid)
+\item Edinburgh (Craig Innes)
+\end{itemize}
+\subsubsection{Ciencias Cognitivas}
+\begin{itemize}
+\item Carnegie Mellon (David Danks)
+\item MIT (Josh Tenenbaum)
+\item UC Berkeley (Tom Griffiths)
+\item Edinburgh University (Andy Clark)
+\item Göttingen Institut für Psychologie (York Hagmayer)
+\end{itemize}
 \newpage
 \bibliographystyle{apalike}
-\bibliography{Bibliografia}
-\appendix
-\section{Teoría de Probabilidad}
-\section{Estadística Bayesiana}
-
+\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
 \end{document}

### 2018-02-08 Segundo_Semestre/Seminario/Semana3/Research_Problem.tex
Cambios en la factibilidad

@@
 \begin{itemize}
-\item (\cite{suppes1970probabilistic})
-\item \cite{holland1986statistics}
+\item \cite{holland1986statistics} hace un recuento de varias definiciones de causalidad a lo largo de la historia, desde David Hume, a Patrick Suppes, Granger.
 \item \cite{sutton1998reinforcement}
@@
 \begin{itemize}
-\item Decision problems and causal models: \cite{dawid2002influence} show how causal Bayesian Nets can be augmented with decision nodes as to model decision problems that use causal information.
-\item \cite{hagmayer2009decision}
-\item \cite{meder2009role}
-\item \cite{meder2010observing}
-\item \cite{kemp2010learning}
-\item On-line learning in causal models(\cite{wellen2012learning})):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop.
-
+\item: \cite{dawid2002influence} estudió problemas de decisión y modelos causales, y mostró cómo las redes bayesianas pueden ser \textit{aumentadas} con nodos de decisión que permitan utilizar información causal.
+\item \cite{hagmayer2009decision} muestra que los humanos entendemos nuestras acciones en el mundo como \textit{intervenciones} sobre este. 
+\item On-line learning in causal models(\cite{wellen2012learning})):  Proponen un modelo de aprendizaje causal \textit{on-line} en el cual se añaden o borran arcos en un modelo causal según un ciclo de predicción-error.
 \item \cite{lopez2015towards}
@@
 \begin{itemize}
-\item \cite{eberhardt2008causal} plantea el problema de \textit{descubrimiento causal} como un juego entre un científico vs la naturaleza en el cual la naturaleza intenta mantener ocultos sus secretos y el científico intenta descubrirlos mientras minimiza un costo.
-\item \cite{eberhardt2012number}
-\item \cite{hauser2012two}
-\item Causal learning through sequential decision making (\cite{hagmayer2013repeated}).
-\item \cite{ortega2014generalized} 
+\item \cite{eberhardt2008causal} plantean el problema de \textit{descubrimiento causal} como un juego entre un científico vs la naturaleza en el cual la naturaleza intenta mantener ocultos sus secretos y el científico intenta descubrirlos mientras minimiza un costo.
+\item \cite{hagmayer2013repeated} proponen explorar el uso de información causal en problemas de toma de decisiones secuenciales en seres humanos.
+\item \cite{ortega2014generalized} exploran el muestreo de Thompson para el descubrimiento de estructura causal en problemas de decisión para los cuales la política está especificada.
 \item \cite{bramley2015staying}
@@
 Es principalmente un paper de estructura de grafos y el \textit{aprendizaje} se define en términos de Teoría de Juegos.
-\item \cite{lattimoreNIPS2016} Quieren descubrir la mejor intervención de modo que se minimice el \textit{regret} después de $T$ rondas.
-\item \cite{zhang2017deeper} Una vuelta a Experience Replay
-\item \cite{albrecht2017autonomous}
-\item \cite{albrecht2016exploiting}
-\item \cite{garnelo2016towards}
+\item \cite{lattimoreNIPS2016} buscan descubrir la mejor intervención, dado un modelo causal, para maximizar un retorno.
+\item \cite{garnelo2016towards} proponen un algoritmo de aprendizaje por refuerzo que construye una representación simbólica del ambiente, la cual es usada para escoger la siguiente acción.
 \item \cite{pearl2018theoretical}
@@
 \\
+Notemos que al hablar de modelos gráficos causales que se actualizan y adaptan a la luz de nueva evidencia estamos tocando terreno Bayesiano, pues tenemos creencias actuales sobre la estructura de un grafo, observaremos evidencia, y actualizaremos estas creencias.
+Existen trabajos de inferencia Bayesiana sobre datos estructurados como grafos (\cite{acar2007adaptive}), además de grafos probabilistas (\cite{pearl1988probabilistic}, \cite{koller2009probabilistic}, \cite{sucar2015probabilistic} ).
+\\
+\\
 El trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
@@
 De la misma manera, \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Por lo tanto, es interesante preguntar si un agente que busque maximizar una recompensa pueda aprender modelos causales mediante la interacción con su entorno.
+Por lo tanto, es interesante preguntar si un agente que busque maximizar una recompensa pueda aprender modelos causales mediante la interacción con su entorno.\\
 \\
@@
 \\
-Además, gracias al trabajo de \cite{lattimoreNIPS2016} se sabe que es posible escoger acciones dado un modelo causal.\\
+Además, gracias al trabajo de \cite{lattimoreNIPS2016} se sabe que es posible escoger acciones dado un modelo causal, mientras que el trabajo de \cite{wellen2012learning} muestra que un modelo causal puede ser aprendido \textit{on-line}; es decir, mediante interacción con el ambiente.
 \\

### 2018-02-08 Segundo_Semestre/Seminario/Semana3/Research_Problem.tex
ultima version

@@
 \begin{document}
+\maketitle
+\tableofcontents
 %\nocite{*}
-\maketitle
 \section{Introducción}
-El problema que estamos tratando de atacar consiste en descubrir relaciones causales que se encuentren presentes en diversos fenómenos.
-El enfoque que se propone es descubrir estas relaciones mediante \textit{interacción} con el entorno.\\
+Mientras jugamos un videojuego, o aprendemos a conducir, estamos obligados a tomar decisiones y a intervenir de manera activa en el mundo.
+Este proceso activo de toma de decisiones modifica el estado del mundo, lo cual nos permite aprender cómo nuestras acciones cambiaron el mundo y qué fue causado por estas.
+De esta manera, a través de la interacción con el entorno, podemos descubrir relaciones causales que luego serán utilizadas en posteriores decisiones.\\
+\\
+ Las intuiciones sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
+Se sabe también que el conocimiento causal adquirido puede ser utilizado para mejorar las futuras decisiones tomadas.\\
+\\
+Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.
+Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones.\\
 \\
@@
 \\
- las intuiciones iniciales sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
-Se sabe también que el conocimiento causal adquirido puede ser utilizado para mejorar las futuras decisiones tomadas.\\
-\\
-Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.
-Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones.
 La pregunta se convierte entonces en cómo obtener e incorporar conocimiento causal en un problema de aprendizaje por refuerzo.
@@
 \begin{itemize}
-\item \cite{holland1986statistics} hace un recuento de varias definiciones de causalidad a lo largo de la historia, desde David Hume, a Patrick Suppes, Granger.
-\item \cite{sutton1998reinforcement}
-\item Causal decision theory (\cite{joyce1999foundations}): Toma de decisiones bajo incertidumbre utilizando información causal; es decir; sé que $A$ \textit{causa} $B$, entonces puedo predecir qué pasará si $A$ es escogido, u observado.
-\item (\cite{spirtes2000causation})
-\item \cite{pearl2009causality},
-\item Causal graphical models ( \cite{koller2009probabilistic}): Semántica causal sobre modelos gráficos probabilistas.
-\item Libro de Teoría de la Decisión bajo incertidumbre (\cite{gilboa2009decision}).
+\item \cite{holland1986statistics} hace un recuento de varias definiciones de causalidad a lo largo de la historia, desde Aristóteles, pasando por David Hume, a Patrick Suppes, Granger.
+\item \cite{sutton1998reinforcement} es el libro clásico en el cual se define como aprendizaje por refuerzo a cualquier técnica orientada a resolver problemas mediante la interacción con un entorno y utilizando \textit{feedback}.
+\item \cite{joyce1999foundations} estudia la toma de decisiones bajo incertidumbre utilizando información causal; es decir; sé que $A$ \textit{causa} $B$, entonces puedo predecir qué pasará si $A$ es escogido, u observado.
+\item \cite{spirtes2000causation} junto con \cite{pearl2009causality} estudian los modelos gráficos causales y exponen la definición de causalidad aquí utilizada.
+\item  El libro de \cite{koller2009probabilistic} tiene un capítulo sobre modelos gráficos causales, y explica que estos consisten en dotar de una semántica causal amodelos gráficos probabilistas.
+\item Libro de Teoría de la Decisión bajo incertidumbre \cite{gilboa2009decision}.
 \item \cite{pearl1988probabilistic}, Libro clásico sobre uso de grafos para representar distribuciones.
-\item \cite{van2012reinforcement}, Estado del arte de aprendizaje por refuerzo.
-\item  \cite{danks2014unifying} Representar modelos mentales como modelos gráficos probabilistas; en particular, el problema de inducción causal. 
+\item \cite{van2012reinforcement} presenta el e stado del arte de aprendizaje por refuerzo hasta esa fecha
+\item  \cite{danks2014unifying} explica cómo es posible representar procesos mentales como operaciones sobre gráficos probabilistas; en particular, el problema de inducción causal. 
 \item \cite{gershman2015reinforcement}. Aprendizaje por Refuerzo en el Cerebro.
@@
 \begin{itemize}
-\item: \cite{dawid2002influence} estudió problemas de decisión y modelos causales, y mostró cómo las redes bayesianas pueden ser \textit{aumentadas} con nodos de decisión que permitan utilizar información causal.
+\item \cite{dawid2002influence} estudió problemas de decisión y modelos causales, y mostró cómo las redes bayesianas pueden ser \textit{aumentadas} con nodos de decisión que permitan utilizar información causal.
 \item \cite{hagmayer2009decision} muestra que los humanos entendemos nuestras acciones en el mundo como \textit{intervenciones} sobre este. 
@@
 \item \cite{krynski2007role} Human reasoning under uncertainty naturally operates over causa mental models and statistical data supports correct Bayesian inference only when they can be incorporated into a causal model.
-\item \cite{van2016deep}
-\item \cite{goudet2017learning}
-\item \cite{leibo2017multi}
-\item \cite{hernandez2017survey}
+\item \cite{goudet2017learning} aprendizaje de modelos gráficos a través de redes neuronales generativas.
+\item \cite{hernandez2017survey} estudia la no-estacionareidad de un entorno.
 \item \cite{li2017deep}. Overview general de Deep Reinforcement Learning
@@
 \item \cite{ortega2014generalized} exploran el muestreo de Thompson para el descubrimiento de estructura causal en problemas de decisión para los cuales la política está especificada.
-\item \cite{bramley2015staying}
+\item \cite{bramley2015staying} presenta heurísticas para el aprendizaje causal en el caso de decisiones secuenciales.
 \item \cite{alon2015online} Estudian problemas de aprendizaje on-line en los cuales en cada ronda un jugador escoje una acción y recibe un \textit{loss} previamente especificado.
@@
 \item \cite{garnelo2016towards} proponen un algoritmo de aprendizaje por refuerzo que construye una representación simbólica del ambiente, la cual es usada para escoger la siguiente acción.
-\item \cite{pearl2018theoretical}
-\item \cite{innes2018reasoning}. 
+\item \cite{innes2018reasoning} proponen un algoritmo para que un tomador de decisiones infiera posibilidades no vistas durante el entrenamiento así como las relaciones causales ente datos. 
 \end{itemize}
@@
 Existen trabajos de inferencia Bayesiana sobre datos estructurados como grafos (\cite{acar2007adaptive}), además de grafos probabilistas (\cite{pearl1988probabilistic}, \cite{koller2009probabilistic}, \cite{sucar2015probabilistic} ).
-\\
+Hacer inferencia bayesiana sobre grafos implicaría utilizar distribuciones de probabilidad sobre un espacio de grafos.
+Por eso, utilizar inferencia Bayesiana no-paramétrica (\cite{phadia2015prior}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}) puede ser útil pues admite modelos de alta complejidad.\\
 \\
@@
 El algoritmo de Lattimore le dirá, después de $T$ observaciones, cuál es la mejor intervención a llevar a cabo con el fin de maximizar esa cantidad.
-Este podría ser un primer paso hacia responder la pregunta \textit{¿dado un modelo causal, qué acción tomar?}
+Este podría ser un primer paso hacia responder la pregunta \textit{¿dado un modelo causal, qué acción tomar?}\\
 \\
@@
 \\
-Además, gracias al trabajo de \cite{lattimoreNIPS2016} se sabe que es posible escoger acciones dado un modelo causal, mientras que el trabajo de \cite{wellen2012learning} muestra que un modelo causal puede ser aprendido \textit{on-line}; es decir, mediante interacción con el ambiente.
+Además, gracias al trabajo de \cite{lattimoreNIPS2016} se sabe que es posible escoger acciones dado un modelo causal, mientras que el trabajo de \cite{wellen2012learning} muestra que un modelo causal puede ser aprendido \textit{on-line}; es decir, mediante interacción con el ambiente.\\
 \\
 Por lo tanto, los diversos elementos existen (en sus ideas de fondo) y pueden ser unidos, aunque no será una tarea trivial. 
-\section{Estado del arte}
-\subsection{Grupos principales}
-\subsubsection{Causalidad}
+\section{Grupos principales}
+\subsection{Causalidad}
 \begin{itemize}
@@
 \end{itemize}
-\subsubsection{Reinforcement Learning}
+\subsection{Reinforcement Learning}
 \begin{itemize}
@@
 \end{itemize}
-\subsubsection{Ciencias Cognitivas}
+\subsection{Ciencias Cognitivas}
 \begin{itemize}
@@
 \end{itemize}
-\newpage
 \bibliographystyle{apalike}

### 2018-02-08 Segundo_Semestre/Seminario/Semana3/Presentacion.Rmd
version final

@@
 ---
-title: "Presentation"
+title: "Research Problem"
 author: "Mauricio Gonzalez"
@@
   ioslides_presentation: default
+  beamer_presentation: default
   slidy_presentation: default
@@
 
+----
+
+- The agent will interact with its environment and receive a feedback signal based on the actions taken by him, which will modify his cur- rent causal knowledge.
+![action-feedback loop](/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/cry.png)
+
+----
+
+- Before the next step is taken, the current causal knowledge will be taken into account when choosing the next action together with the current value of each posible action.
+![action-feedback loop](/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/baby.png)
+
+----
+
+- This new action and its outcome will again modify the causal knowl- edge, which will improve the next action and so on.
+```{r fig.width=5, fig.height=5,echo=FALSE}
+library(png)
+library(grid)
+img <- readPNG("/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/Proposed.png")
+grid.raster(img)
+```
+
 ## Fields involved
@@
 - Causal learning.
+- Bayesian inference.
 
@@
 ## Statement
-- Acquiring causal kwnoledge via interaction with the environment and using this knowledge to guide a decision-making process.
+- **Acquiring causal kwnoledge via interaction with the environment and using this knowledge to guide a decision-making process**
+- Lay a bridge between Reinforcement Learning and Causal Learning: fundamental assumptions.
+- Why reinforcement learning? Interventions are needed in order to acquire causal knowledge and RL provides an action-reaction interactive loop.
 
@@
 - Do not allow high-level reasoning.
+- Learning causal structure will allow the agent to make better decisions because it will be able to consider consequences of his actions, and not only reacting to rewards.
 
@@
 - The Devil is in the details.
+
+## Disclosure
+- Causality as in [@pearl2009causality] and [@spirtes2000causation]
+- Other definitions of Causality exist.
 
@@
     + University College London (Peter Dayan, Watkins, Dawid)
+    + Edinburgh (Craig Innes)
 
@@
 ## Outer Circle
+- Causality
+
 Reference                  | Summary                                |
@@
 [@holland1986statistics]   | General review of different views on Causality |
-[@pearl1988probabilistic]  | Book on Networks for probabilistic reasoning|
-[@heckerman1995decision] ||
+[@heckerman1995decision]   |Decision-theoretic foundations for causal reasoning  |
 [@spirtes2000causation]    | Classic General book on Causality      |
@@
 [@koller2009probabilistic] | Book on Probabilistic Graphical Models |
-[@lopez2015towards]        | Discriminative data-driven framework for distinguishing direction of causality |
-[@goudet2017learning]      | Learning to generate data with neural networks |
 
 ----
+
+- Reinforcement Learning
 
@@
 
+- Cognitive aspects
+
 Reference            | Summary                                                                      |
 ---------------------|------------------------------------------------------------------------------|
-[@krynski2007role]   | 
+[@krynski2007role]   | The role of causality in judgment under uncertainty
 [@danks2014unifying] | Cognitive functions can be modelled as operations on graphical models        |
@@
 [@dawid2002influence] | Influence Diagrams|
-[@eberhardt2008causal] ||
-[@hagmayer2009decision] | |
-[@meder2009role] | |
-[@meder2010observing] | |
-[@kemp2010learning] | |
+[@eberhardt2008causal] |Causal discovery as a game vs Nature|
+[@hagmayer2009decision] | Choices as interventions |
 [@wellen2012learning] | learning causal structure in local maner via precition-error loop |
-[@eberhardt2012number] ||
 
@@
 ----------|---------|
-[@mnih2013playing] | Introducing Deep Q-Network |
+[@mnih2013playing] | Introducing Deep Q-Network with Experience Replay |
 [@silver2014deterministic] | Deterministic Policy Gradients |
-[@mnih2015human] |  Nature report on Deep Q-Networks |
+[@lopez2015towards]        | Discriminative data-driven framework for distinguishing direction of causality |
 [@DBLP:journals/corr/SchulmanLMJA15] | Trust Region Policy Optimization |
-[@DBLP:journals/corr/MnihBMGLHSK16] | Asynchronous Methods for Deep Reinforcement Learning |
-[@van2016deep] ||
-[@lakkarajuNIPS2016] ||
-[@toulisNIPS2016] ||
-[@kulkarniNIPS2016] | Hierarchical Deep Reinforcement Learning |
-[@foersterNIPS2016] | Learning to Communicate with Deep Multi-Agent Reinforcement Learning |
-[@lanctot2017] | Game-Theoretic Approach to Multiagent Reinforcement Learning  |
 [@goudet2017learning] | Functional Causal Models with Generative Neural Networks |
-[@rahmanian2017] | Online Dynamic Programming  |
-[@louizos2017] |Causal Effect Inference with Deep Latent-Variable Models  |
-[@christiano2017] |Deep Reinforcement Learning from Human Preferences|
 [@o2016pgq] | Combining policy gradient and q-learning |
-[@kirsch2017mdp | Environments |
-[@zhang2017deeper] ||
 
@@
 ----------|---------|
-[@hauser2012two] ||
 [@hagmayer2013repeated] | How repeated decision making is guided by causal beliefs. |
@@
 [@alon2015online] | When feedback is specified by graph, the graph structure controls the difficulty of learning. |
-[@bramley2015staying] ||
 [@lattimoreNIPS2016] | Causal Bandits: Given a causal model, choose best intervention. |
-[@albrecht2016exploiting] ||
+
+----
+
+Reference | Summary |
+----------|---------|
+[@albrecht2016exploiting] |Exploiting Causality for Selective Belief Filtering in Dynamic {B}ayesian Networks|
 [@garnelo2016towards] | Reinforcement Learning framework that builds symbolic representation of environment. |
 [@innes2018reasoning] | An agent which, starting unaware of factors on which an optimal policy depends, learns optimal behaviour for single-stage decision problems via direct experience and advice from a domain expert.|
-[@pearl2018theoretical] ||
+[@pearl2018theoretical] |Theoretical Impediments to Machine Learning With Seven Sparks from the Causal Revolution|
 
-## End
-- Thnks.
+----

### 2018-02-14 Segundo_Semestre/Propuesta/Propuesta.tex
Añadiendo td0

@@
 \section{Aprendizaje por Refuerzo}
-El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno.
+\subsection{Definición}
+El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de la interacción con su entorno.
 Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
@@
 \\
-La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
-El objetivo del agente es encontrar la política que maximice la recompensa.\\
+La diferencia fundamental entre el aprendizaje por refuerzo y el aprendizaje supervisado o no supervisado consiste en la existencia de una retroalimentación posiblemente retrasada y la consideración de horizontes de tiempo; un agente de RL busca maximizar ganancias \textit{a largo plazo}.\\
+\\
+La dinámica del agente y su entorno puede esquematizarse de la siguiente manera: en un momento $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
+El objetivo del agente es encontrar la política que maximice la recompensa a largo plazo.\\
 \\
@@
 Veremos ahora con detalle las partes que conforman esta dinámica. 
-\subsubsection{Objetivos y Recompensas}
+\subsection{Objetivos y Recompensas}
 Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada.
@@
 Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
-\subsubsection{Definición de Cadenas de Markov}
+\subsection{Definición de Cadenas de Markov}
 Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
@@
 Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
-\subsubsection{Procesos de Decisión Markovianos con recompensa}
+\subsection{Procesos de Decisión Markovianos con recompensa}
 Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$.
@@
 
-\subsubsection{Función valor}
+\subsection{Políticas y Función valor}
 Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado.
-Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política.
-Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$.
-Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$.
+Debido a que las recompensas obtenidas a largo plazo dependen de toda una secuencia de acciones, tenemos que definir la función valor en términos de una política.
+Se había mencionado que una política era una manera de tomar acciones; formalmente, una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$; esto es, la probabilidad de tomar la acción $a$ estando en el estado $s$.\\
+\\
 La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante.
@@
 \[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
-\subsubsection{Soluciones y Optimalidad}
+\subsection{Soluciones y Optimalidad}
 Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa.
@@
 A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general.
-La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
-\begin{itemize}
-\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
+La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}.\\
+\\
+Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
 Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$.
@@
 Ambos métodos convergen a $V^\pi$.
-\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
+Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
 Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política.
-Entonces, es importante tener un método para estimar $Q^\ast$.
-\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
-El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.
-\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
+Entonces, es importante tener un método para estimar $Q^\ast$.\\
+\\
+Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
+El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.\\
+\\
+Para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
 A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
@@
 \[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
-\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
-\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
+Podemos remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
+En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
 		\begin{itemize}
@@
 		\end{itemize}
-\end{itemize}
 \subsubsection{Diferencias Temporales}
-Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos.
-Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
-\begin{itemize}
-\item Evaluación de política en Diferencias Temporales.
-\item TD(0).
-\item SARSA: Es un método \textit{on-policy}.
-\item Q-Learning: Es un método \textit{off-Policy}.
-\item Métodos Actor-Critic.
-\end{itemize}
+Los métodos de Diferencias Temporales (TD por sus siglas en inglés)  son una idea novedosa, pues combinan ideas de los Métodos Monte Carlo y de los métodos de Programación Dinámica (\cite{sutton1998reinforcement}).\\
+\\
+Los métodos TD son totalmente incrementales y  norequieren un modelo del ambiente, pero son mucho más complejos; los métodos que veremos en esta sección también pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero a diferencia de los métodos Monte-Carlo, utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.\\
+\\
+Tanto los métodos Monte-Carlo como los métodos TD utilizan experiencia para actualizar sus estimaciones del valor de un estado; dada una política $\pi$, los métodos de ambas clases actualizarán la estimación $V(s_t)$ según lo que pace después de visitar ese estado $s_t$, la diferencia está en que los métodos Monte-Carlo deben esperar al final de un episodio para poder hacer las actualizaciones según las recompensas, pues se utilizan los retornos promedio al visitar los estados como aproximaciones del valor de este.
+En el caso de los métodos TD, inmediatamente después de visitar cada estado, la actualización se hará utilizando la recompensa observada $r_{t+1}$ y la estimación de $V(s_{t+1})$.
+Esto da lugar al método más sencillo de todos los TD, el cual se conoce como $TD(0)$.
+El método $TD(0)$  lleva a cabo la siguiente actualización
+\[ V(s_t) \leftarrow V(s_t) + \alpha ( r_{t+1} + \gamma V(s_{t+1}) - V(s_t)). \]
+SARSA: Es un método \textit{on-policy}.
+\\
+Q-Learning: Es un método \textit{off-Policy}.
+\\
+Métodos Actor-Critic.
 \subsection{Value based Reinforcement Learning: Deep Q-Networks}
@@
 Debe hacerse el siguiente cambio en la actualización:
-\[ r_{t+1}  + \gamma  Q (s_{t+1}, \textrm_{ arg max}_a Q(s_{t+1},a_t,\theta_t); \theta^{-}\]
+%\[ r_{t+1}  + \gamma  Q (s_{t+1}, \textrm_{ arg max }_a Q(s_{t+1},a_t,\theta_t); \theta^{-} \]
 donde $\theta$ es el parámetro de la red on-line y $\theta^{-}$ el parámetro de la red objetivo
@@
 \section{Mi Propuesta}
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-\subsection{Objetivos}
-\begin{itemize}
-\item Plantear un modelo causal dado un ambiente.
-\item Aprendizaje estructural del modelo causal.
-\item Evaluación del modelo causa (queries significativas)l.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
-\end{itemize}
 \subsection{Motivación}
@@
 Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).
+\subsection{Pregunta de Investigación}
+¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
+\subsection{Problema}
+Adquirir y utilizar información causal para mejorar el desempeño de un agente que interactua con su entorno con el fin de maximizar una recompensa a largo plazo. 
+\subsection{Objetivos}
+\begin{itemize}
+\item Plantear un modelo causal dado un ambiente.
+\item Aprendizaje estructural del modelo causal.
+\item Evaluación del modelo causa (queries significativas)l.
+\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
+\item Exportar conocimiento causal obtenido para hacer inferencia causal.
+\end{itemize}
 \subsection{La propuesta}
@@
 \\
+Suponiendo que conocieramos el modelo causal, entonces este nos ayudará a escoger acciones.
+Ya no debería usar políticas $\varepsilon$-greedy para explorar el espacio, pues ya conozco relaciones útiles.
+El enfoque de \cite{lattimore2017nips} podría servir aquí.
+\\
+Por otro lado, cómo ir estimando la estructura causal del grafo?
+Parece ser que la estadística Bayesiana es una idea, pues en cada pasao tengo creencias sobre la estructura causal que voy a modificar a la luz de la evidencia, la evidencia está dada por los descubrimientos del agente y las creencias por la estructura actual del grafo.
+\\
 ¿El valor de un estado debe cambiar en el caso causal?

### 2018-02-19 Segundo_Semestre/Propuesta/Propuesta.tex
Añadiendo estado del arte segun alguna taxonomia

@@
 \[ V(s_t) \leftarrow V(s_t) + \alpha ( r_{t+1} + \gamma V(s_{t+1}) - V(s_t)). \]
-SARSA: Es un método \textit{on-policy}.
+Ahora, veremos un método \textit{on-policy} en el cual se consideran transiciones de parejas estado-acción a parejas estado-acción.
+La actualización de valor utilizará toda la información  de la transición; es decir, $(s_t,a_t,r_{t+1},s_{t+1},a_{t+1})$ y de aquí el nombre del método: SARSA.
+La actualización que lleva a cabo este método es
+\[ Q(s_t,a_t) \leftarrow Q(s_t,a_t) + \alpha (r_{t+1} + \gamma Q(s_{t+1},a_{t+1})-Q(s_t,a_r)) \]
 \\
-Q-Learning: Es un método \textit{off-Policy}.
+Por otro lado, tenemos el algoritmo Q-learning (\cite{watkins1992q}) que es \textit{off-policy} y sus actualizaciones las hace como si se hubiera escogido la acción greedy, aunque el agente no esté usando esa política
+\[ Q(s_t,a_t) \leftarrow Q(s_t,a_t)  + \alpha (r_{t+1} + \gamma \max_a Q(s_{t+1},a) - Q(s_t,a_t).  \]
 \\
@@
 \subsection{Towards Deep Symbolic Reinforcement Learning}
-Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
+Deep Reinforcement Learning (DRL) consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
 \\
@@
 \subsection{Experimentos secuenciales: Causal Bandits}
-
+\subsection{Toma de Decisiones Bajo Incertidumbre y Teoría de Juegos}
+Teoría de Juegos (\cite{von1944theory}, \cite{osborne1994course}, \cite{webb2007game}, \cite{gintis2014bounds})
+\subsubsection{Aprendizaje por refuerzo como toma de decisiones}
+\subsubsection{Juegos Bayesianos}
+\subsubsection{Información causal y juegos}
 \section{Mi Propuesta}
+\subsection{Pregunta de Investigación}
+¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
+\subsection{Problema}
+Adquirir y utilizar información causal para mejorar el desempeño de un agente que interactua con su entorno con el fin de maximizar una recompensa a largo plazo. 
 \subsection{Motivación}
-Tomando el cuenta los trabajos mencionados en la sección anterior, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
+Tomando el cuenta los trabajos mencionados en las sección anterior, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
 Este agente buscará aprender un modelo causal de su entorno al mismo tiempo que aprende una \textit{política}.
@@
 Por otro lado, en el caso inverso \cite{ortega2014generalized} exploran un esquema de muestreo llamado Muestreo de Thompson, en el cual un agente muestrea una de entre varias acciones de acuerdo a la probabilidad subjetiva que tenga el agente de que esta sea la mejor acción.
-Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-Adquirir y utilizar información causal para mejorar el desempeño de un agente que interactua con su entorno con el fin de maximizar una recompensa a largo plazo. 
+Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).\\
+\\
+\cite{ho2016reinforcement}\\
+\subsection{Estado del arte}
+De manera más detallada, podemos ver que los diferentes enfoques que se han utilizado para la adquisición, y uso, de información causal en problemas de decisiones secuenciales pueden dividirse de la siguiente manera:
+\subsubsection{Métodos basados en actualización de grafos}
+\cite{heckerman2006bayesian}, \cite{wellen2012learning}, \cite{kummerfeld2013tracking}
+\subsubsection{Métodos basadps en representaciones del ambiente}
+\cite{ho2013knowledge}, \cite{ho2014rapid} \cite{garnelo2016towards}
+\subsubsection{Métodos basados en decisiones secuenciales}
+\cite{hagmayer2008causal}, \cite{hagmayer2013repeated}, \cite{alon2015online}, \cite{albrecht2016exploiting} \cite{lattimoreNIPS2016}, \cite{innes2018reasoning}
 \subsection{Objetivos}

### 2018-02-19 Segundo_Semestre/Seminario/Semana6/State_of_Art.tex
Estado del arte para seminario

--- 2018-02-08 Segundo_Semestre/Propuesta/Propuesta.tex
@@
 \newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Survey}
+\title{State of The Art Circles}
 \author{Mauricio Gonzalez Soto}
 \begin{document}
-\nocite{*}
+%\nocite{*}
 \maketitle
-\tableofcontents
-\newpage
-\section{Introducción}
-El problema que estamos tratando de atacar consiste en descubrir relaciones causales que se encuentren presentes en diversos fenómenos.
-El enfoque que se propone es descubrir estas relaciones mediante \textit{interacción} con el entorno.
-En el contexto de inferencia causal, los datos que, usualmente, se tienen provienen de fuentes ruidosas por lo que la probabilidad será el lenguaje adecuado para razonar sobre estos datos.\\
-\\
-\cite{lake2017building} argumentan que los sistemas inteligentes actuales, a pesar de sus impresionantes logros, aun están lejos de alcanzar una verdadera inteligencia; uno de los requisitos para esto es que puedan construir modelos causales que sean interpretables, y no sólo reconocer patrones estadísticos.
-\\
-La primera parte de este resumen consiste en el estudio clásico de relaciones causales, el cual depende fuertemente de Redes Bayesianas, las cualesson objetos matemáticos utilizados para representar distribuciones de probabilidad de manera eficiente; además, permiten incorporar conocimiento inicial sobre las variables aleatorias representadas en la red.
-Debido a que estas redes son de fácil interpretación, es natural definir modelos causales en términos de estas, pues fácilmente nodos y aristas pueden traducirse en términos de objetos y causas.\\
-\\
-Después, veremos que existe un área del aprendizaje de máquina llamado Aprendizaje por Refuerzo, que es un marco teórico para el aprendizaje automático en el cual un \textit{agente} aprende a realizar una tarea a través de experiencia y de interacción con su entorno.\\
-\\
-Además, las intuiciones iniciales sobre aprendizaje causal por interacción provienen de estudios realizados en Psicología Computacional y Ciencias Cognitivas, de aquí se sabe que los seres humanos obtienen conocimiento causal al encontrarse en situaciones en las cuales deben tomar decisiones de manera secuencial para lograr un objetivo futuro.
-Se sabe también que el conocimiento causal adquirido puede ser utilizado para mejorar las futuras decisiones tomadas.\\
-\\
-Notando que los modelos causales tradicionales requieren de intervenciones y experimentación en el mundo, y que estas intervenciones son una forma de interacción y alteración del mundo, salta a la vista la relación entre Aprendizaje por Refuerzo y aprendizaje causal.
-Entonces, resulta natural preguntar cómo pueden aprenderse relaciones causales de manera interactiva de modo que el conocimiento causal adquirido permita a un agente tomar mejores decisiones.
-La pregunta se convierte entonces en cómo obtener e incorporar conocimiento causal en un problema de aprendizaje por refuerzo.
-Además, uno se pregunta si es posible utilizar este conocimiento causal obtenido al interactuar sobre el mundo para realizar otro tipo de inferencias causales en ciertos dominios de interés.
-\section{Causalidad}
-\subsection{Intro}
-Saber si es posible hablar de relaciones causa-efecto es un tema polémico; los métodos probabilistas tradicionales sólo codifican asociaciones estadísticas, pero la idea
-La diferencia entre modelos probabilistas y modelos causales se da cuando es de interés tomar acciones que modifiquen el entorno y no nada más observar los valores de las variables. \\
-Según \cite{pearl2014probabilistic}, las dos preguntas fundamentales en Causalidad son:
-\begin{itemize}
-\item ¿Qué evidencia empírica es necesaria para hacer inferencias válidas sobre relaciones causa-efecto?
-\item Dado que estamos dispuestos a aceptar información causal sobre un fenómeno, ¿qué inferencias podemos obtener a partir de esta información, y cómo?
-\end{itemize}
-En cuanto al segundo punto, Pearl considera que hay tres tipos de preguntas a considerar
-\begin{itemize}
-\item Predicciones: si sabemos que una persona ingirió alimentos de dudosa procedencia, ¿se enfermará?
-\item Intervenciones: si le damos cierto tratamiento a un paciente, ¿mejorará?
-\item Contrafactuales: ¿el paciente del punto uno se habría enfermado de todos modos si no hubiera consumido esos alimentos?
-\end{itemize}
-Pearl argumenta que aunque causalidad suele entenderse como afirmaciones necesarias, estas son hechas en contexto de incertidumbre, por lo que cualquier teoría causal debe estar en términos de probabilidades.
-\subsection{Un poco de historia}
-Aristóteles definió 4 \textit{causas} para una cosa, una de ellas llamada \textit{causa eficiente} que estaba definida como aquello que hace la cosa.\\
-\\
-En 1690, el filósofo inglés John Locke propuso como definición de causa “aquello que produce alguna idea simple o compleja”, y lo que es producido lo definió como efecto.\\
-\\
-\cite{holland1986statistics} resalta la importante diferencia en que Aristóteles hizo énfasis en las causas de una cosa, pero no en el efecto de estas. \\
-\\
-Fue David Hume quien se considera que hiz una importante contribución, y él definió la causalidad como una relación entre experiencias más que una entre hechos.
-Argumentó que no es verificable empíricamente que la causa produzca el efecto, sino sólo que el evento llamado \textit{causa} es invariablemente seguido por aquella experiencia llamada \textit{efecto}.\\
-\\
-Hume definió 3 criterios básicos para causalidad:
-\begin{itemize}
-\item Contiguidad espacio-temporal: tanto causa y efecto tienen lugar en una entidad común (espacio y tiempo).
-\item Sucesión temporal: La causa tiene que preceder temporalmente al efecto. 
-\item Conjunción constante: Causa y efecto ocurren juntos, o no ocurren juntos.
-\end{itemize}
-John Stuart Mill argumentó tajantemente que observación sin experimentos no puede probar causalidad. Mill identificó cuatro métodos generales para la identificación de relaciones causales. 
-\subsection{Modelo de Rubin}
-El modelo de Rubin parte de una población de \textit{unidades} $U$.
-En general, estas unidades son los objetos de estudio sobre las cuales actuan las causas.
-Para hacer más sencilla la descripción del modelo, supondremos que sólo existen dos causas (o niveles de tratamiento) denotadas por $t$ y $c$.
-Sea $S$ una variable indicadora de la causa a la cual cada unidad $u \in U$ fue expuesta.
-Además, consideramos dos variables respuesta $Y_t$, $Y_c$.
-$Y_t(u)$ reoresenta el valor de la respuesta si la unidad $u$ es expuesta a $t$ y análogo para $c$.\\
-\\
-\begin{defi}
-Se define el efecto de la causa $t$ sobre $u$ relativo a la causa $c$ como
-\[ Y_t(u) - Y_c(u). \]
-Esta expresión se lee de la siguiente manera: $t$ causa el efecto $Y_t(u) - Y_c(u)$ en la unidad $u$ (relativo al control $c$).
-\end{defi}
-Notemos que es imposible observar simultáneamente el efecto de $t$ y de $c$ sobre una misma unidad $u$ y a esto \cite{holland1986statistics} le llama el Problema Fundamental de la Inferencia Causal.
-El problema, según explica Holland, radica en la palabra \textit{observar} pues no podemos saber simultáneamente el efecto de exponer a un enfermo en particular a un nuevo tratamiento y al mismo tiempo el efecto de no exponerlo.
-\subsection{Modelo de Suppes}
-En su libro “Probabilistic Theory of Causality”, (\cite{suppes1970probabilistic}), Patrick Suppes intenta mejorar el análisis de Hume sobre causalidad; en particular, la condición de conjunción constante.
-Para Suppes, no importa qué sean las causas y los efectos, sino que se puedan expresar como eventos que ocurren bajo ciertas condiciones.
-Suppes supone que todas las causas necesariamente preceden en el tiempo a sus efectos.
-Define una causa \textit{prima facie} de un evento como un evento que precede temporalemnte a este y que está asociado de manera positiva con él.
-Define como una causa espuria a una causa prima facie tal que es condicionalmente independiente del efecto dada otra causa la cual es anterior en el tiempo a la causa prima facie y que está asociada con el evento dada la causa.
-Una causa genuina es una causa prima facie que no es espuria.
-\begin{defi}
-Si $r < s$ representan valores de tiempo, se dice que el evento $C_r$ es una causa prima facie del evento $E_s$ si
-\[ P(E_s | C_r ) > P(E_s). \]
-\end{defi}
-\begin{defi}
-Un evento $C_r$ se dice que es una causa espuria de un eventi $E_s$ si $C_r$ es una causa prima facie de él y para alguna $q < r < s$ existe un evento $D_q$ tal que 
-\[P(E_s | C_r, D_q) = P (E_s | D_q) \]
-y además,
-\[P(E_s | C_r , D_q) \geq P(E_s | C_r). \]
-\end{defi}
-Por lo tanto,
-\begin{defi}
-Para tiempos $r < s$, un evento $C_r$ se dice que es una causa genuina de un evento $E_s$ si $C_r$ es una causa prima facie de $E_s$, pero no es una causa espuria.
-\end{defi}
-Notemos que Suppes está definiendo la causa de un efecto, más que el efecto de una causa.
-Además, no hay restricciones a la naturaleza de una causa más allá de que sea un evento que ocurre antes en el tiempo.
-El modelo de Suppes no puede expresar el efecto de una causa en particular, sólo comportamientos promedio (\cite{holland1986statistics}).\\
-\\
-En palabras de Holland, la causalidad de Suppes no es más que correlación entre causa y efecto que no puede ser eliminada al parcializar causas legítimas.
-\subsection{Modelos Causales en Medicina}
-Un caso de aplicación en medicina es determinar que ciertas bacterias causan ciertas enfermedades.
-Existe un conjunto de postulados planteados por Koch para decidir cuándo un micro-organismo está implicado en una enfermedad.
-Aunque no existe un planteamiento formal, estos pueden expresarse como (\cite{yerushalmy1959methodology}):
-\begin{itemize}
-\item El organismo debe estar presente en todos los casos de la enfermedad supuestamente provocada por él.
-\item El organismo debe aislarse de los pacientes y cultivarse en estado puro.
-\item El organismo cultivado en estado puro debe reproducir el padecimiento. 
-\end{itemize}
-Estos postulados aseguran que no haya datos que apoyen una hipótesis nula de cero efecto causal. \\
-\\
-Por otro lado, Hill, quien fue de los primeros en estudiar la asociación entre fumar y cáncer, encontró 9 factores útiles a la hora de decidir que una asociación observada es causal. 
-\subsection{Causalidad de Granger en Economía}
-La causalidad de Granger se utiliza para series de tiempo, lo cual es de alta importancia en Economía.
-\cite{holland1986statistics} opina que las ideas esenciales de Granger no están limitadas a series de tiempo y que pueden ser expresadas en términos de variables que actuan sobre unidades de una población.\\
-\\
-En la teoría de Granger, una variable causa otra variable; es decir ( \cite{holland1986statistics}), las realizaciones de una variable mejoran las predicciones hechas sobre la otra variable.
-La única parte en la que se utilizan series de tiempo en esta definición es para hablar de valores de las variables definidos hasta un cierto tiempo $t$.\\
-\\
-Si $X,Y,Z$ son variables aleatorias definidas sobre cierta población, se dice que $X$ no es una causa de Granger de $Y$ si $X$ y $Y$ son condicionalmente independientes dado $Z$.
-Entonces, $X$ es una causa de Granger de $Y$ si distintos valores de $X$ llevan a distintas distribuciones predictivas de $Y$ dados $X$ y $Z$; esto es, $X$ ayuda a predecir $Y$ aun cuando $Z$ es tomada en cuenta.
-La no-causalidad de Granger es parecida a las causas espurias de Suppes pues ambas consideran que no se pueda predecir un evento dada cierta información.\\
-\\
-En opinión de Holland, la causalidad de Granger falla de la misma manera que la de Suppes, pues a la luz de nueva información una causa Granger puede dejar de serlo.
-\subsection{Modelos Causales en Ciencias Sociales}
-\subsection{Aprendizaje de Modelos Causales a partir de datos}
-En los métodos de estado del arte de Modelos Causales Funcionales suele hacerse los siguientes supuestos (\cite{goudet2017learning}):
-\begin{itemize}
-\item Suficiencia Causal (CSA)
-\item Propiedad Markoviana (CMA)
-\item Independencias Condicionales
-\item Fiabilidad causal (CFA)
-\end{itemize}
-\subsubsection{Aprender el CPDAG}
-Existen tres familias de métodos para extraer el grafo: métodos de restricciones, métodos de score y métodos híbridos.\\
-\\
-Los métodos basados en restricciones explotan las independencias condicionaes para identificar todas las v-estructuras.
-El algoritmo más conocido es el Algoritmo PC (\cite{spirtes2000causation}).
-El algoritmo PC primero construye un esqueleto basado en independencias condicionales; luego, identifica las v-estructuras y posteriormente orienta las aristas.
-El algoritmo FCI (\cite{spirtes1999algorithm}) extiende el algoritmo PC al relajar la suficiencia causal e incorporar variables latentes.
-El algoritmo RFCI de \cite{colombo2012learning} es más rápido que FCI y maneja datos de alta dimensionalidad.
-La debilidad de estos métodos, según \cite{goudet2017learning} es que dependen demasiado en pruebas de independencia, que dependen a su vez de la cantidad disponible de datos.\\
-\\
-Los métodos basados en score exploran el espacio de CPDAG's y minimizan un score global.
-La exploración se realiza en términos de operaciones sobre grafos, por ejemplo añadir o remover aristas.
-El algoritmo Greedy Equivalent Search (GES) de \cite{chickering2002optimal} encuentra la estructura óptima en términos del BIC.\\
-\\
-Los métodos híbridos combinan las técnicas anteriores.
-Por ejemplo el Max-Min Hill Climbing Algorithm de \cite{tsamardinos2006max} primero construye un esqueleto utilizando pruebas de independencia condicional, y luego le da orientación a las aristas utilizando un método greedy sobre un score bayesiano.
-\subsubsection{Explotando la asimetría entre causa y efecto}
-Los métodos anteriores no toman en cuenta la asimetría entre causas y efectos.
-La intuición detrás (\cite{goudet2017learning}) está dada por el siguiente ejemplo: Si consideramos el modelo causal funcional $Y=X+U$, entonces los métodos anteriores basados en grafos no pueden darle orientación a la arista $X - Y$, pues tanto $X \rightarrow Y$ como $Y \rightarrow X$ son Markov equivalentes.\\
-\\
-\cite{stegle2010probabilistic} infiere la dirección causal al construir dos modelos Bayesianos generativos: uno para $X \rightarrow Y$ y otro para $Y \rightarrow X$.
-La dirección causal es determinada a partir del modelo generativo que mejor ajuste a los datos.
-Otros métodos, como el propuesto por \cite{sgouritsa2015inference} se basan en que si $X \rightarrow Y$, entonces la probabilidad marginal de la causa $P(X)$ es independiente del mecanismo causal $P(X | Y)$; por lo tanto, estimar $P(Y | X)$ a partir de $P(X)$ no debería poderse, pero estimar $P(X | Y)$ a partir de $P(Y)$ sí debería ser posible.
-\subsubsection{Como problema de Aprendizaje}
-Inspirados en los concursos de Kaggle de \cite{Guyon2013}, \cite{lopez2015towards} plantean el problema de inferencia causal como un problema de clasificación de medidas de probabilidad.
-Los datos provistos por el concurso son 16,200 parejas de variables $(X_i,Y_i)$.
-Cada par consta de una realización $(x_i,y_i)$ y una etiqueta $l_i$ que indica el valor de la verdadera relación causal; es decir, $l_i$ indica uno de los siguientes casos: $X \rightarrow Y,$ $Y \rightarrow X$, $X_i \perp Y_i$, $X \leftrightarrow Y$.\\
-\\
-Notemos que este método tiene como punto débil que depende de la representatividad de la muestra de entrenamiento y que supone la existencia de una distribución madre
-\subsubsection{Modelos Causales Funcionales Neuronales}
-\cite{goudet2017learning} proponen un esquema que consiste en tres pasos: primero, identificar el esqueleto según propiedades de independencia estándar; luego, orientar las aristas utilizando métodos bivariados estándar y finalmente hacer una optimización sobre todo el grafo examinando relaciones locales.
-Este último paso se realiza al aprender estos modelos locales como redes neuronales generativas; es decir, se modela la distribución $P(X | X_{Pa_i}, U_i)$ de un nodo condicional en sus padres (causas) con una red neuronal.\\
-\\
-Formalmente, se modela cada ecuación
-\[X_i = f_i (X_{Pa_i},U_i) \]
-como una red neuronal generativa global, en la cual cada función $f_i$ es modelada por una red neuronal generativa con una capa oculta.\\
-\\
-Encontrar un modelo causal de este tipo requiere llevar a cabo dos tareas: estimar la estructura (el grafo) de los padres de cada variable, y las redes neuronales $f$
-\subsection{Correlación y causalidad}
-Dos variables aleatorias $X$,$Y$ pueden estar correlacionadas en distintos casos; por ejemplo, si $X$ causa $Y$, o si tanto $X$ como $Y$ son causadas por otra variable desconocida $V$.
-Si e conoce la existencia de $V$ y además podemos observarla.
-
-\section{Modelos Gráficos Probabilistas}
-
-\section{Modelos gráficos causales}
-Sabemos que en una red bayesiana los arcos, aunque dirigidos, no necesariamente tienen una interpretación directa, lo único que importa es que el grafo capture adecuadamente las relaciones de dependencia entre variables aleatorias.
-Nos gustaría que fuera el caso que $X \to Y$ significara $X$ \textit{causa} Y.
-Resulta que existe una familia de modelos causales en los cuales se \textit{interviene} en el mundo.
-La idea clave es utilizar \textit{intervenciones ideales}, las cuales se denotan $do(Z=z)$ que significa forzar a la variable aleatoria $Z$ a tomar el valor $z$, denotado de manera simple como $do(z)$.
-Esta operación corresponde al caso en que un agente directamente modificó el estado del mundo y no sólo observó los valores de este de manera pasiva; por esta razón es muy importante tener en mente que es distinto $P(Y | do(z), X=x)$ que $P(Y | Z=z, X=x)$.
-Las preguntas causales de interés son planteadas de la forma $P(X | do(z))$ y son llamadas \textit{quries intervencionales}.
-Este tipo de intervenciones surgen de manera natural en muchos problemas; por ejemplo, “si el paciente recibe este tratamiento, cuál es la probabilidad de sobrevivir la enfermedad” puede ser planteada como $P(sobreivir | do(tratamiento))$.\\
-\\
-Formalmente, un modelo gráfico causal (\cite{koller2009probabilistic}, \cite{pearl2014causality}) tiene la misma forma que una red Bayesiana; es decir, consiste en una red acíclica dirigida que codifica una distribución entre un conjunto de variables aleatorias.
-La diferencia radica en la interpretación de las flechas, pues cada variable aleatoria está gobernada por un modelo causal en función de sus padres en el grafo.
-Una intervención consiste en forzar uno de los nodos a tomar un valor con probabilidad 1 y a recortar todas las flechas que entran en ese nodo.\\
-\\
-Es importante que se cumpla que si dos variables están correlacionadas, entonces una es padre causal de la otra y que no falten padres causales de ninguna variable.\\
-\\
-Para poder contestar pregunta acerca de intervenciones, se necesita introducir el concepto de \textit{grafo mutilado}
-\begin{defi}
-Sea $\mathcal{B}$ una red sobre un conjunto de variables aleatorias y sea $Z_1=z_1,...,Z_k=z_k$ una instanciación de estas variables denotada por $\mathbf{Z}=\mathbf{z}$ .
-Se define el grafo mutilado $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$ como:
-\begin{itemize}
-\item Cada nodo $Z_i \in \mathbf{Z}$ no tiene padres en $\mathcal{B}_{\mathbf{Z}=\mathbf{z}}$
-\item El modelo local de probabilidad condicional para $Z_i \in \mathbf{Z}$ es tal que asignal probabilidad uno al valor $z_i$ y cero a cualquier otro valor.
-\item No se altera ningún otro nodo $X_i \notin \mathbf{Z}$
-\end{itemize}
-\end{defi}
-Entonces, utilizando este concepto, tenemos que para un modelo causal $\mathbf{C}$ una intervención $P_{\mathcal{C}}(Y|do(z)$ es realmente una \textit{querie} probabilista estándar, pero en el grafo mutilado dado por $\mathcal{C}_{Z=z}$; es decir
-\[ P_{\mathcal{C}}(Y|do(z),x) = P_{\mathcal{C}_{Z=z}}(Y|x).  \] 
-\subsection{Identificabilidad de modelos}
-Aun a pesar de desconocer la totalidad de las variables latentes causales, existen casos en los que es posible responder a preguntas causales en modelos que involucran variables latentes utilizando sólo variables observadas; este es el caso de los modelos identificables.
-Existen formas de simplificar queries intervencionales y convertirlas en queries que sólo requieren de datos observacionales pues existen equivalencias entre unas y otras bajo ciertas condiciones.\\
-\\
-\cite{peters2012identifiability} exploran las condiciones para obtener identificabilidad en modelos causales funcionales. 
-\subsection{Aprendizaje de modelos gráficos causales}
-Según \cite{koller2009probabilistic}, el problema de aprender un modelo causal a partir de datos puede dividirse en varios ejes:
-\begin{itemize}
-\item En primer lugar, definir qué se entiende por un modelo causal, pues si estamos trabajando con modelos gráficos causales entonces el problema se convierte en aprender redes bayesianas, pero con la diferencia de la interpretación que se da a las flechas; por otro lado, en modelos causales funcionales se tienen parametrizaciones mucho más complejas para las variables de respuesta, por lo que se necesitan modelos más complejos como el modelo neuronal presentado anteriormente.
-\item Un segundo eje es determinar si ya se conoce la estructura y sólo se necesita aprender la parametrización, o si también debe ser aprendida la parametrización.
-\item En tercer lugar, es el tipo de datos: observados o intervencionales.
-Datos observados son aquellos que provienen de alguna distribución generadora y datos intervencionales es en aquellos en los que se intervino el modelo y se observó la respuesta de este.
-Los datos intervencionales son difíciles de obtener en la práctica, o incluso pudieran ser inmorales o ilegales.
-\item Finalmente, cómo tratar los factores de confusión (confounding factors). Uno puede suponer que simplemente no existen y en estos casos el problema se vuelve relativamente sencillo.
-\end{itemize}
-\subsubsection{Aprender sin factores de confusión}
-El problema de aprender modelos gráficos causales sin factores de confusión es esencialmente el mismo que aprender redes bayesianas (\cite{koller2009probabilistic}), pero hay que hacer algunas suposiciones importantes, pues existen siempre varios DAG's que representen la misma distribución, pero que tendrían distintas interpretaciones causales.
-Una de las suposiciones necesarias para poder aprender modelos causales a través de datos observacionales es la llamada Propiedad Causal de Markov, que es un análogo a la Propiedad de Markov para redes Bayesianas; y su formulación es la siguiente (\cite{koller2009probabilistic}): \textit{cada variable es condicionalmente independiente de sus no-efectos dadas sus causas directas.} El otro supuesto necesario es que las únicas independencias condicionales sean aquellas que surgen de la d-separación en el grafo causal correspondiente (\textit{faithfulness}).\\
-\\
-Aun con estas suposiciones, lo más que puede aprenderse es una clase de equivalencias para el grafo causal y para llevar esto a cabo existen dos familias de métodos: basados en restricciones y basados en score y son los mismos que se utilizan para aprender el CPDAG en modelos causales funcionales.
-\subsubsection{Aprender de datos intervencionales}
-En el caso en que se tienen datos provenientes de intervenciones, lo cual es lo más común en el contexto de experimentos científicos.
-Se utilizan métodos basados en score para aprender de este tipo de datos.
-\subsubsection{Aprender con variables latentes}
-En el caso en el que suponemos que existen variables latentes, estamos pensando que los datos fueron generados por una distribución que es \textit{parecida}, o parcial, a la verdadera.
-De nuevo, existen familias de métodos basadas en restricciones y métodos basados en scores.
-\subsection{Redes Bayesianas causales: aplicación en genética}
-\cite{su2013using} intentan aprender la estructura de las relaciones entre un conjunto de variables relacionadas a un estudio de cáncer (\cite{karagas1998design},\cite{karagas2004incidence}).
-Específicamente, los autores analizan sólo 11 variables relacionadas a un solo gen.
-Identifican relaciones interesantes, como por ejemplo la variable “fumador” como padre causal de la variable “cancer”, aunque no apareció relación entre la variable cancer y un gen que se sabe está relacionado con el cáncer, pero esto se debe a la manera en la que fueron seleccionados los pacientes.
-\section{Modelos causales funcionales y ecuaciones estructurales}
-En otras áreas, como las ingenierías y la física, las relaciones causales se expresan como relaciones funcionales determinísticas en donde cualquier componente aleatorio se introduce en términos de variables no observadas.
-Esto corresponde (\cite{pearl2009causality}) a la noción Laplaciana de lo aleatorio.
-En este contexto, un modelo causal funcional consiste en un conjunto de ecuaciones de la forma
-\[ X_i = f_i (X_{Pa_i},U_i), \textrm{  } i=1,...n\]
-donde $Pa_i$ corresponde al conjunto de variables que directamente determinan a la variable con índice $i$ y donde $U_i$ representa cualquier error de medición o de aspectos no observados.\\
-\\
-Dado un modelo causal de esta forma, podemos obtener un diagrama de la siguiente manera: para cada variable pintamos un nodo, y desde cada elemento de $Pa_i$ dibujamos una flecha hacia el nodo que corresponde a $X_i$.
-A este diagrama se le llama \textit{diagrama causal}.
-Si cumple que el diagrama es acíclico, se le conoce como semi-Markoviano.
-Si, además de ser acíclico, los términos de error son independientes entonces se dice que es Markoviano.
-El siguiente Teorema muestra la conexión entre el diagrama causal y la distribución de probabilidad asociada.
-\begin{teo}
-Todo modelo causal Markoviano induce una distribución $P(X_1,...,X_n)$ que satisface que cada variable $X_i$ es independiente de todos sus no-descendientes dado sus padres $Pa_i$.
-\end{teo}
-Sabemos que dos DAG's que tienen el mismo esqueleto y las mismas v-estructuras son equivalentes, y a la clase de equivalencia le llamaremos CPDAG (\cite{goudet2017learning})
-\subsection{Intervenciones y efectos causales en modelos funcionales}
-Los modelos funcionales causales proveen una manera de estimar cómo cambiaría la distribución en respuesta a cambios externos.
-Lo que se hace en estos modelos es representar las intervenciones como alteraciones a un conjunto de funciones.
-El efecto total de la intervención se calcularía modificando las ecuaciones correspondientes y con este nuevo modelo calcular las nuevas probabilidades.
-Denotamos una intervención como $do(X=x)$.
-Se dice que una variable $X_i$ es causa directa de $X_j$ si
-\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
-\subsection{Contrafactuales en modelos funcionales}
-
-\section{Teoría de la Decisión}
-\subsection{von Neumann-Morgenstern}
-Según narra \cite{gilboa2009decision}, fue Daniel Bernoulli quién habló por primera vez de la maximización de la utildiad esperada; esto, al estudiar la \textit{paradoja de San Petersburgo}, la cual consiste en lo siguiente: un tomador de decisiones se enfrente a una serie de lanzamientos consecutivos de una moneda justa.
-Se considera un juego basado en los resultados de estas monedas, el cual terminará la primera vez que la moneda arroje una cara, y si esto ocurriera en el $n$-ésimo lanzamiento, el concursante recibirá $2^n$ pesos.
-La \textit{paradoja} consiste en que la mayoría de las personas, según registró Bernoulli, sólo está dispuesta a pagar una cantidad finita de dinero por entrar a este juego, cuyo valor esperado es infinito.
-La paradoja fue resuelta cuando Bernoulli se dio cuenta que las personas estaban maximizando el valor esperado de una \textit{función de utilidad}, pero no directamente el valor esperado del dinero.
-Resulta que si esta función es logarítmica, entonces la utilidad esperada es finita.
-\\
-\\
-El Teorema original de von Neumann y Morgenstern está formulado en términos de árboles de decisión y loterías compuestas.
-Más tarde \cite{jensen1967introduction} lo formuló de manera más compacta al suponer que las loterías compuestas son simplificadas de acuerdo a la fórmula de Bayes.\\
-\\
-Sea $X$ un conjunto de \textit{alternativas}, los objetos a elegir son loterías de soprte finito; de manera formal, definimos
-\[ L = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
-Una mezcla de loterías se puede definir de la siguiente manera: se toma una probabilidad $\alpha$ y se define una lotería $l^\alpha = \alpha P + (1- \alpha) Q$.\\
-\\
-Consideremos una relación  $\succeq L \times L$. El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas:
-\begin{itemize}
-\item A1. Órden débil: $\succeq$ es completa y transitiva
-\item A2. Continuidad: Para toda $P,Q,R \in L$ tales que $P \succ Q \succ R$ existen $\alpha$ y $\beta$ tales que:
-\[  \alpha P + (1-\alpha) R \succ Q \succ \beta P + (1-\beta) R\]
-\item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si $\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R$.
-\end{itemize}
-\begin{teo}
-Una relación $\succeq L \times L$ satisface los axiomas A1 - A3 si y sólo sí existe una función $u: X \to \mathbb{R}$ tal que para toda $P,Q \in L$ se tiene que $P \succeq Q$ si y sólo si 
-\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
-\end{teo}
-Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
-\subsection{De Finnetti}
-¿Cuál es la definición de probabilidad subjetiva?
-\cite{gilboa2009decision} sugiere que para saber qué son debemos explorar cómo se usan estas, y una posible manera de uso es al maximizar una utilidad esperara y esto es lo que hace el Teorema de DeFinnetti.\\
-\\
-De Finetti consideró apuestas sobre \textit{estados del mundo}.
-Consideremos que existen $1,...,n$ estados en el mundo.
-Una apuesta es una función que toma valores en el espacio de estados y los mapea a valores monetarios; es decir, una apuesta $x$ es un vector de $n$ elementos en donde $x_i$ es la ganancia en el caso de que la apuesta resulte en el estado $i$ como el ganador.
-Consideramos una relación $\succeq \subseteq X \times X$.
-Y consideremos los siguientes axiomas:
-\begin{itemize}
-\item A1. Órden débil: $\succeq$ es completa y transitiva.
-\item A2. Continuidad: Para cada $x \in X$, los conjuntos $\{ y | x \succ y \}$ y $\{ y | y \succ x \}$ son abiertos.
-\item A3. Aditividad: Para toda $x,y,z \in X$, $x \succeq y$ si y sólo si $x+z \succeq y+z$.
-\item A4 Monotonicidad: Para toda $x,y \in X$, si se cumple que $x_i \geq y_i$ entonces se tiene que $x \succeq y$.
-\item A5 No-trivialidad: Existen $x,y \in X$ tales que $x \succ y$.
-\end{itemize}
-Notemos que el axioma A3 sólo aplica bajo un esquema de neutralidad al riesgo.
-\begin{teo}
-La relación $\succeq \subseteq X \times X$ satisface A1-A5 si y sólo si existe un vector de probabilidad $p$ de tamaño $n$ tal que para toda $x,y \in X$ se cumple que
-\[ x \succeq y \textrm { si y sólo si } px \geq py. \]
-\end{teo}
-Entonces, la definición de probabilidad subjetiva sería aquella probabilidad utilizada por un tomador de decisiones para escoger una acción mediante el criterio de utilidad esperado.
-\subsection{Savage}
-El Teorema de von Neumann-Morgenstern indica cómo obtener una función de utilidad si las probabilidades están bien definidas; el Teorema de De Finnetti dice cómo obtener probabilidades si conocemos las utiidades de un tomador de decisiones.
-El Teorema de Savage obtiene al mismo tiempo probabilidades y utlidades.
-Su modelo asume sólo estados del mundo y sus resultados.
-Por lo tanto, los axiomas propuestos por Savage son bastante más abstractos y menos interpretables, pero se obtiene la ventaja de que sus axiomas son equivalentes a la existencia de una función de utilidad y de una medida de probabilidad (subjetiva) tales que las decisiones se llevan a cabo al maximizar utilidad esperada respecto a esta función de utilidad y esta medida de probabilidad.\\
-\subsubsection{Estados, resultados y acciones}
-\section{Aprendizaje por Refuerzo}
-El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de interactuar con su entorno.
-Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
-El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente.
-A su vez, el ambiente devuelve información al agente en términos de una recompensa o ganancia según el estado en el que se encuentre.
-El objetivo del agente es encontrar una serie de acciones que lo lleven a lograr su objetivo final.
-La función de recompensa codifica qué es lo que se desea que el agente aprenda, pues esta función devuelve un valor para cada estado y con esto el agente asignará un valor a cada estado según la ganancia a largo plazo que un este representa.\\
-\\
-La dinámica consiste en lo siguiente: en un tiempo $t$, el agente se encuentra en el estado $s_t$, toma una acción $a_t \in \mathcal{A}(s_t)$ de acuerdo a una \textit{policy}, o política, y llega a un nuevo estado $s_{t+1}$ según una probabilidad de transición $\mathcal{T}$ y recibe una recompensa $r_{t+1}$.
-El objetivo del agente es encontrar la política que maximice la recompensa.\\
-\\
-La regla general que se sigue para la definición del ambiente (\cite{sutton1998reinforcement}) es considerar que todo aquello que no pueda ser explícitamente alterado por el agente pertenece al ambiente.
-La frontera agente-ambiente tiene que ver con los límites de lo que el agente puede controlar, no sobre lo que puede \textit{conocer}.
-Veremos ahora con detalle las partes que conforman esta dinámica. 
-\subsubsection{Objetivos y Recompensas}
-Debido a que el objetivo del agente es maximizar la recompensa en el largo plazo, la manera en la que logramos que aprenda una tarea en particular es diseñando una función recompensa adecuada.
-Por ejemplo, un agente que queremos que aprenda a jugar ajedrez debe ser premiado por ganar los juegos, no por lograr sub-objetivos como comer más piezas del oponente o tomar control del centro del tablero (\cite{sutton1998reinforcement}), pues podría suceder que el agente encuentre cómo lograr esto sin ganar el juego necesariamente.
-Es mediante la recompensa que le comunicamos al agente qué queremos de él, no cómo queremos que lo haga.
-\subsubsection{Definición de Cadenas de Markov}
-Una Cadena de Markov discreta es un proceso estocástico en el cual el estado que toma el proceso sólo depende del estado en el tiempo anterior. Esto es, 
-\[P(X_{t+1} \in A | X_0=x_0, X_1=x_1 , ... , X_t=x_t) = P(X_{t+1} \in A | X_t = x_t).\]
-Estos procesos quedan caracterizados por sus probabilidades de transición $p^t_{ij}=P(X_{t+1}=j | X_{t}=i)$.
-Cuando estas probabilidades son iguales para toda $t$, entonces se dice que la cadena es homogénea, y por esta razón podemos almacenar las probabilidades de transición en una matriz $P$ tal que $P_{ij}=p_{ij}$.
-\subsubsection{Procesos de Decisión Markovianos con recompensa}
-Consideremos una Cadena de Markov con una ligera modificación: al tiempo $t$, se toma una \textit{acción} $a_t$ que afecta el estado del proceso al tiempo $t+1$.
-En respuesta a la acción tomada, se recibe una recompensa $r_t = f(a_t)$ para alguna función $f$ desconocida, la cual puede ser aleatoria o determinista.
-Los procesos de decision Markovianos con recompensa (\cite{Puterman:1994:MDP:528623}) están definidos por un espacio de estados $\mathcal{S}$, un espacio de acciones posibles $\mathcal{A}$, una función de transición $f$ y la función recompensa $r$.
-La función transición puede ser estocástica o determinística.
-
-\subsubsection{Función valor}
-Mientras la función recompensa provee la recompensa inmediata de estar en un estado, la función valor asigna una valoración \textit{a largo plazo} de un estado.
-Como las recompensas obtenidas a largo plazo dependen de las acciones que se sigan tomando, tenemos que definir la función valor en términos de una política.
-Una política es un mapeo $\pi : \mathcal{S} \times \mathcal{A} \to \pi(s,a)$.
-Es decir, la probabilidad de tomar la acción $a$ estando en el estado $s$.
-La mayoría de los algoritmos de aprendizaje por refuerzo intentan estimar de una u otra forma la función valor, por lo que la definición de esta es muy importante.
-Formalmente, para un proceso de decisión Markoviano, definimos la \textit{función valor} (value function en la literatura) para un estado $s$, y relativa a una política $\pi$, como el retorno esperado si se comienza en $s$ y a partir de ahí se sigue la política $\pi$
-\[ V^{\pi}(s) = \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]. \]
-De manera similar, definimos la \textit{función acción-valor} $Q^{\pi}(s,a)$ es una valoración de tomar la acción $a$ estandoen el estado $s$, y después seguir la \textit{política} $\pi$
-\[ Q^\pi(s,a) = \mathbb{E}[R_t | s_t =s, a_t = a]. \]
-Por la ley de los grandes números, podemos estimar ambas funciones $V^{\pi}$ y $Q^\pi$ si se promedian los retornos recibidos al llegar a cada estado con el número de veces que se ha llegado al estado.
-Esta es la idea detrás de los llamados Métodos Monte-Carlo que serán expuestos en una sección posterior.\\
-Notemos una propiedad interesante de la función $V^{\pi}$:
-\begin{eqnarray*}
-V^{\pi}(s) &=& \mathbb{E}_{\pi}\left[ R_t | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ \sum_{k=0}^\infty \gamma^k r_{t+k+1} | s_t = s \right]\\
-               &=& \mathbb{E}_{\pi} \left[ r_{t+1} + \gamma \sum_{k=0}^\infty \gamma^k r_{t+k+2} | s_t = s \right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} P(s_{t+1}=s' | s_t = s, a_t = a) \left[ \mathbb{E}[r_{t+1}|s_t=s,a_t=a,s_{t+1}=s'] + \gamma \mathbb{E} \left[ \sum \gamma^k r_{t+k+2} | s_{t+1}=s' \right]\right]\\
-               &=& \sum_a \pi(s,a) \sum_{s'} \mathcal{P}_{ss'}^a \left[ \mathcal{R}_{ss'}^a + \gamma V^{\pi}(s') \right]
-\end{eqnarray*}
-A esta última ecuación se le conoce como la Ecuación de Bellman para $V^\pi$.\\
-\\
-También para la función $Q^{\pi}(s,a)$ se tiene una ecuación de Bellman:
-\[ Q^{\pi}(s,a)=\mathbb{E}_{s',a'}[r+ \gamma Q^{\pi}(s',a') | s',a']. \]
-Además, la función $V^\pi$ puede ser obtenida a partir de la función $Q^{\pi}(s,a)$ de la siguiente manera:
-\[ V^{\pi}(s)=Q^\pi(s,\pi(s)) \]
-\subsubsection{Soluciones y Optimalidad}
-Resolver un problema de aprendizaje por refuerzo consiste en (\cite{sutton1988learning}) encontrar una política que, a la larga, reciba una alta recompensa.
-Resulta que es posible definir un orden parcial entre políticas en términos de la función valor, pues $\pi \geq \pi'$ si $V^{\pi}(s) \geq V^{\pi'}(s)$ para todo estado $s$.
-Para procesos de decisión Markovianos finitos existe siempre una política mejor que las demas, a la cual llamaremos \textit{política óptima}.
-De hecho, pueden existir varias, pero todas ellas comparten la misma función valor, denotada como $V^\ast$.
-Si para una política $\pi^\ast$ se tiene que es mejor que todas las demás, entonces podemos describir su función valor como
-\[  V^{\pi^\ast} = \max_{\pi} V^{\pi}(s).\]
-Por supuesto, $V^\ast = V^{\pi^\ast}$
-\\
-Análogamente, las políticas óptimas tienen la misma función $Q$ óptima, por lo que definimos \textit{función accion-valor óptima} como el retorno esperado de tomar la acción $a$ en el estado $s$ y a partir de ahí seguir una política óptima:
-\[ Q^\ast (s,a) = \max_\pi Q^{\pi}(s,a) = Q^{\pi^\ast}(s,a). \]
-Los métodos modernos de aprendizaje por refuerzo intentan aprender directamente la función valor o la función $Q$, y una vez Q y una vez que tenemos $Q^{\pi^\ast}$ podemos definir una política óptima de la siguiente manera:
-\[ \pi^\ast = \text{argmax}_a Q^\ast (s,a). \]
-Además, utilizando la ecuación de Bellman vimos que se podía escribir $Q$ en términos de $V$, y por lo tanto podemos escribir $Q^\ast$ en términos de $V^\ast$ como
-\[ Q^\ast (s,a) = \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La ecuación de Bellman puede escribirse sin hacer referencia a ninguna política de la siguiente manera:
-\[ V^\ast (s) = \max_a \mathbb{E}[r_{t+1} + \gamma V^\ast (s_{t+1}) | s_t = s, a_t = a]. \]
-La función óptima $ Q^\ast$ lo que hace es maximizar sobre todas las decisiones:
-\begin{eqnarray*}
-Q^\ast &=& r_{t+1} + \gamma \max_{a_{t+1}} r_{t+2} + \gamma^2 \max_{a_{t+2}} r_{t+3}...\\
-&=& r_{t+1} + \gamma \max_{a_{t+1}} Q^{\ast} (s_{t+1},a_{t+1}).
-\end{eqnarray*}
-Formalmente:
-\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
-\subsection{Métodos básicos de solución: Model-based}
-\subsubsection{Programación Dinámica}
-Los métodos de programación dinámica tienen muy buen fundamento teórico, pero requieren una especificación completa del ambiente.
-La idea fundamental en programación dinámica es poder asignar a cada estado del sistema un valor numérico que represente el retorno esperado, a largo plazo, de ese estado; es decir, estamos valorando cada estado según lo que esperemos obtener de él a muy largo plazo.
-\\
-\\
-Recordemos que una política $\pi$ especifica una manera de actuar en cada estado; es decir, a cada estado del sistema le asocia una acción.
-Si ya conocemos una política cualquiera, la pregunta natural es ¿cómo obtener la función $V^\pi$?\\
-\\
-Sabemos que la función $V^\pi$ debe satisfacer la ecuación de Bellman:
-\[ V^\pi (s) = \mathbb{E}[r + \gamma V^\pi (s') | s_t = s]. \]
-Pero esto nos deja con el mismo problema, pues tenemos de ambos lados una función desconocida.
-Para problemas de decisión finitos, este es un sistema de ecuaciones con tantas ecuaciones como estados, lo cual puede ser costoso de resolver.
-Consideramos entonces un método iterativo:
- \[ V_{k+1} (s) = \pi(s) \sum_{s'} \mathcal{P}_{s,s'}^a [ \mathcal{R}_{s,s'}^a + \gamma V_k(s')], \]
-donde, $\mathcal{P}_{s,s'}^a$  es la probabilidad de pasar al estado $s'$ desde $s$ tomando la acción $a$.
-Y $\mathcal{R}_{s,s'}^a$ es el valor esperado de la siguiente recompensa al pasar de $s$ a $s'$ tomando la acción $a$.\\
-\\
-Sabemos que además de la función $V$ tenemos a la función $Q$, la cual valora parejas estado-acción.
-Utilizando esta función, podemos preguntarnos qué pasaría si en vez de escoger la acción $a=\pi(s)$ escogieramos cualquier otra acción $a$.
-Claramente, onsecuencia de hacer esto estaría dada por $Q^{\pi}(s,a)$.
-Entonces, si esto resulta ser mayor que la función valor en ese estado, entonces estaríamos encontrando una mejor política, una que determina $\pi'(s)=a$ en vez de $\pi(s)$.
-Este es un caso particular de un resultado conocido como \textit{Teorema de Mejora de Polítca} que dice que si $\pi$ y $\pi'$ son dos políticas tales que $Q^{\pi} (s, \pi'(s)) \geq V^{\pi}(s)$ entonces $\pi'$ es mejor que $\pi$; esto es, $V^{\pi'} (s) \geq V^{\pi}(s)$ para todo estado $s$.\\
-\\
-Entonces, resulta natural querer buscar estas mejoras para dos los estados posibles y todas las acciones posibles y obtener aquellos que maximicen la función $Q$; es decir, queremos construir una nueva política $\pi'$ como:
-\[ \pi'(s)= \textrm{ arg} \max_a Q^\pi (s,a). \]
-Para proceso de decisión finitos, siempre encontramos una mejor política, a no ser que ya estemos en el óptimo.\\
-\\
-Una vez que una policy $\pi' $ ha sido obtenida de mejorar $\pi$ utilizando $V^\pi$, podemos construir $V^{\pi'}$ y repetir nuevamente el proceso:
-\[ \pi_0 \xrightarrow{E} V^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} V^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} V^\ast  \]
-pues cada política está garantizado que es mejor que la anterior.
-El método anterior puede ser muy tardado pues la convergencia, aunque garantizada teóricamente, sólo se da en el límite.
-Una opción es truncar la etapa de policy evaluación y hacer
-\[ V_{k+1}(s) = \max_a \mathbb{E}[r + \gamma V_k(s')|s_t=s,a_t=a]. \]
-Los métodos de programación dinámica suponen que se cuenta con un modelo completo del sistema, lo cual es prácticamente imposible en la práctica, además de ser muy costosos en términos computacionales.
-Estos métodos sufren de la llamada \textit{maldición de la dimensionalidad}, pues el número de estados crece exponencialmente con el número de variables
-\subsection{Métodos básicos de solución: Model-free}
-Métodos que no necesitan un modelo completo del sistema, sólo necesitan experiencia.
-Encontramos los métodos Monte Carlo que simulan las transiciones y los estados, y los métodos de diferencias temporales, que van actualizando iterativamente el valor de los estados observados, el más famoso de estos es el TD($\lambda$).
-\subsubsection{Métodos Monte-Carlo} 
-Los métodos Monte-Carlo no requieren un modelo del ambiente y son conceptualmente sencillos, pero no son adecuados para hacer actualizaciones incrementales.
-Estos métodos requieren \textit{experiencia} en vez de un modelo completo; es decir, requieren secuencias de estados, acciones y recompensas, ya sean estos obtenidos de simulaciones o de interacción \textit{on-line} con el ambiente.
-Los métodos Monte-Carlo suponen que la experiencia proviene de eventos episódicos, es decir que terminan siempre de una u otra forma y es al final de estos episodios que se alteran las estimaciones de valor así como las políticas.
-En este sentido, los métodos Monte-Carlo son incrementales por episodio, pero no paso a paso dentro de cada episodio.
-A pesar de las diferencias operativas entre los métodos DP y los métodos monte-carlo, las ideas esenciales son las mismas, pues ambos métodos calculan las mismas funciones además de seguir el mismo camino hacia la optimalidad; es decir, evaluación, mejora y finalmente iteración general.
-La siguiente descripción de los métodos Monte-Carlo está basada en el clásico libro de \cite{sutton1998reinforcement}
-\begin{itemize}
-\item Evaluación Monte-Carlo: Como se mencionó antes, una manera de estimar el valor de un estado sería promediar las recompensas obtenidas al visitar un estado entre el número de visitas a este; por la ley de los grandes números este promedio debe converger al valor real.
-Dado un conjunto de episodios obtenidos de seguir una política $\pi$, denotamos como \textit{visita} cada vez que se llega a un estado $s$.
-El algoritmo \textit{every-visit} estima $V^\pi (s)$ como las recompensas promedio de todas las visitas al estado $s$ en todos los episodios; por otro lado, el algoritmo \textit{first-visit} promedia sólo lo obtenido al visitar por primera vez un estado en todos los episodios.
-Ambos métodos convergen a $V^\pi$.
-\item Notemos que al no tener un modelo del ambiente, es más útil estimar el valor de una acción más que el valor de un estado.
-Teniendo un modelo, los valores de los estados son suficientes para determinar una política, pues uno simplemente ve un paso hacia adelante y escoge la acción que tenga la mejor combinación de recompensa y de estado siguiente, pero sin un modelo disponible, uno debe estimar el valor de las acciones para con esto obtener una política.
-Entonces, es importante tener un método para estimar $Q^\ast$.
-\item Estimación de $Q^\pi(s,a)$: Los métodos son esencialmente los mismos que para estimar la función valor; es decir, promediando las recompensas obtenidas al visitar un estado y haber seleccionado cierta acción (todas las visitas) y análogamente para la primer visita.
-El problema que tienen estos métodos es que no todas las parejas acción-estado son visitadas, pues para obtener estimaciones válidas deberíamos tener parejas estado-acción para todas las acciones y no sólo las que se ven en los episodios.
-\item Control Monte-Carlo: para una primera versión Monte-Carlo de iteración de políticas, supondremos que contamos con un número infinito de episodios y que se visitan todas las parejas estado-acción dentro de cada episodio.
-A diferencia de los métodos de DP y por las razones que ya se explicaron, el proceso de evaluación y mejora se hace respecto a la función $Q(s,a)$.
-\[ \pi_0 \xrightarrow{E} Q^{\pi_0} \xrightarrow{I} \pi_1 \xrightarrow{E} Q^{\pi_1} \xrightarrow{E} \pi_2 \xrightarrow{E} \cdots \xrightarrow{I} \pi^\ast \xrightarrow{E} Q^\ast.  \]
-Bajo el supuesto de episodios infinitos y visita a todas las parejas estado-acción se tiene que los métodos Monte-Carlo encuentran $Q^{\pi_k}$ de manera exacta para $\pi_k$.\\
-Para hacer la mejora de una política, se escoge de manera determinística la acción $a$ en la cual se alcanza el máximo de la función $Q$; es decir,
-\[ \pi(s) = \textrm{arg max}_a Q(s,a).\]
-\item Para remover el supuesto de un número infinito de episodios, una opción es abandonar la idea de obtener completamente la función $Q^\pi_k$ en el paso de evaluación antes de proceder a la mejora, sino sólo mover un poco la función $Q$ actual hacia la verdadera.
-\item En cuanto al supuesto de visita a todas las parejas acción-estado la única alternativa a suponer que se visiten todas las parejas es hacer que el agente las seleccione y para esto existen dos familias de métodos:
-		\begin{itemize}
-				\item On-Policy: intentan evaluar mejorar la política que es actualmente utilizada para la toma de decisiones.
-				\item Off-Policy: Evaluan o mejoran una política distinta
-		\end{itemize}
-\end{itemize}
-\subsubsection{Diferencias Temporales}
-Los métodos TD son totalmente incrementales y tampoco requieren un modelo del ambiente, pero son mucho más complejos.
-Estos métodos son una combinación de métodos Monte-Carlo y de Programación Dinámica; estos métodos pueden aprender directamente de la experiencia y no necesitan un modelo del ambiente, pero también utilizan técnicas de \textit{bootstrapping} (estiman usando estimadores) por lo que no necesitan finalizar completamente un episodio para actualizar los valores.
-\begin{itemize}
-\item Evaluación de política en Diferencias Temporales.
-\item TD(0).
-\item SARSA: Es un método \textit{on-policy}.
-\item Q-Learning: Es un método \textit{off-Policy}.
-\item Métodos Actor-Critic.
-\end{itemize}
-\subsection{Value based Reinforcement Learning: Deep Q-Networks}
-Los sorprendentes avances de Deep Learning en visión computacional llevaron de manera natural a plantear la pregunta de cómo utilizar estas nuevas técnicas para el aprendizaje por refuerzo. 
-Las aplicaciones estándar de DL requerían grandes volúmenes de datos etiquetados, además de tener muestras independientes, pero en Reinforcement learning esto es un problema pues las secuencias de estados suelen estar altamente correlacionados.\\
-\cite{mnih2013playing} introducen “Deep Q-Network”, que consiste en una implementación de Q-learning con aproximación de funciones (\cite{sutton1998reinforcement}) en donde el aproximador es una red neuronal convolucional.
-Utilizando esta técnica fue como \cite{mnih2013playing} lograron que un agente aprendiera siete juegos de Atari 2600 utilizando como input únicamente la pantalla del juego.
-\\
-\\
-Estimar la función $Q$ de manera iterativa es muy costoso, pues además de requerir un modelo completo del sistema cada secuencia de estados se utiliza por separado para las estimaciones.
-Por esto, es mejor utilizar alguna función aproximadora para poder generalizar. En este caso, se pretende aproximar la función valor mediante una red neuronal con pesos $\theta$:
-	   \[Q(s,a,\theta) \approx Q^\ast(s,a). \]
-Esta red se entrena al minimizar una sucesión de funciones $L_i$ que está dada por:
-	   \[ L_i (\theta_i) = \mathbb{E}[(y_i - Q(s,a;\theta)^2], \textrm{ donde  }y_i = \mathbb{E}[r + \gamma \max_{a'} Q(s',a';\theta) | s,a]. \]
-Notemos que los objetivos $y_i$ dependen de los pesos de la red neuronal. El gradiente de $L_i (\theta_i)$ está dado por:
-	   \[ \nabla_{\theta_i} L_i (\theta_i) = \mathbb{E}[(r + \gamma \max_{a'} Q(s',a';\theta_{i-1}) - Q(s,a;\theta_i)) \nabla_{\theta_i} Q(s,a;\theta_i).] \]
-Para optimizar la función $L$ se utiliza gradiente estocástico; este es un método \textit{model-free} y \textit{off-policy}.\\
-\subsubsection{Experience Replay}
-Notemos que las secuencias consecutivas de estados y acciones están fuertemente correlacionadas, por lo que el aprendizaje podría ser ineficiente.
-Además, al aprender \textit{on-policy}, los parámetros actuales determinan los siguientes estados sobre los cuales se entrenará, por lo que pueden surgir loops que afecten la convergencia del aprendizaje.
-Para resolver estos problemas se utiliza una técnica conocida como Experience Replay (\cite{Lin1992}); esta técnica consiste en guardar las experiencias del agente para cada tiempo $t$; esto es 
-	   \[ e_t = (s_t, a_t, r_t, s_{t+1}). \]
-Cada $e_t$ se almacena en un conjunto $\mathcal{D} = e_1 , ...
-, e_N$ y en un loop interno del algoritmo, se tomarán aleatoriamente muestras de experiencia las cuales serán utilizadas para la actualización de pesos.
-\subsubsection{Algoritmo completo}
-Deep Q-learning con Experience Replay:
-\subsubsection{Aplicaciones de DQN}
-		\begin{itemize}
-			\item \cite{mnih2013playing}.
-Paper en el que muestran cómo una computadora aprende a jugar Atari 2600 a través de sólo observar los pixeles del juego y obteniendo una 																		recompensa conforme el score aumentaba.
-Utilizando la misma arquitectura, se pudieron aprender 7 juegos
-			\item \cite{mnih2015human}.
-En este paper, aumentaron a 49 juegos aprendidos.
-Utilizan una Deep Q-Network, que consiste en aproximar la función Q mediante una red 																			neuronal que toma como input los estados (las pantallas del juego) y devuelve como output el valor Q para cada posible acción.
-En este caso, 																			utilizaron una red convolucional con 3 capas de convolución y 2 capas totalmente conectadas.
-	     \end{itemize}
-\subsection{Mejoras posteriores a DQN}
-\subsubsection{Double DQN}
-\cite{hasselt2010double}, \cite{van2016deep} propuso una “Double Deep Q-Network”. Recordemos que en DeepQ-learning las actualizaciones que se hacen son:
-\[ \theta_{t+1} = \theta_t + \alpha (r_{t+1} + \gamma \max_{a} Q(s_{t+1},a;\theta_t)) \nabla_\theta Q(s_t,a_t;\theta_t). \]
-Por lo tanto, es más probable escoger acciones que tienen un alto valor, lo que resulta en valores demasiado optimistas.
-Lo que se propone en \cite{van2016deep} es evaluar la política greedy utilizando la red on-line, pero utilizar la red objetivo para estimar el valor.
-Debe hacerse el siguiente cambio en la actualización:
-\[ r_{t+1}  + \gamma  Q (s_{t+1}, \textrm_{ arg max}_a Q(s_{t+1},a_t,\theta_t); \theta^{-}\]
-donde $\theta$ es el parámetro de la red on-line y $\theta^{-}$ el parámetro de la red objetivo
-\subsubsection{Prioritized Replay}
-
-\subsubsection{Duelling Networks}
-\subsection{Método Asíncronos}  
-\subsection{Policy Based DRL}
-\subsubsection{Actor-Critic}
-\subsubsection{Policy Gradients}
-Resulta que Deep Q-Learning no es un gran algoritmo, incluso los autores del paper original (\cite{DBLP:journals/corr/MnihBMGLHSK16}) mencionan sus desventajas. 
-Otra idea es atacar directamente la política.
-La idea es representar una policy mediante una red neuronal profunda con pesos $\textbf{u}$.
-\[ a = \pi (a|s,u) \textrm{ o } a = \pi (s,u). \]
-En este contexto, definimos una función objetivo $L(\textbf{u})$ como:
-		\[ L(\textbf{u}) = \mathbb{E}[r_1 + \gamma r_2 + \gamma^2 r_3 + ... |  \pi(\cdot,\textbf{u})]. \]
-Optimizamos con Stochastic Gradient Descent. El gradiente de una policy estocástica está dado $\pi (a|s,\textbf{u})$ por:
-		\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[  \frac{\partial \log \pi(a|s,\textbf{u})}{\partial \textbf{u}}Q^\pi(s,a)\right].\]
-Por otro lado, el gradiente de una policy determinística $a=\pi(s)$ está dado por:
-\[ \frac{\partial L(\textbf{u})}{\partial \textbf{u}}  = \mathbb{E}\left[ \frac{\partial Q^\pi (s,a)}{\partial a} \frac{\partial a}{\partial \textbf{u}} \right]. \]
- Los métodos \textit{actor-only}, donde \textit{actor} se utiliza como sinónimo de \textit{policy} utilizan familias parametrizadas de policies.
-Las ventajas de tener una policy parametrizada es que podemos generar un espectro continuo de accciones, pero los métodos de optimización (policy gradients) sufren de alta varianza en sus estimaciones del gradiente, lo cual genera un lento aprendizaje (\cite{grondman2012survey},\cite{konda2003onactor},\cite{richter2007natural},\cite{boyan2002technical},\cite{baxter2001infinite},\cite{berenji2003convergent}).\\
- \\
-Los métodos \textit{critic-only}, donde \textit{critic} se utiliza como sinónimo de \textit{value function}, que utilizan diferencias temporales tienen una menor varianza en las estimaciones del retorno esperado (\cite{boyan2002technical},\cite{berenji2003convergent},\cite{sutton1988learning}).
-Una manera de derivar una policy en estos escenarios es seleccionando greedy actions; es decir, acciones praa las cuales la value funtion indique que el retorno esperado es máximo, lo cual puede requerir un alto costo computacional.
-Lo usual es discretizar el espacio continuo de acciones.
-Los métodos \textit{actor-critic}, combinan las vetajas de ambos mundos.
-Mientras que las familias parametrizadas tienen la ventaja de utilizar acciones continuas, los métodos critic-only proveen de conocimiento de baja varianza sobre el desempeño; es decir, la estimación del retorno esperado del crític-only permite al actor-only actualizar los gradientes con información de menor varianza, lo cual acelera todo el proceso.
-La menor varianza es intercambiada por un mayor sesgo al inicio del aprendizaje (\cite{berenji2003convergent}).
-\\
-\\
-Los métodos de Policy Gradients optimizan directamente la recompensa acumulada y pueden utilizarse junto con aproximadores no-lineales como las redes neuronales.
-Este método presenta dos grandes retos: el número de muestras de entrenamiento requeridos y la dificultad de obtener mejoras estables debido a la no-estacionariedad de los datos.
-\cite{DBLP:journals/corr/SchulmanMLJA15} atacan el primer problema se puede atacar utilizando value functions para reducir la varianza de los estimados.
-El segundo problema se ataca utilizando un método de región de confianza tanto para la política como para la value function, los cuales están representados por redes neuronales.\\
-\subsubsection{Trust Region Policy Optimization}
- \cite{DBLP:journals/corr/SchulmanLMJA15} Proponen “Trust Region Policy Optimization”, Procedimiento iterativo para optimizar políticas con mejora monótona garantizada.
-Es un algoritmo similar a Policy Gradients y resulta útil para optimizar políticas con redes neuronales.
-Buen performance en juegos de Atari que utilizan la pantalla como input.
-\section{Reinforcement Learning y Causalidad}
-Aprendemos causalidad mediante intervenciones, interacciones, etc. Reinforcement Learning es aprender a realizar tareas mediante interacción. ¿Habrá algo en medio?
-\subsection{Conocimiento causal y observabilidad parcial}
-Los algoritmos actuales de reinforcement Learning suponen que el problema puede modelarse dentro de un contexto Markoviano de decisión con recompensas.
-Si el supuesto markoviano se elimina, los algoritmos podrían dejar no funcionar, como muestran en \cite{singh1994learning}.
-Además, todo el desarrollo teórico utiliza la propiedad Markoviana.\\
-\\
-Una manera en la que la propiedad Markoviana puede perderse es si el agente no cuenta con información completa sobre el estado; es decir, al igual que en los HMM, suponemos que el ambiente es Markoviano, pero el agente no tiene acceso a él de modo que los datos no le \textit{parecen} Markovianos.
-A diferencia de los HMM, no existe un procedimiento computacionalmente tratable para los POMDP.
-El problema es que una vez que se obtienen las estimaciones del valor para cada estado, se debe hacer programación dinámica en espacios continuos y esto es infactible salvo para casos pequeños\\
-\\
-La manera de generalizar el caso markoviano al caso no observado es considerar un MDP subyacente con estados $\mathcal{S}=\{ s_1,...,s_N\}$ y probabilidades de transición $\mathcal{P}^a_{s,s'}$.
-Luego, en cada tiempo $t$, el agente observa un mensaje $m_t$ el cual proviene de una distribución $P(m | s_t)$ desconocida.\\
-\\
-En \cite{jaakkola1995reinforcement} se describe un procedimiento que evita estimar el valor de los estados y trabaja directamente en el espacio de políticas estocásticas.
-El procedimiento es \textit{directo}, a diferencia de los métodos \textit{indirectos} en los cuales primero se aprenden los parámetros del MDP y luego se utiliza programación dinámica para obtener la política.
-El algoritmo propuesto consiste en dos partes: primero, se utiliza un método Monte-Carlo para calcular un análogo de los Q-values y luego se utiliza un paso de policy-improvement.\\
-\\
-Desde un punto de vista de psicología, una manear de entender la inferencia de estados ocultos es en términos de causas latentes (\cite{gershman2015reinforcement}).
-Siguiendo esta idea, es evidente que un agente de RL que opera en el mundo real no sólo debe hacer inferencia de estados ocultos, sino que debe poder descubrir los estados ocultos que están detrás del mundo que éste observa.
-Esto es una forma de aprendizaje estructural latente (\cite{gershman2010learning}).\\
-\\
-Un primer acercamiento a este problema puede hacerse mediante técnicas de inferencia Bayesiana no paramétrica.
-Por ejemplo, tanto \cite{gershman2010learning} como \cite{gershman2012exploring} utilizan priors no paramétricas para hacer inferencia sobre causas latentes.
-Un modelo causal latente es una buena representación de la estructura causal verdadera bajo condicionamiento Pavloviano, pues es el experimentador quien es una causa latente.\\
-\\
-En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
-\subsection{Towards Deep Symbolic Reinforcement Learning}
-Deep Reinforcement learning consiste, básicamente, en utilizar una red profunda para aproximar la $Q$-function; por lo tanto, los métodos de DRL heredan los problemas del Deep Learning: se requiere un gran volumen de datos, son muy lentas en entrenarse y, según \cite{garnelo2016towards}, son esencialmente \textit{reactivos}, lo que significa que no utilizan procedimientos de alto nivel como planeación o razonamiento causal, además de que no son interpretables.\\
-\\
-\cite{garnelo2016towards} proponen una arquitectura híbrida que, combina redes neuronales con inteligencia artificial simbólica.
-El modelo propuesto consiste en combinar una red neuronal que como entrada directamente lo que se recibe del ambiente y transforma estos datos crudos en una representación simbólica basada en lógica de primer orden; esta representación simbólica es posteriormente utilizada para escoger la acción a tomar.\\
-\\
-Se utiliza un sencillo juego para probar la arquitectura y en este contexto la red neuronal debe generar representaciones simbólicas de secuencias de estados en el juego en las cuales el flujo de pixeles queda codificado en términos de objetos, tipos y sus interaccioes.
-Luego, con esto se aprenden las políticas.\\
-\\
-El primer paso que hacen es utilizar un autoencoder convolucional para generar, de manera no supervisada, un conjunto de símbolos que representarán los objetos en cada escena.
-Luego, para cada pixel, obtienen la feature con mayor activación y se toman los más altos.
-De esta manera se obtiene un pixel representativo para cada objeto a los cuales posteriormente se le asigna un tipo simbólico según el autoencoder.\\
-\\
-Luego, una vez que se tienen estos símbolos de bajo nivel para cada cuadro del juego, se necesita rastrearlos a lo largo del tiempo.
-Utilizan una verosimilitud que es función de proximidad espacial y el tipo de transiciones.
-Para considerar la información de la dinámica de los cuadros utilizan la diferencia entre cuadros y las posiciones relativas entre objetos.\\
-\\
-Para la parte de aprendizaje, entrenan una función $Q$ por separado para cada interacción entre objetos con actualización
-\[Q^{ij}(s_t^{ij},a_t) \leftarrow  Q^{ij}(s_t^{ij},a_t) + \alpha \left( r_{t+1} + \gamma (\max_{a} Q^{ij} (s_{t+1}^{ij},a) - Q^{ij}(s_t^{ij},a_t)) \right). \]
-\\
-Se prueba el framework sobre varias variantes de un sencillo videojuego. \\
-\\
-Llama la atención como estos autores utilizan maquinaria de lo más reciente (deep learning, autoencoders) para luego obtener una representación en términos de lógica de primer orden.
-Este approach ha quedado claro que ya quedó atrás por lo que debería buscarse algo que sustituya este enfoque lógico.
-En particular, pudiera ser esto un modelo gráfico causal el cual vaya aprendiéndose conforme el agente de RL aprende sobre su entorno.
-
-\subsection{Conocimiento causal y toma de decisiones secuenciales: acciones como intervenciones}
-En cuanto al aprendizaje \textit{on-line} de modelos causales, encontramos en los trabajos de \cite{wellen2012learning} y de \cite{kummerfeld2013tracking} acercamientos tangenciales, o un punto de inicio, al aprendizaje de modelos causales al vuelo.\\
-\\
-Según \cite{wellen2012learning} es posible dividir las teorías de cómo las personas usan información estadística para aprender modelos causales en enfoques racionales y heurísticos.
-El enfoque primero modela el aprendizaje causal como inferencia racional, lo que incluye la inferencia Bayesiana; estos modelos se mantienen en el nivel computacional y no intentan explicar el mecanismo cognitivo subyacente.\\
-\\
-Por otro lado, el enfoque heurístico propone que las personas reaccionan a distintos estímulos que los hacen moficiar sus hipótesis causales de una manera no necesariamente racional.
-El modelo de efecto simple de \cite{waldmann2008causal} supone que las personas se enfocan en evaluar una sola relación causal a la vez.
-Es un modelo que aprende la magnitud de la relación causal.
-Si una variable supera cierto umbral, entonces la persona acepta esa relación causal y la incorpora a su bagaje de relaciones.\\
-\\
-El modelo propuesto por \cite{wellen2012learning} sigue esta línea del efecto simple, con la diferencia de que este modelo sí intenta descubrir el proceso cognitivo subyacente, pues el aprender magnitudes causales es un proceso puramente computacional.
-El modelo de efecto simple puede verse como una implementación del modelo de \cite{danks2003dynamical}.
-Además, según narran \cite{wellen2012learning}, el modelo de efecto simple no explica cómo la persona utiliza las magnitudes causales para determinar si existe o no una relación causal.
-En el modelo propuesto, \cite{wellen2012learning}, incorporan un procedimiento de decisión para determinar si existen o no relaciones causales.\\
-\\
-El modelo LPL (\cite{wellen2012learning}) comienza con una estructura causal inicial, la cual consiste en un grafo acíclico dirigido el cual representa las creencias iniciales de un individuo.
-Posteriormente, el modelo altera esta estructura al añadir o remover una sola arista, de manera que el problema se reduce a evaluar relaciones causales entre dos variables.
-Este modelo genera estimaciones de la fuerza de la relación causal para cada posible relación causa-efecto que no sea eliminada a priori.\\
-\\
-Supongamos que un individuo cree a priori que $B$ causa $E$, y está tratando de averiguar si $C$ causa $E$.
-A diferencia del modelo de efecto simple, las estimaciones de la fuerza causal están generadas por un mecanismo similar a los filtros de partículas Bayesianos
-En particular, a partir del conocimiento causal inicial del individuo, el modelo LPL comienza con $n$ partículas.
-Estas partículas $\{V_C^1,...,V_C^n \}$ representan las creencias actuales del individuo sobre si $C$ causa $E$.
-Paralelamente, existen $\{ V_B^1,...V_B^n \}$ para $B$.
-Se define la \textit{capa} $i$ de partículas para el efecto $E$ como la $i$-ésima partícula de cada causa conocida y posible para $E$.
-En el ejemplo, tendríamos que esta capa es $\{ V_B^i, V_C^i \}$.
-Luego, el modelo LPL actualiza cada capa de manera independiente de la siguiente manera
-\[ V_C^{i+1} = V_C^{i,t} + \alpha ( \textrm{observado} - \textrm{esperado}), \]
-Donde “observado” tiene el valor de 1 si el efecto sucede y 0 en caso contrario.
-Por otro lado, “esperado” es el valor esperado del efecto, calculado según la forma funcional de la relación causa-efecto, el cual es calculado por separado para cada causa potencial en una capa.
-El valor esperado de $E$ (para la capa $i$ y la causa potencial $C$) es
-\[ \textrm{esperado} = \prod_{k} (1-V_k^{t-1})(1 - \prod_j (1- V_j^{t-1}).\]
-Conforme el individuo observa más datos, las partículas convergen a los valores verdaderos de los parámetros y las estimaciones para cada capa convergen.
-Este procedimiento no toma en cuenta aspectos estructurales, los cuales se modifican según un criterio de decisión.
-Lo que hace el modelo es una prueba $t$ en la cual la hipótesis nula es que las partículas provienen de una distribución con media cero.
-El criterio de decisión es el siguiente: si actualmente no existe una arista, y la prueba $t$ rechaza la hipótesis nula, entonces se añade la arista.
-Si existe una arista, y la prueba no es significativa, entonces se elimina la arista existente.
-Alterar una arista afectará las futuras estimaciones del valor “esperado” para otras causas.\\
-\\
-\cite{hagmayer2013hierarchical} llaman la atención sobre que los modelos actuales de Decisiones (maximum expected utility) no asumen que los tomadores de decisiones aprenden sobre la estructura causal del problema.
-Por otro lado, las teorías de aprendizaje causal proponen que un tomador de decisiones guían sus acciones a través de conocimiento causal previamente obtenido.
-El modelo de \cite{sloman2006causal} extiende las teorías de aprendizaje causal al contexto de toma de decisiones, la idea básica consiste en que los tomadores de decisiones utilizan la información disponible de su entorno para inducir un modelo causal del problema de decisión al que se enfrentan.
-Se sabe que en los seres humanos el conocimiento causal guía la interpretación de información probabilista y que los humanos somos capaces de aprender la estructura causal del entorno a través de una serie de (traducir cues) (\cite{lagnado2007beyond}).
-La pregunta fundamental que plantean \cite{hagmayer2008causal} es entender qué es lo que aprenden las personas al tomar decisiones de manera secuencial de tal forma que se busque maximizar cierta recompensa a largo plazo y donde las acciones alteran el entorno.
-Al exponerse una persona a estas situaciones de decisión secuencial se obtiene una retroalimentación de cómo las acciones se relacionaban con la obtención de recompensas.
-Las distintas teorías sobre toma de decisiones bajo incertidumbre suponen distintas representaciones para esta retroalimentación.
-Hagmayer propone agrupar estas distintas representaciones en tres niveles:
-\begin{itemize}
-\item Nivel 1: Opciones, resultados y recompensas.
-Sólo los valores esperados de las opciones posibles afectan la toma de decisiones.
-Se conoce como \textit{habit learning} en la literatura de aprendizaje en animales.
-\item Nivel 2: Opciones, relaciones causales de estas opciones a los resultados, relaciones causales entre variables de resultado y relaciones de estas con las recompensas.
-Aquí entra lo que se conoce como \textit{goal-directed learning}.
-La diferencia con el nivel 1 es la capacidad de distinguir entre acciones y los valores asociados a estas.
-\item Nivel 3: Opciones, relaciones causales entre variables de resultado, relaciones causales entre estas variables y relaciones de estas con las recompensas.
-Estos son los modelos causales.
-Las distintas teorías sobre estos modelos suponen que los humanos (o el agente que aprende) adquieren estos modelos mediante procesos de aprendizaje.
-Algunas teorías suponen que la estructura causal del entorno se obtiene a través de \textit{claves}, como la propuesta por \cite{lagnado2007beyond}.
-Por otro lado, otros autores opinan que las personas utilizan propiedades estadísticas de los datos observados, este sería el caso de las teorías de \cite{gopnik2004theory} y de \cite{griffiths2005structure}.
-\end{itemize}
-Aunque las teorías que entran en el nivel 2 son sensibles a las correlaciones entre las respuestas obtenidas, sólo en el nivel 3 se hace una distinción entre causalidad y correlación.\\
-\\
-Los autores hacen énfasis en que aunque las personas sí utilicen modelos causales al tomar decisiones simples, estas decisiones suelen ser sobre escenarios hipotéticos, pero no se han hecho muchos estudios sobre adquisición de conocimiento causal al enfrentarse a problemas de decisión secuencial con el objetivo de maximizar alguna ganancia a largo plazo.
-Además, en los estudios realizados, se ha visto que las personas tienen la capacidad de inducir modelos causales a partir de claves en el entorno; se sabe además que a partir de intervenciones activas se toman mejores decisiones que considerando sólo datos observados de manera pasiva y que las personas tienen la capacidad de integrar relaciones causales obtenidas de manera separada.
-Pero en estos estudios las personas tenían como objetivo descubrir de manera directa estructuras causales.
-No se sabe bien hasta qué punto las personas adquieren conocimiento causal al perseguir objetivos que no sean descubrir directamente estructuras causales a partir del ambiente.\\
-\\
-Los autores realizaron tres experimentos para explorar las representaciones que las personas hacen de las relaciones causales que adquieren al tomar decisiones para maximizar algún objetivo.
-En el primer experimento, el objetivo era estudiar cómo las creencias causales de las personas alteran la toma de decisiones y cómo los participantes utilizaban los resultados de sus decisiones para revisar el modelo causal.
-El experimento consistía en dos etapas de toma de decisiones, la segunda siendo una etapa de prueba.
-En ambas etapas los participantes debían escoger entre una serie de opciones de modo que maximizaran ganancias.
-Se le sugirió a cada participante una posible cadena causal.
-En la primera etapa se les proveía retroalimentación de las consecuencias de sus decisiones, lo cual no sucedió en la segunda etapa sino que ellos mismos debían inferir posibles relaciones causales.
-Se concluyó que los participantes encontraron estructura causal en el problema y la utilizaron para tomar futuras decisiones.\\
-\\
-En el segundo experimento y tercer experimento, a los participantes se les mostró una serie de hipótesis causales para ver si inducían de manera espontánea un modelo causal de la situación.\\
-\\
-Lo que estos experimentos muestran, en conclusión de los autores, es que al intentar maximizar una ganancia a largo plazo, las personas adquieren conocimiento causal del entorno, el cual afecta futuras decisiones.
-Mencionan que estos experimentos ponen en tela de juicio los modelos estándar de toma de decisiones, pues utilizando sólo valores esperados no se llegarían a los mismos resultados.\\
-\\
-Como consecuencia de sus estudios, se obtiene que los métodos basados en los niveles 1 y 2, que es donde caerían los métodos tradicionales de aprendizaje por refuerzo no son suficientes para una adecuada toma de decisiones en un sistema causal, pues estas formulaciones no cuentan con la expresividad suficiente para capturar las relaciones causales.
-Esto confirma la necesidad de pensar en modelos que permitan aprender modelos causales a la par que un agente aprende de su entorno mediante interacción y exploración.\\
-\\
-En cuanto a la formulación de un problema de decisión en sistemas causales, ya desde \cite{dawid2002influence} tenemos una idea de cómo aumentar redes bayesianas con nodos de decisión y variables de estrategia para el modelado causal en problemas de decisión.\\
-\\
-\cite{hagmayer2009decision} plantean que cada decisión tomada por un agente es una intervención directa con el fin de predecir las consecuencias de esta.
-
-\subsection{Experimentos secuenciales: Causal Bandits}
-
-\section{Mi Propuesta}
-\subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
-\subsection{Problema}
-\subsection{Objetivos}
-\begin{itemize}
-\item Plantear un modelo causal dado un ambiente.
-\item Aprendizaje estructural del modelo causal.
-\item Evaluación del modelo causa (queries significativas)l.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
-\end{itemize}
-\subsection{Motivación}
-Tomando el cuenta los trabajos mencionados en la sección anterior, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
+\section{Trabajos previos}
+\subsection{Estado del arte}
+De manera más detallada, podemos ver que los diferentes enfoques que se han utilizado para la adquisición, y uso, de información causal en problemas de decisiones secuenciales pueden dividirse de la siguiente manera:
+\subsubsection{Métodos basados en actualización de grafos}
+\cite{heckerman2006bayesian}, \cite{wellen2012learning}, \cite{kummerfeld2013tracking}
+\subsubsection{Métodos basadps en representaciones del ambiente}
+\cite{ho2013knowledge}, \cite{ho2014rapid} \cite{garnelo2016towards}
+\subsubsection{Métodos basados en decisiones secuenciales}
+\cite{hagmayer2008causal}, \cite{hagmayer2013repeated}, \cite{alon2015online}, \cite{albrecht2016exploiting} \cite{lattimoreNIPS2016}, \cite{innes2018reasoning}
+\section{Limitaciones}
+\section{Motivación}
+Tomando el cuenta los trabajos mencionados, surge la pregunta de si es posible diseñar un agente que aprenda a realizar una tarea por medio de interacción con su entorno y mediante el uso de información causal recolectada a través de esa interacción.
 Este agente buscará aprender un modelo causal de su entorno al mismo tiempo que aprende una \textit{política}.
@@
 Los autores conjeturan que esto pueden utilizarse para hacer inducción causal sobre una serie de acciones establecidad (política conocida).
-\subsection{La propuesta}
-Diseñar un agente que tenga como objetivo aprender a llevar a cabo una tarea.
-El aprendizaje se dará por interacción con el ambiente y además contará con un modelo causal que vaya aprendiendo paso a paso y del cual se apoyará para  tomar mejores decisiones.
-\begin{center}
-\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/diagrama.png}
-\end{center}
-\subsection{Pasos a seguir}
-\begin{itemize}
-\item ¿Por qué esto sería mejor que trabajo de Garnelo?
-\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? Aquí ayuda el artículo de \cite{lattimoreNIPS2016}
-\item Tengo una política aprendida, ¿cómo descubro estructura causal? Aquí puede ayudar el artículo de \cite{ortega2014generalized}
-\item ¿Qué aprendo primero?
-\item ¿Cómo le doy más peso a la información explicativa que a la información reactiva?
-\item Exploración vs explotación causal: ¿cómo decido si seguir la info causal o tomar acciones exploratorias para descubrir relaciones causales? 
-\item ¿Un dominio de interés? 
-\end{itemize}
-\subsection{Ejemplo: Caso de un MDP finito}
-En el caso de un MDP finito completamente especificado; es decir, en el cual conocemos las probabilidades de transición y las recompensas.
-Un modelo causal sobre este dominio opera sobre las acciones, pues son estas las que \textbf{\textit{suponemos}} que \textit{causan} algo; esto es,
-\[ a \xrightarrow[]{\text{causa}} \textrm{estado del mundo}. \]
-Entonces, en un modelo gráfico causal necesitaremos tener \textit{nodos de estado}, así como \textit{nodos de acciones} (\textit{causal decision graphs}).\\
-\\
-Es claro que los estados del mundo afectarán la siguiente decisión, pero no causan nada por sí mismos; en todo caso, el estado afeca la decisión pero sólo a través de la política (el valor del estado).\\
-\\
-El objetivo, ahora, es aprender una política que incorpore conocimiento causal, el cual a su vez fue aprendido mediante interacción.\\
-\\
-¿El valor de un estado debe cambiar en el caso causal?
-Para que el valor de un estado se vea afectado por lo que pueda causar el agente a partir de ese estado se necesita algún tipo de función de utilidad, ¿no?
-No lo sé...
-\newpage
 \bibliographystyle{apalike}
-\bibliography{Bibliografia}
-\appendix
-\section{Teoría de Probabilidad}
-\section{Estadística Bayesiana}
-
+\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
 \end{document}

### 2018-02-22 Segundo_Semestre/Propuesta/Propuesta.tex
problemas de decision

@@
 Se dice que una variable $X_i$ es causa directa de $X_j$ si
-\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')} \]
-\subsection{Contrafactuales en modelos funcionales}
+\[ P_{X_j | Do(X_i = x_i} \neq P_{X_j | do(X_i = x')}. \]
+
+\subsection{El enfoque Bayesiano}
 
@@
 \end{itemize}
-Notemos que el axioma A3 sólo aplica bajo un esquema de neutralidad al riesgo.
+Notemos que el axioma A3 sólo aplica si el tomador de decisiones es neutral al riesgo (\cite{gilboa2009decision}).
 \begin{teo}
@@
 Su modelo asume sólo estados del mundo y sus resultados.
-Por lo tanto, los axiomas propuestos por Savage son bastante más abstractos y menos interpretables, pero se obtiene la ventaja de que sus axiomas son equivalentes a la existencia de una función de utilidad y de una medida de probabilidad (subjetiva) tales que las decisiones se llevan a cabo al maximizar utilidad esperada respecto a esta función de utilidad y esta medida de probabilidad.\\
-\subsubsection{Estados, resultados y acciones}
+Por lo tanto, los axiomas propuestos por Savage son bastante más abstractos y menos interpretables, pero se obtiene la ventaja de que sus axiomas son equivalentes a la existencia de una función de utilidad y de una medida de probabilidad (subjetiva) tales que las decisiones se llevan a cabo al maximizar utilidad esperada respecto a esta función de utilidad y esta medida de probabilidad.
+
+\subsection{Formalismo moderno}
+\subsubsection{Problema de Decisión}
+\begin{defi}
+Un problema de decisión es la situación en la cual un agente (llamado tomador de decisiones) se enfrenta a un conjunto de decisiiones $D$ de entre las cuales tiene que escoger una.
+\end{defi}
+Basado en \cite{Bernardo1985}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, tenemos que los elementos que componen un problema de decisión son:
+\begin{itemize}
+\item Un espacio de medida $(\Omega, \mathcal{E})$
+\item Un conjunto $\{d_i, i \in I\}$ llamado conjunto de decisiones, también llamadas acciones u opciones.
+\item Para cada $d_i \in D$, una partición $E_i= \{E_{ij}, j \in J \}$ de $\mathcal{E}$ de \textit{eventos inciertos}.
+\item Correspondiente a cada $E_i$, un conjunto de consecuencias $C_i = \{ c_{ij}, j \in J \}$.
+\item Una relación de orden $\preceq$ sobre $D \times D$
+\end{itemize}
+Notemos que las particiones asociadas a cada decisión no tienen por qué ser iguales, pero siempre podemos construir
+\[ E = \cap_{i \in I} A_i \textrm{ donde } A_i \in E_i. \]
+De modo que en el futuro supondremos, sin perdida de generalidad, que es una misma partición para todas las decisiones, y por lo tanto un solo conjunto de consecuencias, por lo que la notación utilizada para representar cada posible decisión es la siguiente $d_i = \{ c_j | E_j, j \in J\}$ lo cual se interpreta como que al escoger $d_i$, ocurrirá uno de los eventos inciertos dentro de la partición $E$, digamos, $E_j$ y eso trae como consecuencia $c_j$.
+Además, notemos que el orden de los índices es irrelevante, pues la opción $\{ c_1 | E, c_2 | E^c \}$ es la misma que $\{ c_2 | E^c, c_1 | E \}$.
+\begin{defi}
+Diremos que un problema de decisión es sin incertidumbre si para cada $d_i \in D$ se tiene que $C_i = \{c_i\}$; es decir,  el conjunto de consecuencias posibles  asociadas a la decisión $d_i$ consta de un solo elemento.
+\end{defi}
+\begin{defi}
+Análogamente, un problema de decisión para el cual se tiene que para al menos un elemento $d_i \in D$ existe más de una posible consecuencia se dice que es un problema de decisión con incertidumbre.
+\end{defi}
+Como requerimiento técnico, necesitamos que todos los elementos del conjunto de consecuencias pertenezcan al conjunto de opciones; esto se logra identificando cada consecuencia como la opción $\{c|\Omega \}$.
+Siguiendo esta notación, podemos definir una relación de preferencias entre consecuencias a partir de la relación de preferencia entre opciones de la siguiente manera
+\[ c_1 \leq c_2 \Leftrightarrow \{ c_1 | \Omega \} \preceq \{ c_2 | \Omega \}.\]
+A través de un razonamiento análogo, odemos definir una noción de qué tan \textit{creíble} es un evento incierto para un tomador de decisiones a partir de sus preferencias sobre las acciones de la siguiente manera
+\begin{defi}
+Se dice que un evento más creible que otro como
+\[E \leq F \Leftrightarrow \textrm{ para todo } c_1 < c_2 \{ c_2 | E, c_1 | E^c \} \preceq \{ c_2 | F, c_1 | F^c \} \]
+\end{defi}
+Esto permite definir una relación de preferencia a la luz de la ocurrencia de algún evento incierto
+\begin{defi}
+Para cualquier evento $G$ distinto del vacío, definimos:
+\begin{itemize}
+\item $d_1 \leq_G d_2 \Leftrightarrow \textrm{ para toda opción } d \textrm{ se tiene que } \{d_1 | G, d | G^c \} \preceq \{ d_2 | G, d | G^c\}$
+\item $c_1 \leq_G c_2 \Leftrightarrow \{ c_1 | \Omega \} \leq_G \{ c_2 | \Omega \}$.
+\item $E \leq_G F \Leftrightarrow \textrm{ para } c_1 \leq_G c_2, \{ c_2 | E, c_1 | E^c \} \leq_G \{ c_2 | F, c_1 | F^c \} $
+\end{itemize}
+\end{defi}
+\subsubsection{Axiomas de coherencia}
+
+\subsection{Problemas de Decisión Secuenciales}
 \section{Aprendizaje por Refuerzo}
@@
 El Aprendizaje por Refuerzo (RL) (\cite{sutton1998reinforcement}) consiste en un \textit{framework} general en el cual un \textit{agente} aprende a llevar a cabo una tarea a través de la interacción con su entorno.
-Estas interacciones provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
+Estas interacciones, representadas como decisiones que el agente toma, provocan cambios en el entorno, los cuales a su vez afectarán el comportamiento del agente.
 El comportamiento del agente está caracterizado en términos de acciones y las acciones tomadas por este afectan el \textit{estado} del ambiente.
@@
 Formalmente:
-\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a] \]
+\[ Q^\ast(s,a)=\mathbb{E}_{s'} [r + \gamma \max_{a'} Q^\ast(s',a')|s,a]. \]
+\subsubsection{Relación con problemas de decisión bajo incertidumbre}
 \subsection{Métodos básicos de solución: Model-based}
@@
 Buen performance en juegos de Atari que utilizan la pantalla como input.
+
 \section{Reinforcement Learning y Causalidad}
@@
 \\
-En líneas más avanzadas, \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
+En líneas más avanzadas \cite{courville2006bayesian} exploraron versiones que permiten que múltiples causas latentes estén activas.
+
 \subsection{Towards Deep Symbolic Reinforcement Learning}
@@
 \cite{ho2016reinforcement}\\
+\\
+\cite{ortega2015causal}
 \subsection{Estado del arte}

### 2018-02-28 Segundo_Semestre/Propuesta/Propuesta.tex
Cambios en la motivacion y propuesta para considerar juegos bayesianos

@@
 \subsection{Problemas de Decisión Secuenciales}
+\section{Diagramas de Influencia}
+\section{Juegos Bayesianos}
 \section{Aprendizaje por Refuerzo}
@@
 \subsubsection{Relación con problemas de decisión bajo incertidumbre}
+¿Será que las acciones obtenidas a partir de seguir la política óptima alcanzarán la máxima utilidad esperada? Sí
 \subsection{Métodos básicos de solución: Model-based}
@@
 \subsection{Experimentos secuenciales: Causal Bandits}
+El problema que \cite{lattimoreNIPS2016} estudian consiste en utilizar modelos causales para aprender a escoger intervenciones en un entorno incierto; esto es, dado un modelo causal, se busca escoger la intervención que maximice cierta variable de recompensa sujeto a ciertas restricciones.
+Modelan el escoger intervenciones como \textit{brazos} de una máquina tragamonedas (conocidas como \textit{bandits} en la literatura.
+La influencia de estas intervenciones en la variable de pago se rige por un modelo causal conocido.
+Comentan que una generalización mucho más complicada sería en el caso en el que el modelo causal sea completa o parcialmente desconocido y miembro de una gran clase de modelos.
+
 \subsection{Toma de Decisiones Bajo Incertidumbre y Teoría de Juegos}
-Teoría de Juegos (\cite{von1944theory}, \cite{osborne1994course}, \cite{webb2007game}, \cite{gintis2014bounds})
-\subsubsection{Aprendizaje por refuerzo como toma de decisiones}
+Teoría de Juegos (\cite{von1944theory}, \cite{osborne1994course}, \cite{webb2007game}, \cite{gintis2014bounds}) consiste en el estudio de las interacciones entre tomadores de decisiones en donde cada uno de los cuales busca un objetivo bien definido.
+\\
+\\
+Un juego está formador por (\cite{osborne1994course})
+
 \subsubsection{Juegos Bayesianos}
+Un juego Bayesiano consiste (\cite{osborne1994course}) en 
+\begin{itemize}
+\item Un conjunto finito de jugadores $N$.
+\item Un conjunto finito de estados $S$.
+\end{itemize}
+Para cada jugador $i \in N$
+\begin{itemize}
+\item Un conjunto $A(i)$ de las acciones disponibles al jugador $i$.
+\item Un conjunto finito $T(i)$ se señales observadas por el jugador $i$
+\item Una función de señales $\tau_i : \Omega \to T_i$ del jugador
+\item Una medida de probabilidad $p_i$ sobre los estados $S$ tal que asigna masa estrictamente mayor a cero a los estados que generan las 
+señales.
+\item Una relación de preferencias $\preceq$ sobre el conjunto de medidas de probabilidad que están definidas sobre $A \times S$
+\end{itemize}
 \subsubsection{Información causal y juegos}
+Notemos que los juegos Bayesianos permiten que cada jugador tenga creencias a priori sobre el estado de la naturaleza (toda la información relevante sobre los otros jugadores).
+Entonces, es natural pensar que esta información que tiene sea sobre las relaciones causales que se sostienen en el entorno del juego.\\
+\\
+Para hacer más explícita esta situación, consideremos que la Naturaleza es un jugador más el cual actúa según otros modelos.
+De esta manera, al suponer que existen relaciones causales que rigen el ambiente, estas relaciones causales guiarán los actos del jugador Naturaleza.\\
+
 \section{Mi Propuesta}
 \subsection{Pregunta de Investigación}
-¿Los métodos de aprendizaje por refuerzo pueden aprender un modelo causal del ambiente? Más allá de producir un mapeo estado-acción, queremos obtener un modelo causal del entorno.
+¿Es posible aprender un modelo causal del ambiente en un proceso de toma de decisiones secuencial?
 \subsection{Problema}
-Adquirir y utilizar información causal para mejorar el desempeño de un agente que interactua con su entorno con el fin de maximizar una recompensa a largo plazo. 
+\section{Problema}
+Adquisición y uso de información causal en toma de decisiones secuenciales.
+\subsection{Subproblema}
+Actualización de creencias causales en un juego dinámico de un solo jugador contra la naturaleza. 
 \subsection{Motivación}
@@
 \cite{ortega2015causal}
+
+La toma de decisiones bajo incertidumbre provee un marco de teórico general bajo el cual pueden representarse una gran variedad de problemas.
+Un problema de toma de decisiónes bajo incertidumbre consiste en un conjunto de decisiones, un conjunto de eventos inciertos, y un conjunto de consecuencias asociadas a cada pareja de decisión y eventos inciertos.
+Además, para cada tomador de decisiones se tiene una relación de preferencias para poder cuantificar cuánto \textit{desea} el agente una consecuencia sobre otra.\\
+\\
+A partir de una serie de axiomas, conocidos como axiomas de coherencia (\cite{bernardo2000bayesian}), se puede demostrar formalmente que el único criterio de solución que es coherente con los axiomas es la maximización de la utilidad esperada (\cite{savage1954the}).\\
+\\
+Inicialmente, este marco de trabajo permite incorporar información sobre el ambiente y sobre las creencias iniciales de un tomador de decisiones, lo cual estará codificado mediante el uso de distribuciones de probabilidad.
+Las distribuciones de probabilidad por sí solas no bastan para codificar relaciones causales (\cite{pearl2009causality}) .
+Por ejemplo (\cite{peters2017elements}), se sabe que la altitud y la temperatura promedio de cierta ubicación geográfica dependen una de la otra, y si consideramos su distribución conjunta $P(a,t)$, tenemos que podemos factorizar esta distribución de la siguiente manera $P(a|t)P(t)$ y $P(t|a)P(a)$, pero sólo una de estas puede tener interpretación causal.\\
+\\
+Los modelos gráficos causales (\cite{pearl2009causality}, \cite{koller2009probabilistic}) permiten representar relaciones (dirigidas) causa-efecto.
+Un modelo gráfico causal consta de $n$ nodos y si un vértice $e$ conecta a los nodos $(i,j)$ se dice que $i$ es una causa de $j$.
+
+Los diagramas de influencia (\cite{jensen2001bayesian}, \cite{koller2009probabilistic}) permiten añadir \textit{nodos de decisión} a redes bayesianas.
+En particular, utilizando que las redes bayesianas son los \textit{esqueletos} detrás de los modelos gráficos causales,  \cite{dawid2002influence} considera la información causal que puede obtenerse a partir de un diagrama de influencia.\\
+\\
+Los juegos Bayesianos (\cite{harsanyi1967games}) permiten que los jugadores tengan \textit{creencias subjetivas} sobre el comportamiento de otros jugadores; en particular, un problema de decisiónes secuenciales en un entorno incierto puede verse como un juego secuencial de un solo jugador vs la naturaleza.
+Entonces, podemos pensar en que este jugador tiene creencias sobre el comportamiento de la naturaleza (\cite{monderer1997dynamic}), dentro de la cual opera un modelo causal.
+Entonces, estamos pensando en un juego bayesiano dinámico con información parcial.\\
+\\
+Además, sabemos que el criterio de la utilidad esperada es el único criterio de solución para un problema de decisión con incertidumbre que es fiel a los axiomas de coherencia; esto implica que un modelo que incorpore información causal no está garantizado que sea \textit{teóricamente mejor}
+
+El modelo de \cite{witsenhausen1971information}
+Un ejemplo: 
 \subsection{Estado del arte}
 De manera más detallada, podemos ver que los diferentes enfoques que se han utilizado para la adquisición, y uso, de información causal en problemas de decisiones secuenciales pueden dividirse de la siguiente manera:
-\subsubsection{Métodos basados en actualización de grafos}
-\cite{heckerman2006bayesian}, \cite{wellen2012learning}, \cite{kummerfeld2013tracking}
-\subsubsection{Métodos basadps en representaciones del ambiente}
-\cite{ho2013knowledge}, \cite{ho2014rapid} \cite{garnelo2016towards}
-\subsubsection{Métodos basados en decisiones secuenciales}
-\cite{hagmayer2008causal}, \cite{hagmayer2013repeated}, \cite{alon2015online}, \cite{albrecht2016exploiting} \cite{lattimoreNIPS2016}, \cite{innes2018reasoning}
+\subsubsection{Actualizaciones locales}
+\cite{heckerman2006bayesian}, 
+\cite{wellen2012learning},
+\cite{kummerfeld2013tracking}
+\cite{bramley2017formalizing}
+\subsubsection{Representaciones del ambiente}
+En una serie de trabajos (\cite{ho2013knowledge}, \cite{ho2014rapid}, \cite{ho2016ground}, \cite{ho2016principles}, \cite{ho2016cognitively}) que culminan en \cite{ho2016reinforcement}, los autores explican cómo un sistema podría construir una serie de reglas lógicas que codifiquen las relaciones causales encontradas.
+De esta manera, el agente consultará este conjunto de reglas para la futura toma de decisiones.
+Aunque las intuiciones son interesantes, no proponen un algoritmo sólido para llevar a cabo lo que plantean.
+\\
+\\
+ Un agente que interactua con su entorno con el fin de maximizar una recompensa efectivamente puede aprender aspectos estructurales sobre el entorno, como nos muestran\cite{garnelo2016towards}, quienes construyen un prototipo muy sencillo en el cual un agente construye reglas lógicas a partir de interacciones con el ambiente, las cuales son utilizadas directamente para escoger la siguiente acción a tomar.
+Todo su diseño está orientado en aprender a jugar un sencillo juego a partir de los pixeles de la pantalla.
+Además, muestran la importancia de tener una semántica que permita hablar de la estructura del entorno.
+Ellos utilizaron lógica de primer orden, pues están más interesados en hacer razonamientos \textit{de alto nivel} sobre el entorno, que involocura más que las relaciones causales, y por esto ellos se comprometen con lógica de primer orden, que al restringirnos al caso causal se queda corta pues los modelos gráficos probabilistas son más expresivos, además de modulares y montónicos
+\subsubsection{Decisiones secuenciales}
+El trabajo de \cite{wellen2012learning} muestra cómo un modelo gráfico causal podría ser aprendido en línea al tomar en cuenta las decisiones tomadas y las recompensas obtenidas por un tomador de decisiones.
+Por otro lado, los modelos de \cite{sloman2006causal}, \cite{meder2008inferring}, \cite{hagmayer2009decision} y \cite{hagmayer2013repeated} muestran que estas intuiciones no son erradas y que los seres humanos incorporamos y adquirimos conocimiento causal al enfrentarnos en problemas de toma de decisión simples, aunque aun no existen muchos estudios en los cuales los escenarios de decisión no sean hipotéticos sino que se intente maximizar una recompensa real a largo plazo.
+De la misma manera, \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
+Por lo tanto, es interesante preguntar si un agente que busque maximizar una recompensa pueda aprender modelos causales mediante la interacción con su entorno.
+
+\cite{hagmayer2008causal},
+
+\cite{hauser2012two}
+\cite{hagmayer2013repeated}, 
+\cite{agarwal2014taming}
+\cite{alon2015online}, 
+El problema que \cite{lattimoreNIPS2016} estudian consiste en utilizar modelos causales para aprender a escoger intervenciones en un entorno incierto; esto es, dado un modelo causal, se busca escoger la intervención que maximice cierta variable de recompensa sujeto a ciertas restricciones.
+Modelan el escoger intervenciones como \textit{brazos} de una máquina tragamonedas (conocidas como \textit{bandits} en la literatura.
+La influencia de estas intervenciones en la variable de pago se rige por un modelo causal conocido.
+Comentan que una generalización mucho más complicada sería en el caso en el que el modelo causal sea completa o parcialmente desconocido y miembro de una gran clase de modelos.
+Mientras que \cite{garnelo2016towards} utiliza una representación lógica para escoger una acción, \cite{lattimoreNIPS2016} explora la pregunta de cómo escoger \textit{la mejor intervención} dado un modelo causal a través de una serie de intervenciones y observaciones secuenciales; por ejemplo, en el caso en el que un granjero quiera maximizar sus cosechas, y sabe que ese rendimiento está afectado por el uso de fertilizante, la humedad y la exposición al sol, pero sólo tiene los recursos para llevar a cabo sólo una de estas modificaciones en cada temporada.
+El algoritmo de Lattimore le dirá, después de $T$ observaciones, cuál es la mejor intervención a llevar a cabo con el fin de maximizar esa cantidad.
+Este podría ser un primer paso hacia responder la pregunta \textit{¿dado un modelo causal, qué acción tomar?}\\
+\\
+\cite{thompson1933likelihood} al enfrentarse a un estudio en el cual se trataba de averiguar cuál de dos fármacos era mejor al probarlos en una población, pero con la restricción de exponer a la menor cantidad de personas posibles a la droga menos buena, propuso actualizar las proporciones de muestreo para futuros experimentos, de modo que se exponga a menos personas en el futuro a la droga de menor calidad.
+Actualmente, se conoce como muestreo de Thompson al proceso de aleatorizar acciones según la probabilidad (subjetiva) de que cierta acción sea la óptima en un entorno de incertidumbre.
+Notemos que este esquema de muestreo puede verse como un problema de \textit{bandits}, y de hecho se considera en la literatura como un método para resolver esta clase de problemas.
+\cite{ortega2014generalized} argumentan que el muestreo de Thompson puede utilizarse para hacer inducción causal para agentes que interactuan con un entorno incierto de estructura causa desconocida.
+Lo que hacen es retomar una idea presentada en \cite{ortega2011bayesian} que consiste en usar árboles de probabilidad para modelar estructuras causales sobre eventos inciertos.
+Utilizando el muestreo de Thompson, un tomador de decisiones primero muestrea un parámetro de su modelo de creencias, escoge la acción óptima y a partir de la acción que se escoge se interviene el modelo y actualiza sus creencias de manera bayesiana a partir de la evidencia observada.
+
+\cite{innes2018reasoning} consideran un ejemplo similar al que se mencionó anteriormente, y consideran el caso en el cual aparezcan situaciones sobre las cuales no se sabe nada.
+Proponen un modelo que descubre y aprende a explotar situaciones no vistas antes, esto lo hacen mediante exploración con el mundo y mediante comunicación con un experto.
+Extienden resultados de \cite{bramley2015staying} y \cite{albrecht2016exploiting} quienes suponen que las conjunto de variables aleatorias no cambia durante el entrenamiento.
+Además, incluyen la posibilidad de aprender los pagos potenciales.
+\subsubsection{Juegos Bayesianos con información incompleta}
+
 \subsection{Objetivos}
@@
 \subsection{La propuesta}
-Diseñar un agente que tenga como objetivo aprender a llevar a cabo una tarea.
-El aprendizaje se dará por interacción con el ambiente y además contará con un modelo causal que vaya aprendiendo paso a paso y del cual se apoyará para  tomar mejores decisiones.
+Considerando un entorno incierto y dinámico en el cual existen relaciones causales subyacentes y un agente con acceso información limitada sobre este entorno, se propone que dicho agente descubra y utilice información causal con el fin de resolver una tarea
 \begin{center}

### 2018-02-28 Segundo_Semestre/Propuesta/Propuesta.tex
Cambiando los Objetivos

@@
 \subsubsection{Juegos Bayesianos con información incompleta}
+Un ejemplo:\\
+Supongamos que un paciente que llega a un hospital puede tener un padecimiento $A$ con probabilidad $p$ o $B$ con probabilidad $q$.
+El médico de guardia tiene a su alcance tres tratamientos posibles: cirugía, fármaco o nada.
+En caso de no hacer nada, el paciente morirá con una probabilidad $p_{dn}=0.8$ y sobrevivirá con una probabilidad $1-p_d$.
+Sin importar la enfermedad, el paciente puede morir durante la cirugía con probabilidad $p_{dc}=0.5$.
+De la misma manera, el medicamento producirá una reacción alérgica mortal con probabilidad $p_{df}=0.2$\\
+\\
+Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$.
+Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
+\\
+Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$.
+Si el paciente tenía la enfermedad $B$, entonces ésta lo curará con probabilidad de $0.6$.
+\qed
+\\
+Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza.
+Las acciones disponibles al médico son el fármaco, cirugía, o nada, mientras que las jugadas de la Naturaleza son \textit{volados} que determinarán si el paciente vive o no; el juego va así, primero la naturaleza hace su jugada, lo cual determina qué enfermedad tenía el paciente.
+Después, el médico decide qué hacer, y después la Naturaleza lanzará una moneda cargada que determinará si el paciente sobrevive.
+Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc.
+\\
+Sabemos que el médico fue a la escuela de medicina, de donde posee una gran cantidad de información (causal) a priori, pero puede que se esté enfrentando a una nueva enfermedad, es decir, no conoce realmente las probabilidades de muerte ni los casos en que el fármaco no actua.
+Lo que se desea es que a partir de este caso, o de varios más, pueda actualizar su conocimiento causal para la siguiente ocasión en la que deba enfrentarse a  un problema similar.
 
@@
 \begin{itemize}
-\item Plantear un modelo causal dado un ambiente.
-\item Aprendizaje estructural del modelo causal.
-\item Evaluación del modelo causa (queries significativas)l.
-\item Incorporar el conocimiento causal a la toma de decisiones en un loop de mejora continua.
-\item Exportar conocimiento causal obtenido para hacer inferencia causal.
+\item Estudiar la relación entre un problema de decisión secuencial y un juego vs la Naturaleza.
+\item Definición de creencias subjetivas sobre las relaciones causales de un ambiente.
+\item Definición de un lenguaje que permita incluir información causal en toma de decisiones: diagramas de influencia causales.
+\item Actualización de creencias sobre las relaciones causales a la luz de lo observado por el jugador.
+\item Evaluación del desempeño del jugador según el criterio de la máxima utilidad esperada
 \end{itemize}

### 2018-02-28 Segundo_Semestre/Propuesta/Propuesta.tex
Cambios en los Pasos a Seguir

@@
 \begin{itemize}
-\item ¿Por qué esto sería mejor que trabajo de Garnelo?
+\item Dado un sencillo problema de decisión bajo incertidumbre completamente especificado, plantear un juego y estudiarlo.
+\item Suponer un modelo causal, conocido, que controla las jugadas de la Naturaleza, ¿cómo rastrearlo?
+\item Dotar al jugador de creencias causales, ¿cómo especificarlas?
 \item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? Aquí ayuda el artículo de \cite{lattimoreNIPS2016}

### 2018-02-28 Segundo_Semestre/Seminario/Semana6/State_of_Art_presentation.Rmd
esqueleto presentacion

---
title: "State of the Art"
author: "Mauricio Gonzalez"
date: "28/2/2018"
output:
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
bibliography: /Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib
---

# Introduction
## Problem

# State of The Art

## Local updating

## Environment representations

## Sequential decision making

## Bayesian games

### 2018-03-01 Segundo_Semestre/Propuesta/Propuesta.tex
moviendo lo de decisiones secuenciales y juegos justo despues de decisiones bajo incert

@@
 \section{Teoría de la Decisión}
-\subsection{von Neumann-Morgenstern}
+\subsection{Un poco de historia}
+\subsubsection{von Neumann-Morgenstern}
 Según narra \cite{gilboa2009decision}, fue Daniel Bernoulli quién habló por primera vez de la maximización de la utildiad esperada; esto, al estudiar la \textit{paradoja de San Petersburgo}, la cual consiste en lo siguiente: un tomador de decisiones se enfrente a una serie de lanzamientos consecutivos de una moneda justa.
@@
 Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
-\subsection{De Finnetti}
+\subsubsection{De Finnetti}
 ¿Cuál es la definición de probabilidad subjetiva?
@@
 Entonces, la definición de probabilidad subjetiva sería aquella probabilidad utilizada por un tomador de decisiones para escoger una acción mediante el criterio de utilidad esperado.
-\subsection{Savage}
+\subsubsection{Savage}
 El Teorema de von Neumann-Morgenstern indica cómo obtener una función de utilidad si las probabilidades están bien definidas; el Teorema de De Finnetti dice cómo obtener probabilidades si conocemos las utiidades de un tomador de decisiones.
@@
 
-\subsection{Problemas de Decisión Secuenciales}
+\subsection{Decisiones Secuenciales y Teoría de Juegos}
+Teoría de Juegos (\cite{von1944theory}, \cite{osborne1994course}, \cite{webb2007game}, \cite{gintis2014bounds}) consiste en el estudio de las interacciones entre tomadores de decisiones en donde cada uno de los cuales busca un objetivo bien definido.
+\\
+\\
+Un juego está formador por (\cite{osborne1994course})
+
+\subsubsection{Juegos Bayesianos}
+Un juego Bayesiano consiste (\cite{osborne1994course}) en 
+\begin{itemize}
+\item Un conjunto finito de jugadores $N$.
+\item Un conjunto finito de estados $S$.
+\end{itemize}
+Para cada jugador $i \in N$
+\begin{itemize}
+\item Un conjunto $A(i)$ de las acciones disponibles al jugador $i$.
+\item Un conjunto finito $T(i)$ se señales observadas por el jugador $i$
+\item Una función de señales $\tau_i : \Omega \to T_i$ del jugador
+\item Una medida de probabilidad $p_i$ sobre los estados $S$ tal que asigna masa estrictamente mayor a cero a los estados que generan las 
+señales.
+\item Una relación de preferencias $\preceq$ sobre el conjunto de medidas de probabilidad que están definidas sobre $A \times S$
+\end{itemize}
+\subsubsection{Información causal y juegos}
+Notemos que los juegos Bayesianos permiten que cada jugador tenga creencias a priori sobre el estado de la naturaleza (toda la información relevante sobre los otros jugadores).
+Entonces, es natural pensar que esta información que tiene sea sobre las relaciones causales que se sostienen en el entorno del juego.\\
+\\
+Para hacer más explícita esta situación, consideremos que la Naturaleza es un jugador más el cual actúa según otros modelos.
+De esta manera, al suponer que existen relaciones causales que rigen el ambiente, estas relaciones causales guiarán los actos del jugador Naturaleza.\\
+
 \section{Diagramas de Influencia}
-\section{Juegos Bayesianos}
+
 \section{Aprendizaje por Refuerzo}
@@
 
-\subsection{Toma de Decisiones Bajo Incertidumbre y Teoría de Juegos}
-Teoría de Juegos (\cite{von1944theory}, \cite{osborne1994course}, \cite{webb2007game}, \cite{gintis2014bounds}) consiste en el estudio de las interacciones entre tomadores de decisiones en donde cada uno de los cuales busca un objetivo bien definido.
-\\
-\\
-Un juego está formador por (\cite{osborne1994course})
-
-\subsubsection{Juegos Bayesianos}
-Un juego Bayesiano consiste (\cite{osborne1994course}) en 
-\begin{itemize}
-\item Un conjunto finito de jugadores $N$.
-\item Un conjunto finito de estados $S$.
-\end{itemize}
-Para cada jugador $i \in N$
-\begin{itemize}
-\item Un conjunto $A(i)$ de las acciones disponibles al jugador $i$.
-\item Un conjunto finito $T(i)$ se señales observadas por el jugador $i$
-\item Una función de señales $\tau_i : \Omega \to T_i$ del jugador
-\item Una medida de probabilidad $p_i$ sobre los estados $S$ tal que asigna masa estrictamente mayor a cero a los estados que generan las 
-señales.
-\item Una relación de preferencias $\preceq$ sobre el conjunto de medidas de probabilidad que están definidas sobre $A \times S$
-\end{itemize}
-\subsubsection{Información causal y juegos}
-Notemos que los juegos Bayesianos permiten que cada jugador tenga creencias a priori sobre el estado de la naturaleza (toda la información relevante sobre los otros jugadores).
-Entonces, es natural pensar que esta información que tiene sea sobre las relaciones causales que se sostienen en el entorno del juego.\\
-\\
-Para hacer más explícita esta situación, consideremos que la Naturaleza es un jugador más el cual actúa según otros modelos.
-De esta manera, al suponer que existen relaciones causales que rigen el ambiente, estas relaciones causales guiarán los actos del jugador Naturaleza.\\
-
 \section{Mi Propuesta}
@@
 En una serie de trabajos (\cite{ho2013knowledge}, \cite{ho2014rapid}, \cite{ho2016ground}, \cite{ho2016principles}, \cite{ho2016cognitively}) que culminan en \cite{ho2016reinforcement}, los autores explican cómo un sistema podría construir una serie de reglas lógicas que codifiquen las relaciones causales encontradas.
-De esta manera, el agente consultará este conjunto de reglas para la futura toma de decisiones.
-Aunque las intuiciones son interesantes, no proponen un algoritmo sólido para llevar a cabo lo que plantean.
-\\
-\\
- Un agente que interactua con su entorno con el fin de maximizar una recompensa efectivamente puede aprender aspectos estructurales sobre el entorno, como nos muestran\cite{garnelo2016towards}, quienes construyen un prototipo muy sencillo en el cual un agente construye reglas lógicas a partir de interacciones con el ambiente, las cuales son utilizadas directamente para escoger la siguiente acción a tomar.
+En particular, en \cite{ho2016ground} muestran cómo es posible descubrir relaciones causales a partir de la observación de eventos que ocurren de manera sucesiva en el tiempo, como relámpagos y truenos o disparos y “dolor”.
+Su algoritmo es capaz de distinguir estas relaciones de otros factores externos.
+Es importante mencionar que para ellos basta descubrir causalidad \textit{efectiva}; es decir, relaciones causales que puedan utilizarse sin la necesidad de saber si eran verdaderas o no; es decir, que sirvan a la hora de resolver un problema.
+Posteriormente, en \cite{ho2016reinforcement} utilizan el método desarrollado en \cite{ho2014rapid} para construir \textit{scripts} que son reglas que, a partir de las relaciones causales, describen el resultado de tomar cierta acción.
+Finalmente, utilizan un proceso de encadenamiento hacia atrás para encontrar la mejor serie de acciones a seguir.
+En resumen, un agente primero observará un \textit{stream} de eventos, con lo que construirá reglas locales, luego el agente consultará este conjunto de reglas para la futura toma de decisiones.
+Notemos que no es un procedimiento que se lleve a cabo de manera \textit{on-line}.
+\\
+Como nos muestran\cite{garnelo2016towards} un agente que interactua con su entorno con el fin de maximizar una recompensa efectivamente puede aprender aspectos estructurales sobre el entorno de manera \textit{on-line}.
+Los autores construyen un prototipo muy sencillo en el cual un agente construye reglas lógicas a partir de interacciones con el ambiente, las cuales son utilizadas directamente para escoger la siguiente acción a tomar.
 Todo su diseño está orientado en aprender a jugar un sencillo juego a partir de los pixeles de la pantalla.
 Además, muestran la importancia de tener una semántica que permita hablar de la estructura del entorno.
-Ellos utilizaron lógica de primer orden, pues están más interesados en hacer razonamientos \textit{de alto nivel} sobre el entorno, que involocura más que las relaciones causales, y por esto ellos se comprometen con lógica de primer orden, que al restringirnos al caso causal se queda corta pues los modelos gráficos probabilistas son más expresivos, además de modulares y montónicos
+Ellos utilizaron lógica de primer orden, pues están más interesados en hacer razonamientos \textit{de alto nivel} sobre el entorno, que involocura más que las relaciones causales, y por esto ellos se comprometen con lógica de primer orden, que al restringirnos al caso causal se queda corta pues los modelos gráficos probabilistas son más expresivos, además de modulares y montónicos.
 \subsubsection{Decisiones secuenciales}
@@
 De la misma manera, \cite{hagmayer2013repeated} mencionan que faltan estudios en donde las personas tengan otro objetivo más allá de revelar directamente la estructura causal del entorno y en los cuales vean directamente las consecuencias de sus acciones.
-Por lo tanto, es interesante preguntar si un agente que busque maximizar una recompensa pueda aprender modelos causales mediante la interacción con su entorno.
 
@@
 \item Dotar al jugador de creencias causales, ¿cómo especificarlas?
-\item Tengo un modelo causal \textit{verdadero}, ¿cómo lo incorporo en una política de RL? ¿cómo guío el descubrimiento de acciones? Aquí ayuda el artículo de \cite{lattimoreNIPS2016}
-\item Tengo una política aprendida, ¿cómo descubro estructura causal? Aquí puede ayudar el artículo de \cite{ortega2014generalized}
-\item ¿Qué aprendo primero?
-\item ¿Cómo le doy más peso a la información explicativa que a la información reactiva?
-\item Exploración vs explotación causal: ¿cómo decido si seguir la info causal o tomar acciones exploratorias para descubrir relaciones causales? 
-\item ¿Un dominio de interés? 
-\end{itemize}
+\item ¿cómo actualizar creencias causales?
+\item Diseño de lenguaje para ello
+\item Evaluación
+\end{itemize}
+\subsection{Ejemplo}
+Aquí iría con más detalle el ejemplo del paciente pero haciendo énfasis en el paso a paso para lo causal
 \subsection{Ejemplo: Caso de un MDP finito}

### 2018-03-05 Segundo_Semestre/Seminario/Semana6/State_of_Art_presentation.Rmd
presentacion final

@@
 output:
-  beamer_presentation: default
+  beamer_presentation:
+   includes:
+    in_header: header_pagenrs.tex
   ioslides_presentation: default
@@
 ## Problem
+- Acquiring and using causal knowledge in a sequential decision-making process.
 
 # State of The Art
+## Environment representations
+- In a series of papers, Seng-Beng Ho explains how to build a series of rules that encode causal relations found by exploration.
+- In particular, in [@ho2016ground] they show how to discover (effective) causal relations in a stream of events where chronological order is important (e.g. lightning -> thunder)
+- In [@ho2016reinforcement] they use the results of [@ho2014rapid] to build “scripts” that encode the local causal relations in order to choose actions.
+- It is not an on-line procedure.
 
-## Local updating
+----
 
-## Environment representations
+## Environment representations 2
+- On the other side, it is shown by [@garnelo2016towards] that an agent can build on-line representations of the environment.
+- They build a first-order logic representation which is then used to choose the next action.
+- Not mainly focused on causal learning, but can be realized upon this learned representation.
+- Small prototype designed ad-hoc.
+
+----
 
 ## Sequential decision making
+- A “bandit” is a slot machine in which you pull the lever and observe a reward.
+- Used to mode sequential decision making: choose an action, observe reward.
+- [@lattimoreNIPS2016] model how to choose inverventions in uncertain environment.
+- They model each possible intervention as arms of a bandit.
+- Fully specified causal model needed.
 
-## Bayesian games
+----
+
+## Sequential decision making 2
+- Thompson sampling: sample actions accordingly to (subjective) probability in action's optimality.
+- [@ortega2014generalized] proposes Thompson sampling for causal induction.
+- Agent samples parameter from beliefs, then chooses action as if parameter was true.
+- Intervention and Bayesian updating.
+- Not on-line
+
+----
+
+## Sequential decision making 2
+- [@innes2018reasoning] shows how to incorporate unseen possibilities via environment exploration and expert advise. 
+- Agent revises model of decision problem.
+- Single-stage decision making.
+- Extending results from [bramley2015staying] and [@albrecht2016exploiting] where set of variables was assumed to remain fixed.
+
+----
+
+## Bayesian Games and Belief updating
+- [@bjorndahl2016bayesian] difference between intended and actual strategies
+- [@mansour2015bayesian] & [mansour2016bayesian]: incentive-compatible exploration. When a “social planner” suggests actions.
+- [@larrouy2017mindreading] psychological (but rational) construction of beliefs.
+- [@amir2016endogenous] paying for extra information
+
+# Idea of evaluation
+## Max Expected Utility
+- For a fully specified decision problem under uncertainty and a preference relations that satisfies Savage's axioms there exists a formal criterion for acting: max expected utility.
+- If parameters are unknown, but agent has prior causal knowledge which will be updated in each round, the decisions used this learned model must be similar to MEU.
+- Bayesian games with incomplete information allow for this belief updating.
+
+----
+
+## Simple toy example
+- Patient come to a hospital and he can have eithe disease A or disease B.
+- Medic can choose between doing nothing (N), give some pill (p) or sugery (s).
+- If nothing is done, patient will die with probability 0.8.
+
+----
+
+- If medic choses surgery, patient can die, but if he survives and he had disease A, he will be cured with probability 0.5 or will not have any effect.
+- If he had disease B, surgery saves him with probability 0.6
+
+----
+
+- If pill is given, he can die of allergic reaction, but if he survives and had disease A the pill cures him with probability 0.9.
+- If he has disease B, pill has no effect.
+
+----
+
+- Now, suppose that this is a new disease, so the medic doesn't know exactly the probabilities.
+- But he has previous causal info, from med school (e.g. pills cure anything always).
+- He must update this info from experience in order to act as best as possible (bounded by MEU).
+- Forget about ethics, this is only a mental experiment.
+
+----

### 2018-03-08 Segundo_Semestre/Propuesta/Expected_Contribs.tex
ideas destiladas de planteamiento y contribucines esperadas

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{natbib}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Survey}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\tableofcontents
\newpage
\section{Introducción}
En múltiples ocasiones y contextos diversos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas. A lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto.\\
\\
La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
\\
En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros, pero puede \textit{aprenderlos} del mismo entorno. Un caso particular de esto es cuando el tomador de decisiones sabe con certeza que en el ambiente en el cual se llevan a cabo sus decisiones existen ciertos aspectos estructurales que pueden ser incorporados en la toma de decisiones; el caso en el cual nos vamos a centrar es en el caso en el cual existen relaciones causa-efecto en el ambiente.
\section{Motivación}
\section{Problema de investigación}
Adqusición y uso de información causal en problemas de toma de decisiónes secuenciales bajo incertidumbre.
\section{Planteamiento del Problema}
Supondremos que se conoce de antemano cuando un ambiente está regido por un modelo causal, y suponemos que este es desconocido, pero fijo.\\
\\
Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia. Que un modelo causal exista y opere sobre este ambiente implica que controla la relación entre los eventos inciertos, las causas así como las probabilidades de los eventos inciertos; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dado el evento incierto y las acciones y de esta manera se puede encontrar la acción (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.
\\
Cuando no se conoce el modelo causal, no se puede encontrar el camino a seguir de acuerdo a la maximización de utilidad esperada, y se podría intentar aprender los parámetros a partir de la información que provee el ambiente. Tenemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}). Lo que estas técnicas no toman en cuenta es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esto para aprender los parámetros.\\
\\
Para esto, queremos dotar al agente con \textit{creencias} sobre la estructura de su entorno; entonces, vamos a pensar que el agente está interactuando con otro ente abstracto (llamémosle Naturaleza), y esta interacción la vamos a modelar como un \textit{juego}.\\
\\
Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros. En particular, en esta área existen los llados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias a priori sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).\\
\\
En nuestro caso, vamos a pensar que se está llevando a cabo un juego (bayesiano) entre el tomador de decisiones y la Naturaleza, quien va a muestrear sus acciones a partir del modelo causal que opera en el entorno. Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo. Lo que esperamos tener al final de estas rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
\\
La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que de hecho generalizan los Proceso de Decisión Markovianos, pues un MDP es simplemente un juego estocástico con un solo jugador (\cite{shoham2008multiagent}).
\section{Contribuciones Esperadas}
Para llevar esto a cabo se necesita:
\begin{itemize}
\item Una forma de especificar creencias sobre estructura causal; esto puede ser mediante una distribución probabilista sobre un espacio de grafos. Podemos suponer que el número de variables es conocido de antemano, con lo cual podríamos limitarnos a trabajar sobre distribuciones de matrices de tamaño fijo. Si se permite que el número de variables sea arbitrario/desconocido, tendríamos que operar con modelos Bayesianos no paramétricos.
\item Un criterio de actualización de creencias causales que además debe cumplir la propiedad de que converge al verdadero modelo causal.
\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximan la utilidad máxima esperada.
\item Un diagrama de influencia causal que permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
\end{itemize}
\section{Ejemplo}
Supongamos que un paciente que llega a un hospital puede tener un padecimiento $A$ con probabilidad $p$ o $B$ con probabilidad $q$. El médico de guardia tiene a su alcance tres tratamientos posibles: cirugía, fármaco o nada. En caso de no hacer nada, el paciente morirá con una probabilidad $p_{dn}=0.8$ y sobrevivirá con una probabilidad $1-p_d$. Sin importar la enfermedad, el paciente puede morir durante la cirugía con probabilidad $p_{dc}=0.5$. De la misma manera, el medicamento producirá una reacción alérgica mortal con probabilidad $p_{df}=0.2$\\
\\
Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$. Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
\\
Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$. Si el paciente tenía la enfermedad $B$, entonces ésta lo curará con probabilidad de $0.6$. \qed
\\
\\
Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza. Las acciones disponibles al médico son el fármaco, cirugía, o nada, mientras que las jugadas de la Naturaleza son \textit{volados} que determinarán si el paciente vive o no; el juego va así, primero la naturaleza hace su jugada, lo cual determina qué enfermedad tenía el paciente. Después, el médico decide qué hacer, y después la Naturaleza lanzará una moneda cargada que determinará si el paciente sobrevive. Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc. 
\\
Sabemos que el médico fue a la escuela de medicina, de donde posee una gran cantidad de información (causal) a priori, pero puede que se esté enfrentando a una nueva enfermedad, es decir, no conoce realmente las probabilidades de muerte ni los casos en que el fármaco no actua. Lo que se desea es que a partir de este caso, o de varios más, pueda actualizar su conocimiento causal para la siguiente ocasión en la que deba enfrentarse a  un problema similar.

\bibliographystyle{apalike}
\bibliography{Bibliografia}
\end{document}

### 2018-03-12 Segundo_Semestre/Propuesta/Expected_Contribs.tex
algunas mejores de redaccion

@@
 \newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Survey}
+\title{Problemas de Decisión bajo Incertidumbre: el caso de estructura causal.}
 \author{Mauricio Gonzalez Soto}
@@
 \maketitle
-\tableofcontents
-\newpage
 \section{Introducción}
-En múltiples ocasiones y contextos diversos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas.
+En múltiples ocasiones y contextos diversos los seres humanos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas.
 A lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto.\\
@@
 \\
-En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros, pero puede \textit{aprenderlos} del mismo entorno.
-Un caso particular de esto es cuando el tomador de decisiones sabe con certeza que en el ambiente en el cual se llevan a cabo sus decisiones existen ciertos aspectos estructurales que pueden ser incorporados en la toma de decisiones; el caso en el cual nos vamos a centrar es en el caso en el cual existen relaciones causa-efecto en el ambiente.
-\section{Motivación}
+En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros, pero puede \textit{aprenderlos} del mismo entorno, sobre lo cual los trabajos de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}) muestarn cómo.
+Un caso particular de eventos inciertos es cuando el tomador de decisiones sabe con certeza que en el ambiente en el cual se llevan a cabo sus decisiones existen ciertos aspectos \textit{estructurales} que pueden ser incorporados en la toma de decisiones; el caso en el cual nos vamos a centrar es en el caso en el cual existen relaciones causa-efecto en el ambiente.\\
+\\
+Las relaciones causa-efecto son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated}).\\
 \section{Problema de investigación}
@@
 Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
-Que un modelo causal exista y opere sobre este ambiente implica que controla la relación entre los eventos inciertos, las causas así como las probabilidades de los eventos inciertos; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dado el evento incierto y las acciones y de esta manera se puede encontrar la acción (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.
+Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
+De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
 \\
-Cuando no se conoce el modelo causal, no se puede encontrar el camino a seguir de acuerdo a la maximización de utilidad esperada, y se podría intentar aprender los parámetros a partir de la información que provee el ambiente.
-Tenemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
+En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir.
+Entonces, se podría intentar aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
+Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
 Lo que estas técnicas no toman en cuenta es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esto para aprender los parámetros.\\
@@
 Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
-En particular, en esta área existen los llados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias a priori sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).\\
+En particular, en esta área existen los llamados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
+En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
 \\
-En nuestro caso, vamos a pensar que se está llevando a cabo un juego (bayesiano) entre el tomador de decisiones y la Naturaleza, quien va a muestrear sus acciones a partir del modelo causal que opera en el entorno.
+La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que de hecho generalizan los Proceso de Decisión Markovianos así como los juegos repetiso, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\\
+En nuestro caso, vamos a pensar que se está llevando a cabo un juego (Bayesiano) entre el tomador de decisiones y la Naturaleza, quien va a muestrear sus acciones a partir del modelo causal que opera en el entorno.
 Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
-Lo que esperamos tener al final de estas rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
+Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
+Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
 \\
-La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que de hecho generalizan los Proceso de Decisión Markovianos, pues un MDP es simplemente un juego estocástico con un solo jugador (\cite{shoham2008multiagent}).
+Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
+En este caso el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
+\section{Trabajo Previo}
+Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; como por ejemplo \cite{heckerman1995decision} quien argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
+\\
+Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada).
+El interés del autor es utilizar la representación de diagramas de influencia para \textit{interrogar} el fenómeno.
+En nuestro caso, el objetivo es \textit{construir} este diagrama mediante un ciclo de decisión-observación.\\
+\\
+\cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
+\\
+\cite{lattimoreNIPS2016} atacan un problema de toma de decisiones en entornos causales; en su modelo, las acciones a escoger son intervenciones posibles en el modelo y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
+La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado.
+Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente.\\
+\\
+\cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
+\\
+La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada.
+A lo más, los trabajos plantean el problema per se de inferencia causal \textit{como} un juego.
+En nuestro caso, lo que se propone es jugar un juego para el cual existan relaciones causales de fondo y utilizar lo que se pueda aprender de estas.
 \section{Contribuciones Esperadas}
-Para llevar esto a cabo se necesita:
+Para llevar a cabo lo anterior se necesita:
 \begin{itemize}
@@
 Si se permite que el número de variables sea arbitrario/desconocido, tendríamos que operar con modelos Bayesianos no paramétricos.
-\item Un criterio de actualización de creencias causales que además debe cumplir la propiedad de que converge al verdadero modelo causal.
-\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximan la utilidad máxima esperada.
-\item Un diagrama de influencia causal que permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
+\item Un criterio de actualización de creencias causales que además debe cumplir la propiedad de que converge al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
+\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
+\item Una manera de construir un diagrama de influencia causal que permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
 \end{itemize}
@@
 Después, el médico decide qué hacer, y después la Naturaleza lanzará una moneda cargada que determinará si el paciente sobrevive.
-Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc.
+Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc.\\
 \\
@@
 Lo que se desea es que a partir de este caso, o de varios más, pueda actualizar su conocimiento causal para la siguiente ocasión en la que deba enfrentarse a  un problema similar.
-
 \bibliographystyle{apalike}

### 2018-04-03 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
la version que mandé a sucar y hugo

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{natbib}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage[pdftex]{color,graphicx}
\usepackage{tikz}
\usetikzlibrary{calc}
%\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Problemas de Decisión bajo Incertidumbre: el caso de estructura causal.}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\section{Introducción}
En múltiples ocasiones y contextos diversos los seres humanos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}). A lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto.\\
\\
La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
\\
Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos, la interpretación de la probabilidad es un tema por sí solo interesante, y en este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas; por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida. Finalmente, el Teorema de Savage supone una serie de axiomas de coherencia que debe de cumplir una relación de preferencias y con ellos muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada. Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).\\
\\
En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}). Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente. Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.
\section{Motivación}
Las relaciones causa-efecto\footnote{Entendiendo causalidad como la define \cite{spirtes2000causation}; es decir, como una relación transitiva, irreflexiva y antisimétrica entre eventos, de manera que sucede uno y esto causa que otro evento sea producido.} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}). De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}). En particular, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es muy importante dotar agentes que tienen un comportamiento incoherente con incentivos para recuperar esa coherencia, lo cual se puede hacer si se conocen los \textit{efectos} de sus acciones (\cite{kleinberg2014time}). Además, la información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}). Por otro lado, \cite{lattimoreNIPS2016} muestran que contar con un modelo causal permite tomar decisiones de modo que se alcance cierto objetivo.\\
\\
Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”, pero estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal. \\
\\
Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros. En particular, en esta área existen los llamados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores). En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
\\
La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
\\
A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno. Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}). Además, se pretende operar a la luz de un modelo causal verdadero que rige lo que ocurre en el ambiente; pudiera ocurrir que una señal de recompensa sea construída de modo que oculte estas relaciones al agente, por eso se opta por observar lo que ocurre afuera como jugadas y dejar al agente escoger entre sus acciones y las consecuencias de éstas.
\section{Justificación}
El trabajo de \cite{lattimoreNIPS2016} muestra que conocer información causal sobre cierto entorno ayuda a tomar decisiones de modo que se logre un objetivo externo; en particular, ellos consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa. Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}). De manera similar, el modelo de \cite{lattimoreNIPS2016} concibe las acciones disponibles al agente como posibles \textit{intervenciones} en el entorno; la dinámica que ellos proponen consiste en que posterior a cada intervención, el agente observa el valor de la recompensa así como la realización de las otras variables que no fueron intervenidas; en esto consiste una \textit{ronda}. En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables. Esta dinámica de rondas repetidas lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (bandits en la literatura) y observar el pago. El algoritmo que desarrollan aprende la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente.\\
\\
Este trabajo muestra cómo a la luz de un modelo causal se pueden tomar decisiones utilizando esta información con el objetivo de maximizar cierta variable. La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}). La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.
\section{Problema de investigación}
Adqusición y uso de información causal en problemas de toma de decisiónes secuenciales bajo incertidumbre.
\section{Planteamiento del Problema}
Supondremos que se conoce de antemano cuando un ambiente está regido por un modelo causal, y suponemos que este es desconocido, pero fijo.\\
\\
Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia. Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto. De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
\\
En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir. Entonces, se podría intentar aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este. Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}). Lo que estas técnicas no toman en cuenta es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.\\
\\
Para esto, queremos dotar al agente con una manera de considerar \textit{creencias} sobre la estructura de su entorno las cuales se irán modificando conforme el agente interactue con éste. Para modelar la interacción, a diferencia de lo usual en aprendizaje por refuerzo en donde el agente está “solo con su entorno”, nosotros vamos a asumir que el agente está interactuando con otro ente abstracto (llamémosle Naturaleza), y esta interacción la vamos a modelar como un \textit{juego}.\\
\\
En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables. Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo. Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL. Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
\\
Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular. En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal. En este caso el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
\section{Trabajo Previo}
Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
\\
Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada). El interés del autor es utilizar la representación de diagramas de influencia para \textit{interrogar} el fenómeno. En nuestro caso, el objetivo es \textit{construir} este diagrama mediante un ciclo de decisión-observación.\\
\\
\cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
\\
\cite{board2006equivalence} define la racionalidad causal en juegos, concepto que toma en cuenta lo que se espera que otros jugadores hagan; es decir, toma en cuenta que las acciones de un jugador puedan tener efectos en las estrategias de los otros jugadores. Resulta que este concepto es equivalente a la racionalidad bayesiana estándar bajo el supuesto de independencia causal.\\
\\
\cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.\\
\\
Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal. La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado. Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente. En este trabajo se asume que el modelo causal es conocido y plantean como trabajo futuro el caso en el cual se desconoce el modelo, que es precisamente el caso que se busca atacar aquí.\\
\\
Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales. De manera similar, \cite{javdani2014near} utilizan teoría de la decisión Bayesiana para seleccionar experimentos que proporcionen la mayor información respecto a un fenómeno.\\
\\
\cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
\\
La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada. A lo más, los trabajos plantean el problema per se de inferencia causal \textit{como} un juego. Por ejemplo, \cite{heckerman1995decision}, \cite{eberhardt2008causal}  En nuestro caso, lo que se propone es jugar un juego para el cual existan relaciones causales de fondo y utilizar lo que se pueda aprender de estas.
\section{Objetivos}
Para llevar a cabo lo anterior se necesita desarrollar:
\begin{itemize}
\item Una forma de especificar creencias sobre estructura causal; esto puede ser mediante una distribución probabilista sobre un espacio de grafos. Podemos suponer que el número de variables es conocido de antemano, con lo cual podríamos limitarnos a trabajar sobre distribuciones de matrices de tamaño fijo. Si se permite que el número de variables sea arbitrario/desconocido, tendríamos que operar con modelos Bayesianos no paramétricos.
\item Un criterio de actualización de creencias causales que además debe cumplir la propiedad de que converge al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}). Es importante mencionar que los juegos Bayesianos, en su definición estándar, proveen un mecanismo para la actualización general de creencias, pero se necesita una manera de actualizar creencias que haga uso de la estructura causal misma.
\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
\item Una manera de construir un diagrama de influencia causal que además permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
\end{itemize}
\section{Ejemplo}
Supongamos que un paciente llega a un hospital debido a que tiene la enfermedad $A$ con probabilidad $p$ o la enfermedad $B$ con probabilidad $q$. El médico de guardia tiene a su alcance tres tratamientos posibles: cirugía, fármaco o nada. En caso de no hacer nada, el paciente morirá con una probabilidad $p_{dn}=0.8$ y sobrevivirá con una probabilidad $1-p_d$. Sin importar la enfermedad, el paciente puede morir durante la cirugía con probabilidad $p_{dc}=0.5$. De la misma manera, el medicamento producirá una reacción alérgica mortal con probabilidad $p_{df}=0.2$\\
\\
Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$. Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
\\
Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$. Si el paciente tenía la enfermedad $B$, entonces ésta lo curará con probabilidad de $0.6$. \qed
\\
\\
Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza. Las acciones disponibles al médico son:
\[ A= \{ \textrm{fármaco},  \textrm{cirugía}, \textrm{nada} \}. \]
Por otro lado, las jugadas de la Naturaleza son: 
\[ A= \{ \textrm{A, B, vive, muere} \}. \] 
El juego va así:
\begin{itemize}
\item Primero, la naturaleza hace su jugada, lo cual determina qué enfermedad tenía el paciente. 
\item Después, el médico decide qué hacer. 
\item Posteriormente, la Naturaleza lanzará una moneda cargada que determinará si el paciente sobrevive. Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc.
\end{itemize}
Sabemos que el médico fue a la escuela de medicina, de donde posee una gran cantidad de información (causal) a priori, pero puede que se esté enfrentando a una nueva enfermedad, es decir, no conoce realmente las probabilidades de muerte ni los casos en que el fármaco no actua.\\
\\
Lo que se desea es que a partir de este caso (juego), o de varios más, pueda actualizar su conocimiento causal para la siguiente ocasión en la que deba enfrentarse a  un problema similar.\\
\\
Notemos las similaridades que existen entre este ejemplo y el problema atacado por \cite{lattimoreNIPS2016}, quienes buscan escoger la \textit{mejor intervención} en un modelo causal dado con tal de maximizar cierta recompensa. En su modelo, un agente puede intervenir el modelo, y observar la recompensa y luego empezar desde cero; en nuestro caso, llega un paciente, el médico propone una acción, observa los efectos y procede con el siguiente paciente. El médico no conoce el modelo, que es la situación que \cite{lattimoreNIPS2016} dejan como trabajo futuro al final de su artículo. \\
\\
Para ser coherentes con el planteamiento de \cite{lattimoreNIPS2016} la actualización de creencias debe suceder al final de uno (o varios) juegos; además, el caso secuencial se atacará posteriormente.\\
\\
% Node styles
\tikzset{
% Two node styles for game trees: solid and hollow
solid node/.style={circle,draw,inner sep=1.5,fill=black},
hollow node/.style={circle,draw,inner sep=1.5}
}
\begin{tikzpicture}[scale=1.5,font=\footnotesize]
% Specify spacing for each level of the tree
\tikzstyle{level 1}=[level distance=15mm,sibling distance=35mm]
\tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]
% The Tree
\node(0)[solid node,label=above:{$N$}]{}
child{node(1)[solid node]{}
child{node[hollow node,label=below:{$(a,b)$}]{} edge from parent node[left]{$f$}}
child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$c$}}
child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
edge from parent node[left,xshift=-3]{$A$}
}
child{node(2)[solid node]{}
child{node[hollow node,label=below:{$(e,f)$}]{} edge from parent node[left]{$f$}}
child{node[hollow node,label=below:{$(g,h)$}]{} edge from parent node[right]{$c$}}
child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
edge from parent node[right,xshift=3]{$B$}
};
% information set
\draw[dashed,rounded corners=10]($(1) + (-.2,.25)$)rectangle($(2) +(.2,-.25)$);
% specify mover at 2nd information set
\node at ($(1)!.5!(2)$) {$M$};
\end{tikzpicture}
\section{Metodología}
Para cumplir con los objetivos planteados antes, se propone:
\begin{itemize}
\item Consideramos un problema de decisión bajo incertidumbre que se desarrolla en un entorno regido por un modelo causal. En una primera etapa, consideramos que el número de variables que están contenidas en ese modelo es finito y conocido para el tomador de decisiones.
\item Para este caso finito, se debe encontrar una manera para incorporar conocimiento causal inicial, o previo, el cual debe ser coherente con lo que se conoce, o no, sobre el problema, además de ser coherente con la manera en la que los seres humanos formamos creencias (\cite{larrouy2017mindreading}).
\item Este conocimiento causal inicial debe ser expresado en términos de probabilidades y para esto se construirá un \textit{prior} que consiste en una distribución sobre un espacio de matrices. Recordemos que una matriz (que cumple ciertas características) define un grafo, por lo que tener de esta manera estaremos especificando las probabilidades que un agente asigna, de manera subjetiva, a distintos modelos causales. 
\item De acuerdo a las características del problema de decisión se planteará un juego de dos jugadores, uno de los cuales será el tomador de decisiones inicial, y el otro será un ente abstracto llamado Naturaleza que, suponemos, conoce el modelo causal y lleva a cabo sus jugadas a partir de lo dictado por éste. Los juegos sucederán de manera repetida para poder aprender de éste.
\item Definir un número de juegos de observación, dentro de estos juegos el tomador de decisiones no modificará sus creencias causales, aunque éstas lo obliguen a tener un desempeño no óptimo.
\item Definir una estrategia para el tomador de decisiones a partir del conocimiento causal que tenga hasta ese momento; una \textit{estrategia mixta} permitiría explorar el ambiente con tal de explorar el ambiente de manera amplia.
\item A partir de lo observado en la ronda de juegos de observación, se debe actualizar las creencias causales de manera que se reflejen las relaciones que pudieron ser observadas, las cuales serán almacenadas en un diagrama de influencia.
\item Modificar la estrategia considerando los efectos posibles de cada acción que fueron aprendidos, para esto se usará el Diagrama de Influencia previamente construido. 
\item Comparar el desempeño versus lo que se haría si se conociera el modelo con detalle.
\item Extender al caso de número de variables no conocido.
\end{itemize}
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
\end{document}

### 2018-04-03 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
correcciones en la intro

@@
 \\
-Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos, la interpretación de la probabilidad es un tema por sí solo interesante, y en este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas; por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
+Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+En este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas.
+Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
 Finalmente, el Teorema de Savage supone una serie de axiomas de coherencia que debe de cumplir una relación de preferencias y con ellos muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.

### 2018-04-03 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
correcciones en la motivación

@@
 \\
-Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”, pero estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
+Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
 \\
@@
 A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
-Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}).
-Además, se pretende operar a la luz de un modelo causal verdadero que rige lo que ocurre en el ambiente; pudiera ocurrir que una señal de recompensa sea construída de modo que oculte estas relaciones al agente, por eso se opta por observar lo que ocurre afuera como jugadas y dejar al agente escoger entre sus acciones y las consecuencias de éstas.
+Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
+Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
 \section{Justificación}

### 2018-04-05 Segundo_Semestre/Seminario/Semana12/Proposal.tex
estructura del documento final

--- 2018-04-03 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
@@
 %\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Problemas de Decisión bajo Incertidumbre: el caso de estructura causal.}
+\title{Adquisición y uso de relaciones causales para la toma de decisiones bajo incertidumbre.}
 \author{Mauricio Gonzalez Soto}
@@
 \section{Introducción}
-En múltiples ocasiones y contextos diversos los seres humanos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
-A lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto.\\
-\\
-La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
-\\
-Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
-En este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas.
-Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
-Finalmente, el Teorema de Savage supone una serie de axiomas de coherencia que debe de cumplir una relación de preferencias y con ellos muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
-Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).\\
-\\
-En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}).
-Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente.
-Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.
-\section{Motivación}
-Las relaciones causa-efecto\footnote{Entendiendo causalidad como la define \cite{spirtes2000causation}; es decir, como una relación transitiva, irreflexiva y antisimétrica entre eventos, de manera que sucede uno y esto causa que otro evento sea producido.} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
-De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
-En particular, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es muy importante dotar agentes que tienen un comportamiento incoherente con incentivos para recuperar esa coherencia, lo cual se puede hacer si se conocen los \textit{efectos} de sus acciones (\cite{kleinberg2014time}).
-Además, la información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
-Por otro lado, \cite{lattimoreNIPS2016} muestran que contar con un modelo causal permite tomar decisiones de modo que se alcance cierto objetivo.\\
-\\
-Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
-Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
-\\
-\\
-Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
-En particular, en esta área existen los llamados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
-En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
-\\
-La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
-\\
-A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
-Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
-Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
-\section{Justificación}
-El trabajo de \cite{lattimoreNIPS2016} muestra que conocer información causal sobre cierto entorno ayuda a tomar decisiones de modo que se logre un objetivo externo; en particular, ellos consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa.
-Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
-De manera similar, el modelo de \cite{lattimoreNIPS2016} concibe las acciones disponibles al agente como posibles \textit{intervenciones} en el entorno; la dinámica que ellos proponen consiste en que posterior a cada intervención, el agente observa el valor de la recompensa así como la realización de las otras variables que no fueron intervenidas; en esto consiste una \textit{ronda}.
-En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (bandits en la literatura) y observar el pago.
-El algoritmo que desarrollan aprende la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente.\\
-\\
-Este trabajo muestra cómo a la luz de un modelo causal se pueden tomar decisiones utilizando esta información con el objetivo de maximizar cierta variable.
-La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
-La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.
-\section{Problema de investigación}
-Adqusición y uso de información causal en problemas de toma de decisiónes secuenciales bajo incertidumbre.
-\section{Planteamiento del Problema}
-Supondremos que se conoce de antemano cuando un ambiente está regido por un modelo causal, y suponemos que este es desconocido, pero fijo.\\
-\\
-Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
-Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
-De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
-\\
-En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir.
-Entonces, se podría intentar aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
-Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
-Lo que estas técnicas no toman en cuenta es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.\\
-\\
-Para esto, queremos dotar al agente con una manera de considerar \textit{creencias} sobre la estructura de su entorno las cuales se irán modificando conforme el agente interactue con éste.
-Para modelar la interacción, a diferencia de lo usual en aprendizaje por refuerzo en donde el agente está “solo con su entorno”, nosotros vamos a asumir que el agente está interactuando con otro ente abstracto (llamémosle Naturaleza), y esta interacción la vamos a modelar como un \textit{juego}.\\
-\\
-En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
-Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
-Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
-Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
-\\
-Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
-En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
-En este caso el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
-\section{Trabajo Previo}
-Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
-\\
-Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada).
-El interés del autor es utilizar la representación de diagramas de influencia para \textit{interrogar} el fenómeno.
-En nuestro caso, el objetivo es \textit{construir} este diagrama mediante un ciclo de decisión-observación.\\
-\\
-\cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
-\\
-\cite{board2006equivalence} define la racionalidad causal en juegos, concepto que toma en cuenta lo que se espera que otros jugadores hagan; es decir, toma en cuenta que las acciones de un jugador puedan tener efectos en las estrategias de los otros jugadores.
-Resulta que este concepto es equivalente a la racionalidad bayesiana estándar bajo el supuesto de independencia causal.\\
-\\
-\cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.\\
-\\
-Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
-La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado.
-Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente.
-En este trabajo se asume que el modelo causal es conocido y plantean como trabajo futuro el caso en el cual se desconoce el modelo, que es precisamente el caso que se busca atacar aquí.\\
-\\
-Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales.
-De manera similar, \cite{javdani2014near} utilizan teoría de la decisión Bayesiana para seleccionar experimentos que proporcionen la mayor información respecto a un fenómeno.\\
-\\
-\cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
-\\
-La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada.
-A lo más, los trabajos plantean el problema per se de inferencia causal \textit{como} un juego.
-Por ejemplo, \cite{heckerman1995decision}, \cite{eberhardt2008causal}  En nuestro caso, lo que se propone es jugar un juego para el cual existan relaciones causales de fondo y utilizar lo que se pueda aprender de estas.
+
+\section{Marco Teórico}
+\subsection{Relaciones Causales}
+\subsection{Toma de decisiones}
+\subsection{Teoría de Juegos}
+\subsection{Trabajo Previo}
 \section{Objetivos}
-Para llevar a cabo lo anterior se necesita desarrollar:
-\begin{itemize}
-\item Una forma de especificar creencias sobre estructura causal; esto puede ser mediante una distribución probabilista sobre un espacio de grafos.
-Podemos suponer que el número de variables es conocido de antemano, con lo cual podríamos limitarnos a trabajar sobre distribuciones de matrices de tamaño fijo.
-Si se permite que el número de variables sea arbitrario/desconocido, tendríamos que operar con modelos Bayesianos no paramétricos.
-\item Un criterio de actualización de creencias causales que además debe cumplir la propiedad de que converge al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
-Es importante mencionar que los juegos Bayesianos, en su definición estándar, proveen un mecanismo para la actualización general de creencias, pero se necesita una manera de actualizar creencias que haga uso de la estructura causal misma.
-\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
-\item Una manera de construir un diagrama de influencia causal que además permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
-\end{itemize}
-\section{Ejemplo}
-Supongamos que un paciente llega a un hospital debido a que tiene la enfermedad $A$ con probabilidad $p$ o la enfermedad $B$ con probabilidad $q$.
-El médico de guardia tiene a su alcance tres tratamientos posibles: cirugía, fármaco o nada.
-En caso de no hacer nada, el paciente morirá con una probabilidad $p_{dn}=0.8$ y sobrevivirá con una probabilidad $1-p_d$.
-Sin importar la enfermedad, el paciente puede morir durante la cirugía con probabilidad $p_{dc}=0.5$.
-De la misma manera, el medicamento producirá una reacción alérgica mortal con probabilidad $p_{df}=0.2$\\
-\\
-Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$.
-Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
-\\
-Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$.
-Si el paciente tenía la enfermedad $B$, entonces ésta lo curará con probabilidad de $0.6$.
-\qed
-\\
-\\
-Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza. Las acciones disponibles al médico son:
-\[ A= \{ \textrm{fármaco},  \textrm{cirugía}, \textrm{nada} \}. \]
-Por otro lado, las jugadas de la Naturaleza son: 
-\[ A= \{ \textrm{A, B, vive, muere} \}. \] 
-El juego va así:
-\begin{itemize}
-\item Primero, la naturaleza hace su jugada, lo cual determina qué enfermedad tenía el paciente. 
-\item Después, el médico decide qué hacer. 
-\item Posteriormente, la Naturaleza lanzará una moneda cargada que determinará si el paciente sobrevive.
-Notemos que, de manera latente, es un modelo causal el que determina que el fármaco no cura la enfermedad B, o que éste causa alergias, etc.
-\end{itemize}
-Sabemos que el médico fue a la escuela de medicina, de donde posee una gran cantidad de información (causal) a priori, pero puede que se esté enfrentando a una nueva enfermedad, es decir, no conoce realmente las probabilidades de muerte ni los casos en que el fármaco no actua.\\
-\\
-Lo que se desea es que a partir de este caso (juego), o de varios más, pueda actualizar su conocimiento causal para la siguiente ocasión en la que deba enfrentarse a  un problema similar.\\
-\\
-Notemos las similaridades que existen entre este ejemplo y el problema atacado por \cite{lattimoreNIPS2016}, quienes buscan escoger la \textit{mejor intervención} en un modelo causal dado con tal de maximizar cierta recompensa.
-En su modelo, un agente puede intervenir el modelo, y observar la recompensa y luego empezar desde cero; en nuestro caso, llega un paciente, el médico propone una acción, observa los efectos y procede con el siguiente paciente.
-El médico no conoce el modelo, que es la situación que \cite{lattimoreNIPS2016} dejan como trabajo futuro al final de su artículo.
-\\
-\\
-Para ser coherentes con el planteamiento de \cite{lattimoreNIPS2016} la actualización de creencias debe suceder al final de uno (o varios) juegos; además, el caso secuencial se atacará posteriormente.\\
-\\
-% Node styles
-\tikzset{
-% Two node styles for game trees: solid and hollow
-solid node/.style={circle,draw,inner sep=1.5,fill=black},
-hollow node/.style={circle,draw,inner sep=1.5}
-}
-\begin{tikzpicture}[scale=1.5,font=\footnotesize]
-% Specify spacing for each level of the tree
-\tikzstyle{level 1}=[level distance=15mm,sibling distance=35mm]
-\tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]
-% The Tree
-\node(0)[solid node,label=above:{$N$}]{}
-child{node(1)[solid node]{}
-child{node[hollow node,label=below:{$(a,b)$}]{} edge from parent node[left]{$f$}}
-child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$c$}}
-child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
-edge from parent node[left,xshift=-3]{$A$}
-}
-child{node(2)[solid node]{}
-child{node[hollow node,label=below:{$(e,f)$}]{} edge from parent node[left]{$f$}}
-child{node[hollow node,label=below:{$(g,h)$}]{} edge from parent node[right]{$c$}}
-child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
-edge from parent node[right,xshift=3]{$B$}
-};
-% information set
-\draw[dashed,rounded corners=10]($(1) + (-.2,.25)$)rectangle($(2) +(.2,-.25)$);
-% specify mover at 2nd information set
-\node at ($(1)!.5!(2)$) {$M$};
-\end{tikzpicture}
+
 \section{Metodología}
-Para cumplir con los objetivos planteados antes, se propone:
-\begin{itemize}
-\item Consideramos un problema de decisión bajo incertidumbre que se desarrolla en un entorno regido por un modelo causal.
-En una primera etapa, consideramos que el número de variables que están contenidas en ese modelo es finito y conocido para el tomador de decisiones.
-\item Para este caso finito, se debe encontrar una manera para incorporar conocimiento causal inicial, o previo, el cual debe ser coherente con lo que se conoce, o no, sobre el problema, además de ser coherente con la manera en la que los seres humanos formamos creencias (\cite{larrouy2017mindreading}).
-\item Este conocimiento causal inicial debe ser expresado en términos de probabilidades y para esto se construirá un \textit{prior} que consiste en una distribución sobre un espacio de matrices.
-Recordemos que una matriz (que cumple ciertas características) define un grafo, por lo que tener de esta manera estaremos especificando las probabilidades que un agente asigna, de manera subjetiva, a distintos modelos causales.
-\item De acuerdo a las características del problema de decisión se planteará un juego de dos jugadores, uno de los cuales será el tomador de decisiones inicial, y el otro será un ente abstracto llamado Naturaleza que, suponemos, conoce el modelo causal y lleva a cabo sus jugadas a partir de lo dictado por éste.
-Los juegos sucederán de manera repetida para poder aprender de éste.
-\item Definir un número de juegos de observación, dentro de estos juegos el tomador de decisiones no modificará sus creencias causales, aunque éstas lo obliguen a tener un desempeño no óptimo.
-\item Definir una estrategia para el tomador de decisiones a partir del conocimiento causal que tenga hasta ese momento; una \textit{estrategia mixta} permitiría explorar el ambiente con tal de explorar el ambiente de manera amplia.
-\item A partir de lo observado en la ronda de juegos de observación, se debe actualizar las creencias causales de manera que se reflejen las relaciones que pudieron ser observadas, las cuales serán almacenadas en un diagrama de influencia.
-\item Modificar la estrategia considerando los efectos posibles de cada acción que fueron aprendidos, para esto se usará el Diagrama de Influencia previamente construido. 
-\item Comparar el desempeño versus lo que se haría si se conociera el modelo con detalle.
-\item Extender al caso de número de variables no conocido.
-\end{itemize}
+
+\section{Resultados Preliminares}
+\subsection{Caso en el que se conoce el modelo completo}
+\subsection{Caso en el que se conoce la estructura del modelo}
+\subsection{Caso general}
+\section{Conclusiones}
+
 \bibliographystyle{apalike}

### 2018-04-05 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
añadiendo referencias sobre que los seres humanos utilizan info causal para mejorar decisiones

@@
 \\
-La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
+\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
 \\
-Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+\indent Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
 En este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas.
@@
 \\
-En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}).
+\indent En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}).
 Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente.
@@
 Además, la información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
-Por otro lado, \cite{lattimoreNIPS2016} muestran que contar con un modelo causal permite tomar decisiones de modo que se alcance cierto objetivo.\\
+Además, este conocimiento causal es utilizado de manera correcta por los seres humanos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
+Por otro lado, \cite{lattimoreNIPS2016} muestran que contar con un modelo causal permite encontrar de manera más rápida la acción óptima en un contexto incierto.\\
 \\
-Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+\indent Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
 Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
@@
 \\
-Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
+\indent Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
 En particular, en esta área existen los llamados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
@@
 \\
-La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\indent La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
 \\
-A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
+\indent A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
 Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
@@
 \section{Justificación}
-El trabajo de \cite{lattimoreNIPS2016} muestra que conocer información causal sobre cierto entorno ayuda a tomar decisiones de modo que se logre un objetivo externo; en particular, ellos consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa.
-Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
-De manera similar, el modelo de \cite{lattimoreNIPS2016} concibe las acciones disponibles al agente como posibles \textit{intervenciones} en el entorno; la dinámica que ellos proponen consiste en que posterior a cada intervención, el agente observa el valor de la recompensa así como la realización de las otras variables que no fueron intervenidas; en esto consiste una \textit{ronda}.
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo contar con un modelo causal del ambiente permite encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa.\\
+\\
+\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
+Siguiendo esta idea,  \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
 En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (bandits en la literatura) y observar el pago.
-El algoritmo que desarrollan aprende la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente.\\
+Esta dinámica de rondas repetidas lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura) y observar el pago.
+El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente.\\
 \\
-Este trabajo muestra cómo a la luz de un modelo causal se pueden tomar decisiones utilizando esta información con el objetivo de maximizar cierta variable.
+\indent Este trabajo muestra cómo a la luz de un modelo causal se pueden tomar decisiones utilizando esta información con el objetivo de maximizar cierta variable.
 La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
@@
 \\
-Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
+\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
 Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
@@
 \\
-En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir.
+\indent En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir.
 Entonces, se podría intentar aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
@@
 \\
-Para esto, queremos dotar al agente con una manera de considerar \textit{creencias} sobre la estructura de su entorno las cuales se irán modificando conforme el agente interactue con éste.
+\indent Para esto, queremos dotar al agente con una manera de considerar \textit{creencias} sobre la estructura de su entorno las cuales se irán modificando conforme el agente interactue con éste.
 Para modelar la interacción, a diferencia de lo usual en aprendizaje por refuerzo en donde el agente está “solo con su entorno”, nosotros vamos a asumir que el agente está interactuando con otro ente abstracto (llamémosle Naturaleza), y esta interacción la vamos a modelar como un \textit{juego}.\\
 \\
-En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
+\indent En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
 Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
@@
 \\
-Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
+\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
 En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
@@
 \\
-Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada).
+\indent Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada).
 El interés del autor es utilizar la representación de diagramas de influencia para \textit{interrogar} el fenómeno.
@@
 \\
-\cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
+\indent \cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
 \\
-\cite{board2006equivalence} define la racionalidad causal en juegos, concepto que toma en cuenta lo que se espera que otros jugadores hagan; es decir, toma en cuenta que las acciones de un jugador puedan tener efectos en las estrategias de los otros jugadores.
+\indent \cite{board2006equivalence} define la racionalidad causal en juegos, concepto que toma en cuenta lo que se espera que otros jugadores hagan; es decir, toma en cuenta que las acciones de un jugador puedan tener efectos en las estrategias de los otros jugadores.
 Resulta que este concepto es equivalente a la racionalidad bayesiana estándar bajo el supuesto de independencia causal.\\
 \\
-\cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.\\
+\indent \cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.\\
 \\
-Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
+\indent Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
 La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado.
@@
 \\
-Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales.
+\indent Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales.
 De manera similar, \cite{javdani2014near} utilizan teoría de la decisión Bayesiana para seleccionar experimentos que proporcionen la mayor información respecto a un fenómeno.\\
 \\
-\cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
+\indent \cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
 \\
-La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada.
+\indent La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada.
 A lo más, los trabajos plantean el problema per se de inferencia causal \textit{como} un juego.
@@
 \\
-Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$.
+\indent Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$.
 Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
 \\
-Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$.
+\indent Por otro lado, si el paciente sobrevive a la cirugía, ésta le curará con una probabilidad de $0.5$ si el paciente tenía la enfermedad $A$.
 Si el paciente tenía la enfermedad $B$, entonces ésta lo curará con probabilidad de $0.6$.
@@
 \\
-Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza. Las acciones disponibles al médico son:
+\indent Entonces, en este ejemplo, tenemos que el juego consiste en el médico, y en la Naturaleza. Las acciones disponibles al médico son:
 \[ A= \{ \textrm{fármaco},  \textrm{cirugía}, \textrm{nada} \}. \]
@@
 \[ A= \{ \textrm{A, B, vive, muere} \}. \] 
-El juego va así:
+El juego consiste en lo siguiente:
 \begin{itemize}

### 2018-04-05 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo estructura

@@
 	\subsection{Relaciones Causales}
+		\subsubsection{Definición de Causalidad}
+		\subsubsection{Modelos gráficos causales}
+		\subsubsection{El problema de la identificabilidad}
+		\subsubsection{Cálculo Do}
 	\subsection{Toma de decisiones}
@@
 	\subsection{Trabajo Previo}
+
 \section{Objetivos}
+	\section{Objetivos Generales}
+	\section{Objetivos específicos}
 
 \section{Metodología}
+	\subsection{Hipótesis}
+	\subsection{Metodología de Trabajo}
 
 \section{Resultados Preliminares}
+Para mostrar la factibilidad del modelo planteado, consideramos por separado tres casos cuyo grado de dificultad es ascendente.
+En el primer caso, supondremos que el modelo causal es conocido completamente por el agente y veremos cuál sería su manera de actuar en este caso.
+Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.
+Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este.
+En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo
 \subsection{Caso en el que se conoce el modelo completo}

### 2018-04-05 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo definición formal de causalidad y modelos gráficos causales y operador do

@@
 		\subsubsection{Definición de Causalidad}
+		La definición de Causalidad que estamos considerando es la que aparece en \cite{spirtes2000causation} y que consiste en lo siguiente:
+		\begin{defi}
+		Dado un espacio de probabilidad $(\Omega, \mathcal{F},\mathbb{P})$, consideramos una relación binaria $\to$ definida sobre elementos de $\mathcal{F}$ que sea
+		\begin{itemize}
+		\item Transitiva
+		\item Irefflexiva
+		\item Antisimétrica
+		\end{itemize}
+		\end{defi}
+		De esta manera, si $A \to B$ decimos que $A$ es una \textit{causa directa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
+Abusando de la notación, podemos construir un grafo que 	         represente las relaciones  causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$, entonces se tiene un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
+Además, requerimos las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad que éste expresa:
+		\begin{itemize}
+		\item Markov Causal
+		\item Minimalidad causal
+		\item Fidelidad Causal
+		\end{itemize}
 		\subsubsection{Modelos gráficos causales}
+		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a relaciones causales.
+Además, un operador $do()$ que está definido sobre grafos y cuya acción corresponde a lo siguiente: Dado $\mathbf{X} \subseteq \mathcal{X}$ y $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
+, x_{i_j} \} \in Val(\mathcal{X})$, $do(\mathbf{X} = \mathbf{x} )$ consiste en asignar a cada $X_j \in \mathbf{X}$ el valor $x_{x_{i_j}}$ y eliminar todas las aristas que entran al nodo correspondiente a $X_i$ en el grafo $\mathcal{G}$.
 		\subsubsection{El problema de la identificabilidad}

### 2018-04-05 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo teorema sobre calculo do

@@
 		\begin{itemize}
-		\item Transitiva
-		\item Irefflexiva
-		\item Antisimétrica
+		\item Transitiva: si $A \to B$ y $B \to C$ para $A,B,C \in \mathcal{F}$ entonces $A \to C$.
+		\item Irefflexiva: Para todo $A \in \mathcal{F}$ no es cierto que $A \to A$.
+		\item Antisimétrica: Para $A, B \in \mathcal{F}$, si $A \to B$, entonces no se cumple que $B \to A$.
 		\end{itemize}
@@
 		\item Markov Causal
-		\item Minimalidad causal
+		\item Minimalidad Causal
 		\item Fidelidad Causal
@@
 		\subsubsection{Cálculo Do}
+		\begin{teo}{\cite{peters2017elements}}
+		Los siguientes enunciados se cumplen:
+		\begin{itemize}
+		\item El cálculo do es completo; es decir, para cada intervención identificable existe una manera iterativa de aplicar las tres reglas que resulta en dicha intervención (\cite{huang2006pearl}, \cite{shpitser2006identification})
+		\item Existe un algoritmo que es capaz de encontrar todas las intervenciones identificables (\cite{tian2002}, \cite{huang2006pearl})
+		\item Existe un criterio necesario y suficiente para la identificabilidad de las distribuciones de intervención (\cite{shpitser2006identification}, \cite{huang2006pearl}).
+		\end{itemize}
+		\end{teo}
 	\subsection{Toma de decisiones}
+		\subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
+		\subsubsection{Teorema de DeFinetti: existencia de probabilidad subjetiva}
+		\subsubsection{Teorema de Savage: existencia de utilidad y probabilidad subjetiva}
 	\subsection{Teoría de Juegos}
+		\subsubsection{Juegos en forma normal}
+		\subsubsection{Juegos en forma extensiva}
+
 	\subsection{Trabajo Previo}

### 2018-04-09 Segundo_Semestre/Seminario/Semana12/Proposal.tex
sugerencia sobre el objetigo gral
definicion de causa directa usando refinamientos

@@
 		\end{defi}
-		De esta manera, si $A \to B$ decimos que $A$ es una \textit{causa directa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
-Abusando de la notación, podemos construir un grafo que 	         represente las relaciones  causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$, entonces se tiene un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
+		\subsubsection{Causas Directas e Indirectas}
+		En términos de la definición anterior, si $A \to B$ entonces decimos que $A$ es una \textit{causa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
+Consideremos ahora un \textit{refinamiento} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
+Decimos que $A$ es una \textit{causa directa} de $B$ si se cumple que $A \to B$ y no existe evento $C$ tal que $A \to C \to B$ para todo refinamiento de $\mathcal{F}$.
+Notemos que la definición depende del conjunto de variables, por lo que al decir que un evento es causa directa de otro, es siempre relativo al conjunto de variables.
+
+		\subsubsection{Construcción de un grafo que represente relaciones causales}
+		Abusando de la notación, podemos construir un grafo que represente las relaciones causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
 Además, requerimos las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad que éste expresa:
@@
 		\subsubsection{Modelos gráficos causales}
-		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a relaciones causales.
+		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a causas directas.
 Además, un operador $do()$ que está definido sobre grafos y cuya acción corresponde a lo siguiente: Dado $\mathbf{X} \subseteq \mathcal{X}$ y $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
@@
 		\subsubsection{El problema de la identificabilidad}
+
 		\subsubsection{Cálculo Do}
@@
 \section{Objetivos}
-	\section{Objetivos Generales}
+	\section{Objetivo General (debe ir relacionado con el título)}
 	\section{Objetivos específicos}
@@
 \subsection{Caso en el que se conoce el modelo completo}
+Básicamente lo que quiero es que para cada acción posible se consideren las probabilidades de los efectos y con esto calcular utilidad esperada
 \subsection{Caso en el que se conoce la estructura del modelo}
+Aquí se conocen los efectos de las causas, pero no sus probabilidades, entonces primero debo conocer el subconjunto posible de consecuencias y ahí tal vez hacer exploración greedy.
 \subsection{Caso general}
+No tengo idea aun.
 \section{Conclusiones}

### 2018-04-11 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo reglas de calculo do

@@
 Consideremos ahora un \textit{refinamiento} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
-Decimos que $A$ es una \textit{causa directa} de $B$ si se cumple que $A \to B$ y no existe evento $C$ tal que $A \to C \to B$ para todo refinamiento de $\mathcal{F}$.
+Decimos que $A$ es una \textit{causa directa} de $B$ si se cumple que $A \to B$ y no existe evento $C$ tal que $A \to C \to B$ para todo refinamiento propio de $\mathcal{F}$.
 Notemos que la definición depende del conjunto de variables, por lo que al decir que un evento es causa directa de otro, es siempre relativo al conjunto de variables.
@@
 		Abusando de la notación, podemos construir un grafo que represente las relaciones causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
-Además, requerimos las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad que éste expresa:
+A partir de definición de relación causal se obtiene un grafo acíclico dirigido.
+Además, requerimos las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad $P_{\mathcal{G}}$ que éste expresa:
 		\begin{itemize}
@@
 		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a causas directas.
+Las variables de $\mathcal{X}$ están divididas en dos subconjuntos: las variables endógenas y las variables exógenas, en donde estas últimas son aquellas que no tienen padres en el grafo.
 Además, un operador $do()$ que está definido sobre grafos y cuya acción corresponde a lo siguiente: Dado $\mathbf{X} \subseteq \mathcal{X}$ y $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
@@
 		\subsubsection{Cálculo Do}
-		\begin{teo}{\cite{peters2017elements}}
+		El cálculo do (\cite{pearl1995causal}, \cite{pearl2009causality}) consiste en un conjunto de reglas de inferencia que permiten manipular enunciados sobre observaciones que provienen de 		intervenciones.
+Consideremos un modelo gráfico causal $\mathcal{G}$ y sean $X,Y,Z$ conjuntos disjuntos de nodos de $\mathcal{G}$.
+Como notación, $\mathcal{G}_{\bar{X}}$ será el grafo que 	se obtiene al eliminar de $\mathcal{G}$ todas las aristas que entran en elementos de $X$; análogamente, $\mathcal{G}_{\underline{X}}$ será el grafo que se obtiene al eliminar de $\mathcal{G}$ todas las aristas que salen de elementos de $X$ y por último, $\mathcal{G}_{\underline{Z}\bar{X}}$ al eliminar arcos que entran de uno y que salen de otro conjunto.\\
+	\\
+		Las reglas de inferencia consisten en lo siguiente:
+		\begin{teo}
+		Sea $\mathcal{G}$ un modelo gráfico causal y $P_{\mathcal{G}}$ la medida de probabilidad inducida por el modelo; entonces, para cualesquiera conjuntos disjuntos de variables $X,Y,Z,W$ se cumple que
+		\begin{itemize}
+		\item Si en el grafo $\mathcal{G}_{\bar{X}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$, entonces
+		\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
+		\item Si en el grafo $\mathcal{G}_{\bar{X}\underline{Z}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$, entonces
+		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
+		\item Sea $Z(W)$ el conjunto de nodos en $Z$ tales que no son ancestros de ningún nodo en $W$ en el grafo $\mathcal{G}_{\bar{X}}$.
+Si en el grafo $G_{\bar{X}, \bar{Z(W)}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$ entonces,
+		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
+		\end{itemize}
+		\end{teo}
+		La interpretación de estos puntos, en palabras de \cite{pearl2009causality} se sigue intepretar la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
+El primer punto se obtiene al considerar que eliminar ecuaciones en un sistema no introduce nuevas dependencias y muestra que la d-separación es un criterio válido para la independencia condicional en el nuevo grafo $\mathcal{G}_{\bar{X}}$.
+El punto dos provee una condición para que una intervención tenga el mismo efecto que una observación pasiva y el tercer punto provee una condición para poder introducir, o eliminar una intervención externa sin afectar la probabilidad de observar $Y=y$.
+		\begin{teo}{\cite{peters2017elements}}\\
 		Los siguientes enunciados se cumplen:
@@
 		\subsubsection{Juegos en forma extensiva}
+		\subsubsection{Representaciones más generales}
 
 	\subsection{Trabajo Previo}
+
+\section{Planteamiento y Modelado}
 
@@
 Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este.
-En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo
+En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
+En resumen, en los tres casos el agente contará con \textit{creencias} sobre el modelo causal verdadero; en el primer caso, sus creencias coincidirán plenamente con el modelo; en el segundo caso, sólo con la estructura y en el tercer caso, no tenemos garantía de nada.
 \subsection{Caso en el que se conoce el modelo completo}
-Básicamente lo que quiero es que para cada acción posible se consideren las probabilidades de los efectos y con esto calcular utilidad esperada
+Consideremos el caso en el que el modelo gráfico causal es completamente conocido por el tomador de decisiones, tanto en su estructura como en sus parámetros.
+En este caso, el juego consistiría en lo siguiente:
+
 \subsection{Caso en el que se conoce la estructura del modelo}
@@
 
+\section{Extensiones posibles}
+Más oponentes...la naturaleza tiene intenciones,
+
 \bibliographystyle{apalike}

### 2018-04-12 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo teos de decision y el escenario de prueba

@@
 \section{Introducción}
+	\subsection{Motivación}
+	\subsection{Justificación}
 
@@
 		Las reglas de inferencia consisten en lo siguiente:
-		\begin{teo}
+		\begin{teo}{\label{docalculus}} (\cite{pearl2009causality})\\
 		Sea $\mathcal{G}$ un modelo gráfico causal y $P_{\mathcal{G}}$ la medida de probabilidad inducida por el modelo; entonces, para cualesquiera conjuntos disjuntos de variables $X,Y,Z,W$ se cumple que
@@
 		\end{teo}
+		Ademas
+		\begin{cor} Solución al problema de la identificabilidad
+		Una distribución $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ es identificable en un modelo gráfico causal $\mathcal{G}$ si existe una secuencia finita de transformaciones, donde cada una de las cuales corresponde a una de las reglas del teorema \ref{docalculus} que reduce $q$ a una expresión de probabilidad condicionada sólo en datos observacionales (i.e.
+no aparece el operador $do$)
+		\end{cor}
 	\subsection{Toma de decisiones}
 		\subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
+		Sea $X$ un conjunto de \textit{alternativas}, los objetos a elegir son loterías de soprte finito; de manera formal, definimos
+\[ L = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
+Una mezcla de loterías se puede definir de la siguiente manera: se toma una probabilidad $\alpha$ y se define una lotería $l^\alpha = \alpha P + (1- \alpha) Q$.\\
+\\
+Consideremos una relación  $\succeq L \times L$. El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva
+\item A2. Continuidad: Para toda $P,Q,R \in L$ tales que $P \succ Q \succ R$ existen $\alpha$ y $\beta$ tales que:
+\[  \alpha P + (1-\alpha) R \succ Q \succ \beta P + (1-\beta) R\]
+\item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si $\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R$.
+\end{itemize}
+\begin{teo}
+Una relación $\succeq L \times L$ satisface los axiomas A1 - A3 si y sólo sí existe una función $u: X \to \mathbb{R}$ tal que para toda $P,Q \in L$ se tiene que $P \succeq Q$ si y sólo si 
+\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
+\end{teo}
+Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
 		\subsubsection{Teorema de DeFinetti: existencia de probabilidad subjetiva}
+		De Finetti consideró apuestas sobre \textit{estados del mundo}.
+Consideremos que existen $1,...,n$ estados en el mundo.
+Una apuesta es una función que toma valores en el espacio de estados y los mapea a valores monetarios; es decir, una apuesta $x$ es un vector de $n$ elementos en donde $x_i$ es la ganancia en el caso de que la apuesta resulte en el estado $i$ como el ganador.
+Consideramos una relación $\succeq \subseteq X \times X$.
+Y consideremos los siguientes axiomas:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva.
+\item A2. Continuidad: Para cada $x \in X$, los conjuntos $\{ y | x \succ y \}$ y $\{ y | y \succ x \}$ son abiertos.
+\item A3. Aditividad: Para toda $x,y,z \in X$, $x \succeq y$ si y sólo si $x+z \succeq y+z$.
+\item A4 Monotonicidad: Para toda $x,y \in X$, si se cumple que $x_i \geq y_i$ entonces se tiene que $x \succeq y$.
+\item A5 No-trivialidad: Existen $x,y \in X$ tales que $x \succ y$.
+\end{itemize}
+Notemos que el axioma A3 sólo aplica si el tomador de decisiones es neutral al riesgo (\cite{gilboa2009decision}).
+\begin{teo}
+La relación $\succeq \subseteq X \times X$ satisface A1-A5 si y sólo si existe un vector de probabilidad $p$ de tamaño $n$ tal que para toda $x,y \in X$ se cumple que
+\[ x \succeq y \textrm { si y sólo si } px \geq py. \]
+\end{teo}
+Entonces, la definición de probabilidad subjetiva sería aquella probabilidad utilizada por un tomador de decisiones para escoger una acción mediante el criterio de utilidad esperado.
 		\subsubsection{Teorema de Savage: existencia de utilidad y probabilidad subjetiva}
@@
 \section{Resultados Preliminares}
-Para mostrar la factibilidad del modelo planteado, consideramos por separado tres casos cuyo grado de dificultad es ascendente.
+Para mostrar la factibilidad del modelo planteado, consideramos un escenario particular y tres casos por separado cuyo grado de dificultad es ascendente.
+El escenario es el de un médico que tiene que escoger uno de entre varios tratamientos para un paciente con enfermedad desconocida.
+Los tratamientos posibles conllevan riesgos y ventajas, que están regidos por un modelo causal.\\
+\\
+Los tres casos a considerar son los siguientes:\\
+\\
 En el primer caso, supondremos que el modelo causal es conocido completamente por el agente y veremos cuál sería su manera de actuar en este caso.
-Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.
-Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este.
+Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.\\
+\\ 
+Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este. \\
+\\
 En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
@@
 Consideremos el caso en el que el modelo gráfico causal es completamente conocido por el tomador de decisiones, tanto en su estructura como en sus parámetros.
-En este caso, el juego consistiría en lo siguiente:
+En este caso, el juego consistiría en lo siguiente:\\
+\begin{itemize}
+\item Un modelo gráfico causal conocido para el tomador de decisiones.
+\item Jugadores: Naturaleza y el Tomador de Decisiones.
+\item Acciones: Las acciones de la Naturaleza consisten en asignar cierto estado incierto; en este caso, asigna inicialmente la enfermedad del paciente; posteriormente, selecciona sus acciones a partir del modelo causal.
+\end{itemize}
 

### 2018-04-12 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
cambios redaccion y claridad para posterior copypaste en Proposal

@@
 \\
-\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, define un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
+\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
+Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
 \\
-\indent Para calcular la utilidad esperada, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
-En este contexto existen dos variantes: en el contexto del Teorema de Von Neuman-Morgenstern, si un tomador de decisiones conoce las probabilidades de los eventos, entonces existe una función de utilidad tal que la relación de preferencias del agente es equivalente con la utilidad esperada respecto a esta función y las probabilidades conocidas.
-Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
-Finalmente, el Teorema de Savage supone una serie de axiomas de coherencia que debe de cumplir una relación de preferencias y con ellos muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
+\indent Tal utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
+
+\indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
+De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ‚por dónde empezar?
+\\
+\indent El gran logro de Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
+Lo único que requiere el Teorema de Savage es una serie de axiomas de coherencia sobre la relación de preferencias de un tomador de decisiones; con esto, muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
 Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).\\
@@
 \section{Justificación}
-Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo contar con un modelo causal del ambiente permite encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
-\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa.\\
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
 \\
@@
 En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura) y observar el pago.
-El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente.\\
+Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura) y observar el pago.
+El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
 \\
@@
 La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
-La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.
+La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
+\\
+\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016}t iene sentido pues  existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
+
 \section{Problema de investigación}

### 2018-04-17 Segundo_Semestre/Seminario/Semana12/Proposal.tex
acabe los 3 teoremas de decision

--- 2018-04-11 Segundo_Semestre/Seminario/Semana12/Proposal.tex
@@
 \usepackage{tikz}
+\usetikzlibrary{trees}
 \usetikzlibrary{calc}
@@
 		\subsubsection{Definición de Causalidad}
-		La definición de Causalidad que estamos considerando es la que aparece en \cite{spirtes2000causation} y que consiste en lo siguiente:
+		La definición de Causalidad que estamos considerando, que está inspirada en la definición que da  \cite{spirtes2000causation} consiste en lo siguiente:
 		\begin{defi}
-		Dado un espacio de probabilidad $(\Omega, \mathcal{F},\mathbb{P})$, consideramos una relación binaria $\to$ definida sobre elementos de $\mathcal{F}$ que sea
+		Dado un espacio medible $(\Omega, \mathcal{F})$, consideramos una relación binaria $\to$ definida sobre elementos de $\mathcal{F}$ que sea
 		\begin{itemize}
@@
 		\subsubsection{Causas Directas e Indirectas}
-		En términos de la definición anterior, si $A \to B$ entonces decimos que $A$ es una \textit{causa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
+		En términos de la definición anterior, dado $(\Omega, \mathcal{F})$ si se tiene que $A \to B$ para $A,B \in \mathcal{F}$, entonces decimos que $A$ es una \textit{causa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
 Consideremos ahora un \textit{refinamiento} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
@@
 		Abusando de la notación, podemos construir un grafo que represente las relaciones causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
-A partir de definición de relación causal se obtiene un grafo acíclico dirigido.
-Además, requerimos las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad $P_{\mathcal{G}}$ que éste expresa:
+A partir de definición anterior de relación causal se obtiene un grafo acíclico dirigido.\\
+\\		
+		 Dado un grafo acíclico dirigido, podemos construir una medida de probabilidad que expresa las relaciones que se encuentran establecidas en el grafo, la cual denotaremos $P_{\mathcal{G}}$.
+Requerimos que se cumplan las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad $P_{\mathcal{G}}$ que éste expresa:
 		\begin{itemize}
@@
 		\subsubsection{Modelos gráficos causales}
-		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a causas directas.
-Las variables de $\mathcal{X}$ están divididas en dos subconjuntos: las variables endógenas y las variables exógenas, en donde estas últimas son aquellas que no tienen padres en el grafo.
-Además, un operador $do()$ que está definido sobre grafos y cuya acción corresponde a lo siguiente: Dado $\mathbf{X} \subseteq \mathcal{X}$ y $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
+		Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo acíclico dirigido $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a causas directas.
+Las variables de $\mathcal{X}$ están divididas en dos subconjuntos: las variables endógenas y las variables exógenas, en donde estas últimas son aquellas que no tienen padres en el grafo, esto se deriva naturalmente de la construcción del grafo a partir de las relaciones causales.
+Además, se tiene un operador $do()$ que está definido sobre grafos y cuya acción corresponde a lo siguiente: Dado $\mathbf{X} \subseteq \mathcal{X}$ y $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
 , x_{i_j} \} \in Val(\mathcal{X})$, $do(\mathbf{X} = \mathbf{x} )$ consiste en asignar a cada $X_j \in \mathbf{X}$ el valor $x_{x_{i_j}}$ y eliminar todas las aristas que entran al nodo correspondiente a $X_i$ en el grafo $\mathcal{G}$.
+Más adelante entraremos en detalle respecto a este operador.
 		\subsubsection{El problema de la identificabilidad}
@@
 		\end{teo}
+
 		La interpretación de estos puntos, en palabras de \cite{pearl2009causality} se sigue intepretar la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
@@
 		\end{teo}
+		Las reglas del cálculo do proveen una solución al problema de la identificabilidad:
+		\begin{cor} Solución al problema de la identificabilidad
+		Una distribución $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ es identificable en un modelo gráfico causal $\mathcal{G}$ si existe una secuencia finita de transformaciones, donde cada una de las cuales corresponde a una de las reglas del teorema \ref{docalculus} que reduce $q$ a una expresión de probabilidad condicionada sólo en datos observacionales (i.e.
+no aparece el operador $do$)
+		\end{cor}
 	\subsection{Toma de decisiones}
 		\subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
+		Sea $X$ un conjunto finito de \textit{alternativas} y consideremos una relación de preferencias $\succeq$ sobre $X$, definimos el conjunto de \textit{loterías} sobre elementos de $X$ de la siguiente manera:
+\[ \mathcal{L} = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
+Es decir, una lotería es una distribución de probabilidad sobre elementos de $X$.
+Para $P \in \mathcal{L}$ y $x \in X$.
+Denotamos $P(x)$ como la probabilidad que la lotería $P$ asigna a la alternativa $x$.
+Una forma útil de representar loterías es la siguiente (\cite{sucar2015probabilistic}, \cite{shoham2008multiagent}):
+
+\[ [p_1: x_1, ... , p_k : o_k]; \textrm{  }, x_i \in X, p_i \geq 0, \sum_{i=1}^k  p_i = 1.\]
+
+Podemos extender la relación de preferencias $\succeq$ que está definida sobre $X$ a $\mathcal{L}$ de modo que un tomador de decisiones puede escoger entre alternativas que son elementos de $X$ o loterías de $\mathcal{L}$.
+En particular, notemos que $X \subseteq \mathcal{L}$ pues un elemento $x \in X$ corresponde a la lotería $[1:x]$.
+Dicho esto, podemos considerar sólo relaciones de preferencia sobre $\mathcal{L}$.
+Para cada $P, Q \in \mathcal{L}$ y $\alpha \in [0,1]$ definimos la mezcla de loterías como
+\[ (\alpha P + (1- \alpha)Q)(x) = \alpha P(x) + (1-\alpha)Q(x). \]
+El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas sobre la relación de preferencias $\succeq$ que está definida sobre $\mathcal{L}$:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva
+\item A2. Continuidad: Para toda $P,Q,R \in L$ tales que $P \succ Q \succ R$ existen $\alpha$ y $\beta$ tales que:
+\[  \alpha P + (1-\alpha) R \succ Q \succ \beta P + (1-\beta) R\]
+\item A3. Independencia: Para toda lotería $P,Q,R \in L$ y toda $\alpha \in (0,1)$ se tiene que $P \succeq Q$ si y sólo si 
+\[\alpha P + (1-\alpha) R \succeq \alpha Q  + (1-\alpha) R.\]
+\end{itemize}
+\begin{teo}{según la formulación de \cite{jensen1967introduction}}\\
+Una relación de preferencias $\succeq$ definida sobre un conjunto $\mathcal{L}$ de loterías definidas sobre un conjunto finito de alternativas $X$ satisface los axiomas A1 - A3 si y sólo sí existe una función $u: X \to \mathbb{R}$ tal que para toda $P,Q \in \mathcal{L}$ se tiene que $P \succeq Q$ si y sólo si
+\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
+\end{teo}
+Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
 		\subsubsection{Teorema de DeFinetti: existencia de probabilidad subjetiva}
+		El Teorema de DeFinetti considera el caso opuesto: consideremos un conjunto de  $n$ \textit{estados}.
+Una \textit{apuesta} sobre este conjunto de estados es una función que mapea estados a un número real, y podemos identificar cada una de estas apuestas como un vector de tamaño $n$, cuya $i$-ésima entrada denota el valor que se asigna al estado $i$.
+Denotemos por $X$ el conjunto de apuestas sobre estados del mundo y consideramos una relación $\succeq$ definida sobre $X$.
+De la misma manera que el Teorema de Von Neumann, requerimos de axiomas sobre la relación:
+\begin{itemize}
+\item A1. Órden débil: $\succeq$ es completa y transitiva.
+\item A2. Continuidad: Para cada $x \in X$, los conjuntos $\{ y | x \succ y \}$ y $\{ y | y \succ x \}$ son abiertos en la topología estándar de $\mathbb{R}^n$.
+\item A3. Aditividad: Para toda $x,y,z \in X$, $x \succeq y$ si y sólo si $x+z \succeq y+z$.
+\item A4 Monotonicidad: Para toda $x,y \in X$, si se cumple que $x_i \geq y_i$ entonces se tiene que $x \succeq y$.
+\item A5 No-trivialidad: Existen $x,y \in X$ tales que $x \succ y$.
+\end{itemize}
+Notemos que el axioma A3 sólo aplica si el tomador de decisiones es neutral al riesgo (\cite{gilboa2009decision}).
+\begin{teo}{\cite{definetti1937}}\\
+La relación $\succeq \subseteq X \times X$ satisface A1-A5 si y sólo si existe un vector de probabilidad $p$ de tamaño $n$ tal que para toda $x,y \in X$ se cumple que
+\[ x \succeq y \textrm { si y sólo si } p^t x \geq p^t y. \]
+\end{teo}
+Este Teorema arroja una probabilidad para cada tomador de decisiones tal que la relación de preferencias del éste sea equivalente con la utilidad esperada bajo esta distrbución; por esto, a tal distribución se le conoce como \textit{probabilidad subjetiva}.
 		\subsubsection{Teorema de Savage: existencia de utilidad y probabilidad subjetiva}
+	    Los resultados anteriores parecen estar en polos opuestos, pues el Teorema de De Finetti provee una distribución de probabilidad para tomar decisiones con ella si el tomador de decisiones tiene una manera de medir su utilidad; por otro lado, el Teorema de Von Neumann supone que se conocen las probabilidades y otorga una forma de medir utilidad para tomar decisiones.
+El gran logro del Teorema de Savage es garantizar la existencia tanto de probabildad subjetiva como de una función de utilidad y además hacerlo en tándem (\cite{gilboa2009decision}).\\
+	    \\
+	    El Teorema de Savage requiere dos conceptos básicos: un conjunto de estados $S$ y un conjunto de resultados $X$ y las opciones a escoger en este contexto son \textit{actos}, que son funciones de los estados a los resultados:
+	    \[ F = X^S = \{ f | f: S \to X \}. \]
+	    Consideramos una relación binaria $\succeq$ definida sobre $F$.
+	    Diremos que un conjunto $A \subseteq S$ es un evento, y no se requieren suposiciones en cuanto a la medibilidad de estos.
+	    Consideremos la siguiente notación: para actos $f,g \in F$, y un evento $A \subseteq S$ definimos un acto $f_A^g$ como sigue:
+	    \[ f_A^g(s)=g(s) \textrm{ si  } s \in A, \textrm{  } f(s) \textrm{ si  } s \in A^c   \]
+	    Es decir, $ f_A^g$ es $g$ para valores en $A$. Con esta notación, será más fácil expresar los axiomas que este Teorema requiere:
+	    \begin{itemize}
+	    \item P1. Existen $f,g$ tales que $f \succeq g$
+	    \item P2. Órden débil: $\succeq$ es un órden débil.
+	    \item P3. Para toda $f,g,h,h' \in F$ y todo $A \subseteq S$,
+	    \[ f_{A^c}^h \succeq g_{A^c}^h  \textrm{ si y sólo si }  f_{A^c}^{h'}  \succeq g_{A^c}^{h'} .\]
+	    \item P4. Para toda $f \in F$, para $A \subseteq$ no-nulo y para $x,y \in X$,
+	    \[ x \succeq y \textrm{ si y sólo si } f_A^x \succeq f_A^y.  \]
+	    donde un evento nulo $A$ es aquel evento tal que la restricción de $\succeq$ a $A$ cumple que $f \sim_A g$ para toda $f,g \in F$.
+	    \item P5. Para todo $A,B \subseteq S$ y toda $x,y,z,w \in F$ tales que $x \succeq y$, $z \succeq w$, se cumple que
+	    \[ y_A^x \succeq y_B^x \textrm{ si y sólo si } w_A^z \succeq w_B^z.\]
+	    \item P6. Para toda $f,g,h \in F$ tales que $f \succ g$ existe una partición $\{ A_1, ... , A_n \}$ de $S$ tal que para toda $ i \leq n$,
+	    \[ f_{A_i}^h \succ g, \textrm{ y además } f \succ g_{A_i}^h. \]
+	    \item P7.
+Para toda $f,g \in F$ y eventos $A \subseteq S$, si para toda $s \in A$ se tiene que $f \succeq_A g(s)$, entonces $f \succeq_A g$ y si para toda $s \in A$, $g(s) \succeq_A f$ entonces $g \succeq_A f$
+	    \end{itemize}
+	    \begin{teo}{\cite{savage1954the}}\\
+	    Para un conjunto $X$ finito de resultados y un conjunto $S$ de estados.
+Una relación de preferencias $\succeq$ definida sobre $F=X^S$ sastisface los axiomas P1-P7 si y sólo si existe una medida $\mu$ definida sobre el espacio medible $(S, 2^S)$ que es no-atómica y finitamente aditiva, y existe una función no-constante $u :X \to \mathbb{R}$ tal que para toda $f,g \in F$,
+	    \[ f \succeq g \textrm{ si y sólo si } \int_S u(f(s)) d \mu(s) \geq \int_S u(g(s)) d \mu(s). \]
+	    \end{teo}
+	    El caso para conjuntos de resultados generales:
+	    \begin{teo}
+	    Para un conjunto $X$ (no necesariamente finito) de resultados y un conjunto $S$ de estados.
+Una relación de preferencias $\succeq$ definida sobre $F=X^S$ sastisface los axiomas P1-P7 si y sólo si existe una medida $\mu$ definida sobre el espacio medible $(S, 2^S)$ que es no-atómica y finitamente aditiva, y existe una función no-constante $u :X \to \mathbb{R}$ tal que para toda $f,g \in F$,
+	    \[ f \succeq g \textrm{ si y sólo si } \int_S u(f(s)) d \mu(s) \geq \int_S u(g(s)) d \mu(s). \]
+	    \end{teo}
 	\subsection{Teoría de Juegos}
-		\subsubsection{Juegos en forma normal}
-		\subsubsection{Juegos en forma extensiva}
+		\subsubsection{Comportamiento Racional}
+		\subsubsection{Juegos estratégicos y la forma normal}
+		\subsubsection{Juegos extensivos con información perfecta}
+		\subsubsection{Juegos extensivos con información imperfecta}
 		\subsubsection{Representaciones más generales}
 
-	\subsection{Trabajo Previo}
+\section{Pregunta de investigación}
+\section{Planteamiento y Modelado}
 
-\section{Planteamiento y Modelado}
+\section{Trabajo Previo}
 
@@
 	\subsection{Hipótesis}
+	Tentativa: La incorporación efectiva de información causal en un proceso de toma de decisiones bajo incertidumbre produce un mejora en el desempeño del tomador de decisiones.
+	Otra opción: Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
 	\subsection{Metodología de Trabajo}
@@
 \section{Resultados Preliminares}
-Para mostrar la factibilidad del modelo planteado, consideramos por separado tres casos cuyo grado de dificultad es ascendente.
+Para mostrar la factibilidad del modelo planteado, consideramos un escenario particular y tres casos por separado cuyo grado de dificultad es ascendente.
+El escenario es el de un médico que tiene que escoger uno de entre varios tratamientos para un paciente con enfermedad desconocida.
+Los tratamientos posibles conllevan riesgos y ventajas, que están regidos por un modelo causal.\\
+\\
+Los tres casos a considerar son los siguientes:\\
+\\
 En el primer caso, supondremos que el modelo causal es conocido completamente por el agente y veremos cuál sería su manera de actuar en este caso.
-Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.
-Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este.
+Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.\\
+\\ 
+Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este. \\
+\\
 En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
 En resumen, en los tres casos el agente contará con \textit{creencias} sobre el modelo causal verdadero; en el primer caso, sus creencias coincidirán plenamente con el modelo; en el segundo caso, sólo con la estructura y en el tercer caso, no tenemos garantía de nada.
+\subsection{Escenario}
+Consideremos un paciente que llega a un hospital y puede tener la enfermedad $A$ o la enfermedad $B$, lo cual es desconocido para el médico que lo atiende, éste tiene a su disposición tres opciones: mandar el paciente a cirugía, darle un fármaco, o no hacer nada, con lo cual el paciente morirá casi con certeza.
+Cada una de estas opciones entraña un riesgo: durante la cirugía el paciente puede morir, y al recibir el fármaco el paciente puede tener una reacción alérgica y morir.
+Aun si el paciente sobrevive, no está garantizado con certeza que el procedimiento lo cure, sino que esto depende de cuál enfermedad tenía.
+La cirugía es más probable que cure la enfermedad $A$ y el fármaco la $B$.
+Además, que el fármaco o la cirugía no tengan efecto es equivalente en términos de probabilidades a no hacer nada.
 \subsection{Caso en el que se conoce el modelo completo}
-Consideremos el caso en el que el modelo gráfico causal es completamente conocido por el tomador de decisiones, tanto en su estructura como en sus parámetros.
-En este caso, el juego consistiría en lo siguiente:
-
+Consideremos el caso en el que el modelo gráfico causal es completamente conocido por el tomador de decisiones (el médico en este caso), tanto en su estructura como en sus parámetros.
+En este caso, el juego consistiría en lo siguiente:\\
+\begin{itemize}
+\item Un modelo gráfico causal conocido para el médico.
+\item Jugadores: Naturaleza y el médico.
+\item Acciones: Las acciones de la Naturaleza consisten en asignar cierto estado incierto; en este caso, asigna inicialmente la enfermedad del paciente; posteriormente, selecciona sus acciones a partir del modelo causal.
+\end{itemize}
+Y el juego ocurre de la siguiente forma:
+\begin{itemize}
+\item Primero, la Naturaleza escoge la enfermedad del paciente, pero esto el médico no lo ve.
+\item Debido a que el médico no sabe qué enfermedad escogió la naturaleza, el médico se encuentra en un \textit{conjunto de información}; es decir, los nodos de decisión en los que tiene que escoger son indistinguibles.
+Aquí, debe escoger el médico qué acción llevar a cabo.
+\item Ahora, a partir de la elección del médico, la naturaleza muestrea del modelo causal su siguiente jugada, lo que dará la consecuencia de la acción del médico; es decir, ahora la Naturaleza escoge (a partir del modelo causal) si al paciente le da alergia y muere, si no le da alergia y se cura, si no le da alergia y muere, etc.
+\end{itemize}
+Debido a que el médico tiene a su disposición el modelo causal desde un inicio, puede consultar las probabilidades de $P( \textrm{vivir} | do(\textrm{procedimiento}))$ y escoger como estrategia el procedimiento que arroje la probabilidad más alta; de esta manera, el médico estaría escogiendo su \textit{mejor respuesta} a cualquiera de las jugadas de la naturaleza por lo que esta sería una solución de equilibrio.
+Notemos que para tener una solución de equilibrio también las jugadas de la naturaleza deben ser la mejor respuesta a cualquiera de las acciones posibles del médico,  pero vamos a darle a la naturaleza el mismo pago en todos los resultados posibles para no entrar en discusión de las \textit{intenciones} de la naturaleza.
+Además, darle a la Naturaleza utilidad en los pagos permite dejar espacio para posibles generalizaciones futuras; por ejemplo, en el caso en el que la Naturaleza sí tenga algún tipo de intenciones o de objetivos a cumplir.
+\\
+% Node styles
+\tikzset{
+% Two node styles for game trees: solid and hollow
+solid node/.style={circle,draw,inner sep=1.5,fill=black},
+hollow node/.style={circle,draw,inner sep=1.5}
+}
+\begin{tikzpicture}[scale=1.5,font=\footnotesize]
+% Specify spacing for each level of the tree
+\tikzstyle{level 1}=[level distance=15mm,sibling distance=40mm]
+\tikzstyle{level 2}=[level distance=15mm,sibling distance=15mm]
+\tikzstyle{level 3}=[level distance=15mm,sibling distance=11mm]
+% The Tree
+\node(0)[solid node,label=above:{$N$}]{}
+	child{node(1)[solid node]{}
+		child{node[hollow node,label=left:{$N$}]{} 
+			child{node[hollow node,label=below:{$m$}]{}} 
+			child{node[hollow node,label=below:{$v$}]{}}
+		edge from parent node[left]{$f$}}	
+		child{node[hollow node,label=below:{$N$}]{} 
+			child{node[hollow node,label=below:{$m$}]{}} 
+			child{node[hollow node,label=below:{$v$}]{}}
+		edge from parent node[right]{$c$}}
+		child{node[hollow node,label=below:{$N$}]{} 
+			child{node[hollow node,label=below:{$m$}]{}} 
+			child{node[hollow node,label=below:{$v$}]{}}		
+		edge from parent node[right]{$n$}}
+	edge from parent node[left,xshift=-3]{$A$}
+}
+child{node(2)[solid node]{}
+child{node[hollow node,label=below:{$(e,f)$}]{} edge from parent node[left]{$f$}}
+child{node[hollow node,label=below:{$(g,h)$}]{} edge from parent node[right]{$c$}}
+child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
+edge from parent node[right,xshift=3]{$B$}
+};
+% information set
+\draw[dashed,rounded corners=10]($(1) + (-.2,.25)$)rectangle($(2) +(.2,-.25)$);
+% specify mover at 2nd information set
+\node at ($(1)!.5!(2)$) {$M$};
+\end{tikzpicture}
 \subsection{Caso en el que se conoce la estructura del modelo}
@@
 \subsection{Caso general}
-No tengo idea aun.
+En el caso general en el cual el tomador de decisiones no conoce el modelo causal, lo dotaremos de \textit{creencias} sobre este; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiecia
+
+\section{Plan de trabajo}
+
 \section{Conclusiones}
 
+\section{Limitaciones}
+En este trabajo consideramos un tomador de decisiones que actúa racionalmente.
+
+Consideramos la definición de causalidad que aparece arriba y la consideramos de forma axiomática; es decir, no se pone en tela de juicio si los supuestos son realistas o no, ni es intención encontrar la definición de causalidad que mejor se adapte a una situación particular del mundo \textit{real}.
+
 \section{Extensiones posibles}
-Más oponentes...la naturaleza tiene intenciones,
+Más pasos dentro de un mismo juego, oponentes, intenciones de la naturaleza
 

### 2018-04-17 Segundo_Semestre/Seminario/Semana9/Objectives_and_Methodology.tex
añadiendo papers

@@
 \section{Introducción}
-En múltiples ocasiones y contextos diversos los seres humanos nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
-A lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto.\\
+Debido al enorme número de factores diversos que están involucrados en nuestro entorno, constántemente nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
+Esa misma realidad, altamente compleja, contiene una serie de relaciones causa-efecto que podemos utilizar al momento de tomar decisiones.
+El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) y predecir las consecuencias de nuestros actos.
+Por desgracia, es difícil que contemos con la información necesaria para llevar a cabo estas predicciones de manera correcta, a lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto dada cierta acción.\\
 \\
@@
 \\
-\indent Tal utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+\indent La utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
 El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
@@
 \\
-\indent El gran logro de Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
+\indent El gran logro de J.L.
+Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
 Lo único que requiere el Teorema de Savage es una serie de axiomas de coherencia sobre la relación de preferencias de un tomador de decisiones; con esto, muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
-Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).\\
+Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).
+\cite{gilboa2001theory} propone un nuevo paradigma de toma de decisiones en el cual los agentes toman decisiones haciendo analogías a eventos pasados en los cuales tuvieron un buen desempeño.
+Estas teorías, aunque interesantes, no serán abordadas en este trabajo sino que se mencionan sólo por completez.\\
 \\
@@
 Las relaciones causa-efecto\footnote{Entendiendo causalidad como la define \cite{spirtes2000causation}; es decir, como una relación transitiva, irreflexiva y antisimétrica entre eventos, de manera que sucede uno y esto causa que otro evento sea producido.} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
+Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
 De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
-En particular, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es muy importante dotar agentes que tienen un comportamiento incoherente con incentivos para recuperar esa coherencia, lo cual se puede hacer si se conocen los \textit{efectos} de sus acciones (\cite{kleinberg2014time}).
-Además, la información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
+Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
+\\
+\\
+\indent La información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
 Además, este conocimiento causal es utilizado de manera correcta por los seres humanos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
-Por otro lado, \cite{lattimoreNIPS2016} muestran que contar con un modelo causal permite encontrar de manera más rápida la acción óptima en un contexto incierto.\\
+Aunque, los seres humanos no son perfectos al estimar la fuerza de los efectos, o violan el supuesto Markoviano (\cite{rottman2014reasoning}).
+\\
+\\
+\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto.
+En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.\\
 \\
@@
 \indent Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
-En particular, en esta área existen los llamados \textit{juegos Bayesianos} en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
+En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos Extensivos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
 En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
 \\
-\indent La clase de juegos necesaria para modelar un proceso de decisiones secuenciales bajo incertidumbre y con información incompleta se llaman \textit{Juegos Estocásticos}, que fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores
 \\
-\indent A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
-Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
-Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
+\cite{tingley2010belief}
+
 \section{Justificación}
@@
 En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura) y observar el pago.
+Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
 El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
 \\
-\indent Este trabajo muestra cómo a la luz de un modelo causal se pueden tomar decisiones utilizando esta información con el objetivo de maximizar cierta variable.
+\indent Lo que este trabajo señala es que  la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
 La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
@@
 \\
-\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016}t iene sentido pues  existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
-Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
+\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.\\
+\\
+\indent A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
+Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
+Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
 
@@
 \section{Planteamiento del Problema}
-Supondremos que se conoce de antemano cuando un ambiente está regido por un modelo causal, y suponemos que este es desconocido, pero fijo.\\
+Suponemos un ambiente que está regido por un modelo causal, y suponemos que este es desconocido, pero fijo y además cumple que todas las variables se encuentran contenidas en él (suficiencia causal).\\
 \\
-\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
+\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}) es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
 Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
@@
 \\
-\indent En cambio, cuando no se conoce el modelo causal, no se puede utilizar la utilidad esperada para encontrar el camino a seguir.
-Entonces, se podría intentar aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
+\indent Cuando el modelo causal no es conocido por el tomador de decisiones, tampoco lo son las probabilidades de los efectos dadas las causas, por lo que no se puede calcular la utilidad esperada.
+Entonces, el camino que se propone es aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
 Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
-Lo que estas técnicas no toman en cuenta es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.\\
+Lo que estas técnicas ignoran es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.
+En este caso, la estructura es la que está dada por las relaciones causa-efecto en el ambiente.\\
 \\
-\indent Para esto, queremos dotar al agente con una manera de considerar \textit{creencias} sobre la estructura de su entorno las cuales se irán modificando conforme el agente interactue con éste.
-Para modelar la interacción, a diferencia de lo usual en aprendizaje por refuerzo en donde el agente está “solo con su entorno”, nosotros vamos a asumir que el agente está interactuando con otro ente abstracto (llamémosle Naturaleza), y esta interacción la vamos a modelar como un \textit{juego}.\\
+\indent Para llevar a cabo el proceso aprendizaje dotamos al agente con  \textit{creencias} sobre la estructura causal de su entorno las cuales se irán actualizando conforme el agente interactue con éste.
+Estas creencias pueden considerarse como un modelo causal \textit{propio} del agente, pues el agente utilizará este modelo causal para la toma de decisiones, que por construcción irá acercándose cada vez más al modelo causal verdadero.
+El problema de la actualización de creencias puede estudiarse en dos casos distintos: si el número de variables en el modelo causal es conocido por el agente, entonces podemos representar el modelo causal mediante matrices de tamaño fijo y llevar a cabo actualizaciones en las entradas de estas; pero si el número de variables es desconocido por el tomador de decisiones, un camino a seguir es utilizar un enfoque Bayesiano no-paramétrico y considerar distribuciones sobre un espacio de grafos; este problema parece tener una complejidad alta.
+Enfoques Bayesianos no-paramétricos para descubrimiento causal han sido explorados en \cite{karabatsos2012bayesian}.\\
+\\
+\indent Para modelar la interacción del tomador de decisiones con su entorno, vamos a asumir que el agente está interactuando con un ente abstracto al que le llamaremos Naturaleza y que la dinámica de interacción consiste en un \textit{juego}.
+A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.
+En el contexto de Teoría de Juegos, los juegos que permiten que el agente tenga creencias información incompleta se llaman \textit{Juegos Estocásticos}, los cuales fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
 \\
@@
 \section{Trabajo Previo}
+\subsection{Causalidad y Decisiones}
 Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
@@
 \\
-\indent \cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.\\
-\\
+\indent \cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.
+\subsection{Identificación de intervenciones óptimas}
 \indent Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
@@
 Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente.
+En el artículo consideran dos casos posibles: un caso en el cual hay $N$ causas, independientes entre sí, y una variable a la que estas causas afectan; en el caso general, en el cual las causas pueden afectarse entre sí, los autores utilizan un estimador por importancia truncado para la selección de acciones en cada ronda con un nivel fijo de truncamiento.
 En este trabajo se asume que el modelo causal es conocido y plantean como trabajo futuro el caso en el cual se desconoce el modelo, que es precisamente el caso que se busca atacar aquí.\\
 \\
+\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como bandits; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas, pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
+Este trabajo se encuentra en la intersección de inferencia causal y de descubrimiento de mejor brazo bajo restricciones presupuestales; en este sentido encontramos los trabajos de \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}.
+Por otro lado, en cuanto al aprendizaje de modelos causales a partir de datos o experimentos están los trabajos de \cite{eberhardt2008almost}, \cite{mooij2016distinguishing}, \cite{hyttinen2013experiment}, \cite{hauser2012two}, \cite{loh2014high}, \cite{shanmugam2015learning}.\\
+\\
+\indent El trabajo de \cite{sen2017identifying} considera el problema de la mejor de entre varias posibles intervenciones \textit{suaves} en un modelo causal, del cual sólo se conoce una parte.
+Después de $T$ rondas de intervención, se busca encontrar la mejor intervención sobre alguna una de las variables que está en la parte desconocida del modelo.
+Una gran diferencia de este trabajo respecto a los trabajos clásicos de identificación del mejor brazo como \cite{audibert2010best} es considerar el \textit{filtrado de información} entre brazos, pues gracias al modelo causal los valores muestreados de un brazo (intervención) pueden aportar información sobre los otros
+mejora a \cite{lattimoreNIPS2016} y a \cite{audibert2010best}.
+Uno de las mejores respecto al trabajo de \cite{lattimoreNIPS2016} es que el nivel de truncamiento es adaptativo en cada ronda de modo que balancea sesgo y varianza.
+Aunque los algoritmos de \cite{sen2017identifying} mejoran los de \cite{lattimoreNIPS2016} en el mismo caso, siguen requiriendo conocer las distribuciones condicinales y las marginales
+\subsection{Juegos, decisiones y actualización de creencias}
 \indent Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales.
@@
 Por ejemplo, \cite{heckerman1995decision}, \cite{eberhardt2008causal}  En nuestro caso, lo que se propone es jugar un juego para el cual existan relaciones causales de fondo y utilizar lo que se pueda aprender de estas.
+\\
+\cite{di2013predictive}
+\\
+\cite{tingley2010belief}
+\\
+\cite{10.1007/978-94-010-0189-2_25}
+\\
+\cite{harper1988causal}
 \section{Objetivos}
@@
 \section{Ejemplo}
-Supongamos que un paciente llega a un hospital debido a que tiene la enfermedad $A$ con probabilidad $p$ o la enfermedad $B$ con probabilidad $q$.
+Supongamos que un paciente que llega a un hospital puede tener la nfermedad $A$ con probabilidad $p_A$ o la enfermedad $B$ con probabilidad $p_B=1-p_A$.
 El médico de guardia tiene a su alcance tres tratamientos posibles: cirugía, fármaco o nada.
@@
 \indent Si el paciente tenía la enfermedad $A$, y no muere de una reacción alérgica al fármaco, entonces el fármaco le curará con una probabilidad de $p_{sf}=0.9$.
-Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá.\\
+Si el paciente tenía la enfermedad $B$, entonces el fármaco no le curará y morirá con la misma probabilidad que si no se hubiera hecho nada.\\
 \\
@@
 \end{tikzpicture}
-\section{Metodología}
+\section{Metodología de Trabajo}
 Para cumplir con los objetivos planteados antes, se propone:

### 2018-04-17 Segundo_Semestre/Seminario/Semana12/Proposal.tex
definicion de juego en forma normal

@@
 %\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Adquisición y uso de relaciones causales para la toma de decisiones bajo incertidumbre.}
+\title{Adquisición y uso de relaciones causales para la toma de decisiones bajo incertidumbre utilizando juegos con información imperfecta.}
 \author{Mauricio Gonzalez Soto}
@@
 \section{Marco Teórico}
+En esta sección se mencionarán las definiciones más importantes relativas a esta propuesta de investigación doctoral sin pretender ser exhaustivo.
+El contenido de estas secciones es para dar al lector certeza sobre la existencia y validez de los resultados que se pretenden utilizar.
 	\subsection{Relaciones Causales}
@@
 		\subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
+		El Teorema de Von Neumann-Morgenstern (\cite{von1944theory}) surge en el contexto de Teoría de Juegos, en donde otros resultados de los mismos autores exigen que un jugador busque maximizar su ganancia esperada, lo cual requería una justificación teórica (\cite{gilboa2009decision})
 		Sea $X$ un conjunto finito de \textit{alternativas} y consideremos una relación de preferencias $\succeq$ sobre $X$, definimos el conjunto de \textit{loterías} sobre elementos de $X$ de la siguiente manera:
@@
 	\subsection{Teoría de Juegos}
+	Teoría de Juegos es un área de las matemáticas cuyo objetivo es dar las herramientas necesarias para analizar la interacción estratégica de varios tomadores de decisiones (\cite{osborne1994course}).
+En esta sub-sección veremos varios tipos de juegos posibles.
+Por el momento, sólo veremos distintas clases de juegos para mostrar que el problema a tratar en esta propuesta cuenta con la maquinaria suficiente para ser modelado, pero se dejarán de lado otras cuestiones importantes de Teoría de Juegos como son el Equilibrio de Nash y las Estrategias Mixtas.
 		\subsubsection{Comportamiento Racional}
+		Vamos a asumir que los agentes involucrados tienen preferencias racionales en el sentido de los axiomas de Decisión que fueron mencionados.
 		\subsubsection{Juegos estratégicos y la forma normal}
+		Consideremos una situación en la que dos o más agentes interactúan, y donde cada uno de ellos busca lograr un objetivo.
+Cada agente tomará una y sólo una decisión y de manera simultánea a los demás, a esta clase de situaciones se le conoce como Juegos Estratégicos.
+		\begin{defi}{Juego estratégico en forma normal (\cite{osborne1994course}, \cite{shoham2008multiagent}) }\\
+		Un juego finito en forma normal es una tupla $(N,A,u)$ donde:
+		\begin{itemize}
+		\item $N$ es un conjunto finito de jugadores.
+		\item $A= A_1 \times \cdots \times A_n$ donde cada $A_i$ son las acciones disponibles para el jugador $i$. Cada elemento $a = (a_1,...,a_n) \in A$ es llamado un perfil de acción.
+		\item Para cada jugador $i \in N$, una relación de preferencias $\succeq_i$ definida sobre $A$.
+		\end{itemize}
+		\end{defi}
+		Notemos que por los Teoremas de Decisión que hemos visto, bajo ciertas condiciones podemos sustituir la relación de preferencias $\succeq_i$ por una función de utilidad $u_i$ para cada jugador donde $a \succeq_i b$ si y sólo si $u_i(a) \geq u_i(b)$.
+Notemos que podemos asociar directamente cada acción de cada jugador con un valor numérico (su utilidad y pago), lo cual permite representar el juego en forma matricial, a esta representación matricial se le conoce como la Forma Normal.
+\\
+		\\
+		Es deseable considerar el caso en el que un jugador no conoce todas las características del otro jugador; ya sea sus acciones o sus pagos, y es este desconocimiento lo que se pretende modelar con los Juegos Estratégicos Bayesianos.
+		\begin{defi}{Juego Estratégico Bayesiano en Forma Normal (\cite{osborne1994course})}\\
+		Un juego estratégico Bayesiano consiste en
+		\begin{itemize}
+		\item Un conjunto finito de jugadores $N$
+		\item Un conjunto de \textit{estados} $\Omega$
+		\item Para cada jugador: un conjuto de acciones posibles $A_i$
+		\item Para cada jugador: Un conjunto finito $T_i$ de señales que puede observar el jugador $i$
+		\item Una función de señal $\tau_i : \Omega \to T_i$
+		\item Para cada jugador: una medida de probabilidad $P_i$ sobre $\Omega$ tal que $P( \tau^{-1}_i (t_i)) >0$ para $t_i \in T_i$
+		\end{itemize}
+		\end{defi}
 		\subsubsection{Juegos extensivos con información perfecta}
-		\subsubsection{Juegos extensivos con información imperfecta}
-		\subsubsection{Representaciones más generales}
+		\subsubsection{Juegos extensivos con información incompleta}
 
@@
 \subsection{Caso general}
-En el caso general en el cual el tomador de decisiones no conoce el modelo causal, lo dotaremos de \textit{creencias} sobre este; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiecia
+En el caso general en el cual el tomador de decisiones no conoce el modelo causal, lo dotaremos de \textit{creencias} sobre éste; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
+Estas creencias tendrán la forma de una distribución de probabilidad sobre algún espacio; si suponemos que el tomador de decisiones conoce al menos el número de variables del modelo, entonces las creencias serán distribuciones de probabilidad sobre matrices de tamaño fijo; en cambio, de no conocer el tamaño del modelo causal, debemos recurrir a distribuciones de probabilidad que permitan representar objetos mucho más complejos.
+Una opción, es recurrir a distribuciones sobre espacios de dimensión infinita, como pudiera ser el espacio de los Grafos Acíclicos Dirigidos.
+Estas distribuciones sobre espacios de dimensión infinita se conocen en la literatura como \textit{priors Bayesianos no-paramétricos} y existe toda un área de la Estadística que los estudia, a saber, la Estadística Bayesiana No Paramétrica (\cite{ghosal2017fundamentals}).\\
+\\
+Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
+Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).
 
@@
 
+No estamos considerando que el ente que llamamos Naturaleza tenga intenciones, pero al darle \textit{pagos} en cada juego, aunque iguales, podríamos permitir que busque maximizar algún objetivo.
+
 \section{Extensiones posibles}

### 2018-04-18 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo intro motivacion y justificacion

@@
 \section{Introducción}
+Debido al enorme número de factores diversos que están involucrados en nuestro entorno, constántemente nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
+Esa misma realidad, altamente compleja, contiene una serie de relaciones causa-efecto que podemos utilizar al momento de tomar decisiones.
+El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) y predecir las consecuencias de nuestros actos.
+Por desgracia, es difícil que contemos con la información necesaria para llevar a cabo estas predicciones de manera correcta, a lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto dada cierta acción.\\
+\\
+\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
+Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
+\\
+\indent La utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
+
+\indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
+De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ‚por dónde empezar?
+\\
+\indent El gran logro de J.L.
+Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
+Lo único que requiere el Teorema de Savage es una serie de axiomas de coherencia sobre la relación de preferencias de un tomador de decisiones; con esto, muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
+Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).
+\cite{gilboa2001theory} propone un nuevo paradigma de toma de decisiones en el cual los agentes toman decisiones haciendo analogías a eventos pasados en los cuales tuvieron un buen desempeño.
+Estas teorías, aunque interesantes, no serán abordadas en este trabajo sino que se mencionan sólo por completez.\\
+\\
+\indent En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}).
+Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente.
+Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.
 	\subsection{Motivación}
+	Las relaciones causa-efecto\footnote{Entendiendo causalidad como la define \cite{spirtes2000causation}; es decir, como una relación transitiva, irreflexiva y antisimétrica entre eventos, de manera que sucede uno y esto causa que otro evento sea producido.} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
+Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
+De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
+Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
+\\
+\\
+\indent La información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
+Además, este conocimiento causal es utilizado de manera correcta por los seres humanos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
+Aunque, los seres humanos no son perfectos al estimar la fuerza de los efectos, o violan el supuesto Markoviano (\cite{rottman2014reasoning}).
+\\
+\\
+\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto.
+En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.\\
+\\
+\indent Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
+\\
+\\
+\indent Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
+En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos Extensivos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
+En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
+\\
+Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores
+\\
+\cite{tingley2010belief}
 \subsection{Justificación}
-
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
+\\
+\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
+Siguiendo esta idea,  \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
+En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
+Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
+El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
+\\
+\indent Lo que este trabajo señala es que  la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
+La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
+La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
+\\
+\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.\\
+\\
+\indent A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
+Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
+Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
 \section{Marco Teórico}
@@
 		\item Un conjunto finito de jugadores $N$
-		\item Un conjunto de \textit{estados} $\Omega$
-		\item Para cada jugador: un conjuto de acciones posibles $A_i$
-		\item Para cada jugador: Un conjunto finito $T_i$ de señales que puede observar el jugador $i$
-		\item Una función de señal $\tau_i : \Omega \to T_i$
-		\item Para cada jugador: una medida de probabilidad $P_i$ sobre $\Omega$ tal que $P( \tau^{-1}_i (t_i)) >0$ para $t_i \in T_i$
+		\item Un conjunto de \textit{estados} $\Omega$.
+		\item Para cada jugador: un conjuto de acciones posibles $A_i$, y como en el caso anterior definimos $A = A_1 \times \cdots \times A_n$.
+		\item Para cada jugador: Un conjunto finito $T_i$ de señales que puede observar el jugador $i$.
+		\item Una función de señal $\tau_i : \Omega \to T_i$.
+		\item Para cada jugador: una medida de probabilidad $P_i$ sobre $\Omega$ tal que $P( \tau^{-1}_i (t_i)) >0$ para $t_i \in T_i$.
+		\item Una relación de preferencias $\succeq_i$ definida sobre el conjunto de medidas de probabilidad sobre $A \times \Omega$
 		\end{itemize}
 		\end{defi}
+		La interpretación de esta clase de juegos es que el conjunto de estados $\Omega$ contiene las descripciones de las características relevantes de los jugadores, sobre lo cual cada jugador tiene \textit{creencias a priori} que están dados por la medida $P_i$.
+En cada jugada, se realiza un $\omega \in \Omega$, y cada jugador observa $\tau_i (\omega)$.
+Si un jugador recibe la señal $t_i \in T_i$, entonces deduce que el estado verdadero del mundo se encuentra en el conjunto $\tau^{-1}_i (t_i)$, por eso el requisito de asignar probabilidad mayor a cero.
+El jugador actualiza sus creencias sobre $\omega \in \Omega$ a $P_i(\omega) / P_i(\tau^{-1}_i (t_i))$ si $\omega \in \Omega$ y cero en otro caso.
+Respecto a este tipo de juegos y su aplicabilidad en el problema del uso de relaciones causales, basta decir que lo que se pretende es que al no conocer el modelo causal verdadero, las creencias del tomador de decisiones sean sobre este modelo causal, estas creencias serán utilizadas por el agente como un modelo \textit{propio} que irá actualizando conforme el mundo revela algo de sí.
 		\subsubsection{Juegos extensivos con información perfecta}

### 2018-04-18 Segundo_Semestre/Seminario/Semana12/Proposal.tex
añadiendo planteamiento y modelado

@@
 \section{Planteamiento y Modelado}
-
+Suponemos un ambiente que está regido por un modelo causal, y suponemos que este es desconocido, pero fijo y además cumple que todas las variables se encuentran contenidas en él (suficiencia causal).\\
+\\
+\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}) es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
+Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
+De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
+\\
+\indent Cuando el modelo causal no es conocido por el tomador de decisiones, tampoco lo son las probabilidades de los efectos dadas las causas, por lo que no se puede calcular la utilidad esperada.
+Entonces, el camino que se propone es aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
+Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
+Lo que estas técnicas ignoran es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.
+En este caso, la estructura es la que está dada por las relaciones causa-efecto en el ambiente.\\
+\\
+\indent Para llevar a cabo el proceso aprendizaje dotamos al agente con  \textit{creencias} sobre la estructura causal de su entorno las cuales se irán actualizando conforme el agente interactue con éste.
+Estas creencias pueden considerarse como un modelo causal \textit{propio} del agente, pues el agente utilizará este modelo causal para la toma de decisiones, que por construcción irá acercándose cada vez más al modelo causal verdadero.
+El problema de la actualización de creencias puede estudiarse en dos casos distintos: si el número de variables en el modelo causal es conocido por el agente, entonces podemos representar el modelo causal mediante matrices de tamaño fijo y llevar a cabo actualizaciones en las entradas de estas; pero si el número de variables es desconocido por el tomador de decisiones, un camino a seguir es utilizar un enfoque Bayesiano no-paramétrico y considerar distribuciones sobre un espacio de grafos; este problema parece tener una complejidad alta.
+Enfoques Bayesianos no-paramétricos para descubrimiento causal han sido explorados en \cite{karabatsos2012bayesian}.\\
+\\
+\indent Para modelar la interacción del tomador de decisiones con su entorno, vamos a asumir que el agente está interactuando con un ente abstracto al que le llamaremos Naturaleza y que la dinámica de interacción consiste en un \textit{juego}.
+A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.
+En el contexto de Teoría de Juegos, los juegos que permiten que el agente tenga creencias información incompleta se llaman \textit{Juegos Estocásticos}, los cuales fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\\
+\indent En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
+Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
+Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
+Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
+\\
+\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
+En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
+En este caso el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
 \section{Trabajo Previo}
@@
 \subsection{Caso general}
-En el caso general en el cual el tomador de decisiones no conoce el modelo causal, lo dotaremos de \textit{creencias} sobre éste; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
+En el caso general en el cual el tomador de decisiones no conoce el modelo causal, éste contará con una distribución de probabilidad sobre el modelo causal que represente las de \textit{creencias}, o conocimiento previo, que tiene el agente sobre el modelo; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
 Estas creencias tendrán la forma de una distribución de probabilidad sobre algún espacio; si suponemos que el tomador de decisiones conoce al menos el número de variables del modelo, entonces las creencias serán distribuciones de probabilidad sobre matrices de tamaño fijo; en cambio, de no conocer el tamaño del modelo causal, debemos recurrir a distribuciones de probabilidad que permitan representar objetos mucho más complejos.
@@
 \\
-Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
-Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).
+\indent Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
+Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).\\
+\\
+\indent Tomando esto en cuenta, una de las contribuciones esperadas de esta investigación es construir un procedimiento de actualización de creencias acerca de la estructura causal de cierto entorno sobre el cual un agente está actuando.
 

### 2018-04-18 Segundo_Semestre/Seminario/Semana12/Proposal.tex
objetivos generales y especificos

@@
 Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores
-\\
-\cite{tingley2010belief}
 \subsection{Justificación}
@@
 		\subsubsection{Juegos extensivos con información perfecta}
+		\subsubsection{Juegos extensivos con información impefecta}
+		\cite{10.1007/978-94-010-0189-2_25}
 		\subsubsection{Juegos extensivos con información incompleta}
@@
 \indent Cuando el modelo causal no es conocido por el tomador de decisiones, tampoco lo son las probabilidades de los efectos dadas las causas, por lo que no se puede calcular la utilidad esperada.
-Entonces, el camino que se propone es aprender los parámetros a partir de la información que provee el ambiente posterior a cada interacción con este.
+Por esto, lo que se propone es que el tomador de decisiones \textit{aprenda} los parámetros a partir de la información que provee el ambiente en cada interacción con este.
 Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
@@
 \section{Trabajo Previo}
-
+\subsection{Causalidad y Teoría de Decisión}
+Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
+\\
+\indent Por otro lado, \cite{dawid2002influence} explica cómo plantear un problema de inferencia causal utilizando diagramas de influencia, los cuales son una extensión de las redes Bayesianas para incluir nodos de decisión, y los cuales pueden utilizarse para descubrir las acciones óptimas en un problema de decisión tradicional (sí, aquella serie de acciones que obtiene la máxima utilidad esperada).
+El interés del autor es utilizar la representación de diagramas de influencia para \textit{interrogar} el fenómeno.
+En nuestro caso, el objetivo es \textit{construir} este diagrama mediante un ciclo de decisión-observación.\\
+\\
+\indent \cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
+\\
+\indent \cite{board2006equivalence} define la racionalidad causal en juegos, concepto que toma en cuenta lo que se espera que otros jugadores hagan; es decir, toma en cuenta que las acciones de un jugador puedan tener efectos en las estrategias de los otros jugadores.
+Resulta que este concepto es equivalente a la racionalidad bayesiana estándar bajo el supuesto de independencia causal.\\
+\\
+\cite{di2013predictive} muestran que el uso de contrafactuales no mejora las predicciones de un agente que ya conoce el conjunto de hipótesis posibles (logically omniscient).
+\\
+\indent \cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.
+\subsection{Identificación de intervenciones óptimas}
+\indent Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
+La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado.
+Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente.
+En el artículo consideran dos casos posibles: un caso en el cual hay $N$ causas, independientes entre sí, y una variable a la que estas causas afectan; en el caso general, en el cual las causas pueden afectarse entre sí, los autores utilizan un estimador por importancia truncado para la selección de acciones en cada ronda con un nivel fijo de truncamiento.
+En este trabajo se asume que el modelo causal es conocido y plantean como trabajo futuro el caso en el cual se desconoce el modelo, que es precisamente el caso que se busca atacar aquí.\\
+\\
+\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como bandits; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas, pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
+Este trabajo se encuentra en la intersección de inferencia causal y de descubrimiento de mejor brazo bajo restricciones presupuestales; en este sentido encontramos los trabajos de \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}.
+Por otro lado, en cuanto al aprendizaje de modelos causales a partir de datos o experimentos están los trabajos de \cite{eberhardt2008almost}, \cite{mooij2016distinguishing}, \cite{hyttinen2013experiment}, \cite{hauser2012two}, \cite{loh2014high}, \cite{shanmugam2015learning}.\\
+\\
+\indent El trabajo de \cite{sen2017identifying} considera el problema de la mejor de entre varias posibles intervenciones \textit{suaves} en un modelo causal, del cual sólo se conoce una parte.
+Después de $T$ rondas de intervención, se busca encontrar la mejor intervención sobre alguna una de las variables que está en la parte desconocida del modelo.
+Una gran diferencia de este trabajo respecto a los trabajos clásicos de identificación del mejor brazo como \cite{audibert2010best} es considerar el \textit{filtrado de información} entre brazos, pues gracias al modelo causal los valores muestreados de un brazo (intervención) pueden aportar información sobre los otros
+mejora a \cite{lattimoreNIPS2016} y a \cite{audibert2010best}.
+Uno de las mejores respecto al trabajo de \cite{lattimoreNIPS2016} es que el nivel de truncamiento es adaptativo en cada ronda de modo que balancea sesgo y varianza.
+Aunque los algoritmos de \cite{sen2017identifying} mejoran los de \cite{lattimoreNIPS2016} en el mismo caso, siguen requiriendo conocer las distribuciones condicinales y las marginales
+\subsection{Juegos, decisiones y actualización de creencias}
+\indent Modelar problemas de decisión como un juego de manera que se vaya aprendiendo a realizar una tarea ha sido explorado previamente en \cite{werling2015job}, quienes modelan un problema de clasificación \textit{on-line} como un juego, pero en el cual las acciones del tomador de decisiones consisten en utilizar un oráculo para clasificar las observaciones iniciales.
+De manera similar, \cite{javdani2014near} utilizan teoría de la decisión Bayesiana para seleccionar experimentos que proporcionen la mayor información respecto a un fenómeno.\\
+\\
+\indent \cite{larrouy2017mindreading} estudian la formación de creencias a partir de las acciones observadas y definen el concepto de \textit{mindreading}; es decir, poder adelantarse a las creencias que el otro jugador tomará en cuenta a la hora de tomar sus decisiones.\\
+\\
+\indent La relación entre Teoría de Juegos, toma de decisiones e incorporación de información adicional (causal) parece muy clara, aunque no ha sido particularmente explotada.
+A lo más, los trabajos plantean el problema per se de inferencia causal \textit{como} un juego.
+Por ejemplo, \cite{heckerman1995decision}, \cite{eberhardt2008causal}  En nuestro caso, lo que se propone es jugar un juego para el cual existan relaciones causales de fondo y utilizar lo que se pueda aprender de estas.\\
+\\
+\cite{tingley2010belief} muestran, experimentalmente y para una clase de juegos particular que los agentes involucrados actualizan sus creencias respecto a ciertos aspectos de los otros jugadores; encuentran que los jugadores actualizan sus creencias de manera que no se alejan mucho del \textit{prior} inicial.
 \section{Objetivos}
 	\section{Objetivo General (debe ir relacionado con el título)}
+	La investigación que se propone llevar a cabo tiene dos objetivos generales; por un lado, encontrar la manera en la que un agente que se enfrenta a un entorno incierto pueda usar información causal de manera que ésta sea de ayuda al llevar a cabo una tarea; por otro lado, encontrar la manera en la que un agente adquiera información causal conforme interactúa con su entorno.
 	\section{Objetivos específicos}
-
+	Para llevar satisfacer los objetivos generales, hasta el momento se ha identificado que será necesario lo siguiente:
+\begin{itemize}
+\item Una forma de especificar creencias sobre estructura causal; esto es, encontrar una familia de distribuciones que representen de manera coherente creencias o conocimiento previo acerca de la estructura causal de un entorno.
+Para esto, existen dos casos a considerar; el primero, es suponer que el tomador de decisiones conoce el número máximo de variables que el modelo causal puede tener; en este caso, las creencias del agente pueden estar representadas por una distribución sobre un espacio de matrices de tamaño fijo, aprovechando la relación que existe entre grafos acíclicos dirigidos y matrices.
+Por otro lado, si el agente desconoce el número máximo de variables, entonces el grafo causal puede tener cualquier tamaño.
+En este caso, el camino a explorar es el considerar distribuciones de probabilidad sobre espacios de dimensión infinita, las cuales son conocidas como priors Bayesianos no paramétricos.
+\item Una manera de \textit{utilizar} las creencias causales que el agente tiene en un momento dado.
+\item Un criterio de actualización de creencias causales que debe converger al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
+Es importante mencionar que los juegos Bayesianos, en su definición estándar, proveen un mecanismo para la actualización general de creencias, pero se necesita una manera de actualizar creencias que haga uso de la estructura causal misma.
+\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
+\item Una manera de construir un diagrama de influencia causal que además permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
+\end{itemize}
 \section{Metodología}
@@
 La cirugía es más probable que cure la enfermedad $A$ y el fármaco la $B$.
-Además, que el fármaco o la cirugía no tengan efecto es equivalente en términos de probabilidades a no hacer nada.
+Además, que el fármaco o la cirugía no tengan efecto es equivalente en términos de probabilidades a no hacer nada.\\
+\\
+\indent La tarea a la que se enfrente al médico es qué tratamiento ofrecer al paciente sólo con la información disponible en ese momento.
+El médico puede no estar seguro de la enfermedad que llevó al paciente al hospital, pero dado que ha estudiado muchos años, él tiene un conjunto de creencias y conocimiento previo sobre las enfermedades existentes así como su relación con diversos tratamientos y los riesgos de cada uno.\\
+\\
+\indent Debido a que más pacientes llegarán como este, el médico debe aprender de cada caso para desempeñarse mejor en los eventos futuros. 
 \subsection{Caso en el que se conoce el modelo completo}

### 2018-04-19 Segundo_Semestre/Seminario/Semana12/Beamer_Template.tex
la presentacion

\documentclass{beamer}

\mode<presentation> {

\usetheme{Madrid}

}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{natbib}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Main Topic } % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Mauricio Gonzalez Soto} % Your name
\institute[INAOE] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Instituto Nacional de Astrofísica Óptica y Electrónica \\ % Your institution for the title page
\medskip
\textit{mauricio@inaoep.mx} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Introduction} 
\begin{frame}
\frametitle{Introduction}
\begin{itemize}
\item Given the enormous complexity in our environment, we are constantly faced with making decision in uncertain environments.
\item Causal relations that hold in our environment allows to predict consequences.
\item “Causality” evokes certain regularities, we want to think about the causes of a disease in order to avoid it. 
\item Human beings tend to find causal narratives, sometimes excesively and that leads to incorrect judgements \cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
\item How to effectively use causal knowledge to make decisions?
\end{itemize}
\end{frame}
	\subsection{Motivation} 
	\begin{frame}
	\frametitle{Motivation}
	\begin{itemize}
	\item In decision making processes with a concrete goal to be achieved, causal knowledge is correctly used by human beings (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
	\item Although they sometimes violate some of the axioms required in the formalization of causality (\cite{rottman2014reasoning}).
	\item \cite{lattimoreNIPS2016} show that an autonomous agent can choose good actions if it uses causal information.
	\end{itemize}
	\end{frame}
	\subsection{Justification}
	\begin{frame}
	\frametitle{Justification}
	\begin{itemize}
	\item Human beings conceive their actions upon the world as \textit{interventions} (\cite{hagmayer2009decision})
	\item Following this idea, \cite{lattimoreNIPS2016} consider a decision making process where available actions are interventions in a causal graphical model. Faster identification of best-action.
	\item They require the full model to be known, how to acquire and use causal info on the fly? 
	\item This is, in part, how human beings work (\cite{hagmayer2013repeated}).
	\item Human beings learn local aspects and then unify them in a coherent story (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
	\end{itemize}
	\end{frame}
\section{Framework}
	\subsection{Causal Relations}
		\subsubsection{Definition of Causality}
		\begin{frame}
		\frametitle{So, what is causality?}
		\begin{itemize}
		\item Subject of heavy philosophical debate.
		\item Statisticians have some kind of fear about using this word.
		\item Economists try to find assumptions under an econometric model is causal.
		\item Many definitions exists, (see for example \cite{arlo2016readings}).
		\item Our attitude here is the expressed by \cite{danks2013functions}, \cite{danks2014unifying}:\\
		\textit{...all that matters is that our concept appropriately tracks some relation(s) in the world that can ground and support the learning, reasoning and inferences that we do with our causal knowledge...}
		\end{itemize}
		\end{frame}
		\begin{frame}
		\frametitle{Definition of Causality}
		The working definition of Causality that we will be using is based upon \cite{spirtes2000causation}
		\begin{defi}
		Let $(\Omega, \mathcal{F})$ a finite, measurable space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
		\begin{itemize}
		\item Transitive: if $A \to B$ and $B \to C$ for $A,B,C \in \mathcal{F}$ then $A \to C$.
		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
		\item Antisymetric: For $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
		\end{itemize}
		If $A \to B$ we will say that $A$ is a cause of $B$.
		\end{defi}
		\end{frame}
		\subsubsection{Direct and indirect causes}
		\begin{frame}
		\frametitle{Direct and Indirect Causes}
		Consider a refinement $\mathcal{F}_s$ of $\mathcal{F}$ and extend “$\to$” properly. Notice that it could happen that if $A \to B$, then for some $C_s \in \mathcal{F}$ we have that $A \to C_s \to B$. 
		\begin{defi}
		Under the same setup as the previous definition, we say that $A$ is a direct cause of $B$ if $A \to B$ and if it does not exist any $C$ such that $A \to C \to B$ for all proper refinements $\mathcal{F}_s$ of $\mathcal{F}$.
		\end{defi}
		\end{frame}
		\subsubsection{Causal Graph Construction}
		\begin{frame}
		\frametitle{From Causes to Graphs}
		Abusing notation, notice that we can build a graph that represent all of the direct causes between events that are members of $\mathcal{F}$.\\
		For each pair $A,B \in \mathcal{F}$ such that $A \to B$ and $A$ is a direct cause of $B$, define a node representing each variable, and a directed edge joining them.
		\begin{prop}
		Given a binary relation “$\to$” that satisfies previous requirement, let $\mathcal{G}$ be the directed graph that is obtained by the previous procedure. Then, $\mathcal{G}$ contains no cycles.
		\end{prop}
		\begin{proof}
		Consider a cycle $A \to C_1 \to C_2 \to \cdots \to A$, then we would have that $A \to A$, which contradicts the Irreflexive property.
		\end{proof}
		\end{frame}
		\begin{frame}
		\begin{obs}
		Notice that if the graph is finite, then necesarily there will be nodes without parents, which in the context of causal models will be called exogenous variables
		\end{obs}
		\end{frame}
		\subsubsection{Probability Measure built from the Graph}
		\begin{frame}
		\frametitle{from Graphs to Probability Measures}
		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph.
		An interesting question would be to study the relation of $P_\mathcal{G}$ and any other measure in the \textit{upper level} of the measurable space .
		\end{frame}
	\subsection{Causal Graphical Models}
	\begin{frame}
	\frametitle{Causal Graphical Models}
	\begin{defi}
	A causal graphical model consists of:
	\begin{itemize}
	 \item A set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, 
	 \item A directed acyclic graph $\mathcal{G}$
	 \item An operator $do()$ defined over the space of DAG's whose action consists of: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ... , x_{i_j} \} \in Val(\mathcal{X})$, then $do(\mathbf{X} = \mathbf{x} )$ assigns to each $X_j \in \mathbf{X}$ the value $x_{i_j}$ and deletes any incoming edge on it.
	 \end{itemize}
	 \end{defi}
	 \end{frame}
		\subsubsection{The identifiability issue} 
		\begin{frame}
		\frametitle{Identifiability}
		When can causal questions be answered using only observational data?
		\end{frame}
		\subsubsection{Do Calculus}
		\begin{frame}[allowframebreaks]
		\frametitle{Do calculus}
		The structure of the causal graph produces some equivalence rules between observational an interventional data.
		\begin{teo}{\label{docalculus}} (\cite{pearl2009causality})\\
		Sea $\mathcal{G}$ un modelo gráfico causal y $P_{\mathcal{G}}$ la medida de probabilidad inducida por el modelo; entonces, para cualesquiera conjuntos disjuntos de variables $X,Y,Z,W$ 	se cumple que
		\begin{itemize}
		\item If $Y$ is conditionally independent from $Z$ given $X$ y $W$ in $\mathcal{G}_{\bar{X}}$, then
		\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
		\item Si en el grafo $\mathcal{G}_{\bar{X}\underline{Z}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$, entonces
		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
		\item Sea $Z(W)$ el conjunto de nodos en $Z$ tales que no son ancestros de ningún nodo en $W$ en el grafo $\mathcal{G}_{\bar{X}}$. Si en el grafo $G_{\bar{X}, \bar{Z(W)}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$ entonces,
		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
		\end{itemize}
		\end{teo}
		\end{frame}
		\begin{frame}
		\frametitle{Some results}
		\begin{teo}{\cite{peters2017elements}}\\
 		The next statements hold:
		\begin{itemize}
		\item The do-calculus is complete; i.e., for all identifiable intervention there exists a sequential application of the rules that enables one to compute such intervention (\cite{huang2006pearl}, 				\cite{shpitser2006identification}).
		\item There is an algorithm capable of finding all identifiable interventional distributions (\cite{tian2002}, \cite{huang2006pearl}).
		\item There exits a necessary and sufficient graphical criterion for identifiability of interventional distributions (\cite{shpitser2006identification}, \cite{huang2006pearl}).
		\end{itemize}
		\end{teo}
		\end{frame}
		\begin{frame}
		\frametitle{Answering the identifiability problem}
		\begin{cor} 
		A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each of them corresponds to one of the simplification rules of Theorem \ref{docalculus} which reduces $q$ to a probability statement conditioned only on observations (i.e. the $do$ operator doesn't appear).
\end{cor}
		\end{frame}
	\subsection{Decision Theory}
	\begin{frame}
	\frametitle{Decision Theory: some classic results}
	\end{frame}
		\subsubsection{von Neumann-Morgenstern's Theorem}
		\begin{frame}
		\frametitle{von Neumann-Morgenstern's Theorem: existence of a utility function}
		Let $X$ be a finite set of outomes and consider $\mathcal{L}$ the set of lotteries over elements of $X$ and consider a preference relation $\succeq$ over $\mathcal{L}$.
		\begin{teo}
		A preference relation $\succeq$ over lotteries satisfies the (not shown here) axioms of rational choice if and only if there exists a function $u: X \to \mathbb{R}$ such that for all pair of lotteries $P,Q \in \mathcal{L}$ we have that $P \succeq Q$ if and only if 
		\[ \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x).  \]
		\end{teo}
		\end{frame}
		\subsubsection{De Finetti's Theorem}
		\begin{frame}
		\frametitle{De Finetti's Theorem: existence of subjective probability}
		Consider $n$ states of the world and consider $X$ the set of \textit{bets} over this set, where a bet is an $n$-dimensional vector whose $i$-th entry corresponds to the monetary amount that is being bet on state $n$. Consider a preference relationship $\succeq$ over the $X$. Also, some axioms are required over $\succeq$
		\begin{teo}
		The relationship $\succeq$ satisfies the axioms if an only if there exists a probability vector of size $n$ such that for all $x,y \in X$,
		\[ x \succeq y \textrm { if and only if } p^t x \geq p^t y. \]
		\end{teo}
		\end{frame}
		\subsubsection{Savage's Theorem}
		\begin{frame}
		\frametitle{Savage's Theorem}
		vNM Theorem requires probabilities and guarantees the existence of utility. DeFinetti's Theorem requires utility and guarantees the existence of probability...where do we start? Savage's Theorem gives both in tandem, but requires more abstract (and criticizable) axioms.
		\begin{teo}
		Consider a finite set $X$ of outcomes and a set $S$ of states.  A preference relationship $\succeq$ over $F=X^S$ satisfies the axioms if and only if there exists a finite non-atomic measure $\mu$ defined over the measurable space $(S, 2^S)$ and a non-constant function $u :X \to \mathbb{R}$ such that for all $f,g \in F$ it holds that
		\[ f \succeq g \textrm{ if and only if} \int_S u(f(s)) d \mu(s) \geq \int_S u(g(s)) d \mu(s). \]
		\end{teo}
		\end{frame}

	\subsection{Game Theory}
		\subsubsection{Games in normal form}
		\subsubsection{Games in extensive form}
\section{Research Question}
\begin{frame}
\frametitle{Research Question}
How can an intelligent agent acquire and use causal knowledge from his interaction with an uncertain environment in a decision making process?
\end{frame}

\section{Set up and modelling}
\begin{frame}[allowframebreaks]
\frametitle{Modelling}
\begin{itemize}
\item A (one-stage) decision problem under uncertainty consists in a set of actions, a set of consequences and a set of uncertain events.
\item The standard interpretation (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}) is that an agent will choose some action, then an uncertain event will occur and this will determine the outcome.
\item If all probabilities and utilities of outcomes are known, one can choose by maximizing MEU.

\end{itemize}
\end{frame}
\begin{frame}
\begin{itemize}
\item Now, consider an environment where the relation between the actions, the events, and the consequences is given by a causal model.
\item If this model is known by a decision maker, then it can be used for determining the probabilities of effects given causes (outcomes given actions).
\item If the model is not known, we want to learn it by interaction.
\end{itemize}
\end{frame}
\begin{frame}
\begin{itemize}
\item Interaction will be modelled as a \textit{game} between two players: the original decision maker, and a new abstract entity called Nature.
\item Nature will pick its actions from the causal model.
\end{itemize}
\end{frame}
\begin{frame}
\begin{itemize}
\item If the agent doesn't know the causal model, we must give him some way to reason about it.
\item The agent will have \textit{beliefs} about the causal model which will be used as a \textit{personal} causal model at any given time.
\item Beliefs will be updated from what is observed in each game.
\item Updating must be done in some \textit{consistent} (consistency is a statistical definition for estimators that converge to the true parameter.)
\end{itemize}
\end{frame}

\section{Previous work}
	\subsection{Causality and Decision Theory}
	\subsection{Optimal Interventions}
	\begin{frame}[allowframebreaks]
	\frametitle{Identifying Optimal Interventions}
	\begin{itemize}
		\item \cite{lattimoreNIPS2016} study the problem of decision making under uncertainty in causal models where the causal model is completely known.
		\item In their setup, the actions or choices available to dhe decision maker are interventions in the model.
		\item They look for the minimal-regret intervention that maximices some reward variable. 
		\item They show that optimal intervention can be found faster than if no causal info was considered. Open question of what to do when the causal graph is unknown
		\item Later, \cite{sen2017identifying} start with the same setting than in \cite{lattimoreNIPS2016} but they consider the case when not all of the graph is known. 
		\item They also find the optimal intervention over some variable in the unknown part of the model. They improve the results of \cite{lattimoreNIPS2016} in experiments, but they still require knowing conditional distributions.
	\end{itemize}
\end{frame}
\section{Objectives}
	\subsection{General Objective}
	\begin{frame}
	The proposed research has 2 general objectives: 
	\begin{itemize}
	\item Find the way in which an agent facing an uncertain environment (governed by a causal model) can use causal knowledge in order to make good choices and improve his performance when compared to non-causal ways of choosing.
	\item Find the way in which an agent can acquire causal information while interacting with an uncertain (governed by a causal model) environment.
	\end{itemize}
	\end{frame}
	\subsection{Specific Objectives}
	\begin{frame}[allowframebreaks]
	\frametitle{Specific Objectives}
	\begin{itemize}
\item To find a way to specify beliefs about causal structure; i.e., find (or build) a family of parametric distributions that represent in a coherent way beliefs or previous knowledge about the causal structure of some environment.
\item  We have to consider two cases: if the decision-maker (or the owner of the beliefs) knows the maximum number of variables that the causal model can have.
\item On the other side, if the agent doesn't know the max number of variables.
\item To find a way of effectively using causal knowledge that the agent has at a given moment.
\item Develop an update criterion for the beliefs of the agent based upon what he has observed from the environment while interacting. This updating must be achieved in a way that the sequence of models converges to the true model. In the Bayesian Non parametric setting this is known as \textit{consistency} (\cite{ghosal2017fundamentals}). Also, the beliefs obtained must be coherent with the decision making since we will be using those results. 
\item An aproximate soluton criteria: maybe we can't wait until convergence, so what can be done with the current knowledge.
\end{itemize}
	\end{frame}
\section{Methodology}
	\subsection{Hypothesis}
	\begin{frame}
	\frametitle{Hypothesis}
	Under certain conditions, an agent can acquire and use causal knowledge from interaction with an uncertain environment and effectively use it in a decision problem under uncertainty.
	\end{frame}
	\subsection{Working methodology}
\section{Preliminary Results}
\begin{frame}
\frametitle{Preliminary Results}

\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
\end{frame}

%------------------------------------------------

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

%----------------------------------------------------------------------------------------

\end{document}

### 2018-04-19 Segundo_Semestre/Seminario/Semana12/Proposal.tex
cambios menores

@@
 \\
+\indent En el lenguaje cotidiano, la palabra \textit{causa} evoca cierta regularidad o estabilidad del entorno; cuando hablamos de enfermedades existentes queremos pensar en las causas de esta para tomar acciones preventivas.
+De la misma manera, al estar aprendiendo a realizar cierta tarea prestamos atención a los efectos de alguna acción y asumimos que la acción causó el efecto y lo tendremos en mente para futuros escenarios similares.Tanto así, que los seres humanos formamos explicaciones o interpretaciones causales sin contar con toda la información disponible, lo cual puede llevar a errores sistemáticos al momento de tomar decisiones o emitir juicios (\cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
+La pregunta natural es cómo utilizar esta información causal de manera efectiva para que un agente inteligente tome decisiones de manera correcta.\\
+\\
 \indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
@@
 El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
-
+\\
 \indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
 De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ‚por dónde empezar?
+\\
 \\
@@
 Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente.
-Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.
+Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.\\
+\\
+\indent Aprender del entorno se formula usualmente como un agente autónomo que se encuentra dentro de un entorno o ambiente y que tomar acciones en este entorno modifican el \textit{estado} de éste (\cite{sutton1998reinforcement}).
+A su vez, tal modificación del entorno es percibida por el agente en términos de una señal de recompensa y de encontrarse en otro estado.
+En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un nuevo ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.
 \subsection{Motivación}
-	Las relaciones causa-efecto\footnote{Entendiendo causalidad como la define \cite{spirtes2000causation}; es decir, como una relación transitiva, irreflexiva y antisimétrica entre eventos, de manera que sucede uno y esto causa que otro evento sea producido.} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
+Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
 Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
@@
 \\
-Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores
+\indent Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.
 \subsection{Justificación}
@@
 	\subsection{Relaciones Causales}
+La palabra causa ha tenido múltiples interpretaciones a lo largo de la historia y presenta varias sutilezas filosóficas al momento de querer dar una definición formal.
+En estadística, un mantra muy común en cursos formativos es que \textit{correlación no es causalidad}, pero entonces ¿qué es?\\
+\\
+\indent Existen diversas formulaciones y definiciones de la causalidad; desde el punto de vista computacional llaman la atención la desarrollada por (\cite{pearl2009causality}) y la desarrollada por  (\cite{spirtes2000causation}).
+Aunque, más que preocuparnos por detalles filosóficos/epistemológicos (ver por ejemplo \cite{arlo2016readings}) respecto a cuál definición de causalidad es la mejor o más realista, en este trabajo vamos a proceder con la definición de causalidad que mejor se adapte a las necesidades concretas de este trabajo.
+La postura que estamos adoptando aquí es la que expresada por \cite{danks2013functions}, \cite{danks2014unifying}:\\
+\\
+\indent \textit{...all that matters is that our concept appropriately tracks some relation(s) in the world that can ground and support the learning, reasoning and inferences that we do with our causal knowledge...}
 		\subsubsection{Definición de Causalidad}
-		La definición de Causalidad que estamos considerando, que está inspirada en la definición que da  \cite{spirtes2000causation} consiste en lo siguiente:
-		\begin{defi}
-		Dado un espacio medible $(\Omega, \mathcal{F})$, consideramos una relación binaria $\to$ definida sobre elementos de $\mathcal{F}$ que sea
+		La definición de Causalidad que se propone está inspirada en la definición que aparece en \cite{spirtes2000causation}:
+		\begin{defi}{\label{defcausa}}
+		Sea $(\Omega, \mathcal{F})$ un espacio medible, finito, y consideramos una relación binaria $\to  \subseteq \mathcal{F} \times \mathcal{F}$ que cumple:
 		\begin{itemize}
@@
 		\end{itemize}
+	    Diremos que $A$ es una causa de $B$ si $A \to B$ para $A,B \in \mathcal{F}$.
 		\end{defi}
 		\subsubsection{Causas Directas e Indirectas}
-		En términos de la definición anterior, dado $(\Omega, \mathcal{F})$ si se tiene que $A \to B$ para $A,B \in \mathcal{F}$, entonces decimos que $A$ es una \textit{causa} de $B$ y que $B$ es el efecto de la ocurrencia de $A$.
 Consideremos ahora un \textit{refinamiento} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
@@
 Notemos que la definición depende del conjunto de variables, por lo que al decir que un evento es causa directa de otro, es siempre relativo al conjunto de variables.
+Diremos que la familia de conjuntos sobre la cual tenemos causas directas es un \textit{contexto}.
 
 \subsubsection{Construcción de un grafo que represente relaciones causales}
-		Abusando de la notación, podemos construir un grafo que represente las relaciones causales entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
-A partir de definición anterior de relación causal se obtiene un grafo acíclico dirigido.\\
-\\		
+Abusando de la notación, fijando un contexto, podemos construir un grafo que represente las relaciones (causales) entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto
+\begin{prop}
+Dada una relación causal “$\to$” en el mismo contexto que la Definición \ref{defcausa}.
+El grafo dirigido que se obtiene a partir de añadir nodos para las variables consideraras y arcos para las relaciones entre ellos cumple que es un grafo Acíclico.
+\end{prop}
+\begin{proof}
+La prueba se deriva trivialmente de los requerimientos de la relación binaria, pues si hubiera un camino $A \to B \to C \to A$, se tendría que $A \to A$ por la transitividad, lo cual violaría la irreflexividad.
+\end{proof}
+Notemos que al ser finito el grafo dirigido, existirán necesariamente variables que no tengan padres, las cuales serán conocidas como variables exógenas.
+\subsubsection{Una medida de probabilidad a partir del grafo}
 Dado un grafo acíclico dirigido, podemos construir una medida de probabilidad que expresa las relaciones que se encuentran establecidas en el grafo, la cual denotaremos $P_{\mathcal{G}}$.
@@
 \end{itemize}
-		\subsubsection{Modelos gráficos causales}
+Estos requerimientos vamos a tratarlos de manera axiomática, en el sentido de que se dan por sentado sin cuestionarlos.
+Quien no esté de acuerdo con los axiomas, no lo estará con las conclusiones y queda siempre la invitación de desarrollar la teoría utilizando otras altenativas.
+\subsection{Modelos gráficos causales}
 Un modelo gráfico causal consiste en un conjunto de variables aleatorias $\mathcal{X}=\{ X_1,...,X_n \}$, un grafo acíclico dirigido $\mathcal{G}$ cuyos nodos corresponden a variables en $\mathcal{X}$ y las aristas entre ellos a causas directas.
@@
 \subsubsection{El problema de la identificabilidad} 
-
+¿Bajo qué condiciones es posible contestar preguntas causales en términos de datos observados?
+Se ha demostrado que si se cumple la condición (axioma) de Markov y de Fidelidad caual, entonces el grafo causal puede ser identificado hasta equivalencias Markovianas (arcos sin dirección) \cite{peters2012identifiability}, \cite{mooij2016distinguishing}.
 \subsubsection{Cálculo Do}
@@
 \end{teo}
-
-		La interpretación de estos puntos, en palabras de \cite{pearl2009causality} se sigue intepretar la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
+\indent La interpretación de estos puntos, en palabras de \cite{pearl2009causality} se sigue intepretar la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
 El primer punto se obtiene al considerar que eliminar ecuaciones en un sistema no introduce nuevas dependencias y muestra que la d-separación es un criterio válido para la independencia condicional en el nuevo grafo $\mathcal{G}_{\bar{X}}$.
@@
 \subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
-		El Teorema de Von Neumann-Morgenstern (\cite{von1944theory}) surge en el contexto de Teoría de Juegos, en donde otros resultados de los mismos autores exigen que un jugador busque maximizar su ganancia esperada, lo cual requería una justificación teórica (\cite{gilboa2009decision})
-		Sea $X$ un conjunto finito de \textit{alternativas} y consideremos una relación de preferencias $\succeq$ sobre $X$, definimos el conjunto de \textit{loterías} sobre elementos de $X$ de la siguiente manera:
+El Teorema de Von Neumann-Morgenstern (\cite{von1944theory}) surge en el contexto de Teoría de Juegos, en donde otros resultados de los mismos autores exigen que un jugador busque maximizar su ganancia esperada, lo cual requería una justificación teórica (\cite{gilboa2009decision})\\
+\\
+\indent Sea $X$ un conjunto finito de \textit{alternativas} y consideremos una relación de preferencias $\succeq$ sobre $X$, definimos el conjunto de \textit{loterías} sobre elementos de $X$ de la siguiente manera:
 \[ \mathcal{L} = \{ P:X \to [0,1] | \sum_{x \in X} P(x) = 1, \textrm{ $P$ tiene soporte finito} \}. \]
-Es decir, una lotería es una distribución de probabilidad sobre elementos de $X$.
+\indent Es decir, una lotería es una distribución de probabilidad sobre elementos de $X$.
 Para $P \in \mathcal{L}$ y $x \in X$.
@@
 Una forma útil de representar loterías es la siguiente (\cite{sucar2015probabilistic}, \cite{shoham2008multiagent}):
-
 \[ [p_1: x_1, ... , p_k : o_k]; \textrm{  }, x_i \in X, p_i \geq 0, \sum_{i=1}^k  p_i = 1.\]
-
-Podemos extender la relación de preferencias $\succeq$ que está definida sobre $X$ a $\mathcal{L}$ de modo que un tomador de decisiones puede escoger entre alternativas que son elementos de $X$ o loterías de $\mathcal{L}$.
+\indent Podemos extender la relación de preferencias $\succeq$ que está definida sobre $X$ a $\mathcal{L}$ de modo que un tomador de decisiones puede escoger entre alternativas que son elementos de $X$ o loterías de $\mathcal{L}$.
 En particular, notemos que $X \subseteq \mathcal{L}$ pues un elemento $x \in X$ corresponde a la lotería $[1:x]$.
@@
 \[ (\alpha P + (1- \alpha)Q)(x) = \alpha P(x) + (1-\alpha)Q(x). \]
-El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas sobre la relación de preferencias $\succeq$ que está definida sobre $\mathcal{L}$:
+\indent El Teorema de von Neumann-Morgenstern requiere de los siguientes axiomas sobre la relación de preferencias $\succeq$ que está definida sobre elementos de $\mathcal{L}$:
 \begin{itemize}
@@
 \end{teo}
-Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
+\indent Este Teorema dice que para un tomador de decisiones racional (es decir, que se comporta según los axiomas) existe una función de utilidad tal que las preferencias del tomador de decisiones son equivalentes a maximizar una utilidad esperada.
 \subsubsection{Teorema de DeFinetti: existencia de probabilidad subjetiva}
@@
 \end{teo}
-Este Teorema arroja una probabilidad para cada tomador de decisiones tal que la relación de preferencias del éste sea equivalente con la utilidad esperada bajo esta distrbución; por esto, a tal distribución se le conoce como \textit{probabilidad subjetiva}.
+\indent Este Teorema arroja una probabilidad para cada tomador de decisiones tal que la relación de preferencias del éste sea equivalente con la utilidad esperada bajo esta distrbución; por esto, a tal distribución se le conoce como \textit{probabilidad subjetiva}.
 \subsubsection{Teorema de Savage: existencia de utilidad y probabilidad subjetiva}
@@
 \\
-	    El Teorema de Savage requiere dos conceptos básicos: un conjunto de estados $S$ y un conjunto de resultados $X$ y las opciones a escoger en este contexto son \textit{actos}, que son funciones de los estados a los resultados:
+\indent El Teorema de Savage requiere dos conceptos básicos: un conjunto de estados $S$ y un conjunto de resultados $X$ y las opciones a escoger en este contexto son \textit{actos}, que son funciones de los estados a los resultados:
 \[ F = X^S = \{ f | f: S \to X \}. \]
-	    Consideramos una relación binaria $\succeq$ definida sobre $F$.
-	    Diremos que un conjunto $A \subseteq S$ es un evento, y no se requieren suposiciones en cuanto a la medibilidad de estos.
-	    Consideremos la siguiente notación: para actos $f,g \in F$, y un evento $A \subseteq S$ definimos un acto $f_A^g$ como sigue:
+\indent Diremos que un conjunto $A \subseteq S$ es un evento, y no se requieren suposiciones en cuanto a la medibilidad de estos.
+En el contexto del Teorema de Savage, las preferencias del tomador de decisiones están definidas sobre las funciones que van de los estados a los resultados.
+Entonces, consideramos una relación $\succeq$ definida sobre $F$.\\
+\\
+\indent Consideremos la siguiente notación: para actos $f,g \in F$, y un evento $A \subseteq S$ definimos un acto $f_A^g$ como sigue:
 \[ f_A^g(s)=g(s) \textrm{ si  } s \in A, \textrm{  } f(s) \textrm{ si  } s \in A^c   \]
@@
 \subsubsection{Juegos extensivos con información perfecta}
+Contenido tomado de \cite{shoham2008multiagent}
 \subsubsection{Juegos extensivos con información impefecta}
-		\cite{10.1007/978-94-010-0189-2_25}
+Contenido tomado de \cite{shoham2008multiagent} y \cite{10.1007/978-94-010-0189-2_25}
 \subsubsection{Juegos extensivos con información incompleta}
-
+Contenido tomado de \cite{osborne1994course}
 \section{Pregunta de investigación}
+Cómo puede un agente inteligente adquirir y utilizar conocimiento causal a partir de su interacción con el entorno en un proceso de toma de decisiones bajo incertidumbre? 
 \section{Planteamiento y Modelado}
@@
 \\
-\cite{di2013predictive} muestran que el uso de contrafactuales no mejora las predicciones de un agente que ya conoce el conjunto de hipótesis posibles (logically omniscient).
+\cite{di2013predictive} muestran que el uso de contrafactuales no mejora las predicciones de un agente que ya conoce el conjunto de hipótesis posibles (logically omniscient).\\
 \\
@@
 Después de $T$ rondas de intervención, se busca encontrar la mejor intervención sobre alguna una de las variables que está en la parte desconocida del modelo.
-Una gran diferencia de este trabajo respecto a los trabajos clásicos de identificación del mejor brazo como \cite{audibert2010best} es considerar el \textit{filtrado de información} entre brazos, pues gracias al modelo causal los valores muestreados de un brazo (intervención) pueden aportar información sobre los otros
-mejora a \cite{lattimoreNIPS2016} y a \cite{audibert2010best}.
+Una gran diferencia de este trabajo respecto a los trabajos clásicos de identificación del mejor brazo como \cite{audibert2010best} es considerar el \textit{filtrado de información} entre brazos, pues gracias al modelo causal los valores muestreados de un brazo (intervención) pueden aportar información sobre los otros.
 Uno de las mejores respecto al trabajo de \cite{lattimoreNIPS2016} es que el nivel de truncamiento es adaptativo en cada ronda de modo que balancea sesgo y varianza.
-Aunque los algoritmos de \cite{sen2017identifying} mejoran los de \cite{lattimoreNIPS2016} en el mismo caso, siguen requiriendo conocer las distribuciones condicinales y las marginales
+Aunque los algoritmos de \cite{sen2017identifying} mejoran los de \cite{lattimoreNIPS2016} en el mismo caso, siguen requiriendo conocer las distribuciones condicinales y las marginales.
 \subsection{Juegos, decisiones y actualización de creencias}
@@
 \section{Objetivos}
-	\section{Objetivo General (debe ir relacionado con el título)}
+	\subsection{Objetivo General}
 	La investigación que se propone llevar a cabo tiene dos objetivos generales; por un lado, encontrar la manera en la que un agente que se enfrenta a un entorno incierto pueda usar información causal de manera que ésta sea de ayuda al llevar a cabo una tarea; por otro lado, encontrar la manera en la que un agente adquiera información causal conforme interactúa con su entorno.
-	\section{Objetivos específicos}
+	\subsection{Objetivos específicos}
 	Para llevar satisfacer los objetivos generales, hasta el momento se ha identificado que será necesario lo siguiente:
@@
 	\subsection{Hipótesis}
-	Tentativa: La incorporación efectiva de información causal en un proceso de toma de decisiones bajo incertidumbre produce un mejora en el desempeño del tomador de decisiones.
-	Otra opción: Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
+    Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
 \subsection{Metodología de Trabajo}
-
+\begin{itemize}
+\item Consideramos un problema de decisión bajo incertidumbre que se desarrolla en un entorno regido por un modelo causal.
+\item Un primer paso es modelar las acciones del agente y la respuesta del entorno se modelan como un juego de dos jugadores, uno de los cuales será el tomador de decisiones inicial, y el otro será un ente abstracto llamado Naturaleza que, suponemos, conoce el modelo causal y lleva a cabo sus jugadas a partir de lo dictado por éste.
+Los juegos sucederán de manera repetida para poder aprender de éste.
+\item En este momento, suponemos que la Naturaleza es indiferente a cualquier resultado y para esto se le asignan pagos iguales en cada resultado del juego.
+Esto puede modificarse en un futuro para modelar casos en los que la Naturaleza tenga ciertas \textit{intenciones}.
+\item Como caso inicial, consideramos que el modelo causal es conocido en su totalidad para el tomador de decisiones.
+Para este caso, se encontrará un criterio de solución y se comparará con algoritmos del estado del arte de detección de acciones óptimas en un modelo causal.
+\item En una siguiente etapa, se considera que el tomador de decisiones sólo conoce la \textit{estructura} del modelo causal.
+Para este caso, se encontrará una forma de aprender los valores de los parámetros del modelo así como una manera de usar esta información en futuras decisiones a tomar.
+Para este segundo caso, se debe construir una manera de expresar el conocimiento inicial acerca de la estructura que además de ser coherente con la manera en la que los seres humanos formamos creencias (\cite{larrouy2017mindreading}).
+\item Para llevar a cabo el descubriminto del modelo, se define un número de juegos de observación, dentro de estos juegos el tomador de decisiones no modificará sus creencias causales, aunque éstas lo obliguen a tener un desempeño no óptimo.
+\item Posteriormente, en el caso en el que el agente no conoce nada, sus creencias serán lo único con lo que pueda operar.
+Este conocimiento causal inicial, en el caso general, debe ser expresado en términos de probabilidades y para esto se construirá un \textit{prior} que consiste en una distribución sobre un espacio de matrices.
+Recordemos que una matriz (que cumple ciertas características) define un grafo, por lo que tener de esta manera estaremos especificando las probabilidades que un agente asigna, de manera subjetiva, a distintos modelos causales.
+Todavía no está determinado si consideraremos que el tomador de decisiones conoce el número máximo de variables que puede tener el modelo o no.
+De esto dependerá la \textit{clase} de distribución que se deba usar.
+\item En términos del conocimiento causal obtenido hasta cierto momento dentro de la serie de juegos o interacciones, se debe definir una estrategia para el tomador de decisiones a partir del conocimiento causal que tenga hasta ese momento; una \textit{estrategia mixta} permitiría explorar el ambiente con tal de explorar el ambiente de manera amplia.
+\item A partir de lo observado en la ronda de juegos de observación, se debe actualizar las creencias causales de manera que se reflejen las relaciones que pudieron ser observadas, las cuales serán almacenadas en un diagrama de influencia.
+\item Modificar la estrategia considerando los efectos posibles de cada acción que fueron aprendidos, para esto se usará el Diagrama de Influencia previamente construido. 
+\item Comparar el desempeño versus lo que se haría si se conociera el modelo con detalle.
+\item Extensiones.
+\end{itemize}
 \section{Resultados Preliminares}
@@
 El escenario es el de un médico que tiene que escoger uno de entre varios tratamientos para un paciente con enfermedad desconocida.
-Los tratamientos posibles conllevan riesgos y ventajas, que están regidos por un modelo causal.\\
-\\
-Los tres casos a considerar son los siguientes:\\
-\\
-En el primer caso, supondremos que el modelo causal es conocido completamente por el agente y veremos cuál sería su manera de actuar en este caso.
-Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.\\
-\\ 
-Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este. \\
-\\
-En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
+Los tratamientos posibles conllevan riesgos y ventajas, que están regidos por un modelo causal.
+Los tres casos a considerar son los siguientes
+\begin{itemize}
+\item En el primer caso, supondremos que el modelo causal es conocido completamente por el agente y veremos cuál sería su manera de actuar en este caso.
+Que el modelo sea conocido completamente significa que el agente conoce la estructura del grafo así como las probabilidades asociadas a cada variable.
+\item Posteriormente, en el segundo caso, el agente sólo tendrá a su disposición la estructura del grafo, pero no los parámetros de este. 
+\item En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
 En resumen, en los tres casos el agente contará con \textit{creencias} sobre el modelo causal verdadero; en el primer caso, sus creencias coincidirán plenamente con el modelo; en el segundo caso, sólo con la estructura y en el tercer caso, no tenemos garantía de nada.
+\end{itemize}
 \subsection{Escenario}
@@
 Notemos que para tener una solución de equilibrio también las jugadas de la naturaleza deben ser la mejor respuesta a cualquiera de las acciones posibles del médico,  pero vamos a darle a la naturaleza el mismo pago en todos los resultados posibles para no entrar en discusión de las \textit{intenciones} de la naturaleza.
-Además, darle a la Naturaleza utilidad en los pagos permite dejar espacio para posibles generalizaciones futuras; por ejemplo, en el caso en el que la Naturaleza sí tenga algún tipo de intenciones o de objetivos a cumplir.
+Además, darle a la Naturaleza utilidad en los pagos permite dejar espacio para posibles generalizaciones futuras; por ejemplo, en el caso en el que la Naturaleza sí tenga algún tipo de intenciones o de objetivos a cumplir.\\
 \\
+El juego, en su forma extensiva, queda representado de la siguiente manera:\\
+\begin{itemize}
+\item La naturaleza escoge primero y asigna la enfermedad $A$ o la enfedmedad$B$.
+\item Debido a que el médico no sabe qué enfermedad asignó la Naturaleza, el médico escoge dentro de un \textit{conjunto de información} uno de los tratamientos posibles.
+\item A continuación, la Naturaleza escoge ahora si el paciente vive o muere a partir del modelo causal.
+\item Los pagos, para ambos jugadores en cada hoja son $m=(1,a), v=(0,a)$; es decir, el médico obtiene un pago de 1 si el paciente vive, y en ambos casos la Naturaleza recibe una cantidad cualquiera $a$, de modo que es indiferente ante cualquier resultado.
+\end{itemize}
 % Node styles
@@
 child{node(2)[solid node]{}
-child{node[hollow node,label=below:{$(e,f)$}]{} edge from parent node[left]{$f$}}
-child{node[hollow node,label=below:{$(g,h)$}]{} edge from parent node[right]{$c$}}
-child{node[hollow node,label=below:{$(c,d)$}]{} edge from parent node[right]{$n$}}
+	child{node[hollow node,label=left:{$N$}]{} 
+		child{node[hollow node,label=below:{$m$}]{}} 
+		child{node[hollow node,label=below:{$v$}]{}}
+	edge from parent node[left]{$f$}}
+	child{node[hollow node,label=left:{$N$}]{} 
+		child{node[hollow node,label=below:{$m$}]{}} 
+		child{node[hollow node,label=below:{$v$}]{}}
+	edge from parent node[right]{$c$}}
+	child{node[hollow node,label=left:{$N$}]{} 
+		child{node[hollow node,label=below:{$m$}]{}} 
+		child{node[hollow node,label=below:{$v$}]{}}
+	edge from parent node[right]{$n$}}
 edge from parent node[right,xshift=3]{$B$}
@@
 
-\section{Plan de trabajo}
+\section{Plan de trabajo (cronograma)}
+Por definir.
 
 \section{Conclusiones}
-
+Por añadir.
 \section{Limitaciones}
-En este trabajo consideramos un tomador de decisiones que actúa racionalmente.
-
-Consideramos la definición de causalidad que aparece arriba y la consideramos de forma axiomática; es decir, no se pone en tela de juicio si los supuestos son realistas o no, ni es intención encontrar la definición de causalidad que mejor se adapte a una situación particular del mundo \textit{real}.
-
+En este trabajo consideramos un tomador de decisiones que actúa racionalmente.\\
+\\
+Consideramos la definición de causalidad que aparece arriba y la consideramos de forma axiomática; es decir, no se pone en tela de juicio si los supuestos son realistas o no, ni es intención encontrar la definición de causalidad que mejor se adapte a una situación particular del mundo \textit{real}.\\
+\\
 No estamos considerando que el ente que llamamos Naturaleza tenga intenciones, pero al darle \textit{pagos} en cada juego, aunque iguales, podríamos permitir que busque maximizar algún objetivo.
@@
 \section{Extensiones posibles}
-Más pasos dentro de un mismo juego, oponentes, intenciones de la naturaleza
+Dadas las limitaciones del trabajo propuesto, existen varias extensiones posibles que se dejan como trabajo futuro; entre algunas de estas encontramos:
+\begin{itemize}
+\item Considerar juegos de mayor duración.
+\item Considerar varios jugadores además de la naturaleza.
+\item Considerar que la Naturaleza tiene algún tipo de intenciones o de objetivo a maximizar; por ejemplo, el concepto de Naturaleza que hemos usado podría bien representar a un gobierno o un Banco central.
+\item Estudiar equilibrios en este tipo de juegos en términos de la estructura del modelo causal. 
+\end{itemize}
 

### 2018-04-19 Segundo_Semestre/Seminario/Semana12/Proposal.tex
cambios en intro just mot

@@
 \indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
-De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ‚por dónde empezar?
+De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ¿por dónde empezar?
 \\
@@
 A su vez, tal modificación del entorno es percibida por el agente en términos de una señal de recompensa y de encontrarse en otro estado.
-En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un nuevo ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.
+En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.
 \subsection{Motivación}
@@
 \\
-\indent La información causal por sí sola nos ayuda a los seres humanos a tomar decisiones, pues resulta que los seres humanos utilizan información causal al tomar decisiones; además, modifican la información causal previa que tenían del ambiente conforme interactúan con este (\cite{hagmayer2013repeated}).
-Además, este conocimiento causal es utilizado de manera correcta por los seres humanos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
-Aunque, los seres humanos no son perfectos al estimar la fuerza de los efectos, o violan el supuesto Markoviano (\cite{rottman2014reasoning}).
+\indent Los seres humanos, utilizamos de manera correcta la información causal con la que contamos con tal de escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
+Además, modificamos las creencias previas que se tenían previamente sobre las relaciones causales en su ambiente conforme interactúan con él (\cite{hagmayer2013repeated}).
+Aunque, estos procesos de adquisición y actualización no son perfectos; por ejemplo, al estimar la fuerza de los efectos se cometen errores o el supuesto Markoviano que requieren los modelos causales es violado (\cite{rottman2014reasoning}).
 \\
 \\
-\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto.
+\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto que si no utilizaran información causal del entorno.
 En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.\\
 \\
-\indent Han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+\indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
 Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
-\\
-\\
-\indent Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no solo las acciones y preferencias de un sólo agente sino las de otros.
-En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos Extensivos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene creencias \textit{a priori} sobre algunos aspectos del ambiente (en las aplicaciones de economía estas creencias suelen versar sobre las recompensas o sobre las estrategias de otros jugadores).
-En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.\\
-\\
-\indent Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.
 \subsection{Justificación}
@@
 Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
-Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
+Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.\\
+\\
+\indent Finalmente, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
+\\
+\\
+Entonces, en conclusión, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar una interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
+A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.
+En lo que sigue del documento, veremos con cierto detalle algunas de las herramientas teóricas que serán de utilidad para la realización de la investigación propuesta.
 \section{Marco Teórico}
@@
 	\subsection{Teoría de Juegos}
-	Teoría de Juegos es un área de las matemáticas cuyo objetivo es dar las herramientas necesarias para analizar la interacción estratégica de varios tomadores de decisiones (\cite{osborne1994course}).
-En esta sub-sección veremos varios tipos de juegos posibles.
+Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no sólo las acciones y preferencias de un agente sino también las de otros.
+En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos Extensivos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene \textit{creencias} \textit{a priori} sobre algunos aspectos del ambiente.
+Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.).\\
+\\
+\indent En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.
 Por el momento, sólo veremos distintas clases de juegos para mostrar que el problema a tratar en esta propuesta cuenta con la maquinaria suficiente para ser modelado, pero se dejarán de lado otras cuestiones importantes de Teoría de Juegos como son el Equilibrio de Nash y las Estrategias Mixtas.
@@
 Contenido tomado de \cite{osborne1994course}
-\section{Pregunta de investigación}
-Cómo puede un agente inteligente adquirir y utilizar conocimiento causal a partir de su interacción con el entorno en un proceso de toma de decisiones bajo incertidumbre? 
-\section{Planteamiento y Modelado}
-Suponemos un ambiente que está regido por un modelo causal, y suponemos que este es desconocido, pero fijo y además cumple que todas las variables se encuentran contenidas en él (suficiencia causal).\\
-\\
-\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}) es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
-Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
-De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
-\\
-\indent Cuando el modelo causal no es conocido por el tomador de decisiones, tampoco lo son las probabilidades de los efectos dadas las causas, por lo que no se puede calcular la utilidad esperada.
-Por esto, lo que se propone es que el tomador de decisiones \textit{aprenda} los parámetros a partir de la información que provee el ambiente en cada interacción con este.
-Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
-Lo que estas técnicas ignoran es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.
-En este caso, la estructura es la que está dada por las relaciones causa-efecto en el ambiente.\\
-\\
-\indent Para llevar a cabo el proceso aprendizaje dotamos al agente con  \textit{creencias} sobre la estructura causal de su entorno las cuales se irán actualizando conforme el agente interactue con éste.
-Estas creencias pueden considerarse como un modelo causal \textit{propio} del agente, pues el agente utilizará este modelo causal para la toma de decisiones, que por construcción irá acercándose cada vez más al modelo causal verdadero.
-El problema de la actualización de creencias puede estudiarse en dos casos distintos: si el número de variables en el modelo causal es conocido por el agente, entonces podemos representar el modelo causal mediante matrices de tamaño fijo y llevar a cabo actualizaciones en las entradas de estas; pero si el número de variables es desconocido por el tomador de decisiones, un camino a seguir es utilizar un enfoque Bayesiano no-paramétrico y considerar distribuciones sobre un espacio de grafos; este problema parece tener una complejidad alta.
-Enfoques Bayesianos no-paramétricos para descubrimiento causal han sido explorados en \cite{karabatsos2012bayesian}.\\
-\\
-\indent Para modelar la interacción del tomador de decisiones con su entorno, vamos a asumir que el agente está interactuando con un ente abstracto al que le llamaremos Naturaleza y que la dinámica de interacción consiste en un \textit{juego}.
-A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.
-En el contexto de Teoría de Juegos, los juegos que permiten que el agente tenga creencias información incompleta se llaman \textit{Juegos Estocásticos}, los cuales fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
-\\
-\indent En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
-Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
-Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
-Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
-\\
-\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
-En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
-En este caso el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
+
 \section{Trabajo Previo}
@@
 \cite{tingley2010belief} muestran, experimentalmente y para una clase de juegos particular que los agentes involucrados actualizan sus creencias respecto a ciertos aspectos de los otros jugadores; encuentran que los jugadores actualizan sus creencias de manera que no se alejan mucho del \textit{prior} inicial.
+
+\section{Pregunta de investigación}
+En un entorno incierto para el cual se cumplen una serie de relaciones causales ¿cómo puede un agente inteligente adquirir y utilizar conocimiento sobre estas relaciones a partir de su interacción con el entorno en un proceso de toma de decisiones?
+\section{Planteamiento y Modelado}
+Suponemos un ambiente que está regido por un modelo causal, y suponemos que este es desconocido, pero fijo y además cumple que todas las variables se encuentran contenidas en él (suficiencia causal).\\
+\\
+\indent Un problema de decisión bajo incertidumbre (de una etapa) consiste en un conjunto de acciones, un conjunto de consecuencias y un conjunto de \textit{eventos inciertos}; la interpretación estándar (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}) es que el agente escogerá una de estas acciones, luego ocurrirá uno de estos eventos inciertos y será esto lo que determine la consecuencia.
+Suponer que un modelo causal exista y opere sobre este ambiente implica que éste controle la relación entre los eventos inciertos y las consecuencias de éste, lo cual queda codificado en probabilidades; de este modo, si se conoce el modelo causal, se conocen las probabilidades de ocurrencia de las consecuencias dada la acción escogida y el evento incierto.
+De esta manera, se puede determinar la acción óptima (o serie de acciones) mediante la maximización de la utilidad esperada, pues este es el único criterio de solución que es coherente con los axiomas de decisión de \cite{savage1954the}.\\
+\\
+\indent Cuando el modelo causal no es conocido por el tomador de decisiones, tampoco lo son las probabilidades de los efectos dadas las causas, por lo que no se puede calcular la utilidad esperada.
+Por esto, lo que se propone es que el tomador de decisiones \textit{aprenda} los parámetros a partir de la información que provee el ambiente en cada interacción con este.
+Sabemos que un problema de decisión bajo incertidumbre puede resolverse con algunas de las técnicas ya conocidas en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}), resulta además que una política óptima aprendida con estas técnicas alcanza la MUE (\cite{webb2007game}).
+Lo que estas técnicas ignoran es cuando el agente sabe que su entorno tiene cierta estructura particular, y se desea utilizar esta información para aprender los parámetros del modelo y tomar decisiones de manera correcta.
+En este caso, la estructura es la que está dada por las relaciones causa-efecto en el ambiente.\\
+\\
+\indent Para llevar a cabo el proceso aprendizaje dotamos al agente con  \textit{creencias} sobre la estructura causal de su entorno las cuales se irán actualizando conforme el agente interactue con éste.
+Estas creencias pueden considerarse como un modelo causal \textit{propio} del agente, pues el agente utilizará este modelo causal para la toma de decisiones, que por construcción irá acercándose cada vez más al modelo causal verdadero.
+El problema de la actualización de creencias puede estudiarse en dos casos distintos: si el número de variables en el modelo causal es conocido por el agente, entonces podemos representar el modelo causal mediante matrices de tamaño fijo y llevar a cabo actualizaciones en las entradas de estas; pero si el número de variables es desconocido por el tomador de decisiones, un camino a seguir es utilizar un enfoque Bayesiano no-paramétrico y considerar distribuciones sobre un espacio de grafos; este problema parece tener una complejidad alta.
+Enfoques Bayesianos no-paramétricos para descubrimiento causal han sido explorados en \cite{karabatsos2012bayesian}.\\
+\\
+\indent Para modelar la interacción del tomador de decisiones con su entorno, vamos a asumir que el agente está interactuando con un ente abstracto al que le llamaremos Naturaleza y que la dinámica de interacción consiste en un \textit{juego}.
+A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.
+En el contexto de Teoría de Juegos, los juegos que permiten que el agente tenga creencias información incompleta se llaman \textit{Juegos Estocásticos}, los cuales fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\\
+\indent En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
+Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
+Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
+Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
+\\
+\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
+En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
+En el trabajo de este autor,  el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
+
+\subsection{Hipótesis}
+Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
+
 \section{Objetivos}
@@
 \section{Metodología}
-	\subsection{Hipótesis}
-    Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
 \subsection{Metodología de Trabajo}
@@
 \item Comparar el desempeño versus lo que se haría si se conociera el modelo con detalle.
-\item Extensiones.
+
 \end{itemize}
@@
 \subsection{Caso en el que se conoce la estructura del modelo}
-Aquí se conocen los efectos de las causas, pero no sus probabilidades, entonces primero debo conocer el subconjunto posible de consecuencias y ahí tal vez hacer exploración greedy.
+Consideremos ahora el caso en el cual el tomador de decisiones conoce la estructura del modelo, pero no conoce los valores de los parámetros de éste.
+Entonces, lo que el agente tendrá es una distribución de probabilidad sobre los posibles valores que estos parámetros puedan tener.
 \subsection{Caso general}

### 2018-04-23 Segundo_Semestre/Seminario/Semana12/Beamer_Template.tex
adding hypothesis

@@
 
-\title[Short title]{Main Topic } % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Short title]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page
 
@@
 	\begin{itemize}
-	\item Human beings conceive their actions upon the world as \textit{interventions} (\cite{hagmayer2009decision})
+	\item Human beings conceive their actions upon the world as \textit{interventions} (\cite{hagmayer2009decision}).
 	\item Following this idea, \cite{lattimoreNIPS2016} consider a decision making process where available actions are interventions in a causal graphical model. Faster identification of best-action.
@@
 		\frametitle{Do calculus}
-		The structure of the causal graph produces some equivalence rules between observational an interventional data.
-		\begin{teo}{\label{docalculus}} (\cite{pearl2009causality})\\
-		Sea $\mathcal{G}$ un modelo gráfico causal y $P_{\mathcal{G}}$ la medida de probabilidad inducida por el modelo; entonces, para cualesquiera conjuntos disjuntos de variables $X,Y,Z,W$ 	se cumple que
-		\begin{itemize}
-		\item If $Y$ is conditionally independent from $Z$ given $X$ y $W$ in $\mathcal{G}_{\bar{X}}$, then
-		\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
-		\item Si en el grafo $\mathcal{G}_{\bar{X}\underline{Z}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$, entonces
-		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
-		\item Sea $Z(W)$ el conjunto de nodos en $Z$ tales que no son ancestros de ningún nodo en $W$ en el grafo $\mathcal{G}_{\bar{X}}$.
-Si en el grafo $G_{\bar{X}, \bar{Z(W)}}$ se cumple que $Y$ es condicionalmente independiente de $Z$ dados $X$ y $W$ entonces,
-		\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
-		\end{itemize}
-		\end{teo}
+		The do calculus is a set of rules which allow, under certain conditions, to express interventional statements in terms of observational statements.
 		\end{frame}
@@
 	\subsection{Game Theory}
-		\subsubsection{Games in normal form}
-		\subsubsection{Games in extensive form}
+\begin{frame}
+\frametitle{Game Theory}
+Game Theory is an area of Mathematics that is used to model strategic interaction situations, which can be interpreted as multi-agent decision problems, where the outcomes for each player depend on the actions of the other players.
+Several classes exist:
+\begin{itemize}
+\item Strategic Games
+\item Extensive Games
+\item Stochastic Games
+\end{itemize}
+\end{frame}
 \section{Research Question}
@@
 \frametitle{Research Question}
-How can an intelligent agent acquire and use causal knowledge from his interaction with an uncertain environment in a decision making process?
+In an uncertain environment where the relationship between events is governed by some causal mechanism, how can an intelligent agent acquire and use causal knowledge from his interaction with the environment in a decision making proces in order to learn faster to choose good actions?
+\end{frame}
+
+	\subsection{Hypothesis}
+	\begin{frame}
+	\frametitle{Hypothesis}
+Let $\mathcal{G}$ a causal graphical model whose probability distribution $P_\mathcal{G}$ satisfies (the axioms).
+Let $(D,E,c)$ a decision problem whose uncertain events are related to the variables in $\mathcal{G}$.
+Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, then by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less rounds than if he doesn't consider causal information.
 	\end{frame}
@@
 \section{Previous work}
-	\subsection{Causality and Decision Theory}
 	\subsection{Optimal Interventions}
@@
 \end{frame}
+
 \section{Objectives}
@@
 \section{Methodology}
-	\subsection{Hypothesis}
-	\begin{frame}
-	\frametitle{Hypothesis}
-	Under certain conditions, an agent can acquire and use causal knowledge from interaction with an uncertain environment and effectively use it in a decision problem under uncertainty.
-	\end{frame}
+
 	\subsection{Working methodology}
@@
 \frametitle{Preliminary Results}
-
+For the preliminary results we set up a toy example (the same one that we have been using) and analyze 3 separate cases in order of growing difficulty:
+\begin{itemize}
+\item In the first case, the decision maker knows it all.
+\item In the second case, the decision maker only knows for certain the structure of the model, but not its parameters so he has beliefs about them.
+\item In the third case, the decision maker knows nothing at all, so he has general beliefs about the causal model itself.
+\end{itemize}
 \end{frame}

### 2018-04-25 Segundo_Semestre/Seminario/Semana12/Proposal.tex
Pasando cosas a motivacion y de la intro a marco teorico

@@
 Debido al enorme número de factores diversos que están involucrados en nuestro entorno, constántemente nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
-Esa misma realidad, altamente compleja, contiene una serie de relaciones causa-efecto que podemos utilizar al momento de tomar decisiones.
+Esa misma realidad, altamente compleja e interconectada, contiene una serie de relaciones causa-efecto que podemos utilizar al momento de tomar decisiones.
 El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) y predecir las consecuencias de nuestros actos.
@@
 \indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
-Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.\\
+Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.
+La utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
 \\
-\indent La utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
-El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
-\\
-\indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
-De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ¿por dónde empezar?
-\\
-\\
-\indent El gran logro de J.L.
-Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
-Lo único que requiere el Teorema de Savage es una serie de axiomas de coherencia sobre la relación de preferencias de un tomador de decisiones; con esto, muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.
-Existen otros criterios de optimalidad cuando los axiomas de Savage son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).
-\cite{gilboa2001theory} propone un nuevo paradigma de toma de decisiones en el cual los agentes toman decisiones haciendo analogías a eventos pasados en los cuales tuvieron un buen desempeño.
-Estas teorías, aunque interesantes, no serán abordadas en este trabajo sino que se mencionan sólo por completez.\\
 \\
@@
 A su vez, tal modificación del entorno es percibida por el agente en términos de una señal de recompensa y de encontrarse en otro estado.
-En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.
+En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.\\
+\\
+\indent En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con Nash.}.
+Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.
+Los juegos estocásticos fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
 \subsection{Motivación}
@@
 \\
+\indent Lo que este trabajo señala es que a la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
+La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
+La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
+\\
+\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
+Por otro lado, el conocimiento inicial del tomador de decisiones es un punto delicado, pues a partir de qué tan completo o no sea, así de buenas serán sus acciones iniciales.
+\cite{billot2005probabilities} muestran cómo asignar probabilidades iniciales dado un conjunto de observaciones.\\
+\\
+\indent Encontrar la acción óptima a llevar a cabo en un contexto incierto y de manera tal que el tomador de decisiones pueda aprender algo del ambiente se ha abordado desde el punto de vista de Proceso de Decisión Markovianos y Programación Dinámica, lo que dio lugar al área conocida como Aprendizaje por Refuerzo (RL por sus siglas en inglés).
+En este contexto, el agente que debe aprender a llevar a cabo una tarea se encuentra ante un entorno incierto, y sobre el cual debe aprender mediante interacción.
+El agente lleva a cabo una acción y observa una señal de recompensa del ambiente.
+La señal de recompensa debe ser tal que, al intentar el agente maximizarla en el largo plazo, el agente cumpla su objetivo.\\
+\\
+\indent Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
+En este trabajo, lo que se propone es modelar la interacción agente-entorno como un \textit{juego} de modo que el agente podrá observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
+Además, al ser la señal de recompensa algo que debe diseñarse de acuerdo al problema, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
+Considerar a la naturaleza como un jugador no es una idea nueva, pues desde \cite{milnor1951games}, \cite{papadimitriou1985games}, \cite{Szep1985} y \cite{eiselt2004games} se ha considerado a la Naturaleza como un jugador que \textit{asigna} inicialmente un estado al ambiente.\\
+\\
+\indent Que un agente pueda \textit{aprender} algo de su entorno a lo largo de una serie de interacciones, o juegos, ha sido estudiado ya en el caso en el que el jugador busca aprender los pagos del juego o las estrategias de alguno de los otros agentes \cite{fudenberg1998learning} lo que motiva a pensar que un agente podría aprender sobre el modelo causal que rige su entorno.
+\\
+\\
 \indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
@@
 \\
-\indent Lo que este trabajo señala es que  la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
-La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
-La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
+\indent Posteriormente, el trabajo de \cite{sen2017identifying} extiende el trabajo de \cite{lattimoreNIPS2016} al caso en el cual el modelo causal es parcialmente conocido, pero sigue abierto el problema para el caso en el cual el agente deba aprender el modelo conforme interactúa con su entorno.\\
 \\
-\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
-Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.\\
-\\
-\indent A diferencia de lo que sucede en Aprendizaje por Refuerzo (RL por sus siglas en inglés), en donde el agente lleva a cabo una acción y observa una señal de recompensa del ambiente, lo que se propone es a observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
-Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
-Además, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.\\
-\\
-\indent Finalmente, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
+\indent Además, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
 \\
 \\
-Entonces, en conclusión, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar una interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
-A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.
-En lo que sigue del documento, veremos con cierto detalle algunas de las herramientas teóricas que serán de utilidad para la realización de la investigación propuesta.
+\indent En resumen, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar la interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
+A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.\\
+\\
+En lo que sigue del documento, primero mencionaremos las herramientas teóricas requeridas para el desarrollo de esta propuesta.
+Posteriormente, los trabajos previos que han atacado este mismo problema, o similares.
+Luego, la pregunta de investigación y la hipótesis relacionada, así como los Objetivos generales y específicos.
 \section{Marco Teórico}
 En esta sección se mencionarán las definiciones más importantes relativas a esta propuesta de investigación doctoral sin pretender ser exhaustivo.
-El contenido de estas secciones es para dar al lector certeza sobre la existencia y validez de los resultados que se pretenden utilizar.
+En primer lugar, veremos qué entendemos por causalidad y modelos causales.
+Luego, veremos que la toma de decisiones bajo incertidumbre ya ha sido estudiado por diversos investigadores y finalmente, y utilizando parte de las nociones de toma de decisiones, veremos un poco de Teoría de Juegos.
 \subsection{Relaciones Causales}
@@
 \indent \textit{...all that matters is that our concept appropriately tracks some relation(s) in the world that can ground and support the learning, reasoning and inferences that we do with our causal knowledge...}
+\subsubsection{Circularidad en la definición de causalidad}
+En los trabajos clásicos sobre Causalidad (\cite{spirtes2000causation}, \cite{pearl2009causality}) suele definirse primero el concepto de causalidad haciendo referencia al lenguaje ordinario; posteriormente, se define la maquinaria matemática para construir los modelos causales.
+De esta manera, la única definición formal que se puede obtener de causalidad es aquello que es expresado por modelos causales, pero éstos a su vez se definen como modelos de relaciones causales.
+Esta aparente circularidad es estudiada por \cite{woodward2005making}.\\
+\\
+Por esto, daremos aquí una definición \textit{de trabajo} de causalidad que permita operar con ella de manera conveniente y coherente con el planteamiento.
+
 \subsubsection{Definición de Causalidad}
@@
 \begin{defi}{\label{defcausa}}
-		Sea $(\Omega, \mathcal{F})$ un espacio medible, finito, y consideramos una relación binaria $\to  \subseteq \mathcal{F} \times \mathcal{F}$ que cumple:
+Sea $(\Omega, \mathcal{F}, \mathbb{P})$ un espacio de probabilidad, finito, y consideramos una relación binaria $\to  \subseteq \mathcal{F} \times \mathcal{F}$\footnote{Recordemos que una relación binaria $R$ definida sobre elementos de un conjunto $A$ es un sub-conjunto del producto cartesiano $A \times A$.} que cumple:
 \begin{itemize}
@@
 \end{defi}
+
 \subsubsection{Causas Directas e Indirectas}
-Consideremos ahora un \textit{refinamiento} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
+Consideremos ahora un \textit{refinamiento}\footnote{En el sentido de una Topología más fina \cite{munkres2000topology}.} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
 Decimos que $A$ es una \textit{causa directa} de $B$ si se cumple que $A \to B$ y no existe evento $C$ tal que $A \to C \to B$ para todo refinamiento propio de $\mathcal{F}$.
@@
 \subsubsection{Construcción de un grafo que represente relaciones causales}
-Abusando de la notación, fijando un contexto, podemos construir un grafo que represente las relaciones (causales) entre eventos de $\mathcal{F}$ de la siguiente manera: si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto
+Abusando de la notación, fijando un contexto, podemos construir un grafo que represente las relaciones (causales) entre eventos de $\mathcal{F}$ de la siguiente manera: \\
+si $A \to B$ y $A$ es una causa directa de $B$, entonces se añade un nodo que representa el evento $A$ y un nodo que representa el evento $B$ y una arista dirigida que los une en el sentido de la causa hacia el efecto.
 \begin{prop}
@@
 Notemos que al ser finito el grafo dirigido, existirán necesariamente variables que no tengan padres, las cuales serán conocidas como variables exógenas.
+
 \subsubsection{Una medida de probabilidad a partir del grafo}
 Dado un grafo acíclico dirigido, podemos construir una medida de probabilidad que expresa las relaciones que se encuentran establecidas en el grafo, la cual denotaremos $P_{\mathcal{G}}$.
-Requerimos que se cumplan las siguientes condiciones que relacionan el grafo así construido con la distribución de probabilidad $P_{\mathcal{G}}$ que éste expresa:
+Requerimos que la medida $P_{\mathcal{G}}$ cumpla las siguientes condiciones que la relacionan con el grafo:
 \begin{itemize}
-\item Markov Causal
-\item Minimalidad Causal
-\item Fidelidad Causal
+\item Markov Causal: Un evento $V$ (o nodo en el grafo) es independiente de todo aquel evento (nodo) que no sea ni una causa directa ni un efecto directo dadas las causas directas de $V$.
+\item Minimalidad Causal: el grafo $\mathcal{G}$ es mínimo en el sentido de que ningun sub-grafo propio satisface la condición Markov Causal.
+\item Fidelidad Causal: La condición Markov Causal contiene las independencias condicionales que el grafo $\mathcal{G}$ expresa.
 \end{itemize}
@@
 Quien no esté de acuerdo con los axiomas, no lo estará con las conclusiones y queda siempre la invitación de desarrollar la teoría utilizando otras altenativas.
+
+\subsubsection{Relación entre $\mathbb{P}$ y $P_\mathcal{G}$}{\label{relacion}}
+Es interesante estudiar la relación entre la medida de probabilidad del espacio $\mathbb{P}$ y la medida de probabilidad definida por el grafo que captura las relaciones causales $P_\mathcal{G}$.
+La pregunta concreta que se hace es si la medida $P_\mathcal{G}$ captura por sí sola todas las relaciones causales.
+Sabemos (\cite{koller2009probabilistic}) que una Red Bayesiana captura sólo las (mínimas) independencias de la distribución de probabilidad que representa, por lo que en la extensión al caso causal falta algo más que conecte a la medida de probabilidad que el grafo genera con las relaciones causales que existen en el espacio.
+
 \subsection{Modelos gráficos causales}
@@
 Más adelante entraremos en detalle respecto a este operador.
+
 \subsubsection{El problema de la identificabilidad} 
@@
 Se ha demostrado que si se cumple la condición (axioma) de Markov y de Fidelidad caual, entonces el grafo causal puede ser identificado hasta equivalencias Markovianas (arcos sin dirección) \cite{peters2012identifiability}, \cite{mooij2016distinguishing}.
+
 \subsubsection{Cálculo Do}
@@
 \end{teo}
-\indent La interpretación de estos puntos, en palabras de \cite{pearl2009causality} se sigue intepretar la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
+\indent La interpretación de estos puntos, en palabras de \cite{pearl2009causality} consiste en ver la acción $do(X=x)$ como la imposición de un nuevo mecanismo el cual induce un sub-modelo caracterizado por $\mathcal{G}_{\bar{X}}$.
 El primer punto se obtiene al considerar que eliminar ecuaciones en un sistema no introduce nuevas dependencias y muestra que la d-separación es un criterio válido para la independencia condicional en el nuevo grafo $\mathcal{G}_{\bar{X}}$.
-El punto dos provee una condición para que una intervención tenga el mismo efecto que una observación pasiva y el tercer punto provee una condición para poder introducir, o eliminar una intervención externa sin afectar la probabilidad de observar $Y=y$.
+El punto dos provee una condición para que una intervención tenga el mismo efecto que una observación pasiva y el tercer punto provee una condición para poder introducir, o eliminar una intervención externa sin afectar la probabilidad de observar $Y=y$.\\
+\\
+Este resultado conecta las distribuciones de intervención con los enunciados probabilistas que se obtienen del grafo, con lo que contestamos la pregunta que se había dejado abierta antes en la subsección \ref{relacion}.
 \begin{teo}{\cite{peters2017elements}}\\
@@
 Las reglas del cálculo do proveen una solución al problema de la identificabilidad:
-\begin{cor} Solución al problema de la identificabilidad
+\begin{cor} Solución al problema de la identificabilidad (\cite{pearl2009causality}).\\
 Una distribución $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ es identificable en un modelo gráfico causal $\mathcal{G}$ si existe una secuencia finita de transformaciones, donde cada una de las cuales corresponde a una de las reglas del teorema \ref{docalculus} que reduce $q$ a una expresión de probabilidad condicionada sólo en datos observacionales (i.e.
@@
 \subsection{Toma de decisiones}
+El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
+\\
+\indent Por otro lado, el Teorema de DeFinnetti supone conocida la utilidad para el agente y el resultado garantiza la existencia de una medida de probabilidad (subjetiva) tal que la relación de preferencias es equivalente a maximizar la utilidad respecto a esta medida.
+De esta manera, se resuelve parcialmente el problema de toma de decisiones, pues si por un lado se cuenta con las probabilidades de eventos inciertos el Teorema de von Neumann devuelve la utilidad, y si se cuenta con la utilidad para el tomador de decisiones, DeFinetti devuelve las probabiliddes, pero ¿por dónde empezar?
+\\
+\\
+\indent El gran logro de J.L.
+Savage fue derivar  la utilidad y la probabilidad en conjunto en vez de utilizarlas como primitivas del modelo (\cite{gilboa2009decision}).
+Lo único que requiere el Teorema de Savage es una serie de axiomas de coherencia sobre la relación de preferencias de un tomador de decisiones; con esto, muestra la existencia de una función de utilidad y de una medida de probabilidad tal que la relación de preferencias es equivalente a la maximización de la utilidad esperada.\\
+\\
+ \indent Es importante mencionar que existen otros criterios de optimalidad cuando los axiomas clásicos (vNM-Savage) son cuestionados; por ejemplo, la utilidad maxmin, o la utilidad esperada de Choquet (\cite{gilboa2009decision}).
+\cite{gilboa2001theory} propone un nuevo paradigma de toma de decisiones, llamado \textit{Case-Based Decision Theory}, en el cual los agentes toman decisiones haciendo analogías a eventos pasados en los cuales tuvieron un buen desempeño.
 \subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
@@
 Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no sólo las acciones y preferencias de un agente sino también las de otros.
-En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos Extensivos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene \textit{creencias} \textit{a priori} sobre algunos aspectos del ambiente.
+En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene \textit{creencias} \textit{a priori} sobre algunos aspectos del ambiente.
 Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.).\\
 \\
-\indent En general, los juegos Bayesianos pertenecen a una clase de juegos llamados juegos estratégicos.
-Por el momento, sólo veremos distintas clases de juegos para mostrar que el problema a tratar en esta propuesta cuenta con la maquinaria suficiente para ser modelado, pero se dejarán de lado otras cuestiones importantes de Teoría de Juegos como son el Equilibrio de Nash y las Estrategias Mixtas.
+\indent Por el momento, sólo veremos distintas clases de juegos para mostrar que el problema a tratar en esta propuesta cuenta con la maquinaria suficiente para ser modelado, pero se dejarán de lado otras cuestiones importantes de Teoría de Juegos como son el Equilibrio de Nash, Estrategias Mixtas, etc.
 \subsubsection{Comportamiento Racional}
 Vamos a asumir que los agentes involucrados tienen preferencias racionales en el sentido de los axiomas de Decisión que fueron mencionados.
+En la literatura, el conjunto de axiomas más utilizados para caracterizar comportamiento racional son los utilizados en el Teorema de von Neumann-Morgenstern pues se considera que las probabilidades de eventos están fijas y pueden ser conocidas objetivamente.
 \subsubsection{Juegos estratégicos y la forma normal}
@@
 Es deseable considerar el caso en el que un jugador no conoce todas las características del otro jugador; ya sea sus acciones o sus pagos, y es este desconocimiento lo que se pretende modelar con los Juegos Estratégicos Bayesianos.
-		\begin{defi}{Juego Estratégico Bayesiano en Forma Normal (\cite{osborne1994course})}\\
+\begin{defi}{\label{juegobayesiano}}Juego Estratégico Bayesiano en Forma Normal (\cite{osborne1994course})\\
 Un juego estratégico Bayesiano consiste en
@@
 El jugador actualiza sus creencias sobre $\omega \in \Omega$ a $P_i(\omega) / P_i(\tau^{-1}_i (t_i))$ si $\omega \in \Omega$ y cero en otro caso.
-Respecto a este tipo de juegos y su aplicabilidad en el problema del uso de relaciones causales, basta decir que lo que se pretende es que al no conocer el modelo causal verdadero, las creencias del tomador de decisiones sean sobre este modelo causal, estas creencias serán utilizadas por el agente como un modelo \textit{propio} que irá actualizando conforme el mundo revela algo de sí.
+Respecto a este tipo de juegos y su aplicabilidad en el problema del uso de relaciones causales, basta decir que lo que se pretende es que al no conocer el modelo causal verdadero, las creencias del tomador de decisiones sean sobre este modelo causal, estas creencias serán utilizadas por el agente como un modelo \textit{propio} que irá actualizando conforme el mundo revela algo de sí.\\
+\\
+Existen juegos que son capaces de representar situaciones más complejas; en particular, situaciones en las que no todos los involucrados deciden simultáneamente, sino que existe cierta estructura temporal.
+A estos juegos se les conoce como juegos extensivos, los cuales se dividen a su vez según la información disponible a los jugadores, lo que da lugar a los juegos extensivos con información perfecta, imperfecta, e incompleta según sea el caso.
+Veremos a continuación las definiciones respectivas.
+Referimos al lector a los siguientes textos clásicos\cite{osborne1994course}, \cite{shoham2008multiagent}, \cite{10.1007/978-94-010-0189-2_25}.
 \subsubsection{Juegos extensivos con información perfecta}
-Contenido tomado de \cite{shoham2008multiagent}
-\subsubsection{Juegos extensivos con información impefecta}
-Contenido tomado de \cite{shoham2008multiagent} y \cite{10.1007/978-94-010-0189-2_25}
-\subsubsection{Juegos extensivos con información incompleta}
-Contenido tomado de \cite{osborne1994course}
-
+Un juego extensivo con información incompleta modela situaciones en las que existe estructura \textit{secuencial} en la que los jugadores toman decisiones de manera alternada.
+Se dice que existe información perfecta si en cada ronda cada jugador conoce los eventos que han sucedido hasta ese momento.
+En esta clase de juegos no existe aleatoriedad alguna que genere incertidumbre respecto al estado de las cosas.
+\begin{defi}
+Un juego extensivo con información perfecta consiste de:
+\begin{itemize}
+\item Un conjunto  $N$ de jugadores.
+\item Un conjunto $A$ de acciones.
+\item Un conjunto $H$ de sucesiones (llamadas historias) que satisfacen:
+	\begin{itemize}
+		\item La sucesión vacía pertenece a $H$.
+		\item Si $(a_k)_{k=1}^K \in H$ ($K$ puede ser infinito) y $L < K$ entonces $(a_k)_{k=1}^L \in H$.
+		\item Si $(a_k)_{k=1}^\infty$ satisface que $(a_k)_{k=1}^L \in H$ para toda $L$, entonces $(a_k)_{k=1}^\infty \in H$
+	\end{itemize}
+\item Si una historia $(a_k)_{k=1}^K \in H$ tal que $K$ es finito y si no existe $K+1$ tal que $(a_k)_{k=1}^{K+1} \in H$ entonces la sucesión se dice que es un nodo terminal.
+Al conjunto de nodos terminales se le conoce como $Z$
+\item Una función $\chi : H \to 2^A$ que asignal un conjuntode acciones posibles a tomar después de cada historia $\chi(h)$.
+\item Una función $\rho$ que asigna a cada historia no-terminal tal que $\rho(h)$ es el jugador que escogerá una acción después de suceder la historia $h$.
+\item Una función $\sigma : H \times A \to H \cup Z$ que mapea cada historia y cada acción a una nueva historia tal que si $\sigma(h_1 , a_1) = \sigma (h_2,a_2)$ entonces $h_1 = h_2$ y $a_1 = a_2$
+\item Para cada jugador, una relación de preferencias $\succeq_i$ (equivalentemente, una función de utilidad $u_i$).
+\end{itemize}
+\end{defi}
+\subsubsection{Juegos extensivos con información imperfecta}
+Los juegos extensivos con información perfecta permiten que los jugadores conozcan la historia del juego hasta cualquier momento; en algunos escenarios es importante permitir que los jugadores no sepan cómo se llegó hasta cierto punto; es decir, existe ignorancia respecto a las acciones de otros jugadores.
+\begin{defi}
+Un juego (extensivo) con información imperfecta es una tupa $(N,A,H,Z,\chi,\rho,\sigma, u, I)$ donde
+\begin{itemize}
+\item $(N,A,H,Z,\chi,\rho,\sigma, u)$ es un juego extensivo con información imperfecta.
+\item $I=(I_1,...,I_n)$ donde para cada jugador $i$, se tiene que $I_i = \{ I_{i,1}, I_{i,2}, ...
+, I_{i,k_i} \}$ es una partición del conjunto $\{ h \in H: \rho(h)=i \}$ que satisface que $\chi(h)=\chi(h')$ entonces existe $j$ tal que $h \in I_{i,j}$ y $h \in I_{i,j}$.
+\end{itemize}
+Lo que representa cada partición es que un jugador no pueda distinguir entre nodos.
+\end{defi}
+\subsection{Representaciones más generales}
+\subsubsection{Juegos repetidos}
+Un juego repetido consiste en jugar el mismo juego (uno en forma normal) en repetidas ocasiones por los mismos jugadores.
+Este caso es importante para la propuesta de investigación que aquí se expone porque veremos que un tomador de decisiones debe aprender
+\subsubsection{Juegos estocásticos}
+De manera informal, un juego estocástico es una colección de juegos estratégicos en forma normal, a los que se conoce como juegos base, y los agentes juegan sucesivamente elementos de esta esta colección, y donde el juego que es jugado en cierta etapa depende probabilísticamente del juego anterior.
+Resulta que un juego estocástico generaliza los Proceso de Decisión Markovianos así como los juegos repetidos, pues un MDP es un juego estocástico de un jugador mientras que un juego repetido es un juego estocástico en el que sólo hay un juego base.
+\subsection{Aprendizaje en Juegos}
 \section{Trabajo Previo}
@@
 \\
-\indent \cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más chicos que se resuelven de manera secuencial.\\
+\indent \cite{koller2003multi} utilizan diagramas de influencia para encontrar estrategias relevantes en juegos multi-agente, con lo cual descomponen un juego grande en varios juegos más pequeños que se resuelven de manera secuencial.\\
 \\
@@
 \subsection{Identificación de intervenciones óptimas}
+\cite{bubeck2009pure}
+\cite{gabillon2012best}
+\cite{agarwal2014taming}
+
 \indent Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
@@
 Esta medida de optimalidad no está relacionada con ninguna función de utilidad asociada a un agente y esto restringe que el mismo problema se resuelva con distintos fines en mente.
-En el artículo consideran dos casos posibles: un caso en el cual hay $N$ causas, independientes entre sí, y una variable a la que estas causas afectan; en el caso general, en el cual las causas pueden afectarse entre sí, los autores utilizan un estimador por importancia truncado para la selección de acciones en cada ronda con un nivel fijo de truncamiento.
+En el artículo consideran dos casos posibles: un caso en el cual hay $N$ causas independientes entre sí y una variable a la que estas causas afectan; en el caso general, en el cual las causas pueden afectarse entre sí, los autores utilizan un estimador por importancia truncado para la selección de acciones en cada ronda con un nivel fijo de truncamiento.
 En este trabajo se asume que el modelo causal es conocido y plantean como trabajo futuro el caso en el cual se desconoce el modelo, que es precisamente el caso que se busca atacar aquí.\\
 \\
-\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como bandits; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas, pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
+\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como \textit{bandits}; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas, pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
 Este trabajo se encuentra en la intersección de inferencia causal y de descubrimiento de mejor brazo bajo restricciones presupuestales; en este sentido encontramos los trabajos de \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}.
@@
 \\
-\cite{tingley2010belief} muestran, experimentalmente y para una clase de juegos particular que los agentes involucrados actualizan sus creencias respecto a ciertos aspectos de los otros jugadores; encuentran que los jugadores actualizan sus creencias de manera que no se alejan mucho del \textit{prior} inicial.
+\cite{tingley2010belief} muestran, experimentalmente y para una clase de juegos particular que los agentes involucrados actualizan sus creencias respecto a ciertos aspectos de los otros jugadores; encuentran que los jugadores actualizan sus creencias de manera que no se alejan mucho del \textit{prior} inicial.\\
+\subsection{Aprendizaje activo para descubrimiento causal}
+Otro enfoque propuesto para el descubrimiento de relaciones causales es a través de \textit{active learning}, encontramos, como base, los trabajos de \cite{tong2001active},\cite{murphy2001active}, \cite{meganck2006learning}, \cite{he2008active}, \cite{hauser2012two}, \cite{10.1007/978-3-319-56970-3_9}, \cite{rubenstein2017probabilistic}.
+Estos métodos parten de un grafo inicial, y proceden a seleccionar instancias de los datos que permitan ir refinando este grafo de modo que se orienten todos los arcos.\\
+\subsection{Conclusión}
+A manera de cierre de esta sección, resaltamos la importancia inmediate de los trabajos de \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, pues en ellos lo que se busca es encontrar la mejor acción a tomar en un entorno incierto, pero en el cual existen relaciones causales y se hace uso de estas relaciones para encontrar la mejor acción.
+Ambos trabajos suponen conocido el modelo gráfico causal; en el caso de \cite{lattimoreNIPS2016} ellos consideran que el modelo es completamente conocido, mientras que en \cite{sen2017identifying} sólo requieren conocer una parte.
+Lo que nosotros proponemos es descubrir el modelo a la vez que se utiliza este mismo para la elección de decisiones.
+Además, al modelar la interacción con el entorno, queda abierta la posibilidad de considerar comportamiento estratégico así como otros objetivos a maximizar por parte de la Naturaleza.
+Por ejemplo, si se estuviera modelando un juicio o  una negociación salarial la Naturaleza podría ser un gobierno o un sistema legal, que además de actuar a partir del modelo causal, busca maximizar ciertas cantidades.
 
 \section{Pregunta de investigación}
-En un entorno incierto para el cual se cumplen una serie de relaciones causales ¿cómo puede un agente inteligente adquirir y utilizar conocimiento sobre estas relaciones a partir de su interacción con el entorno en un proceso de toma de decisiones?
+En un entorno incierto para el cual se cumplen una serie de relaciones causales ¿cómo puede un agente inteligente adquirir y utilizar información acerca de estas relaciones a partir de su interacción con el entorno en un proceso de toma de decisiones para aprender más rápido a tomar mejores decisiones?
+\subsection{Hipótesis}
+Sea $\mathcal{G}$ un modelo gráfico causal y consideremos un problema de decisión bajo incertidumbre que involucra variables del modelo causal.
+Entonces, si modelamos el problema de decisión como un juego entre el tomador de decisiones y la Naturaleza, y si el tomador de decisiones actua racionalmente, entonces el uso de información causal por parte del tomador de decisiones lo llevará a escoger sus acciones óptimas (en el sentido de maximización de utilidad) más rápido que si no considerara la información causal del entorno.
+
 \section{Planteamiento y Modelado}
@@
 \indent Para modelar la interacción del tomador de decisiones con su entorno, vamos a asumir que el agente está interactuando con un ente abstracto al que le llamaremos Naturaleza y que la dinámica de interacción consiste en un \textit{juego}.
-A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.
-En el contexto de Teoría de Juegos, los juegos que permiten que el agente tenga creencias información incompleta se llaman \textit{Juegos Estocásticos}, los cuales fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+A diferencia de lo que se hace en Aprendizaje por Refuerzo en donde se considera que el agente está “solo con su entorno”, aquí consideramos que todo lo que proviene de afuera del agente son jugadas de este ente abstracto.\\
 \\
@@
 Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
-Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales.\\
+Lo que esperamos tener al final de un cierto número de rondas de aprendizaje es un \textit{diagrama de influencia causal}, que consistirá en la extensión al caso causal de los diagramas de influencia tradicionales, así como una política óptima en el sentido de la utilidad esperada para el agente.\\
 \\
@@
 En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
-En el trabajo de este autor,  el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
-
-\subsection{Hipótesis}
-Bajo ciertas condiciones, un agente inteligente puede adquirir y utilizar conocimiento causal a partir de su interacción con un entorno incierto.
+En el trabajo \cite{eberhardt2008causal} el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
 
@@
 	\subsection{Objetivo General}
-	La investigación que se propone llevar a cabo tiene dos objetivos generales; por un lado, encontrar la manera en la que un agente que se enfrenta a un entorno incierto pueda usar información causal de manera que ésta sea de ayuda al llevar a cabo una tarea; por otro lado, encontrar la manera en la que un agente adquiera información causal conforme interactúa con su entorno.
+	La investigación que se propone llevar a cabo tiene por objetivo general encontrar la manera en la que un agente que se enfrenta a un entorno incierto descubra y utilice información causal de manera que ésta lo ayude a encontrar la mejor decisión a tomar y de una manera más rápida.
 	\subsection{Objetivos específicos}
-	Para llevar satisfacer los objetivos generales, hasta el momento se ha identificado que será necesario lo siguiente:
+	Para llevar satisfacer el, hasta el momento se ha identificado que será necesario lo siguiente:
 \begin{itemize}
-\item Una forma de especificar creencias sobre estructura causal; esto es, encontrar una familia de distribuciones que representen de manera coherente creencias o conocimiento previo acerca de la estructura causal de un entorno.
-Para esto, existen dos casos a considerar; el primero, es suponer que el tomador de decisiones conoce el número máximo de variables que el modelo causal puede tener; en este caso, las creencias del agente pueden estar representadas por una distribución sobre un espacio de matrices de tamaño fijo, aprovechando la relación que existe entre grafos acíclicos dirigidos y matrices.
+\item Una forma general de especificar un juego que capture los aspectos esenciales de un problema de decisión y cuya solución refleje la mejor decisión para el agente.
+Este juego debe permitir que el jugador tenga creencias sobre aspectos estructurales del ambiente.
+
+\item Una manera de \textit{utilizar} el conocimiento acerca de las relaciones causales que el agente tiene en un momento dado para con ello tomar una buena decisión.
+
+\item Cuando el agente no conoce el modelo causal, se debe encontrar una forma de especificar creencias sobre estructura causal; esto es, encontrar una familia de distribuciones que representen de manera coherente creencias o conocimiento previo acerca de la estructura causal de un entorno.
+
+\item Se debe tener una distribución inicial de creencias, que permitan que el agente sea coherente con los axiomas de decisión.
+En \cite{billot2005probabilities} se muestra, de manera axiomática, cómo asignar probabilidades iniciales dado un conjunto de observaciones pasadas.\\
+
+\item Para esto, existen dos casos a considerar; el primero, es suponer que el tomador de decisiones conoce el número máximo de variables que el modelo causal puede tener; en este caso, las creencias del agente pueden estar representadas por una distribución sobre un espacio de matrices de tamaño fijo, aprovechando la relación que existe entre grafos acíclicos dirigidos y matrices.
 Por otro lado, si el agente desconoce el número máximo de variables, entonces el grafo causal puede tener cualquier tamaño.
-En este caso, el camino a explorar es el considerar distribuciones de probabilidad sobre espacios de dimensión infinita, las cuales son conocidas como priors Bayesianos no paramétricos.
-\item Una manera de \textit{utilizar} las creencias causales que el agente tiene en un momento dado.
-\item Un criterio de actualización de creencias causales que debe converger al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
+En este caso, el camino a explorar es el considerar distribuciones de probabilidad sobre espacios de dimensión infinita, las cuales son conocidas como \textit{priors} Bayesianos no paramétricos \cite{ghosal2017fundamentals}.
+
+\item Una manera de actualización de creencias causales que debe converger al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
 Es importante mencionar que los juegos Bayesianos, en su definición estándar, proveen un mecanismo para la actualización general de creencias, pero se necesita una manera de actualizar creencias que haga uso de la estructura causal misma.
-\item Un criterio de solución del problema: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
+Además, es importante tener en cuenta que no cualquier creencia puede ser modelada por un \textit{prior bayesiano} \cite{gilboa2016ambiguity}.
+
+\item Un criterio de solución del problema de decisión: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
+
 \item Una manera de construir un diagrama de influencia causal que además permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
+
 \end{itemize}
+\subsection{Medición}
 \section{Metodología}
@@
 \item Consideramos un problema de decisión bajo incertidumbre que se desarrolla en un entorno regido por un modelo causal.
-\item Un primer paso es modelar las acciones del agente y la respuesta del entorno se modelan como un juego de dos jugadores, uno de los cuales será el tomador de decisiones inicial, y el otro será un ente abstracto llamado Naturaleza que, suponemos, conoce el modelo causal y lleva a cabo sus jugadas a partir de lo dictado por éste.
-Los juegos sucederán de manera repetida para poder aprender de éste.
-\item En este momento, suponemos que la Naturaleza es indiferente a cualquier resultado y para esto se le asignan pagos iguales en cada resultado del juego.
-Esto puede modificarse en un futuro para modelar casos en los que la Naturaleza tenga ciertas \textit{intenciones}.
-\item Como caso inicial, consideramos que el modelo causal es conocido en su totalidad para el tomador de decisiones.
-Para este caso, se encontrará un criterio de solución y se comparará con algoritmos del estado del arte de detección de acciones óptimas en un modelo causal.
-\item En una siguiente etapa, se considera que el tomador de decisiones sólo conoce la \textit{estructura} del modelo causal.
-Para este caso, se encontrará una forma de aprender los valores de los parámetros del modelo así como una manera de usar esta información en futuras decisiones a tomar.
-Para este segundo caso, se debe construir una manera de expresar el conocimiento inicial acerca de la estructura que además de ser coherente con la manera en la que los seres humanos formamos creencias (\cite{larrouy2017mindreading}).
-\item Para llevar a cabo el descubriminto del modelo, se define un número de juegos de observación, dentro de estos juegos el tomador de decisiones no modificará sus creencias causales, aunque éstas lo obliguen a tener un desempeño no óptimo.
-\item Posteriormente, en el caso en el que el agente no conoce nada, sus creencias serán lo único con lo que pueda operar.
-Este conocimiento causal inicial, en el caso general, debe ser expresado en términos de probabilidades y para esto se construirá un \textit{prior} que consiste en una distribución sobre un espacio de matrices.
-Recordemos que una matriz (que cumple ciertas características) define un grafo, por lo que tener de esta manera estaremos especificando las probabilidades que un agente asigna, de manera subjetiva, a distintos modelos causales.
-Todavía no está determinado si consideraremos que el tomador de decisiones conoce el número máximo de variables que puede tener el modelo o no.
-De esto dependerá la \textit{clase} de distribución que se deba usar.
+\item Primero, para modelar correctamente la interacción entre el agente y su entorno, debemos estudiar qué clase de juegos permite esto.
+Además, estudiar las soluciones a esta clase de juegos y escoger una, o varias, cuya interpretación, así como su valor, sea la más cercana a la maximización de la utilidad esperada.
+
+\item Una vez definido la clase a la que pertenece el juego entre el tomador de decisiones y la naturaleza, se define una manera de asignar pagos, o utilidad, a cada uno de los jugadores.
+Esto permitirá que la Naturaleza también tenga intenciones que desee cumplir.
+
+\item Una vez que se fijan los detalles del juego, pasamos a considerar lo que el tomador de decisiones puede conocer acerca del modelo causal.
+
+\item Relativo a esto, consideramos tres casos:
+
+\item En un primer caso supondremos que el tomador de decisiones conoce completamente el modelo causal y lo usará para conocer los efectos de sus acciones y con esto tomar la acción óptima en el sentido del segundo punto.
+
+\item Posteriormente, se considera que el tomador de decisiones sólo conoce la \textit{estructura} del modelo causal.
+En este caso, se debe establecer una forma para que el jugador tenga \textit{creencias} probabilistas sobre los parámetros del modelo y un criterio de acción teniendo esta información.
+En este punto deben responderse preguntas como: ¿cuántos juegos de aprendizaje?
+además de establecer una forma de comparación respecto a otros algoritmos, por ejemplo los que aparecen en \cite{lattimoreNIPS2016}.
+
+\item Luego, se considera el caso en el que el agente no conoce nada sobre el modelo.
+Debido a que sus creencias serán lo único con lo que pueda operar.
+Se debe definir una manera de establecer creencias sobre la estructura del grafo y una manera de uso de estas creencias en un momento particular.
+Además, utilizar la medición del punto anterior para evaluar desempeño
+
 \item En términos del conocimiento causal obtenido hasta cierto momento dentro de la serie de juegos o interacciones, se debe definir una estrategia para el tomador de decisiones a partir del conocimiento causal que tenga hasta ese momento; una \textit{estrategia mixta} permitiría explorar el ambiente con tal de explorar el ambiente de manera amplia.
-\item A partir de lo observado en la ronda de juegos de observación, se debe actualizar las creencias causales de manera que se reflejen las relaciones que pudieron ser observadas, las cuales serán almacenadas en un diagrama de influencia.
-\item Modificar la estrategia considerando los efectos posibles de cada acción que fueron aprendidos, para esto se usará el Diagrama de Influencia previamente construido. 
-\item Comparar el desempeño versus lo que se haría si se conociera el modelo con detalle.
 
@@
 \item En el tercer caso, el agente no conocerá nada del modelo, pero tendrá \textit{creencias} sobre este; este caso generaliza los anteriores, pues en el primer y segundo caso las creencias con las que inicia el agente es la información que conoce sobre el modelo.
-En resumen, en los tres casos el agente contará con \textit{creencias} sobre el modelo causal verdadero; en el primer caso, sus creencias coincidirán plenamente con el modelo; en el segundo caso, sólo con la estructura y en el tercer caso, no tenemos garantía de nada.
+\item  En resumen, en los tres casos el agente contará con \textit{creencias} sobre el modelo causal verdadero; en el primer caso, sus creencias coincidirán plenamente con el modelo; en el segundo caso, sólo con la estructura y en el tercer caso, no tenemos garantía de nada.
 \end{itemize}
@@
 Entonces, lo que el agente tendrá es una distribución de probabilidad sobre los posibles valores que estos parámetros puedan tener.
+En este caso, en cada etapa (juego) el agente puede tomar una decisión óptima relativa a su conocimiento hasta ese momento, por lo que es posible llevar a cabo cada juego.\\
+\\
+A partir de lo que la Naturaleza lleve a cabo, se contrastarán los valores que el jugador cree contra lo que la evidencia indica y a partir de esto el modelo causal se irá modificando en cada juego.
 \subsection{Caso general}
@@
 \indent Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
-Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).\\
-\\
-\indent Tomando esto en cuenta, una de las contribuciones esperadas de esta investigación es construir un procedimiento de actualización de creencias acerca de la estructura causal de cierto entorno sobre el cual un agente está actuando.
-
+Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).
 \section{Plan de trabajo (cronograma)}
 Por definir.
-
 \section{Conclusiones}

### 2018-04-25 Segundo_Semestre/Seminario/Semana12/Proposal.tex
cambios en la intro

@@
 \\
-\indent Aprender del entorno se formula usualmente como un agente autónomo que se encuentra dentro de un entorno o ambiente y que tomar acciones en este entorno modifican el \textit{estado} de éste (\cite{sutton1998reinforcement}).
-A su vez, tal modificación del entorno es percibida por el agente en términos de una señal de recompensa y de encontrarse en otro estado.
-En este trabajo se propone modelar la interacción entre el agente y el entorno como una interacción entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza; este planteamiento permite, de manera transparente, pensar que el jugador Naturaleza está guiado por un modelo causal.\\
+\indent La clase de problemas en los cuales un agente debe aprender por interacción se formulan, usualmente, como un agente autónomo que se encuentra dentro de un entorno o ambiente el cual es modificado por las acciones del agente (\cite{sutton1998reinforcement}).
+A su vez, tal modificación del entorno afecta al agente en términos de una señal de recompensa.
+En este trabajo se propone modelar la interacción entre el agente y el entorno como una \textit{juego}\footnote{Un juego es una situación estratégica en la que los intereses de dos o más actores se enfrentan y se afectan entre sí.} entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza, cuyas acciones estarán controladas por un modelo causal.\\
 \\
-\indent En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con Nash.}.
+\indent La manera en la que se propone que el agente tome en cuenta, aprenda y utilice la información causal del entorno es dotarlo de \textit{creencias} sobre la estructura causal, las cuales serán utilizadas para la toma de decisiones.
+En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con Nash.}.
 Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.

### 2018-04-25 Segundo_Semestre/Seminario/Semana12/Proposal.tex
hipotesis y game learning

@@
 \\
-\indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del “razonamiento contrafactual” para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+\indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del \textit{razonamiento contrafactual} para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
 Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
@@
 
-\subsubsection{Una medida de probabilidad a partir del grafo}
+\subsubsection{Una medida de probabilidad a partir del grafo}{\label{axiomas}}
 Dado un grafo acíclico dirigido, podemos construir una medida de probabilidad que expresa las relaciones que se encuentran establecidas en el grafo, la cual denotaremos $P_{\mathcal{G}}$.
@@
 Un juego repetido consiste en jugar el mismo juego (uno en forma normal) en repetidas ocasiones por los mismos jugadores.
-Este caso es importante para la propuesta de investigación que aquí se expone porque veremos que un tomador de decisiones debe aprender
+Por ejemplo, jugar de manera secuencial, por los mismos jugadores, el dilema del prisionero o cualquier otro juego conocido.
+En esta clase de situaciones, se puede considerar que el pago final es la suma de los pagos de cada ronda, descontados o no.
+En estas situaciones vale la pena preguntar si los jugadores van a realizar siempre las soluciones de equilibrio, o si en el largo plazo llevarán a cabo otras estrategias.Este caso es importante para la propuesta de investigación que aquí se expone porque veremos que un tomador de decisiones debe aprender
 \subsubsection{Juegos estocásticos}
@@
 \subsection{Aprendizaje en Juegos}
+¿Cómo después de observar una secuencia de juegos los tomadores de decisiones pueden aprender algo sobre estos?
+Tal vez al observar las acciones de los otros agentes puedan inferir sus estrategias y plantear una \textit{mejor respuesta} a la estrategia aprendida.
+Para llevar esto a cabo existen varias técnicas, como el \textit{aprendizaje ficticio}, el \textit{aprendizaje Bayesiano o racional} e incluso algoritmos de aprendizaje por refuerzo.
+Pueden consultarse con más detalle en \cite{shoham2008multiagent}, \cite{fudenberg1998theory}.
+En particular, se tiene el siguiente resultado de interés
+\begin{teo}
+Sea $s$ un perfil de estrategias para un juego repetido de $n$ jugadores y sea $P=P_1,...,P_n$ donde $P_i$ son las crrencias del jugador $i$.
+Sean $\mu_s, \mu_p$ distribuciones sobre las historias infinitas de los juegos inducidas por el perfil $s$ y las probabilidades $P$ Si se cumple que:
+\begin{itemize}
+\item En cada ronda el jugador $i$ juega una \textit{mejor respuesta} dadas sus creencias $P_i$
+\item Después de cada ronda el jugador actualiza Bayesianamente.
+\item La medida $\mu_s$ es absolutamente continua respecto a $\mu_P$.
+\end{itemize}
+Entonces, para toda $\varepsilon>0$ y para casi toda historia en el soporte de $\mu_s$ existe un tiempo $T$ tal que para toda $t \geq T$, la jugada $\mu_{P_i}$ predicha por las creencias del jugador $i$ están $\varepsilon$-cercanas de la distribución $\mu_S$ predicha por las estrategias actuales.
+\end{teo}
 \section{Trabajo Previo}
@@
 \subsection{Hipótesis}
-Sea $\mathcal{G}$ un modelo gráfico causal y consideremos un problema de decisión bajo incertidumbre que involucra variables del modelo causal.
-Entonces, si modelamos el problema de decisión como un juego entre el tomador de decisiones y la Naturaleza, y si el tomador de decisiones actua racionalmente, entonces el uso de información causal por parte del tomador de decisiones lo llevará a escoger sus acciones óptimas (en el sentido de maximización de utilidad) más rápido que si no considerara la información causal del entorno.
+Sea $\mathcal{G}$ un modelo gráfico causal cuya distribución de probabilidad asociada $P_\mathcal{G}$ satisface los axiomas mencionados en la sección \ref{axiomas} y consideremos un problema de decisión bajo incertidumbre $(D,E,C)$ cuyos eventos inciertos están relacionados variables del modelo causal $\mathcal{G}$.
+Además, el tomador de decisiones tiene preferencias racionales sobre las consecuencias de sus decisiones.
+Entonces, mediante la interacción con el entorno el tomador de decisiones podrá adquirir y hacer uso de información causal para  escoger sus acciones óptimas (en el sentido de maximización de utilidad) más rápido que si no considerara la información causal del entorno.
 
@@
 \item Considerar juegos de mayor duración.
-\item Considerar varios jugadores además de la naturaleza.
+\item Considerar varios jugadores además de la Naturaleza.
 \item Considerar que la Naturaleza tiene algún tipo de intenciones o de objetivo a maximizar; por ejemplo, el concepto de Naturaleza que hemos usado podría bien representar a un gobierno o un Banco central.
+Por ejemplo, en una negociación salarial los jugadores podrían ser el empleador, los empleados y el gobierno; en este ejemplo, la Naturaleza sería el gobierno y el modelo causal las leyes actuales que definen el actuar del gobierno, quien podría tener el interés de maximizar el número de personas con empleo.
 \item Estudiar equilibrios en este tipo de juegos en términos de la estructura del modelo causal. 

### 2018-04-26 Segundo_Semestre/Seminario/Semana12/Proposal.tex
metodologia de trabajo en parrafos

@@
 \maketitle
+\tableofcontents
+\newpage
 \section{Introducción}
@@
 \indent La manera en la que se propone que el agente tome en cuenta, aprenda y utilice la información causal del entorno es dotarlo de \textit{creencias} sobre la estructura causal, las cuales serán utilizadas para la toma de decisiones.
-En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con Nash.}.
+En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con John Nash.}.
 Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.
@@
 \end{teo}
-\subsection{Teoría de Juegos}
+\subsection{Teoría de Juegos}{\label{juegos}}
 Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no sólo las acciones y preferencias de un agente sino también las de otros.
@@
 \indent \cite{soares2015toward} compara, desde un punto de vista epistemológico, la toma de decisiones apoyada en evidencia versus el uso de uso de razonamiento contrafactual en la toma de decisiones, aunque conceptualmente esta última permite llevar a cabo razonamientos más complejos el problema que se presenta en la práctica es precisamente la construcción del modelo causal a partir del cual se van a tomar decisiones.
-\subsection{Identificación de intervenciones óptimas}
-\cite{bubeck2009pure}
-\cite{gabillon2012best}
-\cite{agarwal2014taming}
-
-\indent Por otro lado, desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
+\subsection{Identificación de intervenciones óptimas: bandits}
+Desde un punto de vista computacional, \cite{lattimoreNIPS2016} atacan el problema de toma de decisiones en entornos causales pero bajo un modelo causal conocido; en este modelo, las acciones a escoger son intervenciones posibles en el grafo causal y lo que se busca es maximizar cierta variable de recompensa que está influenciada directamente por el modelo causal.
 La optimalidad de su acción está determinada en términos del \textit{regret} mínimo, que es básicamente un valor esperado.
@@
 \\
-\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como \textit{bandits}; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas, pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
+\indent El trabajo de \cite{lattimoreNIPS2016}, en general, considera el aspecto de toma de decisiones en línea en un entorno incierto y modelan esto como \textit{bandits}; al respecto, existen una serie de trabajos que son capaces de encontrar el \textit{mejor brazo} a escoger en un número finito de rondas (\cite{bubeck2009pure}, \cite{gabillon2012best}, \cite{agarwal2014taming}), pero estos modelos son puramente asociativos y no utilizan información adicional del ambiente, que en el caso de los Lattimore esta información adicional está dada por las relaciones causa efecto que ocurren en el ambiente.
 Este trabajo se encuentra en la intersección de inferencia causal y de descubrimiento de mejor brazo bajo restricciones presupuestales; en este sentido encontramos los trabajos de \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}.
@@
 \section{Pregunta de investigación}
-En un entorno incierto para el cual se cumplen una serie de relaciones causales ¿cómo puede un agente inteligente adquirir y utilizar información acerca de estas relaciones a partir de su interacción con el entorno en un proceso de toma de decisiones para aprender más rápido a tomar mejores decisiones?
+En un entorno incierto para el cual se cumplen una serie de relaciones causales que quedan expresadas por un modelo gráfico causal finito $\mathcal{G}$, ¿cómo puede un agente inteligente con prerefencias racionales adquirir y utilizar información acerca de estas relaciones a partir de su interacción con el entorno en un proceso de toma de decisiones para aprender más rápido a tomar decisiones que le devuelvan la máxima utilidad esperada?
 \subsection{Hipótesis}
@@
 
-\section{Planteamiento y Modelado}
+\section{Planteamiento y Modelado}{\label{planteamientoymodelado}}
 Suponemos un ambiente que está regido por un modelo causal, y suponemos que este es desconocido, pero fijo y además cumple que todas las variables se encuentran contenidas en él (suficiencia causal).\\
@@
 \subsection{Objetivo General}
-	La investigación que se propone llevar a cabo tiene por objetivo general encontrar la manera en la que un agente que se enfrenta a un entorno incierto descubra y utilice información causal de manera que ésta lo ayude a encontrar la mejor decisión a tomar y de una manera más rápida.
+La investigación que se propone llevar a cabo tiene por objetivo general encontrar la manera en la que un agente que se enfrenta a un entorno incierto descubra y utilice información causal de manera que permita encontrar la mejor decisión a tomar y de una manera más rápida que si no se hiciera uso de relaciones causales.
 \subsection{Objetivos específicos}
@@
 
-\item Una manera de \textit{utilizar} el conocimiento acerca de las relaciones causales que el agente tiene en un momento dado para con ello tomar una buena decisión.
+\item Garantías teóricas de la posibilidad de aprender  algo del ambiente en este juego.
+
+\item Una manera de \textit{utilizar} el conocimiento acerca de las relaciones causales que el agente tiene en un momento dado para con ello tomar una buena decisión relativo a la utilidad esperada de ésta.
 
@@
 \subsection{Medición}
+¿Cómo concluir que la investigación propuesta cumplió sus objetivos?
+Dado un modelo causal y un agente que se enfrenta a un problema de decisión relativo al modelo causal debe cumplirse que el agente, al aprender con el método propuesto, se desempeñe mejor que sólo aplicando algoritmos asociativos de identificación de acción óptima.
 \section{Metodología}
 \subsection{Metodología de Trabajo}
-\begin{itemize}
-\item Consideramos un problema de decisión bajo incertidumbre que se desarrolla en un entorno regido por un modelo causal.
-\item Primero, para modelar correctamente la interacción entre el agente y su entorno, debemos estudiar qué clase de juegos permite esto.
+Para el desarrollo de este trabajo de investigación se considera un problema de decisión cuyos eventos inciertos están regidos por un modelo causal el cual es fijo, y desconocido para el tomador de decisiones.
+Se asume que el tomador de decisiones no conoce los parámetros del modelo y por lo tanto las probabilidades de los eventos inciertos, por lo que se ve obligado a aprenderlos a través de la interacción con un ente abstracto llamado Naturaleza como quedó explicado en la sección \ref{planteamientoymodelado}.
+Una vez que ha quedado fijo el contexto procedemos a detallar cómo será abordado este problema.\\
+\\
+\indent Un primer paso es definir qué clase de juegos (dentro de los descritos en la sección \ref{juegos}) permite modelar las interacciones mencionadas anteriormente.
 Además, estudiar las soluciones a esta clase de juegos y escoger una, o varias, cuya interpretación, así como su valor, sea la más cercana a la maximización de la utilidad esperada.
+Una vez que se ha encontrado (o, en su caso, definido) el juego que de manera más precisa describe la situación se debe definir una manera para asignar pagos o utilidad a cada uno de los jugadores; esto permitirá que la Naturaleza también tenga intenciones que desee cumplir.\\
+\\
+\indent Ahora, para poder adentrarnos en el aspecto de toma de decisiones, considerando que no se conocen los parámetros de modelo causal, dotamos al agente de \textit{creencias} sobre la estructura causal, las cuales tendrán la forma de una distribución de probabilidad.
+Sobre estas creencias y su construcción es necesario analizar aspectos de aprendizaje con el fin de obtener garantías teóricas de que es posible aprender un modelo causal al utilizar dichas creencias en la manera que han sido asignadas.\\
+\\
+Para simplificar el proceso de solución a este problema, se consideran tres casos de dificultad incremental:
+\begin{enumerate}
+\item El tomador de decisiones conoce completamente el modelo causal y lo usará para conocer los efectos de sus acciones y con esto tomar la acción óptima en el sentido del segundo punto.
 
-\item Una vez definido la clase a la que pertenece el juego entre el tomador de decisiones y la naturaleza, se define una manera de asignar pagos, o utilidad, a cada uno de los jugadores.
-Esto permitirá que la Naturaleza también tenga intenciones que desee cumplir.
-
-\item Una vez que se fijan los detalles del juego, pasamos a considerar lo que el tomador de decisiones puede conocer acerca del modelo causal.
-
-\item Relativo a esto, consideramos tres casos:
-
-\item En un primer caso supondremos que el tomador de decisiones conoce completamente el modelo causal y lo usará para conocer los efectos de sus acciones y con esto tomar la acción óptima en el sentido del segundo punto.
-
-\item Posteriormente, se considera que el tomador de decisiones sólo conoce la \textit{estructura} del modelo causal.
+\item El tomador de decisiones sólo conoce la \textit{estructura} del modelo causal.
 En este caso, se debe establecer una forma para que el jugador tenga \textit{creencias} probabilistas sobre los parámetros del modelo y un criterio de acción teniendo esta información.
-En este punto deben responderse preguntas como: ¿cuántos juegos de aprendizaje?
+En este punto, deben responderse preguntas como: ¿cuántos juegos de aprendizaje?
 además de establecer una forma de comparación respecto a otros algoritmos, por ejemplo los que aparecen en \cite{lattimoreNIPS2016}.
 
-\item Luego, se considera el caso en el que el agente no conoce nada sobre el modelo.
+\item El tomador de decisiones no conoce nada sobre el modelo.
 Debido a que sus creencias serán lo único con lo que pueda operar.
 Se debe definir una manera de establecer creencias sobre la estructura del grafo y una manera de uso de estas creencias en un momento particular.
-Además, utilizar la medición del punto anterior para evaluar desempeño
+Además, utilizar la medición del punto anterior para evaluar desempeño.
+\end{enumerate}
 
-\item En términos del conocimiento causal obtenido hasta cierto momento dentro de la serie de juegos o interacciones, se debe definir una estrategia para el tomador de decisiones a partir del conocimiento causal que tenga hasta ese momento; una \textit{estrategia mixta} permitiría explorar el ambiente con tal de explorar el ambiente de manera amplia.
+\indent Para resolver el primer caso, utilizamos que el tomador de decisiones conoce completamente el modelo causal, con lo cual puede calcular las probabilidades de los eventos.
+Esto se logra mediante un algoritmo que simule a partir de un modelo causal dado, lo que permitirá calcular las probabilidades de efectos dadas las causas.
+Se comparan los resultados obtenidos con los algoritmos de identificación de intervenciones óptimas definidos antes.
+Para el segundo caso, el agente contará creencias sobre los parámetros del modelo, pues conoce la estructura de éste, por lo que paso a paso, a lo largo de varios juegos, utilizará el caso anterior para actuar y modificará sus creencias con la información que reciba.
+La calidad del modelo causal obtenido en el paso $n$ se mide al comparar con lo que obtendrían los algoritmos mencionados de identificación de mejor brazo así como de intervenciones óptimas, ademas de comparar con lo que se obtendría de conocer el modelo causal verdadero.
+Esto se detalla más adelante en la sección \ref{resultados}.\\
+\\
+\indent Para resolver el tercer caso, el agente deberá tener creencias sobre los parámetros y sobre la estructura del modelo.
+En cada juego, utiizará estas creencias \textit{como si} fueran el modelo verdadero y con ello tomará la mejor acción.
+Estas creencias serán modificadas según la información que reciba del ambiente.
+Se debe definir una forma de actualización de creencias que cumpla la propiedad de \textit{consistencia asintótica} (\cite{ghosal2017fundamentals}).
+Además, estas creencias deben ser consistentes con los criterios de actualización racional de creencias detallados en \cite{shoham2008multiagent}.
+Para el tercer caso, debe evaluarse la utilidad que aportan las acciones escogidas con esta información, la cual debe converger a su vez a la utilidad máxima esperada y para esto se usa, como entorno de prueba, el modelo causal verdadero del ambiente.\\
+\\
+\indent Una vez establecida la metodología de trabajo, veremos ahora cómo estos se llevarían a cabo en un caso específico de ejemplo
 
-\end{itemize}
-\section{Resultados Preliminares}
+\section{Resultados Preliminares}{\label{resultados}}
 Para mostrar la factibilidad del modelo planteado, consideramos un escenario particular y tres casos por separado cuyo grado de dificultad es ascendente.
@@
 \indent Debido a que más pacientes llegarán como este, el médico debe aprender de cada caso para desempeñarse mejor en los eventos futuros. 
-\subsection{Caso en el que se conoce el modelo completo}
+\subsubsection{Caso en el que se conoce el modelo completo}
 Consideremos el caso en el que el modelo gráfico causal es completamente conocido por el tomador de decisiones (el médico en este caso), tanto en su estructura como en sus parámetros.
@@
 \begin{itemize}
-\item Un modelo gráfico causal conocido para el médico.
+\item Un modelo gráfico causal conocido para el tomador de decisiones (médico).
 \item Jugadores: Naturaleza y el médico.
@@
 \end{tikzpicture}
-\subsection{Caso en el que se conoce la estructura del modelo}
+\subsubsection{Caso en el que se conoce la estructura del modelo}
 Consideremos ahora el caso en el cual el tomador de decisiones conoce la estructura del modelo, pero no conoce los valores de los parámetros de éste.
@@
 A partir de lo que la Naturaleza lleve a cabo, se contrastarán los valores que el jugador cree contra lo que la evidencia indica y a partir de esto el modelo causal se irá modificando en cada juego.
-\subsection{Caso general}
+\subsubsection{Caso general}
 En el caso general en el cual el tomador de decisiones no conoce el modelo causal, éste contará con una distribución de probabilidad sobre el modelo causal que represente las de \textit{creencias}, o conocimiento previo, que tiene el agente sobre el modelo; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
@@
 \end{itemize}
-
+\newpage
 \bibliographystyle{apalike}

### 2018-04-26 Segundo_Semestre/Seminario/Semana12/Proposal.tex
ultimas correcciones Sucar

@@
 \\
-\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
+\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
 Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.
@@
 Los juegos estocásticos fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\\
+\indent Como resultado de este trabajo se espera desarrollar una manera de adquisición de conocimiento causal a través de la interacción de un agente autónomo con un ambiente incierto.
+Este conocimiento será utilizado dentro de la misma interacción con el fin de que el agente logre un objetivo de manera más rápida, debido a que la incorporación de la estructura causal reduce la incertidumbre pues permite al agente predecir las consecuencias de sus acciones.
+Además, el conocimiento causal aprendido por el agente será exportable para su uso en problemas de decisión relacionados al mismo entorno.
 \subsection{Motivación}
-Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{hagmayer2013repeated},\cite{pearlwhy}).
+Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{spirtes2000causation},\cite{hagmayer2013repeated},\cite{pearlwhy}).
 Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
@@
 \\
-\indent Modificar uno a uno los vértices, o por decirlo de otra forma, hacer actualizaciones locales, como se hace en \cite{lattimoreNIPS2016} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos locales de estructuras causales (\cite{danks2014unifying}) y posteriormente unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+\indent Modificar una a una las relaciones causales que se expresan en un modelo causal como se hace en \cite{wellen2012learning} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos \textit{locales} de las estructuras causal de nuestro entorno (\cite{danks2014unifying}) y posteriormente en unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
 Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
@@
 \end{itemize}
-Estos requerimientos vamos a tratarlos de manera axiomática, en el sentido de que se dan por sentado sin cuestionarlos.
+\indent Estos requerimientos vamos a tratarlos de manera axiomática, en el sentido de que se dan por sentado sin cuestionarlos.
 Quien no esté de acuerdo con los axiomas, no lo estará con las conclusiones y queda siempre la invitación de desarrollar la teoría utilizando otras altenativas.
@@
 \\
-Este resultado conecta las distribuciones de intervención con los enunciados probabilistas que se obtienen del grafo, con lo que contestamos la pregunta que se había dejado abierta antes en la subsección \ref{relacion}.
+\indent Este resultado conecta las distribuciones de intervención con los enunciados probabilistas que se obtienen del grafo, con lo que contestamos la pregunta que se había dejado abierta antes en la subsección \ref{relacion}.
 \begin{teo}{\cite{peters2017elements}}\\
@@
 \\
-Es deseable considerar el caso en el que un jugador no conoce todas las características del otro jugador; ya sea sus acciones o sus pagos, y es este desconocimiento lo que se pretende modelar con los Juegos Estratégicos Bayesianos.
+\indent Es deseable considerar el caso en el que un jugador no conoce todas las características del otro jugador; ya sea sus acciones o sus pagos, y es este desconocimiento lo que se pretende modelar con los Juegos Estratégicos Bayesianos.
 \begin{defi}{\label{juegobayesiano}}Juego Estratégico Bayesiano en Forma Normal (\cite{osborne1994course})\\
@@
 \\
-Existen juegos que son capaces de representar situaciones más complejas; en particular, situaciones en las que no todos los involucrados deciden simultáneamente, sino que existe cierta estructura temporal.
+\indent Existen juegos que son capaces de representar situaciones más complejas; en particular, situaciones en las que no todos los involucrados deciden simultáneamente, sino que existe cierta estructura temporal.
 A estos juegos se les conoce como juegos extensivos, los cuales se dividen a su vez según la información disponible a los jugadores, lo que da lugar a los juegos extensivos con información perfecta, imperfecta, e incompleta según sea el caso.
@@
 Pueden consultarse con más detalle en \cite{shoham2008multiagent}, \cite{fudenberg1998theory}.
-En particular, se tiene el siguiente resultado de interés
+\subsubsection{Aprendizaje Bayesiano}
+Sea $S$ el conjunto de estrategias posibles consideradas de los oponentes consideradas por el jugador $i$ y sea $H$ el conjunto de historias posibles del juego.
+Entonces, la probabilidad asignada por el jugador $i$ al evento en el cual otro jugador está jugando una estrategia particular $s_{-i} \in S^i_{-i}$ dado que se observa la historia $h$ es
+\[ P_i(s_{-1} | h) = \frac{P_i(h|s_{-i})P_i(s_{-i})}{\sum_{s'_{-i}} P_i(h | s'_{-i})P_i(s'_{-i})}. \]
+Si se sigue este proceso de aprendizaje se cumple el siguiente resultado.
 \begin{teo}
@@
 \end{teo}
+Esto significa que las creencias de un jugador eventualmente van a converger a la verdad si hace actualizaciones bayesianas y juega una mejor respuesta dadas sus creencias; pero esto no implica (\cite{shoham2008multiagent}) que el jugador aprenderá la estrategia verdadera utiizada por los otros jugadores
 \section{Trabajo Previo}
 \subsection{Causalidad y Teoría de Decisión}
-Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.\\
+Existen trabajos que utilizan en conjunto inferencia causal junto con toma de decisiones; por ejemplo \cite{heckerman1995decision} argumenta que los trabajos usuales sobre causalidad no contienen primitivas suficientemente básicas, por lo que plantea el problema de inferencia causal como uno de toma de decisiones.
+En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
+En el trabajo \cite{eberhardt2008causal} el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.\\
 \\
@@
 \subsection{Conclusión}
-A manera de cierre de esta sección, resaltamos la importancia inmediate de los trabajos de \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, pues en ellos lo que se busca es encontrar la mejor acción a tomar en un entorno incierto, pero en el cual existen relaciones causales y se hace uso de estas relaciones para encontrar la mejor acción.
-Ambos trabajos suponen conocido el modelo gráfico causal; en el caso de \cite{lattimoreNIPS2016} ellos consideran que el modelo es completamente conocido, mientras que en \cite{sen2017identifying} sólo requieren conocer una parte.
+A manera de cierre de esta sección, hacemos énfasis en los trabajos de \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, pues en ellos lo que se busca es encontrar la mejor acción a tomar en un entorno incierto, pero en el cual existen relaciones causales y se hace uso de estas relaciones para encontrar la mejor acción.
+Ambos trabajos suponen conocido el modelo gráfico causal; en el caso de \cite{lattimoreNIPS2016} ellos consideran que el modelo es completamente conocido, mientras que en \cite{sen2017identifying} sólo requieren conocer una parte del modelo.
 Lo que nosotros proponemos es descubrir el modelo a la vez que se utiliza este mismo para la elección de decisiones.
@@
 \\
+\indent La Naturaleza comenzará escogiendo una jugada que asigna cierto \textit{estado} del ambiente, a lo cual el tomador de decisiones debe responder pero sin saber en qué estado se encuentra (se encuentra en uno de los elementos de la partición al considerar un juego con información imperfecta).
+El tomador de decisiones escogerá su acción y verá el resultado de su acción así como su pago.
+Con este pago y lo que observa del entorno actualiza sus creencias.
+Posteriormente, se procede a un nuevo juego de las mismas características pero con las creencias actualizadas.\\
+\\
 \indent En resumen, el modelo consiste en traducir un problema de toma de decisiones bajo incertidumbre en un juego Bayesiano entre el tomador de decisiones y la Naturaleza donde esta última va a muestrear sus acciones a partir del modelo causal que opera en el entorno y que establece las relaciones causales entre variables así como entre acciones y variables.
-Queremos que a lo largo de las jugadas (o de repeticiones de los juegos) el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
+Queremos que a lo largo sucesivas repeticiones del juego, el jugador vaya actualizando sus creencias causales y descubriendo ese modelo causal; es decir, se necesita que las creencias actualizadas del jugador converjan al verdadero modelo.
 Además, utilizando el concepto de \textit{estrategias mixtas} el jugador puede tomar ciertas acciones de manera aleatoria, lo que introduce el concepto de \textit{exploración} proveniente del área de RL.
@@
 \\
-\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular.
-En \cite{eberhardt2008causal} el problema de descubrimiento causal es planteado también como un juego, en el cual la naturaleza escoge desde el inicio el modelo causal \textit{real} y el jugador debe escoger intervenciones a llevar a cabo con tal de descubrir el modelo causal.
-En el trabajo \cite{eberhardt2008causal} el objetivo es descubrir el modelo causal, no el aspecto de toma de decisiones.
+\indent Un primer planteamiento sería considerar que las acciones de la naturaleza son completamente \textit{desinteresadas}; es decir, que la naturaleza no tiene ningún objetivo en particular, pero al dotar de pagos a la Naturaleza se podrá considerar en un futuro casos en el cual la Naturaleza tenga intenciones.
 
@@
 \begin{itemize}
-\item Una forma general de especificar un juego que capture los aspectos esenciales de un problema de decisión y cuya solución refleje la mejor decisión para el agente.
-Este juego debe permitir que el jugador tenga creencias sobre aspectos estructurales del ambiente.
+\item Definir una forma general de especificar un juego que capture los aspectos esenciales de un problema de decisión lo cual quiere decir que las estrategias óptimas y equilibrios del juego deben estar en correspondencia con los criterios óptimos de solución del problama de decisión.
+Este juego, además, debe permitir que el jugador tenga creencias sobre aspectos estructurales del ambiente.
 
-\item Garantías teóricas de la posibilidad de aprender  algo del ambiente en este juego.
+\item Encontrar una familia de distribuciones que representen de manera coherente creencias o conocimiento previo acerca de la estructura causal de un entorno.
+Asignar una distribución inicial de creencias, que permitan que el agente sea coherente con los axiomas de decisión.
+En \cite{billot2005probabilities} se muestra, de manera axiomática, cómo asignar probabilidades iniciales dado un conjunto de observaciones pasadas.
 
-\item Una manera de \textit{utilizar} el conocimiento acerca de las relaciones causales que el agente tiene en un momento dado para con ello tomar una buena decisión relativo a la utilidad esperada de ésta.
+\item Desarrollar una manera de \textit{utilizar} el conocimiento acerca de las relaciones causales que el agente tiene en un momento dado para con ello tomar una buena decisión relativo a la utilidad esperada de ésta.
 
-\item Cuando el agente no conoce el modelo causal, se debe encontrar una forma de especificar creencias sobre estructura causal; esto es, encontrar una familia de distribuciones que representen de manera coherente creencias o conocimiento previo acerca de la estructura causal de un entorno.
-
-\item Se debe tener una distribución inicial de creencias, que permitan que el agente sea coherente con los axiomas de decisión.
-En \cite{billot2005probabilities} se muestra, de manera axiomática, cómo asignar probabilidades iniciales dado un conjunto de observaciones pasadas.\\
-
-\item Para esto, existen dos casos a considerar; el primero, es suponer que el tomador de decisiones conoce el número máximo de variables que el modelo causal puede tener; en este caso, las creencias del agente pueden estar representadas por una distribución sobre un espacio de matrices de tamaño fijo, aprovechando la relación que existe entre grafos acíclicos dirigidos y matrices.
+%\item Para esto, existen dos casos a considerar; el primero, es suponer que el tomador de decisiones conoce el número máximo de variables que el modelo causal puede tener; en este caso, las creencias del agente pueden estar representadas por una distribución sobre un espacio de matrices de tamaño fijo, aprovechando la relación que existe entre grafos acíclicos dirigidos y matrices.
 Por otro lado, si el agente desconoce el número máximo de variables, entonces el grafo causal puede tener cualquier tamaño.
@@
 
-\item Una manera de actualización de creencias causales que debe converger al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
+\item Establecer una manera de actualización de creencias causales que debe converger al verdadero modelo causal así como ser coherente con estudios de la psicología de la formación de creencias (\cite{larrouy2017mindreading}).
 Es importante mencionar que los juegos Bayesianos, en su definición estándar, proveen un mecanismo para la actualización general de creencias, pero se necesita una manera de actualizar creencias que haga uso de la estructura causal misma.
 Además, es importante tener en cuenta que no cualquier creencia puede ser modelada por un \textit{prior bayesiano} \cite{gilboa2016ambiguity}.
+Una vez actualizadas las creencias, se debe desarrollar una \textit{estrategia} en el juego para el tomador de decisiones que permita explorar otras alternativas que no han sido tomadas en cuenta.
 
-\item Un criterio de solución del problema de decisión: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
-
-\item Una manera de construir un diagrama de influencia causal que además permita \textit{exportar} el conocimiento adquirido para problemas similares; de esta manera, ya sabríamos directamente cómo actuar conociendo el modelo causal.
+\item Establecer un criterio de solución del problema de decisión: Si no es posible esperar hasta la convergencia del modelo, sí debemos garantizar que las decisiones tomadas con el modelo causal aprendido hasta cierta ronda aproximen la utilidad máxima esperada.
 
@@
 Además, estas creencias deben ser consistentes con los criterios de actualización racional de creencias detallados en \cite{shoham2008multiagent}.
-Para el tercer caso, debe evaluarse la utilidad que aportan las acciones escogidas con esta información, la cual debe converger a su vez a la utilidad máxima esperada y para esto se usa, como entorno de prueba, el modelo causal verdadero del ambiente.\\
+Para el tercer caso, debe evaluarse la utilidad que aportan las acciones escogidas con esta información, la cual debe converger a su vez a la utilidad máxima esperada y para esto se usa, como entorno de prueba, el modelo causal verdadero del ambiente, además de comparar contra los algoritmos de identificación de mejor acción y el número de rondas requerido para que la utilidad obtenida supere cierto umbral.\\
 \\
-\indent Una vez establecida la metodología de trabajo, veremos ahora cómo estos se llevarían a cabo en un caso específico de ejemplo
+\indent Una vez establecida la metodología de trabajo, veremos ahora cómo estos se llevarían a cabo en un caso específico de ejemplo.
 
@@
 Los tratamientos posibles conllevan riesgos y ventajas, que están regidos por un modelo causal.
-Los tres casos a considerar son los siguientes
+De los tres casos, sólo dos serán desarrollados como resultados preliminares y dejando el tercer caso como trabajo futuro dentro de la investigación doctoral.
+Los tres casos a considerar son los siguientes:
 \begin{itemize}
@@
 \end{itemize}
+
 \subsection{Escenario}
-Consideremos un paciente que llega a un hospital y puede tener la enfermedad $A$ o la enfermedad $B$, lo cual es desconocido para el médico que lo atiende, éste tiene a su disposición tres opciones: mandar el paciente a cirugía, darle un fármaco, o no hacer nada, con lo cual el paciente morirá casi con certeza.
+Con el fin de ejemplificar las ideas propuestas, consideremos un caso hipotético en el cual se puede aplicar lo propuesto.\\
+\\
+\indent Consideremos un paciente que llega a un hospital y puede tener la enfermedad $A$ o la enfermedad $B$, lo cual es desconocido para el médico que lo atiende, éste tiene a su disposición tres opciones: mandar el paciente a cirugía, darle un fármaco, o no hacer nada, con lo cual el paciente morirá casi con certeza.
 Cada una de estas opciones entraña un riesgo: durante la cirugía el paciente puede morir, y al recibir el fármaco el paciente puede tener una reacción alérgica y morir.
@@
 \end{itemize}
+\begin{figure}
+\centering
 % Node styles
@@
 \end{tikzpicture}
+\caption{Forma extensiva del juego del médico vs la Naturaleza: la Naturaleza comienza asignando la enfermedad del paciente, y debido a que el médico no sabe cuál fue la jugada, debe decidir dentro de un conjunto de información, donde sus opciones son el fármaco, la cirugía o nada.}
+\end{figure}
+La evaluación en este caso se lleva a cabo comparando la utilidad acumulada del médico así como el número de rondas requeridas para llegar esta cantidad con el número de rondas que le toma a los algoritmos de identificación de mejor acción que no incorporan información causal \cite{audibert2010best}.
 \subsubsection{Caso en el que se conoce la estructura del modelo}
@@
 \\
-A partir de lo que la Naturaleza lleve a cabo, se contrastarán los valores que el jugador cree contra lo que la evidencia indica y a partir de esto el modelo causal se irá modificando en cada juego.
-\subsubsection{Caso general}
-En el caso general en el cual el tomador de decisiones no conoce el modelo causal, éste contará con una distribución de probabilidad sobre el modelo causal que represente las de \textit{creencias}, o conocimiento previo, que tiene el agente sobre el modelo; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
+A partir de lo que la Naturaleza lleve a cabo, se contrastarán los valores que el jugador cree contra lo que la evidencia indica y a partir de esto el modelo causal se irá modificando en cada juego.\\
+\\
+De igual forma, la evaluación se da en términos de la utilidad que arroja el uso del modelo causal que se tenga en un determinado momento y el número de rondas requeridas para que la utilidad supere cierto umbral y comparar esto con los algoritmos del estado del arte.
+%\subsubsection{Caso general}
+%El caso general, el cual consiste en que el tomador de decisiones no conoce el modelo causal se dejará abierto para desarrollar posteriormente pues es este el objetivo principal para el trabajo de investigación doctoral.
+En este caso, el tomador de decisiones contará con una distribución de probabilidad sobre el modelo causal que represente las de \textit{creencias}, o conocimiento previo, que tiene el agente sobre el modelo; es decir, el tomador de decisiones tendrá un modelo causal \textit{propio} que usará y modificará con la experiencia.
 Estas creencias tendrán la forma de una distribución de probabilidad sobre algún espacio; si suponemos que el tomador de decisiones conoce al menos el número de variables del modelo, entonces las creencias serán distribuciones de probabilidad sobre matrices de tamaño fijo; en cambio, de no conocer el tamaño del modelo causal, debemos recurrir a distribuciones de probabilidad que permitan representar objetos mucho más complejos.
@@
 Estas distribuciones sobre espacios de dimensión infinita se conocen en la literatura como \textit{priors Bayesianos no-paramétricos} y existe toda un área de la Estadística que los estudia, a saber, la Estadística Bayesiana No Paramétrica (\cite{ghosal2017fundamentals}).\\
-\\
-\indent Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
+%\\
+%\indent Además de la maquinaria matemática para representar dichas creencias, debemos considerar que el proceso de adquisición y uso de creencias debe ser coherente con los axiomas de racionalidad que estamos suponiendo, pues estas creencias afectarán el proceso de toma de decisiones, y como es deseable poder utilizar los resultados ya establecidos, es necesario que la construcción de estas creencias se mantengan en la misma línea.
 Además, para tener cierto grado de realismo e interpretación, es importante apegarse también a lo que dice la psicología (\cite{larrouy2017mindreading}).
+
 \section{Plan de trabajo (cronograma)}

### 2018-04-26 Segundo_Semestre/Seminario/Semana12/Beamer_Template.tex
presentacion

@@
 \item “Causality” evokes certain regularities, we want to think about the causes of a disease in order to avoid it. 
-\item Human beings tend to find causal narratives, sometimes excesively and that leads to incorrect judgements \cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
+\item Human beings tend to find causal narratives, sometimes excesively and that leads to incorrect judgements (\cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}).
 \item How to effectively use causal knowledge to make decisions?
@@
 		\item Statisticians have some kind of fear about using this word.
-		\item Economists try to find assumptions under an econometric model is causal.
+		\item Economists try to find assumptions under which an econometric model is causal.
 		\item Many definitions exists, (see for example \cite{arlo2016readings}).
@@
 		\begin{defi}
-		Let $(\Omega, \mathcal{F})$ a finite, measurable space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
+		Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite, probability space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
 		\begin{itemize}
@@
 		\frametitle{Direct and Indirect Causes}
-		Consider a refinement $\mathcal{F}_s$ of $\mathcal{F}$ and extend “$\to$” properly.
+		Consider a refinement\footnote{In the sense as a refined topology} $\mathcal{F}_s$ of $\mathcal{F}$ and extend “$\to$” properly.
 Notice that it could happen that if $A \to B$, then for some $C_s \in \mathcal{F}$ we have that $A \to C_s \to B$.
@@
 		\frametitle{From Causes to Graphs}
-		Abusing notation, notice that we can build a graph that represent all of the direct causes between events that are members of $\mathcal{F}$.\\
 		For each pair $A,B \in \mathcal{F}$ such that $A \to B$ and $A$ is a direct cause of $B$, define a node representing each variable, and a directed edge joining them.
@@
 		\frametitle{from Graphs to Probability Measures}
-		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph.
-		An interesting question would be to study the relation of $P_\mathcal{G}$ and any other measure in the \textit{upper level} of the measurable space .
+		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
+        \begin{itemize}
+        \item Causal Markov Condition.
+        \item Causal Minimality.
+         \item Causal Faithfulness.
+\end{itemize}
+This requirements will be treated in an axiomatic fashion.
 		\end{frame}
@@
 \frametitle{Game Theory}
-Game Theory is an area of Mathematics that is used to model strategic interaction situations, which can be interpreted as multi-agent decision problems, where the outcomes for each player depend on the actions of the other players.
+Game Theory is an area of Mathematics that is used to model strategic interaction situations, which can be interpreted as multi-agent decision problems, where the outcomes for each player depend on the actions of the other players.\\
+Aumann defined Game Theory as \textit{...optimal decision making in the presence of others with different objectives...}\\
  Several classes exist:
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Strategic Games}
+Consider a situation where two or more agents interact and each one of them seeks to fullfill an objective. Here, each agent will take one and only one decision and simultaneously with other agents.
+\begin{defi}{A strategic normal form game (\cite{osborne1994course}, \cite{shoham2008multiagent}) }\\
+A finite normal form game is a tuple $(N,A,u)$ where:
+\begin{itemize}
+\item $N$ finite set of players.
+\item $A= A_1 \times \cdots \times A_n$ each $A_i$ are the actions available to player $i$. Each element  $a = (a_1,...,a_n) \in A$ es lis called an action or strategy profile.
+\item For each player $i \in N$, a preference relation $\succeq_i$ defined over $A$.
+\end{itemize}
+\end{defi}
+\end{frame}
+
+\begin{frame}
+\frametitle{Bayesian Games}
+\begin{defi}{\label{juegobayesiano}}
+A strategic normal form Bayesian game is composed by:
+\begin{itemize}
+\item finite set of players $N$
+\item Set of states $\Omega$.
+\item For each player: 
+\begin{itemize}
+\item Set of possible actions $A_i$,  and define $A = A_1 \times \cdots \times A_n$.
+\item A finite set $T_i$ of observable signals for player $i$.
+\item Signal function $\tau_i : \Omega \to T_i$.
+\item A probability measure $P_i$ defined over $\Omega$ such that $P( \tau^{-1}_i (t_i)) >0$ for all  $t_i \in T_i$.
+\item Preference relation $\succeq_i$ defined over  $\mathcal{M}(A \times \Omega)$
+\end{itemize}
+\end{itemize}
+\end{defi}
+\end{frame}
+
+\begin{frame}
+\frametitle{Bayesian games: interpretation}
+\begin{itemize}
+\item $\Omega$ contains all of the descriptions of players' relevant characteristics.
+\item Each player has prior beliefs over this characteristics, represented by $P_i$.
+\item At any given play, $\omega \in \Omega$ is realized.
+\item Each player observes $\tau_i(\omega)$ and let $T_i$ all posible values of $\tau_i$.
+\item Player $i$ receives $t_i$, so he  deduces that the true state is in the set $\tau_i^{-1}(t_i)$.
+\item So he updates the probability asigned to $\omega \in \Omega$ by $P_i(\omega) / P(\tau_i^{-1}(t_i))$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}[allowframebreaks]
+\frametitle{Perfect Information Extensive Games}
+\begin{defi}
+A perfect information extensive game:
+\begin{itemize}
+\item A finite set  $N$ of players.
+\item Set $A$ of actions.
+\item Set $H$ of sequences (called histories) such that:
+	\begin{itemize}
+	%	\item The empty sequence belongs to $H$.
+		\item If $(a_k)_{k=1}^K \in H$ ($K$ can be infinite) and $L < K$ then $(a_k)_{k=1}^L \in H$.
+		\item If $(a_k)_{k=1}^\infty$ satisfies $(a_k)_{k=1}^L \in H$ for all $L$ then $(a_k)_{k=1}^\infty \in H$
+	\end{itemize}
+\item If $(a_k)_{k=1}^K \in H$ where $K$ is finite and if no $K+1$ exists such that $(a_k)_{k=1}^{K+1} \in H$ then the sequence is called terminal. The set of terminal histories is called $Z$
+\item A function $\chi : H \to 2^A$  that assigns a set of possible actions to be taken after a history $h$ has ocurred.
+\item A function $\rho: H \to N$ that assigns to every nonterminal history a player $\rho(h) \in N$ who will take an action after history $h$ has ocurred.
+\item For each player, a preference relation $\succeq_i$.
+\end{itemize}
+\end{defi}
+\end{frame}
+
+\begin{frame}
+\frametitle{Imperfect Information Extensive Games}
+\begin{itemize}
+\item $(N,A,H,Z,\chi,\rho,\sigma, u)$ is a perfect information extensive game.
+\item $I=(I_1,...,I_n)$for each player $i$, we have that  $I_i = \{ I_{i,1}, I_{i,2}, ...
+, I_{i,k_i} \}$ is a partition of $\{ h \in H: \rho(h)=i \}$ that satisfies that if $\chi(h)=\chi(h')$ then there exists $j$  such that $h \in I_{i,j}$ and $h \in I_{i,j}$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Repeated Games}
+Playing repeatedly a \textit{base game}: this is interesting because if the sum of payments is considered in the long run, then the one-step equilibria can change.
+For example, in the Prisoner Dilemma, players may realize that the one-stage Nash equilibrium is not optimal: players can \textit{learn}.
+\end{frame}
+
+\begin{frame}
+\frametitle{Learning in Games}
+In a repeated game, what can a rational agent learn? After several rounds he can learn the strategy of others and use it, or play a certain strategy to deceive others that are learning. 
+\end{frame}
+
+\begin{frame}
+\frametitle{Rational Learning}
+Let $S$ be the set of the opponents considered by player $i$ and $H$ the stories of the game.
+Then, the probability assigned by player $i$ to the event in which the opponent is playing a particular strategy $s_{-i} \in S^i_{-i}$ given that story $h$ is observed is:
+\[ P_i(s_{-1} | h) = \frac{P_i(h|s_{-i})P_i(s_{-i})}{\sum_{s'_{-i}} P_i(h | s'_{-i})P_i(s'_{-i})}. \]
+\end{frame}
+\begin{frame}
+\frametitle{Rational Learning and Belief accuracy}
+\begin{teo}
+Let $s$ be a repeated game strategy profile  for a given $n$ player game and let $P=P_1,...,P_n$ be a tuple of probability distributions over such strategy profiles ($P_i$ are player $i$'s beliefs).
+Let $\mu_s$ be de distribution over infinite game histories induced by the profile $s$ and let $\mu_P$ the distribution induced by the belief tuple $P$.
+Then, if
+\begin{itemize}
+\item At each round each payer $i$ plays a best-response strategy given his beliefs $P_i$.
+\item After each round player $i$ updates $P_i$ using rational updating.
+\item $\mu_s$ is absolutely continuous with respect to $\mu_p$.
+\end{itemize}
+Then, for all $\varepsilon > 0$ and for almost every history in the support of $\mu_s$ there is a time $T$ such that for all $t \geq T$ the play $\mu_{P_i}$ predicted by player $i$'s beliefs is $\varepsilon$-close to the distribution of play $\mu_s$ predicted by the actual strategies.
+\end{teo}
+\end{frame}
+\begin{frame}
+This means that a players' beliefs will converge to the truth if he is using Rational Updating and playing a Best Response.
 \end{frame}
@@
 Let $(D,E,c)$ a decision problem whose uncertain events are related to the variables in $\mathcal{G}$.
-Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, then by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less rounds than if he doesn't consider causal information.
+Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$,  by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less rounds than if he doesn't consider causal information.
 	\end{frame}
@@
 	\begin{frame}
-	The proposed research has 2 general objectives: 
+	The proposed research has as a general objective: 
 	\begin{itemize}
-	\item Find the way in which an agent facing an uncertain environment (governed by a causal model) can use causal knowledge in order to make good choices and improve his performance when compared to non-causal ways of choosing.
-	\item Find the way in which an agent can acquire causal information while interacting with an uncertain (governed by a causal model) environment.
+	\item Find the way in which an agent facing an uncertain environment which governed by a causal model can acquire and make use of causal knowledge in order to make good choices and improve his performance when compared to non-causal guided ways of choosing.
 	\end{itemize}
@@
 	\begin{itemize}
+	\item To find a way to specify a game that captures the essential aspects of a decision making problem.
+This means that the strategies and equilibria of the game must be in correspondence with optimal solutions of the decision making problem.
 \item To find a way to specify beliefs about causal structure; i.e., find (or build) a family of parametric distributions that represent in a coherent way beliefs or previous knowledge about the causal structure of some environment.
-\item  We have to consider two cases: if the decision-maker (or the owner of the beliefs) knows the maximum number of variables that the causal model can have.
-\item On the other side, if the agent doesn't know the max number of variables.
-\item To find a way of effectively using causal knowledge that the agent has at a given moment.
+\item A way to specify initial beliefs according to past knowledge in a coherent fashion, previous work in \cite{billot2005probabilities}.
+Not any belief can be modelled by Bayesian priors  \cite{gilboa2016ambiguity}.
+\item Theoretical guarantees of learning in this class of games.
+\item To find a way of effectively using causal knowledge that the agent has at a given moment in order to make a good choice relative to the max utility.
 \item Develop an update criterion for the beliefs of the agent based upon what he has observed from the environment while interacting.
@@
 Also, the beliefs obtained must be coherent with the decision making since we will be using those results.
+\item A way to specify a game strategy given past games and current beliefs.
 \item An aproximate soluton criteria: maybe we can't wait until convergence, so what can be done with the current knowledge.
@@
 	\end{frame}
+
 \section{Methodology}
+\subsection{Working methodology}
+\begin{frame}[allowframebreaks]
+\frametitle{Working Methodology}
+\begin{itemize}
+\item Fix the class of games will be used to model the interactions between a decision maker and his environment.
+\item Define what are the mentioned beliefs and how to represent them using probability distributions.
+\item Update criteria that must converge to the true causal model.
+\item Theoretical guarantees over the conditions and the update criteria.
+\item Analyze, experiment and compare in three cases:
+\begin{itemize}
+	\item Knowing all of the model: how the chosen decision compares to best-action identification algorithm?
+	\item Knowing only the structure of the model: using in each round the current knowledge as the “true” model and choosing with it.
+	\item Not knowing anything: choosing-observing-updating cycle, compare number of rounds for utility above certain threshold
+\end{itemize}
+\end{itemize}
+\end{frame}
 
-	\subsection{Working methodology}
 \section{Preliminary Results}
@@
 \item In the third case, the decision maker knows nothing at all, so he has general beliefs about the causal model itself.
+\end{itemize}
+\end{frame}
+
+\section{Limitations and possible extensions}
+\begin{frame}
+\frametitle{Limitations}
+\begin{itemize}
+\item We are considering a rational decision maker.
+
+\item Causality is used as defined above and it is not questioned if it suits correctly realistic situations.
+We do not consider if other possible definitions of causality correspond better to decision making.
+
+\item Nature has not intentions, but since we are giving her equal payments this could be modified later.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Possible extensions}
+\begin{itemize}
+\item Other forms of decision making, behavioral approach, bounded rationality.
+\item Longer games of unknown structure.
+\item More players besides nature.
+\item Nature has intentions.
+\item Game solutions (equilibria) in terms of the structure of the causal graphical model.
+\item Dynamica systems associated with the learning process.
 \end{itemize}

### 2018-04-26 Segundo_Semestre/Seminario/Semana12/Proposal.tex
pequeño intro a teo decision

@@
 \subsubsection{Causas Directas e Indirectas}
-Consideremos ahora un \textit{refinamiento}\footnote{En el sentido de una Topología más fina \cite{munkres2000topology}.} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
+Consideremos ahora un \textit{refinamiento}\footnote{En el sentido de una Topología más fina (\cite{munkres2000topology}).} $\mathcal{F}_s$ de la sigma-álgebra $\mathcal{F}$ de modo que ahora puedan existir nuevos eventos $C_s \in \mathcal{F}_s$ tales que $A \to C_s \to B$.
 Decimos que $A$ es una \textit{causa directa} de $B$ si se cumple que $A \to B$ y no existe evento $C$ tal que $A \to C \to B$ para todo refinamiento propio de $\mathcal{F}$.
@@
 Dado un grafo acíclico dirigido, podemos construir una medida de probabilidad que expresa las relaciones que se encuentran establecidas en el grafo, la cual denotaremos $P_{\mathcal{G}}$.
-Requerimos que la medida $P_{\mathcal{G}}$ cumpla las siguientes condiciones que la relacionan con el grafo:
+Requerimos que la medida $P_{\mathcal{G}}$ cumpla las siguientes condiciones que la relacionan con el grafo (\cite{spirtes2000causation}):
 \begin{itemize}
@@
 \indent Estos requerimientos vamos a tratarlos de manera axiomática, en el sentido de que se dan por sentado sin cuestionarlos.
-Quien no esté de acuerdo con los axiomas, no lo estará con las conclusiones y queda siempre la invitación de desarrollar la teoría utilizando otras altenativas.
+No cae dentro del enfoque de la investigación propuesta el encontrar definiciones o estructuras causales que se adapten de manera más o menos realista a cierta situación particular.
 
@@
 \subsection{Toma de decisiones}
-El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad si un tomador de decisiones conoce las probabilidades de los eventos, entonces  la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
+En lo siguiente, mencionaremos tres resultados de alta importancia en la formalización de la toma de decisiones bajo incertidumbre.
+Veremos que los elementos comunes en estos resultados son un agente, o tomador de decisiones, que busca escoger una de entre varias opciones, cuyas consecuencias son inciertas, de modo que se maximice cierta cantidad como consecuencia de esta elección.
+Para ello, se requiere que el agente sepa distinguir entre qué opciones le representan un mayor bienestar, el cual llamaremos \textit{utilidad}.
+La característica principal de los resultados que veremos a continuación es que proveen una manera formal de escoger entre las acciones si el tomador de decisiones cumple ciertas condiciones en cuanto a sus preferencias.\\
+\\
+\indent El Teorema de Von Neuman-Morgenstern, garantiza la existencia de una función de utilidad tal que si un tomador de decisiones conoce las probabilidades de los eventos, entonces la relación de preferencias del agente es equivalente al orden que impone la utilidad esperada calculada respecto a esta función y las probabilidades conocidas, por lo que la mejor decisión a tomar es aquella que maximice esta cantidad.\\
 \\

### 2018-04-26 Segundo_Semestre/Seminario/Semana12/Proposal.tex
cambios menores teo juegos

@@
 \subsection{Teoría de Juegos}{\label{juegos}}
-Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no sólo las acciones y preferencias de un agente sino también las de otros.
-En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene \textit{creencias} \textit{a priori} sobre algunos aspectos del ambiente.
-Las creencias que suele tener un jugador suelen ser sobre ciertos aspectos; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.).\\
+Teoría de Juegos (\cite{osborne1994course}) es un área de las matemáticas que los economistas y otros científicos sociales utilizan para modelar situaciones de interacción estratégica; es decir, problemas de decisión en los cuales importa no sólo las acciones y preferencias de un agente sino también de otros y donde además los \textit{pagos} o la utilidad de las acciones de cada jugador se ven afectadas por las acciones de los demás.
+En particular, en esta área existen los llamados \textit{juegos Bayesianos}, también conocidos como Juegos con Información Incompleta (\cite{osborne1994course}, \cite{10.1007/978-94-010-0189-2_25}),  en los cuales cada jugador tiene \textit{creencias} \textit{a priori} sobre algunos aspectos del ambiente las cuales son actualizadas conforme se desarrolla el juego.
+Usualmente, las creencias que suele tener un jugador son sobre ciertas características específicas; por ejemplo, creencias sobre las acciones de los otros jugadores (\cite{costa2008stated}), creencias sobre el estado del mundo (\cite{dominitz2009empirical}) y creencias sobre las creencias de los otros jugadores.\\
 \\

### 2018-05-04 Segundo_Semestre/Seminario/Semana12/Proposal.tex
version presentada en seminario

@@
 \cite{gilboa2001theory} propone un nuevo paradigma de toma de decisiones, llamado \textit{Case-Based Decision Theory}, en el cual los agentes toman decisiones haciendo analogías a eventos pasados en los cuales tuvieron un buen desempeño.
+Para más detalles respecto a los últimos avances en Teoría de la Decisión ver \cite{wakker2010prospect}.
 \subsubsection{Teorema de von Neumann-Morgenstern: existencia de utilidad}
@@
 Los resultados anteriores parecen estar en polos opuestos, pues el Teorema de De Finetti provee una distribución de probabilidad para tomar decisiones con ella si el tomador de decisiones tiene una manera de medir su utilidad; por otro lado, el Teorema de Von Neumann supone que se conocen las probabilidades y otorga una forma de medir utilidad para tomar decisiones.
-El gran logro del Teorema de Savage es garantizar la existencia tanto de probabildad subjetiva como de una función de utilidad y además hacerlo en tándem (\cite{gilboa2009decision}).\\
+El gran logro del Teorema de Savage es garantizar la existencia tanto de probabildad subjetiva como de una función de utilidad y además derivar ambas en conjunto (\cite{gilboa2009decision}).\\
 \\
@@
 \end{teo}
+
+\subsubsection{Formalismo Moderno}
+A partir de los resultados anteriores,
+\begin{defi}
+Un problema de decisión es la situación en la cual un agente (llamado tomador de decisiones) se enfrenta a un conjunto de decisiiones $D$ de entre las cuales tiene que escoger una.
+\end{defi}
+Basado en \cite{Bernardo1985}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, tenemos que los elementos que componen un problema de decisión son:
+\begin{itemize}
+\item Un espacio de medida $(\Omega, \mathcal{E})$
+\item Un conjunto $\{d_i, i \in I\}$ llamado conjunto de decisiones, también llamadas acciones u opciones.
+\item Para cada $d_i \in D$, una partición $E_i= \{E_{ij}, j \in J \}$ de $\mathcal{E}$ de \textit{eventos inciertos}.
+\item Correspondiente a cada $E_i$, un conjunto de consecuencias $C_i = \{ c_{ij}, j \in J \}$.
+\item Una relación de orden $\preceq$ sobre $D \times D$
+\end{itemize}
+Notemos que las particiones asociadas a cada decisión no tienen por qué ser iguales, pero siempre podemos construir
+\[ E = \cap_{i \in I} A_i \textrm{ donde } A_i \in E_i. \]
+De modo que en el futuro supondremos, sin perdida de generalidad, que es una misma partición para todas las decisiones, y por lo tanto un solo conjunto de consecuencias, por lo que la notación utilizada para representar cada posible decisión es la siguiente $d_i = \{ c_j | E_j, j \in J\}$ lo cual se interpreta como que al escoger $d_i$, ocurrirá uno de los eventos inciertos dentro de la partición $E$, digamos, $E_j$ y eso trae como consecuencia $c_j$.
+Además, notemos que el orden de los índices es irrelevante, pues la opción $\{ c_1 | E, c_2 | E^c \}$ es la misma que $\{ c_2 | E^c, c_1 | E \}$.
+\begin{defi}
+Diremos que un problema de decisión es sin incertidumbre si para cada $d_i \in D$ se tiene que $C_i = \{c_i\}$; es decir,  el conjunto de consecuencias posibles  asociadas a la decisión $d_i$ consta de un solo elemento.
+\end{defi}
+\begin{defi}
+Análogamente, un problema de decisión para el cual se tiene que para al menos un elemento $d_i \in D$ existe más de una posible consecuencia se dice que es un problema de decisión con incertidumbre.
+\end{defi}
+Como requerimiento técnico, necesitamos que todos los elementos del conjunto de consecuencias pertenezcan al conjunto de opciones; esto se logra identificando cada consecuencia como la opción $\{c|\Omega \}$.
+Siguiendo esta notación, podemos definir una relación de preferencias entre consecuencias a partir de la relación de preferencia entre opciones de la siguiente manera
+\[ c_1 \leq c_2 \Leftrightarrow \{ c_1 | \Omega \} \preceq \{ c_2 | \Omega \}.\]
+A través de un razonamiento análogo, odemos definir una noción de qué tan \textit{creíble} es un evento incierto para un tomador de decisiones a partir de sus preferencias sobre las acciones de la siguiente manera
+\begin{defi}
+Se dice que un evento más creible que otro como
+\[E \leq F \Leftrightarrow \textrm{ para todo } c_1 < c_2 \{ c_2 | E, c_1 | E^c \} \preceq \{ c_2 | F, c_1 | F^c \} \]
+\end{defi}
+Esto permite definir una relación de preferencia a la luz de la ocurrencia de algún evento incierto
+\begin{defi}
+Para cualquier evento $G$ distinto del vacío, definimos:
+\begin{itemize}
+\item $d_1 \leq_G d_2 \Leftrightarrow \textrm{ para toda opción } d \textrm{ se tiene que } \{d_1 | G, d | G^c \} \preceq \{ d_2 | G, d | G^c\}$
+\item $c_1 \leq_G c_2 \Leftrightarrow \{ c_1 | \Omega \} \leq_G \{ c_2 | \Omega \}$.
+\item $E \leq_G F \Leftrightarrow \textrm{ para } c_1 \leq_G c_2, \{ c_2 | E, c_1 | E^c \} \leq_G \{ c_2 | F, c_1 | F^c \} $
+\end{itemize}
+\end{defi}
+
 \subsection{Teoría de Juegos}{\label{juegos}}

### 2018-05-04 Propuesta/Propuesta.tex
copiando de la de seminario

--- 2018-05-04 Segundo_Semestre/Seminario/Semana12/Proposal.tex
@@
 Debido al enorme número de factores diversos que están involucrados en nuestro entorno, constántemente nos vemos forzados a tomar decisiones cuyas consecuencias son inciertas (\cite{danks2014unifying}).
-Esa misma realidad, altamente compleja e interconectada, contiene una serie de relaciones causa-efecto que podemos utilizar al momento de tomar decisiones.
+Esa misma realidad, altamente compleja e interconectada, contiene una serie de relaciones \textit{causa-efecto} que podemos utilizar al momento de tomar decisiones.
 El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) y predecir las consecuencias de nuestros actos.

### 2018-05-05 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
hasta belief formation

\documentclass{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
\usepackage{icml2018}

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2018}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2018}

\begin{document}

\twocolumn[
\icmltitle{Use and acquisition of causal information for decision making under uncertainty}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2018
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Mauricio Gonzalez-Soto}{inaoe}
\icmlauthor{Luis Enrique Sucar}{inaoe}
\icmlauthor{Hugo Jair Escalante}{inaoe}

\end{icmlauthorlist}

\icmlaffiliation{inaoe}{Department of Computer Science, National Institute of Astrophysics Optics and Electronics (INAOE), Puebla, Mexico}

\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Causality, Causal Inference, Game Theory, Decision under Uncertainty}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
We study how a decision maker can learn and use causal information in an uncertain environment which is governed by a causal model. 
Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
\end{abstract}

\section{Introduction}
A fundamental part of autonomous and intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions. In some cases, the decision maker has enough information about his environment so she can decide what decision to make using this information by predicting consequences and utilities. In some other cases the decision maker does not have enough information to proceed in this way. One possible way is to try to \textit{learn} from the environment by interaction.

We consider the case when the uncertain environment where the decision maker is acting is governed by a \textit{causal mechanism}.

Human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}).

Learning by interaction has been extensively studied in Reinforcement Learning, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment such as causal relations.

\section{Related Work}
Since we are considering one stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it, we could consider this as a bandit problem.

In a bandit problem (\cite{sutton1998reinforcement}) an agent chooses an \textit{arm}, observes a reward, and moves on to the next machine wich is of the same kind.

Several algorithms exists for finding the best arm in a multi-arm bandit, such as \cite{bubeck2009pure}, \cite{gabillon2012best}, \cite{agarwal2014taming} \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of this works consider causal-governed environments.

\cite{lattimoreNIPS2016} is, up to our current knowledge, the first paper to consider causal relations between the effects of actions. They consider a decision maker who must choose the best among several possible interventions on a given, and fully known, causal model. The optimality of the action in this context is in terms of the minimal regret. The case where the causal model is not known is left as future work.

\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part.

\section{Problem setup}
By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ such that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).

A directed acyclic graph (DAG) can be used to represent all of the relations that occur in that space by considering a node for every variable that is related to another and a directed edge to express the causal relation, call this DAG $\mathcal{G}$ and consider a probability measure $P_{\mathcal{G}}$ that expresses the conditional statements from the DAG.

We require that this measure satisfies the Markov Causal Condition, Causal Minimality and Causal Faithfulness as stated in \cite{spirtes2000causation}. The relation between $\mathbb{P}$ and $P_{\mathcal{G}}$ is given by the Manipulation Theorem of \cite{spirtes2000causation} and the Do-Calculus rules from \cite{pearl2009causality}.

Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ where family of uncertain events $E$ and the set of consequences $C$ are causally related. We assume that the decision maker has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$. The decision maker does not know the probabilities underlying the causal model therefore he can not calculate the expected utility of any action.

Instead, the agent will try to learn from succesive rounds of decision making. Each round will be independent from the previous ones.

We propose to model this setting as a repeated game between the original decision maker and a new abstract player called Nature. Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model. The decision maker will have \textit{beliefs} about the causal model and in each round she will use those beliefs as a \textit{personal} causal model.

\section{Belief formation and updating}
For an agent to reason about and modify his causal knowledge we endow him with a probability distribution $p(\theta)$ over a space of causal models.

This beliefs must allow to be used in a given moment as a \textit{true} causal model and to be updated in a coherent way.
\section{Test scenario}

\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
\bibliographystyle{icml2018}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
% by Iain Murray in 2018. It was modified from a version from Dan Roy in
% 2017, which was based on a version from Lise Getoor and Tobias
% Scheffer, which was slightly modified from the 2010 version by
% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
% slightly modified from Prasad Tadepalli's 2007 version which is a
% lightly changed version of the previous year's version by Andrew
% Moore, which was in turn edited from those of Kristian Kersting and
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2018-05-08 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
añadiendo escenario prueba

@@
 
-\begin{abstract}
-We study how a decision maker can learn and use causal information in an uncertain environment which is governed by a causal model. 
-Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
-\end{abstract}
+%\begin{abstract}
+%We study how a decision maker can learn and use causal information in an uncertain environment which is governed by a causal model. 
+%Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
+%The decision maker has beliefs over the causal model which she updates in each round. 
+%\end{abstract}
 
@@
 A fundamental part of autonomous and intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions.
-In some cases, the decision maker has enough information about his environment so she can decide what decision to make using this information by predicting consequences and utilities.
-In some other cases the decision maker does not have enough information to proceed in this way.
-One possible way is to try to \textit{learn} from the environment by interaction.
+In some cases, the decision maker has enough information about his environment to decide which decision to make by predicting consequences and utilities.
+
+In some cases the decision maker does not have enough information to proceed in this way. One possible way is to try to \textit{learn} from the environment by interaction.
 
@@
 \section{Related Work}
-Since we are considering one stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it, we could consider this as a bandit problem.
+Similar one stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it have been modelled as bandit problems.
 
@@
 
-We propose to model this setting as a repeated game between the original decision maker and a new abstract player called Nature.
+We attempt to model this setting as a repeated game between the original decision maker and a new abstract player called Nature.
 Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model.
-The decision maker will have \textit{beliefs} about the causal model and in each round she will use those beliefs as a \textit{personal} causal model.
+Nature having objectives to pursue (non-constant payments) will be left as future work.
+
+The decision maker will have \textit{beliefs} about the causal model and in each round she will use those beliefs as a \textit{personal} causal model and modify them using the observations from each round.
 
 \section{Belief formation and updating}
-For an agent to reason about and modify his causal knowledge we endow him with a probability distribution $p(\theta)$ over a space of causal models.
+For an agent to reason about and modify his causal knowledge we endow him with a probability distribution $p(\theta)$ over a suitable space.
+Since we are considering causal graphical models, and for each DAG there exists a one-to-one correspondence to a matrix, one way to express beliefs about DAG's is using distributions over matrices for the structure of the graph.
+
+Since the beliefs must behave as a \textit{local} model in a given moment, it is preferable to use probability distributions over a graph space, so this distribution gives specific models when sampled upon.
 
@@
 \section{Test scenario}
+Consider a patient who arrives at a hospital and he can have either disease $A$ or $B$. The doctor can either give him some pill or send him into surgery.
+
+Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
+
+Asume that the causal model that governs this situation is given by:
+
+\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}
+
+The node corresponding to the variable lives is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the treatment.
+
+\subsection{If the causal model is completely known}
+If the decision maker knows the causal model, then she can obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made.
+
+The tratment that is the arg max for the highest probability of the desired value for the target variable will be chosen, and this choice will be a \textit{best response} and the max expected utility choice.
+
+Applying Pearl's \cite{pearl2009causality} do-calculus, we see that the interventional distribution $P_{do(Tr)}(V)$ is given by
+\[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E) \]
+\subsection{Knowing only the structure}
+If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) which she must learn by repetitions.
+
+We consider a repeated game between the original decision maker and Nature, which acts without intentions and obeys the causal model by sampling from it given the interventions of the decision maker.
+As usual, Nature is considered to play first to asign a random state of the environment, which in this case is the disease.
 

### 2018-05-09 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
cambios en la intro

@@
 
-\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part.
+\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal
 
@@
 
-Instead, the agent will try to learn from succesive rounds of decision making. Each round will be independent from the previous ones.
+Instead, the agent will try to learn by interaction with the environment succesive rounds of decision making.
+Inside each round, any response from the environment will be independent from the previous rounds.
 
-We attempt to model this setting as a repeated game between the original decision maker and a new abstract player called Nature.
+We attempt to model the interaction between the decision maker and the environment as a repeated game between two players: the original decision maker and a new abstract player called Nature.
 Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model.
@@
 
-Asume that the causal model that governs this situation is given by:
+The causal model that governs this situation is shown in Figure \ref{causal_model}
 
-\includegraphics[scale=0.3]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
+\caption{Causal graphical model: the target variable is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
 

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
redaccion

@@
 \section{Introduction}
-A fundamental part of autonomous and intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions.
-In some cases, the decision maker has enough information about his environment to decide which decision to make by predicting consequences and utilities.
+A fundamental part of intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions.
+In some cases, a decision maker has enough information about his environment to make choices by calculating expected utilities.
 
-In some cases the decision maker does not have enough information to proceed in this way. One possible way is to try to \textit{learn} from the environment by interaction.
+In other cases she may not have enough information to proceed in this way. One possible way is to try to \textit{learn} that missing information from the environment by interacting with it.
 
-We consider the case when the uncertain environment where the decision maker is acting is governed by a \textit{causal mechanism}.
+We consider the case where the uncertain environment is governed by a \textit{causal mechanism} unknown to the agent.
 
@@
 
-Learning by interaction has been extensively studied in Reinforcement Learning, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment such as causal relations.
+Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment such as causal relations.
 
 \section{Related Work}
-Similar one stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it have been modelled as bandit problems.
+One stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it have been modelled as \textit{bandit problems}.
 
-In a bandit problem (\cite{sutton1998reinforcement}) an agent chooses an \textit{arm}, observes a reward, and moves on to the next machine wich is of the same kind.
+In a bandit problem (\cite{sutton1998reinforcement}) an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine wich is of the same kind and whose initial settings are independent of the previous machine and action.
 
@@
 
-\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal
+\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal.
 
 \section{Problem setup}
-By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ such that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).
+By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).
 
@@
 We attempt to model the interaction between the decision maker and the environment as a repeated game between two players: the original decision maker and a new abstract player called Nature.
+The base game will be the original decision problem.
+
 Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model.
@@
 
-The decision maker will have \textit{beliefs} about the causal model and in each round she will use those beliefs as a \textit{personal} causal model and modify them using the observations from each round.
-
 \section{Belief formation and updating}
-For an agent to reason about and modify his causal knowledge we endow him with a probability distribution $p(\theta)$ over a suitable space.
+For an agent to reason about and modify his causal knowledge we endow her with a probability distribution $p(\theta)$ over a suitable space.
 Since we are considering causal graphical models, and for each DAG there exists a one-to-one correspondence to a matrix, one way to express beliefs about DAG's is using distributions over matrices for the structure of the graph.
 
-Since the beliefs must behave as a \textit{local} model in a given moment, it is preferable to use probability distributions over a graph space, so this distribution gives specific models when sampled upon.
+Since the beliefs must behave as a \textit{local} model in a given moment, it is preferable to assign probabilities directly to causal models rather than edges in a model in order to have a fixed model when sampling from this probability assignement.
 
-This beliefs must allow to be used in a given moment as a \textit{true} causal model and to be updated in a coherent way.
+After each round of the base game, the probabilities representing causal beliefs will be updated in a Bayesian way in order to guarantee Bayesian consistency \cite{shoham2008multiagent}.
 \section{Test scenario}
@@
 
+In this scenario, Nature's move will consist in randomly assigning the disease to the patient.
+Then, the medic will asign a treatment, the decision nodes for this play form an information set because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
+Finally, Nature will sample the consequence of the treatment from the causal model and the medic will observe the outcome.
+
 \subsection{If the causal model is completely known}
@@
 \subsection{Knowing only the structure}
-If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) which she must learn by repetitions.
+If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.
 
-We consider a repeated game between the original decision maker and Nature, which acts without intentions and obeys the causal model by sampling from it given the interventions of the decision maker.
-As usual, Nature is considered to play first to asign a random state of the environment, which in this case is the disease.
+As usual in the previous case, Nature is considered to play first to asign a random state of the environment, which in this case is the disease.
+
+The decision maker (medic) doesn't know the probabilities, but since she knows the graph structure, she can explicitly a non interventional expression for the invterventional distribution and the values that appear in the expression can be estimated from succesive observations.
 

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
cambios en la intro y related work segun correcciones de hugo

@@
 A fundamental part of intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions.
-In some cases, a decision maker has enough information about his environment to make choices by calculating expected utilities.
+In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utilities.
+If enough information is not available, the decision maker could try to \textit{learn} from the environment by interacting with it.
 
-In other cases she may not have enough information to proceed in this way. One possible way is to try to \textit{learn} that missing information from the environment by interacting with it.
+Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment.
 
-We consider the case where the uncertain environment is governed by a \textit{causal mechanism} unknown to the agent.
+In this paper we consider the case where the uncertain environment is governed by a \textit{causal mechanism} unknown to the agent.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment.
 
-Human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}).
-
-Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment such as causal relations.
+We model the interaction between the decision maker and his environment a repeated game with incomplete information between a player called Nature whi selects his actions from the causal model and the decision maker.
 
 \section{Related Work}
-One stage decision problems in which the decision maker will choose one action, and one action only, and then observe the consequences of it have been modelled as \textit{bandit problems}.
+Consider a one stage decision problem where the relation between actions and consequences is given by a causal model, \cite{lattimoreNIPS2016} analyzes this situation as a \textit{bandit problem}.
 

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
moviendo refs

@@
 \section{Introduction}
-A fundamental part of intelligent reasoning (\cite{lake2017building}, \cite{danks2014unifying}) is being able to make decisions under uncertain conditions.
+A fundamental part of intelligent reasoning  is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}).
 In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utilities.
@@
 
-Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment.
+Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment beyond what is expressable in an MDP.
 
-In this paper we consider the case where the uncertain environment is governed by a \textit{causal mechanism} unknown to the agent.
+In this paper we consider the case where the uncertain environment is governed by a fixed \textit{causal mechanism} unknown to the agent.
 Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment.

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
mejorando lo propuesto en la intro

@@
 
-We model the interaction between the decision maker and his environment a repeated game with incomplete information between a player called Nature whi selects his actions from the causal model and the decision maker.
+We model the interaction between the decision maker and his environment a repeated game with incomplete information between a player called Nature which selects his actions from the causal model and the decision maker and propose a way for the agent to have \textit{beliefs} about the causal model that governs the environment.
 

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
cambios related work

@@
 
-In a bandit problem (\cite{sutton1998reinforcement}) an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine wich is of the same kind and whose initial settings are independent of the previous machine and action.
+In a bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine wich is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement} .
 
@@
 
-\cite{lattimoreNIPS2016} is, up to our current knowledge, the first paper to consider causal relations between the effects of actions.
+Up to our current knowledge \cite{lattimoreNIPS2016} is the first paper to consider causal relations between the effects of actions.
 They consider a decision maker who must choose the best among several possible interventions on a given, and fully known, causal model.
@@
 
-\cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016} by considering a causal model partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal.
+By considering a causal model which is partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
 

### 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
cambio titulos secciones

@@
 
-We model the interaction between the decision maker and his environment a repeated game with incomplete information between a player called Nature which selects his actions from the causal model and the decision maker and propose a way for the agent to have \textit{beliefs} about the causal model that governs the environment.
+The interaction between the decision maker and his environment is modelled as a repeated game with incomplete information between a player called Nature which selects his actions from the causal model and the decision maker and propose a way for the agent to have \textit{beliefs} about the causal model that governs the environment.
 
@@
 
-We attempt to model the interaction between the decision maker and the environment as a repeated game between two players: the original decision maker and a new abstract player called Nature.
-The base game will be the original decision problem.
+We define a game between the decision maker and a new abstract player called Nature. The base game will be the original decision problem.
 
@@
 
-\subsection{If the causal model is completely known}
+\subsection{The casual model is completely known}
 If the decision maker knows the causal model, then she can obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made.
 
-The tratment that is the arg max for the highest probability of the desired value for the target variable will be chosen, and this choice will be a \textit{best response} and the max expected utility choice.
+The treatment which achieves the highest probability for the desired value of the target variable will be chosen.
+
+The action thus selected will be a \textit{best response} for the decision maker and the max expected utility choice.
 
@@
 \[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E) \]
-\subsection{Knowing only the structure}
+\subsection{Only the structure is known}
 If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.

### 2018-05-11 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
aqui modificare para version larga

--- 2018-05-10 Segundo_Semestre/CausalML_Workshop/Extended_Abstract.tex
@@
 \section{Related Work}
-Consider a one stage decision problem where the relation between actions and consequences is given by a causal model, \cite{lattimoreNIPS2016} analyzes this situation as a \textit{bandit problem}.
+Consider a one stage decision problem where the relation between actions and consequences is given by a causal model, \cite{lattimoreNIPS2016} analyze this situation as a \textit{bandit problem}.
 
-In a bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine wich is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement} .
+In a bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
 
-Several algorithms exists for finding the best arm in a multi-arm bandit, such as \cite{bubeck2009pure}, \cite{gabillon2012best}, \cite{agarwal2014taming} \cite{audibert2010best}, \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of this works consider causal-governed environments.
+Several algorithms exists for finding the best arm in a multi-arm bandit, such as \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of this works consider causal-governed environments.
 
@@
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
-% by Iain Murray in 2018. It was modified from a version from Dan Roy in
-% 2017, which was based on a version from Lise Getoor and Tobias
-% Scheffer, which was slightly modified from the 2010 version by
-% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
-% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
-% slightly modified from Prasad Tadepalli's 2007 version which is a
-% lightly changed version of the previous year's version by Andrew
-% Moore, which was in turn edited from those of Kristian Kersting and
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2018-05-14 Segundo_Semestre/CausalML_Workshop/Paper_short_version.tex
aqui modifico el de 2 pags
cambiando titulo por propuesta de sucar
cambios en intro segun correcciones de Sucar

--- 2018-05-11 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
@@
 \twocolumn[
-\icmltitle{Use and acquisition of causal information for decision making under uncertainty}
+\icmltitle{Playing against Nature: causal discovery for decision making under uncertainty}
 
@@
 
-In this paper we consider the case where the uncertain environment is governed by a fixed \textit{causal mechanism} unknown to the agent.
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}).
+
+In this paper which we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
 
@@
 \section{Related Work}
-Consider a one stage decision problem where the relation between actions and consequences is given by a causal model, \cite{lattimoreNIPS2016} analyze this situation as a \textit{bandit problem}.
-
-In a bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
+One stage decision problem where the relation between actions and consequences is given by a causal model, is analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem}.
+In bandit problems an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
 
@@
 By considering a causal model which is partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
+
+The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked.
+Here we propose to acquire, and use, causal information while interacting with the environment.
 
@@
 Instead, the agent will try to learn by interaction with the environment succesive rounds of decision making.
-Inside each round, any response from the environment will be independent from the previous rounds.
+Inside each round, any response from the environment will be independent from the previous rounds, but the actions of the decision maker, which will be based upon previously acquired causal information are expected to improve the utility for the agent.
 
@@
 \end{document}
+
+% This document was modified from the file originally made available by
+% Pat Langley and Andrea Danyluk for ICML-2K. This version was created
+% by Iain Murray in 2018. It was modified from a version from Dan Roy in
+% 2017, which was based on a version from Lise Getoor and Tobias
+% Scheffer, which was slightly modified from the 2010 version by
+% Thorsten Joachims & Johannes Fuernkranz, slightly modified from the
+% 2009 version by Kiri Wagstaff and Sam Roweis's 2008 version, which is
+% slightly modified from Prasad Tadepalli's 2007 version which is a
+% lightly changed version of the previous year's version by Andrew
+% Moore, which was in turn edited from those of Kristian Kersting and
+% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2018-05-14 Segundo_Semestre/CausalML_Workshop/Paper_short_version.tex
mencionando los 3 casos antes de seccion 5.1
cambio redaccion en only structure is known

@@
 
+We consider three cases of increasing complexity:
+\begin{itemize}
+\item The decision maker fully knows the causal model.
+\item The decision maker knows only the graph structure.
+\item The decision maker knows nothing about the causal model. 
+\end{itemize}
+
 \subsection{The casual model is completely known}
@@
 The decision maker (medic) doesn't know the probabilities, but since she knows the graph structure, she can explicitly a non interventional expression for the invterventional distribution and the values that appear in the expression can be estimated from succesive observations.
+
+\subsection{The model is not known}
+The causal model were unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
+The set of any previous knowledge and considerations will be known as \textit{beliefs}, which will take the form of a probability distributions over causal models.
 

### 2018-05-15 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
copiando related work

@@
 
+The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked.
+Here we propose to acquire, and use, causal information while interacting with the environment.
 \section{Problem setup}

### 2018-05-15 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
copiando subsecciones de cada caso

@@
 
-The decision maker (medic) doesn't know the probabilities, but since she knows the graph structure, she can explicitly a non interventional expression for the invterventional distribution and the values that appear in the expression can be estimated from succesive observations.
+The decision maker (medic) doesn't know the parameters of the model, but since she knows the graph structure, she can explicitly find a non interventional expression for the invterventional distribution and estimate from succesive observations the values that appear in the expression.
+
+\subsection{The model is not known}
+The causal model were unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
+The set of any previous knowledge and considerations will be known as \textit{beliefs}, which will take the form of a probability distributions over causal models.
 

### 2018-05-15 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
comentando caso1

@@
 
-%\begin{abstract}
-%We study how a decision maker can learn and use causal information in an uncertain environment which is governed by a causal model. 
+\begin{abstract}
+We consider decision problems under uncertainty $(A,E,C,u)$ where elements in $C$ and elements in $E$ are related through a causal mechanism which is unknown to the decision maker. 
+We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge in each round.
 %Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
 %The decision maker has beliefs over the causal model which she updates in each round. 
-%\end{abstract}
+\end{abstract}
 
@@
 \section{Related Work}
-One stage decision problem where the relation between actions and consequences is given by a causal model, is analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem}.
-In bandit problems an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
+One stage decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds.
+In a classic bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
 
@@
 Applying Pearl's \cite{pearl2009causality} do-calculus, we see that the interventional distribution $P_{do(Tr)}(V)$ is given by
-\[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E) \]
+\[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E). \]
+
+In fact, noting that the variable to be intervened by the decision maker doesn't has parents, one can explicitly calculate the probability of the target variable given a value for the treatment.
+
+Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.26$.
+
+On the other hand, if the treatment assigned is the pill, then the living probability is $0.4$
 \subsection{Only the structure is known}

### 2018-05-16 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
cambios intro

@@
 
-\icmlcorrespondingauthor{Cieua Vvvvv}{c.vvvvv@googol.com}
-\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}
-
 % You may provide any keywords that you
@@
 A fundamental part of intelligent reasoning  is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}).
-In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utilities.
-If enough information is not available, the decision maker could try to \textit{learn} from the environment by interacting with it.
+In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+On the other hand, if enough information is not available, the decision maker could try to \textit{learn} from the environment by interacting with it.
 
-Learning by interaction has been extensively studied in Reinforcement Learning \cite{sutton1998reinforcement}, but the most common used techniques are purely reactive (\cite{garnelo2016towards}) and do not consider any structure of the environment beyond what is expressable in an MDP.
+Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning  setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely reactive and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
 
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}).
-
-In this paper which we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}) ,we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
 

### 2018-05-16 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
explicando el modelado como un juego en la intro

@@
 
-Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning  setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely reactive and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
+Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely reactive and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
 
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}) ,we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
 
-The interaction between the decision maker and his environment is modelled as a repeated game with incomplete information between a player called Nature which selects his actions from the causal model and the decision maker and propose a way for the agent to have \textit{beliefs} about the causal model that governs the environment.
+The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment in form of a probability distribution over a suitable space.
+This beliefs will be updated according with what is observed.
+
+While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another and observing a reward, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.
 

### 2018-05-17 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
cambios redaccion en intro y related work

@@
 We consider decision problems under uncertainty $(A,E,C,u)$ where elements in $C$ and elements in $E$ are related through a causal mechanism which is unknown to the decision maker. 
-We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge in each round.
-%Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
-%The decision maker has beliefs over the causal model which she updates in each round. 
+We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
 \end{abstract}
@@
 \section{Introduction}
-A fundamental part of intelligent reasoning  is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}).
+A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}, \cite{pearlwhy}).
 In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
-On the other hand, if enough information is not available, the decision maker could try to \textit{learn} from the environment by interacting with it.
+On the other hand, if enough information is not available, the decision maker could attempt to \textit{learn} from the environment by interacting with it.
 
-Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely reactive and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
+Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process \cite{garnelo2016towards}.
 
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
 
@@
 
-While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another and observing a reward, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.
+While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another inside a model of the environemtn and observing a reward as this transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.
 
 \section{Related Work}
-One stage decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds.
+Decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds of action-observation in which only one action can be chosen.
 In a classic bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
@@
 
-By considering a causal model which is partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
+By considering a causal model which is partially known and intervening variables from the unknown part of the model and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
 
 The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked.
-Here we propose to acquire, and use, causal information while interacting with the environment.
+
+In this work we propose to acquire, by repeated interaction, causal information about the environment as well as using the current causal knowledge inside each round to make better decisions.
+By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a causal entity controls the outcomes.
 \section{Problem setup}
@@
 
-In fact, noting that the variable to be intervened by the decision maker doesn't has parents, one can explicitly calculate the probability of the target variable given a value for the treatment.
+In fact, noting that the variable to be intervened by the decision maker doesn't has parents in the graph one can explicitly calculate the probability of the target variable given a value for the treatment using variable elimination (\cite{koller2009probabilistic}) given the treatment as evidence.
 
-Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.26$.
+Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.28$.
 
-On the other hand, if the treatment assigned is the pill, then the living probability is $0.4$
+On the other hand, if the treatment assigned is the pill, then the living probability is $0.37$
 \subsection{Only the structure is known}

### 2018-05-18 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
comentando que en caso1 se tiene el mismo desempeño

@@
 We assume that the decision maker has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$.
-The decision maker does not know the probabilities underlying the causal model therefore he can not calculate the expected utility of any action.
+The decision maker does not know the probabilities nor the structure of the underlying causal model therefore he can not calculate the expected utility of any action.
 
@@
 After each round of the base game, the probabilities representing causal beliefs will be updated in a Bayesian way in order to guarantee Bayesian consistency \cite{shoham2008multiagent}.
+
+Consider the classical rain example from \cite{pearl2009causality} where a floor being wet causes it to be slippery and therefore that causes people to fall over it, but being wet can be caused by either a sprinkler turned on or because it recently rained, which itself is caused by the current season of the year.
+Suppose that a person who recently arrived in the city is about to go out.
+She doesn't know precisely which are the rainy months in this new place, but she believes that rainy dayws are almost the same that in her past home.
+She also noticed a nearby sprinkler which she believes is far away from her house to wet her floor.
+When she goes out and falls and notices the sprinkler turned on, that increases her belief about it affecting her floor.
+
+Even though she didn't knew the causal model itself, some days later she was certain that the sprinkler caused her floor to be wet and since the beliefs she had in her mind were about causal relations.
 \section{Test scenario}
-Consider a patient who arrives at a hospital and he can have either disease $A$ or $B$. The doctor can either give him some pill or send him into surgery.
-
+Consider a patient who arrives at a hospital and he can have either disease $A$ or $B$.
+The doctor can either give him some pill or send him into surgery.
 Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
-
 The causal model that governs this situation is shown in Figure \ref{causal_model}
@@
 \centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
-\caption{Causal graphical model: the target variable is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
 \label{causal_model}
@@
 
+For the sake of explanation we consider three separate cases:
+\begin{itemize}
+\item The decision maker fully knows the causal model.
+\item The decision maker knows only the structure of the causal model.
+\item The decision maker does not know the causal model.
+\end{itemize}
+
 \subsection{The casual model is completely known}
-If the decision maker knows the causal model, then she can obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made.
-
-The treatment which achieves the highest probability for the desired value of the target variable will be chosen.
-
-The action thus selected will be a \textit{best response} for the decision maker and the max expected utility choice.
+If the decision maker knows the causal model, then she can directly obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
+The action selected will be a \textit{best response} for the decision maker as well as the maximum expected utility choice.
 
@@
 
-Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.28$.
+On the other hand, if a decision maker hadn't known the causal nature of the problem, but had he known the probabilities of the possible outomes given his own actions then a choice could be made using expected utilities.
 
-On the other hand, if the treatment assigned is the pill, then the living probability is $0.37$
+What is shown with this gedankenexperiment is that in this full knowledge setting causal information achieves the same performance than simple probabilities.
+
 \subsection{Only the structure is known}
 If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.
-
 As usual in the previous case, Nature is considered to play first to asign a random state of the environment, which in this case is the disease.
@@
 The decision maker (medic) doesn't know the parameters of the model, but since she knows the graph structure, she can explicitly find a non interventional expression for the invterventional distribution and estimate from succesive observations the values that appear in the expression.
+If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm but meanwhile we are assuming that this information is available at the end of each round.
 
@@
 
+For this case we consider two sub-cases depending on the maximum number of variables in the causal model are known or not.
+
 \bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}

### 2018-05-18 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
Caso 2 ya puse una solucion posible

@@
 \subsection{The casual model is completely known}
-If the decision maker knows the causal model, then she can directly obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
+Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed by the realization of a random variable $Y$.
+The decision maker whishes to obtain a value from $Y$ that maximizes his utility.
+
+The relation between values of $Y$ and actions $a \in A$ is expressed through realizations of \textit{uncertain events}, which causally relate $a$ and values $y \in \textrm{Val}(Y)$ and this relations are expressed by a causal graphical model $\mathcal{G}$.
+
+This is the simplest case of the three mentioned because if the decision maker fully knows the causal model, then she can proceed as in classic decision problems by directly obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
 The action selected will be a \textit{best response} for the decision maker as well as the maximum expected utility choice.
 
-Applying Pearl's \cite{pearl2009causality} do-calculus, we see that the interventional distribution $P_{do(Tr)}(V)$ is given by
+For the test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying Pearl's \cite{pearl2009causality} do-calculus, we see that the interventional distribution $P_{do(Tr)}(V)$ is given by
 \[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E). \]
@@
 
-What is shown with this gedankenexperiment is that in this full knowledge setting causal information achieves the same performance than simple probabilities.
+What is shown with this \textit{gedankenexperiment} is that in this full knowledge setting causal information achieves the same performance than simple probabilities.
 
 \subsection{Only the structure is known}
-If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.
-As usual in the previous case, Nature is considered to play first to asign a random state of the environment, which in this case is the disease.
+If only the structure of the graph is known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
 
-The decision maker (medic) doesn't know the parameters of the model, but since she knows the graph structure, she can explicitly find a non interventional expression for the invterventional distribution and estimate from succesive observations the values that appear in the expression.
+The decision maker will attempt to learn from her uncertain environment by forming \textit{beliefs} over the unknown parameters (conditional probabilities) of the model and update them according to the observed outcomes.
+
+We consider a repeated game in which the base game is described by the narrative of the test scenario: 
+\begin{itemize}
+\item \textbf{Players:} The set of players of this game is the set whose elements are the original decision maker, and a new player called Nature.
+\item \textbf{Actions:} The actions for the decision maker are the possible treatments that the medic can give to patient.
+And the actions for the nature are the set of possible diseases, and the set of outcomes for the sick patient.
+\item \textbf{Preferences:} The decision maker satisfies the von Neumann-Morgenstern axioms of rationality and he receives a payment of $1$ if the patient lives and $0$ in other case.
+Nature is indifferent over outcomes.
+\item \textbf{Beliefs:} Since the decision maker doesn't know some aspects about her environment, she will encode any uncertainty about it in a probability distribution $p(\theta)$.
+\end{itemize}
+
+Since she knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
 If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm but meanwhile we are assuming that this information is available at the end of each round.
+
+Given the structure of the model, which is shown in Figure \ref{causal_model}, we see that the involved probabilities in any calculations are:
+\[ P(\textrm{disease}), P(\textrm{treatment}), P(\textrm{reaction} | \textrm{treatment}), \]
+\[P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
+denoting the variables in the model as $X_1 = \textrm{disease}$, $\textrm{treatment}=X_2$, $\textrm{reaction}=X_3$ and $\textrm{lives}=Y$ then a natural notation for the probabilities arises as following:
+\[ p_1 = P(\textrm{disease}), \]
+\[ p_2 = P(\textrm{treatment}), \]
+\[ p_3 = P(\textrm{reaction}),\]
+\[p_y= P(\textrm{lives}), \]
+\[ p_{3 | 2} = P(\textrm{reaction} | \textrm{treatment}), \]
+\[p_{y |1,2,3} = P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
+
+A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
+Since the decision maker pretends to \textit{use} her beliefs and not only update them it is necessary that in each round of the game she can use her current knowledge in order to choose an action.
+
+Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
+Proceeding this way, she will have a fully specified causal graphical model, which will be then used as in Case 1.
 

### 2018-05-18 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
por qué matrices en caso3

@@
 The relation between $\mathbb{P}$ and $P_{\mathcal{G}}$ is given by the Manipulation Theorem of \cite{spirtes2000causation} and the Do-Calculus rules from \cite{pearl2009causality}.
+We also require that the condition known as Causal Sufficiency is satisfied by the model, which means there are not any causes lying \textit{outside} of the model.
 
@@
 
+Notice that the base game is an extensive game with imperfect information since the decision maker makes a choice without knowing the play made by Nature, which is the asigning a disease to the patient, and then assigning the outcome for the patient after the medic has chosen a treatment.
+
 Since she knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
-If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm but meanwhile we are assuming that this information is available at the end of each round.
+If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum} but meanwhile we are assuming that this information is available at the end of each round.
 
@@
 \subsection{The model is not known}
-The causal model were unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
-The set of any previous knowledge and considerations will be known as \textit{beliefs}, which will take the form of a probability distributions over causal models.
+The causal model were fully unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
+Again, consider the set of any previous knowledge and considerations will as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over causal models.
 
-For this case we consider two sub-cases depending on the maximum number of variables in the causal model are known or not.
+For this third case, we make an extra assumption: the maximum number of variables in the model is known to the decision maker.
+This allow to consider only distributions of fixed dimension.
+In this case we will consider distributions over matrixes of fixed size rather than vectors as in the previous case.
 

### 2018-05-18 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
comentarios caso3

@@
 
+As in the previous case, consider a repeated game where the base game consists of Nature assigning a disease to a patient, a medic choosing a treatment for it and Nature once again choosing the final outcome.
+This final choice of Nature is governed by the causal model.
+
+In this game, as well as in the previous one, the decision maker will attempt to learn by updating, and using, beliefs in a suitable way.
+
+The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
+
 \bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}

### 2018-05-21 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
comentando sobre distr Dirichlet

@@
 
-For the test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying Pearl's \cite{pearl2009causality} do-calculus, we see that the interventional distribution $P_{do(Tr)}(V)$ is given by
-\[ \sum_{E,R}P(V|E,Tr,R)P(R|Tr)P(E). \]
+Pearl's  do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed in terms of observational distributions as follows:
+\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)) \]
+
+For the test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the previous formula that the interventional distribution $P_{do(Tr)}(V)$ is given by
+\[ P(V | do(Tr))=P(Y | E, Tr, R)P(R | Tr) P(E). \]
 
@@
 
-A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
-Since the decision maker pretends to \textit{use} her beliefs and not only update them it is necessary that in each round of the game she can use her current knowledge in order to choose an action.
+Since these distributions fully characterize the model, the decision maker will have beliefs over each one of this parameters.
 
-Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
-Proceeding this way, she will have a fully specified causal graphical model, which will be then used as in Case 1.
+Notice that each of this parameter is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$
 
+A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
+
+The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
+
+\[ \frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha-1},\]
+
+where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$
+
+The Dirichlet distribution is useful and practical since it is conjugate for itself \cite{bernardo2000bayesian}.
+
+So, in this way, the decision maker will have beliefs about the CPT's.
+Using the current beliefs, a causal graphical model can be specified.
+Using this fully specified (structure + parameters), the decision maker will make her choice as in Case 1.
+
+Previously we argued that the agent's beliefs were going to be \textit{distributions} over a suitable space, but what is going to be updated are the parameters of such distributions.
+Namely, the $\alpha$ corresponding to the Dirichlet random variable assigned to each CPT.
+
+For the belief updating, given a new data point,  two cases must be considered:
+\begin{itemize}
+\item The variable to update has no parents.
+\item The variable to update has parents.
+\end{itemize}
+
+In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
+\[ \textrm{Dirichlet}(\alpha + c) \]
+where $c$ is a vector of the number of occurrences of that observed data point.
+
+For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
+Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the probability $P(X=i | Pa(X)=j)$ In this case, if the prior of $X$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X$ with parents $Pa(X)$ and an observed data point is given by
+\[ \prod_j \textrm{Dirichlet}(\alpha + u(j)) \]
+where $u(j)$ counts the number of times that in the data point the parents of $X$ had the value $j$.
 \subsection{The model is not known}
@@
 
+A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
+Since the decision maker pretends to \textit{use} her beliefs and not only update them it is necessary that in each round of the game she can use her current knowledge in order to choose an action.
+
+Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
+Proceeding this way, she will have the structure of causal graphical model, which will be then used as in Case 2.
+
 \bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}

### 2018-05-21 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
con figurita

@@
 where $u(j)$ counts the number of times that in the data point the parents of $X$ had the value $j$.
+
+\subsubsection{Experiment}
+As proof of concept we implemented this belief updating. We started with a random assignation of the $\alpha$ parameter for each of the distributions considered.
+
+With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable.
+In the medic scenario, this desired value corresponds to the patient living.
+With this action chosen, we simulate from the causal graphical model using the chosen action as evidence.
+This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
+
+If we plot the average rewards obtained we observe they converge to 1 as shown in the next figure.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/rewar.png}}
+\caption{Average reward obtained by the decision maker in 50 rounds of the decision makin-belief updating cycle.}
+\label{average_reward}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
 \subsection{The model is not known}
 The causal model were fully unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
-Again, consider the set of any previous knowledge and considerations will as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over causal models.
+Again, consider the set of any previous knowledge and considerations will as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over a suitable space.
 
@@
 
-A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
+In the previous case the decision maker knew the form of the CPT's involved in any calculation. In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
+
+If the decision maker knew exactly what variables are involved, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in case 2.
+The Dirichlet Process, which was introduced by \cite{ferguson1973bayesian}, is a stochastic process whose finite-dimensional distributions are Dirichlet random variables \cite{hjort2010bayesian}, \cite{ghosal2017fundamentals}.
+It is the most common \textit{nonparametric prior} in the context of Bayesian Nonparametrics.
+
+Belief updating in this context is yet to be studied.
+
+%A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
 Since the decision maker pretends to \textit{use} her beliefs and not only update them it is necessary that in each round of the game she can use her current knowledge in order to choose an action.
 
-Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
+%Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
 Proceeding this way, she will have the structure of causal graphical model, which will be then used as in Case 2.

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
Correcciones Sucar: abstract

@@
 
-\[ \frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha-1},\]
+\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1},\]
 
@@
 For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
-Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the probability $P(X=i | Pa(X)=j)$ In this case, if the prior of $X$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X$ with parents $Pa(X)$ and an observed data point is given by
-\[ \prod_j \textrm{Dirichlet}(\alpha + u(j)) \]
-where $u(j)$ counts the number of times that in the data point the parents of $X$ had the value $j$.
+Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
+In this case, if the prior of $X_i$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X_i$ with parents $Pa(X)$ and an observed data point is given by
+\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
 
@@
 
-If we plot the average rewards obtained we observe they converge to 1 as shown in the next figure.
-
+The agent obtained a total reward of $47$ out of $50$ posible points
 \begin{figure}[ht]
@@
 
+On the other hand, an agent that is randomly choosing one of the two treatment options obtains a reward of $37$ out of $50$ possible points
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/random_reward.png}}
+\caption{Average reward obtained by the decision maker in 50 rounds of randomly choosing a treatment in the causal model.}
+\label{random_reward}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
 \subsection{The model is not known}
@@
 Again, consider the set of any previous knowledge and considerations will as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over a suitable space.
-
-For this third case, we make an extra assumption: the maximum number of variables in the model is known to the decision maker.
-This allow to consider only distributions of fixed dimension.
-In this case we will consider distributions over matrixes of fixed size rather than vectors as in the previous case.
 
@@
 The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
+In the previous case the decision maker knew the form of the CPT's involved in any calculation.
+In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
 
-In the previous case the decision maker knew the form of the CPT's involved in any calculation. In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
+If the decision maker knew exactly the variables that appear in the true model thar governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in case 2.
+The Dirichlet Process\footnote{with parameters $M,G_0$}, which was introduced by \cite{ferguson1973bayesian}, is random measure defined over a space $S$ such that for each partition $B_1,...,B_k$ the vector $(G(B_1),...,G(B_k))$ follows a Dirichlet distribution \cite{hjort2010bayesian}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}.
 
-If the decision maker knew exactly what variables are involved, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in case 2.
-The Dirichlet Process, which was introduced by \cite{ferguson1973bayesian}, is a stochastic process whose finite-dimensional distributions are Dirichlet random variables \cite{hjort2010bayesian}, \cite{ghosal2017fundamentals}.
-It is the most common \textit{nonparametric prior} in the context of Bayesian Nonparametrics.
-
-Belief updating in this context is yet to be studied.
+Belief updating using causal information is yet to be studied and left as future work.
 

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
Correcciones Hugo: en la intro Nature en italics

@@
 \begin{abstract}
-We consider decision problems under uncertainty $(A,E,C,u)$ where elements in $C$ and elements in $E$ are related through a causal mechanism which is unknown to the decision maker. 
+We consider decision problems under uncertainty where the decisions and the outcomes are related through a causal mechanism which is unknown to the decision maker. 
 We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
@@
 
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
+Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), even though this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
 
@@
 
-While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another inside a model of the environemtn and observing a reward as this transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.
+While the standard setting in RL is to model the interaction agent-environment as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as this transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model.
 

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
Correcciones Hugo: en la intro this por these

@@
 
-While the standard setting in RL is to model the interaction agent-environment as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as this transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model.
+While the standard setting in RL is to model the interaction agent-environment as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model.
 

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
Correcciones Hugo: en related work

@@
 
-Several algorithms exists for finding the best arm in a multi-arm bandit, such as \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of this works consider causal-governed environments.
+Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight},  \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of these works consider causal-governed environments.
 
 Up to our current knowledge \cite{lattimoreNIPS2016} is the first paper to consider causal relations between the effects of actions.
-They consider a decision maker who must choose the best among several possible interventions on a given, and fully known, causal model.
+They consider a decision maker who must choose the best among several possible interventions on a given causal model.
 The optimality of the action in this context is in terms of the minimal regret.

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
pegando primeras secciones

--- 2018-05-22 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
@@
 
-Up to our current knowledge \cite{lattimoreNIPS2016} is the first paper to consider causal relations between the effects of actions.
+As far as we know \cite{lattimoreNIPS2016} is the first paper to consider causal relations between the effects of actions.
 They consider a decision maker who must choose the best among several possible interventions on a given causal model.
@@
 
-By considering a causal model which is partially known and intervening variables from the unknown part of the model and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
+By considering a causal model which is partially known and intervening variables from the unknown part of the model and by avoiding sampling arms that are considered sub-optimal, \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
 
-The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked.
+The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices, but the problem of \textit{acquiring} this causal knowledge is left unatacked.
 
@@
 
-Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ where family of uncertain events $E$ and the set of consequences $C$ are causally related.
+Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ in which an agent has to choose one among several options $a \in A$ which are causally related to the elements of $C$.
 We assume that the decision maker has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$.
@@
 
-We define a game between the decision maker and a new abstract player called Nature. The base game will be the original decision problem.
-
+We define a game between the decision maker and a new abstract player called Nature.
+The base game will be the original decision problem.
 Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model.
@@
 Even though she didn't knew the causal model itself, some days later she was certain that the sprinkler caused her floor to be wet and since the beliefs she had in her mind were about causal relations.
-\section{Test scenario}
-Consider a patient who arrives at a hospital and he can have either disease $A$ or $B$.
-The doctor can either give him some pill or send him into surgery.
-Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
-The causal model that governs this situation is shown in Figure \ref{causal_model}
 
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-
-The node corresponding to the variable lives is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the treatment.
-
-In this scenario, Nature's move will consist in randomly assigning the disease to the patient.
-Then, the medic will asign a treatment, the decision nodes for this play form an information set because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
-Finally, Nature will sample the consequence of the treatment from the causal model and the medic will observe the outcome.
-
-For the sake of explanation we consider three separate cases:
-\begin{itemize}
-\item The decision maker fully knows the causal model.
-\item The decision maker knows only the structure of the causal model.
-\item The decision maker does not know the causal model.
-\end{itemize}
-
-\subsection{The casual model is completely known}
-Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed by the realization of a random variable $Y$.
-The decision maker whishes to obtain a value from $Y$ that maximizes his utility.
-
-The relation between values of $Y$ and actions $a \in A$ is expressed through realizations of \textit{uncertain events}, which causally relate $a$ and values $y \in \textrm{Val}(Y)$ and this relations are expressed by a causal graphical model $\mathcal{G}$.
-
-This is the simplest case of the three mentioned because if the decision maker fully knows the causal model, then she can proceed as in classic decision problems by directly obtain the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
-The action selected will be a \textit{best response} for the decision maker as well as the maximum expected utility choice.
-
-Pearl's  do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed in terms of observational distributions as follows:
-\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)) \]
-
-For the test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the previous formula that the interventional distribution $P_{do(Tr)}(V)$ is given by
-\[ P(V | do(Tr))=P(Y | E, Tr, R)P(R | Tr) P(E). \]
-
-In fact, noting that the variable to be intervened by the decision maker doesn't has parents in the graph one can explicitly calculate the probability of the target variable given a value for the treatment using variable elimination (\cite{koller2009probabilistic}) given the treatment as evidence.
-
-On the other hand, if a decision maker hadn't known the causal nature of the problem, but had he known the probabilities of the possible outomes given his own actions then a choice could be made using expected utilities.
-
-What is shown with this \textit{gedankenexperiment} is that in this full knowledge setting causal information achieves the same performance than simple probabilities.
-
-\subsection{Only the structure is known}
-If only the structure of the graph is known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
-
-The decision maker will attempt to learn from her uncertain environment by forming \textit{beliefs} over the unknown parameters (conditional probabilities) of the model and update them according to the observed outcomes.
-
-We consider a repeated game in which the base game is described by the narrative of the test scenario: 
-\begin{itemize}
-\item \textbf{Players:} The set of players of this game is the set whose elements are the original decision maker, and a new player called Nature.
-\item \textbf{Actions:} The actions for the decision maker are the possible treatments that the medic can give to patient.
-And the actions for the nature are the set of possible diseases, and the set of outcomes for the sick patient.
-\item \textbf{Preferences:} The decision maker satisfies the von Neumann-Morgenstern axioms of rationality and he receives a payment of $1$ if the patient lives and $0$ in other case.
-Nature is indifferent over outcomes.
-\item \textbf{Beliefs:} Since the decision maker doesn't know some aspects about her environment, she will encode any uncertainty about it in a probability distribution $p(\theta)$.
-\end{itemize}
-
-Notice that the base game is an extensive game with imperfect information since the decision maker makes a choice without knowing the play made by Nature, which is the asigning a disease to the patient, and then assigning the outcome for the patient after the medic has chosen a treatment.
-
-Since she knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
-If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum} but meanwhile we are assuming that this information is available at the end of each round.
-
-Given the structure of the model, which is shown in Figure \ref{causal_model}, we see that the involved probabilities in any calculations are:
-\[ P(\textrm{disease}), P(\textrm{treatment}), P(\textrm{reaction} | \textrm{treatment}), \]
-\[P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
-denoting the variables in the model as $X_1 = \textrm{disease}$, $\textrm{treatment}=X_2$, $\textrm{reaction}=X_3$ and $\textrm{lives}=Y$ then a natural notation for the probabilities arises as following:
-\[ p_1 = P(\textrm{disease}), \]
-\[ p_2 = P(\textrm{treatment}), \]
-\[ p_3 = P(\textrm{reaction}),\]
-\[p_y= P(\textrm{lives}), \]
-\[ p_{3 | 2} = P(\textrm{reaction} | \textrm{treatment}), \]
-\[p_{y |1,2,3} = P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
-
-Since these distributions fully characterize the model, the decision maker will have beliefs over each one of this parameters.
-
-Notice that each of this parameter is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$
-
-A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
-
-The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
-
-\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1},\]
-
-where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$
-
-The Dirichlet distribution is useful and practical since it is conjugate for itself \cite{bernardo2000bayesian}.
-
-So, in this way, the decision maker will have beliefs about the CPT's.
-Using the current beliefs, a causal graphical model can be specified.
-Using this fully specified (structure + parameters), the decision maker will make her choice as in Case 1.
-
-Previously we argued that the agent's beliefs were going to be \textit{distributions} over a suitable space, but what is going to be updated are the parameters of such distributions.
-Namely, the $\alpha$ corresponding to the Dirichlet random variable assigned to each CPT.
-
-For the belief updating, given a new data point,  two cases must be considered:
-\begin{itemize}
-\item The variable to update has no parents.
-\item The variable to update has parents.
-\end{itemize}
-
-In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
-\[ \textrm{Dirichlet}(\alpha + c) \]
-where $c$ is a vector of the number of occurrences of that observed data point.
-
-For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
-Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
-In this case, if the prior of $X_i$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X_i$ with parents $Pa(X)$ and an observed data point is given by
-\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
-
-\subsubsection{Experiment}
-As proof of concept we implemented this belief updating. We started with a random assignation of the $\alpha$ parameter for each of the distributions considered.
-
-With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable.
-In the medic scenario, this desired value corresponds to the patient living.
-With this action chosen, we simulate from the causal graphical model using the chosen action as evidence.
-This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
-
-The agent obtained a total reward of $47$ out of $50$ posible points
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/rewar.png}}
-\caption{Average reward obtained by the decision maker in 50 rounds of the decision makin-belief updating cycle.}
-\label{average_reward}
-\end{center}
-\vskip -0.2in
-\end{figure}
-
-On the other hand, an agent that is randomly choosing one of the two treatment options obtains a reward of $37$ out of $50$ possible points
-
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/random_reward.png}}
-\caption{Average reward obtained by the decision maker in 50 rounds of randomly choosing a treatment in the causal model.}
-\label{random_reward}
-\end{center}
-\vskip -0.2in
-\end{figure}
-
-\subsection{The model is not known}
-The causal model were fully unknown, the decision maker will have to deal with the problem using only any previous knowledge and his own intuitions.
-Again, consider the set of any previous knowledge and considerations will as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over a suitable space.
-
-As in the previous case, consider a repeated game where the base game consists of Nature assigning a disease to a patient, a medic choosing a treatment for it and Nature once again choosing the final outcome.
-This final choice of Nature is governed by the causal model.
-
-In this game, as well as in the previous one, the decision maker will attempt to learn by updating, and using, beliefs in a suitable way.
-
-The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
-In the previous case the decision maker knew the form of the CPT's involved in any calculation.
-In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
-
-If the decision maker knew exactly the variables that appear in the true model thar governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in case 2.
-The Dirichlet Process\footnote{with parameters $M,G_0$}, which was introduced by \cite{ferguson1973bayesian}, is random measure defined over a space $S$ such that for each partition $B_1,...,B_k$ the vector $(G(B_1),...,G(B_k))$ follows a Dirichlet distribution \cite{hjort2010bayesian}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}.
-
-Belief updating using causal information is yet to be studied and left as future work.
-
-%A natural way of encoding beliefs about causal relations is to consider all the possible variables and assign probabilities to the edges that connect them in a Directed Aciclyc Graph (DAG).
-Since the decision maker pretends to \textit{use} her beliefs and not only update them it is necessary that in each round of the game she can use her current knowledge in order to choose an action.
-
-%Had she a DAG in mind whose edges have associated weights, she could sample a random variable with a $U(0,1)$ distribution and see if that sample value is less than the probability attached to the edge, if so, then she fixes this edge in the model.
-Proceeding this way, she will have the structure of causal graphical model, which will be then used as in Case 2.
-
-\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
-\bibliographystyle{icml2018}
-
-\end{document}
+\section{Proposed Method}

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
Añadiendo caso 1

@@
 By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a causal entity controls the outcomes.
-\section{Problem setup}
+\section{Problem setup}{\label{problem_setup}}
 By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).
@@
 \section{Proposed Method}
+In this section we describe our approach for studying decision making in causal environments as described in Section \ref{problem_setup}.
+
+For the sake of explanation we consider three separate cases:
+
+\begin{itemize}
+\item The decision maker fully knows the causal model.
+\item The decision maker knows only the structure of the causal model.
+\item The decision maker does not know the causal model.
+\end{itemize}
+
+\subsection{The causal model is completely known}
+Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed by the realization of a random variable $Y$ which we will call \textit{target variable}.
+The decision maker whishes to obtain a value from $Y$ that maximizes his utility.
+
+The relation between values of $Y$ and actions $a \in A$ is expressed through realizations of \textit{uncertain events}, which causally relate $a$ and values $y \in \textrm{Val}(Y)$ and this relations are expressed by a causal graphical model $\mathcal{G}$.
+The action chosen by the decision maker is considered as an \textit{intervention} over some variable in this causal model.
+
+This is the simplest case of the three mentioned because if the decision maker fully knows the causal model, then she can proceed as in classic decision problems by directly obtaining the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
+The action selected will be a \textit{best response} for the decision maker as well as the maximum expected utility choice.
+
+Pearl's do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed in terms of observational distributions as follows:
+\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)) \]
+
+If a decision maker knows the causal model that is governing her environment, then she could use this expression to find out the probability of her desired value for the target variable and choose the action that maximizes this probability.

### 2018-05-22 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
experiments

@@
 
-Consider the classical rain example from \cite{pearl2009causality} where a floor being wet causes it to be slippery and therefore that causes people to fall over it, but being wet can be caused by either a sprinkler turned on or because it recently rained, which itself is caused by the current season of the year.
-Suppose that a person who recently arrived in the city is about to go out.
-She doesn't know precisely which are the rainy months in this new place, but she believes that rainy dayws are almost the same that in her past home.
-She also noticed a nearby sprinkler which she believes is far away from her house to wet her floor.
-When she goes out and falls and notices the sprinkler turned on, that increases her belief about it affecting her floor.
-
-Even though she didn't knew the causal model itself, some days later she was certain that the sprinkler caused her floor to be wet and since the beliefs she had in her mind were about causal relations.
-
 \section{Proposed Method}
@@
 Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed by the realization of a random variable $Y$ which we will call \textit{target variable}.
-The decision maker whishes to obtain a value from $Y$ that maximizes his utility.
+The relation between values of $Y$ and actions $a \in A$ is expressed by a causal graphical model $\mathcal{G}$, which is known by the decision maker.
+The decision maker whishes to obtain a value from $Y$ that maximizes her utility.
+It is assumed that which variable is going to be intervened is known for the decision maker.
 
-The relation between values of $Y$ and actions $a \in A$ is expressed through realizations of \textit{uncertain events}, which causally relate $a$ and values $y \in \textrm{Val}(Y)$ and this relations are expressed by a causal graphical model $\mathcal{G}$.
-The action chosen by the decision maker is considered as an \textit{intervention} over some variable in this causal model.
-
-This is the simplest case of the three mentioned because if the decision maker fully knows the causal model, then she can proceed as in classic decision problems by directly obtaining the probabilities of different values for the target variable given that an intervention on \textit{treatment} is made and choose the treatment which achieves the highest probability for the desired value of the target variable will be chosen.
+This is the simplest case of the three mentioned because if the decision maker fully knows the causal model, then she can proceed as in classic decision problems by directly obtaining the probabilities of different values for the target variable given that an action is made and choose $a$ which achieves the highest probability for the desired value of the target variable.
 The action selected will be a \textit{best response} for the decision maker as well as the maximum expected utility choice.
 
-Pearl's do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed in terms of observational distributions as follows:
-\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)) \]
+Pearl's do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed in terms of \textit{observational} distributions as follows:
+\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)). \]
 
-If a decision maker knows the causal model that is governing her environment, then she could use this expression to find out the probability of her desired value for the target variable and choose the action that maximizes this probability.
+The decision maker can use this expression to find the probabilities for her desired value of the target variable given the possible interventions available to her.
+
+\subsection{Only the structure is known}
+If only the structure of $\mathcal{G}$ this known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
+
+In this case, the decision maker will attempt to learn from her uncertain environment by forming \textit{beliefs} over unknown parameters of the environment and update them according to the observed outcomes.
+
+To model the interaction between the decision maker and the environment, we consider a repeated game in which the base game is described by the narrative of the decision problem: 
+\begin{itemize}
+\item \textbf{Players:} The set of players of this game is the set whose elements are the original decision maker, and a new player called Nature.
+\item \textbf{Actions:} The actions for the decision maker are the available options she has in the decision problem;i.e.
+$A$.
+The actions of Nature are possible realizations of the variables of the causal model.
+This is interpreted as Nature assigning a \textit{state} over the environment that is  either the effect of the action chosen by the agent, or it is assigning some initial random state.
+\item \textbf{Preferences:} The decision maker satisfies the von Neumann-Morgenstern axioms of rationality and therefore it is assumed to be maximizing expected utilities.
+Nature is indifferent over outcomes.
+\item \textbf{Beliefs:} Since the decision maker has uncertainty about her environment, she will encode it in a probability distribution $p(\theta)$ over a suitable space.
+\end{itemize}
+
+In this game we will assume that Nature moves first and assigns some \textit{state} to the environment which is unknown to the decision maker.
+For this reason, the base game is an extensive game with imperfect information since the decision maker makes a choice without knowing the play made by Nature.
+
+Since she knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
+If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum} but meanwhile we are assuming that this information is available at the end of each round.
+
+Given the structure of the model;i.e., the variables in it and the directed edges, we fan factorize the joint distribution as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the DAG corresponding to $\mathcal{G}$.
+
+Since these distributions fully characterize the model, the decision maker will have beliefs over each one of these parameters.
+
+Notice that each of this parameter is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
+As a proof of concept we will later consier only binary variables.
+
+A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
+
+The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
+
+\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1},\]
+
+where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$
+
+The Dirichlet distribution is useful and practical since it is conjugate for itself \cite{bernardo2000bayesian}.
+
+So, in this way, the decision maker will have beliefs about the CPT's.
+Using the current beliefs, a causal graphical model can be specified.
+Using this fully specified (structure + parameters), the decision maker will make her choice as in Case 1.
+When the decision maker observes the value of the target variable, she will update the parameters the specify her beliefs.
+
+Previously we argued that the agent's beliefs were going to be \textit{distributions} over a suitable space, but what is going to be updated are the parameters of such distributions.
+Namely, the $\alpha$ corresponding to the Dirichlet random variable assigned to each CPT.
+
+For the belief updating, given a new data point,  two cases must be considered:
+\begin{itemize}
+\item The variable to update has no parents.
+\item The variable to update has parents.
+\end{itemize}
+
+In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
+\[ \textrm{Dirichlet}(\alpha + c) \]
+where $c$ is a vector of the number of occurrences of that observed data point.
+
+For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
+Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
+In this case, if the prior of $X_i$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X_i$ with parents $Pa(X)$ and an observed data point is given by
+\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
+
+\section{Test scenario}
+To show how the previous ideas could be applied, we consider a hypothetical example.
+
+Consider a patient who arrives at a hospital who can either have disease $A$ or disease $B$.
+The doctor can either give him some pill or send him into surgery.
+Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
+The causal model that governs this situation is shown in Figure \ref{causal_model}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+The node corresponding to the variable lives is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the treatment.
+
+In this scenario, Nature's move will consist in randomly assigning the disease to the patient.
+Then, the medic will asign a treatment, the decision nodes for this play form an \textit{information set} because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
+Finally, Nature will sample the consequence of the treatment from the causal model and the medic will observe the outcome.
+
+For this  test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
+\[ P(Y | do(Tr))=P(Y | E, Tr, R)P(R | Tr) P(E). \]
+
+In fact, from the structure of the model, which is shown in Figure \ref{causal_model}, we see that the involved probabilities in any calculations are:
+\[ P(\textrm{disease}), P(\textrm{treatment}), P(\textrm{reaction} | \textrm{treatment}), \]
+and
+\[P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
+
+\section{Experiment}
+As proof of concept we implemented this belief updating.
+
+\subsection{Case 1: The causal model is completely known}
+If the causal model is completely known to the decision maker, then in one step she can obtain the probability for her desired value of the target variable, which in this case is the value corresponding to the outcome in which the patient lives at the end.
+Using this probability, she can choose which treatment to assign.
+
+\subsection{Case 2}
+
+We started with a random assignation of the $\alpha$ parameter for each of the distributions considered.
+We are considering a Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
+Since each of the variables in the model is binary, then the product of these Dirichlets is Dirichlet.
+
+With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable as in Case 1.
+With this action chosen, we simulate an outcome from the causal graphical model using the chosen action as an intervention.
+This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
+
+\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
+\bibliographystyle{icml2018}
+
+\end{document}

### 2018-05-23 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
cambios

@@
 
-Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), even though this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
+Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect \cite{rottman2014reasoning}, we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
 
@@
 While the standard setting in RL is to model the interaction agent-environment as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model.
+
+The agent, using her current beliefs, will generate a causal model and choose an action from it as if that model was the true one.
+Then, after she observes the consequences of her actions, her beliefs will be updated according to the observed information in order to make a better choice the next time.
+The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true one.
 
@@
 By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a causal entity controls the outcomes.
+
 \section{Problem setup}{\label{problem_setup}}
@@
 The relation between $\mathbb{P}$ and $P_{\mathcal{G}}$ is given by the Manipulation Theorem of \cite{spirtes2000causation} and the Do-Calculus rules from \cite{pearl2009causality}.
-We also require that the condition known as Causal Sufficiency is satisfied by the model, which means there are not any causes lying \textit{outside} of the model.
+We also require that the condition known as \textit{Causal Sufficiency} is satisfied by the model, which means there are not any causes lying \textit{outside} of the model.
 
 Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ in which an agent has to choose one among several options $a \in A$ which are causally related to the elements of $C$.
-We assume that the decision maker has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$.
+The decision maker seeks to maximize her utility and we assume that she has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$ \cite{gilboa2009decision}.
 The decision maker does not know the probabilities nor the structure of the underlying causal model therefore he can not calculate the expected utility of any action.
 
-Instead, the agent will try to learn by interaction with the environment succesive rounds of decision making.
-Inside each round, any response from the environment will be independent from the previous rounds, but the actions of the decision maker, which will be based upon previously acquired causal information are expected to improve the utility for the agent.
+Instead, the agent will try to learn by interacting with the environment through succesive rounds of decision making.
+Inside each round, any response from the environment will be independent from the previous rounds, but the actions of the decision maker will be based upon previously acquired causal information and are expected to improve the utility for the agent.
 
 We define a game between the decision maker and a new abstract player called Nature.
-The base game will be the original decision problem.
-Nature will be indifferent among the different possible outcomes of the game and will select its actions from the causal model.
+The base game's structure will be the same as the original decision problem.
+Nature will be indifferent among the possible outcomes of the game and will select its actions from the causal model.
+This is interpreted as the \textit{causal response} from the environment to the actions of the decision maker.
 Nature having objectives to pursue (non-constant payments) will be left as future work.
 
-\section{Belief formation and updating}
+\subsection{Belief formation and updating}
 For an agent to reason about and modify his causal knowledge we endow her with a probability distribution $p(\theta)$ over a suitable space.
-Since we are considering causal graphical models, and for each DAG there exists a one-to-one correspondence to a matrix, one way to express beliefs about DAG's is using distributions over matrices for the structure of the graph.
+The beliefs must allow to form a \textit{local} model in a given moment to be used for decision making.
 
-Since the beliefs must behave as a \textit{local} model in a given moment, it is preferable to assign probabilities directly to causal models rather than edges in a model in order to have a fixed model when sampling from this probability assignement.
+We will later exploit the fact that causal graphical models can be expressed in terms of conditional distributions, so having beliefs about a causal model is equivalent to having beliefs about these conditional distributions.
+Beliefs about finite discrete-valued distributions can be encoded as probability distributions over probability vectors.
+In particular, the Dirichlet distribution will be very useful.
 
-After each round of the base game, the probabilities representing causal beliefs will be updated in a Bayesian way in order to guarantee Bayesian consistency \cite{shoham2008multiagent}.
+After each round of the game, the beliefs will be updated in a Bayesian way in order to achieve convergence towards the true model \cite{shoham2008multiagent}.
 
@@
 \subsection{The causal model is completely known}
-Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed by the realization of a random variable $Y$ which we will call \textit{target variable}.
+Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed as the outcome of a random variable $Y$ which we will call \textit{target variable}.
 The relation between values of $Y$ and actions $a \in A$ is expressed by a causal graphical model $\mathcal{G}$, which is known by the decision maker.
-The decision maker whishes to obtain a value from $Y$ that maximizes her utility.
-It is assumed that which variable is going to be intervened is known for the decision maker.
+The decision maker whishes to choose an action $a \in A$ such that the observed value of $Y$ maximizes her utility.
+It is assumed that the variable is going to be intervened is known for the decision maker.
 
@@
 \subsection{Only the structure is known}
-If only the structure of $\mathcal{G}$ this known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
+If only the graphical structure of $\mathcal{G}$ this known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
 
 In this case, the decision maker will attempt to learn from her uncertain environment by forming \textit{beliefs} over unknown parameters of the environment and update them according to the observed outcomes.
+In order to make a Bayesian update over the parameters, these must be defined in such way that in each round the decision maker can define a causal model from the parameters in order to make a decision using this model as if it was the true one as in Case 1.
 
@@
 
-Since she knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
+Since the decision maker knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
 If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum} but meanwhile we are assuming that this information is available at the end of each round.
 
-Given the structure of the model;i.e., the variables in it and the directed edges, we fan factorize the joint distribution as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the DAG corresponding to $\mathcal{G}$.
+Given the structure of the model; i.e., the variables in it and the directed edges, the joint distribution of those variables can be expressed as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the underlying DAG in $\mathcal{G}$.
 
@@
 
-So, in this way, the decision maker will have beliefs about the CPT's.
-Using the current beliefs, a causal graphical model can be specified.
-Using this fully specified (structure + parameters), the decision maker will make her choice as in Case 1.
+In this way, the decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
+Using the agent's current beliefs, a causal graphical model can be specified.
+Using this fully specified (structure + parameters) as a true model, the decision maker will make her choice as in Case 1.
 When the decision maker observes the value of the target variable, she will update the parameters the specify her beliefs.
@@
 Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
-In this case, if the prior of $X_i$ is given by a a Dirichlet($\alpha$), then the posterior for a variable $X_i$ with parents $Pa(X)$ and an observed data point is given by
+In this case, if the prior of $X_i$ conditioned on its parents having the value $j$ is given by a a Dirichlet($\alpha$), then the posterior for the variable $X_i$ given an observed data point is given by
 \[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
+
+\subsection{The model is not known}
+The causal model were fully unknown, the decision maker will have to deal with the problem using only any previous knowledge and her own intuitions.
+Again, any previous knowledge and considerations will be expressed as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over a suitable space.
+
+As in the previous case, we consider a repeated game where the base game consists of Nature assigning a random state of the environment and responding to the agents choices with the effects that were caused by her decisions
+
+In this game, as well as in the previous one, the decision maker will attempt to learn by updating, and using, beliefs in a suitable way.
+
+The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
+In the previous case the decision maker knew the form of the CPT's involved in any calculation.
+In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
+
+If the decision maker knew exactly the variables that appear in the true model that governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in Case 2 and updating the parameters of the process using the observed information.
+The Dirichlet Process\footnote{with parameters $M,G_0$}, which was introduced by \cite{ferguson1973bayesian}, is random measure defined over a space $S$ such that for each partition $B_1,...,B_k$ the vector $(G(B_1),...,G(B_k))$ follows a Dirichlet distribution \cite{hjort2010bayesian}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}.
+
+Belief updating using causal information is yet to be studied and left as future work.
 
@@
 Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
+The doctor could be facing a mutation from a known disease, so she has some knowledge about what could happen if a treatment is given to the patient.
+Using her previous knowledge as a true model, she can choose a treatment and observe the outcome from which she will learn about this disease, so she could make a better decision the next time a similar patient arrives.
+
 The causal model that governs this situation is shown in Figure \ref{causal_model}
@@
 
-The node corresponding to the variable lives is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the treatment.
+The variable \textit{Lives} is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the variable \textit{Treatment}.
+The decision maker prefers an outcome in which the patient lives.
 
-In this scenario, Nature's move will consist in randomly assigning the disease to the patient.
-Then, the medic will asign a treatment, the decision nodes for this play form an \textit{information set} because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
+In this scenario, Nature's move will consist in randomly assigning a disease to the patient.
+Then, the medic will asign a treatment using his current beliefs about the disease and the possible outcomes.
+The decision nodes for this play of the medic form an \textit{information set} because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
 Finally, Nature will sample the consequence of the treatment from the causal model and the medic will observe the outcome.
@@
 For this test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
-\[ P(Y | do(Tr))=P(Y | E, Tr, R)P(R | Tr) P(E). \]
+\[ P(Y | do(Tr))=P(Y | D, Tr, R)P(R | Tr) P(D). \]
 
@@
 
+We can also see that the joint distribution for all of the variables can be expressed as
+\[ P(Y | D, Tr, R)P(R | Tr) P(D)P(Tr). \]
+This expression will be useful when specifying beliefs about the model as Dirichlet distributions.
 \section{Experiment}
@@
 Using this probability, she can choose which treatment to assign.
+Since this action maximizes the probability of the occurence desired value, it maximizes the expected utility, and it is also a \textit{best response} to the player Nature.
 
 \subsection{Case 2}
+From the expression of the joint probability we notice that we need to specify a distribution over each one of these distributions, which will be each one a Dirichlet distribution.
 
-We started with a random assignation of the $\alpha$ parameter for each of the distributions considered.
-We are considering a Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
+We begin with a random assignation of the $\alpha$ parameter for each of the distributions considered.
+We use Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
 Since each of the variables in the model is binary, then the product of these Dirichlets is Dirichlet.
@@
 
+We show the results of two experiments.We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
+In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/50_rounds.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{50_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/100_rounds.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{100_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\section{Conclusion and Future Work}
+In this work we have proposed a way to make decisions in uncertain environments which are known to be governed by causal mechanisms.
+The proposed decision making procedure attempts to resemble how human beings act when causal relations are present.
+Human beings are known to use, and modify, causal knowledge when making decisions.
+
+This ideas motivated us to study how an autonomous agent could learn from her environment when it has a particular structure, in this case being causal relations what gives the environment a certain structure.
+
+We assumed here that the decision maker is aware of the causal nature of the environment, but lacks information about its specific parameters which are to be learned by interaction.
+It is reasonable to assume that the variables involved are known, since in many situations we are aware of what we are intervening upon and what do we expect it to affect.
+
+The problem of discovering the variables itself and the connections between them is far more general and it is left as future work.
+
 \bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}

### 2018-05-23 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
cambios de Sucar

@@
 % \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
+%\usepackage[draft]{hyperref}
 \usepackage{hyperref}
@@
 \begin{abstract}
-We consider decision problems under uncertainty where the decisions and the outcomes are related through a causal mechanism which is unknown to the decision maker. 
+We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker.
 We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
@@
 \section{Introduction}
-A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}, \cite{pearlwhy}).
-In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}, \cite{pearlwhy}).
+In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the classic formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
 On the other hand, if enough information is not available, the decision maker could attempt to \textit{learn} from the environment by interacting with it.
@@
 
-The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment in form of a probability distribution over a suitable space.
-This beliefs will be updated according with what is observed.
+The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment and a way to update them after an outcome has been observed.
 
-While the standard setting in RL is to model the interaction agent-environment as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model.
+While the standard setting in RL is to model the agent-environment interaction as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model in response to what the decision maker has chosen.
 
-The agent, using her current beliefs, will generate a causal model and choose an action from it as if that model was the true one.
+The agent, using her current beliefs, will generate a \textit{local} causal model and choose an action from it as if that model was the true one.
 Then, after she observes the consequences of her actions, her beliefs will be updated according to the observed information in order to make a better choice the next time.
-The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true one.
+The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true model.
+
+Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed.
 
 \section{Related Work}
-Decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds of action-observation in which only one action can be chosen.
+Decision problems in which the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds of action-observation in which only one action can be chosen.
 In a classic bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.
@@
 \section{Problem setup}{\label{problem_setup}}
-By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).
+By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric \cite{spirtes2000causation}.
 
@@
 Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ in which an agent has to choose one among several options $a \in A$ which are causally related to the elements of $C$.
+The elements in $E$ are uncertain events which are governed by a causal mechanism.
+This means that when the decision maker choses an action $a$, an uncertain event will occurr in such a way that an outcome $c$ will occur, which is causally related to $a$.
 The decision maker seeks to maximize her utility and we assume that she has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$ \cite{gilboa2009decision}.
@@
 
-\subsection{Belief formation and updating}
 For an agent to reason about and modify his causal knowledge we endow her with a probability distribution $p(\theta)$ over a suitable space.
@@
 \section{Proposed Method}
-In this section we describe our approach for studying decision making in causal environments as described in Section \ref{problem_setup}.
-
-For the sake of explanation we consider three separate cases:
+In this section we describe our approach for studying decision making in causal environments as described in Section \ref{problem_setup}. For the sake of explanation we consider three separate cases:
 
@@
 The decision maker whishes to choose an action $a \in A$ such that the observed value of $Y$ maximizes her utility.
-It is assumed that the variable is going to be intervened is known for the decision maker.
+It is assumed that the variable that is going to be intervened is known by the decision maker; i.e., she knows what variable can she intervene.
 
@@
 \begin{itemize}
-\item \textbf{Players:} The set of players of this game is the set whose elements are the original decision maker, and a new player called Nature.
+\item \textbf{Players:} The set of players of this game is the set whose elements are the original decision maker, and a player called Nature.
 \item \textbf{Actions:} The actions for the decision maker are the available options she has in the decision problem;i.e.
@@
 The actions of Nature are possible realizations of the variables of the causal model.
-This is interpreted as Nature assigning a \textit{state} over the environment that is  either the effect of the action chosen by the agent, or it is assigning some initial random state.
 \item \textbf{Preferences:} The decision maker satisfies the von Neumann-Morgenstern axioms of rationality and therefore it is assumed to be maximizing expected utilities.
@@
 For this reason, the base game is an extensive game with imperfect information since the decision maker makes a choice without knowing the play made by Nature.
+Nature's moves are interpreted as a \textit{response} from the causal model to the actions of the decision maker.
 
-Since the decision maker knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about this unknown quantities.
-If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum} but meanwhile we are assuming that this information is available at the end of each round.
+Since the decision maker knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about these unknown quantities.
+If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum}, but meanwhile we are assuming that this information is available at the end of each round.
 
 Given the structure of the model; i.e., the variables in it and the directed edges, the joint distribution of those variables can be expressed as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the underlying DAG in $\mathcal{G}$.
-
 Since these distributions fully characterize the model, the decision maker will have beliefs over each one of these parameters.
-
-Notice that each of this parameter is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
-As a proof of concept we will later consier only binary variables.
+Notice that each of these parameters is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
+As a proof of concept we will later consider only binary variables.
 
 A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
-
 The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
@@
 
-where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$
-
-The Dirichlet distribution is useful and practical since it is conjugate for itself \cite{bernardo2000bayesian}.
+where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$. The Dirichlet distribution is useful since it is conjugate for itself \cite{bernardo2000bayesian}.
 
@@
 Using this fully specified (structure + parameters) as a true model, the decision maker will make her choice as in Case 1.
-When the decision maker observes the value of the target variable, she will update the parameters the specify her beliefs.
+When the decision maker observes the value of the target variable, she will update the parameters that specify her beliefs.
 
@@
 
-As in the previous case, we consider a repeated game where the base game consists of Nature assigning a random state of the environment and responding to the agents choices with the effects that were caused by her decisions
-
+As in the previous case, we consider a repeated game where the base game consists of Nature assigning a random state of the environment and responding to the agents choices with the effects that were caused by her decisions.
 In this game, as well as in the previous one, the decision maker will attempt to learn by updating, and using, beliefs in a suitable way.
@@
 The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
-In the previous case the decision maker knew the form of the CPT's involved in any calculation.
+In the previous case the decision maker knew the form of the Conditional Probability Tables (CPT) involved in any calculation.
 In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
@@
 
-Belief updating using causal information is yet to be studied and left as future work.
+Belief updating using causal information when the decision maker doesn't know the structure of the model nor its parameters is yet to be studied and left as future work.
 
 \section{Test scenario}
-To show how the previous ideas could be applied, we consider a hypothetical example.
+To illustrate how the previous ideas could be applied, we consider the following hypothetical example.
 
@@
 
-The causal model that governs this situation is shown in Figure \ref{causal_model}
+The causal model that governs this situation is shown in Figure \ref{causal_model}.
+The parameters for this model were fixed intuitively in such a way that each treatment is effective for only one disease, but the most effective treatment is riskier.
 
@@
 
-We show the results of two experiments.We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
-In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making.
+We show the results of two experiments.
+We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
+We show the average perform over $20, 50, 100$ and $200$ rounds.
+
+In Figure \ref{20_rounds} we observe the average rewards for each agent in 20 rounds of decision making.
+Here we notice that Q-learning outperforms our algorithm, which has a similar performance as the random choosin procedure until round 11.
 
@@
 \begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/50_rounds.png}}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{20_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making. Our algorithm follow closely the Q-learning agent and outperform the random agent.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}}
 \caption{Average reward obtained in each round for each agent}
@@
 
-In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds
+In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds, where our algorithm slightly outperforms Q-Learning.
 
@@
 \begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/100_rounds.png}}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}}
 \caption{Average reward obtained in each round for each agent}
@@
 \end{figure}
+
+In Figure \ref{200_rounds} we observe the average reward obtained by the three agents in 200 rounds. The average reward obtained is very similar for Q-learning and our algorithm.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{200_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+We see that our method obtains a very similar reward as the classic Q-learning algorithm for a larger number of rounds, where the random agent is outperformed, but our model offers something extra because it learns a causal model of the environment.
 
@@
 
+The experimental results show that considering causal structure in a decision making process yields a good performance when compared to non-causal classic algorithms, but our model has the extra feature of learning a causal model of the environment which could be exported for problems involving similar scenarios.
+
 The problem of discovering the variables itself and the connections between them is far more general and it is left as future work.

### 2018-05-23 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
añadiendo mails

@@
 
-\icmlaffiliation{inaoe}{Department of Computer Science, National Institute of Astrophysics Optics and Electronics (INAOE), Puebla, Mexico}
+\icmlaffiliation{inaoe}{Department of Computer Science, National Institute of Astrophysics Optics and Electronics (INAOE), Mexico}
+
+\icmlcorrespondingauthor{Mauricio Gonzalez-Soto}{mauricio@inaoep.mx}
+\icmlcorrespondingauthor{Luis Enrique Sucar}{esucar@inaoep.mx}
+\icmlcorrespondingauthor{Hugo Jair Escalante}{hugojair@inaoep.mx}
 
@@
 \section{Experiment}
-As proof of concept we implemented this belief updating.
+As proof of concept we implemented  Case 1 and Case 2 for the test scenario and compared it with an agent performing Q-learning and an agent choosing his actions at random.
+
+For the implementation, we defined a \textit{true} causal model as the one shown in Figure \ref{causal_model} using the library Pgmpy \cite{ankan2015pgmpy}.
+Then, we defined an agent which has beliefs for each of the CPT that appear in the factorization for the model and randomly assigned values for the $\alpha$ parameters for each one of the Dirichlet distributions as mentioned before.
+
+This agent will find the action that maximizes her desired value for the target variable using do-calculus.
+The action thus selected will be used as \textit{evidence} in the true causal model and a MAP inference will be used to simulate the most likely outcome given this action.
+The target variable will output a $1$ value if the patient lives.
+The value of the target variable will be the reward of each round.
 

### 2018-05-23 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
abstract
version enviada

@@
 We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
+We propose a decision making procedure in which an agent holds \textit{beliefs} about her environment which are used to make a choice and then are updated using the observed outcome.
+As proof of concept, we present an implementation of this causal decision making model and apply it to a simple problem.
+We show that the model achieves a performance similar to the classic Q-learning while it also acquires a causal model of the environment.
 \end{abstract}
@@
 \subsection{Only the structure is known}
-If only the graphical structure of $\mathcal{G}$ this known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
+If only the graphical structure of $\mathcal{G}$ is known, then it is not obvious how to find the best action to make since the information required for calculating expected utilities is not available.
 
@@
 Notice that each of these parameters is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
-As a proof of concept we will later consider only binary variables.
 
@@
 
-If the decision maker knew exactly the variables that appear in the true model that governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in Case 2 and updating the parameters of the process using the observed information.
+If the decision maker knew which variables appear in the true model that governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in Case 2 and updating the parameters of the process using the observed information.
 The Dirichlet Process\footnote{with parameters $M,G_0$}, which was introduced by \cite{ferguson1973bayesian}, is random measure defined over a space $S$ such that for each partition $B_1,...,B_k$ the vector $(G(B_1),...,G(B_k))$ follows a Dirichlet distribution \cite{hjort2010bayesian}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}.
@@
 The parameters for this model were fixed intuitively in such a way that each treatment is effective for only one disease, but the most effective treatment is riskier.
+
+The variables in the model are: 
+\begin{itemize}
+\item \textbf{Disease:} Either $A$ or $B$.
+\item \textbf{Treatment:} Either pill or surgery.
+\item \textbf{Reaction:} Either dying or surviving.
+\item \textbf{Lives:} Either living or dying.
+\end{itemize}
+
+The variables are causally related as shown in Figure \ref{causal_model}.
 
@@
 
-\subsection{Case 2}
+\subsection{Case 2: Only the structure is known}
 From the expression of the joint probability we notice that we need to specify a distribution over each one of these distributions, which will be each one a Dirichlet distribution.

### 2018-05-28 Propuesta/Propuesta.tex
empezamos correcciones de Felipe
redaccion

@@
 Esa misma realidad, altamente compleja e interconectada, contiene una serie de relaciones \textit{causa-efecto} que podemos utilizar al momento de tomar decisiones.
-El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) y predecir las consecuencias de nuestros actos.
+El conocimiento causal permite preguntarnos  \textit{qué pasaría si...} (\cite{stalnaker2016knowledge}) e inferir las consecuencias de nuestros actos.
 Por desgracia, es difícil que contemos con la información necesaria para llevar a cabo estas predicciones de manera correcta, a lo más, podemos contar con estimaciones sobre lo probable que es la ocurrencia de algún evento incierto dada cierta acción.\\
@@
 \indent En el lenguaje cotidiano, la palabra \textit{causa} evoca cierta regularidad o estabilidad del entorno; cuando hablamos de enfermedades existentes queremos pensar en las causas de esta para tomar acciones preventivas.
-De la misma manera, al estar aprendiendo a realizar cierta tarea prestamos atención a los efectos de alguna acción y asumimos que la acción causó el efecto y lo tendremos en mente para futuros escenarios similares.Tanto así, que los seres humanos formamos explicaciones o interpretaciones causales sin contar con toda la información disponible, lo cual puede llevar a errores sistemáticos al momento de tomar decisiones o emitir juicios (\cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
+Al estar aprendiendo a realizar cierta tarea prestamos atención a los efectos de alguna acción y asumimos que la acción causó el efecto y lo tendremos en mente para futuros escenarios similares.Tanto así, que los seres humanos formamos explicaciones o interpretaciones causales sin contar con toda la información disponible, lo cual puede llevar a errores sistemáticos al momento de tomar decisiones o emitir juicios (\cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
 La pregunta natural es cómo utilizar esta información causal de manera efectiva para que un agente inteligente tome decisiones de manera correcta.\\

### 2018-05-28 Propuesta/Propuesta.tex
quitando palabra promedio como sinonimo de esperado

@@
 Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.
-La utilidad esperada, como su nombre lo indica es el valor esperado, o promedio, de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
+La utilidad esperada, como su nombre lo indica es el valor esperado de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
 \\

### 2018-06-05 Propuesta/Propuesta.tex
cambios en la intro

@@
 Al estar aprendiendo a realizar cierta tarea prestamos atención a los efectos de alguna acción y asumimos que la acción causó el efecto y lo tendremos en mente para futuros escenarios similares.Tanto así, que los seres humanos formamos explicaciones o interpretaciones causales sin contar con toda la información disponible, lo cual puede llevar a errores sistemáticos al momento de tomar decisiones o emitir juicios (\cite{tversky1977causal}, \cite{tversky1980causal}, \cite{kahneman2011thinking}.
-La pregunta natural es cómo utilizar esta información causal de manera efectiva para que un agente inteligente tome decisiones de manera correcta.\\
+Una pregunta que surge de manera natural es cómo utilizar esta información causal de manera efectiva para que un agente inteligente tome decisiones de manera correcta.\\
 \\
-\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, considera a un tomador de decisiones que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto, el cual influirá en las consecuencias de las acciones.
-Para resolver este problema a partir de una serie de axiomas se propone un criterio formal de elección, el cual consiste en la maximización de la utilidad esperada.
+\indent En este trabajo, consideramos un agente inteligente que debe tomar una decisión en un entorno incierto, el cual está \textit{gobernado} por un modelo causal fijo, pero desconocido para el agente.
+El agente sabe que sus acciones, el impacto de éstas en su entorno, y la consecuencia obtenida, están relacionadas de manera causal, por lo que buscará hacer uso de esta \textit{estructura} en su entorno para la toma de decisiones.
+\\
+\\
+\indent La Teoría de la Decisión bajo Incertidumbre, desarrollada en los trabajos de \cite{von1944theory}, \cite{definetti1930}, \cite{definetti1937}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, considera a un tomador de decisiones racional\footnote{Siguiendo a \cite{bernardo2000bayesian}, \cite{binmore2008rational},\cite{gilboa2009decision} decimos que un tomador de decisiones es racional si su relación de preferencias satisface una serie de axiomas.} que se enfrenta al problema de escoger una de entre varias opciones (o acciones) posibles en un escenario incierto.
+El tomador de decisiones expresa sus preferencias entre dos acciones en una relación binaria $\succeq$.
+Los trabajos de \cite{von1944theory}, \cite{definetti1937}, \cite{savage1954the} proveen de un criterio formal para la toma de decisiones, el cual consiste en la maximización de la utilidad esperada.
 La utilidad esperada, como su nombre lo indica es el valor esperado de la utilidad que le producen al tomador de decisiones las consecuencias de sus acciones; para poder calcularla, es necesario contar con estimaciones de la probabilidad de ocurrencia de ciertos eventos inciertos.
@@
 \\
-\indent En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno, sobre esto ya se ha dicho mucho en el área de Aprendizaje por Refuerzo (\cite{sutton1998reinforcement}).
-Una de las grandes limitaciones de estos métodos es que aprenden formas de actuar que son puramente reactivas; es decir, no capturan, ni intentan, la estructura del ambiente.
-Un caso particular de esta estructura se da en el caso en el cual el tomador de decisiones sabe, o cree, que en el ambiente en el cual se llevan a cabo sus decisiones existen relaciones causa-efecto.\\
+\indent En diversas ocasiones, un tomador de decisiones no tiene a su alcance todos los parámetros (probabilidades de los eventos), pero puede intentar \textit{aprenderlos} del mismo entorno mediante la interacción con éste (\cite{sutton1998reinforcement}).
+Los métodos más comunes de Aprendizaje por Refuerzo aprenden políticas que son puramente reactivas; es decir, se construyen a partir de estimaciones estadísticas, pero no capturan ninguna estructura del entorno más allá de lo que se puede expresar en un Proceso de Decisión Markoviano (MDP por sus siglas en inglés).
+
+\indent Debido a que los seres humanos utilizamos y modificamos la información causal con la que contamos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), proponemos que un agente autónomo puede utilizar y adquirir información causal a partir de su interacción con su entorno, el cual es gobernado por un modelo causal.
+La manera en la que se propone que el agente tome en cuenta, aprenda y utilice la información causal del entorno es dotarlo de \textit{creencias} sobre el modelo causal del entorno, las cuales serán utilizadas y modificadas conforme el agente interactúa con el entorno.
 \\
-\indent La clase de problemas en los cuales un agente debe aprender por interacción se formulan, usualmente, como un agente autónomo que se encuentra dentro de un entorno o ambiente el cual es modificado por las acciones del agente (\cite{sutton1998reinforcement}).
-A su vez, tal modificación del entorno afecta al agente en términos de una señal de recompensa.
-En este trabajo se propone modelar la interacción entre el agente y el entorno como una \textit{juego}\footnote{Un juego es una situación estratégica en la que los intereses de dos o más actores se enfrentan y se afectan entre sí.} entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza, cuyas acciones estarán controladas por un modelo causal.\\
+\indent Mientras que en Aprendizaje por Refuerzo la interacción agente-entorno se modela como transiciones entre estados así como una recompensa asociada a cada transición, en este trabajo se propone modelar la interacción entre el agente y el entorno como una \textit{juego}\footnote{Un juego es una situación estratégica en la que los intereses de dos o más actores se enfrentan y se afectan entre sí.} entre dos actores: el mismo agente o tomador de decisiones y un ente abstracto que llamaremos la Naturaleza, cuyas acciones estarán controladas por un modelo causal.
+Además, como el agente se enfrenta a problemas de decisión de una sola ronda, vamos a considerar \textit{juegos repetidos} en donde el juego a repetir (llamado \textit{juego base} en la literatura) está relacionado con el problema de decisión inicial.
 \\
-\indent La manera en la que se propone que el agente tome en cuenta, aprenda y utilice la información causal del entorno es dotarlo de \textit{creencias} sobre la estructura causal, las cuales serán utilizadas para la toma de decisiones.
-En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con John Nash.}.
-Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.
-Los juegos estocásticos fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+\\
+\indent En cada ronda del juego, el agente utilizará sus \textit{creencias} como si fueran un modelo causal verdadero, el cual utilizará para tomar una buena decisión.
+Posteriormente, a partir del resultado observado, las creencias serán modificadas de modo que se tome una mejor decisión en la siguiente ronda.
+De este modo, además de aprender una política, el agente estará aprendiendo un modelo causal de su entorno.\\
 \\
@@
 Este conocimiento será utilizado dentro de la misma interacción con el fin de que el agente logre un objetivo de manera más rápida, debido a que la incorporación de la estructura causal reduce la incertidumbre pues permite al agente predecir las consecuencias de sus acciones.
-Además, el conocimiento causal aprendido por el agente será exportable para su uso en problemas de decisión relacionados al mismo entorno.
+Además, el conocimiento causal aprendido por el agente será exportable para su uso en problemas de decisión relacionados al mismo entorno.\\
+\\
+\indent  En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con John Nash.}.
+Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.
+Los juegos estocásticos fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+
 \subsection{Motivación}
@@
 Por definir.
-\section{Conclusiones}
-Por añadir.
+\subsection{Plan de publicaciones}
+
 \section{Limitaciones}

### 2018-06-05 Propuesta/Propuesta.tex
intercambiando motivacion y justificacion

@@
 
-\indent Debido a que los seres humanos utilizamos y modificamos la información causal con la que contamos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), proponemos que un agente autónomo puede utilizar y adquirir información causal a partir de su interacción con su entorno, el cual es gobernado por un modelo causal.
+\indent Debido a que las relaciones causa-efecto son un tipo de \textit{estructura} de un ambiente y que los seres humanos utilizamos y modificamos la información causal con la que contamos para escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), y aunque este aprendizaje no es perfecto  (\cite{rottman2014reasoning}), proponemos que un agente autónomo puede utilizar y adquirir información causal a partir de su interacción con su entorno, el cual es gobernado por un modelo causal.
 La manera en la que se propone que el agente tome en cuenta, aprenda y utilice la información causal del entorno es dotarlo de \textit{creencias} sobre el modelo causal del entorno, las cuales serán utilizadas y modificadas conforme el agente interactúa con el entorno.
@@
 \\
-\indent  En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con John Nash.}.
+%\indent  En el contexto de Teoría de Juegos, los modelos que permiten que el agente tenga \textit{incertidumbre} sobre las acciones de otros actores son conocidos como \textit{juegos con información imperfecta}.Por otro lado, los juegos en los que se permite que el agente tenga incertidumbre sobre el \textit{estado del mundo} y por lo tanto tenga \textit{creencias} sobre distintos aspectos de este son conocidos como juegos con información incompleta o Juegos Bayesianos, que fueron introducidos por \cite{harsanyi1967games1}, \cite{harsanyi1968games2}, \cite{harsanyi1968games3} y por lo que le dieron el Nobel en 1994\footnote{Compartido con John Nash.}.
 Finalmente, la clase de los \textit{Juegos Estocásticos}, permiten flexibilizar los escenarios a los que se enfrentan los tomadores de decisiones, pues en esta clase el escenario cambia con el tiempo.
 Los juegos estocásticos fueron introducidos por \cite{shapley1953stochastic} y generalizan los Procesos de Decisión Markovianos (MDP), así como los juegos repetido, pues un MDP es simplemente un juego estocástico con un solo jugador mientras que un juego repetido es un juego estocástico de una sola etapa (\cite{shoham2008multiagent}).\\
+
+\subsection{Justificación}
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
+\\
+\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
+Siguiendo esta idea, \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
+En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
+Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
+El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
+\\
+\indent Posteriormente, el trabajo de \cite{sen2017identifying} extiende el trabajo de \cite{lattimoreNIPS2016} al caso en el cual el modelo causal es parcialmente conocido, pero sigue abierto el problema para el caso en el cual el agente deba aprender el modelo conforme interactúa con su entorno.\\
+\\
+\indent Además, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
+\\
+\\
+\indent En resumen, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar la interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
+A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.\\
+\\
 
@@
 \\
-\indent Los seres humanos, utilizamos de manera correcta la información causal con la que contamos con tal de escoger las acciones que lleven a cierto resultado deseado (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}).
-Además, modificamos las creencias previas que se tenían previamente sobre las relaciones causales en su ambiente conforme interactúan con él (\cite{hagmayer2013repeated}).
-Aunque, estos procesos de adquisición y actualización no son perfectos; por ejemplo, al estimar la fuerza de los efectos se cometen errores o el supuesto Markoviano que requieren los modelos causales es violado (\cite{rottman2014reasoning}).
-\\
-\\
 \indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto que si no utilizaran información causal del entorno.
-En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.\\
-\\
-\indent Lo que este trabajo señala es que a la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
+En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.
+Lo que este trabajo señala es que a la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
 La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
@@
 Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
-\subsection{Justificación}
-Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
-\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
-\\
-\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
-Siguiendo esta idea,  \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
-En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
-El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
-\\
-\indent Posteriormente, el trabajo de \cite{sen2017identifying} extiende el trabajo de \cite{lattimoreNIPS2016} al caso en el cual el modelo causal es parcialmente conocido, pero sigue abierto el problema para el caso en el cual el agente deba aprender el modelo conforme interactúa con su entorno.\\
-\\
-\indent Además, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
-\\
-\\
-\indent En resumen, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar la interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
-A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.\\
-\\
+
 En lo que sigue del documento, primero mencionaremos las herramientas teóricas requeridas para el desarrollo de esta propuesta.

### 2018-06-05 Propuesta/Propuesta.tex
de nuevo intercambiando titulos de motivacion y justificacion

@@
 
-\subsection{Justificación}
-Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera eficiente; en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\subsection{Motivación}
+Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{spirtes2000causation},\cite{hagmayer2013repeated},\cite{pearlwhy}).
+Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
+De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
+Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
+\\
+\\
+
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera más rápida; vemos que en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
 \cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
@@
 
-\subsection{Motivación}
-Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{spirtes2000causation},\cite{hagmayer2013repeated},\cite{pearlwhy}).
-Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
-De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
-Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
-\\
-\\
+\subsection{Justificación}
+
 \indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto que si no utilizaran información causal del entorno.

### 2018-06-05 Propuesta/Propuesta.tex
puse trabajo previo despues de intro

@@
 
-\subsection{Motivación}
-Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{spirtes2000causation},\cite{hagmayer2013repeated},\cite{pearlwhy}).
-Además, las relaciones causales son un tipo de \textit{estructura} del ambiente, la cual puede utilizarse para que un agente pueda razonar sobre su entorno.
-De hecho, se ha estudiado el caso de agentes que son racionales en un entorno causal (\cite{board2006equivalence}).
-Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
-\\
-\\
-
-Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera más rápida; vemos que en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
-\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
-\\
-\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
-Siguiendo esta idea, \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
-En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
-Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
-El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
-\\
-\indent Posteriormente, el trabajo de \cite{sen2017identifying} extiende el trabajo de \cite{lattimoreNIPS2016} al caso en el cual el modelo causal es parcialmente conocido, pero sigue abierto el problema para el caso en el cual el agente deba aprender el modelo conforme interactúa con su entorno.\\
-\\
-\indent Además, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
-\\
-\\
-\indent En resumen, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar la interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
-A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.\\
-\\
-
-\subsection{Justificación}
-
-\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto que si no utilizaran información causal del entorno.
-En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.
-Lo que este trabajo señala es que a la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
-La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
-La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
-\\
-\indent Modificar una a una las relaciones causales que se expresan en un modelo causal como se hace en \cite{wellen2012learning} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos \textit{locales} de las estructuras causal de nuestro entorno (\cite{danks2014unifying}) y posteriormente en unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
-Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
-Por otro lado, el conocimiento inicial del tomador de decisiones es un punto delicado, pues a partir de qué tan completo o no sea, así de buenas serán sus acciones iniciales.
-\cite{billot2005probabilities} muestran cómo asignar probabilidades iniciales dado un conjunto de observaciones.\\
-\\
-\indent Encontrar la acción óptima a llevar a cabo en un contexto incierto y de manera tal que el tomador de decisiones pueda aprender algo del ambiente se ha abordado desde el punto de vista de Proceso de Decisión Markovianos y Programación Dinámica, lo que dio lugar al área conocida como Aprendizaje por Refuerzo (RL por sus siglas en inglés).
-En este contexto, el agente que debe aprender a llevar a cabo una tarea se encuentra ante un entorno incierto, y sobre el cual debe aprender mediante interacción.
-El agente lleva a cabo una acción y observa una señal de recompensa del ambiente.
-La señal de recompensa debe ser tal que, al intentar el agente maximizarla en el largo plazo, el agente cumpla su objetivo.\\
-\\
-\indent Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
-En este trabajo, lo que se propone es modelar la interacción agente-entorno como un \textit{juego} de modo que el agente podrá observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
-Además, al ser la señal de recompensa algo que debe diseñarse de acuerdo al problema, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
-Considerar a la naturaleza como un jugador no es una idea nueva, pues desde \cite{milnor1951games}, \cite{papadimitriou1985games}, \cite{Szep1985} y \cite{eiselt2004games} se ha considerado a la Naturaleza como un jugador que \textit{asigna} inicialmente un estado al ambiente.\\
-\\
-\indent Que un agente pueda \textit{aprender} algo de su entorno a lo largo de una serie de interacciones, o juegos, ha sido estudiado ya en el caso en el que el jugador busca aprender los pagos del juego o las estrategias de alguno de los otros agentes \cite{fudenberg1998learning} lo que motiva a pensar que un agente podría aprender sobre el modelo causal que rige su entorno.
-\\
-\\
-\indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del \textit{razonamiento contrafactual} para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
-Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
-
-En lo que sigue del documento, primero mencionaremos las herramientas teóricas requeridas para el desarrollo de esta propuesta.
-Posteriormente, los trabajos previos que han atacado este mismo problema, o similares.
-Luego, la pregunta de investigación y la hipótesis relacionada, así como los Objetivos generales y específicos.
 \section{Marco Teórico}
@@
 Esto significa que las creencias de un jugador eventualmente van a converger a la verdad si hace actualizaciones bayesianas y juega una mejor respuesta dadas sus creencias; pero esto no implica (\cite{shoham2008multiagent}) que el jugador aprenderá la estrategia verdadera utiizada por los otros jugadores
+
+\subsection{Motivación (por qué quiero hacer esto, hueco en el conocimiento)}
+Las relaciones causa-efecto\footnote{Esto será definido formalmente más adelante} son de gran utilidad en problemas de decisión, pues conociendo los \textit{efectos} de las posibles decisiones a tomar, el agente puede planear mejor sus acciones (\cite{spirtes2000causation}, \cite{hagmayer2013repeated}, \cite{pearlwhy}).
+\\
+
+Contar con información sobre la estructura causal de cierto entorno puede hacer que un proceso de toma de decisiones se lleve a cabo de manera más rápida; vemos que en el trabajo de \cite{lattimoreNIPS2016} se muestra cómo al contar con un modelo causal del ambiente se pueden encontrar acciones óptimas de manera más rápida que si sólo se llevara a cabo una exploración a ciegas con tal de encontrar la mejor acción como llevan a cabo \cite{audibert2010best}.
+\cite{lattimoreNIPS2016} consideran el caso en el cual un tomador de decisiones quiere maximizar cierta variable de recompensa y las acciones que tiene a su disposición el agente están relacionadas de manera causal con la variable recompensa y entre sí.\\
+\\
+\indent Se sabe que los seres humanos conciben sus acciones en el mundo como \textit{intervenciones} sobre este (\cite{hagmayer2009decision}).
+Siguiendo esta idea, \cite{lattimoreNIPS2016} consideran las acciones disponibles a un agente inteligente como \textit{intervenciones} posibles en un modelo causal; la dinámica que proponen consiste en que el agente realice una acción y posterior a esto se observa el valor de la variable recompensa así como la realización de las otras variables del modelo que no fueron intervenidas; en esto consiste una \textit{ronda}.
+En la siguiente ronda, el entorno vuelve a su estado inicial y el agente selecciona otra posible intervención y vuelve a observar estas variables.
+Esta dinámica de rondas repetidas en la que cada ronda inicia sin tomar en cuenta lo sucedido en la anterior lleva a los autores a modelar el problema como escoger brazos en una máquina tragamonedas (conocidas como \textit{bandits} en la literatura), observar el pago y moverse a una máquina nueva.
+El algoritmo que desarrollan identifica la \textit{intervención óptima} después de un número $T$ de rondas fijado por el agente y permite hacerlo de manera más rápida que los algoritmos del estado del arte.\\
+\\
+\indent Posteriormente, el trabajo de \cite{sen2017identifying} extiende el trabajo de \cite{lattimoreNIPS2016} al caso en el cual el modelo causal es parcialmente conocido, pero sigue abierto el problema para el caso en el cual el agente deba aprender el modelo conforme interactúa con su entorno.\\
+\\
+\indent Además, otra ventaja que otorga el tener un modelo causal de un ambiente es que este conocimiento puede ser \textit{transferible} a problemas futuros que también se lleven a cabo en el mismo ambiente.
+\\
+\\
+\indent En resumen, lo que se propone es utilizar la maqunaría de Teoría de Juegos para modelar la interacción entre un tomador de decisiones y un entorno incierto el cual es regido por un modelo causal.
+A través de las interacciones el tomador de decisiones aprenderá del modelo además de usarlo para conseguir su objetivo.\\
+\\
+
+\subsection{Justificación(aquí van como las aplicaciones)}
+Además, en el área de la Economía conocida como Economía Conductual (Behavioral Economics) es un problema interesante el dotar agentes que tienen un comportamiento incoherente en el tiempo (\cite{kleinberg2014time}) con incentivos, lo cual se puede hacer si se conocen una estructura de dependencia entre acciones y recompensas (\cite{albers2016motivating}).
+\\
+
+\indent De la misma manera en la que a los seres humanos nos ayuda contar con información causal, \cite{lattimoreNIPS2016} muestran que un agente, un tomador de decisiones, que cuenta con un modelo causal puede encontrar de manera más rápida la acción óptima en un contexto incierto que si no utilizaran información causal del entorno.
+En su trabajo muestran, empírica y teóricamente, que contar con un modelo causal lleva a una mejor toma de decisiones que si sólo se realizaran búsquedas asociativas.
+Lo que este trabajo señala es que a la luz de un modelo causal se pueden tomar decisiones que maximicen cierta variable de pago de manera más rápida que los métodos que no incorporan relaciones causales.
+La limitación que tiene este trabajo es suponer conocido de antemano el modelo causal, pero para modelar un escenario más realista es necesario suponer no conocido el modelo causal e intentar aprender sobre éste a partir de las interacciones mismas con el ambiente; de hecho es, en parte, como operamos los seres humanos (\cite{hagmayer2013repeated}).
+La importancia del razonamiento causal para el desarrollo de agentes inteligentes está detallada y justificada en \cite{lake2017building}.\\
+\\
+\indent Modificar una a una las relaciones causales que se expresan en un modelo causal como se hace en \cite{wellen2012learning} tiene sentido pues existe evidencia de que los seres humanos se enfocan en aprender aspectos \textit{locales} de las estructuras causal de nuestro entorno (\cite{danks2014unifying}) y posteriormente en unificar estos aspectos en una sola estructura coherente (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}).
+Además, tomar en cuenta la restricción de sólo tener una decisión por ronda es realista, pues problemas de decisión secuencial pueden modelarse como sucesiones de decisiones de una sola acción, además de que siempre es posible considerar varias acciones simultáneas como una sola que las englobe.
+Por otro lado, el conocimiento inicial del tomador de decisiones es un punto delicado, pues a partir de qué tan completo o no sea, así de buenas serán sus acciones iniciales.
+\cite{billot2005probabilities} muestran cómo asignar probabilidades iniciales dado un conjunto de observaciones.\\
+\\
+\indent Encontrar la acción óptima a llevar a cabo en un contexto incierto y de manera tal que el tomador de decisiones pueda aprender algo del ambiente se ha abordado desde el punto de vista de Proceso de Decisión Markovianos y Programación Dinámica, lo que dio lugar al área conocida como Aprendizaje por Refuerzo (RL por sus siglas en inglés).
+En este contexto, el agente que debe aprender a llevar a cabo una tarea se encuentra ante un entorno incierto, y sobre el cual debe aprender mediante interacción.
+El agente lleva a cabo una acción y observa una señal de recompensa del ambiente.
+La señal de recompensa debe ser tal que, al intentar el agente maximizarla en el largo plazo, el agente cumpla su objetivo.\\
+\\
+\indent Diseñar una señal de recompensa es un problema difícil en general (\cite{sutton1998reinforcement},\cite{dewey2014reinforcement}, \cite{DRLnotwork}), pues es la manera de especificarle al agente qué debe hacer, pero no cómo.
+En este trabajo, lo que se propone es modelar la interacción agente-entorno como un \textit{juego} de modo que el agente podrá observar una \textit{jugada} de la Naturaleza; es decir, toda una acción y no sólo una cuantificación de lo devuelto por el entorno.
+Además, al ser la señal de recompensa algo que debe diseñarse de acuerdo al problema, pudiera ocurrir que la señal de recompensa sea construída de modo que oculte las relaciones causales al agente, por eso se opta por observar lo que ocurre afuera como jugadas y que sea el agente quien asigne una utilidad a los resultados finales del juego.
+Considerar a la naturaleza como un jugador no es una idea nueva, pues desde \cite{milnor1951games}, \cite{papadimitriou1985games}, \cite{Szep1985} y \cite{eiselt2004games} se ha considerado a la Naturaleza como un jugador que \textit{asigna} inicialmente un estado al ambiente.\\
+\\
+\indent Que un agente pueda \textit{aprender} algo de su entorno a lo largo de una serie de interacciones, o juegos, ha sido estudiado ya en el caso en el que el jugador busca aprender los pagos del juego o las estrategias de alguno de los otros agentes \cite{fudenberg1998learning} lo que motiva a pensar que un agente podría aprender sobre el modelo causal que rige su entorno.
+\\
+\\
+\indent Desde el punto de vista de la epistemología y la fiolosofía de las ciencias han existido diversos intentos de formalizar una Teoría de Decisión que incorpore información causal, como han sido los trabajos de \cite{joyce1999foundations}, \cite{board2006equivalence}, \cite{joyce2012regret}, \cite{ahmed2012push}, \cite{rottman2014reasoning}, \cite{soares2015toward}, \cite{stalnaker2016knowledge}, los cuales hacen uso del \textit{razonamiento contrafactual} para proponer una manera de tomar decisiones a la luz de “lo que hubiera pasado si...”.
+Estos trabajos tienen ciertas limitaciones filosóficas y conceptuales, pues ignoran el caso en el cual las acciones de un tomador de decisiones están correlacionadas (pero no causalmente conectadas) con algunas partes del ambiente, además de la dificultad que implica la construcción de un modelo causal.
+
 \section{Trabajo Previo}

### 2018-06-06 Propuesta/Proposal_English.tex
inicios de la propuesta en ingles

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
%\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{natbib}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\theoremstyle{plain}
\newtheorem{teo}{Theorem}
\newtheorem{prop}[teo]{Proposition}
\newtheorem{defi}[teo]{Definition}
\newtheorem{obs}[teo]{Observation}
\newtheorem{lem}[teo]{Lemma}
\newtheorem{cor}[teo]{Corolary}
\usepackage[pdftex]{color,graphicx}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}
%\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
\title{Use and acquisition of causal relations for decision making under uncertainty using imperfect information repeated games}
\author{Mauricio Gonzalez Soto}
\begin{document}
%\nocite{*}
\maketitle
\tableofcontents
\newpage
\begin{abstract}
We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker. We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge. We propose a decision making procedure in which an agent holds \textit{beliefs} about her environment which are used to make a choice and then are updated using the observed outcome. As proof of concept, we present an implementation of this causal decision making model and apply it to a simple problem. We show that the model achieves a performance similar to the classic Q-learning while it also acquires a causal model of the environment. 
\end{abstract}
\section{Introduction}
\indent A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}, \cite{pearlwhy}). In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the classic formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}). On the other hand, if enough information is not available, the decision maker could attempt to \textit{learn} from the environment by interacting with it.

Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process \cite{garnelo2016towards}.

Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect \cite{rottman2014reasoning}, we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.

The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment and a way to update them after an outcome has been observed.

While the standard setting in RL is to model the agent-environment interaction as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model in response to what the decision maker has chosen.

The agent, using her current beliefs, will generate a \textit{local} causal model and choose an action from it as if that model was the true one. Then, after she observes the consequences of her actions, her beliefs will be updated according to the observed information in order to make a better choice the next time. The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true model.

Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed.

\section{Main Concepts}
	\subsection{Causal Relations}
	Causality is understood as a relation between \textit{events}. Some event \textit{causes} another event to occurr. The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
	\begin{defi}{\label{causal_relation}}
	Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite probability space, and consider a binary relation $\to \subseteq \mathcal{F} \times \mathcal{F}$ which is:
	\begin{itemize}
	\item Transitivity: If $A \to B$ and $B \to C$ for any $A, B, C \in \mathcal{F}$ then $A \to C$.
	\item Irreflexive: For all $A \in \mathcal{F}$ it doesn't hold that $A \to A$.
	\item Antisymmetric: For $A,B \in \mathcal{F}$ such that $A \neq B$ if $A \to B$ then it doesn't hold that $B \to A$.
	\end{itemize}
	We will say that $A$ is \textit{a cause} of $B$ (or that $A$ causes $B$, $A$ is the cause and $B$ is the effect) if $A \to B$.
	\end{defi}
	We will assume that for any event $A \in \mathcal{F}$ there are no causes lying outside $\mathcal{F}$, and that the relations expressed by $\to$ are the only causal relations in the environment. This assumption can be interpreted as not allowing \textit{intermediate} events between events known to be causally related. For example, if in our space we have that striking a match causes fire, we will not allow into consideration the underlying chemical reactions that cause fire from friction. In this sense, we will say that striking a match is a \textit{direct cause} of fire and this will be the only causes considered (\cite{spirtes2000causation}). 
	\subsection{Representation into a Directed Acyclic Graph}
	Making some notation abuse, the causal relations contained in $\to$ can be summarized in a graph $G=(V,E)$ in the following way: If $A \to B$ then the graph must contain a node $A \in V$ representing $A$, a node  $B \in V$ representing $B$ and a directed edge $e \in E$ connecting the respective nodes in the direction of the causal relation.
	\begin{prop}
	Given a causal relation $\to$ as in Definition \ref{causal_relation} then the graph that is obtained by considering nodes for events and edges for the causal relations as previously described is a Directed Acyclic Graph.
	\end{prop}
	\begin{proof}
	\end{proof}
	\subsection{Causal Graphical Models}
	\subsection{Decision Theory}
	A Decision Problem under Uncertainty is a situation in which an agent must choose one out of many actions with uncertain consequences. We follow \cite{bernardo2000bayesian}.
	\subsubsection{Elements of a Decision Problem}
	A Decision Problem under Uncertainty is composed by:
	\begin{itemize}
	\item A set $\Omega$ 
	\item A set $\mathcal{A}=\{a_i : i \in I \}$ of actions.
	\item For each action $a_i$, a partition $E_i = \{ E_j : j \in J \}$ of $\Omega$. Define $\mathcal{E} = \cup_{i \in I } E_i$.
	\item For each action $a_i$, a set of consequences $C_i = \{ c_j : j \in J \}$. Define $\mathcal{C}=\cup_{i \in I} C_i$
	\item A preference relation $\succeq$ defined over $\mathcal{A}$.
	\end{itemize}
	Since we are considering that a decision maker chooses an action $a \in \mathcal{A}$, and then an uncertain event $E$ will occur which will then produce the consequence $c$, we will identify actions in $\mathcal{A}$ as $a_i = \{ c_j | E_j : j \in J \}$. As a technical assumption required for all of the math to work, we require that elements of $\mathcal{C}$ belong to $\mathcal{A}$. This is achieved by adding into $A$ elements of the form $\{ c | \Omega \}$ for $c \in \mathcal{C}$. This being said, we can extend the preference relation $\succeq$ to elements of $\mathcal{C}$.

	From the prerefence relation $\succeq$ we can derive some other relations:
	\begin{itemize}
	\item $a_1 \sim a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and } a_2 \succeq a_1$
	\item $a_1 \succ a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and it does not hold that } a_2 \succeq a_1$
	\end{itemize}
	\subsubsection{Axioms of Coherence}
	We require the following axioms for the preference relation $\succeq$ which we will call Axioms of Coherence or Rationality, and define a Rational decision maker as a decision maker whose preferences are rational in the sense of the axioms (\cite{bernardo2000bayesian}).\\
	\\
	\textbf{Axiom 1}.
	\begin{itemize}
	\item i) There exists consequences $c_1$, $c_2$ such that $c_1 < c_2$.
	\end{itemize}
	\subsection{Game Theory}

\section{Problem Statement}
	\subsection{Problem Statement}
	\subsection{Motivation}
	Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}). Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent.

	In many real-world applications, the agent doesn't know all of the parameters required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then he could attempt to discover this relations and use them in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning (\cite{sutton1998reinforcement}).

	It is known that human beings conceive their actions on the world as \textit{intervening} in the world. Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model. The agent must choose the intervention that maximizes the value of a \textit{target variable} after a series of learning rounds. They model their problem as if choosing an intervention was choosing an arm of a \textit{slot machine}, in which a gambler chooses an arm and gets some reward. From the rewardsm they estimate probabilities and output an optimal action in the sense of obtaining minimal regret.

	 Their work considers that the causal model is fully known. They mention that the case where the causal model is unknown is left as an open question, and it is precisely what we are proposing to answer. 
	\subsection{Justification}
	Besides obtaining a criteria of choosing actions, our agent will also learn a causal model of his environment. Learning causal models is importante since it allows further questions to be asked. 
	\subsection{Research Question}
	\subsection{Hypothesis}
	Let $\mathcal{G}$ a causal graphical model whose probability distribution $P_\mathcal{G}$ satisfies (the axioms). Let $(D,E,c)$ a decision problem whose uncertain events are related to the variables in $\mathcal{G}$. Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$,  by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less rounds than if he doesn't consider causal information.
	\subsection{Objectives}
		\subsubsection{General Objectives}
		\subsubsection{Specific Objectives}
		\subsubsection{Limitations}

\section{Related Work}

\section{Methodology}
\subsection{Modelling}
\subsection{Working plan}

\newpage
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
\end{document}

### 2018-06-06 Propuesta/Proposal_English.tex
axiomas de coherencia y probabilidad y utilidad

@@
 \section{Main Concepts}
-	\subsection{Causal Relations}
+	\subsection{Attempts to define Causality}
+	Causality has been a difficult concept to define and several attempts have been made.
+In fact, in mainstream statistics  the term \textit{cause} have been nearly banned as documented by \cite{pearl2018why}.
+
+	In the classic works on Causality (\cite{spirtes2000causation}, \cite{pearl2009causality}) the term \textit{causality} is defined in terms of common language and then the mathematical machinery required to model causality is provided.
+In this way, causality becomes defined as what is modeled by causal models, which are themselves defined as to model causality.
+This apparent circularity is deeply studied by \cite{woodward2005making}.
+For this reason, we will give here a formal working definition of causal relations which allows to be used formally.
+	\subsection{Definition of Causality}
 	Causality is understood as a relation between \textit{events}.
@@
 	\begin{proof}
+	The graph is directed since the definition of causality imposes a direction between events; namely, a direction between the cause and the effect.
+To see that the graph is acyclic, suppose that a cycle $A \to B \to C \to A$ exists, this would imply, because of transitivity, that $A \to A$, which can not be since the relation is irreflexive.
 	\end{proof}
+	Notice that since the graph is finite, there exist some nodes that does not have causes, which are called \textit{exogenous}.
+If an $A$ event is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
 	\subsection{Causal Graphical Models}
@@
 This is achieved by adding into $A$ elements of the form $\{ c | \Omega \}$ for $c \in \mathcal{C}$.
-This being said, we can extend the preference relation $\succeq$ to elements of $\mathcal{C}$.
-
-	From the prerefence relation $\succeq$ we can derive some other relations:
+This being said, we can extend the preference relation $\succeq$ to elements of $\mathcal{C}$ and therefore indistinctively write $c_1 \succeq c_2$ for $c_1, c_2 \in \mathcal{C}$ or $a_1 \succeq a_2$ for $a_1,a_2 \in \mathcal{A}$.
+We require that $\mathcal{E}$ is an \textit{algebra}.
+	\begin{defi}
+	From the prerefence relation $\succeq$ we can derive some other relations between actions:
 	\begin{itemize}
 	\item $a_1 \sim a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and } a_2 \succeq a_1$
-	\item $a_1 \succ a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and it does not hold that } a_2 \succeq a_1$
+	\item $a_1 \succ a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and it does not hold that } a_2 \succeq a_1$.
 	\end{itemize}
+	\end{defi}
+	\begin{defi}
+	A relation between events can be derived from $\succeq$ in the following way:
+	\[E \succeq F \Leftrightarrow \textrm{ for all } c_1 \succeq c_2 \textrm{ it holds that } \{ c_2 | E, c_1 | E^c \} \succeq \{ c_2 | F, c_1 | F^c \}. \]
+	In this case, we say that $F$ is not \textit{more likely} than $E$.
+It can easily be shown that $\Omega \succ \emptyset$, and a relation $\sim$ for events is defined in an analog way than for consequences.
+	\end{defi}
+	\begin{defi}
+	Given some $G \succ \emptyset$ we define a conditional preference relation as follows:
+	\begin{itemize}
+	\item i) $a_1 \succeq_G a_2 \Leftrightarrow \textrm{ for all a } \{ a_1 | G, a | G^c \} \succeq \{ a_2 | G, a | G^c \}$.
+	\item ii) $E \succeq_G F \Leftrightarrow \textrm{ for } c_1 \succeq_G c_2 \{ c_2 | E , c_1 | E^c \} \succeq_G \{c_2 | F , c_1  | F^c \}$ 
+	\end{itemize}
+	\end{defi}
+	\begin{defi}
+	Two events $E,F \in \mathcal{E}$ are said to be independent if and only if:
+	\begin{itemize}
+	\item 	
+	\end{itemize}
+	\end{defi}
 	\subsubsection{Axioms of Coherence}
-	We require the following axioms for the preference relation $\succeq$ which we will call Axioms of Coherence or Rationality, and define a Rational decision maker as a decision maker whose preferences are rational in the sense of the axioms (\cite{bernardo2000bayesian}).\\
+	We require the following axioms for the preference relation $\succeq$ which we will call Axioms of Coherence or Rationality, and define a Rational decision maker as a decision maker who will choose according to a preference relation and whose preferences are rational in the sense of the axioms (\cite{bernardo2000bayesian}).\\
 	\\
-	\textbf{Axiom 1}.
+	\indent \textbf{Axiom 1}.
 	\begin{itemize}
-	\item i) There exists consequences $c_1$, $c_2$ such that $c_1 < c_2$.
+	\item i) There exists consequences $c_1$, $c_2$ such that $c_1 \succ c_2$.
+	\item ii) For all consequences $c_1, c_2$ and events $E,F \in \mathcal{E}$ either $\{  c_2 | E , c_1 | E^c\} \succeq \{ c_2 | F, c_1 | F^c \}$ or $\{ c_2 | F, c_1 | F^c \} \succeq \{  c_2 | E , c_1 | E^c\}$
 	\end{itemize}
+
+	\textbf{Axiom 2}.
+	\begin{itemize}
+	\item i) $a \succeq a$ for all $a \in \mathcal{A}$.
+	\item ii) If $a \succeq b$ and $b \succeq c$ for $a,b,c \in \mathcal{A}$, then $a \succeq c$.
+	\end{itemize}
+
+	\textbf{Axiom 3}
+	\begin{itemize}
+	\item i) If $a_1 \succeq a_2$ then for all $G \succ \emptyset$ $a_1 \succeq_G a_2$.
+	\end{itemize}
+
+	\textbf{Axiom 4.}\\
+	There exists a sub-algebra $\mathcal{S}$ of $\mathcal{E}$ and a function $\mu : \mathcal{S} \to [0,1]$ such that:
+	\begin{itemize}
+	\item i) $S_1 \succeq S_2$ if and only if $\mu(S_1) \succeq \mu(S_2)$.
+	\item ii) If $S_1 \cap S_2 = \empty$ then $\mu(S_1 \cup S_2) = \mu(S_1) + \mu(S_2)$.
+	\item iii) For any number $\alpha \in [0,1]$ and independent events $E,F$, there is an event $S$, which is called a standard event, such that $\mu(S)=\alpha$ and $S$ is independent from $E$ and from $F$.
+	\item iv) If $S_1$ is independent from $S_2$ then $\mu(S_1 \cap S_2)=\mu(S_1)\mu(S_2)$.
+	\item v) If $E$ independent of $S$, $S$ independent from $F$ and $E$ independent from $F$ then $E \sim S$ implies that $E \sim_F S$.
+	\end{itemize}
+
+	\textbf{Axiom 5.}
+	\begin{itemize}
+	\item i) If $c_1 \succeq c \succeq c_2$ then there exists a standard event $S$ such that $c \sim \{ c_2 | S, c_1 | S^c \}$.
+	\item ii) For each event $E$ there exists a standard event $S$ such that $E \sim S$.
+	\end{itemize}
+	\subsubsection{Subjective Probability and Utility}
+	\begin{defi}
+	Given a preference relation $\succeq$, we define the (subjective) probability of an event $E \in \mathcal{E}$ as the real number $\mu(S)$ associated with the standard event $S$ such that $E \sim S$.
+	\end{defi}
+	\begin{defi}
+	Given a preference relatin $\succeq$ we define the utility $u(c)$ of a consequence $c \in \mathcal{C}$ relative to the extreme consequences $c_\ast$, $c^\ast$, as the real number $\mu(S)$ associated with any standard event $S$ such that $c \sim \{c^\ast | S, c_\ast | S^c\}$.
+	\end{defi}
+	\subsubsection{Decision Criteria for Decision Problems}
+	\begin{teo}
+	For any bounded decision problem with extreme consequences $c^\ast \succ c_\ast$ and any $G \succ \emptyset$
+	\[ a_1 \succeq_G a_2 \Leftrightarrow \bar{u}(a_1) \geq \bar{u}(a_2). \]
+	\end{teo}
+	\begin{proof}
+	See \cite{bernardo2000bayesian} Chapter 2 Proposition 2.22
+	\end{proof}
 	\subsection{Game Theory}

### 2018-06-11 Propuesta/Proposal_English.tex
acabando teo decision

@@
 	\subsection{Decision Theory}
-	A Decision Problem under Uncertainty is a situation in which an agent must choose one out of many actions with uncertain consequences. We follow \cite{bernardo2000bayesian}.
+	A Decision Problem under Uncertainty is a situation in which an agent must choose one out of many actions with uncertain consequences.
+The most common theories for Decision Making under Uncertainty are those from \cite{von1944theory} and \cite{savage1954the}.
+In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences, and in that case the theory guarantees that the decision maker behaves as if he maximizes the expected value of an utility function.
+On the other hand, if the decision maker doesn't know the probabilities of observing outcomes given a chosen action, then Savage's theory guarantees that the decision maker behaves as if he had in mind a \textit{subjective} probability distribution, a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and utility function.
+
+Both von Neumann and Savage's Decision Under Uncertainty theories are known as the \textit{classic} theories in other areas, mostly in Economics.
+It is important to say that other Decision Making theories exist, such as Prospect Theory (\cite{kahneman1979prospect}), Case-Based Decision Theory (\cite{gilboa1995case}) among others that are out of the scope of this work.
+
+For what remains of this section, we follow \cite{bernardo2000bayesian} in his exposition of Savage's theory since it is the most suitable for our purposes because we will be considering decision makers that do not know all of their environment.
+
 	\subsubsection{Elements of a Decision Problem}
@@
 	\begin{itemize}
-	\item A set $\Omega$ 
+	\item A set $\Omega$ of \textit{states}.
 	\item A set $\mathcal{A}=\{a_i : i \in I \}$ of actions.
@@
 	\item For each action $a_i$, a set of consequences $C_i = \{ c_j : j \in J \}$. Define $\mathcal{C}=\cup_{i \in I} C_i$
-	\item A preference relation $\succeq$ defined over $\mathcal{A}$.
+	\item A preference relation $\succeq$ defined over $\mathcal{A}$ which represent the decision-maker whishes.
+It is assumed that the decision maker will choose among his options according to his preferences, which are represented by $\succeq$.
 	\end{itemize}
@@
 	\end{itemize}
+
+	For interpretations of the axioms in diverse contexts we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}.
 	\subsubsection{Subjective Probability and Utility}
 	\begin{defi}
-	Given a preference relation $\succeq$, we define the (subjective) probability of an event $E \in \mathcal{E}$ as the real number $\mu(S)$ associated with the standard event $S$ such that $E \sim S$.
+	Given a preference relation $\succeq$, we define the (subjective) \textbf{probability} of an event $E \in \mathcal{E}$ as the real number $\mu(S)$ associated with the standard event $S$ such that $E \sim S$.
+	\end{defi}
+	The subjective probability thus defined satisfies all of the Kolmogorov's axioms of probability (\cite{bernardo2000bayesian}).
+So, we can use all of the mathematical machinery for classic probability measures, including the definition of conditional probability, Bayes' theorem, etc.
+	\begin{defi}
+	Consequences $c^\ast$ and $c_\ast$ are called respectively best and worst (extreme consequences) if for any other consequence $c \in \mathcal{C}$ then $c^\ast \succeq c \succeq c_\ast$.
+Decision problems in which we add extreme consequences are called bounded decision problems.
 	\end{defi}
 	\begin{defi}
-	Given a preference relatin $\succeq$ we define the utility $u(c)$ of a consequence $c \in \mathcal{C}$ relative to the extreme consequences $c_\ast$, $c^\ast$, as the real number $\mu(S)$ associated with any standard event $S$ such that $c \sim \{c^\ast | S, c_\ast | S^c\}$.
+	Given a preference relation $\succeq$ we define the canonical \textbf{utility} $u(c)=u(c | c_\ast, c^\ast)$ of a consequence $c \in \mathcal{C}$ relative to the extreme consequences $c_\ast$, $c^\ast$, as the real number $\mu(S)$ associated with any standard event $S$ such that $c \sim \{c^\ast | S, c_\ast | S^c\}$.
 	\end{defi}
 	\subsubsection{Decision Criteria for Decision Problems}
+	\begin{prop}
+	For any bounded decision problem with extreme consequences $c_\ast$ and $c^\ast$ it holds:
+	\begin{itemize}
+	\item i) For all $c \in \mathcal{C}$, $u(c | c_\ast, c^\ast)$ exists and is unique.
+	\item ii) The value of $u(c | c_\ast, c^\ast)$ is unafected by the occurrence of any event $G \succ \emptyset$.
+	\item iii) $0 = u(c_\ast | c_\ast, c^\ast) \leq u(c | c_\ast, c^\ast) \leq u(c^\ast | c_\ast, c^\ast) = 1$.
+	\end{itemize}
+	\end{prop}
+	\begin{defi}
+	For a bounded decision problem and any event $G \succ \emptyset$ and $a = \{ c_j | E_j : j \in J \}$ we define the conditional expected utility of $a$ as
+	\[ \bar{u}(a | c_\ast, c^\ast, G) = \sum_{j \in J} u(c_j | c_\ast, c^\ast)P(E_j | G). \]
+	If $G = \Omega$ we simply denote $\bar{u}(a | c_\ast, c^\ast, G)$ as $\bar{u}(a | c_\ast, c^\ast)$
+	\end{defi}
 	\begin{teo}
@@
 	\end{proof}
+	This result means that for any decision maker whose preferences satisfies axioms 1 through 5 the only election criteria that is compatible with the axioms is the maximization of expected utility.
+Of course, this result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum (\cite{bernardo2000bayesian}).
 	\subsection{Game Theory}
@@
 \bibliographystyle{apalike}
-\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
+\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
 \end{document}

### 2018-06-12 Propuesta/Proposal_English.tex
teo de juegos: normal form games

@@
 	\subsection{Causal Graphical Models}
+	A causal graphical model (CGM) consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, an acyclic directed graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect.
+Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows:
 	\subsection{Decision Theory}
@@
 For what remains of this section, we follow \cite{bernardo2000bayesian} in his exposition of Savage's theory since it is the most suitable for our purposes because we will be considering decision makers that do not know all of their environment.
+Savage's Decision Making theory, together with von Neumann's theory is what it is commonly known as \textit{classical} decision making.
 
@@
 	\[E \succeq F \Leftrightarrow \textrm{ for all } c_1 \succeq c_2 \textrm{ it holds that } \{ c_2 | E, c_1 | E^c \} \succeq \{ c_2 | F, c_1 | F^c \}. \]
-	In this case, we say that $F$ is not \textit{more likely} than $E$.
+	In this case, we say that $F$ is \textbf{not more likely} than $E$.
 It can easily be shown that $\Omega \succ \emptyset$, and a relation $\sim$ for events is defined in an analog way than for consequences.
@@
 	\begin{defi}
-	Given some $G \succ \emptyset$ we define a conditional preference relation as follows:
+	Given some $G \succ \emptyset$ we define a conditional preference relation $\succeq_G$ as follows:
 	\begin{itemize}
@@
 	\subsubsection{Axioms of Coherence}
-	We require the following axioms for the preference relation $\succeq$ which we will call Axioms of Coherence or Rationality, and define a Rational decision maker as a decision maker who will choose according to a preference relation and whose preferences are rational in the sense of the axioms (\cite{bernardo2000bayesian}).\\
+	We require the following axioms for the preference relation $\succeq$ which we will call Axioms of Coherence or Rationality, and define a Rational decision maker as a decision maker who will choose according to a preference relation and whose preferences satisfy the axioms (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).\\
 	\\
@@
 	\begin{proof}
-	See \cite{bernardo2000bayesian} Chapter 2 Proposition 2.22
+	See \cite{bernardo2000bayesian} Chapter 2 “Foundations”, Proposition 2.22.
 	\end{proof}
-	This result means that for any decision maker whose preferences satisfies axioms 1 through 5 the only election criteria that is compatible with the axioms is the maximization of expected utility.
-Of course, this result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum (\cite{bernardo2000bayesian}).
+	This result means that for any decision maker whose preferences satisfies axioms 1 through 5 the only election criteria that is compatible with the axioms is the maximization of expected utility and thus establishes a \textit{normative} criteria for decision making for rational agents.
+Of course, this result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum and further mathematical assumptions are required over the utility function $u$ in order to guarantee the existence of the maxmimum utility option (\cite{bernardo2000bayesian}).
+
+	The result extends previous work of \cite{von1944theory} who assumed that the decision maker knows the probabilties of the uncertain events and guarantees the existence of an utility function such that the decision maker chooses by calculating expected value of this utility function.
 	\subsection{Game Theory}
+	A \textit{game} arises when two or more \textit{rational} decision makers, or players, have to make decisions in a situation in which the outcome for each player is partly determined by the choices of other players (\cite{binmore2008rational}).
+
+	Game Theory is an area of Mathematics which is used to model games.
+The basic assumptions of the theory is that each decision maker is rational, and they take into account their knowledge or expectations of other decision makers behavior.
+
+	In this section we will review the basic models of Game Theory that are required to this thesis proposal, which are the normal-form game, and the extensive-form game.
+
+	The normal-form game models a situation in which two or more agents interact and where each one of them will choose an action simultaneously.
+Also, it is assumed that all of the relevant aspects of the game are known for each player.
+
+	\begin{defi}
+	A normal-form game is a tuple $(N,A,(\succeq_i)_{i \in N})$ where:
+	\begin{itemize}
+	\item $N$ is a finite set of players.
+	\item $A=A_1 \times ... \times A_n$ where each $A_i$ is the set of available actions for player $i$. Each $a=(a_1,...,a_n) \in A$ is called an action profile.
+	\item For each player $i \in N$, a preference relation $\succeq_i$ defined over $A$.
+	\end{itemize}
+	\end{defi}
+
+Because of the decision-making criteria we know that each player can replace his preferences $\succeq_i$ for a utility function $u_i$.
+
+The prisoner's dilemma is the most common example of a normal-form game, where two suspects must either confess or remain silent about a crime and they can't communicate with each other.
+In this game, both of the players know the consequences of each \textit{outcome} and the rewards (utilities) associated with each outcome.
+
+In some cases, the players do not know all of the relevant aspects of a game, such as the payoffs or the available actions to other players.
+It is this situation that is modeled by normal-form Bayesian games (\cite{osborne1994course}, \cite{shoham2008multiagent}).
+
+\begin{defi}
+A normal-form Bayesian game consists of:
+\begin{itemize}
+\item A set of states $\mathcal{S}$.
+\item A finite set of players $N$.
+\end{itemize}
+\end{defi}
 
@@
 	\subsection{Problem Statement}
+	Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
+And consider a rational decision maker who doesn't know the parameters of the causal model which control the probabilities of observing a consequence given an action $a \in \mathcal{A}$, we ask how the decision maker could learn about the causal structure that controls his environment in order to make good choices with respect to the decision makers' preferences.
 	\subsection{Motivation}
@@
 
-	It is known that human beings conceive their actions on the world as \textit{intervening} in the world.
+	It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
 Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
@@
 They model their problem as if choosing an intervention was choosing an arm of a \textit{slot machine}, in which a gambler chooses an arm and gets some reward.
-From the rewardsm they estimate probabilities and output an optimal action in the sense of obtaining minimal regret.
-
+From the rewards they estimate probabilities and output an optimal action in the sense of obtaining minimal regret.
 Their work considers that the causal model is fully known.
@@
 	\subsection{Justification}
-	Besides obtaining a criteria of choosing actions, our agent will also learn a causal model of his environment. Learning causal models is importante since it allows further questions to be asked. 
+	Learning a causal model of some phenomena is important, since it allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+
 	\subsection{Research Question}
+	How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
+How can causal structure help a decision maker in order to guide his learning process?
 	\subsection{Hypothesis}
-	Let $\mathcal{G}$ a causal graphical model whose probability distribution $P_\mathcal{G}$ satisfies (the axioms).
-Let $(D,E,c)$ a decision problem whose uncertain events are related to the variables in $\mathcal{G}$.
-Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$,  by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less rounds than if he doesn't consider causal information.
+	Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
+Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
 	\subsection{Objectives}
@@
 \subsection{Working plan}
+\subsubsection{Publication plan}
+\begin{itemize}
+\item Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
+Accepted for a poster presentation at the CausalML 2018 workshop at ICML 2018.
+\end{itemize}
 
+\section{Preliminary Results}
+To show the factibility of this proposal, we considered a test scenario and two cases in ascending difficulty
 \newpage

### 2018-06-13 Propuesta/Proposal_English.tex
añadiendo figuras

@@
 %\usepackage[spanish, mexico]{babel}
+\usepackage[font=small,labelfont=bf]{caption}
 \usepackage{listings}
@@
 Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed.
+A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions.
 
+\newpage
 \section{Main Concepts}
@@
 	\end{defi}
+
 	\begin{teo}
@@
 
+	In the particular case where the algegra $\Omega$ contains only the set $\Omega$ we say that it is a decision problem without uncertainty and in this case applying the previous Theorem we obtain the decision criteria which consist in maximizing the utility.
+
 	The result extends previous work of \cite{von1944theory} who assumed that the decision maker knows the probabilties of the uncertain events and guarantees the existence of an utility function such that the decision maker chooses by calculating expected value of this utility function.
@@
 	In this section we will review the basic models of Game Theory that are required to this thesis proposal, which are the normal-form game, and the extensive-form game. 
-
+	\subsubsection{Building blocks}
 	The normal-form game models a situation in which two or more agents interact and where each one of them will choose an action simultaneously.
 Also, it is assumed that all of the relevant aspects of the game are known for each player.
+
+	We follow the exposition from \cite{osborne1994course}, \cite{binmore2007playing}, \cite{shoham2008multiagent}
 
@@
 
-Because of the decision-making criteria we know that each player can replace his preferences $\succeq_i$ for a utility function $u_i$.
+Because of the decision-making criteria shown in the last section we know that each player can replace his preferences $\succeq_i$ for a utility function $u_i$.
 
@@
 \item A finite set of players $N$.
+\item For each player: a set of actions $A_i$ and as in the previous definition we define $A= A_1 \times \cdots \times A_n$.
+\item For each player: a finite set $T_i$ of \textit{signals} that are observable to player $i$.
+\item For each player: a signal function $\tau_i : \mathcal{S} \to T_i$.
+\item For each player: a probability measure $P_i$ defined over $\mathcal{S}$ such that $P(\tau^{-1}_i (t_i)) >0$ for $t_i \in T_i$.
+\item For each player: a rational preference relation $\succeq_i$ defined over the set of probability measures defined over $A \times \mathcal{S}$.
 \end{itemize}
@@
 
+\subsubsection{Repeated Games}
+\subsubsection{Learning in Games}
+
+\newpage
 \section{Problem Statement}
@@
 	\subsection{Justification}
-	Learning a causal model of some phenomena is important, since it allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+	Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data.
+For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
+This is highly relevant in real-world applications such as self-driving cars, where it is very important to understand why an accident happened.
+Methods based in Deep Neural Networks aren't supposed to explain why a certain action was chosen since those methods are based in \textit{parallel distributed representations}.
+
+	Learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+It allows to ask \textit{why}.
 
@@
 
+\newpage
 \section{Related Work}
 
+\newpage
 \section{Methodology}
-\subsection{Modelling}
+\subsection{Modelling a Decision Problem under Uncertainty as a Game}
+\subsubsection{Knowing only the graph structure}
+\subsubsection{General case}
+\subsection{Experimental methodology}
 \subsection{Working plan}
@@
 
+\newpage
 \section{Preliminary Results}
-To show the factibility of this proposal, we considered a test scenario and two cases in ascending difficulty
+To show the factibility of this proposal, we considered a test scenario and two cases in ascending difficulty.
+\subsection{Test scenario}
+Consider a sick patient who arrives at a hospital and he either has disease $A$ or disease $B$.
+The doctor can either give him some pill or send him into surgery.
+Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
+The doctor could be facing a mutation from a known disease, so she has some knowledge about what could happen if a treatment is given to the patient.
+Using her previous knowledge as a true model, she can choose a treatment and observe the outcome from which she will learn about this disease, so she could make a better decision the next time a similar patient arrives.
+
+The causal model that governs this situation is shown in Figure \ref{causal_model}.
+The parameters for this model were fixed intuitively in such a way that each treatment is effective for only one disease, but the most effective treatment is riskier.
+
+The variables in the model are: 
+\begin{itemize}
+\item \textbf{Disease:} Either $A$ or $B$.
+\item \textbf{Treatment:} Either pill or surgery.
+\item \textbf{Reaction:} Either dying or surviving.
+\item \textbf{Lives:} Either living or dying.
+\end{itemize}
+
+The variables are causally related as shown in Figure \ref{causal_model}.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+The variable \textit{Lives} is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the variable \textit{Treatment}.
+The decision maker prefers an outcome in which the patient lives.
+
+In this scenario, Nature's move will consist in randomly assigning a disease to the patient.
+Then, the medic will asign a treatment using his current beliefs about the disease and the possible outcomes.
+The decision nodes for this play of the medic form an \textit{information set} because the medic doesn't know how she arrived there since she doesn't know what disease did Nature assign.
+Finally, Nature will sample the consequence of the treatment from the causal model and the medic will observe the outcome.
+
+For this test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
+\[ P(Y | do(Tr))=P(Y | D, Tr, R)P(R | Tr) P(D). \]
+\\
+In fact, from the structure of the model, which is shown in Figure \ref{causal_model}, we see that the involved probabilities in any calculations are:
+\[ P(\textrm{disease}), P(\textrm{treatment}), P(\textrm{reaction} | \textrm{treatment}), \]
+and
+\[P(\textrm{lives} | \textrm{disease, treatment, reaction}). \]
+\\
+We can also see that the joint distribution for all of the variables can be expressed as
+\[ P(Y | D, Tr, R)P(R | Tr) P(D)P(Tr). \]
+This expression will be useful when specifying beliefs about the model as Dirichlet distributions.
+
+\subsection{Case 1: The causal model is completely known}
+If the causal model is completely known to the decision maker, then in one step she can obtain the probability for her desired value of the target variable, which in this case is the value corresponding to the outcome in which the patient lives at the end.
+Using this probability, she can choose which treatment to assign.
+Since this action maximizes the probability of the occurence desired value, it maximizes the expected utility, and it is also a \textit{best response} to the player Nature.
+
+\subsection{Case 2: Only the structure is known}
+From the expression of the joint probability we notice that we need to specify a distribution over each one of these distributions, which will be each one a Dirichlet distribution.
+
+We begin with a random assignation of the $\alpha$ parameter for each of the distributions considered.
+We use Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
+Since each of the variables in the model is binary, then the product of these Dirichlets is Dirichlet.
+
+With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable as in Case 1.
+With this action chosen, we simulate an outcome from the causal graphical model using the chosen action as an intervention.
+This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
+
+We show the results of two experiments.
+We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
+We show the average perform over $20, 50, 100$ and $200$ rounds.
+
+\begin{center}
+\begin{minipage}{0.48\linewidth}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}
+\captionof{figure}{Average reward in 20 rounds}
+\end{minipage}%
+\hfill
+\begin{minipage}{0.49\linewidth}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}
+\captionof{figure}{Average reward in 50 rounds}
+\end{minipage}
+\end{center}
+\begin{center}
+\begin{minipage}{0.49\linewidth}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}
+\captionof{figure}{Average reward in 100 rounds}
+\end{minipage}
+\hfill
+\begin{minipage}{0.49\linewidth}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}
+\captionof{figure}{Average reward in 200 rounds}
+\end{minipage}
+\end{center}
+
+%In Figure \ref{20_rounds} we observe the average rewards for each agent in 20 rounds of decision making.
+Here we notice that Q-learning outperforms our algorithm, which has a similar performance as the random choosin procedure until round 11.
+%
+%\begin{figure}[ht]
+%\vskip 0.2in
+%\begin{center}
+%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}}
+%\caption{Average reward obtained in each round for each agent}
+%\label{20_rounds}
+%\end{center}
+%\vskip -0.2in
+%\end{figure}
+%
+%In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making. Our algorithm follow closely the Q-learning agent and outperform the random agent.
+%
+%\begin{figure}[ht]
+%\vskip 0.2in
+%\begin{center}
+%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}}
+%\caption{Average reward obtained in each round for each agent}
+%\label{50_rounds}
+%\end{center}
+%\vskip -0.2in
+%\end{figure}
+%
+%In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds, where our algorithm slightly outperforms Q-Learning.
+%
+%\begin{figure}[ht]
+%\vskip 0.2in
+%\begin{center}
+%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}}
+%\caption{Average reward obtained in each round for each agent}
+%\label{100_rounds}
+%\end{center}
+%\vskip -0.2in
+%\end{figure}
+%
+%In Figure \ref{200_rounds} we observe the average reward obtained by the three agents in 200 rounds. The average reward obtained is very similar for Q-learning and our algorithm.
+%
+%\begin{figure}[ht]
+%\vskip 0.2in
+%\begin{center}
+%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}}
+%\caption{Average reward obtained in each round for each agent}
+%\label{200_rounds}
+%\end{center}
+%\vskip -0.2in
+%\end{figure}
+
+We see that our method obtains a very similar reward as the classic Q-learning algorithm for a larger number of rounds, where the random agent is outperformed, but our model offers something extra because it learns a causal model of the environment.
+
 \newpage

### 2018-06-14 Propuesta/Proposal_English.tex
actualizando

@@
 \begin{abstract}
-We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker.
-We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
+We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker, although he is aware of the causal nature of his environment.
+We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had he not considered causal knowledge.
 We propose a decision making procedure in which an agent holds \textit{beliefs} about her environment which are used to make a choice and then are updated using the observed outcome.
@@
 
-	In the classic works on Causality (\cite{spirtes2000causation}, \cite{pearl2009causality}) the term \textit{causality} is defined in terms of common language and then the mathematical machinery required to model causality is provided.
+	In the reference works on Causality (\cite{spirtes2000causation}, \cite{pearl2009causality}) the term \textit{causality} is defined in terms of common language and then the mathematical machinery required to model causality is provided.
 In this way, causality becomes defined as what is modeled by causal models, which are themselves defined as to model causality.
@@
 For this reason, we will give here a formal working definition of causal relations which allows to be used formally.
+
+	Several definitions of Causality exist in diverse contexts with diverse applications in mind, such as the Topological Causality for Dynamical Systems (\cite{harnack2017topological}), Lamport's Causality (\cite{lamport1978time}), Granger's Causality for Time Series (\cite{granger1969investigating}), Suppes' Causality (\cite{suppes1970probabilistic}) each with its own mathematical framework.
+A review of several of this theories, including Suppes' and Granger's can be found in \cite{holland1985statistics}.
 	\subsection{Definition of Causality}
-	Causality is understood as a relation between \textit{events}.
+	Causality is understood as a \textit{stochastic} relation between \textit{events}.
 Some event \textit{causes} another event to occurr.
-The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
+For example, The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
 	\begin{defi}{\label{causal_relation}}
@@
 
-	For interpretations of the axioms in diverse contexts we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}.
+	For interpretations and critiques of the axioms we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}, \cite{peterson2017introduction}.
 	\subsubsection{Subjective Probability and Utility}
@@
 
-	In the particular case where the algegra $\Omega$ contains only the set $\Omega$ we say that it is a decision problem without uncertainty and in this case applying the previous Theorem we obtain the decision criteria which consist in maximizing the utility.
+	In the particular case where the algebra $\mathcal{E}$ contains only the set $\Omega$ as its only element we say that it is a decision problem without uncertainty and in this case applying the previous Theorem we obtain the decision criteria which consist in maximizing the utility function.
 
@@
 
-	We follow the exposition from \cite{osborne1994course}, \cite{binmore2007playing}, \cite{shoham2008multiagent}
+	We follow the exposition from \cite{osborne1994course}, \cite{binmore2007playing}, \cite{shoham2008multiagent}.
 
@@
 This is highly relevant in real-world applications such as self-driving cars, where it is very important to understand why an accident happened.
-Methods based in Deep Neural Networks aren't supposed to explain why a certain action was chosen since those methods are based in \textit{parallel distributed representations}.
+
+	For example, as told by \cite{bornstein2016artificial}, at the University of Pittsburgh Medical Center a team of researchers tried to use Machine Learning to predict whether pneumonia patients might develop severe complications.
+For this purpose they trained Neural Networks and Decision Trees using the hospital's own data.
+Neural Networks outperformed Decision Trees, but only by studying the decisions made by the latter did the doctors find out that the algorithms instructed doctors to send home pneumonia patients who already had asthma, despite the fact that asthma sufferers are known to be extremely vulnerable to complications.
+The problem relied in the training data, because the hospital policy was to automatically send asthma sufferers with pneumonia to intensive care, and this policy worked so well that asthma sufferers almost never developed severe complications.
+It was only through the interpretability of the Decision Trees that the doctors didn't send asthma patients with pneumonia home to a certain death.
+
+	Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo (\cite{silver2017mastering}) defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, it was just curve fitting.
 
@@
 \subsection{Modelling a Decision Problem under Uncertainty as a Game}
+\subsubsection{Fully knowing the causal graphical model}
 \subsubsection{Knowing only the graph structure}
-\subsubsection{General case}
+\subsubsection{General case: Model unknown}
+\subsection{Validation of the proposed methods}
 \subsection{Experimental methodology}

### 2018-06-14 Propuesta/Proposal_English.tex
do calculus

@@
 If an $A$ event is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
+	\subsubsection{Relation between the graph and probabilities}
+	Given a Directed Acyclic Graph it is possible to obtain a probability measure that expresses the conditional independence relations that are expressed in the graph (\cite{koller2009probabilistic}), which we will call $P_\mathcal{G}$.
+For the DAG built from the causal relations, we require that its correspondent $P_\mathcal{G}$ satisfies the following conditions (\cite{spirtes2000causation}):
+	\begin{itemize}
+	\item Markov Causality: an event $V$ (or node in $\mathcal{G}$) is independent of every other event $A$ such that $A$ isn't either a cause nor an effect of $V$ given the causes of $V$.
+	\item Causal Minimality: No proper sub-graph of $\mathcal{G}$ satisfies the Markov Causality condition.
+	\item Causal Faithfulness: The Markov Causal condition contains all of the conditional independence statements expressed by the DAG $\mathcal{G}$.
+	\end{itemize}
+	This conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
 	\subsection{Causal Graphical Models}
 	A causal graphical model (CGM) consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, an acyclic directed graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect.
-Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows:
+Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
+, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delet every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
+
+	It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph (\cite{sucar2015probabilistic}).
+	\subsubsection{The identifiability problem}
+	Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
+It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to Markov equivalence (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}).
+	\subsubsection{Do-Calculus}
+	The Do-Calculus (\cite{pearl1995causal}, \cite{pearl2009causality}) is a set of rules for manipulating probabilistic statements that involve interventions.
+
+	Some notation is important to be mentioned: Consider a causal graphical model $\mathcal{G}$ and $X,Y,Z$ disjoint sets of nodes of $\mathcal{G}$.
+We denote by $\mathcal{G}_{\bar{X}}$ the graph that is obtained by deleting from $\mathcal{G}$ all of the edges that enter nodes in $X$.
+In the same way, $\mathcal{G}_{\underline{X}}$ is the graph obtained by deleting the edges that emerge from $X$.
+Finally, $\mathcal{G}_{\underline{Z}\bar{X}}$ is the graph obtained by deleting edges incoming into $X$ and outgoing from $Z$.
+
+	\begin{teo}{\label{docalculus}} (\cite{pearl2009causality})\\
+Let $\mathcal{G}$ a CGM y $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
+\begin{itemize}
+\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ y $W$, then
+\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
+\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ y $W$, then
+\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
+\item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.
+If $Y$ is conditionally independent from $Z$ given $X$ y $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
+\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
+\end{itemize}
+\end{teo}
+
 	\subsection{Decision Theory}

### 2018-06-20 Propuesta/Proposal_English.tex
metdologia

@@
 Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process \cite{garnelo2016towards}.
+
+A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
+A causal structure encondes a series of \textit{cause-effect} relations between events.
 
@@
 	\subsubsection{Do-Calculus}
-	The Do-Calculus (\cite{pearl1995causal}, \cite{pearl2009causality}) is a set of rules for manipulating probabilistic statements that involve interventions.
+	The Do-Calculus (\cite{pearl1995causal}, \cite{pearl2009causality}) is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data.
 
@@
 \end{teo}
+\begin{teo}{\cite{peters2017elements}}\\
+The following statements hold:
+\begin{itemize}
+\item The Do-Calculus is complete; this is, sufficient for deriving every identifiable interventional distribution (\cite{huang2006pearl}, \cite{shpitser2006identification}).
+\item There exists an algorithm capable of finding all of the identifiable interventions (\cite{tian2002}, \cite{huang2006pearl}).
+\item A necessary and sufficient criteria exists for the identifiability of interventional distributions (\cite{shpitser2006identification}, \cite{huang2006pearl}).
+\end{itemize}
+\end{teo}
+Las reglas del cálculo do proveen una solución al problema de la identificabilidad:
+\begin{cor} (\cite{pearl2009causality}).\\
+A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each one of them corresponds to any of the rules of the Do-Calculus, that reduce $q$ to a purely observational expression.
+\end{cor}
 
@@
 So, we can use all of the mathematical machinery for classic probability measures, including the definition of conditional probability, Bayes' theorem, etc.
+
+	As a technical requirement we need to consider two extra options, one that is preferred over every other consequence, which will be denoted as $c^\ast$ and other consequence $c_\ast$ which  the agent prefers every other consequence before $c_\ast$.
+This consequences need not have an interpretation, but they can be thought of as \textit{heaven} and \textit{hell} respectively for obvious (at lest in the Western hemisphery) reasons.
+Critiques of this extra technical consequences can be found in \cite{binmore2008rational}, \cite{peterson2017introduction}.
 	\begin{defi}
@@
 	\begin{defi}
-	Given a preference relation $\succeq$ we define the canonical \textbf{utility} $u(c)=u(c | c_\ast, c^\ast)$ of a consequence $c \in \mathcal{C}$ relative to the extreme consequences $c_\ast$, $c^\ast$, as the real number $\mu(S)$ associated with any standard event $S$ such that $c \sim \{c^\ast | S, c_\ast | S^c\}$.
+	Given a preference relation $\succeq$ in a bounded decision problem we define the canonical \textbf{utility} $u(c)=u(c | c_\ast, c^\ast)$ of a consequence $c \in \mathcal{C}$ relative to the extreme consequences $c_\ast$, $c^\ast$, as the real number $\mu(S)$ associated with any standard event $S$ such that $c \sim \{c^\ast | S, c_\ast | S^c\}$.
 	\end{defi}
@@
 	In this section we will review the basic models of Game Theory that are required to this thesis proposal, which are the normal-form game, and the extensive-form game. 
-	\subsubsection{Building blocks}
+	\subsubsection{Building blocks}{\label{basic_models}}
 	The normal-form game models a situation in which two or more agents interact and where each one of them will choose an action simultaneously.
@@
 
+The interpretation of the elements of this model is as follows: the set $\mathcal{S}$ contains the descriptions of the relevant characteristics for all players, each player has \textit{a priori} beliefs over this characteristics, which are stated by the probability measure $P_i$.
+When a play is to be realized, the world is in some $\omega \in \Omega$ and each player observes his $\tau_i (\omega)$.
+If a player observes the signal $t_i \in T_i$ then he concludes that the real state of the world is in the set $\tau^{-1}_i (t_i)$.
+Player $i$ updates his beliefs over $\omega \in \Omega$ to $P_i(\omega) / P_i(\tau^{-1}_i (t_i))$ if $\omega \in \Omega$ and cero otherwise.
+Bayesian games are also called \textit{incomplete information} games in the literature which is different from the \textit{imperfect information} game which will be soon defined.
+
+The normal-form game models situations where players select their actions simultaneously, or when it does not matter who makes an action first.
+More realistic and complex situations involve some notion of temporal structure, or order among the players' actions.
+The extensive-form game attempts to include this notion of order.
+The standard references for both of this models are \cite{osborne1994course}, \cite{shoham2008multiagent}, \cite{10.1007/978-94-010-0189-2_25}.
+\begin{defi}
+A perfect information extensive game consists of:
+\begin{itemize}
+\item A finite set $N$ of players.
+\item A set $A$ of available action.
+\item A set $H$ of sequences called histories that satisfies:
+	\begin{itemize}
+		\item The epty sequence belongs to $H$.
+		\item If $(a_k)_{k=1}^K \in H$ ($K$ can be infinite) and $L < K$ then $(a_k)_{k=1}^L \in H$.
+		\item If $(a_k)_{k=1}^\infty$ satisfies $(a_k)_{k=1}^L \in H$ for every positive integer $L$, then $(a_k)_{k=1}^\infty \in H$
+	\end{itemize}
+\item If a history $(a_k)_{k=1}^K \in H$ with $K$ finite and if it doesn't exist a $K+1$ such that$(a_k)_{k=1}^{K+1} \in H$ then $(a_k)_{k=1}^K$ is said to be a terminal history.
+The set of terminal histories is called $Z$
+\item A function $\chi : H \to 2^A$ that assigns a set  $\chi(h)$ of possible actions to be taken after a history $h$ has occurred.
+\item A function $\rho$ that assigns to every non terminal history $h$ a player $\rho(h)$ who will choose an action after history $h$ has occurred.
+\item A function $\sigma : H \times A \to H \cup Z$ that maps history-action pairs into a new  history such that  if $\sigma(h_1 , a_1) = \sigma (h_2,a_2)$ then $h_1 = h_2$ y $a_1 = a_2$
+\item For each player $i \in N$ a preference relation $\succeq_i$.
+\end{itemize}
+\end{defi}
+
+In this model, each player knows how did he get up to certain point (knows what history occured).
+To allow more generality, we consider cases where each player doesn't know how he got up to certain point; i.e., he does not know which history happen and he can't distinguish in which node is he.
 \subsubsection{Repeated Games}
@@
 
+	The impact of Causal Inference goes beyond Machine Learning and Computer Science.
+For example, economists are interested in understanding what did certan public policy caused (\cite{athey2017impact}) or how human decision makers that show \textit{inconsistent} preferences over time could be oriented if the causal consequences of this inconsistent behavior is introduced into a decision-making model (\cite{albers2016motivating}).
+In complex adaptive systems, causal relations can be used to clarify the complex interactions between agents in the system (\cite{abbott2017complex}) as well as for prediction and planning (\cite{hunt2016ants}, \cite{brock2018causality}).
+
 	\subsection{Research Question}
@@
 		\subsubsection{General Objectives}
+			The proposed research has as a general objective to find how a rational autonomous agent who faces an uncertain environment can discover and use causal information in order to make good choices that point towards a concise objective such as the maximization of some function.
 		\subsubsection{Specific Objectives}
+		In order to achieve the General Objective, we need to separate it into smaller problems which are required to be solved first.
+
 		\subsubsection{Limitations}
+		The proposed research has some limitations.
+We are considering a \textit{rational} decision maker whose preferences can be represented as if maximizing an expected utility with respect to a subjective probability distribution and an utility that assigns numbers to possible outcomes.
+Maximization of expected utility is not the only decision making criteria as other theories exist, such as Prospect Theory (\cite{kahneman1979prospect}) where a decision maker considers a reference point and from there he considers losses more painfull than gains.
+Other theory is Gilboa's Case-Based Decision Theory, where a decision maker considers how similar is the current decision problem with problems faced in the past.
+
+		We consider a particular definition of Causality, which is Spirtes' definition (\cite{spirtes2000causation}) who defines causality as a stochastic relation between events.
+We take this definition of causality as well as its theoretical requirements as axioms and we do not question if is the definition most suitable for a certain situation.
+
+		We are not considering that the player called Nature has any intentions nor objectives since we are given her constant payments in every outcome.
+A scenario where Nature could have objectives to pursue if it is modelling a player such as a Government, who can only act according to current laws but has general social objectives, such as the maximization of people with jobs in a negotiation with Unions.
+
+		\subsubsection{Extensions}
+		Given the Limitations addressed in the previous section, some possible extensions could be:
+		\begin{itemize}
+		\item Consider longer games.
+		\item Consider more players than an agent and Nature.
+		\item Consider objectives/intentions for Nature.
+		\item Studying game-theoretical properties of the model such as equilibria in terms of the causal structure.
+		\end{itemize}
 
@@
 \section{Related Work}
+Since we are considering decision problems where an agent must choose some available action and then receive a reward from it and after a set of stages the agent is expected to know a good action.
+It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}) if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
+
+On the other hand, if the largest mean reward is desired with a fixed confidence of $\delta$ then an algorithm for finding the optimal action exist in such a way that the number of rounds to find it is within a constant factor of a lower bound (\cite{jamieson2014lil}).
+
+If one does not consider a fixed confidence but instead a fixed number of learning rounds (known as fixed budget), then we face a \textit{fixed budget} problem where a lower bound for the probability of finding the optimal arm is given by \cite{carpentier2016tight}.
+
+Notice that none of the previous works consider any kind of causal structure.
+For this kind of problems, causal relations between actions and consequences could be considered, in this way an action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world (\cite{hagmayer2009decision}).
+Consider a set of variables that are causally related between them and to a \textit{reward} variable whose value has to be maximized by intervening one of the variables.
+It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
+Their work requires that the causal model is fully known to the decision maker, this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part.
+
+Discovering the causal model itself while using current knowledge to make choices is left as future work in both \cite{lattimoreNIPS2016} and \cite{sen2017identifying}, but algorithms to discover causal relations in data can be found in \cite{eberhardt2008almost}, \cite{hauser2012two}, \cite{hyttinen2013experiment}, \cite{loh2014high}, \cite{shanmugam2015learning} \cite{mooij2016distinguishing}.
+Also, an \textit{active learning} approach can be used to learn causal models from data, where one starts with some initial graph and then select the instances in the data that allow to add and orient edges in order to end with a fully oriented graph.
+Active learning algorithms for causal discovery can be found in \cite{tong2001active},\cite{murphy2001active}, \cite{meganck2006learning}, \cite{he2008active}, \cite{hauser2012two}, \cite{10.1007/978-3-319-56970-3_9}, \cite{rubenstein2017probabilistic}.
+
+Notice that the aforementioned papers consider learning while interacting with some (causal) environment.
+We propose to model this interaction as a game, which has been previously considered in \cite{werling2015job}, who model an \textit{on-line} classification problem as a game where an oracle is used to classify initial observations.
 
@@
 \section{Methodology}
+In this section we describe the steps that are to be followed in order to answer the Hypothesis.
 \subsection{Modelling a Decision Problem under Uncertainty as a Game}
+Consider a decision problem under uncertainty $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ where the available actions to a decision maker are causally related to outcomes and this causal relations are captured by a Causal Graphical Model $\mathcal{G}$.
+The Causal model is assumed to remain fixed and unknown to the decision maker, who is assumed to be rational and aware of the causal nature of his problem.
+
+As stated in the Hypothesis, we propose that the agent learns from the environment by interacting with it.
+By interaction we mean a number of rounds where the decision maker will act and then observe the outcome of his action.
+
+Since the observed outcomes, given an action, are guided by the Causal Model, we can think of the environment as a decision maker whose available actions depend on what has been chosen previously by the original decision maker.
+We have mentioned that Game Theory allows to model situations where two or more decision makers interact.
+In this way, we propose to model the interaction of the (original) decision maker and his environment as a game between two players: the decision maker, and a player which will be called Nature, whose actions are to be guided by the causal model.
+Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
+
+After the decision maker makes his play, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
+This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games described in Section \ref{basic_models}.
+
+The extensive-form game considers two possibilities: perfect or imperfect information, where in the latter a player may not know what actions have been played in the past by other players, which is what we need to use because the decision maker does not know what move has Nature selected.
+\subsubsection{Elements of the Game}
+The game which will be used to model the interaction between a decision maker and his environment is composed by:
+\begin{itemize}
+\item \textbf{Players:} The decision maker and a player called Nature, who will move first in order to assign a state.
+\item \textbf{Actions:} The actions available to the decision maker are the same ones as in the original decision problem. Nature must choose according to the causal model.
+\item \textbf{Preferences:} The preference relation to the decision maker is his same rational preference relation as in the decision problem. Nature is indifferente between outcomes.
+\end{itemize}
+
+\subsection{Belief Formation}
+Once a Game that models agent-environment interaction is defined, we must turn to the question about \textit{learning} about the environment, which includes acquiring causal information and also using current causal knowledge in order to obtain new information.
+
+For the agent to learn and reason about the plays of Nature we assume that he is able to observe the final outcome of the game, which includes knowing about the inicial uncertain state that Nature assigned.
+The agent, at the beginning of the game is uncertain about the state of the environment which he is facing, but he still must take an action, so in order to allow the decision maker to use his knowledge to act, he will have \textit{beliefs} about any relevant aspects of the environment which he will use as if they were the true model in any given play.
+These beliefs will be \textit{updated} according to what is observed at the end of each game.
+
+Beliefs represent ignorance about relevant characteristics of certain environment (\cite{bernardo2000bayesian}, \cite{peterson2017introduction}), and thus are encoded as probability distributions, which usually come from a parametric family.
+
+In our proposed research, beliefs will encode current knowledge about the \textit{causal structure} of the environment, and this beliefs will be updated when information about actions-outcomes becomes available, which will happen at the end of each game.
+
+Beliefs must also encode any previous knowledge possesed by the decision maker, and this initial assignment must be coherent with the actions prescribed with the rationality axioms for preferences.
+In \cite{billot2005probabilities} is shown how to build a set of beliefs given a set of past observations in an axiomatic way.
+
+\subsection{Belief updating}
+
+\subsection{Solving first simpler cases}
+In order to make a step towards understanding how to acquire and use causal information in a decision problem under uncertainty we will consider three cases in order of ascending difficulty.
+In all three cases we consider a decision maker who faces an uncertain environment and who must take an action in order to satisfy his preferences.
+The uncertain environment in which the decision maker (or agent) exist is governed by a causal graphical model $\mathcal{G}$ which relates actions taken by the decision maker with outcomes.
+We are assuming that one of the variables of the casusal model is of interest for the decision maker in the sense that he seeks to maximize its value.
+This variable will be known as \textit{target variable} or reward variable.
 \subsubsection{Fully knowing the causal graphical model}
@@
 \item Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
-Accepted for a poster presentation at the CausalML 2018 workshop at ICML 2018.
+Accepted for a poster presentation at the CausalML 2018 workshop at ICML / IJCAI / AAMAS 2018.
 \end{itemize}

### 2018-06-21 Propuesta/Proposal_English.tex
metodologia

@@
 
-Notice that the aforementioned papers consider learning while interacting with some (causal) environment.
-We propose to model this interaction as a game, which has been previously considered in \cite{werling2015job}, who model an \textit{on-line} classification problem as a game where an oracle is used to classify initial observations.
+Notice that these papers consider learning while interacting with a causal environment.
+We propose to model this interaction as a game, which has been previously considered in \cite{werling2015job}, who consideres an \textit{on-line} classification problem and modeled it as a game and where an oracle is used to classify initial observations.
 
@@
 \subsection{Modelling a Decision Problem under Uncertainty as a Game}
-Consider a decision problem under uncertainty $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ where the available actions to a decision maker are causally related to outcomes and this causal relations are captured by a Causal Graphical Model $\mathcal{G}$.
-The Causal model is assumed to remain fixed and unknown to the decision maker, who is assumed to be rational and aware of the causal nature of his problem.
+Consider a decision problem under uncertainty $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ where the available actions $\mathcal{A}$ to a decision maker are causally related to the outcomes $\mathcal{C}$ and this causal relations are captured by a Causal Graphical Model $\mathcal{G}$.
+The Causal model is assumed to remain fixed and unknown to the decision maker, who is assumed to be rational and aware of the causal nature of the decision problem he faces.
 
@@
 We have mentioned that Game Theory allows to model situations where two or more decision makers interact.
-In this way, we propose to model the interaction of the (original) decision maker and his environment as a game between two players: the decision maker, and a player which will be called Nature, whose actions are to be guided by the causal model.
+In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: the decision maker, and a player which will be called Nature, whose actions are to be guided by the causal model.
 Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
@@
 \subsection{Solving first simpler cases}
-In order to make a step towards understanding how to acquire and use causal information in a decision problem under uncertainty we will consider three cases in order of ascending difficulty.
+Once that the model has been established, we proceed to further examine the problem of acquiring and using causal information in a decision problem and we notice three important cases, in order of ascending difficulty.
 In all three cases we consider a decision maker who faces an uncertain environment and who must take an action in order to satisfy his preferences.
@@
 This variable will be known as \textit{target variable} or reward variable.
+
+The three cases that we notice are:
+\begin{itemize}
+\item The decision maker knows the causal model.
+\item The decision maker knows only the graphical \textit{structure} of the causal model.
+\item The decision maker knows nothing about the causal model. 
+\end{itemize}
+In the following we show how the first two cases, which are particular cases of our general problem, can be solved and we show that the proposed solution for the first two cases sheds light on how the general case could be solved.
 \subsubsection{Fully knowing the causal graphical model}
+If the decision maker knows the causal graphical model, then he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
+The decision thus made is the one that maximizes the expected utility, by construction.
+Also, choosing this action is the \textit{best response} action, and therefore a Nash equilibrium for the game.
 \subsubsection{Knowing only the graph structure}
+
 \subsubsection{General case: Model unknown}
+
 \subsection{Validation of the proposed methods}
+
 \subsection{Experimental methodology}
+\subsubsection{Internal Validation: Reproducibility}
+\subsubsection{External Validation: Different problems}
+
 \subsection{Working plan}
+Concrete steps towards solving the problem:
+\begin{itemize}
+\item Literature review, understanding the problem, considering several models and possible solutions.
+\item Formalize the model. 
+\end{itemize}
 \subsubsection{Publication plan}
@@
 \[ P(Y | D, Tr, R)P(R | Tr) P(D)P(Tr). \]
-This expression will be useful when specifying beliefs about the model as Dirichlet distributions.
+This expression will be useful when specifying beliefs about the model as probability distributions.
 
@@
 \subsection{Case 2: Only the structure is known}
-From the expression of the joint probability we notice that we need to specify a distribution over each one of these distributions, which will be each one a Dirichlet distribution.
+Since the decision maker knows the graph structure, he can explicitly find a non interventional expression for the interventional distribution and update his beliefs about these unknown quantities.
+If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum}, but meanwhile we are assuming that this information is available at the end of each round.
 
+Given the structure of the model; i.e., the variables in it and the directed edges, the joint distribution of those variables can be expressed as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the underlying DAG in $\mathcal{G}$.
+Since these distributions fully characterize the model, the decision maker will have beliefs over each one of these parameters.
+Notice that each of these parameters is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
+
+A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
+The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
+
+\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1},\]
+
+where $(x_1,...,x_k)$ are such that $\sum_{i=1}^k x_i =1$ and $\alpha=(\alpha_1,...,\alpha_k)$. The Dirichlet distribution is useful since it is conjugate for itself \cite{bernardo2000bayesian}.
+
+In this way, the decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
+Using the agent's current beliefs, a causal graphical model can be specified.
+Using this fully specified (structure + parameters) as a true model, the decision maker will make her choice as in Case 1.
+When the decision maker observes the value of the target variable, she will update the parameters that specify her beliefs.
+
+Previously we argued that the agent's beliefs were going to be \textit{distributions} over a suitable space, but what is going to be updated are the parameters of such distributions.
+Namely, the $\alpha$ corresponding to the Dirichlet random variable assigned to each CPT.
+
+For the belief updating, given a new data point,  two cases must be considered:
+\begin{itemize}
+\item The variable to update has no parents.
+\item The variable to update has parents.
+\end{itemize}
+
+In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
+\[ \textrm{Dirichlet}(\alpha + c) \]
+where $c$ is a vector of the number of occurrences of that observed data point.
+
+For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
+Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
+In this case, if the prior of $X_i$ conditioned on its parents having the value $j$ is given by a a Dirichlet($\alpha$), then the posterior for the variable $X_i$ given an observed data point is given by
+\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
+
+\subsubsection{Implementation}
 We begin with a random assignation of the $\alpha$ parameter for each of the distributions considered.
@@
 \end{center}
+
+\subsection{Case 3 and future work: The model is not known}
+The causal model were fully unknown, the decision maker will have to deal with the problem using only any previous knowledge and her own intuitions.
+Again, any previous knowledge and considerations will be expressed as \textit{beliefs} about the uncertainties in the environment, which will take the form of a probability distributions over a suitable space.
+
+As in the previous case, we consider a repeated game where the base game consists of Nature assigning a random state of the environment and responding to the agents choices with the effects that were caused by her decisions.
+In this game, as well as in the previous one, the decision maker will attempt to learn by updating, and using, beliefs in a suitable way.
+
+The most notable difference with the previous case is that the \textit{structure} of the model is also to be learned in such a way that both the structure and parameters converge to the true model in the limit.
+In the previous case the decision maker knew the form of the Conditional Probability Tables (CPT) involved in any calculation.
+In this case, she doesn't know the structure of the DAG so which CPT's are involved is unknown.
+
+If the decision maker knew which variables appear in the true model that governs the environment, even though she didn't know how they are connected, she could use a \textit{Dirichlet Process} to generate Dirichlet distributions and generate causal graphical models the same way as in Case 2 and updating the parameters of the process using the observed information.
+The Dirichlet Process\footnote{with parameters $M,G_0$}, which was introduced by \cite{ferguson1973bayesian}, is random measure defined over a space $S$ such that for each partition $B_1,...,B_k$ the vector $(G(B_1),...,G(B_k))$ follows a Dirichlet distribution \cite{hjort2010bayesian}, \cite{muller2016bayesian}, \cite{ghosal2017fundamentals}.
+
+Belief updating using causal information when the decision maker doesn't know the structure of the model nor its parameters is yet to be studied and left as future work.
 

### 2018-06-23 Propuesta/Proposal_English.tex
metodologia experimental

@@
 A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
-A causal structure encondes a series of \textit{cause-effect} relations between events.
+A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).
 
@@
 	\end{itemize}
+	Also, an extra condition, known as Causal Sufficiency, about the nature of the model itself is considered: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ outside of the model $\mathcal{G}$ (\cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{sucar2015probabilistic}).
 	This conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
@@
 
-	Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo (\cite{silver2017mastering}) defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, it was just curve fitting.
+	Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo (\cite{silver2017mastering}) defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, it was just curve fitting, although a very sofisticated method for achieving it.
+Using only associations between variables it is not possible to infer which one is the cause and which one the effect, something extra is required.
 
-	Learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+	On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
 It allows to ask \textit{why}.
@@
 		In order to achieve the General Objective, we need to separate it into smaller problems which are required to be solved first.
+In first place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.
+
+		In second place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.
+
+		Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker.
+A general updating criteria must be provided which uses the causal nature of the environment.
 
@@
 Notice that none of the previous works consider any kind of causal structure.
-For this kind of problems, causal relations between actions and consequences could be considered, in this way an action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world (\cite{hagmayer2009decision}).
+For this kind of problems, causal relations between actions and consequences could be considered, in this way an action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world (\cite{hagmayer2009decision}) and that human beings use and modify causal knowledge during a sequential decision making process (\cite{hagmayer2013repeated}).
 Consider a set of variables that are causally related between them and to a \textit{reward} variable whose value has to be maximized by intervening one of the variables.
@@
 
+We are considering each game as a learning round, where the only information that is carried from a game to another is what is learned by the decision maker via a process of belief updating.
+The causal model is assumed to remain fixed and each initialization at the beginning of each new game is independent from any past outcome.
+
 The extensive-form game considers two possibilities: perfect or imperfect information, where in the latter a player may not know what actions have been played in the past by other players, which is what we need to use because the decision maker does not know what move has Nature selected.
@@
 \subsection{Belief updating}
+The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
+The observed information will modify his current knowledge, which is encoded as probabilistic beliefs, about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world (\cite{hagmayer2013repeated}).
+
+If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms (\cite{bernardo2000bayesian}).
+
+Bayesian updating is also of great importance in the theory of Learning in Games (\cite{fudenberg1998theory}) since it is known (\cite{shoham2008multiagent}) that if all the players in a game perfom Bayesian updating, then the plays generated by these measures converge to the plays that would be generated if all players shared their knowledge.
+
+Even though we know that Bayesian updating is the correct \textit{framework} for belief updating, it remains unanswered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way.
 
@@
 \subsubsection{Knowing only the graph structure}
+If a decision maker knows the structure of the graph, but he does not know the parameters, then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
+In the preliminary results that will be shown later a particular case is considered, where the beliefs have the form of a Dirichlet distribution for each of the Conditional Probability Tables that characterize a Bayesian Network.
 
+In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e.
+in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
 \subsubsection{General case: Model unknown}
@@
 \subsection{Validation of the proposed methods}
+In order to show that an agent has succeeded in acquiring a causal model from his environment via interacting with it and using it in order to make good choices, 
+\subsection{Experimental methodology}
+The proposed methodology will be tested in the following way: A fully specified causal model, which satisfies the conditiones required above, will be given and used as a ground truth.
+The causal model will not be available to the decision maker, since one of the basic assuptions is that the agent does not know the causal model that controls his environment.
+The true causal model will be used to simulate realizations of the other variables in the distribution given any action taken by the decision maker.
 
-\subsection{Experimental methodology}
+One, or more, of the variables is to be selected as \textit{available} to interventions by the decision maker, the set of possible values for the selected variable determines the set of available actions of the decision maker in the formulation as a game.
+Also, another variable is selected as a \textit{target} variable, preferably an endogenous variable in the model in order to be affected by any action taken by the agent.
+The set of possible values of this variable must be in correspondance with the set of outcomes of the decision problem that the agent faces so the value that the target variable outputs has an effect over the preferences of the agent.
+
+A distribution with initial parameters $\theta_0$ will be specified.
+
+Given a number $T$ of rounds, and inside round $t \in \{1,2,...,T \}$ the agent will make an action $a \in \mathcal{A}$ and then observe the state of the other variables of the model.
+Using this information, current beliefs will be updated according to the criteria to be established.
+
+The chosen actions will be compared to classic on-line decision making 
 \subsubsection{Internal Validation: Reproducibility}
+For the results obtained by following the methodology described in previous sections be reproducible, the causal model used to simulate an environment must be fully specified by describing the probabilistic parameters that control it.
+The initial distributions given to the decision maker to encode his beliefs must be shown as well with the updating criteria used when information from the environment becomes available.
 \subsubsection{External Validation: Different problems}
-
+For the proposed method to be tested in different decision problems, it is required that the correspondent causal model satisfies all of the conditions previously described, and any derived algorithm will be designed in such a way that the characteristics of the causal model are an external input.
+In this way, if an external user is able to give a causal model, the method is expected to work independent of the particularities of the causal model which is being tested.
 \subsection{Working plan}
@@
 Accepted for a poster presentation at the CausalML 2018 workshop at ICML / IJCAI / AAMAS 2018.
+July 15 2018, Stockholm, Sweden.
 \end{itemize}
@@
 
-In this way, the decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
+The decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
 Using the agent's current beliefs, a causal graphical model can be specified.
-Using this fully specified (structure + parameters) as a true model, the decision maker will make her choice as in Case 1.
-When the decision maker observes the value of the target variable, she will update the parameters that specify her beliefs.
+Using this fully specified (structure + parameters) as a true model, the decision maker will make his choice as in Case 1.
+When the decision maker observes the value of the target variable, he will update the parameters that specify his beliefs.
 

### 2018-06-24 Propuesta/Proposal_English.tex
megtodologia

@@
 
-	In the reference works on Causality (\cite{spirtes2000causation}, \cite{pearl2009causality}) the term \textit{causality} is defined in terms of common language and then the mathematical machinery required to model causality is provided.
-In this way, causality becomes defined as what is modeled by causal models, which are themselves defined as to model causality.
-This apparent circularity is deeply studied by \cite{woodward2005making}.
-For this reason, we will give here a formal working definition of causal relations which allows to be used formally.
+Attempts to understand the meaning of \textit{cause} has troubled philosophers and scientists, from David Hume and John Stuart Mill to modern scientists as \cite{suppes1970probabilistic}, \cite{cartwright1983laws}, \cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{spohn2012laws}, \cite{sep-causation-probabilistic}.
+In particular, Suppes and Reichenbach tried to define causation as an increase in the probability of occurrence of one event when another event happened, which is stated as $X$ causes $Y$ if $P(Y|X) > P(Y)$, but this definition was flawed because that increase could happen for several reasons, like $Y$ being a cause of $X$, or both having a common cause $Z$ (\cite{pearl2018why}).
 
@@
 A review of several of this theories, including Suppes' and Granger's can be found in \cite{holland1985statistics}.
+
+In some works on Causality (\cite{spirtes2000causation}, \cite{pearl2009causality}) the term \textit{causality} is defined in terms of common language and then the mathematical machinery required to model causality is provided.
+In this way, causality becomes defined as what is modeled by causal models, which are themselves defined as to model causality.
+This apparent circularity is deeply studied by \cite{woodward2005making}.
+In fact, Judea Pearl never gives a formal definition of causality, rather he prefers to build the machinery required to answer \textit{causal queries} (\cite{pearl2009causality}, \cite{pearl2018theoretical}).
+
+	Even though causality is a difficult concept to grasp, we require a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts. 
 	\subsection{Definition of Causality}
@@
 , x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delet every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
+It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ (\cite{pearl2018why}).
 
@@
 \end{teo}
+\begin{proof}
+See \cite{pearl2009causality}.
+\end{proof}
 \begin{teo}{\cite{peters2017elements}}\\
@@
 \end{teo}
-Las reglas del cálculo do proveen una solución al problema de la identificabilidad:
+\begin{proof}
+See \cite{peters2017elements}.
+\end{proof}
+The Do-Calculus rules provide a solution for the identifiability problem:
 \begin{cor} (\cite{pearl2009causality}).\\
@@
 The most common theories for Decision Making under Uncertainty are those from \cite{von1944theory} and \cite{savage1954the}.
-In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences, and in that case the theory guarantees that the decision maker behaves as if he maximizes the expected value of an utility function.
-On the other hand, if the decision maker doesn't know the probabilities of observing outcomes given a chosen action, then Savage's theory guarantees that the decision maker behaves as if he had in mind a \textit{subjective} probability distribution, a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and utility function.
+In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences \footnote{Also known as decision under risk (\cite{binmore2008rational}, \cite{peterson2017introduction}).}, and in that case the theory guarantees that the decision maker behaves as if he maximizes the expected value of an utility function.
+On the other hand, if the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if he had in mind a \textit{subjective}\footnote{Subjective here is a technical term meaning that the function is internal for the agent.} probability distribution, a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and utility function.
 
@@
 
+One way of visualizing this type of games is using a \textit{tree}, in which each internal nodes are identified with non-terminal histories and each one of these corresponds to a decision made by a player.
+Given a history $h$ in the game, which corresponds to a path in the tree, player $\rho(h)$ will choose on action from the set $\chi(h)$
+
 In this model, each player knows how did he get up to certain point (knows what history occured).
@@
 	Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn causal information and use it to find the optimal actions (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
+Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn causal information and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
 	\subsection{Objectives}
@@
 \section{Related Work}
-Since we are considering decision problems where an agent must choose some available action and then receive a reward from it and after a set of stages the agent is expected to know a good action.
+The use of causal information in human beings has been studied by York Hagmayer and his team and they have found that human beings, while facing a sequential decision problem, use and modify causal information (\cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2010spontaneous}, \cite{hagmayer2013repeated}).
+In their studies, all of the subjects were faced with a one-stage decision problem in which an initial causal model was provided.
+In several rounds, the subject made a decision and observed a feedback and was later asked to describe the causal model they thought controlled the situation.
+The subjects knew that they were supposed to learn a causal model, but the case where the final objective is other than directly learning a causal model has not been studied.
+
+It is also known that human beings focus on \textit{local} aspects while learning causal relations that are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}, \cite{danks2014unifying}).
+Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal structure when little prior information is available.
+
+We are considering decision problems where an agent must choose some available action and then receive a reward from it.
+After a certain number of learning rounds, the agent is expected to output a good action.
+Problems of this nature are called \textit{bandit} problems (\cite{sutton1998reinforcement}).
 It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}) if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
@@
 These beliefs will be \textit{updated} according to what is observed at the end of each game.
+The second step is to find a way to specify beliefs about causal structure.
 
@@
 In our proposed research, beliefs will encode current knowledge about the \textit{causal structure} of the environment, and this beliefs will be updated when information about actions-outcomes becomes available, which will happen at the end of each game.
+The representation of the beliefs must allow them to be used in order to make a choice using the causal information they represent in each round.
 
@@
 Even though we know that Bayesian updating is the correct \textit{framework} for belief updating, it remains unanswered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way.
+Therefore, the third step is to find an updating criteria with theoretical properties that guarantee convergence to the true causal model that controls the environment.
+The updating criteria must explicitly use causal information beyond any associative rules.
 
@@
 \subsubsection{General case: Model unknown}
+The previous two cases are per se steps of the methodology, rather simplifications in order to understand what is happening. The real deal is the case where the causal model is unknown.
+
+When the causal model is unknown, two cases of interest appear, depending if the decision maker knows the (maximum) number of variables in the model or not.
+
+If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
+For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model.
+
+If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables.
+In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.
 

### 2018-06-28 Segundo_Semestre/CausalML_Workshop/Playing_Against_Nature.tex
ultimos cambios

@@
 % Use the following line for the initial blind version submitted for review:
-\usepackage{icml2018}
+%\usepackage{icml2018}
 
 % If accepted, instead use the following line for the camera-ready submission:
-%\usepackage[accepted]{icml2018}
+\usepackage[accepted]{icml2018}
 
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Submission and Formatting Instructions for ICML 2018}
+\icmltitlerunning{Playing against Nature: causal discovery for decision making under uncertainty}
 
@@
 \icmlcorrespondingauthor{Mauricio Gonzalez-Soto}{mauricio@inaoep.mx}
-\icmlcorrespondingauthor{Luis Enrique Sucar}{esucar@inaoep.mx}
-\icmlcorrespondingauthor{Hugo Jair Escalante}{hugojair@inaoep.mx}
+\icmlcorrespondingauthor{Luis E. Sucar}{esucar@inaoep.mx}
+\icmlcorrespondingauthor{Hugo J. Escalante}{hugojair@inaoep.mx}
 
@@
 We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker.
-We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
-We propose a decision making procedure in which an agent holds \textit{beliefs} about her environment which are used to make a choice and then are updated using the observed outcome.
+We ask how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge and propose a decision making procedure in which an agent holds \textit{beliefs} about her environment which are used to make a choice and are updated using the observed outcome.
 As proof of concept, we present an implementation of this causal decision making model and apply it to a simple problem.
@@
 
+A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
+A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).
+
 Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect \cite{rottman2014reasoning}, we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
-
-The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment and a way to update them after an outcome has been observed.
 
@@
 
+The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment and a way to update them after an outcome has been observed.
 The agent, using her current beliefs, will generate a \textit{local} causal model and choose an action from it as if that model was the true one.
@@
 Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed.
+A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions.
+Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods can not answer.
+When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action.
 
@@
 
-Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal},\cite{carpentier2016tight}, \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of these works consider causal-governed environments.
+Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best},  \cite{ortega2014generalized}, \cite{chen2015optimal}, \cite{carpentier2016tight}, \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of these works consider causal-governed environments.
 
@@
 
-The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices, but the problem of \textit{acquiring} this causal knowledge is left unatacked.
+The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices, but the problem of \textit{acquiring} this causal knowledge is left unattacked.
 
 In this work we propose to acquire, by repeated interaction, causal information about the environment as well as using the current causal knowledge inside each round to make better decisions.
-By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a causal entity controls the outcomes.
+By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a \textit{causal entity} controls the outcomes.
 
@@
 
-Consider a Decision Problem under Uncertainty $(A,E,C,\succeq)$ in which an agent has to choose one among several options $a \in A$ which are causally related to the elements of $C$.
-The elements in $E$ are uncertain events which are governed by a causal mechanism.
-This means that when the decision maker choses an action $a$, an uncertain event will occurr in such a way that an outcome $c$ will occur, which is causally related to $a$.
-The decision maker seeks to maximize her utility and we assume that she has \textit{rational preferences}, and because of this we can substitute his preferences $\succeq$ for a utility function $u$ \cite{gilboa2009decision}.
-The decision maker does not know the probabilities nor the structure of the underlying causal model therefore he can not calculate the expected utility of any action.
+Let $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ a Decision Problem under Uncertainty in which an agent has to choose one among several options $a \in \mathcal{A}$ which are causally related to the elements of $\mathcal{C}$.
+The elements in $\mathcal{E}$ are uncertain events which are governed by a causal mechanism.
+This means that when the decision maker chooses an action $a$, an uncertain event will occurr in such a way that an outcome $c$ will occur, which is causally related to $a$.
+The decision maker seeks to maximize her utility and we assume that she has \textit{rational preferences}, so we can substitute her preferences $\succeq$ for the expected value of a utility function $u$ \cite{gilboa2009decision}.
+If the decision maker does not know the probabilities nor the structure of the underlying causal model then she can not calculate the expected utility of any action.
+Instead, she will have a \textit{subjective} probability distribution which represents the agent's knowledge and uncertainties which will be updated by interacting with the environment through succesive rounds of decision making.
 
-Instead, the agent will try to learn by interacting with the environment through succesive rounds of decision making.
 Inside each round, any response from the environment will be independent from the previous rounds, but the actions of the decision maker will be based upon previously acquired causal information and are expected to improve the utility for the agent.
@@
 
-For an agent to reason about and modify his causal knowledge we endow her with a probability distribution $p(\theta)$ over a suitable space.
+For an agent to reason about and modify her causal knowledge we endow her with a probability distribution $p(\theta)$ over a suitable space.
 The beliefs must allow to form a \textit{local} model in a given moment to be used for decision making.
@@
 \subsection{The causal model is completely known}
-Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequence, or effect, of his actions is expressed as the outcome of a random variable $Y$ which we will call \textit{target variable}.
+Consider a decision problem under uncertainty where a decision maker has to choose on out of many elements of a set $A$ and where the consequences, or effects, of her actions are expressed as the outcome of a random variable $Y$ which we will call \textit{target variable}.
 The relation between values of $Y$ and actions $a \in A$ is expressed by a causal graphical model $\mathcal{G}$, which is known by the decision maker.
@@
 
-To model the interaction between the decision maker and the environment, we consider a repeated game in which the base game is described by the narrative of the decision problem: 
+To model the interaction between the decision maker and the environment, we consider a game with the following characteristics:
 \begin{itemize}
@@
 For this reason, the base game is an extensive game with imperfect information since the decision maker makes a choice without knowing the play made by Nature.
-Nature's moves are interpreted as a \textit{response} from the causal model to the actions of the decision maker.
+We choose extensive games since Nature's moves are interpreted as a \textit{response} from the causal model to the actions of the decision maker, so a sequential interpretation had to be considered.
 
-Since the decision maker knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about these unknown quantities.
+Since the decision maker knows the graph structure, she can explicitly find a non interventional expression for the interventional distribution and update her beliefs about these unknown quantities from observed data.
 If the decision maker were not allowed to know, at the end of each round, the play of the Nature then this will have to be estimated as a hidden variable using, for example, the EM algorithm \cite{dempster1977maximum}, but meanwhile we are assuming that this information is available at the end of each round.
@@
 
-\bibliography{/Users/MauricioGS1/INAOE/Segundo_Semestre/Propuesta/Bibliografia.bib}
+\section{Acknowledgments}
+We are grateful to the National Institute of Astrophysics Optics and Electronics (INAOE) and to Mrs. Graciela Soto for their generous funding in order to attend ICML 2018.
+
+\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
 \bibliographystyle{icml2018}

### 2018-06-28 Propuesta/Proposal_English.tex
cambios

@@
 
-Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process \cite{garnelo2016towards}.
+Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting (\cite{sutton1998reinforcement}), but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
 
@@
 
-Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect \cite{rottman2014reasoning}, we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
+Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect (\cite{rottman2014reasoning}), my hypothesis is that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
 
@@
 A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions.
+Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods can not answer.
+When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action.
 
@@
 	\begin{itemize}
-	\item Transitivity: If $A \to B$ and $B \to C$ for any $A, B, C \in \mathcal{F}$ then $A \to C$.
+	\item Transitive: If $A \to B$ and $B \to C$ for any $A, B, C \in \mathcal{F}$ then $A \to C$.
 	\item Irreflexive: For all $A \in \mathcal{F}$ it doesn't hold that $A \to A$.
@@
 For example, if in our space we have that striking a match causes fire, we will not allow into consideration the underlying chemical reactions that cause fire from friction.
-In this sense, we will say that striking a match is a \textit{direct cause} of fire and this will be the only causes considered (\cite{spirtes2000causation}).
+In this sense, we will say that striking a match is a \textit{direct cause} of fire and this will be the only causes considered herein (\cite{spirtes2000causation}).
 	\subsection{Representation into a Directed Acyclic Graph}
-	Making some notation abuse, the causal relations contained in $\to$ can be summarized in a graph $G=(V,E)$ in the following way: If $A \to B$ then the graph must contain a node $A \in V$ representing $A$, a node  $B \in V$ representing $B$ and a directed edge $e \in E$ connecting the respective nodes in the direction of the causal relation.
+	The causal relations contained in $\to$ can be summarized in a graph $G=(V,E)$ in the following way: If $A \to B$ then the graph must contain a node $A \in V$ representing $A$, a node  $B \in V$ representing $B$ and a directed edge $e \in E$ connecting the respective nodes in the direction of the causal relation.
 	\begin{prop}
@@
 Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
-, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delet every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
+, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
 It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ (\cite{pearl2018why}).
@@
 	\begin{teo}{\label{docalculus}} (\cite{pearl2009causality})\\
-Let $\mathcal{G}$ a CGM y $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
+Let $\mathcal{G}$ a CGM and $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
 \begin{itemize}
-\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ y $W$, then
+\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ and $W$, then
 \[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
-\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ y $W$, then
+\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ and $W$, then
 \[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
 \item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.
-If $Y$ is conditionally independent from $Z$ given $X$ y $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
+If $Y$ is conditionally independent from $Z$ given $X$ and $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
 \[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
@@
 This being said, we can extend the preference relation $\succeq$ to elements of $\mathcal{C}$ and therefore indistinctively write $c_1 \succeq c_2$ for $c_1, c_2 \in \mathcal{C}$ or $a_1 \succeq a_2$ for $a_1,a_2 \in \mathcal{A}$.
-We require that $\mathcal{E}$ is an \textit{algebra}.
+We require that $\mathcal{E}$ is an \textit{algebra}\footnote{An algebra of sets over $X$ is a family of sets $\mathcal{F} \subseteq \mathcal{P}(X)$ such that $X \in \mathcal{F}$, $A^c \in \mathcal{F}$ for any $A \in \mathcal{F}$ and for $A,B \in \mathcal{F}$ then $A \cup B \in \mathcal{F}$ (\cite{ash2000probability})}.
 	\begin{defi}
@@
 	\begin{defi}
-	Two events $E,F \in \mathcal{E}$ are said to be independent if and only if:
+	Two events $E,F \in \mathcal{E}$ are said to be independent if and only if for all $c, c_1, c_2 \in \mathcal{C}$ it holds:
 	\begin{itemize}
-	\item 	
+	\item $c \succ \{ c_2 | E, c_1 | E^c \} \Rightarrow c \succ_F \{ c_2 | E, c_1 | E^c \}	$.
+	\item $c \succ \{ c_2 | F, c_1 | F^c \} \Rightarrow c \succ_E \{ c_2 | F, c_1 | F^c \}	$
 	\end{itemize}
@@
 	\subsubsection{Subjective Probability and Utility}
+Subjective probability is a well defined mathematical object that, contrary to the frequentist approach to probability, represents the degree of belief an agent has over the occurrence of some phenomena.
+Frequentist probability can not answer questions like \textit{what is the probability of a war in the middle east?} since it is not a phenomena that can be repeated infinite times in an independent way (\cite{gilboa2009decision}, \cite{peterson2017introduction}).
+
+Subjective approachs to probability were developed by Bruno De Finetti (\cite{definetti1930}) and Frank Ramsey in 1926, however a more precise formulation was given by Savage in his famous book “The Foundations of Statistics” (\cite{savage1954the}).
+The formulation shown here is the modern version of Savage's Theory as exposed by \cite{bernardo2000bayesian}.
+
 	\begin{defi}
@@
 So, we can use all of the mathematical machinery for classic probability measures, including the definition of conditional probability, Bayes' theorem, etc.
+
+Not every belief about uncertainties can be represented by a probabiity measure, a subjective degree of belief for an agent must be rational as will be shown soon.
 
@@
 
+	Notice that the expectations are taken with respect to the subjective probability measure, and in fact Savage's general result as stated in \cite{gilboa2009decision} guarantees the existence of a utility function and a probability measure such that the preference relation is represented by the expected utility of that function with respect to that probability measure.
+Here we defined the subjective probability directly from the coherence axioms.
+
 	In the particular case where the algebra $\mathcal{E}$ contains only the set $\Omega$ as its only element we say that it is a decision problem without uncertainty and in this case applying the previous Theorem we obtain the decision criteria which consist in maximizing the utility function.
@@
 	\end{defi}
+
+A \textit{strategy} or action profile is an element $a \in A$.
 
@@
 \item A function $\sigma : H \times A \to H \cup Z$ that maps history-action pairs into a new  history such that  if $\sigma(h_1 , a_1) = \sigma (h_2,a_2)$ then $h_1 = h_2$ y $a_1 = a_2$
-\item For each player $i \in N$ a preference relation $\succeq_i$.
+\item For each player $i \in N$ a preference relation $\succeq_i$ defined over $Z$.
 \end{itemize}
@@
 To allow more generality, we consider cases where each player doesn't know how he got up to certain point; i.e., he does not know which history happen and he can't distinguish in which node is he.
-\subsubsection{Repeated Games}
-\subsubsection{Learning in Games}
+
+Since this model is about sequential decisions, it is useful for a player to define a plan of action for different scenarios, this is called a \textit{strategy}; i.e., a plan that specifies the action chosen by each player for every history after which it is his turn to act.
+
+\begin{defi}
+\end{defi}
+
+%\subsubsection{Repeated Games}
+%\subsubsection{Learning in Games}
 
@@
 \item Formalize the model.
+\item Studying implications of the assumption of rationality in decisiom making under uncertainty in causal environments.
+\item Propose a representation of beliefs that allow causal structure to be considered.
+\item Updating criteria for beliefs.
+\item Discovery of causal structure
+
 \end{itemize}

### 2018-07-02 Segundo_Semestre/CausalML_Workshop/Poster/Playing_Against_Nature_Poster.tex
esqueleto poster icml

--- 2017-12-20 Primer_Semestre/Poster_Dic17/Poster_eng.tex
@@
 \begin{center}
-\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Comentarios/Figures/logos_2.png}
+\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/logos_2.png}
 \end{center}
@@
 \begin{abstract}
-Reinforcement Learning algorithms have seen a huge development in recent years.
-Standard algorithms only provide action-outcome mappings which are difficult to use in order to make high-level reasoning about the environment.
-The question adressed here is how can an agent learn a casual model in terms of his direct experience with the environment.
-It is proposed to develop a learning-by-interaction algorithm that also learns a causal model which is used on-line to make decisions.
+We consider decision problems under uncertainty where the options available to a decision maker and the resulting outcome are related through a causal mechanism which is unknown to the decision maker, although he is aware of the causal nature of his environment.
+We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had he not considered causal knowledge.
+As proof of concept, we present an implementation of this causal decision making model and apply it in a simple scenario.
+We show that the model achieves a performance similar to the classic Q-learning while it also acquires a causal model of the environment.
 \end{abstract}
+
 %----------------------------------------------------------------------------------------
@@
 \section*{Introduction}
-While playing a videogame or learning how to drive, human beings are forced to \textbf{make decisions} and to \textbf{actively intervene} in the world.\\
-\\
-This active \textbf{decision-making} process modifies the state of the world, which allows us to learn about \textbf{causal relations} that hold in our environment\\
-\\
-We ask ourselves how an \textbf{intelligent agent} can acquire and \textbf{use causal knowledge} while engaging with its environment to achieve some task.\\
-\\
-The goal is to learn a \textbf{causal decision model} and not only an action-outcome mapping.
-%\begin{itemize}
-%\item Videogames: after making some decision and we later will use that knowledge to predict outcomes while facing similar situations  (\cite{meder2008inferring}, \cite{hagmayer2009decision}).
-%\item Driving: \cite{danks2014unifying} gives the example that while learning how to drive, we infer that moving the steering wheel will \textit{cause} a reacion in the car.
-%\end{itemize}
+
 %----------------------------------------------------------------------------------------
@@
 \color{DarkSlateGray} % DarkSlateGray color for the rest of the content
-
 \section*{Related work}
-\subsection*{General Background}
-\begin{itemize}
-\item Causal graphical models (\cite{pearl2009causality}, \cite{koller2009probabilistic}): Causal semantics over probabilistic graphical models.
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_example.png}
-\end{center}
-\item Reinforcement Learning (\cite{sutton1998reinforcement}, \cite{gershman2015reinforcement}): An agent learns through exploration and interaction with environment.
-\begin{center}
-\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/RL.png}\footnote{Image taken from \cite{sutton1998reinforcement}}
-\end{center}
-\end{itemize}
-\subsection*{Where are we standing?}
-\begin{itemize}
-\item Causal decision theory (\cite{joyce1999foundations}): Making decisions using causal knowledge; i.e.
-I know that $A$ \textit{causes} $B$, so it can be predicted what will happen if $A$ is chosen, or not, over other options.
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_Decision.png}\footnote{Image taken from \cite{hagmayer2013repeated}}
-\end{center}
-\item Causal learning through sequential decision making (\cite{hagmayer2013repeated}): Causal learning in humans has been studied in the context of sequental decision-making
-\begin{center}
-\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Hagmayer_rats.png}\footnote{Image taken from \cite{hagmayer2013repeated}}
-\end{center}
-\item On-line learning in causal models(\cite{wellen2012learning}), \cite{garnelo2016towards}):  Causal learning in an on-line setting has been framed in terms of causal graphical models, where in a causal graphical model edges are added or removed in a prediction-error loop
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/LPL_model.png}\footnote{Image taken from \cite{wellen2012learning}}
-\end{center}
-\item Decision problems and causal models: \cite{dawid2002influence} show how causal Bayesian Nets can be augmented with decision nodes as to model decision problems that use causal information.
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Influence_diagram.png}\footnote{Image taken from \cite{dawid2002influence}}
-\end{center}
-\end{itemize}
-%And we can think about these areas relating in the following way:
-%\begin{center}
-%\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/diagrama_eng.png}
-%\end{center}
 
 %----------------------------------------------------------------------------------------
-%	MOTIVACIÓN
+%	 PROBLEM SETUP
 %----------------------------------------------------------------------------------------
 
-\section*{Motivation}
-It has been shown by \cite{garnelo2016towards} that causal structure can be learnt from the interaction of an agent with his environment while it learns to perform some task. \\
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Garnelo.png}\footnote{Image taken from \cite{garnelo2016towards}}
-\end{center}
-Instead of a fully expresive casual graphical model, they use first-order logic which is somehow restrictive while showing the importance of having some semantics capable of managing causal structure.\\
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Graph_vs_logic.png}
-\end{center}
-Meanwhile \cite{wellen2012learning} show how a causal graphical model can be learned on-line by a decision maker through local prediction-error learning.\\ 
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/LPL_example_2.png}
-\end{center}
-From the human psychology side \cite{hagmayer2013repeated} show that our intuitions aren't flawed, because human beings acquire and use causal knowledge while engaging in simple decision-making.
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Causal_Feedback.png}
-\end{center}
-%\\
-%\cite{hagmayer2013repeated} asked what would happen with an agent that must learn some particular task instead of only attempting to directly learn causal structure. \\
-Each decision made by the agent can be interpreted as an intervention in his environment (\cite{hagmayer2009decision}) which provides information about the world's causal structure.\\
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Golf.png}
-\end{center}
-Learning causal structure will allow the agent to make better decisions because it will be able to consider \textbf{consequences} of his actions, and \textbf{not only reacting} to rewards.\\
-\begin{center}
-\includegraphics[scale=0.6]{/Users/MauricioGS1/INAOE/Comentarios/Figures/jail.png}
-\end{center}
-Therefore, we ask ourselves how an agent would learn a causal model in terms of his actions and their consequences? How would this knowledge could be used to make better decisions in the future?
-\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out
 %----------------------------------------------------------------------------------------
-%	PROPUESTA
+%	PROPOSED METHOD
 %----------------------------------------------------------------------------------------
-\section*{Proposal}
-\subsection*{Research question}
-Can \textbf{Reinforcement Learning} methods capture a \textbf{causal model} of the environment rather than just an action-outcome mapping as to make \textbf{better decisions}?
-\subsection*{Objectives}
-\begin{itemize}
-\item Defining how to \textbf{learn} causal relations in an \textbf{on-line} setting in terms of actions taken by an agent and their outcomes.
-\item Defining how to \textbf{incorporate} the current causal knowledge in order to choose \textbf{better actions}.
-\item Defining a way to \textbf{evaluate} how the causal knowledge was used in the decision-making process. We do not want to lose the \textbf{interpretability} of causal graphs.
-\item Defining how to \textbf{extract} the acquired causal knowledge to use it in other domains.
-\end{itemize}
-\subsection*{Proposed Methodology}
-My proposal is about incorporating a \textbf{causal model} inside a sequential decision making problem in which an agent has to maximize some reward by \textbf{selecting actions}.\\
-\begin{center}
-\includegraphics[scale=0.9]{/Users/MauricioGS1/INAOE/Comentarios/Figures/choose_actions.png}
-\end{center}
-The agent will \textbf{interact} with its environment and receive a \textbf{feedback} signal based on the actions taken by him, which will modify his current causal knowledge.\\
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/cry.png}
-\end{center}
-%Using this information, as well as the \textbf{new state} of the world, the internal \textbf{causal model} of the agent will be modified (if needed), either adding or deleting edges or modifying the \textit{causal strength} associated.\\
-%\\
-Before the next step is taken, the current \textbf{causal knowledge} will be taken into account when \textbf{choosing the next action} together with the current value of each posible action.\\ 
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/baby.png}
-\end{center}
-This new action and its outcome will again modify the causal knowledge, which will improve the next action and so on.
-\begin{center}
-\includegraphics[scale=0.7]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Proposed.png}
-\end{center}
-\subsection*{Evaluation}
-How is this better? Why should we care about learning anything extra if the task is achieved?\\
-\\
-Causal structures learnt by the agent can be evaluated if compared with an external \textit{real} causal model, or comparing with the initial \textit{prior} causal graph as made by  \cite{hagmayer2013repeated}.
-\subsection*{Expected contributions}
-\begin{itemize}
-\item A framework in which an agent set to learn some task by action-reward can learn causal structure about its environment.
-\item An algorithm that learns causal structure in an on-line way.
-\item A decision-making procedure that uses current causal knowledge together with action-value estimates.
-\end{itemize}
-\subsection*{Expected time-line}
-\begin{center}
-\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Comentarios/Figures/Gantt.png}
-\end{center}
+\section*{Proposed Methodl}
+
 %------------------------------------------------------------------------------------------
@@
 \bibliographystyle{apalike}
-\bibliography{/Users/MauricioGS1/INAOE/Comentarios/Bibliografia.bib}
+\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
 \end{multicols}
- %----------------------------------------------------------------------------------------
-%	REFERENCES
-%----------------------------------------------------------------------------------------
-%\begin{minipage}[b]{0.55\linewidth}
-%\nocite{*} % Print all references regardless of whether they were cited in the poster or not
-%\bibliographystyle{apalike}
-%\bibliography{/Users/MauricioGS1/INAOE/Comentarios/Bibliografia.bib}
-%\end{minipage}
-%\begin{minipage}[b]{0.50\linewidth}
-%\subsection*{Expected contribution}
-%\begin{itemize}
-%\item Build a framework in which an agent set to learn some task by action-reward can learn causal structure about its environment.
-%\item Algorithm that learn causal structure in an on-line way.
-%\item Using causal knowledge in each step to have better insights in the next decision.
-%\end{itemize}
-%\subsection*{Expected time-line}
-%\begin{itemize}
-%\item May 2018: Formal proposal.
-%\item August 2018: Proposal defense.
-%\%item December 2018: On-line causal learning module
-%\end{itemize}
-%\end{minipage}
+
 \end{document}

### 2018-07-03 Segundo_Semestre/CausalML_Workshop/Poster/Playing_Against_Nature_Poster.tex
poster

@@
 \begin{flushleft}
-\LARGE \color{NavyBlue} \textbf{ Acquiring causal models via reinforcement learning} \color{Black}\\ % Title
-\Large\textit{From reinforcement learning to computational psychology}\\ % Subtitle
+\LARGE \color{NavyBlue} \textbf{Playing against Nature: causal discovery for decision making under uncertainty} \color{Black}\\ % Title
 %\small \textbf{Mauricio Gonzalez Soto}\\ % Author(s)
@@
 \begin{center}
-\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/logos_2.png}
+\includegraphics[scale=0.2]{/Users/MauricioGS1/INAOE/Primer_Semestre/Figures/inaoe_logo.jpg}
 \end{center}
@@
 \begin{center}
-\large \textbf{Mauricio Gonzalez Soto}\\ % Author(s)
-\large \textbf{Asesor: Dr. Hugo Jair Escalante}\\
-Email: \texttt{mauricio@inaoep.mx}
+\large \textbf{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante}\\ % Author(s)
+Email: \texttt{mauricio@inaoep.mx, esucar@inaoep.mx, hugojair@inaoep.mx}
 \end{center}
@@
 \section*{Introduction}
+We consider uncertain environments controlled by a fixed causal mechanism and autonomous rational agents operating within.\\
+\\
+Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), we propose that an autonomous agent can learn and use a causal model while interacting an uncertain environment.\\
 
@@
 \section*{Related work}
+We are considering decision problems where after each decision made by a ration agent an outcome is observed and then an action is made over an independent instance of the problem.\\
+\\
+This setting is modelled as a \textit{bandit problem} where one out of many arms of a \textit{slot machine} is pulled, an reward observed and then an agent moves to another machine.\\
+\\
+Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure}, \cite{audibert2010best}, \cite{gabillon2012best}, \cite{agarwal2014taming} , \cite{jamieson2014lil},  \cite{jamieson2014best}, \cite{chen2015optimal}, \cite{carpentier2016tight}, \cite{russo2016simple},  \cite{kaufmann2016complexity}, but none of these works consider causal-governed environments.\\
+\\
+As far as we know \cite{lattimoreNIPS2016} is the first paper to consider causal relations between the effects of actions.They consider a decision maker who must choose the best among several possible interventions on a given causal model.
+
+By considering a causal model which is partially known and intervening variables from the unknown part of the model \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.
 
@@
 %----------------------------------------------------------------------------------------
+\section*{Problem setup}{\label{problem_setup}}
+Let $(\mathcal{A},\mathcal{E},\mathcal{C},\succeq)$ a Decision Problem under Uncertainty in which an agent has to choose one among several options $a \in \mathcal{A}$ which are causally related to the elements of $\mathcal{C}$ according to a Causal Graphical Model $\mathcal{G}$.\\
+\\
+The decision maker seeks to maximize her utility and we assume that she has \textit{rational preferences}, so we can substitute her preferences $\succeq$ for the expected value of a utility function $u$ (\cite{gilboa2009decision}).
 
@@
 %----------------------------------------------------------------------------------------
-\section*{Proposed Methodl}
+\section*{Proposed Method}
+The basic idea underlying our method is that ff a decision maker does not know the underlying causal model then she can learn it by interaction.
+\\
+For the sake of explanation we consider three separate cases:
+\begin{itemize}
+\item The decision maker fully knows the causal model.
+\item The decision maker knows only the structure of the causal model.
+\item The decision maker does not know the causal model.
+\end{itemize}
 
+\subsection{The causal model is completely known}{\label{known}}
+From Pearl's do-calculus \cite{pearl2009causality} says that the effect of setting some variable $X_i$ to a value $x_j$ can be expressed as:
+\[ P(X_1,...,X_n | do(X_i = j )) = \prod_{k \neq i} P(X_k | \textrm{Pa}(X_k)). \]
+
+The decision maker can use this expression to find the probabilities for her desired value of the target variable and choose the action which achieves the highest probability.
+
+\subsection{Only the structure is known}
+To model the interaction between the decision maker and the environment, we consider a game between the decision maker and a player called Nature who will select his actions according to the causal model.\\
+\\
+Since the joint distribution of the variables in the causal model can be expressed  
+\[ \prod_j P(X_j | Pa(X_j))\]
+We can use distributions over distributions to characterize beliefs over the model, we use the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$, which is useful since it is conjugate for itself (\cite{hjort2010bayesian}, \cite{bernardo2000bayesian}).\\
+\\
+In this way, the decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
+Using the agent's current beliefs, a causal graphical model can be specified.
+Using this fully specified (structure + parameters) as a true model, the decision maker will make her choice as in Case 1.
+
+%----------------------------------------------------------------------------------------
+%	Experiment
+%----------------------------------------------------------------------------------------
+
+\section{Test Scenario}
+Consider a patient who arrives at a hospital who can either have disease $A$ or disease $B$.
+The doctor can either give him some pill or send him into surgery.
+Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.\\
+\\
+The Causal Model is shown in Figure \ref{causal_model}.
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
+\captionof{figure}{Causal graphical model for the test scenario.}
+\label{causal_model}
+\end{center}
+
+The variable \textit{Lives} is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the variable \textit{Treatment}. \\
+\\
+By applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
+\[ P(Y | do(Tr))=P(Y | D, Tr, R)P(R | Tr) P(D). \]
+
+\section*{Experiment}
+As proof of concept we implemented  Case 1 and Case 2 for the test scenario and compared it with an agent performing Q-learning \cite{watkins1992q} and an agent choosing her actions at random.\\
+\\
+For the implementation, we defined a \textit{true} causal model as the one shown in Figure \ref{causal_model} using the library Pgmpy \cite{ankan2015pgmpy}.\\
+\\
+This agent will find the action that maximizes her desired value for the target variable using do-calculus.
+The action thus selected will be used as \textit{evidence} in the true causal model and a MAP inference will be used to simulate the most likely outcome given this action.
+\\
+\\
+The target variable will output a $1$ value if the patient lives. The value of the target variable will be the reward of each round.
+\subsection{Case 1: The causal model is completely known}
+
+\subsection{Case 2: Only the structure is known}
+We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
+We show the average performance over $20, 50, 100$ and $200$ rounds.\\
+\\
+In Figure \ref{20_rounds} we observe the average rewards for each agent in 20 rounds of decision making.
+Here we notice that Q-learning outperforms our algorithm, which has a similar performance as the random choosing procedure until round 11.
+
+\begin{center}
+\centerline{\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}}
+\captionof{figure}{Average reward obtained in each round for each agent}
+\label{20_rounds}
+\end{center}
+
+In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making. Our algorithm follow closely the Q-learning agent and outperform the random agent.
+
+\begin{center}
+\centerline{\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}}
+\captionof{figure}{Average reward obtained in each round for each agent}
+\label{50_rounds}
+\end{center}
+
+In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds, where our algorithm slightly outperforms Q-Learning.
+
+\begin{center}
+\centerline{\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}}
+\captionof{figure}{Average reward obtained in each round for each agent}
+\label{100_rounds}
+\end{center}
+
+In Figure \ref{200_rounds} we observe the average reward obtained by the three agents in 200 rounds. The average reward obtained is very similar for Q-learning and our algorithm.
+
+\begin{center}
+\centerline{\includegraphics[scale=0.8]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}}
+\captionof{figure}{Average reward obtained in each round for each agent}
+\label{200_rounds}
+\end{center}
+
+We see that our method obtains a very similar reward as the classic Q-learning algorithm for a larger number of rounds, where the random agent is outperformed, but our model offers something extra because it learns a causal model of the environment.
 %------------------------------------------------------------------------------------------

### 2018-07-03 Propuesta/Proposal_English.tex
correcciones axiomas

@@
 \usepackage{natbib}
+\usepackage{enumitem}
 %\bibliographystyle{stylename}
@@
 
+Learning a causal model of a certain environment allows transferability of knowledge into similar domains.
+\subsection{Outline}
+
 \newpage
@@
 Some event \textit{causes} another event to occurr.
-For example, The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
+The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
 	\begin{defi}{\label{causal_relation}}
@@
 	From the prerefence relation $\succeq$ we can derive some other relations between actions:
-	\begin{itemize}
+	\begin{enumerate}[label=(\roman*)]
 	\item $a_1 \sim a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and } a_2 \succeq a_1$
 	\item $a_1 \succ a_2 \Leftrightarrow a_1 \succeq a_2 \textrm{ and it does not hold that } a_2 \succeq a_1$.
-	\end{itemize}
+	\end{enumerate}
 	\end{defi}
-	\begin{defi}
+	\begin{defi}{\label{more_likely}}
 	A relation between events can be derived from $\succeq$ in the following way:
@@
 	\end{defi}
-	\begin{defi}
+	\begin{defi}{\label{cond_pref}}
 	Given some $G \succ \emptyset$ we define a conditional preference relation $\succeq_G$ as follows:
-	\begin{itemize}
-	\item i) $a_1 \succeq_G a_2 \Leftrightarrow \textrm{ for all a } \{ a_1 | G, a | G^c \} \succeq \{ a_2 | G, a | G^c \}$.
-	\item ii) $E \succeq_G F \Leftrightarrow \textrm{ for } c_1 \succeq_G c_2 \{ c_2 | E , c_1 | E^c \} \succeq_G \{c_2 | F , c_1  | F^c \}$ 
-	\end{itemize}
+	\begin{enumerate}[label=(\roman*)]
+	\item $a_1 \succeq_G a_2 \Leftrightarrow \textrm{ for all a } \{ a_1 | G, a | G^c \} \succeq \{ a_2 | G, a | G^c \}$.
+	\item $E \succeq_G F \Leftrightarrow \textrm{ for } c_1 \succeq_G c_2 \{ c_2 | E , c_1 | E^c \} \succeq_G \{c_2 | F , c_1  | F^c \}$ 
+	\end{enumerate}
 	\end{defi}
@@
 	Two events $E,F \in \mathcal{E}$ are said to be independent if and only if for all $c, c_1, c_2 \in \mathcal{C}$ it holds:
-	\begin{itemize}
+	\begin{enumerate}[label=(\roman*)]
 	\item $c \succ \{ c_2 | E, c_1 | E^c \} \Rightarrow c \succ_F \{ c_2 | E, c_1 | E^c \}	$.
 	\item $c \succ \{ c_2 | F, c_1 | F^c \} \Rightarrow c \succ_E \{ c_2 | F, c_1 | F^c \}	$
-	\end{itemize}
+	\end{enumerate}
 	\end{defi}
@@
 	\\
-	\indent \textbf{Axiom 1}.
-	\begin{itemize}
-	\item i) There exists consequences $c_1$, $c_2$ such that $c_1 \succ c_2$.
-	\item ii) For all consequences $c_1, c_2$ and events $E,F \in \mathcal{E}$ either $\{  c_2 | E , c_1 | E^c\} \succeq \{ c_2 | F, c_1 | F^c \}$ or $\{ c_2 | F, c_1 | F^c \} \succeq \{  c_2 | E , c_1 | E^c\}$
-	\end{itemize}
+	\indent \textbf{Axiom 1: Comparability}.
+	\begin{enumerate}[label=(\roman*)]
+	\item There exists consequences $c_1$, $c_2$ such that $c_1 \succ c_2$.
+	\item For all consequences $c_1, c_2$ and events $E,F \in \mathcal{E}$ either $\{  c_2 | E , c_1 | E^c\} \succeq \{ c_2 | F, c_1 | F^c \}$ or $\{ c_2 | F, c_1 | F^c \} \succeq \{  c_2 | E , c_1 | E^c\}$
+	\end{enumerate}
+	We require that the decision maker is able to strictly prefer at least one action over another, because if this weren't the case then there would not be a decision problem, since any action would lead to the same consequence.
+It is also required that the decision maker has preferences over simple options.
 
-	\textbf{Axiom 2}.
-	\begin{itemize}
-	\item i) $a \succeq a$ for all $a \in \mathcal{A}$.
-	\item ii) If $a \succeq b$ and $b \succeq c$ for $a,b,c \in \mathcal{A}$, then $a \succeq c$.
-	\end{itemize}
+	\textbf{Axiom 2: Transitivity}.
+	\begin{enumerate}[label=(\roman*)]
+	\item $a \succeq a$ for all $a \in \mathcal{A}$.
+	\item If $a \succeq b$ and $b \succeq c$ for $a,b,c \in \mathcal{A}$, then $a \succeq c$.
+	\end{enumerate}
+	Even though transitivity is a very intuitive requirement, its necessity is better explained in terms of \textit{money pumps}: consider $c_1, c_2$ and $c_3$ such that $c_1 \succ c_2$ and $c_2 \succ c_3$ but $c_3 \succ c_1$.
+In this case, the decision maker would be willing to pay a strictly positive ammount to have $c_3$ over $c_1$, and to pay $y>0$ to have $c_2$ rather than $c_3$, and an amount $z>0$ in order to have $c_1$ rather than $c_2$, and then again to pay to have $c_3$ over $c_1$...
 
-	\textbf{Axiom 3}
-	\begin{itemize}
-	\item i) If $a_1 \succeq a_2$ then for all $G \succ \emptyset$ $a_1 \succeq_G a_2$.
-	\end{itemize}
+	\textbf{Axiom 3: Consistency.}
+	\begin{enumerate}[label=(\roman*)]
+	\item If $a_1 \succeq a_2$ then for all $G \succ \emptyset$ $a_1 \succeq_G a_2$.
+	\item If for some $c_1 \succ c_2$ it holds that $\{ c_2 | E, c_1 | E^c \} \succeq \{ c_2 | F, c_1 | F^c \}$ then $E \succeq F$.
+	\item If for some $c \in \mathcal{C}$ and $G \succ \emptyset$ it holds that $\{ c_1 | G, c | G^c \} \succeq \{c_2 | G, c | G^c  \}$ then $c_1 \succeq_G c_2$.
+ 	\end{enumerate}
+    The intuition behind condition (i) is that preferences between \textit{pure} consequences should not be affected by information about the uncertain events in $\mathcal{E}$.
+Conditions (ii) and (iii) make definitions \ref{more_likely} and \ref{cond_pref} well defined concepts.
+In particular, condition (ii) formalises the idea that preferences should not depend on relative likelihood.
+Condition (iii) commonly known as the \textit{sure thing principle} (\cite{bernardo2000bayesian}).
 
-	\textbf{Axiom 4.}\\
+	\textbf{Axiom 4: Existence of Standard Events.}\\
 	There exists a sub-algebra $\mathcal{S}$ of $\mathcal{E}$ and a function $\mu : \mathcal{S} \to [0,1]$ such that:
-	\begin{itemize}
-	\item i) $S_1 \succeq S_2$ if and only if $\mu(S_1) \succeq \mu(S_2)$.
-	\item ii) If $S_1 \cap S_2 = \empty$ then $\mu(S_1 \cup S_2) = \mu(S_1) + \mu(S_2)$.
-	\item iii) For any number $\alpha \in [0,1]$ and independent events $E,F$, there is an event $S$, which is called a standard event, such that $\mu(S)=\alpha$ and $S$ is independent from $E$ and from $F$.
+	\begin{enumerate}[label=(\roman*)]
+	\item $S_1 \succeq S_2$ if and only if $\mu(S_1) \succeq \mu(S_2)$.
+	\item If $S_1 \cap S_2 = \empty$ then $\mu(S_1 \cup S_2) = \mu(S_1) + \mu(S_2)$.
+	\item For any number $\alpha \in [0,1]$ and independent events $E,F$, there is an event $S$, which is called a \textit{standard event}, such that $\mu(S)=\alpha$ and $S$ is independent from $E$ and from $F$.
 	\item iv) If $S_1$ is independent from $S_2$ then $\mu(S_1 \cap S_2)=\mu(S_1)\mu(S_2)$.
 	\item v) If $E$ independent of $S$, $S$ independent from $F$ and $E$ independent from $F$ then $E \sim S$ implies that $E \sim_F S$.
-	\end{itemize}
+	\end{enumerate}
+	This axioms means that the decision maker has a \textit{mental procedure} to generate a probability distribution where one event is more likely than other if and only if it has a higher probability.
+From the previous axioms, the preferences and uncertainties of a decision maker can be quantified up to small intervals, but in order to assign a unique value the following axiom is required.
 
-	\textbf{Axiom 5.}
-	\begin{itemize}
-	\item i) If $c_1 \succeq c \succeq c_2$ then there exists a standard event $S$ such that $c \sim \{ c_2 | S, c_1 | S^c \}$.
-	\item ii) For each event $E$ there exists a standard event $S$ such that $E \sim S$.
-	\end{itemize}
+	\textbf{Axiom 5: Precise measurement.}
+	\begin{enumerate}[label=(\roman*)]
+	\item If $c_1 \succeq c \succeq c_2$ then there exists a standard event $S$ such that $c \sim \{ c_2 | S, c_1 | S^c \}$.
+	\item For each event $E$ there exists a standard event $S$ such that $E \sim S$.
+	\end{enumerate}
 
-	For interpretations and critiques of the axioms we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}, \cite{peterson2017introduction}.
+	For a larger discussion, critiques and extensions of the axioms we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}, \cite{peterson2017introduction}.
 	\subsubsection{Subjective Probability and Utility}
@@
 This consequences need not have an interpretation, but they can be thought of as \textit{heaven} and \textit{hell} respectively for obvious (at lest in the Western hemisphery) reasons.
-Critiques of this extra technical consequences can be found in \cite{binmore2008rational}, \cite{peterson2017introduction}.
+Critiques of these extra technical consequences can be found in \cite{binmore2008rational}, \cite{peterson2017introduction}.
 	\begin{defi}
@@
 	For any bounded decision problem with extreme consequences $c_\ast$ and $c^\ast$ it holds:
-	\begin{itemize}
-	\item i) For all $c \in \mathcal{C}$, $u(c | c_\ast, c^\ast)$ exists and is unique.
-	\item ii) The value of $u(c | c_\ast, c^\ast)$ is unafected by the occurrence of any event $G \succ \emptyset$.
-	\item iii) $0 = u(c_\ast | c_\ast, c^\ast) \leq u(c | c_\ast, c^\ast) \leq u(c^\ast | c_\ast, c^\ast) = 1$.
-	\end{itemize}
+	\begin{enumerate}[label=(\roman*)]
+	\item For all $c \in \mathcal{C}$, $u(c | c_\ast, c^\ast)$ exists and is unique.
+	\item The value of $u(c | c_\ast, c^\ast)$ is unafected by the occurrence of any event $G \succ \emptyset$.
+	\item $0 = u(c_\ast | c_\ast, c^\ast) \leq u(c | c_\ast, c^\ast) \leq u(c^\ast | c_\ast, c^\ast) = 1$.
+	\end{enumerate}
 	\end{prop}
@@
 	This result means that for any decision maker whose preferences satisfies axioms 1 through 5 the only election criteria that is compatible with the axioms is the maximization of expected utility and thus establishes a \textit{normative} criteria for decision making for rational agents.
-Of course, this result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum and further mathematical assumptions are required over the utility function $u$ in order to guarantee the existence of the maxmimum utility option (\cite{bernardo2000bayesian}).
+This result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum and further mathematical assumptions are required over the utility function $u$ in order to guarantee the existence of the maxmimum utility option (\cite{bernardo2000bayesian}).
 
@@
 	A \textit{game} arises when two or more \textit{rational} decision makers, or players, have to make decisions in a situation in which the outcome for each player is partly determined by the choices of other players (\cite{binmore2008rational}).
-
 Game Theory is an area of Mathematics which is used to model games.
@@
 	Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn causal information and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
+Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn a causal model of the environment and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
 	\subsection{Objectives}
 		\subsubsection{General Objectives}
-			The proposed research has as a general objective to find how a rational autonomous agent who faces an uncertain environment can discover and use causal information in order to make good choices that point towards a concise objective such as the maximization of some function.
+			The proposed research has as a general objective to find how a rational autonomous agent who faces an uncertain environment can discover and use causal information in order to make good choices that maximize the expected utility for the decision maker.
 		\subsubsection{Specific Objectives}
-		In order to achieve the General Objective, we need to separate it into smaller problems which are required to be solved first.
+		In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first.
 In first place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.
@@
 July 15 2018, Stockholm, Sweden.
+This can be confirmed at \url{https://sites.google.com/site/faim18wscausalml/program-schedule}
 \end{itemize}

### 2018-07-08 Propuesta/Proposal_English.tex
correcciones y cambios

@@
 %\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Use and acquisition of causal relations for decision making under uncertainty using imperfect information repeated games}
+\title{Use and acquisition of causal relations for decision making under uncertainty using imperfect information games}
 \author{Mauricio Gonzalez Soto}
@@
 \indent A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}, \cite{pearlwhy}).
-In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the classic formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+Uncertain conditions mean that given a set of available choices, a decision maker can not predict the effect of the his choices.
+Because of this, the decision maker must take into consideration the stochastic relation between actions and consequences.
+From the von Neumann-Morgenstern decision making theory we know that when a rational\footnote{To be explained in full detail later.} decision maker knows this stochastic relation, then he must\textit{If rationality is assumed} choose the action that maximizes the \textit{expected utility} of a \textit{utility function} whose existence is guaranteed (\cite{von1944theory}).
+On the other hand, when a rational decision maker does not know the probabilities of the ocurrence of certain outcomes, or consequences, given his actions, then he must encode his personal, subjective, quantifications of any uncertainties in a probability distribution $p$ and choose the action that maximizes expected utility with respect to this probability (\cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+If the current \textit{knowledge} of a decision maker is not enough to make good choices, then the decision maker can \textit{learn} from the environment by interacting with it in order to update his quantifications of uncertainty (\cite{bernardo2000bayesian}, \cite{peterson2017introduction}).
+
+%In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the classic formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
 On the other hand, if enough information is not available, the decision maker could attempt to \textit{learn} from the environment by interacting with it.
@@
 Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting (\cite{sutton1998reinforcement}), but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
-
 A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
@@
 
-The proposed way for an agent to learn from repeated interactions is by giving her \textit{beliefs} about the structure of the environment and a way to update them after an outcome has been observed.
+The proposed way for an agent to learn from repeated interactions is by giving him \textit{beliefs} about the structure of the environment, which are to be used as a \textit{true} causal model, and then are to be updated them after an outcome has been observed.
+While the standard setting in RL is to model the agent-environment interaction as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model in response to what the decision maker has chosen.
+The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true model.
 
-While the standard setting in RL is to model the agent-environment interaction as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model in response to what the decision maker has chosen.
-
-The agent, using her current beliefs, will generate a \textit{local} causal model and choose an action from it as if that model was the true one.
+%The agent, using her current beliefs, will generate a \textit{local} causal model and choose an action from it as if that model was the true one.
 Then, after she observes the consequences of her actions, her beliefs will be updated according to the observed information in order to make a better choice the next time.
@@
 
-Learning a causal model of a certain environment allows transferability of knowledge into similar domains.
+Learning a causal model of a certain environment allows transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
 \subsection{Outline}
+In the Section \ref{Main_Concepts} the main concepts that serve as a foundation for this work will be described; in particular, we will describe in detail what causality means.
+Since this research proposal asumes that a causal model governs an environment, a precise definition of causality will be given along with the mathematical machinery to perform causal inference.
+Also, since the entity that will be facing causal environments is a \textit{rational} decision maker we will precisely define, axiomatically, what does rationality means and which are the main theories of rational decision making.
+Finally, since interactive learning is considered, we will describe a mathematical theory which is suitable to modelling interaction between rational decision makers.
 
+After the main concepts have been exposed, in Section \ref{Problem_Statement} a more precise statement of the problem that we propose to attack will be given, along with the Hypothesis that we attempt to defend and believe that is the answer to the problem.
+
+Once the main problem has been stated, we proceed in Section \ref{Related_Work} to show previous attempts over this and similar problems. 
 \newpage
-\section{Main Concepts}
+\section{Main Concepts}{\label{Main_Concepts}}
 	\subsection{Attempts to define Causality}
@@
 Attempts to understand the meaning of \textit{cause} has troubled philosophers and scientists, from David Hume and John Stuart Mill to modern scientists as \cite{suppes1970probabilistic}, \cite{cartwright1983laws}, \cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{spohn2012laws}, \cite{sep-causation-probabilistic}.
-In particular, Suppes and Reichenbach tried to define causation as an increase in the probability of occurrence of one event when another event happened, which is stated as $X$ causes $Y$ if $P(Y|X) > P(Y)$, but this definition was flawed because that increase could happen for several reasons, like $Y$ being a cause of $X$, or both having a common cause $Z$ (\cite{pearl2018why}).
+In particular, Suppes and Reichenbach tried to define causation as an increase in the probability of occurrence of one event when another event happened, which is stated as $X$ causes $Y$ if $P(Y|X) > P(Y)$, but this definition was flawed because the reason behind such increase could happen for several other reasons, such as $Y$ being a cause of $X$, or both having a common cause $Z$ (\cite{pearl2018why}).
 
@@
 In fact, Judea Pearl never gives a formal definition of causality, rather he prefers to build the machinery required to answer \textit{causal queries} (\cite{pearl2009causality}, \cite{pearl2018theoretical}).
+Causality, as probability, can be thought of either as an objective phenomena that \textit{exists} in the environment, or as a mental construct that enables us to handle complex situations.
 
@@
 	\subsection{Definition of Causality}
-	Causality is understood as a \textit{stochastic} relation between \textit{events}.
-Some event \textit{causes} another event to occurr.
+	Causality is understood as a \textit{stochastic} relation between \textit{events}. Some event (or events) \textit{causes} another event to occurr. 
 	The formal definition of Causality that will be adopted in this work is the definition provided in \cite{spirtes2000causation}.
@@
 	We will say that $A$ is \textit{a cause} of $B$ (or that $A$ causes $B$, $A$ is the cause and $B$ is the effect) if $A \to B$.
+It is important to note that an event may have more than one cause, and that not necesarily each one of this causes is sufficient to produce the effect.
 	\end{defi}
-	We will assume that for any event $A \in \mathcal{F}$ there are no causes lying outside $\mathcal{F}$, and that the relations expressed by $\to$ are the only causal relations in the environment.
+	We will assume that for any event $A \in \mathcal{F}$ there are no causes lying outside $\mathcal{F}$.
+It will also be assumed that the relations expressed by $\to$ are the only causal relations in the environment.
 This assumption can be interpreted as not allowing \textit{intermediate} events between events known to be causally related.
@@
 	Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
-It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to Markov equivalence (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}).
+It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}).
 	\subsubsection{Do-Calculus}
@@
 
-	For a larger discussion, critiques and extensions of the axioms we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}, \cite{peterson2017introduction}.
+	For a richer discussion, critiques and extensions of the coherence axioms we refer the reader to \cite{bernardo2000bayesian}, \cite{binmore2008rational}, \cite{gilboa2009decision}, \cite{wakker2010prospect}, \cite{peterson2017introduction}.
 	\subsubsection{Subjective Probability and Utility}
@@
 
-	The result extends previous work of \cite{von1944theory} who assumed that the decision maker knows the probabilties of the uncertain events and guarantees the existence of an utility function such that the decision maker chooses by calculating expected value of this utility function.
+	The result extends previous work of \cite{von1944theory} where it is assumed that the decision maker knows the probabilties of the uncertain events.
 	\subsection{Game Theory}
@@
 
-\begin{defi}
-\end{defi}
-
 %\subsubsection{Repeated Games}
@@
 \newpage
-\section{Problem Statement}
+\section{Problem Statement}{\label{Problem_Statement}}
 	\subsection{Problem Statement}
 	Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-And consider a rational decision maker who doesn't know the parameters of the causal model which control the probabilities of observing a consequence given an action $a \in \mathcal{A}$, we ask how the decision maker could learn about the causal structure that controls his environment in order to make good choices with respect to the decision makers' preferences.
+Consider a rational decision maker who doesn't know the parameters of the causal model which control the probabilities of observing a consequence given an action $a \in \mathcal{A}$, we ask how the decision maker could learn about the causal structure that controls his environment in order to make good choices with respect to the decision makers' preferences.
 	\subsection{Motivation}
@@
 Using only associations between variables it is not possible to infer which one is the cause and which one the effect, something extra is required.
+The fundamental difference between causal and associative models is that causal models consider \textit{doing} over \textit{observing} (\cite{pearl2018why}).
+Once a causal relation has been established between two events, performing, or intervening, over a cause allows to predict the outcome in a more robust way than if using only correlations between observations.
 
@@
 	How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
-How can causal structure help a decision maker in order to guide his learning process?
+How can a causal structure help a decision maker in order to guide his learning process?
 	\subsection{Hypothesis}
@@
 		\subsubsection{Specific Objectives}
-		In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first.
+		\begin{enumerate}
+		\item In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first.
 In first place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.
 
-		In second place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.
+		\item In second place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.
 
-		Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker.
+		\item Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker.
 A general updating criteria must be provided which uses the causal nature of the environment.
+		\end{enumerate}
 
@@
 \section{Related Work}
-The use of causal information in human beings has been studied by York Hagmayer and his team and they have found that human beings, while facing a sequential decision problem, use and modify causal information (\cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2010spontaneous}, \cite{hagmayer2013repeated}).
+Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction.
+In this setting it is assumed that a Markov Decision Process (MDP) is able to capture the effects of the agent's actions in the environment.
+An MDP is composed by a set of states, a transition function and a reward function, where an agent's action moves the state of the environment according to the transition function and obtaining a reward in such process.
+
+Reinforcement Learning can be roughly divided into two categories: model-based and model-free methods.
+In the former case the agent learns an \textit{optimal policy} from a model of the environment, such as the Dynamic Programming method explained in \cite{sutton1998reinforcement}.
+On the other hand, if the agent does not know the transition function, then he must learn directly values of \textit{state-action} pairs.
+A very well known model-free method is the classic Q-learning algorithm developed by \cite{watkins1992q}.
+In either case, model-free or model-based methods an agent must use information from transitions or from observations.
+For this reason, Reinforcement Learning uses only associative patterns for finding an optimal policy and this methods can not learn beyond what is expressable by an MDP, which is constrained into associative rules and can not express any other kind of structure, such as causal relations.
+
+Since an agent that learns how to act according to any Reinforcement Learning method acquires an optimal policy, optimal in the sense of satisfying the Bellman equations, it can be shown that such policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
+In this way, when only associative information is available, RL is a coherent learning framework.
+On the other hand, it remains to be answered how to incorporate a causal model into an on-line decision making algorithm.
+
+The formation and use of causal knowledge in human beings can provide inspiration for our problem, and it has been extensively studied by York Hagmayer and his team They have found that human beings, while facing a sequential decision problem, use and modify causal information (\cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2010spontaneous}, \cite{hagmayer2013repeated}).
 In their studies, all of the subjects were faced with a one-stage decision problem in which an initial causal model was provided.
@@
 The subjects knew that they were supposed to learn a causal model, but the case where the final objective is other than directly learning a causal model has not been studied.
+\cite{bramley2015conservative} show that people acquire causal structures by selecting the \textit{most informative} options, in the sense of decreasing uncertainty about the true model instead of maximizing expected utility.
 
@@
 Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal structure when little prior information is available.
+
+An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who concluded that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes (\cite{joyce1999foundations}, \cite{peterson2017introduction}).
+His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality, although he captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker.
 
@@
 It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}) if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
-
 On the other hand, if the largest mean reward is desired with a fixed confidence of $\delta$ then an algorithm for finding the optimal action exist in such a way that the number of rounds to find it is within a constant factor of a lower bound (\cite{jamieson2014lil}).
-
 If one does not consider a fixed confidence but instead a fixed number of learning rounds (known as fixed budget), then we face a \textit{fixed budget} problem where a lower bound for the probability of finding the optimal arm is given by \cite{carpentier2016tight}.
@@
 Their work requires that the causal model is fully known to the decision maker, this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part.
+
+In \cite{lattimoreNIPS2016} a causal graphical model $\mathcal{G}$ is assumed to be known and a number of learning rounds $T$ is fixed.
+In round $t \in [1,...,T]$ the decision maker chooses $a_t=do(X_t = x_t)$ and observes a reward $Y_t$.
+After the $T$ learning rounds, the decision maker is expected to choose an optimal action $a^\ast$ that minimizes the \textit{expected regret}, which is defined as $R_T= \mu^\ast - \mathbb{E}[\mu_{a^\ast}]$ where $\mu^\ast = \max \mathbb{E}[a]$.
+They show that the achieved regret is of a smaller order than the regret obtained by non-causal algorithms.
 
@@
 
-After the decision maker makes his play, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
+After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
 This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games described in Section \ref{basic_models}.
@@
 The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
-The observed information will modify his current knowledge, which is encoded as probabilistic beliefs, about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world (\cite{hagmayer2013repeated}).
+The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world (\cite{hagmayer2013repeated}).
 
@@
 Also, choosing this action is the \textit{best response} action, and therefore a Nash equilibrium for the game.
+Choosing the action that is more likely to produce the most desired outcome according to current causal knowledge is required by Joyce's Causal Decision Theory (\cite{joyce1999foundations}).
 \subsubsection{Knowing only the graph structure}
@@
 In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e.
-in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
+in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge and those beliefs will be later updated using what is observed after choosing an action.
 \subsubsection{General case: Model unknown}
@@
 \subsection{Validation of the proposed methods}
-In order to show that an agent has succeeded in acquiring a causal model from his environment via interacting with it and using it in order to make good choices, 
+In order to show that an agent has succeeded in acquiring a causal model from his environment via interacting with it and using it in order to make good choices, it is necessary to compare the average utility obtained in a series of rounds with the theoretical utility that would be obtained if the causal model was fully known, and to compare the performance obtained by a learning agent that is not considering causal information.
+Also, the parameters of the learned causal model must be close to the true ones.
+
 \subsection{Experimental methodology}
@@
 In this way, if an external user is able to give a causal model, the method is expected to work independent of the particularities of the causal model which is being tested.
+\subsubsection{Concurrent Validation}
+Concurrent validation will be tested by comparing the performance obtained by the proposed method against the theoretical optima.
 \subsection{Working plan}
@@
 \end{itemize}
-\subsubsection{Publication plan}
+\subsubsection{Publications plan}
 \begin{itemize}
-\item Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
+\item \textbf{Year 1:} Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
 Accepted for a poster presentation at the CausalML 2018 workshop at ICML / IJCAI / AAMAS 2018.
 July 15 2018, Stockholm, Sweden.
-This can be confirmed at \url{https://sites.google.com/site/faim18wscausalml/program-schedule}
+This can be confirmed at \url{https://sites.google.com/site/faim18wscausalml/program-schedule}.
+The paper can be found at \url{https://arxiv.org/abs/1807.01268}.
+\item \textbf{Year 2:} Conference Paper
+		\begin{itemize}
+			\item Aim: Develop a Bayesian Non-parametric prior and updating procedure for beliefs about causal graphs in order to extend previous paper.
+		\end{itemize}
+\item \textbf{Year 3:} Journal Paper
+		\begin{itemize}
+			\item Aim: Theoretical paper of results for on-line causal learning.
+		\end{itemize}
 \end{itemize}
@@
 We show the results of two experiments.
-We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning.
-We show the average perform over $20, 50, 100$ and $200$ rounds.
+We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning (\cite{watkins1992q}).
+Q-learning was chosen since it learns an \textit{optimal policy} in the sense of the Bellman Equation (\cite{sutton1998reinforcement}) and it is shown in \cite{webb2007game} that such optimal policies also maximize expected utility.
+
+The average perform over $20, 50, 100$ and $200$ rounds are shown.
 

### 2018-08-16 Propuesta/Formato/Proposal_format.tex
Cambiando el Objetivo General

--- 2018-07-08 Propuesta/Proposal_English.tex
@@
-\documentclass[11pt]{article}
+\documentclass[english,letterpaper,12pt,final]{article}
+\usepackage{graphicx}
+\usepackage{etex}
+
+\sloppy
+%\special{papersize=8.5in,11in}% This command prevents the use of other paper size in the dvi2ps conversion
+%------Document properties------------
+\usepackage{setspace}
+\usepackage{pgfgantt}
+\usepackage{enumitem}
+%------math packages-------------------
+\usepackage{amsthm}
+\usepackage{tkz-berge} % Drawing graphs
+%------algorithms packages-------------------
+%\usepackage[ruled,vlined]{basic/algorithm2e}
+\usepackage[noline,linesnumbered,noend]{algorithm2e}
+%\linesnumbered
+%------bibliography packages-----------
+\usepackage{natbib}
+
 \usepackage[utf8]{inputenc}
-%\usepackage[spanish, mexico]{babel}
+
 \usepackage[font=small,labelfont=bf]{caption}
@@
 \usepackage{amssymb,amsthm,amsmath,latexsym}
-\usepackage[margin=1.5cm]{geometry}
+
 \usepackage{natbib}
-\usepackage{enumitem}
-%\bibliographystyle{stylename}
-%\usepackage{fancyhdr}
-%\pagestyle{fancy}
-\theoremstyle{plain}
-\newtheorem{teo}{Theorem}
-\newtheorem{prop}[teo]{Proposition}
-\newtheorem{defi}[teo]{Definition}
-\newtheorem{obs}[teo]{Observation}
-\newtheorem{lem}[teo]{Lemma}
-\newtheorem{cor}[teo]{Corolary}
-\usepackage[pdftex]{color,graphicx}
-\usepackage{tikz}
-\usetikzlibrary{trees}
-\usetikzlibrary{calc}
-%\newcommand{\sgn}{\mathop{\mathrm{sgn}}}
-\title{Use and acquisition of causal relations for decision making under uncertainty using imperfect information games}
-\author{Mauricio Gonzalez Soto}
+
+\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
+\newcommand{\leci}{<\!\!\!\!\!\!\!\bigcirc}
+%\newcommand{\lessint}{$\bigcirc$}
+%------corporative identity of INAOE---
+\usepackage{eso-pic}
+\usepackage{pstricks}
+\usepackage{INAOEMacros}
+
+\usepackage{dsfont}
+\usepackage{lineno,hyperref}
+
+\usepackage{amssymb}
+\usepackage{pifont}
+\usepackage{url}
+\usepackage{array}
+%\usepackage{subfigure}
+\usepackage{lscape}
+\usepackage{amsfonts}
+\usepackage{amsmath}
+
+\usepackage[normalem]{ulem}
+\usepackage{color}
+\usepackage{multirow}
+%\usepackage{mathtools}
+
+%\usepackage[dvipdfmx]{graphicx} 
+%\usepackage{bmpsize}
+
+\usepackage{caption}
+\usepackage{subcaption}
+%------floating -----------------------
+\usepackage{float}
+\floatstyle{plain}
+\newfloat{codigo}{thp}{lop}[section]
+\floatname{codigo}{Algorithm}
+\usepackage{booktabs}
+%
+%usepackage{lscape}
+%\usepackage{pdflscape}
+%\usepackage{rotating}
+\usepackage{tablefootnote}
+%------figure packages--------------
+%\usepackage{subfigure}
+
+%------page layout---------------------
+\oddsidemargin 0.36cm
+\evensidemargin 0.36cm
+\marginparwidth 0.0cm
+\textwidth 16.80cm
+
+\topmargin -1.5cm
+\headheight 1.0cm
+\headsep 1.0cm
+\footskip 1.0cm
+\textheight 22.34cm
+
+%\theoremstyle{plain}
+%\newtheorem{teo}{Theorem}
+%\newtheorem{prop}[teo]{Proposition}
+%\newtheorem{defi}[teo]{Definition}
+%\newtheorem{obs}[teo]{Observation}
+%\newtheorem{lem}[teo]{Lemma}
+%\newtheorem{cor}[teo]{Corolary}
+
+%\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{56}}%
+\theoremstyle{definition}
+\newtheorem{rmk}{Definition}[section]
+\newtheorem{teo}{Theorem}[section]
+\newtheorem{thm}{Theorem}[section]
+\newtheorem{prop}{Proposition}[section]
+\newtheorem{corollary}{Corollary}[section]
+\newtheorem{cor}{Corollary}[section]
+\newtheorem{problem}{Problem}
+\newtheorem{defi}{Definition}[section]
+\newtheorem{pf}{Proof}
+\newtheorem{pot}{Proof of Theorem \ref{thm2}}
+
+%----------myDeclarations------------
+\DeclareMathOperator*{\argmin}{argmin}
+\definecolor{blue1}{rgb}{0.58,0.70,0.84}
+\definecolor{blue2}{rgb}{0.93,0.95,1.0}
+\usepackage{xcolor,colortbl}
+\usepackage{titlesec}
+\setcounter{secnumdepth}{4}
+\titleformat{\paragraph}
+{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+\titlespacing*{\paragraph}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+
+%-----end of preamble------------------
+\pagenumbering{roman}
 \begin{document}
-%\nocite{*}
-\maketitle
+
+% Title page
+\input{gfrontpage}
+%\doublespacing
+\singlespacing
 \tableofcontents
@@
 Once the main problem has been stated, we proceed in Section \ref{Related_Work} to show previous attempts over this and similar problems.
+
 \newpage
@@
 \end{cor}
+
+\subsubsection{Use of Causal Graphical Models over other causal models}
+We choose to use specifically a Causal Graphical model because of the mathematical machinery that exists around them and because Graphs are a model of Cognitive Processes, such as causal reasoning (\cite{glymour2003learning}, \cite{danks2014unifying}).
+In this way, if an agent \textit{knows} that a certain environment is governed by causal relations, then a causal graphical model can be used to model the causal reasoning of the agent.
 
@@
 
-%\subsubsection{Repeated Games}
-%\subsubsection{Learning in Games}
-
 \newpage
@@
 How can a causal structure help a decision maker in order to guide his learning process?
+What does the rationality assumption implies about how to choose when considering causal information?
+How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
 	\subsection{Hypothesis}
@@
 		\subsubsection{General Objectives}
-			The proposed research has as a general objective to find how a rational autonomous agent who faces an uncertain environment can discover and use causal information in order to make good choices that maximize the expected utility for the decision maker.
+			The proposed research has as a general objective to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
 		\subsubsection{Specific Objectives}
@@
 
-\newpage
-\section{Related Work}
 Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction.
@@
 
-An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who concluded that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes (\cite{joyce1999foundations}, \cite{peterson2017introduction}).
+An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes (\cite{joyce1999foundations}, \cite{peterson2017introduction}).
 His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality, although he captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker.
+In his formulation of a Causal Decision Problem, he uses the \textit{probability of causing certain effect with an action}, which we refine here with the machinery of Pearl's Do-Calculus and Spirtes' definition of Causality.
 
-We are considering decision problems where an agent must choose some available action and then receive a reward from it.
-After a certain number of learning rounds, the agent is expected to output a good action.
-Problems of this nature are called \textit{bandit} problems (\cite{sutton1998reinforcement}).
+In the class of decision problems considered for this proposal it is assumed that the a rational agent must choose some available action and then receive a reward from it.
+In the case where the agent does not has enough information available, we consider a certain number of learning rounds and expect the agent to output a good action.
+Problems of this nature have been modelled as \textit{bandit} problems (\cite{sutton1998reinforcement}).
+A bandit is an analogy of a \textit{slot machine} as the ones found in Las Vegas, where a lever is pulled an some monetary reward (probably cero) is observed.
+A bandit problem consists by a set of available actions (the arms of the slot machine) and a reward function.
 It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}) if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
@@
 
+\subsection{Balancing exploitationd and exploration}
+When the model is to be discovered, such as in the general case mentioned before, an agent must choose also between keep making an intervention in a variable that is known to cause a desired output or to explore for new variables that could bring even better rewards, but at the cost of not having best performance while looking out for them; this is known as the exploration vs exploitation dilemma (\cite{sutton1998reinforcement}).
+This question is to be attacked using the concept of \textit{mixed strategy} in a game.
+A mixed strategy is a probability distribution over available actions for an agent, so similarly to the $\varepsilon$-greedy methods in Reinforcement Learning we allow for some actions to be randomly choosen in order to explore further possibilities.
+
 \subsection{Validation of the proposed methods}
@@
 \item Literature review, understanding the problem, considering several models and possible solutions.
-\item Formalize the model.
-\item Studying implications of the assumption of rationality in decisiom making under uncertainty in causal environments.
-\item Propose a representation of beliefs that allow causal structure to be considered.
+\item Formalization of a Causal Decision Problem when causal relations are encoded into a Causal Graphical Model.
+\item Studying implications of the rationality assumption for decisiom making under uncertainty in causal environments.
+\item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered
 \item Updating criteria for beliefs.
-\item Discovery of causal structure
+\item Discovery of causal structure.
+\item Considering interventions in more than one variable.
 
@@
 		\begin{itemize}
-			\item Aim: Develop a Bayesian Non-parametric prior and updating procedure for beliefs about causal graphs in order to extend previous paper.
+			\item When a causal model is unknown, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
+		\end{itemize}
+\item \textbf{Year 2:} Conference Paper
+	\begin{itemize}
+		\item Using game-theoretical aspects for balancing exporation and exploitation for causal discovery. 
 	\end{itemize}
@@
 \begin{center}
-\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/Seminario/Semana12/codigo/causal_graph.png}}
+\centerline{\includegraphics[width=0.7\textwidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/causal_graph.eps}}
 \caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
@@
 \begin{minipage}{0.48\linewidth}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/20_rounds_format.eps}
 \captionof{figure}{Average reward in 20 rounds}
@@
 \begin{minipage}{0.49\linewidth}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/50_rounds_format.eps}
 \captionof{figure}{Average reward in 50 rounds}
@@
 \begin{minipage}{0.49\linewidth}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/100_rounds_format.eps}
 \captionof{figure}{Average reward in 100 rounds}
@@
 \begin{minipage}{0.49\linewidth}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/200_rounds_format.eps}
 \captionof{figure}{Average reward in 200 rounds}
@@
 
-%In Figure \ref{20_rounds} we observe the average rewards for each agent in 20 rounds of decision making.
-Here we notice that Q-learning outperforms our algorithm, which has a similar performance as the random choosin procedure until round 11.
-%
-%\begin{figure}[ht]
-%\vskip 0.2in
-%\begin{center}
-%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/20_rounds_format.png}}
-%\caption{Average reward obtained in each round for each agent}
-%\label{20_rounds}
-%\end{center}
-%\vskip -0.2in
-%\end{figure}
-%
-%In Figure \ref{50_rounds} we observe the average rewards for each agent in 50 rounds of decision making. Our algorithm follow closely the Q-learning agent and outperform the random agent.
-%
-%\begin{figure}[ht]
-%\vskip 0.2in
-%\begin{center}
-%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/50_rounds_format.png}}
-%\caption{Average reward obtained in each round for each agent}
-%\label{50_rounds}
-%\end{center}
-%\vskip -0.2in
-%\end{figure}
-%
-%In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds, where our algorithm slightly outperforms Q-Learning.
-%
-%\begin{figure}[ht]
-%\vskip 0.2in
-%\begin{center}
-%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/100_rounds_format.png}}
-%\caption{Average reward obtained in each round for each agent}
-%\label{100_rounds}
-%\end{center}
-%\vskip -0.2in
-%\end{figure}
-%
-%In Figure \ref{200_rounds} we observe the average reward obtained by the three agents in 200 rounds. The average reward obtained is very similar for Q-learning and our algorithm.
-%
-%\begin{figure}[ht]
-%\vskip 0.2in
-%\begin{center}
-%\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/Segundo_Semestre/CausalML_Workshop/figures/200_rounds_format.png}}
-%\caption{Average reward obtained in each round for each agent}
-%\label{200_rounds}
-%\end{center}
-%\vskip -0.2in
-%\end{figure}
-
-We see that our method obtains a very similar reward as the classic Q-learning algorithm for a larger number of rounds, where the random agent is outperformed, but our model offers something extra because it learns a causal model of the environment.
-
-\newpage
+%------------- References --------------------
+\singlespacing
+%doi=false,isbn=false,url=false
 \bibliographystyle{apalike}
 \bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
+
+%------------- References --------------------
+
 \end{document}

### 2018-08-16 Propuesta/Formato/Proposal_format.tex
añadiendo principio guia

@@
 In this section we describe the steps that are to be followed in order to answer the Hypothesis.
+\subsection{A guiding principle for a Causal Decision Problem}
+We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem that is rational.
+We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
+\begin{prop}
+Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
+Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent.
+Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in mathcal{A}$ that satisfies:
+\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
+\end{prop}
 \subsection{Modelling a Decision Problem under Uncertainty as a Game}

### 2018-08-16 Propuesta/Formato/Proposal_format.tex
modificando los casos segun el principio guia

@@
 \subsubsection{Fully knowing the causal graphical model}
-If the decision maker knows the causal graphical model, then he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
+If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
 The decision thus made is the one that maximizes the expected utility, by construction.
@@
 in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge and those beliefs will be later updated using what is observed after choosing an action.
+The action will be chosen, given a set of beliefs, as in the previous case and by following the guiding principle.
+This means that using current beliefs, the agent forms a causal model which is used as if it were the true causal model.
 \subsubsection{General case: Model unknown}

### 2018-08-16 Propuesta/Formato/Proposal_format.tex
añadiendo nota sobre la palabra explicacion y requiriendo invarianza bajo intervenciones

@@
 \usepackage{setspace}
+\usepackage{pdfpages}
 \usepackage{pgfgantt}
@@
 When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action.
-
-Learning a causal model of a certain environment allows transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
+It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons.
+
+Learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
 \subsection{Outline}
@@
 
+Finally, we state the proposed solution for the problem as well as showing results of experiments performed which implement a small toy example.
+
 \newpage
@@
 
-Even though causality is a difficult concept to grasp, we require a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts. 
+Since causal claims in common language can have different logical form, in this work we give a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts and we will take the given definition as axiomatic.
 	\subsection{Definition of Causality}
@@
 	\end{itemize}
-	Also, an extra condition, known as Causal Sufficiency, about the nature of the model itself is considered: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ outside of the model $\mathcal{G}$ (\cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{sucar2015probabilistic}).
+	Also, an extra pair of conditions, the first one known as Causal Sufficiency, which is about the nature of the model: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ outside of the model $\mathcal{G}$ (\cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{sucar2015probabilistic}).
+The second required condition is that the causal mechanism remains unchanged after interventions, this is called in (\cite{woodward2005making}) as \textit{invariantness}, this basically means that an interventor does not \textit{breaks} the causal mechanism when intervening upon it.
 	This conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
@@
 
+\newpage
+\includepdf[pages=-]{paper_arxiv.pdf}
+
 %------------- References --------------------
@@
 
-%------------- References --------------------
+%------------- Apendice --------------------
 

### 2018-08-24 Propuesta/Formato/Proposal_format.tex
ultima version

@@
 \end{abstract}
+\pagenumbering{arabic}
 \section{Introduction}
 \indent A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}, \cite{pearlwhy}).
-Uncertain conditions mean that given a set of available choices, a decision maker can not predict the effect of the his choices.
-Because of this, the decision maker must take into consideration the stochastic relation between actions and consequences.
-From the von Neumann-Morgenstern decision making theory we know that when a rational\footnote{To be explained in full detail later.} decision maker knows this stochastic relation, then he must\textit{If rationality is assumed} choose the action that maximizes the \textit{expected utility} of a \textit{utility function} whose existence is guaranteed (\cite{von1944theory}).
-On the other hand, when a rational decision maker does not know the probabilities of the ocurrence of certain outcomes, or consequences, given his actions, then he must encode his personal, subjective, quantifications of any uncertainties in a probability distribution $p$ and choose the action that maximizes expected utility with respect to this probability (\cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+Uncertain conditions as when that a set of choices is made available to a decision maker he can not predict the precise consequence of his decision.
+For this reason, the decision maker must take into account the stochastic relation between actions\footnote{Actions, options and choices to be used equivalently.} and consequences.
+From the von Neumann-Morgenstern decision making theory we know that if a rational\footnote{To be explained in full detail later.} decision maker knows this stochastic relation, then he must\textit{If rationality is assumed} choose the action that maximizes the \textit{expected utility} of a \textit{utility function} whose existence is guaranteed (\cite{von1944theory}).
+On the other hand, if a rational decision maker does not know the probabilities of the ocurrence of certain outcomes, or consequences, given his actions, then he must encode his personal, subjective, quantifications of any uncertainties in a probability distribution $p$ and choose the action that maximizes expected utility with respect to this probability distribution(\cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
 If the current \textit{knowledge} of a decision maker is not enough to make good choices, then the decision maker can \textit{learn} from the environment by interacting with it in order to update his quantifications of uncertainty (\cite{bernardo2000bayesian}, \cite{peterson2017introduction}).
@@
 A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
-A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).
-
-Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect (\cite{rottman2014reasoning}), my hypothesis is that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
+A causal structure encondes a series of \textit{cause-effect} relations between events and knowing such relations allows a decision maker to add extra knowledge into the uncertainty of his environment and also allows to plan ahead his actions since he can predict what a certain action will cause (\cite{spirtes2000causation}, \cite{pearl2018theoretical}).\\
+\\
+Causal reasoning is found at the very core of human reasoning, since it allows us to manipulate our environment and being able to predict effects of a given action (\cite{spirtes2000causation}).
+Causal claims also support reasoning of the form \textit{if...hadn't occurred then...wouldn't have happened}.
+It is argued by \cite{pearl2018why} that imagining alternative scenarios is fundamental for humans to make causal reasoning.
+Since human beings are known to learn causal models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), and even though this learning is not perfect (\cite{rottman2014reasoning}), our hypothesis is that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.
 
@@
 Also, since the entity that will be facing causal environments is a \textit{rational} decision maker we will precisely define, axiomatically, what does rationality means and which are the main theories of rational decision making.
-Finally, since interactive learning is considered, we will describe a mathematical theory which is suitable to modelling interaction between rational decision makers.
+At the end of the section, we will describe a mathematical theory which is suitable to modelling interaction between rational decision makers since interactive learning is considered.
 
@@
 If an $A$ event is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
+It is proven in \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.
 	\subsubsection{Relation between the graph and probabilities}
@@
 		\end{itemize}
-
+\newpage
+\section{Related Work}{\label{Related_Work}}
 Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction.
@@
 \subsection{A guiding principle for a Causal Decision Problem}
-We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem that is rational.
+We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational.
 We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
@@
 Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent.
-Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in mathcal{A}$ that satisfies:
+Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
 \[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
@@
 If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
-For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model.
+For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
 
@@
 This question is to be attacked using the concept of \textit{mixed strategy} in a game.
-A mixed strategy is a probability distribution over available actions for an agent, so similarly to the $\varepsilon$-greedy methods in Reinforcement Learning we allow for some actions to be randomly choosen in order to explore further possibilities.
+A mixed strategy is a probability distribution over available actions for an agent, in this way, similar to the $\varepsilon$-greedy methods in Reinforcement Learning, we allow for some actions to be randomly choosen in order to explore further possibilities.
 
@@
 \item Literature review, understanding the problem, considering several models and possible solutions.
-\item Formalization of a Causal Decision Problem when causal relations are encoded into a Causal Graphical Model.
-\item Studying implications of the rationality assumption for decisiom making under uncertainty in causal environments.
+\item Studying implications of the rationality assumption for decision making under uncertainty in causal environments.
+\item Formalization of a Causal Decision Problem: a decision problem under uncertainty where causal relations are encoded into a Causal Graphical Model.
+Propose a solution using the rationality assumption.
 \item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered
-\item Updating criteria for beliefs.
-\item Discovery of causal structure.
+\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information.
+\item Study the problem of on-line causal discovery.
 \item Considering interventions in more than one variable.
@@
 \item \textbf{Year 1:} Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
-Accepted for a poster presentation at the CausalML 2018 workshop at ICML / IJCAI / AAMAS 2018.
+Accepted for poster presentation at the \textbf{CausalML 2018 Workshop at ICML 2018}.
 July 15 2018, Stockholm, Sweden.
@@
 		\begin{itemize}
+		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
+		\end{itemize}
+\item \textbf{Year 2:} Conference Paper
+	\begin{itemize}
 	\item When a causal model is unknown, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
-		\end{itemize}
-\item \textbf{Year 2:} Conference Paper
-	\begin{itemize}
-		\item Using game-theoretical aspects for balancing exporation and exploitation for causal discovery. 
 	\end{itemize}
@@
 
-\newpage
-\includepdf[pages=-]{paper_arxiv.pdf}
-
 %------------- References --------------------

### 2018-09-02 Propuesta/Formato/Proposal_format.tex
cambios menores

@@
 \indent A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}, \cite{pearlwhy}).
-Uncertain conditions as when that a set of choices is made available to a decision maker he can not predict the precise consequence of his decision.
+Uncertain conditions mean that if a set of choices is made available to a decision maker he can not predict the precise consequence of his decision.
 For this reason, the decision maker must take into account the stochastic relation between actions\footnote{Actions, options and choices to be used equivalently.} and consequences.
-From the von Neumann-Morgenstern decision making theory we know that if a rational\footnote{To be explained in full detail later.} decision maker knows this stochastic relation, then he must\textit{If rationality is assumed} choose the action that maximizes the \textit{expected utility} of a \textit{utility function} whose existence is guaranteed (\cite{von1944theory}).
-On the other hand, if a rational decision maker does not know the probabilities of the ocurrence of certain outcomes, or consequences, given his actions, then he must encode his personal, subjective, quantifications of any uncertainties in a probability distribution $p$ and choose the action that maximizes expected utility with respect to this probability distribution(\cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+From the von Neumann-Morgenstern decision making theory we know that if a rational\footnote{To be explained in full detail later.} decision maker knows this stochastic relation, then he \textit{must} choose the action that maximizes the \textit{expected utility} of a \textit{utility function} whose existence is guaranteed (\cite{von1944theory}).
+On the other hand, if a rational decision maker does not know the probabilities of the ocurrence of certain outcomes, or consequences, given his actions, then he must encode his personal subjective quantifications of any uncertainties in a probability distribution $p$ and choose the action that maximizes expected utility with respect to this probability distribution(\cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
 If the current \textit{knowledge} of a decision maker is not enough to make good choices, then the decision maker can \textit{learn} from the environment by interacting with it in order to update his quantifications of uncertainty (\cite{bernardo2000bayesian}, \cite{peterson2017introduction}).
@@
 While the standard setting in RL is to model the agent-environment interaction as an agent that moves from one \textit{state} to another inside a model of the environment and observing a reward as these transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called \textit{Nature} which will select his actions from the causal model in response to what the decision maker has chosen.
-The agent, besides learning a policy to choose actions will also learn a causal model from the environment since the causal model she forms will approximate the true model.
+The agent, besides learning a policy to choose actions, will also learn a causal model from the environment since the causal model she forms will approximate the true model.
 
@@
 	Notice that since the graph is finite, there exist some nodes that does not have causes, which are called \textit{exogenous}.
-If an $A$ event is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
+If an event $A$ is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
 It is proven in \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.
@@
 The second required condition is that the causal mechanism remains unchanged after interventions, this is called in (\cite{woodward2005making}) as \textit{invariantness}, this basically means that an interventor does not \textit{breaks} the causal mechanism when intervening upon it.
-	This conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
+	These conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
 	\subsection{Causal Graphical Models}
@@
 In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences \footnote{Also known as decision under risk (\cite{binmore2008rational}, \cite{peterson2017introduction}).}, and in that case the theory guarantees that the decision maker behaves as if he maximizes the expected value of an utility function.
-On the other hand, if the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if he had in mind a \textit{subjective}\footnote{Subjective here is a technical term meaning that the function is internal for the agent.} probability distribution, a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and utility function.
+On the other hand, if the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if he had in mind a \textit{subjective}\footnote{Subjective here is a technical term meaning that the function is internal for the agent.} probability distribution and a utility function, and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and utility function.
 
 Both von Neumann and Savage's Decision Under Uncertainty theories are known as the \textit{classic} theories in other areas, mostly in Economics.
-It is important to say that other Decision Making theories exist, such as Prospect Theory (\cite{kahneman1979prospect}), Case-Based Decision Theory (\cite{gilboa1995case}) among others that are out of the scope of this work.
-
-For what remains of this section, we follow \cite{bernardo2000bayesian} in his exposition of Savage's theory since it is the most suitable for our purposes because we will be considering decision makers that do not know all of their environment.
+It is important to say that other Decision Making theories exist, such as Prospect Theory (\cite{kahneman1979prospect}), Case-Based Decision Theory (\cite{gilboa1995case}), among others that are out of the scope of this work.
+
+For what remains of this section, we follow \cite{bernardo2000bayesian} in his exposition of Savage's theory since it is the most suitable for our purposes, because we will be considering decision makers that do not know all of their environment.
 Savage's Decision Making theory, together with von Neumann's theory is what it is commonly known as \textit{classical} decision making.
@@
 	\end{proof}
-	This result means that for any decision maker whose preferences satisfies axioms 1 through 5 the only election criteria that is compatible with the axioms is the maximization of expected utility and thus establishes a \textit{normative} criteria for decision making for rational agents.
+	This result means that for any decision maker whose preferences satisfies axioms 1 through 5, the only election criteria that is compatible with the axioms is the maximization of expected utility and thus establishes a \textit{normative} criteria for decision making for rational agents.
 This result establishes a complete ordering of the options, but it does not guarantees the existence of an option for which the expected utility is maximum and further mathematical assumptions are required over the utility function $u$ in order to guarantee the existence of the maxmimum utility option (\cite{bernardo2000bayesian}).
@@
 	Notice that the expectations are taken with respect to the subjective probability measure, and in fact Savage's general result as stated in \cite{gilboa2009decision} guarantees the existence of a utility function and a probability measure such that the preference relation is represented by the expected utility of that function with respect to that probability measure.
-Here we defined the subjective probability directly from the coherence axioms.
+Here we define the subjective probability directly from the coherence axioms.
 
@@
 Game Theory is an area of Mathematics which is used to model games.
-The basic assumptions of the theory is that each decision maker is rational, and they take into account their knowledge or expectations of other decision makers behavior.
-
-	In this section we will review the basic models of Game Theory that are required to this thesis proposal, which are the normal-form game, and the extensive-form game. 
+The basic assumptions of the theory is that each decision maker is rational, and that they take into account their knowledge or expectations of other decision makers behavior.
+
+	In this section we will review the basic models of Game Theory that are required for this thesis proposal, which are the normal-form game, and the extensive-form game. 
 	\subsubsection{Building blocks}{\label{basic_models}}
@@
 	\item $N$ is a finite set of players.
-	\item $A=A_1 \times ... \times A_n$ where each $A_i$ is the set of available actions for player $i$. Each $a=(a_1,...,a_n) \in A$ is called an action profile.
+	\item $A=A_1 \times ... \times A_n$, where each $A_i$ is the set of available actions for player $i$. Each $a=(a_1,...,a_n) \in A$ is called an action profile.
 	\item For each player $i \in N$, a preference relation $\succeq_i$ defined over $A$.

### 2018-09-02 Propuesta/Formato/Proposal_format.tex
correcciones

@@
 Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
-, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
+, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}).
+To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
 It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ (\cite{pearl2018why}).
+In fact, using the \textit{Do} operator we give a definition of a variable $X$ causing $Y$.
+	\begin{defi}
+	Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
+	\[ P_\mathcal{G}(Y | do(X=x) \neq P_\mathcal{G}(Y | do(X=x'). \]
+	\end{defi}
 
@@
 	\subsection{Research Question}
-How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
-How can a causal structure help a decision maker in order to guide his learning process?
-What does the rationality assumption implies about how to choose when considering causal information?
-How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
+The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
+\begin{enumerate}
+\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
+\item How can a causal structure help a decision maker in order to guide his learning process?
+\item What does the rationality assumption implies about how to choose when considering causal information? 
+\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
+\end{enumerate}
 	\subsection{Hypothesis}
@@
 A very well known model-free method is the classic Q-learning algorithm developed by \cite{watkins1992q}.
-In either case, model-free or model-based methods an agent must use information from transitions or from observations.
+In either case, model-free or model-based methods, an agent must use information from transitions or from observations.
 For this reason, Reinforcement Learning uses only associative patterns for finding an optimal policy and this methods can not learn beyond what is expressable by an MDP, which is constrained into associative rules and can not express any other kind of structure, such as causal relations.
@@
 
-The formation and use of causal knowledge in human beings can provide inspiration for our problem, and it has been extensively studied by York Hagmayer and his team They have found that human beings, while facing a sequential decision problem, use and modify causal information (\cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2010spontaneous}, \cite{hagmayer2013repeated}).
+The formation and use of causal knowledge in human beings can provide inspiration for our problem, and it has been extensively studied by York Hagmayer and his team.
+They have found that human beings, while facing a sequential decision problem, use and modify causal information (\cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2010spontaneous}, \cite{hagmayer2013repeated}).
 In their studies, all of the subjects were faced with a one-stage decision problem in which an initial causal model was provided.
@@
 A bandit problem consists by a set of available actions (the arms of the slot machine) and a reward function.
-It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}) if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
-On the other hand, if the largest mean reward is desired with a fixed confidence of $\delta$ then an algorithm for finding the optimal action exist in such a way that the number of rounds to find it is within a constant factor of a lower bound (\cite{jamieson2014lil}).
+It has been shown that the probability of the action selected in the $n$-th round not being optimal has an upper bound (\cite{audibert2010best}), if the optimality criteria is choosen to be the \textit{minimal regret} which is basically the difference of the theoretical mean reward of the optimal action and the chosen action.
+On the other hand, if the largest mean reward is desired with a fixed confidence of $\delta$, then an algorithm for finding the optimal action exist in such a way that the number of rounds to find it is within a constant factor of a lower bound (\cite{jamieson2014lil}).
 If one does not consider a fixed confidence but instead a fixed number of learning rounds (known as fixed budget), then we face a \textit{fixed budget} problem where a lower bound for the probability of finding the optimal arm is given by \cite{carpentier2016tight}.
+For the case of \textit{unobserved confounders} \cite{bareinboim2015bandits} propose a new metric for bandit algorithms to optimize which achieves better experimental performance.
 
 Notice that none of the previous works consider any kind of causal structure.
-For this kind of problems, causal relations between actions and consequences could be considered, in this way an action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world (\cite{hagmayer2009decision}) and that human beings use and modify causal knowledge during a sequential decision making process (\cite{hagmayer2013repeated}).
+For this kind of problems, causal relations between actions and consequences could be considered, in this way an action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world (\cite{hagmayer2009decision}); they use and modify causal knowledge during a sequential decision making process (\cite{hagmayer2013repeated}).
 Consider a set of variables that are causally related between them and to a \textit{reward} variable whose value has to be maximized by intervening one of the variables.
@@
 They show that the achieved regret is of a smaller order than the regret obtained by non-causal algorithms.
-
-Discovering the causal model itself while using current knowledge to make choices is left as future work in both \cite{lattimoreNIPS2016} and \cite{sen2017identifying}, but algorithms to discover causal relations in data can be found in \cite{eberhardt2008almost}, \cite{hauser2012two}, \cite{hyttinen2013experiment}, \cite{loh2014high}, \cite{shanmugam2015learning} \cite{mooij2016distinguishing}.
+Discovering the causal model itself while using current knowledge to make choices is left as future work in both \cite{lattimoreNIPS2016} and \cite{sen2017identifying}.
+
+Algorithms to discover causal relations in data can be found in \cite{spirtes2000causation}, \cite{eberhardt2008almost}, \cite{hauser2012two}, \cite{hyttinen2013experiment}, \cite{loh2014high}, \cite{shanmugam2015learning} \cite{mooij2016distinguishing}.
+The most famous of them being the PC-Algorithm (\cite{spirtes2000causation}) which is a constraint-based method for causal discovery, and has an exponential worst-case running time.
+The PC-Algorithm is based in the Inductive Causation Algorithm proposed in (\cite{judea1990equivalence}).
+In \cite{hauser2012two} an \textit{active learning} approach to prove a conjecture stated in \cite{eberhardt2008causal} about the worst-case number of experiments required to uniquely discover a causal graph (up to Markov equivalencies)
+
 Also, an \textit{active learning} approach can be used to learn causal models from data, where one starts with some initial graph and then select the instances in the data that allow to add and orient edges in order to end with a fully oriented graph.
@@
 
+In this research we are looking for a methodology that will allow to discover causal structure from an uncertain environment while also using it to obtain a reward.
+Being able to discover the causal structure at the same time that it is being used will allow to eliminate the restriction of knowing the full causal model or part of it.
 \newpage
@@
 \subsection{Belief Formation}
-Once a Game that models agent-environment interaction is defined, we must turn to the question about \textit{learning} about the environment, which includes acquiring causal information and also using current causal knowledge in order to obtain new information.
+Once a Game which models the agent-environment interaction is defined, we must turn to the issue about \textit{learning} about the environment, which includes acquiring causal information and also using current causal knowledge in order to obtain new information.
 
@@
 The agent, at the beginning of the game is uncertain about the state of the environment which he is facing, but he still must take an action, so in order to allow the decision maker to use his knowledge to act, he will have \textit{beliefs} about any relevant aspects of the environment which he will use as if they were the true model in any given play.
-These beliefs will be \textit{updated} according to what is observed at the end of each game.
+These beliefs will be \textit{updated} according to what is observed at the end of each move.
 The second step is to find a way to specify beliefs about causal structure.
@@
 The three cases that we notice are:
-\begin{itemize}
+\begin{enumerate}
 \item The decision maker knows the causal model.
@@
 \item The decision maker knows nothing about the causal model. 
-\end{itemize}
+\end{enumerate}
 In the following we show how the first two cases, which are particular cases of our general problem, can be solved and we show that the proposed solution for the first two cases sheds light on how the general case could be solved.
@@
 If a decision maker knows the structure of the graph, but he does not know the parameters, then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
-In the preliminary results that will be shown later a particular case is considered, where the beliefs have the form of a Dirichlet distribution for each of the Conditional Probability Tables that characterize a Bayesian Network.
+In the preliminary results, which will be shown later, a particular case is considered, where the beliefs have the form of a Dirichlet distribution for each of the Conditional Probability Tables that characterize a Bayesian Network.
 
@@
 \subsubsection{General case: Model unknown}
-The previous two cases are per se steps of the methodology, rather simplifications in order to understand what is happening. The real deal is the case where the causal model is unknown.
-
+The previous two cases are per se steps of the methodology, rather simplifications in order to understand what is happening.
+The real deal is the case where the causal model is unknown.
 When the causal model is unknown, two cases of interest appear, depending if the decision maker knows the (maximum) number of variables in the model or not.
@@
 
-Given a number $T$ of rounds, and inside round $t \in \{1,2,...,T \}$ the agent will make an action $a \in \mathcal{A}$ and then observe the state of the other variables of the model.
+Given a number $T$ of rounds, and in round $t \in \{1,2,...,T \}$ the agent will make an action $a \in \mathcal{A}$ and then observe the state of the other variables of the model.
 Using this information, current beliefs will be updated according to the criteria to be established.
@@
 		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
+		\item Possible Forum: Conference on Uncertainty in Artificial Intelligence (UAI), International Conference Machine Learning (ICML)
 		\end{itemize}
@@
 	\item When a causal model is unknown, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
+	\item Possible Forum: International Conference on Bayesian Nonparametrics, International Conference on Probabilistic Graphical Models.
 	\end{itemize}
@@
 
-\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1},\]
+\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1}\]
 
@@
 \subsubsection{Implementation}
-We begin with a random assignation of the $\alpha$ parameter for each of the distributions considered.
+We begin with a random assignment of the $\alpha$ parameter for each of the distributions considered.
 We use Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
@@
 
-The average perform over $20, 50, 100$ and $200$ rounds are shown.
+The average performance over $20, 50, 100$ and $200$ rounds is shown in Figure 2 - 5.
 

### 2018-10-24 Propuesta/Presentacion/Proposal_Presentation.tex
La presentacion de SeminarioTIA

\documentclass{beamer}
\mode<presentation> {

\usetheme{Madrid}

}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{natbib}
\theoremstyle{plain}
\newtheorem{teo}{Teorema}
\newtheorem{prop}[teo]{Proposición}
\newtheorem{defi}[teo]{Definición}
\newtheorem{obs}[teo]{Observación}
\newtheorem{lem}[teo]{Lema}
\newtheorem{cor}[teo]{Corolario}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Mauricio Gonzalez Soto} % Your name
\institute[INAOE] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Instituto Nacional de Astrofísica Óptica y Electrónica \\ % Your institution for the title page
\medskip
\textit{mauricio@inaoep.mx} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%--------SLIDES------------------------------

\section{Introduction}
\subsection{General Overview}
\begin{frame}
\frametitle{General Idea}
\begin{itemize}
\item We live in a causally structured world. 
\item Causal relations allow us to plan and to reason.
\item Causal reasoning allows for \textit{what...if...} claims.
\item In order to engage in causal reasoning we must be able to \textit{imagine} alternative worlds. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Engaging in a Causal Environment}
\begin{itemize}
\item If Causal relations allow for all of this, why don't we use them?
\item In fact, we do, although not necessarily we are aware of this. 
\item \cite{hagmayer2013repeated} show how human beings use and modify causal information.
\item \cite{hohwy2013predictive}, \cite{clark2015surfing}, asserts that the brain itself is a causal hypothesis tester. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our Problem}
\begin{itemize}
\item Engaging with the world has many dimensions: perception, action, agency, etc.
\item We consider only the decision making process where consequences are uncertain.
\item Consequences and actions are causally related.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our setting}
\begin{itemize}
\item Consider a \textit{rational} decision maker in an uncertain environment.
\item This is, consequences can not be predicted, which is equivalent to saying that making the same choice does not has the same consequence.
\item Examples: stock picking.
\item We add an extra to the environment: causal structure. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our Proposal}
\begin{itemize}
\item Since decision-making theories do not consider causal structure in the decision making processes
\item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Rationality and Causality}
\begin{itemize}
\item Rationality is defined in terms of \textit{axioms of coherence}.
\item From the axioms, a formal criteria of choice is obtained: \textbf{maximization of expected utility}. 
\item How is the awareness of causal structure and the rationality assumption related?
\item \cite{joyce1999foundations} stated that a decision maker should consider effects of of decisions, although he didn't said how. 
\item We want to answer how to maximize expected utility in terms of a causal model. 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Our idea}
\begin{itemize}
\item A decision maker who is aware of the causal nature of his environment must use those relations to predict outcomes and take the actions that renders a good (prefered) consequence. 
\item The agent will use current knowledge in order to take best possible action and then update causal \textit{beliefs} after observing outcome.
\item Learning causal information by updating beliefs while interacting with environment.
\end{itemize}
\end{frame}

\subsection{Outline}
\begin{frame}
\frametitle{Outline}
\begin{itemize}
\item We now have two main issues to talk about: Causality and Decision Making. 
\item Once we have set the language, then we will state the problem we are trying to attack.
\item Remember, if you have an idea, then probably someone else already had it: related work.
\item Our attempt, and some results that we have stumbled upon. 
\end{itemize}
\end{frame}

\section{Framework and Main Concepts}
	\subsection{Causality}
	\begin{frame}
		\frametitle{What is causality?}
		\begin{itemize}
		\item We want to think about the causes of certain events; i.e., a disease, an accident. 
		\item Causal claims are associated with what didn't happen. 
		\end{itemize}
	\end{frame}
		\begin{frame}
		\frametitle{Nature of Causality}
		\begin{itemize}
		\item Do causal relations \textit{really} exist? 
		\item Or are they only a mental construct useful for managing certain regularities in our world?
		\item We prefer the former idea. 
		\end{itemize}
		\end{frame}

		\begin{frame}
		\frametitle{Difference between causal and associative relations}
		\begin{itemize}
		\item With associative relations we can't figure out what influenced (or caused) what.
		\item Consider this: A distribution $p(a,t)$ can be factored both as $p(a|t)p(t)$ and $p(t|a)p(a)$. 
		\item In terms of a Bayesian Network, $T \to A$, and $A \to T$, but both can't be causal.
		\item Causal relations need not be in temporal order: consider a barometer and a storm.
		\end{itemize}
		\end{frame}

		\begin{frame}
		\frametitle{The Ladder of Causality (\cite{pearlwhy})}
		Pearl, in his recent book \cite{pearl2018why} considers three levels of scientific inquiry.
		\begin{enumerate}
		\item Observing
		\item Intervening
		\item Counterfactual reasoning
		\end{enumerate}
		\end{frame}

		\begin{frame}
		\frametitle{Definition of Causality}
		The working definition of Causality that we will be using is based upon \cite{spirtes2000causation} and \cite{pearl2009causality}.
		\end{frame}

		\begin{frame}
		\frametitle{Formal Definition}
		\begin{defi}
		Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite, probability space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
		\begin{itemize}
		\item Transitive: if $A \to B$ and $B \to C$ for $A,B,C \in \mathcal{F}$ then $A \to C$.
		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
		\item Antisymetric: For $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
		\end{itemize}
		If $A \to B$ we will say that $A$ is a cause of $B$.
		\end{defi}
		\end{frame}
		\subsubsection{Causal Graph Construction}
		\begin{frame}
		\frametitle{From Causes to Graphs}
		For each pair $A,B \in \mathcal{F}$ such that $A \to B$ and $A$ is a direct cause of $B$, define a node representing each variable, and a directed edge joining them.
		\begin{prop}
		Given a binary relation “$\to$” that satisfies previous requirement, let $\mathcal{G}$ be the directed graph that is obtained by the previous procedure. Then, $\mathcal{G}$ contains no cycles.
		\end{prop}
		\begin{proof}
		Consider a cycle $A \to C_1 \to C_2 \to \cdots \to A$, then we would have that $A \to A$, which contradicts the Irreflexive property.
		\end{proof}
		\end{frame}

		\begin{frame}
		\frametitle{Endogenous and Exogenous variables}
		\begin{obs}
		Notice that if the graph is finite, then necesarily there will be nodes without parents, which in the context of causal models will be called exogenous variables.
		\end{obs}
		\end{frame}

		\begin{frame}
		\frametitle{Causal Sufficiency}
		A requirement needed for the following construction, known as Causal Sufficiency, says that any cause of any event in the model is in the model. 
		\end{frame}

		\subsubsection{Probability Measure built from the Graph}
		\begin{frame}
		\frametitle{from Graphs to Probability Measures}
		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
        \begin{itemize}
        \item Causal Markov Condition.
        \item Causal Minimality.
         \item Causal Faithfulness.
        \end{itemize}
         This requirements will be treated in an axiomatic fashion.
		\end{frame}

	\subsection{Causal Graphical Models}
	\begin{frame}
	\frametitle{Causal Graphical Models}
	\begin{defi}
	A causal graphical model consists of:
	\begin{itemize}
	 \item A set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$.
	 \item A directed acyclic graph (DAG) $\mathcal{G}$.
	 \item An operator $do()$ defined over the space of DAG's whose action consists of: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ... , x_{i_j} \} \in Val(\mathcal{X})$, then $do(\mathbf{X} = \mathbf{x} )$ assigns to each $X_j \in \mathbf{X}$ the value $x_{i_j}$ and deletes any incoming edge on it.
	 \end{itemize}
	 \end{defi}
	 \end{frame}

	 \begin{frame}
	 \frametitle{Causation in terms of the Do operator}
	 \begin{itemize}
	 \item We can use the Do operator to provide a definition of a cause:
	 \begin{defi}
	 We say that $X$ \textit{causes} $Y$ if 
	\[ P_\mathcal{G}(Y | do(X=x)) \neq P_\mathcal{G}(Y | do(X=x')). \]
	 \end{defi}
	 \end{itemize}
	 \end{frame}

	\subsection{Decision Theory}
	\begin{frame}
	\frametitle{Decision Problems}
	A decision problem is composed by:
	\begin{itemize}
	\item A decision maker.
	\item A set of available options $\mathcal{A}$.
	\item A preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
	\item A set of uncertain events $\mathcal{E}$.
    \item A set of possible outcomes $\mathcal{C}$
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Rationality}
	Rationality is a subtle concept with intrinsec philosophical implications. We will understand by \textit{rational} a decision maker whose \textit{preferences} are some how logically consistent.
	In its most basic form:
	\begin{itemize}
	\item The decision maker needs to be able to strictly prefer at least between one action over other.
	\item Transitive preferences.
	\item If $a$ is always prefered over $b$, then it doesn't matter what uncertain event occurs.
	\end{itemize}
	\end{frame}

	\begin{frame}
	\frametitle{Expected Utility Maximizaton.}
	We are considering \textit{bounded decision problems};i.e., where a \textit{best} and \textit{worst} outcomes exist. In that case,
	\begin{teo}
	For any bounded decision problem with extreme consequences $c^\ast \succ c_\ast$ and any $G \succ \emptyset$
	\[ a_1 \succeq_G a_2 \Leftrightarrow \bar{u}(a_1) \geq \bar{u}(a_2). \]
	\end{teo}
	\begin{proof}
	See \cite{bernardo2000bayesian} Chapter 2 “Foundations”, Proposition 2.22.
	\end{proof}
	\end{frame}

\section{Research Problem}
\begin{frame}
\frametitle{Section:Research Problem}
In this section we use the mathematical formalism described above in order to formally enounce the problem we are trying to solve here and the questions that motivated that problem. 
\end{frame}

	\subsection{Research Questions}
	\begin{frame}
	\frametitle{Research Questions}
	The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
	\begin{enumerate}
	\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good    	choices? 
	\item How can a causal structure help a decision maker in order to guide his learning process? 
	\item What does the rationality assumption implies about how to choose when considering causal information? 
\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
\end{enumerate}
\end{frame}
	\subsection{Statement}
	\begin{frame}
	\frametitle{Problem Statement}
	\begin{itemize}
	\item Let  $\mathcal{G}$ a causal graphical model.
	\item Let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. 
	\item Consider a rational decision maker who doesn't know the causal model which control the probabilities of observing a consequence given an action $a \in \mathcal{A}$.
	\item We ask how the decision maker could learn about the causal structure that controls his environment in order to make good choices with respect to the decision makers' preferences.
	\end{itemize}
	\end{frame}
	\subsection{Motivation}
	\begin{frame}
	\frametitle{Motivation}
	\begin{itemize}
	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
	\item What about not knowing the causal model? 
	\end{itemize}
	\end{frame}
	\subsection{Justification}
	\begin{frame}
	\frametitle{Justification}
	\begin{itemize}
	\item Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data.
	\item Current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
	\item This is highly relevant in real-world applications such as self-driving cars.
	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
	\end{itemize}
	\end{frame}
	\begin{frame}
	\begin{itemize}
	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain. 
	\item It allows to ask \textit{why}.
		\end{itemize}
	\end{frame}
	\subsection{Hypothesis}
	\begin{frame}
	\frametitle{Hypothesis}
	\begin{itemize}
	\item Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
	\item Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn a causal model of the environment and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
	\end{itemize}
	\end{frame}
	\subsection{Objectives}
	\begin{frame}
	\frametitle{General Objectives}
	The proposed research has as a general objective:
	\begin{itemize}
 \item to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
	\end{itemize}
	\end{frame}

\section{Approaches and Related Work}
\subsection{Approaches and Related Work: Reinforcement Learinng}
\begin{frame}
\frametitle{Reinforcement Learning (RL)}
\begin{itemize}
\item Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction.
\item  In this setting it is assumed that a Markov Decision Process (MDP) is able to capture the effects of the agent's actions in the environment.
\item Reinforcement Learning uses only associative patterns for finding an optimal policy and this methods can not learn beyond what is expressable by an MDP, which is constrained into associative rules and can not express any other kind of structure, such as causal relations. 
\item it can be shown that optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
\item In this way, when only associative information is available, RL is a coherent learning framework. On the other hand, it remains to be answered how to incorporate a causal model into an on-line decision making algorithm.
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Causal learning in humans}
\begin{frame}
\frametitle{Causal Learning in Human beings}
\begin{itemize}
\item Human beings, while facing a sequential decision problem, use and modify causal information.
\item It is also known that human beings focus on \textit{local} aspects while learning causal relations that are later unified into a single structure.
\item  Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal structure when little prior information is available. 
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Joyce's Causal Decision Theory}
\begin{frame}
\frametitle{Joyce's Causal Decision Theory}
\begin{itemize}
\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
\item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
\item Although he captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker.
\end{itemize}
\end{frame}
\subsection{Approaches and Related Work: Bandits}
\begin{frame}
\frametitle{Bandits}
\begin{itemize}
\item In the class of decision problems considered for this proposal it is assumed that the a rational agent must choose some available action and then receive a reward from it.
\item Problems of this nature have been modelled as \textit{bandit} problems.
\item A bandit is an analogy of a \textit{slot machine}.
\item Best-arm algorithms exist: none consider causal information.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Causal bandits}
\begin{itemize}
\item n action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world
\item they use and modify causal knowledge during a sequential decision making process.
\item It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
\item Their work requires that the causal model is fully known to the decision maker.
\item this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
\end{itemize}
\end{frame}

\section{Proposed Solution and Methodology}
\subsection{Proposed Solution}
\begin{frame}
\frametitle{Proposed Solution}
\begin{itemize}
\item Rational agent in a causal environment.
\item Modelling agent-environment interaction as a \textit{game}.
\item Encoding current causal knowledge as probabilsitic \textit{beliefs}
\item Using current knowledge to make a choice.
\item Updating causal information.
\end{itemize}
\end{frame}
\subsection{A guiding principle}
\begin{frame}
\frametitle{A guiding principle for Causal Decision Problems}
We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational. We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
\end{frame}
\begin{frame}
\begin{prop}
Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent. Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
\end{prop}
\end{frame}

\subsection{Modelling a Causal Decision Problem as a Game}
\begin{frame}
\frametitle{Games}
\begin{itemize}
\item Since the observed outcomes, given an action, are guided by the Causal Model, we can think of the environment as a decision maker whose available actions depend on what has been chosen previously by the original decision maker.
\item Game Theory allows to model situations where two or more decision makers interact.
\item In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
\begin{enumerate}
\item the (original) decision maker, 
\item and a player which will be called Nature, whose actions are to be guided by the causal model.
\end{enumerate}
\end{itemize}
\end{frame}
\begin{frame}
\begin{itemize}
\item  Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
\item After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games
\end{itemize}
\end{frame}

\subsection{Belief updating}
\begin{frame}
\frametitle{Belief Formation and Updating}
\begin{itemize}
\item The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
\item The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world
\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
\item it remains unanswered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
\end{itemize}
\end{frame}

\subsection{Solving first simpler cases}
\subsubsection{Fully knowing the causal graphical model}
\begin{frame}
\frametitle{Fully knowing the causal graphical model}
\begin{itemize}
\item If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
\item Choosing the action that is more likely to produce the most desired outcome according to current causal knowledge is required by Joyce's Causal Decision Theory (\cite{joyce1999foundations}).
\end{itemize}
\end{frame}
\subsubsection{Knowing only the graph structure}
\begin{frame}
\frametitle{Knowing only the causal graph structure}
\begin{itemize}
\item If a decision maker knows the structure of the graph, but he does not know the parameters,
\item then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
\item In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e. in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
\item This means that using current beliefs, the agent forms a causal model which is used as if it were the true causal model.
\end{itemize}
\end{frame}
\subsubsection{General case: Model unknown}
\begin{frame}
\frametitle{Model unknown}
If the model is completely unknown, two ideas appear:
\begin{itemize}
\item If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
\item For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
\item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables. In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.  
\end{itemize}
\end{frame}

\section{Some Results}
\begin{frame}
\frametitle{Some results}
To show the factibility of this proposal, we considered a test scenario and two cases in ascending difficulty.
\end{frame}
\subsection{Test scenario}
\begin{frame}
\frametitle{Test Scenario}
Consider a sick patient who arrives at a hospital and he either has disease $A$ or disease $B$. The doctor can either give him some pill or send him into surgery.  Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally. The doctor could be facing a mutation from a known disease, so she has some knowledge about what could happen if a treatment is given to the patient. Using her previous knowledge as a true model, she can choose a treatment and observe the outcome from which she will learn about this disease, so she could make a better decision the next time a similar patient arrives.
\end{frame}

\begin{frame}
\frametitle{Causal model}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=0.7\textwidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/causal_graph.png}}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Variables of the Model}
The variables in the model are: 
\begin{itemize}
\item \textbf{Disease:} Either $A$ or $B$.
\item \textbf{Treatment:} Either pill or surgery.
\item \textbf{Reaction:} Either dying or surviving.
\item \textbf{Lives:} Either living or dying.
\end{itemize}
The variable \textit{Lives} is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the variable \textit{Treatment}. The decision maker prefers an outcome in which the patient lives.
\end{frame}

\begin{frame}
For this test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
\[ P(Y | do(Tr))=P(Y | D, Tr, R)P(R | Tr) P(D). \]
\end{frame}

\subsection{Case 2: Only the structure is known}
\begin{frame}
\frametitle{Knowing only the graph structure}
\begin{itemize}
\item Given the structure of the model; i.e., the variables in it and the directed edges, the joint distribution of those variables can be expressed as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the underlying DAG in $\mathcal{G}$
\item Since these distributions fully characterize the model, the decision maker will have beliefs over each one of these parameters. Notice that each of these parameters is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
\end{itemize} 
\end{frame}

\begin{frame}
\begin{itemize}
\item A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}. The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by

\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1}\]
\item The decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
\item Using the agent's current beliefs, a causal graphical model can be specified. 
\item Using this fully specified (structure + parameters) as a true model, the decision maker will make his choice as according to the guiding principle.
\item When the decision maker observes the value of the target variable, he will update the parameters that specify his beliefs.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Belief Updating}
For the belief updating, given a new data point,  two cases must be considered:
\begin{itemize}
\item The variable to update has no parents.
\item The variable to update has parents.
\end{itemize}
\end{frame}
\begin{frame}
\frametitle{Belief updating}
\begin{itemize}
\item In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
\[ \textrm{Dirichlet}(\alpha + c) \]
where $c$ is a vector of the number of occurrences of that observed data point. 
\item For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables. Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed. In this case, if the prior of $X_i$ conditioned on its parents having the value $j$ is given by a a Dirichlet($\alpha$), then the posterior for the variable $X_i$ given an observed data point is given by 
\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
\end{itemize}
\end{frame}

\subsection{Implementation}
\begin{frame}
\frametitle{Implementation}
\begin{itemize}
\item We begin with a random assignment of the $\alpha$ parameter for each of the distributions considered.
\item  We use Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
\item With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable.
\item With this action chosen, we simulate an outcome from the causal graphical model using the chosen action as an intervention.
\item This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
\end{itemize}
\end{frame}

\subsection{Experiments}
\begin{frame}
\frametitle{Experiments}
We show the results of two experiments. We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning (\cite{watkins1992q}). Q-learning was chosen since it learns an \textit{optimal policy} in the sense of the Bellman Equation (\cite{sutton1998reinforcement}) and it is shown in \cite{webb2007game} that such optimal policies also maximize expected utility. 
\end{frame}

\begin{frame}
\frametitle{20 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/20_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
\frametitle{50 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/50_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
\frametitle{100 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/100_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
\frametitle{200 rounds}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/200_rounds_format.png}
\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\end{frame}

\begin{frame}
More details in \cite{gonzalez2018playing}
\end{frame}

\section{Conclusions}
\begin{frame}
\frametitle{Conclusions}
We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism can make use of the causal relations that hold in such environment in order to find an action that will take him into the way of a desired result, in terms of his (rational) preferences. Also, causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.
\end{frame} 
\begin{frame}
We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives. Following this intuitions we proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality. From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order. In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment. We are considering. If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
\end{frame}

%------- REFERENCES ----------------------

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
\end{frame}

%------------------------------------------------

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

%------------------------------------------------

\end{document}

### 2018-11-24 Propuesta/Presentacion/Presentacion_Corta.tex
presentacion version sabado

--- 2018-10-24 Propuesta/Presentacion/Proposal_Presentation.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[spanish, mexico]{babel}
+%\usepackage[spanish, mexico]{babel}
 \usepackage{listings}
@@
 \newtheorem{teo}{Teorema}
-\newtheorem{prop}[teo]{Proposición}
-\newtheorem{defi}[teo]{Definición}
-\newtheorem{obs}[teo]{Observación}
-\newtheorem{lem}[teo]{Lema}
-\newtheorem{cor}[teo]{Corolario}
+\newtheorem{prop}[teo]{Proposition}
+\newtheorem{defi}[teo]{Definition}
+\newtheorem{obs}[teo]{Observation}
+\newtheorem{lem}[teo]{Lemma}
+\newtheorem{cor}[teo]{Corolary}
 \usepackage{tikz}
@@
 \item We live in a causally structured world. 
-\item Causal relations allow us to plan and to reason.
-\item Causal reasoning allows for \textit{what...if...} claims.
-\item In order to engage in causal reasoning we must be able to \textit{imagine} alternative worlds. 
+\item Our actions \textit{cause} changes in the environment.
+\item We consider this changes in order to plan our actions
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Engaging in a Causal Environment}
+\frametitle{Human use of Causal Relations}
 \begin{itemize}
-\item If Causal relations allow for all of this, why don't we use them?
-\item In fact, we do, although not necessarily we are aware of this. 
 \item \cite{hagmayer2013repeated} show how human beings use and modify causal information.
-\item \cite{hohwy2013predictive}, \cite{clark2015surfing}, asserts that the brain itself is a causal hypothesis tester. 
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Our Problem}
-\begin{itemize}
-\item Engaging with the world has many dimensions: perception, action, agency, etc.
-\item We consider only the decision making process where consequences are uncertain.
-\item Consequences and actions are causally related.
+\item Both \cite{hohwy2013predictive}, \cite{clark2015surfing}, assert that the brain itself is a causal hypothesis tester. 
 \end{itemize}
@@
 \item Consider a \textit{rational} decision maker in an uncertain environment.
-\item This is, consequences can not be predicted, which is equivalent to saying that making the same choice does not has the same consequence.
-\item Examples: stock picking.
+\item Consequences of actions can not be predicted.
 \item We add an extra to the environment: causal structure. 
@@
 \begin{frame}
+\frametitle{A brief on causal models}
+\begin{itemize}
+\item Causal models encode cause-effect relations. 
+\item Relations are encoded in a graph.
+\item Can be used to predict \textit{effects} of certain causes
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Causal Model: Example}
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/example.png}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+\end{frame}
+
+\begin{frame}
 \frametitle{Our Proposal}
 \begin{itemize}
-\item Since decision-making theories do not consider causal structure in the decision making processes
+\item Current decision-making theories do not consider causal structure in the decision making processes
 \item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
@@
 
+\section{Concepts}
+\subsection{Causal Models}
 \begin{frame}
-\frametitle{Rationality and Causality}
-\begin{itemize}
-\item Rationality is defined in terms of \textit{axioms of coherence}.
-\item From the axioms, a formal criteria of choice is obtained: \textbf{maximization of expected utility}. 
-\item How is the awareness of causal structure and the rationality assumption related?
-\item \cite{joyce1999foundations} stated that a decision maker should consider effects of of decisions, although he didn't said how. 
-\item We want to answer how to maximize expected utility in terms of a causal model. 
-\end{itemize}
+\frametitle{Concepts}
+Concepts and Framework
 \end{frame}
-
-\begin{frame}
-\frametitle{Our idea}
-\begin{itemize}
-\item A decision maker who is aware of the causal nature of his environment must use those relations to predict outcomes and take the actions that renders a good (prefered) consequence. 
-\item The agent will use current knowledge in order to take best possible action and then update causal \textit{beliefs} after observing outcome.
-\item Learning causal information by updating beliefs while interacting with environment.
-\end{itemize}
-\end{frame}
-
-\subsection{Outline}
-\begin{frame}
-\frametitle{Outline}
-\begin{itemize}
-\item We now have two main issues to talk about: Causality and Decision Making. 
-\item Once we have set the language, then we will state the problem we are trying to attack.
-\item Remember, if you have an idea, then probably someone else already had it: related work.
-\item Our attempt, and some results that we have stumbled upon. 
-\end{itemize}
-\end{frame}
-
-\section{Framework and Main Concepts}
-	\subsection{Causality}
-	\begin{frame}
-		\frametitle{What is causality?}
-		\begin{itemize}
-		\item We want to think about the causes of certain events; i.e., a disease, an accident. 
-		\item Causal claims are associated with what didn't happen. 
-		\end{itemize}
-	\end{frame}
-		\begin{frame}
-		\frametitle{Nature of Causality}
-		\begin{itemize}
-		\item Do causal relations \textit{really} exist? 
-		\item Or are they only a mental construct useful for managing certain regularities in our world?
-		\item We prefer the former idea. 
-		\end{itemize}
-		\end{frame}
-
 		\begin{frame}
@@
 		\begin{frame}
-		\frametitle{The Ladder of Causality (\cite{pearlwhy})}
-		Pearl, in his recent book \cite{pearl2018why} considers three levels of scientific inquiry.
-		\begin{enumerate}
-		\item Observing
-		\item Intervening
-		\item Counterfactual reasoning
-		\end{enumerate}
-		\end{frame}
-
-		\begin{frame}
-		\frametitle{Definition of Causality}
-		The working definition of Causality that we will be using is based upon \cite{spirtes2000causation} and \cite{pearl2009causality}.
-		\end{frame}
-
-		\begin{frame}
-		\frametitle{Formal Definition}
+		\frametitle{Formal Definition of Causality according to \cite{spirtes2000causation}}
 		\begin{defi}
@@
 		\end{frame}
-		\subsubsection{Causal Graph Construction}
-		\begin{frame}
-		\frametitle{From Causes to Graphs}
-		For each pair $A,B \in \mathcal{F}$ such that $A \to B$ and $A$ is a direct cause of $B$, define a node representing each variable, and a directed edge joining them.
-		\begin{prop}
-		Given a binary relation “$\to$” that satisfies previous requirement, let $\mathcal{G}$ be the directed graph that is obtained by the previous procedure. Then, $\mathcal{G}$ contains no cycles.
-		\end{prop}
-		\begin{proof}
-		Consider a cycle $A \to C_1 \to C_2 \to \cdots \to A$, then we would have that $A \to A$, which contradicts the Irreflexive property.
-		\end{proof}
-		\end{frame}
 
 		\begin{frame}
-		\frametitle{Endogenous and Exogenous variables}
-		\begin{obs}
-		Notice that if the graph is finite, then necesarily there will be nodes without parents, which in the context of causal models will be called exogenous variables.
-		\end{obs}
-		\end{frame}
-
-		\begin{frame}
-		\frametitle{Causal Sufficiency}
-		A requirement needed for the following construction, known as Causal Sufficiency, says that any cause of any event in the model is in the model. 
-		\end{frame}
-
-		\subsubsection{Probability Measure built from the Graph}
-		\begin{frame}
-		\frametitle{from Graphs to Probability Measures}
+		\frametitle{From Graphs to Probability Measures}
 		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
@@
          \item Causal Faithfulness.
+         \item Causal Sufficiency
         \end{itemize}
@@
 
-	\subsection{Causal Graphical Models}
 	\begin{frame}
@@
 	\begin{defi}
-	A causal graphical model consists of:
+	A causal graphical model (\cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{koller2003multi}, \cite{sucar2015probabilistic}) consists of:
 	\begin{itemize}
@@
 
-	 \begin{frame}
-	 \frametitle{Causation in terms of the Do operator}
-	 \begin{itemize}
-	 \item We can use the Do operator to provide a definition of a cause:
-	 \begin{defi}
-	 We say that $X$ \textit{causes} $Y$ if 
-	\[ P_\mathcal{G}(Y | do(X=x)) \neq P_\mathcal{G}(Y | do(X=x')). \]
-	 \end{defi}
-	 \end{itemize}
-	 \end{frame}
-
-	\subsection{Decision Theory}
+\subsection{Decision Problems}
 	\begin{frame}
 	\frametitle{Decision Problems}
-	A decision problem is composed by:
+	A decision problem is composed by (\cite{bernardo2000bayesian}):
 	\begin{itemize}
-	\item A decision maker.
+	\item A (rational) decision maker.
 	\item A set of available options $\mathcal{A}$.
@@
 	\begin{frame}
-	\frametitle{Rationality}
-	Rationality is a subtle concept with intrinsec philosophical implications. We will understand by \textit{rational} a decision maker whose \textit{preferences} are some how logically consistent.
-	In its most basic form:
+	\frametitle{Choosing rationally}
+	Savage's Theorem (\cite{savage1954the}) says that:
+\begin{teo}	
+	If a rational decision maker does not know the probabilities of uncertain events ocurring, then he must choose according to the expected utility with respect to a \textit{subjective} probability measure and a \textit{subjective} utility function.
+\end{teo}
+Details can be found at \cite{bernardo2000bayesian}, \cite{gilboa2009decision}.
+	\end{frame}
+
+\section{Research Problem}
+\begin{frame}
+About my Proposed Research
+\end{frame}
+	\subsection{Motivation}
+	\begin{frame}
+	\frametitle{Motivation: 1}
 	\begin{itemize}
-	\item The decision maker needs to be able to strictly prefer at least between one action over other.
-	\item Transitive preferences.
-	\item If $a$ is always prefered over $b$, then it doesn't matter what uncertain event occurs.
+	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
+	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
+	\end{itemize}
+	\end{frame}
+	\begin{frame}
+	\frametitle{Motivation: 2}
+	\begin{itemize}
+	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
+	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
+	\item What about not knowing the causal model? 
+	\end{itemize}
+	\end{frame}
+
+	\subsection{Justification}
+	\begin{frame}
+	\frametitle{Justification: Un-interpretable models}
+	\begin{itemize}
+	\item Many real-world applications of decision making are solved by \textit{associative} methods.
+	\item Associative methods capture only statistical patterns that are found in data.
+	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
 	\end{itemize}
@@
 	\begin{frame}
-	\frametitle{Expected Utility Maximizaton.}
-	We are considering \textit{bounded decision problems};i.e., where a \textit{best} and \textit{worst} outcomes exist. In that case,
-	\begin{teo}
-	For any bounded decision problem with extreme consequences $c^\ast \succ c_\ast$ and any $G \succ \emptyset$
-	\[ a_1 \succeq_G a_2 \Leftrightarrow \bar{u}(a_1) \geq \bar{u}(a_2). \]
-	\end{teo}
-	\begin{proof}
-	See \cite{bernardo2000bayesian} Chapter 2 “Foundations”, Proposition 2.22.
-	\end{proof}
+	\frametitle{Justification: Need for nterpretable decision-making models}
+	\begin{itemize}
+	\item Current methods for Decision Making are based in Reinforcement Learning, and specifically in Deep Reinforcement Learning. 
+	\item Although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
+	\item This is highly relevant in real-world applications such as self-driving cars.
+	\end{itemize}
 	\end{frame}
 
-\section{Research Problem}
 	\begin{frame}
-\frametitle{Section:Research Problem}
-In this section we use the mathematical formalism described above in order to formally enounce the problem we are trying to solve here and the questions that motivated that problem. 
+	\frametitle{Justification: Explaining why}
+	\begin{itemize}
+	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+	\item It allows to ask \textit{why}.
+	\end{itemize}
 	\end{frame}
@@
 \end{frame}
+
 \subsection{Statement}
@@
 	\end{frame}
-	\subsection{Motivation}
-	\begin{frame}
-	\frametitle{Motivation}
-	\begin{itemize}
-	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
-	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
-	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
-	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
-	\item What about not knowing the causal model? 
-	\end{itemize}
-	\end{frame}
-	\subsection{Justification}
-	\begin{frame}
-	\frametitle{Justification}
-	\begin{itemize}
-	\item Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data.
-	\item Current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
-	\item This is highly relevant in real-world applications such as self-driving cars.
-	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
-	\end{itemize}
-	\end{frame}
-	\begin{frame}
-	\begin{itemize}
-	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
-	\item It allows to ask \textit{why}.
-		\end{itemize}
-	\end{frame}
+
 	\subsection{Hypothesis}
@@
 	\end{frame}
-	\subsection{Objectives}
+
+\section{Approaches and Related Work}
 \begin{frame}
-	\frametitle{General Objectives}
-	The proposed research has as a general objective:
+\frametitle{Approaches and Related Work}
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Related_Work.png}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+\end{frame} 
+\subsection{Approaches and Related Work: Reinforcement Learning}
+\begin{frame}
+\frametitle{Reinforcement Learning (RL)}
 \begin{itemize}
- \item to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
+\item Reinforcement Learning uses only associative patterns for finding an optimal policies.
+\item Optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
+\item In this way, when only associative information is available, RL is a coherent learning framework. 
 \end{itemize}
@@
 
-\section{Approaches and Related Work}
-\subsection{Approaches and Related Work: Reinforcement Learinng}
-\begin{frame}
-\frametitle{Reinforcement Learning (RL)}
-\begin{itemize}
-\item Reinforcement Learning (RL) has been the standard setting for decision-making and learning by interaction.
-\item  In this setting it is assumed that a Markov Decision Process (MDP) is able to capture the effects of the agent's actions in the environment.
-\item Reinforcement Learning uses only associative patterns for finding an optimal policy and this methods can not learn beyond what is expressable by an MDP, which is constrained into associative rules and can not express any other kind of structure, such as causal relations.
-\item it can be shown that optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
-\item In this way, when only associative information is available, RL is a coherent learning framework.
-On the other hand, it remains to be answered how to incorporate a causal model into an on-line decision making algorithm.
-\end{itemize}
-\end{frame}
 \subsection{Approaches and Related Work: Causal learning in humans}
@@
 \item Human beings, while facing a sequential decision problem, use and modify causal information.
-\item It is also known that human beings focus on \textit{local} aspects while learning causal relations that are later unified into a single structure.
-\item  Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal structure when little prior information is available.
+\item \cite{hagmayer2009decision}, \cite{hagmayer2013repeated} show that human beings can learn causal information in simple decision-making tasks, but remains open to study causal learning while seeking for a specific objective.
+\item 
 \end{itemize}
@@
 \item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
-\item Although he captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Bandits}
+\frametitle{Bandits: Non-causal}
 \begin{itemize}
@@
 \item Problems of this nature have been modelled as \textit{bandit} problems.
-\item A bandit is an analogy of a \textit{slot machine}.
 \item Best-arm algorithms exist: none consider causal information.
@@
 \begin{frame}
-\frametitle{Causal bandits}
+\frametitle{Bandits: Causal bandits}
 \begin{itemize}
-\item n action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world
-\item they use and modify causal knowledge during a sequential decision making process.
+\item An action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world
+\item They use and modify causal knowledge during a sequential decision making process.
 \item It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
@@
 \section{Proposed Solution and Methodology}
-\subsection{Proposed Solution}
 \begin{frame}
-\frametitle{Proposed Solution}
+\frametitle{Proposed Solution and Methodology}
+\end{frame}
+	\subsection{Objectives}
+	\begin{frame}
+	\frametitle{General Objectives}
+	The proposed research has as a general objective:
+	\begin{itemize}
+ \item Understand the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
+	\end{itemize}
+	\end{frame}
+
+\subsection{Methodology}
+\begin{frame}
+\frametitle{Proposed Solution: Conceptual step-by-step methodology}
 \begin{itemize}
@@
 \end{frame}
-\subsection{A guiding principle}
+
+\subsection{Validation}
+\begin{frame}
+\frametitle{Validation}
+How to know if objective has been achieved?
+\end{frame}
+
+\begin{frame}
+\frametitle{Concurrent Validation}
+We will be comparing against standard Reinforcement Learning methods such as Q-Learning (\cite{watkins1992q}) since it has been shown that optimal policies achieve the objective of Max Expected Utility.
+\end{frame}
+
+\begin{frame}
+\frametitle{A conjecture}
+It needs to be shown that given the rewards obtained by a decision-making procedure that achieves the MEU, then my proposed algorithm must stay close to it.
+\begin{teo}
+Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$.
+\end{teo}
+\begin{proof}
+To be proved.
+\end{proof}
+\end{frame}
+
+\begin{frame}
+In this case, if we succeed in finding an action whose reward equals that of RL algorithms, then we will have chosen the optimal action using Causal Information as proposed in the Hypothesis.
+\end{frame}
+
+\begin{frame}
+\frametitle{How to}
+How to
+\end{frame}
+
+\subsection{Proposed Solutuion: A guiding principle}
 \begin{frame}
@@
 We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
+It remains to be refined and extended for a general case.
 \end{frame}
+
 \begin{frame}
@@
 
-\subsection{Modelling a Causal Decision Problem as a Game}
+\subsection{First Step: Modelling a Causal Decision Problem as a Game}
 \begin{frame}
@@
 \begin{itemize}
-\item Since the observed outcomes, given an action, are guided by the Causal Model, we can think of the environment as a decision maker whose available actions depend on what has been chosen previously by the original decision maker.
-\item Game Theory allows to model situations where two or more decision makers interact.
 \item In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
@@
 \end{frame}
+
 \begin{frame}
@@
 
-\subsection{Belief updating}
+\subsection{Second Step: Belief formation}
 \begin{frame}
-\frametitle{Belief Formation and Updating}
+\frametitle{Belief Formation}
 \begin{itemize}
@@
 \item The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world
-\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
-\item it remains unanswered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
 \end{itemize}
@@
 
-\subsection{Solving first simpler cases}
-\subsubsection{Fully knowing the causal graphical model}
+\subsection{Third Step: Making choices}
+\begin{frame}
+\frametitle{Making Choices}
+\begin{itemize}
+\item The decision maker will use his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
+\item With this information he will choose the action that maximize the Expected Utility, according with our guiding principle.
+\end{itemize}
+\end{frame}
+
+\subsection{Third Step: Belief updating}
+\begin{frame}
+\frametitle{Belief Updating}
+\begin{itemize}
+\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
+\item it remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
+\end{itemize}
+\end{frame}
+
+\section{Methodology divided into simpler cases}
+\begin{frame}
+We will show how the methodology applies in simpler and easier cases where the decision maker has some exta info about his environment.
+\end{frame}
+\subsection{Fully knowing the causal graphical model}
 \begin{frame}
@@
 \item If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
-\item Choosing the action that is more likely to produce the most desired outcome according to current causal knowledge is required by Joyce's Causal Decision Theory (\cite{joyce1999foundations}).
 \end{itemize}
 \end{frame}
-\subsubsection{Knowing only the graph structure}
+
+\subsection{Knowing only the graph structure}
 \begin{frame}
@@
 \end{frame}
-\subsubsection{General case: Model unknown}
+
+\subsection{General case: Model unknown}
 \begin{frame}
@@
 If the model is completely unknown, two ideas appear:
+\begin{enumerate}
+\item If the max number of variables is known.
+\item If it is not.
+\end{enumerate}
+\end{frame}
+\begin{frame}
+\frametitle{Idea 1 for the General Case}
 \begin{itemize}
@@
 \item For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Idea 2 for the General Case}
+\begin{itemize}
 \item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables.
@@
 
+\section{Working and Publications Plan}
+\subsection{Working Plan}
+\subsection{Publications}
+
 \section{Some Results}
@@
 \frametitle{Some results}
-To show the factibility of this proposal, we considered a test scenario and two cases in ascending difficulty.
+To show the factibility of this proposal, we considered a test scenario and two of the previously mentioned cases.
 \end{frame}
@@
 \begin{itemize}
-\item A distribution suitable to modelling discrete probability vectors is the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}.
-The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
-
+\item We use the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}. 
+\item The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
 \[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1}\]
 \item The decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
-\item Using the agent's current beliefs, a causal graphical model can be specified. 
 \item Using this fully specified (structure + parameters) as a true model, the decision maker will make his choice as according to the guiding principle.

### 2018-11-25 Propuesta/Presentacion/Presentacion_Corta.tex
Indice sin subsecciones

@@
 \begin{frame}[allowframebreaks]
-\frametitle{Overview} % Table of contents slide, comment this block out to remove it
-\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
+\frametitle{Contents} % Table of contents slide, comment this block out to remove it
+\tableofcontents[hideallsubsections] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
 \end{frame}
+
+%
+%\begin{frame}
+%    \frametitle{Outline}
+%    \begin{columns}[t]
+%        \begin{column}{.3\textwidth}
+%            \tableofcontents[sections={1-3}]
+%        \end{column}
+%        \begin{column}{.3\textwidth}
+%            \tableofcontents[sections={4-5}]
+%        \end{column}
+%         \begin{column}{.3\textwidth}
+%            \tableofcontents[sections={6-9}]
+%        \end{column}
+%    \end{columns}
+%\end{frame}
 

### 2018-11-25 Propuesta/Presentacion/Presentacion_Corta.tex
añadiendo objetivos específicos

@@
 \item Our actions \textit{cause} changes in the environment.
-\item We consider this changes in order to plan our actions
+\item We consider this changes in order to plan our actions when facing similar situations.
 \end{itemize}
@@
 \item Relations are encoded in a graph.
-\item Can be used to predict \textit{effects} of certain causes
+\item Can be used to predict \textit{effects} of certain causes.
 \end{itemize}
@@
 \includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/example.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
 \end{center}
@@
 \section{Concepts}
-\subsection{Causal Models}
 \begin{frame}
 \frametitle{Concepts}
-Concepts and Framework
+Concepts and Framework:
+\begin{enumerate}
+\item Causal Models
+\item Decision Problems
+\end{enumerate}
 \end{frame}
+
+\subsection{Causal Models}
 		\begin{frame}
@@
 	\begin{itemize}
-	\item A (rational) decision maker.
 	\item A set of available options $\mathcal{A}$.
-	\item A preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
-	\item A set of uncertain events $\mathcal{E}$.
-    \item A set of possible outcomes $\mathcal{C}$
+	\item A (rational) decision maker with preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
+	\item A family of uncertain events $\mathcal{E}$.
+    \item A family of possible outcomes $\mathcal{C}$
 	\end{itemize}
@@
 	\item It allows to ask \textit{why}.
+	\item Transferability
 	\end{itemize}
@@
 \begin{itemize}
-\item An action could be conceived as an \textit{intervention} over some environment, which is in fact how humans consider actions in the world
+\item An action can be conceived as an \textit{intervention} over an environment, which is in fact how humans consider their actions in the world.
 \item They use and modify causal knowledge during a sequential decision making process.
@@
 	\end{itemize}
+	\end{frame}
+
+	\subsection{Specific Objectives}
+	\begin{frame}
+	\frametitle{Specific Objectives}
+	In order to reach the General Objective we must:
+	\begin{enumerate}
+
+	\end{enumerate}
 	\end{frame}

### 2018-11-25 Propuesta/Presentacion/Presentacion_Corta.tex
Plan de trabajo publicaciones y gantt

@@
 	\begin{enumerate}
-
+	\item Game between agent and Nature in order to modell agent-environment interaction.
+	\item Find an adequate family of probability distributions that encode causal knowledge.
+	\item General belief-updating criteria in terms of causal information.
 	\end{enumerate}
@@
 \section{Working and Publications Plan}
+\begin{frame}
+Working and Publications Plan
+\end{frame}
 \subsection{Working Plan}
-\subsection{Publications}
+\begin{frame}
+\frametitle{Working Plan}
+\begin{enumerate}
+\item Literature Review.
+\item Studying implications of the rationality assumption for decision making under uncertainty in causal environments.
+\item Formalization of a Causal Decision Problem.
+\item Propose a representation of beliefs that allow causal structure to be considered.
+\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information when involved variables are known.
+\item Study the problem of on-line causal discovery (fully unknown model). Convergence of proposed solution.
+\end{enumerate}
+\end{frame}
+
+\subsection{Publications Plan}
+\begin{frame}
+\frametitle{Publications Plan}
+\begin{itemize}
+\item \textbf{Year 1:} Gonzalez-Soto, M., Sucar, L.E., Escalante, H.J., \textbf{Playing against Nature: causal discovery for decision making under uncertainty}.
+Presented at \textbf{CausalML 2018 Workshop at ICML 2018}.
+July 15 2018, Stockholm, Sweden.
+The paper can be found at \url{https://arxiv.org/abs/1807.01268}.
+\item \textbf{Year 2:} Conference Paper 2 
+		\begin{itemize}
+		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
+		\item Possible Forum: Beyond Curve Fitting: Causation, Counterfactuals, and Imagination-based AI. AAAI Spring Symposium, March 25-27, 2019, Stanford, CA
+		\end{itemize}
+\item \textbf{Year 2:} Conference Paper 3 and 4
+	\begin{itemize}
+	\item When a causal model is unknown but the involved variables are known, how to encode uncertainties about causal structure in a probabilistic distribution?
+Bayesian Nonparametric prior that explicitly represents causal structure.
+	\item When variables of the environment are unknown how to encode prior beliefs and how to update them when data is not fully observed.
+	\item Possible Forum: International Conference on Bayesian Nonparametrics, International Conference on Probabilistic Graphical Models.
+	\end{itemize}
+\item \textbf{Year 3:} Journal Paper
+		\begin{itemize}
+			\item Aim: Theoretical paper of results for on-line causal learning.
+		\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Gantt Diagram}
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Gantt.png}
+\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+\end{frame}
 

### 2018-11-27 Propuesta/Presentacion/Presentacion_Corta.tex
ultimas correcciones

@@
 \theoremstyle{plain}
-\newtheorem{teo}{Teorema}
+\newtheorem{teo}{Theorem}
 \newtheorem{prop}[teo]{Proposition}
@@
 
-\title[Short title]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page
+\title[Causally Choosing]{Use and acquisition of causal knowledge in a decision making process} % The short title appears at the bottom of every slide, the full title is only on the title page
 
@@
 \item Both \cite{hohwy2013predictive}, \cite{clark2015surfing}, assert that the brain itself is a causal hypothesis tester. 
+\item While Danks \cite{danks2014unifying} argues that cognitive processes (including causal reasoning) take place in the mind as graphical models, Tenenbaum's school argues that the mind uses hierarchical bayesian inference (\cite{kemp2010learning}).
 \end{itemize}
@@
 \item We add an extra to the environment: causal structure.
+\item This means that there exists a fix but unknown causal model that governs the environment where the agent lives.
 \end{itemize}
@@
 \item Relations are encoded in a graph.
-\item Can be used to predict \textit{effects} of certain causes.
+\item Can be used to predict \textit{effects} of certain causes without necessarily performing the action.
+\item They allow to \textit{imagine} what would happen if...
 \end{itemize}
@@
 \begin{itemize}
-\item Current decision-making theories do not consider causal structure in the decision making processes
+\item Current decision-making theories do not consider causal structure in decision making processes
 \item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
@@
 \begin{enumerate}
-\item Causal Models
-\item Decision Problems
+\item Causal Models.
+\item Decision Problems.
 \end{enumerate}
@@
 \subsection{Causal Models}
+		\begin{frame}
+		\frametitle{Causal Models}
+		Causality and Causal Models:
+		\end{frame}
 		\begin{frame}
@@
 		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
-		\item Antisymetric: For $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
+		\item Antisymetric: for $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
 		\end{itemize}
-		If $A \to B$ we will say that $A$ is a cause of $B$.
+		If $A \to B$ we will say that $A$ is \textbf{a cause} of $B$.
 		\end{defi}
@@
 		\frametitle{From Graphs to Probability Measures}
-		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
+		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph (\cite{koller2009probabilistic}). \\
+		We require, as axioms, that this measure satisfies:
         \begin{itemize}
@@
         \end{itemize}
-         This requirements will be treated in an axiomatic fashion.
 		\end{frame}
@@
 	\begin{frame}
+	Decision Problems and Rational Thinking.
+	\end{frame}
+	\begin{frame}
 	\frametitle{Decision Problems}
-	A decision problem is composed by (\cite{bernardo2000bayesian}):
+	A decision problem is composed by (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}):
 	\begin{itemize}
@@
 \begin{frame}
-About my Proposed Research
+About my Proposed Research:
+\begin{enumerate}
+\item Motivation.
+\item Justification.
+\item Research Questions.
+\item Problem Statement.
+\item Hypothesis.
+\end{enumerate}
 \end{frame}
@@
 	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
-	\item What about not knowing the causal model? 
+	\item What about \textbf{not knowing} the causal model? 
 	\end{itemize}
@@
 	\item Associative methods capture only statistical patterns that are found in data.
-	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
+	\item As an example, methods based in Deep Neural Networks aren't even supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
 	\end{itemize}
@@
 	\begin{frame}
-	\frametitle{Justification: Need for nterpretable decision-making models}
+	\frametitle{Justification: Need for interpretable decision-making models}
 	\begin{itemize}
@@
 	\begin{itemize}
-	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
+	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be able to explain.
 	\item It allows to ask \textit{why}.
@@
 \frametitle{Approaches and Related Work}
+Approaches and Related Work:
+\end{frame}
+\begin{frame}
+\frametitle{Approaches and Related Work}
 \begin{figure}[ht]
@@
 \includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Related_Work.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
+\caption{Related Work}
 \end{center}
@@
 \end{frame}
+
 \subsection{Approaches and Related Work: Reinforcement Learning}
@@
 \item \cite{hagmayer2009decision}, \cite{hagmayer2013repeated} show that human beings can learn causal information in simple decision-making tasks, but remains open to study causal learning while seeking for a specific objective.
-\item 
 \end{itemize}
@@
 \begin{itemize}
-\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
+\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce (\cite{joyce1999foundations}), who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
 \item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
@@
 \item Their work requires that the causal model is fully known to the decision maker.
-\item this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
+\item Relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
 \end{itemize}
 \end{frame}
-
 \section{Proposed Solution and Methodology}
@@
 \frametitle{Proposed Solution and Methodology}
+Proposed Solution and Methodology:
+\begin{enumerate}
+\item General and Specific Objectives.
+\item Methodology.
+\item Validation.
+\item Detailed Methodology.
+\end{enumerate}
 \end{frame}
+
 	\subsection{Objectives}
@@
 	\begin{enumerate}
-	\item Game between agent and Nature in order to modell agent-environment interaction.
+	\item Define a Game between agent and Nature in order to model agent-environment interaction.
 	\item Find an adequate family of probability distributions that encode causal knowledge.
-	\item General belief-updating criteria in terms of causal information.
+	\item Develop a general belief-updating criteria in terms of causal information.
 	\end{enumerate}
@@
 \begin{teo}
-Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$.
+Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information in the way we propose, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$.
 \end{teo}
@@
 \frametitle{A guiding principle for Causal Decision Problems}
-We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational.
-We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
+We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational.\\
+We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.\\
 It remains to be refined and extended for a general case. 
@@
 Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
-\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
+\[ P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
 \end{prop}
@@
 \begin{frame}
-\frametitle{Games}
+\frametitle{First Step: Modelling a Causal Decision Problem as a Game}
 \begin{itemize}
-\item In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
+\item The next step is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
 \begin{enumerate}
-\item the (original) decision maker, 
-\item and a player which will be called Nature, whose actions are to be guided by the causal model.
+\item The (original) decision maker.
+\item A player which will be called \textbf{Nature}, whose actions are to be guided by the causal model.
+\item The game is over when the \textit{outcome} of the original decision problem is realized. This we will call a \textit{round}.
 \end{enumerate}
@@
 \begin{itemize}
-\item  Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
+\item  We assume that player Nature has the first move and she assigns some state to any, or all, of the variable(s) in the causal model $\mathcal{G}$.
 \item After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
-\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games
+\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games.
+Details in \cite{osborne1994course}.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Belief Formation}
+\frametitle{Second Step: Belief Formation}
 \begin{itemize}
-\item The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
-\item The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world
+\item The whole idea is that a decision maker is able to \textbf{learn} and \textbf{reason} about his environment by observing the effects of his actions.
+\item The agent encodes any previous knowledge or uncertainties in a probability distribution.
+\item The observed information will be used to modify his current knowledge in a way that is similar to how human beings modify causal knowledge while intervening in the world
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Making Choices}
+\frametitle{Third Step: Making Good Choices}
 \begin{itemize}
-\item The decision maker will use his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
+\item The decision maker uses his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
 \item With this information he will choose the action that maximize the Expected Utility, according with our guiding principle.
@@
 
-\subsection{Third Step: Belief updating}
+\subsection{Fourth Step: Belief updating}
 \begin{frame}
@@
 \begin{itemize}
-\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
-\item it remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
+\item It is known that belief updating using Bayes' Theorem is the only way rational way of updating uncertainty.
+\item It remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way.
 \end{itemize}
@@
 \begin{frame}
-We will show how the methodology applies in simpler and easier cases where the decision maker has some exta info about his environment.
+We will show how the methodology applies in simpler and easier cases where the decision maker has some extra info about his environment.
 \end{frame}
@@
 \begin{itemize}
-\item If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
+\item Following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has.
+\item Choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
 \end{itemize}
@@
 \begin{itemize}
-\item If a decision maker knows the structure of the graph, but he does not know the parameters,
-\item then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
-\item In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e.
-in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
-\item This means that using current beliefs, the agent forms a causal model which is used as if it were the true causal model.
+\item The agent encodes uncertainty over the parameters of the model as a probability distribution over a suitable space.
+\item Done in such way that a \textit{local} (i.e. in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
+\item The agent forms a causal model which is used as, according to the guiding principle, as if it were the true causal model.
+\item Belief updating.
 \end{itemize}
@@
 \begin{enumerate}
-\item If the max number of variables is known.
+\item If the maximum number of variables is known.
 \item If it is not.
@@
 \item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables. 
-In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.
+\item In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.  
+\item The graph generating process must converge to the true model.
 \end{itemize}
@@
 \subsection{Publications Plan}
-\begin{frame}
+\begin{frame}[allowframebreaks]
 \frametitle{Publications Plan}
@@
 		\item The implications of Causal Relations for Decision Making: a general principle for Decision under Uncertainty.
-		\item Possible Forum: Beyond Curve Fitting: Causation, Counterfactuals, and Imagination-based AI. AAAI Spring Symposium, March 25-27, 2019, Stanford, CA
+		\item Possible Forum: \textbf{Beyond Curve Fitting: Causation, Counterfactuals, and Imagination-based AI. AAAI Spring Symposium}, March 25-27, 2019, Stanford, CA
 		\end{itemize}
-\item \textbf{Year 2:} Conference Paper 3 and 4
+\item \textbf{Year 2:} Conference Paper 3 
 	\begin{itemize}
@@
 \includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Gantt.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
 \end{center}
@@
 \frametitle{Some results}
-To show the factibility of this proposal, we considered a test scenario and two of the previously mentioned cases.
+To show the factibility of this proposal, we considered a simple test scenario and two of the previously mentioned cases.
 \end{frame}
@@
 \caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
 \end{center}
@@
 \section{Conclusions}
+\begin{frame}[allowframebreaks]
+\frametitle{Conclusions}
+\begin{itemize}
+\item We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism who can make use of the causal relations that hold in such environment in order to find an optmal action.
+\item Results show that causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.
+\end{itemize}
+\end{frame} 
+\begin{frame}[allowframebreaks]
+\begin{itemize}
+\item We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives. 
+\item We proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality. 
+\item From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order. 
+\item In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment.
+\item If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
+\end{itemize}
+\end{frame}
+
+\section{Future Work}
 \begin{frame}
-\frametitle{Conclusions}
-We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism can make use of the causal relations that hold in such environment in order to find an action that will take him into the way of a desired result, in terms of his (rational) preferences.
-Also, causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.
-\end{frame} 
-\begin{frame}
-We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives.
-Following this intuitions we proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality.
-From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order.
-In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment.
-We are considering.
-If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
+\frametitle{Future Work}
+\begin{itemize}
+\item Our immediate next work is to consider the \textit{2.5} case where the decision maker partially knows the \textit{structure} of the graph.
+\item Study the case 1 for the General Case stated above: where the agent knows the max number of variables. 
+\end{itemize}
 \end{frame}

### 2018-12-05 WHY19/Draft/Paper.tex
Abstract y estructura

\def\year{2018}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
\pdfinfo{
/Title (A Guiding Principle for Causal Decision Problems)
/Author (M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante)
}

\title{A Guiding Principle for Causal Decision Problems}
\author{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante\\
Address line
}

\begin{document}
\maketitle
\begin{abstract}
We define a Causal Decision Problem as a Decision Problem where available actions and outcomes are connected through the variables of a Causal Graphical Model \mathcal{G} and propose a solution criteria based on Pearl's Do-Calculus and the Expected Utility criteria for rational preferences. The implementation of this criteria leads to an on-line decision making procedure that has been shown to have similar performance to classic Reinforcement Learning algorithms while allowing for a causal model of an environment to be learned. 
\end{abstract}

\section{Introduction}
\noindent

\section{Rationality and Expected Utility}

\section{Optimal Policies}

\section{Causal Decision Problems}

\section{Proposed Solution}

\bibliography{Bibliografia.bib}
\bibliographystyle{aaai}
\end{document}

### 2018-12-05 WHY19/Draft/Paper.tex
añadiendo seccion

@@
 \begin{abstract}
-We define a Causal Decision Problem as a Decision Problem where available actions and outcomes are connected through the variables of a Causal Graphical Model \mathcal{G} and propose a solution criteria based on Pearl's Do-Calculus and the Expected Utility criteria for rational preferences.
+We define a Causal Decision Problem as a Decision Problem where available actions and outcomes are connected through the variables of a Causal Graphical Model $\mathcal{G}$ and propose a solution criteria based on Pearl's Do-Calculus and the Expected Utility criteria for rational preferences.
 The implementation of this criteria leads to an on-line decision making procedure that has been shown to have similar performance to classic Reinforcement Learning algorithms while allowing for a causal model of an environment to be learned.
@@
 
+\section{On-line decision making and causal learning}
+
 \bibliography{Bibliografia.bib}

### 2018-12-07 WHY19/Draft/Paper.tex
primeras secciones

@@
 \author{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante\\
-Address line
+National Institute of Astrophysics Optics and Electronics (INAOE)\\
+Luis Enrique Erro No. 1, Santa Maria Tonanzintla, Puebla, Mexico.
 }
@@
 \section{Rationality and Expected Utility}
-
+Rationality in a Decision Making setting is defined axiomatically in a way that the \textit{preferences} of a decision maker are logically consistent.
+If rational preferences are assumed, then it is known that the coherent criteria for making choices is the maximization of expected utility, either with respect to a known utility function and probability distribution or a pair of \textit{subjective} objects (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+Rational decision making has been the standard theory in economics both as a \textit{normative} and a \textit{descriptive} theory of human behavior and it has been the subject of multiple debates (\cite{tversky1974judgment}, \cite{kahneman1982judgment}).
+In this work we pretend to take a \textit{normative} view for a rational decision maker who faces an uncertain environment which is controlled by some unknown causal mechanism.
 \section{Optimal Policies}
-
+Given a Reinforcement Learning Problem defined over a Markov Decision Process, a \textit{policy} is a function from the space of \textit{states} to the space of \textit{actions} which is interpreted as what should an agent do in a given state (\cite{sutton1998reinforcement}).
+An \textit{optimal policy} is a policy which is optimal in the sense of achieiving the maximum possible reward, which is equivalent to finding the optimal action in the sense of the maximum expected utility (\cite{webb2007game}).
 \section{Causal Decision Problems}
-
+We define a Causal Decision Problem under Uncertainty as a tuple $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \mathcal{G}, \succeq)$ where  $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \succeq)$ is a classical Decision Problem under Uncertainty (\cite{bernardo2000bayesian} and $\mathcal{G}$ is a Causal Graphical Model (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) such that the set of available actions $\mathcal{A}$ and the set of outcomes $\mathcal{C}$ are related through the variables of the Causal Model $\mathcal{G}$; i.e., the events in the family $\mathcal{E}$ correspond to variables in $\mathcal{G}$.
+It is assumed that the agent does not know the Causal Model $\mathcal{G}$, which is equivalent to not knowing the probabilities of the events $E \in \mathcal{E}$.
+The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions (\cite{woodward2005making}) and to satisfy the conditions expressed in \cite{spirtes2000causation}.
 \section{Proposed Solution}
+Since rationality is assumed we must seek to maximize the expected utility of
 
 \section{On-line decision making and causal learning}
-
+In \cite{gonzalez2018playing} a decision-making procedure was proposed using the intuition of the Proposed Solution and a Bayesian \textit{belief updating} procedure in order to learn an optimal action in the simpler case where the decision maker knows the \textit{structure} of the model $\textit{G}$.
+The actions learned after a series of rounds obtained similar performance as an agent learning using the classical Q-Learning procedure.
+Experimentally, this
 \bibliography{Bibliografia.bib}

### 2018-12-07 WHY19/Draft/Paper.tex
añadiendo solucion propuesta

@@
 The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions (\cite{woodward2005making}) and to satisfy the conditions expressed in \cite{spirtes2000causation}.
+The variable in $\mathcal{G}$ which encodes the consequence of the action taken by the agent will be referred as \textit{target variable} since it is the variable where the agent whishes to obtain a desired result.
 \section{Proposed Solution}
-Since rationality is assumed we must seek to maximize the expected utility of
+Since rationality is assumed we must seek to maximize the expected utility of the agent in terms of his current knowledge, which is expressed as a (subjective) probability distribution.
+Using the awareness of the agent about a causal mechanism governing the environment, intuition encourages to use this causal relations to cause some desirable action, as expressed by \cite{joyce1999foundations}
+
+Consider a Causal Decision Problem as stated above where the target variable, call it $Y$, takes its values in the set $\{0,1\}$ and without loss of generality assume that $1$ is the desired output for the agent, then he must choose the action $a^\ast \in \mathcal{A}$ such that
+\[ P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a^\ast)) \textrm{ for all } a \in \mathcal{A}. \]
+Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent.
+If an action $a_0 \in \mathcal{A}$ yielded a higher expected utility, that would imply that it has higher probability of \textit{causing} the same desired action, but this is not possible because of how $a^\ast$ is obtained.
 
 \section{On-line decision making and causal learning}
-In \cite{gonzalez2018playing} a decision-making procedure was proposed using the intuition of the Proposed Solution and a Bayesian \textit{belief updating} procedure in order to learn an optimal action in the simpler case where the decision maker knows the \textit{structure} of the model $\textit{G}$.
-The actions learned after a series of rounds obtained similar performance as an agent learning using the classical Q-Learning procedure.
-Experimentally, this
+In \cite{gonzalez2018playing} a decision-making procedure was proposed using the Proposed Solution together with Bayesian \textit{belief updating} procedure in order to learn an optimal action while acquiring a causal model in the environment, this was applied in the simpler case where the decision maker knows the \textit{structure} of the model $\textit{G}$.
+In the referred work, a decision maker held \textit{beliefs} about the causal information of the environment, which were encoded into probability distributions.
+Those beliefs were used \textit{as if} they were the true model governing the environment in order to choose an action according to the Solution proposed here.
+Beliefs were updated in a Bayesian fashion after observing the outcome.
+
+The actions learned after a series of decision rounds obtained similar performance (in terms of average reward) as an agent learning using the classical Q-Learning procedure.
+Experimentally, this shows that causal information allows an agent to learn an optimal action, in the sense of expected utility, as well as learning a causal model of the environment.
 \bibliography{Bibliografia.bib}

### 2018-12-14 WHY19/Draft/Draft.tex
version enviada

\def\year{2018}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{
/Title (A Guiding Principle for Causal Decision Problems)
/Author (M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{A Guiding Principle for Causal Decision Problems}
\author{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante\\
National Institute of Astrophysics Optics and Electronics (INAOE)}
\maketitle
\begin{abstract}
Extending previous work 
\end{abstract}

\noindent Congratulations on having a paper selected for inclusion in an AAAI Press proceedings or technical report! This document details the requirements necessary to get your accepted paper published using \LaTeX{}. If you are using Microsoft Word, instructions are provided in a different document. If you want to use some other formatting software, you must obtain permission from AAAI Press first.

The instructions herein are provided as a general guide for experienced \LaTeX{} users. If you do not know how to use \LaTeX{}, do not use it to format your paper. AAAI cannot provide you with support and the accompanying style files are \textbf{not} guaranteed to work. If the results you obtain are not in accordance with the specifications you received, you must correct your source file to achieve the correct result.

These instructions are generic. Consequently, they do not include specific dates, page charges, and so forth. Please consult your specific written conference instructions for details regarding your submission. Please review the entire document for specific instructions that might apply to your particular situation. All authors must comply with the following:

\begin{itemize}
\item You must use the 2018 AAAI Press \LaTeX{} style file and bib file, which are located in the 2018 author kit.
\item You must complete, sign, and return by the deadline the AAAI copyright form (proceedings authors) or distribution license (technical report authors).
\item You must read and format your paper source and PDF according to the formatting instructions for authors.
\item You must submit your electronic files and abstract using our electronic submission form \textbf{on time.}
\item You must pay any required page or formatting charges to AAAI Press so that they are received by the deadline.
\item You must check your paper before submitting it, ensuring that it compiles without error, and complies with the guidelines found in the author kit.
\end{itemize}

\section{Copyright}
All papers submitted for publication by AAAI Press must be accompanied by a valid signed copyright form or, in the case of technical reports, by a valid signed permission to distribute form. There are no exceptions to this requirement. You must send us the original version of this form. However, to meet the deadline, you may fax (1-650-321-4457) or scan and e-mail the form (pubforms18@aaai.org) to AAAI by the submission deadline, and then mail the original via postal mail to the AAAI office. \textbf{If you fail to send in a signed copyright or permission form, we will be unable to publish your paper. There are no exceptions to this policy.} You will find PDF versions of the AAAI copyright and permission to distribute forms in the author kit.

\section{Formatting Requirements in Brief}
We need source and PDF files that can be used in a variety of ways and can be output on a variety of devices. The design and appearance of the paper is governed by the aaai style file. 
\textbf{You must not make any changes to the aaai style file, nor use any commands, packages, style files, or macros within your own paper that alter that design, including, but not limited to spacing, floats, margins, fonts, font size, and appearance.} AAAI imposes  requirements on your source and PDF files that must be followed. Most of these requirements are based on our efforts to standardize conference manuscript properties and layout. All papers submitted to AAAI for publication will be recompiled for standardization purposes. Consequently, every paper submission must comply with the following requirements:

\begin{itemize}
\item Your .tex file must compile in PDF\LaTeX{} --- \textbf{ no .ps or .eps figure files.}
\item All fonts must be embedded in the PDF file --- \textbf{ this includes your figures.}
\item Modifications to the style file, whether directly or via commands in your document may not be made, most especially when made in an effort to avoid extra page charges or make your paper fit in a specific number of pages.
\item No type 3 fonts may be used (even in illustrations).
\item You may not alter the spacing above and below captions, figures, headings, and subheadings.
\item You may not alter the font sizes of text elements, footnotes, heading elements, captions, or title information (for references and tables and mathematics, please see the the limited exceptions provided herein).
\item You may not alter the line spacing of text.
\item Your title must follow Title Case capitalization rules (not sentence case).
\item Your .tex file must include completed metadata to pass-through to the PDF (see PDFINFO below)
\item \LaTeX{} documents must use the Times or Nimbus font package (do not use Computer Modern for the text of your paper).
\item No \LaTeX{} 209 documents may be used or submitted.
\item Your source must not require use of fonts for non-Roman alphabets within the text itself. If your paper includes symbols in other languages (such as, but not limited to, Arabic, Chinese, Hebrew, Japanese, Thai, Russian and other Cyrillic languages), you must restrict their use to bit-mapped figures.
\item Fonts that require non-English language support (CID and Identity-H) must be converted to outlines or 300 dpi bitmap or removed from the document (even if they are in a graphics file embedded in the document). 
\item Two-column format in AAAI style is required for all papers.
\item The paper size for final submission must be US letter without exception.
\item The source file must exactly match the PDF.
\item The document margins must be as specified in the formatting instructions.
\item The number of pages and the file size must be as specified for your event.
\item No document may be password protected.
\item Neither the PDFs nor the source may contain any embedded links or bookmarks.
\item Your source and PDF must not have any page numbers, footers, or headers.
\item Your PDF must be compatible with Acrobat 5 or higher.
\item Your \LaTeX{} source file (excluding references) must consist of a \textbf{single} file (use of the ``input" command is not allowed.
\item Your graphics must be sized appropriately outside of \LaTeX{} (do not use the ``clip" command) .
\end{itemize}

If you do not follow the above requirements, your submission will be subject to reformatting and special handling fees that can easily exceed the extra page fee.

\section{What Files to Submit}
You must submit the following items to ensure that your paper is published:
\begin{itemize}
\item A fully-compliant PDF file.
\item Your  \LaTeX{}  source file submitted as a \textbf{single} .tex file (do not use the ``input" command to include sections of your paper --- every section must be in the single source file). The only exception is the reference list, which you should include separately. Your source must compile on our system, which includes the standard \LaTeX{} support files.
\item Only the graphics files used in compiling paper.
\item The \LaTeX{}-generated files (e.g. .aux and .bib file, etc.) for your compiled source.
\item If you have used an old installation of \LaTeX{}, you should include algorithm style files). If in doubt, include it.
\end{itemize}

Your \LaTeX{} source will be reviewed and recompiled on our system (if it does not compile, you may incur late fees).   \textbf{Do not submit your source in multiple text files.} Your single \LaTeX{} source file must include all your text, your bibliography (formatted using aaai.bst), and any custom macros. Accompanying this source file, you must also supply any nonstandard (or older) referenced style files and all your referenced graphics files.

Your files should work without any supporting files (other than the program itself) on any computer with a standard \LaTeX{} distribution. Place your PDF and source files in a single tar, zipped, gzipped, stuffed, or compressed archive. Name your source file with your last (family) name.

\textbf{Do not send files that are not actually used in the paper.} We don't want you to send us any files not needed for compiling your paper, including, for example, this instructions file, unused graphics files, standard style files, additional material sent for the purpose of the paper review, and so forth.

\textbf{Obsolete style files.}  The commands for some common packages (such as some used for algorithms), may have changed. Please be certain that you are not compiling your paper using old or obsolete style files.

\section{Using \LaTeX{} to Format Your Paper}

The latest version of the AAAI style file is available on AAAI's website. Download this file and place it in  the \TeX\ search path. Placing it in the same directory as the paper should also work. You must download the latest version of the complete author kit so that you will have the latest instruction set and style file.

\subsection{Document Preamble}

In the \LaTeX{} source for your paper, you \textbf{must} place the following lines as shown in the example in this subsection. This command set-up is for three authors. Add or subtract author and address lines as necessary, and uncomment the portions that apply to you. In most instances, this is all you need to do to format your paper in the Times font. The helvet package will cause Helvetica to be used for sans serif. These files are part of the PSNFSS2e package, which is freely available from many Internet sites (and is often part of a standard installation).

Leave the setcounter for section number depth commented out and set at 0 unless you want to add section numbers to your paper. If you do add section numbers, you must uncomment this line and change the number to 1 (for section numbers), or 2 (for section and subsection numbers). The style file will not work properly with numbering of subsubsections, so do not use a number higher than 2.

If (and only if) your author title information will not fit within the specified  height allowed, put \textbackslash setlength \textbackslash titlebox{2.5in} in your preamble. Increase the height until the height error disappears from your log. You may not use the  \textbackslash setlength command elsewhere in your paper, and it may not be used to reduce the height of the author-title box.

\subsubsection{The Following Must Appear in Your Preamble}
\begin{quote}
\begin{scriptsize}\begin{verbatim}
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{url}
\usepackage{graphicx}
\frenchspacing
% Add additional packages here. The following
% packages may NOT be used  (this list
% is not exhaustive:
% authblk, caption, CJK, float, fullpage, geometry, 
%hyperref, layout, nameref, natbib, savetrees, 
%setspace, titlesec, tocbibind, ulem
%
%US Lettersize Paper Is Required
\setlength{pdfpagewidth}{8.5in}
\setlength{pdfpageheight}{11in}\\
%
%
% PDFINFO
% You are required to complete the following
% for pass-through to the PDF. 
% No LaTeX commands of any kind may be
% entered. The parentheses and spaces 
% are an integral part of the 
% pdfinfo script and must not be removed.
%
\pdfinfo{
/Title (Input Your Paper Title Here)
/Author (John Doe, Jane Doe)
/Keywords (Input your keywords in this optional area)
}
%
%Section Numbers
% Uncomment if you want to use section numbers
% and change the 0 to a 1 or 2
% \setcounter{secnumdepth}{0}

% Title and Author Information Must Immediately Follow
% the pdfinfo within the preamble
%
\title{Title}\\
\author\{Author 1 \ and Author 2\\
Address line\\
Address line\\
\ And\\
Author 3\\
Address line\\
Address line
}\\
%
\end{verbatim}\end{scriptsize}
\end{quote}

\subsection{Preparing Your Paper}

After the preamble above, you should prepare your paper as follows:
\begin{quote}
\begin{scriptsize}\begin{verbatim}
%
\begin{document}
\maketitle
\begin{abstract}
%...
\end{abstract}\end{verbatim}\end{scriptsize}
\end{quote}

\subsubsection{The Following Must Conclude Your Document}
\begin{quote}
\begin{scriptsize}\begin{verbatim}
%References and End of Paper
%These lines must be placed at the end of your paper
\bibliography{Bibliography-File}
\bibliographystyle{aaai}
\end{document}
\end{verbatim}\end{scriptsize}
\end{quote}

\subsection{Inserting Document Metadata with \LaTeX{}}
PDF files contain document summary information that enables us to create an Acrobat index (pdx) file, and also allows search engines to locate and present your paper more accurately. \textit{Document metadata for author and title are REQUIRED.} You may not apply any script or macro to implementation of the title, author, and metadata information in your paper.

\textit{Important:} Do not include \textit{any} \LaTeX{} code or nonascii characters (including accented characters) in the metadata. The data in the metadata must be completely plain ascii. It may not include slashes, accents, linebreaks, unicode, or any \LaTeX{} commands. Type the title exactly as it appears on the paper (minus all formatting). Input the author names in the order in which they appear on the paper (minus all accents), separating each author by a comma. You may also include keywords in the Keywords field.

\begin{quote}
\begin{scriptsize}\begin{verbatim}
\begin{document}\\
\maketitle\\
...\\
\bibliography{Bibliography-File}\\
\bibliographystyle{aaai}\\
\end{document}\\
\end{verbatim}\end{scriptsize}
\end{quote}

\subsection{Illegal Commands}
There are a number of packages, commands, scripts, and macros that are incompatable with aaai18.sty. The common ones are provided later on in this document (See some Common  Errors to Avoid).  Generally, if a command, package, script, or macro alters floats, margins, fonts, sizing, linespacing, or the presentation of the references and citations, it is unacceptable. The following commands are some of the more common ones that may not be used in your paper (this list is not exhaustive --- there are others):
\begin{itemize}
\item \textbackslash renewcommand (in almost all instances)
\item \textbackslash baselinestretch
\item \textbackslash setlength (except for titlebox)
\item \textbackslash input
\item \textbackslash vspace or vskip (when used before or after a section or subsection or figure or table)
\item \textbackslash addtolength 
\item \textbackslash columnsep
\item \textbackslash top margin (or text height or addsidemargin or even side margin)
\item trim or clip (used to crop figures)
\item any command that globally alters floats, space above and below figures and tables
\end{itemize}

\subsection{Removing Certain Commands in Your Final Paper}
For your final camera ready copy, you must not use any page break commands, including, but not limited to:
\begin{itemize}
\item \textbackslash newpage
\item \textbackslash break
\item \textbackslash clearpage
\item \textbackslash pagebreak
\end{itemize}
(References must flow directly after the text without breaks.) Note that this may \textit{not} the case when submitting a paper for review. Some conferences require references to be on a separate page during the review process. AAAI Press, however, does not require this condition for the final paper. You must also be certain that your \textbackslash pdfino author field includes all your authors. The latter step is often forgotten.

\subsection{Paper Size, Margins, and Column Width}
Papers must be formatted to print in two-column format on 8.5 x 11 inch US letter-sized paper. The margins must be exactly as follows: 
\begin{itemize}
\item Top margin: .75 inches
\item Left margin: .75 inches
\item Right margin: .75 inches
\item Bottom margin: 1.25 inches
\end{itemize}

The default paper size in most installations of \LaTeX{} is A4. However, because we require that your electronic paper be formatted in US letter size, the preamble we have provided includes commands that alter the default to US letter size. Please note that using any other package to alter page size (such as, but not limited to the Geometry package) will result in your final paper being rejected.

\subsubsection{Column Width and Margins.}
To ensure maximum readability, your paper must include two columns. Each column should be 3.3 inches wide (slightly more than 3.25 inches), with a .375 inch (.952 cm) gutter of white space between the two columns. The aaai18.sty file will automatically create these columns for you.

\subsection{Overlength Papers}
If your paper is too long, turn on \textbackslash frenchspacing, which will reduce the space after periods. Next,  shrink the size of your graphics. Use \textbackslash centering instead of \textbackslash begin\{center\} in your figure environment. For mathematical environments, you may reduce fontsize {\bf but not below 6.5 point}. You may also alter the size of your bibliography by inserting \textbackslash fontsize\{9.5pt\}\{10.5pt\} \textbackslash selectfont
right before the bibliography (the minimum size is \textbackslash fontsize\{9.0pt\}\{10.0pt\}.

Commands that alter page layout are forbidden. These include \textbackslash columnsep, \textbackslash topmargin, \textbackslash topskip, \textbackslash textheight, \textbackslash textwidth, \textbackslash oddsidemargin, and \textbackslash evensizemargin (this list is not exhaustive). If you alter page layout, you will be required to pay the page fee \textit{plus} a reformatting fee. Other commands that are questionable and may cause your paper to be rejected include  \textbackslash parindent, and \textbackslash parskip. Commands that alter the space between sections are forbidden. The title sec package is not allowed. Regardless of the above, if your paper is obviously ``squeezed" it is not going to to be accepted. Options for reducing the length of a paper include reducing the size of your graphics, cutting text, or paying the extra page charge (if it is offered).

\subsection{Figures}
Your paper must compile in PDF\LaTeX{}. Consequently, all your figures must be .jpg, .png, or .pdf. You may not use  the .gif (the resolution is too low), .ps, or .eps file format for your figures.

When you include your figures, you must crop them \textbf{outside} of \LaTeX{}. The command \textbackslash includegraphics*[clip=true, viewport 0 0 10 10]{...} might result in a PDF that looks great, but the image is \textbf{not really cropped.} The full image can reappear (and obscure whatever it is overlapping) when page numbers are applied or color space is standardized.

\subsection{Type Font and Size}
Your paper must be formatted in Times Roman or Nimbus. We will not accept papers formatted using Computer Modern or Palatino or some other font as the text or heading typeface. Sans serif, when used, should be Courier. Use Symbol or Lucida or Computer Modern for \textit{mathematics only. }

Do not use type 3 fonts for any portion of your paper, including graphics. Type 3 bitmapped fonts are designed for fixed resolution printers. Most print at 300 dpi even if the printer resolution is 1200 dpi or higher. They also often cause high resolution imagesetter devices and our PDF indexing software to crash. Consequently, AAAI will not accept electronic files containing obsolete type 3 fonts. Files containing those fonts (even in graphics) will be rejected.

Fortunately, there are effective workarounds that will prevent your file from embedding type 3 bitmapped fonts. The easiest workaround is to use the required times, helvet, and courier packages with \LaTeX{}2e. (Note that papers formatted in this way will still use Computer Modern for the mathematics. To make the math look good, you'll either have to use Symbol or Lucida, or you will need to install type 1 Computer Modern fonts --- for more on these fonts, see the section ``Obtaining Type 1 Computer Modern.")

If you are unsure if your paper contains type 3 fonts, view the PDF in Acrobat Reader. The Properties/Fonts window will display the font name, font type, and encoding properties of all the fonts in the document. If you are unsure if your graphics contain type 3 fonts (and they are PostScript or encapsulated PostScript documents), create PDF versions of them, and consult the properties window in Acrobat Reader.

The default size for your type should be ten-point with twelve-point leading (line spacing). Start all pages (except the first) directly under the top margin. (See the next section for instructions on formatting the title page.) Indent ten points when beginning a new paragraph, unless the paragraph begins directly below a heading or subheading.

\subsubsection{Obtaining Type 1 Computer Modern for \LaTeX{}.}

If you use Computer Modern for the mathematics in your paper (you cannot use it for the text) you may need to download type 1 Computer fonts. They are available without charge from the American Mathematical Society:
http://www.ams.org/tex/type1-fonts.html.

\subsection{Title and Authors}
Your title must appear in mixed case (nouns, pronouns, and verbs are capitalized) near the top of the first page, centered over both columns in sixteen-point bold type (twenty-four point leading). This style is called ``mixed case." Author's names should appear below the title of the paper, centered in twelve-point type (with fifteen point leading), along with affiliation(s) and complete address(es) (including electronic mail address if available) in nine-point roman type (the twelve point leading). (If the title is long, or you have many authors, you may reduce the specified point sizes by up to two points.) You should begin the two-column format when you come to the abstract.

\subsubsection{Formatting Author Information}
Author information can be set in a number of different styles, depending on the number of authors and the number of affiliations you need to display. In formatting your author information, however, you may not use a table nor may you employ the \textbackslash authorblk.sty package. For several authors from the same institution, use \textbackslash and:

\begin{quote}\begin{scriptsize}\begin{verbatim}
\author{Author 1 \and ... \and Author n\\ 
Address line \\ ... \\ Address line}
\end{verbatim}\end{scriptsize}\end{quote}

\noindent If the names do not fit well on one line use:

\begin{quote}\begin{scriptsize}\begin{verbatim}
\author{Author 1} ... \\
{\bf \Large Author  ...  Author}\\ 
Address line \\ ... \\ Address line
}
\end{verbatim}\end{scriptsize}\end{quote}

\noindent For authors from different institutions, use \textbackslash And:

\begin{quote}\begin{scriptsize}\begin{verbatim}
\author{Author 1\\ Address line \\ ... \\ Address line 
\And ... \And Author n\\
Address line\\ ... \\ Address line}
\end{verbatim}\end{scriptsize}\end{quote}

\noindent To start a separate ``row" of authors, use \textbackslash AND:

\noindent If the title and author information does not fit in the area allocated, place
\textbackslash setlength\textbackslash titlebox\{\textit{height}\}
after the \textbackslash documentclass line where \{\textit{height}\} is  2.5in or greater.

\subsubsection{Formatting Author Information --- Alternative Method}
If your paper has a large number of authors from different institutions, you may use the following alternative method for displaying the author information.

\begin{quote}\begin{scriptsize}\begin{verbatim}
\author{AuthorOne},\textsuperscript{1}
\author{AuthorTwo},\textsuperscript{2}
\author{AuthorThree},\textsuperscript{3}
\author{AuthorFour},\textsuperscript{4}
\author{AuthorFive}, \textsuperscript{5}\\
\textsuperscript{1}AffiliationOne}\\
\textsuperscript{2}AffiliationTwo}\\
\textsuperscript{3}AffiliationThree}\\
\textsuperscript{4}AffiliationFour}\\
\textsuperscript{5}AffiliationFive}\\
\{email, email\}@affiliation.com,
email@affiliation.com,
email@affiliation.com, 
email@affiliation.com
\end{verbatim}\end{scriptsize}\end{quote}

\subsection{\LaTeX{} Copyright Notice}
The copyright notice automatically appears if you use aaai.sty. If you are creating a technical report, it is not necessary to include this notice. You may disable the copyright line using the \verb+\+nocopyrightcommand. To change the entire text of the copyright slug, use:
\textbackslash copyrighttext \{\emph{text}\}.
Either of these must appear before \textbackslash maketitle. Please be advised, however, that \textit{if you disable or change the copyright line and transfer of copyright is required, your paper will not be published.}

\subsection{Credits}
Any credits to a sponsoring agency should appear in the acknowledgments section, unless the agency requires different placement. If it is necessary to include this information on the front page, use
\textbackslash thanks in either the \textbackslash author or \textbackslash title commands.
For example:
\begin{quote}
\begin{small}
\textbackslash title\{Very Important Results in AI\textbackslash thanks\{This work is
 supported by everybody.\}\}
\end{small}
\end{quote}
Multiple \textbackslash thanks commands can be given. Each will result in a separate footnote indication in the author or title with the corresponding text at the botton of the first column of the document. Note that the \textbackslash thanks command is fragile. You will need to use \textbackslash protect.

Please do not include \textbackslash pubnote commands in your document.

\subsection{Abstract}
Follow the example commands in this document for creation of your abstract. Further indentation is not required. {Do not include references in your abstract!}

\subsection{Page Numbers}

Do not \textbf{ever} print any page numbers on your paper.

\subsection{Text }
The main body of the paper must be formatted in ten-point with twelve-point leading (line spacing).

\subsection{Citations}
Citations within the text should include the author's last name and year, for example (Newell 1980). Append lower-case letters to the year in cases of ambiguity. Multiple authors should be treated as follows: (Feigenbaum and Engelmore 1988) or (Ford, Hayes, and Glymour 1992). In the case of four or more authors, list only the first author, followed by et al. (Ford et al. 1997).

\subsection{Extracts}
Long quotations and extracts should be indented ten points from the left and right margins.

\begin{quote}
This is an example of an extract or quotation. Note the indent on both sides. Quotation marks are not necessary if you offset the text in a block like this, and properly identify and cite the quotation in the text.

\end{quote}

\subsection{Footnotes}
Avoid footnotes as much as possible; they interrupt the reading of the text. When essential, they should be consecutively numbered throughout with superscript Arabic numbers. Footnotes should appear at the bottom of the page, separated from the text by a blank line space and a thin, half-point rule.

\subsection{Headings and Sections}
When necessary, headings should be used to separate major sections of your paper. Remember, you are writing a short paper, not a lengthy book! An overabundance of headings will tend to make your paper look more like an outline than a paper. The aaai.sty package will create headings for you. Do not alter their size nor their spacing above or below.

\subsubsection{Section Numbers}
The use of section numbers in AAAI Press papers is optional. To use section numbers in \LaTeX{}, uncomment the setcounter line in your document preamble and change the 0 to a 1 or 2. Section numbers should not be used in short poster papers.

\subsubsection{Section Headings.}
Sections should be arranged and headed as follows:

\subsubsection{Acknowledgments.}
The acknowledgments section, if included, appears after the main body of text and is headed ``Acknowledgments." This section includes acknowledgments of help from associates and colleagues, credits to sponsoring agencies, financial support, and permission to publish. Please acknowledge other contributors, grant support, and so forth, in this section. Do not put acknowledgments in a footnote on the first page. If your grant agency requires acknowledgment of the grant on page 1, limit the footnote to the required statement, and put the remaining acknowledgments at the back. Please try to limit acknowledgments to no more than three sentences.

\subsubsection{Appendices.}
Any appendices follow the acknowledgments, if included, or after the main body of text if no acknowledgments appear.

\subsubsection{References}
The references section should be labeled ``References" and should appear at the very end of the paper (don't end the paper with references, and then put a figure by itself on the last page). A sample list of references is given later on in these instructions. Please use a consistent format for references. Poorly prepared or sloppy references reflect badly on the quality of your paper and your research. Please prepare complete and accurate citations.

\subsection{Illustrations and Figures}
Figures, drawings, tables, and photographs should be placed throughout the paper near the place where they are first discussed. Do not group them together at the end of the paper. If placed at the top or bottom of the paper, illustrations may run across both columns. Figures must not invade the top, bottom, or side margin areas. Figures must be inserted using the \textbackslash usepackage\{graphicx\}. Number figures sequentially, for example, figure 1, and so on.

The illustration number and caption should appear under the illustration. Labels, and other text with the actual illustration must be at least nine-point type.

If your paper includes illustrations that are not compatible with PDF\TeX{} (such as .eps or .ps documents), you will need to convert them. The epstopdf package will usually work for eps files. You will need to convert your ps files to PDF however.

\subsubsection{Low-Resolution Bitmaps.}
You may not use low-resolution (such as 72 dpi) screen-dumps and GIF files---these files contain so few pixels that they are always blurry, and illegible when printed. If they are color, they will become an indecipherable mess when converted to black and white. This is always the case with gif files, which should never be used. The resolution of screen dumps can be increased by reducing the print size of the original file while retaining the same number of pixels. You can also enlarge files by manipulating them in software such as PhotoShop. Your figures should be 300 dpi when incorporated into your document.

\subsubsection{\LaTeX{} Overflow.}
\LaTeX{} users please beware: \LaTeX{} will sometimes put portions of the figure or table or an equation in the margin. If this happens, you need to scale the figure or table down, or reformat the equation.{ \bf Check your log file!} You must fix any overflow into the margin (that means no overfull boxes in \LaTeX{}). If you don't, the overflow text will simply be eliminated. \textbf{Nothing is permitted to intrude into the margin or gutter.}

\subsubsection{Using Color.}
Use of color is restricted to figures only. It may never be used for any portion of the text of your paper. The archival version of your paper will be printed in black and white and grayscale. Consequently, because conversion to grayscale can cause undesirable effects (red changes to black, yellow can disappear, and so forth), we strongly suggest you avoid placing color figures in your document. If you do include color figures, you must (1) use the CMYK (not RGB) colorspace and (2)  be mindful of readers who may happen to have a color deficiency. Your paper must be decipherable without using color for distinction.

\subsubsection{Drawings.}
We suggest you use computer drawing software (such as Adobe Illustrator or, (if unavoidable), the drawing tools in Microsoft Word) to create your illustrations. Do not use Microsoft Publisher. These illustrations will look best if all line widths are uniform (half- to two-point in size), and you do not create labels over shaded areas. Shading should be 133 lines per inch if possible. Use Times Roman or Helvetica for all figure call-outs. \textbf{Do not use hairline width lines} --- be sure that the stroke width of all lines is at least .5 pt. Zero point lines will print on a laser printer, but will completely disappear on the high-resolution devices used by our printers.

\subsubsection{Photographs and Images.}
Photographs and other images should be in grayscale (color photographs will not reproduce well; for example, red tones will reproduce as black, yellow may turn to white, and so forth) and set to a minimum of 300 dpi. Do not prescreen images.

\subsubsection{Resizing Graphics.}
Resize your graphics \textbf{before} you include them with LaTeX. You may \textbf{not} use trim or clip options as part of your \textbackslash includegraphics command. Resize the media box of your PDF using a graphics program instead.

\subsubsection{Fonts in Your Illustrations}
You must embed all fonts in your graphics before including them in your LaTeX document.

\subsection{References} 
The AAAI style includes a set of definitions for use in formatting references with BibTeX. These definitions make the bibliography style fairly close to the one specified below. To use these definitions, you also need the BibTeX style file ``aaai.bst," available in the author kit on the AAAI web site. Then, at the end of your paper but before \textbackslash end{document}, you need to put the following lines:

\begin{quote}
\begin{small}
\textbackslash bibliographystyle\{aaai\}
\textbackslash bibliography\{bibfile1,bibfile2,...\}
\end{small}
\end{quote}

Please note that you are required to use \textbackslash bibliographystyle\{aaai\} for your references. You may not use named, plain, apalike, acm, ieeetr, siam, chicago, or any other style. Use of natbib is also not acceptable.

The list of files in the \textbackslash  bibliography command should be the names of your BibTeX source files (that is, the .bib files referenced in your paper).

The following commands are available for your use in citing references:
\begin{quote}
\begin{small}
\textbackslash cite: Cites the given reference(s) with a full citation. This appears as ``(Author Year)'' for one reference, or ``(Author Year; Author Year)'' for multiple references.\\
\textbackslash shortcite: Cites the given reference(s) with just the year. This appears as ``(Year)'' for one reference, or ``(Year; Year)'' for multiple references.\\
\textbackslash citeauthor: Cites the given reference(s) with just the author name(s) and no parentheses.\\
\textbackslash citeyear: Cites the given reference(s) with just the date(s) and no parentheses.
\end{small}
\end{quote}

\textbf{Warning:} The aaai.sty file is incompatible with the hyperref and natbib packages. If you use either, your references will be garbled and your paper will not be published.

Formatted bibliographies should look like the following examples.

\smallskip \noindent \textit{Book with Multiple Authors}\\
Engelmore, R., and Morgan, A. eds. 1986. \textit{Blackboard Systems.} Reading, Mass.: Addison-Wesley.

\smallskip \noindent \textit{Journal Article}\\
Robinson, A. L. 1980a. New Ways to Make Microcircuits Smaller. \textit{Science} 208: 1019--1026.

\smallskip \noindent \textit{Magazine Article}\\
Hasling, D. W.; Clancey, W. J.; and Rennels, G. R. 1983. Strategic Explanations in Consultation. \textit{The International Journal of Man-Machine Studies} 20(1): 3--19.

\smallskip \noindent \textit{Proceedings Paper Published by a Society}\\
Clancey, W. J. 1983. Communication, Simulation, and Intelligent Agents: Implications of Personal Intelligent Machines for Medical Education. In \textit{Proceedings of the Eighth International Joint Conference on Artificial Intelligence,} 556--560. Menlo Park, Calif.: International Joint Conferences on Artificial Intelligence, Inc.

\smallskip \noindent \textit{Proceedings Paper Published by a Press or Publisher}\\
Clancey, W. J. 1984. Classification Problem Solving. In \textit{Proceedings of the Fourth National Conference on Artificial Intelligence,} 49--54. Menlo Park, Calif.: AAAI Press.

\smallskip \noindent \textit{University Technical Report}\\
Rice, J. 1986. Poligon: A System for Parallel Problem Solving, Technical Report, KSL-86-19, Dept. of Computer Science, Stanford Univ.

\smallskip \noindent \textit{Dissertation or Thesis}\\
Clancey, W. J. 1979. Transfer of Rule-Based Expertise through a Tutorial Dialogue. Ph.D. diss., Dept. of Computer Science, Stanford Univ., Stanford, Calif.

\smallskip \noindent \textit{Forthcoming Publication}\\
Clancey, W. J. 1986. The Engineering of Qualitative Models. Forthcoming.

\section{Some Common Errors to Avoid}

The following list includes a number of the most common mistakes made by authors when formatting their paper and the reponses that have been sent to the author regarding them:

\begin{small}
\begin{quote}
PDFINFO\\
The following required elements are missing from the preamble to your LaTeX source (or are malformed):
\begin{small}\begin{verbatim}
\pdfinfo{ 
/Title () 
/Author () 
}
\end{verbatim}\end{small}

A4\\
The PDF you submitted is A4. All PDFs must be US Letter sized. Submissions that do not conform to this requirement cannot be published. 
\smallskip

ABOVEDISPLAY\\
You've used \textbackslash abovedisplay, \textbackslash belowdisplay, \textbackslash above caption, and \textbackslash belowcaption to alter aaai18.sty. These commands are illegal and must be stripped from your source.
\smallskip

ABSTRACT INDENTATION\\
Remove the \textbackslash begin\{quote\} and \textbackslash \{endquote\} around your abstract. It might buy you some space.
\smallskip

ASPECT RATIO OF FIGURES CHANGED\\
You may scale the height OR the width of your figures, but not both.
\smallskip

AUTHOR LIST MALFORMED (AUTHBLK)\\
The authblk.sty package cannot be used. 
\smallskip

AUTHOR LIST MALFORMED\\
You did not use the built-in commands in aaai18.sty to format your authors. You need to follow the formatting instructions for authors using LaTeX. 
\smallskip

AUTHOR LIST TOO WIDE\\
Your author list goes beyond the accepted page width. Please break the line and insert \{\textbackslash bf \textbackslash Large\} around the authors on the second line. The author and title information must be presented as specified in the author instructions.
\smallskip

AUTHOR AFFLIATIONS TOO WIDE\\
Your author's affiliations go beyond the accepted page width. Please break the line.
\smallskip

BASELINESTRETCH\\
You've used \textbackslash baselinestretch --- a command specifically prohibited in the author formatting instructions. Please remove it.
\smallskip

CAPTION STYLE ALTERED (CAPTION.STY)\\
You've altered the font and style of captions. Captions must be rendered as specified by aaai.sty. They cannot be altered by changing the font, fontsize, or position, whether manually or by using caption.sty. Such revisions must be removed.
\smallskip

CJK\\
The CJK package cannot be used. Restrict use of non-roman alphabets to figures and tables, which must be compiled and converted to PDF, then imported as a figure into LaTeX.
\smallskip

CLIP\\
You may not use the clip to import only a portion of a graphic. It only places a fragile layer that screens the remained of the figure from view, but imports the entire figure into the PDF. The layer is fragile, and often fails when flattened. The full content of all figures must be imported. In addition and for the same reason, you may not use vspace to reposition a figure. All figures must be completely cropped and adjusted outside of LaTeX before being imported. 
\smallskip

COLUMN WIDTH\\
Unfortunately, your paper contains material that exceeds the column width. Please review your LaTeX log file and correct all the overfull boxes.
\smallskip

COMPILE ERRORS\\
Unfortunately, the source file you submitted will not compile. We are getting undefined control sequence errors. There are either missing packages in your preamble, or missing definitions, or perhaps you are using an old or modified style. For your information, we are using a complete 2017 install of TeXLive, complete with all updates as of September 1. The packages listed in your preamble are all installed on all our systems. If you are using a style file that is not in that package, you will need to include it. Whatever the reason, we need you to correct your LaTeX and resubmit your source files. (If your paper contains graphics, be certain that the fonts are embedded.) 
\smallskip

EPSF / EPSFIG / PSFIG\\
The epsf, epsfig, and psfig packages are obsolete. Use the graphicx package instead.
\smallskip

EULER\\
The euler package is obsolte
\smallskip

FIGURES HAVE BEEN CROPPED WITH LATEX\\
You have used trim and clip commands in your \textbackslash includegraphics statement. Please crop your graphics appropriately using a standard graphics program (not Preview). The masks created by LaTeX are fragile, and disappear when PDFs are combined together. This will result in text in your paper becoming obscured by the fully restored graphic you have imported.
\smallskip

FIGURES MISSING\\
Unfortunately, the source file you submitted are incomplete. You did not include the graphics files. As a result, your paper will not compile. You must upload a new compressed archive that contains all the files required to compile your paper on a different computer and network. (If your paper contains graphics, be certain that the fonts are embedded.) 
\smallskip

FLOATS ALTERED\\
You've altered floats.  The \textbackslash renewcommand cannot be used. The float package cannot be used. Please remove them.
\smallskip

FOOTNOTE STYLE ALTERED\\
Footnotes must follow the built-in style.
\smallskip

FONTS (EMBEDDING)\\
Unfortunately, one or more of the fonts in your paper (most likely the figures) were not embedded in your PDF. This can cause hidden and silent changes to characters once the paper is published, especially when it is combined with other PDFs in the proceedings. If you need to know how to verify the fonts in your figures, please search for ``font embedding pdf." You will find a number of methods, one or more of which may be suitable for your operating system and available software. We recommend using Acrobat Reader to check document properties in your PDF, as it is standards-based software. Consequently, you must embed all the fonts in all your graphics files. If you do not know how to do this, change them to png or jpg. 
\smallskip

FONTS (LANGUAGE)\\
We cannot accept files that require custom installation of non-roman font sets, including CJK and arabic.
 \smallskip

FONTS (TYPE 3)\\
Unfortunately, one or more of the fonts in your paper (most likely the figures) are type 3 PDF. If in your figures, you must either change them to type 1 embedded fonts or change them to png or jpg. If the problem is in your text, you will need to switch the package you are using that is calling type 3 fonts to one that uses type 1 fonts. (Blackboard fonts are often the problem in these cases.)
\smallskip

FULL PAGE USED\\
You used \textbackslash usepackage\{full page\}, which is a package specifically disallowed. You must remove it.
\smallskip

GEOMETRY USED\\
You used \textbackslash usepackage\{geometry\}, which is a package specifically disallowed. You must remove it.
\smallskip

GRAPHICS\\
The graphics package is obsolete. Use the graphicx package instead.
\smallskip

HYPERREF\\
The hyperref package may never be used. If you are using it for URLs, use url.sty instead
\smallskip

INCOMPATIBLE SCRIPT OR MACRO\\
You've used an incompatible script (such as \textbackslash maltepaper.sty) to automate completion of the title and author information on your paper and in the metadata. Such scripts impede correction of your paper and often are incompatible with Acrobat (producing corrupted metadata fields). Scripts and macros cannot be used to automate population of metadata, author or title information, or insertion of figures or tables.
\smallskip

LAYOUT\\
The layout package cannot be used. Use url.sty instead.
\smallskip

LINESPREAD\\
\textbackslash linespread alters aaai.sty. This command (along with \textbackslash baselineshift and others that alter aaai.sty) cannot be used. You must remove it.
\smallskip

LMODERN\\
You used \textbackslash usepackage\{lmodern\}, which is a package specifically disallowed. You must remove it.
\smallskip

LAYOUT USED\\
You used \textbackslash usepackage\{layout\}, which is a package specifically disallowed. You must remove it.
\smallskip

MBOX USED FOR FIGURES\\
You cannot use mboxes to insert figures. They circumvent the spacing requirements for aaai.sty. Use subfigures instead.
\smallskip

MULTIPLE TEX FILES\\
You did not submit a single .tex file as specified in the instructions. We require this to avoid production errors and to facilitate debugging. The \textbackslash input command is not allowed in your source. Please combine all your tex files into a single .tex file (references can remain in .bbl or .bib file).
\smallskip

NAMEREF\\
The nameref package cannot be used.
\smallskip

NATBIB\\
Natbib is incompatible with aaai18.sty and aaai.bst. It must be removed. See References Malformed below.
\smallskip
\smallskip

PAGE BREAKS \\
You've included pagebreaks in your final submission. (They are not required before your references accepted final papers and must be removed).
\smallskip

PAGE NUMBERS\\
Your paper contains page numbers. You need to remove them.
\smallskip

PDFLaTeX NOT USED\\
Your source will not compile in PDFLaTeX. You must convert your eps graphics to PDF.
\smallskip

PDFCOMMENT\\
You used pdfcomment, which is a disallowed package. You must remove it.
\smallskip

PSTRICKS\\
You used pstricks, which is a package specifically disallowed. You must remove it. (Use Tikz instead). If you are simply adding labels, use pinlabel.sty.
\smallskip

REFERENCES ARE INDENTED\\
You've indented your references. Remove the \textbackslash begin\{quote\} and \textbackslash end{quote}
\smallskip

REFERENCES MALFORMED\\
You've used natbib, which is incompatible with aaai.bst.

As an alternate to this, try this workaround:

\begin{quote}\begin{scriptsize}\begin{verbatim}% \usepackage{natbib}
\newcommand{\citet}[1]
{\citeauthor{#1}~\shortcite{#1}}
\newcommand{\citep}{\cite}
\newcommand{\citealp}[1]
{\citeauthor{#1}~\citeyear{#1}}\end{verbatim}\end{scriptsize}\end{quote}
\smallskip

REFERENCES TOO SMALL\\
Your references are too small. They cannot be any smaller than 9 pt (\textbackslash small). Please correct this.
 \smallskip

SAVETREES\\
The savetrees may not be used.
\smallskip

SETLENGTH\\
You've used \textbackslash setlength to alter textfloats. This command is specifically disallowed. Please strip it from your source and recompile.
\smallskip

SETSPACE\\
The setspace package may not be used.
\smallskip

SPACE ABOVE, BELOW ELEMENTS ALTERED\\
You've adjusted the space above or below tables, figures, captions, floats, and/or section/subsection/subsubsections. These changes alter aaai.sty and are not allowed. Please remove them and either edit your text or resize figures or both.
\smallskip

SOURCE FILES MISSING\\
Your compressed archive does not contain *ALL* your source files (a stand-alone archive containing all the files necessary to recompile your paper in LaTeX). Consequently, your LaTeX file will not compile. You must create another compressed archive in which you have included all the files necessary to compile your paper on a separate computer. We cannot publish your paper without these files.
\smallskip

SPACE ALTERED ABOVE, BELOW FLOATS AND SECTIONS\\
You've adjusted the space above or below tables, figures, captions, and/or section/subsection/subsubsections. These changes alter aaai.sty and have resulted in a paper that is more difficult to read. Please remove them and either edit your text or resize figures or both.
\smallskip

STYLE FILE OF BIBLIOGRAPHY IS INCORRECT\\
You did not use aaai.bst (you used named.bst instead). This is not an IJCAI conference. You need to recompile your bibliography using the correct style for AAAI-18.
\smallskip

STYLE FILE IS INCORRECT\\
You did not use aaai18.sty. This style file is required. No other style file (including previous versions of aaai.sty) may be used.
\smallskip

STYLE FILE NOT IN CTAN\\
All styles must be available in the CTAN archive. If they are custom, you need to remove them and put the contents of the custom style directly in your preamble so that our source checker will allow compilation of your file.
\smallskip

TIMES PACKAGE NOT USED\\
Your paper is not formatted using the Times package, either because it is missing completely from your preamble, or because you are using another package that is cancelling use of times. This package is required.
\smallskip

TITLESEC\\
The title sec package may not be used.
\smallskip

VSPACE USED\\
You've used negative vspace around section / subsections. These must be removed.
\smallskip

TIMES PACKAGE MISSING\\
You neglected to use \textbackslash usepackage\{times\}. This font is required.
\smallskip

TITLE AND AUTHOR INFORMATION MALFORMED\\
You need to remove the formatting commands from your title and author information. The style for this information is governed by the conference style file. It cannot be altered.
\smallskip

TITLE BOX TOO SMALL\\
Your author-title-affiliation data exceeds the limit of the title box. To
fix this, you need to increase the size of the title box. You can do this by
placing
\textbackslash setlength\textbackslash titlebox\{2.5in\} in your preamble. You may need to increase the measurement until
the log file error goes away. (Please note: This command cannot be used elsewhere
in the paper, and cannot be used to shrink the size of the title box.)
\smallskip

TITLE IN SENTENCE CASE\\
Your title is formatted in sentence case. Title Case (Mixed Case) is required. Please capitalize the initial letters of all words in your title except for conjunctions and prepositions. You need to do this in your source, in the \textbackslash pdfinfo in your source, and on the submission website.
\smallskip

TITLE MALFORMED (TITLESEC)\\
You've used titlesec.sty, which alters the AAAI style. It must be removed.
\smallskip

TOCBIBIND\\
The tocbibind package may not be used.
\smallskip

ULEM\\
The ulem package may not be used.
\smallskip

T1ENC\\
The T1enc package is obsolete. Us the CM Super Fonts package instead.

\end{quote}
\end{small}

\section{Producing Reliable PDF\\Documents with \LaTeX{}}
Generally speaking, PDF files are platform independent and accessible to everyone. When creating a paper for a proceedings or publication in which many PDF documents must be merged and then printed on high-resolution PostScript RIPs, several requirements must be met that are not normally of concern. Thus to ensure that your paper will look like it does when printed on your own machine, you must take several precautions:
\begin{itemize}
\item Use type 1 fonts (not type 3 fonts)
\item Use only standard Times, Nimbus, and CMR font packages (not fonts like F3 or fonts with tildes in the names or fonts---other than Computer Modern---that are created for specific point sizes, like Times\~{}19) or fonts with strange combinations of numbers and letters
\item Embed all fonts when producing the PDF
\item Do not use the [T1]{fontenc} package (install the CM super fonts package instead)
\end{itemize}

\subsection{Creating Output Using PDF\LaTeX{} Is Required}
By using the PDF\TeX{} program instead of straight \LaTeX{} or \TeX{}, you will probably avoid the type 3 font problem altogether (unless you use a package that calls for metafont). PDF\LaTeX{} enables you to create a PDF document directly from \LaTeX{} source. The one requirement of this software is that all your graphics and images must be available in a format that PDF\LaTeX{} understands (normally PDF, jpg, or png).

PDF\LaTeX{}'s default is to create documents with type 1 fonts. If you find that it is not doing so in your case, it is likely that one or more fonts are missing from your system or are not in a path that is known to PDF\LaTeX{}.

\subsubsection{dvipdf Script}
Scripts such as dvipdf which ostensibly bypass the Postscript intermediary should not be used since they generally do not instruct dvips to use the config.pdf file.

\subsubsection{dvipdfm}
Do not use this dvi-PDF conversion package.

\subsection{Ghostscript}
\LaTeX{} users should not use GhostScript to create their PDFs.

\subsection{Graphics}
If you are still finding type 3 fonts in your PDF file, look at your graphics! \LaTeX{} users should check all their imported graphics files as well for font problems.

\section{Proofreading Your PDF}
Please check all the pages of your PDF file. Is the page size A4? Are there any type 3, Identity-H, or CID fonts? Are all the fonts embedded? Are there any areas where equations or figures run into the margins? Did you include all your figures? Did you follow mixed case capitalization rules for your title? Did you include a copyright notice? Do any of the pages scroll slowly (because the graphics draw slowly on the page)? Are URLs underlined and in color? You will need to fix these common errors before submitting your file.

\section{Improperly Formatted Files }
In the past, AAAI has corrected improperly formatted files submitted by the authors. Unfortunately, this has become an increasingly burdensome expense that we can no longer absorb. Consequently, if your file is improperly formatted, it will not be included in the publication. If time allows, however, you will be notified via e-mail of the problems with your file and given the option of correcting the file yourself. A resubmission fee (minimum \$50.00, and likely higher) will be required for this service). Optionally, you may ask that AAAI have the file corrected for you, for an additional fee. If you opt to correct the file yourself, please note that we cannot provide you with any additional advice beyond that given in your packet. Files that are not corrected after a second attempt will not be included in the publication.

\subsection{\LaTeX{} 209 Warning}
If you use \LaTeX{} 209 we will not be able to publish your paper. Convert your paper to \LaTeX{}2e.

\section{Naming Your Electronic File}
We request that you name your \LaTeX{} source file with your last name (family name) so that it can easily be differentiated from other submissions. If you name your files with the name of the event or ``aaai" or ``paper" or ``camera-ready" or some other generic or indecipherable name, you bear all risks of loss --- it is extremely likely that your file may be overwritten.

\section{Submitting Your Electronic Files to AAAI}
Submitting your files to AAAI is a two-step process. It is explained fully in the author registration and submission instructions. Please consult this document for details on how to submit your paper.

\section{Inquiries} 
If you have any questions about the preparation or submission of your paper as instructed in this document, please contact AAAI Press at the address given below. If you have technical questions about implementation of the aaai style file, please contact an expert at your site. We do not provide technical support for \LaTeX{} or any other software package. To avoid problems, please keep your paper simple, and do not incorporate complicated macros and style files.

\begin{quote}
\noindent AAAI Press\\
2275 East Bayshore Road, Suite 160\\
Palo Alto, California 94303\\ 
\textit{Telephone:} (650) 328-3123\\ 
\textit{E-mail:} See the submission instructions for your particular conference or event.
\end{quote}

\section{Additional Resources}
\LaTeX{} is a difficult program to master. If you've used that software, and this document didn't help or some items were not explained clearly, we recommend you read Michael Shell's excellent document (testflow doc.txt V1.0a 2002/08/13) about obtaining correct PS/PDF output on \LaTeX{} systems. (It was written for another purpose, but it has general application as well). It is available at www.ctan.org in the tex-archive.

\section{ Acknowledgments}
AAAI is especially grateful to Peter Patel Schneider for his work in implementing the aaai.sty file, liberally using the ideas of other style hackers, including Barbara Beeton. We also acknowledge with thanks the work of George Ferguson for his guide to using the style and BibTeX files --- which has been incorporated into this document  --- and Hans Guesgen, who provided several timely modifications, as well as the many others who have, from time to time, sent in suggestions on improvements to the AAAI style.

The preparation of the \LaTeX{} and Bib\TeX{} files that implement these instructions was supported by Schlumberger Palo Alto Research, AT\&T Bell Laboratories, Morgan Kaufmann Publishers, The Live Oak Press, LLC, and AAAI Press. Bibliography style changes were added by Sunil Issar. \verb+\+pubnote was added by J. Scott Penberthy. George Ferguson added support for printing the AAAI copyright slug. Additional changes to aaai.sty and aaai.bst have been made by the AAAI staff.

\bigskip
\noindent Thank you for reading these instructions carefully. We look forward to receiving your electronic files!

\end{document}

### 2019-02-13 WHY19/Draft/Paper.tex
con comentarios de Danks

--- 2018-12-05 WHY19/Draft/Paper.tex
@@
 \usepackage{graphicx}  %Required
+\usepackage{amssymb,amsthm,amsmath,latexsym}
 \frenchspacing  %Required
@@
 \setlength{\pdfpageheight}{11in}  %Required
+\theoremstyle{plain}
+\newtheorem{teo}{Theorem}
+\newtheorem{prop}[teo]{Proposition}
+\newtheorem{defi}[teo]{Definition}
+\newtheorem{obs}[teo]{Observation}
+\newtheorem{lem}[teo]{Lemma}
+\newtheorem{cor}[teo]{Corolary}
+\newtheorem{conj}{Conjecture}
 \pdfinfo{
@@
 \author{M. Gonzalez-Soto, L.E. Sucar, H.J. Escalante\\
-Address line
+\texttt{\{mauricio,esucar,hugojair\}@inaoep.mx}\\
+Coordinaci\'on de Ciencias Computacionales,\\
+Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica (INAOE)\\
+Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla, M\'exico.
 }
@@
 \begin{abstract}
-We define a Causal Decision Problem as a Decision Problem where available actions and outcomes are connected through the variables of a Causal Graphical Model $\mathcal{G}$ and propose a solution criteria based on Pearl's Do-Calculus and the Expected Utility criteria for rational preferences.
-The implementation of this criteria leads to an on-line decision making procedure that has been shown to have similar performance to classic Reinforcement Learning algorithms while allowing for a causal model of an environment to be learned.
+We define a Causal Decision Problem as a Decision Problem Under Uncertainty where the available actions, the family of uncertain events and the set of outcomes are related through the variables of a Causal Graphical Model $\mathcal{G}$.
+We assume rational preferences for the decision maker and based from the Expected Utility criteria we propose an optimality criteria which uses makes use of the causal nature of the environment from the agent's point of view (his beliefs).
+The implementation of this criteria leads to an on-line decision making procedure that has been shown to have similar performance to classic Reinforcement Learning algorithms while also allowing for a causal model of an environment to be learned.
+Thus, we aim to provide the theoretical guarantees of the usefulness and optimality of a decision making procedure based on causal information.
 \end{abstract}
@@
 \section{Introduction}
-\noindent
+Decision making under uncertainty is a fundamental part of intelligent reasoning \cite{lake2017building} and many real-world applications rely on decisions made by an autonomous agent, such as self-driving cars.
+Current decision making methods rely on associative methods, which find only statistical patterns in data.
+On the contrary, causal knowledge allows both for planning and counterfactual reasoning as well as interpretability and explainability \cite{spirtes2000causation}, \cite{woodward2005making}, \cite{pearl2018why}.We propose in this work a way of considering causal information for decision making under uncertainty with rational preferences in such a way that optimal actions are chosen according to the principle of Maximum Expected Utility, which is the formal criteria for making choices under uncertain conditions if rationality is assumed \cite{bernardo2000bayesian}.
 
 \section{Rationality and Expected Utility}
+Rationality in a Decision Making setting is defined axiomatically in a way that the \textit{preferences} of a decision maker are logically consistent.
+If rational preferences are assumed, then it is known that the coherent criteria for making choices is the maximization of expected utility, either with respect to a known utility function and probability distribution or a pair of \textit{subjective} objects \cite{bernardo2000bayesian}, \cite{gilboa2009decision}.
+Rational decision making has been the standard theory in economics both as a \textit{normative} and a \textit{descriptive} theory of human behavior and it has been the subject of multiple debates \cite{tversky1974judgment}, \cite{kahneman1982judgment}.
+In this work we pretend to take a \textit{normative} view for a rational decision maker who faces an uncertain environment which is controlled by some unknown causal mechanism.
 
 \section{Optimal Policies}
+Given a Reinforcement Learning Problem defined over a Markov Decision Process, a \textit{policy} is a function from the space of \textit{states} to the space of \textit{actions} which is interpreted as what should an agent do in a given state \cite{sutton1998reinforcement}.
+An \textit{optimal policy} is a policy which is optimal in the sense of achieiving the maximum possible expected reward.
+Optimal policies can be characterized by the Bellman equations \cite{puterman2005markov}, from where it can be shown to beequivalent to finding the optimal action in the sense of the maximum expected utility \cite{webb2007game}.
+
+\section{Related Work}
+Human use and learning of Causal Relations has been extensively studied by Cognitive Scientists.
+In particular, \cite{hagmayer2009decision}, \cite{wellen2012learning} \cite{hagmayer2013repeated} show that human beings conceive their actions in their environment as interventions over it and they are able to learn, use, and modify previous causal knowledge during a sequential decision process.
+
+From the Machine Learning point of view,  a \textit{bandit problem} where the actions available to an autonomous agent are interventions over a known causal model have been previously considered and it has been shown that an optimal actions, for a finite number of learning rounds, can be learned faster than if causal information hadn't been considered \cite{lattimoreNIPS2016}.
+In this work it is required that the causal model is known, this assumption later relaxed by \cite{sen2017identifying} who considers as unknown only a part of the causal model.
+
+Our formulation of a Causal Decision Problem attempts to give a framework for an agent to learn optimal actions where a causal model controls his environment and the agent is aware of this.
 
 \section{Causal Decision Problems}
+We define a Causal Decision Problem under Uncertainty (CDPU) as a tuple $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \mathcal{G}, \succeq)$ where  $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \succeq)$ is a classical Decision Problem under Uncertainty \cite{bernardo2000bayesian} and $\mathcal{G}$ is a Causal Graphical Model \cite{sucar2015probabilistic} such that the set of available actions $\mathcal{A}$ and the set of outcomes $\mathcal{C}$ are related through the variables of the Causal Model $\mathcal{G}$; i.e., the events in the family $\mathcal{E}$ correspond to variables in $\mathcal{G}$.
+It is assumed that the agent does not know the Causal Model $\mathcal{G}$, which is equivalent to not knowing the probabilities of the events $E \in \mathcal{E}$.
+The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions \cite{woodward2005making} and to satisfy the conditions expressed in \cite{spirtes2000causation}.
+The variable in $\mathcal{G}$ which encodes the consequence of the action taken by the agent will be referred as \textit{target variable} since it is the variable where the agent whishes to obtain a desired result.
+
+In this way, in a CDPU we have a rational agent who chooses an action $a$ among the many available in $\mathcal{A}$, then this action will produce some random effect in the environment which will \textit{cause} a certain consequence, or outcome $c \in \mathcal{C}$.
+Since
 
 \section{Proposed Solution}
+Since rationality is assumed we must seek to maximize the expected utility of the agent in terms of his current knowledge, which is expressed as a (subjective) probability distribution.
+Using the awareness of the agent about a causal mechanism governing the environment, intuition encourages to use causal relations to cause some desirable action, as expressed by \cite{joyce1999foundations}
+
+Consider a Causal Decision Problem as stated above where the target variable, call it $Y$, takes its values in the set $\{0,1\}$ and without loss of generality assume that $Y=1$ is the desired output for the agent, then, according to the maximum expected utility principle, as well with Joyce intuition (\cite{joyce1999foundations}), he must choose the action $a^\ast \in \mathcal{A}$ such that
+\[ P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}, \]
+where $P$ is a probability distribution which encodes the agent knowledge, or beliefs, about the causal relations that hold in his evironment. This is equivalent to finding $a^\ast$  as
+\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g). \]
+Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent.
+If an action $a_0 \in \mathcal{A}$ yielded a higher expected utility, that would imply that it has higher probability of \textit{causing} the same desired action, but this is not possible because of how $a^\ast$ is obtained.
+
+The procedure for selecting $a^\ast$ connects both the utility for the decision maker (because of the most desired outcome) and his beliefs (the measure $P$) about \textit{causing} a desired result.
+Since the causal mechanism and the uncertainties in the environment are connected, $a^\ast$ to maximizes expected utility.
+This procedure allows to choose a \textit{better action} in terms of what is currently known about causal relations.
+
+Since $a^\ast$ was found by asking the probability of \textit{causing} the most desired action for the agent, it is also a counter-factual distribution since we know that the probability of observing the most desired outcome is lower for any other action $a \in \mathcal{A}$.
+
+While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
+\[ U(x)=\sum_y P(y | do(x)) u(y), \]
+we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment.
+Also, the only formal criteria for making choices in an uncertain environment is the maximization of expected utility, as shown by vonNeumann and Morgenstern, De Finetti and Savage.
+The expected utility is taken with respect to a subjective utility function and to either the known probabilities, or a subjective probability measure which encodes the agents \textit{beliefs}.
+It is not shown by Pearl that his function $U$ is equivalent to the maximization of expected utility in the context of Savage's Theorem.
 
 \section{On-line decision making and causal learning}
+In \cite{gonzalez2018playing} a decision-making procedure was proposed using the Proposed Solution together with Bayesian \textit{belief updating} procedure in order to learn an optimal action while parallely acquiring a causal model in the environment while using the current causal knowledge to make choices.
+We note that our proposal makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.
 
+This procedure was applied in the simpler case where the decision maker knows the \textit{structure} of the model $\textit{G}$.
+In the referred work, a decision maker held \textit{beliefs} about the causal information of the environment, which were encoded into probability distributions.
+Those beliefs were used \textit{as if} they were the true model governing the environment in order to choose an action according to the Solution proposed here.
+Beliefs were updated in a Bayesian way after observing the causally produced outcome, or consequence, of the action chosen by the agent.
+
+The actions learned after a series of decision rounds obtained similar performance (in terms of average reward) as an agent learning using the classical Q-Learning procedure.
+Experimentally, this shows that causal information allows an agent to learn an optimal action, in the sense of expected utility, as well as learning a causal model of the environment.
+
+\section{Experiments}
+In \cite{gonzalez2018playing} a test scenario about a medic who is trying to learn an optimal treatment for a sick patient was used in order to show how causal information could effectively guide a decision making process while also allowing for learning of a causal model.
+
+We reproduce here the results obtained in \cite{gonzalez2018playing}, which show the average reward of a \textit{causal agent}, where the agent knows the structure of the graphical model and helds beliefs about the \textit{parameters} of the true causal model.
+The beliefs of the causal agent are used in each decision round as if they were the truth about the causal relations and used as stated above.
+Comparison against an agent simply choosing at random is also shown.
+
+In Figure \ref{100_rounds} we observe the average reward obtained by the three agents in 100 rounds, where our algorithm slightly outperforms Q-Learning.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{100_rounds_format.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{100_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+In Figure \ref{200_rounds} we observe the average reward obtained by the three agents in 200 rounds. The average reward obtained is very similar for Q-learning and our algorithm.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{200_rounds_format.png}}
+\caption{Average reward obtained in each round for each agent}
+\label{200_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\section{Future Work}
+Numerical results show that the trayectories of the average rewards both from our Causal Agent and the Q-agent seem to stay together after some number of rounds, while leaving behind the random-agent.
+In order to get a valid form of concurrent validation it is required that this behavior will remain like that from a certain point.
+\\
+We thus state a conjecture which is yet to be proven:
+\begin{conj}
+Let $(X_1,X_2,...) \in \mathbb{R}^\infty$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...) \in \mathbb{R}\infty$ are the rewards obtained by a decision-making procedure that uses causal information in the way we propose, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$.
+\end{conj}
+
+\section{Conclusions}
+We have proposed an optimality criterion for decision making under uncertainty when the environment where the agent is situated is governed by an unknown causal mechanism.
+This criteria, when used to build a decision making procedure yields similar performance as classical algorithms which aim towards the same objective: maximizing expected utility, thus showing that causal information, and causal-based decision making is at least as useful as associative-based case, while it also allows to learn a causal model of the environment.
+The experiments shown serve as a form of concurrent validation, where our proposed method is compared to a decision-making method that it is known to learn optimal policies (i.e., maximize expected utility)
+
+Learning a causal model is useful because of the interpretability and explainability that it provides when analyzing \textit{why} a particular decision was made.
+For the hypothetical scenario used in the experiments, the causal model allows for further inquiries about the choices made by the medic.
+As \cite{pearl2018why} mention, the three levels of causal learning are observing, intervening and counterfactual reasoning.
+Our proposed decision-making method based on the solution stated above allows for each of the three levels to be used.
+First of all, it allows to observe (and learn) of effects of interventions.
+In second place, it allow to intervene, and in third place to have the ability of explaining \textit{why} a particular choice was made in terms of the effects it would produce given a certain level of causal knowledge.
+
+It is left as future work to provide a decision making procedure when the causal model is completely unknown for the agent and to provide theoretical convergence results both to the true causal model and the optimal action.
+
+\section{Acknowledgements}
+The authors are deeply thankful for the useful comments and observations from Dr. David Danks. 
 \bibliography{Bibliografia.bib}

### 2019-02-13 Reporte_Tecnico_1st_year/Technical_Report_1st_Year.tex
cambios recientes

--- 2018-09-02 Propuesta/Formato/Proposal_format.tex
@@
 
-Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting (\cite{sutton1998reinforcement}), but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
+Learning to perform tasks by interaction has been extensively studied in Computer Science using Reinforcement Learning (RL) (\cite{sutton1998reinforcement}), but the most common used techniques in this field are based on purely associative relations and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).
 A particular case of a \textit{higher level structure}; i.e., beyond associative patterns, is the case of \textit{causal} structure.
@@
 \\
-Causal reasoning is found at the very core of human reasoning, since it allows us to manipulate our environment and being able to predict effects of a given action (\cite{spirtes2000causation}).
+Causal reasoning is found at the very core of human reasoning (\cite{danks2014unifying}, \cite{lake2017building}), since it allows us to manipulate our environment and being able to predict effects of a given action (\cite{spirtes2000causation}).
+It is even argued that in a more basic level the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} (\cite{friston2010free}, \cite{hohwy2013predictive}, \cite{clark2015surfing}).
 Causal claims also support reasoning of the form \textit{if...hadn't occurred then...wouldn't have happened}.
@@
 It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons.
-
-Learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
+Also, learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
 \subsection{Outline}
@@
 \section{Main Concepts}{\label{Main_Concepts}}
-	\subsection{Attempts to define Causality}
+
+\subsection{Causality}
+Since Causality is at the core of this proposal, we must make a careful treatment of it, both in its definition and use.
+Leaving aside deeper philosophical and epistemological issues, we consider \textit{causal structre} as part of of the world itself, something that holds deep inside the tissue of the physical world.
+We consider that the notion of causal relations does not contradict the  stochastic nature of the world in the sense of \citep{gilboa2009decision}, since we can both know that throwing a dice is both a stochastic phenomenon while also acknowledging that throwing it caused a stochastic quantity to be realized.
+
+\subsection{Difference between Causal Relations and Associative Relations}
+As \cite{pearl2018why} mentions, there are three levels of causal reasoning: observing, manipulating and counterfactual reasoning.
+Associative relations belong to the first level, while causal reasoning encompasses the second and third levels the core of reasoning with causal relations.
+Associative relations are based on correlation between events, or patterns of occurence to be found in data, while causal relations are based in cause-efect patterns, which result from some mechanism which is rooted in the nature of the observed events.
+As an example, although a very simple one, consider the following: a dog that lives as a pet in some household has observed that every day he has food so he does not worry about eating or not, but any kid knows that whether he eats or not depends on certain grown-up being present.
+The dog uses only observed occurrences in order to obtain conclusions, while the kid can imagine what would happen if his parents arrive home late.
+
+It is important to mention that although in the usual examples of causal relations the cause precedes temporally the effect this need not be a general rule.
+It suffices to show one example of a causal relation in which the effect precedes temporally the cause in order to conclude that cause-effect relations need not be in temporal order.
+For example, consider the example of a barometer and a storm; in this case, the barometer drops first, but it is the storm which actually causes this event (\cite{pearl2009causality}).
+
+\subsection{About definitions of Causality}
 Causality has been a difficult concept to define and several attempts have been made.
@@
 Since causal claims in common language can have different logical form, in this work we give a \textit{working} definition of causality that allows a formal language for stating and manipulating causal concepts and we will take the given definition as axiomatic.
+From this point, the words \textit{causality}, \textit{causal relation} and similar terms will be understood in terms of the following formal definition and not to be confused with the ordinary-language words that are spelled in the same way.
 	\subsection{Definition of Causality}
@@
 \subsubsection{The identifiability problem}
-	Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
-It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}).
+The identifiability problem is to find under what are the conditions under which causal inquiries can be answered in terms of purely \textit{observational} data?
+It has been shown (\cite{peters2012identifiability}, \cite{mooij2016distinguishing}) that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence).
 \subsubsection{Do-Calculus}
@@
 We choose to use specifically a Causal Graphical model because of the mathematical machinery that exists around them and because Graphs are a model of Cognitive Processes, such as causal reasoning (\cite{glymour2003learning}, \cite{danks2014unifying}).
-In this way, if an agent \textit{knows} that a certain environment is governed by causal relations, then a causal graphical model can be used to model the causal reasoning of the agent.
+In this way, if an agent \textit{knows} that a certain environment is governed by causal relations, then a causal graphical model can be used to model the process of causal reasoning of the agent in a way that resembles human reasoning.
+
+\subsection{Advantages of Causal Models}
+If a Causal Model is learned from data, and if it is a \textit{good} model, then according to Pearl's levels, the model can be used to predict the effect of certain interventions as well to perform counterfactual reasoning in order to evaluate an intervention that has been made; i.e., what if another intervention had been performed?
+
+A Causal Model can allow for transferability of knowledge to similar domains.
 
@@
 It allows to ask \textit{why}.
+Since a causal model, once it is learned, does not depend neither on the agent nor its preferences, the gained knowledge could be easily transfered for problems of similar domain where the use of the acquired causal relations is useful.
 
@@
 \subsection{Research Question}
-The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
+The proposed research attempts to answer a series of questions about the nature of causal relations and about their possible use.
 \begin{enumerate}
@@
 \item How can a causal structure help a decision maker in order to guide his learning process?
+\item When does causal information will be helpful?
 \item What does the rationality assumption implies about how to choose when considering causal information? 
@@
 \subsubsection{General Objectives}
-			The proposed research has as a general objective to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
+The proposed research has as a general objective to provide the understanding and knowledge about what are the implications of causality for rational decision maker who faces an uncertain environment and about how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
 \subsubsection{Specific Objectives}
+In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first. Given a scenario which can be modelled as a causal decision problem:
 \begin{enumerate}
-		\item In order to achieve the general objective, we need to separate it into smaller problems which are required to be solved first.
-In first place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.
-
-		\item In second place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.
+\item In first place, we require to characterize the problems where this approach can be used.
+
+\item  In second place, it is required to find a general way of defining a game that captures the interaction between a decision maker and a causal-governed environment.
+
+\item In third place, to find a family of distributions that represent any past knowledge and current beliefs about the causal structure of the environment and a way to use these beliefs in order to make choices according to the decision maker's preferences.
 
 \item Beliefs must be updated after external information is provided by the environment in response to any given action that was taken by the decision maker.
-A general updating criteria must be provided which uses the causal nature of the environment.
+A general updating criteria must be provided which explicitly uses the causal nature of the environment.
 \end{enumerate}
@@
 
+\subsubsection{Domain of application}
+Even if, as argued in the Introduction, causal relations are present everywhere, we do not pretend that our proposed method can be applied anywhere without further considerations.
+For the proposed method to be applied, it is necessary that a Decision Problem exists (i.e., a non-empty set of available actions exist which is related to a well-defined goal to be achieved), and that the elements of the decision problem can be causally related through a causal model.
+This is, we require:
+\begin{enumerate}
+\item The main task is to choose an action in order to achieve some objective.
+\item The set of available actions must have at least one element which is known by the decision maker. 
+\item Rational preferences for the decision maker. 
+\item Known environment.
+\item Causal relations in the environment. 
+\item Decision-making steps in discrete time. 
+\end{enumerate}
+
 \subsubsection{Extensions}
@@
 \end{itemize}
+
 \newpage
+
 \section{Related Work}{\label{Related_Work}}
@@
 Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent.
+Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $Y=1$ is the most desired outcome for the agent.
 Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
-\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
+\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}, \]
+where $P$ is a probability distribution which encodes the agent knowledge, or beliefs, about the causal relations that hold in his evironment. This is equivalent to finding $a^\ast$  as
+\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g). \]
+Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent.\\
+\\
+While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
+\[ U(x)=\sum_y P(y | do(x)) u(y), \]
+we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment.
+We note that our proposal makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.
 \end{prop}
+
 \subsection{Modelling a Decision Problem under Uncertainty as a Game}
@@
 Propose a solution using the rationality assumption.
-\item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered
-\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information.
-\item Study the problem of on-line causal discovery.
+\item Propose a representation of beliefs that allow causal structure to be considered. Three subcases considered. Conference Paper No. 1
+\item Propose and evaluate an updating criteria for causal beliefs that can be updated according to new information when involved variables are known. Conference Paper No. 2
+\item Study the problem of on-line causal discovery (fully unknown model). Convergence of proposed solution. Conference Paper No. 3 and Journal Paper.
 \item Considering interventions in more than one variable.
@@
 		\end{itemize}
-\item \textbf{Year 2:} Conference Paper
-	\begin{itemize}
-	\item When a causal model is unknown, how to encode uncertainties about causal structure in a probabilistic distribution? Bayesian Nonparametric prior that explicitly represents causal structure.
+\item \textbf{Year 2:} Conference Paper 2 and 3
+	\begin{itemize}
+	\item When a causal model is unknown but the involved variables are known, how to encode uncertainties about causal structure in a probabilistic distribution?
+Bayesian Nonparametric prior that explicitly represents causal structure.
+	\item When variables of the environment are unknown how to encode prior beliefs and how to update them when data is not fully observed.
 	\item Possible Forum: International Conference on Bayesian Nonparametrics, International Conference on Probabilistic Graphical Models.
@@
 \end{itemize}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\includegraphics[scale=0.5]{Gantt.eps}
+\caption{Gantt Diagram for the proposed research}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
 
@@
 
+We observe that when the number of rounds is small, as in Figure 2, Q-learning outperforms our method, which achieves similar performance than the random-acting agent.
+On the other hand, in 50 rounds (Figure 3), our Causal Agent and the Q-Agent have left behind the Random Agent, which is even more noticeable by 200 rounds.
+This means that our agent (Causal Agent) is both learning a good action while learning a Causal Model of the environment that he is facing.
+
 \subsection{Case 3 and future work: The model is not known}
@@
 
-Belief updating using causal information when the decision maker doesn't know the structure of the model nor its parameters is yet to be studied and left as future work.
-
+Belief updating using causal information when the decision maker current knowledge is specified as a Bayesian Nonparametric Prior (BNP)  is yet to be studied and left as future work, since the construction and use of this kind of priors is a hard problem, this is because a BNP is basically a probability distribution over infinite-dimensional spaces, such as the space of continuous functions or the space of unbounded graphs (\cite{hjort2010bayesian}, \cite{phadia2015prior}, \cite{ghosal2017fundamentals}).
+\newpage
+\section{Conclusions}
+We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism can make use of the causal relations that hold in such environment in order to find an action that will take him into the way of a desired result, in terms of his (rational) preferences.
+Also, causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.
+
+We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives.
+Following this intuitions we proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality.
+From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order.
+In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment.
+We are considering.
+If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
+\newpage
 %------------- References --------------------

### 2019-02-27 UAI_2019/UAI_Draft.tex
lo que va de paper UAI

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[letterpaper]{article}
\usepackage{uai2019}
\usepackage[margin=1in]{geometry}

% Set the typeface to Times Roman
\usepackage{times}

\title{Structure Learning for Causal Decision Making}

\author{Anonymous authors.} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\begin{document}

\maketitle

\begin{abstract}
In a recent paper by \cite{2019arXiv190202279G} Decision Problems under Uncertainty where the uncertain events are causally related where studied. For such problems, a criteria for choosing an optimal action was proposed for rational preferences. It had been shown by such authors in \cite{gonzalez2018playing} that by implicitly considering a distribution over causal structures and using such criteria within each structre then an on-line decision-making algorithm could be obtained, which achieved a similar performance than the classical Q-learning. Their work has a limitation, since they only considered the case where the decision maker knows the structure of the causal model that controls the environment. In this paper, we propose to extend such limitation by proposing a local structure learning procedure in terms of belief updating. 
\end{abstract}

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning (\cite{danks2014unifying}, \cite{lake2017building}). It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes, either by knowing the probabilities in the environment, or by considering a \textit{subjective} probability measure (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}). We consider cases where the uncertainty in the environment is controlled by a \textit{causal mechanism}, represented as a Causal Graphical Model which is always unknown by the agent.

\section{On Causality and Learning}
Several definitions of Causality have been attempted by philosophers and scientists, and we acknowledge that such task is a complex one. We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}. We consider Causal Graphical Models (\cite{koller2009probabilistic}) as representations of the causal relations that hold in an environment.

From the human learning perspective, it has been shown that humans acquire and use causal information for decision-making and that humans even consider their actions as interventions in the world (\cite{hagmayer2009decision}, \cite{wellen2012learning} \cite{hagmayer2013repeated}). It has also been argued that humans tend to ignore probabilistic information and assign more importance to causal information (\cite{tversky1980causal}, \cite{pearl2009causality}). Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}, \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learna causal.

\section{Rationality, Decision Making and Expected Utility}

\section{Related Work}
\cite{gonzalez2018playing} propose that by holding \textit{beliefs} about causal relations and using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be learned. Such procedure implicitly considers a distribution over causal structures, where each structure is used locally to make a choice. In \cite{2019arXiv190202279G}, such authors propose a formal criteria which has the form:
\[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker. We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent

While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
\[ U(x)=\sum_y P(y | do(x)) u(y), \]
we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment. We note that the criteria proposed by \cite{2019arXiv190202279G} makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.

Using a version of the criteria described above, in \cite{gonzalez2018playing} a decision-making procedure is proposed for the case when the agent knows the \textit{structure} of the Causal Model which controls his environment. It remains an open question to tackle the problem of an unknown structure.

\section{Structure Learning: a local approach}
We consider the problem of extending the work of \cite{gonzalez2018playing} in the setting of \cite{2019arXiv190202279G}: this is, a $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \mathcal{G}, \succeq)$ where $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \succeq)$ is a classical Decision Problem under Uncertainty as defined in \cite{bernardo2000bayesian} and $\mathcal{G}$ is a Causal Graphical Model (\cite{spirtes2000causation}, \cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) such that the set of available actions $\mathcal{A}$ and the set of outcomes $\mathcal{C}$ are related through the variables of the Causal Model $\mathcal{G}$; i.e., the events in the family $\mathcal{E}$ correspond to variables in $\mathcal{G}$. It is assumed that the agent does not know the Causal Model $\mathcal{G}$, which is equivalent to not knowing the probabilities of the events $E \in \mathcal{E}$. The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions \cite{woodward2005making} and to satisfy the conditions expressed in \cite{spirtes2000causation}. The variable in $\mathcal{G}$ which encodes the consequence of the action taken by the agent will be referred as \textit{target variable} since it is the variable where the agent whishes to obtain a desired result.

Causal information use and learning when the underlying model is completely unknown is a difficult task (\cite{lattimoreNIPS2016}). We will consider a series of assumptions in order to simplify the problem. We consider:
\begin{enumerate}
\item The decision maker knows the variables in the model, let $V$ be the set of variables.
\item The decision maker knows which variable can he intervene.
\item The decision maker knows which variable he wants to affect with his interventions. 
\item The decision maker can fully observe realizations of the variables from the true causal model $\mathcal{G}$, but such realizations are independent.
\item Variables are binary-valued $\{0,1 \}$ and interpreted as \textit{acting} on them or not. 
\end{enumerate}

\subsection{The form of uncertainty}
Our approach begins by noticing that if the decision maker does not know the relations between the (known) variables, then the his uncertainty has the following \textit{form}: Does $A$ cause $B$? For any pair of variables $A,B$ in V. Since we are following a Bayesian approach, we translate this statement into:
\[p_{AB}=P(A \textrm{ cause } B). \]
Now, the idea is to update $p_{AB}$ according to the manipulationist idea stated in \cite{woodward2005making} where a manipulation in $A$ should translate in a change of value in $B$. Since we are assuming that the variables are binary and interpreted as acting on them or not (consider for example giving a certain treatment or not), the agent will increase his probability of $A$ causing $B$ if the pair $(A=1,B=1)$ is observed. Since the probability $p_{AB}$ is considered for any pair $(A,B)$ we must be careful in order not to update also $p_{BA}$ since from \cite{spirtes2000causation} we have that if $A$ causes $B$, then $B$ can not cause $A$. We add an extra assumption: the decision maker has a partial ordering of the variables which corresponds to variables that can not cause each other. This extra assumption, although a strong one, can be thought of expert knowledge in the domain of the decision problem.

\section{Conclusions}

\section{Future Work}

\newpage

\subsubsection*{Acknowledgements}
The authors would like to thank \texttt{<Anonymous>} for valuable comments and insights.

\subsubsection*{References}
%doi=false,isbn=false,url=false
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}

\end{document}

### 2019-03-01 UAI_2019/UAI_Draft.tex
primer version

@@
 \usepackage[margin=1in]{geometry}
-
+\usepackage{breakcites}
+\usepackage{graphicx}
+\usepackage{subfigure}
 % Set the typeface to Times Roman
@@
 \begin{abstract}
-In a recent paper by \cite{2019arXiv190202279G} Decision Problems under Uncertainty where the uncertain events are causally related where studied.
+In a recent paper by \cite{2019arXiv190202279G}, Decision Problems under Uncertainty where the uncertain events are causally related where studied.
 For such problems, a criteria for choosing an optimal action was proposed for rational preferences.
-It had been shown by such authors in \cite{gonzalez2018playing} that by implicitly considering a distribution over causal structures and using such criteria within each structre then an on-line decision-making algorithm could be obtained, which achieved a similar performance than the classical Q-learning.
+It had been shown by such authors in \cite{gonzalez2018playing} that by implicitly considering a distribution over causal structures and using such criteria within each structre then an on-line decision-making algorithm could be obtained, which achieved a similar performance than the classical Q-learning while also acquiring a Causal Model of the environment.
 Their work has a limitation, since they only considered the case where the decision maker knows the structure of the causal model that controls the environment.
-In this paper, we propose to extend such limitation by proposing a local structure learning procedure in terms of belief updating.
+In this paper, we propose to extend such limitation by proposing a local structure learning procedure in terms of belief updating, where the beliefs are probabilities over the events of a variable having causal influence over another one.
+We show experimentally that such method is capable of finding the causal links that hold in a true causal model after a series of observations are drawn from such true model.
 \end{abstract}
@@
 \section{Introduction}
-Decision making under uncertain conditions is a fundamental task for intelligent reasoning (\cite{danks2014unifying}, \cite{lake2017building}).
-It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes, either by knowing the probabilities in the environment, or by considering a \textit{subjective} probability measure (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
-We consider cases where the uncertainty in the environment is controlled by a \textit{causal mechanism}, represented as a Causal Graphical Model which is always unknown by the agent.
+Decision making under uncertain conditions is a fundamental task for intelligent reasoning and a any day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}).
+It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+We consider the case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, represented as a Causal Graphical Model which is always unknown by the agent and assumed to remain fixed and invariant to interventions (\cite{woodward2005making}).
+It has been proposed a mechanism for a decision maker to find good actions by holding beliefs about the causal relations in the environment, using them to make an action and then updating those beliefs in terms of what is caused by a chosen action.
+Such mechanism, which is described in \cite{gonzalez2018playing} is limited since it is assumed that the decision maker knows the \textit{structure} of the Causal Graphical Model.
+Here we attempt to relax such assumption and allowint the decision maker to know only the variables in the model, with the help of an \textit{ordering} in the variables in such a way that it is know which variables can not cause each other.
 
 \section{On Causality and Learning}
-Several definitions of Causality have been attempted by philosophers and scientists, and we acknowledge that such task is a complex one.
+Several definitions of Causality have been attempted by philosophers and scientists, such as those described in \cite{holland1986statistics}.
 We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}.
-We consider Causal Graphical Models (\cite{koller2009probabilistic}) as representations of the causal relations that hold in an environment.
+We consider Causal Graphical Models (\cite{koller2009probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution.
 
@@
 Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}, \cite{danks2014unifying}).
-Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learna causal.
+Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations.
 
 \section{Rationality, Decision Making and Expected Utility}
+In this work we begin with a Decision Problem under Uncertainty where an agent, who has rational preferences, faces the problem of choosing one among many available actions with uncertain consequences.
+From the work of \cite{von1944theory} and \cite{savage1954the}, we know that the only criteria for making choices is the maximization of expected utility with respect to the agent's preferences and a probability distribution which represents either the probabilities of the uncertain events, or the agents own beliefs about such uncertainty.
+In \cite{gonzalez2018playing} and \cite{2019arXiv190202279G} the case where the uncertainty of the environment is controlled by a causal mechanism is considered.
+Since the (unknown) causal mechanism is itself a stochastic one, and the agent's preferences are rational, Savage's theorem state how an action must be chosen: maximizing the expected utility.
+If the decision maker does not know the causal mechanism, this is equivalent to not knowing the probabilities of the environment, so a subjective distribution must be used.
 
 \section{Related Work}
-\cite{gonzalez2018playing} propose that by holding \textit{beliefs} about causal relations and using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be learned.
-Such procedure implicitly considers a distribution over causal structures, where each structure is used locally to make a choice.
+In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered, their assumption of knowing the causal moder is later relaxed by \cite{sen2017identifying} who considers interventions over a partially unknown causal model.
+By following \cite{joyce1999foundations} intuitions, in \cite{gonzalez2018playing} it is proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found.
+Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used locally (in each step) to make a choice.
 In \cite{2019arXiv190202279G}, such authors propose a formal criteria which has the form:
@@
 where $\mathcal{A}$ is the set of available actions to the decision maker and $1$ is the most desired value of the \textit{objective variable} $Y$, for the decision maker.
-We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent
+We notice that since the action chosen is the action with the highest probability of causally producing the most desired action, then $a^\ast$ necesarilly is the action that maximizes the expected utility for the agent.
+The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.
 
@@
 Using a version of the criteria described above, in \cite{gonzalez2018playing} a decision-making procedure is proposed for the case when the agent knows the \textit{structure} of the Causal Model which controls his environment.
-It remains an open question to tackle the problem of an unknown structure.
+It remains an open question to tackle the problem of a fully unknown structure.
+It is known that from observational data alone, a Causal Graphical Model is identifiable up to Markov equivalence without any the assumption of a specific functional model with given error distributions (\cite{hauser2012characterization}, \cite{hauser2012two}, \cite{peters2011identifiability}).
+We pretend to extend here the work of \cite{gonzalez2018playing} to the case of a rational decision maker who knows the variables that control his environment and when he has extra knowledge of variables that can not be cause of each other.
 
@@
 It is assumed that the agent does not know the Causal Model $\mathcal{G}$, which is equivalent to not knowing the probabilities of the events $E \in \mathcal{E}$.
-The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions \cite{woodward2005making} and to satisfy the conditions expressed in \cite{spirtes2000causation}.
+The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions \cite{woodward2005making} and to satisfy the conditions expressed in \cite{spirtes2000causation} together with Causal Sufficiency.
 The variable in $\mathcal{G}$ which encodes the consequence of the action taken by the agent will be referred as \textit{target variable} since it is the variable where the agent whishes to obtain a desired result.
@@
 \begin{enumerate}
-\item The decision maker knows the variables in the model, let $V$ be the set of variables.
-\item The decision maker knows which variable can he intervene.
+\item The decision maker knows the variables in the model, let $V=\{ X_1,...,X_n \}$ be the set of variables.
+\item The decision maker knows which variable can he intervene and this variable does not depend on others.
 \item The decision maker knows which variable he wants to affect with his interventions. 
 \item The decision maker can fully observe realizations of the variables from the true causal model $\mathcal{G}$, but such realizations are independent.
-\item Variables are binary-valued $\{0,1 \}$ and interpreted as \textit{acting} on them or not. 
+\item Variables are binary-valued $\{0,1 \}$ and interpreted as \textit{acting} on them or not. For example, giving a patient a drug or not. 
 \end{enumerate}
+The last assumption appears to be a very strong one, but the multi-valued case can be dealt with by replacing the variable with an artificial one which is binary-valued, where one of the states correspond to the values of the original variable, and the other to doing nothing, as in the binary case.
 
 \subsection{The form of uncertainty}
-Our approach begins by noticing that if the decision maker does not know the relations between the (known) variables, then the his uncertainty has the following \textit{form}: Does $A$ cause $B$?
+Our approach begins by noticing that if the decision maker does not know the relations between the (known) variables, then his uncertainty about the causal structure has the following \textit{form}: Does $A$ cause $B$?
 For any pair of variables $A,B$ in V.
@@
 Since we are assuming that the variables are binary and interpreted as acting on them or not (consider for example giving a certain treatment or not), the agent will increase his probability of $A$ causing $B$ if the pair $(A=1,B=1)$ is observed.
+We know from \cite{holland1985statistics} that manipulation is fundamental to the notion of Causality, and although we notice that our decision maker is merely \textit{observing}, we are interpreting the binary-nature of our variables as observing manipulations over them in the form of \textit{acting} or them or not.
 Since the probability $p_{AB}$ is considered for any pair $(A,B)$ we must be careful in order not to update also $p_{BA}$ since from \cite{spirtes2000causation} we have that if $A$ causes $B$, then $B$ can not cause $A$.
-We add an extra assumption: the decision maker has a partial ordering of the variables which corresponds to variables that can not cause each other.
+We add an extra assumption: the decision maker has a partial ordering $\succeq$ of the variables which corresponds to variables that can not cause each other.
 This extra assumption, although a strong one, can be thought of expert knowledge in the domain of the decision problem.
 
+\subsubsection{Uncertainty Updating}
+Given that the decision maker can observe a full realization of $X_t=(X_1=x_1,...,X_n=x_n)$ of the variables from the true causal model, where $x_i \in \{0,1 \}$, then for any pair $X_i,X_j$ with $X_j \succeq X_i$, if $(X_i=1, X_j=1)$ is observed on stage $t$, then the belief $p_{X_i X_j}$ is increased by de decision maker to:
+\[  p_{X_i X_j}^{(t+1)} = p_{X_i X_j}^{(t)}+\alpha (1-p_{X_i X_j}^{(t)}) \textrm{ for some } \alpha \in (0,1).  \]
+On the other hand, if $(X_i=1, X_j=0)$ is observed on stage $t$, then we decrease $p_{X_i X_j}$ to:
+\[ p_{X_i X_j}^{(t+1)} =   \alpha p_{X_i X_j}^{(t)}  \textrm{ for some } \alpha \in (0,1). \]
+Once the values  $p_{X_i X_j}$ are updated for every $i,j$ such that $X_j \succeq X_i$, the procedure is once again started with a new observation $X_{t+1}$.
+\subsubsection{Obtaining a structure and decision making}
+After a series of rounds of observation and belief updating it is expected that the information provided by the environment, together with the partial ordering, will allow the decision maker to update his beliefs about the existence of a causal relation between variables in such a way that the stronger (bigger than some threshold) values are considered as a relation existing between a pair of variables.
+Once a causal structure $\bar{\mathcal{G}}$ is obtained, the decision maker uses this structure \textit{as if} it were the true causal model and then use the procedure described in \cite{gonzalez2018playing} in order to choose an action and observe the causal consequence of the action; this procedure ends with a vector parameters $\alpha$ over the Conditional Probability Tables of the Causal Model which has structure $\bar{\mathcal{G}}$, and such parameters can be used in order to start again the structure-learning procedure, since a low-value parameter points to a possible non-existence of a causal link in $\bar{\mathcal{G}}$ when the outcome of the action is considered.
+
+\section{Experiments and Results}
+In order to test our method, we use the simple example in \cite{gonzalez2018playing}, in which a sick patient arrives to a hospital and a medic who does not know his disease must give him some treatment, which can have or not a reaction, which can kill him.
+The Causal Graphical Model that controls what happens to the patient is shown in Figure \ref{causal_model}.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Propuesta/causal_graph.png}}
+\caption{The true Causal Graphical Model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
+\label{causal_model}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+We test our method by starting with a random initialization of the probabilities of a variable causing another and updating as mentioned.
+As an external input we included a list of pairs of variables in which the first element can not cause the second one.
+Using this list, a second list of \textit{possible causes} was obtained and to the elements of this list random probabilities were assigned.
+A value of $\alpha=0.8$ was used, and $\alpha$ was decreased by a factor of $1/k$ where $k$ was the current round.
+After 10 rounds of observing realizations from the true model we see in Figure \ref{Results }that the beliefs about causal relations that do hold in Figure \ref{10_rounds} tend to $1$ while non-existing links go down to $0$
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_10_2_ylabel}}
+\caption{Beliefs about the existence of causal links between variables.}
+\label{10_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_25_2.png}}
+\caption{Beliefs about the existence of causal links between variables in 25 rounds.}
+\label{25_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
 \section{Conclusions}
+We have proposed a method for finding a causal structure based on the updating of beliefs by a rational agent about the existence of causal relations between variables.
+The need of a causal ordering is required since it is known that based on mere observations a Directed Graph can not be obtained.
+Our method allows the use of a previously shown method for decision making based on causal information in the case where the decision maker only knows the variables of the model.
+In a simple case, we tested our method and it was shown experimentally that the proposed method could find the causal links that hold in the true causal model, while forgettin about non-existing causal links.
 
 \section{Future Work}
-
 \newpage
-
 \subsubsection*{Acknowledgements}
@@
 
-\subsubsection*{References}
-%doi=false,isbn=false,url=false
 \bibliographystyle{apalike}

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambiando redaccion del abstract para anonimizacion

@@
 \begin{abstract}
-In a recent paper by \cite{2019arXiv190202279G}, Decision Problems under Uncertainty where the uncertain events are causally related where studied.
-For such problems, a criteria for choosing an optimal action was proposed for rational preferences.
-It had been shown by such authors in \cite{gonzalez2018playing} that by implicitly considering a distribution over causal structures and using such criteria within each structre then an on-line decision-making algorithm could be obtained, which achieved a similar performance than the classical Q-learning while also acquiring a Causal Model of the environment.
-Their work has a limitation, since they only considered the case where the decision maker knows the structure of the causal model that controls the environment.
+We consider a Decision Problem under Uncertainty where the uncertain events and the outcomes causally related.
+It is a well known fact that under rational preferences the maximization of expected utility is the only criteria to be used.
+We address the problem where the decision maker knows the variables in his environment but does not know the causal relations between them, which includes the variable to be intervened when making a decision.
+Previous works that have studied decision making in causally-controled environments have left open the case where the causal mechanism which controls the environment is unknown, thus advancing knowledge in this direction.
 In this paper, we propose to extend such limitation by proposing a local structure learning procedure in terms of belief updating, where the beliefs are probabilities over the events of a variable having causal influence over another one.
@@
 
-\section{Conclusions}
+\section{Conclusions and Future Work}
 We have proposed a method for finding a causal structure based on the updating of beliefs by a rational agent about the existence of causal relations between variables.
@@
 Our method allows the use of a previously shown method for decision making based on causal information in the case where the decision maker only knows the variables of the model.
-In a simple case, we tested our method and it was shown experimentally that the proposed method could find the causal links that hold in the true causal model, while forgettin about non-existing causal links.
+In a simple case, we tested our method and it was shown experimentally that the proposed method could find the causal links that hold in the true causal model, while gradually stop believing in the existence of links not present in the true causal model.
 
-\section{Future Work}
+There are still several issues that must be adressed in the near future, such as the case of multi-valued variables or considering a dynamically changing causal model.
+Also, the generale case involving variable discovery must be solved in order to solve the problem of causal decision making when an unknown causal mechanism controls the environment.
+
 \newpage

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambiando intro y related work

@@
 \section{Introduction}
-Decision making under uncertain conditions is a fundamental task for intelligent reasoning and a any day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}).
+Decision making under uncertain conditions is a fundamental task for intelligent reasoning and an any-day task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}).
 It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
-We consider the case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, represented as a Causal Graphical Model which is always unknown by the agent and assumed to remain fixed and invariant to interventions (\cite{woodward2005making}).
-It has been proposed a mechanism for a decision maker to find good actions by holding beliefs about the causal relations in the environment, using them to make an action and then updating those beliefs in terms of what is caused by a chosen action.
-Such mechanism, which is described in \cite{gonzalez2018playing} is limited since it is assumed that the decision maker knows the \textit{structure} of the Causal Graphical Model.
+The case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, has been previously studied in \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{2019arXiv190202279G}.
+In particular, \cite{joyce1999foundations} provided the intuition that a decision maker should consider what is \textit{caused} by his actions, and \cite{pearl2009causality} proposed decision-making by maximizing a function involving the well known \textit{do} operator.
+
+In this work we consider that the causal mechanism controlling the environment is represented as a Causal Graphical Model which is unknown by the agent and assumed to remain fixed and invariant to interventions (\cite{woodward2005making}).
+In recent papers, such as those by \cite{lattimoreNIPS2016}, \cite{sen2017identifying} and \cite{gonzalez2018playing} it is assumed that the decision maker knows either all of, or a certain aspect such as the structure, of the causal model.
 Here we attempt to relax such assumption and allowint the decision maker to know only the variables in the model, with the help of an \textit{ordering} in the variables in such a way that it is know which variables can not cause each other.
@@
 We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}.
-We consider Causal Graphical Models (\cite{koller2009probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution.
+We consider Causal Graphical Models (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution.
 
@@
 \section{Related Work}
-In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered, their assumption of knowing the causal moder is later relaxed by \cite{sen2017identifying} who considers interventions over a partially unknown causal model.
-By following \cite{joyce1999foundations} intuitions, in \cite{gonzalez2018playing} it is proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found.
-Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used locally (in each step) to make a choice.
-In \cite{2019arXiv190202279G}, such authors propose a formal criteria which has the form:
+In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved.
+Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model.
+In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found.
+Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used within each step in otder to make a choice by considering his causal knowledge as it if were true, as proposed by \cite{joyce1999foundations}.
+Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model.
+Later, \cite{2019arXiv190202279G}, such authors formalize that procedure by stating a formal criteria which has the form:
 \[  a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g), \]
@@
 we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment.
-We note that the criteria proposed by \cite{2019arXiv190202279G} makes use of an implicit probability distribution over causal structures, since the decision maker holds beliefs about the causal nature of his environment and within each realization of those beliefs the maximization for the optimal action is performed, while Pearl's function assumes a single causal structure.
+Pearl's function assumes a single causal structure.
+as opposed by \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures.
 
-Using a version of the criteria described above, in \cite{gonzalez2018playing} a decision-making procedure is proposed for the case when the agent knows the \textit{structure} of the Causal Model which controls his environment.
 It remains an open question to tackle the problem of a fully unknown structure.
 It is known that from observational data alone, a Causal Graphical Model is identifiable up to Markov equivalence without any the assumption of a specific functional model with given error distributions (\cite{hauser2012characterization}, \cite{hauser2012two}, \cite{peters2011identifiability}).
-We pretend to extend here the work of \cite{gonzalez2018playing} to the case of a rational decision maker who knows the variables that control his environment and when he has extra knowledge of variables that can not be cause of each other.
+We pretend to relax the mentioned limitations to the case of a rational decision maker who knows the variables that control his environment and he has extra knowledge of variables that, in principle can not be cause of each other.
 

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambios en la seccion de rationality

@@
 It is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).
+
 The case where the uncertainty component in the decision problem is controlled by a \textit{causal mechanism}, has been previously studied in \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{2019arXiv190202279G}.
@@
 It has also been argued that humans tend to ignore probabilistic information and assign more importance to causal information (\cite{tversky1980causal}, \cite{pearl2009causality}).
-Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal}, \cite{wellen2012learning}, \cite{danks2014unifying}).
-Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations.
+Human beings focus on \textit{local} aspects while learning causal relationswhich are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}).
+Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning.
+Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.
 
-\section{Rationality, Decision Making and Expected Utility}
+\section{Rationality and Decision Making under}
 In this work we begin with a Decision Problem under Uncertainty where an agent, who has rational preferences, faces the problem of choosing one among many available actions with uncertain consequences.
 From the work of \cite{von1944theory} and \cite{savage1954the}, we know that the only criteria for making choices is the maximization of expected utility with respect to the agent's preferences and a probability distribution which represents either the probabilities of the uncertain events, or the agents own beliefs about such uncertainty.
-In \cite{gonzalez2018playing} and \cite{2019arXiv190202279G} the case where the uncertainty of the environment is controlled by a causal mechanism is considered.
-Since the (unknown) causal mechanism is itself a stochastic one, and the agent's preferences are rational, Savage's theorem state how an action must be chosen: maximizing the expected utility.
-If the decision maker does not know the causal mechanism, this is equivalent to not knowing the probabilities of the environment, so a subjective distribution must be used.
+
+Since \cite{joyce1999foundations} it has been argued that a decision maker should consider all of his beliefs about what will his decisions cause, while \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as
+\[ U(x)=\sum_y P(y | do(x)) u(y), \]
+we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment as opposed to \cite{gonzalez2018playing} and \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures, while Pearl's function assumes a single causal structure.
 
 \section{Related Work}
+
 In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved.
@@
 The function above described is indeed a counterfactual distribution, since it allows to compare the probability of causing a desired actions by alternative actions that were not chosen.
-
-While \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as 
-\[ U(x)=\sum_y P(y | do(x)) u(y), \]
-we notice that this function does not consider the agent's beliefs, or current knowledge, about the causal structure that holds in his environment.
-Pearl's function assumes a single causal structure.
-as opposed by \cite{2019arXiv190202279G} who make use of an implicit probability distribution over causal structures.
 

### 2019-03-03 UAI_2019/UAI_Draft.tex
cambios en related work

@@
 \section{Related Work}
-
 In \cite{lattimoreNIPS2016}, the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by considering causal information a \textit{faster} optimal-action learning can be achieved.
 Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model.
+
 In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused} good actions can be found.
@@
 Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model.
+
 Later, \cite{2019arXiv190202279G}, such authors formalize that procedure by stating a formal criteria which has the form:

### 2019-03-05 UAI_2019/Draft_v2.tex
version que incluye los 2 papers

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[letterpaper]{article}
\usepackage{uai2019}
\usepackage[margin=1in]{geometry}
\usepackage{breakcites}
\usepackage{graphicx}
\usepackage{subfigure}
% Set the typeface to Times Roman
\usepackage{times}

\title{Causal Structure Learning for Decision Making under Uncertainty}

\author{Anonymous authors.} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI  reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an
%alternate address.}} \\
%Computer Science Dept. \\
%Cranberry University\\
%Pittsburgh, PA 15213 \\
%\And
%{\bf Coauthor}  \\
%Affiliation          \\
%Address \\
%\And
%{\bf Coauthor}   \\
%Affiliation \\
%Address    \\
%(if needed)\\
%}

\begin{document}

\maketitle

\begin{abstract}
We define a Causal Decision Problem as a Decision Problem Under Uncertainty where the available actions and the set of outcomes are related stochastically through a Causal Graphical Model $\mathcal{G}$, which is assumed to be unknown by the decision maker. Based on the Expected Utility principle we propose a decision-making criteria which allows a distribution over causal structures which encode what the decision maker believes. Previous works that have studied decision making in causal environments have studied cases where the causal model is known, or parts of it such as the graphical structure. We propose to extend such works by considering the case where the decision maker knows the variables in his environment but not the causal relations between them. We then propose a local structure-learning procedure in terms of belief updating, where such beliefs represent whether the agent believes there is a causal relation between pairs of variables. We show experimentally that our method is capable of finding the causal links that hold in a true causal model after a series of observations are drawn from the true causal model. Our paper contributes to the design of learning algorithm in unknown causal environments in both a theoretical and experimental way.
\end{abstract}

\section{Introduction}
Decision making under uncertain conditions is a fundamental task for intelligent reasoning and a common task for human beings and organizations (\cite{savage1954the}, \cite{danks2014unifying}, \cite{lake2017building}). Many real-world applications rely on decisions made by an autonomous agent, such as self-driving cars. Current decision making methods rely on associative methods, which find only statistical patterns in data. On the contrary, causal knowledge allows both for planning and counterfactual reasoning as well as interpretability and explainability \cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{woodward2005making}, \cite{pearl2018why}. We propose in this work a criteria which must be fulfilled while making decisions under uncertain conditions which are causal-controlled.

In Decision Theory it is a well known fact that if an agent has \textit{rational preferences} then he must choose according to the Maximum Expected Utility (MEU) criteria, in which the agent chooses an action by considering both his preferences, which are represented by an utility function, and the probability of occurrence of outcomes: with respect to this point, the agent either knows the probabilities in the environment, or, if he does not, by considering a \textit{subjective} probability measure (\cite{von1944theory}, \cite{savage1954the}, \cite{bernardo2000bayesian}, \cite{gilboa2009decision}).

Our criteria acknowledges that causal-decision making is a particular case of uncertain decision making, so the MEU criteria must be satisfied, but we will make explicit the causal component in it in terms of Pearl's $do$ operator. The case when the uncertainty component in a decision problem is controlled by a \textit{causal mechanism} has been previously studied, for example in: \cite{joyce1999foundations}, \cite{pearl2009causality}, \cite{bottou2013counterfactual}, \cite{ortega2014generalized}, \cite{bareinboim2015bandits}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}, \cite{acharya2018learning},. In particular, \cite{joyce1999foundations} provided the intuition that a decision maker should consider what is \textit{caused} by his actions, and \cite{pearl2009causality} proposed decision-making by maximizing a function involving the well known \textit{do} operator.

In this work we consider that the causal mechanism which controls the environment is represented as a Causal Graphical Model $\mathcal{G}$ and that it is unknown by the agent. $\mathcal{G}$ is also assumed to remain fixed and invariant to interventions (\cite{woodward2005making}). In recent papers, such as those by \cite{lattimoreNIPS2016}, \cite{sen2017identifying} and \cite{gonzalez2018playing} it is assumed that the decision maker knows either all of, or a certain aspect such as the structure, of the causal model.

In this paper we propose an optimality criteria for causal decision making based on the Maximization of Expected Utility principle, in Joyce's intuitions and in Pearl's $do$ operator. Our criteria allows the decision maker to have in mind a distribution over causal structures and then maximize as previously proposed in \cite{pearl2009causality} \textit{within} each structure. We also attempt to relax the assumption of the agent knowing the causal model by allowing him to only know the variables in the model, with the help of an \textit{ordering} in the variables in such a way that it is know which variables can not cause each other and we attempt to learn a causal structure by \textit{local} belief updating.

We call our method \textit{local}\footnote{Inspired by the use of the term in \cite{wellen2012learning}} since we use the uncertainty (from the agent's point of view) about the existence of a causal relation between pairs of variables, which are local components of the structure of a graph. Such uncertainty is to be updated in terms of what is \textit{observed} from realizations of the true causal mechanism that controls the environment. Since observations alone do not suffice to uniquely determine a DAG, we make use of a partial ordering between the variables in the model that, at least in principle, can not be a cause of each other. This ordering must be an input product from expert knowledge.

The main contributions of this paper are a decision-making criteria for causal decision making in terms of an agent's beliefs, and a method for discovering a causal structure by considering the beliefs of the decision maker in terms of the possible existence of causal relations between variables.

\section{On Causality and Learning}
Several definitions of Causality have been proposed by philosophers and scientists, such as those described in \cite{holland1986statistics}. We use here the formal definition of Causality given by \cite{spirtes2000causation}, which belongs to the family of \textit{manipulationist} theories as described by \cite{woodward2005making}. We consider Causal Graphical Models (\cite{koller2009probabilistic}, \cite{sucar2015probabilistic}) as representations of the causal relations that hold in an environment and the resulting interventional distribution since graphs are plausible representations of cognitive processes in human beings; in particular, causal reasoning (\cite{danks2014unifying}).

In fact, it has been shown that humans acquire and use causal information for decision-making and that humans even consider their actions as interventions in the world (\cite{hagmayer2009decision}, \cite{wellen2012learning} \cite{hagmayer2013repeated}). It has also been argued that humans tend to ignore probabilistic information and assign more importance to causal information (\cite{tversky1980causal}, \cite{pearl2009causality}). Human beings focus on \textit{local} aspects while learning causal relations which are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning. Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.

\section{Related Work}
We begin with a Decision Problem under Uncertainty in which an agent, or decision maker, who has rational preferences, faces the problem of choosing one among many available actions with uncertain consequences. From the work of \cite{von1944theory} and \cite{savage1954the}, we know that the only criteria for making choices is the maximization of expected utility with respect to the agent's preferences and a probability distribution which represents either the probabilities of the uncertain events, or the agents own beliefs about such uncertainty. In this work we pretend to take a \textit{normative} view for a rational decision maker who faces an uncertain environment which is controlled by some unknown causal mechanism.

Since \cite{joyce1999foundations} it has been argued that a decision maker should consider all of his beliefs about what will his decisions cause, while \cite{pearl2009causality} proposes that a rational agent in the presence of a causal environment should optimize the function $U(x)$ defined as: 
\begin{equation}{\label{pearl}} 
U(x)=\sum_y P(y | do(x)) u(y).
\end{equation}

Notice that this function does not consider the agent's beliefs about the causal structure that holds in his environment. We will later define formally a Causal Decision Problem and provide a solution that explicitly includes the decision maker belief's about the existence of causal relations in the environment.

On the problem of discovering causal structure it was conjectured in \cite{eberhardt2005number} \cite{eberhardt2008almost} and late proven by \cite{hauser2012two} that the worst case number of interventions required for full causal identifiability is a function of the maximal clique in the true DAG, but in order to apply such result an agent must be able to perform such interventions, while the method we are proposing here attempts to rely only in observations with the help of external expert knowledge.

In \cite{ortega2014generalized} propose that Thompson Sampling can be adapted to the problem of causal induction; this is, the problem of discovering a causal structure. \cite{bareinboim2015bandits} apply the setting of Bandit problems to decision making, while \cite{lattimoreNIPS2016}, who also use bandits, consider the problem of finding a best intervention over a known causal model for a fixed number of learning trials is considered and it was found that by them that by considering causal information a \textit{faster} optimal-action learning can be achieved. Their assumption of knowing the causal model was relaxed by \cite{sen2017identifying} who considers a partially known causal model and allows interventions over a the unknown part of the model.

In \cite{gonzalez2018playing} it was proposed that a rational decision maker can find an optimal action by holding \textit{beliefs} about the existence and strength of causal relations and by using those beliefs \textit{as if} they were the true relations in order to make a choice and then updating those beliefs after observing what the chosen action \textit{caused}. Such procedure implicitly considers a distribution over Causal Graphical Models, where each structure is used within each step in otder to make a choice by considering his causal knowledge as it if were true, as proposed by \cite{joyce1999foundations}. Their work assumes that the decision maker knows the structure, but not the associated parameters, of the model. Their work was able to find good actions in a way comparable to that of the classic Q-Learning algorithm for Reinforcement Learning (\cite{watkins1992q}).

It remains an open question to tackle the problem of a fully unknown structure. It is known that from observational data alone, a Causal Graphical Model is identifiable up to Markov equivalence without any the assumption of a specific functional model with given error distributions (\cite{hauser2012characterization}, \cite{hauser2012two}, \cite{peters2011identifiability}). We pretend to relax the mentioned limitations to the case of a rational decision maker who knows the variables that control his environment and he has extra knowledge of variables that, in principle can not be cause of each other. We pretend to find a causal structure after a series of observations.

Once such a structure is found, then the procedure proposed in \cite{gonzalez2018playing} could be applied in order to learn optimal actions.

\section{Causal Decision Problems and Optimal Actions}{\label{CDPU}}
We define a Causal Decision Problem under Uncertainty (CDPU) as a tuple $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \mathcal{G}, \succeq)$ where  $(\mathcal{A}, \mathcal{E}, \mathcal{C}, \succeq)$ is a classical Decision Problem under Uncertainty (\cite{bernardo2000bayesian}) and $\mathcal{G}$ is a Causal Graphical Model (\cite{sucar2015probabilistic}) such that the set of available actions $\mathcal{A}$ and the set of outcomes $\mathcal{C}$ are related through the variables of the Causal Model $\mathcal{G}$; i.e., the events in the family $\mathcal{E}$ correspond to variables in $\mathcal{G}$. It is assumed that the agent does not know the Causal Model $\mathcal{G}$, which is equivalent to not knowing the probabilities of the events $E \in \mathcal{E}$. The model $\mathcal{G}$ is also assumed to remain fixed and to be invariant under interventions \cite{woodward2005making} and to satisfy the conditions expressed in \cite{spirtes2000causation}. The variable in $\mathcal{G}$ which encodes the consequence of the action taken by the agent will be referred as \textit{target variable} since it is the variable where the agent wishes to obtain a desired result.

In this way, in a CDPU we have a rational agent who chooses an action $a$ among the many available in $\mathcal{A}$ and then this action will produce some random effect in the environment which will \textit{cause} a certain consequence, or outcome $c \in \mathcal{C}$. The causal relation between $a \in \mathcal{A}$ and $c \in \mathcal{C}$ is encoded in $\mathcal{G}$.

Since rational preferences are assumed, the agent must choose as if maximizing his expected utility in terms of his current knowledge, or \textit{beliefs}, which is expressed as a probability distribution. Using the awareness of the agent about a causal mechanism governing the environment, intuition encourages to use causal relations to cause some desirable action, as expressed by \cite{joyce1999foundations} and by what Pearl calls \textit{commonsensical decision making}.

Consider a Causal Decision Problem as stated above where the target variable, call it $Y$, takes its values in the set $\{0,1\}$ and without loss of generality assume that $Y=1$ is the desired output for the agent; then, according to the maximum expected utility principle, as well with Joyce intuition (\cite{joyce1999foundations}), he must choose the action $a^\ast \in \mathcal{A}$ such that

\[ P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}, \]

where $P$ is a probability distribution which encodes the agent knowledge, or beliefs, about the causal relations that hold in his evironment. This is equivalent to finding $a^\ast$  as
\begin{equation}{\label{principle}}
 a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}\sum_g P(Y=1 | do(a), g)P(g).
\end{equation}

Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent. If an action $a_0 \in \mathcal{A}$ yielded a higher expected utility, that would imply that it has higher probability of \textit{causing} the same desired action, but this is not possible because of how $a^\ast$ is obtained.

The procedure for selecting $a^\ast$ connects both the utility for the decision maker (because of the most desired outcome) and his beliefs (the measure $P$) about \textit{causing} a desired result. Since the causal mechanism and the uncertainties in the environment are connected, $a^\ast$ to maximizes expected utility. This procedure allows to choose a \textit{better action} in terms of what is currently known about causal relations. In fact, the distribution in Equation \ref{principle} includes an implicit probability distribution over causal structures as opposed to Equation \ref{pearl} which assumes a single causal structure rather than having a distribution over structures.

Since $a^\ast$ was found by asking the probability of \textit{causing} the most desired action for the agent, it is also a counter-factual distribution since we know that the probability of observing the most desired outcome is lower for any other action $a \in \mathcal{A}$ and the probability of not choosing $a$ can also be easily obtained.

\section{Structure Learning: a local approach}
\subsection{Setting and assumptions}
In the context of the CDPU's defined in Section \ref{CDPU} we consider the problem of a decision maker who faces a causal-controlled environment and does not know the causal relations between variables. It is known that learning and using a causal model when the underlying model is completely unknown is a difficult task (\cite{lattimoreNIPS2016}). We will consider a series of assumptions in order to simplify the problem. We consider:
\begin{enumerate}
\item The decision maker knows the variables in the model, let $V=\{ X_1,...,X_n \}$ be the set of variables.
\item The decision maker knows which variable can he intervene and this variable does not depend on others.
\item The decision maker knows which variable he wants to affect with his interventions. 
\item The decision maker can fully observe realizations of the variables from the true causal model $\mathcal{G}$, but such realizations are independent.
\item Variables are binary-valued $\{0,1 \}$ where the value $1$ interpreted as \textit{acting} on them or not, or as being in either in \textit{on} or \textit{off} states.
\end{enumerate}
The last assumption appears to be a very strong one, but the multi-valued case can be dealt with by replacing the variable with an artificial one which is binary-valued, where one of the states correspond to the values of the original variable, and the other to doing nothing, as in the binary case. As for the meaning of the states, consider for example giving a patient a drug or not.

\subsection{The form of uncertainty}
Our approach begins by noticing that if the decision maker does not know the relations between the variables, then, as shown in Figure \ref{form_ofl} his uncertainty about the causal structure has the following \textit{form}: Does $A$ cause $B$? For any pair of variables $A,B$ in $V$ we translate this statement into:
\begin{equation}
p_{AB}=P(A \textrm{ cause } B).
\end{equation}
Now, the idea is to update $p_{AB}$ according to the manipulationist idea stated in \cite{woodward2005making} where a manipulation in $A$ should translate in a change of value in $B$. Since we are assuming that the variables are binary and interpreted as acting on them or not, or being on an \textit{on} or \textit{off} states, (consider for example giving a certain treatment or not), the agent will increase his probability of $A$ causing $B$ if the pair $(A=1,B=1)$ is observed. It is known from \cite{holland1985statistics} that manipulation is fundamental to the notion of Causality, and although we notice that our decision maker is merely \textit{observing}, we are interpreting the binary-nature of our variables as observing manipulations over them in the form of \textit{acting} or them or not.

Since the probability $p_{AB}$ is considered for any pair $(A,B)$ we must be careful in order not to update also $p_{BA}$ since from the definition of Causality in \cite{spirtes2000causation} we have that if $A$ causes $B$, then $B$ can not cause $A$. Here we use the partial ordering previously mentioned: the decision maker has a partial ordering $\succeq$ of the variables which corresponds to variables that can not cause each other. This extra assumption, although a strong one, can be thought of expert knowledge in the domain of the decision problem.

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/uncertainties.png}}
\caption{The probabilistic form of the uncertainties from the agent's point of view.}
\label{form_ofl}
\end{center}
\vskip -0.2in
\end{figure}

\subsubsection{Uncertainty Updating}
Given that the decision maker can observe a full realization of $X_t=(X_1=x_1,...,X_n=x_n)$ of the variables from the true causal model, where $x_i \in \{0,1 \}$, then for any pair $X_i,X_j$ with $X_j \succeq X_i$, if $(X_i=1, X_j=1)$ is observed on stage $t$, then the belief $p_{X_i X_j}$ is increased by de decision maker to:
\begin{equation}
 p_{X_i X_j}^{(t+1)} = p_{X_i X_j}^{(t)}+\alpha (1-p_{X_i X_j}^{(t)}) \textrm{ for } \alpha \in (0,1). 
 \end{equation}
On the other hand, if $(X_i=1, X_j=0)$ is observed on stage $t$, then we decrease $p_{X_i X_j}$ to:
\begin{equation}
p_{X_i X_j}^{(t+1)} =   \alpha p_{X_i X_j}^{(t)}  \textrm{ for } \alpha \in (0,1).  \end{equation}
Once the values  $p_{X_i X_j}$ are updated for every $i,j$ such that $X_j \succeq X_i$, the procedure is once again started with a new observation $X_{t+1}$.

\subsubsection{Obtaining a structure and decision making}
After a series of rounds of observation and belief updating it is expected that the information provided by the environment, together with the partial ordering, will allow the decision maker to update his beliefs about the existence of a causal relation between variables in such a way that the stronger (bigger than some threshold) values are considered as a relation existing between a pair of variables. Once a causal structure $\bar{\mathcal{G}}$ is obtained, the decision maker this structure \textit{as if} it were the true causal model in order to choose an action by using our proposed criteria and then updating beliefs by observe the causal consequence of the action; this procedure ends with a vector parameters $\alpha$ over the Conditional Probability Tables of the Causal Model which has structure $\bar{\mathcal{G}}$, and such parameters can be used in order to start again the structure-learning procedure, since a low-value parameter points to a possible non-existence of a causal link in $\bar{\mathcal{G}}$ when the outcome of the action is considered.

\section{Experiments and Results}
In order to test our method, we use the simple example in \cite{gonzalez2018playing}, in which a sick patient arrives to a hospital and a medic who does not know his disease must give him some treatment, which can have or not a reaction, which can kill him. The Causal Graphical Model that controls what happens to the patient is shown in Figure \ref{causal_model}.

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/Propuesta/causal_graph.png}}
\caption{The true Causal Graphical Model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}

We test our method by starting with a random initialization of the probabilities of a variable causing another and updating as mentioned. As an external input we included a list of pairs of variables in which the first element can not cause the second one. Using this list, a second list of \textit{possible causes} was obtained and to the elements of this list random probabilities were assigned. A value of $\alpha=0.8$ was used, and $\alpha$ was decreased by a factor of $1/k$ where $k$ was the current round. After 10 rounds of observing realizations from the true model we see in Figure \ref{10_rounds} that the beliefs about causal relations that do hold in Figure \ref{causal_model} tend to $1$ while non-existing links go down to $0$. In Figure \ref{25_rounds} the same behavior is observed.

We repeated the experiments of $10$ and $25$ rounds $100$ times

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_10_5marzo.png}}
\caption{Beliefs about the existence of causal links between variables in 10 rounds.}
\label{10_rounds}
\end{center}
\vskip -0.2in
\end{figure}

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_25_5marzo.png}}
\caption{Beliefs about the existence of causal links between variables in 25 rounds.}
\label{25_rounds}
\end{center}
\vskip -0.2in
\end{figure}

\section{Conclusions and Future Work}
We have proposed an optimality criteria of decision making under causal-controlled uncertainty which makes use of a decision maker belief's about the existence of causal relations in his environment. The proposed method makes use of an implicit distribution over causal structure and applies a simpler, but intuitive, criteria within each structure. We also proposed a method for finding a causal structure based in beliefs about the existence of causal relations between variables and the updating of those beliefs in terms of what was observed from the true model. The need of a causal ordering is required since it is known that based on mere observations a Directed Graph can not be obtained.

Our method allows the use of a previously shown method for decision making based on causal information in the case where the decision maker only knows the variables of the model. In a simple case, we tested our method and it was shown experimentally that the proposed method could find the causal links that hold in the true causal model, while gradually stop believing in the existence of links not present in the true causal model.

There are still several issues that must be adressed in the near future, such as the case of multi-valued variables or considering a dynamically changing causal model. Also, the generale case involving variable discovery must be solved in order to solve the problem of causal decision making when an unknown causal mechanism controls the environment.

\subsubsection*{Acknowledgements}
The authors would like to thank \texttt{<Anonymous>} for his valuable comments on certain important issues.

\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}

\end{document}

### 2019-03-06 UAI_2019/Draft_v2.tex
ultima version

@@
 
-Since the action chosen is the action with the highest probability of producing the most desired action, then it is the action that maximizes the expected utility for the agent.
-If an action $a_0 \in \mathcal{A}$ yielded a higher expected utility, that would imply that it has higher probability of \textit{causing} the same desired action, but this is not possible because of how $a^\ast$ is obtained.
+Notice that, by construction, $a^\ast$ is the action with the highest probability of producing the most desired action; therefore, we argue that $a^\ast$ it is in fact the action that maximizes the expected utility for the agent.
+In order to show this, let $a \in \mathcal{A}$ an action that has higher expected utility; this is, $a$ produces the outcome $Y=1$ with a higher probability than $a^\ast$, which is a contradiction considering how $a^\ast$ is obtained.
 
@@
 Given that the decision maker can observe a full realization of $X_t=(X_1=x_1,...,X_n=x_n)$ of the variables from the true causal model, where $x_i \in \{0,1 \}$, then for any pair $X_i,X_j$ with $X_j \succeq X_i$, if $(X_i=1, X_j=1)$ is observed on stage $t$, then the belief $p_{X_i X_j}$ is increased by de decision maker to:
-\begin{equation}
+\begin{equation}{\label{increase}}
  p_{X_i X_j}^{(t+1)} = p_{X_i X_j}^{(t)}+\alpha (1-p_{X_i X_j}^{(t)}) \textrm{ for } \alpha \in (0,1). 
@@
 
+\subsubsection{Smoothing}
+Notice that the method proposed in the previous section allows for big jumps to happen when current beliefs are small, since the difference between $1$ and the current quantification of uncertainty $p$ can ve big if $p$ is small, so we slightly modify Equation \ref{increase} to:
+\begin{equation}
+ p_{X_i X_j}^{(t+1)} = p_{X_i X_j}^{(t)}+\alpha (1-p_{X_i X_j}^{(t)}) \textrm{ for } \alpha \in (0,1),
+\end{equation}
+When $p_{X_i X_j}^{(t)} \geq 1/2$ and,
+\begin{equation}
+ p_{X_i X_j}^{(t+1)} = p_{X_i X_j}^{(t)}+\alpha (\frac{1}{2}-p_{X_i X_j}^{(t)}) \textrm{ for } \alpha \in (0,1),
+\end{equation}
+when $p_{X_i X_j}^{(t)} \leq 1/2$. 
 \subsubsection{Obtaining a structure and decision making}
@@
 \section{Experiments and Results}
+\subsection{Sick Patient Problem}
 In order to test our method, we use the simple example in \cite{gonzalez2018playing}, in which a sick patient arrives to a hospital and a medic who does not know his disease must give him some treatment, which can have or not a reaction, which can kill him.
@@
 A value of $\alpha=0.8$ was used, and $\alpha$ was decreased by a factor of $1/k$ where $k$ was the current round.
+
+\subsection{One-trial experiments}
+In order to show how the beliefs are being updated inside the agent's head we show here the results of one-trial realizations; this is, the process of starting with random initial beliefs and update for $k$ rounds of observing, with $k \in \{ 10,25,50 \}$.
+
 After 10 rounds of observing realizations from the true model we see in Figure \ref{10_rounds} that the beliefs about causal relations that do hold in Figure \ref{causal_model} tend to $1$ while non-existing links go down to $0$.
+
 In Figure \ref{25_rounds} the same behavior is observed.
 
-We repeated the experiments of $10$ and $25$ rounds $100$ times
+In Figure \ref{50_rounds} we observe a similar behavior
+
+\subparagraph{Averages of multiple trials}
+Both of the previous experiments were a single realization of our method; this is, a single structure would be obtained.
+In order to explore our results in a more general way, we repeated the experiments of $10$ and $25$ and $50$ rounds $500$ times and we observe the averages in the probabilities (beliefs).
+These results are shown in Figure \ref{500_10_rounds},, Figure \ref{500_25_rounds} and Figure \ref{500_50_rounds}.
 
@@
 \begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_10_5marzo.png}}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_10.png}}
 \caption{Beliefs about the existence of causal links between variables in 10 rounds.}
@@
 \begin{center}
-\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_25_5marzo.png}}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_25.png}}
 \caption{Beliefs about the existence of causal links between variables in 25 rounds.}
@@
 
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_50.png}}
+\caption{Beliefs about the existence of causal links between variables in 50 rounds.}
+\label{50_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_500_10.png}}
+\caption{Repeating 500 times the 10 round learning.}
+\label{500_10_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_500_25.png}}
+\caption{Repeating 500 times the 25 round learning.}
+\label{500_25_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_500_50.png}}
+\caption{Repeating 500 times the 50 round learning.}
+\label{500_50_rounds}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+\subsection{Alarm Problem}
+We tested our method in the well known Alarm problem described in \cite{barber2012bayesian}, where al Alarm can be triggered by a Burglar or by an Earthquake, and we introduced the situation in which the Burglary can be caused by the occurrence of a Recession.
+The causal model for this example is shown in Figure \ref{alarm_problem}.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/alarm}}
+\caption{Causal model for the Alarm Problem.}
+\label{alarm_problem}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
+In this example, the only elements in the ordering were the pairs $(Alarm, Burglary)$ and $(Alarm,Earthquake)$, so there were $10$ possible causes that the method had to consider.
+We show only the $500$ repetitions of $25$ rounds in Figure \ref{500_25_alarm_problem} where we observe that the method correctly discarded 5 non-existing causal links and correctly captured 4, and we had only 1 false-positive which corresponds to the link Recession causing Alarm, which is not an edge present in the graph, but since Causality is a transitive relation, it is in fact a causal relation.
+
+\begin{figure}[ht]
+\vskip 0.2in
+\begin{center}
+\centerline{\includegraphics[width=\columnwidth]{/Users/MauricioGS1/INAOE/UAI_2019/Code/results_500_50_alarm.png}}
+\caption{500 repetitions of 25 learning rounds.}
+\label{500_25_alarm_problem}
+\end{center}
+\vskip -0.2in
+\end{figure}
+
 \section{Conclusions and Future Work}
-We have proposed an optimality criteria of decision making under causal-controlled uncertainty which makes use of a decision maker belief's about the existence of causal relations in his environment.
+We have proposed an optimality criteria for decision making under a causal-controlled uncertain component which makes use of a decision maker belief's about the existence of causal relations in his environment.
 The proposed method makes use of an implicit distribution over causal structure and applies a simpler, but intuitive, criteria within each structure.
-We also proposed a method for finding a causal structure based in beliefs about the existence of causal relations between variables and the updating of those beliefs in terms of what was observed from the true model.
-The need of a causal ordering is required since it is known that based on mere observations a Directed Graph can not be obtained.
+Such decision criteria, when implemented, leads to a decision-making algorithm in which a decision maker uses beliefs \textit{as if} they were the true causal model and then updates after observing an outcome.
+It has been experimentally shown that such algorithm has a similar performance to the classic Q-learning.
+
+We have also proposed a method for obtaining a causal structure based in a decision maker beliefs about the existence of causal relations between pairs variables and then updating of those beliefs in terms of what was observed from the true model.
+The need of a causal ordering is required since it is known that based on mere observations a Directed Graph can not be obtained as previously argued.
 
@@
 There are still several issues that must be adressed in the near future, such as the case of multi-valued variables or considering a dynamically changing causal model.
-Also, the generale case involving variable discovery must be solved in order to solve the problem of causal decision making when an unknown causal mechanism controls the environment.
+Also, the general case involving variable discovery must be solved in order to solve the problem of causal decision making when an unknown causal mechanism controls the environment.
 

### 2019-05-17 NeurIPS_2019/Draft.tex
casi final

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
 \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
 %    \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
     %\usepackage[final]{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts,amssymb,amsthm,amsmath,latexsym}     % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\theoremstyle{plain}
\newtheorem{teo}{Theorem}
\newtheorem{prop}[teo]{Proposition}
\newtheorem{defi}[teo]{Definition}
\newtheorem{obs}[teo]{Observation}
\newtheorem{lem}[teo]{Lemma}
\newtheorem{cor}[teo]{Corolary}

\title{von Neumann-Morgenstern and Savage Theorems for Causal Decision Making}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{
  Mauricio Gonzalez-Soto \\
  Coordinación de Ciencias Computacionales\\
  Instituto Nacional de Astrofisica Optica y Electronica (INAOE)\\
  Mexico \\
  \texttt{mauricio@inaoep.mx}
   \And
   Luis E. Sucar\\
   Coordinación de Ciencias Computacionales \\
   Instituto Nacional de Astrofisica Optica y Electronica (INAOE) \\
   Mexico\\
   \texttt{esucar@inaoep.mx}
   \And
  Hugo J. Escalante Balderas \\
  Coordinación de Ciencias Computacionales \\
  Instituto Nacional de Astrofisica Optica y Electronica (INAOE) \\ \\
   \texttt{hugojair@inaoep.mx}
}

\begin{document}

\maketitle

\begin{abstract}
Decision making under uncertain conditions has been well studied when uncertainty can only be considered at the associative level of information. The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rationally making choices using associative information. We provide here a previous result from Pearl and show that it can be considered as a causal version of the von Neumann-Morgenstern Theorem; furthermore, we consider the case when the true causal mechanism that controlls the environment is unknown to the decision maker and propose a causal version of the Savage Theorem. As applications, we argue how previous optimal action learning methods for causal environments fit within the Causal Savage Theorem we present thus showing the utility of our result in the justification and design of learning algorithms; furthermore, we define a Causal Nash Equilibria for a strategic game in a causal environment in terms of the Causal Savage Theorem.
\end{abstract}
\section{Introduction}
Causal reasoning is a constant element in our lives as it is human nature to constantly ask \textit{why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds (\cite{waldmann2013causal}, \cite{danks2014unifying}). It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world (\cite{friston2010free}, \cite{clark2015surfing}). Acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices (\cite{tversky1980causal}, \cite{Garcia-Retamero2006}, \cite{hagmayer2008causal}, \cite{hagmayer2009decision}, \cite{hagmayer2013repeated}, \cite{hagmayer2017causality}).

An important aspect of acting in the world is being able to make decisions under uncertain conditions (\cite{danks2014unifying}, \cite{lake2017building}). \cite{von1944theory} gave an answer for how to make choices if \textit{rational} preferences are assumed and the decision maker knows the stochastic relation between actions and outcomes: maximize expected utility. If no such relation is known, then \cite{savage1954the} showed that a rational decision maker must choose \textit{as if} she is maximizing expected utility using a \textit{subjective} probability distribution. Such theorems provide formal criteria for decision making if rationality is assumed. This criterion is the basis for many of the techniques used in Artificial Intelligence; for example, Reinforcement Learning algorithms learn \textit{optimal policies} in such a way that any action prescribed by the optimal policy achieves the maximum expected utility (\cite{sutton1998reinforcement}, \cite{Puterman:1994:MDP:528623}). Algorithms which rely on the von Neumann-Morgenstern or Savage Theorems are based on associative relations which are expressed using correlations or probability distributions. It is a natural question how to formalize rational decision making when causal information is present. Such question has been previously considered by \cite{nozick1969newcomb}, \cite{lewis1981causal}, \cite{joyce1999foundations} without an explicit optimality criterion for decision making and by \cite{pearl2009causality} who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known.

In this work we extend such criterion to the case where a decision maker does not know the causal mechanism so she holds \textit{beliefs} about possible causal model and uses such beliefs \textit{as if} such beliefs were true, as prescribed by \cite{joyce1999foundations} in order to attempt to make a good choice given her beliefs. We provide an explicit way of making a good choice given what the decision maker believes about the causal structure of her environment, thus providing a causal version of Savage's Theorem. We take a normative point of view to study how causal relations should be used by a \textit{rational} agent when making decisions with uncertain consequences.

Our goal is to establish the foundations for decision making algorithms which rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found in \cite{bareinboim2015bandits}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing}.

\section{The Classical Decision Making Theorems}
A Decision problem under uncertainty is a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

The most well-known theories for Decision Making are those from \cite{von1944theory} and \cite{savage1954the}. In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences, which is also known as decision under risk (\cite{binmore2008rational}, \cite{peterson2017introduction}), and in that case the theory guarantees that the decision maker behaves \textit{as if} she maximizes the expected value of a utility function.

If the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if she has in mind a \textit{subjective} probability distribution and a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and the utility function. It is important to say that other Decision Making theories exist, such as Prospect Theory (\cite{kahneman1979prospect}), Case-Based Decision Theory (\cite{gilboa1995case}), among others that are out of the scope of this work. We will later talk about another theory: Causal Decision Making (\cite{joyce1999foundations}). For further details on classical decision making, see \cite{bernardo2000bayesian}, \cite{gilboa2009decision}, \cite{wakker2010prospect}.

\subsection{von Neumann-Morgenstern}
\cite{von1944theory} had the objective of justifying strategies in which players in a game maximized expected utility. This theorem considers a scenario of \textit{decision under risk} and rational preferences; this is, choosing between uncertain outcomes with known probabilities. Formally, we consider a set $X$ of alternatives. Let $L$ be the set of lotteries (with finite support) over $X$. The actual object of choice are the elements $l \in L$, which are known to the decision maker; we represent the decision maker's preferences by a preference relation $\succeq \subseteq L \times L$ which satisfies being complete, transitive, continuous and a technical condition called \textit{independence}. This family of conditions is called \textit{von Neumann-Morgenstern rationality axioms}. \cite{gilboa2009decision}.
\begin{teo}{\label{vNM}}
A preference relation $\succeq \subseteq L \times L$ where $L$ is a set of lotteries with finite support over a set $X$ satisfies the von Neumann-Morgenstern rationality axioms if and only if there exists a function $u: X \to \mathbb{R}$ such that for every $P, Q \in L$ we have that
\begin{equation}
P \succeq Q \textrm{ if and only if } \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x). 
\end{equation}
\end{teo}
The theorem states that if a rational decision maker knows the probabilities of obtaining a certain outcome, then she must choose \textit{as if} maximizing the expected value of some function $u$ whose existence is guaranteed by Theorem \ref{vNM}. See \cite{gilboa2009decision} for details on the proof.
\subsection{Savage's Theorem}
If a rational decision maker does not know the probabilities of obtaining certain outcomes and does not have a precise quantification of her preferences (utility function), then it is \cite{savage1954the} who gives a formal choosing criterion. Savage's Theorem extends von Neumann-Morgenstern Theorem since it considers the case in which a rational decision maker does not know either her utility function nor the probabilities to be used in order to obtain the expected values required for making choices according to the von Neumann-Morgenstern Theorem.

The notation used here is the one used by \cite{bernardo2000bayesian}. Consider a set $\Omega$. Consider a countable set $\mathcal{A}$ of available actions; for each action $a_i \in \mathcal{A}$, a partition $E_i$ of $\Omega$ and a set of consequences $C_i$. Let $\mathcal{E}$ the union of every $E_i$, we assume $\mathcal{E}$ to be an algebra of events, and $\mathcal{C}$ the union of all $C_i$.
\begin{defi}
An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$. Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$. 
\end{defi}
When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty:
\begin{defi}
A Decision Problem under Uncertainty is an uncertain environment $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ plus a preference relation $\succeq$ defined over $\mathcal{C}$ 
\end{defi}
In \cite{bernardo2000bayesian} the existence of a subjective probability measure is derived from a set of \textit{rationality axioms}, where a decision maker has some mechanism of quantifying uncertainty in terms of real numbers within the $[0,1]$ interval, and then classical Kolmogorov axioms are derived from such construction, and therefore all of the known machinery of Probability Theory. We now state Savage's Theorem. Details in \cite{kreps1988choice} and \cite{bernardo2000bayesian}.
\begin{teo}
In a Decision Problem under Uncertainty $(\mathcal{A}, \mathcal{C}, \mathcal{E}, \succeq)$, the preference relation $\succeq$ satisfies the Savage rationality axioms if and only if there exists: 
\begin{itemize}
\item A \textit{probability measure} $P$, called a subjective probability, that associates with each uncertain event $E \in \mathcal{E}$ a real number $P(E)$. 
\item A utility function $u : \mathcal{C} \to \mathbb{R}$ such that it associates each outcome with a real number $u(c)$.
\end{itemize}
Such that for $c_1$ and $c_2$ consequences in $\mathcal{C}$,
\begin{equation}
 c_1 \succeq c_2 \textrm{ if and only if } \mathbb{E}_P[u(c_1)] \geq  \mathbb{E}_P[u(c_2)].
 \end{equation}
\end{teo}
This theorem states that if a rational decision maker does not know the precise probabilities of outcomes given that an action is taken, then she chooses \textit{as if} having in mind a probability assignement to the uncertainties in her environment and using such probabilities to calculate the expected utility with respect to a subjective utility function that represents her preferences. This result also gives a precise definition of \textit{subjective probability} as a quantification of uncertainty which is used to make good decisions. See \cite{hens1992note}, \cite{gilboa2009decision} for further details. See \cite{ellsberg1961risk} and \cite{binmore2008rational} for critiques of the Savage axioms. 
%\subsection{Application: Reinforcement Learning}
%The classical setting for an agent to learn by interaction is that of a Markov Decision Process, in which at time $t$ an agent is situated in state $s_t$ and takes an action $a_t$, he then observes a new state $s_{t+1}$ and a reward signal $r_{t+1}$. The objective is to find the \textit{policy} which maximizes the long-run reward. It can be shown that this is equivalent to solving the Bellman Equations (\cite{sutton1998reinforcement}), and that the actions prescribed by such optimal policy maximize expected utility (\cite{webb2007game}). This shows the connection between the modern field of Reinforcement Learning and classical results in Decision Theory which justify choosing actions that maximize expected utility when certain \textit{rationality} assumptions are made. 
\section{Causal Decision Problems}
\subsection{On causality}
The concept of Causality deals with regularities found in a given environment which are stronger than probabilistic (or associative) relations in the sense that a causal relation allows for evaluating a change in the \textit{consequence} given that a change in the \textit{cause} is performed. The \textit{manipulationist} interpretation of Causality (\cite{woodward2005making}) is adopted here. The main paradigm is clearly expressed by \cite{campbell1979quasi} as \textit{manipulation of a cause will result in a manipulation of the effect}. Consider the following example from \cite{woodward2005making}: manually forcing a barometer to go down won't cause a storm, whereas the occurrence of a storm will cause the barometer to go down.

We adopt the formal definition of Causality given by \cite{spirtes2000causation} as a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}. Similar descriptions of the manipulationist approach can be found in \cite{holland1986statistics} and \cite{freedman1997association}. Causal inference tools, such as Pearl's do-calculus (\cite{pearl2009causality}) allows to find the effect of an intervention in terms of probabilistic information when certain conditions are met. For what remains, we assume the \textit{causal axioms} found in \cite{spirtes2000causation} with the condition known as \textit{causal sufficiency}.
\subsection{Causal Environments and Causal Decision Problems}
We define a Causal Environment to be an \textit{uncertain environment} for which there exists a causal graphical model $\mathcal{G}$ which controls the environment.
\begin{defi}
A Causal Environment is a tuple $(\Omega, \mathcal{A},\mathcal{G},\mathcal{C},\mathcal{E})$ where $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ is an uncertain environment and $\mathcal{G}$ is a Causal Graphical Model such that the set of variables of $\mathcal{G}$ correspond to the uncertain events in $\mathcal{E}$.
\end{defi}
For a Causal Environment we will distinguish two particular variables: one corresponding to the available actions, and one corresponding to the produced (caused) outcome. We are considering that only one variable can be intervened upon and that the values of such variable represent the actions available to the decision maker.
\begin{defi}
We define a Causal Decision Problem as a tuple $(\mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$ where $(\mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C})$ is a Causal Environment and $\succeq$ is a preference relation. 
\end{defi}
The intuition behind the definition of a Causal Decision Problem is this: a decision maker chooses an action $a \in \mathcal{A}$, which is automatically inputed into the model $\mathcal{G}$, which outputs the \textit{causal outcome} $c \in \mathcal{C}$.
\subsection{Related work: Causal Decision Making and Decision-Theoretic foundations of Causal Inference}
A previous attempt to formalize Decision Theory in the presence of Causal Information is given in \cite{lewis1981causal}, \cite{joyce1999foundations}. According to such formulation, a decision maker must choose whatever action is more likely to (causally) produce desired outcomes while keeping any beliefs about causal relations fixed (\cite{peterson2017introduction}). This is stated by the Stalnaker (\cite{stalnaker1968}) equation
\begin{equation}
u(a)=\sum_{x} P(a \square \to x)u(x),
\end{equation}
where $a \square \to x$ is to be read as \textit{if the decision maker does} $a$ \textit{then} $x$ \textit{would be the case} (\cite{kleinberg2013causality}, \cite{peterson2017introduction}). Lewis' and Joyce's work captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker. In Section \ref{section_savage} we refine the $\square \to$ operator by an explicit way of calculating the probability of causing an outcome by doing a certain action in terms of Pearl's do-calculus.

\cite{heckerman1995decision} provides a framework for defining the notions of cause and effect in terms of decision theoretical concepts, such as states and outcomes and gives a theoretical basis for graphical description of causes and effects, such as causal influence diagrams (\cite{dawid2002influence}). Heckerman gave an elegant definition of causality, but did not addressed how to actually make choices using causal information.

In \cite{dawid2012decision} we are presented with a decision-theoretic approach to causal inference in which a decision maker must take into account how alternatives compare against the other in terms of the \textit{average causal effect}, such approach uses the well-known influence diagrams (\cite{dawid2002influence}, \cite{dawid2003causal}) in order to derive formulas that allow an explicit calculation of the average causal effect. Influence diagrams have the ability of expressing both intervention variables and chance variables into a single graphical structure in such a way that the standard techniques for probabilistic DAG's still apply. In \cite{dawid2008identifying} an optimality criterion for sequential interventions is obtained by maximizing the expectation of outcomes.
\subsection{A von Neumann-Morgenstern type theorem for causal environments}
Consider a rational decision maker who faces a causal environment in which she knows the causal model controlling the relation between actions and outcomes. She can use the known causal model in order to find the probabilities of \textit{causing} a desired outcome given she takes a certain action. The following theorem is found in \cite{pearl2009causality}, but the intuitions that lie behind can be traced back to \cite{lewis1981causal} and \cite{joyce1999foundations}.

Consider a Causal Model $G$ and its associated distribution $P_G$ and let $C$ the set of consequences of interest for a decision maker. Then,
\begin{teo}
If a rational decision maker faces a Causal Environment and if the causal model is known, then the preference relation $\succeq$ satisfies the von Neumann-Morgenstern rationality axioms if and only if:
\begin{equation} 
a \succeq b \textrm{ if and only if } \sum_{c \in C} P(c | do(a))u(c) \geq \sum_{c \in C} P(c | do(b))u(c).
\end{equation}
\end{teo}
Equivalently, the action that must be chosen is 
\[ a^\ast = \textrm{argmax}_{a \in \mathcal{A}} \sum_{c \in C} P(c | do(a))u(c). \]
We argue that Pearl's result can be considered as a causal version of the von Neumann-Morgenstern result since it assumes that a causal model is known. If the causal model that controls an environment is known to a decision maker, then this is equivalent of being able to know the probabilities of outcomes given actions. As stated in Section 4.1 of \cite{pearl2009causality}, the utility function $u$ is considered as given, even though the Theorem \ref{vNM} guarantees its existence. Pearl argues that decision making by maximizing such function is \textit{commonsensical} without appealing to the original results in rational decision making. We avoid entering in the long-standing debate between \textit{causal} and \textit{evidential} decision making. We only note that both von Neumann-Morgenstern's and Savage's Theorems have no a priori causal interpretation.
\subsection{A Savage type theorem for causal environments}{\label{section_savage}}
When the decision maker does not know the causal model which controls her environment we argue that the decision maker is facing a particular case of Savage's Theorem. The difference here being that the subjective probability must make use of the causal nature of the environment. The idea we will follow is that Savage's Theorem gives the decision maker a subjective quantification of her uncertainty about the environment and the associated probabilities which must be used \textit{as if} they were true. We will assume that the causal environment is controlled by a Causal Graphical Model (CGM), which is a very particular causal model.

In this case, where a Causal Graphical Model controls the relation between actions and outcomes, such subjective information about the environment must consider such causal structures. For this reason, we assert that the probability distribution that the decision maker has in mind is in fact a distribution over causal structures where the decision maker uses each structure as if it were the true one in order to choose the best action within each structure by using Pearl's result. We assume a finite set of actions and outcomes. Formally:
\begin{teo}{\label{causal_savage}}
In a Causal Decision Problem  $(\mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$, where $\mathcal{G}$ is a Causal Graphical Model, we have that the preferences $\succeq$ of a decision maker are Savage-rational if and only if there exists a probability distribution $P_C$ over a family $\mathcal{F}$ of causal structures such that 
\begin{equation}
a \succeq b \textrm{ iff } \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right) \geq \sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right),
\end{equation}
where $P_g$ a the probability distribution associated with the causal structure $g$. 
\end{teo}
\begin{proof}
The decision maker is facing an environment in which any action she takes will stochastically cause an outcome $c \in \mathcal{C}$. For this reason, the decision making is facing a very particular case of decision making under uncertainty, we use Savage's Theorem to obtain a utility function $u^S$ and a probability measure $P^S$ which satisfy that the preference relation is represented by the expectation of $u^S$ with respect to $P^S$.

In such a causal environment, the Causal Graphical Model $\mathcal{G}$ contains all of the information which connects actions to outcomes, and noting that we can identify any action $a$ with $\{ c_j | E_j : j \in J  \}$ with $J$ a countable set of indexes (\cite{bernardo2000bayesian}), then:
\[\mathbb{E}_{P^S}[u(c)] = \sum_{j \in J} u(c_j)P^S(E_j).\]
For each action $a=\{ c_j | E_j \}$, $P^S(E_j)$ is the probability of \textit{causing} consequence $c_j$ by choosing action $a$. In order for the decision maker to find the probability of a certain consequence $c_j$ given that an action $a$ is performed then she must have in mind a single causal model $g$ and a way to assign probabilities over a family of causal models: let $P_C$ a probability measure over a family of causal models $\mathcal{F}$, and call $g \in \mathcal{F}$ a causal model which is considered by the decision maker with probability $P_C(g)$ and call $P_g$ the probability distribution that is generated by $g$.

This is, the uncertainty component $P^S(E_j)$ is formed by two parts: a distribution $P_C$ which represents the degree of belief of the decision maker about a specific model $g$ being the true one, and within $g$, a distribution $P_g$ used to calculate the probability of causing some consequence $c_j$ given that action $a$ is chosen.

Then,
\begin{eqnarray}
\mathbb{E}_{P^S}[u(c)] &=& \sum_{j \in J} u(c_j)P^S(E_j)\\
                                      &=& \sum_{j \in J} u(c_j) \left( \sum_g P_g(c_j | do(a))P_C(g) \right).
\end{eqnarray}
\end{proof}
\subsection{Interpretation}
Theorem \ref{causal_savage} says that a decision maker who faces a Causal Decision Problem is considering a probability distribution $P_C$ over a family $\mathcal{F}$ and, within each structure, using the term $P_g(c|do(a))$ in order to find the probability of obtaining a certain consequence given that the intervention $do(a)$ is performed; in this way, the optimal action $a^\ast$ is given by:
\begin{equation}
a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}  \sum_c u(c) \left( \sum_g P_g(c | do(a))P_C(g) \right). 
\end{equation}
We note that $a^\ast$ is obtained by taking into account the utility obtained by every possible consequences weighted using both the probability of causing such action within a specific causal model $g$ and the probability that the decision maker assign to such $g \in \mathcal{F}$.

We are considering a \textit{normative} interpretation for Theorem \ref{causal_savage} according to which a decision maker must use any causal information in order to obtain the best possible action. Such action must be obtained by considering the \textit{beliefs} of the decision maker about the causal relations that hold in her environment (the distribution $P_C$), how such relations could produce the best action when considered \textit{as if} they were true (distribution $P_g$), and the satisfaction (utility $u$) produced by the consequences of actions.

Several studies have considered how human beings use causal information when making choices. Human tend to ignore pure probabilistic information over causal information (\cite{tversky1980causal}. Humans are able to learn, and use, causal models in sequential decision making processes, even though such learning is not perfect (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{wellen2012learning}, \cite{danks2014unifying}, \cite{rottman2014reasoning}).

\section{Application: Formalizing an optimal action learning method for causal environments}
In recent papers by \cite{bareinboim2015bandits}, \cite{lattimoreNIPS2016}, \cite{sen2017identifying}, \cite{gonzalez2018playing} a decision maker is given the task to learn an optimal action after a series of learning rounds in a causal environment.

By relying on the bandit framework for causal inference developed in \cite{bareinboim2015bandits}, it is developed by \cite{lattimoreNIPS2016} an optimal action learning procedure given that the conditional interventional distributions are known and the decision maker must learn an optimal action, in the sense of minimal regret after a fixed number of learning rounds. It is shown that considering causal information yields a faster rate of optimal action identification. Later, \cite{sen2017identifying} allows a partially known causal graph.

In \cite{gonzalez2018playing} a decision maker has the task of learning an optimal action when only the \textit{structure} of such model is known. In their method, the decision maker holds beliefs about the parameters of the model, which are used in order to build a causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round. In such paper, the authors use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action.

The mentioned works fit within our Theorem \ref{causal_savage} because the assumptions can be thought of as a prior assigning probability mass of $1$ to models in a very specific class of causal models; i.e., only models that have the known conditionals in the case of \cite{lattimoreNIPS2016}, or the known partial structure in \cite{sen2017identifying} or the graphical structure in \cite{gonzalez2018playing} are considered by the prior $P_C$.

We have provided the extension to the causal case of the formal criterion (expected utility) to be used when a causal model is present in the environment and therefore we hope our result to be of interest to anyone using decision making under uncertainty.
\section{Application: Causal Nash equilibria}
We define a \textit{causal strategic game} as a strategic game within a causal environment: consider a \textit{strategic} game between $N$ rational players who are situated in a causal environment. We assume that it is \textit{common knowledge} the causal nature of the environment as well as the rationality assumption for each player. We also assume that the causal mechanism, which represented by a Causal Graphical Model $\mathcal{G}$, remains fixed and it is unknown for each player. Any knowledge or beliefs of a player $i$ about the causal structure is encoded in a probability distribution $P_i$.  This is, we have:
\begin{itemize}
\item A set of $N$ rational players; for each player $i$, a finite set $A_i$ of available actions. Let $A=A_1 \times \cdots \times A_N$ the set of action profiles.
\item For each player, a preference relation $\succeq_i$ defined over $A$.
\item A causal model $\mathcal{G}$ which produces outcomes in $C$ for each of the actions taken by players. Such model acts independently for each player.
\item A set $C$ of consequences and for each player a function $h^{\mathcal{G}}_i:A \to C$ that maps each action profile into the outcome $c \in C$ caused by action $a_i$.
\item A probability measure $P_i$ for each player which encodes any knowledge about the (real) causal structure of the environment.
\end{itemize}
We assume that each $\succeq_i$ can be represented by a Savage utility function $u_i: A \to \mathbb{R}$. According to the \textit{deductive} interpretation of an equilibrium (\cite{binmore1987modeling}, \cite{binmore1988modeling}), each player, because each one of them is rational and knows that every other player is rational as well, will use her knowledge about the causal nature of the environment in order to cause a desired outcome. We argue that a Nash equilibrium in this setting should reflect this information. Recall that a Nash Equilibrium is an action profile $a^\ast \in A$ such that for any player $i$ there is no other action $a_i \in A_i$ that yields a better outcome than the one produced by $a_i^\ast$ when the other players choose $a^\ast_{-i}$  (\cite{osborne1994course}) .

For each player, we define $u^C_i : A \to \mathbb{R}^{+}$ as follows:
\begin{equation}{\label{causal_ut}}
u^C_i (a) = \sum_{f \in A}  u_i(a) \left( \sum_g P^g_i (h_i(f) | do(a_i), a_{-i}) P_C(g) \right) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
where $P^g_i$ is the probability of causing a certain consequence within a causal structure $g$ and $P_C$ are the players beliefs about the causal structure that controls her environment, $u_i$ is the Savage utility of player $i$, and $do()$ is the well known intervention operator from \cite{pearl2009causality}. Notice that $u^C_i$ evaluates an action profile $a \in A$ in terms of: The knowledge about the causal structure of each player represented by $P_i$, which allows each player to evaluate the probability of causing outcomes in terms of actions by using the $do$ operator as well as the other actions taken by the other players, given by $a_{-i}$ and the preferences of each player $u_i$
\begin{defi}
We define a Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
\end{defi}
This is, an action profile is a Nash equilibrium if and only if each player uses her current knowledge about the causal structure of the environment in order to (causally) produce the best possible outcome given the actions taken by the other players. The existence of the Causal Nash Equilibrium is guaranteed if every $A_i$ is a nonempty compact convex set in some $\mathbb{R}^n$ and if the preference relation induced by $u^C_i$ is continuous and quasi-concave.
\section{Limitations}
We are working within the classical rationality assumption. Rationality can be ultimately though of as a \textit{consistent} or coherent way of making choices, but the precise definition has been a subject of debate. See \cite{ellsberg1961risk}, \cite{binmore2008rational}, \cite{gilboa2009decision} and \cite{machina2014ambiguity} for critiques of the Savage Rationality Axioms. Another line of decision making, which attempts to develop models of how human beings actually make choices has been developed by psychologists and economists; see \cite{TverskyKahneman74}, \cite{kahneman1979prospect}, \cite{kahneman1982judgment}, \cite{kahneman2000choices}. From a descriptive perspective it is an interesting line of research to find an adequate definition of rationality in a way that best allows for causal interpretations and from there develop the formal machinery of causal decision making.

We are considering Causal Graphical Models as the representation of causal relations in the environment, and we are considering stochastic causal relations according to the manipulationist interpretation, which is one of many. We have favoured the Causal Graphical Models over other alternatives since it has been argued that several cognitive processes, such as causal reasoning, can be best represented as graphical models (\cite{chater2013programs}, \cite{danks2014unifying}, \cite{sloman2015causality}). As alternative frameworks for causality we have: Topological Causality for Dynamical Systems (\cite{harnack2017topological}), Lamport's Causality (\cite{lamport1978time}), Granger's Causality for Time Series (\cite{granger1969investigating}), Suppes' Causality (\cite{suppes1970probabilistic}), each with its own mathematical framework. A review of several theories of Causality can be found in \cite{holland1985statistics}.

In our formulation of a Causal Decision Problem we are assuming Causal Sufficiency; this is, any cause and effect are considered within our causal graphical model $\mathcal{G}$. Even though we are considering how to make causal choices when a decision maker does not know the true causal model, we are assuming that any model that he \textit{believes} may be controlling the environment satisfies the causal sufficiency condition.
\section{Conclusion and Future Work}
We have defined a Causal Decision Problem in terms of a classical Decision Problem under Uncertainty provided of a Causal Mechanism which controls the relation between actions and outcomes. In the case in which a rational decision maker knows such relations, \cite{pearl2009causality} provides a causal version of Theorem \ref{vNM} for rational decision making.

On the other hand, when a decision maker does not know the causal mechanism, in Theorem \ref{causal_savage} we have provided a causal version of Savage's Theorem; our result explicitly states how a decision maker uses any subjective beliefs, encoded as a probability distribution over causal models, as well as the causal inference machinery within any causal structure considered in order to find an optimal action. Such subjective probabilities over causal models can be updated using any causal evidence provided after a decision has been made. Learning algorithms for causal environments fit within the machinery stated in Theorem \ref{causal_savage}, so we are confident that further implementations of such result will show that Causality is a fundamental concept for Machine Learning and Artificial Intelligence as proposed by \cite{lake2017building} and \cite{pearl2018why}.

\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
\end{document}

### 2019-05-22 Expo_TIA/TIA_Mayo_2019.tex
estructura de presentacion

--- 2018-11-25 Propuesta/Presentacion/Presentacion_Corta.tex
@@
 \begin{document}
-
-\begin{frame}
-\titlepage % Print the title page as the first slide
-\end{frame}
-
-\begin{frame}[allowframebreaks]
-\frametitle{Contents} % Table of contents slide, comment this block out to remove it
-\tableofcontents[hideallsubsections] % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
-\end{frame}
-
-%
-%\begin{frame}
-%    \frametitle{Outline}
-%    \begin{columns}[t]
-%        \begin{column}{.3\textwidth}
-%            \tableofcontents[sections={1-3}]
-%        \end{column}
-%        \begin{column}{.3\textwidth}
-%            \tableofcontents[sections={4-5}]
-%        \end{column}
-%         \begin{column}{.3\textwidth}
-%            \tableofcontents[sections={6-9}]
-%        \end{column}
-%    \end{columns}
-%\end{frame}
-
-%--------SLIDES------------------------------
-
-\section{Introduction}
-\subsection{General Overview}
-\begin{frame}
-\frametitle{General Idea}
-\begin{itemize}
-\item We live in a causally structured world. 
-\item Our actions \textit{cause} changes in the environment.
-\item We consider this changes in order to plan our actions when facing similar situations.
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Human use of Causal Relations}
-\begin{itemize}
-\item \cite{hagmayer2013repeated} show how human beings use and modify causal information.
-\item Both \cite{hohwy2013predictive}, \cite{clark2015surfing}, assert that the brain itself is a causal hypothesis tester. 
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Our setting}
-\begin{itemize}
-\item Consider a \textit{rational} decision maker in an uncertain environment.
-\item Consequences of actions can not be predicted.
-\item We add an extra to the environment: causal structure. 
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{A brief on causal models}
-\begin{itemize}
-\item Causal models encode cause-effect relations. 
-\item Relations are encoded in a graph.
-\item Can be used to predict \textit{effects} of certain causes.
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Causal Model: Example}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/example.png}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{Our Proposal}
-\begin{itemize}
-\item Current decision-making theories do not consider causal structure in the decision making processes
-\item We \textbf{propose} an autonomous rational agent who can use and learn causal information while interacting with its environment.
-\end{itemize}
-\end{frame}
-
-\section{Concepts}
-\begin{frame}
-\frametitle{Concepts}
-Concepts and Framework:
-\begin{enumerate}
-\item Causal Models
-\item Decision Problems
-\end{enumerate}
-\end{frame}
-
-\subsection{Causal Models}
-		\begin{frame}
-		\frametitle{Difference between causal and associative relations}
-		\begin{itemize}
-		\item With associative relations we can't figure out what influenced (or caused) what.
-		\item Consider this: A distribution $p(a,t)$ can be factored both as $p(a|t)p(t)$ and $p(t|a)p(a)$. 
-		\item In terms of a Bayesian Network, $T \to A$, and $A \to T$, but both can't be causal.
-		\item Causal relations need not be in temporal order: consider a barometer and a storm.
-		\end{itemize}
-		\end{frame}
-
-		\begin{frame}
-		\frametitle{Formal Definition of Causality according to \cite{spirtes2000causation}}
-		\begin{defi}
-		Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite, probability space and consider a binary relation  $\to  \subseteq \mathcal{F} \times \mathcal{F}$ such that
-		\begin{itemize}
-		\item Transitive: if $A \to B$ and $B \to C$ for $A,B,C \in \mathcal{F}$ then $A \to C$.
-		\item  Irefflexive: for each $A \in \mathcal{F}$ it is not true that $A \to A$.
-		\item Antisymetric: For $A, B \in \mathcal{F}$, if $A \to B$, then it is not true that $B \to A$.
-		\end{itemize}
-		If $A \to B$ we will say that $A$ is a cause of $B$.
-		\end{defi}
-		\end{frame}
-
-		\begin{frame}
-		\frametitle{From Graphs to Probability Measures}
-		Given a Directed Acyclic Graph, one can build the probability distribution $P_\mathcal{G}$ that expresses the relations of the graph. We require that this measure satisfies:
-        \begin{itemize}
-        \item Causal Markov Condition.
-        \item Causal Minimality.
-         \item Causal Faithfulness.
-         \item Causal Sufficiency
-        \end{itemize}
-         This requirements will be treated in an axiomatic fashion.
-		\end{frame}
-
-	\begin{frame}
-	\frametitle{Causal Graphical Models}
-	\begin{defi}
-	A causal graphical model (\cite{spirtes2000causation}, \cite{pearl2009causality}, \cite{koller2003multi}, \cite{sucar2015probabilistic}) consists of:
-	\begin{itemize}
-	 \item A set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$.
-	 \item A directed acyclic graph (DAG) $\mathcal{G}$.
-	 \item An operator $do()$ defined over the space of DAG's whose action consists of: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
-, x_{i_j} \} \in Val(\mathcal{X})$, then $do(\mathbf{X} = \mathbf{x} )$ assigns to each $X_j \in \mathbf{X}$ the value $x_{i_j}$ and deletes any incoming edge on it.
-	 \end{itemize}
-	 \end{defi}
-	 \end{frame}
-
-\subsection{Decision Problems}
-	\begin{frame}
-	\frametitle{Decision Problems}
-	A decision problem is composed by (\cite{bernardo2000bayesian}):
-	\begin{itemize}
-	\item A set of available options $\mathcal{A}$.
-	\item A (rational) decision maker with preference relation $\succeq$ defined over $\mathcal{A} \times \mathcal{A}$.
-	\item A family of uncertain events $\mathcal{E}$.
-    \item A family of possible outcomes $\mathcal{C}$
-	\end{itemize}
-	\end{frame}
-
-	\begin{frame}
-	\frametitle{Choosing rationally}
-	Savage's Theorem (\cite{savage1954the}) says that:
-\begin{teo}	
-	If a rational decision maker does not know the probabilities of uncertain events ocurring, then he must choose according to the expected utility with respect to a \textit{subjective} probability measure and a \textit{subjective} utility function.
-\end{teo}
-Details can be found at \cite{bernardo2000bayesian}, \cite{gilboa2009decision}.
-	\end{frame}
-
-\section{Research Problem}
-\begin{frame}
-About my Proposed Research
-\end{frame}
-	\subsection{Motivation}
-	\begin{frame}
-	\frametitle{Motivation: 1}
-	\begin{itemize}
-	\item Decision making under uncertain conditions is a fundamental part of intelligent reasoning (\cite{lake2017building}).
-	\item Intelligent agents often face situations where an action must be chosen in the presence of uncertain conditions.
-	\end{itemize}
-	\end{frame}
-	\begin{frame}
-	\frametitle{Motivation: 2}
-	\begin{itemize}
-	\item It is known that human beings conceive their actions on the world as \textit{intervening} in the world (\cite{hagmayer2009decision}).
-	\item Following this idea, \cite{lattimoreNIPS2016} consider decision problems where the action to be chosen is an intervention over a known causal graphical model.
-	\item What about not knowing the causal model? 
-	\end{itemize}
-	\end{frame}
-
-	\subsection{Justification}
-	\begin{frame}
-	\frametitle{Justification: Un-interpretable models}
-	\begin{itemize}
-	\item Many real-world applications of decision making are solved by \textit{associative} methods.
-	\item Associative methods capture only statistical patterns that are found in data.
-	\item Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations}.
-	\end{itemize}
-	\end{frame}
-
-	\begin{frame}
-	\frametitle{Justification: Need for nterpretable decision-making models}
-	\begin{itemize}
-	\item Current methods for Decision Making are based in Reinforcement Learning, and specifically in Deep Reinforcement Learning. 
-	\item Although good performance in the task, they can not explain \textit{why} a specific trajectory was chosen by the algorithm.
-	\item This is highly relevant in real-world applications such as self-driving cars.
-	\end{itemize}
-	\end{frame}
-
-	\begin{frame}
-	\frametitle{Justification: Explaining why}
-	\begin{itemize}
-	\item On the other hand, learning a causal model of an environment and using it to act upon the environment allows to \textit{explain} aspects of the model that a purely associative model would not be abel to explain.
-	\item It allows to ask \textit{why}.
-	\item Transferability
-	\end{itemize}
-	\end{frame}
-
-	\subsection{Research Questions}
-	\begin{frame}
-	\frametitle{Research Questions}
-	The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
-	\begin{enumerate}
-	\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good    	choices? 
-	\item How can a causal structure help a decision maker in order to guide his learning process? 
-	\item What does the rationality assumption implies about how to choose when considering causal information? 
-\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
-\end{enumerate}
-\end{frame}
-
-\subsection{Statement}
-	\begin{frame}
-	\frametitle{Problem Statement}
-	\begin{itemize}
-	\item Let  $\mathcal{G}$ a causal graphical model.
-	\item Let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-	\item Consider a rational decision maker who doesn't know the causal model which control the probabilities of observing a consequence given an action $a \in \mathcal{A}$.
-	\item We ask how the decision maker could learn about the causal structure that controls his environment in order to make good choices with respect to the decision makers' preferences.
-	\end{itemize}
-	\end{frame}
-
-	\subsection{Hypothesis}
-	\begin{frame}
-	\frametitle{Hypothesis}
-	\begin{itemize}
-	\item Let $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-	\item Then, if a decision maker doesn't know the probabilities of the uncertain events in $E$, by repeatedly making decisions he can learn a causal model of the environment and use it to find the optimal action (in the sense of expected utility theory) in less, or equal, rounds than if he doesn't consider causal information.
-	\end{itemize}
-	\end{frame}
-
-\section{Approaches and Related Work}
-\begin{frame}
-\frametitle{Approaches and Related Work}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/presentacion/Related_Work.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame} 
-\subsection{Approaches and Related Work: Reinforcement Learning}
-\begin{frame}
-\frametitle{Reinforcement Learning (RL)}
-\begin{itemize}
-\item Reinforcement Learning uses only associative patterns for finding an optimal policies.
-\item Optimal policies actually achieve the maximum expected utility as required by the rationality condition (\cite{webb2007game}).
-\item In this way, when only associative information is available, RL is a coherent learning framework. 
-\end{itemize}
-\end{frame}
-
-\subsection{Approaches and Related Work: Causal learning in humans}
-\begin{frame}
-\frametitle{Causal Learning in Human beings}
-\begin{itemize}
-\item Human beings, while facing a sequential decision problem, use and modify causal information.
-\item \cite{hagmayer2009decision}, \cite{hagmayer2013repeated} show that human beings can learn causal information in simple decision-making tasks, but remains open to study causal learning while seeking for a specific objective.
-\item 
-\end{itemize}
-\end{frame}
-\subsection{Approaches and Related Work: Joyce's Causal Decision Theory}
-\begin{frame}
-\frametitle{Joyce's Causal Decision Theory}
-\begin{itemize}
-\item An attempt to formalize Decision Theory in the presence of Causal Information was attempted by Joyce, who stated that a decision maker must choose whatever action is more likely to (causally) produce desired outcomes.
-\item  His formulation falls short since he does not consider any kind of formal causal model beyond what is commonly understood by causality
-\end{itemize}
-\end{frame}
-\subsection{Approaches and Related Work: Bandits}
-\begin{frame}
-\frametitle{Bandits: Non-causal}
-\begin{itemize}
-\item In the class of decision problems considered for this proposal it is assumed that the a rational agent must choose some available action and then receive a reward from it.
-\item Problems of this nature have been modelled as \textit{bandit} problems.
-\item Best-arm algorithms exist: none consider causal information.
-\end{itemize}
-\end{frame}
-\begin{frame}
-\frametitle{Bandits: Causal bandits}
-\begin{itemize}
-\item An action can be conceived as an \textit{intervention} over an environment, which is in fact how humans consider their actions in the world.
-\item They use and modify causal knowledge during a sequential decision making process.
-\item It is shown by \cite{lattimoreNIPS2016} that adding causal information in a fixed budget decision problem allows the decision maker to learn \textit{faster} had he not considered causal information.
-\item Their work requires that the causal model is fully known to the decision maker.
-\item this requirement is relaxed later by \cite{sen2017identifying} who requires only that some part of the causal model is known and allow interventions over the unknown part. 
-\end{itemize}
-\end{frame}
-
-\section{Proposed Solution and Methodology}
-\begin{frame}
-\frametitle{Proposed Solution and Methodology}
-\end{frame}
-	\subsection{Objectives}
-	\begin{frame}
-	\frametitle{General Objectives}
-	The proposed research has as a general objective:
-	\begin{itemize}
- \item Understand the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
-	\end{itemize}
-	\end{frame}
-
-	\subsection{Specific Objectives}
-	\begin{frame}
-	\frametitle{Specific Objectives}
-	In order to reach the General Objective we must:
-	\begin{enumerate}
-
-	\end{enumerate}
-	\end{frame}
-
-\subsection{Methodology}
-\begin{frame}
-\frametitle{Proposed Solution: Conceptual step-by-step methodology}
-\begin{itemize}
-\item Rational agent in a causal environment.
-\item Modelling agent-environment interaction as a \textit{game}.
-\item Encoding current causal knowledge as probabilsitic \textit{beliefs}
-\item Using current knowledge to make a choice.
-\item Updating causal information.
-\end{itemize}
-\end{frame}
-
-\subsection{Validation}
-\begin{frame}
-\frametitle{Validation}
-How to know if objective has been achieved?
-\end{frame}
-
-\begin{frame}
-\frametitle{Concurrent Validation}
-We will be comparing against standard Reinforcement Learning methods such as Q-Learning (\cite{watkins1992q}) since it has been shown that optimal policies achieve the objective of Max Expected Utility.
-\end{frame}
-
-\begin{frame}
-\frametitle{A conjecture}
-It needs to be shown that given the rewards obtained by a decision-making procedure that achieves the MEU, then my proposed algorithm must stay close to it.
-\begin{teo}
-Let $(X_1,X_2,...)$ the rewards obtained by a decision-making procedure which is known to converge to the max expected utility (or an optimal policy), then, if $(Y_1, Y_2,...)$ are the rewards obtained by a decision-making procedure that uses causal information, then for all $\varepsilon > 0$ there exists an $N_\varepsilon \in \mathbb{N}$ such that $ | X_t - Y_t | < \varepsilon$ for any $t > N_\varepsilon$.
-\end{teo}
-\begin{proof}
-To be proved.
-\end{proof}
-\end{frame}
-
-\begin{frame}
-In this case, if we succeed in finding an action whose reward equals that of RL algorithms, then we will have chosen the optimal action using Causal Information as proposed in the Hypothesis.
-\end{frame}
-
-\begin{frame}
-\frametitle{How to}
-How to
-\end{frame}
-
-\subsection{Proposed Solutuion: A guiding principle}
-\begin{frame}
-\frametitle{A guiding principle for Causal Decision Problems}
-We propose the following principle as a normative guide for how causal information must be used in order to find a solution to a decision problem where the preferences of the decision maker are assumed to be rational.
-We consider only the case where the decision maker's utility is given by the outcome of a $\{0,1 \}$-binary variable.
-It remains to be refined and extended for a general case.
-\end{frame}
-
-\begin{frame}
-\begin{prop}
-Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
-Let $Y$ a variable in $\mathcal{G}$ the variable whose realizations correspond to the consequences of the decision problem once the agent has chosen an action $a \in \mathcal{A}$ and assume that $Y$ only takes values in the set $\{ 0,1\}$ where $1$ is a more desired outcome for the agent.
-Then, the $\textit{solution}$ for this decision problem is given by the action $a^\ast \in \mathcal{A}$ that satisfies:
-\[ P(Y=1 | do(a^\ast) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A}. \]
-\end{prop}
-\end{frame}
-
-\subsection{First Step: Modelling a Causal Decision Problem as a Game}
-\begin{frame}
-\frametitle{Games}
-\begin{itemize}
-\item In this way, the first step in order to solve the stated problem is to model the interaction of the (original) decision maker and his environment as a \textbf{game} between two players: 
-\begin{enumerate}
-\item the (original) decision maker, 
-\item and a player which will be called Nature, whose actions are to be guided by the causal model.
-\end{enumerate}
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item  Since in decision problems under uncertainty it is assumed that the \textit{state} of the environment is unknown, we will assume that player Nature has the first move and he assigns some state to any of the variable in the causal model $\mathcal{G}$.
-\item After the decision maker makes his \textit{play}, Nature will \textit{respond} to the action selected according to the causal relations expressed by $\mathcal{G}$.
-\item This action-response dynamic forces to consider some notion of order between the players' moves, and because of this reason we must use the \textit{extensive-form} games
-\end{itemize}
-\end{frame}
-
-\subsection{Second Step: Belief formation}
-\begin{frame}
-\frametitle{Belief Formation}
-\begin{itemize}
-\item The whole idea is that a decision maker is able to learn from his environment by observing, and reasoning about, the effects of his actions.
-\item The observed information will modify his current knowledge, which is encoded as probabilistic beliefs about the environment in a way that is similar to how human beings modify causal knowledge while intervening in the world
-\end{itemize}
-\end{frame}
-
-\subsection{Third Step: Making choices}
-\begin{frame}
-\frametitle{Making Choices}
-\begin{itemize}
-\item The decision maker will use his current knowledge \textit{as if} it were the truth about the causal relations of his environment.
-\item With this information he will choose the action that maximize the Expected Utility, according with our guiding principle.
-\end{itemize}
-\end{frame}
-
-\subsection{Third Step: Belief updating}
-\begin{frame}
-\frametitle{Belief Updating}
-\begin{itemize}
-\item If a decision maker encodes his current knowledge, and ignorance, about some relevant characteristic of his environment, then it is known that belief updating using Bayes' Theorem is the only way of updating that is coherent with the rational preference axioms.
-\item it remains to be answered how to concretely use causal information in order to update the parameters that control the distributions used to express beliefs in a tractable way
-\end{itemize}
-\end{frame}
-
-\section{Methodology divided into simpler cases}
-\begin{frame}
-We will show how the methodology applies in simpler and easier cases where the decision maker has some exta info about his environment.
-\end{frame}
-\subsection{Fully knowing the causal graphical model}
-\begin{frame}
-\frametitle{Fully knowing the causal graphical model}
-\begin{itemize}
-\item If the decision maker knows the causal graphical model, then following the guiding principle, he could easily calculate the effect on the target variable that any of his actions has and then choose the action that has the highest probability of causing a \textit{desired} value of the target variable.
-\end{itemize}
-\end{frame}
-
-\subsection{Knowing only the graph structure}
-\begin{frame}
-\frametitle{Knowing only the causal graph structure}
-\begin{itemize}
-\item If a decision maker knows the structure of the graph, but he does not know the parameters,
-\item then he encodes this ignorance, along with any useful previous knowledge, as a probability distribution over a suitable space.
-\item In general, what must be specified is a way of encoding current causal knowledge into a probability distribution in such a way that a \textit{local} (i.e.
-in each round) causal model can be obtained from current beliefs in order to make a decision using current causal knowledge.
-\item This means that using current beliefs, the agent forms a causal model which is used as if it were the true causal model.
-\end{itemize}
-\end{frame}
-
-\subsection{General case: Model unknown}
-\begin{frame}
-\frametitle{Model unknown}
-If the model is completely unknown, two ideas appear:
-\begin{enumerate}
-\item If the max number of variables is known.
-\item If it is not.
-\end{enumerate}
-\end{frame}
-\begin{frame}
-\frametitle{Idea 1 for the General Case}
-\begin{itemize}
-\item If the maximum number of variables is assumed to be known by the decision maker, then the situation simplifies since a decision maker could generate distributions that represent causal relations among variables.
-\item For example, a Dirichlet Process (\cite{ferguson1973bayesian}, \cite{ghosal2017fundamentals}) could be used to generate Dirichlet distributions that are used generate Conditional Probability Tables in order to specify a causal model as in the previous case.
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Idea 2 for the General Case}
-\begin{itemize}
-\item If the number of variables is unknown, then the beliefs that the decision maker holds must allow unbounded cardinality on the number of variables.
-In that case, a distribution over a \textit{graph space} could be used in such a way that when sampled a graph structure is obtained and used as in the previous case.
-\end{itemize}
-\end{frame}
-
-\section{Working and Publications Plan}
-\subsection{Working Plan}
-\subsection{Publications}
-
-\section{Some Results}
-\begin{frame}
-\frametitle{Some results}
-To show the factibility of this proposal, we considered a test scenario and two of the previously mentioned cases.
-\end{frame}
-\subsection{Test scenario}
-\begin{frame}
-\frametitle{Test Scenario}
-Consider a sick patient who arrives at a hospital and he either has disease $A$ or disease $B$.
-The doctor can either give him some pill or send him into surgery.
-Both treatments entail risks and whether the treatment cures the patient or not depends on which disease it had originally.
-The doctor could be facing a mutation from a known disease, so she has some knowledge about what could happen if a treatment is given to the patient.
-Using her previous knowledge as a true model, she can choose a treatment and observe the outcome from which she will learn about this disease, so she could make a better decision the next time a similar patient arrives.
-\end{frame}
-
-\begin{frame}
-\frametitle{Causal model}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=0.7\textwidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/causal_graph.png}}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{Variables of the Model}
-The variables in the model are: 
-\begin{itemize}
-\item \textbf{Disease:} Either $A$ or $B$.
-\item \textbf{Treatment:} Either pill or surgery.
-\item \textbf{Reaction:} Either dying or surviving.
-\item \textbf{Lives:} Either living or dying.
-\end{itemize}
-The variable \textit{Lives} is the \textit{target variable} and, in this example, the only variable that can be intervened upon is the variable \textit{Treatment}.
-The decision maker prefers an outcome in which the patient lives.
-\end{frame}
-
-\begin{frame}
-For this test scenario whose causal graphical model is shown in Figure \ref{causal_model}, we see  by applying the Pearl's do-calculus that the interventional distribution $P_{do(Tr)}(Y)$ is given by
-\[ P(Y | do(Tr))=P(Y | D, Tr, R)P(R | Tr) P(D). \]
-\end{frame}
-
-\subsection{Case 2: Only the structure is known}
-\begin{frame}
-\frametitle{Knowing only the graph structure}
-\begin{itemize}
-\item Given the structure of the model; i.e., the variables in it and the directed edges, the joint distribution of those variables can be expressed as a product of the form $P(X_j | Pa(X_j))$ where $Pa(X_j)$ are the parents of $X_j$ in the underlying DAG in $\mathcal{G}$
-\item Since these distributions fully characterize the model, the decision maker will have beliefs over each one of these parameters.
-Notice that each of these parameters is itself a distribution of length equal to the number of possible values of the variable which is being conditioned, call the maximum number of possible values $k$ .
-\end{itemize} 
-\end{frame}
-
-\begin{frame}
-\begin{itemize}
-\item We use the $k$-dimensional Dirichlet distribution, whose support is the set of probability vectors of length $k$ \cite{hjort2010bayesian}. 
-\item The $k$ dimensional Dirichlet distribution has a density $f$ with respect to the Lebesgue measure given by
-\[ f(x_1,...,x_k | \alpha_i,...,\alpha_k)=\frac{1}{B(\alpha)}  \prod_{i=1}^k x_i^{\alpha_i-1}\]
-\item The decision maker will have beliefs about the CPT's in the form of parameters of several Dirichlet distributions.
-\item Using this fully specified (structure + parameters) as a true model, the decision maker will make his choice as according to the guiding principle.
-\item When the decision maker observes the value of the target variable, he will update the parameters that specify his beliefs.
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Belief Updating}
-For the belief updating, given a new data point,  two cases must be considered:
-\begin{itemize}
-\item The variable to update has no parents.
-\item The variable to update has parents.
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Belief updating}
-\begin{itemize}
-\item In the first case, if a prior Dirichlet($\alpha$) is used, then the posterior is given by
-\[ \textrm{Dirichlet}(\alpha + c) \]
-where $c$ is a vector of the number of occurrences of that observed data point. 
-\item For the second case, we must consider both the occurrences of that data point as well as the parents for each of the variables.
-Following \cite{barber2012bayesian} we denote as $\theta_i(X,j)$ the number of times the event $\{X=i | Pa(X)=j\}$ is observed.
-In this case, if the prior of $X_i$ conditioned on its parents having the value $j$ is given by a a Dirichlet($\alpha$), then the posterior for the variable $X_i$ given an observed data point is given by
-\[ \prod_j \textrm{Dirichlet}(\alpha + \theta_i(\cdot,j)). \]
-\end{itemize}
-\end{frame}
-
-\subsection{Implementation}
-\begin{frame}
-\frametitle{Implementation}
-\begin{itemize}
-\item We begin with a random assignment of the $\alpha$ parameter for each of the distributions considered.
-\item  We use Dirichlet distribution for each of the conditional probability tables that appear in the factorization of the joint probability for the graph of $\mathcal{G}$.
-\item With this parameters, the decision maker forms a causal model and chooses the action that maximizes the probability of the desired value for target variable.
-\item With this action chosen, we simulate an outcome from the causal graphical model using the chosen action as an intervention.
-\item This evidence is used to update the parameters, which then will be used to generate a new causal model, and so on.
-\end{itemize}
-\end{frame}
-
-\subsection{Experiments}
-\begin{frame}
-\frametitle{Experiments}
-We show the results of two experiments.
-We compare the performance obtained by the causal agent, a \textit{random agent} who selects his actions at random, and an agent performing Q-learning (\cite{watkins1992q}).
-Q-learning was chosen since it learns an \textit{optimal policy} in the sense of the Bellman Equation (\cite{sutton1998reinforcement}) and it is shown in \cite{webb2007game} that such optimal policies also maximize expected utility.
-\end{frame}
-
-\begin{frame}
-\frametitle{20 rounds}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/20_rounds_format.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{50 rounds}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/50_rounds_format.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{100 rounds}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/100_rounds_format.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-\frametitle{200 rounds}
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\includegraphics[width=\linewidth]{/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/200_rounds_format.png}
-\caption{Causal graphical model for the test scenario: the target variable \textit{Lives} is causally influenced by the disease the patient has, the treatment assigned and the survival to the secondary effects of treatment.}
-\label{causal_model}
-\end{center}
-\vskip -0.2in
-\end{figure}
-\end{frame}
-
-\begin{frame}
-More details in \cite{gonzalez2018playing}
-\end{frame}
-
-\section{Conclusions}
-\begin{frame}
-\frametitle{Conclusions}
-We have proposed that an autonomous, rational agent who faces an uncertain environment which is governed by an unknown causal mechanism can make use of the causal relations that hold in such environment in order to find an action that will take him into the way of a desired result, in terms of his (rational) preferences.
-Also, causal information can be acquired while interacting with the environment and this newly acquired information will enrich the decision making procedure.
-\end{frame} 
-\begin{frame}
-We have taken inspiration in how human beings use causal information when interacting with the world, which has a complex causal structure, in order to pursue objectives.
-Following this intuitions we proposed a \textit{guiding principle} to exploit causal information that must be followed by a rational agent in order to be coherent with the definition of rationality.
-From the guiding principle we stated a systematic approach into causal decision making and considered three cases to be solved in order.
-In two of the three cases we showed that a performance similar to classical decision-making algorithms can be achieved while also learning a causal model from the environment.
-We are considering.
-If this research is allowed to continue, we will delve into general settings that resemble realistic situations in which the causal model that controls an environment is barely known.
-\end{frame}
-
-%------- REFERENCES ----------------------
-
-\begin{frame}[allowframebreaks]
-\frametitle{References}
-\bibliographystyle{apalike}
-\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
-\end{frame}
-
-%------------------------------------------------
-
-\begin{frame}
-\Huge{\centerline{The End}}
-\end{frame}
-
-%------------------------------------------------
+\section{Causality}
+\section{Rationality}
+\section{Decision Making under Uncertainty}
+\subsection{von Neumann-Morgenstern}
+\subsection{Savage's Theorem}
+\section{Causal Decision Making}
+\subsection{Causal version of vNM}
+\subsection{Causal version of Savage}
 

### 2020-08-25 3rd_year_eval/Formato/3rd_year_short.tex
reporte tercer año

\documentclass[english,letterpaper,12pt,final]{article}
\usepackage{graphicx}
\usepackage{etex}

\sloppy
%\special{papersize=8.5in,11in}% This command prevents the use of other paper size in the dvi2ps conversion
%------Document properties------------
\usepackage{setspace}
%\usepackage{pdfpages}
\usepackage{pgfgantt}
\usepackage{enumitem}
%------math packages-------------------
\usepackage{amsthm}
\usepackage{tkz-berge} % Drawing graphs
%------algorithms packages-------------------
%\usepackage[ruled,vlined]{basic/algorithm2e}
\usepackage[noline,linesnumbered,noend]{algorithm2e}
%\linesnumbered
%------bibliography packages-----------
\usepackage{natbib}

\usepackage[utf8]{inputenc}

\usepackage[font=small,labelfont=bf]{caption}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}

\usepackage{natbib}

\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
\newcommand{\leci}{<\!\!\!\!\!\!\!\bigcirc}
%\newcommand{\lessint}{$\bigcirc$}
%------corporative identity of INAOE---
\usepackage{eso-pic}
\usepackage{pstricks}
\usepackage{INAOEMacros}

\usepackage{dsfont}
\usepackage{lineno,hyperref}

\usepackage{amssymb}
\usepackage{pifont}
\usepackage{url}
\usepackage{array}
%\usepackage{subfigure}
\usepackage{lscape}
\usepackage{amsfonts}
\usepackage{amsmath}

\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{multirow}
%\usepackage{mathtools}

%\usepackage[dvipdfmx]{graphicx} 
%\usepackage{bmpsize}

\usepackage{caption}
\usepackage{subcaption}
%------floating -----------------------
\usepackage{float}
\floatstyle{plain}
\newfloat{codigo}{thp}{lop}[section]
\floatname{codigo}{Algorithm}
\usepackage{booktabs}
%
%usepackage{lscape}
%\usepackage{pdflscape}
%\usepackage{rotating}
\usepackage{tablefootnote}
%------figure packages--------------
%\usepackage{subfigure}

%------page layout---------------------
\oddsidemargin 0.36cm
\evensidemargin 0.36cm
\marginparwidth 0.0cm
\textwidth 16.80cm

\topmargin -1.5cm
\headheight 1.0cm
\headsep 1.0cm
\footskip 1.0cm
\textheight 22.34cm

%\theoremstyle{plain}
%\newtheorem{teo}{Theorem}
%\newtheorem{prop}[teo]{Proposition}
%\newtheorem{defi}[teo]{Definition}
%\newtheorem{obs}[teo]{Observation}
%\newtheorem{lem}[teo]{Lemma}
%\newtheorem{cor}[teo]{Corolary}

%\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{56}}%
\theoremstyle{definition}
\newtheorem{rmk}{Definition}[section]
\newtheorem{teo}{Theorem}[section]
\newtheorem{thm}{Theorem}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{cor}{Corollary}[section]
\newtheorem{problem}{Problem}
\newtheorem{defi}{Definition}[section]
\newtheorem{pf}{Proof}
\newtheorem{pot}{Proof of Theorem \ref{thm2}}

%----------myDeclarations------------
\DeclareMathOperator*{\argmin}{argmin}
\definecolor{blue1}{rgb}{0.58,0.70,0.84}
\definecolor{blue2}{rgb}{0.93,0.95,1.0}
\usepackage{xcolor,colortbl}
\usepackage{titlesec}
\setcounter{secnumdepth}{4}
\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

%-----end of preamble------------------
\pagenumbering{roman}
\begin{document}

% Title page
\input{gfrontpage}
%\doublespacing
\singlespacing
\tableofcontents
\newpage
\begin{abstract}
In order to answer how a rational decision maker \textit{should} use causal information, we defined a Causal Decision Problem together with an optimality criterion; as an extension, the definition of a causal decision problem was brought into to a multi-agent setting where each agent is affected by his decisions and the decisions of the other agents, this situation is called a \textit{strategic game} in the literature. For such games we stablished the notion of a Causal Nash Equilibria. Since our framework allows for learning by interaction, we used an analogy between non-isomorphic mathematical structures defined over the same set and the algebras induced by associative and causal levels of information in order to argue that Reinforcement Learning, in its current formulation, is not a causal problem, independently if the motivation behind it has to do with an agent taking actions. A natural consequence is the need for decision making criteria that merge causal information and rational decision making. In the same line of learning by interaction, we propose an algorithm for learning a causal structure by interaction using random graphs. 
\end{abstract}
\pagenumbering{arabic}

\section{Context}
\subsection{Introduction}
Given the relevance of decision making and causal inference in artificial intelligence \citep{lake2017building,pearl2018why}, we study the problem of \textit{how to make good decisions} when causal information is available to an intelligent agent. This problem has been previously approached both by philosophers \citep{lewis1981causal,stalnaker1968,joyce1999foundations, waldmann2013causal,danks2014unifying} as well as computer scientists \citep{pearl2009causality,bareinboim2015bandits,ortega2015causal,lattimoreNIPS2016,sen2017identifying}. We have noticed that such works either focus on obtaining purely theoretical implications of making decisions this way, or algorithmically learning a good action without explicitly stating the required assumptions. With such limitations in mind, we have proposed a formal criterion for decision making under uncertainty making use of causal information in such a way that learning algorithms can be derived from our criterion.

\subsection{Problem Statement}
Consider the problem of choosing one among several available actions with uncertain consequences; this means that once an action is chosen, then some uncertain event will produce an outcome, such setting is known as Classical Decision Making. Here, we consider that actions and outcomes, or consequences, are related in terms of a Causal Graphical Model $\mathcal{G}$ in such a way that actions, uncertain events and consequences are \textit{causally related}. We consider the problem of making a good choice when the decision maker is aware of the causal nature of his environment. A formal statement of such problem is given in the Technical Report \cite{gonzalez2019causal}.

\subsection{Relevance}
The classical theorems for decision making under uncertainty provide formal criterion for decision making if rationality is assumed. This criterion is the basis for many of the techniques used in Artificial Intelligence; for example, Reinforcement Learning algorithms learn \textit{optimal policies} in such a way that any action prescribed by the optimal policy achieves the maximum expected utility \citep{sutton1998reinforcement,Puterman:1994:MDP:528623}. Algorithms which rely on the von Neumann-Morgenstern or Savage Theorems are based on associative relations which are expressed using correlations or probability distributions. It is a natural question how to formalize rational decision making when causal information is present. Such question has been previously considered by \cite{nozick1969newcomb}, \cite{lewis1981causal}, \cite{joyce1999foundations} without an explicit optimality criterion for decision making and by \cite{pearl2009causality} who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known.

\subsection{Research Question}
The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
\begin{enumerate}
\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices? \item How can a causal structure help a decision maker in order to guide his learning process? 
\item What does the rationality assumption implies about how to choose when considering causal information? 
\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
\end{enumerate}

\subsection{General Objectives}
			The proposed research has as a general objective to understand what are the implications of causality for rational decision maker who faces an uncertain environment and how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker.
\newpage
\section{First Year}{\label{first_year}}
\begin{itemize}
\item During the 2017-2018 academic year, the research problem to be studied was defined, and from the general research problem we defined a mathematical problem which we called a Causal Decision Problem.
\item We proposed a preliminar solution which allowed for an optimal action to be found by a decision maker under some assumptions:
\begin{equation}{\label{first_version}}
P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A},
\end{equation}
\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility \citep{gonzalez2018playing}.
\item Poster presented at Causality Workshop ICML 2018.
\end{itemize}

\section{Second Year}
\begin{itemize}
\item The solution criterion given in Equation \ref{first_version} for causal decision making was further developed and extended for the case when a decision maker does not know the causal model which controls his environment:
\begin{equation}{\label{second_version}}
a \succeq b \textrm{ iff } \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right) \geq \sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right).
\end{equation}
\item Using \ref{second_version} we were able to define the notion of a Causal Game and the idea of a Nash equilibria within such game. We define a function $u^C_i : A \to \mathbb{R}^{+}$, for each player:
\begin{equation}{\label{games_causal_ut}}
u^C_i (a) = \sum_{f \in A}  u_i(a) \left( \sum_g P^g_i (h_i(f) | do(a_i), a_{-i}) P_C(g) \right) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
\item We define a Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}{\label{nash_eq}}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
 \item We argued that Reinforcement Learning is not a Causal Problem in itself in \cite{gonzalez2019reinforcement}.
 \item Technical report CCC-19-002 which we cite here as \cite{gonzalez2019causal}. 
\end{itemize}

\section{Third Year}
During the 2019-2020 academic year, we further refined the concept of a Causal Game and that of a Nash Equilibrium for Causal Games, work which was presented at a MICAI 2019 Workshop and at a NeurIPS 2019 Workshop, both as oral presentations. We also studied causal-structure learning via interaction with a stochastic causally controled environment, similar as we had done in the first year; the novelty being that now we are considering that the decision-maker does not know the graphical structure of the model which controls the environment, which has been left as an open problem in various works such as \cite{lattimoreNIPS2016} and \cite{sen2017identifying}. To tackle this new difficulty, we use an extension of our previous approach, this extension relies on random graphs.
\subsection{Causal Games}
We refined the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:\\
\\
For each player $i \in N$ in the strategic game, we define the following probability distribution over consequences:
\begin{equation}{\label{causal_utility}}
p^a_i (c) =  p^\omega_i (c | do(a_i), a_{-i}) p_i(\omega)\textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
where $p^\omega_i$ is the probability of causing a certain consequence within a causal structure $\omega$ and $p_i$ are the player's \textit{posterior beliefs} about the causal structure that controls the environment, and $do()$ is the well known intervention operator from \cite{pearl2009causality}. We now define:
\begin{equation}
u^C_i (a) = \sum_{c \in C}  u_i(c) p^a_i (c) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
Notice that $u^C_i$ evaluates an action profile $a \in A$ in terms of: The knowledge about the causal structure of each player represented by $p_i$, which allows each player to evaluate the probability of causing outcomes in terms of actions by using the $do$ operator as well as the other actions taken by the other players, given by $a_{-i}$ and the preferences of each player $u_i$. Using this new function, we define the equilibria for a strategic game with causal information and Bayesian players as:
\begin{defi}
A Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
\end{defi}
This is, an action profile is a Nash equilibrium if and only if each player uses her current knowledge about the causal structure of the environment in order to (causally) produce the best possible outcome given the actions taken by the other players. The existence of the Causal Nash Equilibrium is guaranteed if every $A_i$ is a nonempty compact convex set in some $\mathbb{R}^n$ and if the preference relation induced by $u^C_i$ is continuous and quasi-concave.

\subsection{Causal-Structure Learning: A Random Graph Approach}{\label{random_graph}}
Human beings focus on \textit{local} aspects while learning causal relations which are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning. Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.\\
\\
Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$. Our extension will make use both of our decision making criterion, random graphs and the \textit{as if} intuition present in our work.\\
\\
We call our method \textit{local} since we use the uncertainty (from the agent's point of view) about the existence of a causal relation between pairs of variables, which are local components of the structure of a graph. Such uncertainty is to be updated in terms of what is \textit{observed} from realizations of the true causal mechanism that controls the environment. Since observations alone do not suffice to uniquely determine a DAG, we make use of a partial ordering between the variables in the model that, at least in principle, can not be a cause of each other. This ordering must be an input product from expert knowledge.\\
\\
Let a rational agent consider the following set of variables $\mathcal{X}=\{ X_1,...,X_n \}$ which are causally related, even though the agent does not now how; the agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$. Let $p_{ij}$ be the belief that the agent has over a causal relation (directed link) existing between variable with index $i$ and variable with index $j$. Let $G$ an initial {\em random} DAG formed as follows: node set is $N=\{1,...,n\}$ and a link exists between $i$ and $j$ with probability $p_{ij}$. Now, we use the methodology found in \cite{gonzalez2018playing}  as well as our results presented in \cite{gonzalez2019theorems} in order to find the best action $a^\ast$ for the obtained graph $G$. The best action is taken, and a full realization $X_1=x_1,...,X_n=x_n$ is observed.\\
\\
Now, we update the $p_{ij}$'s using Bayes Theorem as follows: for each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$, so we have:
\begin{equation}{\label{bayesian_updating}}
p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
\end{equation}

\subsubsection{Implementation}
The way in which we implemented the previously described algorithm is as follows:\\
We are using the same setting as in the experiments described in Year 1; this is, a patient sick with an unknown disease and a doctor who must choose between two possible treatments, each one with some risk of causing a life-threatening allergic reaction. Our agent (the doctor) must now learn which variables (disease, treatment, reaction, dying) are causally related.\\
\begin{enumerate}
\item Initialize $p_{ij}$ randomly
\item Initialize graph $G$ with link from node $i$ to node $j$ with probability $p_{ij}$
\item Find optimal action $a^\ast$ for graph $G$ and probabilities given by a count of the observations.
\item Calculate probability of the approximate model $G$ producing the full observation $X_1 = a^\ast, X_2=x_2,...,X_n=x_n$. 
\item Update $p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.$
\end{enumerate}
\subsubsection{Results}
Ground truth:\\
We know that the true causal model is a causal graph in which it holds that:
\begin{itemize}
\item Disease causes Final
\item Treatment causes Final
\item Treatment causes Reaction
\item Reaction causes Final
\end{itemize}

As seen in Figure \ref{causal_model}.

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=0.7\textwidth]{/Users/MauricioGS1/INAOE/3rd_year_eval/Figures/eps_files/causal_model_espanol.eps}}
\caption{Causal graphical model for the test scenario: the target variable \textit{Final} is causally influenced by the \textit{Enfermedad} (disease) the patient has, the \textit{Tratamiento} (treatment) assigned and the survival to the secondary effects of treatment.}
\label{causal_model}
\end{center}
\vskip -0.2in
\end{figure}
\newpage
Experiment:\\
We ran 200 times an experiment in which the agent attempts to learn the optimal action for 10 rounds. For each of these 10 rounds the agent updates the probabilities of two variables being causally connected. We show the behavior of the probabilities in Figure \ref{All_beliefs}.\\
\\
Conclusion:\\
We notice that the connections which do exist in the true model (Figure \ref{causal_model}) do go up, which means that our agent is in fact learning connections that hold. On the other side, connections that are not present do not go down, this could be due to the relatively small number of learning rounds, or to the size of the model. Work in progress to improve these results.
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/3rd_year_eval/Figures/eps_files/all_connection_beliefs_exp_10_rounds_200_[Tratamiento].eps}}
\caption{Behavior for all connections in the graph}
\label{All_beliefs}
\end{center}
\vskip -0.2in
\end{figure}
We also notice that our algorithm learns that Disease is causally connected with the variable which represents the patient living or dying as seen in Figure \ref{disease_live}.
\newpage
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/3rd_year_eval/Figures/eps_files/connection_beliefs_(Enfermedad_Final)_exp_10_rounds_200_[Tratamiento].eps}}
%/Users/MauricioGS1/INAOE/Propuesta/Formato/figures/50_rounds_format.eps
\caption{Behavior of the probability of disease causing the patient to live or die}
\label{disease_live}
\end{center}
\vskip -0.2in
\end{figure}

\section{Publications}
\subsection{First Year}
\begin{itemize}
\item The learning procedure described in Section \ref{first_year} was presented at the CausalML Workshop during ICML 2018 in Stockholm, Sweden. Link to Pre-print: \url{https://arxiv.org/abs/1807.01268}
\end{itemize}
\subsection{Second Year}
\begin{itemize}
\item A Technical Report containing the overall details of the research in progress. Cited as \cite{gonzalez2019causal}. \url{https://ccc.inaoep.mx/archivos/CCC-19-002.pdf}
\item A paper in which via an algebraic argument we state that Reinforcement Learning, in its current mathematical definition, can not be considered as a causal problem. Co-authored with F. Orihuela-Espina. Implied by this result is the need for a Causal version of learning-by-interaction. Cited as \cite{gonzalez2019reinforcement}
\end{itemize}
\subsection{Third Year}
\begin{itemize}
\item A paper including the first version of the Causal Nash Equilibrium was presented at a MICAI 2019 Workshop  \url{https://ccc.inaoep.mx/~bio/2019_MICAI_CausalityWorkshop/}. Cited as \cite{gonzalez2019games}
\item A paper including the decision-making results was orally presented at a NeurIPS 2019 Workshop  \url{https://nehzux.github.io/NewInML2019/}.  Cited as \cite{gonzalez2019theorems}
\item Journal paper submitted to Theory and Decision \url{https://www.springer.com/journal/11238/}
\end{itemize}

\section{Future Work}
\subsection{Learning graphical structure}
\begin{itemize}
\item More complex causal model.
\item Intervening more than one variable.
\item Limitations on the learning proposal.
\item OpenAI Gym scenario.
\item Theoretical guarantees on our structure-learning proposal.
\end{itemize}

\subsection{Bayesian Non parametrics}
Bayesian non parametrics is a field within Bayesian statistics in which priors of infinite dimension are allowed, so models with a greater expressibility can be developed.
\begin{itemize}
\item Priors over complex graphs.
\item Fully Bayesian learning procedure.
\end{itemize}

%\subsection{Extra: Optimal Transport and Causality}
%Optimal Transport \citep{villani2009ot} is a field of mathematics which deals with probabilistic objects and transformations within. The basic intuition is to \textit{transport} one such object into another, which is something similar to what it is done in causality, where the user \textit{transforms} an observational distribution into a \textit{interventional} distribution via the rules of Do-Calculus.\\
%\\
%Given a pair of measures $\mu$, $\nu$ both defined over polish spaces $(X,d)$ and $(Y,d')$ and a cost function $c: X \times Y \to \mathbb{R}$, we want to solve
%\[ \textrm{Minimize } \int_{X \times Y} c(x,y) d \pi(x,y) \textrm{ for } \pi \in \Pi (\mu,\nu), \]
%where $\pi$ is a probability measure having as \textit{marginals} $\mu$ and $\nu$. This is called the Kantorovich Problem, which is in fact a relaxed version of the \textit{Monge Problem}, in which we require that $y=T(x)$; this is also known as finding a deterministic coupling between $\mu$ and $\nu$\\
%\\
%Within causal graphical models, one has a joint distribution 
%\[ P(X_1 , ... , X_n) \]
%which is expressed as
%\[ P(X_1 , ... , X_n ) = \prod P(X_i | pa(X_i)) \] 
%and an interventional distribution
%\[ P(X_1 , ... , X_{j-1}, X_{j+1} , ... , X_n | do(X_j = x_j)), \]
%which can be expressed, if some conditions hold, using the rules of Pearl's do-calculus.\\
%\\
%If we call $\mu=P(X_1 , ... , X_n)$ and $\nu=P(X_1 , ... , X_{j-1}, X_{j+1} , ... , X_n | do(X_j = x_j))$, some questions arise:
%\begin{itemize}
%\item Is it possible to obtain the $do$ operator as a particular case of mass transportation? This question seems natural since the $do$ operation reconfigures $\mu$ into $\nu$.
%\item What cost can be associated in a natural way to the transport from $\mu$ to $\nu$? Is it related to the underlying graphical structure? 
%\item What is the distance between $\mu$ and $\nu$ in terms of the graphical structure
%\end{itemize}

\subsection{Timetable}
\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\includegraphics[width=0.9\textwidth]{/Users/MauricioGS1/INAOE/2nd_year_eval/Formato/file.eps}
\caption{Gantt Diagram for the proposed research}
\label{Gantt}
\end{center}
\vskip -0.2in
\end{figure}
\newpage
%------------- References --------------------
\singlespacing
%doi=false,isbn=false,url=false
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}

%------------- Apendice --------------------

\end{document}

### 2020-08-25 3rd_year_eval/Slides/3rd_year_slides.tex
slides

\documentclass{beamer}
\mode<presentation> {

\usetheme{Madrid}

}
\usepackage{booktabs}
\usepackage[utf8]{inputenc}
%\usepackage[spanish, mexico]{babel}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{amssymb,amsthm,amsmath,latexsym}
\usepackage{natbib}
\theoremstyle{plain}
\newtheorem{teo}{Theorem}
\newtheorem{prop}[teo]{Proposition}
\newtheorem{defi}[teo]{Definition}
\newtheorem{obs}[teo]{Observation}
\newtheorem{lem}[teo]{Lemma}
\newtheorem{cor}[teo]{Corolary}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}

\title{Use and acquisition of causal relations for decision making under uncertainty using imperfect information games} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Mauricio Gonzalez Soto} % Your name
\institute[INAOE] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Instituto Nacional de Astrofísica Óptica y Electrónica \\ % Your institution for the title page
\medskip
\textit{mauricio@inaoep.mx} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\tableofcontents
\end{frame}

\section{Introduction and Problem Statement}
\begin{frame}
\begin{center}
\LARGE{Introduction}
\end{center}
\end{frame}
\subsection{Introduction}
\begin{frame}
\frametitle{Main idea}
How to learn and use causal information in order to make a good choice in an uncertain environment?
\end{frame}

\begin{frame}
\frametitle{Two components}
We notice two components from the main idea:
\begin{itemize}
\item Learn.
\item Use.
\end{itemize}
\end{frame}

\subsection{Relevance}
\begin{frame}
\frametitle{Why?}
\begin{itemize}
\item Decision making under uncertainty is fundamental for Artificial Intelligence (\cite{lake2017building}).
\item Optimal policies in Reinforcement Learning are the best action for each state.
\item Reinforcement Learning (RL) is \textit{not} a Causal Problem.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Why?}
\begin{itemize}
\item Mainstream ideas in DMU work at the \textit{associative}\footnote{probabilities, correlations...} level of information.
\item In fact, Optimal Policies in RL satisfy the classical criterion of Maximum Expected Utility.
\item Given the importance of Causation in science \textbf{we consider decision making theory that relies on causal information.}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{What do we know?}
\begin{itemize}
\item Rationality $+$ probabilistic information $=$ maximize expected utility.
\item Rationality $+$ \textbf{known} causal information $=$ maximize expected utility considering what is caused by actions (\cite{pearl2009causality}, \cite{lattimoreNIPS2016})
\item Rationality $+$ \textbf{unknown} causal information $=$ ¿? 
\end{itemize}
\end{frame}

\subsection{Objectives}
\begin{frame}
\frametitle{General Objective}
We recall the General Objective stated in \cite{gonzalez2019causal}:
\\
\textit{...to provide the understanding and knowledge about what are the implications of causality for a rational decision maker who faces an uncertain environment and about how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker...}
\end{frame}

\subsection{Research Problem}
\begin{frame}
\frametitle{Research Problem}
We recall here the Research Problem described in \cite{gonzalez2019causal}:
\begin{itemize}
\item Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$. We ask how a decision maker could \textit{learn} about the causal structure that controls his environment in order to \textit{make good choices} with respect to his preferences.
\end{itemize}
\end{frame}

\section{First year}
\begin{frame}
\begin{center}
\LARGE{First Year}
\end{center}
\end{frame}

\begin{frame}
\frametitle{First Year}
During the 2017-2018 academic year:
\begin{itemize}
\item The general objective and research problem to be studied were defined, as well as a mathematical problem which we called a Causal Decision Problem.
\item We proposed a preliminar solution which allowed for an optimal action to be found by a decision maker under some assumptions:
\begin{equation}{\label{first_version}}
P(Y=1 | do(a^\ast)) \geq P(Y=1 | do(a)) \textrm{ for all } a \in \mathcal{A},
\end{equation}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{First Year}
\begin{itemize}
\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility \citep{gonzalez2018playing}.
\item Poster presented at Causality Workshop ICML 2018.
\item Proposal Defense (27-Nov-2018)
\end{itemize}
\end{frame}

\section{Second Year}
\begin{frame}
\begin{center}
\LARGE{Second Year}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Second Year}
During the 2018-2019 academic year:
\begin{itemize}
\item The solution criterion given in Equation \ref{first_version} for causal decision making was further developed and extended for the case when a decision maker does not know the causal model which controls his environment:
\begin{equation}{\label{second_version}}
a \succeq b \textrm{ iff } \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right) \geq \sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right).
\end{equation}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Second Year}
\begin{itemize}
\item Using \ref{second_version} we were able to define the notion of a Causal Game and the idea of a Nash equilibria within such game. We define a function $u^C_i : A \to \mathbb{R}^{+}$, for each player:
\begin{equation}{\label{games_causal_ut}}
u^C_i (a) = \sum_{f \in A}  u_i(a) \left( \sum_g P^g_i (h_i(f) | do(a_i), a_{-i}) P_C(g) \right) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
\item We define a Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}{\label{nash_eq}}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
 \item We argued that Reinforcement Learning is not a Causal Problem in itself in \cite{gonzalez2019reinforcement}.
 \item Technical report CCC-19-002 which we cite here as \cite{gonzalez2019causal}. 
\end{itemize}
\end{frame}

\section{Third Year}
\begin{frame}
\begin{center}
\LARGE{Third Year}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Third Year}
During the 2019-2020 academic year:
\begin{itemize}
\item We further refined the concept of a Causal Game and that of a Nash Equilibrium for Causal Games.
\item We also studied causal-structure learning via interaction with a causally controled environment.
\end{itemize}
\end{frame}

\subsection{Causal Games}
\begin{frame}
\frametitle{Causal Games: New work}
We refined the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:
\begin{itemize}
\item For each player $i \in N$ in the strategic game, we define the following probability distribution over consequences:
\begin{equation}{\label{causal_utility}}
p^a_i (c) =  p^\omega_i (c | do(a_i), a_{-i}) p_i(\omega)\textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
\item We now define:
\begin{equation}
u^C_i (a) = \sum_{c \in C}  u_i(c) p^a_i (c) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
\item A Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
\item This work, \cite{gonzalez2019games}, was presented at a MICAI 2019 Workshop.
\end{itemize}
\end{frame}

\subsection{Causal-Structure Learning: A Random Graph Approach}{\label{random_graph}}
\begin{frame}
\frametitle{Causal-Structure Learning: A Random Graph Approach}
\begin{itemize}
\item Human beings focus on \textit{local} aspects while learning causal relations which are later unified into a single structure \citep{fernbach2009causal,waldmann2008causal, danks2014unifying}.
\item Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning.
\item Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Main Idea}
\begin{itemize}
\item Let a rational agent consider the following set of variables $\mathcal{X}=\{ X_1,...,X_n \}$ which are causally related, even though the agent does not now how.
\item the agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$.
\item Let $p_{ij}$ be the belief that the agent has over a causal relation (directed link) existing between variable with index $i$ and variable with index $j$.
\item Let $G$ an initial {\em random} DAG formed as follows: node set is $N=\{1,...,n\}$ and a link exists between $i$ and $j$ with probability $p_{ij}$.
\item Now, we use the methodology found in \cite{gonzalez2018playing}  as well as our results presented in \cite{gonzalez2019theorems} in order to find the best action $a^\ast$ for the obtained graph $G$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Main Idea}
\begin{itemize}
\item The best action is taken, and a full realization $X_1=x_1,...,X_n=x_n$ is observed.
\item Now, we update the $p_{ij}$'s using Bayes Theorem as follows:
\item For each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$
\item So we have:
\begin{equation}{\label{bayesian_updating}}
p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
\end{equation}
\end{itemize}
\end{frame}

\subsection{Implementation}
\begin{frame}
\frametitle{Implementation}
\begin{itemize}
\item We are using the same setting as in the experiments described in Year 1; this is, a patient sick with an unknown disease and a doctor who must choose between two possible treatments, each one with some risk of causing a life-threatening allergic reaction. Our agent (the doctor) must now learn which variables (disease, treatment, reaction, dying) are causally related.
\end{itemize}
\begin{enumerate}
\item Initialize $p_{ij}$ randomly
\item Initialize graph $G$ with link from node $i$ to node $j$ with probability $p_{ij}$
\item Find optimal action $a^\ast$ for graph $G$ and probabilities given by a count of the observations.
\item Calculate probability of the approximate model $G$ producing the full observation $X_1 = a^\ast, X_2=x_2,...,X_n=x_n$. 
\item Update $p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.$
\end{enumerate}
\end{frame}

\subsection{Results}
\begin{frame}
\frametitle{Ground Truth}
We know that the true causal model is a causal graph in which it holds that:
\begin{itemize}
\item Disease causes Final
\item Treatment causes Final
\item Treatment causes Reaction
\item Reaction causes Final
\end{itemize}
\end{frame}

\begin{frame}
\includegraphics[width=0.7\textwidth]{/Users/MauricioGS1/INAOE/3rd_year_eval/Figures/causal_model_espanol.png}
\end{frame}

\begin{frame}
\frametitle{Results}
\includegraphics[scale=0.5]{/Users/MauricioGS1/INAOE/3rd_year_eval/Figures/figura.png}
\end{frame}

\begin{frame}
\frametitle{What are we seeing?}
\begin{itemize}
\item Our method correctly learned three edges.
\item It did not learned a possible wrong edge. 
\item Variance needs to be reduced. 
\end{itemize}
\end{frame}

\section{Publications}
\begin{frame}
\begin{center}
\LARGE{Publications}
\end{center}
\end{frame}

\subsection{First Year}
\begin{frame}
\frametitle{First Year}
\begin{itemize}
\item The learning procedure described in Section \ref{first_year} was presented at the CausalML Workshop during ICML 2018 in Stockholm, Sweden. Link to Pre-print: \url{https://arxiv.org/abs/1807.01268}
\end{itemize}
\end{frame}

\subsection{Second Year}
\begin{frame}
\frametitle{Second Year}
\begin{itemize}
\item A Technical Report containing the overall details of the research in progress. Cited as \cite{gonzalez2019causal}. \url{https://ccc.inaoep.mx/archivos/CCC-19-002.pdf}
\item A paper in which via an algebraic argument we state that Reinforcement Learning, in its current mathematical definition, can not be considered as a causal problem. Co-authored with F. Orihuela-Espina. Implied by this result is the need for a Causal version of learning-by-interaction. Cited as \cite{gonzalez2019reinforcement}
\end{itemize}
\end{frame}

\subsection{Third Year}
\begin{frame}
\frametitle{Third Year}
\begin{itemize}
\item A paper including the first version of the Causal Nash Equilibrium was presented at a MICAI 2019 Workshop  \url{https://ccc.inaoep.mx/~bio/2019_MICAI_CausalityWorkshop/}. Cited as \cite{gonzalez2019games}
\item A paper including the decision-making results was orally presented at a NeurIPS 2019 Workshop  \url{https://nehzux.github.io/NewInML2019/}. Cited as \cite{gonzalez2019theorems}
\item Journal paper submitted to Theory and Decision  \url{https://www.springer.com/journal/11238/}
\end{itemize}
\end{frame}

\section{Future Work}
\begin{frame}
\begin{center}
\LARGE{Future Work}
\end{center}
\end{frame}

\subsection{Learning graphical structure}
\begin{frame}
\frametitle{Learning graphical structure}
\begin{itemize}
\item More complex causal model.
\item Intervening more than one variable.
\item Limitations on the learning proposal.
\item OpenAI Gym scenario.
\item Theoretical guarantees on our structure-learning proposal.
\end{itemize} 
\end{frame}

\subsection{Bayesian Non parametrics}
\begin{frame}
\frametitle{Bayesian Nonparametrics}
Bayesian non parametrics is a field within Bayesian statistics in which priors of infinite dimension are allowed, so models with a greater expressibility can be developed.
\begin{itemize}
\item Priors over complex graphs.
\item Fully Bayesian learning procedure.
\end{itemize}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{/Users/MauricioGS1/INAOE/Propuesta/Bibliografia.bib}
\end{frame}
\end{document}

### 2020-08-29 Theory_and_decision/Draft/Theory_and_decision_draft.tex
draft

\RequirePackage{fix-cm}

\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
%
%
\usepackage{mathptmx}      
\usepackage[utf8]{inputenc}
\usepackage[pdftex]{color,graphicx}
\usepackage{listings}
\usepackage{breakcites}
\usepackage{dsfont}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amssymb,amsmath,latexsym}
\usepackage[margin=1.5cm]{geometry}
\usepackage{enumitem}
%\bibliographystyle{stylename}
%\usepackage{fancyhdr}
%\pagestyle{fancy}
\usepackage{tikz}
\usetikzlibrary{trees}
\usetikzlibrary{calc}
\usepackage{color}
\usepackage{todonotes}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making}

%\titlerunning{Short form of title}        % if too long for running head
%
%\author{Mauricio Gonzalez-Soto         \and
%        Luis E. Sucar \and
%        Hugo J. Escalante
%}
\author{Anonymous}
%\authorrunning{Short form of author list} % if too long for running head

%\institute{Mauricio Gonzalez-Soto \at
%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
%              Mexico\\
%              \email{mauricio@inaoep.mx}           \\
%              \url{https://orcid.org/0000-0003-2668-9013}        
%           \and
%           Luis E. Sucar \at
%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
%              M\'exico\\
%              \email{esucar@inaoep.mx}\\
%              \url{https://orcid.org/0000-0002-3685-5567}
%           \and
%           Hugo J. Escalante \at
%           Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
%              M\'exico\\
%              \email{hugojair@inaoep.mx}\\
%              \url{https://orcid.org/0000-0003-4603-3513}
%}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle
% modificar abstract para reflejar cambios en el paper
\begin{abstract} 
Decision making under uncertainty has been well studied when information is considered at the associative (probabilistic) level. The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rational choice using purely associative information. In this work, we consider decision problems in which available actions and consequences are causally connected. We define a Causal Decision Problem and state a previous result from J. Pearl, showing that it can be considered as a causal version of the classical von Neumann-Morgenstern Theorem. Furthermore, we consider the case when the causal mechanism that controls the environment is unknown to the decision maker, and propose and prove a causal version of Savage's Theorem. Then, we describe two applications for which these theorems provide theoretical foundations: causal games and optimal action learning in causal environments. These results highlight the importance of causal models in decision making and the variety of potential applications.

\keywords{Causality \and Decision Making \and Game Theory}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}
\tableofcontents
\newpage
\section{Introduction}
\label{intro}
Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}. Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{waldmann2013causal,danks2014unifying,neil2019causality}. It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}. 
Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{tversky1977causal, tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.

Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent, located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward. Optimal policies, which are prescriptions of which action must be taken at some state $s$, satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem. Such solution is equivalent to maximizing \textit{expected utility} \citep{webb2007game}, which is the decision criterion prescribed by the most commonly used decision-making theories: von Neumann-Morgenstern's and Savage's. Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations.

On the other hand, an important aspect of acting in the world is being able to make decisions under uncertain conditions \citep{bernardo2000bayesian,kahneman2000choices,gilboa2009decision,lake2017building}. \cite{von1944theory} gave an answer for how to make choices if rational preferences are assumed and if the decision maker knows the stochastic relation between actions and outcomes: maximize expected utility. If no such relation is known, then \cite{savage1954the} showed that a rational decision maker must choose \textit{as if} she is maximizing expected utility using a \textit{subjective} probability distribution. Such theorems provide formal criteria for decision making if rationality is assumed.

These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms learn \textit{optimal policies} in such a way that any action prescribed by such optimal policy achieves the maximum expected utility \citep{sutton1998reinforcement,webb2007game}. In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on associative relations which are expressed using correlations or probability distributions.

It is a natural question how to formalize rational decision making when causal information is present; answering such question is relevant given the importance of causal relations in everyday life. Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage.

The previous question has been already considered by \cite{nozick1969newcomb}, \cite{lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making and also by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known by the decision maker.

\subsection{Motivation}
Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent. In many real-world applications, the agent doesn't have access to all of the parameters required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover this relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.

Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data. For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, they cannot explain \textit{why} a specific trajectory was chosen by the algorithm. This is highly relevant in real-world applications, e.g. self-driving cars, where it is very important to understand why an accident happened.

%For example, as told by \cite{bornstein2016artificial}, at the University of Pittsburgh Medical Center a team of researchers tried to use Machine Learning to predict whether pneumonia patients might develop severe complications. For this purpose they trained Neural Networks and Decision Trees using the hospital's own data. Neural Networks outperformed Decision Trees, but only by studying the decisions made by the latter the doctors did find out that the algorithms instructed doctors to send home pneumonia patients who already had asthma, despite the fact that asthma sufferers are known to be extremely vulnerable to complications. The problem relied in the training data, because the hospital policy was to automatically send asthma sufferers with pneumonia to intensive care, and this policy worked so well that asthma sufferers almost never developed severe complications. It was only through the interpretability of the Decision Trees that the doctors didn't send asthma patients with pneumonia home to a certain death.

%Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo \citep{silver2017mastering} defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, although it was a very sophisticated method for achieving it. Using only associations between variables it is not possible to infer which one is the cause and which one the effect, something extra is required. The fundamental difference between causal and associative models is that causal models consider \textit{doing} over \textit{observing} \citep{pearl2018why}. Once a causal relation has been established between two events, performing, or intervening, over a cause allows one to predict the outcome in a more robust way than if using only correlations between observations.

Learning a causal model of an environment and using it to act upon the environment allows us to \textit{explain} aspects of the model that a purely associative model would not be able to explain. It allows one to ask \textit{why}. Since a causal model, once it is learned, does not depend neither on the agent nor its preferences, the gained knowledge could be easily transferred for problems of similar domain where the use of the acquired causal relations is useful.  Our work provides the foundations of decision making algorithms which are based on causal relations and, as will be shown, are able to learn causal models of the environment.

%The impact of Causal Inference goes beyond Machine Learning and Computer Science. For example, economists are interested in understanding what did certain public policy caused \citep{athey2017impact} or how human decision makers that show \textit{inconsistent} preferences over time could be oriented if the causal consequences of this inconsistent behavior is introduced into a decision-making model \citep{albers2016motivating}. In complex adaptive systems, causal relations can be used to clarify the complex interactions between agents in the system \citep{abbott2017complex} as well as for prediction and planning \citep{hunt2016ants,brock2018causality}.

%cerrar esta sección sobre cómo mi trabajo impacta en la motivación

\subsection{Aim}
Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying}. We provide here an implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl. With our result we expect to highlight the importance and applicability of causal relations in decision making and the wide variety of possible applications.

\subsection{Main Result}
In this work we extend Pearl's result to the case where a decision maker \textit{does not know the causal mechanism} that controls her environment, so she holds \textit{beliefs} about possible causal models and uses such beliefs \textit{as if} they were true, as prescribed by \cite{joyce1999foundations}, in order to attempt to make a good choice given her beliefs. We provide an explicit way of choosing an optimal decision given what the decision maker believes about the causal structure of her environment, thus providing a causal version of Savage's Theorem. We take a normative point of view to study how causal relations should be used by a \textit{rational} agent when making decisions with uncertain consequences.

In Section \ref{on_causality} we define Causality and state the differences between causal and associative information; having shown this, we state two of the most important decision making results, the von Neumann-Morgenstern and Savage Theorems, which rely on purely associative information; then, our objective is to present two analogous theorems: J. Pearl's result, Theorem \ref{causal_ut} which relies on a known causal model, and our main contribution: Theorem \ref{causal_savage}, which is able to use a set of \textit{beliefs} about possible causal models (i.e., a distribution) instead of a known one.

\subsection{Applications}
Game Theory deals with situations in which several rational decision makers, or players, \textit{interact} while pursuing some well-defined objective \citep{osborne1994course}; the case in which decision makers make a choice simultaneously without knowing the choice made by the other players is called a \textit{strategic game}; a well-known strategic game is the famous \textit{prisoners' dilemma} in which two detainees must choose between confessing or remaining silent and both know the consequences of any combination of actions, what is ignored by each player is the decision made by the other. When players ignore both the actions made by other players as well as the knowledge that made them choose a certain action, is called a game with incomplete information, or a \textit{Bayesian Game} which was introduced by \cite{harsanyi1967games1,harsanyi1968games2,harsanyi1968games3}.

In this work we use the Bayesian Game model in order to study what happens when several decision makers have certain knowledge about an environment which is controlled by some, unknown but fixed, causal mechanism. We study games in which the players actions \textit{cause} some consequence according to some unknown causal model. We apply our a rational choice criterion to formulate a Nash equilibrium for Causal Games by defining a probability updating for Bayesian Causal Games.

Furthermore, we consider the problem of learning optimal actions in a causal environment. A decision maker is given the task of learning an optimal action in a causal environment in which a causal model controls the relation between actions and outcomes. We consider the case in which the causal model is unknown or partially known to the decision maker, so she has to use beliefs about possible causal models. From the decision maker point of view, having in mind a causal model allows one to ask what would have happened if another action was taken without actually performing the alternative action; therefore, being able of choosing the best action in terms of the probabilities of causing a good outcome.  Also, learning a causal model of a certain environment will facilitate knowledge transfer to similar domains, as the underlying cause-effect structure has been captured and could later be used.

% añadir motivación del paper y de las teorías

\subsection{Organization of the paper}
\label{organization}
This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which use only associative information in their formulation; then, in Section \ref{causal_decision_problems} a brief introduction to Causation and Causal Graphical Models is given. In  Section \ref{related_work}  we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning. In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure, as well as our main result, Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure. Finally two applications of Theorem \ref{causal_savage} are described. First, a causal analogue of Nash Equilibrium for games in which actions and consequences are causally connected is defined. Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works. Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider;  and in Section \ref{conclusion} we summarize the main contributions of our work.

\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
\label{classical_theorems}
A decision problem under uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors. Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

The most well-known theories for Decision Making are those from \cite{von1944theory} and \cite{savage1954the}. In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences, which is also known as \textit{decision under risk}, and in that case the theory guarantees that the decision maker behaves \textit{as if} she maximizes the expected value of a utility function. If the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if she has in mind a \textit{subjective} probability distribution and a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and the utility function, this is known in the literature as \textit{decision under uncertainty}.

Other Decision Making theories exist, such as Kahneman and Tversky's Prospect Theory \citep{kahneman1979prospect}, Gilboa's Case-Based Decision Theory \citep{gilboa1995case}, among others that are out of the scope of this work. In Section \ref{joyce_theory} we make reference to another theory, Joyce's Causal Decision Making \citep{joyce1999foundations}.
%We will later talk about another theory: Joyce's Causal Decision Making \citep{joyce1999foundations}. 
For further details on classical (non-causal) decision making, see \cite{bernardo2000bayesian} and \cite{gilboa2009decision}.

\subsection{Von Neumann-Morgenstern Theorem}
\label{vNM-M}
\cite{von1944theory} had the objective of justifying strategies in which players in a game maximized their expected utility. This theorem considers a scenario of \textit{decision under risk} and rational preferences; this is, choosing between uncertain outcomes with known probabilities.

Formally, using the notation from \cite{gilboa2009decision}, we consider a set $X$ of alternatives. Let $L$ be the set of lotteries with finite support over $X$. The object of choice are the elements $l \in L$ that are known to the decision maker; we represent the decision maker's preferences by a preference relation $\succeq \subseteq L \times L$ that  satisfies being complete, transitive, continuous and a technical condition called \textit{independence}. This family of conditions is called \textit{von Neumann-Morgenstern rationality axioms} as described in \cite{gilboa2009decision}.
\begin{theorem}[von Neumann-Morgenstern]{\label{vNM}}
A preference relation $\succeq \subseteq L \times L$ where $L$ is a set of lotteries with finite support over a set $X$ satisfies the von Neumann-Morgenstern rationality axioms if and only if there exists a function $u: X \to \mathbb{R}$ such that for every $P, Q \in L$ we have that
\begin{equation}
P \succeq Q \textrm{ if and only if } \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x). 
\end{equation}
\end{theorem}
The theorem states that if a rational decision maker knows the probabilities of obtaining a certain outcome, then she must choose \textit{as if} maximizing the expected value of some function $u$ whose existence is guaranteed by Theorem \ref{vNM}. See \cite{gilboa2009decision} for details on the proof.

\subsection{Savage's Theorem}
\label{savage_theorem}
If a rational decision maker does not know the probabilities of obtaining certain outcomes and does not have a precise, objective quantification of her preferences (utility function), then it is Savage's Theorem \citep{savage1954the}, which gives a formal choosing criterion. Savage's result extends von Neumann-Morgenstern Theorem since it considers the case in which a rational decision maker does not know neither her utility function nor the probabilities to be used in order to obtain the expected values required for making choices according to the von Neumann-Morgenstern Theorem. The notation and setting we use here is the one used by \cite{bernardo2000bayesian}. Consider a set $\Omega$. Consider a countable set $\mathcal{A}$ of available actions; for each action $a_i \in \mathcal{A}$, a partition $E_i$ of $\Omega$ and a set of consequences $C_i$. Let $\mathcal{E}$ the union of every $E_i$, we assume $\mathcal{E}$ to be a $\sigma$-algebra of events, and $\mathcal{C}$ the union of all $C_i$.

\begin{definition}
An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$. Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$. 
\end{definition}

When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty, \cite{bernardo2000bayesian}:

\begin{definition}
A Decision Problem under Uncertainty is an uncertain environment $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ plus a preference relation $\succeq$ defined over $\mathcal{C}$ 
\end{definition}

\begin{definition}
A Decision Problem under Uncertainty is said to be bounded if there exists a pair of consequences $c_\ast$ and $c^\ast$ such that for every $c \in \mathcal{C}$, $c^{\ast} \succeq c \succeq c_\ast$
\end{definition}

\begin{definition}
A Decision Problem under Uncertainty is said to be finite if the set $\mathcal{A}$ of available actions is finite.
\end{definition}

\cite{bernardo2000bayesian} derive the existence of a subjective probability measure from a set of \textit{rationality axioms}, where a decision maker has some mechanism of quantifying uncertainty in terms of real numbers within the $[0,1]$ interval, and then the classical Kolmogorov axioms are derived from such construction, and therefore all of the known machinery of Probability Theory. We now state Savage's Theorem. Details can be found in \citep{kreps1988choice,bernardo2000bayesian,gilboa2009decision}.

\begin{theorem}[Savage]{\label{savage}}
In a finite, bounded Decision Problem under Uncertainty $(\mathcal{A}, \mathcal{C}, \mathcal{E}, \succeq)$, the preference relation $\succeq$ satisfies the Savage rationality axioms if and only if there exists: 
\begin{itemize}
\item A \textit{probability measure} $P$, called a \textit{subjective probability}, that associates with each uncertain event $E \in \mathcal{E}$ a real number $P(E)$. 
\item A utility function $u : \mathcal{C} \to \mathbb{R}$ such that it associates each consequence with a real number $u(c)$. 
\end{itemize}
Such that for $a_1$ and $a_2$ actions in $\mathcal{A}$,
\begin{eqnarray*}
 &a_1 \succeq a_2&\\
 & \textrm{ if and only if }&\\
 &\mathbb{E}_P[u(a_1)] \geq  \mathbb{E}_P[u(a_2)],&
\end{eqnarray*}
where $\mathbb{E}_P[u(a)]$ is given by $\sum_{j \in J} u(c_j) P(E_j)$ since any action $a$ can be identified as $\{ c_j | E_j ; j \in J \}.$ Details in Section 2.5.2 from \cite{bernardo2000bayesian}.
\end{theorem}
\begin{proof}
Proposition 2.22 in \cite{bernardo2000bayesian}.
\end{proof}
This theorem states that if a rational decision maker does not know the precise probabilities of outcomes given that an action is taken, then she chooses \textit{as if} having in mind a probability assignment to the uncertainties in her environment and using such probabilities to calculate the expected utility with respect to a subjective utility function that represents her preferences. This result also gives a precise definition of \textit{subjective probability} as a quantification of uncertainty which is used to make good decisions. See \cite{hens1992note} and \cite{gilboa2009decision} for further details. See \cite{ellsberg1961risk} and \cite{binmore2008rational} for critiques of the Savage axioms.

\subsection{Limitations of the Theorems}
\label{theorems_limitations}
The mentioned theorems have the limitations of relying only on associative information; this is, on correlations and patterns found in data rather than strong causal relations. For example, as mentioned by \cite{pearl2009causality}, using only associative information one would be tempted to stop going to the doctor in order to stop being diagnosed as sick. On the other hand, our results provide an easy to implement way of making choices using causal relations and beliefs about such relations as used by humans. Human beings \textit{know} (this is, use causal relations) that going to the physician does not \textit{cause} being sick and this is what Pearl calls \textit{commonsensical} decision making. Also, as we will mention later in the paper, associative information lacks an \textit{ordering} thus failing to determine \textit{if $X$ is performed, then $Y$ will be caused}.

Formally, in this paper we consider the actions chosen by a decision maker as an \textit{interventions} over a causal graphical model; this is, in the classical von Neumann-Morgenstern and Savage Theorems, once a decision maker chooses an action $a$, the observed consequence has a probability $P(c|a)$; if causal information is considered, then the probability of a certain consequence is $P(c|do(a))$. In the next section we state some basic facts on causality and why $P(\cdot |x)$ is different from $P(\cdot | do(x))$ for some value $x$.

\section{Causal Decision Problems}
\label{causal_decision_problems}
\subsection{On causality}
\label{on_causality}
The concept of Causality deals with regularities found in a given environment which are stronger than probabilistic (or associative) relations in the sense that a causal relation allows for evaluating a change in the \textit{consequence} given that a change in the \textit{cause} is performed. The \textit{manipulationist} interpretation of Causality (details in \cite{woodward2005making}) is adopted here. The main paradigm is clearly expressed by \cite{campbell1979quasi} as \textit{manipulation of a cause will result in a manipulation of the effect}. Consider the following example from \cite{woodward2005making}: manually forcing a barometer to go down won't cause a storm, whereas the occurrence of a storm will cause the barometer to go down.

We know from \cite{granger1969investigating} and \cite{lamport1978time} that there are two sufficient and necessary conditions for causality: a context and an order. We will adopt here the formal definition of Causality given by \cite{spirtes2000causation}; i.e., a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}. Similar descriptions of the manipulationist approach can be found in \cite{holland1986statistics}. Causal inference tools, such as Pearl's do-calculus \citep{pearl2009causality} allow to find the effect of an intervention in terms of probabilistic information when certain conditions are met. For what remains, we assume the \textit{causal axioms} found in \citep{spirtes2000causation} as well as the condition known as \textit{causal sufficiency} \citep{pearl2009causality}.

\subsection{A definition of Causality}
The formal definition of Causality that will be adopted in this work is the definition provided by \cite{spirtes2000causation} which states that causation is a \textit{stochastic} relation between \textit{events} in a probability space; this is, some event (or events) \textit{causes} another event to occurr.
\begin{definition}{\label{causal_relation}}
Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite probability space, and consider a binary relation $\to \subseteq \mathcal{F} \times \mathcal{F}$ which is:
\begin{itemize}
\item Transitive: If $A \to B$ and $B \to C$ for any $A, B, C \in \mathcal{F}$ then $A \to C$.
\item Irreflexive: For all $A \in \mathcal{F}$ it doesn't hold that $A \to A$.
\item Antisymmetric: For $A,B \in \mathcal{F}$ such that $A \neq B$ if $A \to B$ then it doesn't hold that $B \to A$.
\end{itemize}
We say that $A$ is \textit{a cause} of $B$ (or that $A$ causes $B$, $A$ is the cause and $B$ is the effect) if $A \to B$. It is important to note that an event may have more than one cause, and that not necessarily each one of this causes is sufficient to produce the effect. 
\end{definition}

We assume that for any event $A \in \mathcal{F}$ there are no causes lying outside $\mathcal{F}$ (causal sufficiency). It is also be assumed that the relations expressed by $\to$ are the only causal relations in the environment. This assumption can be interpreted as not allowing \textit{intermediate} events between events known to be causally related. For example, if in our space we have that striking a match causes fire, we do not allow into consideration the underlying chemical reactions that cause fire from friction. In this sense, we say that striking a match is a \textit{direct cause} of fire and this are the only causes considered herein \citep{spirtes2000causation}.

\subsection{Representation into a Directed Acyclic Graph}
The causal relations contained in $\to$ can be summarized in a graph $G=(V,E)$ in the following way: If $A \to B$ then the graph must contain a node $A \in V$ representing $A$, a node  $B \in V$ representing $B$ and a directed edge $e \in E$ connecting the respective nodes in the direction of the causal relation.

\begin{proposition}
Given a causal relation $\to$ as in Definition \ref{causal_relation} then the graph that is obtained by considering nodes for events and edges for the causal relations as previously described is a Directed Acyclic Graph.
\end{proposition}

\begin{proof}
The graph is directed since the definition of causality imposes a direction between events; namely, a direction between the cause and the effect. To see that the graph is acyclic, suppose that a cycle $A \to B \to C \to A$ exists, this would imply, because of transitivity, that $A \to A$, which cannot be since the relation is irreflexive.
\end{proof}

Notice that since the graph is finite, there exist some nodes that do not have causes, which are called \textit{exogenous}. If an event $A$ is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$. It is proven in \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.

\subsubsection{Relations between a causal graph and probabilities}
Given a Directed Acyclic Graph (DAG) it is possible to obtain a probability measure that expresses the conditional independence relations that are expressed in the graph \citep{koller2009probabilistic}, which we will call $P_\mathcal{G}$. For the DAG built from the causal relations, we require that its correspondent $P_\mathcal{G}$ satisfies the following conditions \citep{spirtes2000causation}:
\begin{itemize}
\item Markov Causality: an event $V$ (or node in $\mathcal{G}$) is independent of every other event $A$ such that $A$ isn't either a cause nor an effect of $V$ given the causes of $V$.
\item Causal Minimality: No proper sub-graph of $\mathcal{G}$ satisfies the Markov Causality condition.
\item Causal Faithfulness: The Markov Causal condition contains all of the conditional independence statements expressed by the DAG $\mathcal{G}$.
\end{itemize}

Also, we consider an extra pair of conditions. The first one, known as Causal Sufficiency, is about the \textit{nature} of the model: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ \textit{outside} of the model $\mathcal{G}$\citep{spirtes2000causation,pearl2009causality,sucar2015probabilistic}. The second required condition is that the causal mechanism remains unchanged after interventions, this is called in \citep{woodward2005making} as \textit{invariantness}, this basically means that an intervention does not \textit{break} the causal mechanism when intervening upon it. These conditions are required in an axiomatic fashion so they will be taken as they are without questioning.

\subsection{Causal Graphical Models}
A Causal Graphical Model (CGM) consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect. Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ... , x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ \citep{pearl2009causality,koller2009probabilistic,sucar2015probabilistic}. To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable. It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ \citep{pearl2018why}. In fact, using the \textit{Do} operator we give a definition of a variable $X$ causing $Y$.
\begin{definition}
Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
\[ P_\mathcal{G}(Y | do(X=x) \neq P_\mathcal{G}(Y | do(X=x'). \]
\end{definition}
It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph \citep{sucar2015probabilistic}.
\subsubsection{The identifiability problem}
Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data? It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) \citep{peters2012identifiability,mooij2016distinguishing}.

\subsubsection{Do-Calculus}
The Do-Calculus is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data \citep{pearl1995causal,pearl2009causality}.

It is important to introduce some notation. Consider a causal graphical model $\mathcal{G}$ and $X,Y,Z$ disjoint sets of nodes of $\mathcal{G}$. We denote by $\mathcal{G}_{\bar{X}}$ the graph that is obtained by deleting from $\mathcal{G}$ all of the edges that enter nodes in $X$. In the same way, $\mathcal{G}_{\underline{X}}$ is the graph obtained by deleting the edges that emerge from $X$. Finally, $\mathcal{G}_{\underline{Z}\bar{X}}$ is the graph obtained by deleting edges incoming into $X$ and outgoing from $Z$.

\begin{theorem}[Do-Calculus, \cite{pearl2009causality}]{\label{docalculus}}
Let $\mathcal{G}$ a CGM and $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
\begin{itemize}
\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ and $W$, then
\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ and $W$, then
\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
\item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.  If $Y$ is conditionally independent from $Z$ given $X$ and $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
\end{itemize}
\end{theorem}
\begin{proof}
See \cite{pearl2009causality}.
\end{proof}
\begin{theorem}[Properties of Do-calculus]{\cite{peters2017elements}}\\
The following statements hold:
\begin{itemize}
\item The Do-Calculus is complete; this is, sufficient for deriving every identifiable interventional distribution \citep{huang2006pearl,shpitser2006identification}.
\item There exists an algorithm capable of finding all of the identifiable interventions \citep{tian2002,huang2006pearl}.
\item A necessary and sufficient criteria exists for the identifiability of interventional distributions \citep{shpitser2006identification,huang2006pearl}.
\end{itemize}
\end{theorem}
\begin{proof}
See \cite{peters2017elements}.
\end{proof}
The Do-Calculus rules provide a solution for the identifiability problem:
\begin{corollary} \cite{pearl2009causality}.\\
A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each one of them corresponds to any of the rules of the Do-Calculus, that reduce $q$ to a purely observational expression.
\end{corollary}

\subsection{Further limitations of associative information}
As we have said, one of the fundamental aspects of Causality is the presence of an order and a context; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
\begin{equation}
    p(a,t)=p(a|t)p(t)
\end{equation}
and
\begin{equation}
    p(a,t)=p(t|a)p(a)
\end{equation}
which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$. This situation does not happen while considering causal relations.

Also, we recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$. If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false.
\subsection{Causal Environments and Causal Decision Problems}{\label{causal_problems}}
We consider decision making with causal information. We define a Causal Environment to be an \textit{uncertain environment} such that there exists a Causal Graphical Model (CGM) $\mathcal{G}$ \citep{koller2009probabilistic} which controls the environment; this is, there exists a causal relation between available actions and consequences in the sense that any chosen action will stochastically \textit{cause} a consequence. The role of the CGM is to encode all of the causal relations present in the environment, not only between actions and consequences but also between any other variables in the environment. 
\begin{definition}
A Causal Environment is a tuple $(\Omega, \mathcal{A},\mathcal{G},\mathcal{C},\mathcal{E})$ where $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ is an uncertain environment and $\mathcal{G}$ is a CGM such that the set of variables of $\mathcal{G}$ correspond to the uncertain events in $\mathcal{E}$.
\end{definition}
\begin{definition}
We define a Causal Decision Problem (CDP) as a tuple $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$ where $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C})$ is a Causal Environment and $\succeq$ is a preference relation. 
\end{definition}

For the CGM in a CDP we distinguish two particular variables: one corresponding to the available actions, and one corresponding to the produced (caused) outcome. We are considering that only one variable can be intervened upon and that the values of such variable represent the actions available to the decision maker; i.e., the value forced upon such variable under an intervention represents the action taken by the decision maker. The intuition behind the definition of a Causal Decision Problem is this: a decision maker chooses an action $a \in \mathcal{A}$, which is automatically inputed into the model $\mathcal{G}$, which outputs the \textit{causal outcome} $c \in \mathcal{C}$. We say a CDP is \textit{finite} if the set $\mathcal{A}$ is finite.

\section{Related Work}
\label{related_work}
\subsection{Causal Decision Making and Decision-Theoretic foundations of Causal Inference}
\label{joyce_theory}
A previous attempt to formalize Decision Theory in the presence of Causal Information was given by \cite{lewis1981causal,joyce1999foundations}. According to such formulation, a decision maker must choose whatever action is more likely to (causally) produce a desired outcome while keeping any beliefs about causal relations fixed. This is stated by the Stalnaker equation \citep{stalnaker1968})
\begin{equation}
u(a)=\sum_{x} P(a \square \to x)u(x),
\end{equation}
\noindent
where $a \square \to x$ is to be read as \textit{if the decision maker does} $a$ \textit{then} $x$ \textit{would be the case} \citep{gibbard1978counterfactuals,kleinberg2013causality}. Lewis' and Joyce's work captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker. In Section \ref{section_savage} we will refine the $\square \to$ operator by an explicit way of calculating the probability of causing an outcome by doing a certain action in terms of Pearl's do-calculus.

\cite{heckerman1995decision} provide a framework for defining the notions of cause and effect in terms of decision theoretical concepts, such as states and outcomes and gives a theoretical basis for graphical description of causes and effects, such as the causal influence diagrams from \citep{dawid2002influence}. Heckerman gave an elegant definition of causality, but did not address how to actually make choices using causal information.

\cite{dawid2012decision} presented a decision-theoretic approach to causal inference in which a decision maker must take into account how alternatives compare against the other in terms of the \textit{average causal effect}, such approach uses the well-known influence diagrams \citep{dawid2002influence,dawid2003causal} in order to derive formulas that allow an explicit calculation of the average causal effect. Influence diagrams have the ability of expressing both intervention variables and chance variables into a single graphical structure in such a way that the standard techniques for probabilistic DAG's still apply. An optimality criterion for sequential interventions is obtained by \cite{dawid2008identifying} by maximizing the expectation of outcomes.

\subsection{Causal Reinforcement Learning and Bandit Problems}
It is known that human beings conceive their actions on the world as \textit{intervening} in the world \citep{hagmayer2009decision}. Following this idea, \cite{lattimoreNIPS2016} consider decision problems in which the action to be chosen is an intervention over a known causal graphical model. The agent must choose the intervention that maximizes the value of a \textit{target variable} after a series of learning rounds. They model their problem by considering that choosing an intervention is choosing an arm of a \textit{slot machine}, usually known as Bandit Problems, in which a gambler chooses an arm and gets some reward. From the rewards they estimate probabilities and output an optimal action in the sense of obtaining minimal regret. Their work considers that the causal model is fully known, restriction which is partially relaxed by \cite{sen2017identifying}, and mention that the case in which the causal model is unknown is left as an open question, and this is precisely what we are proposing to answer.

In general, in a classic Bandit Problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \citep{sutton1998reinforcement}. Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure,audibert2010best,gabillon2012best,agarwal2014taming,jamieson2014lil,jamieson2014best,chen2015optimal,carpentier2016tight,russo2016simple,kaufmann2016complexity}, but none of these works consider causal-governed environments. \cite{ortega2014generalized} gives results on sequential decision-making using Generalized Thompson Sampling that could be extended into causal inference problems. The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices, but the problem of \textit{acquiring} this causal knowledge is left unattacked.

Bandit Problems are a particular case of a Reinforcement Learning problem, which as mentioned in the introduction, is a general framework for interactive learning in which an agent performs an action and receives some feedback from his environment. The classical setting for an agent to learn by interaction is that of a Markov Decision Process, in which at time $t$ an agent is situated in state $s_t$ and takes an action $a_t$, he then observes a new state $s_{t+1}$ and a reward signal $r_{t+1}$. The objective is to find the \textit{policy} which maximizes the long-run reward. It can be shown that this is equivalent to solving the Bellman Equations \citep{sutton1998reinforcement}, and that the actions prescribed by such optimal policy maximize the expected utility. This shows the connection between the modern field of Reinforcement Learning and classical results in Decision Theory which justify choosing actions that maximize expected utility when certain \textit{rationality} assumptions are made.

It has been recently studied the connection between Reinforcement Learning and Causation in a series of papers by \cite{gershman2015reinforcement}, \cite{lattimoreNIPS2016}, \cite{bareinboim2015bandits} among others such as \citep{zhang2016markov, zhang2017transfer, lee2019structural, correa2019statistical, zhang2019near}. Even though the connection between Reinforcement Learning and causal reasoning is intuitive, a proper  framework for Reinforcement Learning based on Causal Decision Making (see \cite{joyce1999foundations} for an introduction to Causal Decision Making and \cite{pearl2009causality} for a more recent computational version) has not been explored.

In this paper we present a causal decision making criterion useful when the probabilities that hold in an environment are not previously known to the decision maker, which is usually the case in Reinforcement Learning. Therefore, the findings of this paper can have an important impact into decision making because of both, the importance of causal relations in artificial intelligence and the realistic assumption of an agent not having full information of her environment.

\section{Main Result: Causal Decision Making}
\label{causal_results}
\subsection{A von Neumann-Morgenstern type theorem for causal environments}
Consider a rational decision maker who faces a causal environment in which she knows the causal model controlling the relation between actions and outcomes. She can use the known causal model in order to find the probabilities of \textit{causing} a desired outcome given she takes a certain action. The following theorem is found in Section 4.1 of \cite{pearl2009causality}, but the intuitions that lie behind can be traced back to \cite{lewis1981causal} and \cite{joyce1999foundations}.

Consider a Causal Model $G$ and its associated distribution $P_G$ and let $C$ be the set of consequences of interest for a decision maker. Then,
\begin{theorem}[Causal von Neumann-Morgenstern Theorem, \cite{pearl2009causality}]{\label{causal_ut}}
If a rational decision maker faces a Causal Environment and if the causal model is known, then the preference relation $\succeq$ satisfies the von Neumann-Morgenstern rationality axioms if and only if:
\begin{eqnarray*}
&a \succeq b&\\
& \textrm{ if and only if }&\\
 &\sum_{c \in C} P(c | do(a))u(c) \geq \sum_{c \in C} P(c | do(b))u(c).&
\end{eqnarray*} 
\end{theorem}
Equivalently, the action that must be chosen is 
\begin{equation}
a^\ast = \textrm{argmax}_{a \in \mathcal{A}} \sum_{c \in C} P(c | do(a))u(c).
\end{equation}
\begin{proof}
This Theorem assumes the existence of a Causal Graphical Model known by the decision maker; then, by Theorem \ref{vNM}, we know that the action that must be chosen is the one that maximizes the expected utility; we must now ask the expected utility of what? The answer lies on the Causal Graphical Model which controls the relation between actions and consequences; if this is the case, then the probability of a consequence given an action $a$ is in fact $P(c|do(a))$. The rest follows by Theorem \ref{vNM}. Details in Section 4.1 of \cite{pearl2009causality}.
\end{proof}

As shown in the proof of Theorem \ref{causal_ut}, we claim that Pearl's result can be considered as a causal version of the von Neumann-Morgenstern result since it assumes that a causal model is known. If the causal model that controls an environment is known to a decision maker, then this is equivalent of being able to know the probabilities of outcomes given actions. As stated in Section 4.1 of \cite{pearl2009causality}, the utility function $u$ is considered as given, even though the Theorem \ref{vNM} guarantees its existence. Pearl argues that decision making by maximizing such function is \textit{commonsensical} without appealing to the original results in rational decision making. We avoid entering in the long-standing debate between \textit{causal} and \textit{evidential} decision making. We only note that both von Neumann-Morgenstern's and Savage's Theorems have no {\em a priori} causal interpretation and therefore rely on associative information.

\subsection{A Savage Theorem for causal environments}
\label{section_savage}
When the decision maker does not know the causal model which controls her environment we argue that the decision maker is facing a particular case of Savage's Theorem. The difference here being that the subjective probability must make use of the causal nature of the environment. The idea we will follow is that Savage's Theorem gives the decision maker a subjective quantification of her uncertainty about the environment and the associated probabilities which must be used \textit{as if} they were true.

In this case, where a Causal Graphical Model controls the relation between actions and outcomes, such subjective information about the environment must consider such causal structures. For this reason, we assert that the probability distribution that the decision maker has in mind is in fact a distribution over causal structures where the decision maker uses each structure as if it were the true one in order to choose the best action within each structure by using Pearl's result. We assume a finite set of actions and outcomes. Formally:

\begin{theorem}[Causal Savage Theorem]
\label{causal_savage}
In a finite, bounded Causal Decision Problem  $(\mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$, where $\mathcal{G}$ is a Causal Graphical Model, we have that the preferences $\succeq$ of a decision maker are Savage-rational if and only if there exists a probability distribution $P_C$ over a family $\mathcal{F}$ of causal structures such that for $a,b \in \mathcal{A}$:
\begin{eqnarray*}
&a \succeq b&\\ 
&\textrm{ if and only if }&\\
&\sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right)&\\
&\geq& \\
&\sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right)&
\end{eqnarray*}
where $P_g$ is the probability distribution associated with the causal structure $g$. 
\end{theorem}

\begin{proof}
The decision maker is facing an environment in which any action she takes will stochastically cause an outcome $c \in \mathcal{C}$. For this reason, the decision making is facing a very particular case of decision making under uncertainty. Assuming Savage-rationality, by Theorem \ref{savage} we obtain a utility function $u^S$ and a probability measure $P^S$ which satisfy that the preference relation is represented by the expectation of $u^S$ with respect to $P^S$.

In such a causal environment, the CGM $\mathcal{G}$ contains all of the information which connects actions, uncertain events and outcomes, and noting that we can identify any action $a$ with $\{ c_j | E_j : j \in J  \}$ where $J$ a countable set of indexes (\cite{bernardo2000bayesian}) we have that:
\[\mathbb{E}_{P^S}[u(c)] = \sum_{j \in J} u(c_j)P^S(E_j).\]
For each action $a=\{ c_j | E_j : j \in J \}$, $P^S(E_j)$ is the probability of \textit{causing} consequence $c_j$ by choosing action $a$. In order for the decision maker to find the probability of a certain consequence $c_j$ given that an action $a$ is performed then she must have in mind a single causal model $g$ and a way to assign probabilities over a family of causal models; i.e., the uncertainty component $P^S(E_j)$ is formed by two parts: a distribution $P_C$ which represents the degree of belief of the decision maker about a specific model $g$ being the true one, and within $g$, a distribution $P_g$ used to calculate the probability of causing some consequence $c_j$ given that action $a$ is chosen. Using the Caratheodory Extension Theorem \citep{ash2000probability}, a probability measure  $P_C$ whose support is a sufficiently general family of causal models $\mathcal{F}$ can be shown to exist. For $g \in \mathcal{F}$, the decision maker considers $g$ to be the true causal model with probability $P_C(g)$ and will use Theorem \ref{causal_ut} in order to find the best action \textit{within} the causal graphical model $g$ as if it were the true one. Let $P_g$ the probability distribution associated with $g$.

Then:
\begin{eqnarray*}
\mathbb{E}_{P^S}[u(c)] &=& \sum_{j \in J} u(c_j)P^S(E_j)\\
                                      &=& \sum_{j \in J} u(c_j) \left( \sum_{g \in \mathcal{F}} P_g(c_j | do(a))P_C(g) \right).
\end{eqnarray*}
We have shown what is the expected utility for some action $a \in \mathcal{A}$, and again by Theorem \ref{savage} the result follows. 
\end{proof}

\subsection{Interpretation}
Theorem \ref{causal_savage} says that a decision maker who faces a Causal Decision Problem is considering a probability distribution $P_C$ over a family $\mathcal{F}$ and, within each structure, using the term $P_g(c|do(a))$ in order to find the probability of obtaining a certain consequence given that the intervention $do(a)$ is performed; in this way, the optimal action $a^\ast$ is given by:
\begin{equation}
a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}  \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right). 
\end{equation}
We note that $a^\ast$ is obtained by taking into account the utility obtained by every possible consequences weighted using both the probability of causing such action within a specific causal model $g$ and the probability that the decision maker assign to such $g \in \mathcal{F}$.

We are considering a \textit{normative} interpretation for Theorem \ref{causal_savage} according to which a decision maker must use any causal information in order to obtain the best possible action. Such action must be obtained by considering the \textit{beliefs} of the decision maker about the causal relations that hold in her environment (the distribution $P_C$), how such relations could produce the best action when considered \textit{as if} they were true (distribution $P_g$), and the satisfaction (utility $u$) produced by the consequences of actions.

Humans tend to ignore pure probabilistic information over causal information \citep{tversky1980causal}, and in fact are able to learn, and use, causal models in sequential decision making processes \citep{lagnado2007beyond,sloman2006causal,nichols2007decision,meder2010observing,hagmayer2013repeated,wellen2012learning}, even though such learning is not perfect \citep{rottman2014reasoning}. Therefore, this theorem provides the basis for a much stronger, and computationally implementable, framework for decision making in which causal information is preferred over associative information, even though complete causal information may not be available to the decision maker.

\section{Application: Causal Games and Nash Equilibrium}
\label{causal_games}
In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment. We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence. With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information. Therefore, provide motivation to further extend classical results, which have been shown useful, to use causal information as a basis.

A game is a model of a situation in which several players must take an action and afterwards they will be affected both by the outcome of their own action as well as the actions of the other players \citep{osborne1994course}. In a strategic game it is assumed that no player knows the action taken by any other players; we also assume that the causal mechanism, which represented by a Causal Graphical Model $\mathcal{G}$, remains fixed and it is unknown for each player.

In this game, players ignore the actions taken by any other player, and since the causal model which controls the environment is unknown by every player, the players also ignore the information that players will use in order to take their respective actions: strategic games of this type are called \textit{Bayesian Games}, introduced by Harsanyi \citep{harsanyi1967games1,harsanyi1968games2,harsanyi1968games3}. In the games we will consider, the uncertainty of every player consists of two levels: on a first level, the true causal model $\mathcal{G}$; on a second level, what an action $do(a)$ causes if a certain Causal Graphical Model $\omega$ is considered to be the causal model. 
\begin{definition}
A Bayesian strategic game(\cite{osborne1994course}), consists of:
\begin{itemize}
\item A finite set $N$ of players.
\item A finite set $\Omega$ of \textit{states of nature}.
\item For each player, a nonempty set $A_i$ of actions.
\item For each player, a finite set $T_i$ and a function $\tau_i : \Omega \mapsto T_i$ the signal function of the player
\item For each player, a probability measure $p_i$ over $\Omega$ such that $p_i (\tau^{-1}_i (t_i))>0$ for all $t_i \in T_i$.
\item A preference relation $\succeq_i$ defined over the set of probability measures over $A \times \Omega$ where $A= A_1 \times \cdots A_n$
\end{itemize}
\end{definition}

We will consider $\Omega$ to be a family of admissible causal models; in this way,  $\omega \in \Omega$ being the true state of nature fixes a causal model which controls the environment in which the players make their choices. In classical Bayesian games, once $\omega \in \Omega$ is realized as the true state, then each player receives a signal $t_i=\tau_i (\omega)$ and the posterior belief $p_i(\omega | \tau^{-1}_i (t_i) )$ given by $p_i(\omega) / p_i (\tau^{-1}_i (t_i))$ if $\omega \in \tau^{-1}_i (t_i)$. In the case for causal bayesian games, we must consider both the probability $p_i$ of $\omega$ being the true state as well as the probability $p^\omega_i$ of observing a certain consequence when doing some action $a_i$ if $\omega$ is the true model.

Following \cite{osborne1994course}, we define a new game $G^\ast$ in which its players are all of the possible combinations $(i, t_i) \in N \times T_i$, where the possible actions for $(i,T_i)$ is $A_i$. We see that fixing a player $i \in N$, the posterior probability $p(\omega | \tau^{-1}_i (t_i))$ induces a lottery over the pairs $(a^\ast(j,\tau_j(\omega)))_j,\omega)$ for some other $j \in N$. This lottery assigns to $(a^\ast(j,\tau_j(\omega)))_j,\omega)$ the probability $p_i(\omega) / p_i (\tau^{-1}_i (t_i))$ if $\omega \in \tau^{-1}_i (t_i)$. The classical Bayesian game will simply call a Nash equilibrium for the game $G^\ast$ a Nash equilibrium of the original game; but we have the second level of uncertainty: the consequences caused by some action $a$ through a causal model $\omega$. We notice that the posterior probability itself induces a probability distribution defined over \textit{actions} for each player once a \textit{desired consequence} is fixed, this distribution, according to Theorem \ref{causal_savage} is given by $p^\omega_i (c | do(a^\ast_i), a^\ast_{-i}) p_i(\omega | \tau^{-1}_i (t_i))$. This motivates the following definition of a \textit{Causal Nash equilibrium}.
\subsection{Causal Nash Equilibrium}
For each player $i \in N$ in the strategic game, we define the following probability distribution over consequences:
\begin{equation}{\label{causal_utility}}
p^a_i (c) =  p^\omega_i (c | do(a_i), a_{-i}) p_i(\omega)\textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
where $p^\omega_i$ is the probability of causing a certain consequence within a causal structure $\omega$ and $p_i$ are the player's \textit{posterior beliefs} about the causal structure that controls the environment, and $do()$ is the well known intervention operator from \cite{pearl2009causality}. We now define:
\begin{equation}
u^C_i (a) = \sum_{c \in C}  u_i(c) p^a_i (c) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
\end{equation}
Notice that $u^C_i$ evaluates an action profile $a \in A$ in terms of: The knowledge about the causal structure of each player represented by $p_i$, which allows each player to evaluate the probability of causing outcomes in terms of actions by using the $do$ operator as well as the other actions taken by the other players, given by $a_{-i}$ and the preferences of each player $u_i$. Using this new function, we define the equilibria for a strategic game with causal information and Bayesian players as:
\begin{definition}
A Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
\begin{equation}
 u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
 \end{equation}
\end{definition}
This is, an action profile is a Nash equilibrium if and only if each player uses her current knowledge about the causal structure of the environment in order to (causally) produce the best possible outcome given the actions taken by the other players. The existence of the Causal Nash Equilibrium is guaranteed if every $A_i$ is a nonempty compact convex set in some $\mathbb{R}^n$ and if the preference relation induced by $u^C_i$ is continuous and quasi-concave.

\section{Application: Formalizing an optimal action learning method for causal environments}
\label{optimal_action}
%lo que estoy planteando complementa estos trabajos que se hicieron de manera empirica
In recent papers by \cite{lattimoreNIPS2016,sen2017identifying} and \cite{gonzalez2018playing}, a decision maker is given the task to learn an optimal action from a series of learning rounds in a causal environment when the true causal model which controls the environment is unknown for the decision maker. \cite{lattimoreNIPS2016} assume that the conditional distributions for each variable given its parents is known, while \cite{sen2017identifying} assume as known only a part of the causal graphical model and allows for interventions in the unknown part of the model. \cite{gonzalez2018playing}  assume that the \textit{graphical structure} of such model is known. In their method, the decision maker holds beliefs about the parameters of the model, which are used in order to build a causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round.

\cite{gonzalez2018playing}, assume that the graphical structure of the causal model is known and use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action; the best action is defined in terms of the maximum expected utility as required by both the von Neumann-Morgenstern and Savage Theorems. Their method fits within our Theorem \ref{causal_savage} because in this paper the distribution $P_C(g)$ is a \textit{distribution over causal models with the same structure}, and when a model is specified from $P_C$, it is then used as if it were the true one in order to find the best action for such causal model. In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper.

Such learning method, which has been shown to fit in our causal decision making theorem achieves a similar performance than the classical Q-Learning algorithm with the extra of learning a causal model of a very simple environment. Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed. A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions. Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods cannot answer.

When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action. It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons. Also, learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used. We have provided the extension to the causal case of the formal criteria (expected utility) to be used when a causal model is present in the environment, and therefore we hope that our result will be of interest to anyone using decision making under uncertainty.

%Queda la duda de si se necesita especificar un orden previo
\subsection{A random graph approach to causal structure learning}
\label{random_graphs}

For the sake of completeness, we now provide a relaxation of the restriction found in \cite{gonzalez2018playing}; in such work, it is required that the graphical structure of the causal model is known, which is a strong assumption.

Human beings focus on \textit{local} aspects while learning causal relations which are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}). Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning. Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.

Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$. Our extension will make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.

We call our method \textit{local} since we use the uncertainty (from the agent's point of view) about the existence of a causal relation between pairs of variables, which are local components of the structure of a graph. Such uncertainty is to be updated in terms of what is \textit{observed} from realizations of the true causal mechanism that controls the environment. Since observations alone do not suffice to uniquely determine a DAG, we make use of a partial ordering between the variables in the model that, at least in principle, can not be a cause of each other. This ordering must be an input product from expert knowledge.

Let a rational agent consider the following set of variables $\mathcal{X}=\{ X_1,...,X_n \}$ which are causally related, even though the agent does not now how; the agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$. Let $p_{ij}$ be the belief that the agent has over a causal relation (directed link) existing between variable with index $i$ and variable with index $j$. Let $G$ an initial {\em random} DAG formed as follows: node set is $N=\{1,...,n\}$ and a link exists between $i$ and $j$ with probability $p_{ij}$. Now, we use the methodology found in \cite{gonzalez2018playing}  as well as our Theorem \ref{causal_savage} in order to find the best action $a^\ast$ for the obtained graph $G$. The best action is taken, and a full realization $X_1=x_1,...,X_n=x_n$ is observed.

Now, we update the $p_{ij}$'s using Bayes Theorem as follows: for each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$, so we have
\begin{equation}{\label{bayesian_updating}}
p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
\end{equation}
Then, we update the model generating a new graph according to $p_{ij}$. We do not provide an implementation of Equation \ref{bayesian_updating} as it is not in the scope of this paper.

\section{Limitations}
\label{limitations}
We are working within the classical rationality assumption. Rationality can be ultimately thought of as a \textit{consistent} or coherent way of making choices, but the precise definition has been a subject of debate. See \cite{ellsberg1961risk}, \cite{gilboa2009decision} and \cite{machina2014ambiguity} for critiques of the Savage Rationality Axioms. Another line of decision making, from a descriptive point of view, has been developed by psychologists and economists; see \cite{TverskyKahneman74,kahneman1979prospect,kahneman2000choices}.

We are considering Causal Graphical Models as the representation of causal relations in the environment, and we are considering stochastic causal relations according to the manipulationist interpretation, which is one of many. We have favoured the Causal Graphical Models over other alternatives since it has been argued that several cognitive processes, such as causal reasoning, can be best represented as graphical models \cite{danks2014unifying,sloman2015causality,hagmayer2016causal}.

\section{Conclusion}
\label{conclusion}
We have defined a Causal Decision Problem in terms of a classical Decision Problem under Uncertainty provided of a Causal Mechanism which controls the relation between actions and outcomes, which is represented by a Causal Graphical Model in this work. In the case in which a rational decision maker knows such causal relations, \cite{pearl2009causality} provides a causal version of the von Neumann-Morgenstern Theorem for decision making.

On the other hand, when a decision maker does not know the causal mechanism, in Theorem \ref{causal_savage} we have provided a causal version of the Savage Theorem; our result explicitly states how a rational decision maker should use subjective beliefs, encoded as a probability distribution over causal models, as well as the causal inference machinery within the considered causal structures in order to find an optimal action.

Using Theorem \ref{causal_savage} and taking as a basis Harsanyi's model of a Bayesian Game in which every player has incomplete information about both the actions taken by other players as well as the information that made each player take his action we have been able to provide a definition of a Causal Nash Equilibrium in which every player is aware that there exists a Causal Mechanism that will produce some consequence once he takes an action. Our decision making result (i.e., Theorem \ref{causal_savage}), besides motivating the Causal Nash Equilibrium, also provides an optimality criterion for learning algorithms in causal settings such as those presented in~\cite{lattimoreNIPS2016,sen2017identifying}. Our definition of Causal Equilibrium takes into account classical game theory through the incorporation of the classical von Neumann-Morgenstern utility function as well as the fundamental notion in Causation of Pearl's $do$ operator.

We consider optimal action selection in a causal environment, providing an extension of the formal criteria (expected utility) to be used when a causal model is present in the environment. Our results provide a theoretical foundation for previous works that address this problem, considering a known or partially known causal model. Additionally, we consider the case in which the structure of the causal model is not known, proposing a novel approach based on Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition.

We are confident that further developments based on our results will show that Causality is a fundamental concept for decision making under uncertainty.

\section*{Declarations}

\section*{Funding}
The present research has been funded by the Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT).

\section*{Availability of data and material}
Not applicable

\section*{Code availability}
Not applicable

\section*{Conflict of interest} 
The authors declare that they have no conflict of interest.

\section*{Authors' contributions}
Equal contribution

% BibTeX users please use one of
\bibliographystyle{apalike}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliographystyle{spbasic}
\bibliography{Bibliografia.bib}
\end{document}
% end of file template.tex

### 2020-08-29 3rd_year_eval/Slides/3rd_year_slides.tex
slides

@@
 
+\begin{frame}
+Previously we focused on the \textit{use} part.
+\end{frame}
+
 \subsection{Relevance}
@@
 \begin{itemize}
-\item Decision making under uncertainty is fundamental for Artificial Intelligence (\cite{lake2017building}).
-\item Optimal policies in Reinforcement Learning are the best action for each state.
+\item Decision Making Under uncertainty (DMU) is fundamental for Artificial Intelligence (\cite{lake2017building}).
+\item Any Bayesian procedure is based on Decision Making
+\item Optimal policies in Reinforcement Learning are the \textit{best} action for each state.
 \item Reinforcement Learning (RL) is \textit{not} a Causal Problem.
@@
 \item Rationality $+$ probabilistic information $=$ maximize expected utility.
-\item Rationality $+$ \textbf{known} causal information $=$ maximize expected utility considering what is caused by actions (\cite{pearl2009causality}, \cite{lattimoreNIPS2016})
+\item Rationality $+$ \textbf{known} causal information $=$ maximize expected utility considering what is caused by actions \citep{pearl2009causality,lattimoreNIPS2016}
 \item Rationality $+$ \textbf{unknown} causal information $=$ ¿? 
@@
 
+\subsection{Research Questions}
+\begin{frame}
+\frametitle{Research Questions}
+We recall the Research Questions stated in \cite{gonzalez2019causal}:\\
+The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
+\begin{enumerate}
+\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
+\item How can a causal structure help a decision maker in order to guide his learning process?
+\item What does the rationality assumption implies about how to choose when considering causal information? 
+\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
+\end{enumerate}
+\end{frame}
+
 \subsection{Research Problem}
@@
 We ask how a decision maker could \textit{learn} about the causal structure that controls his environment in order to \textit{make good choices} with respect to his preferences.
+\end{itemize}
+\end{frame}
+
+\subsection{Some Background}
+\begin{frame}
+\frametitle{Some Background}
+\begin{itemize}
+\item Rationality.
+\item Decision Theory: rational decision making if and only if utility maximization
+\item Causality: Context and direction. Interventions. 
 \end{itemize}
@@
 \begin{itemize}
-\item The general objective and research problem to be studied were defined, as well as a mathematical problem which we called a Causal Decision Problem.
+\item The general objective, research problem and research questions to be studied were stated, as well as a mathematical problem which we called a Causal Decision Problem.
 \item We proposed a preliminar solution which allowed for an optimal action to be found by a decision maker under some assumptions:
@@
 \item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility \citep{gonzalez2018playing}.
-\item Poster presented at Causality Workshop ICML 2018.
-\item Proposal Defense (27-Nov-2018)
+\item Poster presented at Causality Workshop ICML 2018 at Stockholm, Sweden.
+\item Proposal Defense (27-Nov-2018).
 \end{itemize}
@@
 \item The solution criterion given in Equation \ref{first_version} for causal decision making was further developed and extended for the case when a decision maker does not know the causal model which controls his environment:
-\begin{equation}{\label{second_version}}
-a \succeq b \textrm{ iff } \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right) \geq \sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right).
-\end{equation}
+\begin{eqnarray*}{\label{second_version}}
+a &\succeq & b\\ 
+   &\textrm{ iff }&\\ 
+\sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right) \\
+&\geq & \\ \sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right).
+\end{eqnarray*}
 \end{itemize}
@@
 \begin{itemize}
-\item Using \ref{second_version} we were able to define the notion of a Causal Game and the idea of a Nash equilibria within such game.
+\item Using the previous inequality we were able to define the notion of a Causal Game and the idea of a Nash equilibria within such game.
 We define a function $u^C_i : A \to \mathbb{R}^{+}$, for each player:
 \begin{equation}{\label{games_causal_ut}}
-u^C_i (a) = \sum_{f \in A}  u_i(a) \left( \sum_g P^g_i (h_i(f) | do(a_i), a_{-i}) P_C(g) \right) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
+u^C_i (a) = \sum_{f \in A}  u_i(a) \left( \sum_g P^g_i (h_i(f) | do(a_i), a_{-i}) P_C(g) \right).
 \end{equation}
@@
 \begin{itemize}
-\item We further refined the concept of a Causal Game and that of a Nash Equilibrium for Causal Games.
-\item We also studied causal-structure learning via interaction with a causally controled environment.
+\item We further developed the machinery required for the Causal Game and that of a Nash Equilibrium for Causal Games.
+\item We also studied causal-structure learning via interaction with a causally controled environment using a purely Bayesian approach based on random graphs.
 \end{itemize}
@@
 \frametitle{Causal Games: New work}
-We refined the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:
+We improved the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:
 \begin{itemize}
@@
 \begin{frame}
+\frametitle{Random Graphs}
+\begin{itemize}
+\item What is a Random Graph? 
+\item Were invented by Erdös and Renyi. 
+\item They are random objects which take its values in the space of graphs (or adjacency matrices).
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Random Graphs: most common models}
+\begin{itemize}
+\item We will be using the following model:
+\item Among $N$ vertices, draw an edge between vertices $i$ and $j$ with probability $p_{ij}$.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Assumptions}
+\begin{itemize}
+\item Rational agent.
+\item Causal sufficiency
+\item Knowledge of the \textit{variables} of the model.
+\item Convergence to true parameter of Bayesian estimators \cite{autzen2018bayesian}.
+\item Existence of a causal ordering of the variables.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{Main Idea}
 \begin{itemize}
+\item In Bayesian inference, one starts by focusing in the source of uncertainty and defining a model over it.
+\item Since we are trying to learn causal structure between known variables, our problem reduces to find causal relations between pairs of variables.
+\item The source of our uncertainty is the existence or not of a link (causal relation) between nodes in a graph. 
+\item We will model such uncertainty using random graphs.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Main Idea: How}
+\begin{itemize}
 \item Let a rational agent consider the following set of variables $\mathcal{X}=\{ X_1,...,X_n \}$ which are causally related, even though the agent does not now how.
-\item the agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$.
+\item The agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$.
 \item Let $p_{ij}$ be the belief that the agent has over a causal relation (directed link) existing between variable with index $i$ and variable with index $j$.
@@
 \item Now, we use the methodology found in \cite{gonzalez2018playing}  as well as our results presented in \cite{gonzalez2019theorems} in order to find the best action $a^\ast$ for the obtained graph $G$
+\item Update $p_{ij}$.
 \end{itemize}
@@
 \item Now, we update the $p_{ij}$'s using Bayes Theorem as follows:
-\item For each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$
-\item So we have:
-\begin{equation}{\label{bayesian_updating}}
-p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
-\end{equation}
+\item For each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$.
+\item Update accordingly.
 \end{itemize}
@@
 
+\begin{frame}
+\frametitle{Reproducibility}
+In order to guarantee reproducibility, we declare that:
+\begin{itemize}
+\item $p_{ij}$ are initialized randomly
+\item The probability tables for the \textit{local} causal models are initialized randomly
+\item The probabilities of the \textit{true} causal model are stated. 
+\end{itemize}
+\end{frame}
+
 \section{Publications}
@@
 \item A paper including the decision-making results was orally presented at a NeurIPS 2019 Workshop  \url{https://nehzux.github.io/NewInML2019/}. Cited as \cite{gonzalez2019theorems}
-\item Journal paper submitted to Theory and Decision  \url{https://www.springer.com/journal/11238/}
+\item Journal paper \textbf{submitted} to Theory and Decision  \url{https://www.springer.com/journal/11238/}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Learning graphical structure}
+\frametitle{Future Work: Learning graphical structure}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Bayesian Nonparametrics}
+\frametitle{Future Work: Bayesian Nonparametrics}
 Bayesian non parametrics is a field within Bayesian statistics in which priors of infinite dimension are allowed, so models with a greater expressibility can be developed.
@@
 \item Fully Bayesian learning procedure.
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusions}
+\begin{itemize}
+\item This year we have focused on learning an unknown \textit{causal structure} by interacting with a causally controlled environment.
+\item Previously, we had addressed the question on how to \textit{use} causal information in order to make good choices. 
+\item Together, this two tasks answer questions 1, 2 and 3 of the mentioned Research Questions.
 \end{itemize}

### 2020-08-29 Theory_and_decision/Draft/Theory_and_decision_draft.tex
subseccion sobre teoria de decision antes de los teoremas

@@
 Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations.
+In fact, it can be argued that RL is not a causal problem in itself \cite{gonzalezsoto2019reinforcement}.
 
@@
 \subsection{Aim}
-Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying}.
-We provide here an implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl.
+Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying, gonzalez2018playing}.
+We provide here an easily implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl.
 With our result we expect to highlight the importance and applicability of causal relations in decision making and the wide variety of possible applications.
@@
 In Section \ref{joyce_theory} we make reference to another theory, Joyce's Causal Decision Making \citep{joyce1999foundations}.
+For further details on classical (non-causal) decision making, see \cite{bernardo2000bayesian} and \cite{gilboa2009decision}.
 %We will later talk about another theory: Joyce's Causal Decision Making \citep{joyce1999foundations}.
-For further details on classical (non-causal) decision making, see \cite{bernardo2000bayesian} and \cite{gilboa2009decision}.
+
+\subsection{Some Theory}
+\begin{definition}
+An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$.
+Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$.
+\end{definition}
+
+When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty, \cite{bernardo2000bayesian}:
+
+\begin{definition}
+A Decision Problem under Uncertainty is an uncertain environment $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ plus a preference relation $\succeq$ defined over $\mathcal{C}$ 
+\end{definition}
+
+\begin{definition}
+A Decision Problem under Uncertainty is said to be bounded if there exists a pair of consequences $c_\ast$ and $c^\ast$ such that for every $c \in \mathcal{C}$, $c^{\ast} \succeq c \succeq c_\ast$
+\end{definition}
+
+\begin{definition}
+A Decision Problem under Uncertainty is said to be finite if the set $\mathcal{A}$ of available actions is finite.
+\end{definition}
 
@@
 Let $\mathcal{E}$ the union of every $E_i$, we assume $\mathcal{E}$ to be a $\sigma$-algebra of events, and $\mathcal{C}$ the union of all $C_i$.
-
-\begin{definition}
-An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$.
-Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$.
-\end{definition}
-
-When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty, \cite{bernardo2000bayesian}:
-
-\begin{definition}
-A Decision Problem under Uncertainty is an uncertain environment $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ plus a preference relation $\succeq$ defined over $\mathcal{C}$ 
-\end{definition}
-
-\begin{definition}
-A Decision Problem under Uncertainty is said to be bounded if there exists a pair of consequences $c_\ast$ and $c^\ast$ such that for every $c \in \mathcal{C}$, $c^{\ast} \succeq c \succeq c_\ast$
-\end{definition}
-
-\begin{definition}
-A Decision Problem under Uncertainty is said to be finite if the set $\mathcal{A}$ of available actions is finite.
-\end{definition}
-
-\cite{bernardo2000bayesian} derive the existence of a subjective probability measure from a set of \textit{rationality axioms}, where a decision maker has some mechanism of quantifying uncertainty in terms of real numbers within the $[0,1]$ interval, and then the classical Kolmogorov axioms are derived from such construction, and therefore all of the known machinery of Probability Theory.
+\cite{bernardo2000bayesian} derives the existence of a subjective probability measure from a set of \textit{rationality axioms}, where a decision maker has some mechanism of quantifying uncertainty in terms of real numbers within the $[0,1]$ interval, and then the classical Kolmogorov axioms are derived from such construction, and therefore all of the known machinery of Probability Theory.
 We now state Savage's Theorem.

### 2020-08-29 Theory_and_decision/Draft/Theory_and_decision_draft.tex
añadiendo citas

@@
 
-\subsection{Some Theory}
+\subsection{Preliminary Definitions}
 \begin{definition}
@@
 \subsection{Causal Graphical Models}
-A Causal Graphical Model (CGM) consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect.
+A Causal Graphical Model (CGM) \citep{koller2009probabilistic,sucar2015probabilistic} consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect.
 Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
@@
 Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
-\[ P_\mathcal{G}(Y | do(X=x) \neq P_\mathcal{G}(Y | do(X=x'). \]
+\[ P_\mathcal{G}(Y | do(X=x)) \neq P_\mathcal{G}(Y | do(X=x')). \]
 \end{definition}

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en el abstract

@@
 Decision making under uncertainty has been well studied when information is considered at the associative (probabilistic) level.
-The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rational choice using purely associative information.
+The classical Theorems of von Neumann-Morgenstern and Savage provide a formal criterion for rational choice using purely associative information, which form the basis of many learning algorithms such as those used in Reinforcement Learning.
 In this work, we consider decision problems in which available actions and consequences are causally connected.
-We define a Causal Decision Problem and state a previous result from J.
-Pearl, showing that it can be considered as a causal version of the classical von Neumann-Morgenstern Theorem.
+We define a Causal Decision Problem and within this framework we state a previous result from J.
+Pearl, which relies on a known causal model and thus showing that it can be considered as a causal version of the classical von Neumann-Morgenstern Theorem.
 Furthermore, we consider the case when the causal mechanism that controls the environment is unknown to the decision maker, and propose and prove a causal version of Savage's Theorem.

### 2020-08-30 Theory_and_decision/Draft/title_page.tex
cambios en abstract

--- 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
@@
 \RequirePackage{fix-cm}
-
-\documentclass{svjour3}                     % onecolumn (standard format)
+%
+\documentclass{svjour3} 
 %\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
@@
 %
+\smartqed  % flush right qed marks, e.g. at end of proof
 %
+\usepackage{graphicx}
+\usepackage{hyperref}
 %
-\usepackage{mathptmx}      
-\usepackage[utf8]{inputenc}
-\usepackage[pdftex]{color,graphicx}
-\usepackage{listings}
-\usepackage{breakcites}
-\usepackage{dsfont}
-\usepackage{hyperref}
-\usepackage{natbib}
-\usepackage{amssymb,amsmath,latexsym}
-\usepackage[margin=1.5cm]{geometry}
-\usepackage{enumitem}
-%\bibliographystyle{stylename}
-%\usepackage{fancyhdr}
-%\pagestyle{fancy}
-\usepackage{tikz}
-\usetikzlibrary{trees}
-\usetikzlibrary{calc}
-\usepackage{color}
-\usepackage{todonotes}
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
 % etc.
@@
 %\titlerunning{Short form of title}        % if too long for running head
-%
-%\author{Mauricio Gonzalez-Soto         \and
-%        Luis E. Sucar \and
-%        Hugo J. Escalante
-%}
-\author{Anonymous}
+
+\author{Mauricio Gonzalez-Soto         \and
+        Luis E. Sucar \and
+        Hugo J. Escalante
+}
+
 %\authorrunning{Short form of author list} % if too long for running head
 
-%\institute{Mauricio Gonzalez-Soto \at
-%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              Mexico\\
-%              \email{mauricio@inaoep.mx}           \\
-%              \url{https://orcid.org/0000-0003-2668-9013}        
-%           \and
-%           Luis E. Sucar \at
-%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              M\'exico\\
-%              \email{esucar@inaoep.mx}\\
-%              \url{https://orcid.org/0000-0002-3685-5567}
-%           \and
-%           Hugo J. Escalante \at
-%           Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              M\'exico\\
-%              \email{hugojair@inaoep.mx}\\
-%              \url{https://orcid.org/0000-0003-4603-3513}
-%}
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
+\institute{Mauricio Gonzalez-Soto \at
+              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
+              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
+              Mexico\\
+              \email{mauricio@inaoep.mx}           \\
+              \url{https://orcid.org/0000-0003-2668-9013}        
+           \and
+           Luis E. Sucar \at
+              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
+              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
+              M\'exico\\
+              \email{esucar@inaoep.mx}\\
+              \url{https://orcid.org/0000-0002-3685-5567}
+           \and
+           Hugo J. Escalante \at
+           Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
+              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
+              M\'exico\\
+              \email{hugojair@inaoep.mx}\\
+              \url{https://orcid.org/0000-0003-4603-3513}
+}
 
 \maketitle
-% modificar abstract para reflejar cambios en el paper
+
 \begin{abstract} 
@@
 
-\keywords{Causality \and Decision Making \and Game Theory}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 \end{abstract}
-\tableofcontents
-\newpage
-\section{Introduction}
-\label{intro}
-Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}.
-Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{waldmann2013causal,danks2014unifying,neil2019causality}.
-It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}.
-Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{tversky1977causal, tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.
 
-Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent, located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward.
-Optimal policies, which are prescriptions of which action must be taken at some state $s$, satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem.
-Such solution is equivalent to maximizing \textit{expected utility} \citep{webb2007game}, which is the decision criterion prescribed by the most commonly used decision-making theories: von Neumann-Morgenstern's and Savage's.
-Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations.
-In fact, it can be argued that RL is not a causal problem in itself \cite{gonzalezsoto2019reinforcement}.
-
-On the other hand, an important aspect of acting in the world is being able to make decisions under uncertain conditions \citep{bernardo2000bayesian,kahneman2000choices,gilboa2009decision,lake2017building}.
-\cite{von1944theory} gave an answer for how to make choices if rational preferences are assumed and if the decision maker knows the stochastic relation between actions and outcomes: maximize expected utility.
-If no such relation is known, then \cite{savage1954the} showed that a rational decision maker must choose \textit{as if} she is maximizing expected utility using a \textit{subjective} probability distribution.
-Such theorems provide formal criteria for decision making if rationality is assumed.
-
-These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms learn \textit{optimal policies} in such a way that any action prescribed by such optimal policy achieves the maximum expected utility \citep{sutton1998reinforcement,webb2007game}.
-In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on associative relations which are expressed using correlations or probability distributions.
-
-It is a natural question how to formalize rational decision making when causal information is present; answering such question is relevant given the importance of causal relations in everyday life.
-Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage.
-
-The previous question has been already considered by \cite{nozick1969newcomb}, \cite{lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making and also by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known by the decision maker.
-
-\subsection{Motivation}
-Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent.
-In many real-world applications, the agent doesn't have access to all of the parameters required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover this relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.
-
-Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data.
-For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, they cannot explain \textit{why} a specific trajectory was chosen by the algorithm.
-This is highly relevant in real-world applications, e.g.
-self-driving cars, where it is very important to understand why an accident happened.
-
-%For example, as told by \cite{bornstein2016artificial}, at the University of Pittsburgh Medical Center a team of researchers tried to use Machine Learning to predict whether pneumonia patients might develop severe complications.
-For this purpose they trained Neural Networks and Decision Trees using the hospital's own data.
-Neural Networks outperformed Decision Trees, but only by studying the decisions made by the latter the doctors did find out that the algorithms instructed doctors to send home pneumonia patients who already had asthma, despite the fact that asthma sufferers are known to be extremely vulnerable to complications.
-The problem relied in the training data, because the hospital policy was to automatically send asthma sufferers with pneumonia to intensive care, and this policy worked so well that asthma sufferers almost never developed severe complications.
-It was only through the interpretability of the Decision Trees that the doctors didn't send asthma patients with pneumonia home to a certain death.
-
-%Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo \citep{silver2017mastering} defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, although it was a very sophisticated method for achieving it.
-Using only associations between variables it is not possible to infer which one is the cause and which one the effect, something extra is required.
-The fundamental difference between causal and associative models is that causal models consider \textit{doing} over \textit{observing} \citep{pearl2018why}.
-Once a causal relation has been established between two events, performing, or intervening, over a cause allows one to predict the outcome in a more robust way than if using only correlations between observations.
-
-Learning a causal model of an environment and using it to act upon the environment allows us to \textit{explain} aspects of the model that a purely associative model would not be able to explain.
-It allows one to ask \textit{why}.
-Since a causal model, once it is learned, does not depend neither on the agent nor its preferences, the gained knowledge could be easily transferred for problems of similar domain where the use of the acquired causal relations is useful.
-Our work provides the foundations of decision making algorithms which are based on causal relations and, as will be shown, are able to learn causal models of the environment.
-
-%The impact of Causal Inference goes beyond Machine Learning and Computer Science.
-For example, economists are interested in understanding what did certain public policy caused \citep{athey2017impact} or how human decision makers that show \textit{inconsistent} preferences over time could be oriented if the causal consequences of this inconsistent behavior is introduced into a decision-making model \citep{albers2016motivating}.
-In complex adaptive systems, causal relations can be used to clarify the complex interactions between agents in the system \citep{abbott2017complex} as well as for prediction and planning \citep{hunt2016ants,brock2018causality}.
-
-%cerrar esta sección sobre cómo mi trabajo impacta en la motivación
-
-\subsection{Aim}
-Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying, gonzalez2018playing}.
-We provide here an easily implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl.
-With our result we expect to highlight the importance and applicability of causal relations in decision making and the wide variety of possible applications.
-
-\subsection{Main Result}
-In this work we extend Pearl's result to the case where a decision maker \textit{does not know the causal mechanism} that controls her environment, so she holds \textit{beliefs} about possible causal models and uses such beliefs \textit{as if} they were true, as prescribed by \cite{joyce1999foundations}, in order to attempt to make a good choice given her beliefs.
-We provide an explicit way of choosing an optimal decision given what the decision maker believes about the causal structure of her environment, thus providing a causal version of Savage's Theorem.
-We take a normative point of view to study how causal relations should be used by a \textit{rational} agent when making decisions with uncertain consequences.
-
-In Section \ref{on_causality} we define Causality and state the differences between causal and associative information; having shown this, we state two of the most important decision making results, the von Neumann-Morgenstern and Savage Theorems, which rely on purely associative information; then, our objective is to present two analogous theorems: J.
-Pearl's result, Theorem \ref{causal_ut} which relies on a known causal model, and our main contribution: Theorem \ref{causal_savage}, which is able to use a set of \textit{beliefs} about possible causal models (i.e., a distribution) instead of a known one.
-
-\subsection{Applications}
-Game Theory deals with situations in which several rational decision makers, or players, \textit{interact} while pursuing some well-defined objective \citep{osborne1994course}; the case in which decision makers make a choice simultaneously without knowing the choice made by the other players is called a \textit{strategic game}; a well-known strategic game is the famous \textit{prisoners' dilemma} in which two detainees must choose between confessing or remaining silent and both know the consequences of any combination of actions, what is ignored by each player is the decision made by the other.
-When players ignore both the actions made by other players as well as the knowledge that made them choose a certain action, is called a game with incomplete information, or a \textit{Bayesian Game} which was introduced by \cite{harsanyi1967games1,harsanyi1968games2,harsanyi1968games3}.
-
-In this work we use the Bayesian Game model in order to study what happens when several decision makers have certain knowledge about an environment which is controlled by some, unknown but fixed, causal mechanism.
-We study games in which the players actions \textit{cause} some consequence according to some unknown causal model.
-We apply our a rational choice criterion to formulate a Nash equilibrium for Causal Games by defining a probability updating for Bayesian Causal Games.
-
-Furthermore, we consider the problem of learning optimal actions in a causal environment.
-A decision maker is given the task of learning an optimal action in a causal environment in which a causal model controls the relation between actions and outcomes.
-We consider the case in which the causal model is unknown or partially known to the decision maker, so she has to use beliefs about possible causal models.
-From the decision maker point of view, having in mind a causal model allows one to ask what would have happened if another action was taken without actually performing the alternative action; therefore, being able of choosing the best action in terms of the probabilities of causing a good outcome.
-Also, learning a causal model of a certain environment will facilitate knowledge transfer to similar domains, as the underlying cause-effect structure has been captured and could later be used.
-
-% añadir motivación del paper y de las teorías
-
-\subsection{Organization of the paper}
-\label{organization}
-This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which use only associative information in their formulation; then, in Section \ref{causal_decision_problems} a brief introduction to Causation and Causal Graphical Models is given.
-In Section \ref{related_work}  we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning.
-In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure, as well as our main result, Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure.
-Finally two applications of Theorem \ref{causal_savage} are described.
-First, a causal analogue of Nash Equilibrium for games in which actions and consequences are causally connected is defined.
-Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works.
-Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider;  and in Section \ref{conclusion} we summarize the main contributions of our work.
-
-\section{The Classical Decision Making Theorems: von Neumann-Morgenstern and Savage}
-\label{classical_theorems}
-A decision problem under uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors.
-Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.
-
-The most well-known theories for Decision Making are those from \cite{von1944theory} and \cite{savage1954the}.
-In the former theory it is assumed that the decision maker knows the stochastic relation between actions and consequences, which is also known as \textit{decision under risk}, and in that case the theory guarantees that the decision maker behaves \textit{as if} she maximizes the expected value of a utility function.
-If the decision maker doesn't know the probabilities of observing an outcome given a chosen action, then Savage's theory guarantees that the decision maker behaves as if she has in mind a \textit{subjective} probability distribution and a utility function and chooses the action which maximizes the expected utility with respect to that subjective probability distribution and the utility function, this is known in the literature as \textit{decision under uncertainty}.
-
-Other Decision Making theories exist, such as Kahneman and Tversky's Prospect Theory \citep{kahneman1979prospect}, Gilboa's Case-Based Decision Theory \citep{gilboa1995case}, among others that are out of the scope of this work.
-In Section \ref{joyce_theory} we make reference to another theory, Joyce's Causal Decision Making \citep{joyce1999foundations}.
-For further details on classical (non-causal) decision making, see \cite{bernardo2000bayesian} and \cite{gilboa2009decision}.
-%We will later talk about another theory: Joyce's Causal Decision Making \citep{joyce1999foundations}.
-
-\subsection{Preliminary Definitions}
-\begin{definition}
-An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$.
-Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$.
-\end{definition}
-
-When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty, \cite{bernardo2000bayesian}:
-
-\begin{definition}
-A Decision Problem under Uncertainty is an uncertain environment $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ plus a preference relation $\succeq$ defined over $\mathcal{C}$ 
-\end{definition}
-
-\begin{definition}
-A Decision Problem under Uncertainty is said to be bounded if there exists a pair of consequences $c_\ast$ and $c^\ast$ such that for every $c \in \mathcal{C}$, $c^{\ast} \succeq c \succeq c_\ast$
-\end{definition}
-
-\begin{definition}
-A Decision Problem under Uncertainty is said to be finite if the set $\mathcal{A}$ of available actions is finite.
-\end{definition}
-
-\subsection{Von Neumann-Morgenstern Theorem}
-\label{vNM-M}
-\cite{von1944theory} had the objective of justifying strategies in which players in a game maximized their expected utility.
-This theorem considers a scenario of \textit{decision under risk} and rational preferences; this is, choosing between uncertain outcomes with known probabilities.
-
-Formally, using the notation from \cite{gilboa2009decision}, we consider a set $X$ of alternatives.
-Let $L$ be the set of lotteries with finite support over $X$.
-The object of choice are the elements $l \in L$ that are known to the decision maker; we represent the decision maker's preferences by a preference relation $\succeq \subseteq L \times L$ that  satisfies being complete, transitive, continuous and a technical condition called \textit{independence}.
-This family of conditions is called \textit{von Neumann-Morgenstern rationality axioms} as described in \cite{gilboa2009decision}.
-\begin{theorem}[von Neumann-Morgenstern]{\label{vNM}}
-A preference relation $\succeq \subseteq L \times L$ where $L$ is a set of lotteries with finite support over a set $X$ satisfies the von Neumann-Morgenstern rationality axioms if and only if there exists a function $u: X \to \mathbb{R}$ such that for every $P, Q \in L$ we have that
-\begin{equation}
-P \succeq Q \textrm{ if and only if } \sum_{x \in X} P(x) u(x) \geq \sum_{x \in X} Q(x) u(x). 
-\end{equation}
-\end{theorem}
-The theorem states that if a rational decision maker knows the probabilities of obtaining a certain outcome, then she must choose \textit{as if} maximizing the expected value of some function $u$ whose existence is guaranteed by Theorem \ref{vNM}.
-See \cite{gilboa2009decision} for details on the proof.
-
-\subsection{Savage's Theorem}
-\label{savage_theorem}
-If a rational decision maker does not know the probabilities of obtaining certain outcomes and does not have a precise, objective quantification of her preferences (utility function), then it is Savage's Theorem \citep{savage1954the}, which gives a formal choosing criterion.
-Savage's result extends von Neumann-Morgenstern Theorem since it considers the case in which a rational decision maker does not know neither her utility function nor the probabilities to be used in order to obtain the expected values required for making choices according to the von Neumann-Morgenstern Theorem.
-The notation and setting we use here is the one used by \cite{bernardo2000bayesian}.
-Consider a set $\Omega$.
-Consider a countable set $\mathcal{A}$ of available actions; for each action $a_i \in \mathcal{A}$, a partition $E_i$ of $\Omega$ and a set of consequences $C_i$.
-Let $\mathcal{E}$ the union of every $E_i$, we assume $\mathcal{E}$ to be a $\sigma$-algebra of events, and $\mathcal{C}$ the union of all $C_i$.
-\cite{bernardo2000bayesian} derives the existence of a subjective probability measure from a set of \textit{rationality axioms}, where a decision maker has some mechanism of quantifying uncertainty in terms of real numbers within the $[0,1]$ interval, and then the classical Kolmogorov axioms are derived from such construction, and therefore all of the known machinery of Probability Theory.
-We now state Savage's Theorem.
-Details can be found in \citep{kreps1988choice,bernardo2000bayesian,gilboa2009decision}.
-
-\begin{theorem}[Savage]{\label{savage}}
-In a finite, bounded Decision Problem under Uncertainty $(\mathcal{A}, \mathcal{C}, \mathcal{E}, \succeq)$, the preference relation $\succeq$ satisfies the Savage rationality axioms if and only if there exists:
-\begin{itemize}
-\item A \textit{probability measure} $P$, called a \textit{subjective probability}, that associates with each uncertain event $E \in \mathcal{E}$ a real number $P(E)$. 
-\item A utility function $u : \mathcal{C} \to \mathbb{R}$ such that it associates each consequence with a real number $u(c)$. 
-\end{itemize}
-Such that for $a_1$ and $a_2$ actions in $\mathcal{A}$,
-\begin{eqnarray*}
- &a_1 \succeq a_2&\\
- & \textrm{ if and only if }&\\
- &\mathbb{E}_P[u(a_1)] \geq  \mathbb{E}_P[u(a_2)],&
-\end{eqnarray*}
-where $\mathbb{E}_P[u(a)]$ is given by $\sum_{j \in J} u(c_j) P(E_j)$ since any action $a$ can be identified as $\{ c_j | E_j ; j \in J \}.$ Details in Section 2.5.2 from \cite{bernardo2000bayesian}.
-\end{theorem}
-\begin{proof}
-Proposition 2.22 in \cite{bernardo2000bayesian}.
-\end{proof}
-This theorem states that if a rational decision maker does not know the precise probabilities of outcomes given that an action is taken, then she chooses \textit{as if} having in mind a probability assignment to the uncertainties in her environment and using such probabilities to calculate the expected utility with respect to a subjective utility function that represents her preferences.
-This result also gives a precise definition of \textit{subjective probability} as a quantification of uncertainty which is used to make good decisions.
-See \cite{hens1992note} and \cite{gilboa2009decision} for further details.
-See \cite{ellsberg1961risk} and \cite{binmore2008rational} for critiques of the Savage axioms.
-
-\subsection{Limitations of the Theorems}
-\label{theorems_limitations}
-The mentioned theorems have the limitations of relying only on associative information; this is, on correlations and patterns found in data rather than strong causal relations.
-For example, as mentioned by \cite{pearl2009causality}, using only associative information one would be tempted to stop going to the doctor in order to stop being diagnosed as sick.
-On the other hand, our results provide an easy to implement way of making choices using causal relations and beliefs about such relations as used by humans.
-Human beings \textit{know} (this is, use causal relations) that going to the physician does not \textit{cause} being sick and this is what Pearl calls \textit{commonsensical} decision making.
-Also, as we will mention later in the paper, associative information lacks an \textit{ordering} thus failing to determine \textit{if $X$ is performed, then $Y$ will be caused}.
-
-Formally, in this paper we consider the actions chosen by a decision maker as an \textit{interventions} over a causal graphical model; this is, in the classical von Neumann-Morgenstern and Savage Theorems, once a decision maker chooses an action $a$, the observed consequence has a probability $P(c|a)$; if causal information is considered, then the probability of a certain consequence is $P(c|do(a))$.
-In the next section we state some basic facts on causality and why $P(\cdot |x)$ is different from $P(\cdot | do(x))$ for some value $x$.
-
-\section{Causal Decision Problems}
-\label{causal_decision_problems}
-\subsection{On causality}
-\label{on_causality}
-The concept of Causality deals with regularities found in a given environment which are stronger than probabilistic (or associative) relations in the sense that a causal relation allows for evaluating a change in the \textit{consequence} given that a change in the \textit{cause} is performed.
-The \textit{manipulationist} interpretation of Causality (details in \cite{woodward2005making}) is adopted here.
-The main paradigm is clearly expressed by \cite{campbell1979quasi} as \textit{manipulation of a cause will result in a manipulation of the effect}.
-Consider the following example from \cite{woodward2005making}: manually forcing a barometer to go down won't cause a storm, whereas the occurrence of a storm will cause the barometer to go down.
-
-We know from \cite{granger1969investigating} and \cite{lamport1978time} that there are two sufficient and necessary conditions for causality: a context and an order.
-We will adopt here the formal definition of Causality given by \cite{spirtes2000causation}; i.e., a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}.
-Similar descriptions of the manipulationist approach can be found in \cite{holland1986statistics}.
-Causal inference tools, such as Pearl's do-calculus \citep{pearl2009causality} allow to find the effect of an intervention in terms of probabilistic information when certain conditions are met.
-For what remains, we assume the \textit{causal axioms} found in \citep{spirtes2000causation} as well as the condition known as \textit{causal sufficiency} \citep{pearl2009causality}.
-
-\subsection{A definition of Causality}
-The formal definition of Causality that will be adopted in this work is the definition provided by \cite{spirtes2000causation} which states that causation is a \textit{stochastic} relation between \textit{events} in a probability space; this is, some event (or events) \textit{causes} another event to occurr.
-\begin{definition}{\label{causal_relation}}
-Let $(\Omega, \mathcal{F}, \mathbb{P})$ a finite probability space, and consider a binary relation $\to \subseteq \mathcal{F} \times \mathcal{F}$ which is:
-\begin{itemize}
-\item Transitive: If $A \to B$ and $B \to C$ for any $A, B, C \in \mathcal{F}$ then $A \to C$.
-\item Irreflexive: For all $A \in \mathcal{F}$ it doesn't hold that $A \to A$.
-\item Antisymmetric: For $A,B \in \mathcal{F}$ such that $A \neq B$ if $A \to B$ then it doesn't hold that $B \to A$.
-\end{itemize}
-We say that $A$ is \textit{a cause} of $B$ (or that $A$ causes $B$, $A$ is the cause and $B$ is the effect) if $A \to B$.
-It is important to note that an event may have more than one cause, and that not necessarily each one of this causes is sufficient to produce the effect.
-\end{definition}
-
-We assume that for any event $A \in \mathcal{F}$ there are no causes lying outside $\mathcal{F}$ (causal sufficiency).
-It is also be assumed that the relations expressed by $\to$ are the only causal relations in the environment.
-This assumption can be interpreted as not allowing \textit{intermediate} events between events known to be causally related.
-For example, if in our space we have that striking a match causes fire, we do not allow into consideration the underlying chemical reactions that cause fire from friction.
-In this sense, we say that striking a match is a \textit{direct cause} of fire and this are the only causes considered herein \citep{spirtes2000causation}.
-
-\subsection{Representation into a Directed Acyclic Graph}
-The causal relations contained in $\to$ can be summarized in a graph $G=(V,E)$ in the following way: If $A \to B$ then the graph must contain a node $A \in V$ representing $A$, a node  $B \in V$ representing $B$ and a directed edge $e \in E$ connecting the respective nodes in the direction of the causal relation.
-
-\begin{proposition}
-Given a causal relation $\to$ as in Definition \ref{causal_relation} then the graph that is obtained by considering nodes for events and edges for the causal relations as previously described is a Directed Acyclic Graph.
-\end{proposition}
-
-\begin{proof}
-The graph is directed since the definition of causality imposes a direction between events; namely, a direction between the cause and the effect.
-To see that the graph is acyclic, suppose that a cycle $A \to B \to C \to A$ exists, this would imply, because of transitivity, that $A \to A$, which cannot be since the relation is irreflexive.
-\end{proof}
-
-Notice that since the graph is finite, there exist some nodes that do not have causes, which are called \textit{exogenous}.
-If an event $A$ is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
-It is proven in \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.
-
-\subsubsection{Relations between a causal graph and probabilities}
-Given a Directed Acyclic Graph (DAG) it is possible to obtain a probability measure that expresses the conditional independence relations that are expressed in the graph \citep{koller2009probabilistic}, which we will call $P_\mathcal{G}$.
-For the DAG built from the causal relations, we require that its correspondent $P_\mathcal{G}$ satisfies the following conditions \citep{spirtes2000causation}:
-\begin{itemize}
-\item Markov Causality: an event $V$ (or node in $\mathcal{G}$) is independent of every other event $A$ such that $A$ isn't either a cause nor an effect of $V$ given the causes of $V$.
-\item Causal Minimality: No proper sub-graph of $\mathcal{G}$ satisfies the Markov Causality condition.
-\item Causal Faithfulness: The Markov Causal condition contains all of the conditional independence statements expressed by the DAG $\mathcal{G}$.
-\end{itemize}
-
-Also, we consider an extra pair of conditions.
-The first one, known as Causal Sufficiency, is about the \textit{nature} of the model: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ \textit{outside} of the model $\mathcal{G}$\citep{spirtes2000causation,pearl2009causality,sucar2015probabilistic}.
-The second required condition is that the causal mechanism remains unchanged after interventions, this is called in \citep{woodward2005making} as \textit{invariantness}, this basically means that an intervention does not \textit{break} the causal mechanism when intervening upon it.
-These conditions are required in an axiomatic fashion so they will be taken as they are without questioning.
-
-\subsection{Causal Graphical Models}
-A Causal Graphical Model (CGM) \citep{koller2009probabilistic,sucar2015probabilistic} consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect.
-Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ...
-, x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ \citep{pearl2009causality,koller2009probabilistic,sucar2015probabilistic}.
-To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable.
-It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ \citep{pearl2018why}.
-In fact, using the \textit{Do} operator we give a definition of a variable $X$ causing $Y$.
-\begin{definition}
-Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
-\[ P_\mathcal{G}(Y | do(X=x)) \neq P_\mathcal{G}(Y | do(X=x')). \]
-\end{definition}
-It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph \citep{sucar2015probabilistic}.
-\subsubsection{The identifiability problem}
-Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
-It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) \citep{peters2012identifiability,mooij2016distinguishing}.
-
-\subsubsection{Do-Calculus}
-The Do-Calculus is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data \citep{pearl1995causal,pearl2009causality}.
-
-It is important to introduce some notation.
-Consider a causal graphical model $\mathcal{G}$ and $X,Y,Z$ disjoint sets of nodes of $\mathcal{G}$.
-We denote by $\mathcal{G}_{\bar{X}}$ the graph that is obtained by deleting from $\mathcal{G}$ all of the edges that enter nodes in $X$.
-In the same way, $\mathcal{G}_{\underline{X}}$ is the graph obtained by deleting the edges that emerge from $X$.
-Finally, $\mathcal{G}_{\underline{Z}\bar{X}}$ is the graph obtained by deleting edges incoming into $X$ and outgoing from $Z$.
-
-\begin{theorem}[Do-Calculus, \cite{pearl2009causality}]{\label{docalculus}}
-Let $\mathcal{G}$ a CGM and $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
-\begin{itemize}
-\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ and $W$, then
-\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
-\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ and $W$, then
-\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
-\item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.
-If $Y$ is conditionally independent from $Z$ given $X$ and $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
-\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
-\end{itemize}
-\end{theorem}
-\begin{proof}
-See \cite{pearl2009causality}.
-\end{proof}
-\begin{theorem}[Properties of Do-calculus]{\cite{peters2017elements}}\\
-The following statements hold:
-\begin{itemize}
-\item The Do-Calculus is complete; this is, sufficient for deriving every identifiable interventional distribution \citep{huang2006pearl,shpitser2006identification}.
-\item There exists an algorithm capable of finding all of the identifiable interventions \citep{tian2002,huang2006pearl}.
-\item A necessary and sufficient criteria exists for the identifiability of interventional distributions \citep{shpitser2006identification,huang2006pearl}.
-\end{itemize}
-\end{theorem}
-\begin{proof}
-See \cite{peters2017elements}.
-\end{proof}
-The Do-Calculus rules provide a solution for the identifiability problem:
-\begin{corollary} \cite{pearl2009causality}.\\
-A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each one of them corresponds to any of the rules of the Do-Calculus, that reduce $q$ to a purely observational expression.
-\end{corollary}
-
-\subsection{Further limitations of associative information}
-As we have said, one of the fundamental aspects of Causality is the presence of an order and a context; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
-\begin{equation}
-    p(a,t)=p(a|t)p(t)
-\end{equation}
-and
-\begin{equation}
-    p(a,t)=p(t|a)p(a)
-\end{equation}
-which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$. This situation does not happen while considering causal relations.
-
-Also, we recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$.
-If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false.
-\subsection{Causal Environments and Causal Decision Problems}{\label{causal_problems}}
-We consider decision making with causal information.
-We define a Causal Environment to be an \textit{uncertain environment} such that there exists a Causal Graphical Model (CGM) $\mathcal{G}$ \citep{koller2009probabilistic} which controls the environment; this is, there exists a causal relation between available actions and consequences in the sense that any chosen action will stochastically \textit{cause} a consequence.
-The role of the CGM is to encode all of the causal relations present in the environment, not only between actions and consequences but also between any other variables in the environment.
-\begin{definition}
-A Causal Environment is a tuple $(\Omega, \mathcal{A},\mathcal{G},\mathcal{C},\mathcal{E})$ where $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ is an uncertain environment and $\mathcal{G}$ is a CGM such that the set of variables of $\mathcal{G}$ correspond to the uncertain events in $\mathcal{E}$.
-\end{definition}
-\begin{definition}
-We define a Causal Decision Problem (CDP) as a tuple $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$ where $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C})$ is a Causal Environment and $\succeq$ is a preference relation.
-\end{definition}
-
-For the CGM in a CDP we distinguish two particular variables: one corresponding to the available actions, and one corresponding to the produced (caused) outcome.
-We are considering that only one variable can be intervened upon and that the values of such variable represent the actions available to the decision maker; i.e., the value forced upon such variable under an intervention represents the action taken by the decision maker.
-The intuition behind the definition of a Causal Decision Problem is this: a decision maker chooses an action $a \in \mathcal{A}$, which is automatically inputed into the model $\mathcal{G}$, which outputs the \textit{causal outcome} $c \in \mathcal{C}$.
-We say a CDP is \textit{finite} if the set $\mathcal{A}$ is finite.
-
-\section{Related Work}
-\label{related_work}
-\subsection{Causal Decision Making and Decision-Theoretic foundations of Causal Inference}
-\label{joyce_theory}
-A previous attempt to formalize Decision Theory in the presence of Causal Information was given by \cite{lewis1981causal,joyce1999foundations}.
-According to such formulation, a decision maker must choose whatever action is more likely to (causally) produce a desired outcome while keeping any beliefs about causal relations fixed.
-This is stated by the Stalnaker equation \citep{stalnaker1968})
-\begin{equation}
-u(a)=\sum_{x} P(a \square \to x)u(x),
-\end{equation}
-\noindent
-where $a \square \to x$ is to be read as \textit{if the decision maker does} $a$ \textit{then} $x$ \textit{would be the case} \citep{gibbard1978counterfactuals,kleinberg2013causality}.
-Lewis' and Joyce's work captured the intuition that causal relations may be used to control the environment and to predict what is caused by the actions of a decision maker.
-In Section \ref{section_savage} we will refine the $\square \to$ operator by an explicit way of calculating the probability of causing an outcome by doing a certain action in terms of Pearl's do-calculus.
-
-\cite{heckerman1995decision} provide a framework for defining the notions of cause and effect in terms of decision theoretical concepts, such as states and outcomes and gives a theoretical basis for graphical description of causes and effects, such as the causal influence diagrams from \citep{dawid2002influence}.
-Heckerman gave an elegant definition of causality, but did not address how to actually make choices using causal information.
-
-\cite{dawid2012decision} presented a decision-theoretic approach to causal inference in which a decision maker must take into account how alternatives compare against the other in terms of the \textit{average causal effect}, such approach uses the well-known influence diagrams \citep{dawid2002influence,dawid2003causal} in order to derive formulas that allow an explicit calculation of the average causal effect.
-Influence diagrams have the ability of expressing both intervention variables and chance variables into a single graphical structure in such a way that the standard techniques for probabilistic DAG's still apply.
-An optimality criterion for sequential interventions is obtained by \cite{dawid2008identifying} by maximizing the expectation of outcomes.
-
-\subsection{Causal Reinforcement Learning and Bandit Problems}
-It is known that human beings conceive their actions on the world as \textit{intervening} in the world \citep{hagmayer2009decision}.
-Following this idea, \cite{lattimoreNIPS2016} consider decision problems in which the action to be chosen is an intervention over a known causal graphical model.
-The agent must choose the intervention that maximizes the value of a \textit{target variable} after a series of learning rounds.
-They model their problem by considering that choosing an intervention is choosing an arm of a \textit{slot machine}, usually known as Bandit Problems, in which a gambler chooses an arm and gets some reward.
-From the rewards they estimate probabilities and output an optimal action in the sense of obtaining minimal regret.
-Their work considers that the causal model is fully known, restriction which is partially relaxed by \cite{sen2017identifying}, and mention that the case in which the causal model is unknown is left as an open question, and this is precisely what we are proposing to answer.
-
-In general, in a classic Bandit Problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \citep{sutton1998reinforcement}.
-Several algorithms exists for finding the best arm in a multi-arm bandit, such as those described in \cite{bubeck2009pure,audibert2010best,gabillon2012best,agarwal2014taming,jamieson2014lil,jamieson2014best,chen2015optimal,carpentier2016tight,russo2016simple,kaufmann2016complexity}, but none of these works consider causal-governed environments.
-\cite{ortega2014generalized} gives results on sequential decision-making using Generalized Thompson Sampling that could be extended into causal inference problems.
-The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices, but the problem of \textit{acquiring} this causal knowledge is left unattacked.
-
-Bandit Problems are a particular case of a Reinforcement Learning problem, which as mentioned in the introduction, is a general framework for interactive learning in which an agent performs an action and receives some feedback from his environment.
-The classical setting for an agent to learn by interaction is that of a Markov Decision Process, in which at time $t$ an agent is situated in state $s_t$ and takes an action $a_t$, he then observes a new state $s_{t+1}$ and a reward signal $r_{t+1}$.
-The objective is to find the \textit{policy} which maximizes the long-run reward.
-It can be shown that this is equivalent to solving the Bellman Equations \citep{sutton1998reinforcement}, and that the actions prescribed by such optimal policy maximize the expected utility.
-This shows the connection between the modern field of Reinforcement Learning and classical results in Decision Theory which justify choosing actions that maximize expected utility when certain \textit{rationality} assumptions are made.
-
-It has been recently studied the connection between Reinforcement Learning and Causation in a series of papers by \cite{gershman2015reinforcement}, \cite{lattimoreNIPS2016}, \cite{bareinboim2015bandits} among others such as \citep{zhang2016markov, zhang2017transfer, lee2019structural, correa2019statistical, zhang2019near}.
-Even though the connection between Reinforcement Learning and causal reasoning is intuitive, a proper  framework for Reinforcement Learning based on Causal Decision Making (see \cite{joyce1999foundations} for an introduction to Causal Decision Making and \cite{pearl2009causality} for a more recent computational version) has not been explored.
-
-In this paper we present a causal decision making criterion useful when the probabilities that hold in an environment are not previously known to the decision maker, which is usually the case in Reinforcement Learning.
-Therefore, the findings of this paper can have an important impact into decision making because of both, the importance of causal relations in artificial intelligence and the realistic assumption of an agent not having full information of her environment.
-
-\section{Main Result: Causal Decision Making}
-\label{causal_results}
-\subsection{A von Neumann-Morgenstern type theorem for causal environments}
-Consider a rational decision maker who faces a causal environment in which she knows the causal model controlling the relation between actions and outcomes.
-She can use the known causal model in order to find the probabilities of \textit{causing} a desired outcome given she takes a certain action.
-The following theorem is found in Section 4.1 of \cite{pearl2009causality}, but the intuitions that lie behind can be traced back to \cite{lewis1981causal} and \cite{joyce1999foundations}.
-
-Consider a Causal Model $G$ and its associated distribution $P_G$ and let $C$ be the set of consequences of interest for a decision maker. Then,
-\begin{theorem}[Causal von Neumann-Morgenstern Theorem, \cite{pearl2009causality}]{\label{causal_ut}}
-If a rational decision maker faces a Causal Environment and if the causal model is known, then the preference relation $\succeq$ satisfies the von Neumann-Morgenstern rationality axioms if and only if:
-\begin{eqnarray*}
-&a \succeq b&\\
-& \textrm{ if and only if }&\\
- &\sum_{c \in C} P(c | do(a))u(c) \geq \sum_{c \in C} P(c | do(b))u(c).&
-\end{eqnarray*} 
-\end{theorem}
-Equivalently, the action that must be chosen is 
-\begin{equation}
-a^\ast = \textrm{argmax}_{a \in \mathcal{A}} \sum_{c \in C} P(c | do(a))u(c).
-\end{equation}
-\begin{proof}
-This Theorem assumes the existence of a Causal Graphical Model known by the decision maker; then, by Theorem \ref{vNM}, we know that the action that must be chosen is the one that maximizes the expected utility; we must now ask the expected utility of what?
-The answer lies on the Causal Graphical Model which controls the relation between actions and consequences; if this is the case, then the probability of a consequence given an action $a$ is in fact $P(c|do(a))$.
-The rest follows by Theorem \ref{vNM}.
-Details in Section 4.1 of \cite{pearl2009causality}.
-\end{proof}
-
-As shown in the proof of Theorem \ref{causal_ut}, we claim that Pearl's result can be considered as a causal version of the von Neumann-Morgenstern result since it assumes that a causal model is known.
-If the causal model that controls an environment is known to a decision maker, then this is equivalent of being able to know the probabilities of outcomes given actions.
-As stated in Section 4.1 of \cite{pearl2009causality}, the utility function $u$ is considered as given, even though the Theorem \ref{vNM} guarantees its existence.
-Pearl argues that decision making by maximizing such function is \textit{commonsensical} without appealing to the original results in rational decision making.
-We avoid entering in the long-standing debate between \textit{causal} and \textit{evidential} decision making.
-We only note that both von Neumann-Morgenstern's and Savage's Theorems have no {\em a priori} causal interpretation and therefore rely on associative information.
-
-\subsection{A Savage Theorem for causal environments}
-\label{section_savage}
-When the decision maker does not know the causal model which controls her environment we argue that the decision maker is facing a particular case of Savage's Theorem.
-The difference here being that the subjective probability must make use of the causal nature of the environment.
-The idea we will follow is that Savage's Theorem gives the decision maker a subjective quantification of her uncertainty about the environment and the associated probabilities which must be used \textit{as if} they were true.
-
-In this case, where a Causal Graphical Model controls the relation between actions and outcomes, such subjective information about the environment must consider such causal structures.
-For this reason, we assert that the probability distribution that the decision maker has in mind is in fact a distribution over causal structures where the decision maker uses each structure as if it were the true one in order to choose the best action within each structure by using Pearl's result.
-We assume a finite set of actions and outcomes.
-Formally:
-
-\begin{theorem}[Causal Savage Theorem]
-\label{causal_savage}
-In a finite, bounded Causal Decision Problem  $(\mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$, where $\mathcal{G}$ is a Causal Graphical Model, we have that the preferences $\succeq$ of a decision maker are Savage-rational if and only if there exists a probability distribution $P_C$ over a family $\mathcal{F}$ of causal structures such that for $a,b \in \mathcal{A}$:
-\begin{eqnarray*}
-&a \succeq b&\\ 
-&\textrm{ if and only if }&\\
-&\sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right)&\\
-&\geq& \\
-&\sum_{c \in \mathcal{C}}  u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(b))P_C(g) \right)&
-\end{eqnarray*}
-where $P_g$ is the probability distribution associated with the causal structure $g$. 
-\end{theorem}
-
-\begin{proof}
-The decision maker is facing an environment in which any action she takes will stochastically cause an outcome $c \in \mathcal{C}$.
-For this reason, the decision making is facing a very particular case of decision making under uncertainty.
-Assuming Savage-rationality, by Theorem \ref{savage} we obtain a utility function $u^S$ and a probability measure $P^S$ which satisfy that the preference relation is represented by the expectation of $u^S$ with respect to $P^S$.
-
-In such a causal environment, the CGM $\mathcal{G}$ contains all of the information which connects actions, uncertain events and outcomes, and noting that we can identify any action $a$ with $\{ c_j | E_j : j \in J  \}$ where $J$ a countable set of indexes (\cite{bernardo2000bayesian}) we have that:
-\[\mathbb{E}_{P^S}[u(c)] = \sum_{j \in J} u(c_j)P^S(E_j).\]
-For each action $a=\{ c_j | E_j : j \in J \}$, $P^S(E_j)$ is the probability of \textit{causing} consequence $c_j$ by choosing action $a$.
-In order for the decision maker to find the probability of a certain consequence $c_j$ given that an action $a$ is performed then she must have in mind a single causal model $g$ and a way to assign probabilities over a family of causal models; i.e., the uncertainty component $P^S(E_j)$ is formed by two parts: a distribution $P_C$ which represents the degree of belief of the decision maker about a specific model $g$ being the true one, and within $g$, a distribution $P_g$ used to calculate the probability of causing some consequence $c_j$ given that action $a$ is chosen.
-Using the Caratheodory Extension Theorem \citep{ash2000probability}, a probability measure  $P_C$ whose support is a sufficiently general family of causal models $\mathcal{F}$ can be shown to exist.
-For $g \in \mathcal{F}$, the decision maker considers $g$ to be the true causal model with probability $P_C(g)$ and will use Theorem \ref{causal_ut} in order to find the best action \textit{within} the causal graphical model $g$ as if it were the true one.
-Let $P_g$ the probability distribution associated with $g$.
-
-Then:
-\begin{eqnarray*}
-\mathbb{E}_{P^S}[u(c)] &=& \sum_{j \in J} u(c_j)P^S(E_j)\\
-                                      &=& \sum_{j \in J} u(c_j) \left( \sum_{g \in \mathcal{F}} P_g(c_j | do(a))P_C(g) \right).
-\end{eqnarray*}
-We have shown what is the expected utility for some action $a \in \mathcal{A}$, and again by Theorem \ref{savage} the result follows. 
-\end{proof}
-
-\subsection{Interpretation}
-Theorem \ref{causal_savage} says that a decision maker who faces a Causal Decision Problem is considering a probability distribution $P_C$ over a family $\mathcal{F}$ and, within each structure, using the term $P_g(c|do(a))$ in order to find the probability of obtaining a certain consequence given that the intervention $do(a)$ is performed; in this way, the optimal action $a^\ast$ is given by:
-\begin{equation}
-a^\ast = \textrm{ argmax }_{a \in \mathcal{A}}  \sum_{c \in \mathcal{C}} u(c) \left( \sum_{g \in \mathcal{F}} P_g(c | do(a))P_C(g) \right). 
-\end{equation}
-We note that $a^\ast$ is obtained by taking into account the utility obtained by every possible consequences weighted using both the probability of causing such action within a specific causal model $g$ and the probability that the decision maker assign to such $g \in \mathcal{F}$.
-
-We are considering a \textit{normative} interpretation for Theorem \ref{causal_savage} according to which a decision maker must use any causal information in order to obtain the best possible action.
-Such action must be obtained by considering the \textit{beliefs} of the decision maker about the causal relations that hold in her environment (the distribution $P_C$), how such relations could produce the best action when considered \textit{as if} they were true (distribution $P_g$), and the satisfaction (utility $u$) produced by the consequences of actions.
-
-Humans tend to ignore pure probabilistic information over causal information \citep{tversky1980causal}, and in fact are able to learn, and use, causal models in sequential decision making processes \citep{lagnado2007beyond,sloman2006causal,nichols2007decision,meder2010observing,hagmayer2013repeated,wellen2012learning}, even though such learning is not perfect \citep{rottman2014reasoning}.
-Therefore, this theorem provides the basis for a much stronger, and computationally implementable, framework for decision making in which causal information is preferred over associative information, even though complete causal information may not be available to the decision maker.
-
-\section{Application: Causal Games and Nash Equilibrium}
-\label{causal_games}
-In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment.
-We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence.
-With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information.
-Therefore, provide motivation to further extend classical results, which have been shown useful, to use causal information as a basis.
-
-A game is a model of a situation in which several players must take an action and afterwards they will be affected both by the outcome of their own action as well as the actions of the other players \citep{osborne1994course}.
-In a strategic game it is assumed that no player knows the action taken by any other players; we also assume that the causal mechanism, which represented by a Causal Graphical Model $\mathcal{G}$, remains fixed and it is unknown for each player.
-
-In this game, players ignore the actions taken by any other player, and since the causal model which controls the environment is unknown by every player, the players also ignore the information that players will use in order to take their respective actions: strategic games of this type are called \textit{Bayesian Games}, introduced by Harsanyi \citep{harsanyi1967games1,harsanyi1968games2,harsanyi1968games3}.
-In the games we will consider, the uncertainty of every player consists of two levels: on a first level, the true causal model $\mathcal{G}$; on a second level, what an action $do(a)$ causes if a certain Causal Graphical Model $\omega$ is considered to be the causal model.
-\begin{definition}
-A Bayesian strategic game(\cite{osborne1994course}), consists of:
-\begin{itemize}
-\item A finite set $N$ of players.
-\item A finite set $\Omega$ of \textit{states of nature}.
-\item For each player, a nonempty set $A_i$ of actions.
-\item For each player, a finite set $T_i$ and a function $\tau_i : \Omega \mapsto T_i$ the signal function of the player
-\item For each player, a probability measure $p_i$ over $\Omega$ such that $p_i (\tau^{-1}_i (t_i))>0$ for all $t_i \in T_i$.
-\item A preference relation $\succeq_i$ defined over the set of probability measures over $A \times \Omega$ where $A= A_1 \times \cdots A_n$
-\end{itemize}
-\end{definition}
-
-We will consider $\Omega$ to be a family of admissible causal models; in this way,  $\omega \in \Omega$ being the true state of nature fixes a causal model which controls the environment in which the players make their choices.
-In classical Bayesian games, once $\omega \in \Omega$ is realized as the true state, then each player receives a signal $t_i=\tau_i (\omega)$ and the posterior belief $p_i(\omega | \tau^{-1}_i (t_i) )$ given by $p_i(\omega) / p_i (\tau^{-1}_i (t_i))$ if $\omega \in \tau^{-1}_i (t_i)$.
-In the case for causal bayesian games, we must consider both the probability $p_i$ of $\omega$ being the true state as well as the probability $p^\omega_i$ of observing a certain consequence when doing some action $a_i$ if $\omega$ is the true model.
-
-Following \cite{osborne1994course}, we define a new game $G^\ast$ in which its players are all of the possible combinations $(i, t_i) \in N \times T_i$, where the possible actions for $(i,T_i)$ is $A_i$.
-We see that fixing a player $i \in N$, the posterior probability $p(\omega | \tau^{-1}_i (t_i))$ induces a lottery over the pairs $(a^\ast(j,\tau_j(\omega)))_j,\omega)$ for some other $j \in N$.
-This lottery assigns to $(a^\ast(j,\tau_j(\omega)))_j,\omega)$ the probability $p_i(\omega) / p_i (\tau^{-1}_i (t_i))$ if $\omega \in \tau^{-1}_i (t_i)$.
-The classical Bayesian game will simply call a Nash equilibrium for the game $G^\ast$ a Nash equilibrium of the original game; but we have the second level of uncertainty: the consequences caused by some action $a$ through a causal model $\omega$.
-We notice that the posterior probability itself induces a probability distribution defined over \textit{actions} for each player once a \textit{desired consequence} is fixed, this distribution, according to Theorem \ref{causal_savage} is given by $p^\omega_i (c | do(a^\ast_i), a^\ast_{-i}) p_i(\omega | \tau^{-1}_i (t_i))$.
-This motivates the following definition of a \textit{Causal Nash equilibrium}.
-\subsection{Causal Nash Equilibrium}
-For each player $i \in N$ in the strategic game, we define the following probability distribution over consequences:
-\begin{equation}{\label{causal_utility}}
-p^a_i (c) =  p^\omega_i (c | do(a_i), a_{-i}) p_i(\omega)\textrm{ for } a \in A=A_1 \times \cdots \times A_N.
-\end{equation}
-where $p^\omega_i$ is the probability of causing a certain consequence within a causal structure $\omega$ and $p_i$ are the player's \textit{posterior beliefs} about the causal structure that controls the environment, and $do()$ is the well known intervention operator from \cite{pearl2009causality}.
-We now define:
-\begin{equation}
-u^C_i (a) = \sum_{c \in C}  u_i(c) p^a_i (c) \textrm{ for } a \in A=A_1 \times \cdots \times A_N.
-\end{equation}
-Notice that $u^C_i$ evaluates an action profile $a \in A$ in terms of: The knowledge about the causal structure of each player represented by $p_i$, which allows each player to evaluate the probability of causing outcomes in terms of actions by using the $do$ operator as well as the other actions taken by the other players, given by $a_{-i}$ and the preferences of each player $u_i$.
-Using this new function, we define the equilibria for a strategic game with causal information and Bayesian players as:
-\begin{definition}
-A Nash equilibrium for this \textit{causal strategic game} is an action profile $a^\ast \in A$ if and only if
-\begin{equation}
- u^C_i(a^\ast) \geq u^C_i(a_i, a^\ast_{-i}) \textrm{ for any other } a_i \in A_i. 
- \end{equation}
-\end{definition}
-This is, an action profile is a Nash equilibrium if and only if each player uses her current knowledge about the causal structure of the environment in order to (causally) produce the best possible outcome given the actions taken by the other players.
-The existence of the Causal Nash Equilibrium is guaranteed if every $A_i$ is a nonempty compact convex set in some $\mathbb{R}^n$ and if the preference relation induced by $u^C_i$ is continuous and quasi-concave.
-
-\section{Application: Formalizing an optimal action learning method for causal environments}
-\label{optimal_action}
-%lo que estoy planteando complementa estos trabajos que se hicieron de manera empirica
-In recent papers by \cite{lattimoreNIPS2016,sen2017identifying} and \cite{gonzalez2018playing}, a decision maker is given the task to learn an optimal action from a series of learning rounds in a causal environment when the true causal model which controls the environment is unknown for the decision maker.
-\cite{lattimoreNIPS2016} assume that the conditional distributions for each variable given its parents is known, while \cite{sen2017identifying} assume as known only a part of the causal graphical model and allows for interventions in the unknown part of the model.
-\cite{gonzalez2018playing}  assume that the \textit{graphical structure} of such model is known.
-In their method, the decision maker holds beliefs about the parameters of the model, which are used in order to build a causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round.
-
-\cite{gonzalez2018playing}, assume that the graphical structure of the causal model is known and use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action; the best action is defined in terms of the maximum expected utility as required by both the von Neumann-Morgenstern and Savage Theorems.
-Their method fits within our Theorem \ref{causal_savage} because in this paper the distribution $P_C(g)$ is a \textit{distribution over causal models with the same structure}, and when a model is specified from $P_C$, it is then used as if it were the true one in order to find the best action for such causal model.
-In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper.
-
-Such learning method, which has been shown to fit in our causal decision making theorem achieves a similar performance than the classical Q-Learning algorithm with the extra of learning a causal model of a very simple environment.
-Learning a causal model of the environment allows to extract high-level insights of a phenomena beyond associative descriptions of what is observed.
-A causal model is able to \textit{explain} why a particular decision was made since it allows to extract the causes and effects of an agent's actions.
-Once a causal model is acquired, an external user is able to reason about \textit{what...if...} statements that associative methods cannot answer.
-
-When a decision maker chooses an action out of many, a causal model allows to ask what would've happened if another action was taken without actually performing the alternative action.
-It is important to be careful when using the word \textit{explanation} since it allows for several uses in language, here we understand \textit{explanation} as used by \cite{woodward2005making} who considers explanations where what is explained depends on other factors via some relationship that holds as a matter of empirical fact rather than \textit{logical} reasons.
-Also, learning a causal model of a certain environment would allow transferability of knowledge into similar domains because the underlying cause-effect structure has been captured and could later be used.
-We have provided the extension to the causal case of the formal criteria (expected utility) to be used when a causal model is present in the environment, and therefore we hope that our result will be of interest to anyone using decision making under uncertainty.
-
-%Queda la duda de si se necesita especificar un orden previo
-\subsection{A random graph approach to causal structure learning}
-\label{random_graphs}
-
-For the sake of completeness, we now provide a relaxation of the restriction found in \cite{gonzalez2018playing}; in such work, it is required that the graphical structure of the causal model is known, which is a strong assumption.
-
-Human beings focus on \textit{local} aspects while learning causal relations which are later unified into a single structure (\cite{fernbach2009causal}, \cite{waldmann2008causal},  \cite{danks2014unifying}).
-Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning.
-Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.
-
-Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
-Our extension will make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.
-
-We call our method \textit{local} since we use the uncertainty (from the agent's point of view) about the existence of a causal relation between pairs of variables, which are local components of the structure of a graph.
-Such uncertainty is to be updated in terms of what is \textit{observed} from realizations of the true causal mechanism that controls the environment.
-Since observations alone do not suffice to uniquely determine a DAG, we make use of a partial ordering between the variables in the model that, at least in principle, can not be a cause of each other.
-This ordering must be an input product from expert knowledge.
-
-Let a rational agent consider the following set of variables $\mathcal{X}=\{ X_1,...,X_n \}$ which are causally related, even though the agent does not now how; the agent knows that she can only intervene one variable, and does so in order to alter the value of some  identified reward variable; without loss of generality assume that the agent can only intervene on $X_1$ wishing to affect $X_n$.
-Let $p_{ij}$ be the belief that the agent has over a causal relation (directed link) existing between variable with index $i$ and variable with index $j$.
-Let $G$ an initial {\em random} DAG formed as follows: node set is $N=\{1,...,n\}$ and a link exists between $i$ and $j$ with probability $p_{ij}$.
-Now, we use the methodology found in \cite{gonzalez2018playing}  as well as our Theorem \ref{causal_savage} in order to find the best action $a^\ast$ for the obtained graph $G$.
-The best action is taken, and a full realization $X_1=x_1,...,X_n=x_n$ is observed.
-
-Now, we update the $p_{ij}$'s using Bayes Theorem as follows: for each pair of indexes $i,j$ we consider the subgraph containing only $1,i,j,n$ as nodes, either connected or not, and we ask for the probability of such graph producing the output $(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n)$, which will be used as the likelihood of data, and as a prior probability we simply use $p_{ij}$, so we have
-\begin{equation}{\label{bayesian_updating}}
-p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
-\end{equation}
-Then, we update the model generating a new graph according to $p_{ij}$. We do not provide an implementation of Equation \ref{bayesian_updating} as it is not in the scope of this paper.
-
-\section{Limitations}
-\label{limitations}
-We are working within the classical rationality assumption.
-Rationality can be ultimately thought of as a \textit{consistent} or coherent way of making choices, but the precise definition has been a subject of debate.
-See \cite{ellsberg1961risk}, \cite{gilboa2009decision} and \cite{machina2014ambiguity} for critiques of the Savage Rationality Axioms.
-Another line of decision making, from a descriptive point of view, has been developed by psychologists and economists; see \cite{TverskyKahneman74,kahneman1979prospect,kahneman2000choices}.
-
-We are considering Causal Graphical Models as the representation of causal relations in the environment, and we are considering stochastic causal relations according to the manipulationist interpretation, which is one of many.
-We have favoured the Causal Graphical Models over other alternatives since it has been argued that several cognitive processes, such as causal reasoning, can be best represented as graphical models \cite{danks2014unifying,sloman2015causality,hagmayer2016causal}.
-
-\section{Conclusion}
-\label{conclusion}
-We have defined a Causal Decision Problem in terms of a classical Decision Problem under Uncertainty provided of a Causal Mechanism which controls the relation between actions and outcomes, which is represented by a Causal Graphical Model in this work.
-In the case in which a rational decision maker knows such causal relations, \cite{pearl2009causality} provides a causal version of the von Neumann-Morgenstern Theorem for decision making.
-
-On the other hand, when a decision maker does not know the causal mechanism, in Theorem \ref{causal_savage} we have provided a causal version of the Savage Theorem; our result explicitly states how a rational decision maker should use subjective beliefs, encoded as a probability distribution over causal models, as well as the causal inference machinery within the considered causal structures in order to find an optimal action.
-
-Using Theorem \ref{causal_savage} and taking as a basis Harsanyi's model of a Bayesian Game in which every player has incomplete information about both the actions taken by other players as well as the information that made each player take his action we have been able to provide a definition of a Causal Nash Equilibrium in which every player is aware that there exists a Causal Mechanism that will produce some consequence once he takes an action.
-Our decision making result (i.e., Theorem \ref{causal_savage}), besides motivating the Causal Nash Equilibrium, also provides an optimality criterion for learning algorithms in causal settings such as those presented in~\cite{lattimoreNIPS2016,sen2017identifying}.
-Our definition of Causal Equilibrium takes into account classical game theory through the incorporation of the classical von Neumann-Morgenstern utility function as well as the fundamental notion in Causation of Pearl's $do$ operator.
-
-We consider optimal action selection in a causal environment, providing an extension of the formal criteria (expected utility) to be used when a causal model is present in the environment.
-Our results provide a theoretical foundation for previous works that address this problem, considering a known or partially known causal model.
-Additionally, we consider the case in which the structure of the causal model is not known, proposing a novel approach based on Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition.
-
-We are confident that further developments based on our results will show that Causality is a fundamental concept for decision making under uncertainty.
-
-\section*{Declarations}
-
-\section*{Funding}
-The present research has been funded by the Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT).
-
-\section*{Availability of data and material}
-Not applicable
-
-\section*{Code availability}
-Not applicable
-
-\section*{Conflict of interest} 
-The authors declare that they have no conflict of interest.
-
-\section*{Authors' contributions}
-Equal contribution
-
-% BibTeX users please use one of
-\bibliographystyle{apalike}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliographystyle{spbasic}
-\bibliography{Bibliografia.bib}
 \end{document}

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en abstract

@@
 \begin{abstract} 
+Causal reasoning is a fundamental aspect of intelligent reasoning since it allows to consider interventions and counterfactuals.
 Decision making under uncertainty has been well studied when information is considered at the associative (probabilistic) level.
@@
 Causal reasoning is a constant element in our lives as it is in human nature to constantly ask \textit{why} \citep{spirtes2000causation,pearl2018why}.
-Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{waldmann2013causal,danks2014unifying,neil2019causality}.
+Looking for causes is an everyday task and, in fact, causal reasoning is to be found at the very core of our minds \citep{tversky1977causal,waldmann2013causal,danks2014unifying,neil2019causality}.
 It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}.
-Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{tversky1977causal, tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.
+Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.
 

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en intro

@@
 On the other hand, an important aspect of acting in the world is being able to make decisions under uncertain conditions \citep{bernardo2000bayesian,kahneman2000choices,gilboa2009decision,lake2017building}.
-\cite{von1944theory} gave an answer for how to make choices if rational preferences are assumed and if the decision maker knows the stochastic relation between actions and outcomes: maximize expected utility.
+\cite{von1944theory} gave an answer for how to make choices if rational preferences are assumed, utilities are known, and if the decision maker knows the stochastic relation (i.e., probabilities of events) between actions and outcomes: maximize expected utility.
 If no such relation is known, then \cite{savage1954the} showed that a rational decision maker must choose \textit{as if} she is maximizing expected utility using a \textit{subjective} probability distribution.

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en intro

@@
 
-These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms learn \textit{optimal policies} in such a way that any action prescribed by such optimal policy achieves the maximum expected utility \citep{sutton1998reinforcement,webb2007game}.
-In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on associative relations which are expressed using correlations or probability distributions.
+These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms attempt to learn \textit{optimal policies} in such a way that the actions prescribed by such optimal policy achieve the maximum expected utility by satisfying the Bellman Equations \citep{sutton1998reinforcement,webb2007game}.
+In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on \textit{associative} relations which are expressed using patterns found in the data in terms of correlations encoded in probability distributions.
 
-It is a natural question how to formalize rational decision making when causal information is present; answering such question is relevant given the importance of causal relations in everyday life.
-Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage.
+It is a natural question \textit{how to formalize rational decision making when causal information is present}; answering such question is relevant given the importance of causal relations in everyday life as well as in science.
+Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage in order to provide the foundations for Causal Reinforcement Learning.
 
-The previous question has been already considered by \cite{nozick1969newcomb}, \cite{lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making and also by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known by the decision maker.
+The previous question has been already considered by \cite{nozick1969newcomb,lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making as well as by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty \textit{when the causal mechanism which controls the environment is known} by the decision maker.
 

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en applications

@@
 Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent.
-In many real-world applications, the agent doesn't have access to all of the parameters required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover this relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.
+In many real-world applications, the agent doesn't have access to all of the parameters, or even the distribution itself, required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover these relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.
 
@@
 We provide an explicit way of choosing an optimal decision given what the decision maker believes about the causal structure of her environment, thus providing a causal version of Savage's Theorem.
-We take a normative point of view to study how causal relations should be used by a \textit{rational} agent when making decisions with uncertain consequences.
+We take a normative point of view to study how causal relations \textit{should} be used by a rational agent when making decisions which have uncertain consequences.
 
@@
 
-In this work we use the Bayesian Game model in order to study what happens when several decision makers have certain knowledge about an environment which is controlled by some, unknown but fixed, causal mechanism.
-We study games in which the players actions \textit{cause} some consequence according to some unknown causal model.
-We apply our a rational choice criterion to formulate a Nash equilibrium for Causal Games by defining a probability updating for Bayesian Causal Games.
+In this work we use the Bayesian Game model in order to study what happens when several decision makers have certain knowledge about an environment which is controlled by some, unknown but fixed, causal mechanism in the form of a causal graphical model.
+In our model, the players actions' \textit{cause} some consequence according to the unknown causal model.
+We apply our rational choice criterion to formulate a Nash equilibrium for Causal Games.
 

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en organization of the paper

@@
 We consider the case in which the causal model is unknown or partially known to the decision maker, so she has to use beliefs about possible causal models.
-From the decision maker point of view, having in mind a causal model allows one to ask what would have happened if another action was taken without actually performing the alternative action; therefore, being able of choosing the best action in terms of the probabilities of causing a good outcome.
+From the decision maker's point of view, having in mind a causal model allows her to ask \textit{what would have happened if another action was taken} without actually performing the alternative action and thus being able to obtain the best overall action; i.e., being able of choosing the best action in terms of the probabilities of causing a desired outcome.
 Also, learning a causal model of a certain environment will facilitate knowledge transfer to similar domains, as the underlying cause-effect structure has been captured and could later be used.
-
-% añadir motivación del paper y de las teorías
 
@@
 \label{organization}
-This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which use only associative information in their formulation; then, in Section \ref{causal_decision_problems} a brief introduction to Causation and Causal Graphical Models is given.
+This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which rely on associative information in their formulation; then, in Section \ref{causal_decision_problems} we present a brief introduction to Causation and Causal Graphical Models.
 In Section \ref{related_work} we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning.
-In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure, as well as our main result, Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure.
+In Section \ref{causal_results} we state a von Neumann-Morgenstern Theorem for Causal Decision Making that is due to Pearl, which assumes a known causal structure.
+Our main result is Theorem \ref{causal_savage} which extends Pearl's result to the case of an unknown causal structure.
 Finally two applications of Theorem \ref{causal_savage} are described.
@@
 Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works.
-Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider;  and in Section \ref{conclusion} we summarize the main contributions of our work.
+Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider.
+Finally, in Section \ref{conclusion} we summarize the main contributions of our work.
 
@@
 \label{classical_theorems}
-A decision problem under uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors.
+A Decision Droblem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}.
 Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
unifique las limitaciones de info asociativa

@@
 \label{organization}
-This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which rely on associative information in their formulation; then, in Section \ref{causal_decision_problems} we present a brief introduction to Causation and Causal Graphical Models.
+This paper is organized as follows: first, in Section \ref{classical_theorems} we present the classical decision making theorems which rely on associative information in their formulation: Theorems \ref{vNM} and \ref{savage}; then, in Section \ref{causal_decision_problems} we present a brief introduction to Causation and Causal Graphical Models.
 In Section \ref{related_work} we present previous approaches both to Causal Decision Making as well as its applications to interactive optimal action learning: causal reinforcement learning.
@@
 
-When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty, \cite{bernardo2000bayesian}:
+When we consider the preferences of some decision maker over the set of consequences of some uncertain environment we have a Decision Problem under Uncertainty \citep{bernardo2000bayesian}:
 
@@
 This family of conditions is called \textit{von Neumann-Morgenstern rationality axioms} as described in \cite{gilboa2009decision}.
+
 \begin{theorem}[von Neumann-Morgenstern]{\label{vNM}}
@@
 Savage's result extends von Neumann-Morgenstern Theorem since it considers the case in which a rational decision maker does not know neither her utility function nor the probabilities to be used in order to obtain the expected values required for making choices according to the von Neumann-Morgenstern Theorem.
+
 The notation and setting we use here is the one used by \cite{bernardo2000bayesian}.
@@
 We now state Savage's Theorem.
-Details can be found in \citep{kreps1988choice,bernardo2000bayesian,gilboa2009decision}.
+Details can be found in \citep{savage1954the,kreps1988choice,bernardo2000bayesian,gilboa2009decision}.
 
@@
 \end{proof}
-This theorem states that if a rational decision maker does not know the precise probabilities of outcomes given that an action is taken, then she chooses \textit{as if} having in mind a probability assignment to the uncertainties in her environment and using such probabilities to calculate the expected utility with respect to a subjective utility function that represents her preferences.
+This theorem states that if a rational decision maker does not know the precise probabilities of outcomes given that an action is taken, then she must choose \textit{as if} having in mind a probability assignment to the uncertainties in her environment and use such probabilities to calculate the expected utility with respect to a subjective utility function that represents her preferences.
 This result also gives a precise definition of \textit{subjective probability} as a quantification of uncertainty which is used to make good decisions.
@@
 In the next section we state some basic facts on causality and why $P(\cdot |x)$ is different from $P(\cdot | do(x))$ for some value $x$.
+
+One of the fundamental aspects of Causality is the presence of an order and a context; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
+\begin{equation}
+    p(a,t)=p(a|t)p(t)
+\end{equation}
+and
+\begin{equation}
+    p(a,t)=p(t|a)p(a)
+\end{equation}
+which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$. This situation does not happen while considering causal relations.
+
+Also, we recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$.
+If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false, see \cite{pearl2009causality}.
 
@@
 
-\subsection{Further limitations of associative information}
-As we have said, one of the fundamental aspects of Causality is the presence of an order and a context; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
-\begin{equation}
-    p(a,t)=p(a|t)p(t)
-\end{equation}
-and
-\begin{equation}
-    p(a,t)=p(t|a)p(a)
-\end{equation}
-which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$. This situation does not happen while considering causal relations.
-
-Also, we recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$.
-If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false.
 \subsection{Causal Environments and Causal Decision Problems}{\label{causal_problems}}

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en limitations of theorems

@@
 Human beings \textit{know} (this is, use causal relations) that going to the physician does not \textit{cause} being sick and this is what Pearl calls \textit{commonsensical} decision making.
-Also, as we will mention later in the paper, associative information lacks an \textit{ordering} thus failing to determine \textit{if $X$ is performed, then $Y$ will be caused}.
 
-Formally, in this paper we consider the actions chosen by a decision maker as an \textit{interventions} over a causal graphical model; this is, in the classical von Neumann-Morgenstern and Savage Theorems, once a decision maker chooses an action $a$, the observed consequence has a probability $P(c|a)$; if causal information is considered, then the probability of a certain consequence is $P(c|do(a))$.
-In the next section we state some basic facts on causality and why $P(\cdot |x)$ is different from $P(\cdot | do(x))$ for some value $x$.
-
-One of the fundamental aspects of Causality is the presence of an order and a context; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
+We know from \cite{granger1969investigating} and \cite{lamport1978time} that there are two sufficient and necessary conditions for causality: a context and an order.; the latter of such elements not being present in associative information; this is, consider a distribution $p(a,t)$, which we can factorize in the two following ways:
 \begin{equation}
@@
 \end{equation}
-which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$. This situation does not happen while considering causal relations.
+which can be represented indistinguishably as Bayesian Networks as $a \to t$ or as $t \to a$.
+This situation does not happen while considering causal relations.
+Also, associative information lacks an \textit{ordering} thus failing to determine \textit{if $X$ is performed, then $Y$ will be caused}.
+We recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$.
+If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false, see \cite{pearl2009causality}.
 
-Also, we recall the difference between an observation and an intervention: $p(x|y) \neq p(x | do(y))$.
-If this was not the case, then correlation would equal causation which can be shown in a variety of examples to be false, see \cite{pearl2009causality}.
+Formally, in this paper we consider the actions chosen by a decision maker as an \textit{interventions} over a causal graphical model; on the other hand, in the classical von Neumann-Morgenstern and Savage Theorems, once a decision maker chooses an action $a$, the observed consequence has a probability $P(c|a)$; if causal information is considered, then the probability of a certain consequence is $P(c|do(a))$.
+In the next section we state some basic facts on causality and why $P(\cdot |x)$ is different from $P(\cdot | do(x))$ for some value $x$.
 
@@
 \label{on_causality}
-The concept of Causality deals with regularities found in a given environment which are stronger than probabilistic (or associative) relations in the sense that a causal relation allows for evaluating a change in the \textit{consequence} given that a change in the \textit{cause} is performed.
-The \textit{manipulationist} interpretation of Causality (details in \cite{woodward2005making}) is adopted here.
+The concept of Causality deals with regularities found in a given environment (context) which are stronger than probabilistic (or associative) relations in the sense that a causal relation allows for evaluating a change in the \textit{consequence} given a change in the \textit{cause}.
+We adopt here the \textit{manipulationist} interpretation of Causality (details in \cite{woodward2005making}).
 The main paradigm is clearly expressed by \cite{campbell1979quasi} as \textit{manipulation of a cause will result in a manipulation of the effect}.
@@
 
-We know from \cite{granger1969investigating} and \cite{lamport1978time} that there are two sufficient and necessary conditions for causality: a context and an order.
 We will adopt here the formal definition of Causality given by \cite{spirtes2000causation}; i.e., a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}.

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios menores

@@
 \subsection{Preliminary Definitions}
+\label{definitions}
 \begin{definition}
@@
 
-We will adopt here the formal definition of Causality given by \cite{spirtes2000causation}; i.e., a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}.
+We will adopt here the formal definition of Probabilistic Causality given by \cite{spirtes2000causation}; i.e., a stochastic relation between events which is \textit{irreflexive, antisymmetric} and \textit{transitive}.
 Similar descriptions of the manipulationist approach can be found in \cite{holland1986statistics}.
@@
 \subsection{A definition of Causality}
-The formal definition of Causality that will be adopted in this work is the definition provided by \cite{spirtes2000causation} which states that causation is a \textit{stochastic} relation between \textit{events} in a probability space; this is, some event (or events) \textit{causes} another event to occurr.
+The formal definition of Causality that will be adopted in this work is the definition provided by \cite{spirtes2000causation} which states that causation is a \textit{stochastic} relation between \textit{events} within a probability space; this is, some event (or events) \textit{causes} another event to occurr.
 \begin{definition}{\label{causal_relation}}
@@
 We consider decision making with causal information.
-We define a Causal Environment to be an \textit{uncertain environment} such that there exists a Causal Graphical Model (CGM) $\mathcal{G}$ \citep{koller2009probabilistic} which controls the environment; this is, there exists a causal relation between available actions and consequences in the sense that any chosen action will stochastically \textit{cause} a consequence.
+In this section we define a Causal Environment to be an \textit{uncertain environment}, as defined in Section \ref{definitions}, with the extra condition that there exists a CGM $\mathcal{G}$ which controls in the following sense: there exists a causal relation between available actions and consequences in the sense that any chosen action will stochastically \textit{cause} a consequence.
 The role of the CGM is to encode all of the causal relations present in the environment, not only between actions and consequences but also between any other variables in the environment.
-\begin{definition}
+
+\begin{definition}{\label{causal_environment}}
 A Causal Environment is a tuple $(\Omega, \mathcal{A},\mathcal{G},\mathcal{C},\mathcal{E})$ where $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$ is an uncertain environment and $\mathcal{G}$ is a CGM such that the set of variables of $\mathcal{G}$ correspond to the uncertain events in $\mathcal{E}$.
 \end{definition}
+
 \begin{definition}
@@
 
-\cite{heckerman1995decision} provide a framework for defining the notions of cause and effect in terms of decision theoretical concepts, such as states and outcomes and gives a theoretical basis for graphical description of causes and effects, such as the causal influence diagrams from \citep{dawid2002influence}.
+\cite{heckerman1995decision} provides a framework for defining the notions of cause and effect in terms of decision theoretical concepts, such as states and outcomes and gives a theoretical basis for graphical description of causes and effects, such as the causal influence diagrams from \citep{dawid2002influence}.
 Heckerman gave an elegant definition of causality, but did not address how to actually make choices using causal information.
 
-\cite{dawid2012decision} presented a decision-theoretic approach to causal inference in which a decision maker must take into account how alternatives compare against the other in terms of the \textit{average causal effect}, such approach uses the well-known influence diagrams \citep{dawid2002influence,dawid2003causal} in order to derive formulas that allow an explicit calculation of the average causal effect.
+\cite{dawid2012decision} presents a decision-theoretic approach to causal inference in which a decision maker must take into account how alternatives compare against the other in terms of the \textit{average causal effect}, such approach uses the well-known influence diagrams \citep{dawid2002influence,dawid2003causal} in order to derive formulas that allow an explicit calculation of the average causal effect.
 Influence diagrams have the ability of expressing both intervention variables and chance variables into a single graphical structure in such a way that the standard techniques for probabilistic DAG's still apply.
@@
 In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment.
-We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence.
+We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence \cite{shoham2008multiagen,wooldridge2009introductiont}.
 With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information.
@@
 \begin{definition}
-A Bayesian strategic game(\cite{osborne1994course}), consists of:
+A Bayesian strategic game \citep{osborne1994course} consists of:
 \begin{itemize}
@@
 \label{random_graphs}
-
-For the sake of completeness, we now provide a relaxation of the restriction found in \cite{gonzalez2018playing}; in such work, it is required that the graphical structure of the causal model is known, which is a strong assumption.
+For the sake of completeness, we now provide a relaxation of the restriction found in \cite{lattimoreNIPS2016}, \cite{sen2017identifying} and \cite{gonzalez2018playing}: knowing the causal model; in particular, in \cite{gonzalez2018playing} it is required that the graphical structure of the causal model is known, which is a strong assumption.
+Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
+Our extension will take on a purely Bayesian approach and make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.
 
@@
 Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.
-
-Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
-Our extension will make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.
 

### 2020-09-01 3rd_year_eval/Slides/3rd_year_slides.tex
slides 5 6 7 en una sola

@@
 
-\begin{frame}
+\begin{frame}[allowframebreaks]
 \tableofcontents
@@
 \frametitle{Main idea}
-How to learn and use causal information in order to make a good choice in an uncertain environment?
-\end{frame}
-
-\begin{frame}
-\frametitle{Two components}
-We notice two components from the main idea:
 \begin{itemize}
-\item Learn.
-\item Use.
+\item How to learn and use causal information in order to make a good choice in an uncertain environment?
+\item We notice two components from the main idea:
+	\begin{enumerate}
+		\item Learn
+		\item Use
+	\end{enumerate}
 \end{itemize}
-\end{frame}
-
-\begin{frame}
 Previously we focused on the \textit{use} part.
@@
 \item Optimal policies in Reinforcement Learning are the \textit{best} action for each state.
-\item Reinforcement Learning (RL) is \textit{not} a Causal Problem.
+\item Reinforcement Learning (RL) is \textit{not} a Causal Problem \citep{gonzalez2019reinforcement}.
 \end{itemize}
@@
 \begin{itemize}
-\item Mainstream ideas in DMU work at the \textit{associative}\footnote{probabilities, correlations...} level of information.
-\item In fact, Optimal Policies in RL satisfy the classical criterion of Maximum Expected Utility.
+\item Mainstream ideas in DMU work at the \textit{associative} level of information.
+\item In fact, Optimal Policies in RL satisfy the classical criterion of Maximum Expected Utility \citep{webb2007game}.
 \item Given the importance of Causation in science \textbf{we consider decision making theory that relies on causal information.}
@@
 \end{itemize}
+Using a Bayesian Nonparametric prior we could define probabilities over a space of \textit{graphs} and automate what we have done here.
 \end{frame}
@@
 \item Previously, we had addressed the question on how to \textit{use} causal information in order to make good choices. 
+\item We extended the \textit{use} part to a multi-agent setting; Causal Games.
 \item Together, this two tasks answer questions 1, 2 and 3 of the mentioned Research Questions.

### 2020-09-02 3rd_year_eval/Slides/3rd_year_slides.tex
quitando research problem

@@
 
-\subsection{Research Problem}
-\begin{frame}
-\frametitle{Research Problem}
-We recall here the Research Problem described in \cite{gonzalez2019causal}:
-\begin{itemize}
-\item Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
+%\subsection{Research Problem}
+%\begin{frame}
+%\frametitle{Research Problem}
+%We recall here the Research Problem described in \cite{gonzalez2019causal}:
+%\begin{itemize}
+%\item Let  $\mathcal{G}$ a causal graphical model and let $(\mathcal{A},\mathcal{E},\mathcal{C})$ a decision problem under uncertainty whose actions $a_i = \{ c_j | E_j : j \in J \}$  are causally related to consequences $c \in \mathcal{C}$ through the uncertain events $E \in \mathcal{E}$ which correspond to variables in $\mathcal{G}$.
 We ask how a decision maker could \textit{learn} about the causal structure that controls his environment in order to \textit{make good choices} with respect to his preferences.
-\end{itemize}
-\end{frame}
+%\end{itemize}
+%\end{frame}
 
@@
 \begin{itemize}
-\item Using the previous inequality we were able to define the notion of a Causal Game and the idea of a Nash equilibria within such game.
+\item Using the previous inequality we were able to define the notion of a Causal Game and propose a first version of a Nash equilibria within such game.
 We define a function $u^C_i : A \to \mathbb{R}^{+}$, for each player:
@@
 \frametitle{Causal Games: New work}
-We improved the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:
+We continued the work done in the Second Year by modifying Equation \ref{games_causal_ut} in the following way:
 \begin{itemize}
@@
 \item Following this idea, \cite{wellen2012learning} propose a model to explain how observations  and interventions are used by human beings to learn causal relations in terms of a local prediction-error learning.
-\item Following this line of thought, we propose here a local probabilistic encoding of the uncertainty that a decision maker has over the existence or not of causal relations between variables.
+\item We propose here a local probabilistic encoding of the uncertainty that a decision maker has over \textit{the existence or not of causal relations} between variables.
 \end{itemize}
@@
 \item They are random objects which take its values in the space of graphs (or adjacency matrices).
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Random Graphs: most common models}
-\begin{itemize}
-\item We will be using the following model:
-\item Among $N$ vertices, draw an edge between vertices $i$ and $j$ with probability $p_{ij}$.
+\item For example, fix a set of nodes, and connect two nodes with some probability $p$.
 \end{itemize}
@@
 \item Rational agent.
-\item Causal sufficiency
+\item Causal sufficiency.
 \item Knowledge of the \textit{variables} of the model.
@@
 \item Find optimal action $a^\ast$ for graph $G$ and probabilities given by a count of the observations.
-\item Calculate probability of the approximate model $G$ producing the full observation $X_1 = a^\ast, X_2=x_2,...,X_n=x_n$. 
-\item Update $p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.$
+\item Observe the full observation $X_1 = a^\ast, X_2=x_2,...,X_n=x_n$. 
+\item Using current $G$, update $p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.$
 \end{enumerate}

### 2020-09-02 3rd_year_eval/Slides/3rd_year_slides.tex
cambios menores

@@
 \end{frame}
+
 \subsection{Introduction}
 \begin{frame}
-\frametitle{Main idea}
+\frametitle{Introduction: Main idea}
 \begin{itemize}
@@
 \end{itemize}
-Previously we focused on the \textit{use} part.
+Previously we had focused on the \textit{use} part.
 \end{frame}
@@
 \begin{frame}
-\frametitle{Why?}
+\frametitle{Introduction: Relevance}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Why?}
+\frametitle{Introduction: Relevance}
 \begin{itemize}
 \item Mainstream ideas in DMU work at the \textit{associative} level of information.
-\item In fact, Optimal Policies in RL satisfy the classical criterion of Maximum Expected Utility \citep{webb2007game}.
+\item In fact, Optimal Policies in RL satisfy the classical (associative) criterion of Maximum Expected Utility \citep{webb2007game}.
 \item Given the importance of Causation in science \textbf{we consider decision making theory that relies on causal information.}
+\end{itemize}
+\end{frame}
+
+\subsection{Some Background}
+\begin{frame}
+\frametitle{Some Background}
+\begin{itemize}
+\item Rationality.
+\item Decision Theory: rational decision making if and only if utility maximization
+\item Causality: Interventions. Context and order.
 \end{itemize}
@@
 \begin{itemize}
-\item Rationality $+$ probabilistic information $=$ maximize expected utility.
+\item Rationality $+$ probabilistic information either known or unknown $=$ maximize expected utility.
 \item Rationality $+$ \textbf{known} causal information $=$ maximize expected utility considering what is caused by actions \citep{pearl2009causality,lattimoreNIPS2016}
-\item Rationality $+$ \textbf{unknown} causal information $=$ ¿? 
+\item Rationality $+$ \textbf{unknown} causal information $=$ \cite{gonzalez2019theorems} 
 \end{itemize}
+\end{frame}
+
+\subsection{Research Questions}
+\begin{frame}
+\frametitle{Research Questions}
+We recall the Research Questions stated in \cite{gonzalez2019causal}:
+\begin{enumerate}
+\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
+\item How can a causal structure help a decision maker in order to guide his learning process?
+\item What does the rationality assumption implies about how to choose when considering causal information? 
+\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
+\end{enumerate}
 \end{frame}
@@
 \textit{...to provide the understanding and knowledge about what are the implications of causality for a rational decision maker who faces an uncertain environment and about how causal relations can be discovered and used in order to make good choices that maximize the expected utility for the decision maker...}
-\end{frame}
-
-\subsection{Research Questions}
-\begin{frame}
-\frametitle{Research Questions}
-We recall the Research Questions stated in \cite{gonzalez2019causal}:\\
-The proposed research is ultimately trying to answer a series of questions about the nature of causal relations and about their possible use.
-\begin{enumerate}
-\item How a rational decision maker who faces an uncertain environment which is governed by a causal mechanism can learn and make use of this causal structure in order to make good choices?
-\item How can a causal structure help a decision maker in order to guide his learning process?
-\item What does the rationality assumption implies about how to choose when considering causal information? 
-\item How to trade off exploration and exploitation when trying to learn about the causal structure of an environment while also trying to make good choices?
-\end{enumerate}
 \end{frame}
@@
 %\end{frame}
-
-\subsection{Some Background}
-\begin{frame}
-\frametitle{Some Background}
-\begin{itemize}
-\item Rationality.
-\item Decision Theory: rational decision making if and only if utility maximization
-\item Causality: Context and direction. Interventions. 
-\end{itemize}
-\end{frame}
 
@@
 \begin{itemize}
-\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility \citep{gonzalez2018playing}.
-\item Poster presented at Causality Workshop ICML 2018 at Stockholm, Sweden.
+\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility.
+\item \cite{gonzalez2018playing} Presented as a poster at Causality Workshop ICML 2018 at Stockholm, Sweden.
 \item Proposal Defense (27-Nov-2018).
@@
 \item We also studied causal-structure learning via interaction with a causally controled environment using a purely Bayesian approach based on random graphs.
+\item Experiments
 \end{itemize}
@@
  \end{equation}
-\item This work, \cite{gonzalez2019games}, was presented at a MICAI 2019 Workshop.
+\item This work, \cite{gonzalez2019games}, was orally presented at a MICAI 2019 Workshop.
 \end{itemize}

### 2020-09-03 3rd_year_eval/Slides/3rd_year_slides.tex
ultimos detalles

@@
 \item In fact, Optimal Policies in RL satisfy the classical (associative) criterion of Maximum Expected Utility \citep{webb2007game}.
-\item Given the importance of Causation in science \textbf{we consider decision making theory that relies on causal information.}
+\item Given the importance of Causation in science\footnote{\cite{lake2017building,pearl2018why}} \textbf{we consider decision making theory that relies on causal information.}
 \end{itemize}
@@
 \item Rationality.
-\item Decision Theory: rational decision making if and only if utility maximization
+\item Decision Theory: rational decision making if and only if expected utility maximization
 \item Causality: Interventions. Context and order.
@@
 \begin{itemize}
-\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility.
+\item Using Equation \ref{first_version}, we devised a learning algorithm which assumed known the causal structure and learned both the parameters of the causal model of the environment, and an optimal action, yielding a performance similar to classic Reinforcement Learning algorithms which do not use causal information, but rather rely on the maximization of expected utility based on associative information.
 \item \cite{gonzalez2018playing} Presented as a poster at Causality Workshop ICML 2018 at Stockholm, Sweden.
@@
 \begin{frame}
-\frametitle{Causal-Structure Learning: A Random Graph Approach}
+\frametitle{Causal-Structure Learning: A purely Bayesian + Random Graphs Approach}
 \begin{itemize}

### 2020-09-10 SIGecom/Mauricio_Gonzalez_Profile.tex
profile for sigecom

\documentclass[acmsigecom]{acmtrans2m_nocategories}

\usepackage{enumitem,amsmath,amsfonts,setspace}
\usepackage{graphicx}
\usepackage{makeidx}
\makeindex

\usepackage{savesym}
\savesymbol{pdfbookmark}
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}
\restoresymbol{HR}{pdfbookmark}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\l@section}
  {\hfil}
  {\leaders\hbox{\normalfont$\m@th\mkern \@dotsep mu\hbox{.}\mkern \@dotsep mu$}\hfill}
  {}{}
\makeatother

\makeatletter
\def\paragraph{\@startsection{paragraph}{4}%
  \z@\z@{-\fontdimen2\font}%
  {\normalfont\itshape}}
\makeatother

\markboth{}{Job Market Candidate Profiles 2021}

\title[Job Market Candidate Profiles 2021]{SIGecom Job Market Candidate Profiles 2021}

\begin{abstract} 
This is the sixth annual collection of profiles of the junior faculty job market candidates 
of the SIGecom community. The (number TBA) candidates for 2021 are listed alphabetically and 
indexed by research areas that define the interests of the community. The candidates can be 
contacted individually or via the moderated mailing list ecom-candidates2021@acm.org.\\

--Vasilis Gkatzelis and Jason Hartline
\vspace{-15pt}
\end{abstract}

\category{}{}{}
\setcounter{secnumdepth}{0}

\begin{document}

\maketitle

\tableofcontents
\newpage

% Replace CandidateFirstName and CandidateLastName with your first and last name.
% Choose some keywords (as many as you like, as long as it all fits in one line
% under the table of contents (which is, roughly, up to 70 characters).
% Replace the CandidateHomepage_URL and the CandidateCV_URL with the URL to your
% homepage and CV respectively. You can also add links to your Google scholar
% profile using the commented-out link. This is recommended, but not required.
\section[Mauricio Gonzalez-Soto\\ 
{\small keyword1, keyword2, keyword3, keyword4}]
{Mauricio Gonzalez-Soto 
(\href{https://www.researchgate.net/profile/Mauricio_Gonzalez-Soto}{Homepage},
\href{CandidateCV_URL}{CV}
%\href{CandidateGoogleScholar_URL}{Google Scholar}
)}

% Replace ThesisTitleCapitalized with the title of your thesis (if this is not finalized yet,
% then just choose some tentative one) and add the information of your advisor(s).
\textbf{Thesis:} ThesisTitleCapitalized

\vspace{3pt}
\noindent\textbf{Advisor:} Luis Enrique Sucar and Hugo J. Escalante, Instituto Nacional de Astrofísica Óptica y Electrónica.

% Distribute the remaining space however you like, but make sure that your whole profile fits
% within a page once you compile.
\vspace{3pt}
\noindent\textbf{Brief Biography:}
Mauricio received an Applied Mathematics Bachelors Degree from the Instituto Tecnológico Autónomo de México (ITAM) as well as a Masters Degree in Data Science from the same university, where he got awarded a Special Mention for each of his theses, he is currently studying a PhD in Computer Science under the supervision of Luis Enrique Sucar and Hugo J. Escalante.

% In your research summary, you can cite papers that are listed in the itemized environment below
% using static references: [1], [2],... Make sure that you cite the right paper if you later change
% the order in which your papers appear.
\vspace{3pt}
\noindent\textbf{Research Summary:}
Mauricio is mainly interested in Causality and its implications for Decision Making under Uncertainty

% In the itemized section below, you can list however many paper you like, but please use
% the same format that we have used in previous years for each publication. If you are not
% sure, just send an email to gkatz@drexel.edu
\vspace{3pt}
\noindent\textbf{Representative Papers:}
\begin{enumerate}[label={[\arabic*]}]
\item Causal Structure Learning using Random Graphs (Work in Progress: to be submitted to The 2020 NeurIPS Workshop on Causal Discovery and Causality-Inspired Machine Learning)\\
with Ivan Feliciano Avelino, L. Enrique Sucar and Hugo J. Escalante 
\item Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making (Submitted to Artificial Intelligence Journal and oral presentation at New in ML Forum co-located with NeurIPS 2019) \href{https://arxiv.org/abs/1907.11752}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante 
\item Causal Games and Causal Nash Equilibrium (1st Causal Reasoning Workshop at the Mexican International Conference on Artificial Intelligence 2019) \href{https://arxiv.org/abs/1910.06729}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante 
\item Playing against Nature: causal discovery for decision making under uncertainty (Machine Learning for Causal Inference, Counterfactual Prediction and Autonomous Action (CausalML) Workshop at ICML 2018) \href{https://arxiv.org/abs/1807.01268}{Link to Pre-print}\\
with L. Enrique Sucar and Hugo J. Escalante

\end{enumerate}

\printindex

\end{document}

### 2020-09-10 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
draft para AIJ

--- 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
@@
-\RequirePackage{fix-cm}
+\documentclass[review]{elsarticle}
 
-\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-%
-%
-%
-\usepackage{mathptmx}      
-\usepackage[utf8]{inputenc}
-\usepackage[pdftex]{color,graphicx}
-\usepackage{listings}
-\usepackage{breakcites}
-\usepackage{dsfont}
-\usepackage{hyperref}
+\usepackage{lineno,hyperref}
+\usepackage{amsfonts,amssymb,amsthm,amsmath,latexsym}  
 \usepackage{natbib}
 \usepackage{amssymb,amsmath,latexsym}
-\usepackage[margin=1.5cm]{geometry}
-\usepackage{enumitem}
-%\bibliographystyle{stylename}
-%\usepackage{fancyhdr}
-%\pagestyle{fancy}
-\usepackage{tikz}
-\usetikzlibrary{trees}
-\usetikzlibrary{calc}
-\usepackage{color}
-\usepackage{todonotes}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
-% \journalname{myjournal}
-%
+\modulolinenumbers[5]
+\newtheorem{theorem}{Theorem}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{obs}[theorem]{Observation}
+\newtheorem{lem}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corolary}
+\journal{Artificial Intelligence Journal}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
+
+\begin{frontmatter}
 
@@
 
-%\titlerunning{Short form of title}        % if too long for running head
-%
-%\author{Mauricio Gonzalez-Soto         \and
-%        Luis E. Sucar \and
-%        Hugo J. Escalante
-%}
-\author{Anonymous}
-%\authorrunning{Short form of author list} % if too long for running head
+%% Group authors per affiliation:
+\author{Mauricio Gonzalez Soto}
+\author{Luis E. Sucar Succar}
+\author{Hugo J. Escalante Balderas}
+\address{Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica, Luis Enrique Erro 1, Santa Maria Tonanzintla, Mexico}
 
-%\institute{Mauricio Gonzalez-Soto \at
-%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              Mexico\\
-%              \email{mauricio@inaoep.mx}           \\
-%              \url{https://orcid.org/0000-0003-2668-9013}        
-%           \and
-%           Luis E. Sucar \at
-%              Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              M\'exico\\
-%              \email{esucar@inaoep.mx}\\
-%              \url{https://orcid.org/0000-0002-3685-5567}
-%           \and
-%           Hugo J. Escalante \at
-%           Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica \\
-%              Luis Enrique Erro 1, Santa Maria Tonanzintla, Puebla\\
-%              M\'exico\\
-%              \email{hugojair@inaoep.mx}\\
-%              \url{https://orcid.org/0000-0003-4603-3513}
-%}
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-\maketitle
-% modificar abstract para reflejar cambios en el paper
 \begin{abstract}
@@
 These results highlight the importance of causal models in decision making and the variety of potential applications.
+\end{abstract}
 
-\keywords{Causality \and Decision Making \and Game Theory}
-% \PACS{PACS code1 \and PACS code2 \and more}
-% \subclass{MSC code1 \and MSC code2 \and more}
-\end{abstract}
-\tableofcontents
-\newpage
+\begin{keyword}
+Causality \sep Decision Making \sep Game Theory
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+
+%\tableofcontents
 \section{Introduction}
@@
 
-%For example, as told by \cite{bornstein2016artificial}, at the University of Pittsburgh Medical Center a team of researchers tried to use Machine Learning to predict whether pneumonia patients might develop severe complications.
-For this purpose they trained Neural Networks and Decision Trees using the hospital's own data.
-Neural Networks outperformed Decision Trees, but only by studying the decisions made by the latter the doctors did find out that the algorithms instructed doctors to send home pneumonia patients who already had asthma, despite the fact that asthma sufferers are known to be extremely vulnerable to complications.
-The problem relied in the training data, because the hospital policy was to automatically send asthma sufferers with pneumonia to intensive care, and this policy worked so well that asthma sufferers almost never developed severe complications.
-It was only through the interpretability of the Decision Trees that the doctors didn't send asthma patients with pneumonia home to a certain death.
-
-%Methods based in Deep Neural Networks aren't supposed to explain why a certain output was produced since those methods are based in \textit{parallel distributed representations} and the same goes for any other learning algorithm that uses Deep Neural Networks, like Deep Reinforcement learning; when AlphaGo \citep{silver2017mastering} defeated the world-champion, humanity didn't learn anything new about the game of Go, because the algorithm was not designed to explain its moves, although it was a very sophisticated method for achieving it.
-Using only associations between variables it is not possible to infer which one is the cause and which one the effect, something extra is required.
-The fundamental difference between causal and associative models is that causal models consider \textit{doing} over \textit{observing} \citep{pearl2018why}.
-Once a causal relation has been established between two events, performing, or intervening, over a cause allows one to predict the outcome in a more robust way than if using only correlations between observations.
-
 Learning a causal model of an environment and using it to act upon the environment allows us to \textit{explain} aspects of the model that a purely associative model would not be able to explain.
@@
 Our work provides the foundations of decision making algorithms which are based on causal relations and, as will be shown, are able to learn causal models of the environment.
-
-%The impact of Causal Inference goes beyond Machine Learning and Computer Science.
-For example, economists are interested in understanding what did certain public policy caused \citep{athey2017impact} or how human decision makers that show \textit{inconsistent} preferences over time could be oriented if the causal consequences of this inconsistent behavior is introduced into a decision-making model \citep{albers2016motivating}.
-In complex adaptive systems, causal relations can be used to clarify the complex interactions between agents in the system \citep{abbott2017complex} as well as for prediction and planning \citep{hunt2016ants,brock2018causality}.
-
-%cerrar esta sección sobre cómo mi trabajo impacta en la motivación
 
@@
 Then, in Section \ref{optimal_action} we describe some optimal-action learning methods for causal environments which are able to learn an \textit{optimal intervention} under different assumptions and conditions; in particular, we describe and provide an extension for one of such works.
-Finally, in Section \ref{limitations} we comment on the limitations of the assumptions we consider.
+In Section \ref{limitations} we comment on the limitations of the assumptions we consider.
 Finally, in Section \ref{conclusion} we summarize the main contributions of our work.
@@
 
-\section*{Declarations}
+\section*{References}
+\bibliographystyle{model5-names}\biboptions{authoryear}
+\bibliography{mybibfile}
 
-\section*{Funding}
-The present research has been funded by the Consejo Nacional de Ciencia y Tecnolog\'ia (CONACyT).
-
-\section*{Availability of data and material}
-Not applicable
-
-\section*{Code availability}
-Not applicable
-
-\section*{Conflict of interest} 
-The authors declare that they have no conflict of interest.
-
-\section*{Authors' contributions}
-Equal contribution
-
-% BibTeX users please use one of
-\bibliographystyle{apalike}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliographystyle{spbasic}
-\bibliography{Bibliografia.bib}
 \end{document}
-% end of file template.tex

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambio en nombre del doc
cambios en la intro

@@
 %% Group authors per affiliation:
-\author{Mauricio Gonzalez Soto}
-\author{Luis E. Sucar Succar}
-\author{Hugo J. Escalante Balderas}
-\address{Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica, Luis Enrique Erro 1, Santa Maria Tonanzintla, Mexico}
+\author[1]{Mauricio Gonzalez-Soto}
+\ead{mauricio@inaoep.mx}
+
+\author[1]{L. Enrique Sucar}
+\ead{esucar@inaoep.mx}
+
+\author[1]{Hugo J. Escalante}
+\ead{hugojair@inaoep.mx}
+
+\address[1]{Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica ,Luis Enrique Erro 1, Santa Maria Tonanzintla, M\'exico.}
 
@@
 It has been argued that the brain is a causal inference machine which uses \textit{effects} to figure out \textit{causes} in order to engage with the world \citep{friston2010free,clark2015surfing}.
-Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal, lagnado2007beyond, hagmayer2008causal, hagmayer2009decision, hagmayer2017causality}.
+Furthermore, acting in the world is conceived by human beings as \textit{intervening} the world and in fact humans are able to learn and use causal relations while making choices \citep{ tversky1980causal,lagnado2007beyond,hagmayer2008causal,hagmayer2009decision,hagmayer2017causality}.
 
-Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent, located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward.
-Optimal policies, which are prescriptions of which action must be taken at some state $s$, satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem.
-Such solution is equivalent to maximizing \textit{expected utility} \citep{webb2007game}, which is the decision criterion prescribed by the most commonly used decision-making theories: von Neumann-Morgenstern's and Savage's.
+Reinforcement Learning (RL) \citep{sutton1998reinforcement} is the standard framework for automated learning by interaction, in which an agent located at some time $t$ and some state $s$ has to make a choice in order to maximize her long-run discounted accumulated reward.
+Optimal policies are prescriptions of which action must be taken at some state $s$ that satisfy the Bellman Equations \citep{Puterman:1994:MDP:528623} and are the standard solution to the problem.
+Such solution is equivalent to maximizing the \textit{expected utility} \citep{shoham2008multiagent,webb2007game}.
+Maximization of expected utility is the criterion prescribed by the most common used decision-making theories: von Neumann-Morgenstern's and Savage's.
 Such theories rely on \textit{associative} information, which is based on patterns found in data, rather than stronger causal relations.

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambios menores en motivation

@@
 In many real-world applications, the agent doesn't have access to all of the parameters, or even the distribution itself, required to calculate the maximum utility, but if the agent knew that his actions and the possible consequences were \textit{causally related}, then she could attempt to discover these relations and \textit{use them} in order to predict consequences of actions better than if he only observes multiple action-outcome pairs as done in Reinforcement Learning \citep{sutton1998reinforcement}.
+As it is mentioned by \cite{pearl2009causality}, avoiding going to the doctor's office would be the best decision in order to avoid getting sick.
 
 Many real-world applications of decision making are solved by \textit{associative} methods which capture only statistical patterns that are found in data.
-For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, they cannot explain \textit{why} a specific trajectory was chosen by the algorithm.
+For example, current methods in Reinforcement Learning, and specifically in Deep Reinforcement Learning, although they have good performance in the task that were supposed to solve, such methods were not designed to explain \textit{why} a specific trajectory was chosen.
 This is highly relevant in real-world applications, e.g.
-self-driving cars, where it is very important to understand why an accident happened.
+self-driving cars, where it is very important to understand, for example, why an accident happened.
 
@@
 \subsection{Aim}
-Our goal is to establish the foundations for decision making algorithms that rely on the existence of causal relations even though such relations are not previously known or fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying, gonzalez2018playing}.
+Our goal is to establish the foundations for decision making algorithms that rely on causal relations even though such relations are not previously known nor fully observable to a decision maker such as the algorithms found, for example, in \citep{bareinboim2015bandits,lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}.
 We provide here an easily implementable decision making criterion which we build over known results from von Neumman and Morgenstern, Savage and Pearl.

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambios de redaccion en main result

@@
 \subsection{Main Result}
-In this work we extend Pearl's result to the case where a decision maker \textit{does not know the causal mechanism} that controls her environment, so she holds \textit{beliefs} about possible causal models and uses such beliefs \textit{as if} they were true, as prescribed by \cite{joyce1999foundations}, in order to attempt to make a good choice given her beliefs.
+In this work we extend a result from \cite{pearl2009causality} about decision making using known causal information to the case in which a decision maker \textit{does not know the causal mechanism} that controls her environment.
+In our result, the decision maker holds \textit{beliefs}, encoded as a probability distribution over possible causal models, and uses such beliefs \textit{as if} they were true, as prescribed by \cite{joyce1999foundations} and using Pearl's result, in order to attempt to make a good choice given her current beliefs.
 We provide an explicit way of choosing an optimal decision given what the decision maker believes about the causal structure of her environment, thus providing a causal version of Savage's Theorem.

### 2020-09-15 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
Application: Formalizing an optimal action learning method for causal environments

@@
 In Section \ref{on_causality} we define Causality and state the differences between causal and associative information; having shown this, we state two of the most important decision making results, the von Neumann-Morgenstern and Savage Theorems, which rely on purely associative information; then, our objective is to present two analogous theorems: J.
-Pearl's result, Theorem \ref{causal_ut} which relies on a known causal model, and our main contribution: Theorem \ref{causal_savage}, which is able to use a set of \textit{beliefs} about possible causal models (i.e., a distribution) instead of a known one.
+Pearl's result, Theorem \ref{causal_ut} which relies on a known causal model, and our main contribution: Theorem \ref{causal_savage}, which is able to use a set of \textit{beliefs} about possible causal models (i.e., a distribution) instead of a known model.
 
@@
 
-We will consider $\Omega$ to be a family of admissible causal models; in this way,  $\omega \in \Omega$ being the true state of nature fixes a causal model which controls the environment in which the players make their choices.
+We will consider $\Omega$ to be a family of admissible causal models; in this way,  $\omega \in \Omega$ being the true state of Nature fixes a causal model which controls the environment in which the players make their choices.
 In classical Bayesian games, once $\omega \in \Omega$ is realized as the true state, then each player receives a signal $t_i=\tau_i (\omega)$ and the posterior belief $p_i(\omega | \tau^{-1}_i (t_i) )$ given by $p_i(\omega) / p_i (\tau^{-1}_i (t_i))$ if $\omega \in \tau^{-1}_i (t_i)$.
@@
 \label{optimal_action}
-%lo que estoy planteando complementa estos trabajos que se hicieron de manera empirica
-In recent papers by \cite{lattimoreNIPS2016,sen2017identifying} and \cite{gonzalez2018playing}, a decision maker is given the task to learn an optimal action from a series of learning rounds in a causal environment when the true causal model which controls the environment is unknown for the decision maker.
+In recent papers by \cite{lattimoreNIPS2016,sen2017identifying} and \cite{gonzalez2018playing}, a decision maker is given the task to learn an optimal action from a series of learning rounds in a causal environment when the causal model which controls the environment is unknown to the decision maker.
 \cite{lattimoreNIPS2016} assume that the conditional distributions for each variable given its parents is known, while \cite{sen2017identifying} assume as known only a part of the causal graphical model and allows for interventions in the unknown part of the model.
-\cite{gonzalez2018playing}  assume that the \textit{graphical structure} of such model is known.
-In their method, the decision maker holds beliefs about the parameters of the model, which are used in order to build a causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round.
+Meanwhile, \cite{gonzalez2018playing} assume that the \textit{graphical structure} of the true model is known.
 
+In the method described in \cite{gonzalez2018playing}, the decision maker holds probabilistic beliefs about the parameters of the model, which are used in order to build a \textit{local} causal model that is used \textit{as if} it were the true one in order to obtain the \textit{best action} in each round.
 \cite{gonzalez2018playing}, assume that the graphical structure of the causal model is known and use a Dirichlet distribution in order to sample the Conditional Probability Tables required to fully specify a Causal Graphical Model; once such model is specified the authors use an expression derived from Pearl's do-calculus in order to find a best action; the best action is defined in terms of the maximum expected utility as required by both the von Neumann-Morgenstern and Savage Theorems.
@@
 In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper.
+Our result Theorem \ref{causal_savage} formalizes their method, thus providing solid ground for learning algorithms which rely on causal information.
 

### 2020-09-16 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
ultima version

@@
 \label{random_graphs}
-For the sake of completeness, we now provide a relaxation of the restriction found in \cite{lattimoreNIPS2016}, \cite{sen2017identifying} and \cite{gonzalez2018playing}: knowing the causal model; in particular, in \cite{gonzalez2018playing} it is required that the graphical structure of the causal model is known, which is a strong assumption.
-Such assumption is to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
+For the sake of completeness, we now provide a relaxation of the restrictions found in \cite{lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}: knowing the causal model, or some parts of it; in particular, we have mentione that in \cite{gonzalez2018playing} it is required that the graphical structure of the causal model is known, which is a strong assumption.
+Such assumption is now to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
 Our extension will take on a purely Bayesian approach and make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.
@@
 \begin{equation}{\label{bayesian_updating}}
-p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{subgraph formed by nodes 1,i,j,n})p_{ij}^t.
+p_{ij}^{t+1} \propto p(X_1 = a^\ast, X_i = x_i, X_j=x_j, X_n=x_n | \textrm{current graph})p_{ij}^t.
 \end{equation}
@@
 We have favoured the Causal Graphical Models over other alternatives since it has been argued that several cognitive processes, such as causal reasoning, can be best represented as graphical models \cite{danks2014unifying,sloman2015causality,hagmayer2016causal}.
+
+\section{Future Work}
+In Section \ref{random_graphs} we considered the problem of a decision maker situated in a causal environment who does not fully knows the causal model which controls her environment; previous works have considered different levels of uncertainty over the causal model: knowing conditional distributions, knowing a part of the model, knowing the structure of the model or the variables involved.
+We leave as future work the case in which the variables of the model are not known and we consider that this scenario could be attacked using tools from Bayesian Nonparametrics \citep{hjort2010bayesian,phadia2013prior} in the form of a prior over causal models, which generates causal graphs of distinct sizes and complexity.
 

### 2020-09-16 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
añadiendo que permite contrafactuales

@@
 \subsection{Motivation}
+\label{Motivation}
 Intelligent agents often face situations in which an action must be chosen in the presence of uncertain conditions; this means that an outcome will be observed according to some probability distribution given the action chosen by the agent.
@@
 It allows one to ask \textit{why}.
+On the other hand, a causal model allows to \textit{imagine} what would have happened if another course of action had been taken, this is known as counterfactual reasoning \citep{pearl2018why}.
 Since a causal model, once it is learned, does not depend neither on the agent nor its preferences, the gained knowledge could be easily transferred for problems of similar domain where the use of the acquired causal relations is useful.

### 2020-09-16 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
cambios menores

@@
 \label{classical_theorems}
-A Decision Droblem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}.
+A Decision Problem under Uncertainty is the mathematical model of a situation in which an agent must choose one out of many available actions with uncertain consequences which depend on different, possibly unknown, factors \citep{bernardo2000bayesian,gilboa2009decision}.
 Such consequences are ordered in terms of the \textit{satisfaction} that they produce in the decision maker, and such ordering is represented by a \textit{preference relation} denoted by $\succeq$, where $a \succeq b$ is read as $a$ being preferred to $b$.
@@
 In this section we know consider an application in the domain of game theory: we consider a \textit{strategic game} between $N$ rational players who are situated in a causal environment.
-We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence \cite{shoham2008multiagen,wooldridge2009introductiont}.
+We explore the implications of causal decision making in game theory given the importance of multi-agent systems for computational intelligence \citep{shoham2008multiagent,wooldridge2009introduction}.
 With this contribution we expect to show that standard notions of game theory such as Nash Equilibrium can be extended to the case in which causal information is considered over associative information.
@@
 In the above mentioned paper, the possible outcomes are only $1$ and $0$, so once a model $g$ is sampled, the product $u(c)P_g(c|do(a))$ simply becomes $P(C=1 | do(a))$ which is stated only as an intuitive solution in the mentioned paper.
-Our result Theorem \ref{causal_savage} formalizes their method, thus providing solid ground for learning algorithms which rely on causal information.
+Our result, Theorem \ref{causal_savage}, formalizes their method, thus providing solid ground for learning algorithms which rely on causal information.
 
@@
 \label{random_graphs}
-For the sake of completeness, we now provide a relaxation of the restrictions found in \cite{lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}: knowing the causal model, or some parts of it; in particular, we have mentione that in \cite{gonzalez2018playing} it is required that the graphical structure of the causal model is known, which is a strong assumption.
+For the sake of completeness, we now provide a relaxation of the restrictions found in \cite{lattimoreNIPS2016,sen2017identifying,gonzalez2018playing}: knowing the causal model, or some parts of it; in particular, as stated before in \cite{gonzalez2018playing}, it is required that the graphical structure of the causal model is known, which is a strong assumption.
 Such assumption is now to be relaxed to the case in which the graphical structure of the causal model is unknown, but the variables of the model are known; also, we assume the decision maker has an ordering $\gg$ according to which if $A \gg B$, then $B$ can not cause $A$.
-Our extension will take on a purely Bayesian approach and make use both of Theorem \ref{causal_savage}, random graphs and the \textit{as if} intuition present in this paper.
+Our extension will take on a purely Bayesian approach and make use both, Theorem \ref{causal_savage} and random graphs as well as the \textit{as if} intuition present in this paper.
 
@@
 \section{Future Work}
-In Section \ref{random_graphs} we considered the problem of a decision maker situated in a causal environment who does not fully knows the causal model which controls her environment; previous works have considered different levels of uncertainty over the causal model: knowing conditional distributions, knowing a part of the model, knowing the structure of the model or the variables involved.
-We leave as future work the case in which the variables of the model are not known and we consider that this scenario could be attacked using tools from Bayesian Nonparametrics \citep{hjort2010bayesian,phadia2013prior} in the form of a prior over causal models, which generates causal graphs of distinct sizes and complexity.
+We have left out the implementation of the algorithm stated in Section \ref{random_graphs} we considered the problem of a decision maker situated in a causal environment who does not fully knows the causal model which controls her environment; previous works have considered different levels of uncertainty over the causal model: knowing conditional distributions, knowing a part of the model, knowing the structure of the model or the variables involved.
+We leave as future work the case in which the variables of the model are not known and we consider that this scenario could be attacked using tools from Bayesian Nonparametrics \citep{hjort2010bayesian,phadia2013prior} in the form of a prior over causal models.
+If only a part of the set of variables is known, then a Bayesian Nonparametric prior, which generates causal graphs of distinct sizes and complexity, could capture the patterns of causal interaction between known and possible but unknown variables.
 

### 2020-09-18 Artificial_Intelligence_Journal/Draft/AIJ_Draft.tex
version final

@@
 
-\address[1]{Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica ,Luis Enrique Erro 1, Santa Maria Tonanzintla, M\'exico.}
+\address[1]{Coordinaci\'on de Ciencias Computacionales, Instituto Nacional de Astrof\'isica \'Optica y Electr\'onica, Luis Enrique Erro 1, Santa Maria Tonanzintla, M\'exico.}
 
@@
 If an event $A$ is caused by some other event, then we say it is \textit{endogenous} and we denote the set of its causes as $Pa(A)$.
-It is proven in \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.
+It is proven by \cite{kiiveri1984recursive} that at least one exogenous node exists in a causal graph.
 
@@
 Also, we consider an extra pair of conditions.
-The first one, known as Causal Sufficiency, is about the \textit{nature} of the model: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ \textit{outside} of the model $\mathcal{G}$\citep{spirtes2000causation,pearl2009causality,sucar2015probabilistic}.
+The first one, known as Causal Sufficiency, is about the \textit{nature} of the model: for any variable $X$ in the model $\mathcal{G}$ there are no causes of $X$ \textit{outside} of the model $\mathcal{G}$ \citep{spirtes2000causation,pearl2009causality,sucar2015probabilistic}.
 The second required condition is that the causal mechanism remains unchanged after interventions, this is called in \citep{woodward2005making} as \textit{invariantness}, this basically means that an intervention does not \textit{break} the causal mechanism when intervening upon it.
@@
 It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph \citep{sucar2015probabilistic}.
-\subsubsection{The identifiability problem}
-Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
+%\subsubsection{The identifiability problem}
+%Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data?
 It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) \citep{peters2012identifiability,mooij2016distinguishing}.
-
-\subsubsection{Do-Calculus}
-The Do-Calculus is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data \citep{pearl1995causal,pearl2009causality}.
-
-It is important to introduce some notation.
+%
+%\subsubsection{Do-Calculus}
+%The Do-Calculus is a set of rules for manipulating probabilistic statements that involve interventions and, under certain conditions, allow them to be transformed into statements that do not involve interventional data \citep{pearl1995causal,pearl2009causality}.
+%	
+%It is important to introduce some notation.
 Consider a causal graphical model $\mathcal{G}$ and $X,Y,Z$ disjoint sets of nodes of $\mathcal{G}$.
@@
 Finally, $\mathcal{G}_{\underline{Z}\bar{X}}$ is the graph obtained by deleting edges incoming into $X$ and outgoing from $Z$.
-
-\begin{theorem}[Do-Calculus, \cite{pearl2009causality}]{\label{docalculus}}
-Let $\mathcal{G}$ a CGM and $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
-\begin{itemize}
-\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ and $W$, then
-\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
-\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ and $W$, then
-\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
-\item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.
+%	
+%\begin{theorem}[Do-Calculus, \cite{pearl2009causality}]{\label{docalculus}}
+%Let $\mathcal{G}$ a CGM and $P_{\mathcal{G}}$ the probability measure induced by the model; then, for disjoint sets of nodes $X,Y,Z,W$ it holds:
+%\begin{itemize}
+%\item If for the graph $\mathcal{G}_{\bar{X}}$ it holds that $Y$ is conditionally independent from  $Z$ given $X$ and $W$, then
+%\[ P_{\mathcal{G}}(Y=y | do(X=x), Z=z, W=w) = P_{\mathcal{G}}(Y=y | do(X=x), W=w). \]
+%\item If for the graph $\mathcal{G}_{\bar{X}\underline{Z}}$ it holds that $Y$ is conditionally independent from $Z$ given $X$ and $W$, then
+%\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), Z = z, W=w). \]
+%\item Let $Z(W)$ the set of nodes in  $Z$ that aren't ancestors of any node in  $W$ in the graph $\mathcal{G}_{\bar{X}}$.
 If $Y$ is conditionally independent from $Z$ given $X$ and $W$ in the graph $G_{\bar{X}, \bar{Z(W)}}$, then
-\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
-\end{itemize}
-\end{theorem}
-\begin{proof}
-See \cite{pearl2009causality}.
-\end{proof}
-\begin{theorem}[Properties of Do-calculus]{\cite{peters2017elements}}\\
-The following statements hold:
-\begin{itemize}
-\item The Do-Calculus is complete; this is, sufficient for deriving every identifiable interventional distribution \citep{huang2006pearl,shpitser2006identification}.
-\item There exists an algorithm capable of finding all of the identifiable interventions \citep{tian2002,huang2006pearl}.
-\item A necessary and sufficient criteria exists for the identifiability of interventional distributions \citep{shpitser2006identification,huang2006pearl}.
-\end{itemize}
-\end{theorem}
-\begin{proof}
-See \cite{peters2017elements}.
-\end{proof}
-The Do-Calculus rules provide a solution for the identifiability problem:
-\begin{corollary} \cite{pearl2009causality}.\\
-A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each one of them corresponds to any of the rules of the Do-Calculus, that reduce $q$ to a purely observational expression.
-\end{corollary}
+%\[ P(Y=y | do(X=x), do(Z=z), W=w) = P(Y=y | do(X=x), W=w). \]
+%\end{itemize}
+%\end{theorem}
+%\begin{proof}
+%See \cite{pearl2009causality}.
+%\end{proof}
+%\begin{theorem}[Properties of Do-calculus]{\cite{peters2017elements}}\\
+%The following statements hold:
+%\begin{itemize}
+%\item The Do-Calculus is complete; this is, sufficient for deriving every identifiable interventional distribution \citep{huang2006pearl,shpitser2006identification}.
+%\item There exists an algorithm capable of finding all of the identifiable interventions \citep{tian2002,huang2006pearl}.
+%\item A necessary and sufficient criteria exists for the identifiability of interventional distributions \citep{shpitser2006identification,huang2006pearl}.
+%\end{itemize}
+%\end{theorem}
+%\begin{proof}
+%See \cite{peters2017elements}.
+%\end{proof}
+%The Do-Calculus rules provide a solution for the identifiability problem:
+%\begin{corollary} \cite{pearl2009causality}.\\
+%A distribution $q=P(y_1,...,y_k | do(x_1),...,do(x_n))$ is identifiable in a causal graphical model $\mathcal{G}$ if there exists a finite sequence of transformations, where each one of them corresponds to any of the rules of the Do-Calculus, that reduce $q$ to a purely observational expression.
+%\end{corollary}
 
@@
 \begin{definition}
-We define a Causal Decision Problem (CDP) as a tuple $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$ where $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C})$ is a Causal Environment and $\succeq$ is a preference relation.
+We define a Causal Decision Problem (CDP) as a tuple\\ $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C},\succeq)$ where $(\Omega, \mathcal{A}, \mathcal{G},\mathcal{E},\mathcal{C})$ is a Causal Environment and $\succeq$ is a preference relation.
 \end{definition}
@@
 It is known that human beings conceive their actions on the world as \textit{intervening} in the world \citep{hagmayer2009decision}.
-Following this idea, \cite{lattimoreNIPS2016} consider decision problems in which the action to be chosen is an intervention over a known causal graphical model.
+Following this idea, \cite{lattimoreNIPS2016} consider decision problems in which the action to be chosen is an intervention over a causal graphical model in which the conditional probabilities are assumed to be known.
 The agent must choose the intervention that maximizes the value of a \textit{target variable} after a series of learning rounds.
@@
 From the rewards they estimate probabilities and output an optimal action in the sense of obtaining minimal regret.
-Their work considers that the causal model is fully known, restriction which is partially relaxed by \cite{sen2017identifying}, and mention that the case in which the causal model is unknown is left as an open question, and this is precisely what we are proposing to answer.
+The assumption by \cite{lattimoreNIPS2016} of the causal model being known, is later partially relaxed by \cite{sen2017identifying}.
+In the former, it is mentioned that the case in which the causal model is unknown is left as future work.
+In our work, we study decision making over unknown causal information.
 
@@
 & \textrm{ if and only if }&\\
- &\sum_{c \in C} P(c | do(a))u(c) \geq \sum_{c \in C} P(c | do(b))u(c).&
+ &\sum_{c \in C} P_G(c | do(a))u(c) \geq \sum_{c \in C} P_G(c | do(b))u(c).&
 \end{eqnarray*} 
@@
 \begin{equation}
-a^\ast = \textrm{argmax}_{a \in \mathcal{A}} \sum_{c \in C} P(c | do(a))u(c).
+a^\ast = \textrm{argmax}_{a \in \mathcal{A}} \sum_{c \in C} P_G(c | do(a))u(c).
 \end{equation}
@@
 This Theorem assumes the existence of a Causal Graphical Model known by the decision maker; then, by Theorem \ref{vNM}, we know that the action that must be chosen is the one that maximizes the expected utility; we must now ask the expected utility of what?
-The answer lies on the Causal Graphical Model which controls the relation between actions and consequences; if this is the case, then the probability of a consequence given an action $a$ is in fact $P(c|do(a))$.
+The answer lies on the Causal Graphical Model which controls the relation between actions and consequences; if this is the case, then the probability of a consequence given an action $a$ is in fact $P_G(c|do(a))$.
 The rest follows by Theorem \ref{vNM}.
-Details in Section 4.1 of \cite{pearl2009causality}.
+Details and discussion in Section 4.1 of \cite{pearl2009causality}.
 \end{proof}
@@
 If only a part of the set of variables is known, then a Bayesian Nonparametric prior, which generates causal graphs of distinct sizes and complexity, could capture the patterns of causal interaction between known and possible but unknown variables.
+This would allow to relax the requirement of causal sufficiency.
 
